### 2009-04-12 kth/branches/year1/doc/cogx-kth-git/source/cogx-kth-git.tex
Added manual for using Git with CogX-KTH code. Squashed commit of the following:
commit 1d0b60ab61411de698ed1c4b2c21e0c73c471507
Author: Andrzej Pronobis <pronobis@csc.kth.se>
Date:   Sun Apr 12 15:27:58 2009 +0200

    Spellchecking.

commit 1e7bae03d882a0c9646fdd6cbae24fe133111c84
Author: Andrzej Pronobis <pronobis@csc.kth.se>
Date:   Sun Apr 12 15:26:28 2009 +0200

    Generated PDF file for the document.

commit 2ed083cdc043c1cde09d0d691a931ac9e55f44b8
Author: Andrzej Pronobis <pronobis@csc.kth.se>
Date:   Sun Apr 12 15:25:41 2009 +0200

    Updated the document.

commit e9543bfbeb3381a308

\documentclass[a4paper,10pt]{article}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{color,soul}

% ----------------------------------------------------
\definecolor{hlcolor}{rgb}{1,1,0}
\sethlcolor{hlcolor}

\definecolor{light-gray}{gray}{0.90}
\lstset{language=bash}
\lstset{backgroundcolor=\color{light-gray}}

% ----------------------------------------------------
\title{Using Git to Manage CogX KTH Code}
\author{Andrzej Pronobis}

% ----------------------------------------------------
\begin{document}
\maketitle

% ----------------------------------------------------
\section{Benefits of Using Git}

\begin{itemize}
\item \textit{Branches and tags are managed by Git}. Merging can be performed easily without the need to manually track the origins of the code and worry about the same code being merged multiple times.
\item \textit{Distributed repositories}. The work can be performed locally and off-line. Branching/merging/committing does not require connection to any central server.
\item \textit{Cheap (topic) branches}. Creating local branches e.g. for testing/implementing a feature, maintenance etc. is quick and easy. A workflow based on topic branches allows to create a higher quality code and maintain stability of the master branch.
\item \textit{Commits are uniquely identified by an SHA hash code}. The SHA code is calculated based on the contents of the commit and therefore is the same even for unrelated repositories.
\item \textit{Non-linear history}. Git tracks the history of all branches and merges.
\item \textit{Performance}. Git is fast.
\end{itemize}

% ----------------------------------------------------
\section{Documentation}
\begin{itemize}
  \item Git Community Book \\
       \texttt{http://book.git-scm.com/}
  \item Git for the lazy \\
       \texttt{http://www.spheredev.org/wiki/Git\_for\_the\_lazy}
  \item Git workflow for agile teams \\
       \texttt{http://reinh.com/blog/2009/03/02/a-git-workflow-for-agile-teams.html}
  \item Git for computer scientists \\
        \texttt{http://eagain.net/articles/git-for-computer-scientists/}
\end{itemize}

% ----------------------------------------------------
\section{Installing Git}
\subsection{Ubuntu Linux}
  \begin{lstlisting}
apt-get install git-core git-gui gitk
  \end{lstlisting}

\subsection{Windows}
  Full windows installer can be found at \\
  \texttt{http://code.google.com/p/msysgit/downloads/list}

% ----------------------------------------------------
\section{Setting Up the Local Repository}
\begin{itemize}
  \item Get access to the remote repository. Generate your public SSH key:
    \begin{lstlisting}
ssh-keygen -t rsa
    \end{lstlisting}
    and send the file \texttt{$\sim$/.ssh/id\_rsa.pub} to the admin of the repository.
  \item Clone the remote repository into a local directory:
    \begin{lstlisting}
git clone git@cogvis.nada.kth.se:cogx-kth.git
    \end{lstlisting}
  \item Create local branches tracking the remote branches e.g. \texttt{year1}:
    \begin{lstlisting}
git branch --track year1 origin/year1
    \end{lstlisting}
\end{itemize}

% ----------------------------------------------------
\section{Workflow for the CogX-KTH Code}
\subsection{Overview}
\includegraphics[width=\textwidth]{diagram}

\subsection{Feature Development}
\begin{itemize}
  \item Checkout the master branch:
    \begin{lstlisting}
git checkout master
    \end{lstlisting}
  \item Pull updates from the central server and merge them to master:
    \begin{lstlisting}
git pull
    \end{lstlisting}
  \item Create and checkout a topic branch for the new feature:
    \begin{lstlisting}
git checkout -b topic
    \end{lstlisting}
  \item Do some work in the topic branch:
    \begin{lstlisting}
echo "contents" > file
git add file
git commit -m "First commit"
echo "contents2" > file2
git add file2
git commit -m "Second commit"
    \end{lstlisting}
  \item Merge with the master branch:
    \begin{lstlisting}
git checkout master
git merge --squash topic
git commit
    \end{lstlisting}
  \item Push the changes upstream:
    \begin{lstlisting}
git push
    \end{lstlisting}
  \item Do more work in the topic branch:
    \begin{lstlisting}
git checkout topic
echo "new contents" > file
git add file
git commit -m "Third commit"
    \end{lstlisting}
  \item Pull some new updates from the central server and merge them to master:
    \begin{lstlisting}
git checkout master
git pull
    \end{lstlisting}
  \item Rebase the topic branch against the new changes:
    \begin{lstlisting}
git checkout topic
git rebase master
    \end{lstlisting}
  \item Merge the changes with the master branch:
    \begin{lstlisting}
git checkout master
git merge --squash topic
git commit
    \end{lstlisting}
  \item Push the changes upstream:
    \begin{lstlisting}
git push
    \end{lstlisting}

\subsection{Pushing Changes to SVN}
\hl{Missing}

\subsection{Pulling Changes from SVN}
\hl{Missing}

\end{itemize}

% ----------------------------------------------------
\section{FAQ}
\begin{enumerate}
  \item How to list branches and check which branch is checked out?
    \begin{lstlisting}
git branch
    \end{lstlisting}
  \item How to list all branches including remote branches?
    \begin{lstlisting}
git branch -a
    \end{lstlisting}
  \item How to view history of a branch or all branches?
    \begin{lstlisting}
gitk
gitk --all
    \end{lstlisting}

\end{enumerate}

\end{document}

### 2009-05-14 code/binding/development/specifications/bham_meeting/scenarios/playmate/bham_scripts.tex
very very first commit

\documentclass{article}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdwlist}
\usepackage{graphicx}

\title{Bham scenario specification}
\date{\today}
\author{CoSy}

\begin{document}
\maketitle
\begin{abstract}
Specification of the Bham scenarios...
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Script -- level 1}
Some of the science demostrated through this scenario:
\begin{enumerate}
\item Planning of sensing actions and visual processing.
\item Planning of clarifying questions.
\item Ability to recognise objects from a variety of views.
\item Ability to recognise complex actions (composed of sequences of
  actions).  
\item Ability to use learned categories for shape, colour and
  projective spatial relations.
\item Ability to understand which events in a sequence an utterance
  refers to.
\item Ability to reason about episodes and sequences of events.
\end{enumerate}

Before the demo we could record movies of the robot learning about
shape, colour and projective relations using the objects to be used in
the demonstration. Also, before the demo the games will be learned
off-line, using the complex action recognition system developed at
Birmingham. Learning on-line will be too challenging.

The objects to be used in the demonstration are (other than the human):
\begin{itemize}
\item \emph{part1}: a red triangle, a blue triangle, a blue square,
  and a red square, a blue car and a red car (cars being used as
  beacons/flags to which the shapes will be moved to during the
  color/shape game).
\item \emph{part2}: a car, and an occluding object, say a wall (car
  placed behind the wall) or a box (car placed within the box). 
\end{itemize}

In the shape game triangles go to the left, squares to the right.  In
the color game shapes go to the car of their color, wherever it is.

For Part1, the human sits next to the robot. There are two cars as
landmarks either side of the table in front of the robot. A red one is
on the left. A blue one is on the right. The script is as follows:

Recognition of complex action sequences from vision + recognizing the
game being played:
\begin{enumerate}
\item \label{lvl1:step1} 
  H: ``I am playing a game.''\\
  R: ``OK.''

\item \label{lvl1:step2} The human puts a red triangle in front of the
  robot. He retracts his arm.

\item \label{lvl1:step3} The human reaches out, grasps the triangle,
  and pushes it to the left car. He then pulls back the hand.

\item \label{lvl1:step4} The human withdraws the shape away from the
  left car (so that it is not in view).

\item \label{lvl1:step5}
  H: ``Which game am I playing?''\\
  R: ``I do not know, the shape game or the colour game.''

\item \label{lvl1:step6} The human puts a red square in front of the
  robot.  Same as Step~\ref{lvl1:step2}

\item \label{lv1:step7} The human moves the red square to the right
  car.  Same as Step~\ref{lvl1:step3}.

\item \label{lvl1:step8} The human withdraws the red square from the
  right car.  Same as Step~\ref{lvl1:step4}.

\item \label{lvl1:step9}
  H: ``Which game am I playing?''\\
  Same as Step~\ref{lvl1:step5}.

\item \label{lvl1:step10}
  R: ``You are playing the shape game.''\\
  We could instead have Step~\ref{lvl1:step6} happen here.

\item \label{lvl1:step11}
  The human withdraws the red square from the right car.
  same as Step~\ref{lvl1:step8}.

\item \label{lvl1:step12}
  The human swaps the cars.

\item \label{lvl1:step13}
  Human puts a red triangle down and withdraws hand.

\item \label{lvl1:step14}
  H reaches and pushes red triangle to the right car.

\item \label{lvl1:step15}
  H: ``Which game?''\\
  R: ``The color game.''

\item \label{lvl1:step16}
  H withdraws the red triangle.
  Same as in Step~\ref{lvl1:step4}

\item \label{lvl1:step17} (We now start on the ability to play the
  game, including the planning of sensory processing to support
  action).\\
  The human puts down the red triangle in front of the robot.
  Analysis similar to Step~\ref{lvl1:step2}

\item \label{lvl1:step18}  
  H: ``Play the colour game.''

\item \label{lvl1:step19} 
  The robot picks up the shape and puts it next to the red car on
  the right.

\item \label{lvl1:step20}
  H removes the shape (red triangle) from the car.\\
  See Step~\ref{lvl1:step4}.

\item \label{lvl1:step21}
  Human puts the blue square in front of the robot.\\
  H: ``Play the shape game.''\\
  The robot picks up the shape and puts it next to the right hand
  (red) car.
  See Step~\ref{lvl1:step17}-\ref{lvl1:step20}.
\end{enumerate}

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{example1_msc.png}
% \caption{
%   Some simple msc-graph can be generated to clarify interaction
%   between things in the the scenario. Use the
%   \texttt{../gen\_msc.sh}-script to generate pngs from all
%   msc-files. mscgen is documented here:
%   \texttt{http://www.mcternan.me.uk/mscgen/}. It's VERY simple}
% \label{fig:lvl1}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Script -- level 2}
The scripts are discussed here in 'slightly' more detail...

\begin{enumerate}
\item \label{lvl2:step1}
  H: ``I am playing a game.''\\
  R: ``OK.''
  \begin{enumerate}
  \item \label{lvl1:step1:detail0} The robot produces color and
    position representations of the cars + Car recognition (viewpoint
    planning may be involved here -- more details later).
  \item \label{lvl2:step1:detail1} Robot produces a representation
    (where?) that it is now ``in'' a game episode.
  \item \label{lvl2:step1:detail2} Robot produces as acknowledgement
    utterance.
  \end{enumerate}

\item \label{lvl2:step2}
  The human puts a red triangle in front of the robot.
  \begin{enumerate}
  \item \label{lvl2:step2:detail} The robot has to notice the object
    (and track the hand before this?).
  \item \label{lvl2:step2:detail2} The robot has to extract color,
    shape and position of objects.
  \end{enumerate}

\item \label{lvl2:step3}
  The human reaches out, grasps the triangle, and pushes it to the
  left car. He then pulls back the hand.
  \begin{enumerate}
  \item \label{lvl2:step3:detail1}
    The robot tracks the hand and the object throughout the movement.
  \item \label{lvl2:step3:detail2} The robot has to recognize the
    ``reach'', ``push to car'', ``retract'' sequence as a 'game
    move'.
  \item \label{lvl2:step3:detail3} The robot should recognize that it
    is playing the color game or the shape game.
  \item \label{lvl2:step3:detail4} The robot extracts shape, color and
    position in the new scene.
  \end{enumerate}

\item \label{lvl2:step4}
  The human withdraws the shape from the left car.
  \begin{enumerate}
  \item \label{lvl2:step4:detail1} The robot recognizes complex
    actions such as ``reach'', ``pull to human''.
  \item \label{lvl2:step4:detail2} The robot extracts color and
    position for the cars.
  \end{enumerate}

\item \label{lvl2:step5}
  H: ``Which game am I playing?''\\
  R: ``I do not know, the shape game or the color game.''
  \begin{enumerate}
  \item \label{lvl2:step5:detail1}Understand the utterance, and that
    it is being asked a question, and that it refers to a game.
  \item \label{lvl2:step5:detail2} Ascertain the answer and deliver
    it.
  \end{enumerate}

\item \label{lvl2:step6}
  The human puts a red square in front of the robot.
  Same as Step~\ref{lvl2:step2}

\item \label{lv2:step7}
  The human moves the red square to the right car.\\
  Same as Step~\ref{lvl2:step3}.

\item \label{lvl2:step8}
  The human withdraws the red square from the right car.\\
  Same as Step~\ref{lvl2:step4} \emph{except} that it concludes that it is
  playing the shape game.

\item \label{lvl2:step9}
  H: ``Which game am I playing?''\\
  Same as Step~\ref{lvl2:step5}.

\item \label{lvl2:step10}
  R: "You are playing the shape game."\\
  We could instead have Step~\ref{lvl2:step6} happen here.

\item \label{lvl2:step11}
  The human withdraws the red square from the right car. \\
  Instead we could have Step~\ref{lvl2:step8} happen here.

\item \label{lvl2:step12}
  The human swaps the cars.
  \begin{enumerate}
  \item \label{lvl2:step12:detail1} The robot will track the hand and
    cars. ``Reach'', ``pull'' actions.
  \item \label{lvl2:step12:details2} The robot re-extracts the new
    car positions, ``recognize car''.
  \end{enumerate}

\item \label{lvl2:step13}
  Human puts a red triangle down and withdraws hand.
  \begin{enumerate}
  \item \label{lvl2:step13:detail1} Robot extracts color and shape.
  \end{enumerate}

\item \label{lvl2:step14}
  H reaches and pushed red triangle to the right hand car.
  \begin{enumerate}
  \item \label{lvl2:step14:detail1} Robot recognizes the component
    actions ``reach'', ``push to car'', and the complex actions.
  \item \label{lvl2:step14:detail2} The robto recognizes that it is
    playing the color game.
  \end{enumerate}

\item \label{lvl2:step15}
  H: ``Which game?''\\
  R: ``The color game.''
  Same as Step~\ref{lvl2:step5} and Step~\ref{lvl2:step6}.

\item \label{lvl2:step16}
  H withdraws the red triangle.
  Same as in Step~\ref{lvl2:step4}

\item \label{lvl2:step17} (We now start on the ability to play the
  game, including the planning of sensory processing to support
  action).\\
  The human puts down the red triangle in front of the robot.
  Same as in Step~\ref{lvl2:step2}

\item \label{lvl2:step18}  
  H: ``Play the colour game.''
  \begin{enumerate}
  \item \label{lvl2:step18:detail1} Recognizes utteranc and
    understands that it is an imperative to play the color game.
  \item \label{lvl2:step18:detail2} Produces a ``move the red triangle
    to the red car (on the right)'' plan.
  \end{enumerate}

\item \label{lvl2:step19} The robot picks up the shape and puts it
  next to the red car on the right.
  \begin{enumerate}
  \item \label{lvl2:step19:detail1} Recognize the grasp
    location/surface (might involve moving the hand camera).
  \item \label{lvl2:step19:detail2} Plan a trajectory to grasp it.
  \item \label{lvl2:step19:detail3} Execute the visual servoing.
  \item \label{lvl2:step19:detail4} Check that it has succeeded in
    grasping the red triangle.
  \item \label{lvl2:step19:detail5} Move the red triangle to the red
    car and put it down.
  \item \label{lvl2:step19:detail6} Retract the arm.
  \item \label{lvl2:step19:detail7} Re-extract the shape, color and
    position of the red triangle.
  \item \label{lvl2:step19:detail8} Check that the goal has been
    reached.
  \end{enumerate}

\item \label{lvl2:step20}
  H removes red triangle.\\
  Analysis similar to Step~\ref{lvl2:step4}

\item \label{lvl2:step21}
  Human puts the blue square in front of the robot.\\
  H: ``Play the shape game.''\\
  The robot picks up the shape and puts it next to the right hand
  (red) car.\\
  See Step~\ref{lvl2:step17}-Step~\ref{lvl2:step20}.

\end{enumerate}

% \begin{figure}[!htcb]
% \centering
% \includegraphics[width=\linewidth]{example2_msc.png}
% \caption{
% A more detailed msc-graph could be used on this or next level...
% }
% \label{fig:lvl2}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Subarchitecture Level}

%%----------------------------------------------------------
\subsection{Step~\ref{lvl2:step1}}
Cars are in place when the system is switched on.

\subsubsection{Vision SA}
\begin{itemize}
\item Change detection (Bham-done).
\item ROIs + Scene Objects (Bham-done).
\item Feature information from ROIs (UoL) matlab wrapped + (Bham-done) Mohan.
\item ROIs - color classifier - Scene Object color (UoL) Alen + (Bham-done) Mohan.
\item ROIs - Car recognizer - Scene Object 'type' (TUD) Nico.
\end{itemize}

\subsubsection{Comsys SA}
\begin{itemize}
\item Speech Signal - speech recognition - text string (DFKI-done) Pierre.
\item Text - OpenCCG + Moloko mk4 - logical forms (DFKI-done) GJ, Trevor.
\item Text - Dialogue analysis - type of dialogue, say assertion (DFKI-done).
\end{itemize}

\subsubsection{Spatial SA}
\begin{itemize}
\item Scene Objects (VSA) - positions on the ground plane.
\item Positions - spatial relations.
\end{itemize}

\subsubsection{Binding SA}
\begin{itemize}
\item Proxies for car 1 and car2.
\item Unions created for car1 and car2.
\end{itemize}

\subsubsection{Motivation SA}
\begin{itemize}
\item Logical forms - posts motivation that I am playing a game.
\item Posts acknowledgement.
\end{itemize}

%%----------------------------------------------------------
\subsection{Step~\ref{lvl2:step2}}
Human puts red triangle in front of the robot.

\subsubsection{Vision SA}
\begin{itemize}
\item Tracking information (Bham-done) Somboon.
\item Change detection (Bham-done).
\item ROIs + Scene Objects (Bham-done).
\item Feature information: color, shape (Bham-done), possibly use
  (UoL) when ready.
\end{itemize}

\subsubsection{Spatial SA}
\begin{itemize}
\item Recognize ``putting'' action based on hand (and object)
  tracking.
\item Realize that something is in front of the robot.
\item Cars 1 and 2 recognized.
\end{itemize}

\subsubsection{Binding SA}
\begin{itemize}
\item Proxies for cars and objects.
\item Unions.
\end{itemize}

%%----------------------------------------------------------
\subsection{Step~\ref{lvl2:step3}}

%%----------------------------------------------------------
\subsection{Step~\ref{lvl2:step4}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Dependency Level}

\textit{
  What else is required to generate the above information and
  behaviour. Remember the already existing specs prepared prior to the
  meeting: \texttt{http://www.dfki.de/\~\ henrikj/doc\_test/specifications/main/html/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Component Level}

\textit{
  First pass as describing the behaviours in terms of the components
  that are involved.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Responsibility Level}
Some details specified at the subarchitecture-level specifications (level 3)...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2009-05-14 code/binding/development/specifications/bham_meeting/scenarios/template/template.tex
very very first commit

\documentclass{article}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdwlist}
\usepackage{graphicx}

\title{Template scenario specification for Bham meeting}
\date{\today}
\author{CoSy}

\begin{document}
\maketitle
\begin{abstract}
This template defines a structure for the documentation and workflow
of the Wednesday meeting. Look at the \texttt{../yr3example/} for a
more detailed example of what is expected. 'svn cp' this entire
template dir to your own directory before you start (one per
\emph{sub}scenario is desired, i.e.\ more than one for explorer and
playmate respectively). Rename template.tex to sth. sensible and then
get going and delete the instructions as you progress. Good luck and
keep things simple! At least initially :-)
\end{abstract}

\section{Script -- level 1}
\textit{
  The script should describe the ``surface'' level of the interaction
  between the robot, the user and the environment of the robot. This
  serves as a basis for the more detailed account of the processes in
  the architecture in the following sections.}

\textit{
  make sure you keep track of the \texttt{label}s... it will make it easier for
  you to refer back to the script later on. }

\begin{enumerate}
\item \label{lvl1:sth}
  sth happens
\item \label{lvl1:sth2}
  sth else happens
\end{enumerate}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{example1_msc.png}
\caption{
  Some simple msc-graph can be generated to clarify interaction
  between things in the the scenario. Use the
  \texttt{../gen\_msc.sh}-script to generate pngs from all
  msc-files. mscgen is documented here:
  \texttt{http://www.mcternan.me.uk/mscgen/}. It's VERY simple}
\label{fig:lvl1}
\end{figure}

\section{Script -- level 2}
\textit{
  The level 1 script describes what can obviously be seen when the
  robot is interacting, for example in a video uptake of the
  scenario. The next level is to describe, without any technical
  detail, what activities take place within the robot during this
  interaction. }

\textit{
  make sure the enumeration from level 1 is preserved... it makes the
  whole thing easier to read. Tip: finish level 1 first, and copy
  paste the enumeration and make subenumerations as below.}

\begin{enumerate}
\item \label{lvl2:sth}
  sth happens
\begin{enumerate}
\item \label{lvl2:sth:detail}
    sth happens inside the robot
\item \label{lvl2:sth:detail2}
    sth else happens inside the robot
\end{enumerate}
\item \label{lvl2:sth2}
  sth else happens
\begin{enumerate}
\item \label{lvl2:sth2:detail}
    sth happens inside the robot
\item \label{lvl2:sth2:detail2}
    sth else happens inside the robot
\end{enumerate}
\end{enumerate}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{example2_msc.png}
\caption{
A more detailed msc-graph could be used on this or next level...
}
\label{fig:lvl2}
\end{figure}

\section{Subarchitecture Level}
\textit{
  Describe for each step in the detailed lvl2 script what the involved
  subarchitectures are and how they are involved.}

\subsection{\ref{lvl2:sth}}
\subsubsection{Subarchitecture X}
\begin{itemize}
\item do this
\item do that
\end{itemize}

\subsubsection{Subarchitecture Y}
\begin{itemize}
\item do this
\item do that
\end{itemize}

\subsection{\ref{lvl2:sth:detail}}
...
\subsection{\ref{lvl2:sth:detail2}}
...
\subsection{\ref{lvl2:sth2}}
...
\subsection{\ref{lvl2:sth2:detail}}
...
\subsection{\ref{lvl2:sth2:detail2}}
...

\section{Dependency Level}

\textit{
  What else is required to generate the above information and
  behaviour. Remember the already existing specs prepared prior to the
  meeting: \texttt{http://www.dfki.de/\~\ henrikj/doc\_test/specifications/main/html/}}

\section{Component Level}

\textit{
  First pass as describing the behaviours in terms of the components
  that are involved.}

\section{Responsibility Level}

\textit{
  Who's doing what!}

\end{document}

\end{document}

### 2009-05-14 code/binding/development/specifications/bham_meeting/scenarios/yr3example/example.tex
very very first commit

\documentclass{article}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdwlist}
\usepackage{graphicx}

\title{Example Specification Process for Year 3 PlayMate}
\date{\today}
\author{Nick and Henrik}

\begin{document}

\maketitle

\section{Script -- level 1}

\begin{enumerate}
\item \label{lvl1:redbox}
  human puts down red box
\item \label{lvl1:bluebox}
  human puts down blue box to left of right box
\item \label{lvl1:human}
  human says "put the blue thing to the right of the red thing"
\item \label{lvl1:success}
  robot does this successfully.
\end{enumerate}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{example1_msc.png}
\caption{
A sketch of the interaction between the involved systems.
}
\label{fig:lvl1}
\end{figure}

\section{Script -- level 2}

\begin{enumerate}

\item \label{lvl2:redbox}
  human puts down red box
\begin{enumerate}
\item \label{lvl2:redbox:notice}
    robot notices object,
\item \label{lvl2:redbox:extract}
    extracts colour and position information
\end{enumerate}

\item \label{lvl2:bluebox}
  human puts down blue box to left of right box
\begin{enumerate}
\item \label{lvl2:bluebox:notice}
    robot notices object,
\item \label{lvl2:bluebox:extract}
    extracts colour and position information,
\item \label{lvl2:bluebox:spatial}
    generates spatial relations between objects (blue object left of
    right object, blue object near red object, red object near blue
    object)
\end{enumerate}

\item \label{lvl2:human}
  human says "put the blue thing to the right of the red thing"
\begin{enumerate}
\item \label{lvl2:human:speech}
    robot recognises speech and creates a logical interpretation of it
\item \label{lvl2:human:thing}
    robot resolves the "thing" references in interpretation to the
    objects in the scene
\end{enumerate}

\item \label{lvl2:success}
  robot does this successfully.
\begin{enumerate}
\item \label{lvl2:success:goalgen}
    robot detects that the utterance is a command, and represents the
    the content of the command as a behaviour to perform.
\item \label{lvl2:success:plangen}
    robot generates a plan to pick up the blue object and move it to
    the right of the red object
\item \label{lvl2:success:manip}
    robot picks up the blue object and moves it to the right of the
    red object
\item \label{lvl2:success:verify}
    robot looks at the world to verify that the plan has been executed
    correctly
\end{enumerate}

\end{enumerate}

\section{Subarchitecture Level}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\linewidth]{example2_msc.png}
\end{center}
\caption{
A sketch of the interaction between the involved subsystems.
}
\label{fig:lvl2}
\end{figure}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:redbox:notice}}

\subsubsection{Visual Subarchitecture (VSA)}

\begin{itemize}

\item notification of scene change

\item representation of red box generated

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:redbox:extract}}

\subsubsection{VSA}

\begin{itemize}

\item colour value of red box

\item 3D position of red box

\end{itemize}

\subsubsection{Spatial Subarchitecture (SSA)}

\begin{itemize}

\item 2D position of red box on table

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:bluebox:notice}}

\subsubsection{VSA}

\begin{itemize}

\item notification of scene change

\item representation of blue box generated

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:bluebox:extract}}

\subsubsection{VSA}

\begin{itemize}

\item colour value of blue box

\item 3D position of blue box

\end{itemize}

\subsubsection{SSA}

\begin{itemize}

\item 2D position of blue box on table

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:bluebox:spatial}}

\subsubsection{SSA}

\begin{itemize}

\item left-of, blue object, red object

\item right-of, red object, blue object

\item near, blue object, red object

\item near, red object, blue object

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:human:speech}}

\subsubsection{Communication Subsubsection (CSA)}

\begin{itemize}

\item recognised speech string

\item logical form representing speech string

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:human:thing}}

\subsubsection{CSA}

\begin{itemize}

\item linguistic representation of ``red thing''

\item linguistic representation of ``blue thing''

\end{itemize}

\subsubsection{Binding Subarchitecture (BSA)}

\begin{itemize}

\item union binding together linguistic and visual representations of
  the red box

\item union binding together linguistic and visual representations of
  the blue box

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:goalgen}}

\subsubsection{BSA}

\begin{itemize}

\item Representation of command structure including arguments.

\end{itemize}

\subsubsection{Mediation Subarchitecture (MSA)}

\begin{itemize}

\item Description of the desired state of the world: right-of, blue
  object, red object

\item Goal to achieve this state

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:plangen}}

\subsubsection{MSA}

\begin{itemize}

\item Motivation to achieve goal state

\end{itemize}

\subsubsection{Planning Subarchitecture (PSA)}

\begin{itemize}

\item Plan to pick up blue thing, place it to the right of red thing

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:manip}}

\subsubsection{Manipulation Subarchitecture (ASA)}

\begin{itemize}

\item Actions to pick up blue box, place down at position to right of
  red box.

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:verify}}

\subsubsection{VSA}

\begin{itemize}

\item New position information for blue box

\end{itemize}

\subsubsection{SSA}

\begin{itemize}

\item new 2D position of blue box

\item right-of, blue object, red object

\item left-of, red object, blue object

\item near, blue object, red object

\item near, red object, blue object

\end{itemize}

\subsubsection{PSA}

\begin{itemize}

\item Representation of current state and desired state, and
  comparison

\end{itemize}

\section{Dependency Level}

What else is required to generate the above information and behaviour.

\section{Component Level}

First pass as describing the behaviours in terms of the components
that are involved.

\section{Responsibility Level}

Who's doing what!

\end{document}

% LocalWords:  CSA MSA PSA

### 2009-05-14 code/binding/development/subarchitectures/comsys.mk4/doc/comsys.mk4/main.incrsitdial.tex
very very first commit

%% ==============================================================
%% DOCUMENTCLASS 
%% ==============================================================

\documentclass[onecolumn,11pt]{elsart}

%% ==============================================================
%% PACKAGES 
%% ==============================================================

%% ==============================================================
%% NEW COMMANDS
%% ==============================================================

\newcommand{\cmpn}[1]{\textsf{#1}}
\newcommand{\comsys}{\textsc{comsys.mk4}}
\newcommand{\data}[1]{\texttt{#1}}

%% ==============================================================
%% DOCUMENT - FRONTMATTER
%% ==============================================================

\begin{document}

\begin{frontmatter}
\title{Incremental processing of situated human-robot dialogue \thanksref{cosy}}
\author{Geert-Jan M. Kruijff}
\address{DFKI GmbH, Saarbr\"ucken Germany\\ $\langle$\emph{\texttt{gj@dfki.de}}$\rangle$}
\thanks[cosy]{Thanks to ... CoSy, Kerstin Hadelich, Maria Staudte}

\input{abstract.incrsitdial}

\end{frontmatter}

%% ==============================================================
%% DOCUMENT - BODY MATTER
%% ==============================================================

\section{Introduction}

\section{Implementation}

%% ==============================================================
%% DOCUMENT - APPENDIX MATTER
%% ==============================================================

\appendix

\section{Code documentation}

\subsection{IDL structures / ontological data types}

The \comsys\ working memory contains data structures of the following ontological types.

\data{LexicalRequest} is produced by the \cmpn{GrammaticalInference} component, and contains fields for specifying the word form for which lexical data is being requested.

\paragraph{LexicalData}

\end{document}

### 2009-05-14 code/binding/development/subarchitectures/comsys.mk4/grammars/openccg/moloko.v6/docs/comsys.moloko/dual-relation.moloko.tex
very very first commit

\section{Dual-Relation Words} \label{sec-Dual-Relation-Words}

In section \ref{sec-Semantic-Relations}, we saw that some word classes need to be able to play the role of both modifiers and dependents. We will first consider how this is handled in 'traditional' \textsc{CCG} and outline its limitations.  We will then describe how these were overcome in Moloko.

 \subsection{Dual-Relation words in standard \textsc{CCG} }

In standard \textsc{CCG}, the dual functionality of such words is handled by assigning them two entirely separate lexical families. So, for example, the adjective \emph{bigger} would receive:

\begin{enumerate}
\item \catg{adj\id{M}}    :    \atsign{b1}{size}( \wand\  \prop{big} \wand\ \diam{Degree}(comparative) )
\item \catg{n\id{T} \fwdsl{} n\id{T} }  : \\ \atsign{t1}{entity}(  \diam{Modifier}(b1:size \wand\  \prop{big} \wand\ \diam{Degree}(comparative) ) )
\end{enumerate}
In the first case, \emph{bigger} is given an atomic category \catg{adj} and thus provides it with its own semantic head M. This variable can then be 'selected for', and then used to fill a dependency role, allowing \emph{bigger} to act as a dependent. Consider again the sentence \emph{I made it bigger} from above. At the point in the incremental parsing of this sentence, just before the word \emph{bigger} is reached, we have this partial parse \\

parse: \catg{s \fwdsl{} adj } 
\vspace{-1.5em} %% some negative vertical spacing

\begin{align*}
\atsign{m1}{action-non-motion}(& \prop{make}  \\ 
      & \wand\ \diam{Mood}(ind) \\
      & \wand\ \diam{Actor}(i1:person \wand\ \prop{I} \wand\ \diam{Num}(sg))  \\ 
      & \wand\ \diam{Patient}(b1:thing \wand\ \prop{it} \wand\ \diam{Num}(sg))  \\
      & \wand\ \diam{Result}x1:quality )
\end{align*}
In order for this sentence to recieve a complete parse, syntactically it 'needs' an adjective (atomic cat \catg{adj} ) to its left, an adjective whose nominal variable would semantically fill its currently empty  \diam{Result} role. When the word \emph{bigger} is encountered next, lexical reading 1) above is employed, and the variable M fills this missing role.

As for the second reading above, \emph{bigger} simply adds its semantic content to the modifee (in this case T, the entity designated by the noun). This is much clearer if we consider how \textsc{CCG}, being a dependency grammar, treats modifiers. Syntactically, modifiers are treated  as identity functions: they take an argument of category X and return the same category X. This is done by giving modifiers a complex category, one which specifies both the modifee's syntactic category and its combinatorial possible locations (via its slash direction and mode). The intuition here, is that despite their semantic differences, syntactically \emph{dog} behaves identically to \emph{big dog},  or \emph{big scary dog} or even \emph{big scary dog running right towards you}.  It is instead on the semantic side that modifiers do their work: they take the nominal variable associated with their modifee and attach some 'new' semantic content to it, that is they modify it.

Unfortunately, by creating two separate readings--one for each relation--we are forced into a few unexpected limitations. First, consider the following two sentences, where the examples from section \ref{sec-Semantic-Relations} have been slightly altered by the addition of the word \emph{much}
\begin{itemize}
\item	 I made it much bigger
\item	 I wanted the much bigger picture
\end{itemize}
Cases like these can be handled quite simply by treating the word \emph{much} as an adjective pre-modifier (Note: of type 'degree')\\

\catg{adj\id{M} \fwdsl{} adj\id{M} }  : \atsign{m1}{modifier}(  \diam{Modifier}(m2:degree \wand\  \prop{much}  ) ) \\\\
This then combines with \emph{bigger} creating \emph{much bigger}, a 'new' adjective whose nominal variable M is given the semantic content of \emph{much}:

\vspace{-1.5em} %% some negative vertical spacing
\begin{align*}
\atsign{b1}{size}(& \prop{big}  \\ 
      & \wand\ \diam{Degree}(comparative) \\
      & \wand\ \diam{Modifier}(m2:degree \wand\  \prop{much} )) 
\end{align*} 
This can then act as a dependent, e.g. filling the \diam{Resultant} role in \emph{made}, in exactly the same way as a simple, unmodified adjective.

It is in the second sentence that we enounter a problem. As we saw above, modifiers require a nominal variable to attach their 'modifying' semantic content to. This is true of any linguistic unit acting as modifier, whether it is a single word like \emph{big} or \emph{quickly}, or a larger constituent like \emph{on the table} , \emph{that I told you about} or \emph{when you come back}. It is also true regardless of the type of the modifee. This was the case in section \ref{sec-Semantic-Relations}, and also with \emph{much} directly above: because the atomic category version of adjectives already has a nominal variable (to be selected for as a dependent), it can easily be modified. However, this is not the case for the 'modifier' reading of adjectives, prepositions and adverbs: currently, complex categories do not receive their own nominal variables in \textsc{OpenCCG}. So in the case of \emph{bigger} above, the nominal variable T can be used to 'refer' to the modified entity, but unlike the dependent-version in 1), there is no separate nominal variable refers to the adjective as a whole. Consequently, there is no way to modify an adjective operating in this way; or more generally, it is impossible to modify a modifier.

This is a signicant problem. It rules out the possibility of parsing, or generating sentences like these:
\begin{itemize}
\item	Adverb Modifiers: go to the kitchen \underline{very slowly}, please don't do it \underline{so quickly}
\item	Adjective Modifiers:	I want the \underline{really big} one
\item Preposition (pp) Modifiers: walk \underline{over to GJ}, walk \underline{up to the table} 
\end{itemize}									 
A second highly similar problem is that without semantic heads, modifiers performing the modifying role cannot be conjoined, i.e. sentences like \emph{the \underline{green and yellow} ones  }, \emph{put it on the table \underline{quickly but carefully}  } and \emph{I walked \underline{in the room and around the desk}} cannot be handled.

\subsection{Dual-Relation Words in Moloko} \label{sec-Dual-Relation-in-Moloko}

These limitations were addressed in the Moloko grammar through the use a single reading for dual-relation words in combination with type-changing rules. Adjectives, prepositions and adverbs are given only an atomic category reading, and consequently always have a nominal variable, allowing them to serve as dependents (including being modified). To perform their modifying function, these basic categories can be 'transformed' or 'exploded' into their complex combinatorial form. For example, Moloko contains this type-changing rule for adjectives: \\

\catg{adj\id{M}}   \arrow \catg{n\id{T}\fwdsl{} n\id{T} }  : \atsign{t1}{entity}(  \diam{Modifier}(m1) ) ) \\ \\
The result is the familiar pre-nominal adjective reading, but instead of the  \diam{Modifier} relation receiving its contents off-line (directly from the lexicon), this rule dynamically copies the contents of the atomic version's nominal variable (M) into the 'open' dependency relation: in some sense it turns modification into some form of bizarre dependency. What is crucial here is that this rule acts on any linguistic constituent with category adj encountered during parsing, including already modified adjs.
Thus, while parsing \emph{I want the really big one}, the rule can operate either on {big} or on {really big}\footnote{ in actuality, it operates on \emph{really} alone. See the second example in the section on Incrementalilty ( \ref{sec-Incrementality} ).}

By continually using adjectives as an example, we have masked a very important fact: most modifiers have a variety of combinatorial possibilities. Consider the adverb \emph{normally}. It can placed in a variety of locations within the clause:
\begin{itemize}
\item	\emph{normally} you can get it from GJ
\item	you \emph{normally} can get it from GJ
\item	you can \emph{normally} get it from GJ
\item	you can get it from GJ \emph{normally}
\end{itemize}
These different syntactic permutations are handled by giving \emph{normally} multiple readings, each with a different syntactic form. In the standard \textsc{CCG} approach, this is done by giving these words multiple lexical entries. Of course this leads to the same problems mentioned above. In the Moloko grammar, we handle this variability by creating a type-changing rule for each syntactic possibility. So for example, here are two adverb type-changing rules:

\begin{itemize}
\item \catg{adv\id{M}}   \arrow \catg{s\id{E} \fwdsl{} s\id{E} }  : \atsign{e1}{event}(  \diam{Modifier}(m1) ) ) 
\item \catg{adv\id{M}}   \arrow \catg{s\id{E} \backsl{} s\id{E} }  : \atsign{e1}{event}(  \diam{Modifier}(m1) ) )
\end{itemize}

However, this still leaves one major issue unresolved: different words within the same
syntactic class can behave differently. Compare the adverb \emph{normally} to the adverb \emph{too}
(with the meaning of  \emph{also}):
\begin{itemize}
\item	* too you can get it from GJ
\item	you too can get it from GJ
\item	? you can too get it from GJ
\item	you can get it from GJ too
\end{itemize}
To account for these 'idiosyncracies', each word must somehow specify its behaviour. Again, in standard \textsc{CCG} this can be done easily by assigning a word to only its appropriate families. In Moloko, this is done instead by 'naming' each type-changing rule and for each word, specifying lexically which rules can apply to it. We could have created a binary syntactic feature for each rule, and thus basically would specify + values for each feature, and too would specify - for the rules it wishes to block. However, to
avoid this explosion of features, all of this information is instead contained in a single syntactic feature called cc-type (complex-category type). Here is a more detailed look at some of the rules used for adverbs:
\begin{itemize}
\item \catg{adv\fb{\fid{M}, cc-type=pre-s} }    \arrow  \\\catg{s\id{E} \fwdsl{} s\id{E}}  : \atsign{e1}{event}(  \diam{Modifier}(m1) ) ) 
\item \catg{adv\fb{\fid{M}, cc-type=pre-s}}   \arrow \\\catg{s\id{E} \backsl{} s\id{E}}  : \atsign{e1}{event}(  \diam{Modifier}(m1) ) )
\end{itemize}
These individual, rule-specific feature values are then grouped (via multiple inheritance) in the syntactic feature heirarchy. This creates a set of 'syntactic classes', each  specifying exactly which rules can apply (and can't apply) to words belonging to this class, and hence enforces the appropriate combinatorially behaviours.

Each of the Dual-Relation parts of speech (\catg{adj}, \catg{adv} and \catg{prep}) has its own set of syntactic classes (e.g. adv1, adv2, adv3, etc), and each word from these POS belongs to one of these classes. This is specified lexically via macro.

### 2009-05-14 code/binding/development/subarchitectures/binding.sa/src/java/BindingFeatures/SingularHelper.java

@@ -1,0 +1,90 @@
+package BindingFeatures;
+
+
+/**
+* BindingFeatures/SingularHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from BindingFeatures.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class SingularHelper
+{
+  private static String  _id = "IDL:BindingFeatures/Singular:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, BindingFeatures.Singular that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static BindingFeatures.Singular extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [3];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_alias_tc (cast.cdl.WorkingMemoryIDHelper.id (), "WorkingMemoryID", _tcOf_members0);
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "m_groupID",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().get_primitive_tc (org.omg.CORBA.TCKind.tk_long);
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "m_elementNumber",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = BindingFeaturesCommon.ParentFeatureHelper.type ();
+          _members0[2] = new org.omg.CORBA.StructMember (
[…]

### 2009-05-14 code/binding/development/subarchitectures/binding.sa/src/java/BindingQueries/AdvancedQueryHelper.java

@@ -1,0 +1,149 @@
+package BindingQueries;
+
+
+/**
+* BindingQueries/AdvancedQueryHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from BindingQueries.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class AdvancedQueryHelper
+{
+  private static String  _id = "IDL:BindingQueries/AdvancedQuery:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, BindingQueries.AdvancedQuery that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static BindingQueries.AdvancedQuery extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [9];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = BindingQueries.QueryParametersHelper.type ();
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "m_parameters",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = BindingData.FeaturePointerHelper.type ();
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "m_featurePointer",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
[…]

### 2009-05-14 code/binding/development/subarchitectures/coma.sa/src/java/coma/scraps/ComaBindingMonitor_old.java

@@ -1,0 +1,409 @@
+package coma.scraps;
+
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.Properties;
+
+import org.cognitivesystems.reasoner.base.OntologyMemberFactory;
+
+import ComaData.ComaConcept;
+import ComaData.ComaInstance;
+import ComaData.ComaReasonerFunction;
+import coma.aux.ComaFunctionWriter;
+import coma.aux.ComaHelper;
+import coma.aux.InstanceUnknownException;
+import coma.aux.ComaFunctionWriter.ConceptQueryRestrictor;
+
+import cast.architecture.abstr.ChangeFilterFactory;
+import cast.architecture.abstr.WorkingMemoryChangeReceiver;
+import cast.architecture.subarchitecture.SubarchitectureProcessException;
+import cast.cdl.WorkingMemoryChange;
+import cast.cdl.WorkingMemoryChangeQueueBehaviour;
+import cast.cdl.WorkingMemoryOperation;
+import cast.cdl.WorkingMemoryPointer;
+import cast.core.CASTUtils;
+import cast.core.data.CASTData;
+import BindingData.BINDING_SUBARCH_CONFIG_KEY;
+import BindingFeatures.AreaID;
+import BindingFeatures.Concept;
+import BindingFeatures.DebugString;
+import BindingFeaturesCommon.TruthValue;
+import binding.abstr.AbstractMonitor;
+import binding.common.BindingComponentException;
+
+/**
+ * The new ComaBindingMonitor.
+ * 
+ * Right now this monitor waits for AreaID feature 
+ * working memory changes.
+ * TODO make sure only the binding WM is taken into account!
+ * 
+ * When a new AreaID is added, this monitor queries the ComaReasoner
+ * for properties of the corresponding Area instance.
+ * These properties include the following list:
+ * the known most specific concepts of the instance
+ * ??? relations ??? 
+ * 
+ * @author Hendrik Zender (zender@dfki.de)
+ * @started 2008-07-07
+ * @version 2008-07-07
+ *
+ */
+public class ComaBindingMonitor_old extends AbstractMonitor {
+
+	String m_bindingSA;
+	
+	private ComaFunctionWriter m_funcWriter;
+	// the following 2 shld be inverse to each other
+	private HashMap<String, String> m_myProxies2myAreas;
+	private HashMap<String, String> m_myAreas2myProxies;
+	private String m_ontologyNamespace;
[…]

### 2009-05-14 code/binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/comsys/autogen/ComsysEssentials/AspectualRelation.java

@@ -1,0 +1,28 @@
+package org.cognitivesystems.comsys.autogen.ComsysEssentials;
+
+
+/**
+* org/cognitivesystems/comsys/autogen/ComsysEssentials/AspectualRelation.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from ComsysEssentials.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class AspectualRelation implements org.omg.CORBA.portable.IDLEntity
+{
+  public String eventSourceId = null;
+  public String stateTargetId = null;
+  public String aspect = null;
+
+  public AspectualRelation ()
+  {
+  } // ctor
+
+  public AspectualRelation (String _eventSourceId, String _stateTargetId, String _aspect)
+  {
+    eventSourceId = _eventSourceId;
+    stateTargetId = _stateTargetId;
+    aspect = _aspect;
+  } // ctor
+
+} // class AspectualRelation

### 2009-05-14 code/binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/comsys/util/uttplan/UtterancePlanningGrammar.java

@@ -1,0 +1,250 @@
+//=================================================================
+// Copyright (C) 2005 Geert-Jan M. Kruijff (gj@acm.org)
+//
+// This library is free software; you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License
+// as published by the Free Software Foundation; either version 2.1 of
+// the License, or (at your option) any later version.
+//
+// This library is distributed in the hope that it will be useful, but
+// WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+// Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU Lesser General Public
+// License along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA
+// 02111-1307, USA.
+//=================================================================
+
+//=================================================================
+// PACKAGE DEFINITION 
+//=================================================================
+
+package org.cognitivesystems.comsys.util.uttplan;
+
+
+//=================================================================
+// IMPORTS
+//=================================================================
+
+import java.io.*;
+import java.util.*; 
+
+// ORG.COGNITIVESYSTEMS packages
+
+import org.cognitivesystems.repr.lf.autogen.LFEssentials.*;
+
+
+//=================================================================
+// JAVADOC CLASS DOCUMENTATION 
+//=================================================================
+
+/** 
+ * 
+ *  @version 050208 (started 050208)
+ *  @author  Geert-Jan M. Kruijff
+ */ 
+
+//=================================================================
+// CLASS DEFINITION 
+//=================================================================
+
+public class UtterancePlanningGrammar {
+    
+      
+    //=================================================================
+    // CLASS-INTERNAL GLOBAL VARIABLES
+    //=================================================================
+
+    private Vector choosersfiles;
+    private Vector systemsfiles;
+
+    private HashMap systems;
+    private HashMap choosers;
+
+    private String grammarname;
+
+    boolean logging = true;
+
+    //=================================================================
[…]

### 2009-05-14 code/binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/repr/lf/autogen/LFPacking/PackedFeaturesHelper.java

@@ -1,0 +1,64 @@
+package org.cognitivesystems.repr.lf.autogen.LFPacking;
+
+
+/**
+* org/cognitivesystems/repr/lf/autogen/LFPacking/PackedFeaturesHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from LFPacking.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+
+// collection of packed featurs
+abstract public class PackedFeaturesHelper
+{
+  private static String  _id = "IDL:org/cognitivesystems/repr/lf/autogen/LFPacking/PackedFeatures:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature[] that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature[] extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      __typeCode = org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeatureHelper.type ();
+      __typeCode = org.omg.CORBA.ORB.init ().create_sequence_tc (0, __typeCode);
+      __typeCode = org.omg.CORBA.ORB.init ().create_alias_tc (org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeaturesHelper.id (), "PackedFeatures", __typeCode);
+    }
+    return __typeCode;
+  }
+
+  public static String id ()
+  {
+    return _id;
+  }
+
+  public static org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature[] read (org.omg.CORBA.portable.InputStream istream)
+  {
+    org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature value[] = null;
+    int _len0 = istream.read_long ();
+    value = new org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature[_len0];
[…]

### 2009-05-14 code/binding/development/subarchitectures/explorer/src/c++/explorer/components/ExplorerBindingMotiveGenerator/ExplorerBindingMotiveGenerator.hpp

@@ -1,0 +1,35 @@
+#include <motivation/components/BindingMotiveGenerator/BindingMotiveGenerator.hpp>
+
+
+/**
+ * An example of how to use the binding motive generator. This is not
+ * necessarily just for he explorer, or all the explorer requires.
+ */
+class ExplorerBindingMotiveGenerator : public BindingMotiveGenerator {
+public:
+  ExplorerBindingMotiveGenerator(const std::string &_id);
+  /**
+   * Start the motive generator.
+   */
+  virtual void start();
+
+
+protected:
+
+  /**
+   * Provide details of the agent for the planning process/
+   */
+  virtual std::string domainName() const {
+    return "explorer";
+  }
+
+  /**
+   * Provide details of the agent for the planning process/
+   */
+  virtual std::string domainFile() const {
+    return "subarchitectures/planning.sa/src/python/mapsim/domains/demo_explorer/mapl_files/domain.mapl";
+  }
+
+
+
+}; 

### 2009-05-14 code/binding/development/subarchitectures/motivation/src/java/motivation/idl/FeatureGenerationCompetenceHelper.java

@@ -1,0 +1,99 @@
+package motivation.idl;
+
+
+/**
+* motivation/idl/FeatureGenerationCompetenceHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from MotivationData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class FeatureGenerationCompetenceHelper
+{
+  private static String  _id = "IDL:motivation/idl/FeatureGenerationCompetence:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, motivation.idl.FeatureGenerationCompetence that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static motivation.idl.FeatureGenerationCompetence extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [4];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().get_primitive_tc (org.omg.CORBA.TCKind.tk_boolean);
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "m_isRelationLabel",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_alias_tc (cast.cdl.OntologyEntryHelper.id (), "OntologyEntry", _tcOf_members0);
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "m_type",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
[…]

### 2009-05-14 code/binding/development/subarchitectures/motivation/src/java/motivation/idl/FeatureGenerationCompetenceHolder.java

@@ -1,0 +1,38 @@
+package motivation.idl;
+
+/**
+* motivation/idl/FeatureGenerationCompetenceHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from MotivationData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class FeatureGenerationCompetenceHolder implements org.omg.CORBA.portable.Streamable
+{
+  public motivation.idl.FeatureGenerationCompetence value = null;
+
+  public FeatureGenerationCompetenceHolder ()
+  {
+  }
+
+  public FeatureGenerationCompetenceHolder (motivation.idl.FeatureGenerationCompetence initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = motivation.idl.FeatureGenerationCompetenceHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    motivation.idl.FeatureGenerationCompetenceHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return motivation.idl.FeatureGenerationCompetenceHelper.type ();
+  }
+
+}

### 2009-05-14 code/binding/development/subarchitectures/nav.sa/src/java/NavData/CompletionHolder.java

@@ -1,0 +1,38 @@
+package NavData;
+
+/**
+* NavData/CompletionHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from NavData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class CompletionHolder implements org.omg.CORBA.portable.Streamable
+{
+  public NavData.Completion value = null;
+
+  public CompletionHolder ()
+  {
+  }
+
+  public CompletionHolder (NavData.Completion initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = NavData.CompletionHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    NavData.CompletionHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return NavData.CompletionHelper.type ();
+  }
+
+}

### 2009-05-14 code/binding/development/subarchitectures/nav.sa/src/java/NavData/NavCtrlCommand.java

@@ -1,0 +1,42 @@
+package NavData;
+
+
+/**
+* NavData/NavCtrlCommand.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from NavData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class NavCtrlCommand implements org.omg.CORBA.portable.IDLEntity
+{
+  public NavData.CtrlCommand m_command = null;
+  public double m_x = (double)0;
+  public double m_y = (double)0;
+  public double m_r = (double)0;
+  public double m_theta = (double)0;
+  public double m_distance = (double)0;
+  public int m_areaID = (int)0;
+  public int m_nodeID = (int)0;
+  public NavData.StatusError m_status = null;
+  public NavData.Completion m_completion = null;
+
+  public NavCtrlCommand ()
+  {
+  } // ctor
+
+  public NavCtrlCommand (NavData.CtrlCommand _m_command, double _m_x, double _m_y, double _m_r, double _m_theta, double _m_distance, int _m_areaID, int _m_nodeID, NavData.StatusError _m_status, NavData.Completion _m_completion)
+  {
+    m_command = _m_command;
+    m_x = _m_x;
+    m_y = _m_y;
+    m_r = _m_r;
+    m_theta = _m_theta;
+    m_distance = _m_distance;
+    m_areaID = _m_areaID;
+    m_nodeID = _m_nodeID;
+    m_status = _m_status;
+    m_completion = _m_completion;
+  } // ctor
+
+} // class NavCtrlCommand

### 2009-05-14 code/binding/development/subarchitectures/nav.sa/src/java/NavData/ObjSequenceHolder.java

@@ -1,0 +1,39 @@
+package NavData;
+
+
+/**
+* NavData/ObjSequenceHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from NavData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class ObjSequenceHolder implements org.omg.CORBA.portable.Streamable
+{
+  public NavData.Obj value[] = null;
+
+  public ObjSequenceHolder ()
+  {
+  }
+
+  public ObjSequenceHolder (NavData.Obj[] initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = NavData.ObjSequenceHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    NavData.ObjSequenceHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return NavData.ObjSequenceHelper.type ();
+  }
+
+}

### 2009-05-14 code/binding/development/subarchitectures/planning.sa/src/java/Planner/ModalityEnumHelper.java

@@ -1,0 +1,53 @@
+package Planner;
+
+
+/**
+* Planner/ModalityEnumHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from Planner.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class ModalityEnumHelper
+{
+  private static String  _id = "IDL:Planner/ModalityEnum:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, Planner.ModalityEnum that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static Planner.ModalityEnum extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      __typeCode = org.omg.CORBA.ORB.init ().create_enum_tc (Planner.ModalityEnumHelper.id (), "ModalityEnum", new String[] { "NO_MODALITY", "K_MODALITY"} );
+    }
+    return __typeCode;
+  }
+
+  public static String id ()
+  {
+    return _id;
+  }
+
+  public static Planner.ModalityEnum read (org.omg.CORBA.portable.InputStream istream)
+  {
+    return Planner.ModalityEnum.from_int (istream.read_long ());
+  }
+
+  public static void write (org.omg.CORBA.portable.OutputStream ostream, Planner.ModalityEnum value)
+  {
+    ostream.write_long (value.value ());
+  }
+
+}

### 2009-05-14 code/binding/development/subarchitectures/planning.sa/src/python/mapsim/planning/plan_monitor/VAL/Validator.cpp

@@ -1,0 +1,3535 @@
+/*-----------------------------------------------------------------------------
+  VAL - The Automatic Plan Validator for PDDL+
+
+  $Date: 2005/06/07 14:00:00 $
+  $Revision: 4 $
+
+  Maria Fox, Richard Howey and Derek Long - PDDL+ and VAL
+  Stephen Cresswell - PDDL Parser
+
+  maria.fox@cis.strath.ac.uk
+  derek.long@cis.strath.ac.uk
+  stephen.cresswell@cis.strath.ac.uk
+  richard.howey@cis.strath.ac.uk
+
+  By releasing this code we imply no warranty as to its reliability
+  and its use is entirely at your own risk.
+
+  Strathclyde Planning Group
+  http://planning.cis.strath.ac.uk
+  ----------------------------------------------------------------------------*/
+#include <iostream>
+#include <string>
+#include <algorithm>
+#include <math.h>
+#include "Validator.h"
+#include "Action.h"
+#include "Utils.h"
+#include "typecheck.h"
+#include "Exceptions.h"
+#include "RobustAnalyse.h"
+#include "random.h"
+#include <sstream>
+
+
+
+
+
+
+using std::stringstream;
+using std::make_pair;
+
+//#define map std::map
+//#define vector std::vector
+
+namespace VAL {
+
+	string getName(plan_step* ps)
+	{
+		string actionName = ps->op_sym->getName();
+
+		for(typed_symbol_list<const_symbol>::const_iterator j = ps->params->begin();
+			j != ps->params->end(); ++j)
+      		{
+      			actionName += (*j)->getName();
+      		};
+
+		return actionName;
+	};
+
+
+	struct compareActionEndPoints {
+
+		compareActionEndPoints() {};
+
+		bool operator()(pair<double,string> a1,pair<double,string> a2) const
+		{
+
+			if(a1.first < a2.first) return true;
+
+			return false;
+		};
+	};
+
+	void changeActionTime(const plan * aPlan, string actionName, double newTime)
+	{
+
+		for(pc_list<plan_step*>::const_iterator i = aPlan->begin(); i != aPlan->end(); ++i)
+			{
+				if(getName(*i) == actionName)
+					{
+						(*i)->start_time = newTime;  (*i)->start_time_given = true;
+					};
+
+			};
+
+	};
+
+
+
[…]

### 2009-05-14 code/binding/development/subarchitectures/spatial.sa/src/java/spatial/workingmemory/SpatialWorkingMemory.java

@@ -1,0 +1,264 @@
+/**
+ * 
+ */
+package spatial.workingmemory;
+
+import java.awt.Color;
+
+import org.cognitivesystems.common.autogen.Math.Pose3D;
+
+import spatial.autogen.SpatialLocation;
+import spatial.autogen.SpatialScene;
+import spatial.util.SpatialDrawingUtils;
+import cast.architecture.subarchitecture.SubarchitectureProcessException;
+import cast.architecture.subarchitecture.SubarchitectureWorkingMemory;
+import cast.cdl.guitypes.Point2D;
+import cast.core.CASTUtils;
+import cast.core.data.CASTWorkingMemoryItem;
+
+/**
+ * @author nah
+ */
+public class SpatialWorkingMemory extends SubarchitectureWorkingMemory {
+
+	private SpatialScene m_currentScene;
+
+	/**
+	 * @param _id
+	 */
+	public SpatialWorkingMemory(String _id) {
+		super(_id);
+
+		// determines whether this wm should broadcast to oher
+		// sub-architectures
+		setSendXarchChangeNotifications(true);
+
+	}
+
+	@Override
+	protected boolean addToWorkingMemory(String _id, CASTWorkingMemoryItem _data) {
+		boolean result = super.addToWorkingMemory(_id, _data);
+		// log("Working memory contains [" + m_workingMemory.size()
+		// + "] items");
+		// log("Forwarding: " + isForwardingXarchChangeNotifications());
+
+		try {
+
+			// if a new scene is added, render it to the gui
+			if (result
+					&& _data.getType().equals(
+							CASTUtils.typeName(SpatialScene.class))) {
+				m_currentScene = getItem(_id, SpatialScene.class);
+				redrawGraphicsNow();
+			}
+		}
+		catch (SubarchitectureProcessException e) {
+			e.printStackTrace();
+			System.exit(1);
+		}
+
+		return result;
+	}
+
+	/**
+	 * @param _scene
+	 * @throws SubarchitectureProcessException
+	 */
+	private void drawScene(SpatialScene _scene)
+			throws SubarchitectureProcessException {
+
+		drawWorkspace();
+		drawViewer(_scene.m_viewerPose);
+
+//		// draw objects
+//		ArrayList<SpatialLocation> objects = getW(_scene.m_locations,
+//				SpatialLocation.class);
[…]

### 2009-05-14 code/binding/development/subarchitectures/spatial.sa/src/java_old/spatial/components/SceneManager.java

@@ -1,0 +1,564 @@
+/**
+ * 
+ */
+package spatial.components;
+
+import java.util.ArrayList;
+import java.util.Hashtable;
+
+import org.cognitivesystems.common.autogen.Math.Vector3D;
+
+import spatial.autogen.SpatialLocation;
+import spatial.autogen.SpatialRelationship;
+import spatial.autogen.SpatialScene;
+import spatial.components.abstr.SpatialSubarchitectureManagedProcess;
+import spatial.ontology.SpatialOntologyFactory;
+import spatial.util.SpatialGoals;
+import spatial.util.SpatialSubarchitectureException;
+import Vision.Camera;
+import Vision.SceneChanged;
+import Vision.SceneObject;
+import balt.corba.autogen.FrameworkBasics.BALTTime;
+import balt.jni.NativeProcessLauncher;
+import balt.management.ProcessLauncher;
+import cast.architecture.abstr.ChangeFilterFactory;
+import cast.architecture.abstr.WorkingMemoryChangeReceiver;
+import cast.architecture.subarchitecture.SubarchitectureProcessException;
+import cast.cdl.FilterRestriction;
+import cast.cdl.WorkingMemoryAddress;
+import cast.cdl.WorkingMemoryChange;
+import cast.cdl.WorkingMemoryOperation;
+import cast.cdl.WorkingMemoryPermissions;
+import cast.cdl.WorkingMemoryPointer;
+import cast.core.CASTUtils;
+import cast.core.data.CASTTypedData;
+import cast.core.data.CachedCASTData;
+
+/**
+ * Component to manage the structs that relate to the current scene. Also adds
+ * viewer pose information into the scene struct.
+ * 
+ * 
+ * When the visual scene is static, this component creates a
+ * {@link SpatialScene} struct to represent the current static scene. Currently
+ * old scenes are not deleted, but they easily could be if things get too nasty.
+ * 
+ * @author nah
+ */
+public class SceneManager extends SpatialSubarchitectureManagedProcess {
+
+	/**
+	 * A class that monitors a visual object for changes and updates a local
+	 * representation of it.
+	 * 
+	 * @author nah
+	 * 
+	 */
+	private class VisualObjectMonitor {
+
+		private boolean m_deleted;
+		private boolean m_needsUpdate;
+		private SpatialLocation m_spatialObject;
[…]

### 2009-05-14 code/binding/development/subarchitectures/vision/src/c++/vision/components/CameraServer/CameraServer.cpp

@@ -1,0 +1,278 @@
+/**
+ * Camera
+ * Provides internal and external (= 3D pose) camera parameters.
+ * Setting 3D pose is a job for either the StereoHead or the Manipulator,
+ * whoever carries a camera.
+ *
+ * @author Michael Zillich
+ * @date October 2006
+ */
+
+#include <fstream>
+#include <vision/VisionGoals.h>
+#include <vision/utils/VisionUtils.h>
+#include "CameraServer.h"
+#include <cast/architecture/ChangeFilterFactory.hpp>
+using namespace Vision;
+
+extern "C" {
+  FrameworkProcess* newComponent(const string &_id) {
+    return new CameraServer(_id);
+  }
+}
+
+
+CameraServer::CameraServer(const string &_id)
+  : WorkingMemoryAttachedComponent(_id), 
+    ManagedProcess(_id)
+{
+  setReceiveXarchChangeNotifications(true);
+
+  // default values
+  cam_num = DEFAULT_CAMERA;
+  head_num = DEFAULT_HEAD;
+  downsample = DEFAULT_DOWNSAMPLE;
+  SetIdentity(eye_pose);
+  SetIdentity(head_pose);
+  SetIdentity(ground_pose);  // HACK
+}
+
+void CameraServer::start()
+{
+  ManagedProcess::start();
+  
+  MemberFunctionChangeReceiver<CameraServer> * headChangeReceiver = 
+      new MemberFunctionChangeReceiver<CameraServer>(this,
+						     &CameraServer::handleUpdateHead);
+  
+
+  addChangeFilter(createGlobalTypeFilter<Head>(cdl::OVERWRITE),
+		  headChangeReceiver);  
+  
+}
+
+CameraServer::~CameraServer()
+{
+}
+
+void CameraServer::taskAdopted(const string &_taskID)
+{
+}
+
+void CameraServer::taskRejected(const string &_taskID)
+{
+}
+
+void CameraServer::handleUpdateHead(const cdl::WorkingMemoryChange &_wmc)
+{
+  shared_ptr<const CASTTypedData<Head> > blorb =
+    getWorkingMemoryEntry<Head>(_wmc.m_address);
+
+  // if our head was updated, get new pose
+  if(blorb->getData()->m_num == head_num)
+  {
+    UpdateHeadPose(blorb->getData()->m_pose);
+    UpdateWorkingMemory();
+  }
+}
+
[…]

### 2009-05-14 code/binding/development/subarchitectures/vision/src/c++/vision/components/ObjectTracker/BTComponents/ObjectTypeManage.cpp

@@ -1,0 +1,47 @@
+#include "ObjectTypeManage.hpp"
+#include "vision/components/common/SystemUtils/Common.h"
+
+using namespace std; 
+using namespace Common;
+
+map<Vision::ObjType, string> ObjectTypeManage::m_mapObjectTypes;
+
+ObjectTypeManage::ObjectTypeManage()
+{
+    
+}
+
+ObjectTypeManage::~ObjectTypeManage()
+{
+}
+
+
+void ObjectTypeManage::Initialize()
+{
+    m_mapObjectTypes[Vision::UNDEF] = Vision::TYPE_UNDEF;
+    m_mapObjectTypes[Vision::HAND] = Vision::TYPE_HAND;
+    m_mapObjectTypes[Vision::SHAPE] = Vision::TYPE_SHAPE;
+    m_mapObjectTypes[Vision::FLAG] = Vision::TYPE_FLAG;
+    m_mapObjectTypes[Vision::CAR] = Vision::TYPE_CAR;
+    m_mapObjectTypes[Vision::BOX] = Vision::TYPE_BOX;
+    m_mapObjectTypes[Vision::BALL] = Vision::TYPE_BALL;    
+}
+
+Vision::ObjType ObjectTypeManage::StringToObjtype(std::string _str)
+{
+    map<Vision::ObjType, string>::iterator it;
+    for (it=m_mapObjectTypes.begin(); it!=m_mapObjectTypes.end(); it++) {
+	if (it->second == _str)
+	    return it->first;
+    }
+    throw user_error(__HERE__, "ObjType %s does not match a string in Vision.idl.", _str.c_str());
+}
+
+string ObjectTypeManage::ObjtypeToString(Vision::ObjType _type)
+{
+    map<Vision::ObjType, string>::iterator it = m_mapObjectTypes.find(_type);
+    if (it != m_mapObjectTypes.end())	
+	return it->second;
+    else
+	throw user_error(__HERE__, "ObjType error");
+}

### 2009-05-14 code/binding/development/subarchitectures/vision/src/c++/vision/components/SIFTComponent/SiftCore/DatabaseEntry.cpp

@@ -1,0 +1,131 @@
+#include <iostream>
+#include <fstream>
+#include "DatabaseEntry.h"
+#include "ioutil.h"
+#include "siftGlobal.h"
+
+using namespace std;
+
+
+DatabaseEntry::DatabaseEntry() {
+
+}
+
+DatabaseEntry::DatabaseEntry(string _strObject) {
+  strObject = _strObject;
+}
+
+
+DatabaseEntry::~DatabaseEntry() {
+
+  map<string, Representation *>::iterator iter;  
+  for (iter=map_viewreps.begin(); iter!=map_viewreps.end(); ++iter) {
+    Representation *prep = iter->second;
+    delete(prep);
+  }
+  map_viewreps.clear();
+}
+
+
+
+int DatabaseEntry::add_viewrep(string viewName, Representation * prep) {
+  
+  map_viewreps[viewName] = prep;
+  return RETURN_OK;
+}
+
+
+// should make sure viewname exist
+void DatabaseEntry::add_features(string viewName, FeatureVector *pfeatures) {
+  
+  map<string, Representation *>::iterator iter =  map_viewreps.find(viewName);
+  
+  if (iter ==  map_viewreps.end()) { // this view does not exist
+    
+    Representation *prep = new Representation();
+    prep->push_back(pfeatures);
+    add_viewrep(viewName, prep);    
+
+    cout << "adding new view " << viewName.c_str() << endl;
+  }
+  else { 
+    //Representation *prep = map_viewreps[viewName];
+    Representation *prep = (*iter).second;
+    prep->push_back(pfeatures);
+
+    cout << "adding features to view " << viewName.c_str() << endl;
+  }
+}
+
+//-----------------------------
+int DatabaseEntry::save(ofstream &file) {
+
+  //file << strObject << " "; // obejctname
+  writeString(file, strObject);
+
+  int nViews = map_viewreps.size();
+  //file << nViews << " "; // number of views
+  file.write((char*)&nViews, sizeof(nViews));
+
+
+  map<string, Representation *>::iterator iter;
+  for (iter=map_viewreps.begin(); iter!=map_viewreps.end(); iter++) {
+
+    //file << (*iter).first << " "; // viewname    
+    writeString(file, (*iter).first);
+    
+    Representation *prep= (*iter).second;
+    prep->save(file);
+  }
+
+  return RETURN_OK;
+}
+
+
+
+//--------------------------------
+int DatabaseEntry::load(ifstream &file) {
+  
[…]

### 2009-05-14 code/binding/development/tools/balt/src/java/balt/corba/connectors/push/primitive/LongRemotePushSenderImpl.java

@@ -1,0 +1,81 @@
+/*
+ * BALT - The Boxes and Lines Toolkit for component communication.
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+/**
+ * 
+ */
+package balt.corba.connectors.push.primitive;
+
+import balt.corba.autogen.RemoteConnectors.RemotePushConnector;
+import balt.corba.autogen.RemoteConnectors.RemotePushSenderPOA;
+import balt.corba.data.RemoteDataTranslator;
+import balt.corba.data.translation.FrameworkDataTranslatorException;
+import balt.core.connectors.push.primitive.interfaces.LongPushInterface;
+
+/**
+ * @author nah
+ */
+public class LongRemotePushSenderImpl extends RemotePushSenderPOA
+        implements LongPushInterface.LongPushConnectorOut {
+
+    private RemotePushConnector m_out;
+
+    /**
+     * 
+     */
+    public LongRemotePushSenderImpl() {
+        super();
+    }
+
+    //these are all out of date because they're not threaded
+    public void flush() {
+        
+    }
+
+    
+    /*
+     * (non-Javadoc)
+     * 
+     * @see balt.core.connectors.push.PushConnectorOut.LongPushConnectorOut#push(java.lang.String,
+     *      int)
+     */
+    public void push(String _src, long _data) {
+        try {
+            m_out
+                .push(_src, RemoteDataTranslator.translateToAny(_data));
+        }
+        catch (FrameworkDataTranslatorException e) {
+            e.printStackTrace();
+            System.exit(1);
+        }
+    }
+
+    /*
+     * (non-Javadoc)
+     * 
[…]

### 2009-05-14 code/binding/development/tools/cast/src/c++/cast/architecture/AlwaysPositiveTaskManager.cpp

@@ -1,0 +1,112 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+#include "AlwaysPositiveTaskManager.hpp"
+
+
+using namespace std;
+using namespace cast::cdl;
+
+/**
+ * The function called to create a new instance of our component.
+ *
+ * Taken from zwork
+ */
+extern "C" {
+  FrameworkProcess* newComponent(const string &_id) {
+    return new cast::AlwaysPositiveTaskManager(_id);
+  }
+}
+
+namespace cast {
+
+  AlwaysPositiveTaskManager::AlwaysPositiveTaskManager(const string &_id)
+    : //InspectableComponent(_id), 
+    //CASTProcessingComponent(_id),
+    WorkingMemoryAttachedComponent(_id),  
+    SubarchitectureTaskManager(_id) {
+  
+    receiveNoChanges();
+  }
+
+  AlwaysPositiveTaskManager::~AlwaysPositiveTaskManager() {
+  }
+
+  void AlwaysPositiveTaskManager::processGoalQueue() {
+
+    while(!m_taskQueue.empty()) {
+
+      InformationProcessingTask task = m_taskQueue.front();
+      m_taskQueue.pop();
+    
+      //send decision
+      sendTaskDecision(task,cdl::GOAL_ADOPTED);    
+
+    
+    }
+  
+  }
+
+  void AlwaysPositiveTaskManager::runComponent() {
+
+    while(m_status == STATUS_RUN) {
+      lockProcess();
+      processGoalQueue();
+      unlockProcess();
+      //sleepProcess(1);
[…]

### 2009-05-14 code/binding/development/tools/cast/src/c++/cast/architecture/ConsistencyException.cpp

@@ -1,0 +1,54 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+#include "ConsistencyException.hpp"
+#include <cstdarg>
+
+using namespace std;
+
+namespace cast {
+
+/**
+ * Except constructor.
+ * \param file     (in) __FILE__ macro
+ * \param function (in) __FUNCTION__ macro
+ * \param line     (in) __LINE__ macro
+ * \param format   (in) printf-style format string
+ */
+ConsistencyException::ConsistencyException(const cast::cdl::WorkingMemoryAddress& _wma, 
+					   const char *file, 
+					   const char *function, 
+					   int line,
+					   const char *format, ...) throw() 
+  : WMException(_wma)
+{
+  static char what[1024];
+  static char msg[1024];
+  va_list arg_list;
+  va_start(arg_list, format);
+  vsnprintf(what, 1024, format, arg_list);
+  va_end(arg_list);
+  snprintf(msg, 1024, "ConsistencyException: %s:%s:%d: %s", file, function, line, what);
+  _what = msg;
+}
+
+}

### 2009-05-14 code/binding/development/tools/cast/src/java/cast/cdl/EXISTS_QUERY_PREFIX.java

@@ -1,0 +1,14 @@
+package cast.cdl;
+
+
+/**
+* cast/cdl/EXISTS_QUERY_PREFIX.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from CAST.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public interface EXISTS_QUERY_PREFIX
+{
+  public static final String value = "ex";
+}

### 2009-05-14 code/binding/development/tools/cast/src/java/cast/cdl/TaskNameListHelper.java

@@ -1,0 +1,63 @@
+package cast.cdl;
+
+
+/**
+* cast/cdl/TaskNameListHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from CAST.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class TaskNameListHelper
+{
+  private static String  _id = "IDL:cast/cdl/TaskNameList:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, String[] that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static String[] extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      __typeCode = org.omg.CORBA.ORB.init ().create_string_tc (0);
+      __typeCode = org.omg.CORBA.ORB.init ().create_alias_tc (cast.cdl.TaskNameHelper.id (), "TaskName", __typeCode);
+      __typeCode = org.omg.CORBA.ORB.init ().create_sequence_tc (0, __typeCode);
+      __typeCode = org.omg.CORBA.ORB.init ().create_alias_tc (cast.cdl.TaskNameListHelper.id (), "TaskNameList", __typeCode);
+    }
+    return __typeCode;
+  }
+
+  public static String id ()
+  {
+    return _id;
+  }
+
+  public static String[] read (org.omg.CORBA.portable.InputStream istream)
+  {
+    String value[] = null;
+    int _len0 = istream.read_long ();
+    value = new String[_len0];
+    for (int _o1 = 0;_o1 < value.length; ++_o1)
+      value[_o1] = cast.cdl.TaskNameHelper.read (istream);
+    return value;
+  }
+
+  public static void write (org.omg.CORBA.portable.OutputStream ostream, String[] value)
+  {
+    ostream.write_long (value.length);
+    for (int _i0 = 0;_i0 < value.length; ++_i0)
+      cast.cdl.TaskNameHelper.write (ostream, value[_i0]);
+  }
+
+}

### 2009-05-14 code/binding/development/tools/cast/src/java/cast/ui/UnifiedGUI.java

@@ -1,0 +1,1522 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+/**
+ * 
+ */
+package cast.ui;
+
+import java.io.*;
+import java.nio.ByteBuffer;
+import java.util.*;
+
+import javax.media.opengl.*;
+import javax.media.opengl.glu.GLU;
+
+import org.eclipse.swt.SWT;
+import org.eclipse.swt.SWTException;
+import org.eclipse.swt.custom.*;
+import org.eclipse.swt.events.SelectionEvent;
+import org.eclipse.swt.events.SelectionListener;
+import org.eclipse.swt.graphics.Rectangle;
+import org.eclipse.swt.layout.*;
+import org.eclipse.swt.opengl.GLCanvas;
+import org.eclipse.swt.opengl.GLData;
+import org.eclipse.swt.widgets.*;
+
+import balt.core.connectors.FrameworkConnectionException;
+import balt.core.data.FrameworkQuery;
+import cast.cdl.guitypes.*;
+import cast.cdl.ui.*;
+import cast.configuration.ArchitectureConfiguration;
+import cast.ui.architecture.*;
+import cast.ui.architecture.interfaces.ComponentStatusPullInterface.ComponentStatusPullConnectorOut;
+import cast.ui.architecture.views.ArchitectureView;
[…]

### 2009-05-14 code/binding/development/tools/cast/src/java/cast/ui/architecture/ComponentMonitorChangeEvent.java

@@ -1,0 +1,72 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+/**
+ * 
+ */
+package cast.ui.architecture;
+
+import java.util.EventObject;
+
+import cast.cdl.ui.ComponentEventType;
+
+/**
+ * @author nah
+ */
+public class ComponentMonitorChangeEvent extends EventObject {
+
+    private ComponentEventType m_changeType;
+
+    /**
+     * 
+     */
+    private static final long serialVersionUID = -4424014569147917361L;
+
+    /**
+     * @param _source
+     */
+    public ComponentMonitorChangeEvent(ComponentMonitor _source,
+            ComponentEventType _changeType) {
+        super(_source);
+        m_changeType = _changeType;
+    }
+
+    /*
+     * (non-Javadoc)
+     * 
+     * @see java.util.EventObject#getSource()
+     */
+    @Override
+    public ComponentMonitor getSource() {
+        // TODO Auto-generated method stub
+        return (ComponentMonitor) super.getSource();
+    }
+
+    
+    /**
+     * @return the changeType
+     */
+    public ComponentEventType getChangeType() {
+        return m_changeType;
+    }
+    
+}

### 2009-05-14 code/binding/development/tools/cast/src/java/cast/ui/architecture/interfaces/ComponentEventPullInterface.java

@@ -1,0 +1,64 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+package cast.ui.architecture.interfaces;
+
+
+import balt.core.connectors.FrameworkConnectionException;
+import balt.core.connectors.pull.*;
+import balt.core.data.FrameworkQuery;
+import cast.cdl.ui.ComponentEvent;
+
+public interface ComponentEventPullInterface {
+
+    public interface ComponentEventPullSender extends PullSender {
+
+        public abstract void setPullConnector(String _connectionID,
+                ComponentEventPullConnectorOut _senderAdaptor);
+    }
+
+    public interface ComponentEventPullReceiver extends PullReceiver {
+
+        public abstract ComponentEvent receivePullQueryComponentEvent(
+                FrameworkQuery _query);
+    }
+
+    public interface ComponentEventPullConnectorRegister extends
+            PullConnectorRegister {
+
+        public abstract void setPullReceiver(ComponentEventPullReceiver _pr);
+    }
+
+    public interface ComponentEventPullConnectorOut extends
+            PullConnectorOut {
+
+        public abstract ComponentEvent pull(FrameworkQuery _query)
+                throws FrameworkConnectionException;
+    }
+
+    public interface ComponentEventPullConnector extends
+            ComponentEventPullConnectorOut,
+            ComponentEventPullConnectorRegister, PullConnector {
+
+    }
+
+}

### 2009-05-14 code/binding/development/tools/cosyidl/src/java/org/cognitivesystems/idl/constExpr/Divide.java

@@ -1,0 +1,118 @@
+/*
+ * @(#)Divide.java	1.12 03/12/19
+ *
+ * Copyright 2004 Sun Microsystems, Inc. All rights reserved.
+ * SUN PROPRIETARY/CONFIDENTIAL. Use is subject to license terms.
+ */
+/*
+ * COMPONENT_NAME: idl.parser
+ *
+ * ORIGINS: 27
+ *
+ * Licensed Materials - Property of IBM
+ * 5639-D57 (C) COPYRIGHT International Business Machines Corp. 1997, 1999
+ * RMI-IIOP v1.0
+ * US Government Users Restricted Rights - Use, duplication or
+ * disclosure restricted by GSA ADP Schedule Contract with IBM Corp.
+ *
+ * @(#)Divide.java	1.12 03/12/19
+ */
+
+package org.cognitivesystems.idl.constExpr;
+
+// NOTES:
+// -D52042<daz> Added protypical code for computing mixed-operand binary
+//  expressions, which promotes result to Double only when the target type
+//  is floating-point. Code violates spec, but may be usable at some future
+//  time.
+
+import java.math.BigInteger;
+
+import org.cognitivesystems.idl.Util;
+
+/**
+ *
+ **/
+public class Divide extends BinaryExpr
+{
+  /**
+   * Constructor: set operation and operands.
+   **/
+  protected Divide (Expression leftOperand, Expression rightOperand)
+  {
+    super ("/", leftOperand, rightOperand);
+  } // ctor
+
+  /**
+   *
+   **/
+  public Object evaluate () throws EvaluationException
+  {
+    try
+    {
+      Number l = (Number)left ().evaluate ();
+      Number r = (Number)right ().evaluate ();
+
+      boolean lIsNonInteger = l instanceof Float || l instanceof Double;
+      boolean rIsNonInteger = r instanceof Float || r instanceof Double;
+
+      if (lIsNonInteger && rIsNonInteger)
+        value (new Double (l.doubleValue () / r.doubleValue ()));
+      else if (lIsNonInteger || rIsNonInteger)
+      {
+        String[] parameters = {Util.getMessage ("EvaluationException.divide"),
+            left ().value ().getClass ().getName (),
+            right ().value ().getClass ().getName ()};
+        throw new EvaluationException (Util.getMessage ("EvaluationException.1", parameters));
+      }
+      else
+      {
[…]

### 2009-05-14 code/binding/development/tools/cosyidl/src/java/org/cognitivesystems/idl/toJavaPortable/MethodGen.java

@@ -1,0 +1,1065 @@
+/*
+ * @(#)MethodGen.java	1.17 02/01/15
+ *
+ * Copyright 2004 Sun Microsystems, Inc. All rights reserved.
+ * SUN PROPRIETARY/CONFIDENTIAL. Use is subject to license terms.
+ */
+/*
+ * COMPONENT_NAME: idl.toJava
+ *
+ * ORIGINS: 27
+ *
+ * Licensed Materials - Property of IBM
+ * 5639-D57 (C) COPYRIGHT International Business Machines Corp. 1997, 1999
+ * RMI-IIOP v1.0
+ * US Government Users Restricted Rights - Use, duplication or
+ * disclosure restricted by GSA ADP Schedule Contract with IBM Corp.
+ *
+ * @(#)MethodGen.java	1.9 99/08/18
+ */
+
+package org.cognitivesystems.idl.toJavaPortable;
+
+// NOTES:
+// -After demarshalling an IOR, think about how to deal with the exceptions.
+// -catching Exception throws a string which should be in a properties file.
+// -30jul1997<daz> Modified to write comment immediately preceding method signature.
+// -07May1998<ktp> Modified to support RMI Portable Stub
+// -26Aug1998<klr> Modified to pass helper instance to read_Value.
+// -F46082.51<daz> Remove -stateful feature; javaStatefulName() obsolete.
+// -D56554   <klr> Port bounded string checks from toJava to toJavaPortable
+// -D58549   <klr> bounded string checks on in/inout parms throw BAD_PARAM
+// -D57112<daz> Valuetype initializers map to ctor, regardless of name, and
+//  "void _init(...)" methods now mapped correctly.
+// -D59297   <klr> pass context parm when Remarshalling
+// -D59560   <klr> call read/write_Context
+// -D60929   <klr> Update for RTF2.4 changes
+// -D61056   <klr> Use Util.helperName
+// -D61650<daz> Remove '\n' from generated strings; use println()'s. 
+
+import java.io.PrintWriter;
+import java.util.Enumeration;
+import java.util.Hashtable;
+
+import org.cognitivesystems.idl.*;
+import org.cognitivesystems.idl.constExpr.Expression;
+
+
+
+/**
+ *
+ **/
+public class MethodGen implements org.cognitivesystems.idl.MethodGen
+{
+  private static final String ONE_INDENT   = "    ";
+  private static final String TWO_INDENT   = "        ";
+  private static final String THREE_INDENT = "            ";
[…]

### 2009-05-14 code/binding/development/tools/cosyidl/src/java/org/cognitivesystems/idl/toJavaPortable/ModuleGen.java

@@ -1,0 +1,64 @@
+/*
+ * @(#)ModuleGen.java	1.13 03/12/19
+ *
+ * Copyright 2004 Sun Microsystems, Inc. All rights reserved.
+ * SUN PROPRIETARY/CONFIDENTIAL. Use is subject to license terms.
+ */
+/*
+ * COMPONENT_NAME: idl.toJava
+ *
+ * ORIGINS: 27
+ *
+ * Licensed Materials - Property of IBM
+ * 5639-D57 (C) COPYRIGHT International Business Machines Corp. 1997, 1999
+ * RMI-IIOP v1.0
+ * US Government Users Restricted Rights - Use, duplication or
+ * disclosure restricted by GSA ADP Schedule Contract with IBM Corp.
+ *
+ * @(#)ModuleGen.java	1.13 03/12/19
+ */
+
+package org.cognitivesystems.idl.toJavaPortable;
+
+// NOTES:
+
+import java.io.File;
+import java.io.PrintWriter;
+import java.util.Enumeration;
+import java.util.Hashtable;
+
+import org.cognitivesystems.idl.ModuleEntry;
+import org.cognitivesystems.idl.SymtabEntry;
+
+
+/**
+ *
+ **/
+public class ModuleGen implements org.cognitivesystems.idl.ModuleGen
+{
+  /**
+   * Public zero-argument constructor.
+   **/
+  public ModuleGen ()
+  {
+  } // ctor
+
+  /**
+   * Generate Java code for all members of an IDL module.
+   **/
+  public void generate (Hashtable symbolTable, ModuleEntry entry, PrintWriter stream)
+  {
+    // Generate the package directory
+    String name = Util.containerFullName( entry ) ;
+    Util.mkdir (name);
+
+    // Generate all of the contained types
+    Enumeration e = entry.contained ().elements ();
+    while (e.hasMoreElements ())
+    {
+      SymtabEntry element = (SymtabEntry)e.nextElement ();
+      if (element.emit ())
+        element.generate (symbolTable, stream);
+    }
+  } // generate
+} // class ModuleGen

### 2009-05-14 code/binding/development/tools/reasoning/src/java/org/cognitivesystems/reasoner/base/OntologyInstance.java

@@ -1,0 +1,39 @@
+package org.cognitivesystems.reasoner.base;
+
+public class OntologyInstance implements ReasonerInstance {
+
+	private String m_namespace;
+	private String m_sep;
+	private String m_name;
+	
+	protected OntologyInstance(String _namespace, String _sep, String _name) {
+		m_name = _name;
+		m_namespace = _namespace;
+		m_sep = _sep;
+	}
+	
+	public String getFullName() {
+		return m_namespace + m_sep + m_name;
+	}
+
+	public String getName() {
+		return m_name;
+	}
+
+	public String getNamespace() {
+		return m_namespace;
+	}
+
+	public int compareTo(Object _ins) {
+		return this.getFullName().compareTo(((ReasonerInstance) _ins).getFullName());
+	}
+	
+	public String toString() {
+		return getName();
+	}
+
+	public String getSep() {
+		return m_sep;
+	}
+
+}

### 2009-05-14 code/binding/development/tools/reasoning/src/java/org/cognitivesystems/reasoner/base/OntologyRelation.java

@@ -1,0 +1,59 @@
+package org.cognitivesystems.reasoner.base;
+
+public class OntologyRelation implements ReasonerRelation {
+
+	private String m_namespace;
+	private String m_sep;
+	private String m_name;
+
+	private ReasonerInstance m_arg1;
+	private ReasonerInstance m_arg2;
+	
+	protected OntologyRelation(String _namespace, String _sep, String _name, ReasonerInstance _arg1, ReasonerInstance _arg2) {
+		m_namespace = _namespace;
+		m_sep = _sep;
+		m_name = _name;
+		m_arg1 = _arg1;
+		m_arg2 = _arg2;
+	}
+	 
+	public ReasonerInstance getArg1() {
+		return m_arg1;
+	}
+
+	public ReasonerInstance getArg2() {
+		return m_arg2;
+	}
+
+	public String getFullRelationName() {
+		return m_namespace + m_sep + m_name;
+	}
+
+	public String getRelationName() {
+		return m_name;
+	}
+
+	/* (non-Javadoc)
+	 * @see java.lang.Object#toString()
+	 */
+	public String toString() {
+		
+		return getRelationName()+"("+(m_arg1!=null ? m_arg1.getFullName() : "")+","+(m_arg2!=null ? m_arg2.getFullName() : "")+")";	}
+
+	/* (non-Javadoc)
+	 * @see java.lang.Comparable#compareTo(java.lang.Object)
+	 */
+	public int compareTo(Object _rel) {
+		return this.toString().compareTo(((ReasonerRelation) _rel).toString());
+	}
+
+	public String getRelationNamespace() {
+		return m_namespace;
+	}
+
+	public String getRelationSep() {
+		return m_sep;
+	}
+
+
+}

### 2009-05-14 code/binding/development/tools/reasoning/src/java/org/cognitivesystems/reasoner/pellet/PelletWrapper.java

@@ -1,0 +1,1235 @@
+package org.cognitivesystems.reasoner.pellet;
+
+// Java API
+import java.io.BufferedReader;
+import java.io.FileReader;
+import java.io.IOException;
+import java.io.PrintStream;
+import java.io.StringWriter;
+import java.net.URI;
+import java.util.HashSet;
+import java.util.Iterator;
+import java.util.LinkedList;
+import java.util.Map;
+import java.util.Queue;
+import java.util.List;
+import java.util.Set;
+import java.util.TreeSet;
+
+// Reasoner API
+import org.cognitivesystems.reasoner.*;
+import org.cognitivesystems.reasoner.base.ReasonerConcept;
+import org.cognitivesystems.reasoner.base.ReasonerInstance;
+import org.cognitivesystems.reasoner.base.OntologyMemberFactory;
+import org.cognitivesystems.reasoner.base.ReasonerException;
+import org.cognitivesystems.reasoner.base.ReasonerInterface;
+import org.cognitivesystems.reasoner.base.ReasonerRelation;
+
+// External API: pellet, owlapi, aterm-java-1.6
+import org.mindswap.pellet.EdgeList;
+import org.mindswap.pellet.Individual;
+import org.mindswap.pellet.Literal;
+import org.mindswap.pellet.Node;
+import org.mindswap.pellet.Role;
+import org.mindswap.pellet.exceptions.InternalReasonerException;
+import org.mindswap.pellet.output.OutputFormatter;
+import org.mindswap.pellet.owlapi.ConceptConverter;
+import org.mindswap.pellet.owlapi.Reasoner;
+import org.semanticweb.owl.apibinding.OWLManager;
+import org.semanticweb.owl.io.StringInputSource;
+import org.semanticweb.owl.model.OWLClass;
+import org.semanticweb.owl.model.OWLConstant;
+import org.semanticweb.owl.model.OWLDataProperty;
+import org.semanticweb.owl.model.OWLDescription;
+import org.semanticweb.owl.model.OWLIndividual;
+import org.semanticweb.owl.model.OWLObjectProperty;
+import org.semanticweb.owl.model.OWLObjectPropertyExpression;
+import org.semanticweb.owl.model.OWLOntology;
+import org.semanticweb.owl.model.OWLOntologyCreationException;
+import org.semanticweb.owl.model.OWLOntologyManager;
+import org.mindswap.pellet.utils.ATermUtils;
+
[…]

### 2009-05-14 binding/development/specifications/bham_meeting/scenarios/explorer/explorer.tex
very very first commit

\documentclass{article}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdwlist}
\usepackage{graphicx}

\title{Explorer scenario specification for Bham meeting}
\date{\today}
\author{CoSy Explorer group, Hendrik}

\begin{document}
\maketitle

\section{Script -- level 1}
\subsection{Scenario 1}
\label{scen1}
\textit{Scenario 1, cf. Specs}

\begin{enumerate}
\item Robot is idling in the corridor. 
\item User approaches the robot.
\item \label{lvl1:user_asks_1}
	User asks robot: ``Could you go and find the Borland book, please?''
\item \label{lvl1:rob_answs_1}
  	Robot says: ``I will go and look for it in a library where books are normally found.''
\item \label{lvl1:user_answs_2}
	User answers: ``Ok.''
\item \label{lvl1:rob_goes_lib}
	Robot moves to the library.
\item \label{lvl1:look_for_book}
	Robot looks for the book.

\emph{deviation from original scenario script!}
\item \label{lvl1:not_find_book}
	Robot does not find it.
\item \label{lvl1:look_for_person}
	Robot looks for a person (to ask).
\item \label{lvl1:}
	Robot asks: ``Do you know where the Borland book is?''
\item 
	Person answers: ``I believe it is in GJ's office.''
\item
	Robot says: ``Okay.'' and takes off to GJ's office.

\item 
	Robot enters GJ's office and finds the Borland book.
\item
	Robot goes back to the user.
\item
	Robot finds user and says ``The book is in GJ's office near the Starbucks mug.''

\end{enumerate}

%\begin{figure}[ht]
%\centering
%\includegraphics[width=\linewidth]{example1_msc.png}
%\caption{
%  Some simple msc-graph can be generated to clarify interaction
%  between things in the the scenario. Use the
%  \texttt{../gen\_msc.sh}-script to generate pngs from all
%  msc-files. mscgen is documented here:
%  \texttt{http://www.mcternan.me.uk/mscgen/}. It's VERY simple}
%\label{fig:lvl1}
%\end{figure}

\section{Script -- level 2}
\textit{
  The level 1 script describes what can obviously be seen when the
  robot is interacting, for example in a video uptake of the
  scenario. The next level is to describe, without any technical
  detail, what activities take place within the robot during this
  interaction. }

\begin{enumerate}
\item \label{lvl2:1}
	The robot is idle (with some initial info).
	\begin{enumerate}
	\item Robot is in Hendrik's office and so is Hendrik.
	\item It has a map of the environment with H's office, 
			GJ's office, a corridor, and the library.
	\item Robot has ontological knowledge that books are 
			in libraries and can recognize the Borland book.
	\end{enumerate}
\item \label{lvl2:2} 
	User comes up to the robot.
	\begin{enumerate}
	\item Robot detects a person approaching
	\item Robot looks at the ``user''
	\item 	
	\end{enumerate}

\item \label{lvl2:3} 
	User asks robot...
	\begin{enumerate}
	\item Utterance gets interpreted and translated into a command
	\item Motive generated on basis of command
	\item ``BorlandBook'' is on the binder\footnote{could be represented
		as mental note tag; maybe associated with a probability}
	\item Robot does inference and planning
		\begin{itemize}
		\item plan: where is location of BorlandBook
		\item result: does not know location of this instance
		\item infer: type of BorlandBook is book; books are usually found in libraries
		\item plan: path to library

		\end{itemize}
	\end{enumerate}

\item \label{lvl2:4} 
	Robot gives feedback (answer) based on generated plan
	\begin{enumerate}
	\item Robot plans a response
	\item Robot utters the response
	\end{enumerate}

\item \label{lvl2:5} 
	Users confirms (``Ok.'')
	\begin{enumerate}
	\item Robot interprets utterance as a confirmation and 
		allows the plan to be executed.
	\end{enumerate}

\item \label{lvl2:6} 
	Robot moves to library
	\begin{enumerate}
	\item Robot moves through a door into corridor, follows it and enters library through a door
	\item (Robot verifies that it is the library)
	\end{enumerate}

\item \label{vis_search_lib} \label{lvl2:7} Robot looks for book
	\begin{enumerate}
	\item Robot creates a plan to search the room for finding the 
		book\footnote{this might correspond to planning of visual operators done by Mohan;
			KTH should clarify potential overlap or contradictions with Mohan}
	\item Executes the visual search plan
	\end{enumerate}

\item \label{vis_search_lib_fail} \label{lvl2:8} Robot does not find it
	\begin{enumerate}
	\item Plan ends; no book found\footnote{need to represent fact that we have not
		succeeded finding the book instance}
	\item (Adds explicit negation ``book is not in library'' to planning 
		state)\footnote{probabilities of typical knowledge could be linked to threshold
		function -- problem: probabilities will be very low if there are many rooms.}
	\end{enumerate}

\item \label{lvl2:9}
	Robot looks for person to ask
	\begin{enumerate}
	\item Visual search (plan) failure triggers re-planning
	\item New plan involves finding and asking a person
	\item looks for a person and finds one
	\item drives up to the person
	\end{enumerate}

\item \label{lvl2:10} 
	Robot asks person ``Do you know where the Borland book is?''
	\begin{enumerate}
	\item Plans a dialogue and utters question
	\end{enumerate}

\item Persons answers: ``I believe it is in GJ's office.''
	\begin{enumerate}
	\item Robot interprets that as an answer to its question
	\item resolves ``it'' to BorlandBook
	\item Robot extracts the position info
	\item (BorlandBook + pos info represented on binder (updated) with modality)
	\end{enumerate}

\item Robot says ``Okay.'' and takes off to GJ's office
	\begin{enumerate}
	\item planner plans moveTo(GJ's office) action
	\item starts moving to GJ's office
	\end{enumerate}

\item Robot enters GJ's office and finds BB.
	\begin{enumerate}
	\item see \ref{vis_search_lib} and \ref{vis_search_lib_fail}, but now finding it
	\item adds book position
	\end{enumerate}

\item Robot goes back to user
	\begin{enumerate}
	\item continues execution of original plan, which involves
		\begin{itemize}
		\item assumes user in same place
		\item moves back there
		\end{itemize}
	\end{enumerate}

\item Robot finds user and says ``book is in GJ's office near the mug...''
	\begin{enumerate}
	\item looks for a person and assumes it is the user
	\item robot verbalizes opsition of the book
	\end{enumerate}

\end{enumerate}

\section{List of subarchitectures}

\subsection{motivator}
look for people to interact with 
(passively! get ready for interaction, 
pay attention to people; replace Y2 ``wake up function'')

listen for event structures on the binder: ``find'' can be transormed
into a planning goal

\subsection{vision.sa}

\subsection{navigation.sa}
SLAM, graph building, people detection (laser-based, ?vision?)

KTH; overall responsible: Patric

\subsection{planner}
get goal from MotiveGenerator and Handler

\subsection{binder}

\subsection{comsys}

\subsection{coma.sa}

\subsection{spatial}
same as in PlayMate!
near, left of, right of

\section{Subarchitecture Level}
\textit{
  Describe for each step in the detailed lvl2 script what the involved
  subarchitectures are and how they are involved.}

\begin{figure}[bth]
\centering
	\includegraphics[angle=-90,width=\linewidth]{ex-img_0051.jpg}
	\caption{Final drawing of SA interaction: Orange arrows denote stepwise
	processing. The first number is the index of the level 2 description, the second
	number is an incremental global index over all processes.} 
\label{fig:subarch_interaction}
\end{figure}

\subsection{\ref{lvl2:1}} smth in the binder called ``me'' (ie the robot),
	which has a position (by ?nav.sa?); cf. Fig. \ref{fig:nav_orig}

\subsection{\ref{lvl2:2}} User detection triggered by laser and 
	verified by vision
	person represented on the binder (put there by the person detector/tracker), cf. Fig. \ref{fig:nav_orig}
	open question: how to trigger look at user

\begin{figure}[bth]
\centering
	\includegraphics[width=\linewidth]{explorer_1.jpg}
	\caption{Original state of the binder: nav.sa represents the robot and the user on the binder.
	Not shown is the representation of the robot's and the user's current (topological) positions.} 
\label{fig:nav_orig}
\end{figure}

\subsection{\ref{lvl2:3}} User asks robot...
	utterance parsed by comsys; produces a logical form that expresses that there
	is a request for finding the object
	proxy for ``you'' = addressee/hearer resolved via binder to robot proxy;
	proxy for ``BorlandBook'' by comsys;
	proxy for event structure ``find the book'' by comsys
	behind the scenes: transform that into knowledge goal by MotiveGenerator
	generate goal for planner: user\_knows(pos(book))
	This is shown in Fig. \ref{fig:comsys_plan}.

	coma.sa notices mention of "borland book"; since it does not know
	concrete information, it decides to provide default knowledge:
	coma.sa creates a proxy ``Book" (or ``Books'' if plural handling works!)
	and a proxy ``Library'' and links them with a ``position'' relation proxy.
	This relation proxy is tagged as ``default''.
	Coma.sa furthermore provides proxies for all libraries that it knows about,
	in this case a single proxy for ``area 42''.
	Nav.sa notices that and provides a proxy for its internal ``area 42''
	Now comes the binder: 
	``BorlandBook'' and ``Book'' get bound to the same union (u1)
	``Library'' (abstract) and ``Library / area 42'' (concrete) from coma.sa get bound (u2);
	``area 42'' from nav.sa ends up in the same union (u2)!
	Those unions (u1 and u2) will also be linked by the ``position'' (default) relation.
	This is shown in Fig. \ref{fig:coma_default}.

	Planner works out a plan (cf. Fig. \ref{fig:planner})
	(moveto area42)
	(sense BB)
	(moveto area26)
	(tell user pos(book))

\begin{figure}[bth]
\centering
	\includegraphics[width=\linewidth]{explorer_2.jpg}
	\caption{add caption!} 
\label{fig:comsys_plan}
\end{figure}

\begin{figure}[bth]
\centering
	\includegraphics[width=\linewidth]{explorer_3.jpg}
	\caption{add caption!} 
\label{fig:coma_default}
\end{figure}

\begin{figure}[bth]
\centering
	\includegraphics[width=\linewidth]{explorer_4.jpg}
	\caption{add caption!} 
\label{fig:planner}
\end{figure}

\subsection{\ref{lvl1:rob_answs_1}} Planner can come up with a satisfiable plan. Planner can provide
	MotiveHandler tells comsys about the outcome of planning to give verbal feedback
	-- MotiveMonitor between motive.sa and comsys.sa
	\begin{enumerate}
	\item satisfiability (``yes'')
	\item goal state (``will go to Y'')
	\item based on instance or default knowledge (``X is in Y'' vs ``X'es are typically found in Y's'')
	\end{enumerate}

\subsection{\ref{lvl2:5}} Comsys interprets confirmation

\subsection{\ref{lvl2:6}} MotiveHandler/monitor notices acknowledgement by comsys
	MotiveHandler triggers execution of the plan by the planner
	Planner passes (moveto area42) action to nav.sa, which executes it (mark in struct ``ACK")
	-- and succeeds (mark in struct ``DONE")

\subsection{\ref{lvl2:7}} Planner sends (sense BB) action to nav.sa, which runs 
	visual search planning + execution (mark in struct ``ACK")

\subsection{\ref{lvl2:8}} Visual search fails; 
	nav.sa adds NEGATED ``position'' relation between BB and area42;
	BB will be removed from the ``Book(s)'' (abstract) union;
	``area42/lib'' (concrete, coma.sa) and area42 (concrete, nav.sa) will be removed from the ``Library'' (abstract) union;
	\footnote{coma.sa could take the fact that BB is not in the Book(s) union anymore as an
		indication to remove the default proxies -- because those were provided because of
		/ triggered by BB}

\subsection{\ref{lvl2:9}} Robot needs to look for person to ask
	should people tracking run constantly throughout the experiment? Or only on-demand?

	If the motive is ``look for person and ask'', people tracker can
	provide knowledge about persons to the binder.

	The planner could also have an action ``find person'' that will
	trigger a specific behavior by the people recognition.

	This could also be handled by ``information request'' / ``clarification
	request'': create a proxy ``Person'' w/o any additional information and
	request position information.

	go to person = goto(person)

\subsection{\ref{lvl2:10}}
	Comsys plans a dialogue to fulfill askPersonAboutBook command.

\subsection{\ref{lvl2:11}}
	Person answers;
	comsys analyzes the answer and provides the conveyed information 
	to the binder;

	modal qualification:
	personBelieves(pos(book,officeGJ))

	might result in assertion in the planner -- ask Michael?
	is that parallel to the standard assumption that persons are where they were before?

\subsection{\ref{lvl2:16}} planner gives tell(pos(BB,xy)) to comsys; comsys figures out via 
	coma.sa how to generate an appropriate referential description!!!

\section{Discussion, Day 2}
GJ leads discussion about Episodes.

Scenario 2++ -- i.e. route descriptions instead of / together with autonomous exploration.

\subsection{Episodes / episodic structure}
\begin{enumerate} 
	\item Spatio-temporal structure: ``temporal flow''
	\item Spatio-temporal structure:
	\begin{itemize}
		\item ``inclusion''
		\item ``zooming in \& out'' -- local spatial scenes w/ multi-level maps
	\end{itemize}
	\begin{enumerate}
		\item Route description:\\
				(this constitutes info about the future that has to be instantiated later;
				creates expectations -- ``priming'' of perceptual processes)
		\item ``semantic tags'' / active vision in exploration and HAM:\\
				robot drops ``tag'' when it cannot act upon asserted info in that point in time
				(when exploring later: info from the past)

	\end{enumerate}

\end{enumerate}

Question is how the sequence of ``episodes'' is cunstructed.
Motives and plans. Two levels: (Nick)
1) overall task (go to the XXX)
2) chunk plan into small steps (goals) -- action segmentation / kind of episodic memory (Michael B)

plan results from sts on EM
EM is the counter-part of planner / planning WM

Question is how the change of the binding WM can serve as an explicit or implicit 
discretization of states into episodes...

Nick very simple 1st pass implementation; write a series of planning exec steps in the motivegen
by hand; second step: motive handler can read a different WM in which some goals are
structured in a specific way

expectations are goal states

Nick: this could be fleshed out to a planning / motivation level.
GJ thinks this can be largely considered independent of hard-core robotics.

2 names: GJ and Dennis S

\section{Responsibility Level}

\textit{Who's doing what!}

General rules:
\begin{itemize}
	\item maintain own proxies on the binder as long as it is
	part of the current scene as observed by that SA; remove proxy
	unless it is in a union with some other proxy, which means that
	it is part of the ``scene''/``situation'' for another SA.
	\item Binder needs comparators; whoever is responsible for putting
		stuff in the binder is also responsible for adding features and
		provide the respective comparators
	\item nav.sa tells the planner what it can do (``planning actions'');
		set some status flag in WM struct (this has been set-up by Nick) -- Chandana
	\item planner needs to offer the possibility to report back failure to find a plan
		(or rather: failure to achieve goal state) to the user!
\end{itemize}

\subsection{Step 1}
Robot in binder + its location in binder (linked by CURRENT\_POS relation) -- nav.sa.
this changes when area changes;
KTH, Kristoffer

\subsection{Step 2}
\subsubsection{Robot is idle; maintenance goal by Motivation;}
Motive: if you see a person, you look at it; as long as you don't do anything else,
you look at it. (Status MotiveGenerator: abstract classes, need to fill in the gaps)
Nick work on the general bits that are used across scenarios.
motive detection and generation (Patrick)

\subsubsection{User detection triggered by laser}
3 things: 
\begin{itemize} 
	\item ALU-FR: laser-based person detection (Daniel)
	\item KTH: people tracker
	\item TUD: visual person detection (Niko)
	\item TUD: face detection (Niko)
\end{itemize}
laser-based detector \& tracker can be combined; responsible: Daniel;
co-ordination with TUD / Nikodem;

\subsubsection{look at the user}
simple way: send position/angle to PT(Z)U
if the camera is moving: can it still detect???
move camera: Niko + Daniel

\subsubsection{represent person on binder}
person + topological position on the binder (Daniel)
maintain in the same way as with the robot's position

\subsection{Step 3}
User asks robot

\subsubsection{Comsys analyzes utterance}
produce logical form; represent addressee on the binder (should bind to robot);
represent goal; represent ``BorlandBook'' being talked about on binder
-- GJ

\subsubsection{Motive generation}
Motive Generator notices this structure (from comsys) on the binder (Patrick)

\subsubsection{Coma listens to binder and provides default knowledge}
\begin{itemize}
	\item to do!
\end{itemize}
-- Hendrik

\subsubsection{nav.sa listens and provides its LTM knowledge on the binder}
puts a proxy for its internal ``BorlandBook'' object model
puts a proxy for its internal ``area42'' spatial unit
-- Kristoffer

\subsubsection{Binder does its magic}
\begin{itemize}
	\item Plural handling
	\item tag relations as DEFAULT
\end{itemize}
-- Henrik J

\subsubsection{Plan generation}
MotiveHandler / Planner requests state from binder (Michael B)
Plan generation inside the planner (Michael B)

prelim. steps: dummy plan? extend planning domain (Michael B)

\subsection{Step 4}
Planner has satisfiable plan;

\subsubsection{MotiveMonitor between motive.sa and comsys.sa}
There is a MotiveMonitor between motive.sa and comsys.sa; via this monitor
the outcome of planning is provided so that verbal feedback can be given
-- ?

\subsubsection{Utterance generation}
comsys produces feedback about plan -- GJ

\subsection{Step 5}
User confirms ``Ok.''; comsys analyzes -- GJ

\subsection{Step 6}
MotiveHandler recognizes confirmation of plan (otherwise it could just simply time-out)
-- ? or Michael B / Patrick

Planner starts execution (Michael)

sends ``gotoArea'' command to nav.sa; nav.sa receives and executes commands -- Chandana

\subsection{Step 7}
nav.sa reports DONE for ``gotoArea'' command (Chandana)

planner sends ``find(BB)'' command to nav.sa (Michael)

do view planning \& execute view planning -- Kristoffer (co-operation KTH and TUD)
view planning fails, update struct ``Done''

put negated position relation between BB and area42 into the binder (Kristoffer)
-- input from Henrik and Hendrik

in the planner, replanning is triggered (Michael)

complex scenario:
new plan involves asking a person; plan step lookForPerson: turn on people detection, tracking,
recognition; when recognized, move up to person (people following?)

simplified scenario:
assume there is a person, goto person (Chandana), ask person

planner sends command to comsys.sa (Michael, GJ)
comsys produces question -- GJ

\subsection{Step 8}
Visual

\subsection{Step 10}
Dialogue planning -- Ivana \& GJ

\section{Details about subarchitectures}

\subsection{\ref{lvl2:sth}}
\subsubsection{Subarchitecture X}
\begin{itemize}
\item do this
\item do that
\end{itemize}

\subsubsection{Subarchitecture Y}
\begin{itemize}
\item do this
\item do that
\end{itemize}

\subsection{\ref{lvl2:sth:detail}}
...
\subsection{\ref{lvl2:sth:detail2}}
...
\subsection{\ref{lvl2:sth2}}
...
\subsection{\ref{lvl2:sth2:detail}}
...
\subsection{\ref{lvl2:sth2:detail2}}
...

\section{Dependency Level}

\textit{
  What else is required to generate the above information and
  behaviour. Remember the already existing specs prepared prior to the
  meeting: \texttt{http://www.dfki.de/\~\ henrikj/doc\_test/specifications/main/html/}}

\section{Component Level}

\textit{
  First pass as describing the behaviours in terms of the components
  that are involved.}

\end{document}

\end{document}

### 2009-05-14 binding/development/specifications/bham_meeting/scenarios/playmate/uol_scripts.tex
very very first commit

\documentclass{article}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdwlist}
\usepackage{graphicx}

\title{UoL scenario specification}
\date{\today}
\author{CoSy}

\begin{document}
\maketitle
\begin{abstract}
Specification of the UoL scenario...
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Script -- level 1}
Some of the science demostrated through this scenario:
\begin{enumerate}
\item Tutor driven learning.
\item Tutor supervised and exploratory learning.
\item Unlearning.
\item Colearning.
\item Determination of salience.
\item Implicit learning.
\item Clarification.
\end{enumerate}

\noindent
Tutor driven learning.
\begin{enumerate}
\item \label{lvl1:step1} The human puts an object in front of the
  robot and retracts his arm.

\item \label{lvl1:step2} 
  H: ``This is a red triangular object.''\\
  R: ``OK.''

\item \label{lvl1:step3} H puts down a second object in front of the
  robot and retracts his arm.

\item \label{lvl1:step4} 
  H: ``This is a blue and triangular object.''\\
  R: ``OK.''

\item \label{lvl1:step5} The H reaches and retracts all the objects in
  the scene.

\item \label{lvl1:step6} The H puts three green objects and one yellow
  object down.

\item \label{lv1:step7}
  H: ``These are green objects.''\\
  R: ``OK.''

\item \label{lvl1:step8}
  H takes away the three given objects. \\
  OR \\
  H points to the yellow object.

\item \label{lvl1:step9}
  H: ``This object is not green.''\\
  R: ``OK.''

\item \label{lvl1:step10}
  H: ``Is this object green?''\\
  R: ``No.''

\item \label{lvl1:step11}
  H removes the yellow object.

\item \label{lvl1:step12}
  H puts down a green triangle.

\item \label{lvl1:step13}
  H: ``What is this object?''

\item \label{lvl1:step14}
  R: ``It is a green triangle.'' \\
  OR\\
  R: ``Is it a green triangle?''

\end{enumerate}

\noindent
Next, some examples on ambiguous reference to an object while
learning, and verbalizing learned knowledge.
\begin{enumerate}
\item \label{lvl1a:step1} H takes away all previous objects in the
  scene.

\item \label{lvl1a:step2} H puts a red triangle and a blue square
  together.

\item \label{lvl1a:step3} 
  H: ``This is a blue square.''\\

\item \label{lvl1a:step4} 
  R: ``Which thing do you mean?''

\item \label{lvl1a:step5} 
  H: ``The object on the right.''

\item \label{lvl1a:step6}
  R updates its representation.

\item \label{lvl1a:step7} 
  R: ``OK. So the object on the right is a blue square.''

\item \label{lvl1a:step8} 
  H picks up the red triangle and the blue square.

\item \label{lvl1a:step9}
  H puts down a blue square.

\item \label{lvl1a:step10}
  H: ``What color is this object?''

\item \label{lvl1a:step11}
  R: ``Is it a blue thing?''

\item \label{lvl1a:step12}
  H: ``Yes.''

\item \label{lvl1a:step13}
  H takes away the blue square.

\item \label{lvl1a:step14}
  H puts down the green triangle.

\item \label{lvl1a:step15}
  H: ``What do you see?''

\item \label{lvl1a:step16}
  R: ``I see a green thing. Is it a triangle?''

\item \label{lvl1a:step17}
  H: ``Yes.''

\item \label{lvl1a:step18}
  H picks up the green triangle.

\item \label{lvl1a:step19}
  H puts down a green circle.

\item \label{lvl1a:step20}
  R: ``I think this object is green, but what shape is it?''

\end{enumerate}

\noindent
Next, a look at the changes in the subarchitecture that accompany the
list of ambiguous reference steps above (just a few steps).
\begin{enumerate}
\item \label{lvl2a:step1} H removes all previous objects in the scene (trivial case:)

\item \label{lvl2a:step2} H puts a red triangle and a blue square
  together.

\begin{itemize}
\item \label{lvl2a:step2:detail1} VideoServer:VSA $\to$
  ChangeDetector:VSA $\to$ SceneChanged (stopped changing):VSA. \\
  Responsibility: UoL.

\item \label{lvl2a:step2:detail2} SceneChanged:VSA $\to$
  Segmentor:VSA $\to$ ROIs:VSA, SceneObjects:VSA. \\
  Responsibility: UoL.

\item \label{lvl2a:step2:detail3} SceneObjects:VSA $\to$
  BindingMonitor:VSA $\to$ Proxy:VSA. \\
  Responsibility: UoL/Bham (Binding Monitor needs to be resolved).

\item \label{lvl2a:step2:detail4} ROI:VSA $\to$
  FeatureExtractor:VSA $\to$ ROI:VSA. \\
  Responsibility: UoL.

\item \label{lvl2a:step2:detail5} Proxy:VSA $\to$ BindingComponent:BSA
  $\to$ Union:BSA. \\
  Responsibility: DFKI (Binding).

\item \label{lvl2a:step2:detail6} ROI:VSA (feature vector) $\to$
  Learner-Recognizer:VSA $\to$ ROI (attribute\_vector\_possible:VSA,
  attribute\_vector\_probable:VSA) $\to$ SceneObject (color,
  shape):VSA.\\
  Responsibility: UoL (Learner-Recognizer).

\item \label{lvl2a:step2:detail7} SceneObject:VSA $\to$
  BindingMonitor:VSA $\to$ Proxy:BSA (features added).

\item \label{lvl2a:step2:detail8} Proxy:BSA $\to$
  BindingComponent:BSA $\to$ Union:BSA.

\end{itemize}

\end{enumerate}

\noindent
Issues to be handled across the playmate demonstrator:
\begin{enumerate}
\item Proxy updating in Binding SA can fail if interrupted.
\item Learner-recognizer in VSA needs new synchronization/notification
  model.
\item Locking-invalidate-on-write inconsistency model.
\item Attribute vector needs to contain the posterior distributions.
\end{enumerate}

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{example1_msc.png}
% \caption{
%   Some simple msc-graph can be generated to clarify interaction
%   between things in the the scenario. Use the
%   \texttt{../gen\_msc.sh}-script to generate pngs from all
%   msc-files. mscgen is documented here:
%   \texttt{http://www.mcternan.me.uk/mscgen/}. It's VERY simple}
% \label{fig:lvl1}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Script -- level 2}
Here is where the scripts are discussed in 'slightly' more detail...

\begin{enumerate}
\item \label{lvl2:step1} The human puts an object in front of the
  robot and retracts his arm.
  \begin{enumerate}
  \item \label{lvl2:step1:detail0} The robot tracks the hand and the
    object: recognizes ``put'' and ``retract'' actions.
  \item \label{lvl2:step1:detail1} The robot recognizes that the scene
    has changes.
  \item \label{lvl2:step1:detail2} Robot creates a (set of)
    placeholder(s) for the object(s).
  \end{enumerate}

\item \label{lvl2:step2} 
  H: ``This is a red triangular object.''\\
  R: ``OK.''
  \begin{enumerate}
  \item \label{lvl2:step2:detail} The robot recognizes the utterance +
    creates a representation for the context of the utterance.
  \item \label{lvl2:step2:detail2} The robot recognizes that it is
    being given a ``learning'' instruction.
  \item \label{lvl2:step2:detail3} It connects the thing being talked
    about with the thing being seen.
  \item \label{lvl2:step2:detail4} It initiates/triggers a learning
    action which creates the ``concepts'' of ``red'' and ``triangle''.
  \item \label{lvl2:step2:detail5} It generates an acknowledegment.
  \end{enumerate}

\item \label{lvl2:step3} H puts down a second object in front of the
  robot and retracts his arm.
  \begin{enumerate}
  \item \label{lvl2:step3:detail1} The robot tracks the object and the
    human hand + recognizes the ``put'' and ``retract'' actions.
  \item \label{lvl2:step3:detail2} The robot recognizes that the scene
    has been changed.
  \item \label{lvl2:step3:detail3} It attempts to recognize the shape,
    color of the new palceholder object.
  \item \label{lvl2:step3:detail4} 
  \end{enumerate}

\item \label{lvl2:step4} 
  H: ``This is a blue and triangular object.''\\
  R: ``OK.''
  \begin{enumerate}
  \item \label{lvl2:step4:detail1} As in Step~\ref{lvl2:step2} but it
    uses saliency to select the visual object being referred to (to
    learn about).
  \item \label{lvl2:step4:detail2} Also creates a new concept for
    ``blue''.
  \end{enumerate}

\item \label{lvl2:step5} The H reaches and retracts all the objects in
  the scene.
  \begin{enumerate}
  \item \label{lvl2:step5:detail1} The robot tracks hand and objects:
    ``reach for'', ``pull'' action recognized.
  \item \label{lvl2:step5:detail2} The robot analyzes the new
    ``blank'' scene.
  \end{enumerate}

\item \label{lvl2:step6} The H puts three green objects and one yellow
  object down.
  Analysis is similar to Step~\ref{lvl2:step1}.

\item \label{lv2:step7}
  H: ``These are green objects.''\\
  R: ``OK.''
  \begin{enumerate}
  \item \label{lvl2:step7:detail1} The robot does speech recognition
    and creates a representation of the context of the utterance.
  \item \label{lvl2:step7:detail2} The robot recognizes it has a
    learning assertion.
  \item \label{lvl2:step7:detail3} Connects the linguistic reference
    to the whole group.
  \item \label{lvl2:step7:detail4} Creates a concept for ``green'' and
    updates it three more times (one update erroneous -- yellow
    object).
  \item \label{lvl2:step7:detail5} Generates an acknowledgement.
  \end{enumerate}

\item \label{lvl2:step8} 
  H takes away the three given objects. \\
  OR \\
  H points to the yellow object.
  \begin{enumerate}
  \item \label{lvl2:step8a:detail1} Recognize the ``reach'' and
    ``pull'' actions.
  \item \label{lvl2:step8a:detail2} Re-analyzes the scenes -- matches
    the yellow object to the green concept.
  \end{enumerate}

  OR:
  \begin{enumerate}
  \item \label{lvl2:step8b:detail1} Robot recognizes the ``point''
    action.b
  \item \label{lvl2:step8b:detail2} The yellow placeholder becomes
    more salient.
  \end{enumerate}

\item \label{lvl2:step9}
  H: ``This object is not green.''\\
  R: ``OK.''
  See Step~\ref{lvl2:step7:detail1} and Step~\ref{lvl2:step7:detail2}.
  \begin{enumerate}
  \item \label{lvl2:step9:detail1} Connects the reference to the
    salient (yellow) object.
  \item \label{lvl2:step9:detail2} Updates the concept of green by
    removing the yellow component.
  \item \label{lvl2:step9:detail3} Generates an acknowledgement.
  \end{enumerate}

\item \label{lvl2:step10}
  H: ``Is this object green?''\\
  R: ``No.''
  \begin{enumerate}
  \item \label{lvl2:step10:detail1} Parse the utterance as a question,
    and its context ``green object'' is represented.
  \item \label{lvl2:step10:detail2} The reference is bound to the
    salient visual object.
  \item \label{lvl2:step10:detail3} A goal is raised to answer the
    question.
  \item \label{lvl2:step10:detail4} The match to the concept of
    ``green'' is interrogated for the salient visual regions.
  \item \label{lvl2:step10:detail5} Robot returns the answer about a
    lack of match.
  \end{enumerate}

\item \label{lvl2:step11}
  H removes the yellow object.
  Analysis similar to Step~\ref{lvl2:step5}.

\item \label{lvl2:step12}
  H puts down a green triangle.
  Analysis similar to Step~\ref{lvl2:step1} or Step~\ref{lvl2:step3}.

\item \label{lvl2:step13}
  H: ``What is this object?''
  \begin{enumerate}
  \item \label{lvl2:step13:detail1} Comprehend the utterance and bind
    the reference to the visual object.
  \item \label{lvl2:step13:detail2} Raises a goal to answer the
    question.
  \end{enumerate}

\item \label{lvl2:step14}
  R: ``It is a green triangle.'' \\
  OR\\
  R: ``Is it a green triangle?''
  \begin{enumerate}
  \item \label{lvl2:step14:detail1} Generates an answer based on the
    recognized properties.
  \item \label{lvl2:step14:detail2} Generates the utterance.
  \end{enumerate}
\end{enumerate}

% \begin{figure}[!htcb]
% \centering
% \includegraphics[width=\linewidth]{example2_msc.png}
% \caption{
% A more detailed msc-graph could be used on this or next level...
% }
% \label{fig:lvl2}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Subarchitecture Level}

%%----------------------------------------------------------
\subsection{Step~\ref{lvl1a:step1}-\ref{lvl1a:step2}}
H takes away all previous objects in the scene.  H puts a red triangle
and a blue square together.

\subsubsection{Vision SA}
\begin{itemize}
\item Change detection.
\item ROIs, SOs (two of them).
\item No ``knowledge''.
\end{itemize}

\subsubsection{Binding SA}
\begin{itemize}
\item Proxies for ROIs and SOs.
\item Unions and relationships between the proxies for the objects.
\item Proxy for the discourse-referent blue square thing.
\end{itemize}

\subsubsection{Spatial SA}
\begin{itemize}
\item Proxies for ROIs and Sos.
\item Spatial relationships between the two objects.
\end{itemize}

\subsubsection{Comsys SA}
\begin{itemize}
\item A discourse-referent for the blue square thing.
\end{itemize}

Nick will fill in the subsequent steps...

%%----------------------------------------------------------
\subsection{Step~\ref{lvl1:step3}}

%%----------------------------------------------------------
\subsection{Step~\ref{lvl1:step4}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Dependency Level}

\textit{
  What else is required to generate the above information and
  behaviour. Remember the already existing specs prepared prior to the
  meeting: \texttt{http://www.dfki.de/\~\ henrikj/doc\_test/specifications/main/html/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Component Level}

\textit{
  First pass as describing the behaviours in terms of the components
  that are involved.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Responsibility Level}

Responsibilitites for the wide range of vision components (yes, we
have people assigned to tasks finally in the fourth year:)
\begin{enumerate}
\item VideoServer: Mohan.
\item Change Detector: Alen.
\item Segmentor: Alen.
\item Binding Monitor: Alen/Nick.
\item Feature Extractor: Alen.
\item Learner-Recognizer: Alen.
\item Discirminant Classifier: Nikodem.
\item Hand-tracker: Somboon.
\item Object Tracker: Somboon.
\item Action Recognizer: Somboon.
\item Pointing Analyzer: Peter.
\item Designing of IDLs: Jeremy, and only Jeremy!
\item Design of Synchronization: Nick.
\item Coordination over matlab/visual WM interface: Jeremy.
\end{enumerate}

\noindent
The entries in the various data structures in the visual WM:
\begin{itemize}
\item Region-Of-Interest (ROI):
  \begin{itemize}
  \item Time/frame number.
  \item 2D bounding box.
  \item Image raster.
  \item Feature vector.
  \item Color histogram (mean + standard deviation/range along each
    channel).
  \item Attribute-matrix-posterior (posterior distribution over the
    color/shape classes).
  \item Salience.
  \item Contour points (envelope of the objects in the ROI).
  \item Type (Hand/Thing).
  \end{itemize}

\item Scene Object (SOb):
  \begin{itemize}
  \item Color.
  \item Shape.
  \item 3D pose.
  \item Type (hand/thing).
  \end{itemize}

\item Scene Changed:
  \begin{itemize}
    \item Scene changing.
    \item Scene changed.
  \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2009-05-14 binding/development/subarchitectures/binding.sa/config/test/generate_fake_comsys_cast_files.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+for ((a=$1; a < $2 ; a++))
+do
+  ./generate_fake_comsys_cast_file.sh $a > ./test-fake-comsys_$a.cast;
+done  
+./generate_fake_comsys_xml_file.sh $1 $2 > test-fake-comsys.xml;

### 2009-05-14 binding/development/subarchitectures/binding.sa/src/java/BindingData/MonitorToken.java

@@ -1,0 +1,26 @@
+package BindingData;
+
+
+/**
+* BindingData/MonitorToken.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from BindingData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class MonitorToken implements org.omg.CORBA.portable.IDLEntity
+{
+
+  /// not used for anything
+  public boolean m_dummy = false;
+
+  public MonitorToken ()
+  {
+  } // ctor
+
+  public MonitorToken (boolean _m_dummy)
+  {
+    m_dummy = _m_dummy;
+  } // ctor
+
+} // class MonitorToken

### 2009-05-14 binding/development/subarchitectures/binding.sa/src/java/BindingQueries/FeatureAssertionHelper.java

@@ -1,0 +1,83 @@
+package BindingQueries;
+
+
+/**
+* BindingQueries/FeatureAssertionHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from BindingQueries.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class FeatureAssertionHelper
+{
+  private static String  _id = "IDL:BindingQueries/FeatureAssertion:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, BindingQueries.FeatureAssertion that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static BindingQueries.FeatureAssertion extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [2];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = BindingQueries.BasicQueryHelper.type ();
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "m_assertion",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_alias_tc (cast.cdl.SubarchitectureIDHelper.id (), "SubarchitectureID", _tcOf_members0);
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "m_fromSA",
+            _tcOf_members0,
+            null);
+          __typeCode = org.omg.CORBA.ORB.init ().create_struct_tc (BindingQueries.FeatureAssertionHelper.id (), "FeatureAssertion", _members0);
[…]

### 2009-05-14 binding/development/subarchitectures/coma.sa/src/java/ComaData/ComaInstanceListHolder.java

@@ -1,0 +1,41 @@
+package ComaData;
+
+
+/**
+* ComaData/ComaInstanceListHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from ComaData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+
+// typedef sequence < ComaInstance > ComaInstanceList;
+public final class ComaInstanceListHolder implements org.omg.CORBA.portable.Streamable
+{
+  public cast.cdl.WorkingMemoryPointer value[] = null;
+
+  public ComaInstanceListHolder ()
+  {
+  }
+
+  public ComaInstanceListHolder (cast.cdl.WorkingMemoryPointer[] initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = ComaData.ComaInstanceListHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    ComaData.ComaInstanceListHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return ComaData.ComaInstanceListHelper.type ();
+  }
+
+}

### 2009-05-14 binding/development/subarchitectures/coma.sa/src/java/coma/components/ComaNavgraphMonitor.java

@@ -1,0 +1,544 @@
+package coma.components;
+
+import java.util.HashMap;
+import java.util.Properties;
+
+import NavData.AEdge;
+import NavData.Area;
+import NavData.FNode;
+import NavData.ONode;
+import NavData.RobotPose;
+import NavData.TopologicalRobotPos;
+import NavData.VEdge;
+
+import org.cognitivesystems.reasoner.base.OntologyMemberFactory;
+import org.cognitivesystems.reasoner.base.ReasonerInstance;
+import org.cognitivesystems.reasoner.base.ReasonerRelation;
+
+import coma.aux.ComaFunctionWriter;
+
+import cast.architecture.abstr.ChangeFilterFactory;
+import cast.architecture.abstr.WorkingMemoryChangeReceiver;
+import cast.architecture.subarchitecture.PrivilegedManagedProcess;
+import cast.architecture.subarchitecture.SubarchitectureProcessException;
+import cast.cdl.WorkingMemoryAddress;
+import cast.cdl.WorkingMemoryChange;
+import cast.cdl.WorkingMemoryChangeQueueBehaviour;
+import cast.cdl.WorkingMemoryOperation;
+import cast.core.data.CASTData;
+
+public class ComaNavgraphMonitor extends PrivilegedManagedProcess {
+
+	private ComaFunctionWriter m_funcWriter;
+	private HashMap<WorkingMemoryAddress, Integer> m_wma2nodeID;
+	private HashMap<WorkingMemoryAddress, Integer> m_wma2areaID;
+	private HashMap<WorkingMemoryAddress, ReasonerRelation> m_wma2rel;
+	private String m_ontologyNamespace;
+	private String m_ontologySep;
+	private OntologyMemberFactory m_oeMemberMaker;
+
+	public ComaNavgraphMonitor(String _id) {
+		super(_id);
+        m_queueBehaviour = WorkingMemoryChangeQueueBehaviour.QUEUE;
+        m_wma2nodeID = new HashMap<WorkingMemoryAddress, Integer>();
+        m_wma2areaID = new HashMap<WorkingMemoryAddress, Integer>();
+        m_wma2rel = new HashMap<WorkingMemoryAddress, ReasonerRelation>();
+	}
+
+    public void configure(Properties _config) {
+    	super.configure(_config);
+        
+        if (_config.containsKey("--onto_ns")) {
+            this.m_ontologyNamespace = _config.getProperty("--onto_ns");
+        }
+        else {
[…]

### 2009-05-14 binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/comsys/autogen/ComsysEssentials/EventHolder.java

@@ -1,0 +1,38 @@
+package org.cognitivesystems.comsys.autogen.ComsysEssentials;
+
+/**
+* org/cognitivesystems/comsys/autogen/ComsysEssentials/EventHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from ComsysEssentials.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class EventHolder implements org.omg.CORBA.portable.Streamable
+{
+  public org.cognitivesystems.comsys.autogen.ComsysEssentials.Event value = null;
+
+  public EventHolder ()
+  {
+  }
+
+  public EventHolder (org.cognitivesystems.comsys.autogen.ComsysEssentials.Event initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = org.cognitivesystems.comsys.autogen.ComsysEssentials.EventHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    org.cognitivesystems.comsys.autogen.ComsysEssentials.EventHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return org.cognitivesystems.comsys.autogen.ComsysEssentials.EventHelper.type ();
+  }
+
+}

### 2009-05-14 binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/comsys/autogen/ComsysEssentials/StateDiscRefRelationHolder.java

@@ -1,0 +1,38 @@
+package org.cognitivesystems.comsys.autogen.ComsysEssentials;
+
+/**
+* org/cognitivesystems/comsys/autogen/ComsysEssentials/StateDiscRefRelationHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from ComsysEssentials.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class StateDiscRefRelationHolder implements org.omg.CORBA.portable.Streamable
+{
+  public org.cognitivesystems.comsys.autogen.ComsysEssentials.StateDiscRefRelation value = null;
+
+  public StateDiscRefRelationHolder ()
+  {
+  }
+
+  public StateDiscRefRelationHolder (org.cognitivesystems.comsys.autogen.ComsysEssentials.StateDiscRefRelation initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = org.cognitivesystems.comsys.autogen.ComsysEssentials.StateDiscRefRelationHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    org.cognitivesystems.comsys.autogen.ComsysEssentials.StateDiscRefRelationHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return org.cognitivesystems.comsys.autogen.ComsysEssentials.StateDiscRefRelationHelper.type ();
+  }
+
+}

### 2009-05-14 binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/comsys/monitors/proxyfactories/ActionTakeProxyFactory.java

@@ -1,0 +1,352 @@
+//=================================================================
+// Copyright (C) 2008 Geert-Jan M. Kruijff (gj@dfki.de)
+//
+// This library is free software; you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License
+// as published by the Free Software Foundation; either version 2.1 of
+// the License, or (at your option) any later version.
+//
+// This library is distributed in the hope that it will be useful, but
+// WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+// Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU Lesser General Public
+// License along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA
+// 02111-1307, USA.
+//=================================================================
+
+//=================================================================
+// PACKAGE DEFINITION 
+//=================================================================
+
+package org.cognitivesystems.comsys.monitors.proxyfactories;
+
+//=================================================================
+// IMPORTS
+//=================================================================
+
+// ----------------------------------------------------------------
+// BINDING imports
+// ----------------------------------------------------------------
+import binding.common.BindingComponentException;
+
+// ----------------------------------------------------------------
+// CAST imports
+// ----------------------------------------------------------------
+import cast.architecture.subarchitecture.SubarchitectureProcessException;
+
+// ----------------------------------------------------------------
+// COMSYS imports
+// ----------------------------------------------------------------
+import org.cognitivesystems.comsys.monitors.AbstractProxyFactory;
+import org.cognitivesystems.comsys.monitors.PendingProxyRelation;
+import org.cognitivesystems.comsys.monitors.ProxyFactoryResults;
+import org.cognitivesystems.comsys.monitors.LocalProxyStructure;
+
+// ----------------------------------------------------------------
+// JAVA imports
+// ----------------------------------------------------------------
[…]

### 2009-05-14 binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/comsys/util/uttplan/UPGSystemsReader.java

@@ -1,0 +1,768 @@
+//=================================================================
+// Copyright (C) 2005 Geert-Jan M. Kruijff (gj@acm.org)
+//
+// This library is free software; you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License
+// as published by the Free Software Foundation; either version 2.1 of
+// the License, or (at your option) any later version.
+//
+// This library is distributed in the hope that it will be useful, but
+// WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+// Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU Lesser General Public
+// License along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA
+// 02111-1307, USA.
+//=================================================================
+
+//=================================================================
+// PACKAGE DEFINITION 
+//=================================================================
+
+package org.cognitivesystems.comsys.util.uttplan;
+
+//=================================================================
+// IMPORTS
+//=================================================================
+
+import java.io.*;
+import java.util.*;
+
+// SAX parser
+
+import org.xml.sax.*;
+import org.xml.sax.helpers.*;
+import javax.xml.parsers.*;
+
+
+// ORG.COGNITIVESYSTEMS packages
+
+import org.cognitivesystems.repr.lf.autogen.LFEssentials.*;
+
+
+
+//=================================================================
+// JAVADOC CLASS DOCUMENTATION 
+//=================================================================
+
+/** 
+ *  The class <b>UPGSystemsReader</b> constructs the systems for the
+ *  systemic network, as specified in the systems XML file for the
+ *  grammar.
+ * 
+ *  @version 050719
+ *  @author  Geert-Jan M. Kruijff
+ */ 
+
+//=================================================================
+// CLASS DEFINITION 
+//=================================================================
+
+public class UPGSystemsReader 
+    extends DefaultHandler
+{
+    
+    //=================================================================
+    // CLASS-INTERNAL GLOBAL VARIABLES
+    //=================================================================
+
+    String  filename; 
+    HashMap systems;
+
+    UPGSystem currentSystem; 
[…]

### 2009-05-14 binding/development/subarchitectures/comsys.mk4/src/java/org/cognitivesystems/repr/lf/autogen/LFPacking/PackedFeatureHelper.java

@@ -1,0 +1,92 @@
+package org.cognitivesystems.repr.lf.autogen.LFPacking;
+
+
+/**
+* org/cognitivesystems/repr/lf/autogen/LFPacking/PackedFeatureHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from LFPacking.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class PackedFeatureHelper
+{
+  private static String  _id = "IDL:org/cognitivesystems/repr/lf/autogen/LFPacking/PackedFeature:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static org.cognitivesystems.repr.lf.autogen.LFPacking.PackedFeature extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [3];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "feat",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "value",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
[…]

### 2009-05-14 binding/development/subarchitectures/nav.sa/src/java/NavData/FNodeHolder.java

@@ -1,0 +1,38 @@
+package NavData;
+
+/**
+* NavData/FNodeHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from NavData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class FNodeHolder implements org.omg.CORBA.portable.Streamable
+{
+  public NavData.FNode value = null;
+
+  public FNodeHolder ()
+  {
+  }
+
+  public FNodeHolder (NavData.FNode initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = NavData.FNodeHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    NavData.FNodeHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return NavData.FNodeHelper.type ();
+  }
+
+}

### 2009-05-14 binding/development/subarchitectures/nav.sa/src/java/NavData/NavCommandHelper.java

@@ -1,0 +1,138 @@
+package NavData;
+
+
+/**
+* NavData/NavCommandHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from NavData.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class NavCommandHelper
+{
+  private static String  _id = "IDL:NavData/NavCommand:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, NavData.NavCommand that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static NavData.NavCommand extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [10];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = NavData.CommandHelper.type ();
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "m_command",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = NavData.PriorityHelper.type ();
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "m_priority",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _members0[2] = new org.omg.CORBA.StructMember (
+            "m_target",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().get_primitive_tc (org.omg.CORBA.TCKind.tk_double);
+          _members0[3] = new org.omg.CORBA.StructMember (
+            "m_xpos",
+            _tcOf_members0,
+            null);
[…]

### 2009-05-14 binding/development/subarchitectures/planning.sa/src/java/Planner/Failure.java

@@ -1,0 +1,33 @@
+package Planner;
+
+
+/**
+* Planner/Failure.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from Planner.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class Failure extends org.omg.CORBA.UserException
+{
+  public String message = null;
+
+  public Failure ()
+  {
+    super(FailureHelper.id());
+  } // ctor
+
+  public Failure (String _message)
+  {
+    super(FailureHelper.id());
+    message = _message;
+  } // ctor
+
+
+  public Failure (String $reason, String _message)
+  {
+    super(FailureHelper.id() + "  " + $reason);
+    message = _message;
+  } // ctor
+
+} // class Failure

### 2009-05-14 binding/development/subarchitectures/planning.sa/src/python/mapsim/plans.py

@@ -1,0 +1,808 @@
+#! /usr/bin/env python
+# -*- coding: latin-1 -*-
+
+from collections import defaultdict
+import itertools
+import string
+import sys
+
+
+from config import log
+from constants import *
+import config
+import graph
+import operator
+import os.path
+import state
+import utils
+from utils import cached_attribute
+
+def pddl_str2action_dict (astring):
+    arguments = astring.split()
+    assert len(arguments) >= 2
+    return dict(operator=arguments[0], agent=arguments[1], arguments=arguments[2:])
+
+class Action(object):
+    def __init__(self, agent=None, operator=None, arguments=None, rw_description=None, planning_agent=None):
+        assert not isinstance(operator, Action)
+        self.operator = operator
+        self.agent = agent
+        self.arguments = arguments
+        self.rw_description = rw_description
+        self.planning_agent = planning_agent
+
+    def copy(self):
+        return Action(self.agent, self.operator, self.arguments, self.rw_description, self.planning_agent)
+
+    @cached_attribute
+    def mapl_op(self):
+        try:
+            domain = self.planning_agent.mapl_domain
+
+            if self.is_sensing_action():
+                return domain.get_sensor_by_name(self.operator[len(PDDL_SENSOR_PREFIX):])
+            return domain.get_action_by_name(self.operator)
+        except AttributeError:
+            log("MAPL op access hack doesn't seem to work. Sorry.  Contact Michael B to fix it!")
+            return None
+
+    @cached_attribute
+    def description(self):
+        return " ".join([self.operator, self.agent] + self.arguments)
+
+    @cached_attribute
+    def action_type(self):
+            if self.is_hidden():
+                return VIRTUAL_ACTION
+            if self.is_sensing_action():
+                return SENSING_ACTION
+            if self.operator == NEGOTIATE_PLAN_OP:
+                return NEGOTIATION_ACTION
+            if self.is_speech_act():
+                return OTHER_SPEECH_ACT
+            else:
+                return PHYS_ACTION
+
+    def grounded_mapl_action(self, vars_in_parens=True):
+        arity = len(self.mapl_op.parameters)
+        grounded_mapl = "%s: %s %s" % (self.agent, self.operator, " ".join(self.arguments[:arity]))
+        if vars_in_parens and arity < len(self.arguments):
+            grounded_mapl += " (%s)" % " ".join(self.arguments[arity:])
+        return grounded_mapl
+
+    def is_assertion(self):
+        if not self.mapl_op or self.is_sensing_action():
+            return False
[…]

### 2009-05-14 binding/development/subarchitectures/vision/src/c++/vision/components/common/GeomUtils/Polygon2D.h

@@ -1,0 +1,62 @@
+/** @file Polygon2D.h
+ *  @brief Definition of a 2D polygon.
+ *
+ *
+ *  @author Somboon Hongeng
+ *  @bug No known bugs.
+ */
+#ifndef _POLYGON2D_H_
+#define _POLYGON2D_H_
+
+#include <iostream>
+#include <vector>
+#include "Vector2D.h"
+#include "Polygon.h"
+
+namespace Geom {
+    
+class Polygon2D {
+ private:
+    POLYGONARC* poly; 
+    Vector2D center;  
+    
+    void compute_center(void);
+	
+ public:
+    // Creation.
+    Polygon2D(void);
+    Polygon2D(const Polygon2D& p2D);
+    ~Polygon2D(void);
+  
+    Polygon2D& operator=(const Polygon2D& rhs);
+
+    // Operations.
+    const POLYGONARC& g_poly(void) const { return(*poly); };
+    const Vector2D& g_center(void) const { return(center); };
+    
+    /** @brief Initialize polygonarc by points pol[].
+     */ 
+    void init(int *pol, int nb_pts);
+
+    
+    void init(std::vector<Vector2D> points);
+
+    int is_pt_in_zone(Vector2D* pr) const;
+    int is_pt_close_to(Vector2D* pr) const;
+    double distance(Vector2D* pr) const; /* diff_x + diff_y */
+    void chgt_ref(void);
+
+    int is_pt_within_range(Vector2D* pr, float thres_dist=200) const;
+    
+    // Affichage.
+    void draw(void);
+    void print(std::ostream *os) const { *os << poly->nb_vertices();};
+};
+ 
+}
+
+
+//std::ostream& operator<<(std::ostream& out, const Geom::Polygon2D& p);
+
+
+#endif /* _POLYGON2D_H_ */

### 2009-05-14 binding/development/subarchitectures/vision/src/c++/vision/components/common/SystemUtils/Common.cpp

@@ -1,0 +1,109 @@
+/** @file Common.cpp
+ *  @brief Common utility and error handling functions.
+ *
+ *
+ *  @author Somboon Hongeng
+ *  @bug No known bugs.
+ */
+#include <cstdarg>
+#include <iostream>
+#include <libgen.h>
+#include "Common.h"
+
+using namespace std;
+
+void Common::user_assert(bool expression, 
+			 const char *file, const char *function, int line,
+			 const char* message) // may throw exception
+{
+  if (!expression) 
+  {
+      std::cout << "Assertion failed!";
+      throw user_error(file, function, line, message);
+//       abort();
+  }
+}
+
+void Common::user_printf(const char *file, const char *function, int line,
+			 const char *format, ...) 
+{
+    char what[1024];
+    char msg[1024];
+    va_list arg_list;
+    va_start(arg_list, format);
+    vsnprintf(what, 1024, format, arg_list);
+    va_end(arg_list);
+    
+    string pathname = file;
+    int pos = pathname.find_last_of("/");
+    string classname = pathname.substr(pos+1);
+
+    snprintf(msg, 1024, "[%s:%s:%d]: %s", classname.c_str(), function, line, what);
+    std::cout << msg;
+}
+
+
+Common::user_error::user_error(const char *file, 
+			       const char *function, int line,
+			       const char *format, ...) throw() // don't throw anything
+{
+    static char what[1024];
+    static char msg[1024];
+    va_list arg_list;
+    va_start(arg_list, format);
+    vsnprintf(what, 1024, format, arg_list);
+    va_end(arg_list);
+    snprintf(msg, 1024, "%s:%s:%d: %s", file, function, line, what);
+    _what = msg;
+    printf(msg);  
+}
+
+void Common::user_error::addInfo(const std::string& info) {
+    printf(info.c_str());
+}
+
+
+/**
+ *  Construct a complete filename by prefix the filename with
+ *  the current directory from "cpath_filename".
+ */
+string Common::completeFilename(string cpath_filename, string filename) 
+{
+    char cpath_file[1024];
+    snprintf(cpath_file, 1024, "%s", cpath_filename.c_str());
+    string path = dirname(cpath_file);
+    string complete_filename = path + "/" + filename.c_str();
+    return complete_filename;
+}
+
+void Common::ParseConfigFile(string cfgFilename, map<string,string> &config)
+{
[…]

### 2009-05-14 binding/development/subarchitectures/vision/src/matlab/showLmodels.m

@@ -1,0 +1,32 @@
+function showLmodels(mC,Fnames,Cnames,sphw,hp)
+%showLmodels(mC,Fnames,Cnames,hp)
+%Show learned models
+
+if nargin<4
+   sphw=[2 ceil(length(mC)/2)];
+end;   
+if nargin<5
+   hp=gcf;
+end;
+
+
+
+numC=length(mC);
+
+for i=1:numC
+   if ~isempty(mC(1).name)
+      ha=subplot(sphw(1),sphw(2),i,'Parent',hp);
+      cla(ha);
+      showDecPdf(mC(i),ha);
+      %set(ha,'FontSize',16);
+      title(ha,['{\bf' Cnames(mC(i).name,:) '}\leftrightarrow' Fnames(mC(i).Fb,:) ' (' num2str(mC(i).conf) ')']);
+      set(ha,'ytick',[]);
+      
+      alim=axis(ha);
+      xticks=[alim(1),(alim(1)+alim(2))/2,alim(2)];
+      set(ha,'XTick',xticks);
+      set(ha,'XTickLabel',{num2str(xticks(1),'%3.3f'),num2str(xticks(2),'%3.3f'),num2str(xticks(3),'%3.3f')})    
+      
+   end
+end;
+

### 2009-05-14 binding/development/tools/balt/src/java/balt/corba/autogen/FrameworkBasics/ProcessDescriptionHelper.java

@@ -1,0 +1,103 @@
+package balt.corba.autogen.FrameworkBasics;
+
+
+/**
+* balt/corba/autogen/FrameworkBasics/ProcessDescriptionHelper.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from FrameworkBasics.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+abstract public class ProcessDescriptionHelper
+{
+  private static String  _id = "IDL:FrameworkBasics/ProcessDescription:1.0";
+
+  public static void insert (org.omg.CORBA.Any a, balt.corba.autogen.FrameworkBasics.ProcessDescription that)
+  {
+    org.omg.CORBA.portable.OutputStream out = a.create_output_stream ();
+    a.type (type ());
+    write (out, that);
+    a.read_value (out.create_input_stream (), type ());
+  }
+
+  public static balt.corba.autogen.FrameworkBasics.ProcessDescription extract (org.omg.CORBA.Any a)
+  {
+    return read (a.create_input_stream ());
+  }
+
+  private static org.omg.CORBA.TypeCode __typeCode = null;
+  private static boolean __active = false;
+  synchronized public static org.omg.CORBA.TypeCode type ()
+  {
+    if (__typeCode == null)
+    {
+      synchronized (org.omg.CORBA.TypeCode.class)
+      {
+        if (__typeCode == null)
+        {
+          if (__active)
+          {
+            return org.omg.CORBA.ORB.init().create_recursive_tc ( _id );
+          }
+          __active = true;
+          org.omg.CORBA.StructMember[] _members0 = new org.omg.CORBA.StructMember [5];
+          org.omg.CORBA.TypeCode _tcOf_members0 = null;
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _members0[0] = new org.omg.CORBA.StructMember (
+            "m_processName",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = org.omg.CORBA.ORB.init ().create_string_tc (0);
+          _members0[1] = new org.omg.CORBA.StructMember (
+            "m_className",
+            _tcOf_members0,
+            null);
+          _tcOf_members0 = balt.corba.autogen.FrameworkBasics.ProcessLanguageHelper.type ();
+          _members0[2] = new org.omg.CORBA.StructMember (
+            "m_language",
+            _tcOf_members0,
+            null);
[…]

### 2009-05-14 binding/development/tools/balt/src/java/balt/corba/data/translation/CharTranslator.java

@@ -1,0 +1,67 @@
+/*
+ * BALT - The Boxes and Lines Toolkit for component communication.
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+/**
+ * 
+ */
+package balt.corba.data.translation;
+
+import org.omg.CORBA.Any;
+import org.omg.CORBA.BAD_OPERATION;
+import org.omg.CORBA.ORB;
+import org.omg.CORBA.TCKind;
+
+/**
+ * @author nah
+ */
+public class CharTranslator {
+
+    public CharTranslator() {}
+
+    public Any translate(char _data)
+            throws FrameworkDataTranslatorException {
+        Any a = ORB.init().create_any();
+        a.insert_char(_data);
+        return a;
+    }
+
+    public char translate(Any _data)
+            throws FrameworkDataTranslatorException {
+        try {
+            return _data.extract_char();
+        }
+        catch (BAD_OPERATION e) {
+            throw new FrameworkDataTranslatorException(
+                "Unable to extract String from Any", e);
+        }
+    }
+
+    /*
+     * (non-Javadoc)
+     * 
+     * @see sandbox.FrameworkDataTranslator#getTCKind()
+     */
+    public TCKind getTCKind() {
+        return TCKind.tk_char;
+    }
+
+}

### 2009-05-14 binding/development/tools/balt/src/java/balt/core/connectors/push/PushConnectorOut.java

@@ -1,0 +1,42 @@
+/*
+ * BALT - The Boxes and Lines Toolkit for component communication.
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+
+package balt.core.connectors.push;
+
+/**
+ * Interface to define a connector that can be used for push output from
+ * a process. This is separated from the PushConnectorRegister interface
+ * because the remote connectors do not need both aspects of the
+ * functionality.
+ * 
+ * @author nah
+ */
+public interface PushConnectorOut {
+
+    public static final String PUSH_CONNECTOR_OUT_METHOD = "push";
+
+    /**
+     * Flush the connector.
+     */
+    void flush();
+}

### 2009-05-14 binding/development/tools/cast/src/java/cast/cdl/WorkingMemoryChangeFilter.java

@@ -1,0 +1,63 @@
+package cast.cdl;
+
+
+/**
+* cast/cdl/WorkingMemoryChangeFilter.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from CAST.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class WorkingMemoryChangeFilter implements org.omg.CORBA.portable.IDLEntity
+{
+
+  /**
+         * 
+         */
+  public cast.cdl.WorkingMemoryOperation m_operation = null;
+
+  /**
+         * Source of change. If not "" then adhered to.
+         */
+  public String m_src = null;
+
+  /**
+         * If either of these are not "" then the filter will adhere to
+         * them.
+         */
+  public cast.cdl.WorkingMemoryAddress m_address = null;
+  public String m_type = null;
+
+  /**
+         * Whether this filter should only accept changes from local
+         * sources.
+         */
+  public cast.cdl.FilterRestriction m_restriction = null;
+
+  /**
+         * What to do with this filter object.
+         */
+  public cast.cdl.WorkingMemoryOperation m_filterChange = null;
+
+  /**
+         * The component that originally created the change. Used to
+         * distinguish otherwise identical filters when aggregated.
+         */
+  public String m_origin = null;
+
+  public WorkingMemoryChangeFilter ()
+  {
+  } // ctor
+
+  public WorkingMemoryChangeFilter (cast.cdl.WorkingMemoryOperation _m_operation, String _m_src, cast.cdl.WorkingMemoryAddress _m_address, String _m_type, cast.cdl.FilterRestriction _m_restriction, cast.cdl.WorkingMemoryOperation _m_filterChange, String _m_origin)
+  {
+    m_operation = _m_operation;
+    m_src = _m_src;
+    m_address = _m_address;
+    m_type = _m_type;
+    m_restriction = _m_restriction;
+    m_filterChange = _m_filterChange;
+    m_origin = _m_origin;
+  } // ctor
+
+} // class WorkingMemoryChangeFilter

### 2009-05-14 binding/development/tools/cast/src/java/cast/cdl/ui/OutputTypeHolder.java

@@ -1,0 +1,38 @@
+package cast.cdl.ui;
+
+/**
+* cast/cdl/ui/OutputTypeHolder.java .
+* Generated by the IDL-to-Java compiler (portable), version "3.2"
+* from CAST.idl
+* Freitag, 13. MŠrz 2009 17.31 Uhr CET
+*/
+
+public final class OutputTypeHolder implements org.omg.CORBA.portable.Streamable
+{
+  public cast.cdl.ui.OutputType value = null;
+
+  public OutputTypeHolder ()
+  {
+  }
+
+  public OutputTypeHolder (cast.cdl.ui.OutputType initialValue)
+  {
+    value = initialValue;
+  }
+
+  public void _read (org.omg.CORBA.portable.InputStream i)
+  {
+    value = cast.cdl.ui.OutputTypeHelper.read (i);
+  }
+
+  public void _write (org.omg.CORBA.portable.OutputStream o)
+  {
+    cast.cdl.ui.OutputTypeHelper.write (o, value);
+  }
+
+  public org.omg.CORBA.TypeCode _type ()
+  {
+    return cast.cdl.ui.OutputTypeHelper.type ();
+  }
+
+}

### 2009-05-14 binding/development/tools/cast/src/java/cast/ui/UIUtils.java

@@ -1,0 +1,165 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+/**
+ * 
+ */
+package cast.ui;
+
+import balt.clients.ClientUtils;
+import balt.jni.NativeProcessLauncher;
+import cast.cdl.guitypes.*;
+import cast.cdl.ui.*;
+import cast.ui.architecture.interfaces.ComponentEventPullInterface.ComponentEventPullConnector;
+import cast.ui.architecture.interfaces.ComponentEventPushInterface.ComponentEventPushConnector;
+import cast.ui.architecture.interfaces.ComponentStatusPullInterface.ComponentStatusPullConnector;
+import cast.ui.architecture.interfaces.ComponentStatusPushInterface.ComponentStatusPushConnector;
+import cast.ui.architecture.interfaces.TextOutputPullInterface.TextOutputPullConnector;
+import cast.ui.architecture.interfaces.TextOutputPushInterface.TextOutputPushConnector;
+import cast.ui.inspectable.GDBPullInterface.GDBPullConnector;
+import cast.ui.inspectable.GDBPushInterface.GDBPushConnector;
+
+/**
+ * @author nah
+ */
+public class UIUtils {
+
+	private static boolean m_bInit = false;
+
+	public static void init() {
+		if (!m_bInit) {
+
+			ClientUtils.addObjectDatatype(DrawBatch.class,
[…]

### 2009-05-14 binding/development/tools/cast/src/java/cast/ui/architecture/views/ReaderViewAnnotator.java

@@ -1,0 +1,107 @@
+/*
+ * CAST - The CoSy Architecture Schema Toolkit
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+/**
+ * 
+ */
+package cast.ui.architecture.views;
+
+import org.eclipse.swt.SWT;
+import org.eclipse.swt.widgets.Composite;
+import org.eclipse.swt.widgets.Label;
+
+import cast.ui.architecture.ComponentMonitor;
+
+/**
+ * @author nah
+ */
+public class ReaderViewAnnotator extends ComponentViewAnnotator {
+
+    private Label m_changesTypeLabel;
+    private Label m_changesValueLabel;
+    private Label m_readsTypeLabel;
+    private Label m_readsValueLabel;
+
+    private UpdateRunnable m_changesValueRunnable;
+    private UpdateRunnable m_readsValueRunnable;
+    
+    /**
+     * @param _view
+     * @param _monitor
+     */
+    public ReaderViewAnnotator(Composite _view,
+            ComponentMonitor _monitor) {
+        super(_monitor,_view.getDisplay());
+        
+        
+        String changesText = "queued changes/total filtered/total received";
+        m_changesTypeLabel = new Label(_view, SWT.NONE);
+        m_changesTypeLabel.setText("changes: ");
+        m_changesTypeLabel.setToolTipText(changesText);
+        m_changesValueLabel = new Label(_view, SWT.NONE);
+        m_changesValueLabel.setText("  0/  0/  0");
+        m_changesValueLabel.setToolTipText(changesText);
+        m_changesValueRunnable = new UpdateRunnable(m_changesValueLabel);
+        
[…]

### 2009-05-14 binding/development/tools/comedian-architecture/src/c++/comedyarch/NotLaughingNow.cpp

@@ -1,0 +1,217 @@
+/*
+ * Comedian example code to demonstrate CAST functionality.
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+#include "NotLaughingNow.h"
+#include "ComedyGoals.h"
+#include "idl/ComedyEssentials.hh"
+
+#include <cast/architecture/ChangeFilterFactory.hpp>
+using namespace comedyarch::autogen;
+
+using namespace std;
+using namespace boost;
+using namespace cast;
+using namespace comedyarch::autogen;
+
+/**
+ * The function called to create a new instance of our component.
+ *
+ * Taken from zwork
+ */
+extern "C" {
+  FrameworkProcess* newComponent(const string &_id) {
+    return new NotLaughingNow(_id);
+  }
+}
+
+
+NotLaughingNow::NotLaughingNow(const string &_id) : 
+  WorkingMemoryAttachedComponent(_id),
+  ManagedProcess(_id) {
+
+}
+
+NotLaughingNow::~NotLaughingNow() {
+}
+
+void NotLaughingNow::start() {
+  
+  ManagedProcess::start();
+
+  MemberFunctionChangeReceiver<NotLaughingNow> * pReceiver = 
+    new MemberFunctionChangeReceiver<NotLaughingNow>(this,
+					       &NotLaughingNow::newJokeAdded);
+  
+  if(m_bLogOutput) {//lazy!
[…]

### 2009-05-14 binding/development/tools/cosyidl/src/java/org/cognitivesystems/idl/ExceptionEntry.java

@@ -1,0 +1,76 @@
+/*
+ * @(#)ExceptionEntry.java	1.12 03/12/19
+ *
+ * Copyright 2004 Sun Microsystems, Inc. All rights reserved.
+ * SUN PROPRIETARY/CONFIDENTIAL. Use is subject to license terms.
+ */
+/*
+ * COMPONENT_NAME: idl.parser
+ *
+ * ORIGINS: 27
+ *
+ * Licensed Materials - Property of IBM
+ * 5639-D57 (C) COPYRIGHT International Business Machines Corp. 1997, 1999
+ * RMI-IIOP v1.0
+ * US Government Users Restricted Rights - Use, duplication or
+ * disclosure restricted by GSA ADP Schedule Contract with IBM Corp.
+ *
+ * @(#)ExceptionEntry.java	1.12 03/12/19
+ */
+
+package org.cognitivesystems.idl;
+
+// NOTES:
+
+import java.io.PrintWriter;
+import java.util.Enumeration;
+import java.util.Hashtable;
+import java.util.Vector;
+
+/**
+ * This is the symbol table entry for exceptions.  An exception is simply
+ * a struct by a different name.
+ **/
+public class ExceptionEntry extends StructEntry
+{
+  protected ExceptionEntry ()
+  {
+    super ();
+  } // ctor
+
+  protected ExceptionEntry (ExceptionEntry that)
+  {
+    super (that);
+  } // ctor
+
+  protected ExceptionEntry (SymtabEntry that, IDLID clone)
+  {
+    super (that, clone);
+   } // ctor
+
+  public Object clone ()
+  {
+    return new ExceptionEntry (this);
+  } // clone
+
+  /** Invoke the exception generator.
+      @param symbolTable the symbol table is a hash table whose key is
+       a fully qualified type name and whose value is a SymtabEntry or
+       a subclass of SymtabEntry.
+      @param stream the stream to which the generator should sent its output.
+      @see SymtabEntry */
+  public void generate (Hashtable symbolTable, PrintWriter stream)
+  {
+    exceptionGen.generate (symbolTable, this, stream);
+  } // generate
+
+  /** Access the exception generator.
+      @returns an object which implements the ExceptionGen interface.
+      @see ExceptionGen */
+  public Generator generator ()
+  {
+    return exceptionGen;
+  } // generator
+
+  static ExceptionGen exceptionGen;
+} // class ExceptionEntry

### 2009-05-14 binding/development/tools/potential-fields/src/c++/main.cpp

@@ -1,0 +1,427 @@
+#include <iostream>
+#include <stdio.h>
+#include <math.h>
+#include <float.h>
+#include <fstream>
+#include <stdlib.h>
+//#include <conio.h>
+
+
+#include "myVector.h"
+#include "potentialField.h"
+#include "scene.h"
+#include <math.h> //we need cos(..) and sin(..)
+
+#include "ProximityMap.h"
+#include "ProjectiveMap.h"
+
+#ifdef __APPLE__
+
+#include <OpenGL/gl.h>
+#include <OpenGL/glu.h>
+#include <GLUT/glut.h>
+
+#else
+
+#include <GL/gl.h>
+#include <GL/glu.h>
+#include <GL/glut.h>
+
+
+#endif
+
+
+using namespace std;
+
+
+/**********************************************************************
+ ** graphics Functions
+ **********************************************************************/
+
+typedef myVector<float> vector_type;
+typedef potentialField<float> pf_type;
+typedef scene<float> scene_type;
+
+pf_type* renderedField;
+scene_type* scene_pf;
+
+vector_type* viewerPosition;
+vector_type* landmarkPosition;
+vector_type* originPosition;
+
+bool sweetSpot;
+string sName;
+
+int width;
+int height;
+  
+const float DEG2RAD = 3.14159/180;
+
+ProximityMap * proxMap;
+//ProjectiveMap * proxMap;
+ 
+void drawCircle(float radius, float x, float y)
+{
+  //   glBegin(GL_LINE_LOOP);
+  glBegin(GL_LINES);
+ 
+  for (int i=0; i < 361; i++)
+    {
+      float degInRad = i*DEG2RAD;
+      glVertex2f(x, y);
+      glVertex2f(x + cos(degInRad)*radius, y + sin(degInRad)*radius);
+    }
+ 
+  glEnd();
+}
+
+void renderScene(void) {
+
+  float scene_row = renderedField->getRow();
+  float scene_column = renderedField->getColumn();
+	
+  float** DrawingArray = renderedField->getPotentialField();
+
+  if(sweetSpot) {
+    for(int i = 0; i < 2; i++) {
+      float x,y,val;
+      
+      proxMap->nextSweetSpot(x,y,val);
+      
[…]

### 2009-06-09 code/kth/trunk/tools/hardware/laser/src/c++/components/tests/PushClientLaserRaw.hpp
SpatialPeekabotControl now uses Place structs instead of FNodes

@@ -16,7 +16,7 @@
 
 #include <Laser.hpp>
 
-#include <CASTComponent.hpp>
+#include <cast/core/CASTComponent.hpp>
 
 class PushClientLaserRaw : public cast::CASTComponent {
 

### 2009-06-09 kth/trunk/tools/hardware/tests/src/c++/components/LaserStop.hpp
SpatialPeekabotControl now uses Place structs instead of FNodes

@@ -16,7 +16,7 @@
 
 #include <Robotbase.hpp>
 
-#include <CASTComponent.hpp>
+#include <cast/core/CASTComponent.hpp>
 #include <OdometryReceiver.hpp>
 #include <Scan2dReceiver.hpp>
 

### 2009-06-09 kth/trunk/tools/hardware/tts/src/c++/components/TTSServerDoubleTalk.hpp

@@ -1,7 +1,7 @@
 #ifndef TTS_SERVER_DOUBLETALK_HPP_
 #define TTS_SERVER_DOUBLETALK_HPP_
 
-#include <ManagedComponent.hpp>
+#include <cast/architecture/ManagedComponent.hpp>
 
 class TTSServerDoubleTalk : 
   public cast::ManagedComponent {

### 2009-07-03 comsys/tags/comsys.mk5/subarchitectures/comsys/grammars/openccg/moloko.v6/docs/comsys.moloko/the-be-verb.moloko.tex
first release

\section{The \emph{be} Verb} \label{sec-The-Be-Verb}

This section discusses Moloko's treatment of the \emph{be} verb. To begin, here are a few examples which can be used to isolate its broad functions:
\begin{enumerate}
\item  \emph{the cat \ul{was} sitting on the table} 
\item \emph{the cat \ul{was} picked up} 
\item \emph{the cat \ul{was} willing to sit next to me} 
\item \emph{the cat \ul{was} on the table} 
\item \emph{the cat \ul{was} happy}
\item \emph{this \ul{is} a cat}
\item \emph{a cat \ul{is} a kind of animal}
\item \emph{there \ul{was} a cat}
\item \emph{on the table \ul{was} a cat} 
\end{enumerate}

1 and 2 are clear cases of auxilliary verbs, the \emph{ing} \textbf{continuous/progressive} and \textbf{passive} respectively. We have treated the usage in 3, with what are sometimes call predicatival or adjectival verbs, as a auxiliary as well, i.e. we treat \emph{willing} as the main verb/event:\footnote{in this and all the other examples in this section, I have removed all semantic features (\diam{Num}, \diam{Degree}, etc.)}
\begin{verbatim}
  @w1:modal(willing ^ 
            <Mood>ind ^ 
            <Tense>past ^ 
            <Actor>(c1:animate ^ cat )^ 
            <Event>(s1:action-non-motion ^ sit ^ 
                    <Actor>c1:animate ^ 
                    <Modifier>(n1:m-location ^ next ^ 
                               <Anchor>(i1:person ^ I ) ^ 
            <Subject>m1:person)
\end{verbatim}
Thus, for uses 1-3, see section \ref{sec-TAM} for semantics and \ref{sec-Auxiliary-and-Modal-Verbs} for families.

4-7 are various instances of  the \textbf{copular/ascriptive} usage of the verb. The last two are \textbf{presentational/existential}. We will discuss the semantics of each of these in turn. For their families, see \ref{sec-Verbs}

\subsection{Ascription}

The ascriptive use of \emph{be} has this basic semantic structure:
\vspace{-1.5em}

\begin{align*}
\atsign{b1}{ascription}(& \prop{be}  \\ 
      & \wand\ \diam{Cop-Restr}(X)  \\ 
      & \wand\ \diam{Cop-Scope}(Y) \ )
\end{align*}
The \diam{Cop-Restr} is the entity which is being ascribed, and the \diam{Cop-Scope} is what is being ascribed of it. Here are the three main classes of \diam{Cop-Scope}:
\begin{enumerate}
\item \textsc{Nominal}:  \emph{it is \ul{a ball} / \ul{GJ} / \ul{me} / \ul{a kind of coffee} / \ul{the color of the mug}}
\item \textsc{Adjectival}: \emph{it is \ul{blue} / \ul{bigger than the mug} / \ul{happy} /  \ul{off} / \ul{correct} / \ul{ok}} 
\item \textsc{Prepositional}: \emph{it is \ul{on the table} / \ul{here} / \ul{with me} / \ul{for me}}
\end{enumerate}
Class 1 includes but does not distinguish between \textbf{category description}\footnote{Note that we currently do not have a proper treatment  of generic entities. However, it can be 'hacked' as in sentence 5 above.} and \textbf{identificational} uses. Class 2 allows the ascription of any \textbf{quality}, physical, attitudinal or other. Class 3 currently handles \textbf{static location, accompaniment} and \textbf{benefactor}.

In addition to the Indicative uses illustrated above, it can also be used for Imperatives, both both positive ( \emph{be happy}) and negative (\emph{don't be so sad}).
\begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>imp ^ 
                 <Cop-Restr>(a1:entity ^ addressee) ^ 
                 <Cop-Scope>(h1:q-attitude ^ happy) ^ 
                 <Subject>a1:entity)

  @b1:ascription(be ^ 
                 <Mood>imp ^ 
                 <Polarity>neg ^ 
                 <Cop-Restr>(a1:entity ^ addressee) ^ 
                 <Cop-Scope>(s1:q-attitude ^ sad) ^ 
                 <Subject>a1:entity)
\end{verbatim}

It can also be used for Closed (Y/N) Interrogatives, again positive ( \emph{is it a ball}) and negative (\emph{wasn't it here earlier}).
\begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(i1:thing ^ it ) ^ 
                 <Cop-Scope>(b2:thing ^ ball ) 
                 <Subject>i1:thing)

  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Polarity>neg ^ 
                 <Tense>past ^ 
                 <Cop-Restr>(i1:thing ^ it ) ^ 
                 <Cop-Scope>(c1:m-location ^ context ) ^ 
                 <Modifier>(e1:m-time ^ early ) ^ 
                 <Subject>i1:thing)
\end{verbatim}

Finally, it can be used in Open (Wh) Interrogatives to question both the \diam{Cop-Scope} role and the \diam{Cop-Restr}. Like all events, it can also be questioned for modifiers. These will be handled in turn. \\ \\
\textbf{ \ul{Questioning the \diam{Cop-Restr} }} \\ \\
An entity can be given with a prompt for 'filling in' some 'property' specified by the Wh-Word. Here are some examples sorted by the classes specified above:

\begin{tabbing}
\textsc{1. Nominal} \\ \\
\emph{\ul{what} is this thing on the table} \=  it is \ul{a ball} \\
\emph{\ul{who} is it} \> it is \ul{me} (see below)\\
\emph{\ul{who} is that guy} \> it is \ul{my dad} \\
\emph{\ul{what} is a cat}  \> a cat is \ul{a kind of animal} \\
\emph{\ul{which ball} is mine}  \> yours is \ul{the one over there} \\
\end{tabbing}
\begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(i1:entity ^ it)
                 <Cop-Scope>(w1:entity ^ what) ^ 
                 <Subject>c1:entity ^ 
                 <Wh-Restr>w1:entity)
\end{verbatim}

\begin{tabbing}
\textsc{2. Adjectival} \\ \\
\emph{\ul{what shape} is it}  \= it is \ul{round} \\
\emph{\ul{what color} is it} \> it is \ul{red} (see below) \\
\emph{\ul{what size} is it}  \> it is \ul{small} \\
\emph{\ul{how} are you} \>  I am \ul{good}\footnote{\emph{how} does not restrict the kind quality given} (see below) \\
\emph{\ul{how big} is it} \> it is \ul{really big}\footnote{this is ambiguous between the 'degree' reading and the standard 'scale' reading (c.f. \emph{what size}). We have not distinguished between these two readings.} \\
\emph{\ul{how cold} is it} \> it is \ul{ok}
\end{tabbing}
\begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(y1:person ^ you)
                 <Cop-Scope>(h1:quality ^ how) ^ 
                 <Subject>y1:person ^ 
                 <Wh-Restr>h1:quality)

  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(i1:thing ^ it ) ^ 
                 <Cop-Scope>(c1:quality ^ color) ^ 
                 <Subject>i1:thing ^ 
                 <Wh-Restr>(w1:specifier ^ what ^ 
                            <Scope>c1:quality))

\end{verbatim}

\begin{tabbing}
\textsc{1. Prepositional} \\ \\
\emph{\ul{where} is the big green mug} \=  it is \ul{on your table} (see below) \\
\emph{\ul{what} is it under} \>  it is under \ul{the table} (see below) \\
\emph{\ul{which room} is it in} \>  it is in \ul{GJ's office}  \\
\emph{\ul{who} is it for} \> it is for \ul{me} \\
\emph{\ul{which cat} is this milk for} \> it is for \ul{mine} (see below)  \\
\emph{\ul{who} is he with} \> he is with \ul{GJ} \\
\end{tabbing}
\begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(m1:thing ^ mug ^ 
                             <Modifier>(b2:q-size ^ big) ^ 
                             <Modifier>(g1:q-color ^ green)) ^ 
                 <Cop-Scope>(w1:m-location ^ where) ^ 
                 <Subject>m1:thing ^ 
                 <Wh-Restr>w1:m-location)

   @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(i1:thing ^ it ) 
                 <Cop-Scope>(u1:m-location ^ under ^ 
                             <Anchor>w1:physical) ^ 
                 <Subject>i1:thing ^ 
                 <Wh-Restr>(w1:physical ^ what))

   @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(m1:thing ^ milk ) 
                 <Cop-Scope>(f1:m-benefactor ^ for ^ 
                             <Anchor>(c2:animate ^ cat ^ )
                 <Subject>c1:thing ^ 
                 <Wh-Restr>(w1:specifier ^ which ^ 
                            <Scope>c2:animate))

 \end{verbatim}
A note on questioning \textbf{physical properties} (color, size, shape, etc.) In addition \emph{what \ul {color/size/shape} is the mug}, there is the option of \emph{what is the \ul{color/size/shape} of the mug}. The former is treated as an 'adjectival' questioning, scoping over \diam{Cop-Scope}\emph{(q:quality)} (see above). The latter is treated very naively,  scoping over the whole \diam{Cop-Scope} AND as some sort of entity. We hope unify these two readings sometime soon.
 \begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>pres ^ 
                 <Cop-Restr>(c1:xxxx ^ color ^ 
                             <Delimitation>unique ^ 
                             <Num>sg ^ 
                             <Quantification>specific ^ 
                             <Owner>(b2:thing ^ ball ^ 
                                     <Delimitation>unique ^ 
                                     <Num>sg ^ 
                                     <Quantification>specific)) ^ 
                 <Cop-Scope>(w1:xxxx ^ what) ^ 
                 <Subject>c1:entity ^ 
                 <Wh-Restr>w1:xxxx)
 \end{verbatim} \\ 
\textbf{ \ul{Questioning the \diam{Cop-Restr} }} \\ \\
In addition,  a 'property' can be given with a prompt for 'filling in' what entity(s) this applies to. So, for example, \emph{what is big} or \emph{what is on the table}. Of course all of the other Wh-words are available:  \emph{who is nicer},\footnote{we have not yet handled \textbf{constrained option} questions, like \emph{who is nicer, me or him}} \emph{which coffee is black}, \emph{how\_many balls are in here}, etc. So, for example, here is \emph{who was in there}: 
\begin{verbatim}
  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>past ^ 
                 <Cop-Restr>(w1:animate ^ who) ^ 
                 <Cop-Scope>(i1:m-location ^ in ^ 
                             <Anchor>(c1:e-location ^ context ) 
                 <Subject>w1:animate ^ 
                 <Wh-Restr>w1:animate)
\end{verbatim}
The \diam{Cop-Restr} for class 1 (nominals) is 'shut off' from questioning, i.e. \emph{what is this} does not expect an answer like \emph{a ball is this}. There maybe a sub-class of cases where such questioning would make sense, but this is for future work.
\\Ê\\ \textbf{ \ul{Questioning the Ascription Itself }} \\ \\
Currently, the time and place of the ascription can be questioned. For example, \emph{where was he happy} and \emph{when was he here}:
\begin{verbatim}
@b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>past ^ 
                 <Cop-Restr>(h1:person ^ he ^) ^ 
                 <Cop-Scope>(h2:q-attitude ^ happy) ^ 
                 <Subject>h1:person ^ 
                 <Wh-Restr>(w1:m-location ^ where ^ 
                            <Scope>b1:ascription))

  @b1:ascription(be ^ 
                 <Mood>int ^ 
                 <Tense>past ^ 
                 <Cop-Restr>(h1:person ^ he) ^ 
                 <Cop-Scope>(c1:m-location ^ context ^ 
                             <Proximity>m-proximal) ^ 
                 <Subject>h1:person ^ 
                 <Wh-Restr>(w1:m-time-point ^ when ^ 
                            <Scope>b1:ascription))
\end{verbatim}

\subsection{Presentation}

The \emph{be} verb can also be used to present entities 'on the scene' and/or to assert/deny their existence. Here are some indicative examples: \\

      \emph{there is \ul{a mug} on the table}

     \emph{there are \ul{two mugs and a book}}

      \emph{there were \ul{some good books} yesterday}

      \emph{there has never been \ul{a table} there} \\ \\
Here, for example, is the first:
\begin{verbatim}
  @b1:presentational(be ^ 
                     <Mood>ind ^ 
                     <Tense>pres ^ 
                     <Presented>(m1:thing ^ mug) 
                     <Modifier>(o1:m-location ^ on ^ 
                              <Anchor>(t1:thing ^ table ^)) 
                     <Subject>(t2:dummy ^ there) )
\end{verbatim}
The entity is placed in the \diam{Presented} role, and additional modifiers (location, time, etc) modify the event itself.\footnote{in fact, this example is treated as ambiguous. It also has a reading where the mug is modified. If there is any difference in meaning, we certainly can't figure it out. Note that the possibility of other modifiers (time, etc) and the possibility of questions like \emph{where were there balls } suggests that this location should be able to modify the event.} The \diam{Subject} is filled by the 'dummy' word \emph{there}. This does not contribute to the semantics, but is required for proper generation and for handling the syntax of questions.

In addition, there is a special indicative construction exemplified by \emph{in the room were some cats and a dog}:
\begin{verbatim}
  @b1:presentational(be ^ 
                     <Mood>ind ^ 
                     <Tense>past ^ 
                     <Modifier>(i1:m-location ^ in ^ 
                                <Anchor>(r1:e-place ^ room) ^
                     <Presented>(a1:entity ^ and ^ 
                                <First>(c1:animate ^ cat) 
                                 <Next>(d1:animate ^ dog ))
\end{verbatim}
Note that this leads to two readings for \emph{there is my ball}, with with a 'real' locational \emph{there} and one with a dummy.

There is no imperative form for the presentational, but it can be used for both Open and Closed Interrogatives. For example, \emph{were there any dogs there} and \emph{what was there yesterday}\footnote{this is also ambiguous between \emph{there} as dummy subject or as the \diam{Cop-Scope} of ascription.}:
\begin{verbatim}
  @b1:presentational(be ^ 
                     <Mood>int ^ 
                     <Tense>past ^ 
                     <Modifier>(c1:m-location ^ context ^ 
                                <Proximity>m-distal) ^ 
                     <Presented>(d1:animate ^ dog) ^ 
                     <Subject>(t1:dummy ^ there))

  @b1:presentational(be ^ 
                     <Mood>int ^ 
                     <Tense>past ^ 
                     <Modifier>(y1:m-time-point ^ yesterday) ^ 
                     <Presented>(w1:entity ^ what) ^ 
                     <Subject>(t1:dummy ^ there) ^ 
                     <Wh-Restr>w1:entity)
\end{verbatim}

### 2009-07-03 comsys/tags/comsys.mk5/tools/cast/src/c++/cast/examples/comedyarch/AssistantDirector.cpp

@@ -1,0 +1,122 @@
+/*
+ * Comedian example code to demonstrate CAST functionality.
+ *
+ * Copyright (C) 2006-2007 Nick Hawes
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public License
+ * as published by the Free Software Foundation; either version 2.1 of
+ * the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
+ * 02110-1301 USA
+ *
+ */
+
+#include "AssistantDirector.hpp"
+#include <ComedyEssentials.hpp>
+#include <ChangeFilterFactory.hpp>
+using namespace comedyarch::autogen;
+
+using namespace std;
+using namespace boost;
+using namespace cast;
+
+
+/**
+ * The function called to create a new instance of our component.
+ *
+ * Taken from zwork
+ */
+extern "C" {
+  cast::CASTComponentPtr newComponent() {
+    return new AssistantDirector();
+  }
+}
+
+
+
+void AssistantDirector::start() {
+  
+  MemberFunctionChangeReceiver<AssistantDirector> * jokeReceiver = 
+    new MemberFunctionChangeReceiver<AssistantDirector>(this,
+							&AssistantDirector::handleTwoLiner);
+    
+  addChangeFilter(createChangeFilter<TwoLiner>(cdl::OVERWRITE, "", "", "stage.subarch",
+					   cdl::ALLSA),		  
+		  jokeReceiver);  
+  
+
+  MemberFunctionChangeReceiver<AssistantDirector> * reactionReceiver = 
+    new MemberFunctionChangeReceiver<AssistantDirector>(this,
+					       &AssistantDirector::handleReaction);
+
+  addChangeFilter(createGlobalTypeFilter<Reaction>(cdl::ADD),
+		  reactionReceiver);  
+
+
+}
+
+void AssistantDirector::handleTwoLiner(const cdl::WorkingMemoryChange & _wmc) {
+  try {
+    // get the joke from working memory to see what
+    // we can do
+
[…]

### 2009-07-03 code/comsys/tags/comsys.mk5/subarchitectures/comsys/src/java/comsys/processing/parse/GrammaticalInference.java
first release

@@ -1,0 +1,107 @@
+// =================================================================
+// Copyright (C) 2007 Geert-Jan M. Kruijff (gj@dfki.de)
+//
+// This library is free software; you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License
+// as published by the Free Software Foundation; either version 2.1 of
+// the License, or (at your option) any later version.
+//
+// This library is distributed in the hope that it will be useful, but
+// WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU
+// Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU Lesser General Public
+// License along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA
+// 02111-1307, USA.
+// =================================================================
+
+// =================================================================
+// PACKAGE DEFINITION
+// =================================================================
+
+package comsys.processing.parse;
+
+// =================================================================
+// IMPORTS
+// =================================================================
+
+// -----------------------------------------------------------------
+// CAST IMPORTS
+// -----------------------------------------------------------------
+
+import cast.core.CASTData;
+
+// -----------------------------------------------------------------
+// COMSYS IMPORTS
+// -----------------------------------------------------------------
+
+import comsys.datastructs.comsysEssentials.*;
+import comsys.arch.ComsysException;
+
+
+// -----------------------------------------------------------------
+// JAVA IMPORTS
+// -----------------------------------------------------------------
+import java.util.List;
+import java.util.Properties;
+
+// -----------------------------------------------------------------
+// OPENCCG IMPORTS
+// -----------------------------------------------------------------
+import opennlp.ccg.parse.ParseResults;
+import opennlp.ccg.parse.ParseException;
+
+
+// =================================================================
+// INTERFACE DOCUMENTATION
+// =================================================================
+
+/**
+ * The interface <b>GrammaticalInference</b> defines methods that 
+ * any parse process should provide. 
+ *
+ * @version 080906 (Started: 070810)
+ * @author Geert-Jan M. Kruijff (gj@dfki.de)
+*/
+
+public interface GrammaticalInference { 
+
+	/** 
[…]

### 2009-09-21 kth/trunk/doc/cogx-kth-git/source/cogx-kth-git.tex
Docs for svn/git maintenance (KTH only)

--- 2009-04-12 kth/branches/year1/doc/cogx-kth-git/source/cogx-kth-git.tex
@@
        \texttt{http://book.git-scm.com/}
+  \item Git FAQ \\
+       \texttt{http://git.or.cz/gitwiki/GitFaq}
   \item Git for the lazy \\
@@
     \end{lstlisting}
+\end{itemize}
 
@@
 \subsection{Pulling Changes from SVN}
-\hl{Missing}
+Setup a joint GIT/SVN local repository:
+\begin{itemize}
+\item setup another local repository for instance by making a copy of your current 
+git repository: cp -r my\_git\_repo svn-git\_repo
+\item enter the directory with the repository: cd svn-git\_repo
+\item checkout branch master: git checkout master
+\item remove all files apart from the .git directory
+\item run svn co https://codex.cs.bham.ac.uk/svn/nah/cogx/code/kth/trunk
+\item move all the files from trunk to your git working copy: mv trunk/* .
+\end{itemize}
 
+Now you have a joint git/svn working copy. To pull new changes from svn and 
+merge them to git, do as follows:
+\begin{itemize}
+\item if you messed up sth with the files, you can do git reset --hard
+\item now, pull the most recent version of the master branch: git pull
+\item update the changes from the SVN: svn up
+\item commit the changes to git: git add/git commit
+\item push the changes to the shared repository: git push
 \end{itemize}
@@
     \end{lstlisting}
+  \item How to throw away uncommitted change in a single file?
+    \begin{lstlisting}
+git checkout -- file
+    \end{lstlisting}
 

### 2009-10-23 code/subarchitectures/comsys/branches/stable-for-binder-7/src/java/comsys/utils/Pair.java
made branch to go with binder 7

@@ -1,0 +1,39 @@
+package comsys.utils;
+
+public class Pair<A, B> {
+	
+    public A fst;
+    public B snd;
+ 
+    public Pair() {
+    	this.fst = null;
+    	this.snd = null;
+    }
+
+    public Pair(A _fst, B _snd) {
+        this.fst = _fst;
+        this.snd = _snd;
+    }
+    
+    public boolean equals(Object o) {
+        if (!(o instanceof Pair))
+            return false;
+ 
+        Pair<?, ?> other = (Pair) o;
+        return equal(fst, other.fst) && equal(snd, other.snd);
+    }
+    
+    private static final boolean equal(Object o1, Object o2) {
+        if (o1 == null) {
+            return o2 == null;
+        }
+        return o1.equals(o2);
+    }
+ 
+    public int hashCode() {
+        int hFst = fst == null ? 0 : fst.hashCode();
+        int hSnd = snd == null ? 0 : snd.hashCode();
+ 
+        return hFst + (57 * hSnd);
+    }
+}

### 2009-10-23 code/subarchitectures/comsys/branches/stable-for-binder-7/src/mercury-c++/abduction/server-mk2/ForwardedAbducerServer.cc

@@ -1,0 +1,322 @@
+#include "common.h"
+#include "ForwardedAbducerServer.h"
+
+#include "Tokens.h"
+#include "TermTokeniser.h"
+#include "SliceToString.h"
+#include "StringToSlice.h"
+
+#include "TtyUtils.h"
+
+using namespace std;
+using namespace Abducer;
+
+#include <vector>
+
+static const size_t bufsize = 8192;
+
+static char buf[bufsize];
+
+ForwardedAbducerServer::ForwardedAbducerServer()
+{
+	cerr << tty::green << "* initialising abducer context" << tty::dcol << endl;
+	cout << "init_ctx." << endl;
+}
+
+void
+ForwardedAbducerServer::loadFile(const string& filename, const Ice::Current&)
+{
+	cerr << tty::green << "* loading file `" << filename << "'" << tty::dcol << endl;
+	cout << "load_file(\"" << filename << "\")." << endl;
+
+	if (cin) {
+		cin.getline(buf, bufsize);
+		cerr << tty::red << "  [unimplemented: check success]" << tty::dcol << endl;
+		debug(cerr << "  load file reply: " << buf << endl);
+	}
+
+/*
+	MR_Word w_result;
+	char * args;
+	int argi;
+
+	load_file(s, &w_result, ctx, &ctx);
+
+	if (load_result_is_ok(w_result)) {
+		cout << tty::green << "  result: ok" << tty::dcol << endl;
+	}
+	else if (load_result_is_file_read_error(w_result)) {
+		cout << tty::red << "  file read error" << tty::dcol << endl;
+		throw FileReadErrorException(filename);
+	}
+	else if (load_result_is_syntax_error(w_result, &args, &argi)) {
+		cout << tty::red << "  syntax error: " << args << " on line " << argi << tty::dcol << endl;
+		throw SyntaxErrorException(filename, args, argi);
+	}
+*/
+}
+
+void
+ForwardedAbducerServer::clearRules(const Ice::Current&)
+{
+	cerr << tty::green << "* clearing rules" << tty::dcol << endl;
+	cout << "clear_rules." << endl;
+}
+
+void
+ForwardedAbducerServer::clearFacts(const Ice::Current&)
+{
+	cerr << tty::green << "* clearing all facts" << tty::dcol << endl;
+	cout << "clear_facts." << endl;
+}
+
+void
[…]

### 2010-06-09 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: documentation

% vim:ft=tex:sw=3:ts=8:et:fileencoding=utf-8

\section{Installation}

\begin{itemize}
   \item Add visualization.sa to SVN externals:
   \begin{verbatim}
      $ svn pe svn:externals .
   \end{verbatim}
   Add the line:
   \begin{verbatim}
      ^/code/subarchitectures/visualization.sa/trunk   subarchitectures/visualization.sa
   \end{verbatim}

   % TODO: the script may be in another location
   \item install the required libraries; there is a script in visualization.sa/config
   that you can run:
   \begin{verbatim}
      sudo ubuntu-install-deps.sh
   \end{verbatim}

   \item enable BUILD\_VISUALIZATION\_SA in cmake
\end{itemize}

\section{General operation}

The library provieds a display server that displays objects created by its
clients.  The objects that are on the display server are groupped into views.
Initially every new object is displayed in its own view that has the same name
as the object.
% (TODO: Additional views can be defined on the server that display multiple objects).
The display server creates rendering contexts in which the views are displayed.
There are two types of contexts: 2D and OpenGL.
% (TODO: new contexts: tabular; textual; ...)

The client that wants to display an object connects to the server and sends the
serialized object to the server. Each object has a name. If an object with the
same name already exists on the srver it is updated with the new data. Some
objects are composed of multiple parts where each part can be manipulated
independently.

Currently the following types of objects are suported: images, SVG graphics and
OpenGL objects define with LuaGl script or as TomGine render-models.

The display server can also display simple user interface elements. If a view
is active (visible) all the UI element associated with the view are also
displayed. When the state of a UI element changes, eg.  because of a muse
click, the client that owns the UI element is notified and it receives the new
state of the object.

(2010-06-09: at this time only two types of UI elements are provided: a button
and a checkbox).

\section{Sending data to the server}

The library provides a client component that interacts with the server. The
client is devised in such a way that its use is as trivial as the use of
OpenCV's function cvShowImage(). For simple objects the server can also
be used without a client if the developer is willing to use the file system
as a communication channel. This involves a CAST component that monitors
the file system and sends the newly created files to the server.

First we will show how to use the file system monitor and then we will give
instructions for the use of the real client.

\subsection{File system monitor}

The File system monitor is a CAST component that monitors the filesystem for
changes. Usualy it observes a directory on the disk and waits for files in that
directory that have been closed after they were modified. If the modified file
is one of the files the monitor is waiting for, it reads the file processes it
and sends it to the server. A single component is able to monitor multiple files
in multiple directories.

The files that the monitor can process are images, SVG files and files created
for GraphWiz.

Here is an example CAST setup of a DisplayServer and a FileMonitorDC. Note that
all text after fileviewer.cli is on the same line.

\begin{verbatim}
   HOST localhost
   SUBARCHITECTURE visualization.sa 
   CPP WM SubarchitectureWorkingMemory
   CPP TM AlwaysPositiveTaskManager

   CPP MG display.srv DisplayServer

   CPP MG fileviewer.cli FileMonitorDC --display-server "display.srv"
      --setvars "dir=subarchitectures/visualization.sa/src/c++/clients/fileviewer/test/xdata"
      --monitor "
         CurrentPlan=dot:%(dir)/{*.dot}
         AltPlan=twopi:%(dir)/{*.dot}
         ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
         SvgFile=%(dir)/{*.svg}
         %(dir)/{*}"
         --log true --debug true
\end{verbatim}

The file monitor first connects to the server using the parameter
\code{--display-server}. If the server is not present, a display client will
operate normally except it won't be able to display its resutls.

The parameter \code{--setvars} sets some variables thata are used later when
the component configures the monitored files. The format of the parameter
string is \code{"var1=value1 var2=value2 ..."}. Later in the paramteter
\code{--monitor} a variable can be inserted with \code{\%(var1)}.

The most important and complex part is the definition of the files to be
monitored, the \code{--monitor} parameter. It is composed of multiple entries
that are separated by spaces. Each entry defines the directory where the files
are, the names of the files to monitor, the converters for the files and the
name of the object that will be created from the files. The general format of
each entry is:
\begin{verbatim}
   ObjectName=converter:directory/{filemasks}
\end{verbatim}

If the converter is omitted, one of the known converters will be selected.
If the object name is omitted, the name of the object will be the filename.

We can now decode the meanings of the above definitions:

\begin{verbatim}
   CurrentPlan=dot:%(dir)/{*.dot}
   AltPlan=twopi:%(dir)/{*.dot}
\end{verbatim}

Every new dot file in the directory will be converted with two converters: dot
and twopi. The files converted with dot will be displayed on the server as
CurrentPlan while the files converted with twopi well be displayed as AltPlan.
The file monitor also knows the following converters for dot: neato, circo and
fdp.

\begin{verbatim}
   ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
   SvgFile=%(dir)/{*.svg}
\end{verbatim}

Images and SVG files can be sent without conversion if the file extension is
one of the known ones. If a SVG file is stored in a file with an extension
other than \code{.svg}, the converter must be specified although no conversion
will be performed:

\begin{verbatim}
   SvgFile=svg:%(dir)/{*.xml}
\end{verbatim}

The final entry will pick up any files that change in the directory. If the
type of the file (deduced from the extension) is one of the known types, the
file will be sent to the server. The name of the object on the server will be
the name of the file.

\begin{verbatim}
   %(dir)/{*}"
\end{verbatim}

\subsection{A Passive Display Client}

The library provides two types of clients: a passive client that only sends
data to the server and an active client that interacts with the server using UI
elements. We will first show how to create a passive client and then convert it
to an active client. We will use the VideoViewer component as an example. Only
the relevant parts of the code will be shown. Full code is in the VideoViewer
component in vision.sa.

Before we can start we need to prepare the project file to use the display server.
For every component that is defined in CMakeLists.txt we add the following:

\begin{verbatim}
   include_directories(${VISUALIZATION_INCLUDE_DIRS})
   link_cast_component(${CAST_COMPONENT_NAME} Video ${VISUALIZATION_LIBRARIES})
\end{verbatim}

If the component uses OpenCV's IplImage we can let the client convert the image
before it is sent to the server. To do this we need to enable the client's OpenCV
interface with

\begin{verbatim}
   add_definitions(-DFEAT_VISUALIZATION_OPENCV)
\end{verbatim}

Now the component is ready to interact with the display server. The component
will communicate with the servet through a display client:

\begin{verbatim}
   class VideoViewer : public ManagedComponent, public VideoClient
   {
   private:
   #ifdef FEAT_VISUALIZATION
     cogx::display::CDisplayClient m_display;
   #endif
   };
\end{verbatim}

Note that all the code related to visualization.sa is enclosed in conditional
blocks. The variable \code{FEAT\_VISUALIZATION} is defined when
viusalisation.sa is being built. If the visualization.sa is not used in the
project, the componetns that depend on it will continue to work, but they won't
display their data unless an alternative is provided (see VideoViewer.cpp which
uses OpenCV when visualization.sa is not used).

The client must connect to the server and for that it needs the parameters of the
connection (the parameter \code{--displaysever}):

\begin{verbatim}
   void VideoViewer::configure(const map<string,string> & _config)
   {
   #ifdef FEAT_VISUALIZATION
     m_display.configureDisplayClient(_config);
   #endif
   }
\end{verbatim}

The connection between the client and the server is established in the component's
\code{start()} method:
\begin{verbatim}
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
      m_display.connectIceClient(*this);
   #endif
   }
\end{verbatim}

When a new image is available, the VideoServer sends the image to the VideoViewer which
passes the image to the DisplayServer:

\begin{verbatim}
   void VideoViewer::receiveImages(const std::vector<Video::Image>& images)
   {
   #ifdef FEAT_VISUALIZATION
     m_display.setImage(getComponentID(), images[0]);
   #endif
   }
\end{verbatim}

Here \code{getComponentID()} is used for the name of the object since the component
is displaying only one object. Any other string can be used instead.

\subsection{An Active Display Client}

An active display client is needed when the component wants to present some UI
elements with which it can be configured. The client receives notifications and
requests related to these elements from the display server. The component must
define two callback functions that will handle the events sent by the server.

\begin{verbatim}
   class VideoViewer : public ManagedComponent, public VideoClient
   {
   private:
   #ifdef FEAT_VISUALIZATION
     cogx::display::CActiveDisplayClient<VideoViewer> m_display;
     void handleGuiEvent(const Visualization::TEvent &event);
     std::string getControlState(const std::string& ctrlId);
   #endif
   };
\end{verbatim}

The event path from the server to the client must be initiated. This is done right
after the component connects to the client. After the connections are established,
the callback functions are registered:

\begin{verbatim}
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
     m_display.connectIceClient(*this);
     m_display.installEventReceiver();
     m_display.setEventCallback(this, &VideoViewer::handleGuiEvent);
     m_display.setStateQueryCallback(this, &VideoViewer::getControlState);
   #endif
   }
\end{verbatim}

We define a checkbox element that will start or stop the flow of images
from the server:

\begin{verbatim}
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
   #endif
   }
\end{verbatim}

Now all that is missing are the callback functions:

\begin{verbatim}
   #ifdef FEAT_VISUALIZATION
   void VideoViewer::handleGuiEvent(const Visualization::TEvent &event)
   {
     if (event.type == Visualization::evCheckBoxChange) {
       if (event.sourceId == "toggle.viewer.running") {
         bool newrcv = (event.data != "0");
         if (newrcv != receiving) {
           if(receiving) {
             videoServer->stopReceiveImages(getComponentID());
           }
           else {
             videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
           }
           receiving = !receiving;
         }
       }
     }
   }
   #endif
\end{verbatim}

\begin{verbatim}
   #ifdef FEAT_VISUALIZATION
   std::string VideoViewer::getControlState(const std::string& ctrlId)
   {
     if (ctrlId == "toggle.viewer.running") {
       return receiving ? "2" : "0";
     }
     return "";
   }
   #endif
\end{verbatim}

### 2010-06-10 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs updated

@@
    \item Add visualization.sa to SVN externals:
-   \begin{verbatim}
+   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
       $ svn pe svn:externals .
-   \end{verbatim}
+   \end{Verbatim}
    Add the line:
-   \begin{verbatim}
+   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
       ^/code/subarchitectures/visualization.sa/trunk subarchitectures/visualization.sa
-   \end{verbatim}
+   \end{Verbatim}
 
@@
    that you can run:
-   \begin{verbatim}
+   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
       sudo ubuntu-install-deps.sh
-   \end{verbatim}
+   \end{Verbatim}
 
@@
 
-The library provieds a display server that displays objects created by its
-clients.  The objects that are on the display server are groupped into views.
+The library provides a display server that displays objects created by its
+clients.  The objects that are on the display server are grouped into views.
 Initially every new object is displayed in its own view that has the same name
 as the object.
-% (TODO: Additional views can be defined on the server that display multiple objects).
+\todo{(TODO: Additional views that display multiple objects can be defined on the server)}
 The display server creates rendering contexts in which the views are displayed.
@@
 serialized object to the server. Each object has a name. If an object with the
-same name already exists on the srver it is updated with the new data. Some
+same name already exists on the server it is updated with the new data. Some
 objects are composed of multiple parts where each part can be manipulated
@@
 
-Currently the following types of objects are suported: images, SVG graphics and
+Currently the following types of objects are supported: images, SVG graphics and
 OpenGL objects define with LuaGl script or as TomGine render-models.
@@
 
-The File system monitor is a CAST component that monitors the filesystem for
-changes. Usualy it observes a directory on the disk and waits for files in that
+The File system monitor is a CAST component that monitors the file system for
+changes. Usually it observes a directory on the disk and waits for files in that
 directory that have been closed after they were modified. If the modified file
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    HOST localhost
@@
 
-   CPP MG fileviewer.cli FileMonitorDC --display-server "display.srv"
+   CPP MG fileviewer.cli FileMonitorDC --displayserver "display.srv"
       --setvars "dir=subarchitectures/visualization.sa/src/c++/clients/fileviewer/test/xdata"
@@
          --log true --debug true
-\end{verbatim}
+\end{Verbatim}
 
-The file monitor first connects to the server using the parameter
-\code{--display-server}. If the server is not present, a display client will
-operate normally except it won't be able to display its resutls.
+The file monitor first connects to the server using the \code{--displayserver}
+parameter. If the server is not present, a display client will operate normally
+except it won't be able to display its results.
 
-The parameter \code{--setvars} sets some variables thata are used later when
+The parameter \code{--setvars} sets some variables that are used later when
 the component configures the monitored files. The format of the parameter
-string is \code{"var1=value1 var2=value2 ..."}. Later in the paramteter
+string is \code{"var1=value1 var2=value2 ..."}. Later in the parameter
 \code{--monitor} a variable can be inserted with \code{\%(var1)}.
@@
 each entry is:
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    ObjectName=converter:directory/{filemasks}
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    CurrentPlan=dot:%(dir)/{*.dot}
    AltPlan=twopi:%(dir)/{*.dot}
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
    SvgFile=%(dir)/{*.svg}
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    SvgFile=svg:%(dir)/{*.xml}
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    %(dir)/{*}"
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    include_directories(${VISUALIZATION_INCLUDE_DIRS})
    link_cast_component(${CAST_COMPONENT_NAME} Video ${VISUALIZATION_LIBRARIES})
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    add_definitions(-DFEAT_VISUALIZATION_OPENCV)
-\end{verbatim}
+\end{Verbatim}
 
 Now the component is ready to interact with the display server. The component
-will communicate with the servet through a display client:
+will communicate with the server through a display client:
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    class VideoViewer : public ManagedComponent, public VideoClient
@@
    };
-\end{verbatim}
+\end{Verbatim}
 
@@
 blocks. The variable \code{FEAT\_VISUALIZATION} is defined when
-viusalisation.sa is being built. If the visualization.sa is not used in the
-project, the componetns that depend on it will continue to work, but they won't
+visualization.sa is being built. If the visualization.sa is not used in the
+project, the components that depend on it will continue to work, but they won't
 display their data unless an alternative is provided (see VideoViewer.cpp which
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    void VideoViewer::configure(const map<string,string> & _config)
@@
    }
-\end{verbatim}
+\end{Verbatim}
 
@@
 \code{start()} method:
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    void VideoViewer::start()
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    void VideoViewer::receiveImages(const std::vector<Video::Image>& images)
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    class VideoViewer : public ManagedComponent, public VideoClient
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    void VideoViewer::start()
@@
    }
-\end{verbatim}
+\end{Verbatim}
 
-We define a checkbox element that will start or stop the flow of images
-from the server:
+The flow of images from the VideoServer to the VideoViewer can be stopped at
+any time. To let the user control the flow the component has to add a UI
+element to the display server.  An appropriate type of UI element for this
+purpose is a chekcbox:
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    void VideoViewer::start()
@@
    }
-\end{verbatim}
+\end{Verbatim}
 
-Now all that is missing are the callback functions:
+The first parameter is the name of the {\em view} that will display the UI
+element when it is active, not the name of object. This means that the UI
+element will be displayed only when an object is displayed in its default view.
+When an object is displayed in another view, the UI element won't be available.
+\todo{TODO: user-defined views will be able to specify which elements to show}
 
-\begin{verbatim}
+The \code{handleGuiEvent()} callback function will be called every time the
+UI element emits a known event. For the chekcbox element we need to respond
+to the \code{evCheckBoxChange} event:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    #ifdef FEAT_VISUALIZATION
@@
    #endif
-\end{verbatim}
+\end{Verbatim}
 
-\begin{verbatim}
+Note that the data for UI elements passed between the client and the server is
+serialized into strings. The checkbox element knows only 2 values: \code{"0"}
+and everything else.
+\todo{TODO: Add more element types, define values}
+
+A UI element may be displayed in multiple views. When a UI element is first
+displayed it must be put in a state that reflects the current state of the
+component.  Since a UI element may be created at any time it wouldn't be enough
+to set the initial state of the element at creation time. Instead, we define
+the \code{getControlState()} callback function:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    #ifdef FEAT_VISUALIZATION
@@
    #endif
-\end{verbatim}
+\end{Verbatim}
+
+When the function returns \code{""} the display server won't change the state
+of the UI element.
+
+\section{Displaying objects}
+
+The server can display various types of objects. Initially it was created to
+display images, but with time other objects (supported by Qt) were added:
+SVG images and OpenGL scenes.
+
+\subsection{Images}
+
+The display client can send various types of images to the server: raw images, 
+\code{Video::Image}, OpenCV's \code{IplImage} and encoded images that are
+supported by Qt. An \code{IplImage} can only be sent if OpenCV was enabled
+during compilation.
+The clients interface for sending images is:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setImage(const std::string& id, int width, int height, int channels,
+     const std::vector<unsigned char>& data);
+   void setImage(const std::string& id, const Video::Image& image); 
+   void setImage(const std::string& id, const std::vector<unsigned char>& data,
+      const std::string &format=""); 
+   #ifdef FEAT_VISUALIZATION_OPENCV
+      void setImage(const std::string& id, const IplImage* pImage); 
+   #endif
+\end{Verbatim}
+
+Raw images can have 1 or 3 channels and the specified image size must match the
+size of \code{data}. The supported formats of the compressed image are: bmp,
+gif, jpeg, jpg, png, pbm, pgm, ppm, tiff, xbm, xpm. It the format is not
+specified, Qt will try to deduce the type from \code{data}. An \code{IplImage}
+is first converted to a contiguous array of bytes and sent to the server as a
+raw image.
+
+Every image has an id which uniquely identifies an image on the server. This way
+an image (as any other object) can be replaced on the server.

### 2010-06-10 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs updated
v11n: docs

--- 2010-06-10 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
@@
 
-\begin{itemize}
-   \item Add visualization.sa to SVN externals:
-   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
+To get the visualization from the svn, edit the SVN externals in the root
+of your CogX system directory:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    $ svn pe svn:externals .
 \end{Verbatim}
-   Add the line:
-   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
+
+Add the following line, save and do an svn update:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
    ^/code/subarchitectures/visualization.sa/trunk subarchitectures/visualization.sa
@@
 
-   % TODO: the script may be in another location
-   \item install the required libraries; there is a script in visualization.sa/config
-   that you can run:
-   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
-      sudo ubuntu-install-deps.sh
+The subarchitecture depends on some external libraries. If they are not
+installed, the system may still compile, but some of the features won't be
+available. The dependencies can be installed with the script in
+visualization.sa/bin:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   sudo visualization.sa/bin/ubuntu-install-deps.sh
 \end{Verbatim}
 
-   \item enable BUILD\_VISUALIZATION\_SA in cmake
-\end{itemize}
+The new subarchitecture must be added to the main file CMakeLists.txt of the 
+project. It should be added before any subarchitectures that use visualization
+so that the dependent libraries will build correctly. Add the following
+block before the first subarchitecture in CMakeLists.txt:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   option(BUILD_SA_VISUALIZATION "Build visualization libraries (Requires QT4)" NO)
+   if(BUILD_SA_VISUALIZATION)
+      set(VISUALIZATION_INCLUDE_DIRS
+         ${COGX_ROOT}/subarchitectures/visualization.sa/src/c++/autogen
+         ${COGX_ROOT}/subarchitectures/visualization.sa/src/c++/core
+         )
+      set(VISUALIZATION_LIBRARIES
+         DisplayClient VisualizationData
+         )
+      add_subdirectory(subarchitectures/visualization.sa)
+      add_definitions(-DFEAT_VISUALIZATION)
+   endif(BUILD_SA_VISUALIZATION)
+\end{Verbatim}
+
+Finally run \code{ccmake} and set \code{BUILD\_VISUALIZATION\_SA} to \code{ON}
+to enable the visualization subarchitecture.
 
@@
 When an object is displayed in another view, the UI element won't be available.
-\todo{TODO: user-defined views will be able to specify which elements to show}
+\todo{TODO: user-defined views will be able to specify which UI elements to show}
 
@@
 
+The objects can be geometrically transformed on the server when they are
+displayed. A client side interface is provided to set display transformations
+of objects and their parts.
+
 \subsection{Images}
@@
 during compilation.
-The clients interface for sending images is:
+The client-side interface for sending images is:
 
@@
 size of \code{data}. The supported formats of the compressed image are: bmp,
-gif, jpeg, jpg, png, pbm, pgm, ppm, tiff, xbm, xpm. It the format is not
+gif, jpeg, jpg, png, pbm, pgm, ppm, tiff, xbm, xpm. If the format is not
 specified, Qt will try to deduce the type from \code{data}. An \code{IplImage}
@@
 an image (as any other object) can be replaced on the server.
+
+Images can only be displayed in 2D contexts. Image objects can be transformed
+by defining a $3 \times 3$ transformation matrix that is applied before
+the image is drawn.
+
+\subsection{SVG objects}
+
+SVG objects are sent to the server as XML documents which are then interpreted
+by the Qt SVG subsystem. Every SVG object can have multiple parts where each
+part is defined in a separate XML document.
+\todo{TODO: the interface to define the order of drawing for objects and parts
+is missing}
+The client-side interface for defining SVG objects is:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setObject(const std::string& id, const std::string& partId,
+         const std::string& xmlData); 
+\end{Verbatim}
+
+The function is named \code{setObject} because initially it was planned that
+the server will support multiple types of objects. Currently the server
+supports only SVG and doesn't check if the SVG file is valid.
+
+SVG objects can only be displayed in 2D contexts. Every part of an object can
+be transformed by defining a $3 \times 3$ transformation matrix that is applied
+before the part is drawn.

### 2010-06-10 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 an image (as any other object) can be replaced on the server.
+
+Images can only be displayed in 2D contexts. Image objects can be transformed
+by defining a $3 \times 3$ transformation matrix that is applied before
+the image is drawn.
+
+\subsection{SVG objects}
+
+SVG objects are sent to the server as XML documents which are then interpreted
+by the Qt SVG subsystem. Every SVG object can have multiple parts where each
+part is defined in a separate XML document.
+\todo{TODO: the interface to define the order of drawing for objects and parts
+is missing}
+The client-side interface for defining SVG objects is:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setObject(const std::string& id, const std::string& partId,
+         const std::string& xmlData); 
+\end{Verbatim}
+
+The function is named \code{setObject} because initially it was planned that
+the server will support multiple types of objects. Currently the server
+supports only SVG and doesn't check if the SVG file is valid.
+
+SVG objects can only be displayed in 2D contexts. Every part of an object can
+be transformed by defining a $3 \times 3$ transformation matrix that is applied
+before the part is drawn.
+
+\subsection{OpenGL objects}
+
+OpenGL objects can be rendered in two ways. The more flexible option is the use
+of the GlLuaScript subsystem where normal OpenGL commands are written to a Lua
+script and sent to the server. The other option is to build TomGine models on
+the client, and send them to the server using boost serialization.
+
+\subsubsection{LuaGlScript objects}
+
+A LuaGlScript objects are multipart objects where each part of the object is a
+Lua script engine into which Lua scripts are loaded. The scripts are sent to
+the server with a display client function:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setLuaGlObject(const std::string& id, const std::string& partId,
+         const std::string& script);
+\end{Verbatim}
+
+Each script has a method called \code{render()} that is called every time the
+scene needs to be rendered.  This method then calls various OpenGL functions to
+render the scene. Most of the OpenGL functions should be available (textures
+are not supported at this time). An example of a script that renders a colored
+spiral could look like this (the tables \code{tabcos} and \code{tabsin} were
+defined in another part of the same script):
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   N=36
+   function render()
+      glBegin(GL_LINES)
+      for i=1,10000 do
+         glColor(i%23/23.0,i%17/17.0,i%11/11.0)
+         glVertex(tabcos[i%N+1], tabsin[i%N+1], -i/100.0)
+      end
+      glEnd()
+   end
+\end{Verbatim}
+
+A script is loaded into the Lua script engine every time the server receives it
+from a client. This way the appearance of the object can be changed during its
+lifetime. Complex objects can use display lists to make rendering simpler and 
+faster. Lists are created for each OpenGL context separately so if an object
+is visible in multiple contexts, each context will have a copy of every
+display list defined by the object.
+\todo{VERIFY: diasplay lists are local to every LuaGlScript part}
+The following commands are available for display list management:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   DispList:create(name)
+   DispList:replace(name)
+   DispList:endList()
+   DispList:delete(name)
+   DispList:exists(name)
+   DispList:draw(name)
+\end{Verbatim}
+
+In the following example we create a function that creates a display list named
+{\em myobject}. The function \code{initlists()} calls it when {\em myobject}
+doesn't exist in the current context. Finally \code{render()} intializes the
+lists before doing the rest of the rendering.
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   function make_the_thing()
+      DispList:create('myobject')
+        glBegin(GL_TRIANGLES)
+        glVertex(0.1, 0.1, 0.1)
+        glVertex(0.3, 0.1, 0.2)
+        glVertex(0.3, 0.4, 0.3)
+        glEnd()
+      DispList:endList()
+   end
+
+   function initlists()
+      if not DispList:exists('myobject') then
+         make_the_thing()
+      end
+   end
+
+   function render()
+      initlists()
+      DispList:draw('myobject')
+      glPushMatrix()
+         glTranslate(0.5,0.5,0.0)
+         DispList:draw('myobject')
+      glPopMatrix()
+   end
+\end{Verbatim}
+
+With this approach the lists are created only once in an OpenGL context and
+only when a context is active.  When the code is loaded into the Lua script
+engine, the script is locked and the rendering of the object is not possible.
+The OpenGL context might not even be available at the time or a wrong one might
+be active. That is why all OpenGL calls should be enclosed in functions that
+are accessible only from \code{render()}.
+
+It is also possible for a programmer to change only a part of the object by
+sending a new definition for \code{make\_the\_thing()}:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   function make_the_thing()
+      DispList:create('myobject')
+        glBegin(GL_QUADS)
+        glVertex(0.1, 0.1, 0.1)
+        glVertex(0.3, 0.1, 0.2)
+        glVertex(0.3, 0.4, 0.3)
+        glVertex(0.0, 0.5, 0.4)
+        glEnd()
+      DispList:endList()
+   end
+\end{Verbatim}
+
+There is one problem with the new \code{make\_the\_thing()}: if a display list
+for {\em myobject} already exists, the function will not be executed and the
+object will keep its current shape. This can be fixed if user closes and
+reopens the OpenGL view which will create a new OpenGL context. A better
+approach is to use a global variable with a list of display list names that
+have to be recreated. We change \code{initlists()} so that it verifies if any
+objects should be deleted.
+\todo{VERIFY: code from this point on is not verified}
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   renewal = {} 
+   function initlists()
+      if #renewal > 0 then
+         local ren = renewal; renewal = {}
+         for _,name in pairs(ren) do DispList:delete(name) end
+      end
+      if not DispList:exists('myobject') then make_the_thing() end
+   end
+\end{Verbatim}
+
+When we define a new version of \code{make\_the\_thing()} we add the name of
+the display list to be recreated ({\em myobject}) to the \code{renewal} table:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   function make_the_thing()
+      ...
+   end
+   table.insert(renewal, 'myobject')
+\end{Verbatim}

### 2010-06-10 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs, minor

@@
 
-\subsection{An Active Display Client}
+\section{Interaction with the display server}
 
@@
 
-Every image has an id which uniquely identifies an image on the server. This way
-an image (as any other object) can be replaced on the server.
+Every image has an id which uniquely identifies the image on the server. This
+way the image (as any other object) can be replaced on the server.
 
@@
 before the part is drawn.
+
+\subsection{OpenGL objects}
+
+OpenGL objects can be rendered in two ways. The more flexible option is the use
+of the GlLuaScript subsystem where normal OpenGL commands are written to a Lua
+script and sent to the server. The other option is to build TomGine models on
+the client, and send them to the server using boost serialization.
+
+\subsubsection{LuaGlScript objects}
+
+A LuaGlScript objects are multipart objects where each part of the object is a
+Lua script engine into which Lua scripts are loaded. The scripts are sent to
+the server with a display client function:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setLuaGlObject(const std::string& id, const std::string& partId,
+         const std::string& script);
+\end{Verbatim}
+
+Each script has a method called \code{render()} that is called every time the
+scene needs to be rendered.  This method then calls various OpenGL functions to
+render the scene. Most of the OpenGL functions should be available (textures
+are not supported at this time). An example of a script that renders a colored
+spiral could look like this (the tables \code{tabcos} and \code{tabsin} were
+defined in another part of the same script):
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   N=36
+   function render()
+      glBegin(GL_LINES)
+      for i=1,10000 do
+         glColor(i%23/23.0,i%17/17.0,i%11/11.0)
+         glVertex(tabcos[i%N+1], tabsin[i%N+1], -i/100.0)
+      end
+      glEnd()
+   end
+\end{Verbatim}
+
+A script is loaded into the Lua script engine every time the server receives it
+from a client. This way the appearance of the object can be changed during its
+lifetime. Complex objects can use display lists to make rendering simpler and 
+faster. Lists are created for each OpenGL context separately so if an object
+is visible in multiple contexts, each context will have a copy of every
+display list defined by the object.
+\todo{VERIFY: diasplay lists are local to every LuaGlScript part}
+The following commands are available for display list management:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   DispList:create(name)
+   DispList:replace(name)
+   DispList:endList()
+   DispList:delete(name)
+   DispList:exists(name)
+   DispList:draw(name)
+\end{Verbatim}
+
+In the following example we create a function that creates a display list named
+{\em myobject}. The function \code{initlists()} calls it when {\em myobject}
+doesn't exist in the current context. Finally \code{render()} intializes the
+lists before doing the rest of the rendering.
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   function make_the_thing()
+      DispList:create('myobject')
+        glBegin(GL_TRIANGLES)
+        glVertex(0.1, 0.1, 0.1)
+        glVertex(0.3, 0.1, 0.2)
+        glVertex(0.3, 0.4, 0.3)
+        glEnd()
+      DispList:endList()
+   end
+
+   function initlists()
+      if not DispList:exists('myobject') then
+         make_the_thing()
+      end
+   end
+
+   function render()
+      initlists()
+      DispList:draw('myobject')
+      glPushMatrix()
+         glTranslate(0.5,0.5,0.0)
+         DispList:draw('myobject')
+      glPopMatrix()
+   end
+\end{Verbatim}
+
+With this approach the lists are created only once in an OpenGL context and
+only when a context is active.  When the code is loaded into the Lua script
+engine, the script is locked and the rendering of the object is not possible.
+The OpenGL context might not even be available at the time or a wrong one might
+be active. That is why all OpenGL calls should be enclosed in functions that
+are accessible only from \code{render()}.
+
+It is also possible for a programmer to change only a part of the object by
+sending a new definition for \code{make\_the\_thing()}:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   function make_the_thing()
+      DispList:create('myobject')
+        glBegin(GL_QUADS)
+        glVertex(0.1, 0.1, 0.1)
+        glVertex(0.3, 0.1, 0.2)
+        glVertex(0.3, 0.4, 0.3)
+        glVertex(0.0, 0.5, 0.4)
+        glEnd()
+      DispList:endList()
+   end
+\end{Verbatim}
+
+There is one problem with the new \code{make\_the\_thing()}: if a display list
+for {\em myobject} already exists, the function will not be executed and the
+object will keep its current shape. This can be fixed if user closes and
+reopens the OpenGL view which will create a new OpenGL context. A better
+approach is to use a global variable with a list of display list names that
+have to be recreated. We change \code{initlists()} so that it verifies if any
+objects should be deleted.
+\todo{VERIFY: code from this point on is not verified}
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   renewal = {} 
+   function initlists()
+      if #renewal > 0 then
+         local ren = renewal; renewal = {}
+         for _,name in pairs(ren) do DispList:delete(name) end
+      end
+      if not DispList:exists('myobject') then make_the_thing() end
+   end
+\end{Verbatim}
+
+When we define a new version of \code{make\_the\_thing()} we add the name of
+the display list to be recreated ({\em myobject}) to the \code{renewal} table:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   function make_the_thing()
+      ...
+   end
+   table.insert(renewal, 'myobject')
+\end{Verbatim}
+
+\subsection{TODO: Transformations}

### 2010-06-11 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs updated

@@
 display images, but with time other objects (supported by Qt) were added:
-SVG images and OpenGL scenes.
+SVG images, HTML documents and OpenGL scenes.
 
@@
 before the part is drawn.
+
+\subsection{HTML documents}
+
+The display server renders HTML documents with Qt-WebKit. It provides the
+sturcture of the HTML document while display clients provide chunks of
+HTML code that goes into the body of the document.
+\todo{TODO: Add support to modify head of document so that CSS can be used}
+The client-side interface for defining HTML document chunks is:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setHtml(const std::string& id, const std::string& partId,
+         const std::string& htmlData); 
+\end{Verbatim}
 
@@
 \end{Verbatim}
-
-\subsection{TODO: Transformations}

### 2010-06-11 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 before the part is drawn.
+
+\subsection{HTML documents}
+
+The display server renders HTML documents with Qt-WebKit. It provides the
+sturcture of the HTML document while display clients provide chunks of
+HTML code that goes into the head or the body of the document.
+The client-side interface for defining HTML document chunks is:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setHtml(const std::string& id, const std::string& partId,
+         const std::string& htmlData); 
+   void setHtmlHead(const std::string& id, const std::string& partId,
+         const std::string& htmlData); 
+\end{Verbatim}
 

### 2010-06-11 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 The display server creates rendering contexts in which the views are displayed.
-There are two types of contexts: 2D and OpenGL.
+There are three types of contexts: 2D, OpenGL and HTML.
 % (TODO: new contexts: tabular; textual; ...)
@@
 
-Currently the following types of objects are supported: images, SVG graphics and
-OpenGL objects define with LuaGl script or as TomGine render-models.
+Currently the following types of objects are supported: images, HTML documents,
+SVG graphics and OpenGL objects defined with LuaGl script or as TomGine
+render-models.
 

### 2010-06-11 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

--- 2010-06-10 subarchitectures/visualization.sa/trunk/doc/v11n.tex
@@
 before the part is drawn.
+
+\subsection{HTML documents}
+
+The display server renders HTML documents with Qt-WebKit. It provides the
+sturcture of the HTML document while display clients provide chunks of
+HTML code that goes into the head or the body of the document.
+The client-side interface for defining HTML document chunks is:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void setHtml(const std::string& id, const std::string& partId,
+         const std::string& htmlData); 
+   void setHtmlHead(const std::string& id, const std::string& partId,
+         const std::string& htmlData); 
+\end{Verbatim}
+
+Chunks for the head and the body part of a document are independent. This means
+that a chunk in the head may have the same partId as a chunk in the body of
+a document. When a part is removed from a document, the chunks for the head
+and the body with the requested partId are removed.
+\todo{TODO: ICE interface for removing (parts of) objects is still missing}
 

### 2010-06-14 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 
+Chunks for the head and the body part of a document are independent. This means
+that a chunk in the head may have the same partId as a chunk in the body of
+a document. When a part is removed from a document, the chunks for the head
+and the body with the requested partId are removed.
+\todo{TODO: ICE interface for removing (parts of) objects is still missing}
+
 \subsection{OpenGL objects}
@@
 \begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   DispList:create(name)
-   DispList:replace(name)
+   DispList:newList(name)
    DispList:endList()
@@
    DispList:draw(name)
+   DispList:setDirty(name)
+   DispList:getDirty(name_table)
 \end{Verbatim}
@@
    function make_the_thing()
-      DispList:create('myobject')
+      DispList:newList('myobject')
         glBegin(GL_TRIANGLES)
@@
    function initlists()
-      if not DispList:exists('myobject') then
+      local dirty = DispList:getDirty( {'myobject'} )
+      if dirty['myobject'] then
          make_the_thing()
@@
    end
+   DispList:setDirty('myobject')
 \end{Verbatim}
 
-There is one problem with the new \code{make\_the\_thing()}: if a display list
-for {\em myobject} already exists, the function will not be executed and the
-object will keep its current shape. This can be fixed if user closes and
-reopens the OpenGL view which will create a new OpenGL context. A better
-approach is to use a global variable with a list of display list names that
-have to be recreated. We change \code{initlists()} so that it verifies if any
-objects should be deleted.
-\todo{VERIFY: code from this point on is not verified}
+The function \code{setDirty()} tells the display list manager that the object
+was modified. The renderer will detect the change the next time
+\code{getDirty()} is called, in this case from \code{initlists()}.
+\code{getDirty(namelist)} returns a table with keys that represent the names of
+the objects from {\em namelist} that need to be recreated. It will contain the
+names of the objects that were not yet created and the objects that were marked
+as dirty. The list of dirty objects is reset after every call to
+\code{getDirty()}.
+
+The module \code{StdModel} defines the following functions that create simple shapes:
 
 \begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   renewal = {} 
-   function initlists()
-      if #renewal > 0 then
-         local ren = renewal; renewal = {}
-         for _,name in pairs(ren) do DispList:delete(name) end
-      end
-      if not DispList:exists('myobject') then make_the_thing() end
-   end
+   StdModel:box(sx, sy, sz)
+   StdModel:cylinder(rx, ry, np)
 \end{Verbatim}
 
-When we define a new version of \code{make\_the\_thing()} we add the name of
-the display list to be recreated ({\em myobject}) to the \code{renewal} table:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   function make_the_thing()
-      ...
-   end
-   table.insert(renewal, 'myobject')
-\end{Verbatim}
+The box and the cylinder are created with the center in $(0, 0, 0)$. The
+parameters for the box define the size of the box. The cylinder has an elliptic
+shape with the given radii and is approximated with {\em np} sides.

### 2010-06-14 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: cmakelists - new options

@@
 \begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   option(BUILD_SA_VISUALIZATION "Build visualization libraries (Requires QT4)" NO)
-   if(BUILD_SA_VISUALIZATION)
+   option(BUILD_SA_V11N "Build visualization libraries (Requires QT4)" NO)
+   if(BUILD_SA_V11N)
       set(VISUALIZATION_INCLUDE_DIRS
@@
       add_definitions(-DFEAT_VISUALIZATION)
-   endif(BUILD_SA_VISUALIZATION)
+   endif(BUILD_SA_V11N)
 \end{Verbatim}
 
-Finally run \code{ccmake} and set \code{BUILD\_VISUALIZATION\_SA} to \code{ON}
-to enable the visualization subarchitecture.
+Finally run \code{ccmake} and set \code{BUILD\_SA\_V11N} to \code{ON} to enable
+the visualization subarchitecture.
 
@@
 as the object.
-\todo{(TODO: Additional views that display multiple objects can be defined on the server)}
+\todo{(TODO: It will be posible to define additional views that display multiple objects)}
 The display server creates rendering contexts in which the views are displayed.
 There are three types of contexts: 2D, OpenGL and HTML.
-% (TODO: new contexts: tabular; textual; ...)
+% (TODO: additional contexts: eg. tabular)
 

### 2010-06-14 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: filemonitor, new converters: html, htmlhead

@@
 
-The files that the monitor can process are images, SVG files and files created
-for GraphWiz.
+The files that the monitor can process are images, SVG files, dot files created
+for GraphWiz, LuaGl scripts, HTML chunks.
 
@@
 \end{Verbatim}
+
+The converters {\em luagl}, {\em html} and {\em htmlhead} send the unprocessed
+contents of the monitored file to the server.
 

### 2010-06-14 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 
-The library provides a client component that interacts with the server. The
-client is devised in such a way that its use is as trivial as the use of
-OpenCV's function cvShowImage(). For simple objects the server can also
-be used without a client if the developer is willing to use the file system
-as a communication channel. This involves a CAST component that monitors
-the file system and sends the newly created files to the server.
+The C++ client library provides a client interface that enables the interaction
+with the server. It lets the programmer send objects to be dispayed to the
+server and receive notifications about the events emited by the user interface
+that was built on the server by the client component.
+\todo{TODO: add interfaces for Java and Python}
 
-First we will show how to use the file system monitor and then we will give
-instructions for the use of the real client.
+For simple objects the server can be used without linking to the display client
+because the file system as a communication channel. To do this a CAST
+component FileMonitor has to be added to the cast file and configured to
+monitor certain files that are created by other componetns.
 

### 2010-06-18 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: html plugins work if nspluginwrapper is uninstalled; disabled -- a nasty bug needs more work

@@
 
-The new subarchitecture must be added to the main file CMakeLists.txt of the 
+The new subarchitecture must be added to the main CMakeLists.txt file of the 
 project. It should be added before any subarchitectures that use visualization
@@
 \begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   option(BUILD_SA_V11N "Build visualization libraries (Requires QT4)" NO)
-   if(BUILD_SA_V11N)
-      set(VISUALIZATION_INCLUDE_DIRS
-         ${COGX_ROOT}/subarchitectures/visualization.sa/src/c++/autogen
-         ${COGX_ROOT}/subarchitectures/visualization.sa/src/c++/core
-         )
-      set(VISUALIZATION_LIBRARIES
-         DisplayClient VisualizationData
-         )
-      add_subdirectory(subarchitectures/visualization.sa)
-      add_definitions(-DFEAT_VISUALIZATION)
-   endif(BUILD_SA_V11N)
+   include(subarchitectures/visualization.sa/cmake/v11n.cmake)
 \end{Verbatim}

### 2010-06-21 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 elements with which it can be configured. The client receives notifications and
-requests related to these elements from the display server. The component must
-define two callback functions that will handle the events sent by the server.
+requests related to these elements from the display server. To achieve this 
+a few methods must be overridden in a class derived from the CDisplayClient:
 
@@
    #ifdef FEAT_VISUALIZATION
+     class CVvDisplayClient: public cogx::display::CDisplayClient
+     {
+       VideoViewer* pViewer;
+     public:
+       CVvDisplayClient() { pViewer = NULL; }
+       void setClientData(VideoViewer* pVideoViewer) { pViewer = pVideoViewer; }
+       void handleEvent(const Visualization::TEvent &event); /*override*/
+       std::string getControlState(const std::string& ctrlId); /*override*/
+     };
      cogx::display::CActiveDisplayClient<VideoViewer> m_display;
-     void handleGuiEvent(const Visualization::TEvent &event);
-     std::string getControlState(const std::string& ctrlId);
    #endif
@@
 
-The event path from the server to the client must be initiated. This is done right
-after the component connects to the client. After the connections are established,
-the callback functions are registered:
+In the constructor of the VideoViewer we call \code{setClientData()} so that
+the new display client will be able to access the VideoViewer:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+  VideoViewer() : camId(0)
+  {
+#ifdef FEAT_VISUALIZATION
+    m_display.setClientData(this);
+#endif
+  }
+\end{Verbatim}
+
+The event path from the server to the client must be initialized. This is done
+right after the component connects to the client:
 
@@
      m_display.installEventReceiver();
-     m_display.setEventCallback(this, &VideoViewer::handleGuiEvent);
-     m_display.setStateQueryCallback(this, &VideoViewer::getControlState);
    #endif
@@
 
-The \code{handleGuiEvent()} callback function will be called every time the
-UI element emits a known event. For the chekcbox element we need to respond
-to the \code{evCheckBoxChange} event:
+When the UI element is first displayed it has to obtain the current state from
+the component that created it. This is done with the method
+\code{getControlState()}:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   std::string VideoViewer::CVvDisplayClient::getControlState(
+         const std::string& ctrlId)
+   {
+     if (ctrlId == "toggle.viewer.running") {
+       return pViewer->receiving ? "1" : "0";
+     }
+     return "";
+   }
+\end{Verbatim}
+
+The method \code{handleGuiEvent()} will be called every time the UI element
+emits a known event. For the chekcbox element we need to respond to the
+\code{evCheckBoxChange} event:
 
@@
    #ifdef FEAT_VISUALIZATION
-   void VideoViewer::handleGuiEvent(const Visualization::TEvent &event)
+   void VideoViewer::CVvDisplayClient::handleGuiEvent(
+         const Visualization::TEvent &event)
    {

### 2010-07-07 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: minor doc update

@@
 \begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   #ifdef FEAT_VISUALIZATION
+   #include <CDisplayClient.hpp>
+   #endif
+   ...
    class VideoViewer : public ManagedComponent, public VideoClient
@@
    #ifdef FEAT_VISUALIZATION
-     cogx::display::CActiveDisplayClient<VideoViewer> m_display;
-     void handleGuiEvent(const Visualization::TEvent &event);
-     std::string getControlState(const std::string& ctrlId);
+     class CVvDisplayClient: public cogx::display::CDisplayClient
+     {
+       VideoViewer* pViewer;
+     public:
+       CVvDisplayClient() { pViewer = NULL; }
+       void setClientData(VideoViewer* pVideoViewer) { pViewer = pVideoViewer; }
+       void handleEvent(const Visualization::TEvent &event); /*override*/
+       std::string getControlState(const std::string& ctrlId); /*override*/
+     };
+     CVvDisplayClient m_display;
    #endif

### 2010-07-07 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: manual updated

@@
 
-\subsection{A Passive Display Client}
+\subsection{A Passive Display Client}\label{section:PassiveClient}
 
@@
 is displaying only one object. Any other string can be used instead.
-
-\section{Interaction with the display server}
-
-An active display client is needed when the component wants to present some UI
-elements with which it can be configured. The client receives notifications and
-requests related to these elements from the display server. To achieve this 
-a few methods must be overridden in a class derived from the CDisplayClient:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   class VideoViewer : public ManagedComponent, public VideoClient
-   {
-   private:
-   #ifdef FEAT_VISUALIZATION
-     class CVvDisplayClient: public cogx::display::CDisplayClient
-     {
-       VideoViewer* pViewer;
-     public:
-       CVvDisplayClient() { pViewer = NULL; }
-       void setClientData(VideoViewer* pVideoViewer) { pViewer = pVideoViewer; }
-       void handleEvent(const Visualization::TEvent &event); /*override*/
-       std::string getControlState(const std::string& ctrlId); /*override*/
-     };
-     cogx::display::CActiveDisplayClient<VideoViewer> m_display;
-   #endif
-   };
-\end{Verbatim}
-
-In the constructor of the VideoViewer we call \code{setClientData()} so that
-the new display client will be able to access the VideoViewer:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-  VideoViewer() : camId(0)
-  {
-#ifdef FEAT_VISUALIZATION
-    m_display.setClientData(this);
-#endif
-  }
-\end{Verbatim}
-
-The event path from the server to the client must be initialized. This is done
-right after the component connects to the client:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   void VideoViewer::start()
-   {
-   #ifdef FEAT_VISUALIZATION
-     m_display.connectIceClient(*this);
-     m_display.installEventReceiver();
-   #endif
-   }
-\end{Verbatim}
-
-The flow of images from the VideoServer to the VideoViewer can be stopped at
-any time. To let the user control the flow the component has to add a UI
-element to the display server.  An appropriate type of UI element for this
-purpose is a chekcbox:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   void VideoViewer::start()
-   {
-   #ifdef FEAT_VISUALIZATION
-     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
-   #endif
-   }
-\end{Verbatim}
-
-The first parameter is the name of the {\em view} that will display the UI
-element when it is active, not the name of object. This means that the UI
-element will be displayed only when an object is displayed in its default view.
-When an object is displayed in another view, the UI element won't be available.
-\todo{TODO: user-defined views will be able to specify which UI elements to show}
-
-When the UI element is first displayed it has to obtain the current state from
-the component that created it. This is done with the method
-\code{getControlState()}:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   std::string VideoViewer::CVvDisplayClient::getControlState(
-         const std::string& ctrlId)
-   {
-     if (ctrlId == "toggle.viewer.running") {
-       return pViewer->receiving ? "1" : "0";
-     }
-     return "";
-   }
-\end{Verbatim}
-
-The method \code{handleGuiEvent()} will be called every time the UI element
-emits a known event. For the chekcbox element we need to respond to the
-\code{evCheckBoxChange} event:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   #ifdef FEAT_VISUALIZATION
-   void VideoViewer::CVvDisplayClient::handleGuiEvent(
-         const Visualization::TEvent &event)
-   {
-     if (event.type == Visualization::evCheckBoxChange) {
-       if (event.sourceId == "toggle.viewer.running") {
-         bool newrcv = (event.data != "0");
-         if (newrcv != receiving) {
-           if(receiving) {
-             videoServer->stopReceiveImages(getComponentID());
-           }
-           else {
-             videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
-           }
-           receiving = !receiving;
-         }
-       }
-     }
-   }
-   #endif
-\end{Verbatim}
-
-Note that the data for UI elements passed between the client and the server is
-serialized into strings. The checkbox element knows only 2 values: \code{"0"}
-and everything else.
-\todo{TODO: Add more element types, define values}
-
-A UI element may be displayed in multiple views. When a UI element is first
-displayed it must be put in a state that reflects the current state of the
-component.  Since a UI element may be created at any time it wouldn't be enough
-to set the initial state of the element at creation time. Instead, we define
-the \code{getControlState()} callback function:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   #ifdef FEAT_VISUALIZATION
-   std::string VideoViewer::getControlState(const std::string& ctrlId)
-   {
-     if (ctrlId == "toggle.viewer.running") {
-       return receiving ? "2" : "0";
-     }
-     return "";
-   }
-   #endif
-\end{Verbatim}
-
-When the function returns \code{""} the display server won't change the state
-of the UI element.
 
@@
          const std::string& htmlData); 
+   void setHtmlForm(const std::string& id, const std::string& partId,
+         const std::string& htmlData); 
 \end{Verbatim}
@@
 \todo{TODO: ICE interface for removing (parts of) objects is still missing}
+
+For a more detailed description of HTML form chuks see section
+\ref{section:HtmlForms}.
 
@@
 shape with the given radii and is approximated with {\em np} sides.
+
+\section{Interaction with the Display Server}
+
+In section \ref{section:PassiveClient} we described a simple client that is
+able to send data to the Display Server. The server also supports some simple
+user interface interaction with remote components. A client component that
+wants to interact with the Display Server and receive notifications about UI
+events can do it through an active display client. We derive a class from
+\code{"CDisplayClient"} and override a few methods:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   class VideoViewer : public ManagedComponent, public VideoClient
+   {
+   private:
+   #ifdef FEAT_VISUALIZATION
+     class CVvDisplayClient: public cogx::display::CDisplayClient
+     {
+       VideoViewer* pViewer;
+     public:
+       CVvDisplayClient() { pViewer = NULL; }
+       void setClientData(VideoViewer* pVideoViewer) { pViewer = pVideoViewer; }
+       void handleEvent(const Visualization::TEvent &event); /*override*/
+       std::string getControlState(const std::string& ctrlId); /*override*/
+     };
+     CVvDisplayClient m_display;
+   #endif
+   };
+\end{Verbatim}
+
+In the constructor of the VideoViewer we call \code{setClientData()} so that
+the new display client will be able to access the VideoViewer:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+  VideoViewer() : camId(0)
+  {
+#ifdef FEAT_VISUALIZATION
+    m_display.setClientData(this);
+#endif
+  }
+\end{Verbatim}
+
+The event path from the server to the client must be initialized. This is done
+right after the component connects to the client:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void VideoViewer::start()
+   {
+   #ifdef FEAT_VISUALIZATION
+     m_display.connectIceClient(*this);
+     m_display.installEventReceiver();
+   #endif
+   }
+\end{Verbatim}
+
+The flow of images from the VideoServer to the VideoViewer can be stopped at
+any time. To let the user control the flow the component has to add a UI
+element to the display server.  An appropriate type of UI element for this
+purpose is a chekcbox:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void VideoViewer::start()
+   {
+   #ifdef FEAT_VISUALIZATION
+     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
+   #endif
+   }
+\end{Verbatim}
+
+The first parameter is the name of the {\em view} that will display the UI
+element when it is active, not the name of object. This means that the UI
+element will be displayed only when an object is displayed in its default view.
+When an object is displayed in another view, the UI element won't be available.
+\todo{TODO: user-defined views will be able to specify which UI elements to show}
+
+When the UI element is first displayed it has to obtain the current state from
+the component that created it. This is done with the method
+\code{getControlState()}:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   std::string VideoViewer::CVvDisplayClient::getControlState(
+         const std::string& ctrlId)
+   {
+     if (ctrlId == "toggle.viewer.running") {
+       return pViewer->receiving ? "1" : "0";
+     }
+     return "";
+   }
+\end{Verbatim}
+
+The method \code{handleGuiEvent()} will be called every time the UI element
+emits a known event. For the chekcbox element we need to respond to the
+\code{evCheckBoxChange} event:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   #ifdef FEAT_VISUALIZATION
+   void VideoViewer::CVvDisplayClient::handleGuiEvent(
+         const Visualization::TEvent &event)
+   {
+     if (event.type == Visualization::evCheckBoxChange) {
+       if (event.sourceId == "toggle.viewer.running") {
+         bool newrcv = (event.data != "0");
+         if (newrcv != pViewer->receiving) {
+           if(pViewer->receiving) {
+             pViewer->videoServer->stopReceiveImages(getComponentID());
+           }
+           else {
+             pViewer->videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
+           }
+           pViewer->receiving = !pViewer->receiving;
+         }
+       }
+     }
+   }
+   #endif
+\end{Verbatim}
+
+Note that the data for UI elements passed between the client and the server is
+serialized into strings. The checkbox element knows only 2 values: \code{"0"}
+and everything else.
+\todo{TODO: Add more element types, define values}
+
+A UI element may be displayed in multiple views. When a UI element is first
+displayed it must be put in a state that reflects the current state of the
+component.  Since a UI element may be created at any time it wouldn't be enough
+to set the initial state of the element at creation time. Instead, we define
+the \code{getControlState()} callback function:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   #ifdef FEAT_VISUALIZATION
+   std::string VideoViewer::CVvDisplayClient::getControlState(const std::string& ctrlId)
+   {
+     if (ctrlId == "toggle.viewer.running") {
+       return pViewer->receiving ? "2" : "0";
+     }
+     return "";
+   }
+   #endif
+\end{Verbatim}
+
+When the function returns \code{""} the display server won't change the state
+of the UI element.
+
+\subsection{Support for HTML forms}\label{section:HtmlForms}
+
+The display server can present HTML forms and send data from a form to a client
+when the form is submitted.
+
+To add support for form processing in an active client we override the methods
+\code{"handleForm"} and \code{"getFormData"}. We also add a convenience
+function that will create the form and send it to the server:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+#ifdef FEAT_VISUALIZATION
+   class COrDisplayClient: public cogx::display::CDisplayClient
+   {
+      CObjectRecognizer* pRecognizer;
+   public:
+      void createForms();
+      void handleForm(const std::string& id, const std::string& partId,
+            const std::map<std::string, std::string>& fields); /*override*/
+      bool getFormData(const std::string& id, const std::string& partId,
+            std::map<std::string, std::string>& fields); /*override*/
+   };
+\end{Verbatim}
+
+A form can be defined with standard HTML form elements but it must not be
+enclosed in the \code{"<form>"} tag which is added by the Display Server.
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void CObjectRecognizer::COrDisplayClient::createForms()
+   {
+      std::ostringstream ss;
+      ss << "Min score: <select name='minScore'>";
+      for (float f = 0.7; f <= 1.0; f += 0.01) {
+         ss << "<option>" << sfloat(f, 2) << "</option>";
+      }
+      ss << "</select>";
+      ss << "<input type=\"submit\" name=\"submit\" value=\"Apply\"/>";
+      setHtmlForm("Settings", "ObjectRecognizer", ss.str());
+   }
+   std::string sfloat(double f, int precision)
+   {
+      std::ostringstream out;
+      out << std::fixed << std::setprecision(precision) << f;
+      return out.str();
+   }
+\end{Verbatim}
+
+The getter and the setter functions pass the form parameters in a dictionary.
+If a HTML field supports multiple values (like a multivalue select field, a
+group of check boxes or radio buttons), the values are stored as a delimited
+string where the delimiter is the newline character \Verb|"\n"|.
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void CObjectRecognizer::COrDisplayClient::handleForm(
+         const std::string& id, const std::string& partId,
+         const std::map<std::string, std::string>& fields)
+   {
+      if (! pRecognizer) return;
+      if (id != "Settings" || partId != "ObjectRecognizer") retrun;
+      float f;
+      typeof(fields.begin()) it;
+      it = fields.find("minScore");
+      if (it != fields.end()) {
+         f = atof(it->second.c_str());
+         if (f < 0.7) f = 0.7;
+         if (f > 1.0) f = 1.0;
+         pRecognizer->m_minScore = f;
+      }
+   }
+
+   bool CObjectRecognizer::COrDisplayClient::getFormData(
+         const std::string& id, const std::string& partId,
+         std::map<std::string, std::string>& fields)
+   {
+      if (! pRecognizer) return false;
+      if (id != "Settings" || partId != "ObjectRecognizer") retrun false;
+      fields["minScore"] = sfloat(pRecognizer->m_minScore, 2);
+      fields["maxAmbiguity"] = sfloat(pRecognizer->m_maxAmbiguity, 2);
+      return true;
+   }
+\end{Verbatim}
+
+A form can be shown in a separate view or as a part of an HTML page composed of
+multiple chunks. When a page contains a form it is slower to display because
+some javascript has to be evaluated every time the page is reloaded. Therfore
+the forms should not be a part of a page that is changed frequently (eg. more
+than once per second).

### 2010-07-09 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: manual

@@
 
-\section{Interaction with the display server}
-
-An active display client is needed when the component wants to present some UI
-elements with which it can be configured. The client receives notifications and
-requests related to these elements from the display server. To achieve this 
-a few methods must be overridden in a class derived from the CDisplayClient:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   class VideoViewer : public ManagedComponent, public VideoClient
-   {
-   private:
-   #ifdef FEAT_VISUALIZATION
-     class CVvDisplayClient: public cogx::display::CDisplayClient
-     {
-       VideoViewer* pViewer;
-     public:
-       CVvDisplayClient() { pViewer = NULL; }
-       void setClientData(VideoViewer* pVideoViewer) { pViewer = pVideoViewer; }
-       void handleEvent(const Visualization::TEvent &event); /*override*/
-       std::string getControlState(const std::string& ctrlId); /*override*/
-     };
-     CVvDisplayClient m_display;
-   #endif
-   };
-\end{Verbatim}
-
-In the constructor of the VideoViewer we call \code{setClientData()} so that
-the new display client will be able to access the VideoViewer:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-  VideoViewer() : camId(0)
-  {
-#ifdef FEAT_VISUALIZATION
-    m_display.setClientData(this);
-#endif
-  }
-\end{Verbatim}
-
-The event path from the server to the client must be initialized. This is done
-right after the component connects to the client:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   void VideoViewer::start()
-   {
-   #ifdef FEAT_VISUALIZATION
-     m_display.connectIceClient(*this);
-     m_display.installEventReceiver();
-   #endif
-   }
-\end{Verbatim}
-
-The flow of images from the VideoServer to the VideoViewer can be stopped at
-any time. To let the user control the flow the component has to add a UI
-element to the display server.  An appropriate type of UI element for this
-purpose is a chekcbox:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   void VideoViewer::start()
-   {
-   #ifdef FEAT_VISUALIZATION
-     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
-   #endif
-   }
-\end{Verbatim}
-
-The first parameter is the name of the {\em view} that will display the UI
-element when it is active, not the name of object. This means that the UI
-element will be displayed only when an object is displayed in its default view.
-When an object is displayed in another view, the UI element won't be available.
-\todo{TODO: user-defined views will be able to specify which UI elements to show}
-
-When the UI element is first displayed it has to obtain the current state from
-the component that created it. This is done with the method
-\code{getControlState()}:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   std::string VideoViewer::CVvDisplayClient::getControlState(
-         const std::string& ctrlId)
-   {
-     if (ctrlId == "toggle.viewer.running") {
-       return pViewer->receiving ? "1" : "0";
-     }
-     return "";
-   }
-\end{Verbatim}
-
-The method \code{handleGuiEvent()} will be called every time the UI element
-emits a known event. For the chekcbox element we need to respond to the
-\code{evCheckBoxChange} event:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   #ifdef FEAT_VISUALIZATION
-   void VideoViewer::CVvDisplayClient::handleGuiEvent(
-         const Visualization::TEvent &event)
-   {
-     if (event.type == Visualization::evCheckBoxChange) {
-       if (event.sourceId == "toggle.viewer.running") {
-         bool newrcv = (event.data != "0");
-         if (newrcv != receiving) {
-           if(receiving) {
-             videoServer->stopReceiveImages(getComponentID());
-           }
-           else {
-             videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
-           }
-           receiving = !receiving;
-         }
-       }
-     }
-   }
-   #endif
-\end{Verbatim}
-
-Note that the data for UI elements passed between the client and the server is
-serialized into strings. The checkbox element knows only 2 values: \code{"0"}
-and everything else.
-\todo{TODO: Add more element types, define values}
-
-A UI element may be displayed in multiple views. When a UI element is first
-displayed it must be put in a state that reflects the current state of the
-component.  Since a UI element may be created at any time it wouldn't be enough
-to set the initial state of the element at creation time. Instead, we define
-the \code{getControlState()} callback function:
-
-\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
-   #ifdef FEAT_VISUALIZATION
-   std::string VideoViewer::getControlState(const std::string& ctrlId)
-   {
-     if (ctrlId == "toggle.viewer.running") {
-       return receiving ? "2" : "0";
-     }
-     return "";
-   }
-   #endif
-\end{Verbatim}
-
-When the function returns \code{""} the display server won't change the state
-of the UI element.
-
 \section{Displaying objects}
@@
 shape with the given radii and is approximated with {\em np} sides.
+
+\section{Interaction with the Display Server}
+
+In section \ref{section:PassiveClient} we described a simple client that is
+able to send data to the Display Server. The server also supports some simple
+user interface interaction with remote components. A client component that
+wants to interact with the Display Server and receive notifications about UI
+events can do it through an active display client. We derive a class from
+\code{"CDisplayClient"} and override a few methods:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   class VideoViewer : public ManagedComponent, public VideoClient
+   {
+   private:
+   #ifdef FEAT_VISUALIZATION
+     class CVvDisplayClient: public cogx::display::CDisplayClient
+     {
+       VideoViewer* pViewer;
+     public:
+       CVvDisplayClient() { pViewer = NULL; }
+       void setClientData(VideoViewer* pVideoViewer) { pViewer = pVideoViewer; }
+       void handleEvent(const Visualization::TEvent &event); /*override*/
+       std::string getControlState(const std::string& ctrlId); /*override*/
+     };
+     CVvDisplayClient m_display;
+   #endif
+   };
+\end{Verbatim}
+
+In the constructor of the VideoViewer we call \code{setClientData()} so that
+the new display client will be able to access the VideoViewer:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+  VideoViewer() : camId(0)
+  {
+#ifdef FEAT_VISUALIZATION
+    m_display.setClientData(this);
+#endif
+  }
+\end{Verbatim}
+
+The event path from the server to the client must be initialized. This is done
+right after the component connects to the client:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void VideoViewer::start()
+   {
+   #ifdef FEAT_VISUALIZATION
+     m_display.connectIceClient(*this);
+     m_display.installEventReceiver();
+   #endif
+   }
+\end{Verbatim}
+
+The flow of images from the VideoServer to the VideoViewer can be stopped at
+any time. To let the user control the flow the component has to add a UI
+element to the display server.  An appropriate type of UI element for this
+purpose is a chekcbox:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void VideoViewer::start()
+   {
+   #ifdef FEAT_VISUALIZATION
+     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
+   #endif
+   }
+\end{Verbatim}
+
+The first parameter is the name of the {\em view} that will display the UI
+element when it is active, not the name of object. This means that the UI
+element will be displayed only when an object is displayed in its default view.
+When an object is displayed in another view, the UI element won't be available.
+\todo{TODO: user-defined views will be able to specify which UI elements to show}
+
+When the UI element is first displayed it has to obtain the current state from
+the component that created it. This is done with the method
+\code{getControlState()}:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   std::string VideoViewer::CVvDisplayClient::getControlState(
+         const std::string& ctrlId)
+   {
+     if (ctrlId == "toggle.viewer.running") {
+       return pViewer->receiving ? "1" : "0";
+     }
+     return "";
+   }
+\end{Verbatim}
+
+The method \code{handleGuiEvent()} will be called every time the UI element
+emits a known event. For the chekcbox element we need to respond to the
+\code{evCheckBoxChange} event:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   #ifdef FEAT_VISUALIZATION
+   void VideoViewer::CVvDisplayClient::handleGuiEvent(
+         const Visualization::TEvent &event)
+   {
+     if (event.type == Visualization::evCheckBoxChange) {
+       if (event.sourceId == "toggle.viewer.running") {
+         bool newrcv = (event.data != "0");
+         if (newrcv != pViewer->receiving) {
+           if(pViewer->receiving) {
+             pViewer->videoServer->stopReceiveImages(getComponentID());
+           }
+           else {
+             pViewer->videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
+           }
+           pViewer->receiving = !pViewer->receiving;
+         }
+       }
+     }
+   }
+   #endif
+\end{Verbatim}
+
+Note that the data for UI elements passed between the client and the server is
+serialized into strings. The checkbox element knows only 2 values: \code{"0"}
+and everything else.
+\todo{TODO: Add more element types, define values}
+
+A UI element may be displayed in multiple views. When a UI element is first
+displayed it must be put in a state that reflects the current state of the
+component.  Since a UI element may be created at any time it wouldn't be enough
+to set the initial state of the element at creation time. Instead, we define
+the \code{getControlState()} callback function:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   #ifdef FEAT_VISUALIZATION
+   std::string VideoViewer::CVvDisplayClient::getControlState(const std::string& ctrlId)
+   {
+     if (ctrlId == "toggle.viewer.running") {
+       return pViewer->receiving ? "2" : "0";
+     }
+     return "";
+   }
+   #endif
+\end{Verbatim}
+
+When the function returns \code{""} the display server won't change the state
+of the UI element.
+
+\subsection{Support for HTML forms}\label{section:HtmlForms}
+
+The display server can present HTML forms and send data from a form to a client
+when the form is submitted. The data from a form can also be saved into a
+persistent storage and reloaded later. This way the forms can be used to
+configure the components from the user interface instead of using CAST
+configuration files.
+%
+\todo{TODO: Apply the settings from persistent storage at startup when a component
+registers a form}
+%
+\todo{TODO: Add config param to DS to change the location of persistent storage}
+%
+\todo{TODO: Add menu entry to change the location of persistent storage}
+%
+\todo{TODO: Add widgets to load/save all forms on a page}
+
+To add support for form processing in an active client we override the methods
+\code{"handleForm"} and \code{"getFormData"}. We also add a convenience
+function that will create the form and send it to the server:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   #ifdef FEAT_VISUALIZATION
+   class COrDisplayClient: public cogx::display::CDisplayClient
+   {
+      CObjectRecognizer* pRecognizer;
+   public:
+      void createForms();
+      void handleForm(const std::string& id, const std::string& partId,
+            const std::map<std::string, std::string>& fields); /*override*/
+      bool getFormData(const std::string& id, const std::string& partId,
+            std::map<std::string, std::string>& fields); /*override*/
+   };
+   #endif
+\end{Verbatim}
+
+A form can be defined with standard HTML form elements but it must not be
+enclosed in the \code{"<form>"} tag which is added by the Display Server.
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void CObjectRecognizer::COrDisplayClient::createForms()
+   {
+      std::ostringstream ss;
+      ss << "Min score: <select name='minScore'>";
+      for (float f = 0.7; f <= 1.0; f += 0.01) {
+         ss << "<option>" << sfloat(f, 2) << "</option>";
+      }
+      ss << "</select>";
+      ss << "<input type=\"submit\" name=\"submit\" value=\"Apply\"/>";
+      setHtmlForm("Settings", "ObjectRecognizer", ss.str());
+   }
+   std::string sfloat(double f, int precision)
+   {
+      std::ostringstream out;
+      out << std::fixed << std::setprecision(precision) << f;
+      return out.str();
+   }
+\end{Verbatim}
+
+The getter and the setter functions pass the form parameters in a dictionary.
+If a HTML field supports multiple values (like a multivalue select field, a
+group of check boxes or radio buttons), the values are stored as a delimited
+string where the delimiter is the newline character \Verb|"\n"|.
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   void CObjectRecognizer::COrDisplayClient::handleForm(
+         const std::string& id, const std::string& partId,
+         const std::map<std::string, std::string>& fields)
+   {
+      if (! pRecognizer) return;
+      if (id != "Settings" || partId != "ObjectRecognizer") retrun;
+      float f;
+      typeof(fields.begin()) it;
+      it = fields.find("minScore");
+      if (it != fields.end()) {
+         f = atof(it->second.c_str());
+         if (f < 0.7) f = 0.7;
+         if (f > 1.0) f = 1.0;
+         pRecognizer->m_minScore = f;
+      }
+   }
+
+   bool CObjectRecognizer::COrDisplayClient::getFormData(
+         const std::string& id, const std::string& partId,
+         std::map<std::string, std::string>& fields)
+   {
+      if (! pRecognizer) return false;
+      if (id != "Settings" || partId != "ObjectRecognizer") retrun false;
+      fields["minScore"] = sfloat(pRecognizer->m_minScore, 2);
+      fields["maxAmbiguity"] = sfloat(pRecognizer->m_maxAmbiguity, 2);
+      return true;
+   }
+\end{Verbatim}
+
+A form can be shown in a separate view or as a part of an HTML page composed of
+multiple chunks. When a page contains a form it is slower to display because
+some javascript has to be evaluated every time the page is reloaded. Therfore
+the forms should not be a part of a page that is changed frequently (eg. more
+than once per second).

### 2010-07-09 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

@@
 configuration files.
-%
-\todo{TODO: Apply the settings from persistent storage at startup when a component
-registers a form}
-%
-\todo{TODO: Add config param to DS to change the location of persistent storage}
-%
-\todo{TODO: Add menu entry to change the location of persistent storage}
-%
-\todo{TODO: Add widgets to load/save all forms on a page}
 

### 2010-07-14 subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: manual updated (installation)

@@
 the visualization subarchitecture.
+
+\subsection{Java client interface}
+
+To build the Java client interface add the following to the main build.xml file:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   <target name="template" depends="init">
+      ...
+      <ant target="${target}" dir="subarchitectures/visualization.sa" />
+      ...
+   </target>
+\end{Verbatim}
+
+\subsection{Python client interface}
+
+The Python client interface is installed as part of the CMake installation
+process. To enable the installation the following steps should be performed:
+
+\begin{enumerate}
+   \item Add to \verb|svn:externals| of the CogX root directory:
+
+   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
+      ^/code/tools/python/trunk         tools/python
+   \end{Verbatim}
+
+   \item Perform an update get tools/python and cmake/PythonBuild.cmake:
+
+   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
+      svn update
+   \end{Verbatim}
+
+   \item Add at the beginning of the main CMakeLists.txt file, \\
+   somewhere {\em after} {\scriptsize \verb|set(CMAKE_INSTALL_PREFIX ...)|}:
+
+   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
+      include(cmake/PythonBuild.cmake)
+   \end{Verbatim}
+
+   \item Check that the option \verb|BUILD_PYTHON_COMPONENTS| is ON when
+   running \verb|ccmake|.
+
+\end{enumerate}
 
@@
 The display server can present HTML forms and send data from a form to a client
-when the form is submitted.
+when the form is submitted. The data from a form can also be saved into a
+persistent storage and reloaded later. This way the forms can be used to
+configure the components from the user interface instead of using CAST
+configuration files.
 
@@
    };
+   #endif
 \end{Verbatim}
@@
 A form can be defined with standard HTML form elements but it must not be
-enclosed in the \code{"<form>"} tag which is added by the Display Server.
+enclosed in the \code{"<form>"} tag which is added by the Display Server.  The
+"submit" button will also be added by the Display Server.  It will be placed
+above the form along with the other buttons that handle form data.  If you want
+the button placed in a different location, simply add another one like in the
+example below:
 

### 2010-07-20 code/subarchitectures/dialogue.sa/trunk/resources/grammars/openccg/moloko.v6/docs/comsys.moloko/event-modifier.tex
readding the trunk

 \section{Event Modifier Restriction} \label{sec-Event-Modifier-Restriction}
We mentioned earlier that the Moloko is quite promiscous when it comes to  semantic or pragmatic 'un-acceptability', notions which must be handled grammar externally. There is, however, one major exception to this design principle: verbs are able to lexically specify what kinds of event modifiers they can recieve. This is the counterpart to treating most constituents as modifiers instead of as dependencies (see section X). In this section, first, we will give some examples illustrating the utility of this design feature, and then go into detail on how this was implemented.
\subsection{Motivation}
Consider first this pair of sentences:
\begin{itemize}
\item I played in the room
\item I walked in the room
\end{itemize}
Despite their similarity, these two sentences differ in that the second allows a reading which the first does not. 
In both cases, \emph{in the room} can be a static locational modifier, specifying the place where the event occured.\footnote{for the second, imagine a person pacing.} However, only in the second can \emph{in the room} specify the dynamic 'goal' of the event, i.e., walking \emph{into} the room. There is a case for arguing that this difference in motificational behaviour is inherit in the verb 
and should thus be lexically specified (c.f. Goldberg).  As another example, consider
\begin{itemize}
\item* I am a ball to the door
\end{itemize}
The unacceptability of this sentence can be accommodated by lexically specifying that the copular verb (be) blocks the class of 'dynamic modifiers'. Similarly, in the sentence
\begin{itemize}
\item* I want you to play in the room.
\end{itemize}
we can easily block the reading where \emph{in the room} modifies \emph{want}, i.e., the reading where it was the 'wanting' that occured within the room, instead of the true reading where it is the playing in the room that is wanted.

Another, perhaps less obvious, application of this feature is the blocking of certain questions.  It can be used to correctly allow the first to have 2 readings, the second 1, and the third none. 
\begin{itemize}
\item where did you walk
\item where did you play
\item * where did you want
\end{itemize}

It is clear that we can not and should not attempt to grammatically specify every allowable type of event modification. For one, the type of information that underlies these differences certainly falls under the realm of world knowledge. More importantly, for nearly any restriction we could specify, we could most likely find a set of contexts where these restrictions would no longer apply--language is after all, a tool used by humans who have a remarkable gift for construing situations in novel and unexpected ways. However, practically, the ability to grammar-internally 'rule out' some bizarre or highly unlikely readings is a powerful one. It can eliviate a great deal of burden on parse pruning by limiting what readings we are willing to even consider within the context the task for which the grammar is being employed.

\subsection{Implementation}

The restriction of event modifiers is handled almost identically to the way the combinatorial behaviour of modifiers was handled (see \ref{sec-Dual-Relation-Words} above). It is done through the use of 
\begin{enumerate}
\item  a pair syntatic features which are attached lexically to modifiers and verbs
\item  the inclusion of these features within the syntactic category responsible for 'attaching' the  modifier to its modifee. This includes the type-changing rules used to handle  the modifier function of adverbs and prepositions (XXXXX), and also subordinate clause modifiers.\footnote{although event modifier restriction is surely a matter of semantics and not syntax, there are two reasons why it is better handled via feature. First, OpenCCG  much better at 'accessing' syntactic features than semantic ones. In fact, once semantic content has been added the semantic representation of a parse, it is no longer 'visible' to the parsing at all. Second, as discussed above (XXXXX), one of the design features for Moloko was to keep the semantic sorts as grammar external as possible. Using the event-type to control modification would violate this.}
\end{enumerate}
\subsubsection{Syntactic Features}
The first syntactic feature, m-type (modifier type), is attached to the event modifier and essentially parrots the modifier's semantic sort. For example, the preposition 'into', of semantic type 'm-whereto' is assigned a mod-type value of 's-whereto',

The second feature, m-class, is attached to the verb, and specifies exactly which 'types' of modifiers can attach to it. Just like complexcat-class, this is done in the syntactic type heirarchy by grouping the mod-type values (via multiple inheritance). For example, consider this type definition:\\

       \fv{ m-class-4     [s-frequency s-probability s-comment s-time]} \\\\
This class is very restrictive, allowing only frequency (\emph{always}), probability (\emph{certainly}), time (\emph{on Monday}) and comment (\emph{please}) modifiers, blocking all others (location, dynamic, manner, etc.) (NOTE: m- vs s-) Both of these features are specified lexically via macro. For a full list of the current mod-types and mod-classes see types-feature.ccg

\subsubsection{Syntatic Category Modification}
In the case of prepositions and adverbs (i.e. the dual-relation words able to serve as event modifiers), we simply expanded the type-changing rules to enforce these restrictions. For each of the n different syntactic variants, we added a rule for each of the m different  m-types. Thus, instead of their being only n rules, there are now n x m rules, each specifying a particular syntactic variant and a specific modifer type. Here is an example for event modifying prepositions:\\

\catg{pp \fb{ \fid{M},  cc-type=post-s , m-type=m-whereto} \$_{1} }   
\\ \arrow 
\catg{s\id{E} \fwdsl{} s \fb{ \fid{E},  m-class:s-whereto}  \$_{1} }  : \atsign{e1}{event}(  \diam{Modifier}(m1) ) ) \\

This rule handles dynamic 'whereto' prepositions that modify a sentence on its right, e.g., the preposition \emph{into} in the sentence \emph{I went into the room}. Note as well, that the modified sentence is marked with the feature m-class:s-whereto. This insures that only those verbs which belong to a mod-class which allow (can unify with) whereto modifers can actually be modified by this.

For other event modifier categories, such as subordinate clause markers (when, while, etc), again, the appropriate modifier type was assigned to the modifee's mod-class value.

It is important to remember, that because the semantic classes used for modifiers have actually been built into the grammar signature (by creating features and rules), any further modification to this portion of the semantic type heirarchy would lead to potential inconsistencies within the grammar. Thus, any such changes would need to be reflected within the grammar itself as well. See types-feature.ccg for details.

### 2010-11-16 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
More paper drafting.
 - Charles

% File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\include{macros}

\nocopyright

\pdfinfo{
/Title (Switching in Continual Planning for Practical Robot Control)
/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
/Author (NA)}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
%\title{Switching in Continual Planning for Practical Robot Control}
\title{A Continual Planner that Switches Between Serial and
  Decision-Theoretic Planning} 
\author{NA}
\setcounter{secnumdepth}{0}

\begin{document} 
\maketitle

\begin{abstract}

The problem of practical robot control poses many important challenges
to automated planning.

Planning and plan execution must be robust, seamlessly accommodating
exogenous events and the underlying unpredictability of the
environment.

Plans with a reasonable chance of exhibiting good behaviours must be
synthesised in a timely manner.

interleaved planning and execution.

, reason about degrees of belief and uncertainty about the world,

\end{abstract}

\section{Introduction}

We describe, \pcogx, the planning component in a robotic system that
continuously deliberates in a stochastic dynamic environment in order
to achieve objectives set by the user, and acquire a good
understanding of its environment.
%%

Have a good understanding of the contents and function of rooms, as
well as the linguistic referrents to rooms, widgets, and their visual
qualities.

Having confidence in its beliefs about the linguistic terms for
spaces, and the function of those spaces.

sometimes difficult to predict, with exogenous events, such as the
changing of the objective,

the speed and scalability of software for serial planning in
deterministic

We developed a first-order declarative language, called DTPDDL, for
describing domains of planning problems that correspond to POMDPs.  To
accompany that language we have developed an information-state
\laostar\ procedure for solving problems expressed in DTPDDL. We have
an automated determinisation procedure that generates MAPL
descriptions given DTPDDL input. To accompany that, we have extended
MAPSIM to simulate DTPDDL instances, and modified \fastdownward\ so
that it can find useful serial executions given DTPDDL and MAPL models
of the problem at hand.

 called DTPDDL, along
the lines of PDDL for the partially observable case, an \laostar\
solution procedure, a determinisation of the DTPDDL problem in MAPL,
and modify the \fastdownward\ system to find high quality serial plans
.

We model the environment as a partially observable Markov decision
process. Although planning in that model is undecidable in
general~\cite{}, an optimal finite-horizon plan corresponds to a
contingent plan, that is a function mapping action-observation
histories to actions.

our continual planner is reactive, replanning whenever the underlying
domain and problem models change. For example, replanning occurs if
the motivational component alters the objectives, and if an assumed
outcome of a sensing action is not realised.

brittle sensing model,

the evaluation of a fluent at a state is either known. Moreover, there
is a sensing process that run-time variables  after-which

%% For $\prop \in \state$ we say proposition $\prop$ characterises state
%% $\state$. There is always a unique starting state $\state_0$. The goal
%% $\goal$ is a set of propositions, and we say that state $\state$ is a
%% goal state iff $\goal \subseteq \state$.

%%  To keep this exposition
%% simple, for any two distinct actions $\stochAction_i \neq
%% \stochAction_j$, if outcome $\detAct$ is a possibility for
%% $\stochAction_i$ then it cannot also be a possibility for
%% $\stochAction_j$ -- i.e., if $\detAct \in
%% \detActions(\stochAction_i)$ then $\detAct \not\in
%% \detActions(\stochAction_j)$.

%% The solution to a probabilistic planning problem is a contingency
%% plan. This consists of an assignment of actions to states at each
%% discrete timestep up to the planning horizon $n$. The optimal
%% contingency plan is one which prescribes actions to states that
%% maximise the probability that the goal is achieved within $n$ steps
%% from the starting state $\state_0$. For the purposes of this paper
%% we say a plan fails, i.e. achieves the goal with probability $0$, in
%% situations where it does not prescribe an action. Computing the
%% optimal plan for a problem is computationally intractable, and an
%% important direction for research in the field is to develop
%% heuristic mechanisms for generating small linear plans
%% quickly~\cite{littman:etal:98}

An outline of the paper is as follows. We describe POMDPs with a
propositionally factored representation of states and observations,
and describe how to evaluate plans when they correspond to
finite-state controllers.

\section{Propositional Decision-Theoretic Planning}

%% {\em runtime variables}
%% {\em decision variables}
%% {\em uninitialised variables}
%% {\em omitted variables}

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying world
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.  A problem state
$\state$ is a set of propositions $\state \subseteq \props$, and we
denote \states\ the set of reachable states. The perceptual model is
given in terms of a finite set of stochastic senses $\stochSenses$,
deterministic sensing outcomes $\detSenses$, and perceptual
propositions $\percepts$, called percepts. Here, an observation
$\observ$ is a set of percepts $\observ \subseteq \percepts$, and we
denote \observations\ the set of reachable observations.

Every stochastic action $\stochAction \in \stochActions$ has a
precondition $\poss(\stochAction)$, which is a set of propositions. An
action is applicable at a state $\state$ when $\poss(\stochAction)
\subseteq \state$. We denote $\stochActions(\state)$ the set of
actions applicable at state $\state$.  When $\stochAction \in
\stochActions(\state)$ is applied, nature chooses one elements amongst
a small set of deterministic outcomes $\detActions(\stochAction)
\equiv \{\detAction_1, \ldots, \detAction_k \}$. We denote
$\mu_{\stochAction}(\detAction_i)$ the probability that nature takes
outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. That outcome so chosen has an
effect on the underlying state given in terms of two lists of
propositions. The add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
chosen by nature when \stochAction\ is applied at state $\state$, then
the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
added to $\state$, and those from $\delete(\detAction)$ are removed
from $\state$.

We are concerned with problems that feature partial
observability. Here the underlying state of the world cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in \observations$ that informs what
should be believed about the world. In the POMDP model, if $\action$
is applied effecting a transition to a successor state $\state'$, then
an observation occurs according to the active senses
$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
precept is active, written $\stochSense \in
\stochSenses(\stochAction, \state')$, if the senses
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition
$\poss_\states(\stochSense) \subseteq \propositions$ is satisfied by
the state $\state'$; In the usual sense that
$\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

A POMDP has a starting configuration that corresponds to a subjective
Bayesian belief state. Written $\bstate_0$, this is a probability
distribution over states. We use the notation $\bstate_0(\state)$ to
denote the probability that the starting state is equal to
$\state$. We suppose this is given in a factored tree format.

%% Let $\detSense_i \subseteq \percepts$ for each $i$, so that if
%% we apply $\stochAction$ and transition to $\state'$, the observation
%% received will be a union of choices from the active percepts, which
%% will occur with the probability according to $\psi$ that nature can
%% make those choices.

\subsection{Costs, Rewards, and Plan Evaluation}

Whereas until now we have considered a POMDP model factored in terms
of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of states, \actions\ is the
finite set of actions, and \observations\ is the finite set of
observations.  Where $\state, \state' \in \states$, $a \in \actions$,
from $\mu$ we have a state transition function $\transProb(\state,
\action, \state')$ giving the probability of a transition from state
$s$ to $s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

A general formalism for representing solutions to POMDPs is the
finite-state controller (FSC). This is a three-tuple $\langle \nodes,
\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
nodes, $\selection_\node(\action) = P(\action | \node)$, and
$\transition_\node(\action, \observation, \node') = P(\node'|\node,
\action, \observation)$. The value of state $\state$ at node $\node$
given a problem model, written $V_\node(\state)$, is:

\begin{equation}\label{eq:evaluation}
V_\node(\state) = \sum_{\action \in \actions}
\selection_\node(\action) \reward(\state, \action) + \beta \sum_{\action, \observation,
\state', \node'} \transition_\node(\action, \observation, \node')
\transProb(\state, \action, \state') \obsDist_\observ(\state',
\action) V_{\node'}(\state')
\end{equation}

Using the notation \bstate\ for a belief-state, and writing
$\bstate(\state)$ for the probability the robot believes it is in
state \state, we have that the value of that belief $\bstate$ according to a
controller is:

\begin{equation} \label{eq:valueBelief}
V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
\end{equation}

\section{Serial Planning in Determinisations}

\section{Decision-Theoretic PDDL}

the effects of a sensing schema are perceptual, whereas the effects of
an operator schema are on the underlying state space.

knowledge objectives and  sensory effects.

probabilistic models of the sensing consequences of acting to
quantitatively capture the unreliability of sensing, and thus reason
precisely about the utility of a plan.

planning in practical sized POMDPs is intractable. One approach is to
described the task hierarchically, thereby constrain and simplify the
policy search task. The approach ~\cite{}.

It can arise that in some propositional states an action is
applicable, and in others it is not. Moreover, a belief-state can
assign non-zero probability to states where that applicability holds,
and states where it does not. We forbid execution of actions that are
not applicable in any state of the current belief, and otherwise
suppose that when a non-applicable action is executed, that it has
no effect.

\section{Evaluation}

\section{Discussion}

There have been a number of developments recently towards planning
under uncertainty using systems that were intended for serial planning
in deterministic problems.  Notably,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition,
uses the fast satisficing procedure
\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
corresponding execution trace. The resulting plan is then executed
until the observed trace deviates from the planned trace.

 ---i.e., the values of {\em
runtime variables} in the continual paradigm--- are treated , and
therefore plan for a specific eventuality, replanning from scratch if
something

More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust serial plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}.

%%
Recently the system
\system{Conformant-FF}~\cite{hoffmann:brafman:2006}, demonstrated how
conformant planning ---i.e., serial planning in unobservable worlds---
can be modelled as a deterministic problem, and therefore solved using
serial systems. In this conformant setting, advances have been towards
compact representations and lazy evaluations of belief-states.

\section{Concluding Remarks}

The switching continual planning system we have described serves as
the underlying planner for CogX.

\bibliography{papers}
\bibliographystyle{aaai}
\end{document}

\subsection{PPDDL}

We briefly discuss the description language PPDDL, a fairly
straight-forward extension of the Planning Domain Definition Language
for describing fully observable Markov decision processes.

\small{
\begin{tabtt}
(\= :action look-at-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - label ?p - place) \\
  \> :precondition (and (= (is-in ?r) ?p) \\
  \> \> (= (label ?o) ?l) ) \\
  \> :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
}

\small{
\begin{tabtt}
(\= :observe visual-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - label ?p - place) \\
  \> :execution (look-at-object ?r ?v ?l ?p) \\
  \> :effect (when (= (is-in ?o) ?p) (probabilistic 0.8)) \\
\end{tabtt}
}

\small{
\begin{tabtt}
(\= :sensor look-for-object \\
  \> :agent (?a - robot) \\
  \> :parameters (?l - label ?p - place) \\
  \> :variables (?o - visual-object) \\
  \> :precondition (\=and (= (is-in ?r) ?p) \\
  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?p) ) \\
  \> :effect () \\
  \> :sense (= (is-in ?o) ?p) \\
\end{tabtt}
}

\subsection{MAPL}

### 2010-11-16 code/subarchitectures/dialogue.sa/trunk/apis/dialogue/src/java/de/dfki/lt/tr/dialogue/cplan/matches/GlobalVar.java
syncing with development trunk

@@ -13,7 +13,7 @@ public class GlobalVar extends Match implements MatchLVal {
 
   @Override
   public String toString() {
-    return "##" + _varName;
+    return (_negated ? "!" : "") + "##" + _varName;
   }
 
   /** This constitutes a new global match ?? Is there a scenario where it would

### 2010-11-16 subarchitectures/dialogue.sa/trunk/resources/grammars/openccg/prosodic_moloko.v6.0.2/docs/comsys.moloko/using-moloko.moloko.tex
syncing with development trunk

\section{Using Moloko} \label{sec-Using-Moloko}

This section contains 
\begin{enumerate}
\item help on getting Moloko running on your machine
\item information concerning Moloko's file structure 
\item a general overview of the layout and organization of the grammar.
\item instructions for adding words, or more substantial grammatical alterations
\item instructions for compiling the grammar and testing it.
\end{enumerate}

\subsection{Getting Started}

Moloko is written in \textsc{DotCCG}, the 'higher level' grammar writing language which extends \textsc{OpenCCG}.\footnote{for an introduction to \textsc{DotCCG} see, and for \textsc{OpenCCG} see} Consequently, in order

\subsection{Folder Contents}

To begin, Moloko (like any \textsc{OpenCCG} grammar) consists of two types of files: \textbf{generated} and  \textbf{non-generated}. \\

DO NOT MAKE CHANGES TO GENERATED FILES DIRECTLY.

THESE CHANGES WILL NOT BE SAVED.

FIND THE APPROPRIATE NON-GENERATED FILE AND CHANGE THAT.\\ \\
\lab{Generated Files}
\begin{itemize}
\item the complete \textsc{Dot-CCG}  file:   \ \ \ \  \filename{moloko.ccg}
\item the compiled \textsc{Open-CCG} files:   \filename{grammar.xml, rules.xml,} etc
\end{itemize}
\lab{Non-Generated Files}
\begin{itemize}
\item a batch file for compiling the grammar called \filename{build-moloko} (see below) \\
         (this should be placed in your \textsc{OpenCCG} \filename{bin} directory)
\item the folder \filename{ccg-files} containing:
                  \begin{itemize}
		\item all of the \textsc{Dot-CCG} component files forming the grammar (see below) 
		\item a perl script \filename{merge.pl} used to combine these \filename{.ccg} files to create \filename{moloko.ccg}
		\end{itemize}
\end{itemize}		
\subsection{Grammar Layout}

The Moloko Grammar, contained in \filename{./ccg-files} is divided into 3 parts:
 \begin{enumerate}   
 \item the grammar signature \\ all of the syntactic categories, semantics, lexical families, hierarchies and rules
 \item the dictionary \\ all of the words and their assignment to lexical families, i.e. the entries
 \item the testbed \\ a list of examples used in testing the grammar, and illustrating its features
\end{enumerate}

\subsubsection{The Grammar Signature}

The grammar proper is divided into a number of different files of the form: \\

 \filename{\#\_description-of-contents.ccg}. \\ \\   
The reason the files are numbered like this is because the \textsc{Dot-CCG} compiler is ordered, and consequently any macro-definitions that are used must be defined prior to that.

The first two files \filename{types-ontology.ccg} and \filename{types-features.ccg} contain hierarchies used in the grammar, and in the case of the ontology, for outside inferencing as well. The ontology is a hierarchy of semantic sorts (see section \ref{sec-Ontology}). The feature hierarchy contains both syntactic and semantic feature categories and a listing of their possible (hierarchically organized) values. These definitions also function as 'macros' (in the old \textsc{OpenCCG} sense of the word), i.e., values that  can be 'used' within the grammar and dictionary (i.e. lexically specified).

For a syntactic example, it contains a feature category \textsc(vform) for verbal forms, including values like verb-ing, base, past-particple, infinitive, etc. In the grammar, this is used, for example, to create the family \textbf{v+verb-inf} i.e., the family of verbs which take a verbal compliment in infinitive form (ex. \emph{I want to go home}). It is also used in dictionary entries to specify, e.g., that the past-partiple form of \emph{sing} is \emph{sung}.

For a semantic example, it contains the feature \diam{Aspect} with values perfect and imperfect. These values are attached to events via auxillary verb entries.

The remainder of the files in the Grammar Signature contain the syntactic and semantic categories, families, rules and  dictionary entry creating macros. Syntactic and semantic categories are combinined to make lexical families, like ditranstive verbs. They are also used in rules which are used for a variety of purposes in the grammar. Dictionary-entry creating macros are a new component of the grammar. They utilize the power of the \textsc{Dot-CCG} language to massively simplify the task of adding new words to the grammar. For example, here are some examples of noun entries from dictionary-open.ccg (see below for more details): \\

         \ingram{noun-irr(man, men, person,)}

         \ingram{noun(box, thing,)}

         \ingram{name(GJ, person,)} \\ \\
These grammar components have been divided among the files according to two general organizational principles. First, encapsulation, e.g. \filename{\#\_adj.ccg} contains everything particular to adjectives. you want to find the families, rules, and dictionary macros for adjectives, see this file. That is, everything \textbf{but} the adjective words themselves, which are in the dictionary files (see the next section). Second, generality and efficiency: minimize overlap and put generalized components together, e.g., adjectives, prepositions and adverbs overlap a lot in their grammatical information (like semantics and their syntactic categories). Thus, this common info is contained in \filename{\#\_modifiers.ccg}. \\ \\
Within any of these files, there is an order to the components. 
\begin{enumerate}		
\item  syntactic and semantic categories. 
\item  def-macros used for simplifying the building of lexical families
\item  the various families themselves
\item  associated type-changing rules
\item  dictionary-entry macros
\end{enumerate}
The syntatic and semantic categories created via def-macros are typically those used frequently within
its file, and those used externally. For example, the noun file contains a syntactic category \ingram{n()} which corresponds to 'any old generic noun', i.e. a non-bound n . It can be further specified by
adding feature values to its parameters, e.g. \ingram{n(3rd s-sg)}. Similarly, \ingram{ ENTITY() } is the corresponding semantic representation (both share index) . Again, features and arguments can be further specified by adding them to the parameters.

\subsubsection{The Dictionary}

   The dictionary, i.e. the words in the grammar, are divided into \textbf{closed} and \textbf{open} class entries, located in \filename{X\_dictionary-open.ccg} and \filename{X\_dictionary-closed.ccg} respectively. Each of these files  contains a sorted listing of all the entries currently contained in the grammar. The vast majority are simply 'calls' to (instances of) the various dictionary entry macros specified in the grammar signiture files. For example:

             \ingram{noun(box, thing,)}

              \ingram{verb(give, giving, gave, given, m-class-1, action-non-motion, tv dtv) }

              \ingram{pronoun(I,    1st, sg, I    , me   , my   , mine   , person,)} \\ \\
Some of the more irregular or singular words (specifically closed class function words) are simply
build using the default \textsc{Dot-CCG} syntax. Here, e.g., is the entry for \emph{most} as in \emph{the most ugly}:  \\

\textbf{ word form: Family: \{other forms: args;\}  }

\ingram{ word most : Most-adj: superlative;} \\ \\
Furthermore, some 'higher level' dictionary macros are located within the dictionary files themselves.
They combine a series of other 'base level' macros to create the appropriate lexical entries. For example, the macro \ingram{number} takes the ordinal and cardinal forms of numbers ( \emph{two, second})  and then 'calls' a large number of other macros, like \ingram{adj} (\emph{those \ul{two} balls}), \ingram{context-n} (\emph{those \ul{two}}), \ingram{sgroup-np} (\emph{\ul{two} of the balls}) etc. to account for all of the 'uses' of numbers. This eliminates redundany by collecting all of these entries into a meaningful conceptual category,  instead of having them spread throughout the dictionary.

\subsubsection{The Testbed}

The testbed, contained in \filename{X\_testbed.ccg} is a listing of sample sentences used to test and to 'showcase' the grammar. Each line in the testbed consists of a sentence and the expected number of parses. This listing can  be tested for both parsing and generation using the command \textbf{ccg-test} (see below)

We have organized these test-sentences in an attempt to illustrate the capability of the grammar. See the file itself for more comments.

\subsection{Modifying the Grammar}

Obviously, we can only give a brief overview of some of the most common ways that the grammar is modified.

\subsubsection{Adding New Words (i.e. dictionary entries) to the grammar}

The simplest and most common way of modifying the grammar is to add new words using the current dictionary-entry building macros. In most cases, simply parroting or mimicing one of the currently existing entries will suffice. However, as discussed above, these macros are defined throughout the grammar and for a full description of the parameters involved, you'll have to look into these files (or better yet section \ref{sec-Families}.)

If none of the existing dictionary macros fit your exact needs, this does not necessarily mean that you underlying lexical families, rules and features cannot support what you want. In some cases, you can overload one of the more general dictionary macros by adding extra feature values in the argument list. 
If this still won't work you may be able to combine these appropriately by resorting to the \textsc{Dot-CCG} word entry syntax: \\

	\textbf{unique-id: families (class, pred): \{ }

          \textbf{form1: associated feature-values; }

          \textbf{form2: associated feature-values; }

	...

          \}

\subsubsection{Changing the Ontological Hierarchy}

   The semantic sorts/classes/types attached to the semantic objects produced in the semantic
   representations used in the grammar are specified in \filename{X\_types-ontology.ccg}

   If you wish to modify this hierarchy by either collapsing, expanding, or altering its
   components and their relations, in some cases, this can be done by simply modifying this file. 
   In other cases, however, this will require further changes within other parts of the grammar.

   See \filename{\#\_types-ontology.ccg} for specifics

\subsubsection{Classes}

It is relatively easy to extend the various classes defined in the grammar. See \filename{\#\_types-features.ccg} for details.

\subsubsection{Other changes (features, new lexical families, rules, etc)}

We have tried to organize the grammar in such a way that it can easily be navigated and  extended. Moreover, some of the grammar files include instructions on extending the more 'open' elements of the grammar. Good Luck!

\subsection{Compiling and Testing the Grammar}

Compiling the grammar actually consists of 2 steps:
\begin{enumerate}   
\item merging all of the \filename{.ccg} files to create \filename{moloko.ccg}  (using the \filename{merge.pl} script)
\item converting \filename{moloko.ccg} into the \textsc{OpenCCG} \filename{.xml} files   (using the \textsc{Dot-CCG} Parser \filename{ccg2xml})
\end{enumerate}
These have been combined into a single batch file called \filename{build-moloko}. \\
This file should be moved to your openccg bin director. \\

\smallskip

There are two ways of testing the grammar. These are identical to the old \textsc{OpenCCG} methods:
\begin{enumerate}  
\item Command line parsing and generation using the \filename{tccg} tool
\item Running through the testbed using the \filename{ccg-test} tool
\end{enumerate}

### 2010-11-16 subarchitectures/dialogue.sa/trunk/resources/grammars/openccg/prosodic_moloko.v6/docs/comsys.moloko/families.moloko.tex
syncing with development trunk

\sect{Families}

The Moloko grammar is written in the \textsc{DotCCG} language. Due to the power of its definitional macro function, a large amount of redundancy has been removed from the underlying families. We have aimed to reduce the number of families and move a lot of the variability to the level of the 'dictionary'. Consequently, in what follows, we have included a list of the various 'dictionary forms' used in addition to the traditional families and rules. Section \ref{sec-Higher-Order} discusses the use of 'higher order' dictionary macros to group multiple uses of a single word together. \\
\\ \\
Each dictionary form has an additional form \_XXXXX which allows for alternate word-forms (i.e. separate word-forms which map to the same predicate),  e.g. \\

      \ingram{ noun(mom, person,)}

      \ingram{ \_noun(mom, mamma, person,)}

      \ingram{ \_noun(mom, mommy, person,)} \\ \\
Examples of each family and dictionary form are given. \\
For a discussion of the organization of the grammar files themselves, see section \ref{sec-Using-Moloko}

\subsect{Coordination}
Coordination families are defined locally, i.e. noun and np coordination is with the other nominal families, adjective coordination with adjectives, \footnote{adjectival, adverbial and prepositional coordination are possible in Moloko, see section \ref{sec-Dual-Relation-Words}.} etc. The general form for such a family is: \\

\family{Coord-X-}    {and, but, then}

\entry {\cxx{X}{R}{COM=yes} \fwdsl{} \cx{X}{N} \backsl{*} \cxx{X}{F}{COM=no}}  {\atsign{R}{}( \prop{*} \wand\ \diam{First}(F) \wand\ \diam{Next}(N) )} \\

The feature COM is used to guarantee only a single reading for multiple coordination ( \emph{ball and cup and mug} ), in particular the left-branching or 'staircase' reading. The \diam{First}-\diam{Next} semantics is also used for prepositional chains (section \ref{XXX}) and discourse markers (section \ref{sec-Discourse-Markers}).

Naturally, individual families have control over which syntactic features are enforced on the conjuncts and which are inherited by their 'result'. For example, verb coordination specifies that the conjuncts share vform and agreement features and that these are inherited by the result. Thus, e.g. \emph{I am blue and want a drink} parses, but  \emph{* I am blue and wants a drink} doesn't, similarly \emph{I am sitting and looking at the table} but \emph{* I am sitting and looked at the table}. As another example, sentential coordination specifies that each of the conjuncts are either indicatives or interrogatives. This forces imperatives to combine using verbal coordination, resulting in the nice single mood representation discussed in \ref{sec-Mood}. \\ \\
\lab{Dictionary Forms} \\
\begin{itemize}
\item
 \dict{coord}{form, pos, class}{but, adj, quality}
 \dict{coord+}{form, pos, class, features}{and, n, entity, s-pl pl}
- creates an entry for the specified part-of-speech. The + version allows features to be added to the result.
\end{itemize}

\subsect{Nouns} 
\lab{Syntactic Features}
\begin{itemize}
\item \feat{num}{s-sg  s-pl \{s-pl-sp  s-pl-unsp\}  s-mass} \\
       \fv{s-pl-sp} and \fv{s-pl-unsp} are used to mark, e.g. \emph{the balls} and {the two balls} respectively. See determiners below.
\item \feat{pers}{non-3rd \{1st 2nd \} 3rd}
\item \feat{case}{nom acc acc-loc} \\
          \fv{acc-loc} is used for prepositional compliments (see preps below)
\item \feat{nform}{ nf-real \{full  pro  nf-ctxt \} nf-dummy \{dummy-there \}} \\
	First, specifies if the noun is 'real', i.e. referential, or 'dummy', i.e. purely grammatical e.g. \emph{ \ul{there} is a ball on the table}. Real are further subdivided into full (lexical) (\emph{the ball, GJ}), pronominal or contextualized (\emph{the green ) }. 
\item \feat{cc-type}{  compound-1st  compound-head .... n-all  n-1 ...}\\    
allows the lexical specification of a nouns behavior in the noun-noun compound construction (see rules below). \fv{compound-1st} and \fv{compound-2nd} are atomic values and the rest define syntactic classes like those outline in section X above.

\end{itemize}
\lab{Families}
\begin{enumerate}
\item \family{Noun}    {ball, men, library, water}
          \entry { \cx{n}{T}}  { \atsign{t1}{entity}( \prop{*} )  }
\item \family{Noun+of-np}  {edge, corners, side} 
          \entry { \cx{n}{T} \fwdsl{} \cxx{obl}{T}{of}}
                     { \atsign{t1}{entity}( \prop{*} \wand \  \diam{Owner}(o1:entity))  }
\item \family{Context-n+modifier}  {green, second, big}
           \entry {\cx{n}{T}}  
                      { \atsign{t1}{entity}( \prop{context} \wand\  \diam{Modifier}(m1:modifier \prop{*}))  }
\item \family{Owned-np}  {mine, yours, hers}
           \entryl { \cxx{np}{T}{ s-sg, 3rd, full } }
                      { \atsign{t1}{entity}( \prop{context} \wand\  \diam{Spec}(sg, specific, unique) \wand\ \diam{Owner}(o1:entity \prop{*}))  }
\item \family{Owner-pro}  {my, your, her}
           \entryl { \cxx{np}{T}{ 3rd } \fwdsl{} \cxx{np}{T}{s-sg} }
                   { \atsign{t1}{entity}(   \ \  \diam{Spec}(sg, specific, unique) \wand\ \diam{Owner}(o1:entity \prop{*}) )  }
\item \family{event-np}  {it, this , that}
           \entry {\cx{s}{E} }  
                      { \atsign{e1}{}( \prop{*} )}
\end{enumerate}
for 1 - 3, there is a corresponding 'bare-np' family, i.e. replace \catg{n\id{T}} with \catg{np\id{T}} \\
for 4 - 5, there is also a plural entry, i.e. replace syntactic and sg features with pl
\\ \\
\lab{Rules} \\
1 and 2 act like determiners, specifying certain special types of nouns and changing them to nps.
As these words begin lexically  as nouns, they are able to do what all nouns can do (be modified, act as parts of compounds, etc). 3 handles the noun-noun compound construction (see section X) \\ 
\begin{enumerate}
\item  any plural, unspecific noun into a generic plural \catg{np} e.g. \emph{balls are round} \\
               \tc { \cxx{n}{T}{ s-pl-unsp } }
               { \cxx{np}{T}{ 3rd} }
               {\atsign{t1}{entity}( \diam{Spec}(variable, unspecific) )}
\item  nouns marked as mass into a full \catg{np} e.g. \emph{I want coffee}  \\
           \tc{ \cxx{n}{T}{ s-mass}}
               { \cxx{np}{T}{  3rd, sg}}
               {\atsign{t1}{entity}( \diam{Spec}(variable, uncountable) )}
\item  A noun's ability to function as 1st part or head can  restricted lexically via the cc-type feature.
           Setting the 'resulting' n to \fv{cc-none} blocks the recursion of this rule \\
            \tcl{ \cxx{n}{C}{full, cc-type=compound-1st} }
               { \cxx{n}{T}{cc-type=cc-none} \fwdsl{*} \cx{n}{T cc-type=compound-head} }
               {\atsign{t1}{entity}( \  \diam{Compound}(c:entity) \ )}
\end{enumerate}
\lab{Dictionary Forms} \\
\begin{itemize}
\item
\dict{noun}{sg-form, class, args}{ball, thing,}
\dict{noun-irr}{sg-form, pl-form, class, args}{man, men, person,}
\dict{noun+of-np}{sg-form, class, args}{edge, e-location}
\dict{noun-irr+of-np}{sg-form, pl-form, class, args}{}
- these create a \catg{ \cxx{n}{T}{ s-sg, full} } and\catg{ \cxx{n}{T}{ s-pl, full} }, each with \atsign{T}{class}(\prop{form}). All specification, including \emph{semantic} number is added by the determiner. \item \dict{name}{form, class, args}{GJ, person} 
- creates a  \catg{\cxx{np}{T}{ s-sg, 3rd, full}} with \atsign{T}{class}(\prop{form})   
\item \dict{noun-mass}{form, class}{water, thing}
- creates a  \catg{ \cxx{n}{T}{s-mass, 3rd, full} } with \atsign{T}{class}(\prop{form}). \\ The latter is transformed into the \catg{np} form via rule (see rules)  
\item \dict{pronoun}{pred, pers, num, nom-form, acc-form, owner, owned, class,}{I, sg, 1st, I, me, my, mine, person}
-  creates 6 forms: \catg{ \cxx{np}{T}{num, pers, pro, nom}} and \catg{ \cxx{np}{T}{ num, pers, pro, acc}} and a sg and pl for owner and owned (see families)
\item 
\dict{context-n}{form, class, args}{one, entity, s-sg}
\dict{context-np}{form, class, args}{this, entity, sg s-sg proximal unique specific}
\dict{context-n+modifier}{form, class, args}{green,q-color, s-sg }
- these create \catg{\cxx{n}{T}{nf-ctxt}} and \catg{\cxx{np}{T}{3rd, nf-ctxt}}, each with \atsign{T}{class}(\prop{context})
\item \dict{context-s}{form, args}{that, distal}
- creates \catg{\cxx{s}{E}{fin-deictic, m-class-none}} with \atsign{E}{}(\prop{context})
\end{itemize}

\subsect{Determiners}

\lab{Families}
\begin{enumerate}
\item \family{Det}    {a, the, these}
          \entry {\cxx{np}{T}{ 3rd} \fwdsl{} \cx{n}{T} }  { no-semantics  }
\item \family{SDet}   {every, a\_few, more }
          \entry {\cxx{np}{T}{ 3rd} \fwdsl{} \cx{n}{T}}  { \atsign{T}{entity}( \ \diam{Modifier}(\prop{*} ) \ ) }
\item \family{Un-to-Spec-Det}    {three, four } 
           \entry {\cxx{n}{T}{ s-pl-sp} \fwdsl{} \cxx{n}{T}{s-pl-unsp} } { \atsign{T}{entity}( \ \diam{Modifier}(\prop{*} ) ) }
\item \family{Group-np}    {some, any} 
          \entry {\cx{np}{T} \fwdsl{} \cxx{obl}{S}{ of} }  { \atsign{T}{entity}( \prop{group} \wand\ \diam{Set}(s:entity) )  }
\item \family{SGroup-np}    {some, any} 
          \entryl {\cx{np}{T} \fwdsl{} \cxx{obl}{S}{ of} }  { \atsign{T}{entity}( \prop{group} \wand\ \diam{Modifier}(\prop{*} ) \wand\ \diam{Set}(s:entity) )  }
\item \family{SGroup-n}    {first, second} 
          \entryl {\cx{n}{T} \fwdsl{} \cxx{obl}{S}{ of} }  { \atsign{T}{entity}( \prop{group} \wand\ \diam{Modifier}(\prop{*} ) \wand\ \diam {Set}(s:entity ) )  }
\item \family{Det-poss-s}    { 's }
          \entry {\cxx{np}{T}{ 3rd} \fwdsl{} \cx{n}{T} \backsl{*} \cx{np}{O} }  { \atsign{T}{entity}( \ \diam{Owner}(o:entity ) \ )  }
\item \family{Numer-id}    { 101, three }
          \entry {\cx{nid}{N} }  { \atsign{N}{number-id}(\prop{*} )  }

\end{enumerate}
Note: 3 does not result in an \catg{np}, another determiner is required, e.g. \emph{\ul{the} three balls} \\
See section \ref{sec-Groups} for a discussion of groups. \\ 
7 is used with the rule below to create a 'post-determiner' for \emph{floor 3}, \emph{office 101}, etc. \\
Possessive pronouns (e.g. \emph{my, your, his}) are handled in the noun family \family{Owner-pro}{} \\
\lab{Rules} \\
\begin{enumerate}
\item  turns a number-id ( \catg{nid}, see above ) into a post-determiner, creating a specific, singular and unique entity which is 'identified' by this number: e.g. \emph{go to \ul{floor 3}} \\
               \tcll{ \cx{nid}{N} }
               { \cxx{np}{T}{3rd} \backsl{*} \cxx{n}{T}{s-sg} }
               {\atsign{t1}{entity}( \diam{Spec}(unique, specific, sg) \wand\ \diam{Modifier}(N:number-id) )}
               \footnote{this rule is not incremental. It should be replaced by a rule which turns a \catg{n} into an \catg{np} \fwdsl{*} \catg{nid}}
\end{enumerate}
\lab{Dictionary Forms} 
\begin{itemize}
\item
 \dict{det}{form, num, args}{a, sg, existential specific}
 \dict{un-to-spec-det}{form, class}{two, number-cardinal}
 \dict{sdet}{form, num, class, args}{three, pl, number-cardinal, existential specific}
- specifies the syn-num of the \catg{n} compliment, adds specification to T through features and/or \diam{Modifier}
\item
\dict{group-np}{form, args}{any, variable specific}
\dict{sgroup-np}{form, class, args}{three, number-cardinal, existential specific} 
\dict{sgroup-n}{form, class, args}{third, number-ordinal, sg s-sg} 
- creates a group with head \catg{n} or \catg{np}, subset specified by feats and/or \diam{Modifier}
\item
\dict{number-id}{form}{three} 
- creates an entry, e.g. \atsign{t1}{number-id}(\prop{three})
\end{itemize}

\subsect{Verbs}

\lab{Syntactic Features}
\begin{itemize}
\item \feat{num and pers}{see Nouns} 
\item \feat{mood}{s-major\{ s-ind s-imp s-int s-ind-ell \} s-minor} \\
	\fv{s-ind-ell} is given to indicatives with contextualized (ellighted/dropped) subjects \\
	\fv{s-minor} is given to clauses which are selected for. In this grammar, this includes
	all lexical verbs (see Mood).
\item \feat{pol}{s-pos s-prov-pos s-neg} \\
          \fv{p-prov-pos} is for verbs which are provisionally positive, i.e. they are able to negated (e.g. \emph{walks} is \fv{s-pos} but \emph{can} is \fv{s-prov-pos} )
\item \feat{vform}{fin\{ fin-clause\{ fin-full fin-ctxt\} fin-deictic\} vf-base vf-to-imp inf ing pp  vf-be } \\
	\fv{pp} is past-particple: \emph{I haven't \ul{seen} him} \\
	\fv{vf-be} is for 'adjectival verbs' \emph{I am \ul{able} to go} \\
	\fv{vf-to-imp} is the same form as \fv{vf-base}, but exists to allow lexical control of imperative.\\
	\fv{fin-deictic} is for pronominal verbs \emph{I said \ul{it} } \\
	\fv{fin-ctxt} is for subj+finite clauses \emph{I did, I should, he couldn't } \\
	\fv{fin-full} is for all other finite clauses \emph{I am hungry, I walked } \\ 
	The separation of \fv{fin} into two levels is necessary because general sentential complements allow deictics, but the indicative mood rules requires a clausal compliment (i.e. \emph{ * I this} ) 
\item \feat{fin}{ be do can should will could would must have} \\
	this feature is currently not being used, but it will allow for handling tag questions 
         \emph{I can do it \ul{can't I} }
\item \feat{mclass}{ s-manner s-instrumental,  ... m-class-1, m-class-2, ...} \\
	the \fv{s-....} values are grouped into classes \fv{m-class-x}. These classes are lexically selected and specify the types of modifiers which can modify this clause (see \ref{sec-Event-Modifier-Restriction} for description, see \filename{types-feature.ccg} for full list of values) 
\end{itemize}

\subsubsection{Basic Verbs}

\lab{Basic Families}	\\ 
All standard verb family entries receive: \\ 
\entryl {s\fb{\fid{E}, num:num, pers:pers, mood = s-minor} \backsl{!} np \fb{\fid{S}, num:num, pers:pers, case = nom} \ \ \ . . .} {\atsign{E}{event}( \prop{*} \wand\  \diam {Actor}(S:entity) . . . )} \\  \\
In the following descriptions, only the additional structure (i.e. that of the verbs compliments) will be given 
Also, in the remainder of this document \fv {num:num, pers:pers} will be abbreviated as \fv{agr}

\begin{enumerate}
\item \family{iv} {I walked}
         \entry { - }  {  -  }
\item \family{tv}    {I saw the ball}
         \entry { \fwdsl{} \cxx{np}{P}{acc}}  {  \wand\ \diam{Patient}(P:entity)   }
\item \family{v+at-np}    {I looked at the ball}
          \entry { \fwdsl{} \cxx{obl}{P}{at} }  { \wand\ \diam{Patient}(P:entity)   }  
	\\ \\ also families for at-np, to-np, with-np, for-np 
\item \family{v+np+prt}    {I pick up the ball / picked it up}
          \entry { \fwdsl{} \cxx{np}{P}{acc} \fwdsl{} \cx{prt}{R} } 
            { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Particle}(R) }
           \entry {\fwdsl{} \cx{prt}{R} \fwdsl{} \cxx{np}{P}{acc} } 
            { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Particle}(R) }
\item \family{v+adj}    {I feel happy}
         \entry {\fwdsl{} \cx{adj}{R}}  { \wand\ \diam{Result}(R:quality) }
\item \family{v+pp-loc}    {it goes on the table}
         \entry {\fwdsl{} \cx{pp}{R}}  { \wand\ \diam{Result}(R:m-location) }
	\\ \\ also family for pp-whereto
\item \family{v+np+pp-whereto}    {I put it on the table}
          \entry { \fwdsl{} \cx{pp}{R} \fwdsl{} \cxx{np}{P}{acc} } 
            { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Result}(R:m-whereto) }
         \\ \\ also family for pp-loc  (\emph{I want it on the table})
\item \family{v+np+adj}    {I made it bigger}
          \entry { \fwdsl{} \cx{adj}{R} \fwdsl{} \cxx{np}{P}{acc} } 
            { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Result}(R:property) }         
\item \family{dtv}      {I gave him the ball}
          \entry {\cxx{np}{P}{acc} \fwdsl{} \cxx{np}{R}{acc} }  { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Recipient}(R:entity) }
\item \family{dtv-to} {I gave it to him}
          \entry { \cxx{obl}{R}{to} \fwdsl{} \cxx{np}{P}{acc} }  { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Recipient}(R:entity) }
\item \family{v+sent-ind}    {I thought I was hungry}
         \entry { \fwdsl{} \cxx{s}{V}{ fin s-ind }}  {  \wand\ \diam{Event}(V:event)   }
\item \family{v+deictic-event}    {I did it}
         \entry { \fwdsl{} \cxx{s}{V}{ fin-deictic}}  {  \wand\ \diam{Event}(V:event)   }
\item \family{v+verb-inf}    {I need to eat}
         \entry { \fwdsl{} ( \cxx{s}{V}{ inf } \backsl{!} \cx{np}{S} ) }  {  \wand\ \diam{Event}(V:event)   }
      \\ \\ also family for verb-ing  (\emph{I kept going})         
\item \family{v+np+verb-inf}    {I wanted him to eat}
         \entry { \fwdsl{} ( \cxx{s}{V}{ inf } \backsl{!} \cx{np}{P} )  \fwdsl{} \cx{np}{P}}  { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Event}(V:event)   }
      \\ \\ also family for verb-ing  (\emph{I kept it going}) and verb-base  (\emph{I helped him eat})        
\item \family{v+instrumental-np+verb-inf}    {you use it to make coffee}
         \entry { \fwdsl{} ( \cxx{s}{V}{ inf } \backsl{!} \cx{np}{P} )  \fwdsl{} \cx{np}{P}}  
         { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Event}(V:event \diam{Modifier}(m-instrumental \wand\ \prop{with} \wand\ \diam{Anchor}(P) ) )   }
\item \family{v+np+from-np-result}    {I made it from plastic, it is made from plastic}\footnote{this is actually as passivization of this family}
          \entry { \fwdsl{} \cxx{obl}{R}{from} \fwdsl{} \cxx{np}{P}{acc} } 
           { \wand\ \diam{Patient}(P:entity) \wand\ \diam{Result}(R:entity) }
           \\ \\ also family for of+np and out\_of-np  (\emph{I helped him eat})                  
\end{enumerate}
\lab{Special Families}	
\begin{enumerate}
\item \family{imp-do} {don't be so loud}
         \entryl {\cxx{s}{E}{ s-imp} \fwdsl{} ( \cxx{s}{E}{ vf-base } \backsl{!} \cx{np}{S} ) } 
                     {  \atsign{E}{event}( \diam{Mood}(imp) \wand\ \diam{Subject}(S:entity \prop{addressee}) \ )}
\item \family{imp-lets} {let's go get some coffee}
         \entryl {\cxx{s}{E}{ s-imp} \fwdsl{} ( \cxx{s}{E}{ vf-base } \backsl{!} \cx{np}{S} ) } 
                     {  \atsign{E}{event}( \diam{Mood}(imp) \wand\ \diam{Subject}(S:entity \prop{addressee+speaker}) \ )}
\item Thanks and Thank-you (see verbs.ccg)
\end{enumerate}
\lab{Dictionary Forms} \\ \\
For each of these dictionary forms, these is a corresponding version xxxx-no-imp which 'blocks' the imperative form of this verb, i.e. it does not create a \fv{vf-to-imp} entry.
\begin{itemize}
\item \dict{verb}{stem, ing, pasttense, pastpart, modifier-class, class, families}{give, giving, gave, given, m-class-2, action-non-motion, tv dtv dtv-to }
- for each family specified, creates an entry for each form \\ (i.e. a past tense form, past-particple form, etc)

\item \dict{verb-reg}{stem, modifier-class, class, families}{want, m-class-3, xxxxx, tv v+verb-inf v+np+verb-inf v+np+pp-loc v+np+adj}
- works identically, but for verbs that are perfectly regular \\ (i.e. stem, stem+ing, stem+ed, stem+ed)

\item \dict{adjectival-verb}{stem, modifier-class, class}{able, m-class-3, ability}
- creates one form which is selected by the auxilliary \emph{be} verb
\end{itemize}

\subsubsection{Copular \emph{be} }

Recall that the copula \emph{be} has 3 basic argument structures (see section \ref{sec-The-Be-Verb} ): 
\begin{enumerate}
\item \emph {it is \ul{a ball}} \ : \ an \catg{np} argument  which agrees in number and person with the subject \catg{np}, blocking readings like \emph{they are a ball}.
\item \emph {it is \ul{on the table ball}} \ : \ a \catg{pp} argument, which must have a semantic type which can modify entities.\footnote{currently, m-location \emph{I am in the room}, m-accompaniment \emph{I am with you}, and m-benefactor \emph{this is for you}. There is a separate entry for each. } 
\item  \emph {it is \ul{blue}} \ : \ an \catg{adj} argument.
\end{enumerate}

Each of these has an entry for the minor form, y/n interrogative form,  minor form  and compliment-questioning form . 
\begin{enumerate}
\item Minor Entries (used in indicatives and questioning the \diam{Cop-Restr} ) each with: 
         \\ \atsign{E}{event}( \prop{*} \diam{Cop-Restr}(S:entity) \diam{Cop-Scope}(X)  ) \\
         \entry  { \cxx{s}{E}{ agr,   s-minor} \backsl{!} \cxx{np}{S}{ agr, nom}  \fwdsl{} \cx{adj}{X}}   { }
         \entry { \cxx{s}{E}{ agr,   s-minor} \backsl{!} \cxx{np}{S}{ agr, nom}  \fwdsl{} \cxx{pp}{} }    { }
         \entry { \cxx{s}{E}{ agr,   s-minor} \backsl{!} \cxx{np}{S}{ num:num, pers:pers, nom, questionable=no}  \fwdsl{} \cxx{np}{ num:num} }  { }      \footnote{the subject in this entry is blocked from being questioned, i.e. what is this receives only one reading, questioning the \diam{Cop-Scope} }
\item Yes/No Interrogative Entries each with:
          \\ \atsign{E}{event}(\prop{*} \diam{Mood}(int) \wand\ \diam{Subject}(S:entity)  \\ \wand\ \diam{Cop-Restr}(S:entity) \diam{Cop-Scope}(X)  )\\
         \entry{ \cxx{s}{E}{ agr,   s-minor} \fwdsl{} \cx{adj}{X} \fwdsl{} \cxx{np}{S}{ agr, nom}  } {}              
         \entry{ \cxx{s}{E}{ agr,   s-minor} \fwdsl{} \cxx{pp}{}  \fwdsl{} \cxx{np}{S}{ agr, nom}  } {}           
         \entry{ \cxx{s}{E}{ agr,   s-minor}  \fwdsl{} \cxx{np}{ num:num} \fwdsl{} \cxx{np}{S}{ num:num, pers:pers, nom}  } {}
\item Entries for questioning the \diam{Cop-Scope}, selected for by the wh-word (see below):
          \\ \atsign{E}{event}(\prop{*}  \wand\ \diam{Subject}(S:entity)  \\ \wand\ \diam{Cop-Restr}(S:entity) \diam{Cop-Scope}(X)  )\\
         \entry{ \cxx{cop}{E}{ agr } \fwdsl{} \cx{adj}{X} \fwdsl{} \cxx{np}{S}{ agr, nom}  } {}              
         \entry{ \cxx{cop}{E}{ agr } \fwdsl{} \cxx{pp}{}  \fwdsl{} \cxx{np}{S}{ agr, nom}  } {}           
         \entry{ \cxx{cop}{E}{ agr } \fwdsl{} \cxx{np}{ num:num} \fwdsl{} \cxx{np}{S}{ num:num, pers:pers, nom}  } {}

\item Negation: There is a negation entry for each of these entries. See verbs.ccg for details
\end{enumerate}

\subsubsection{Presentational \emph{be} }
The presentational \emph{be} construction has three entries corresponding to its different uses:
\begin{enumerate}
\item Minor Entry \emph{there are some balls / what were there} 
\entryl {\cxx{s}{E}{s-minor} \backsl{!} \cxx{np}{D}{dummy-there} \fwdsl{} \cxx{np}{S}{agr} }
            {\atsign{E}{event}( \prop{*} \wand\ \diam{Presented}(S:entity) )} 
\item Y-N Interrogative \emph{were there any balls}    
\entryl {\cxx{s}{E}{s-int} \fwdsl{} \cxx{np}{S}{agr} \backsl{!} \cxx{np}{D}{dummy-there}  }
            {\atsign{E}{event}( \prop{*} \wand \diam{Mood}(int) \wand \diam{Subject}(D:dummy)  \wand  \diam{Presented}(S:entity))} 
\item Inverted Locational \emph{on the table was a ball} 
\entryl {\cxx{s}{E}{s-int} \fwdsl{} \cxx{np}{S}{agr} \backsl{} \cxx{M}{pp}{cc-type:post-n}  }
            {\atsign{E}{event}( \prop{*} \wand \diam{Mood}(ind) \wand \diam{Modifier}(M:m-location)  \wand  \diam{Presented}(S:entity))} 
\end{enumerate}
It is the presented object which determines agreement, i.e. \emph{* there is some balls}. The subject \emph{there} is a 'dummy' contributing nothing to the propositional semantics of event. It consequently receives no semantic role beyond the standard \diam{Subject}. Note that having both 1 and 3 means that the sentence \emph{there was a ball} correctly receives two readings:
\begin{verbatim}
  @b1:presentational(be ^ 
                     <Mood>ind ^ 
                     <Tense>past ^ 
                     <Modifier>(c1:m-location ^ context ^ 
                                <Proximity>m-distal) ^ 
                     <Presented>(b2:thing ^ ball ^ 
                                 <Delimitation>existential ^ 
                                 <Num>sg ^ 
                                 <Quantification>specific))

  @b1:presentational(be ^ 
                     <Mood>ind ^ 
                     <Tense>past ^ 
                     <Presented>(b2:thing ^ ball ^ 
                                 <Delimitation>existential ^ 
                                 <Num>sg ^ 
                                 <Quantification>specific) ^ 
                     <Subject>(t1:dummy ^ there))
\end{verbatim}

\subsubsection{Auxiliary and Modal Verbs}  \label{sec-Auxiliary-and-Modal-Verbs}
These are the five entries for the families handling modal verbs : \\ \\
\family{Modal-vf-base}    {can, should, would}
\begin{enumerate}
\item \ul{Indicative}   \ \ \ \emph{I can walk} 
     \\  acts like an adverbial modifier which also "changes" the vform of the verb.
      \entry { ( \cxx{s}{E}{vform:fin , mclass:mclass, pol:pol} \backsl{!} \cxx{np}{S}{agr} ) 
                 \fwdsl{\wedge} ( \cxx{s}{E}{vf-base , mclass:mclass, pol:pol} \backsl{!} \cxx{np}{S}{agr} )} 
                 { \atsign{E}{event}( \ \diam{Modifier}(\prop{*}) \ )  }

\item \ul{Indicative Contextual-Event}  \ \ \ \emph{I can} 
           \\ the main event is set to \prop{context}. m-class restricted to avoid crazy post mod
\entryl { \cxx{s}{E}{fin-ctxt ,  m-class-3} \backsl{!} \cxx{np}{S}{agr} } { \atsign{E}{event}( \prop{context} \wand\ \diam{Modifier}(\prop{*}) )  }

\item \ul{Yes/No Interrogative}   \ \ \ \emph{can you walk}     
      \entryl {  \cxx{s}{E}{ s-int } \ \fwdsl{} ( \cxx{s}{E}{ vf-base }  \backsl{!} \cxx{np}{S}{agr} ) \ \fwdsl{} \cxx{np}{S}{agr} }
                  { \atsign{E}{event}( \  \diam{Modifier}(\prop{*}) \wand\ \diam{Mood}(int) \wand\ \diam{Subject}(S:entity) \ )  }

\item \ul{Yes/No Interrogative Contextual-Event}   \ \ \ \emph{can you}     
       \\ the main event is set to \prop{context}. m-class restricted to avoid crazy post mod
      \entryl {  \cxx{s}{E}{ s-int, fin-ctxt, m-class-3}  \ \fwdsl{} \cxx{np}{S}{agr} }
                  { \atsign{E}{event}( \prop{context}  \wand\ \diam{Modifier}(\prop{*}) \wand\ \diam{Mood}(int) \wand\ \diam{Subject}(S:entity))}
\item \ul{Atomic}   \ \ \ \emph{can}     
       \\ this is selected for by wh-words and other UDCs (see below).
      \entry {  \cxx{aux}{E}{vf-base} } { \atsign{E}{event}(  \ \diam{Modifier}(\prop{*}) \ )  }
\end{enumerate}
There are only two differences between the treatment of modal verbs and auxiliary verbs (e.g. \emph{be, have, do} etc.). First, whereas modals add their own propositional head using \diam{Modifier} (see section X), auxiliaries do not, adding only tense, aspect and voice features lexically (see secton X). Second modals always select a verb in base form whereas auxiliaries can choose different forms ( e.g. \emph{am walking, did walk} ). Thus, there are 4 other families essentially identical to this except they have no \diam{Modifier}(*) and instead of having \fv{vf-base} the selected verbal group has \fv{pp, ing, vf-be} and \fv{vf-base}.

One exception is passive auxiliary \emph{be}. This requires more complex structure. There are two indicative and two yes/no interrogative entries. The first is for an unexpressed \diam{Actor}. This is marked with \atsign{S}{entity}( \prop{context} ). The second is for an \diam{Actor} specified by an oblique \emph{by} phrase. Note that the feature \diam{Voice}(passive) is added lexically and that the object becomes the \diam{Subject} \\ \\
\family{Aux-passive}  {the ball was picked up , the ball was picked up by GJ}
\begin{enumerate}
\item \entry { ( \cxx{s}{E}{vform:fin , mclass:mclass, pol:pol} \backsl{!} \cx{np}{X} ) 
                 \fwdsl{\wedge} ( \cxx{s}{E}{pp , mclass:mclass, pol:pol} \backsl{} \cxx{np}{S}{agr} \fwdsl{} \cx{np}{X} )} 
                  {\atsign{S}{entity}( \prop{context} ) }
\item \entry { ( \cxx{s}{E}{vform:fin , mclass:mclass, pol:pol} \backsl{!} \cx{np}{X} ) \fwdsl{\wedge} \cxx{obl}{S}{by} 
                 \fwdsl{\wedge} ( \cxx{s}{E}{pp , mclass:mclass, pol:pol} \backsl{} \cxx{np}{S}{agr} \fwdsl{} \cx{np}{X} )} 
                 { no semantics }
\item \entryl {  \cxx{s}{E}{ s-int } \ \fwdsl{} ( \cxx{s}{E}{ pp }  \backsl{} \cxx{np}{S}{agr} \fwdsl{} \cx{np}{X}) \ \fwdsl{} \cxx{np}{X}{nom} }
                  { \atsign{E}{event}( \  \diam{Mood}(int) \wand\ \diam{Subject}(X:entity)  \ ) \\
                  \atsign{S}{entity}( \prop{context} ) }
\item \entryl {  \cxx{s}{E}{ s-int } \ \fwdsl{\wedge} \cxx{obl}{S}{by}  \ \fwdsl{} ( \cxx{s}{E}{ pp }  \backsl{} \cxx{np}{S}{agr} \fwdsl{} \cx{np}{X}) \ \fwdsl{} \cxx{np}{X}{nom} }
                  { \atsign{E}{event}( \   \diam{Mood}(int) \wand\ \diam{Subject}(X:entity)  \ ) }
\end{enumerate}
The \emph{get} passive takes only the indicative entries, not the y-n interrogatives, \\ i.e. \emph{\ul{did} you get hit} not \emph{* \ul{got} you hit} \\ \\
\lab{Dictionary Forms} 
\begin{itemize}
\item \dict{modal}{form, class, args}{can, ability, }
- creates two forms, a positive form with \fv{s-prov-pos} and a negative form (form + nt) with \fv{s-neg} and \diam{Polarity}(neg)
\end{itemize}
All of the auxiliary verbs are individually specified in \filename{dictionary-closed.ccg}

\subsection{Mood Rules} \label{sec-Mood-Rules}\\ \\
These are the Mood Rules described in section \ref{sec-Mood}. 
\begin{enumerate}
\item 
               \tcll {\cxx{np}{S}{ nom, agr } \ \$_{1} }
               { \cxx{s}{E}{s-ind} \ \fwdsl{} ( \cxx{s}{E}{ VFORM:fin-clause,  s-minor, agr} \backsl{!} \cx{np}{S} )  \ \$_{1} }
              { \atsign{E}{event}( \ \wand\  \diam{Mood}(ind) \wand\ \diam{Subject}(S:entity) \ )}
\item 
                \tcll {np \fb{\fid{F} acc } \ \$_{1} }
               { \cxx{s}{E}{s-ind} \ \fwdsl{} ( \cxx{s}{E}{ VFORM:fin-clause,  s-minor , agr} \backsl{!} \cx{np}{S} \fwdsl{} \cx{np}{F}) \ \fwdsl{} \cxx{np}{S}{ nom, agr } \ \$_{1} }
              { \atsign{E}{event}( \ \wand\  \diam{Mood}(ind) \wand\ \diam{Subject}(S:entity) \wand\ \diam{Fronted}(F:entity) \ )}
\item 
               \tcll {\cxx{s}{E}{ agr, vf-to-imp} \backsl{!} \cxx{np}{S}{agr, nom} \ \ \ . . .} 
              { \cxx{s}{E}{ s-imp} \ \ \ . . .}
              { \atsign{E}{event}( \ \wand\  \diam{Mood}(imp) \wand\ \diam{Subject}(S:entity \wand\ \prop{addressee}) \ )} \footnote{ \fv{vf-to-imp} is the same form as \fv{vf-base}. It was added to allow verbs to lexically determine whether or not they can generate imperative forms}
\item 
               \tcll {\cxx{s}{E}{ agr , fin} \backsl{!} \cxx{np}{S}{agr, nom} \ \ \ . . .} 
              { \cxx{s}{E}{ s-ind-ell} \ \ \ . . .}
              { \atsign{E}{event}( \ \wand\  \diam{Mood}(imp) \wand\ \diam{Subject}(S:entity \wand\ \prop{context}) \ )}
\end{enumerate}
1 and 2 require their verbal compliment to be finite and clausal (i.e. \emph{ * I this} ) and also of mood \fv{s-minor} \\
3 and 4 are actually a family of rules, one per argument structure i.e. one for each verb family. \\ 
The need for multiple rules is the result of the combination of 1) the manner in which verbal arguments are ordered in (most?) CCG grammars and 2) the specific workings of the \$ operator. Despite 'coming' first, subjects are built as the last complement of verbs, i.e. a transitive verb receives: \catg{ \cx{s}{E} \backsl{} \cx{np}{S} \fwdsl{} \cx{np}{X} }, not the more 'natural' (i.e. incrementally iconic): \catg{ \cx{s}{E} \fwdsl{} \cx{np}{X} \backsl{} \cx{np}{S}  }. If the latter were employed, a single rule of this nature would handle all argument structures:

\tc {\cxx{s}{E}{ ...}  \ \ \$_{1} \ \backsl{!} \cxx{np}{S}{...} } 
              { \cxx{s}{E}{ ... }  \$_{1} } {\atsign{E}{event}{...}} \\
However, because the  \$ operator attaches to the \emph{immediately previous} category, in this analog rule

\tc {\cxx{s}{E}{ ...}  \  \backsl{!} \cxx{np}{S}{...} \ \ \$_{1} } 
              { \cxx{s}{E}{ ... }  \$_{1} } {\atsign{E}{event}{...}} \\ 
the \$ would attach to the preceding subject \catg{np}, i.e. it would not 'stand for'  the (potential) compliments of the verb, but of the subject. This is clearly not what we want, hence the multiple rules.

\subsection{Open-Class Modifiers} \label{sec-Open-Class-Modifiers}

In this section I will outline some of the common features of adjectives, prepositions and adverbs.
This will allow simplification of these individual sections. \\ \\
\lab{Syntactic Features}
\begin{itemize}
\item \feat{mod-type}{x-manner x-instrumental ....} \\
	 this syntactic feature parrots the semantic sort of the modifier. It is used in modifier restriction	 (currently only event modifiers) 
\item \feat{cc-type}{ post-s pre-s pre-vp post-vp  post-n pre-n  +  classes (see below) }
            these 'atomic' values are collected into classes using multiple inheritence, e.g.\fv{ prep-1 [post-s post-vp post-n] }. Modifiers are lexically assigned a class which specifies the rules it can undergo, i.e. it's    combinatorial possibilities.
\item \feat{degree}{s-no-degrees s-degree \{s-degree-base s-comparative s-superlative\}} \\
           \fv{s-no-degree} is given to modifiers which cannot be given degrees, e.g. stative adjectives
           \fv{s-degree-base} is given to the base form of modifiers, which can then be changed,  e.g. \emph{big} \arrow \emph{more big}
\end{itemize}
\lab{Rules} \\ \\
Recall from section \ref{sec-Dual-Relation-Word} that all open-class modifiers receive an initial lexical entry containing an atomic syntactic  head with its associated semantic index, e.g. generic prepositions begin as \catg{ \cxx{pp}{M}{mod-type,  cc-type } \fwdsl{} \cx{np}{T} }. In general then we have \catg{ \cxx{cat}{M}{mod-type,  cc-type } \ ( + comps) }. \\ \\
Each of the values of cc-type corresponds to a complex syntactic category: \footnote{note that the placement of the comps is dependent on the positioning of the modifier : pre vs post modifee. This is essential for the proper incremental parsing of modification. }
\begin{enumerate}
\item \fv{pre-n} \ \ \ \ : \ \ \catg{ \cx{n}{T} \fwdsl{} \cx{n}{T} \  ( + comps) }
\item \fv{pre-s}  \ \ \ \ : \ \  \catg{ \cx{s}{E} \fwdsl{} \cx{s}{E} \ ( + comps)}
\item \fv{pre-vp}  \ \ \ : \ \  \catg{ ( \cx{s}{E} \backsl{!} \cx{np}{S} ) \ \fwdsl{} ( \cx{s}{E} \backsl{!} \cx{np}{S} ) \ ( + comps)}
\item \fv{post-n} \ \ \ : \ \ \catg{ \cx{n}{T}  \  ( + comps) \ \backsl{*} \cx{n}{T}  }
\item \fv{post-s}  \ \ \ : \ \  \catg{ \cx{s}{E}  \  ( + comps) \ \backsl{*} \cx{s}{E} }
\item \fv{post-vp}  \ \ : \ \  \catg{ ( \cx{s}{E} \backsl{!} \cx{np}{S} ) \  ( + comps) \  \backsl{*} ( \cx{s}{E} \backsl{!} \cx{np}{S} ) }
\item \fv{pre-cop-comp}  \ \ : \ \  \emph{that is \ul{also} a ball, I am \ul{certainly} bigger, he is \ul{never} here }\footnote{see modifiers.ccg for the specific details of these three categories}
\end{enumerate}
Each of the values of mod-type corresponds to one of the 'atomic' values of mclass (see verbs above), 
e.g. \fv {x-location} (modifier feature) \ : \fv{s-location} (\catg{s} feature).

For each category and each combination of cc-type and mclass, there is a rule which changes the atomic lexical cat into a complex modifying cat which imposes its restriction on its modifiee , e.g. \\ \\
   \tcl{\cxx{cat}{M}{x-location,  post-s } \ \$_{1}} 
            { \cx{s}{E}  \  \$_{1} \ \backsl{*} \cxx{s}{E}{ \ mclass:s-location} }
            {\atsign{E}{event}( \ \diam{Modifier}{M:m-location} \ )} 
\\ \\ The use of \ensuremath{ \$_{1} } ensures that any compliments, lexically specified or 'picked up' in some other way, are carried through after the change. \\ \\  Because all of the atomic-to-complex modifier rules follow the same basic pattern, I will not list them individually. \\ \\
\lab{Modifier Modifiers} \\ \\
Each open class modifier has two families for handling pre and post modifiers: \\
\entry {\cx{pos}{M} \fwdsl{} \cx{pos}{M} }  { \atsign{M}{modifier}( \ \diam{Modifier}(X:m-class \prop{*})  \ )  } \\
\entry {\cx{pos}{M} \backsl{*} \cx{pos}{M} }  { \atsign{M}{modifier}( \ \diam{Modifier}(X:m-class \prop{*})  \ )  } \\ \\
There is a single dictionary form which uses these families to create entries: \\ \\
\dict{modifier}{form, pos, side, class}{really, adj, pre, intensity}
Note that each form requires its own dictionary entry

\subsect{Adjectives}
\lab{Families}
\begin{enumerate}
\item \family{Adj}  {big, red, better}
          \entry {\cx{adj}{M}}  { \atsign{M}{property}( \prop{*} )  }
\item \family{Mod-pre-adj-comparative}    {much}
          \entry {\cx{adj}{M} \fwdsl{} \cxx{adj}{M}{ comparative}}  { \atsign{M}{modifier}( \ \diam{Modifier}(\prop{*})  \ )  }
\item \family{More-adj}    {more}
          \entryl {\cxx{adj}{M}{ comparative} \fwdsl{} \cxx{adj}{M}{ degree-base}}  { \atsign{M}{modifier}( \ \diam{Degree}(comparative) \ )  }
\item \family{Most-adj}    {most}
          \entryl {\cxx{adj}{M}{ superlative} \fwdsl{} \cxx{adj}{M}{ degree-base}}  { \atsign{M}{modifier}( \ \diam{Degree}(superlative) \ )  }
\item \family{Adj-er-than}    {than}
          \entryl {\cx{adj}{M} \fwdsl{} np\id{A} \backsl{*} \cxx{adj}{M}{ comparative}}  { \atsign{M}{modifier}( \ \diam{Modifier}(\prop{*} \wand\ \diam{Anchor}(a:entity) ) \ )  }
\end{enumerate}
\lab{Rules} see above \\ \\
\lab{Dictionary Forms} \\
\begin{itemize}
\item \dict{adj-none}{base, class, args}{wrong, q-attitude,}
- creates  \catg{ \cxx{adj}{M}{s-degree-base}} with \atsign{M}{class}(\prop{form})   
\item \dict{adj-deg}{base, comp, sup, class, args}{big, bigger, biggest, q-size, }
- creates an entry for each degree value, e.g.  \catg{\cxx{adj}{M}{s-comparative} } with \\ \atsign{M}{class}(\prop{form} \wand\ \diam{Degree}(comparative) )
\end{itemize}

\subsect{Adverbs}
\lab{Families}
\begin{enumerate}
\item \family{Adv}    {quickly, now, forward}
          \entry { \cx{adv}{M}}  { \atsign{M}{modifier}( \prop{*} )  }
\item \family{Adv+dep-clause}    {when, while, if}
          \entry { \cx{adv}{M} \fwdsl{} \cxx{s}{V}{fin, s-ind} }  { \atsign{M}{Modifier}( \prop{*} \wand\ \diam{Event}(V:event) )  }
\end{enumerate}
\lab{Rules}  see above \\ \\
\lab{Dictionary Forms} \\
\begin{itemize}
\item \dict{adverb}{base, class, cc-class, args}{quickly, manner, adv-all, }
- creates  \catg{ \cxx{adv}{M}{s-degree-base}} with \atsign{M}{class}(\prop{form})   
\item \dict{adverb-deg}{base, comparative, superlative, class, cc-class, args}{soon, sooner, soonest, time, adv-all, }
- creates an entry for each degree value, e.g.  \catg{\cxx{adv}{M}{s-comparative} } with \\ \atsign{M}{class}
\item \dict{adverb+dep-clause}{base, class, cc-class, args}{when, time, adv-1, }
- see family above.
\end{itemize}

\subsect{Prepositions}
\lab{Families} \\ \\
The \diam{Anchor} complement sub-categorizes for entities of sort physical (man, table, kitchen) , time-unit (minute, second, hour) or section(right, left, edge, corner, side). This guarantees that \emph{in the office} and \emph{in five minutes} receive disjoint readings and that \emph{I am to your right} parses but |\emph{I  am to the kitchen} doesn't.\footnote{this sentence of course does have uses, e.g. when giving feedback indicating the current progress of a route travelled, but this is behind the scope of the current grammar} The standard families subcat for physical. \\
The compliment of prepositions is marked as case \fv{acc-loc}. This is also marked on the pro-locational \catg{np}s \emph{here} and \emph{there}. Thus, \emph{in here, up there} are permitted but \emph{I took here and here walked} are not
\begin{enumerate}
\item \family{Prep}    {in , through, with, for}
          \entry { \cx{pp}{M} \fwdsl{} \cxx{np}{A}{acc-loc} }  { \atsign{M}{modifier}( \prop{*} \wand\ \diam{Anchor}(A:physical) )  )  }
\item \family{Prep+of-np-}   {right, in\textunderscore{}front, on\textunderscore{}top }
          \entry { \cx{pp}{M} \fwdsl{} \cxx{obl}{A}{of, acc-loc} }  { \atsign{M}{modifier}( \prop{*} \wand\ \diam{Anchor}(A:physical) )  )  } \\
\\ there are similar families for \emph{to} and \emph{from}, i.e. \emph{next \ul{to} } and \emph{away \ul{from}}          
\item \family{Prep+no-arg-}   { here, there, somewhere }
          \entry { \cx{pp}{M} }  { \atsign{M}{modifier}( \prop{*}  )  }
\item \family{Prep--time-unit}    {for five minutes, in two weeks}
          \entry { \cx{pp}{M} \fwdsl{} \cxx{np}{A}{acc-loc} }  { \atsign{M}{modifier}( \prop{*} \wand\ \diam{Anchor}(A:e-time-unit) )  )  }
\end{enumerate}
\lab{Rules} \\ \\
These rules turn any locational/dynamic preposition into a first conjunct. This is required to handle location chains (see section X). The rule are projective (they add a new 'complement' to the parse), and because they apply to any such preposition, will always 'fire'. Forcing the conjuncts and the result to have the same cc-type weeds out a lot of unnecessary parses.
\begin{enumerate}
\item 
               \tcll{ \cxx{pp}{F}{x-dynamic , cc-type:cc-type }}
               {\cxx{pp}{R}{x-dynamic , cc-type:cc-type} \fwdsl{} \cxx{pp}{N}{x-dynamic , cc-type:cc-type }}
               {\atsign{R}{m-dynamic}( \prop{list} \diam{First}(F:m-dynamic) \wand\ \diam{Next}(N:m-dynamic) ) } 
\item 
               \tcll{ \cxx{pp}{F}{x-location , cc-type:cc-type }}
               {\cxx{pp}{R}{x-location , cc-type:cc-type} \fwdsl{} \cxx{pp}{N}{x-location , cc-type:cc-type }}
               {\atsign{R}{m-location}( \prop{list} \diam{First}(F:m-location) \wand\ \diam{Next}(N:m-location) ) } 
\end{enumerate}
\lab{Dictionary Forms} \\
\begin{itemize}
\item \dict{prp}{form, class, cc-class, args}{into, whereto, prep-2,}
- any prep with a physical \diam{Anchor}
\item \dict{prp--}{subcat, form, class, cc-class, args}{time-unit, for, time-interval, prep-2,}
- time unit \diam{Anchor}
\item \dict{prp+}{comp, form, class, args}{of-np, right, location, prep-1, }
- This also handles no-arg entries 
\end{itemize}

\subsect{Wh-Words}
All wh-word entries control the ordering of the clausal elements (1) , set the syntactic and semantic mood of the clause,  and add  the semantic role \diam{Wh-Restr} which specifies the nature and the scope of the questioned item (3). In some cases, they also have their own compliments which are used to further build up the \diam{Wh-Restr} (2). Thus, all Wh-words have the same top level structure and in what follows we specify only these three components: \\  
\entryl {  \cxx{E}{s}{s-int} \ \  + 1  \Ê\ \ ( +  2 )  }
            { \atsign{E}{event}( \ \diam{Mood}(int) \wand\ \diam{Subject}(S) \wand\ \diam{Wh-Restr}( + 3 ) \  ) } \\ \\
There are a wealth of wh-word families each with a potentially large number of entries. We will begin by separating these families into two broad groups: those which question a role (i.e. they 'fill' an extracted argument) and those which question an event itself (they do no fill an argument slot).

\subsubsection{Questioning a Role}
In this section,we will refer to the questioned role as \emph{item}. Within this sub-set of wh families, we can further divide the entries along two lines. First, the nature of item (subject vs. other) determines a lot about the structure of the entry. 
\begin{itemize}
\item Each subject entry has:  (1) =  \catg{ \fwdsl{} ( \cxx{s}{E}{fin} \backsl{!} \cxx{np}{S}{3rd, s-sg} ) } and (3) involves \atsign{S}{entity}, .i.e. there is a  co-indexing with the subject. 
\item Each other entry involves a verbal group 'missing' a complement, i.e. \catg{ ( \cxx{s}{E}{vform:vform} \backsl{!} \cx{np}{S}  \fwdsl{} \emph{item} ) }. This will be abbreviated to \catg{ v-minus(\emph{item}) }
\end{itemize}
Second, the nature of the clause (copular/presentational-be vs. auxiliary/modal verb + other-verb) is important.  Whereas the copular verb in \emph{where is the ball} controls the syntax and semantics of the clause, the aux in \emph{what is he looking at } and the modal in \emph{where can I put it}  do not: they merely add some semantic information to the clause.\footnote{as well as determining various syntactic features, e.g. verbal agreement, vform of the main verb, etc.} Note that this distinction is not important for subject questions: there is no need for a separate aux/modal entry corresponding to \emph{who went to bathroom} 
\begin{itemize}
\item Each copula entry has: (1) =  \catg{ \fwdsl{} ( \cxx{cop}{E}{agr}   \ \emph{item} ) } \\
          Note that in this case, the subject is handled by the copular verb itself. This 1) allows questioning \diam{Anchor} in, e.g. \emph{who is he with} \footnote{although this is non-incremental} and 2) blocks it in  \emph{who is on xx the table}. 
\item There are two aux/modal entries: 
     \begin{enumerate}
     \item has (1) =  \catg{ \fwdsl{} v-minus (\emph{item}) \fwdsl{} \cxx{np}{S}{3rd, s-sg} \fwdsl{} \cxx{aux}{E}        {pol:pol , fin:fin, . . . } } 
      \item has (1) =  \catg{ \fwdsl{} \ (\cx{s}{G} \ \$_{1} \ \emph{item}) \ \fwdsl{} v-minus ( \fwdsl{} \ \cx{s}{G} \$_{1} ) \fwdsl{} \cxx{np}{S}{3rd, s-sg} \fwdsl{} \cxx{aux}{E}{pol:pol , fin:fin, . . . } } 
\end{enumerate}      
The first handles cases where the questioned item occurs in the main clause, e.g. \emph{what did he pick up \ul{xxx}} and the second where it occurs in a subordinate clause, e.g. \emph{what did he want you to pick up \ul{xxx} }. In the latter case, note that this subordinate clause compliment (\catg{ \fwdsl{} \cx{s}{G} \ \$_{1} \ \emph{item} } ) must be 'repeated' as a compliment of this wh-construction following the verbal group compliment ( \catg{v-minus (...)} )
\end{itemize}
\lab{Families}
\begin{enumerate}
\item \family{Wh-np-}    {who, what}
\item \family{Wh-np-spec-}    {\ul{which} ball, ul{what} ball}
          \\ \emph{item} = \catg{ \fwdsl{} \cx{np}{F}}
          \\ (2)  = \catg{ \fwdsl{} \cxx{n}{F}{s-sg}} 
          \\ (3)  = \atsign{V}{??}( \prop{*} \wand\ \diam{Scope}(F:entity) \ ) 
\item \family{Wh-np-spec-ctxt-}    {\ul{which} }
          \\ \emph{item} = \catg{ \fwdsl{} \cx{np}{F}}
          \\ (2)  = no compliments (the semantic head is 'contextualized' as in \emph{which do you want}
          \\ (3)  =  \atsign{V}{??}( \prop{*} \wand\ \diam{Scope}(F:entity \prop{context}) \ ) \\ \\
There are also two families for questioning the quantity of referents which are countable (i.e. \emph{how\textunderscore{}many} ) or uncountable (i.e. \emph{how\textunderscore{}much} ). \footnote{this could and undoubtedly should be handled using a single form of \emph{how} and two separate lexical entries for \emph{much} and \emph{many} } The only difference is  the number feature value of (2). Each of these families contain entries for both nominally specified for contextualized semantic head. The reason why these have been separated into two families in the case np-spec above is that whereas we want the nominally specified specifier reading for \emph{what} (i.e. \emph{\ul{what} ball did you pick up} ) we did not want the contextualized reading, i.e. \emph{\ul{what} did you pick up} should not receive two readings with differing \diam{Wh-Restr}.
\item \family{Wh-np-qclass-}    { \ul{which} color ball }
          \\ \emph{item} = \catg{ \fwdsl{} \cx{np}{F}}
          \\ (2)  = \catg{ \fwdsl{} \cx{n}{F} \  \fwdsl{} \cx{qclass}{Q} }
          \footnote{The category \catg{qclass}, which stands for Quality Class, refers to classes of properties such: \emph{color, size, shape, temperature, etc.} } 
          \\ (3)  = \atsign{V}{??}( \prop{*} \wand\ \diam{Scope}(Q  \diam{Scope}(F:entity) ) \ )  \\Ê\\
This family also includes a set of entries for contextualized entities allowing the more common \emph{\ul{which} color do you want}. 
\begin{verbatim}
  @w1:cognition(want ^ 
                <Mood>int ^ 
                <Tense>pres ^ 
                <Actor>(y1:person ^ you ^  <Num>sg) ^ 
                <Patient>(c1:entity ^ context) ^ 
                <Subject>y1:person ^ 
                <Wh-Restr>(w2:specifier ^ which ^ 
                           <Scope>(c2:quality ^ color ^ 
                                   <Scope>c1:entity)))
\end{verbatim}
This corresponds to the situation where, say, the recipient is being asked to choose between a blue, a green and a red ball, an answer like \emph{blue} would be referring to \ul{the blue ball}. Consequently, it also makes sense to treat the question as referring to an entity with a contextualized semantic head, in this case \prop{ball}.  In other words, this reading is a better representation of this question's semantics than:
\begin{verbatim}
  @w1:cognition(want ^ 
                <Mood>int ^ 
                <Tense>pres ^ 
                <Actor>(y1:person ^ you ^  <Num>sg) ^ 
                <Patient>(c1:entity ^ color) ^ 
                <Subject>y1:person ^ 
                <Wh-Restr>(w2:specifier ^ which ^ 
                           <Scope>c1:entity))
\end{verbatim} \\
\item \family{Wh-pp-}    {where}
          \\ \emph{item} = \catg{ \fwdsl{} \cx{pp}{F}} 
          \\ (2)  = no compliments
          \\ (3)  =  \atsign{F}{modifier}( \prop{*} ) 
          \entry {}  { \atsign{}{}( \prop{} )  }
\item \family{Wh-sent-}    {what}
          \\ \emph{item} = \catg{ \fwdsl{} \cx{s}{F}}  \ \ ..OR.. \ \ \emph{item} = \catg{ ( \fwdsl{} \cxx{s}{F}{inf} \backsl{!} \cx{np}{S} )} 
          \\ (2)  = no compliments
          \\ (3)  =  \atsign{F}{event}( \prop{*} ) \\ \\
The first handles questioning full sentence complements \emph{what did you say} and the second infinitival compliments \emph{what did you want} where \emph{want} is the same as in \emph{I want to play with the ball}, i.e. the question is more or less \emph{what did you want to do}.          
\item \family{Wh-adj-}    {how}
          \\ \emph{item} = \catg{ \fwdsl{} \cx{adj}{F}} 
          \\ (2)  = no compliments
          \\ (3)  =  \atsign{F}{quality}( \prop{*} ) 
\item \family{Wh-adj-degree}    {\ul{how} big, \ul{how} strong }
          \\ \emph{item} = \catg{ \fwdsl{} \cx{adj}{F}} 
          \\ (2)  = \catg{ \fwdsl{} \cx{adj}{F}} 
          \\ (3)  =  \atsign{V}{??}( \prop{*} \wand\ \diam{Scope}(F:quality) ) 
\item \family{Wh-adj-qclass}    {\ul{which} color, \ul{what} shape }
          \\ \emph{item} = \catg{ \fwdsl{} \cx{adj}{F}} 
          \\ (2)  = \catg{ \fwdsl{} \cx{qclass}{F}} 
          \\ (3)  =  \atsign{V}{??}( \prop{*} \wand\ \diam{Scope}(F:quality) ) \\ \\
Consider the resultative use of the verb \emph{make} exemplified in \emph{I made it big}. These three families provide three ways of questioning this \diam{Result} role, i.e. \emph{\ul{how} / \ul{how big} / \ul{what size} did you make it}. Of course they also provide alternate ways of questioning the adjectival \diam{Cop-Restr} role in the copular verb, i.e. \emph{\ul{how} / \ul{how big} / \ul{what size} is it}.
\end{enumerate}

\subsubsection{Questioning an Event Modifier} \label{sec-Questioning-an-Event-Modifier}
The last family of wh-words contains entries which do not question one of the events 'required' roles, but instead some optional aspect, i.e. a modifier. For example, in the questions 
\begin{itemize}
\item \emph{ \ul{where} are you \prop{sitting}} 
\item \emph{ \ul{where} are you \prop{going}} 
\item \emph{ \ul{when} did you \prop{come} in} 
\item \emph{ \ul{how} did you want me to \prop{walk}} 
\end{itemize}
the wh word is questioning the bolded verb/event in terms of its static location, dynamic (where-to) location, time, and its manner respectively. \\ \\
\family{Wh-sent-modifier}{where, how} \\ 
         (1) = \catg{ \fwdsl{}  ( \cxx{s}{F}{vform:vform mclass:mclass} \backsl{!} \cx{np}{S}) \ \fwdsl{} \cxx{np}{S}{3rd, s-sg} \fwdsl{} \cxx{aux}{F}{pol:pol , fin:fin, . . . } } \\ 
         ..OR.. \\
         \catg{ \fwdsl{} \ ( \cx{s}{F} \ \$_{1} ) \ \fwdsl{} v-minus ( \fwdsl{} \ \cxx{s}{F}{mclass:mclass} \$_{1} ) \fwdsl{} \cxx{np}{S}{3rd, s-sg} \fwdsl{} \cxx{aux}{E}{pol:pol , fin:fin, . . .}} \\
         (2) = no compliments \\
         (3) = \atsign{V}{modifier}( \prop{*} \wand\ \diam{Scope}(F:event) ) \\ \\
The first entry is used for questioning (i.e. scoping over) the main clause (first two examples above), and the 2nd some subordinate clause (third example). The constraint on this clause's \fv{mclass} is lexically specified by the wh-word (see section \ref{sec-Event-Modifier-Restriction}). \\ \\
\lab{Dictionary Forms} \\
\begin{itemize}
\item \dict{wh-word}{form, class, families}{which, specificer, Wh-np-spec- Wh-np-qclass- ...} 
Note that like the dictionary forms for verbs, multiple families can be evoked from a single entry.
\item \dict{wh-word+feats}{form, class, families, feats}{where, Wh-sent-modifier, s-dynamic}
This parrots wh-word but allows the addition of lexical macros. Note that this is how \emph{where} sets restricts its scoped-over event to mclass = \fv{s-dynamic}
\item \dict{quality-class}{form}{size} 
This creates a single entry of the form \cxx{qclass}{Q} with \atsign{Q}{quality}. It is currently only used in questions
\end{itemize}

\subsection{Discourse Markers} \label{sec-Discourse-Markers}
Currently, discourse markers\footnote{also referred to as cue words/phrases, discourse particles/conenctives} (DMs) are given quite naive treatment. They are ordered linearly using the standard \diam{First} and \diam{Next} structure used in coordination. Thus, respectively, \emph{right ok} and \emph{right ok put it over there} receive:
\begin{verbatim}
  @l1:d-units(list ^ 
              <First>(r1:marker ^ right) ^ 
              <Next>(o1:marker ^ ok))

  @l1:d-units(list ^ 
              <First>(r1:marker ^ right) ^ 
              <Next>(l2:d-units ^ list ^ 
                     <First>(o1:marker ^ ok) ^ 
                     <Next>(p1:action-non-motion ^ put ^ 
                            <Mood>imp ^ 
                            <Actor>(a1:entity ^ addressee) ^ 
                            <Patient>(i1:thing ^ it ^ <Num>sg) ^ 
                            <Result>(c1:m-whereto ^ context ^ 
                                     <Proximity>m-distal ^ 
                                     <Modifier>(o2:direction ^ over)) ^ 
                            <Subject>a1:entity)))
\end{verbatim}
A required extension is the handling of post positioned DMs, e.g. \emph{it's in the kitchen \ul{right} / \ul{isn't it} ... }. \\ \\
\lab{Families}
\begin{enumerate}
\item \family{DM}    {yes, no, okay, right}
          \entry {\cxx{du}{R}{COM=no}}  { \atsign{D}{marker}( \prop{*} )  }\footnote{see \ref{sec-Coordination} below for a description of COM}
          \entryl {\cxx{du}{R}{COM=yes} \fwdsl{} \cxx{du}{N}{COM=no}}
                     { \atsign{R}{d-units}( \prop{list} \wand\ \diam{First}(F:marker \prop{*} ) \wand\ \diam{First}(N:marker) }
          \entryl {\cxx{du}{R}{COM=yes} \fwdsl{} \cxx{s}{N}{mood:s-major}}
                     { \atsign{R}{d-units}( \prop{list} \wand\ \diam{First}(F:marker \prop{*} ) \wand\ \diam{First}(N:event ) } \\ \\
There is also a family DM+np which takes an \catg{np} compliment with role \diam{Addressee}. It is used in greetings and closings, e.g. \emph{hello Robot, bye GJ}. 
\end{enumerate}
Note that the pre-sentential entry restricts to major mood. Also, note that the parse for a sentence preceded by a DM will receive category \catg{du} and not \catg{s} \\ \\
Vocative NPs are handled by giving all names a DM reading. Thus, \emph{Robot can you come get it for me} receives: 
\begin{verbatim}
  @l1:d-units(list ^ 
              <First>(r1:animate ^ Robot) ^ 
              <Next>(g1:action-non-motion ^ get ^ 
                     <Mood>int ^ 
                     <Tense>pres ^ 
                     <Actor>(y1:person ^ you ^ <Num>sg) ^ 
                     <Modifier>(c1:modal ^ can) ^ 
                     <Modifier>(c2:m-manner ^ come) ^ 
                     <Modifier>(f1:m-benefactor ^ for ^ 
                                <Anchor>(i1:person ^ I ^  <Num>sg)) ^ 
                     <Patient>(i2:thing ^ it ^ <Num>sg) ^ 
                     <Subject>y1:person))
 \end{verbatim}
\lab{Dictionary Forms} \\
\begin{itemize}
\item
 \dict{dis-marker}{form, class}{yes, alignment}
 \dict{dis-marker+np}{form, class}{hi, greeting}
- these create three items corresponding to the three entries above.
\end{itemize}

\subsect{Other Minor Families}
\lab{Families}
\begin{enumerate}
\item \family{NP-marker}    {I gave it \ul{to} him}
          \entry {\cxx{obl}{T}{MARK=*} \fwdsl{} \cxx{np}{T}{acc} }  { }
\item \family{Infinitive-to}    {I want \ul{to} help you}
          \entry { ( \cxx{s}{E}{inf} \backsl{!} \cx{np}{s} ) \fwdsl{} ( \cxx{s}{E}{vf-base} \backsl{!} \cx{np}{s} )}  {}
          \entry { \cxx{s}{E}{inf} \backsl{!} \cx{np}{s} }  { \atsign{E}{event}( \prop{context} )  }
\item \family{For-verb-ing}    {thank you \ul{for} helping me}
          \entry {( \cxx{s}{E}{for-ing} \backsl{!} \cx{np}{s} ) \fwdsl{} ( \cxx{s}{E}{ing} \backsl{!} \cx{np}{s} )}  { }
\end{enumerate}
1 marks oblique entity compliments. 2 and 3 mark verbs. \\
The second entry in 2 contextualizes event compliments, as in \emph{I wanted to}:
\begin{verbatim}
@w1:cognition(want ^ 
                <Mood>ind ^ 
                <Tense>past ^ 
                <Actor>(i1:person ^ I ^ <Num>sg) ^ 
                <Event>(c1:event ^ context) ^ 
                <Subject>i1:person)
\end{verbatim}

\subsection{Higher Order Dictionary Forms} \label{sec-Higher-Order}
Moloko has made use of \textsc{DotCCG} \ingram{def} macros by creating 'higher order' dictionary forms to group the various, sometimes disparate, entries of single words. For example, here is definition for direction words \emph{right, left, front, back}:
\begin{verbatim}
def direction-word(loc) {
  prp+(of-np, loc, location, prep-all)    
  prp+(of-np, loc, whereto, prep-2) 
  prp+(no-arg, loc, direction, prep-2)  
  noun+of-np(loc, e-region, cc-none)
  noun(loc, e-region, cc-none)
  adj-none(loc, q-location, )
}
\end{verbatim}
Thus, the macro call: \ingram{direction-word(right)} 'calls' each of these 'atomic' macros creating the entire gamut of entries needed for handling the various behaviours of this word:  \\

\emph{I am \ul{right} of the table}

\emph{go \ul{right} of the table}

\emph{go \ul{right}}

\emph{I am to the \ul{right} of the table}

\emph{I am to the \ul{right}}

\emph{I want the \ul{right} one} (i.e. not the left one) \\ \\
Similar macros exist for numbers and determiners/context-nps/group-heads. In fact, the macro \ingram{pronoun} described in \ref{sec-Nouns} above also operates in this way.

A useful future task would be to re-organize other areas of the dictionary in this way. Thinking in this way would also help in removing the redundancy in the ontology (etc. e-location, q-location, m-location).

### 2010-11-16 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Tiny changes
 - Charles

--- 2010-11-16 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
 describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have developed an information-state
+accompany that language we have implemented an information-state
 \laostar\ procedure for solving problems expressed in DTPDDL. We have
@@
 $\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.  A problem state
-$\state$ is a set of propositions $\state \subseteq \props$, and we
-denote \states\ the set of reachable states. The perceptual model is
-given in terms of a finite set of stochastic senses $\stochSenses$,
-deterministic sensing outcomes $\detSenses$, and perceptual
-propositions $\percepts$, called percepts. Here, an observation
-$\observ$ is a set of percepts $\observ \subseteq \percepts$, and we
-denote \observations\ the set of reachable observations.
+state-characterising propositions $\propositions$.
 
-Every stochastic action $\stochAction \in \stochActions$ has a
-precondition $\poss(\stochAction)$, which is a set of propositions. An
+A problem state $\state$ is a set of propositions $\state \subseteq
+\props$. Every stochastic action $\stochAction \in \stochActions$ has
+a precondition $\poss(\stochAction)$, also a set of propositions. An
 action is applicable at a state $\state$ when $\poss(\stochAction)
@@
 actions applicable at state $\state$.  When $\stochAction \in
-\stochActions(\state)$ is applied, nature chooses one elements amongst
+\stochActions(\state)$ is applied, nature chooses one element amongst
 a small set of deterministic outcomes $\detActions(\stochAction)
@@
 We are concerned with problems that feature partial
-observability. Here the underlying state of the world cannot be
+observability. These have a perceptual model given in terms of a
+finite set of stochastic senses $\stochSenses$, deterministic sensing
+outcomes $\detSenses$, and perceptual propositions $\percepts$, called
+percepts. Here, an observation $\observ$ is a set of percepts $\observ
+\subseteq \percepts$, and we denote \observations\ the set of
+reachable observations. The underlying state of the world cannot be
 observed directly, rather, senses $\stochSense \in \stochSenses$
@@
 $\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
-precept is active, written $\stochSense \in
-\stochSenses(\stochAction, \state')$, if the senses
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition
-$\poss_\states(\stochSense) \subseteq \propositions$ is satisfied by
-the state $\state'$; In the usual sense that
-$\poss_\states(\stochSense) \subseteq \state'$.
+precept is active, written $\stochSense \in \stochSenses(\stochAction,
+\state')$, if the senses action-precondition,
+$\poss_\stochActions(\stochSense)$, is equal to $\stochAction$, and
+the state-precondition $\poss_\states(\stochSense) \subseteq
+\propositions$ is satisfied by the state $\state'$; In the usual sense
+that $\poss_\states(\stochSense) \subseteq \state'$.
 %%
@@
 \observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of states, \actions\ is the
-finite set of actions, and \observations\ is the finite set of
-observations.  Where $\state, \state' \in \states$, $a \in \actions$,
-from $\mu$ we have a state transition function $\transProb(\state,
-\action, \state')$ giving the probability of a transition from state
-$s$ to $s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
+belief-state, \states\ is the finite set of reachable propositional
+states, \actions\ is the finite set of actions, and \observations\ is
+the finite set of reachable observations.  Where $\state, \state' \in
+\states$, $a \in \actions$, from $\mu$ we have a state transition
+function $\transProb(\state, \action, \state')$ giving the probability
+of a transition from state $s$ to $s'$ if $a$ is applied. For any
+$\state$ and $\action$ we have $\sum_{\state' \in \states}
+\transProb(\state, \action, \state') = 1$.
 %%

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Few small changes.
 -Charles

@@
 
+Finally, we make a number of fairly standard assumptions. First, that
+action execution and sensing occurs instantaneously, and that only one
+action can be applied at a plan-step. Second, it can arise that in
+some propositional states an action is applicable, and in others it is
+not. Moreover, a belief-state can assign non-zero probability to
+states where that applicability holds, and states where it does
+not. We differ
+from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
+because we forbid the execution of actions that are not applicable in
+any state of the current belief.  Also,
+unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
+we do incorporate a PPDDL-like default semantics, supposing that when
+an action \stochAction\ is executed at some \bstate, that it has no
+effect on states $\bstate(\state) > 0$ where $\poss(\stochAction)
+\not\subseteq \state$.
+
 \subsection{Costs, Rewards, and Plan Evaluation}
@@
 
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
+knows where the {\em cornflakes} are--- are treated as
+control-knowledge by the planner. Here, CP is not allowed to start
+executing a plan, unless in the finial stet the goal condition is
+satisfied.
+
 \section{Decision-Theoretic PDDL}
 
-the effects of a sensing schema are perceptual, whereas the effects of
-an operator schema are on the underlying state space.
+The modelling language of choice for planning in probabilistic
+problems is the Probabilistic Planning Domain Definition
+Language~\cite{younes:littman:04}. PPDDL has been used to model
+benchmarks in all 3 of the International Planning Competitions since
+2004. A variation on PDDL for describing domains with stochastic
+actions and uncertain starting configurations, PPDDL is a declarative
+first-order (a.k.a., relational) that facilitates factored description
+of domains and problem instances. There are straightforward
+compilations from problems expressed compactly in PPDDL to
+propositional representations amenable to state-of-the-art planning
+procedures.  Because this language cannot model domains that feature
+partial observability, for our setting we have developed DTPDDL, an
+extension of PPDDL to partial observability.
+
+For example, $(:\pp{predicates} (\pp{location} ?p - VisualObject))$
+says that there is a unary predicate
+
+the effects of a {\em sense} schema are perceptual, whereas the
+effects of an {\em operator} schema are over state propositions.
 
@@
 policy search task. The approach ~\cite{}.
-
-It can arise that in some propositional states an action is
-applicable, and in others it is not. Moreover, a belief-state can
-assign non-zero probability to states where that applicability holds,
-and states where it does not. We forbid execution of actions that are
-not applicable in any state of the current belief, and otherwise
-suppose that when a non-applicable action is executed, that it has
-no effect.
 

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Belief revision.
 - Charles

@@
 
-Whereas until now we have considered a POMDP model factored in terms
-of propositions and percepts, in order to discuss utilities and
+Whereas until now we have considered a POMDP model that is factored in
+terms of propositions and percepts, in order to discuss utilities and
 policies it is convenient to consider the underlying decision process
@@
 \observations} \obsDist_\observ(\state, \action) = 1$.
+
+Successive state estimation is available if we applying Bayes' rule.
+Taking the current belief $\bstate$ as our {\em prior}, and supposing
+action $\action$ is executed with perceptive outcome $\observ$, then
+the probability that we are in $\state$ in the successive belief-state
+$\bstate'$ is given by:
+
+\begin{equation}\label{eq:revision}
+\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
+  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
+\end{equation}
+
+\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
+normalising factor, that is, the probability of getting observation
+$\observ$ given we execute $\action$ in $\bstate$.
 
@@
 $(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes} are--- are treated as
-control-knowledge by the planner. Here, CP is not allowed to start
-executing a plan, unless in the finial stet the goal condition is
-satisfied.
+knows where the {\em cornflakes}--- are treated as control-knowledge
+by the planner. Here, a plan is not considered valid unless in the
+finial stet the goal condition is satisfied.
 
@@
 problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04}. PPDDL has been used to model
-benchmarks in all 3 of the International Planning Competitions since
-2004. A variation on PDDL for describing domains with stochastic
-actions and uncertain starting configurations, PPDDL is a declarative
-first-order (a.k.a., relational) that facilitates factored description
-of domains and problem instances. There are straightforward
-compilations from problems expressed compactly in PPDDL to
-propositional representations amenable to state-of-the-art planning
-procedures.  Because this language cannot model domains that feature
-partial observability, for our setting we have developed DTPDDL, an
-extension of PPDDL to partial observability.
+Language~\cite{younes:littman:04}. PPDDL was used in all 3 of the
+International Planning Competitions since 2004. A variation on PDDL
+for describing domains with stochastic actions and uncertain starting
+configurations, PPDDL is a declarative first-order (a.k.a.,
+relational) that facilitates factored description of domains and
+problem instances. There are straightforward compilations from
+problems expressed compactly in PPDDL to propositional representations
+amenable to state-of-the-art planning procedures.  Because this
+language cannot model domains that feature partial observability, for
+our setting we have developed DTPDDL, an extension of PPDDL to partial
+observability.
 

### 2010-11-17 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Equation fix.
 - Charles

@@
 
+Finally, we make a number of fairly standard assumptions. First, that
+action execution and sensing occurs instantaneously, and that only one
+action can be applied at a plan-step. Second, it can arise that in
+some propositional states an action is applicable, and in others it is
+not. Moreover, a belief-state can assign non-zero probability to
+states where that applicability holds, and states where it does
+not. We differ
+from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
+because we forbid the execution of actions that are not applicable in
+any state of the current belief.  Also,
+unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
+we do incorporate a PPDDL-like default semantics, supposing that when
+an action \stochAction\ is executed at some \bstate, that it has no
+effect on states $\bstate(\state) > 0$ where $\poss(\stochAction)
+\not\subseteq \state$.
+
 \subsection{Costs, Rewards, and Plan Evaluation}
 
-Whereas until now we have considered a POMDP model factored in terms
-of propositions and percepts, in order to discuss utilities and
+Whereas until now we have considered a POMDP model that is factored in
+terms of propositions and percepts, in order to discuss utilities and
 policies it is convenient to consider the underlying decision process
@@
 
+Successive state estimation is available if we applying Bayes' rule.
+Taking the current belief $\bstate$ as our {\em prior}, and supposing
+action $\action$ is executed with perceptive outcome $\observ$, then
+the probability that we are in $\state$ in the successive belief-state
+$\bstate'$ is given by:
+
+\begin{equation}\label{eq:revision}
+\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
+  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
+\end{equation}
+
+\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
+normalising factor, that is, the probability of getting observation
+$\observ$ given we execute $\action$ in $\bstate$.
+
 A general formalism for representing solutions to POMDPs is the
@@
 \begin{equation}\label{eq:evaluation}
-V_\node(\state) = \sum_{\action \in \actions}
-\selection_\node(\action) \reward(\state, \action) + \beta \sum_{\action, \observation,
+\begin{array}{lcl}
+V_\node(\state) & = & \sum_{\action \in \actions}
+\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+
+&& \hspace{-10ex} \beta \sum_{\action, \observation,
 \state', \node'} \transition_\node(\action, \observation, \node')
@@
 \action) V_{\node'}(\state')
+\end{array}
 \end{equation}
@@
 
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
+knows where the {\em cornflakes}--- are treated as control-knowledge
+by the planner. Here, a plan is not considered valid unless in the
+finial stet the goal condition is satisfied.
+
 \section{Decision-Theoretic PDDL}
 
-the effects of a sensing schema are perceptual, whereas the effects of
-an operator schema are on the underlying state space.
+The modelling language of choice for planning in probabilistic
+problems is the Probabilistic Planning Domain Definition
+Language~\cite{younes:littman:04}. PPDDL was used in all 3 of the
+International Planning Competitions since 2004. A variation on PDDL
+for describing domains with stochastic actions and uncertain starting
+configurations, PPDDL is a declarative first-order (a.k.a.,
+relational) that facilitates factored description of domains and
+problem instances. There are straightforward compilations from
+problems expressed compactly in PPDDL to propositional representations
+amenable to state-of-the-art planning procedures.  Because this
+language cannot model domains that feature partial observability, for
+our setting we have developed DTPDDL, an extension of PPDDL to partial
+observability.
+
+For example, $(:\pp{predicates} (\pp{location} ?p - VisualObject))$
+says that there is a unary predicate
+
+the effects of a {\em sense} schema are perceptual, whereas the
+effects of an {\em operator} schema are over state propositions.
 

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
New notations.
 - Charles

@@
 continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire a good
-understanding of its environment.
+to achieve objectives set by the user, and acquire knowledge about its
+surroundings.
 %%
@@
 We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying world
+planning problem, with costs and rewards. The underlying process
 dynamics are modelled in terms of a finite set of stochastic actions
@@
 
-A problem state $\state$ is a set of propositions $\state \subseteq
-\props$. Every stochastic action $\stochAction \in \stochActions$ has
-a precondition $\poss(\stochAction)$, also a set of propositions. An
-action is applicable at a state $\state$ when $\poss(\stochAction)
-\subseteq \state$. We denote $\stochActions(\state)$ the set of
-actions applicable at state $\state$.  When $\stochAction \in
-\stochActions(\state)$ is applied, nature chooses one element amongst
-a small set of deterministic outcomes $\detActions(\stochAction)
-\equiv \{\detAction_1, \ldots, \detAction_k \}$. We denote
-$\mu_{\stochAction}(\detAction_i)$ the probability that nature takes
-outcome $\detAct_i$, and for all \stochAction\ we require
+We model a process state $\state$ as the set of propositions that are
+true of the state. Notationally, we have $\state \subseteq
+\props$. State change is induced by application of actions. A
+stochastic action $\stochAction \in \stochActions$ is applicable if
+its precondition $\poss(\stochAction)$, a set of propositions, are
+satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
+\state$. We write $\stochActions(\state)$ for set of actions
+applicable at state $\state$.  If action $\stochAction \in
+\stochActions(\state)$ is applied at \state, nature chooses one
+element amongst a small set of deterministic outcomes
+$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
+\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
+nature takes outcome $\detAct_i$, and for all \stochAction\ we require
 $\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. That outcome so chosen has an
-effect on the underlying state given in terms of two lists of
-propositions. The add-list $\add(\detAction)$ and delete-list
+\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
+effect on the state given in terms of two lists of propositions. The
+add-list $\add(\detAction)$ and delete-list
 $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
@@
 observability. These have a perceptual model given in terms of a
-finite set of stochastic senses $\stochSenses$, deterministic sensing
-outcomes $\detSenses$, and perceptual propositions $\percepts$, called
-percepts. Here, an observation $\observ$ is a set of percepts $\observ
-\subseteq \percepts$, and we denote \observations\ the set of
-reachable observations. The underlying state of the world cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in \observations$ that informs what
-should be believed about the world. In the POMDP model, if $\action$
-is applied effecting a transition to a successor state $\state'$, then
-an observation occurs according to the active senses
-$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
-precept is active, written $\stochSense \in \stochSenses(\stochAction,
-\state')$, if the senses action-precondition,
-$\poss_\stochActions(\stochSense)$, is equal to $\stochAction$, and
-the state-precondition $\poss_\states(\stochSense) \subseteq
-\propositions$ is satisfied by the state $\state'$; In the usual sense
-that $\poss_\states(\stochSense) \subseteq \state'$.
+finite set of stochastic {\em senses} $\stochSenses$, deterministic
+sensing outcomes $\detSenses$, and perceptual propositions
+$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
+a set of percepts $\observ \subseteq \percepts$, and we denote
+\observations\ the set of reachable observations. The underlying state
+of the process cannot be observed directly, rather, senses
+$\stochSense \in \stochSenses$ effect an observation $\observ \in
+\observations$ that informs what should be believed about the state a
+process is in. In detail, if $\action$ is applied effecting a
+transition to a successor state $\state'$, then an observation occurs
+according to the active senses $\stochSenses(\stochAction, \state')
+\subseteq \stochSenses$. A precept is active, written $\stochSense \in
+\stochSenses(\stochAction, \state')$, if the senses'
+action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
+$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
+\subseteq \propositions$ is satisfied by the state $\state'$; In the
+usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
 %%
@@
 
-A POMDP has a starting configuration that corresponds to a subjective
-Bayesian belief state. Written $\bstate_0$, this is a probability
-distribution over states. We use the notation $\bstate_0(\state)$ to
-denote the probability that the starting state is equal to
-$\state$. We suppose this is given in a factored tree format.
-
-%% Let $\detSense_i \subseteq \percepts$ for each $i$, so that if
-%% we apply $\stochAction$ and transition to $\state'$, the observation
-%% received will be a union of choices from the active percepts, which
-%% will occur with the probability according to $\psi$ that nature can
-%% make those choices.
+A POMDP has a starting configuration that corresponds to a Bayesian
+belief-state. Intuitively, this is the robot's subjective belief about
+its environment. Technically, a belief-state $\bstate$ is a
+probability distribution over process states. We write
+$\bstate(\state)$ to denote the probability that the process is in
+$\state$ according to $\bstate$, and $\bstate_0$ when discussing the
+starting configuration. We suppose the latter is given in a factored
+tree format.
 
@@
 we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at some \bstate, that it has no
-effect on states $\bstate(\state) > 0$ where $\poss(\stochAction)
-\not\subseteq \state$.
+an action \stochAction\ is executed at $\state_i$ where
+$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
+$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
+0$.
 
@@
 states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations.  Where $\state, \state' \in
-\states$, $a \in \actions$, from $\mu$ we have a state transition
-function $\transProb(\state, \action, \state')$ giving the probability
-of a transition from state $s$ to $s'$ if $a$ is applied. For any
-$\state$ and $\action$ we have $\sum_{\state' \in \states}
-\transProb(\state, \action, \state') = 1$.
+the finite set of reachable observations (i.e., perceptual states).
+Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
+have a state transition function $\transProb(\state, \action,
+\state')$ giving the probability of a transition from state $s$ to
+$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
+$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
 %%
@@
 \action) V_{\node'}(\state')
+\end{array}
 \end{equation}
 
-Using the notation \bstate\ for a belief-state, and writing
-$\bstate(\state)$ for the probability the robot believes it is in
-state \state, we have that the value of that belief $\bstate$ according to a
-controller is:
+\noindent The value of $\bstate$ according to a controller is
+therefore:
 

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Forgot the bibliography last time.
 - Charles

@@
 according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A precept is active, written $\stochSense \in
+\subseteq \stochSenses$. A percept is active, written $\stochSense \in
 \stochSenses(\stochAction, \state')$, if the senses'
@@
 belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Technically, a belief-state $\bstate$ is a
-probability distribution over process states. We write
-$\bstate(\state)$ to denote the probability that the process is in
-$\state$ according to $\bstate$, and $\bstate_0$ when discussing the
-starting configuration. We suppose the latter is given in a factored
-tree format.
+its environment. Formally, a belief-state $\bstate$ is a probability
+distribution over process states. We write $\bstate(\state)$ to denote
+the probability that the process is in $\state$ according to
+$\bstate$, and $\bstate_0$ when discussing the starting
+configuration.
 
@@
 problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04}. PPDDL was used in all 3 of the
-International Planning Competitions since 2004. A variation on PDDL
-for describing domains with stochastic actions and uncertain starting
-configurations, PPDDL is a declarative first-order (a.k.a.,
-relational) that facilitates factored description of domains and
-problem instances. There are straightforward compilations from
+Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
+all 3 of the International Planning Competitions since 2004. A
+variation on PDDL for describing domains with stochastic actions and
+uncertain starting configurations, PPDDL is a declarative first-order
+(a.k.a., relational) that facilitates factored description of domains
+and problem instances. There are straightforward compilations from
 problems expressed compactly in PPDDL to propositional representations
@@
 observability.
+
+We suppose $\bstate_0$ is given in a factored tree format of the
+form:
+
+\def\oom{$^{\tt +}$}
+\def\zom{$^*$}
+\def\bump{\hspace{1cm}}
+\def\req#1{$^{\tt #1}$}
+\def\noteme#1{}%{[[#1]]}
+\def\notecoauth#1{\ }
+\def\notereader#1{[[#1]]}
+\def\meta{$\uparrow\uparrow$}
+\def\la{\langle}
+\def\ra{\rangle}
+
+\scriptsize
+\begin{nopagebreak}\begin{tabtt}
+\bump \= <observation-def> \bump\=::= \=(:obs\=erve <o-symbol> \+\+\+\+\\
+                        %%{[}<extension-def>]  \\
+                        :parameters (<typed list (variable)>)  \\
+                        <o-def body>) \-\-\- \\
+  <o-def body> \> ::= [:state <emptyOr (pre-GD)>] \\
+  \> \> [:action <atomic action(term)> ] \\
+    \> \> [:effect <emptyOr (o-effect)>] \\
+  <o-symbol> \> ::= <name> \\
+  <atomic action(t)> \> ::= (<action-symbol> t\zom) \\
+  <o-effect> \> ::= (and <c-o-effect>\zom) \\
+  <o-effect> \> ::= <c-o-effect> \\
+  <c-o-effect> \> ::= \req{:conditional-effects} (when <GD>
+                        <o-effect>) \\  
+  <c-o-effect> \> ::= \req{:universal-effects} (forall (<typed list (var)>) <o-effect>) \\
+  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic
+  <prob> <o-effect>) \\
+  <c-o-effect> \> ::=  \req{:probabilistic-effects}\req{:universal-unwinding} \\
+%% 
+\hspace{3ex}(probabilistic (for-each (<typed list (x)>) <s-f-head>  <o-effect> )) \\
+  <c-o-effect> \> ::= <p-o-effect> \\
+  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
+  <p-o-effect> \> ::= <atomic o-formula(term)> \\
+  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
+  <p-o-effect> \> ::= \req{:fluents}(<assign-op> <o-f-head> <o-f-exp>) \\
+  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
+  <o-f-exp> \> ::= <number>\\
+  <o-f-exp> \> ::= (<binary-op>  <o-f-exp> <o-f-exp>)\\
+  <o-f-exp> \> ::= (<multi-op>  <o-f-exp> <o-f-exp>\oom)\\
+  <o-f-exp> \> ::= (- <o-f-exp>)\\
+  <o-f-exp> \> ::= <o-f-head>\\
+  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
+  <o-f-head> \> ::= <o-function-symbol>\\
+\end{tabtt}\end{nopagebreak}
+\noteme{<structure-def> ::= <attach-def>}
+\normalsize
+
+%% \small{
+%% \begin{tabtt}
+%% (:init  (and \=f_{11} f_{12} ..
+%%  \> (probabilistic p_{}) 
+%%  \> (probabilistic ..) ..) \\
+%% \end{tabtt}
+%% }
+
+%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
+
+\noindent Nodes of the tree have one of three types: (1) atomic nodes
+are labelled with a state proposition, e.g., $(=
+(\pp{location}~\pp{cornflakes}) \pp{kitchen})$, (2) probabilistic
+nodes, have the form $(r_1, C_1;..;r_n, C_n)$, and (3) constructive
+nodes, take the form of a set of probabilistic and atomic nodes. The
+root node is conjunctive, and leaves are always atomic. A convenient
+way to give semantics to the tree, is to treat nodes as if they were
+actions. An atomic node is an action with a positive effect
+corresponding to the node label, .
 
@@
 
+\bibliographystyle{aaai}
 \bibliography{papers}
-\bibliographystyle{aaai}
+
 \end{document}

### 2010-11-17 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
I'll try and work on this again tonight. I'm not happy with how I think the paper is going to end up, but I don't know what else to do. I think we need to make the introduction and discussion sections as strong as possible. Just make it an interesting discussion of the literature here.
And make sure the experiments show our sensing is better!!

 - Charles

@@
 observability. These have a perceptual model given in terms of a
-finite set of stochastic senses $\stochSenses$, deterministic sensing
-outcomes $\detSenses$, and perceptual propositions $\percepts$, called
-percepts. Here, an observation $\observ$ is a set of percepts $\observ
-\subseteq \percepts$, and we denote \observations\ the set of
-reachable observations. The underlying state of the world cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in \observations$ that informs what
-should be believed about the world. In the POMDP model, if $\action$
-is applied effecting a transition to a successor state $\state'$, then
-an observation occurs according to the active senses
-$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
-precept is active, written $\stochSense \in \stochSenses(\stochAction,
-\state')$, if the senses action-precondition,
-$\poss_\stochActions(\stochSense)$, is equal to $\stochAction$, and
-the state-precondition $\poss_\states(\stochSense) \subseteq
-\propositions$ is satisfied by the state $\state'$; In the usual sense
-that $\poss_\states(\stochSense) \subseteq \state'$.
+finite set of stochastic {\em senses} $\stochSenses$, deterministic
+sensing outcomes $\detSenses$, and perceptual propositions
+$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
+a set of percepts $\observ \subseteq \percepts$, and we denote
+\observations\ the set of reachable observations. The underlying state
+of the process cannot be observed directly, rather, senses
+$\stochSense \in \stochSenses$ effect an observation $\observ \in
+\observations$ that informs what should be believed about the state a
+process is in. In detail, if $\action$ is applied effecting a
+transition to a successor state $\state'$, then an observation occurs
+according to the active senses $\stochSenses(\stochAction, \state')
+\subseteq \stochSenses$. A percept is active, written $\stochSense \in
+\stochSenses(\stochAction, \state')$, if the senses'
+action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
+$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
+\subseteq \propositions$ is satisfied by the state $\state'$; In the
+usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
 %%
@@
 
-A POMDP has a starting configuration that corresponds to a subjective
-Bayesian belief state. Written $\bstate_0$, this is a probability
-distribution over states. We use the notation $\bstate_0(\state)$ to
-denote the probability that the starting state is equal to
-$\state$. We suppose this is given in a factored tree format.
-
-%% Let $\detSense_i \subseteq \percepts$ for each $i$, so that if
-%% we apply $\stochAction$ and transition to $\state'$, the observation
-%% received will be a union of choices from the active percepts, which
-%% will occur with the probability according to $\psi$ that nature can
-%% make those choices.
+A POMDP has a starting configuration that corresponds to a Bayesian
+belief-state. Intuitively, this is the robot's subjective belief about
+its environment. Formally, a belief-state $\bstate$ is a probability
+distribution over process states. We write $\bstate(\state)$ to denote
+the probability that the process is in $\state$ according to
+$\bstate$, and $\bstate_0$ when discussing the starting
+configuration.
 
@@
 not. We differ
-from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
+from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
 because we forbid the execution of actions that are not applicable in
 any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
+unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
 we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at some \bstate, that it has no
-effect on states $\bstate(\state) > 0$ where $\poss(\stochAction)
-\not\subseteq \state$.
+an action \stochAction\ is executed at $\state_i$ where
+$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
+$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
+0$.
 
@@
 
-Using the notation \bstate\ for a belief-state, and writing
-$\bstate(\state)$ for the probability the robot believes it is in
-state \state, we have that the value of that belief $\bstate$ according to a
-controller is:
+\noindent The value of $\bstate$ according to a controller is
+therefore:
 
@@
 problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04}. PPDDL was used in all 3 of the
-International Planning Competitions since 2004. A variation on PDDL
-for describing domains with stochastic actions and uncertain starting
-configurations, PPDDL is a declarative first-order (a.k.a.,
-relational) that facilitates factored description of domains and
-problem instances. There are straightforward compilations from
-problems expressed compactly in PPDDL to propositional representations
-amenable to state-of-the-art planning procedures.  Because this
-language cannot model domains that feature partial observability, for
-our setting we have developed DTPDDL, an extension of PPDDL to partial
-observability.
+Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
+all 3 of the International Planning Competitions since 2004. A
+variation on PDDL for describing domains with stochastic actions and
+uncertain starting configurations, PPDDL is a declarative first-order
+(a.k.a., relational) language that facilitates factored description of
+domains and problem instances. There are straightforward compilations
+from problems expressed compactly in PPDDL to propositional
+representations amenable to state-of-the-art planning procedures.
+Because this language cannot model domains that feature partial
+observability, for our setting we have developed DTPDDL, an extension
+of PPDDL to partial observability.
 
-For example, $(:\pp{predicates} (\pp{location} ?p - VisualObject))$
+Our mobile robot is able to look for a visual object, for example a
+box of cornflakes, at a given place. We can model that deterministic
+action using the following schema:
+
+\def\oom{$^{\tt +}$}
+\def\zom{$^*$}
+\def\bump{\hspace{1cm}}
+\def\req#1{$^{\tt #1}$}
+\def\noteme#1{}%{[[#1]]}
+\def\notecoauth#1{\ }
+\def\notereader#1{[[#1]]}
+\def\meta{$\uparrow\uparrow$}
+\def\la{\langle}
+\def\ra{\rangle}
+
+\small{
+\begin{tabtt}
+(\= :action look-for-object \\
+  \> :parameters \=(?r - robot ?v - visual-object\\
+  \> \> ?l - location) \\
+  \> :precondition (and (= (is-in ?r) ?l) ) \\
+  \> :effect (and (assign (reward) -3) ) ) \\
+\end{tabtt}
+}
+
+\noindent That is, the robot can always look for a visual object at
+its current location. The only effect this has on the state, is that
+the robot incurs a penalty, $3$, that corresponds to the {\em cost} of
+performing a visual search. Instances of this operator are useful,
+because their application activates an the following sense schema:
+
+\small{
+\begin{tabtt}
+(\= :sense vision \\
+  \> :parameters \=(?r - robot ?v - visual-object\\
+  \> \> ?l - location) \\
+  \> :execution (look-for-object ?r ?v ?l) \\
+  \> :effect ( and (when (= (is-in ?v) ?l) \\
+  \> \bump (probabilistic 0.8 \\
+  \> \bump (= (observe-is-in ?v) ?l))) \\
+  \> (when (not (= (is-in ?v) ?l)) \\
+  \> \bump (probabilistic 0.1 \\
+  \> \bump (= (observe-is-in ?v) ?l))))) \\
+\end{tabtt}
+}
+
+\noindent Here, the {\em execution} clause is a lifted description of
+the sense action-precondition -- I.e,
+$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
+\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. No
+other conditions guard the activation of this sense, therefore if
+$(\pp{look-for-object} \pp{Dora} \pp{cornflakes} \pp{bedroom})$ is
+executed, we suppose there is a $0.8$ chance of observing the
+cornflakes if they are in the $\pp{bedroom}$, and otherwise a $0.1$ of
+observing them.\footnote{We have simplified the schemata for
+illustrative purposes. In practice, the probabilities of a particular
+observation here should be parametrised by the type of visual object
+the robot is searching for, and category of the location.}
+
+\scriptsize
+\begin{nopagebreak}\begin{tabtt}
+<observation-def> \=::= \=(:sense <observation symbol> \\
+                        \> \> :parameters (<typed list (variable)>)  \\
+                        \> \> <o-def body>) \\
+  <o-symbol> \> ::= <name> \\
+  <o-def body> \> ::= [:precondition <GD>] \\
+  \> \> [:execution <atomic action(term)> ] \\
+    \> \> [:effect <o-effect>] \\
+  <atomic action(t)> \> ::= (<action symbol> t\zom) \\
+  <o-effect> \> ::= (and <c-o-effect>\zom) \\
+  <o-effect> \> ::= <c-o-effect> \\
+  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
+  <c-o-effect> \> ::= <p-o-effect> \\
+  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
+  <p-o-effect> \> ::= <atomic o-formula(term)> \\
+  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
+  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
+  <o-f-exp> \> ::= <number>\\
+  <o-f-exp> \> ::= (- <o-f-exp>)\\
+  <o-f-exp> \> ::= <o-f-head>\\
+  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
+  <o-f-head> \> ::= <o-function-symbol>\\
+\end{tabtt}\end{nopagebreak}
+\noteme{<structure-def> ::= <attach-def>}
+\normalsize
+
+We suppose $\bstate_0$ is given in a factored tree format of the
+form:
+
+%% \small{
+%% \begin{tabtt}
+%% (:init  (and \=f_{11} f_{12} ..
+%%  \> (probabilistic p_{}) 
+%%  \> (probabilistic ..) ..) \\
+%% \end{tabtt}
+%% }
+
+%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
+
+\noindent Nodes of the tree have one of three types: (1) atomic nodes
+are labelled with a state proposition, e.g., $(=
+(\pp{location}~\pp{cornflakes}) \pp{kitchen})$, (2) probabilistic
+nodes, have the form $(r_1, C_1;..;r_n, C_n)$, and (3) constructive
+nodes, take the form of a set of probabilistic and atomic nodes. The
+root node is conjunctive, and leaves are always atomic. A convenient
+way to give semantics to the tree, is to treat nodes as if they were
+actions. An atomic node is an action with a positive effect
+corresponding to the node label, .
+
+For example, $(:\pp{predicates} (\pp{location} ?l - VisualObject))$
 says that there is a unary predicate
@@
   \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?p - place) \\
-  \> :precondition (and (= (is-in ?r) ?p) \\
+  \> \> ?l - label ?l - location) \\
+  \> :precondition (and (= (is-in ?r) ?l) \\
   \> \> (= (label ?o) ?l) ) \\
@@
   \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?p - place) \\
-  \> :execution (look-at-object ?r ?v ?l ?p) \\
-  \> :effect (when (= (is-in ?o) ?p) (probabilistic 0.8)) \\
+  \> \> ?l - label ?l - location) \\
+  \> :execution (look-at-object ?r ?v ?l ?l) \\
+  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
 \end{tabtt}
@@
   \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?p - place) \\
+  \> :parameters (?l - label ?l - location) \\
   \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?p) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?p) ) \\
+  \> :precondition (\=and (= (is-in ?r) ?l) \\
+  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
   \> :effect () \\
-  \> :sense (= (is-in ?o) ?p) \\
+  \> :sense (= (is-in ?o) ?l) \\
 \end{tabtt}

### 2010-11-18 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Rough outline for the switching planner part

@@
 
+\section{Switching Planner Outline}
+
+Outline the high level approach here:
+\begin{itemize}
+\item Problem input (factored initial state, DTPDDL observations)
+\item Determinisation
+\item Problem generation for DT
+\item Belief revision
+\end{itemize}
+
+\section{Determinisation}
+
+\begin{itemize}
+\item Generation of deterministic sensing models
+\item Determinisation of the initial state
+\item Compiling reward model into (soft-)goal model
+\item Cost function in FD (maybe this could mentioned in the outline?)
+\end{itemize}
+
+\section{Subtask Generation}
+\begin{itemize}
+\item Goal generation from CP trace
+\item Disconfirm actions
+\item State pruning according to CP trace
+\end{itemize}
+
+\section{Belief revision}
+Not sure if this needs its own section
+\begin{itemize}
+\item Convert inital state into bayesian network
+\item Belief updates with observation nodes
+\item Create new network for next planning runs
+\end{itemize}
+
 \section{Propositional Decision-Theoretic Planning}
@@
 not. We differ
-from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
+from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
 because we forbid the execution of actions that are not applicable in
 any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
+unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
 we do incorporate a PPDDL-like default semantics, supposing that when
@@
 uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) that facilitates factored description of domains
-and problem instances. There are straightforward compilations from
-problems expressed compactly in PPDDL to propositional representations
-amenable to state-of-the-art planning procedures.  Because this
-language cannot model domains that feature partial observability, for
-our setting we have developed DTPDDL, an extension of PPDDL to partial
-observability.
+(a.k.a., relational) language that facilitates factored description of
+domains and problem instances. There are straightforward compilations
+from problems expressed compactly in PPDDL to propositional
+representations amenable to state-of-the-art planning procedures.
+Because this language cannot model domains that feature partial
+observability, for our setting we have developed DTPDDL, an extension
+of PPDDL to partial observability.
 
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
+Our mobile robot is able to look for a visual object, for example a
+box of cornflakes, at a given place. We can model that deterministic
+action using the following schema:
 
@@
 
+\small{
+\begin{tabtt}
+(\= :action look-for-object \\
+  \> :parameters \=(?r - robot ?v - visual-object\\
+  \> \> ?l - location) \\
+  \> :precondition (and (= (is-in ?r) ?l) ) \\
+  \> :effect (and (assign (reward) -3) ) ) \\
+\end{tabtt}
+}
+
+\noindent That is, the robot can always look for a visual object at
+its current location. The only effect this has on the state, is that
+the robot incurs a penalty, $3$, that corresponds to the {\em cost} of
+performing a visual search. Instances of this operator are useful,
+because their application activates an the following sense schema:
+
+\small{
+\begin{tabtt}
+(\= :sense vision \\
+  \> :parameters \=(?r - robot ?v - visual-object\\
+  \> \> ?l - location) \\
+  \> :execution (look-for-object ?r ?v ?l) \\
+  \> :effect ( and (when (= (is-in ?v) ?l) \\
+  \> \bump (probabilistic 0.8 \\
+  \> \bump (= (observe-is-in ?v) ?l))) \\
+  \> (when (not (= (is-in ?v) ?l)) \\
+  \> \bump (probabilistic 0.1 \\
+  \> \bump (= (observe-is-in ?v) ?l))))) \\
+\end{tabtt}
+}
+
+\noindent Here, the {\em execution} clause is a lifted description of
+the sense action-precondition -- I.e,
+$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
+\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. No
+other conditions guard the activation of this sense, therefore if
+$(\pp{look-for-object} \pp{Dora} \pp{cornflakes} \pp{bedroom})$ is
+executed, we suppose there is a $0.8$ chance of observing the
+cornflakes if they are in the $\pp{bedroom}$, and otherwise a $0.1$ of
+observing them.\footnote{We have simplified the schemata for
+illustrative purposes. In practice, the probabilities of a particular
+observation here should be parametrised by the type of visual object
+the robot is searching for, and category of the location.}
+
 \scriptsize
 \begin{nopagebreak}\begin{tabtt}
-\bump \= <observation-def> \bump\=::= \=(:obs\=erve <o-symbol> \+\+\+\+\\
-                        %%{[}<extension-def>]  \\
-                        :parameters (<typed list (variable)>)  \\
-                        <o-def body>) \-\-\- \\
-  <o-def body> \> ::= [:state <emptyOr (pre-GD)>] \\
-  \> \> [:action <atomic action(term)> ] \\
-    \> \> [:effect <emptyOr (o-effect)>] \\
+<observation-def> \=::= \=(:sense <observation symbol> \\
+                        \> \> :parameters (<typed list (variable)>)  \\
+                        \> \> <o-def body>) \\
   <o-symbol> \> ::= <name> \\
-  <atomic action(t)> \> ::= (<action-symbol> t\zom) \\
+  <o-def body> \> ::= [:precondition <GD>] \\
+  \> \> [:execution <atomic action(term)> ] \\
+    \> \> [:effect <o-effect>] \\
+  <atomic action(t)> \> ::= (<action symbol> t\zom) \\
   <o-effect> \> ::= (and <c-o-effect>\zom) \\
   <o-effect> \> ::= <c-o-effect> \\
-  <c-o-effect> \> ::= \req{:conditional-effects} (when <GD>
-                        <o-effect>) \\  
-  <c-o-effect> \> ::= \req{:universal-effects} (forall (<typed list (var)>) <o-effect>) \\
-  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic
-  <prob> <o-effect>) \\
-  <c-o-effect> \> ::=  \req{:probabilistic-effects}\req{:universal-unwinding} \\
-%% 
-\hspace{3ex}(probabilistic (for-each (<typed list (x)>) <s-f-head>  <o-effect> )) \\
+  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
   <c-o-effect> \> ::= <p-o-effect> \\
@@
   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-  <p-o-effect> \> ::= \req{:fluents}(<assign-op> <o-f-head> <o-f-exp>) \\
   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
   <o-f-exp> \> ::= <number>\\
-  <o-f-exp> \> ::= (<binary-op>  <o-f-exp> <o-f-exp>)\\
-  <o-f-exp> \> ::= (<multi-op>  <o-f-exp> <o-f-exp>\oom)\\
   <o-f-exp> \> ::= (- <o-f-exp>)\\
@@
 \normalsize
+
+We suppose $\bstate_0$ is given in a factored tree format of the
+form:
 
@@
 
-For example, $(:\pp{predicates} (\pp{location} ?p - VisualObject))$
+For example, $(:\pp{predicates} (\pp{location} ?l - VisualObject))$
 says that there is a unary predicate

### 2010-11-18 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Writers block is aweful on this one. I think it is somewhat because I've never written an applications paper before. Hope I can get over it soon...
 - Charles

@@
 
-The problem of practical robot control poses many important challenges
-to automated planning.
+From an automated planning perspective, the problem of practical
+mobile robot control poses important and contrary challenges.
+%%
+On the one hand, planning and execution monitoring must be
+lightweight, robust, timely, and should span the lifetime of the
+robot. Those processes must seamlessly accommodate exogenous events,
+changing objectives, and the underlying unpredictability of the
+environment.
+%%
+On the other hand, robot planning should perform computationally
+expensive reasoning about contingencies, and possible revisions of
+subjective belief according to quantitatively modelled uncertainty in
+acting and sensing.
 
-Planning and plan execution must be robust, seamlessly accommodating
-exogenous events and the underlying unpredictability of the
-environment.
+In this paper we address these challenges, developing a continual
+planner that switches between sequential and decision-theoretic
+planning. Given a POMDP model of the environment, serial planning is
+responsible for an initial deterministic sequential plan and
+complementary runtime evolution of the decision process. That plan is
+executed until an assumption about a runtime proposition
 
-Plans with a reasonable chance of exhibiting good behaviours must be
-synthesised in a timely manner.
+posted by a motivational component of the underlying robotic
+architecture.
+
+contingent sensory plans that are tailored to current
+objectives.
+
+In this paper we present an approach to continual planing that uses
+two planning systems. The first
+
+ to a distinct class of
+challenges. We suppose 
+%%
+The underlying environment is modelled as a POMDP. We use the fast
+classical satisficing system FastDownward to find a deterministic
+sequential plan and complementary runtime evolution of that
+process. This corresponds to a generalisation of replanning in
+probabilistic planning to problems with partial observability.
+
+Interaction between the serial planner and execution proceeds
+more-or-less analogously to popular replanning approaches
+
+Addressing these challenges in a monolithic framework, we present a
+{\em switching} continual planner, that uses the fast serial
+satisficing procedure FastDownward to perform net-benefit
+%%
+makes reasonable assumptions about the evolution of the runtime state
+given a POMDP model of the environment.
+
+contingent sensory plans that are tailored to current
+objectives.
+
+The decision-theoretic planner is able to tailor sensory processing on
+a robot platform to the current objective, while FastDownward  quickly
+
+In this paper we develop a continual planing system that uses two
+planning systems. The first, is a state-of-the-art domain independent
+planner for deterministic problems. The second is a information-state
+contingency planning the information-state space of
+
+Continual planning is a powerful technique that goes some way to
+addressing those challenges. That approach interleaves planning and
+execution, deliberately postponing planing for contingencies unless
+they eventuate during execution.
+
+computing a single serial plan and
+eventuality
+
+To these challenges it seems a continual planning approach is best,
+
+where the runtime state evolves during plan execution
 
@@
 
+the latter is able to tailor sensory processing on a robot platform,
+in order that it.
+
+{\em ad-hoc}
+
 , reason about degrees of belief and uncertainty about the world,
+
+probabilistic sequential decision making in practical sized problems
+is intractable.
+
+quantitative probabilistic models of the perception and action .
 
@@
 finite-state controllers.
+
+\section{Switching Planner Outline}
+
+Outline the high level approach here:
+\begin{itemize}
+\item Problem input (factored initial state, DTPDDL observations)
+\item Determinisation
+\item Problem generation for DT
+\item Belief revision
+\end{itemize}
+
+\section{Determinisation}
+
+\begin{itemize}
+\item Generation of deterministic sensing models
+\item Determinisation of the initial state
+\item Compiling reward model into (soft-)goal model
+\item Cost function in FD (maybe this could mentioned in the outline?)
+\end{itemize}
+
+\section{Subtask Generation}
+\begin{itemize}
+\item Goal generation from CP trace
+\item Disconfirm actions
+\item State pruning according to CP trace
+\end{itemize}
+
+\section{Belief revision}
+Not sure if this needs its own section
+\begin{itemize}
+\item Convert inital state into bayesian network
+\item Belief updates with observation nodes
+\item Create new network for next planning runs
+\end{itemize}
 

### 2010-11-19 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
New abstract from Richard.
 -0 Charles

@@
 %\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Continual Planner that Switches Between Serial and
-  Decision-Theoretic Planning} 
+\title{A Switching Planner for Combined Task and Observation Planning}
+%\title{A Continual Planner that Switches Between Serial and
+%  Decision-Theoretic Planning} 
 \author{NA}
@@
 
-The problem of practical robot control poses many important challenges
-to automated planning.
+Realistic robot planning problems in uncertain environments often
+require achieving tasks while also finding out about the
+world. Because the world state cannot be observed directly, these
+problems are naturally represented as partially observable Markov
+decision problems (POMDPs). However, these are typically intractable
+for realistic problems.
 
-Planning and plan execution must be robust, seamlessly accommodating
-exogenous events and the underlying unpredictability of the
-environment.
-
-Plans with a reasonable chance of exhibiting good behaviours must be
-synthesised in a timely manner.
-
-interleaved planning and execution.
-
-, reason about degrees of belief and uncertainty about the world,
+We present a \emph{switching planner} that employs fast sequential
+planning to decide on the overall strategy, and uses a
+decision-theoretic planner to solve the subproblems where partial
+observability will significantly impact the quality of the plan. We
+demonstrate the approach in a realistic robot exploration domain.
 
@@
 
+From an automated planning perspective, the problem of practical
+mobile robot control poses important and contrary challenges.
+%%
+On the one hand, planning and execution monitoring must be
+lightweight, robust, timely, and should span the lifetime of the
+robot. Those processes must seamlessly accommodate exogenous events,
+changing objectives, and the underlying unpredictability of the
+environment.
+%%
+On the other hand, robot planning should perform computationally
+expensive reasoning about contingencies, and possible revisions of
+subjective belief according to quantitatively modelled uncertainty in
+acting and sensing.
+
+In this paper we address these challenges, developing a continual
+planner that switches between sequential and decision-theoretic
+planning. Given a POMDP model of the environment, serial planning is
+responsible for an initial deterministic sequential plan and
+complementary runtime evolution of the decision process. That plan is
+executed until an assumption about a runtime proposition
+
+posted by a motivational component of the underlying robotic
+architecture.
+
+contingent sensory plans that are tailored to current
+objectives.
+
+In this paper we present an approach to continual planing that uses
+two planning systems. The first
+
+ to a distinct class of
+challenges. We suppose 
+%%
+The underlying environment is modelled as a POMDP. We use the fast
+classical satisficing system FastDownward to find a deterministic
+sequential plan and complementary runtime evolution of that
+process. This corresponds to a generalisation of replanning in
+probabilistic planning to problems with partial observability.
+
+Interaction between the serial planner and execution proceeds
+more-or-less analogously to popular replanning approaches
+
+Addressing these challenges in a monolithic framework, we present a
+{\em switching} continual planner, that uses the fast serial
+satisficing procedure FastDownward to perform net-benefit
+%%
+makes reasonable assumptions about the evolution of the runtime state
+given a POMDP model of the environment.
+
+contingent sensory plans that are tailored to current
+objectives.
+
+The decision-theoretic planner is able to tailor sensory processing on
+a robot platform to the current objective, while FastDownward  quickly
+
+In this paper we develop a continual planing system that uses two
+planning systems. The first, is a state-of-the-art domain independent
+planner for deterministic problems. The second is a information-state
+contingency planning the information-state space of
+
+Continual planning is a powerful technique that goes some way to
+addressing those challenges. That approach interleaves planning and
+execution, deliberately postponing planing for contingencies unless
+they eventuate during execution.
+
+computing a single serial plan and
+eventuality
+
+To these challenges it seems a continual planning approach is best,
+
+where the runtime state evolves during plan execution
+
+interleaved planning and execution.
+
+the latter is able to tailor sensory processing on a robot platform,
+in order that it.
+
+{\em ad-hoc}
+
+, reason about degrees of belief and uncertainty about the world,
+
+probabilistic sequential decision making in practical sized problems
+is intractable.
+
+quantitative probabilistic models of the perception and action .
+
 \bibliographystyle{aaai}

### 2010-11-19 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Want to remove missleading text I had in the introduction.
 - Charles

@@
 accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL. We have
-an automated determinisation procedure that generates MAPL
-descriptions given DTPDDL input. To accompany that, we have extended
-MAPSIM to simulate DTPDDL instances, and modified \fastdownward\ so
-that it can find useful serial executions given DTPDDL and MAPL models
-of the problem at hand.
+\laostar\ procedure for solving problems expressed in DTPDDL.
+
+We have extended MAPSIM to simulate DTPDDL instances, and modified
+\fastdownward\ so that it can find useful serial executions given
+DTPDDL models of the problem at hand.
 

### 2010-11-19 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
New bibliography, and reference to Andrew Ng. work with Mike Jordan...

@@
 \section{Introduction}
+
+Any (PO)MDP with stochastic actions can be compiled into an equivalent
+deterministic-action POMDP with all the action uncertainty expressed
+in the starting-state distribution~\cite{ng:Jordan:2000}. In our
+setting, that of finite-horizon planning, this compilation yields a
+finite flat representation of the POMDP. Therefore, although our
+current implementation does not explicitly support it, our approach
+generalises to POMDPs with stochastic action models.
 
@@
 %%
-Recently the system
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006}, demonstrated how
-conformant planning ---i.e., serial planning in unobservable worlds---
-can be modelled as a deterministic problem, and therefore solved using
-serial systems. In this conformant setting, advances have been towards
-compact representations and lazy evaluations of belief-states.
+Recently the systems
+\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
+planning ---i.e., serial planning in unobservable worlds--- can be
+modelled as a deterministic problem, and therefore solved using serial
+systems. In this conformant setting, advances have been towards
+compact representations and lazy evaluations of beliefs.
 

### 2010-11-19 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Completed the Hoffman+palacios thread... I could not bring myself to move forward on this until I had, as in my mind it is _hugely_ related
 - Charles

@@
 systems. In this conformant setting, advances have been towards
-compact representations and lazy evaluations of beliefs.
+compact representations of beliefs amenable to existing best-first
+search planning procedures, and lazy evaluations of beliefs. This
+research thread has also been extended recently to contingent planning
+in fully observable non-deterministic
+environments~\cite{albore:etal:2009}.
 

### 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
abstract
More DTPDDL

@@
 
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
+Realistic robot planning problems in uncertain environments often
+require achieving tasks while also finding out about the
+world. Because the world state cannot be observed directly, these
+problems are naturally represented as partially observable Markov
+decision problems (POMDPs). However, these are typically intractable
+for realistic problems.
 
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, serial planning is
-responsible for an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until an assumption about a runtime proposition
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the serial planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast serial
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planing for contingencies unless
-they eventuate during execution.
-
-computing a single serial plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
+We present a \emph{switching planner} that employs fast sequential
+planning to decide on the overall strategy, and uses a
+decision-theoretic planner to solve the subproblems where partial
+observability will significantly impact the quality of the plan. We
+demonstrate the approach in a realistic robot exploration domain.
 
@@
 representations amenable to state-of-the-art planning procedures.
-Because this language cannot model domains that feature partial
-observability, for our setting we have developed DTPDDL, an extension
-of PPDDL to partial observability.
 
-Our mobile robot is able to look for a visual object, for example a
-box of cornflakes, at a given place. We can model that deterministic
-action using the following schema:
-
-\def\oom{$^{\tt +}$}
-\def\zom{$^*$}
-\def\bump{\hspace{1cm}}
-\def\req#1{$^{\tt #1}$}
-\def\noteme#1{}%{[[#1]]}
-\def\notecoauth#1{\ }
-\def\notereader#1{[[#1]]}
-\def\meta{$\uparrow\uparrow$}
-\def\la{\langle}
-\def\ra{\rangle}
+Because PPDDL cannot model domains that feature partial observability,
+we developed an extension of that language, called DTPDDL. This can
+express probabilistic models of the sensing consequences of acting to
+quantitatively capture unreliability in perception. That expressive
+power is achieved by incorporating perceptual analogues of fluent and
+predicate signatures, and of action definitions. In detail, we have
+declarations of state characterising predicate and fluent symbols as
+usual. In addition, we allow two other declarative blocks, for
+perceptual predicates and fluents respectively. In detail, suppose
+our robot is tasked with exploring {\em locations} in order to
+identify the whereabouts of a {\em visual-object}. We must describe
+state and perceptual facts that model the {\em true}, resp. perceived,
+locations of objects. In DTPDDL, these declarations appear as:
 
@@
 \begin{tabtt}
-(\= :action look-for-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
+(:\=functions  ;; state fluents\\
+  \> (is-in ?v - visual-object) - location )\\
+(:observable-functions  ;; perceptual fluents\\
+  \> (o-is-in ?v - visual-object) - location )
 \end{tabtt}
@@
 
-\noindent That is, the robot can always look for a visual object at
-its current location. The only effect this has on the state, is that
-the robot incurs a penalty, $3$, that corresponds to the {\em cost} of
-performing a visual search. Instances of this operator are useful,
-because their application activates an the following sense schema:
+\noindent To model the sensing capabilities of the agent, we have
+operator-like descriptive elements, with preconditions in both action
+and state symbols, and uniformly positive effects over perceptual
+predicates -- I.e, the negation normal form of a sensing effect
+formula cannot contain negation. Moreover, and without a loss of
+generality, we find it convenient to separate the actional
+precondition from the state precondition. For example, our mobile
+robot is able to look for a visual object, such as a box of
+cornflakes, at a given place. We can model that deterministic action
+using the following operator schema:
 
@@
 \begin{tabtt}
-(\= :sense vision \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - location) \\
-  \> :execution (look-for-object ?r ?v ?l) \\
-  \> :effect ( and (when (= (is-in ?v) ?l) \\
-  \> \bump (probabilistic 0.8 \\
-  \> \bump (= (observe-is-in ?v) ?l))) \\
+(\=:action look-for-object \+ \\
+   :parameters (\=?r - robot ?v - visual-object\\
+   \> ?l - location) \\
+   :precondition (and (= (is-in ?r) ?l) ) \\
+   :effect (and (assign (reward) -3) ) ) \\
+\end{tabtt}
+}
+
+\noindent In a slight departure from PPDDL, we suppose that
+$\pp{look-for-object}$ has an effect effect on the state. That is,
+execution incurs a penalty, $3$, corresponding to the {\em cost} of
+performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
+reserved word}, and occurs in {\em increased} and {\em decreased}
+clauses of operator schemata. In that setting, {\em reward} is
+interpreted as accumulated rewards and penalties.}
+
+Completing the example, we model the sensing outcome that results from
+instantiating the $\pp{look-for-object}$ operator:
+
+\small{
+\begin{tabtt}
+(\= :sense vision \+\\
+ :parameters \= (\= ?r - robot ?v - visual-object\\
+ \>\>  ?l - location) \\
+ :execution \> ( \> look-for-object ?r ?v ?l) \\
+ :precondition (and (= (is-in ?r) ?l) ) \\
+ :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
+   \> \> (probabilistic 0.8 \\
+   \>  \>(= (observe-is-in ?v) ?l))) \\
   \> (when (not (= (is-in ?v) ?l)) \\
-  \> \bump (probabilistic 0.1 \\
-  \> \bump (= (observe-is-in ?v) ?l))))) \\
+   \>  \> (probabilistic 0.1 \\
+   \>  \> (= (observe-is-in ?v) ?l))))) \\
 \end{tabtt}
@@
 $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. No
-other conditions guard the activation of this sense, therefore if
-$(\pp{look-for-object} \pp{Dora} \pp{cornflakes} \pp{bedroom})$ is
-executed, we suppose there is a $0.8$ chance of observing the
-cornflakes if they are in the $\pp{bedroom}$, and otherwise a $0.1$ of
-observing them.\footnote{We have simplified the schemata for
-illustrative purposes. In practice, the probabilities of a particular
-observation here should be parametrised by the type of visual object
-the robot is searching for, and category of the location.}
+\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. We
+also include a redundant state precondition, $(=(\pp{is-in}~?r)?l)$,
+in order to fully demonstrate the syntax. Interpreting the schema, if
+action
+$(\pp{look-for-object}~\pp{robot}~\pp{cornflakes}~\pp{bedroom})$ is
+executed, there is a $0.8$ chance of perceiving the cornflakes if they
+are in the $\pp{bedroom}$, and otherwise a $0.1$ of perceiving them.
 
-\scriptsize
-\begin{nopagebreak}\begin{tabtt}
-<observation-def> \=::= \=(:sense <observation symbol> \\
-                        \> \> :parameters (<typed list (variable)>)  \\
-                        \> \> <o-def body>) \\
-  <o-symbol> \> ::= <name> \\
-  <o-def body> \> ::= [:precondition <GD>] \\
-  \> \> [:execution <atomic action(term)> ] \\
-    \> \> [:effect <o-effect>] \\
-  <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-  <o-effect> \> ::= (and <c-o-effect>\zom) \\
-  <o-effect> \> ::= <c-o-effect> \\
-  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-  <c-o-effect> \> ::= <p-o-effect> \\
-  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-  <p-o-effect> \> ::= <atomic o-formula(term)> \\
-  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-  <o-f-exp> \> ::= <number>\\
-  <o-f-exp> \> ::= (- <o-f-exp>)\\
-  <o-f-exp> \> ::= <o-f-head>\\
-  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-  <o-f-head> \> ::= <o-function-symbol>\\
-\end{tabtt}\end{nopagebreak}
-\noteme{<structure-def> ::= <attach-def>}
-\normalsize
+%% \footnote{We have simplified the schemata for illustrative
+%% purposes. In practice, the probabilities of a particular observation
+%% here should be parametrised by the type of visual object the robot is
+%% searching for, and category of the location.}
 
@@
 
+planning in practical sized POMDPs is intractable. One approach is to
+described the task hierarchically, thereby constrain and simplify the
+policy search task. The approach ~\cite{}.
+
 \noindent Nodes of the tree have one of three types: (1) atomic nodes
@@
 corresponding to the node label, .
-
-For example, $(:\pp{predicates} (\pp{location} ?l - VisualObject))$
-says that there is a unary predicate
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-knowledge objectives and  sensory effects.
-
-probabilistic models of the sensing consequences of acting to
-quantitatively capture the unreliability of sensing, and thus reason
-precisely about the utility of a plan.
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
 
@@
 %%
-Recently the system
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006}, demonstrated how
-conformant planning ---i.e., serial planning in unobservable worlds---
-can be modelled as a deterministic problem, and therefore solved using
-serial systems. In this conformant setting, advances have been towards
-compact representations and lazy evaluations of belief-states.
+Recently the systems
+\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
+planning ---i.e., serial planning in unobservable worlds--- can be
+modelled as a deterministic problem, and therefore solved using serial
+systems. In this conformant setting, advances have been towards
+compact representations of beliefs amenable to existing best-first
+search planning procedures, and lazy evaluations of beliefs. This
+research thread has also been extended recently to contingent planning
+in fully observable non-deterministic
+environments~\cite{albore:etal:2009}.
 
@@
 the underlying planner for CogX.
+
+the effects of a {\em sense} schema are perceptual, whereas the
+effects of an {\em operator} schema are over state propositions.
+
+From an automated planning perspective, the problem of practical
+mobile robot control poses important and contrary challenges.
+%%
+On the one hand, planning and execution monitoring must be
+lightweight, robust, timely, and should span the lifetime of the
+robot. Those processes must seamlessly accommodate exogenous events,
+changing objectives, and the underlying unpredictability of the
+environment.
+%%
+On the other hand, robot planning should perform computationally
+expensive reasoning about contingencies, and possible revisions of
+subjective belief according to quantitatively modelled uncertainty in
+acting and sensing.
+
+In this paper we address these challenges, developing a continual
+planner that switches between sequential and decision-theoretic
+planning. Given a POMDP model of the environment, serial planning is
+responsible for an initial deterministic sequential plan and
+complementary runtime evolution of the decision process. That plan is
+executed until an assumption about a runtime proposition
+
+posted by a motivational component of the underlying robotic
+architecture.
+
+contingent sensory plans that are tailored to current
+objectives.
+
+In this paper we present an approach to continual planing that uses
+two planning systems. The first
+
+ to a distinct class of
+challenges. We suppose 
+%%
+The underlying environment is modelled as a POMDP. We use the fast
+classical satisficing system FastDownward to find a deterministic
+sequential plan and complementary runtime evolution of that
+process. This corresponds to a generalisation of replanning in
+probabilistic planning to problems with partial observability.
+
+Interaction between the serial planner and execution proceeds
+more-or-less analogously to popular replanning approaches
+
+Addressing these challenges in a monolithic framework, we present a
+{\em switching} continual planner, that uses the fast serial
+satisficing procedure FastDownward to perform net-benefit
+%%
+makes reasonable assumptions about the evolution of the runtime state
+given a POMDP model of the environment.
+
+contingent sensory plans that are tailored to current
+objectives.
+
+The decision-theoretic planner is able to tailor sensory processing on
+a robot platform to the current objective, while FastDownward  quickly
+
+In this paper we develop a continual planing system that uses two
+planning systems. The first, is a state-of-the-art domain independent
+planner for deterministic problems. The second is a information-state
+contingency planning the information-state space of
+
+Continual planning is a powerful technique that goes some way to
+addressing those challenges. That approach interleaves planning and
+execution, deliberately postponing planing for contingencies unless
+they eventuate during execution.
+
+computing a single serial plan and
+eventuality
+
+To these challenges it seems a continual planning approach is best,
+
+where the runtime state evolves during plan execution
+
+interleaved planning and execution.
+
+the latter is able to tailor sensory processing on a robot platform,
+in order that it.
+
+{\em ad-hoc}
+
+, reason about degrees of belief and uncertainty about the world,
+
+probabilistic sequential decision making in practical sized problems
+is intractable.
+
+quantitative probabilistic models of the perception and action .
 
@@
 \subsection{MAPL}
+
+%% \scriptsize
+%% \begin{nopagebreak}\begin{tabtt}
+%% <observation-def> \=::= \=(:sense <observation symbol> \\
+%%                         \> \> :parameters (<typed list (variable)>)  \\
+%%                         \> \> <o-def body>) \\
+%%   <o-symbol> \> ::= <name> \\
+%%   <o-def body> \> ::= [:precondition <GD>] \\
+%%   \> \> [:execution <atomic action(term)> ] \\
+%%     \> \> [:effect <o-effect>] \\
+%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
+%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
+%%   <o-effect> \> ::= <c-o-effect> \\
+%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
+%%   <c-o-effect> \> ::= <p-o-effect> \\
+%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
+%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
+%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
+%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
+%%   <o-f-exp> \> ::= <number>\\
+%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
+%%   <o-f-exp> \> ::= <o-f-head>\\
+%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
+%%   <o-f-head> \> ::= <o-function-symbol>\\
+%% \end{tabtt}\end{nopagebreak}
+%% \noteme{<structure-def> ::= <attach-def>}
+%% \normalsize

### 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Initial state

@@
 we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting to
+express probabilistic models of the sensing consequences of acting, to
 quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent and
-predicate signatures, and of action definitions. In detail, we have
-declarations of state characterising predicate and fluent symbols as
-usual. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. In detail, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
+power is achieved by incorporating perceptual analogues of fluent,
+predicate, and action definitions. In detail, we have declarations of
+state characterising predicate and fluent symbols according to PPDDL
+syntax. In addition, we allow two other declarative blocks, for
+perceptual predicates and fluents respectively. For example, suppose our
+robot is tasked with exploring {\em locations} in order to identify
+the whereabouts of a {\em visual-object}. We must describe state and
+perceptual facts that model the {\em true}, resp. perceived, locations
+of objects. In DTPDDL, these declarations appear as:
 
@@
 \begin{tabtt}
-(:\=functions  ;; state fluents\\
+(\=:functions  ;; state fluents\\
   \> (is-in ?v - visual-object) - location )\\
@@
 predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Moreover, and without a loss of
-generality, we find it convenient to separate the actional
-precondition from the state precondition. For example, our mobile
-robot is able to look for a visual object, such as a box of
-cornflakes, at a given place. We can model that deterministic action
-using the following operator schema:
+formula cannot contain negation. Whereas the effects of an {\em
+operator} schema are over state propositions, and describe how the
+states change under application of actions, the effects of a {\em
+sense} schema are perceptual, and describe the composition of an
+observation following the execution of an action. Also, and without a
+loss of generality, we find it convenient to separate the actional
+precondition from the state precondition. For example, our robot is
+able to look for a visual object, such as a box of cornflakes, at a
+given place. We can model that deterministic action using the
+following operator schema:
 
@@
 \noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect effect on the state. That is,
+$\pp{look-for-object}$ has an effect on the state. That is, its
 execution incurs a penalty, $3$, corresponding to the {\em cost} of
@@
 clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated rewards and penalties.}
+interpreted as accumulated (resp. instantaneous) rewards and
+penalties.}
 
@@
    \> \> (probabilistic 0.8 \\
-   \>  \>(= (observe-is-in ?v) ?l))) \\
+   \>  \>(= (o-is-in ?v) ?l))) \\
   \> (when (not (= (is-in ?v) ?l)) \\
    \>  \> (probabilistic 0.1 \\
-   \>  \> (= (observe-is-in ?v) ?l))))) \\
+   \>  \> (= (o-is-in ?v) ?l))))) \\
 \end{tabtt}
@@
 the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. We
-also include a redundant state precondition, $(=(\pp{is-in}~?r)?l)$,
-in order to fully demonstrate the syntax. Interpreting the schema, if
-action
-$(\pp{look-for-object}~\pp{robot}~\pp{cornflakes}~\pp{bedroom})$ is
-executed, there is a $0.8$ chance of perceiving the cornflakes if they
-are in the $\pp{bedroom}$, and otherwise a $0.1$ of perceiving them.
+$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
+\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
+we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
+order to fully demonstrate the syntax. Interpreting the above schema,
+if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
+executed, there is a $0.8$ chance of perceiving the box if it is in
+the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
 
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
+The syntax and semantics for describing initial state distributions in
+DTPDDL is taken verbatim from PPDDL. It is useful to present that
+factored tree-like structure here, in order to aid our discussion of
+sequential planning in POMDP determinisations. The distribution is
+expressed in a tree-like structure of terms. Each term is either: (1)
+atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
+probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$, or
+(3) a conjunct over probabilistic and atomic terms. The root term is
+conjunctive, and the leaves are atomic. For example, the starting
+distribution for $\pp{R2D2}$ might be described thusly:
 
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
+\small{
+\begin{tabtt}
+(\=:init (= (is-in R2D2) kitchen) \+ \\
+       (probabilistic \=.8 (= (is-in box) office)  \\
+		      \>.2 (= (is-in box) kitchen)) \\
+       (probabilistic .3 (= (is-in cup) office)  \\
+		      \>.7 (= (is-in cup) kitchen))) \\
+\end{tabtt}
+}
 
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
-
-\noindent Nodes of the tree have one of three types: (1) atomic nodes
-are labelled with a state proposition, e.g., $(=
-(\pp{location}~\pp{cornflakes}) \pp{kitchen})$, (2) probabilistic
-nodes, have the form $(r_1, C_1;..;r_n, C_n)$, and (3) constructive
-nodes, take the form of a set of probabilistic and atomic nodes. The
-root node is conjunctive, and leaves are always atomic. A convenient
-way to give semantics to the tree, is to treat nodes as if they were
-actions. An atomic node is an action with a positive effect
-corresponding to the node label, .
+\noindent The interpretation of such an expression can be given
+according to a recursive descent.
 
@@
 %% \normalsize
+
+%% \footnote{We have simplified the schemata for illustrative
+%% purposes. In practice, the probabilities of a particular observation
+%% here should be parametrised by the type of visual object the robot is
+%% searching for, and category of the location.}
+
+We suppose $\bstate_0$ is given in a factored tree format of the
+form:
+
+%% \small{
+%% \begin{tabtt}
+%% (:init  (and \=f_{11} f_{12} ..
+%%  \> (probabilistic p_{}) 
+%%  \> (probabilistic ..) ..) \\
+%% \end{tabtt}
+%% }
+
+%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
+
+planning in practical sized POMDPs is intractable. One approach is to
+described the task hierarchically, thereby constrain and simplify the
+policy search task. The approach ~\cite{}.

### 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
initial state

@@
 atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$, or
-(3) a conjunct over probabilistic and atomic terms. The root term is
-conjunctive, and the leaves are atomic. For example, the starting
-distribution for $\pp{R2D2}$ might be described thusly:
+probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
+where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
+atomic terms. The root term is always conjunctive, and the leaves are
+atomic. For example, the starting distribution for $\pp{R2D2}$ might
+be described thusly:
 
@@
 \noindent The interpretation of such an expression can be given
-according to a recursive descent.
+according to a recursive descent. Here, an atom is visited as soon as
+its conjunctive parent is visited, and a conjunctive term is visited
+if all its immediate subterms are visited. A probabilistic term is
+visited if exactly one of its subterms, $T_i$, is visited. Each valid
+visitation of the root term according to this recursive definition
+encapsulates a starting state, along with the probability it
+occurs. The former corresponds to the union of all visited atoms, and
+the latter corresponds to the product of $p_i$ entries on the visited
+subterms of probabilistic elements. Making these ideas concrete, our
+example yields the following flat state distribution:
+
+\begin{tabular}{cccc}
+\hline
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+\hline
+.24 & kitchen & office & office \\
+.06 & kitchen & kitchen & office \\
+.56 & kitchen & office & kitchen \\
+.14 & kitchen & kitchen & kitchen \\
+\hline
+\end{tabular}
 

### 2010-11-19 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex

@@
 
+\section{Moritz's Initial Outline}
+
+\subsection{Switching Planner Outline}
+
+Outline the high level approach here:
+\begin{itemize}
+\item Problem input (factored initial state, DTPDDL observations)
+\item Determinisation
+\item Problem generation for DT
+\item Belief revision
+\end{itemize}
+
+\subsection{Determinisation}
+
+\begin{itemize}
+\item Generation of deterministic sensing models
+\item Determinisation of the initial state
+\item Compiling reward model into (soft-)goal model
+\item Cost function in FD (maybe this could mentioned in the outline?)
+\end{itemize}
+
+\subsection{Subtask Generation}
+\begin{itemize}
+\item Goal generation from CP trace
+\item Disconfirm actions
+\item State pruning according to CP trace
+\end{itemize}
+
+\subsection{Belief revision}
+Not sure if this needs its own section
+\begin{itemize}
+\item Convert inital state into bayesian network
+\item Belief updates with observation nodes
+\item Create new network for next planning runs
+\end{itemize}
+
 \section{Introduction}
 
-Any (PO)MDP with stochastic actions can be compiled into an equivalent
+Any POMDP with stochastic actions can be compiled into an equivalent
 deterministic-action POMDP with all the action uncertainty expressed
@@
 finite-state controllers.
-
-\section{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\section{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\section{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\section{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
 
@@
 
-\section{Serial Planning in Determinisations}
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes}--- are treated as control-knowledge
-by the planner. Here, a plan is not considered valid unless in the
-finial stet the goal condition is satisfied.
-
 \section{Decision-Theoretic PDDL}
@@
 representations amenable to state-of-the-art planning procedures.
-Because this language cannot model domains that feature partial
-observability, for our setting we have developed DTPDDL, an extension
-of PPDDL to partial observability.
 
-Our mobile robot is able to look for a visual object, for example a
-box of cornflakes, at a given place. We can model that deterministic
-action using the following schema:
-
-\def\oom{$^{\tt +}$}
-\def\zom{$^*$}
-\def\bump{\hspace{1cm}}
-\def\req#1{$^{\tt #1}$}
-\def\noteme#1{}%{[[#1]]}
-\def\notecoauth#1{\ }
-\def\notereader#1{[[#1]]}
-\def\meta{$\uparrow\uparrow$}
-\def\la{\langle}
-\def\ra{\rangle}
+Because PPDDL cannot model domains that feature partial observability,
+we developed an extension of that language, called DTPDDL. This can
+express probabilistic models of the sensing consequences of acting, to
+quantitatively capture unreliability in perception. That expressive
+power is achieved by incorporating perceptual analogues of fluent,
+predicate, and action definitions. In detail, we have declarations of
+state characterising predicate and fluent symbols according to PPDDL
+syntax. In addition, we allow two other declarative blocks, for
+perceptual predicates and fluents respectively. For example, suppose our
+robot is tasked with exploring {\em locations} in order to identify
+the whereabouts of a {\em visual-object}. We must describe state and
+perceptual facts that model the {\em true}, resp. perceived, locations
+of objects. In DTPDDL, these declarations appear as:
 
@@
 \begin{tabtt}
-(\= :action look-for-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
+(\=:functions  ;; state fluents\\
+  \> (is-in ?v - visual-object) - location )\\
+(:observable-functions  ;; perceptual fluents\\
+  \> (o-is-in ?v - visual-object) - location )
 \end{tabtt}
@@
 
-\noindent That is, the robot can always look for a visual object at
-its current location. The only effect this has on the state, is that
-the robot incurs a penalty, $3$, that corresponds to the {\em cost} of
-performing a visual search. Instances of this operator are useful,
-because their application activates an the following sense schema:
+\noindent To model the sensing capabilities of the agent, we have
+operator-like descriptive elements, with preconditions in both action
+and state symbols, and uniformly positive effects over perceptual
+predicates -- I.e, the negation normal form of a sensing effect
+formula cannot contain negation. Whereas the effects of an {\em
+operator} schema are over state propositions, and describe how the
+states change under application of actions, the effects of a {\em
+sense} schema are perceptual, and describe the composition of an
+observation following the execution of an action. Also, and without a
+loss of generality, we find it convenient to separate the actional
+precondition from the state precondition. For example, our robot is
+able to look for a visual object, such as a box of cornflakes, at a
+given place. We can model that deterministic action using the
+following operator schema:
 
@@
 \begin{tabtt}
-(\= :sense vision \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - location) \\
-  \> :execution (look-for-object ?r ?v ?l) \\
-  \> :effect ( and (when (= (is-in ?v) ?l) \\
-  \> \bump (probabilistic 0.8 \\
-  \> \bump (= (observe-is-in ?v) ?l))) \\
+(\=:action look-for-object \+ \\
+   :parameters (\=?r - robot ?v - visual-object\\
+   \> ?l - location) \\
+   :precondition (and (= (is-in ?r) ?l) ) \\
+   :effect (and (assign (reward) -3) ) ) \\
+\end{tabtt}
+}
+
+\noindent In a slight departure from PPDDL, we suppose that
+$\pp{look-for-object}$ has an effect on the state. That is, its
+execution incurs a penalty, $3$, corresponding to the {\em cost} of
+performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
+reserved word}, and occurs in {\em increased} and {\em decreased}
+clauses of operator schemata. In that setting, {\em reward} is
+interpreted as accumulated (resp. instantaneous) rewards and
+penalties.}
+
+Completing the example, we model the sensing outcome that results from
+instantiating the $\pp{look-for-object}$ operator:
+
+\small{
+\begin{tabtt}
+(\= :sense vision \+\\
+ :parameters \= (\= ?r - robot ?v - visual-object\\
+ \>\>  ?l - location) \\
+ :execution \> ( \> look-for-object ?r ?v ?l) \\
+ :precondition (and (= (is-in ?r) ?l) ) \\
+ :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
+   \> \> (probabilistic 0.8 \\
+   \>  \>(= (o-is-in ?v) ?l))) \\
   \> (when (not (= (is-in ?v) ?l)) \\
-  \> \bump (probabilistic 0.1 \\
-  \> \bump (= (observe-is-in ?v) ?l))))) \\
+   \>  \> (probabilistic 0.1 \\
+   \>  \> (= (o-is-in ?v) ?l))))) \\
 \end{tabtt}
@@
 the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. No
-other conditions guard the activation of this sense, therefore if
-$(\pp{look-for-object} \pp{Dora} \pp{cornflakes} \pp{bedroom})$ is
-executed, we suppose there is a $0.8$ chance of observing the
-cornflakes if they are in the $\pp{bedroom}$, and otherwise a $0.1$ of
-observing them.\footnote{We have simplified the schemata for
-illustrative purposes. In practice, the probabilities of a particular
-observation here should be parametrised by the type of visual object
-the robot is searching for, and category of the location.}
+$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
+\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
+we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
+order to fully demonstrate the syntax. Interpreting the above schema,
+if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
+executed, there is a $0.8$ chance of perceiving the box if it is in
+the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
 
-\scriptsize
-\begin{nopagebreak}\begin{tabtt}
-<observation-def> \=::= \=(:sense <observation symbol> \\
-                        \> \> :parameters (<typed list (variable)>)  \\
-                        \> \> <o-def body>) \\
-  <o-symbol> \> ::= <name> \\
-  <o-def body> \> ::= [:precondition <GD>] \\
-  \> \> [:execution <atomic action(term)> ] \\
-    \> \> [:effect <o-effect>] \\
-  <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-  <o-effect> \> ::= (and <c-o-effect>\zom) \\
-  <o-effect> \> ::= <c-o-effect> \\
-  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-  <c-o-effect> \> ::= <p-o-effect> \\
-  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-  <p-o-effect> \> ::= <atomic o-formula(term)> \\
-  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-  <o-f-exp> \> ::= <number>\\
-  <o-f-exp> \> ::= (- <o-f-exp>)\\
-  <o-f-exp> \> ::= <o-f-head>\\
-  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-  <o-f-head> \> ::= <o-function-symbol>\\
-\end{tabtt}\end{nopagebreak}
-\noteme{<structure-def> ::= <attach-def>}
-\normalsize
+The syntax and semantics for describing initial state distributions in
+DTPDDL is taken verbatim from PPDDL. It is useful to present that
+factored tree-like structure here, in order to aid our discussion of
+sequential planning in POMDP determinisations. The distribution is
+expressed in a tree-like structure of terms. Each term is either: (1)
+atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
+probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
+where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
+atomic terms. The root term is always conjunctive, and the leaves are
+atomic. For example, the starting distribution for $\pp{R2D2}$ might
+be described thusly:
 
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
+\small{
+\begin{tabtt}
+(\=:init (= (is-in R2D2) kitchen) \+ \\
+       (probabilistic \=.8 (= (is-in box) office)  \\
+		      \>.2 (= (is-in box) kitchen)) \\
+       (probabilistic .3 (= (is-in cup) office)  \\
+		      \>.7 (= (is-in cup) kitchen))) \\
+\end{tabtt}
+}
 
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
+\noindent The interpretation of such an expression can be given
+according to a recursive descent. Here, an atom is visited as soon as
+its conjunctive parent is visited, and a conjunctive term is visited
+if all its immediate subterms are visited. A probabilistic term is
+visited if exactly one of its subterms, $T_i$, is visited. Each valid
+visitation of the root term according to this recursive definition
+encapsulates a starting state, along with the probability it
+occurs. The former corresponds to the union of all visited atoms, and
+the latter corresponds to the product of $p_i$ entries on the visited
+subterms of probabilistic elements. Making these ideas concrete, our
+example yields the following flat state distribution:
 
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
+\begin{tabular}{cccc}
+\hline
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+\hline
+.24 & kitchen & office & office \\
+.06 & kitchen & kitchen & office \\
+.56 & kitchen & office & kitchen \\
+.14 & kitchen & kitchen & kitchen \\
+\hline
+\end{tabular}
 
-\noindent Nodes of the tree have one of three types: (1) atomic nodes
-are labelled with a state proposition, e.g., $(=
-(\pp{location}~\pp{cornflakes}) \pp{kitchen})$, (2) probabilistic
-nodes, have the form $(r_1, C_1;..;r_n, C_n)$, and (3) constructive
-nodes, take the form of a set of probabilistic and atomic nodes. The
-root node is conjunctive, and leaves are always atomic. A convenient
-way to give semantics to the tree, is to treat nodes as if they were
-actions. An atomic node is an action with a positive effect
-corresponding to the node label, .
+\section{Serial Planning in Determinisations}
 
-For example, $(:\pp{predicates} (\pp{location} ?l - VisualObject))$
-says that there is a unary predicate
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-knowledge objectives and  sensory effects.
-
-probabilistic models of the sensing consequences of acting to
-quantitatively capture the unreliability of sensing, and thus reason
-precisely about the utility of a plan.
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
+knows where the {\em cornflakes}--- are treated as control-knowledge
+by the planner. Here, a plan is not considered valid unless in the
+finial stet the goal condition is satisfied.
 
@@
 the underlying planner for CogX.
+
+the effects of a {\em sense} schema are perceptual, whereas the
+effects of an {\em operator} schema are over state propositions.
 
@@
 \subsection{MAPL}
+
+%% \scriptsize
+%% \begin{nopagebreak}\begin{tabtt}
+%% <observation-def> \=::= \=(:sense <observation symbol> \\
+%%                         \> \> :parameters (<typed list (variable)>)  \\
+%%                         \> \> <o-def body>) \\
+%%   <o-symbol> \> ::= <name> \\
+%%   <o-def body> \> ::= [:precondition <GD>] \\
+%%   \> \> [:execution <atomic action(term)> ] \\
+%%     \> \> [:effect <o-effect>] \\
+%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
+%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
+%%   <o-effect> \> ::= <c-o-effect> \\
+%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
+%%   <c-o-effect> \> ::= <p-o-effect> \\
+%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
+%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
+%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
+%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
+%%   <o-f-exp> \> ::= <number>\\
+%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
+%%   <o-f-exp> \> ::= <o-f-head>\\
+%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
+%%   <o-f-head> \> ::= <o-function-symbol>\\
+%% \end{tabtt}\end{nopagebreak}
+%% \noteme{<structure-def> ::= <attach-def>}
+%% \normalsize
+
+%% \footnote{We have simplified the schemata for illustrative
+%% purposes. In practice, the probabilities of a particular observation
+%% here should be parametrised by the type of visual object the robot is
+%% searching for, and category of the location.}
+
+We suppose $\bstate_0$ is given in a factored tree format of the
+form:
+
+%% \small{
+%% \begin{tabtt}
+%% (:init  (and \=f_{11} f_{12} ..
+%%  \> (probabilistic p_{}) 
+%%  \> (probabilistic ..) ..) \\
+%% \end{tabtt}
+%% }
+
+%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
+
+planning in practical sized POMDPs is intractable. One approach is to
+described the task hierarchically, thereby constrain and simplify the
+policy search task. The approach ~\cite{}.

### 2010-11-20 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
I think, sequential, rather than serial, is a better term.

@@
 
-\include{macros}
+\include{macros}%
 
@@
 \title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Serial and
+%\title{A Continual Planner that Switches Between Sequential and
 %  Decision-Theoretic Planning} 
@@
 
-the speed and scalability of software for serial planning in
+the speed and scalability of software for sequential planning in
 deterministic
@@
 
-We have extended MAPSIM to simulate DTPDDL instances, and modified
-\fastdownward\ so that it can find useful serial executions given
-DTPDDL models of the problem at hand.
+We have extended MAPSIM to parse and simulate DTPDDL problems, and
+modified \fastdownward\ so that it can find useful sequential executions
+given DTPDDL models of the problem at hand.
 
@@
 solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality serial plans
+and modify the \fastdownward\ system to find high quality sequential plans
 .
@@
 
-\section{Serial Planning in Determinisations}
+\section{Sequential Planning in Determinisations}
 
@@
 There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for serial planning
+under uncertainty using systems that were intended for sequential planning
 in deterministic problems.  Notably,
@@
 uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
+\system{FF}~\cite{hoffmann:nebel:2001} to compute a sequential plan and
 corresponding execution trace. The resulting plan is then executed
@@
 deterministic planning procedures. The outcome is typically a more
-robust serial plan~\cite{yoon:etal:2008}, or contingent
+robust sequential plan~\cite{yoon:etal:2008}, or contingent
 plan~\cite{majercik:2006}.
@@
 $T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
-planning ---i.e., serial planning in unobservable worlds--- can be
-modelled as a deterministic problem, and therefore solved using serial
+planning ---i.e., sequential planning in unobservable worlds--- can be
+modelled as a deterministic problem, and therefore solved using sequential
 systems. In this conformant setting, advances have been towards
@@
 planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, serial planning is
+planning. Given a POMDP model of the environment, sequential planning is
 responsible for an initial deterministic sequential plan and
@@
 
-Interaction between the serial planner and execution proceeds
+Interaction between the sequential planner and execution proceeds
 more-or-less analogously to popular replanning approaches
@@
 Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast serial
+{\em switching} continual planner, that uses the fast sequential
 satisficing procedure FastDownward to perform net-benefit
@@
 
-computing a single serial plan and
+computing a single sequential plan and
 eventuality

### 2010-11-20 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Better POMDP section.
 - Charles

@@
 If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
-chosen by nature when \stochAction\ is applied at state $\state$, then
-the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
-\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
-added to $\state$, and those from $\delete(\detAction)$ are removed
-from $\state$.
+..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
+..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
+applied at state $\state$, then the resultant state is $ ( \state \cup
+\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
+propositions from $\add(\detAction)$ are added to $\state$, and those
+from $\delete(\detAction)$ are removed from $\state$.
 
@@
 according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A percept is active, written $\stochSense \in
-\stochSenses(\stochAction, \state')$, if the senses'
+\subseteq \stochSenses$. A sense $\stochSense$ is active, written
+$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
 action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
@@
 we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$ where
-$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
-$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
-0$.
+an action \stochAction\ is executed at $\state_i$, that it has no
+effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
+\state_i$.
 
-\subsection{Costs, Rewards, and Plan Evaluation}
+\subsection{Costs, Rewards, and Belief Revision}
 
@@
 Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function.  Consequently there is a positive constant $c$
-so that for all $\state \in \states$ and $\action \in \actions$,
-$|\reward(\state, \action)| < c$.
+valued reward function. Therefore a finite positive constant $c$
+exists so that for all $\state \in \states$ and $\action \in
+\actions$, $|\reward(\state, \action)| < c$. We model costs are
+negative rewards.
 %%
@@
 
-Successive state estimation is available if we applying Bayes' rule.
-Taking the current belief $\bstate$ as our {\em prior}, and supposing
-action $\action$ is executed with perceptive outcome $\observ$, then
-the probability that we are in $\state$ in the successive belief-state
-$\bstate'$ is given by:
+In the POMDP model successive state estimation is available by
+application of Bayes' rule.  Taking the current belief $\bstate$ as
+the {\em prior}, and supposing action $\action$ is executed with
+perceptive outcome $\observ$, then the probability that we are in
+$\state$ in the successive belief-state $\bstate'$ is given by:
 
@@
 
-A general formalism for representing solutions to POMDPs is the
-finite-state controller (FSC). This is a three-tuple $\langle \nodes,
-\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
-nodes, $\selection_\node(\action) = P(\action | \node)$, and
-$\transition_\node(\action, \observation, \node') = P(\node'|\node,
-\action, \observation)$. The value of state $\state$ at node $\node$
-given a problem model, written $V_\node(\state)$, is:
+\subsection{Plan Evaluation}
 
-\begin{equation}\label{eq:evaluation}
-\begin{array}{lcl}
-V_\node(\state) & = & \sum_{\action \in \actions}
-\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+An optimal solution to a finite-horizon POMDP is a contingent plan,
+and can be expressed as a mapping from observation histories to
+actions. Although suboptimal in general, useful plans can also take a
+classical sequential format. This is the case in {\em conformant}
+planning, where the objective is to find a sequence of actions that
+achieves a goal ---I.e., reaches a state that satisfies some given
+Boolean condition--- with probability $1$.  More generally, a useful
+formalism for representing solutions to POMDPs is the finite-state
+controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+\transition \rangle$ where: $\node \in \nodes$ is a set of controller
+states, $\selection_\node(\action) = P(\action | \node)$ gives the
+probability that the controller prescribes \action\ when in state
+\node, and $\transition_\node(\action, \observation, \node') =
+P(\node'|\node, \action, \observation)$ gives the probability the
+controller transitions to state $\node'$, supposing we execute
+\action\ in state \node\ receiving observation \observation. When
+acting for finite $N$ steps, the value of a controller corresponds to
+the expected cumulative reward:
 
-&& \hspace{-10ex} \beta \sum_{\action, \observation,
-\state', \node'} \transition_\node(\action, \observation, \node')
-\transProb(\state, \action, \state') \obsDist_\observ(\state',
-\action) V_{\node'}(\state')
-\end{array}
-\end{equation}
+\[
+V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
+\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
+\]
 
-\noindent The value of $\bstate$ according to a controller is
-therefore:
+\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
+useful to note that there is a corresponding deterministic FSCs for
+both sequential (i.e., conformant) and contingent plans.
 
-\begin{equation} \label{eq:valueBelief}
-V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-\end{equation}
+%% More generally, a useful
+%% formalism for representing solutions to POMDPs is the finite-state
+%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
+%% $\selection_\node(\action) = P(\action | \node)$, and
+%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
+%% \action, \observation)$. Where the value of acting is the discounted
+%% accumulated reward over an infinite horizon, then each controller can
+%% be assigned a value, written $V_\node(\state)$, expressed in the usual
+%% way according to the Bellman equation:
+
+%% \begin{equation}\label{eq:evaluation}
+%% \begin{array}{lcl}
+%% V_\node(\state) & = & \sum_{\action \in \actions}
+%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+
+%% && \hspace{-10ex} \beta \sum_{\action, \observation,
+%% \state', \node'} \transition_\node(\action, \observation, \node')
+%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
+%% \action) V_{\node'}(\state')
+%% \end{array}
+%% \end{equation}
+
+%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
+%% then:
+
+%% \begin{equation} \label{eq:valueBelief}
+%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
+%% \end{equation}
+
+%% \noindent There is a corresponding deterministic FSCs for both
+%% sequential (i.e., conformant) and contingent
+%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
+%% case, if we take $\beta = 1$, and alter the process dynamics so there
+%% is a compulsory transition to a zero utility sink state when that
+%% horizon is reached.
 
@@
 
-\small{
+\small
 \begin{tabtt}
@@
 \end{tabtt}
-}
+\normalsize
 
@@
 
+\small
 \begin{tabular}{cccc}
@@
 \end{tabular}
+\normalsize
 

### 2010-11-20 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Fixed my related work text.
 - Charles

@@
 
-\[
+\begin{equation}\label{eq:expectedvalue}
 V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
 \sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
-\]
+\end{equation}
 
@@
 both sequential (i.e., conformant) and contingent plans.
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
-%% $\selection_\node(\action) = P(\action | \node)$, and
-%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
-%% \action, \observation)$. Where the value of acting is the discounted
-%% accumulated reward over an infinite horizon, then each controller can
-%% be assigned a value, written $V_\node(\state)$, expressed in the usual
-%% way according to the Bellman equation:
-
-%% \begin{equation}\label{eq:evaluation}
-%% \begin{array}{lcl}
-%% V_\node(\state) & = & \sum_{\action \in \actions}
-%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-%% && \hspace{-10ex} \beta \sum_{\action, \observation,
-%% \state', \node'} \transition_\node(\action, \observation, \node')
-%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
-%% \action) V_{\node'}(\state')
-%% \end{array}
-%% \end{equation}
-
-%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
-%% then:
-
-%% \begin{equation} \label{eq:valueBelief}
-%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-%% \end{equation}
-
-%% \noindent There is a corresponding deterministic FSCs for both
-%% sequential (i.e., conformant) and contingent
-%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
-%% case, if we take $\beta = 1$, and alter the process dynamics so there
-%% is a compulsory transition to a zero utility sink state when that
-%% horizon is reached.
 
@@
 predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to PPDDL
-syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose our
-robot is tasked with exploring {\em locations} in order to identify
-the whereabouts of a {\em visual-object}. We must describe state and
-perceptual facts that model the {\em true}, resp. perceived, locations
-of objects. In DTPDDL, these declarations appear as:
+state characterising predicate and fluent symbols according to the
+PPDDL syntax. In addition, we allow two other declarative blocks, for
+perceptual predicates and fluents respectively. For example, suppose
+our robot is tasked with exploring {\em locations} in order to
+identify the whereabouts of a {\em visual-object}. We must describe
+state and perceptual facts that model the {\em true}, resp. perceived,
+locations of objects. In DTPDDL, these declarations appear as:
 
@@
 loss of generality, we find it convenient to separate the actional
-precondition from the state precondition. For example, our robot is
-able to look for a visual object, such as a box of cornflakes, at a
-given place. We can model that deterministic action using the
-following operator schema:
+precondition from the state precondition.
+
+Making the above ideas concrete with an example, suppose a robot is
+able to look for a visual-object, such as a box, at a given place. We
+can model that deterministic action using the following operator
+schema:
 
@@
 
-\section{Sequential Planning in Determinisations}
+\section{Sequential Planning in POMDP Determinisations}
 
 Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes}--- are treated as control-knowledge
-by the planner. Here, a plan is not considered valid unless in the
-finial stet the goal condition is satisfied.
+$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
+location of the {\em box}--- are treated as control-knowledge by the
+planner. Here, a plan is not considered valid unless in the finial
+stet the goal condition is satisfied.
 
@@
 
-\section{Discussion}
+\section{Discussion and Related Work}
 
 There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential planning
-in deterministic problems.  Notably,
+under uncertainty using systems that were intended for sequential
+planning in deterministic problems.  Notably,
 \system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
 probabilistic track of the 2004 International Planning Competition,
-uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a sequential plan and
-corresponding execution trace. The resulting plan is then executed
-until the observed trace deviates from the planned trace.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something
-
+that in a continual framework uses the fast satisficing procedure
+\system{FF}~\cite{hoffmann:nebel:2001} to compute a sequential plan
+and corresponding execution trace. 
+%%
 More computationally expensive approaches in this vein combine
@@
 plan~\cite{majercik:2006}.
-
 %%
-Recently the systems
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
-planning ---i.e., sequential planning in unobservable worlds--- can be
-modelled as a deterministic problem, and therefore solved using sequential
+Also leveraging deterministic planners in problems that feature
+uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
+---i.e., sequential planning in unobservable worlds--- can be modelled
+as a deterministic problem, and therefore solved using sequential
 systems. In this conformant setting, advances have been towards
 compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. This
-research thread has also been extended recently to contingent planning
+search planning procedures, and lazy evaluations of beliefs. Most
+recently this research thread has been extended to contingent planning
 in fully observable non-deterministic
@@
 policy search task. The approach ~\cite{}.
+
+%% More generally, a useful
+%% formalism for representing solutions to POMDPs is the finite-state
+%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
+%% $\selection_\node(\action) = P(\action | \node)$, and
+%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
+%% \action, \observation)$. Where the value of acting is the discounted
+%% accumulated reward over an infinite horizon, then each controller can
+%% be assigned a value, written $V_\node(\state)$, expressed in the usual
+%% way according to the Bellman equation:
+
+%% \begin{equation}\label{eq:evaluation}
+%% \begin{array}{lcl}
+%% V_\node(\state) & = & \sum_{\action \in \actions}
+%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+
+%% && \hspace{-10ex} \beta \sum_{\action, \observation,
+%% \state', \node'} \transition_\node(\action, \observation, \node')
+%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
+%% \action) V_{\node'}(\state')
+%% \end{array}
+%% \end{equation}
+
+%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
+%% then:
+
+%% \begin{equation} \label{eq:valueBelief}
+%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
+%% \end{equation}
+
+%% \noindent There is a corresponding deterministic FSCs for both
+%% sequential (i.e., conformant) and contingent
+%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
+%% case, if we take $\beta = 1$, and alter the process dynamics so there
+%% is a compulsory transition to a zero utility sink state when that
+%% horizon is reached.
+
+ ---i.e., the values of {\em
+runtime variables} in the continual paradigm--- are treated , and
+therefore plan for a specific eventuality, replanning from scratch if
+something

### 2010-11-20 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex

@@
 \system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition,
-that in a continual framework uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a sequential plan
-and corresponding execution trace. 
+probabilistic track of the 2004 International Planning Competition.
+In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
+procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
+plans and corresponding execution traces.
 %%
@@
 
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
 From an automated planning perspective, the problem of practical
@@
 planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning is
-responsible for an initial deterministic sequential plan and
+planning. Given a POMDP model of the environment, sequential planning
+is used to compute an initial deterministic sequential plan and
 complementary runtime evolution of the decision process. That plan is
-executed until an assumption about a runtime proposition
+executed until a validation of an assumption about a runtime
+proposition is requested.
+
+The switching continual planning system we have described serves as
+the underlying planner for CogX.
+
+the effects of a {\em sense} schema are perceptual, whereas the
+effects of an {\em operator} schema are over state propositions.
 
@@
 addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planing for contingencies unless
+execution, deliberately postponing planning for contingencies unless
 they eventuate during execution.

### 2010-11-20 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Still working on a vocabulary that I am happy with..
 - Charles

@@
 \noindent The interpretation of such an expression can be given
-according to a recursive descent. Here, an atom is visited as soon as
-its conjunctive parent is visited, and a conjunctive term is visited
-if all its immediate subterms are visited. A probabilistic term is
-visited if exactly one of its subterms, $T_i$, is visited. Each valid
-visitation of the root term according to this recursive definition
-encapsulates a starting state, along with the probability it
-occurs. The former corresponds to the union of all visited atoms, and
-the latter corresponds to the product of $p_i$ entries on the visited
-subterms of probabilistic elements. Making these ideas concrete, our
-example yields the following flat state distribution:
+according to a recursive note of {\em visiting} terms. Here, an atom
+is visited if its conjunctive parent is visited, and a conjunctive
+term is visited if all its immediate subterms are visited. A
+probabilistic term is visited if exactly one of its subterms, $T_i$,
+is visited. Each valid visitation of the root term according to this
+recursive definition encapsulates a starting state, along with the
+probability it occurs. The former corresponds to the union of all
+visited atoms, and the latter corresponds to the product of $p_i$
+entries on the visited subterms of probabilistic elements. Making
+these ideas concrete, our example yields the following flat state
+distribution:
 
@@
 
-\section{Sequential Planning in POMDP Determinisations}
+\section{Switching Planner}
+
+RICHARD, I HAVE STOLEN A FEW LINES FROM YOUR HIPPO THREAD TO SKETCH
+THIS STUFF. I HOPE IT'S TAKEN AS FLATTERY :)
+
+We now describe a continual planning approach that switches between
+sequential and decision-theoretic planning. The problem at hand is a
+POMDP described in DTPDDL.  Our approach uses sequential planning to
+compute a linear plan, and one trace of that plan's execution. The
+latter encapsulates assumptions about: (1) the true underlying state,
+(2) how the plan's execution will progress, and (3) the possibility of
+eventually holding strong beliefs about the truth values of specific
+state propositions. Sequential planning proceeds in a continual manner
+until the validation of an assumption about a runtime state
+proposition is scheduled. At that point the system switches to a
+decision-theoretic planner that is used to tailor sensory processing
+by the robot to test the validity of the assumptions so far.
+
+The problem posed to the decision-theoretic system is an abstraction
+of the problem at hand. It is induced according to the executing
+sequential plan. Specifically, the abstract process is derived by
+first omitting all problem variables that are not mentioned in the
+linear trace, and then refining that preliminary model, by
+progressively adding more problem variables while the resulting state
+space remains sufficiently small. Refinement is a simple and
+inexpensive step, guided by the level at which variable symbols occur
+in the description of the initial conditions. The abstract process is
+also given a utility model that captures the reward (resp. penalty) of
+correctly (resp. falsely) in/validating the assumptions made in the
+linear trace. In detail, for each assumption on a state variable, we
+can have a reward of $1$ for correctly rejecting it, and a penalty of
+$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
+a more-or-less continual manner until an assumption is rejected, or
+until it returns control to the sequential system because it can see
+no benefit from performing further actions in the abstraction.
+
+Addressing briefly the specifics of our mobile robot's
+architecture. In our scenario the utility model for the POMDP posed to
+decision-theoretic planning is sourced from a motivational component
+of our robotic architecture. This proposes a non-zero reward
+(resp. penalty) for sequential planning with a true (resp. false)
+assumption about a state proposition.
+
+\subsection{Sequential Planning in POMDP Determinisations}
 
@@
 stet the goal condition is satisfied.
+
+\subsection{Abstractions for Decision-Theoretic Planner}
 

### 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Few tiny mistakes from yesterday.
 - Charles

@@
 useful to note that there is a corresponding deterministic FSCs for
-both sequential (i.e., conformant) and contingent plans.
+both sequential (e.g., conformant) and contingent plans.
 
@@
 DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid our discussion of
+factored tree-like structure here, in order to aid us when discussing
 sequential planning in POMDP determinisations. The distribution is
@@
 \noindent The interpretation of such an expression can be given
-according to a recursive note of {\em visiting} terms. Here, an atom
-is visited if its conjunctive parent is visited, and a conjunctive
-term is visited if all its immediate subterms are visited. A
-probabilistic term is visited if exactly one of its subterms, $T_i$,
-is visited. Each valid visitation of the root term according to this
-recursive definition encapsulates a starting state, along with the
-probability it occurs. The former corresponds to the union of all
-visited atoms, and the latter corresponds to the product of $p_i$
-entries on the visited subterms of probabilistic elements. Making
-these ideas concrete, our example yields the following flat state
-distribution:
+according to a recursive notion of {\em visitation} over terms. Here,
+an atom is {\em visited} if its conjunctive parent is visited, and a
+conjunctive term is visited if all its immediate subterms are
+visited. A probabilistic term is visited if exactly one of its
+subterms, $T_i$, is visited. Each valid visitation of the root term
+according to this recursive definition encapsulates a starting state,
+along with the probability it occurs. The former corresponds to the
+union of all visited atoms, and the latter corresponds to the product
+of $p_i$ entries on the visited subterms of probabilistic
+elements. Making these ideas concrete, our example yields the
+following flat state distribution:
 

### 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
I have to go to a lunch. Writers block was just starting to subside too, what a shame...
 - Charles

@@
 
-RICHARD, I HAVE STOLEN A FEW LINES FROM YOUR HIPPO THREAD TO SKETCH
-THIS STUFF. I HOPE IT'S TAKEN AS FLATTERY :)
-
 We now describe a continual planning approach that switches between
 sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.  Our approach uses sequential planning to
-compute a linear plan, and one trace of that plan's execution. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how the plan's execution will progress, and (3) the possibility of
-eventually holding strong beliefs about the truth values of specific
-state propositions. Sequential planning proceeds in a continual manner
-until the validation of an assumption about a runtime state
-proposition is scheduled. At that point the system switches to a
-decision-theoretic planner that is used to tailor sensory processing
-by the robot to test the validity of the assumptions so far.
+POMDP described in DTPDDL.
+
+\subsection{Sequential Planning in POMDPs}
+
+Our approach uses sequential planning to compute a linear plan, and
+one trace of that plan's execution. The latter encapsulates
+assumptions about: (1) the true underlying state, (2) how the plan's
+execution will progress, and (3) the possibility of eventually holding
+strong beliefs about the truth values of specific state
+propositions. Sequential planning proceeds in a continual manner until
+the verification of an assumption about a runtime state proposition is
+scheduled. At that point the system switches to a decision-theoretic
+planner that is used to tailor sensory processing by the robot to test
+the validity of the assumptions so far.
+
+The first actions sequenced by the planner are virtual, and correspond
+to identifying an abstract start state. That prefix corresponds to a
+{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
+block. In the relaxed case, we can modify the initial definition so
+that a conjunctive term is visited provided all its atomic subterms
+are visited, and zero or more of its probabilistic terms are
+visited. Writing \#\ if the value of a proposition is unspecified by a
+the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
+following abstract starting states:
+
+\small
+\begin{tabular}{cccc}
+\hline
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+\hline
+.24 & kitchen & office & office \\
+.3 & kitchen & \# & office \\
+.8 & kitchen & office & \# \\
+.2 & kitchen & kitchen & \# \\
+.7 & kitchen & \# &  kitchen\\
+.06 & kitchen & kitchen & office \\
+.56 & kitchen & office & kitchen \\
+.14 & kitchen & kitchen & kitchen \\
+\hline
+\end{tabular}
+\normalsize
+
+Sequential planning in our approach proceeds in two phases. In the
+first phase, the goal is to identify a plan prefix
+
+In sequential planning, the plan prefix corresponds to choosing a
+valid abstract starting state.
+
+moreover, for the purpose of decision-theoretic abstractions,
+
+A relaxed form of {\em visitation} we call {\em abstract
+  visitation}. In the abstract case a conjunctive term is visited
+  provided all its atomic subterms are visited, and zero or more of
+  its probabilistic terms are visited.
+
+The planner has a compulsory starting action, that visits the root
+term, and then has to execute visitation actions until that plan
+prefix corresponds to a valid {\em abstract visitation} of the {\em
+:init} block. Beyond such a prefix, the process actions and senses are
+available to the planner
+
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
+location of the {\em box}--- are treated as control-knowledge by the
+planner. Here, a plan is not considered valid unless in the finial
+stet the goal condition is satisfied.
+
+\subsection{Decision-Theoretic Planning in Abstractions}
 
 The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand. It is induced according to the executing
-sequential plan. Specifically, the abstract process is derived by
-first omitting all problem variables that are not mentioned in the
-linear trace, and then refining that preliminary model, by
-progressively adding more problem variables while the resulting state
-space remains sufficiently small. Refinement is a simple and
+of the problem at hand with a distinct utility model that focuses the
+agents activities on assumption verification. The abstraction is
+induced according to the executing sequential plan. Specifically, we
+first omit all problem variables that are not mentioned in the linear
+trace, and then refine that preliminary model, by progressively adding
+more problem variables while the resulting state space remains of a
+practicable size. Refinement in our approach is a simple and
 inexpensive step, guided by the level at which variable symbols occur
@@
 assumption about a state proposition.
-
-\subsection{Sequential Planning in POMDP Determinisations}
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
-
-\subsection{Abstractions for Decision-Theoretic Planner}
 

### 2010-11-21 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Paper do far
Initial state description to actions, sketched.
 - Charels

@@
 
+\section{Moritz's Initial Outline}
+
+\subsection{Switching Planner Outline}
+
+Outline the high level approach here:
+\begin{itemize}
+\item Problem input (factored initial state, DTPDDL observations)
+\item Determinisation
+\item Problem generation for DT
+\item Belief revision
+\end{itemize}
+
+\subsection{Determinisation}
+
+\begin{itemize}
+\item Generation of deterministic sensing models
+\item Determinisation of the initial state
+\item Compiling reward model into (soft-)goal model
+\item Cost function in FD (maybe this could mentioned in the outline?)
+\end{itemize}
+
+\subsection{Subtask Generation}
+\begin{itemize}
+\item Goal generation from CP trace
+\item Disconfirm actions
+\item State pruning according to CP trace
+\end{itemize}
+
+\subsection{Belief revision}
+Not sure if this needs its own section
+\begin{itemize}
+\item Convert inital state into bayesian network
+\item Belief updates with observation nodes
+\item Create new network for next planning runs
+\end{itemize}
+
 \section{Introduction}
-
-Any (PO)MDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
 
@@
 
-\section{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\section{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\section{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\section{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
 \section{Propositional Decision-Theoretic Planning}
@@
 
-A general formalism for representing solutions to POMDPs is the
-finite-state controller (FSC). This is a three-tuple $\langle \nodes,
-\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
-nodes, $\selection_\node(\action) = P(\action | \node)$, and
-$\transition_\node(\action, \observation, \node') = P(\node'|\node,
-\action, \observation)$. The value of state $\state$ at node $\node$
-given a problem model, written $V_\node(\state)$, is:
+\subsection{Plan Evaluation}
 
-\begin{equation}\label{eq:evaluation}
-\begin{array}{lcl}
-V_\node(\state) & = & \sum_{\action \in \actions}
-\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+An optimal solution to a finite-horizon POMDP is a contingent plan,
+and can be expressed as a mapping from observation histories to
+actions. Although suboptimal in general, useful plans can also take a
+classical sequential format. This is the case in {\em conformant}
+planning, where the objective is to find a sequence of actions that
+achieves a goal ---I.e., reaches a state that satisfies some given
+Boolean condition--- with probability $1$.  More generally, a useful
+formalism for representing solutions to POMDPs is the finite-state
+controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+\transition \rangle$ where: $\node \in \nodes$ is a set of controller
+states, $\selection_\node(\action) = P(\action | \node)$ gives the
+probability that the controller prescribes \action\ when in state
+\node, and $\transition_\node(\action, \observation, \node') =
+P(\node'|\node, \action, \observation)$ gives the probability the
+controller transitions to state $\node'$, supposing we execute
+\action\ in state \node\ receiving observation \observation. When
+acting for finite $N$ steps, the value of a controller corresponds to
+the expected cumulative reward:
 
-&& \hspace{-10ex} \beta \sum_{\action, \observation,
-\state', \node'} \transition_\node(\action, \observation, \node')
-\transProb(\state, \action, \state') \obsDist_\observ(\state',
-\action) V_{\node'}(\state')
-\end{array}
+\begin{equation}\label{eq:expectedvalue}
+V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
+\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
 \end{equation}
 
-\noindent The value of $\bstate$ according to a controller is
-therefore:
-
-\begin{equation} \label{eq:valueBelief}
-V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-\end{equation}
-
-\section{Serial Planning in Determinisations}
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes}--- are treated as control-knowledge
-by the planner. Here, a plan is not considered valid unless in the
-finial stet the goal condition is satisfied.
+\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
+useful to note that there is a corresponding deterministic FSCs for
+both sequential (e.g., conformant) and contingent plans.
 
@@
 predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to PPDDL
-syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose our
-robot is tasked with exploring {\em locations} in order to identify
-the whereabouts of a {\em visual-object}. We must describe state and
-perceptual facts that model the {\em true}, resp. perceived, locations
-of objects. In DTPDDL, these declarations appear as:
+state characterising predicate and fluent symbols according to the
+PPDDL syntax. In addition, we allow two other declarative blocks, for
+perceptual predicates and fluents respectively. For example, suppose
+our robot is tasked with exploring {\em locations} in order to
+identify the whereabouts of a {\em visual-object}. We must describe
+state and perceptual facts that model the {\em true}, resp. perceived,
+locations of objects. In DTPDDL, these declarations appear as:
 
-\small{
+\small
 \begin{tabtt}
@@
 loss of generality, we find it convenient to separate the actional
-precondition from the state precondition. For example, our robot is
-able to look for a visual object, such as a box of cornflakes, at a
-given place. We can model that deterministic action using the
-following operator schema:
+precondition from the state precondition.
 
-\small{
+Making the above ideas concrete with an example, suppose a robot is
+able to look for a visual-object, such as a box, at a given place. We
+can model that deterministic action using the following operator
+schema:
+
+\small
 \begin{tabtt}
@@
 \end{tabtt}
-}
+\normalsize
 
 \noindent The interpretation of such an expression can be given
-according to a recursive descent. Here, an atom is visited as soon as
-its conjunctive parent is visited, and a conjunctive term is visited
-if all its immediate subterms are visited. A probabilistic term is
-visited if exactly one of its subterms, $T_i$, is visited. Each valid
-visitation of the root term according to this recursive definition
-encapsulates a starting state, along with the probability it
-occurs. The former corresponds to the union of all visited atoms, and
-the latter corresponds to the product of $p_i$ entries on the visited
-subterms of probabilistic elements. Making these ideas concrete, our
-example yields the following flat state distribution:
+according to a recursive notion of {\em visitation} over terms. Here,
+an atom is {\em visited} iff its conjunctive parent is visited, and a
+conjunctive term is visited iff all its immediate subterms are
+visited. A probabilistic term is visited iff its conjunctive parent is
+visited, and exactly one of its subterms, $T_i$, is visited. Each
+valid visitation of the root term according to this recursive
+definition encapsulates a starting state, along with the probability
+it occurs. The former corresponds to the union of all visited atoms,
+and the latter corresponds to the product of $p_i$ entries on the
+visited subterms of probabilistic elements. Making these ideas
+concrete, our example yields the following flat state distribution:
 
+\small
 \begin{tabular}{cccc}
@@
 \end{tabular}
+\normalsize
+
+\section{Switching Planner}
+
+We now describe a continual planning approach that switches between
+sequential and decision-theoretic planning.
+
+In this work, the problem at hand is a POMDP described in DTPDDL. From
+hereon, we restrict our attention to POMDPs with deterministic
+actions. Therefore it is worthwhile noting that any POMDP with
+stochastic actions can be compiled into an equivalent
+deterministic-action POMDP, where all the original action uncertainty
+is expressed in the starting-state
+distribution~\cite{ng:Jordan:2000}. In our setting, that of
+finite-horizon planning, such a compilation yields a finite flat
+representation of the original POMDP, only with deterministic
+actions.
+
+\Omit{
+  Therefore, although our current implementation does not
+  explicitly support it, our approach generalises to POMDPs with
+  stochastic action models.
+}
+
+\subsection{Sequential Planning in POMDPs}
+
+Our approach uses sequential planning to compute a linear plan, and
+one trace of that plan's execution. The latter encapsulates
+assumptions about: (1) the true underlying state, (2) how the plan's
+execution will progress, and (3) the possibility of eventually holding
+strong beliefs about the truth values of specific state
+propositions. Sequential planning proceeds in a continual manner until
+the verification of an assumption about a runtime state proposition is
+scheduled. At that point the system switches to a decision-theoretic
+planner that is used to tailor sensory processing by the robot to test
+the validity of the assumptions made so far.
+
+In a deterministic-action POMDP, all the uncertainty in state and
+action is expressed in the $(\pp{:init})$ declaration. Our approach
+uses the structure of that, as it occurs in the problem description,
+do define a set of abstract state-assumptions available to 
+sequential planning.  
+%%
+Writing \#\ if the value of a proposition is unspecified in the
+assumption, for the $\pp{R2D2}$ example we have that the following
+assumptions are available:
+
+\small
+\begin{tabular}{cccc}
+\hline
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+\hline
+.24 & kitchen & office & office \\
+.3 & kitchen & \# & office \\
+.8 & kitchen & office & \# \\
+.2 & kitchen & kitchen & \# \\
+.7 & kitchen & \# &  kitchen\\
+.06 & kitchen & kitchen & office \\
+.56 & kitchen & office & kitchen \\
+.14 & kitchen & kitchen & kitchen \\
+1.0 & kitchen & \# & \# \\
+\hline
+\end{tabular}
+\normalsize
+
+\noindent A valid assumption in our setting corresponds to a {\em
+relaxed} visitation of the root term of $(:init)$. In this relaxed
+case, a conjunctive term is visited iff its atomic subterms are
+visited, and zero or more of its immediate probabilistic subterms are
+visited. The starting state, $s_0$, for sequential planning is the
+abstract state with probability $1.0$ -- E.G., for $\pp{R2D2}$ $s_0
+\equiv \{(= (is-in R2D2) kitchen)\}$.
+
+In order to make assumptions available to the sequential planner, we
+add one action $\actions(p_i;(T_i))$ to the problem for each element,
+$p_i (T_i)$, of each probabilistic term from $(:init)$. The physics
+for these actions is as follows. $\actions(p_i;(T_i))$ can be executed
+if no $\actions(p_j;(T_j))$, $j \neq i$, has been executed from the
+same probabilistic term, and, either
+$(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct, or it
+occurs in $T_k$ for some executed $\actions(p_k;(T_k))$.
+%%
+Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
+successor state $s^v$ with probability $p_i$, and $s^\bot$ with
+probability $1 - p_i$. Here, $s^v$ is characterised by the union of
+$s$ and atomic terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
+$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
+$\transProb(\state^\bot, \action, \state^\bot) = 1$.
+
+\subsection{Decision-Theoretic Planning in Abstractions}
+
+The problem posed to the decision-theoretic system is an abstraction
+of the problem at hand with a distinct utility model that focuses the
+agents activities on assumption verification. The abstraction is
+induced according to the executing sequential plan. Specifically, we
+first omit all problem variables that are not mentioned in the linear
+trace, and then refine that preliminary model, by progressively adding
+more problem variables while the resulting state space remains of a
+practicable size. Refinement in our approach is a simple and
+inexpensive step, guided by the level at which variable symbols occur
+in the description of the initial conditions. The abstract process is
+also given a utility model that captures the reward (resp. penalty) of
+correctly (resp. falsely) in/validating the assumptions made in the
+linear trace. In detail, for each assumption on a state variable, we
+can have a reward of $1$ for correctly rejecting it, and a penalty of
+$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
+a more-or-less continual manner until an assumption is rejected, or
+until it returns control to the sequential system because it can see
+no benefit from performing further actions in the abstraction.
+
+Addressing briefly the specifics of our mobile robot's
+architecture. In our scenario the utility model for the POMDP posed to
+decision-theoretic planning is sourced from a motivational component
+of our robotic architecture. This proposes a non-zero reward
+(resp. penalty) for sequential planning with a true (resp. false)
+assumption about a state proposition.
 
@@
 
-\section{Discussion}
+\section{Discussion and Related Work}
 
 There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for serial planning
-in deterministic problems.  Notably,
+under uncertainty using systems that were intended for sequential
+planning in deterministic problems.  Notably,
 \system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition,
-uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
-corresponding execution trace. The resulting plan is then executed
-until the observed trace deviates from the planned trace.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something
-
+probabilistic track of the 2004 International Planning Competition.
+In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
+procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
+plans and corresponding execution traces.
+%%
 More computationally expensive approaches in this vein combine
@@
 deterministic planning procedures. The outcome is typically a more
-robust serial plan~\cite{yoon:etal:2008}, or contingent
+robust sequential plan~\cite{yoon:etal:2008}, or contingent
 plan~\cite{majercik:2006}.
-
 %%
-Recently the systems
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
-planning ---i.e., serial planning in unobservable worlds--- can be
-modelled as a deterministic problem, and therefore solved using serial
+Also leveraging deterministic planners in problems that feature
+uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
+---i.e., sequential planning in unobservable worlds--- can be modelled
+as a deterministic problem, and therefore solved using sequential
 systems. In this conformant setting, advances have been towards
 compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. This
-research thread has also been extended recently to contingent planning
+search planning procedures, and lazy evaluations of beliefs. Most
+recently this research thread has been extended to contingent planning
 in fully observable non-deterministic
@@
 \section{Concluding Remarks}
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
 
@@
 planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, serial planning is
-responsible for an initial deterministic sequential plan and
+planning. Given a POMDP model of the environment, sequential planning
+is used to compute an initial deterministic sequential plan and
 complementary runtime evolution of the decision process. That plan is
-executed until an assumption about a runtime proposition
+executed until a validation of an assumption about a runtime
+proposition is requested.
+
+The switching continual planning system we have described serves as
+the underlying planner for CogX.
+
+the effects of a {\em sense} schema are perceptual, whereas the
+effects of an {\em operator} schema are over state propositions.
 
@@
 addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planing for contingencies unless
+execution, deliberately postponing planning for contingencies unless
 they eventuate during execution.
 
-computing a single serial plan and
+computing a single sequential plan and
 eventuality
@@
 policy search task. The approach ~\cite{}.
+
+%% More generally, a useful
+%% formalism for representing solutions to POMDPs is the finite-state
+%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
+%% $\selection_\node(\action) = P(\action | \node)$, and
+%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
+%% \action, \observation)$. Where the value of acting is the discounted
+%% accumulated reward over an infinite horizon, then each controller can
+%% be assigned a value, written $V_\node(\state)$, expressed in the usual
+%% way according to the Bellman equation:
+
+%% \begin{equation}\label{eq:evaluation}
+%% \begin{array}{lcl}
+%% V_\node(\state) & = & \sum_{\action \in \actions}
+%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+
+%% && \hspace{-10ex} \beta \sum_{\action, \observation,
+%% \state', \node'} \transition_\node(\action, \observation, \node')
+%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
+%% \action) V_{\node'}(\state')
+%% \end{array}
+%% \end{equation}
+
+%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
+%% then:
+
+%% \begin{equation} \label{eq:valueBelief}
+%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
+%% \end{equation}
+
+%% \noindent There is a corresponding deterministic FSCs for both
+%% sequential (i.e., conformant) and contingent
+%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
+%% case, if we take $\beta = 1$, and alter the process dynamics so there
+%% is a compulsory transition to a zero utility sink state when that
+%% horizon is reached.
+
+ ---i.e., the values of {\em
+runtime variables} in the continual paradigm--- are treated , and
+therefore plan for a specific eventuality, replanning from scratch if
+something
+
+Sequential planning in our approach proceeds in two phases. In the
+first phase, the goal is to identify a plan prefix
+
+In sequential planning, the plan prefix corresponds to choosing a
+valid abstract starting state.
+
+moreover, for the purpose of decision-theoretic abstractions,
+
+A relaxed form of {\em visitation} we call {\em abstract
+  visitation}. In the abstract case a conjunctive term is visited
+  provided all its atomic subterms are visited, and zero or more of
+  its probabilistic terms are visited.
+
+The planner has a compulsory starting action, that visits the root
+term, and then has to execute visitation actions until that plan
+prefix corresponds to a valid {\em abstract visitation} of the {\em
+:init} block. Beyond such a prefix, the process actions and senses are
+available to the planner
+
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
+location of the {\em box}--- are treated as control-knowledge by the
+planner. Here, a plan is not considered valid unless in the finial
+stet the goal condition is satisfied.

### 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
I am burnt out. Perhaps I will get some more energy later tonight.
Moritz, I think I am representing your work correctly, but part of the pain is mulling over your Python, and staring at a cloud...

 - Charles

@@
 \section{Introduction}
-
-Any POMDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
 
@@
 according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} if its conjunctive parent is visited, and a
-conjunctive term is visited if all its immediate subterms are
-visited. A probabilistic term is visited if exactly one of its
-subterms, $T_i$, is visited. Each valid visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
+an atom is {\em visited} iff its conjunctive parent is visited, and a
+conjunctive term is visited iff all its immediate subterms are
+visited. A probabilistic term is visited iff its conjunctive parent is
+visited, and exactly one of its subterms, $T_i$, is visited. Each
+valid visitation of the root term according to this recursive
+definition encapsulates a starting state, along with the probability
+it occurs. The former corresponds to the union of all visited atoms,
+and the latter corresponds to the product of $p_i$ entries on the
+visited subterms of probabilistic elements. Making these ideas
+concrete, our example yields the following flat state distribution:
 
@@
 We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.
+sequential and decision-theoretic planning.
+
+In this work, the problem at hand is a POMDP described in DTPDDL. From
+hereon, we restrict our attention to POMDPs with deterministic
+actions. Therefore it is worthwhile noting that any POMDP with
+stochastic actions can be compiled into an equivalent
+deterministic-action POMDP, where all the original action uncertainty
+is expressed in the starting-state
+distribution~\cite{ng:Jordan:2000}. In our setting, that of
+finite-horizon planning, such a compilation yields a finite flat
+representation of the original POMDP, only with deterministic
+actions.
+
+PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
+HOW WE DO BELIEF REVISION ACROSS SWITCHING.
+
+\Omit{
+  Therefore, although our current implementation does not
+  explicitly support it, our approach generalises to POMDPs with
+  stochastic action models.
+}
 
@@
 assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of eventually holding
-strong beliefs about the truth values of specific state
-propositions. Sequential planning proceeds in a continual manner until
-the verification of an assumption about a runtime state proposition is
-scheduled. At that point the system switches to a decision-theoretic
-planner that is used to tailor sensory processing by the robot to test
-the validity of the assumptions so far.
+execution will progress, and (3) the possibility of the agent
+eventually holding strong beliefs about the truth values of specific
+state propositions. Sequential planning proceeds in a continual manner
+until the verification of an assumption about a runtime state
+proposition is scheduled. At that point the system switches to a
+decision-theoretic planner that is used to tailor sensory processing
+by the robot to test the validity of the assumptions made so far.
 
-The first actions sequenced by the planner are virtual, and correspond
-to identifying an abstract start state. That prefix corresponds to a
-{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
-block. In the relaxed case, we can modify the initial definition so
-that a conjunctive term is visited provided all its atomic subterms
-are visited, and zero or more of its probabilistic terms are
-visited. Writing \#\ if the value of a proposition is unspecified by a
-the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
-following abstract starting states:
+In a deterministic-action POMDP, all the uncertainty in state and
+action is expressed in the $(\pp{:init})$ declaration. Our approach
+uses the structure of that, as it occurs in the problem description,
+do define a set of abstract state-assumptions available to 
+sequential planning.  
+%%
+Writing \#\ if the value of a proposition is unspecified in the
+assumption, for the $\pp{R2D2}$ example we have that the following
+assumptions are available:
 
@@
 .24 & kitchen & office & office \\
+.06 & kitchen & kitchen & office \\
+.56 & kitchen & office & kitchen \\
+.14 & kitchen & kitchen & kitchen \\
+.7 & kitchen & \# &  kitchen\\
 .3 & kitchen & \# & office \\
@@
 .2 & kitchen & kitchen & \# \\
-.7 & kitchen & \# &  kitchen\\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
+1.0 & kitchen & \# & \# \\
 \hline
@@
 
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
+\noindent A valid assumption in our setting corresponds to a {\em
+relaxed} visitation of the root term of $(:init)$. In this relaxed
+case, a conjunctive term is visited iff its atomic subterms are
+visited, and zero or more of its immediate probabilistic subterms are
+visited. The starting state, $\state_0$, for sequential planning is
+the abstract state with probability $1.0$ -- E.G., for $\pp{R2D2}$
+$s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
 
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
+In order to make assumptions available to the sequential planner, we
+add one action $\actions(p_i;(T_i))$ to the problem for each element,
+$p_i (T_i)$, of each probabilistic term from $(:init)$. The physics
+of these actions is as follows. $\actions(p_i;(T_i))$ can be executed
+if no $\actions(p_j;(T_j))$, $j \neq i$, has been executed from the
+same probabilistic term, and, either
+$(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct, or it
+occurs in $T_k$ for some executed $\actions(p_k;(T_k))$.
+%%
+Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
+successor state $s^v$ with probability $p_i$, and $s^\bot$ with
+probability $1 - p_i$. Here, $s^v$ is characterised by the union of
+$s$ and atomic terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
+$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
+$\transProb(\state^\bot, \action, \state^\bot) = 1$.
 
-moreover, for the purpose of decision-theoretic abstractions,
+Because $\state_0$ is abstract, and because actions can have
+preconditions that are not necessarily true in the belief-state, we
+must make two further adjustments to our sequential continual
+planning. First, if $\action\in\actions$ has a precondition that is
+not true with probability $1$ in the current state, then before it is
+executed blindly, continual planning switches to a decision-theoretic
+mode, that shall be discussed in detail below. Therefore, by
+scheduling $\action$ the sequential planner makes a naive assumption
+about the possibility of the agent eventually holding strong beliefs
+about the truth values of specific state propositions. Second, if the
+truth value of $\prop$ is not known, and no assumption is made about
+it in the current abstract state, then if $\prop \in \poss(\action)$
+we have that action $\action$ is not applicable.
 
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
+Finally, we come to the issue of optimisation criteria. Here, we
+propose the value of an $N$ horizon trace ---something we expect the
+sequential planner to maximise--- corresponds to the maximal value of
+that trace. In detail, where $p_i$ is the probability that the
+$i^{th}$ execution is not a transition to $s^\bot$, we value a trace
+according to:
 
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
+\begin{equation}\label{eq:tracevalue}
+V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
+\end{equation}
 
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
+\noindent In goal directed problems, where the only non-zero reward is
+received on the first transition to a goal state,
+Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
+infamous degenerate hindsight system
+\system{FFR$_a$}~\cite{yoon:etal:2007}. Moreover, form the perspective
+of gradient-based reinforcement learning systems, such as William's
+REINFORCE and GPOMDP, our optimal trace yields a maximal gradient
+contribution under reasonable assumptions.
 
@@
 something
+
+Sequential planning in our approach proceeds in two phases. In the
+first phase, the goal is to identify a plan prefix
+
+In sequential planning, the plan prefix corresponds to choosing a
+valid abstract starting state.
+
+moreover, for the purpose of decision-theoretic abstractions,
+
+A relaxed form of {\em visitation} we call {\em abstract
+  visitation}. In the abstract case a conjunctive term is visited
+  provided all its atomic subterms are visited, and zero or more of
+  its probabilistic terms are visited.
+
+The planner has a compulsory starting action, that visits the root
+term, and then has to execute visitation actions until that plan
+prefix corresponds to a valid {\em abstract visitation} of the {\em
+:init} block. Beyond such a prefix, the process actions and senses are
+available to the planner
+
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
+location of the {\em box}--- are treated as control-knowledge by the
+planner. Here, a plan is not considered valid unless in the finial
+stet the goal condition is satisfied.

### 2010-11-22 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Only committing changes to yesterday's text.
 - Charles

@@
 In order to make assumptions available to the sequential planner, we
-add one action $\actions(p_i;(T_i))$ to the problem for each element,
-$p_i (T_i)$, of each probabilistic term from $(:init)$. The physics
-of these actions is as follows. $\actions(p_i;(T_i))$ can be executed
-if no $\actions(p_j;(T_j))$, $j \neq i$, has been executed from the
-same probabilistic term, and, either
-$(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct, or it
-occurs in $T_k$ for some executed $\actions(p_k;(T_k))$.
+add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
+for each element, $p_i (T_i)$, of each probabilistic term from
+$(:init)$. The physics of these actions is as
+follows. $\actions(p_i;(T_i))$ can be executed if no
+$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
+probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
+is in the root conjunct, or it occurs in $T_k$ for some executed
+$\actions(p_k;(T_k))$.
 %%
@@
 successor state $s^v$ with probability $p_i$, and $s^\bot$ with
-probability $1 - p_i$. Here, $s^v$ is characterised by the union of
-$s$ and atomic terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
+probability $1 - p_i$. Here, $s^v$ is the union of $s$ and atomic
+terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
 $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
@@
 Because $\state_0$ is abstract, and because actions can have
-preconditions that are not necessarily true in the belief-state, we
-must make two further adjustments to our sequential continual
-planning. First, if $\action\in\actions$ has a precondition that is
-not true with probability $1$ in the current state, then before it is
-executed blindly, continual planning switches to a decision-theoretic
-mode, that shall be discussed in detail below. Therefore, by
-scheduling $\action$ the sequential planner makes a naive assumption
-about the possibility of the agent eventually holding strong beliefs
-about the truth values of specific state propositions. Second, if the
-truth value of $\prop$ is not known, and no assumption is made about
-it in the current abstract state, then if $\prop \in \poss(\action)$
-we have that action $\action$ is not applicable.
+preconditions that are not uniformly true in the underlying
+(resp. abstract) belief-state, we must make two further adjustments to
+the problem posed to the sequential planner. First, if
+$\action\in\actions$ has a precondition that is not true with
+probability $1$, then before it is executed blindly, the sequential
+continual planning switches to a decision-theoretic mode, discussed in
+detail below. Overall, by scheduling $\action$ the sequential planner
+makes a naive assumption about the possibility of the agent eventually
+committing ---for the purpose of the current task--- to the truth
+values of specific state propositions. Second, if the truth value of
+$\prop$ is not known, and no assumption is made about it in the
+current abstract state, then if $\prop \in \poss(\action)$ we have
+that action $\action$ is not applicable.
 
-Finally, we come to the issue of optimisation criteria. Here, we
-propose the value of an $N$ horizon trace ---something we expect the
-sequential planner to maximise--- corresponds to the maximal value of
-that trace. In detail, where $p_i$ is the probability that the
-$i^{th}$ execution is not a transition to $s^\bot$, we value a trace
-according to:
+%% In other words, the value of a proposition in an abstract
+%% state is either {\em true}, {\em false}, or {\em unspecified} (i.e.,
+%% \#). In the last case actions whose execution is preconditioned on the
+%% term being truth functional are not available.
+
+Finally, we come to the issue of valuing sequential plans, and
+therefore giving the criterion that the sequential planner seeks to
+optimise. Where $p_i$ is the probability that the $i^{th}$ action
+execution is not a transition to $s^\bot$, we value a trace according
+to:
 
@@
 \noindent In goal directed problems, where the only non-zero reward is
-received on the first transition to a goal state,
+received at the first transition to a goal state,
 Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
 infamous degenerate hindsight system
-\system{FFR$_a$}~\cite{yoon:etal:2007}. Moreover, form the perspective
-of gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP, our optimal trace yields a maximal gradient
-contribution under reasonable assumptions.
+\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
+gradient-based reinforcement learning systems, such as William's
+REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
+under reasonable assumptions the optimal trace yields a maximal
+gradient step from a uniform soft-max policy.
 
@@
 proposition is requested.
+
+The sequential mode of our system always schedules decision-theoretic
+planning before executing actions that are not applicable with
+certainty. This is vastly inefficient if the utility of that plan is
+not dependent on the actions successful execution. This is the case
+where a plan is conformant. Therefore, in the future we must develop a
+cheap (heuristic) procedure to evaluate the utility of sensing at
+steps in sequential plan.
 

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
Split file into sections to aid multiple editors.

--- 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}%
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Sequential and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Moritz's Initial Outline}
-
-\subsection{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\subsection{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\subsection{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\subsection{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Introduction}
-
-Any POMDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for sequential planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
-..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
-applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-propositions from $\add(\detAction)$ are added to $\state$, and those
-from $\delete(\detAction)$ are removed from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$, that it has no
-effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
-\state_i$.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs are
-negative rewards.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition.
-
-Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a box, at a given place. We
-can model that deterministic action using the following operator
-schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{R2D2}$ might
-be described thusly:
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} if its conjunctive parent is visited, and a
-conjunctive term is visited if all its immediate subterms are
-visited. A probabilistic term is visited if exactly one of its
-subterms, $T_i$, is visited. Each valid visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-\section{Switching Planner}
-
-We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.
-
-\subsection{Sequential Planning in POMDPs}
-
-Our approach uses sequential planning to compute a linear plan, and
-one trace of that plan's execution. The latter encapsulates
-assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of eventually holding
-strong beliefs about the truth values of specific state
-propositions. Sequential planning proceeds in a continual manner until
-the verification of an assumption about a runtime state proposition is
-scheduled. At that point the system switches to a decision-theoretic
-planner that is used to tailor sensory processing by the robot to test
-the validity of the assumptions so far.
-
-The first actions sequenced by the planner are virtual, and correspond
-to identifying an abstract start state. That prefix corresponds to a
-{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
-block. In the relaxed case, we can modify the initial definition so
-that a conjunctive term is visited provided all its atomic subterms
-are visited, and zero or more of its probabilistic terms are
-visited. Writing \#\ if the value of a proposition is unspecified by a
-the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
-following abstract starting states:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-.7 & kitchen & \# &  kitchen\\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
-
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
-
-moreover, for the purpose of decision-theoretic abstractions,
-
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
-
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
-
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
-
-\section{Evaluation}
-
-\section{Discussion and Related Work}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
 From an automated planning perspective, the problem of practical
@@
 quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
-%% $\selection_\node(\action) = P(\action | \node)$, and
-%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
-%% \action, \observation)$. Where the value of acting is the discounted
-%% accumulated reward over an infinite horizon, then each controller can
-%% be assigned a value, written $V_\node(\state)$, expressed in the usual
-%% way according to the Bellman equation:
-
-%% \begin{equation}\label{eq:evaluation}
-%% \begin{array}{lcl}
-%% V_\node(\state) & = & \sum_{\action \in \actions}
-%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-%% && \hspace{-10ex} \beta \sum_{\action, \observation,
-%% \state', \node'} \transition_\node(\action, \observation, \node')
-%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
-%% \action) V_{\node'}(\state')
-%% \end{array}
-%% \end{equation}
-
-%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
-%% then:
-
-%% \begin{equation} \label{eq:valueBelief}
-%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-%% \end{equation}
-
-%% \noindent There is a corresponding deterministic FSCs for both
-%% sequential (i.e., conformant) and contingent
-%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
-%% case, if we take $\beta = 1$, and alter the process dynamics so there
-%% is a compulsory transition to a zero utility sink state when that
-%% horizon is reached.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/discussion.tex
Split file into sections to aid multiple editors.

--- 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Serial and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Any (PO)MDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
+There have been a number of developments recently towards planning
+under uncertainty using systems that were intended for sequential
+planning in deterministic problems.  Notably,
+\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
+probabilistic track of the 2004 International Planning Competition.
+In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
+procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
+plans and corresponding execution traces.
 %%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for serial planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to simulate DTPDDL instances, and modified
-\fastdownward\ so that it can find useful serial executions given
-DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality serial plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\section{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\section{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\section{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
-chosen by nature when \stochAction\ is applied at state $\state$, then
-the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
-\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
-added to $\state$, and those from $\delete(\detAction)$ are removed
-from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A percept is active, written $\stochSense \in
-\stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$ where
-$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
-$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
-0$.
-
-\subsection{Costs, Rewards, and Plan Evaluation}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function.  Consequently there is a positive constant $c$
-so that for all $\state \in \states$ and $\action \in \actions$,
-$|\reward(\state, \action)| < c$.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-Successive state estimation is available if we applying Bayes' rule.
-Taking the current belief $\bstate$ as our {\em prior}, and supposing
-action $\action$ is executed with perceptive outcome $\observ$, then
-the probability that we are in $\state$ in the successive belief-state
-$\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-A general formalism for representing solutions to POMDPs is the
-finite-state controller (FSC). This is a three-tuple $\langle \nodes,
-\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
-nodes, $\selection_\node(\action) = P(\action | \node)$, and
-$\transition_\node(\action, \observation, \node') = P(\node'|\node,
-\action, \observation)$. The value of state $\state$ at node $\node$
-given a problem model, written $V_\node(\state)$, is:
-
-\begin{equation}\label{eq:evaluation}
-\begin{array}{lcl}
-V_\node(\state) & = & \sum_{\action \in \actions}
-\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-&& \hspace{-10ex} \beta \sum_{\action, \observation,
-\state', \node'} \transition_\node(\action, \observation, \node')
-\transProb(\state, \action, \state') \obsDist_\observ(\state',
-\action) V_{\node'}(\state')
-\end{array}
-\end{equation}
-
-\noindent The value of $\bstate$ according to a controller is
-therefore:
-
-\begin{equation} \label{eq:valueBelief}
-V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-\end{equation}
-
-\section{Serial Planning in Determinisations}
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes}--- are treated as control-knowledge
-by the planner. Here, a plan is not considered valid unless in the
-finial stet the goal condition is satisfied.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to PPDDL
-syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose our
-robot is tasked with exploring {\em locations} in order to identify
-the whereabouts of a {\em visual-object}. We must describe state and
-perceptual facts that model the {\em true}, resp. perceived, locations
-of objects. In DTPDDL, these declarations appear as:
-
-\small{
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-}
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition. For example, our robot is
-able to look for a visual object, such as a box of cornflakes, at a
-given place. We can model that deterministic action using the
-following operator schema:
-
-\small{
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small{
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-}
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid our discussion of
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$, or
-(3) a conjunct over probabilistic and atomic terms. The root term is
-conjunctive, and the leaves are atomic. For example, the starting
-distribution for $\pp{R2D2}$ might be described thusly:
-
-\small{
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-}
-
-\noindent The interpretation of such an expression can be given
-according to a recursive descent.
-
-\section{Evaluation}
-
-\section{Discussion}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for serial planning
-in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition,
-uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
-corresponding execution trace. The resulting plan is then executed
-until the observed trace deviates from the planned trace.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something
-
 More computationally expensive approaches in this vein combine
@@
 deterministic planning procedures. The outcome is typically a more
-robust serial plan~\cite{yoon:etal:2008}, or contingent
+robust sequential plan~\cite{yoon:etal:2008}, or contingent
 plan~\cite{majercik:2006}.
-
 %%
-Recently the systems
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
-planning ---i.e., serial planning in unobservable worlds--- can be
-modelled as a deterministic problem, and therefore solved using serial
+Also leveraging deterministic planners in problems that feature
+uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
+---i.e., sequential planning in unobservable worlds--- can be modelled
+as a deterministic problem, and therefore solved using sequential
 systems. In this conformant setting, advances have been towards
 compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. This
-research thread has also been extended recently to contingent planning
+search planning procedures, and lazy evaluations of beliefs. Most
+recently this research thread has been extended to contingent planning
 in fully observable non-deterministic
 environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, serial planning is
-responsible for an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until an assumption about a runtime proposition
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the serial planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast serial
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planing for contingencies unless
-they eventuate during execution.
-
-computing a single serial plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Split file into sections to aid multiple editors.

--- 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Serial and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Any (PO)MDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
 We describe, \pcogx, the planning component in a robotic system that
@@
 finite-state controllers.
-
-\section{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\section{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\section{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\section{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
-chosen by nature when \stochAction\ is applied at state $\state$, then
-the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
-\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
-added to $\state$, and those from $\delete(\detAction)$ are removed
-from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A percept is active, written $\stochSense \in
-\stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$ where
-$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
-$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
-0$.
-
-\subsection{Costs, Rewards, and Plan Evaluation}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function.  Consequently there is a positive constant $c$
-so that for all $\state \in \states$ and $\action \in \actions$,
-$|\reward(\state, \action)| < c$.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-Successive state estimation is available if we applying Bayes' rule.
-Taking the current belief $\bstate$ as our {\em prior}, and supposing
-action $\action$ is executed with perceptive outcome $\observ$, then
-the probability that we are in $\state$ in the successive belief-state
-$\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-A general formalism for representing solutions to POMDPs is the
-finite-state controller (FSC). This is a three-tuple $\langle \nodes,
-\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
-nodes, $\selection_\node(\action) = P(\action | \node)$, and
-$\transition_\node(\action, \observation, \node') = P(\node'|\node,
-\action, \observation)$. The value of state $\state$ at node $\node$
-given a problem model, written $V_\node(\state)$, is:
-
-\begin{equation}\label{eq:evaluation}
-\begin{array}{lcl}
-V_\node(\state) & = & \sum_{\action \in \actions}
-\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-&& \hspace{-10ex} \beta \sum_{\action, \observation,
-\state', \node'} \transition_\node(\action, \observation, \node')
-\transProb(\state, \action, \state') \obsDist_\observ(\state',
-\action) V_{\node'}(\state')
-\end{array}
-\end{equation}
-
-\noindent The value of $\bstate$ according to a controller is
-therefore:
-
-\begin{equation} \label{eq:valueBelief}
-V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-\end{equation}
-
-\section{Serial Planning in Determinisations}
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes}--- are treated as control-knowledge
-by the planner. Here, a plan is not considered valid unless in the
-finial stet the goal condition is satisfied.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to PPDDL
-syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose our
-robot is tasked with exploring {\em locations} in order to identify
-the whereabouts of a {\em visual-object}. We must describe state and
-perceptual facts that model the {\em true}, resp. perceived, locations
-of objects. In DTPDDL, these declarations appear as:
-
-\small{
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-}
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition. For example, our robot is
-able to look for a visual object, such as a box of cornflakes, at a
-given place. We can model that deterministic action using the
-following operator schema:
-
-\small{
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small{
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-}
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid our discussion of
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$, or
-(3) a conjunct over probabilistic and atomic terms. The root term is
-conjunctive, and the leaves are atomic. For example, the starting
-distribution for $\pp{R2D2}$ might be described thusly:
-
-\small{
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-}
-
-\noindent The interpretation of such an expression can be given
-according to a recursive descent.
-
-\section{Evaluation}
-
-\section{Discussion}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for serial planning
-in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition,
-uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
-corresponding execution trace. The resulting plan is then executed
-until the observed trace deviates from the planned trace.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something
-
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust serial plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%
-Recently the systems
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
-planning ---i.e., serial planning in unobservable worlds--- can be
-modelled as a deterministic problem, and therefore solved using serial
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. This
-research thread has also been extended recently to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, serial planning is
-responsible for an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until an assumption about a runtime proposition
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the serial planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast serial
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planing for contingencies unless
-they eventuate during execution.
-
-computing a single serial plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Split file into sections to aid multiple editors.

--- 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
 
-\section{Introduction}
+\section{Moritz's Initial Outline}
 
-Any (PO)MDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for serial planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to simulate DTPDDL instances, and modified
-\fastdownward\ so that it can find useful serial executions given
-DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality serial plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Switching Planner Outline}
+\subsection{Switching Planner Outline}
 
@@
 
-\section{Determinisation}
+\subsection{Determinisation}
 
@@
 
-\section{Subtask Generation}
+\subsection{Subtask Generation}
 \begin{itemize}
@@
 
-\section{Belief revision}
+\subsection{Belief revision}
 Not sure if this needs its own section
@@
 
+\section{Introduction}
+\input{intro.tex}
+
 \section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
-chosen by nature when \stochAction\ is applied at state $\state$, then
-the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
-\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
-added to $\state$, and those from $\delete(\detAction)$ are removed
-from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A percept is active, written $\stochSense \in
-\stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$ where
-$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
-$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
-0$.
-
-\subsection{Costs, Rewards, and Plan Evaluation}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function.  Consequently there is a positive constant $c$
-so that for all $\state \in \states$ and $\action \in \actions$,
-$|\reward(\state, \action)| < c$.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-Successive state estimation is available if we applying Bayes' rule.
-Taking the current belief $\bstate$ as our {\em prior}, and supposing
-action $\action$ is executed with perceptive outcome $\observ$, then
-the probability that we are in $\state$ in the successive belief-state
-$\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-A general formalism for representing solutions to POMDPs is the
-finite-state controller (FSC). This is a three-tuple $\langle \nodes,
-\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
-nodes, $\selection_\node(\action) = P(\action | \node)$, and
-$\transition_\node(\action, \observation, \node') = P(\node'|\node,
-\action, \observation)$. The value of state $\state$ at node $\node$
-given a problem model, written $V_\node(\state)$, is:
-
-\begin{equation}\label{eq:evaluation}
-\begin{array}{lcl}
-V_\node(\state) & = & \sum_{\action \in \actions}
-\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-&& \hspace{-10ex} \beta \sum_{\action, \observation,
-\state', \node'} \transition_\node(\action, \observation, \node')
-\transProb(\state, \action, \state') \obsDist_\observ(\state',
-\action) V_{\node'}(\state')
-\end{array}
-\end{equation}
-
-\noindent The value of $\bstate$ according to a controller is
-therefore:
-
-\begin{equation} \label{eq:valueBelief}
-V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-\end{equation}
-
-\section{Serial Planning in Determinisations}
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
-knows where the {\em cornflakes}--- are treated as control-knowledge
-by the planner. Here, a plan is not considered valid unless in the
-finial stet the goal condition is satisfied.
+\input{prop-dtp.tex}
 
 \section{Decision-Theoretic PDDL}
+\input{dtpddl.tex}
 
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to PPDDL
-syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose our
-robot is tasked with exploring {\em locations} in order to identify
-the whereabouts of a {\em visual-object}. We must describe state and
-perceptual facts that model the {\em true}, resp. perceived, locations
-of objects. In DTPDDL, these declarations appear as:
-
-\small{
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-}
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition. For example, our robot is
-able to look for a visual object, such as a box of cornflakes, at a
-given place. We can model that deterministic action using the
-following operator schema:
-
-\small{
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small{
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-}
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid our discussion of
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$, or
-(3) a conjunct over probabilistic and atomic terms. The root term is
-conjunctive, and the leaves are atomic. For example, the starting
-distribution for $\pp{R2D2}$ might be described thusly:
-
-\small{
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-}
-
-\noindent The interpretation of such an expression can be given
-according to a recursive descent.
+\section{Switching Planner}
+\input{switching.tex}
 
 \section{Evaluation}
+\input{evaluation.tex}
 
-\section{Discussion}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for serial planning
-in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition,
-uses the fast satisficing procedure
-\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
-corresponding execution trace. The resulting plan is then executed
-until the observed trace deviates from the planned trace.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something
-
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust serial plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%
-Recently the systems
-\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
-planning ---i.e., serial planning in unobservable worlds--- can be
-modelled as a deterministic problem, and therefore solved using serial
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. This
-research thread has also been extended recently to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
+\section{Discussion and Related Work}
+\input{discussion.tex}
 
 \section{Concluding Remarks}
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, serial planning is
-responsible for an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until an assumption about a runtime proposition
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the serial planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast serial
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planing for contingencies unless
-they eventuate during execution.
-
-computing a single serial plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
+\input{conc.tex}
 

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
Split file into sections to aid multiple editors.

--- 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}%
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Sequential and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Moritz's Initial Outline}
-
-\subsection{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\subsection{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\subsection{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\subsection{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Introduction}
-
-Any POMDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for sequential planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Propositional Decision-Theoretic Planning}
-
 %% {\em runtime variables}
@@
 both sequential (e.g., conformant) and contingent plans.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition.
-
-Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a box, at a given place. We
-can model that deterministic action using the following operator
-schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{R2D2}$ might
-be described thusly:
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} if its conjunctive parent is visited, and a
-conjunctive term is visited if all its immediate subterms are
-visited. A probabilistic term is visited if exactly one of its
-subterms, $T_i$, is visited. Each valid visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-\section{Switching Planner}
-
-We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.
-
-\subsection{Sequential Planning in POMDPs}
-
-Our approach uses sequential planning to compute a linear plan, and
-one trace of that plan's execution. The latter encapsulates
-assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of eventually holding
-strong beliefs about the truth values of specific state
-propositions. Sequential planning proceeds in a continual manner until
-the verification of an assumption about a runtime state proposition is
-scheduled. At that point the system switches to a decision-theoretic
-planner that is used to tailor sensory processing by the robot to test
-the validity of the assumptions so far.
-
-The first actions sequenced by the planner are virtual, and correspond
-to identifying an abstract start state. That prefix corresponds to a
-{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
-block. In the relaxed case, we can modify the initial definition so
-that a conjunctive term is visited provided all its atomic subterms
-are visited, and zero or more of its probabilistic terms are
-visited. Writing \#\ if the value of a proposition is unspecified by a
-the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
-following abstract starting states:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-.7 & kitchen & \# &  kitchen\\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
-
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
-
-moreover, for the purpose of decision-theoretic abstractions,
-
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
-
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
-
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
-
-\section{Evaluation}
-
-\section{Discussion and Related Work}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the sequential planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast sequential
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planning for contingencies unless
-they eventuate during execution.
-
-computing a single sequential plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
-%% $\selection_\node(\action) = P(\action | \node)$, and
-%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
-%% \action, \observation)$. Where the value of acting is the discounted
-%% accumulated reward over an infinite horizon, then each controller can
-%% be assigned a value, written $V_\node(\state)$, expressed in the usual
-%% way according to the Bellman equation:
-
-%% \begin{equation}\label{eq:evaluation}
-%% \begin{array}{lcl}
-%% V_\node(\state) & = & \sum_{\action \in \actions}
-%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-%% && \hspace{-10ex} \beta \sum_{\action, \observation,
-%% \state', \node'} \transition_\node(\action, \observation, \node')
-%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
-%% \action) V_{\node'}(\state')
-%% \end{array}
-%% \end{equation}
-
-%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
-%% then:
-
-%% \begin{equation} \label{eq:valueBelief}
-%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-%% \end{equation}
-
-%% \noindent There is a corresponding deterministic FSCs for both
-%% sequential (i.e., conformant) and contingent
-%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
-%% case, if we take $\beta = 1$, and alter the process dynamics so there
-%% is a compulsory transition to a zero utility sink state when that
-%% horizon is reached.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Split file into sections to aid multiple editors.

--- 2010-11-22 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}%
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Sequential and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Moritz's Initial Outline}
-
-\subsection{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\subsection{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\subsection{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\subsection{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Introduction}
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for sequential planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
-..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
-applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-propositions from $\add(\detAction)$ are added to $\state$, and those
-from $\delete(\detAction)$ are removed from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$, that it has no
-effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
-\state_i$.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs are
-negative rewards.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition.
-
-Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a box, at a given place. We
-can model that deterministic action using the following operator
-schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{R2D2}$ might
-be described thusly:
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} iff its conjunctive parent is visited, and a
-conjunctive term is visited iff all its immediate subterms are
-visited. A probabilistic term is visited iff its conjunctive parent is
-visited, and exactly one of its subterms, $T_i$, is visited. Each
-valid visitation of the root term according to this recursive
-definition encapsulates a starting state, along with the probability
-it occurs. The former corresponds to the union of all visited atoms,
-and the latter corresponds to the product of $p_i$ entries on the
-visited subterms of probabilistic elements. Making these ideas
-concrete, our example yields the following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-\section{Switching Planner}
-
 We now describe a continual planning approach that switches between
@@
 assumption about a state proposition.
-
-\section{Evaluation}
-
-\section{Discussion and Related Work}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
-
-The sequential mode of our system always schedules decision-theoretic
-planning before executing actions that are not applicable with
-certainty. This is vastly inefficient if the utility of that plan is
-not dependent on the actions successful execution. This is the case
-where a plan is conformant. Therefore, in the future we must develop a
-cheap (heuristic) procedure to evaluate the utility of sensing at
-steps in sequential plan.
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the sequential planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast sequential
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planning for contingencies unless
-they eventuate during execution.
-
-computing a single sequential plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
-%% $\selection_\node(\action) = P(\action | \node)$, and
-%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
-%% \action, \observation)$. Where the value of acting is the discounted
-%% accumulated reward over an infinite horizon, then each controller can
-%% be assigned a value, written $V_\node(\state)$, expressed in the usual
-%% way according to the Bellman equation:
-
-%% \begin{equation}\label{eq:evaluation}
-%% \begin{array}{lcl}
-%% V_\node(\state) & = & \sum_{\action \in \actions}
-%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-%% && \hspace{-10ex} \beta \sum_{\action, \observation,
-%% \state', \node'} \transition_\node(\action, \observation, \node')
-%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
-%% \action) V_{\node'}(\state')
-%% \end{array}
-%% \end{equation}
-
-%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
-%% then:
-
-%% \begin{equation} \label{eq:valueBelief}
-%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-%% \end{equation}
-
-%% \noindent There is a corresponding deterministic FSCs for both
-%% sequential (i.e., conformant) and contingent
-%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
-%% case, if we take $\beta = 1$, and alter the process dynamics so there
-%% is a compulsory transition to a zero utility sink state when that
-%% horizon is reached.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something
-
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
-
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
-
-moreover, for the purpose of decision-theoretic abstractions,
-
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
-
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.

### 2010-11-22 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
Split file into sections to aid multiple editors.

--- 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}%
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Sequential and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Moritz's Initial Outline}
-
-\subsection{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\subsection{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\subsection{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\subsection{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Introduction}
-
-Any POMDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for sequential planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
-..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
-applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-propositions from $\add(\detAction)$ are added to $\state$, and those
-from $\delete(\detAction)$ are removed from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$, that it has no
-effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
-\state_i$.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs are
-negative rewards.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
-
-\section{Decision-Theoretic PDDL}
-
 The modelling language of choice for planning in probabilistic
@@
 \normalsize
-
-\section{Switching Planner}
-
-We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.
-
-\subsection{Sequential Planning in POMDPs}
-
-Our approach uses sequential planning to compute a linear plan, and
-one trace of that plan's execution. The latter encapsulates
-assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of eventually holding
-strong beliefs about the truth values of specific state
-propositions. Sequential planning proceeds in a continual manner until
-the verification of an assumption about a runtime state proposition is
-scheduled. At that point the system switches to a decision-theoretic
-planner that is used to tailor sensory processing by the robot to test
-the validity of the assumptions so far.
-
-The first actions sequenced by the planner are virtual, and correspond
-to identifying an abstract start state. That prefix corresponds to a
-{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
-block. In the relaxed case, we can modify the initial definition so
-that a conjunctive term is visited provided all its atomic subterms
-are visited, and zero or more of its probabilistic terms are
-visited. Writing \#\ if the value of a proposition is unspecified by a
-the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
-following abstract starting states:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-.7 & kitchen & \# &  kitchen\\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
-
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
-
-moreover, for the purpose of decision-theoretic abstractions,
-
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
-
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
-
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
-
-\section{Evaluation}
-
-\section{Discussion and Related Work}
-
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the sequential planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast sequential
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planning for contingencies unless
-they eventuate during execution.
-
-computing a single sequential plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
-%% $\selection_\node(\action) = P(\action | \node)$, and
-%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
-%% \action, \observation)$. Where the value of acting is the discounted
-%% accumulated reward over an infinite horizon, then each controller can
-%% be assigned a value, written $V_\node(\state)$, expressed in the usual
-%% way according to the Bellman equation:
-
-%% \begin{equation}\label{eq:evaluation}
-%% \begin{array}{lcl}
-%% V_\node(\state) & = & \sum_{\action \in \actions}
-%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-%% && \hspace{-10ex} \beta \sum_{\action, \observation,
-%% \state', \node'} \transition_\node(\action, \observation, \node')
-%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
-%% \action) V_{\node'}(\state')
-%% \end{array}
-%% \end{equation}
-
-%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
-%% then:
-
-%% \begin{equation} \label{eq:valueBelief}
-%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-%% \end{equation}
-
-%% \noindent There is a corresponding deterministic FSCs for both
-%% sequential (i.e., conformant) and contingent
-%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
-%% case, if we take $\beta = 1$, and alter the process dynamics so there
-%% is a compulsory transition to a zero utility sink state when that
-%% horizon is reached.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Tiny changes to what Richard posted.
 - Charles

@@
 
-Finally, we come to the issue of valuing sequential plans, and
-therefore giving the criterion that the sequential planner seeks to
-optimise. Where $p_i$ is the probability that the $i^{th}$ action
-execution is not a transition to $s^\bot$, we value a trace according
-to:
+Finally, we come to valuing sequential plans, and therefore giving a
+criterion that the sequential planner seeks to optimise. Where $p_i$
+is the probability that the $i^{th}$ action execution is not a
+transition to $s^\bot$, we value a trace according to:
 
@@
 
-\noindent In goal directed problems, where the only non-zero reward is
-received at the first transition to a goal state,
+\noindent In other words, the value of a trace captures its
+significance, in relation to other possible realisations of the plan,
+to the overall expected value of the plan in
+Eq~\ref{eq:expectedvalue}. In goal directed problems, where the only
+non-zero reward is received at the first transition to a goal state,
 Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
@@
 REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace yields a maximal
-gradient step from a uniform soft-max policy.
+under reasonable assumptions the optimal trace identifies a maximal
+gradient step from a uniformly random soft-max policy.
+
+%% Finally, we come to the issue of valuing sequential plans, and
+%% therefore giving the criterion that the sequential planner seeks to
+%% optimise. Where $p_i$ is the probability that the $i^{th}$ action
+%% execution is not a transition to $s^\bot$, we value a trace according
+%% to:
+
+%% \begin{equation}\label{eq:tracevalue}
+%% V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
+%% \end{equation}
+
+%% \noindent In goal directed problems, where the only non-zero reward is
+%% received at the first transition to a goal state,
+%% Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
+%% infamous degenerate hindsight system
+%% \system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
+%% gradient-based reinforcement learning systems, such as William's
+%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
+%% under reasonable assumptions the optimal trace yields a maximal
+%% gradient step from a uniform soft-max policy.
 

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
spelling
the value stuff clarrified
 - Charles

@@
 Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
-successor state $s^v$ with probability $p_i$, and $s^\bot$ with
-probability $1 - p_i$. Here, $s^v$ is the union of $s$ and atomic
+successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
+probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
 terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
@@
 Finally, we come to valuing sequential plans, and therefore giving a
-criterion that the sequential planner seeks to optimise. Where $p_i$
-is the probability that the $i^{th}$ action execution is not a
-transition to $s^\bot$, we value a trace according to:
+criteria for the sequential planner to optimise. Where $p_i$ is the
+probability that the $i^{th}$ sequenced action does not transition to
+$s^\bot$, we value a trace as follows:
 
@@
 
-\noindent In other words, the value of a trace captures its
-significance, in relation to other possible realisations of the plan,
-to the overall expected value of the plan in
-Eq~\ref{eq:expectedvalue}. In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
-infamous degenerate hindsight system
+\noindent The optimal trace given a plan, i.e., the non-assumptive
+actions in sequence, therefore has value:
+
+\[
+V^* = \max_{p_1, .., p_n} \prod_{i=1..N} p_i \sum_{i=1..N-1}
+\reward(\state_i, \action_i),
+\]
+
+\noindent which is equal to the maximal contribution a trace of the plan can
+make to Eq~\ref{eq:expectedvalue}. In other words, writing $V_i$ for
+the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
+s_1, a_1,.., s_N$, we have:
+
+\[
+V^* = \arg\min_{V_i} \Expect \bigg{[} 
+\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate = \bstate_0
+\bigg{]} - V_i
+\]
+
+In goal directed problems, where the only non-zero reward is received
+at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
+criteria corresponds exactly to that of the infamous degenerate
+hindsight system
 \system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of

### 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Richard missed some of my changes...
 - Charles

@@
 makes a naive assumption about the possibility of the agent eventually
-committing ---for the purpose of the current task--- to the truth
-values of specific state propositions. Second, if the truth value of
-$\prop$ is not known, and no assumption is made about it in the
-current abstract state, then if $\prop \in \poss(\action)$ we have
-that action $\action$ is not applicable.
+``committing'' to the truth values of specific state
+propositions. Thus, the role of decision-theoretic plans is to verify
+the validity of assumptions about the runtime state in sequential
+planning. Second, if the truth value of $\prop$ is not known, and no
+assumption is made about it in the current abstract state, then if
+$\prop \in
+\poss(\action)$ we have that action $\action$ is not applicable.
+
+%% Because $\state_0$ is abstract, and because actions can have
+%% preconditions that are not uniformly true in the underlying
+%% (resp. abstract) belief-state, we must make two further adjustments to
+%% the problem posed to the sequential planner. First, if
+%% $\action\in\actions$ has a precondition that is not true with
+%% probability $1$, then before it is executed blindly, the sequential
+%% continual planning switches to a decision-theoretic mode, discussed in
+%% detail below. Overall, by scheduling $\action$ the sequential planner
+%% makes a naive assumption about the possibility of the agent eventually
+%% committing ---for the purpose of the current task--- to the truth
+%% values of specific state propositions. Second, if the truth value of
+%% $\prop$ is not known, and no assumption is made about it in the
+%% current abstract state, then if $\prop \in \poss(\action)$ we have
+%% that action $\action$ is not applicable.
 
@@
 the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
-s_1, a_1,.., s_N$, we have:
+s_1, a_1,.., s_N$, we have
 
@@
 V^* = \arg\min_{V_i} \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate = \bstate_0
-\bigg{]} - V_i
+\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate_0
+\bigg{]} - V_i.
 \]

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
Might still have some algebra wrong, but I think I did some repairing of the reward sketch.
 - Charles

@@
 V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
+\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
 \end{equation}
 
-\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
+\noindent Where $b_t$ is the belief-state at step $t$, $\pp{FSC}_t$ is
+the action prescribed at step $t$, and
+
+\[\reward(\bstate, \action) = \sum_{\state \in \states}
+\bstate(\state)\reward(\state, \action).\]
+
+\noindent Finally, it is
 useful to note that there is a corresponding deterministic FSCs for

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
grammar
I think I have understoof Moritz's code, but I may not have...
 - Charles

--- 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
@@
 visited. The starting state, $\state_0$, for sequential planning is
-the abstract state with probability $1.0$ -- E.G., for $\pp{R2D2}$
-$s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
+the unique abstract state with probability $1.0$ -- E.G., for
+$\pp{R2D2}$ $s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
 
@@
 \begin{equation}\label{eq:tracevalue}
-V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
+V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N-1} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
 \end{equation}
@@
 \[
-V^* = \max_{p_1, .., p_n} \prod_{i=1..N} p_i \sum_{i=1..N-1}
+V^* = \max_{p_1, .., p_n} \prod_{i=1..N-1} p_i \sum_{i=1..N-1}
 \reward(\state_i, \action_i),
@@
 V^* = \arg\min_{V_i} \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate = \bstate_0
-\bigg{]} - V_i
+\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
+\bigg{]} - V_i.
 \]
@@
 under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy.
+gradient step from a uniformly random soft-max policy. Finally, it is
+worth clarifying that although the above discussion has been in terms
+of a trace of length $N$, there is nothing in our approach that
+artificially limits the length of candidate sequential plans.
 
@@
 \subsection{Decision-Theoretic Planning in Abstractions}
+
+In the decision-theoretic mode, the switching planner solves a sensing
+problem in an abstract processes defined in terms of the assumptive
+actions in the current sequential plan. Therefore, this mode uses a
+different utility model, that reflects the value of performing sensing
+actions.
+
+We construct a {\em relaxed} $(\pp{:init})$ block by first
+constructing a tree which only includes terms from the {\em relaxed}
+visitation characterised by the {\em assumptive} actions in the
+current sequential plan. For example, if $\pp{R2D2}$'s plan is:
+
+\[
+\begin{array}{c}
+\actions(.8,(=(\pp{is-in}~\pp{box})\pp{office}));\\
+(pp{reportlook-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
+(\pp{report-is-in}~\pp{box}~\pp{office})
+\end{array}
+\]
+
+\noindent Then the {\em relaxed} block will be,
+
+\small
+\begin{tabtt}
+(\=:init (= (is-in R2D2) kitchen) \+ \\
+       (probabilistic \=.8 (= (is-in box) office)  \\
+		      \>.2 (= (is-in box) kitchen))) \\
+\end{tabtt}
+\normalsize
+
+\noindent and therefore, the belief-state $\bstate_0$ is:
+
+\small
+\begin{tabular}{cccc}
+\hline
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+\hline
+.8 & kitchen & office & \# \\
+.2 & kitchen & kitchen & \# \\
+\hline
+\end{tabular}
+\normalsize
+
+\noindent At this point, decision-theoretic planning would proceed in a cupless
+abstraction of the environment, where the inclusion (exclusion) of
+POMDP actions is the same as we defined it for the sequential mode.
+
+In the second step, we iteratively add terms from the original
+$(\pp{:init})$ block while the resulting state space remains of a
+practicable size. Of course, it is expensive to measure size exactly,
+thus in practice we estimate it to be $2^{|\propositions^\#|}$, where
+$\propositions^\#$ is the set of propositions in the abstraction
+induced by the proposed tree. Terms are proposed for addition, because
+they are probabilistic, and because they were omitted in the {\em
+relaxed} visitation characterised by {\em assumptive} actions in the
+sequential plan. Each iteration, we add one candidate term with the
+largest number of parent terms.
+
+Terms a proposed sequentially, with preference given to
+terms that are closest to, or which otherwise are, the conjunctive
+leaves.
 

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
typos

@@
 \[
-\begin{array}{c}
-\actions(.8,(=(\pp{is-in}~\pp{box})\pp{office}));\\
-(pp{reportlook-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
+\begin{array}{l}
+\actions(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
+(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
 (\pp{report-is-in}~\pp{box}~\pp{office})

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Another macro, and some more sketching for DTP

@@
 In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
-for each element, $p_i (T_i)$, of each probabilistic term from
+add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
+each element, $p_i (T_i)$, of each probabilistic term from
 $(:init)$. The physics of these actions is as
-follows. $\actions(p_i;(T_i))$ can be executed if no
-$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
-probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
-is in the root conjunct, or it occurs in $T_k$ for some executed
-$\actions(p_k;(T_k))$.
+follows. $\assumptiveS(i)$ can be executed if no $\assumptiveS(j)$,
+$j \neq i$, has been executed from the same probabilistic term, and,
+either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
+or it occurs in $T_k$ for some executed $\assumptiveS(k)$.
 %%
-Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
+Executing $\assumptiveS(i)$ in $s$ effects a transition to a
 successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
@@
 problem in an abstract processes defined in terms of the assumptive
-actions in the current sequential plan. Therefore, this mode uses a
-different utility model, that reflects the value of performing sensing
-actions.
+actions in the current sequential plan, and the $(\pp{:init})$ block
+from the original problem description.
 
-We construct a {\em relaxed} $(\pp{:init})$ block by first
+The focus for decision-theoretic planning is on sensing, rather than
+the achievement of utility according to the given objectives. The
+process in this mode has a different reward model that reflects the
+value of performing sensing actions in the context of a trace proposed
+for the preceding sequential deliberation. In detail, for each
+$\assumptiveS{i}$ action scheduled in the current sequential plan, we
+have a {\em dual} $\assumptiveDT{i}$ so that
+$\forall\state\in\states$:
+
+\[
+\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
+\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
+-\$(T_i) & \pp{otherwise} \\
+\end{array}
+\]
+
+\noindent Where $\$(T_i)$ is a positive numeric quantity which
+rewards the agent for correctly jettisoning a false assumption, and
+otherwise penalises the agent. Propositionally these dual actions have
+no preconditions or effects -- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
+\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Additionally, for the action,
+$\action_i$, whose scheduled execution triggered the switch to the
+decision-theoretic mode, we include an action
+$\actions.\poss(\action_i)$ with the reward property:
+
+\[
+\reward(\state, \actions.\poss(\action_i)) = \bigg\{ \begin{array}{ll}
+\$(\poss(\action_i)) & \pp{if}~\;\; \poss(\action_i) \subseteq \state \\
+-\$(\poss(\action_i)) & \pp{otherwise} \\
+\end{array}
+\]
+
+\noindent Execution of either a $\assumptiveDT{i}$
+action, or $\actions.\poss(\action_i)$, returns control to the
+sequential planner. If a dual is executed, then the sequential planner
+must replan with $\bstate_0$ equal to the current belief. Otherwise,
+if $\actions.\poss(\action_i)$ is executed, then the current
+sequential plan is executed until further sensing is scheduled, or to
+completion. In our project the rewards, $\$:2^\propositions\to\Re$,
+for the POMDP posed to decision-theoretic planning is sourced from a
+motivational component of our robotic architecture.
+
+The starting belief-state for the abstract process is given in terms
+of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
 constructing a tree which only includes terms from the {\em relaxed}
@@
 \begin{array}{l}
-\actions(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
 (\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
@@
 
-In the second step, we iteratively add terms from the original
-$(\pp{:init})$ block while the resulting state space remains of a
-practicable size. Of course, it is expensive to measure size exactly,
-thus in practice we estimate it to be $2^{|\propositions^\#|}$, where
-$\propositions^\#$ is the set of propositions in the abstraction
-induced by the proposed tree. Terms are proposed for addition, because
-they are probabilistic, and because they were omitted in the {\em
-relaxed} visitation characterised by {\em assumptive} actions in the
-sequential plan. Each iteration, we add one candidate term with the
-largest number of parent terms.
-
-Terms a proposed sequentially, with preference given to
-terms that are closest to, or which otherwise are, the conjunctive
-leaves.
-
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
-
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
+In a second step, we iteratively refine the {\em relaxed} block by
+adding terms from the original block while the resulting state space
+remains of a practicable size. Of course, it is expensive to measure
+size exactly, thus in practice we estimate it to be
+$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
+propositions in the abstraction induced by the proposed tree. Terms
+are proposed for addition, because they are probabilistic, and because
+they were omitted in the {\em relaxed} visitation characterised by
+{\em assumptive} actions in the sequential plan. Each iteration, we
+add one candidate term with the largest number of parent terms already
+in the candidate tree.

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/discussion.tex
New section

@@
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
+A theoretical criticism of switching continual planning, concerns
+interleaved sequential and decision-theoretic sessions failing to make
+any progress towards the objective.

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
New section

--- 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-
-\include{macros}%
-
-\nocopyright
-
-\pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Sequential and
-%  Decision-Theoretic Planning} 
-\author{NA}
-\setcounter{secnumdepth}{0}
-
-\begin{document} 
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Moritz's Initial Outline}
-
-\subsection{Switching Planner Outline}
-
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
-
-\subsection{Determinisation}
-
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
-
-\subsection{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
-
-\subsection{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
-
-\section{Introduction}
-
-Any POMDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for sequential planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
-..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
-applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-propositions from $\add(\detAction)$ are added to $\state$, and those
-from $\delete(\detAction)$ are removed from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$, that it has no
-effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
-\state_i$.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs are
-negative rewards.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition.
-
-Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a box, at a given place. We
-can model that deterministic action using the following operator
-schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{R2D2}$ might
-be described thusly:
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} if its conjunctive parent is visited, and a
-conjunctive term is visited if all its immediate subterms are
-visited. A probabilistic term is visited if exactly one of its
-subterms, $T_i$, is visited. Each valid visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-\section{Switching Planner}
-
-We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.
-
-\subsection{Sequential Planning in POMDPs}
-
-Our approach uses sequential planning to compute a linear plan, and
-one trace of that plan's execution. The latter encapsulates
-assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of eventually holding
-strong beliefs about the truth values of specific state
-propositions. Sequential planning proceeds in a continual manner until
-the verification of an assumption about a runtime state proposition is
-scheduled. At that point the system switches to a decision-theoretic
-planner that is used to tailor sensory processing by the robot to test
-the validity of the assumptions so far.
-
-The first actions sequenced by the planner are virtual, and correspond
-to identifying an abstract start state. That prefix corresponds to a
-{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
-block. In the relaxed case, we can modify the initial definition so
-that a conjunctive term is visited provided all its atomic subterms
-are visited, and zero or more of its probabilistic terms are
-visited. Writing \#\ if the value of a proposition is unspecified by a
-the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
-following abstract starting states:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-.7 & kitchen & \# &  kitchen\\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
-
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
-
-moreover, for the purpose of decision-theoretic abstractions,
-
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
-
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
-
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
-
-\section{Evaluation}
-
-\section{Discussion and Related Work}
-
 There have been a number of developments recently towards planning
@@
 environments~\cite{albore:etal:2009}.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the sequential planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast sequential
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planning for contingencies unless
-they eventuate during execution.
-
-computing a single sequential plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
-
-\bibliographystyle{aaai}
-\bibliography{papers}
-
-\end{document}
-
-\subsection{PPDDL}
-
-We briefly discuss the description language PPDDL, a fairly
-straight-forward extension of the Planning Domain Definition Language
-for describing fully observable Markov decision processes.
-
-\small{
-\begin{tabtt}
-(\= :action look-at-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :precondition (and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) ) \\
-  \> :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :observe visual-object \\
-  \> :parameters \=(?r - robot ?v - visual-object\\
-  \> \> ?l - label ?l - location) \\
-  \> :execution (look-at-object ?r ?v ?l ?l) \\
-  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
-\end{tabtt}
-}
-
-\small{
-\begin{tabtt}
-(\= :sensor look-for-object \\
-  \> :agent (?a - robot) \\
-  \> :parameters (?l - label ?l - location) \\
-  \> :variables (?o - visual-object) \\
-  \> :precondition (\=and (= (is-in ?r) ?l) \\
-  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
-  \> :effect () \\
-  \> :sense (= (is-in ?o) ?l) \\
-\end{tabtt}
-}
-
-\subsection{MAPL}
-
-%% \scriptsize
-%% \begin{nopagebreak}\begin{tabtt}
-%% <observation-def> \=::= \=(:sense <observation symbol> \\
-%%                         \> \> :parameters (<typed list (variable)>)  \\
-%%                         \> \> <o-def body>) \\
-%%   <o-symbol> \> ::= <name> \\
-%%   <o-def body> \> ::= [:precondition <GD>] \\
-%%   \> \> [:execution <atomic action(term)> ] \\
-%%     \> \> [:effect <o-effect>] \\
-%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
-%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
-%%   <o-effect> \> ::= <c-o-effect> \\
-%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
-%%   <c-o-effect> \> ::= <p-o-effect> \\
-%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
-%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
-%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
-%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <number>\\
-%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
-%%   <o-f-exp> \> ::= <o-f-head>\\
-%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
-%%   <o-f-head> \> ::= <o-function-symbol>\\
-%% \end{tabtt}\end{nopagebreak}
-%% \noteme{<structure-def> ::= <attach-def>}
-%% \normalsize
-
-%% \footnote{We have simplified the schemata for illustrative
-%% purposes. In practice, the probabilities of a particular observation
-%% here should be parametrised by the type of visual object the robot is
-%% searching for, and category of the location.}
-
-We suppose $\bstate_0$ is given in a factored tree format of the
-form:
-
-%% \small{
-%% \begin{tabtt}
-%% (:init  (and \=f_{11} f_{12} ..
-%%  \> (probabilistic p_{}) 
-%%  \> (probabilistic ..) ..) \\
-%% \end{tabtt}
-%% }
-
-%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
-
-planning in practical sized POMDPs is intractable. One approach is to
-described the task hierarchically, thereby constrain and simplify the
-policy search task. The approach ~\cite{}.
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
-%% $\selection_\node(\action) = P(\action | \node)$, and
-%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
-%% \action, \observation)$. Where the value of acting is the discounted
-%% accumulated reward over an infinite horizon, then each controller can
-%% be assigned a value, written $V_\node(\state)$, expressed in the usual
-%% way according to the Bellman equation:
-
-%% \begin{equation}\label{eq:evaluation}
-%% \begin{array}{lcl}
-%% V_\node(\state) & = & \sum_{\action \in \actions}
-%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
-
-%% && \hspace{-10ex} \beta \sum_{\action, \observation,
-%% \state', \node'} \transition_\node(\action, \observation, \node')
-%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
-%% \action) V_{\node'}(\state')
-%% \end{array}
-%% \end{equation}
-
-%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
-%% then:
-
-%% \begin{equation} \label{eq:valueBelief}
-%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
-%% \end{equation}
-
-%% \noindent There is a corresponding deterministic FSCs for both
-%% sequential (i.e., conformant) and contingent
-%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
-%% case, if we take $\beta = 1$, and alter the process dynamics so there
-%% is a compulsory transition to a zero utility sink state when that
-%% horizon is reached.
-
- ---i.e., the values of {\em
-runtime variables} in the continual paradigm--- are treated , and
-therefore plan for a specific eventuality, replanning from scratch if
-something

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
I have to go home to cook dinner for Katrina. Still have to write the section about DTP from serial abstraction..
 - Charles
Might still have some algebra wrong, but I think I did some repairing of the reward sketch.
 - Charles
Indexing error.
 - Charles
New section

@@
 visited. The starting state, $\state_0$, for sequential planning is
-the abstract state with probability $1.0$ -- E.G., for $\pp{R2D2}$
-$s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
+the unique abstract state with probability $1.0$ -- E.G., for
+$\pp{R2D2}$ $s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
 
 In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
-for each element, $p_i (T_i)$, of each probabilistic term from
+add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
+each element, $p_i (T_i)$, of each probabilistic term from
 $(:init)$. The physics of these actions is as
-follows. $\actions(p_i;(T_i))$ can be executed if no
-$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
-probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
-is in the root conjunct, or it occurs in $T_k$ for some executed
-$\actions(p_k;(T_k))$.
+follows. $\assumptiveS(i)$ can be executed if no $\assumptiveS(j)$,
+$j \neq i$, has been executed from the same probabilistic term, and,
+either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
+or it occurs in $T_k$ for some executed $\assumptiveS(k)$.
 %%
-Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
+Executing $\assumptiveS(i)$ in $s$ effects a transition to a
 successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
@@
 V^* = \arg\min_{V_i} \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{PLAN}, \bstate_0
+\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
 \bigg{]} - V_i.
@@
 under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy.
-
-%% Finally, we come to the issue of valuing sequential plans, and
-%% therefore giving the criterion that the sequential planner seeks to
-%% optimise. Where $p_i$ is the probability that the $i^{th}$ action
-%% execution is not a transition to $s^\bot$, we value a trace according
-%% to:
-
-%% \begin{equation}\label{eq:tracevalue}
-%% V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-%% \end{equation}
-
-%% \noindent In goal directed problems, where the only non-zero reward is
-%% received at the first transition to a goal state,
-%% Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
-%% infamous degenerate hindsight system
-%% \system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
-%% gradient-based reinforcement learning systems, such as William's
-%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-%% under reasonable assumptions the optimal trace yields a maximal
-%% gradient step from a uniform soft-max policy.
+gradient step from a uniformly random soft-max policy. Finally, it is
+worth clarifying that although the above discussion has been in terms
+of a trace of length $N$, there is nothing in our approach that
+artificially limits the length of candidate sequential plans.
 
@@
 
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
+In the decision-theoretic mode, the switching planner solves a sensing
+problem in an abstract processes defined in terms of the assumptive
+actions in the current sequential plan, and the $(\pp{:init})$ block
+from the original problem description.
 
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
+The focus for decision-theoretic planning is on sensing, rather than
+the achievement of utility according to the given objectives. We
+therefor give the process in this mode a different reward model, which
+reflects the value of performing sensing actions in the context of the
+trace proposed for the preceding sequential deliberation. In detail,
+all positive rewards from the initial POMDP are removed. Then, for
+each $\assumptiveS{i}$ action scheduled in the current sequential
+plan, we have a {\em dual} $\assumptiveDT{i}$ so that
+$\forall\state\in\states$:
+
+\[
+\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
+\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
+-\$(T_i) & \pp{otherwise} \\
+\end{array}
+\]
+
+\noindent Where $\$(T_i)$ is a positive numeric quantity which
+captures the reward the agent receives for correctly jettisoning a
+false assumption. The added dual actions are propositionally trivial
+-- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
+\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
+$\action_i$, whose scheduled execution triggered the switch to the
+decision-theoretic mode, we include another propositionally trivial
+action $\actions.\poss(\action_i)$ with the reward property:
+
+\[
+\reward(\state, \actions.\poss(\action_i)) = \bigg\{ \begin{array}{ll}
+\$(\poss(\action_i)) & \pp{if}~\;\; \poss(\action_i) \subseteq \state \\
+-\$(\poss(\action_i)) & \pp{otherwise} \\
+\end{array}
+\]
+
+For the management of switching, execution of either a
+$\assumptiveDT{i}$ action, or $\actions.\poss(\action_i)$, returns
+control to the sequential planner. At that point, if a dual is
+executed, then the sequential planner must replan with $\bstate_0$
+equal to the current belief. Otherwise, if $\actions.\poss(\action_i)$
+is executed, then the current sequential plan is executed until
+further sensing is scheduled, or to completion. In our project the
+rewards, i.e., function $\$:2^\propositions\to\Re$, for the POMDP
+posed to decision-theoretic planning are sourced from a motivational
+component of our robotic architecture.
+
+The starting belief-state for the abstract process is given in terms
+of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
+constructing a tree which only includes terms from the {\em relaxed}
+visitation characterised by the {\em assumptive} actions in the
+current sequential plan. For example, if $\pp{R2D2}$'s plan is:
+
+\[
+\begin{array}{l}
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
+(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
+(\pp{report-is-in}~\pp{box}~\pp{office})
+\end{array}
+\]
+
+\noindent Then the {\em relaxed} block will be,
+
+\small
+\begin{tabtt}
+(\=:init (= (is-in R2D2) kitchen) \+ \\
+       (probabilistic \=.8 (= (is-in box) office)  \\
+		      \>.2 (= (is-in box) kitchen))) \\
+\end{tabtt}
+\normalsize
+
+\noindent and therefore, the belief-state $\bstate_0$ is:
+
+\small
+\begin{tabular}{cccc}
+\hline
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+\hline
+.8 & kitchen & office & \# \\
+.2 & kitchen & kitchen & \# \\
+\hline
+\end{tabular}
+\normalsize
+
+\noindent At this point, decision-theoretic planning would proceed in an
+abstraction of the environment that does not contain a $\pp{cup}$,
+where the inclusion (exclusion) of POMDP actions is the same as we
+defined it for the sequential mode.
+
+In a second step, we iteratively refine the {\em relaxed} block by
+adding terms from the original block while the resulting state space
+remains of a practicable size. Of course, it is expensive to measure
+size exactly, thus in practice we estimate it to be
+$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
+propositions in the abstraction induced by the proposed tree. Terms
+are proposed for addition because they are probabilistic, and because
+they were omitted in the {\em relaxed} visitation characterised by
+{\em assumptive} actions in the sequential plan. Each iteration, we
+add one candidate term with the largest number of parent terms already
+in the candidate tree.

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
New structure again.

@@
 steps in sequential plan.
+
+One theoretical criticism of switching continual planning concerns
+interleaved sequential and decision-theoretic sessions failing to make
+any progress towards the POMDP objectives. For example, in the worst
+case we could have each sequential session producing an identical
+plan, and each decision-theoretic session rejecting it without further
+sensing. Although this has not been an issue in our work so far, it
+must be dealt with rigorously moving forward. We propose a good way of
+mitigation this problem is by using a {\em motivational} component
+that maintains a {\em dynamic} reward model for sensing.
 

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
New structure again.

--- 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
 \section{Introduction}
-
-Any POMDP with stochastic actions can be compiled into an equivalent
-deterministic-action POMDP with all the action uncertainty expressed
-in the starting-state distribution~\cite{ng:Jordan:2000}. In our
-setting, that of finite-horizon planning, this compilation yields a
-finite flat representation of the POMDP. Therefore, although our
-current implementation does not explicitly support it, our approach
-generalises to POMDPs with stochastic action models.
-
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
-%%
-
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
-
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
-
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
-
-the speed and scalability of software for sequential planning in
-deterministic
-
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
-
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
-
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
-
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
-
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
-
-brittle sensing model,
-
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We describe POMDPs with a
-propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
+\input{intro.tex}
 
 \section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
-state-characterising propositions $\propositions$.
-
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
-..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
-applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-propositions from $\add(\detAction)$ are added to $\state$, and those
-from $\delete(\detAction)$ are removed from $\state$.
-
-We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$, that it has no
-effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
-\state_i$.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs are
-negative rewards.
-%%
-From $\psi$ we have that for each $\state \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
-\action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
-
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N}  \reward(\bstate_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief state at time $t$. Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
+\input{prop-dtp.tex}
 
 \section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
-
-Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarative blocks, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition.
-
-Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a box, at a given place. We
-can model that deterministic action using the following operator
-schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
-
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} clause is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{R2D2}$ might
-be described thusly:
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} if its conjunctive parent is visited, and a
-conjunctive term is visited if all its immediate subterms are
-visited. A probabilistic term is visited if exactly one of its
-subterms, $T_i$, is visited. Each valid visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
+\input{dtpddl.tex}
 
 \section{Switching Planner}
-
-We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning. The problem at hand is a
-POMDP described in DTPDDL.
-
-\subsection{Sequential Planning in POMDPs}
-
-Our approach uses sequential planning to compute a linear plan, and
-one trace of that plan's execution. The latter encapsulates
-assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of eventually holding
-strong beliefs about the truth values of specific state
-propositions. Sequential planning proceeds in a continual manner until
-the verification of an assumption about a runtime state proposition is
-scheduled. At that point the system switches to a decision-theoretic
-planner that is used to tailor sensory processing by the robot to test
-the validity of the assumptions so far.
-
-The first actions sequenced by the planner are virtual, and correspond
-to identifying an abstract start state. That prefix corresponds to a
-{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
-block. In the relaxed case, we can modify the initial definition so
-that a conjunctive term is visited provided all its atomic subterms
-are visited, and zero or more of its probabilistic terms are
-visited. Writing \#\ if the value of a proposition is unspecified by a
-the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
-following abstract starting states:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-.7 & kitchen & \# &  kitchen\\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-Sequential planning in our approach proceeds in two phases. In the
-first phase, the goal is to identify a plan prefix
-
-In sequential planning, the plan prefix corresponds to choosing a
-valid abstract starting state.
-
-moreover, for the purpose of decision-theoretic abstractions,
-
-A relaxed form of {\em visitation} we call {\em abstract
-  visitation}. In the abstract case a conjunctive term is visited
-  provided all its atomic subterms are visited, and zero or more of
-  its probabilistic terms are visited.
-
-The planner has a compulsory starting action, that visits the root
-term, and then has to execute visitation actions until that plan
-prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
-available to the planner
-
-Goal expressions in the problem description ---for example
-$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
-location of the {\em box}--- are treated as control-knowledge by the
-planner. Here, a plan is not considered valid unless in the finial
-stet the goal condition is satisfied.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-The problem posed to the decision-theoretic system is an abstraction
-of the problem at hand with a distinct utility model that focuses the
-agents activities on assumption verification. The abstraction is
-induced according to the executing sequential plan. Specifically, we
-first omit all problem variables that are not mentioned in the linear
-trace, and then refine that preliminary model, by progressively adding
-more problem variables while the resulting state space remains of a
-practicable size. Refinement in our approach is a simple and
-inexpensive step, guided by the level at which variable symbols occur
-in the description of the initial conditions. The abstract process is
-also given a utility model that captures the reward (resp. penalty) of
-correctly (resp. falsely) in/validating the assumptions made in the
-linear trace. In detail, for each assumption on a state variable, we
-can have a reward of $1$ for correctly rejecting it, and a penalty of
-$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
-a more-or-less continual manner until an assumption is rejected, or
-until it returns control to the sequential system because it can see
-no benefit from performing further actions in the abstraction.
-
-Addressing briefly the specifics of our mobile robot's
-architecture. In our scenario the utility model for the POMDP posed to
-decision-theoretic planning is sourced from a motivational component
-of our robotic architecture. This proposes a non-zero reward
-(resp. penalty) for sequential planning with a true (resp. false)
-assumption about a state proposition.
+\input{switching.tex}
 
 \section{Evaluation}
+\input{evaluation.tex}
 
-\section{Discussion and Related Work}
+%% \section{Discussion}
+%% \input{discussion.tex}
 
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
+\section{Related Work}
+\input{relatex.tex}
 
 \section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
-
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
-
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
-
-posted by a motivational component of the underlying robotic
-architecture.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
-
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
-
-Interaction between the sequential planner and execution proceeds
-more-or-less analogously to popular replanning approaches
-
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast sequential
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
-
-contingent sensory plans that are tailored to current
-objectives.
-
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
-
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
-
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planning for contingencies unless
-they eventuate during execution.
-
-computing a single sequential plan and
-eventuality
-
-To these challenges it seems a continual planning approach is best,
-
-where the runtime state evolves during plan execution
-
-interleaved planning and execution.
-
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
-
-{\em ad-hoc}
-
-, reason about degrees of belief and uncertainty about the world,
-
-probabilistic sequential decision making in practical sized problems
-is intractable.
-
-quantitative probabilistic models of the perception and action .
+\input{conc.tex}
 

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
Name change

@@
 the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
+$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office}) =
+\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
 we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
 order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
+if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
 executed, there is a $0.8$ chance of perceiving the box if it is in
@@
 atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{R2D2}$ might
+atomic. For example, the starting distribution for $\pp{DORA}$ might
 be described thusly:
@@
 \begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
+(\=:init (= (is-in DORA) kitchen) \+ \\
        (probabilistic \=.8 (= (is-in box) office)  \\
@@
 \hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
 \hline

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Name change

@@
 We now describe a continual planning approach that switches between
-sequential and decision-theoretic planning.
+sequential and decision-theoretic planning sessions.
 
@@
 uses the structure of that, as it occurs in the problem description,
-do define a set of abstract state-assumptions available to 
-sequential planning.  
+do define a set of state-assumptions available to sequential planning.
 %%
-Writing \#\ if the value of a proposition is unspecified in the
-assumption, for the $\pp{R2D2}$ example we have that the following
+Writing \#\ if the value of a proposition is unspecified in an
+assumption, for our $\pp{DORA}$ example we have that the following
 assumptions are available:
@@
 the unique abstract state with probability $1.0$ -- E.G., for
-$\pp{R2D2}$ $s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
+$\pp{DORA}$ $s_0 \equiv \{(= (is-in DORA) kitchen)\}$.
 
@@
 probability $1$, then before it is executed blindly, the sequential
-continual planning switches to a decision-theoretic mode, discussed in
+continual planning switches to a decision-theoretic session, discussed in
 detail below. Overall, by scheduling $\action$ the sequential planner
@@
 
-In the decision-theoretic mode, the switching planner solves a sensing
+In a decision-theoretic session, the switching planner solves a sensing
 problem in an abstract processes defined in terms of the assumptive
@@
 the achievement of utility according to the given objectives. We
-therefor give the process in this mode a different reward model, which
+therefor give the process in this session a different reward model, which
 reflects the value of performing sensing actions in the context of the
@@
 $\action_i$, whose scheduled execution triggered the switch to the
-decision-theoretic mode, we include another propositionally trivial
+decision-theoretic session, we include another propositionally trivial
 action $\actions.\poss(\action_i)$ with the reward property:
@@
 visitation characterised by the {\em assumptive} actions in the
-current sequential plan. For example, if $\pp{R2D2}$'s plan is:
+current sequential plan. For example, if $\pp{DORA}$'s plan is:
 
@@
 \actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office});\\
+(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
 (\pp{report-is-in}~\pp{box}~\pp{office})
@@
 where the inclusion (exclusion) of POMDP actions is the same as we
-defined it for the sequential mode.
+defined it for the sequential session.
 

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
New controller spec

--- 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
@@
 applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
+\add(\detAction) ) \backslash \delete(\detAction)$ -- I.e.,
 propositions from $\add(\detAction)$ are added to $\state$, and those
@@
 exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs are
+\actions$, $|\reward(\state, \action)| < c$. We model costs as
 negative rewards.
@@
 \action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
+acting for finite $N$ steps from controller state $n$, the value of a
+controller corresponds to the expected cumulative reward:
 
 \begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
+V_{\pp{FSC}}(\bstate, n) = \Expect \bigg{[} 
+\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, n_0 = n, \bstate_0 = \bstate \bigg{]}
 \end{equation}

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
New controller spec

--- 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
@@
 formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, and describe how the
+operator} schema are over state propositions, specifying how the
 states change under application of actions, the effects of a {\em
-sense} schema are perceptual, and describe the composition of an
+sense} schema are perceptual, specifying the composition of an
 observation following the execution of an action. Also, and without a
@@
 Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a box, at a given place. We
+able to look for a visual-object, such as a $\pp{box}$, at a given place. We
 can model that deterministic action using the following operator

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
relational is gone

@@
 uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
+ language that facilitates factored description of
 domains and problem instances. There are straightforward compilations

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Notation

@@
 assumption, for our $\pp{DORA}$ example we have that the following
-assumptions are available:
+assumptions:
 
@@
 visited. The starting state, $\state_0$, for sequential planning is
-the unique abstract state with probability $1.0$ -- E.G., for
-$\pp{DORA}$ $s_0 \equiv \{(= (is-in DORA) kitchen)\}$.
+the unique abstract state with probability $1.0$ -- E.g., for
+$\pp{DORA}$ $s_0 \equiv \{(= (\pp{is-in}~\pp{DORA})~\pp{kitchen})\}$.
 
@@
 $(:init)$. The physics of these actions is as
-follows. $\assumptiveS(i)$ can be executed if no $\assumptiveS(j)$,
+follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
 $j \neq i$, has been executed from the same probabilistic term, and,
 either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
-or it occurs in $T_k$ for some executed $\assumptiveS(k)$.
+or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
 %%
-Executing $\assumptiveS(i)$ in $s$ effects a transition to a
+Executing $\assumptiveS{i}$ in $s$ effects a transition to a
 successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with

### 2010-11-23 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
small corrections

@@
 uses the structure of that, as it occurs in the problem description,
-do define a set of abstract state-assumptions available to 
-sequential planning.  
+do define a set of state-assumptions available to sequential planning.
 %%
-Writing \#\ if the value of a proposition is unspecified in the
-assumption, for the $\pp{R2D2}$ example we have that the following
-assumptions are available:
+Writing \#\ if the value of a proposition is unspecified in an
+assumption, for our $\pp{DORA}$ example we have that the following
+assumptions:
 
@@
 visited. The starting state, $\state_0$, for sequential planning is
-the unique abstract state with probability $1.0$ -- E.G., for
-$\pp{R2D2}$ $s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.
+the unique abstract state with probability $1.0$ -- E.g., for
+$\pp{DORA}$ $s_0 \equiv \{(= (\pp{is-in}~\pp{DORA})~\pp{kitchen})\}$.
 
@@
 $\action\in\actions$ has a precondition that is not true with
-probability $1$, then before it is executed blindly, the sequential
-continual planning switches to a decision-theoretic mode, discussed in
-detail below. Overall, by scheduling $\action$ the sequential planner
-makes a naive assumption about the possibility of the agent eventually
-``committing'' to the truth values of specific state
-propositions. Thus, the role of decision-theoretic plans is to verify
-the validity of assumptions about the runtime state in sequential
-planning. Second, if the truth value of $\prop$ is not known, and no
-assumption is made about it in the current abstract state, then if
-$\prop \in
+probability $1$ in $\bstate_0$, then before it is executed blindly,
+the sequential continual planning switches to a decision-theoretic
+session, discussed in detail below. Overall, by scheduling $\action$
+the sequential planner makes a naive assumption about the possibility
+of the agent eventually ``committing'' to the truth values of specific
+state propositions. Thus, the role of decision-theoretic plans is to
+examine the validity of assumptions about the runtime state of
+sequential planning. Second, if the truth value of $\prop$ is not
+known, and no assumption is made about it in the current abstract
+state, then if $\prop \in
 \poss(\action)$ we have that action $\action$ is not applicable.
-
-%% Because $\state_0$ is abstract, and because actions can have
-%% preconditions that are not uniformly true in the underlying
-%% (resp. abstract) belief-state, we must make two further adjustments to
-%% the problem posed to the sequential planner. First, if
-%% $\action\in\actions$ has a precondition that is not true with
-%% probability $1$, then before it is executed blindly, the sequential
-%% continual planning switches to a decision-theoretic mode, discussed in
-%% detail below. Overall, by scheduling $\action$ the sequential planner
-%% makes a naive assumption about the possibility of the agent eventually
-%% committing ---for the purpose of the current task--- to the truth
-%% values of specific state propositions. Second, if the truth value of
-%% $\prop$ is not known, and no assumption is made about it in the
-%% current abstract state, then if $\prop \in \poss(\action)$ we have
-%% that action $\action$ is not applicable.
-
-%% In other words, the value of a proposition in an abstract
-%% state is either {\em true}, {\em false}, or {\em unspecified} (i.e.,
-%% \#). In the last case actions whose execution is preconditioned on the
-%% term being truth functional are not available.
 
@@
 at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
-criteria corresponds exactly to that of the infamous degenerate
-hindsight system
+criteria corresponds exactly to that of
 \system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
@@
 
-%% Finally, we come to the issue of valuing sequential plans, and
-%% therefore giving the criterion that the sequential planner seeks to
-%% optimise. Where $p_i$ is the probability that the $i^{th}$ action
-%% execution is not a transition to $s^\bot$, we value a trace according
-%% to:
-
-%% \begin{equation}\label{eq:tracevalue}
-%% V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-%% \end{equation}
-
-%% \noindent In goal directed problems, where the only non-zero reward is
-%% received at the first transition to a goal state,
-%% Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
-%% infamous degenerate hindsight system
-%% \system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
-%% gradient-based reinforcement learning systems, such as William's
-%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-%% under reasonable assumptions the optimal trace yields a maximal
-%% gradient step from a uniform soft-max policy.
-
 \subsection{Decision-Theoretic Planning in Abstractions}
 
-In the decision-theoretic mode, the switching planner solves a sensing
+In a decision-theoretic session, the switching planner solves a sensing
 problem in an abstract processes defined in terms of the assumptive
@@
 The focus for decision-theoretic planning is on sensing, rather than
-the achievement of utility according to the given objectives. The
-process in this mode has a different reward model that reflects the
-value of performing sensing actions in the context of a trace proposed
-for the preceding sequential deliberation. In detail, for each
-$\assumptiveS{i}$ action scheduled in the current sequential plan, we
-have a {\em dual} $\assumptiveDT{i}$ so that
+the achievement of utility according to the given objectives. We
+therefor give the process in this session a different reward model, which
+reflects the value of performing sensing actions in the context of the
+trace proposed for the preceding sequential deliberation. In detail,
+all positive rewards from the initial POMDP are removed. Then, for
+each $\assumptiveS{i}$ action scheduled in the current sequential
+plan, we have a {\em dual} $\assumptiveDT{i}$ so that
 $\forall\state\in\states$:
@@
 \noindent Where $\$(T_i)$ is a positive numeric quantity which
-rewards the agent for correctly jettisoning a false assumption, and
-otherwise penalises the agent. Propositionally these dual actions have
-no preconditions or effects -- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Additionally, for the action,
+captures the reward the agent receives for correctly jettisoning a
+false assumption. The added dual actions are propositionally trivial
+-- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
+\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
 $\action_i$, whose scheduled execution triggered the switch to the
-decision-theoretic mode, we include an action
-$\actions.\poss(\action_i)$ with the reward property:
+decision-theoretic session, we include another propositionally trivial
+action $\actions.\poss(\action_i)$ with the reward property:
 
@@
 
-\noindent Execution of either a $\assumptiveDT{i}$
-action, or $\actions.\poss(\action_i)$, returns control to the
-sequential planner. If a dual is executed, then the sequential planner
-must replan with $\bstate_0$ equal to the current belief. Otherwise,
-if $\actions.\poss(\action_i)$ is executed, then the current
-sequential plan is executed until further sensing is scheduled, or to
-completion. In our project the rewards, $\$:2^\propositions\to\Re$,
-for the POMDP posed to decision-theoretic planning is sourced from a
-motivational component of our robotic architecture.
+For the management of switching, execution of either a
+$\assumptiveDT{i}$ action, or $\actions.\poss(\action_i)$, returns
+control to the sequential planner. At that point, if a dual is
+executed, then the sequential planner must replan with $\bstate_0$
+equal to the current belief. Otherwise, if $\actions.\poss(\action_i)$
+is executed, then the current sequential plan is executed until
+further sensing is scheduled, or to completion. In our project the
+rewards, i.e., function $\$:2^\propositions\to\Re$, for the POMDP
+posed to decision-theoretic planning are sourced from a motivational
+component of our robotic architecture.
 
 The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
-constructing a tree which only includes terms from the {\em relaxed}
-visitation characterised by the {\em assumptive} actions in the
-current sequential plan. For example, if $\pp{R2D2}$'s plan is:
+of a {\em relaxed} $(\pp{:init})$ block.  We first construct a
+tree which only includes terms from the {\em relaxed} visitation
+characterised by the {\em assumptive} actions in the current
+sequential plan. For example, if $\pp{DORA}$'s plan is:
 
@@
 
-\noindent At this point, decision-theoretic planning would proceed in a cupless
-abstraction of the environment, where the inclusion (exclusion) of
-POMDP actions is the same as we defined it for the sequential mode.
+\noindent At this point, decision-theoretic planning would proceed in an
+abstraction of the environment that does not contain a $\pp{cup}$,
+where the inclusion (exclusion) of POMDP actions is the same as we
+defined it for the sequential session.
 
@@
 remains of a practicable size. Of course, it is expensive to measure
-size exactly, thus in practice we estimate it to be
+size exactly, and in practice we estimate this to be
 $2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
-propositions in the abstraction induced by the proposed tree. Terms
-are proposed for addition, because they are probabilistic, and because
-they were omitted in the {\em relaxed} visitation characterised by
-{\em assumptive} actions in the sequential plan. Each iteration, we
-add one candidate term with the largest number of parent terms already
-in the candidate tree.
+propositions in the abstraction that were not assigned the {\em
+unknown} value \#. Terms are proposed for addition because they are
+probabilistic, and because they were omitted in the {\em relaxed}
+visitation characterised by {\em assumptive} actions in the sequential
+plan. Each iteration, we add one candidate term with the largest
+number of parent terms already in the candidate tree.

### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
Relaxed conclusion

@@
 In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
+planner that switches between fast sequential and expensive
+decision-theoretic planning. Given a POMDP model of the environment,
+sequential planning is used to compute an initial deterministic
+sequential plan and complementary runtime evolution of the decision
+process. That plan is executed until a validation of an assumption
+about a runtime proposition is schedule. At that point a
+decision-theoretic session performs some sensing which determines
+whether or not the sequential session should be allowed to continue
+without plan repair.
 
@@
 planning before executing actions that are not applicable with
-certainty. This is vastly inefficient if the utility of that plan is
-not dependent on the actions successful execution. This is the case
-where a plan is conformant. Therefore, in the future we must develop a
-cheap (heuristic) procedure to evaluate the utility of sensing at
-steps in sequential plan.
+certainty. This is inefficient if the utility of the plan is not
+dependent on that actions successful execution. This situation always
+arises for conformant plans.
 

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
NOt more "blocks". Also, I got rid of the ambiguity with uninitialised variables.

@@
 uncertain starting configurations, PPDDL is a declarative first-order
-(a.k.a., relational) language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
+language that facilitates factored description of domains and problem
+instances. There are straightforward compilations from problems
+expressed compactly in PPDDL to propositional representations amenable
+to state-of-the-art planning procedures.
 
@@
 state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarative blocks, for
+PPDDL syntax. In addition, we allow two other declarations, for
 perceptual predicates and fluents respectively. For example, suppose
@@
 reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
+terms in operator schemata. In that setting, {\em reward} is
 interpreted as accumulated (resp. instantaneous) rewards and
@@
 
-\noindent Here, the {\em execution} clause is a lifted description of
+\noindent Here, the {\em execution} declaration is a lifted description of
 the sense action-precondition -- I.e,

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
New section
NOt more "blocks". Also, I got rid of the ambiguity with uninitialised variables.

@@
 
-\section{Discussion and Related Work}
-\input{discussion.tex}
+%% \section{Discussion}
+%% \input{discussion.tex}
+
+\section{Related Work}
+\input{relatex.tex}
 
@@
 prefix corresponds to a valid {\em abstract visitation} of the {\em
-:init} block. Beyond such a prefix, the process actions and senses are
+:init} decleration. Beyond such a prefix, the process actions and senses are
 available to the planner

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
NOt more "blocks". Also, I got rid of the ambiguity with uninitialised variables.

@@
 \noindent A valid assumption in our setting corresponds to a {\em
-relaxed} visitation of the root term of $(:init)$. In this relaxed
+relaxed} visitation of the root term of $(\pp{:init})$. In this relaxed
 case, a conjunctive term is visited iff its atomic subterms are
@@
 the unique abstract state with probability $1.0$ -- E.g., for
-$\pp{DORA}$ $s_0 \equiv \{(= (\pp{is-in}~\pp{DORA})~\pp{kitchen})\}$.
+$\pp{DORA}$:
+
+\[
+\begin{array}{l}
+s_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
+\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
+\end{array}
+\]
 
@@
 each element, $p_i (T_i)$, of each probabilistic term from
-$(:init)$. The physics of these actions is as
+$(\pp{:init})$. The physics of these actions is as
 follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
@@
 %%
-Executing $\assumptiveS{i}$ in $s$ effects a transition to a
-successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
-probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
-terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
+Executing $\assumptiveS{i}$ in $s$ effects a transition to a successor
+state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with probability
+$1 - p_i$. Here, $s^{T_i}$ is the union of $s$ with atomic terms from
+$T_i$ and conflicting propositions from $s$ removed. For example, if
+the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
+delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is a sink, where
 $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
@@
 of the agent eventually ``committing'' to the truth values of specific
-state propositions. Thus, the role of decision-theoretic plans is to
-examine the validity of assumptions about the runtime state of
-sequential planning. Second, if the truth value of $\prop$ is not
-known, and no assumption is made about it in the current abstract
+state propositions in $\poss{\action}$. The role of decision-theoretic
+plans is to examine the validity of assumptions about the runtime
+state of sequential planning. Second, if the truth value of $\prop$ is
+not known, and no assumption is made about it in the current abstract
 state, then if $\prop \in
@@
 problem in an abstract processes defined in terms of the assumptive
-actions in the current sequential plan, and the $(\pp{:init})$ block
+actions in the current sequential plan, and the $(\pp{:init})$ declaration
 from the original problem description.
@@
 The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ block.  We first construct a
+of a {\em relaxed} $(\pp{:init})$ declaration.  We first construct a
 tree which only includes terms from the {\em relaxed} visitation
@@
 
-\noindent Then the {\em relaxed} block will be,
+\noindent Then the {\em relaxed} declaration will be,
 
@@
 
-In a second step, we iteratively refine the {\em relaxed} block by
-adding terms from the original block while the resulting state space
-remains of a practicable size. Of course, it is expensive to measure
-size exactly, and in practice we estimate this to be
-$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
-propositions in the abstraction that were not assigned the {\em
-unknown} value \#. Terms are proposed for addition because they are
-probabilistic, and because they were omitted in the {\em relaxed}
-visitation characterised by {\em assumptive} actions in the sequential
-plan. Each iteration, we add one candidate term with the largest
-number of parent terms already in the candidate tree.
+In a second step, we iteratively refine the above{\em relaxed}
+declaration by adding terms from the original statement of
+$(\pp{:init})$ while the resulting state space remains of a
+practicable size. Of course, it is expensive to measure size exactly,
+and in practice we estimate this to be $2^{|\propositions^\#|}$, where
+$\propositions^\#$ is the set of propositions in the abstraction that
+were not assigned the {\em unknown} value \#. Terms are proposed for
+addition because they are probabilistic, and because they were omitted
+in the {\em relaxed} visitation characterised by {\em assumptive}
+actions in the sequential plan. Each iteration, we add one candidate
+term with the largest number of parent terms already in the candidate
+tree.

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
typo

@@
 $\action\in\actions$ has a precondition that is not true with
-probability $1$, then before it is executed blindly, the sequential
-continual planning switches to a decision-theoretic session, discussed in
-detail below. Overall, by scheduling $\action$ the sequential planner
-makes a naive assumption about the possibility of the agent eventually
-``committing'' to the truth values of specific state
-propositions. Thus, the role of decision-theoretic plans is to verify
-the validity of assumptions about the runtime state in sequential
-planning. Second, if the truth value of $\prop$ is not known, and no
-assumption is made about it in the current abstract state, then if
-$\prop \in
+probability $1$ in $\bstate_0$, then before it is executed blindly,
+the sequential continual planning switches to a decision-theoretic
+session, discussed in detail below. Overall, by scheduling $\action$
+the sequential planner makes a naive assumption about the possibility
+of the agent eventually ``committing'' to the truth values of specific
+state propositions in $\poss(\action)$. The role of decision-theoretic
+plans is to examine the validity of assumptions about the runtime
+state of sequential planning. Second, if the truth value of $\prop$ is
+not known, and no assumption is made about it in the current abstract
+state, then if $\prop \in
 \poss(\action)$ we have that action $\action$ is not applicable.
-
-%% Because $\state_0$ is abstract, and because actions can have
-%% preconditions that are not uniformly true in the underlying
-%% (resp. abstract) belief-state, we must make two further adjustments to
-%% the problem posed to the sequential planner. First, if
-%% $\action\in\actions$ has a precondition that is not true with
-%% probability $1$, then before it is executed blindly, the sequential
-%% continual planning switches to a decision-theoretic mode, discussed in
-%% detail below. Overall, by scheduling $\action$ the sequential planner
-%% makes a naive assumption about the possibility of the agent eventually
-%% committing ---for the purpose of the current task--- to the truth
-%% values of specific state propositions. Second, if the truth value of
-%% $\prop$ is not known, and no assumption is made about it in the
-%% current abstract state, then if $\prop \in \poss(\action)$ we have
-%% that action $\action$ is not applicable.
-
-%% In other words, the value of a proposition in an abstract
-%% state is either {\em true}, {\em false}, or {\em unspecified} (i.e.,
-%% \#). In the last case actions whose execution is preconditioned on the
-%% term being truth functional are not available.
 
@@
 The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ block.  This is constructed by first
-constructing a tree which only includes terms from the {\em relaxed}
-visitation characterised by the {\em assumptive} actions in the
-current sequential plan. For example, if $\pp{DORA}$'s plan is:
+of a {\em relaxed} $(\pp{:init})$ declaration.  We first construct a
+tree which only includes terms from the {\em relaxed} visitation
+characterised by the {\em assumptive} actions in the current
+sequential plan. For example, if $\pp{DORA}$'s plan is:
 
@@
 
-In a second step, we iteratively refine the {\em relaxed} block by
-adding terms from the original block while the resulting state space
-remains of a practicable size. Of course, it is expensive to measure
-size exactly, thus in practice we estimate it to be
-$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
-propositions in the abstraction induced by the proposed tree. Terms
-are proposed for addition because they are probabilistic, and because
-they were omitted in the {\em relaxed} visitation characterised by
-{\em assumptive} actions in the sequential plan. Each iteration, we
-add one candidate term with the largest number of parent terms already
-in the candidate tree.
+In a second step, we iteratively refine the above{\em relaxed}
+declaration by adding terms from the original statement of
+$(\pp{:init})$ while the resulting state space remains of a
+practicable size. Of course, it is expensive to measure size exactly,
+and in practice we estimate this to be $2^{|\propositions^\#|}$, where
+$\propositions^\#$ is the set of propositions in the abstraction that
+were not assigned the {\em unknown} value \#. Terms are proposed for
+addition because they are probabilistic, and because they were omitted
+in the {\em relaxed} visitation characterised by {\em assumptive}
+actions in the sequential plan. Each iteration, we add one candidate
+term with the largest number of parent terms already in the candidate
+tree.

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
MOre small fixes

@@
 sequential and decision-theoretic planning sessions.
+
+Sequential planning proceeds in a continual manner until the
+verification of an assumption about runtime state is scheduled. At
+that point the system switches to a decision-theoretic planner that is
+used to tailor sensory processing by the robot to test the validity of
+the assumptions made so far.
 
@@
 eventually holding strong beliefs about the truth values of specific
-state propositions. Sequential planning proceeds in a continual manner
-until the verification of an assumption about a runtime state
-proposition is scheduled. At that point the system switches to a
-decision-theoretic planner that is used to tailor sensory processing
-by the robot to test the validity of the assumptions made so far.
+state propositions.
 
@@
 \begin{array}{l}
-s_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
+\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
 \;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
@@
 of the agent eventually ``committing'' to the truth values of specific
-state propositions in $\poss{\action}$. The role of decision-theoretic
+state propositions in $\poss(\action)$. The role of decision-theoretic
 plans is to examine the validity of assumptions about the runtime
-state of sequential planning. Second, if the truth value of $\prop$ is
-not known, and no assumption is made about it in the current abstract
-state, then if $\prop \in
-\poss(\action)$ we have that action $\action$ is not applicable.
+state of sequential planning. Second, if in the current state \state\
+the value of $\prop$ is $\#$ ---because a truth value is not specified
+in $\state_0$ and no assumption is made about it in the plan
+to \state--- then if $\prop \in \poss(\action)$, or if $\prop$ is the
+subject of a positive or negative effect of $\action$, we have that
+action $\action$ is not applicable in \state.
 
@@
 \noindent which is equal to the maximal contribution a trace of the plan can
-make to Eq~\ref{eq:expectedvalue}. In other words, writing $V_i$ for
-the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
-s_1, a_1,.., s_N$, we have
-
-\[
-V^* = \arg\min_{V_i} \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
-\bigg{]} - V_i.
-\]
-
+make to Eq~\ref{eq:expectedvalue}. %% In other words, writing $V_i$ for
+%% the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
+%% s_1, a_1,.., s_N$, we have
+%%
+%% \[
+%% V^* = \arg\min_{V_i} \Expect \bigg{[} 
+%% \sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
+%% \bigg{]} - V_i.
+%% \]
+%%
 In goal directed problems, where the only non-zero reward is received

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
tiny change.
I need to make some big changes now. I need to deal with the turnery nature of sequential assumptions. And then I have to make the distinction between a plan and a trace clear. Heaps of work.

@@
 plans is to examine the validity of assumptions about the runtime
-state of sequential planning. Second, if the truth value of $\prop$ is
-not known, and no assumption is made about it in the current abstract
-state, then if $\prop \in
-\poss(\action)$ we have that action $\action$ is not applicable.
+state of sequential planning. Second, if in the current state \state\
+the value of $\prop$ is $\#$ ---because a truth value is not specified
+in $\state_0$ and no assumption is made about it in the plan
+to \state--- then if $\prop \in \poss(\action)$, or if $\prop$ is the
+subject of a positive or negative effect of $\action$, we have that
+action $\action$ is not applicable in \state.
 
@@
 
-For the management of switching, execution of either a
-$\assumptiveDT{i}$ action, or $\actions.\poss(\action_i)$, returns
-control to the sequential planner. At that point, if a dual is
+For the management of switching, execution of either a dual action of
+the form $\assumptiveDT{i}$, or the $\actions.\poss(\action_i)$,
+returns control to the sequential planner. At that point, if a dual is
 executed, then the sequential planner must replan with $\bstate_0$

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
c v.s. our

@@
-We now describe a continual planning approach that switches between
+We now describe our continual planning approach that switches between
 sequential and decision-theoretic planning sessions.

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
Edits to early sections

@@
 useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
+both sequential (e.g., conformant) and contingent plan formats.

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
Edits to early sections

@@
 \noindent To model the sensing capabilities of the agent, we have
-operator-like descriptive elements, with preconditions in both action
-and state symbols, and uniformly positive effects over perceptual
-predicates -- I.e, the negation normal form of a sensing effect
-formula cannot contain negation. Whereas the effects of an {\em
-operator} schema are over state propositions, specifying how the
-states change under application of actions, the effects of a {\em
-sense} schema are perceptual, specifying the composition of an
-observation following the execution of an action. Also, and without a
-loss of generality, we find it convenient to separate the actional
-precondition from the state precondition.
+operator-like declarations, with preconditions expressed using state
+predicate and action function symbols, and uniformly positive effects
+over perceptual predicates. A sense declaration corresponds to a
+lifted description of proposition {\em senses}. Whereas the effects of
+an {\em operator} schema describe how states change under application
+of actions, the effects of a {\em sense} schema are perceptual,
+specifying the composition of an observation following the execution
+of an action.
 
-Making the above ideas concrete with an example, suppose a robot is
-able to look for a visual-object, such as a $\pp{box}$, at a given place. We
-can model that deterministic action using the following operator
-schema:
+%% Also, and without a loss of generality, we find it
+%% convenient to separate the actional precondition from the state
+%% precondition.
+
+Making the above ideas concrete with an example, suppose a robot
+called $\pp{DORA}$ is able to look for a visual-object, such as a
+$\pp{box}$, at a given place. We can model that deterministic action
+using the following operator schema:
 
@@
 $\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-terms in operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
+execution incurs an instantaneous penalty, $3$, that corresponds to
+the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
+reward} is a {\em reserved word}, and occurs in {\em increased} and
+{\em decreased} terms in operator schemata. In that setting, {\em
+reward} is interpreted as accumulated (resp. instantaneous) rewards
+and penalties.}
 
@@
 factored tree-like structure here, in order to aid us when discussing
-sequential planning in POMDP determinisations. The distribution is
-expressed in a tree-like structure of terms. Each term is either: (1)
-atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
+our switching continual planner. The distribution is expressed in a
+tree-like structure of terms. Each term is either: (1) atomic, e.g., a
+state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$,
+(2) probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
 where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
@@
 atomic. For example, the starting distribution for $\pp{DORA}$ might
-be described thusly:
+be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
+expresses the conjunctive root of the tree -- I.e., the root node
+$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
+$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
+subterm.}
 

### 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
forgot to save

@@
 \noindent The interpretation of such an expression can be given
-according to a recursive notion of {\em visitation} over terms. Here,
-an atom is {\em visited} iff its conjunctive parent is visited, and a
-conjunctive term is visited iff all its immediate subterms are
-visited. A probabilistic term is visited iff its conjunctive parent is
-visited, and exactly one of its subterms, $T_i$, is visited. Each
-valid visitation of the root term according to this recursive
-definition encapsulates a starting state, along with the probability
-it occurs. The former corresponds to the union of all visited atoms,
-and the latter corresponds to the product of $p_i$ entries on the
-visited subterms of probabilistic elements. Making these ideas
-concrete, our example yields the following flat state distribution:
+according to a {\em visitation} of terms. An atom is {\em visited} iff
+its conjunctive parent is visited, and a conjunctive term is visited
+iff all its immediate subterms are visited. A probabilistic term is
+visited iff its conjunctive parent is visited, and exactly one of its
+subterms, $T_i$, is visited. Each visitation of the root term
+according to this recursive definition encapsulates a starting state,
+along with the probability it occurs. The former corresponds to the
+union of all visited atoms, and the latter corresponds to the product
+of $p_i$ entries on the visited subterms of probabilistic
+elements. Making these ideas concrete, our example yields the
+following flat state distribution:
 

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
Not sure what this change is
Forgot this one last change

@@
 uncertain starting configurations, PPDDL is a declarative first-order
- language that facilitates factored description of
-domains and problem instances. There are straightforward compilations
-from problems expressed compactly in PPDDL to propositional
-representations amenable to state-of-the-art planning procedures.
+language that facilitates factored description of domains and problem
+instances. There are straightforward compilations from problems
+expressed compactly in PPDDL to propositional representations amenable
+to state-of-the-art planning procedures.
 
@@
 $\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs a penalty, $3$, corresponding to the {\em cost} of
-performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
-reserved word}, and occurs in {\em increased} and {\em decreased}
-clauses of operator schemata. In that setting, {\em reward} is
-interpreted as accumulated (resp. instantaneous) rewards and
-penalties.}
+execution incurs an instantaneous penalty, $3$, that corresponds to
+the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
+reward} is a {\em reserved word}, and occurs in {\em increased} and
+{\em decreased} terms in operator schemata. In that setting, {\em
+reward} is interpreted as accumulated, whereas in DTPDDL it is
+instantaneous.}
 

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
overline was silly for this paper

@@
 If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\overline\prop_{n+1},
-..,\overline\prop_m]$ is chosen by nature when \stochAction\ is
+..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
+..,\prop_m]$ is chosen by nature when \stochAction\ is
 applied at state $\state$, then the resultant state is $ ( \state \cup

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
tiny changes.

@@
 uncertain starting configurations, PPDDL is a declarative first-order
-language that facilitates factored description of domains and problem
+language that facilitates factored descriptions of domains and problem
 instances. There are straightforward compilations from problems
@@
 Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarations, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
+we developed an extension of that language, called decision-theoretic
+(DT)PDDL. This can express probabilistic models of the sensing
+consequences of acting, to quantitatively capture unreliability in
+perception. That expressive power is achieved by incorporating
+perceptual analogues of fluent, predicate, and action definitions. In
+detail, we have declarations of state characterising predicate and
+fluent symbols according to the PPDDL syntax. In addition, we allow
+two other declarations, for perceptual predicates and fluents
+respectively. For example, suppose our robot is tasked with exploring
+{\em locations} in order to identify the whereabouts of a {\em
+visual-object}. We must describe state and perceptual facts that model
+the {\em true}, resp. perceived, locations of objects. In DTPDDL,
+these declarations appear as:
 
@@
 {\em decreased} terms in operator schemata. In that setting, {\em
-reward} is interpreted as accumulated (resp. instantaneous) rewards
-and penalties.}
+reward} is interpreted as accumulated, whereas in DTPDDL it is
+instantaneous.}
 
@@
 \noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- I.e,
+the sense action-precondition -- i.e,
 $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office}) =
@@
 executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
+the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
 
@@
 be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
-expresses the conjunctive root of the tree -- I.e., the root node
+expresses the conjunctive root of the tree -- i.e., the root node
 $(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
tiny changes.

@@
 uses the structure of that, as it occurs in the problem description,
-do define a set of state-assumptions available to sequential planning.
+to define a set of state-assumptions available to sequential planning.
 %%
 Writing \#\ if the value of a proposition is unspecified in an
-assumption, for our $\pp{DORA}$ example we have that the following
+assumption, for our $\pp{DORA}$ example we have the following
 assumptions:
@@
 visited. The starting state, $\state_0$, for sequential planning is
-the unique abstract state with probability $1.0$ -- E.g., for
+the unique abstract state with probability $1.0$ -- e.g., for
 $\pp{DORA}$:
@@
 criteria corresponds exactly to that of
-\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
+\system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
 gradient-based reinforcement learning systems, such as William's
@@
 false assumption. The added dual actions are propositionally trivial
--- I.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
+-- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
 \subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
@@
 in the {\em relaxed} visitation characterised by {\em assumptive}
-actions in the sequential plan. Each iteration, we add one candidate
-term with the largest number of parent terms already in the candidate
-tree.
+actions in the sequential plan. In each iteration, we add one
+candidate term with the largest number of parent terms already in the
+candidate tree.

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
tiny changes.

--- 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
@@
 sensing. Although this has not been an issue in our work so far, it
-must be dealt with rigorously moving forward. We propose a good way of
-mitigation this problem is by using a {\em motivational} component
-that maintains a {\em dynamic} reward model for sensing.
+must be dealt with rigorously moving forward. We suggest that a good
+way of mitigate this problem is by using a {\em motivational}
+component that maintains a {\em dynamic} reward model for sensing.
 
@@
 
-In this paper we present an approach to continual planing that uses
+In this paper we present an approach to continual planning that uses
 two planning systems. The first

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
tiny changes.

@@
 its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
+satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
 \state$. We write $\stochActions(\state)$ for set of actions
@@
 applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- I.e.,
+\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
 propositions from $\add(\detAction)$ are added to $\state$, and those
@@
 unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, supposing that when
-an action \stochAction\ is executed at $\state_i$, that it has no
-effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
-\state_i$.
+we do incorporate a PPDDL-like default semantics, treating an action
+\stochAction\ executed at $\state_i$ as if it has no effect at
+states $\state$ if $\poss(\stochAction) \not\subseteq
+\state$.
 
@@
 planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
+achieves a goal ---i.e., reaches a state that satisfies some given
 Boolean condition--- with probability $1$.  More generally, a useful

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
I have to really rewrite the Switching section. Am doing that tonight and tomorrow.
 - Charles

@@
-We now describe our continual planning approach that switches between
-sequential and decision-theoretic planning sessions.
+We describe our continual planning approach that switches between
+sequential and decision-theoretic sessions. Continual planning and
+execution proceeds in a POMDP described in DTPDDL. During sequential
+sessions, planning is performed by a ``classical'' system,\footnote{By
+``classical'', we mean a system the is designed to solve fully
+observable deterministic tasks.}  and execution proceeds according to
+the {\em trace} computed by that system. In the format of a classical
+plan, that {\em trace} specifies a sequence of POMDP actions that
+achieves the agent's objectives for a deterministic approximation,
+i.e., {\em determinisation}, of the problem at hand. More precisely,
+the trace is an interleaved sequence of POMDP actions and {\em
+assumptive} actions. The latter give some of the assumptions the
+planner makes about the truth value of {\em runtime state variables}
+-- e.g. that a box of cornflakes is located in the kitchen at the
+third plan step. Another kind of assumption, called an {\em
+executability} assumption, is made when the trace includes an action
+$\action$ that is not applicable with probability $1$ at the
+belief-state
+\bstate\ that the system is projected to be in when \action\ is executed, 
+i.e., $\exists\state\in\states
+\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
+scheduling
+\action,  the serial planner makes an assumption about the
+observability of the precondition $\poss(\action)$.
 
-Sequential planning proceeds in a continual manner until the
-verification of an assumption about runtime state is scheduled. At
-that point the system switches to a decision-theoretic planner that is
-used to tailor sensory processing by the robot to test the validity of
-the assumptions made so far.
+A sequential session always proceeds first. Planning and execution are
+interleaved according to the continual paradigm until an executability
+assumption about an action, $\action$, scheduled for execution,
+occurs. At this point our approach switches to using a
+decision-theoretic solver which tailors sensory processing by the
+agent to test the validity of the assumptions made in the trace,
+including the truth value of $\poss(\action)$.
+
+%% Sequential planning proceeds in a continual manner until the
+%% verification of an assumption about runtime state is scheduled. At
+%% that point the system switches to a decision-theoretic planner that is
+%% used to tailor sensory processing by the robot to test the validity of
+%% the assumptions made so far.
 

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
starting paragraphs

@@
 \normalsize
+
+From hereon, in order to simplify the discussion (and implementation)
+we shall restrict our attention to POMDPs with deterministic
+actions. It is worthwhile noting that any POMDP with stochastic
+actions can be compiled into an equivalent deterministic-action POMDP,
+where all the original action uncertainty is expressed in the
+starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
+that of finite-horizon planning, such a compilation yields a finite
+flat representation of the original POMDP, only with deterministic
+actions.

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
starting paragraphs

@@
-We now describe our continual planning approach that switches between
-sequential and decision-theoretic planning sessions.
+We describe our continual planning approach that switches between
+sequential and contingent sessions. Continual planning and execution
+proceeds in a deterministic-action POMDP described in DTPDDL. During a
+sequential session, planning is performed by a ``classical''
+system,\footnote{That is, a system designed to solve fully observable
+deterministic tasks.}  and execution proceeds according to the {\em
+trace} computed by that system. Taking the form of a classical plan,
+the {\em trace} specifies a sequence of POMDP actions that achieve the
+agent's objectives in a deterministic approximation, i.e., {\em
+determinisation}, of the problem at hand. More precisely, the trace is
+an interleaved sequence of POMDP actions and {\em assumptive}
+actions. The latter correspond to assumptions the planner makes about
+the truth value of {\em runtime state variables} -- e.g. that a box of
+cornflakes is located in the kitchen at the third plan step. Another
+kind of assumption, called an {\em applicability} assumption, is made
+when the trace includes an action $\action$ that is not applicable
+with probability $1$ at the belief-state
+\bstate\ that the system is projected to be in when \action\ is executed, 
+i.e., $\exists\state\in\states
+\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
+scheduling
+\action,  the serial planner makes an assumption about the
+observability of the precondition $\poss(\action)$.
 
-Sequential planning proceeds in a continual manner until the
-verification of an assumption about runtime state is scheduled. At
-that point the system switches to a decision-theoretic planner that is
-used to tailor sensory processing by the robot to test the validity of
-the assumptions made so far.
-
-In this work, the problem at hand is a POMDP described in DTPDDL. From
-hereon, we restrict our attention to POMDPs with deterministic
-actions. Therefore it is worthwhile noting that any POMDP with
-stochastic actions can be compiled into an equivalent
-deterministic-action POMDP, where all the original action uncertainty
-is expressed in the starting-state
-distribution~\cite{ng:Jordan:2000}. In our setting, that of
-finite-horizon planning, such a compilation yields a finite flat
-representation of the original POMDP, only with deterministic
-actions.
+Our approach always begins with a sequential session. Non-assumptive
+actions from the trace are executed until the applicability of the
+next scheduled action, here \action, is questionable. A contingent
+session then begins, which tailors sensory processing by the agent to
+perceive whether the assumptions made in the trace hold. For each
+active assumption in the trace, we add a {\em disconfirm} action to
+the POMDP whose execution is rewarding (resp. costly) if the
+assumption is false (resp. true). We also add one {\em confirm} action
+that is rewarding (resp. costly) to execute if $\poss(\action)$ is
+true (resp. false). If execution of the contingent plan applies a
+disconfirm action, then a new sequential session begins at the current
+underlying belief-state. If the confirm action is executed, the
+sequential session is resumed, and \action\ is executed. Otherwise,
+control rests in the continual session.
+%%
+Finally, because contingent planning is only practical in relatively
+small POMDPs, contingent sessions proceed in an abstract decision
+process determined by the current trace. That abstract process is
+constructed by first omitting all variables from the model that are
+not featured in the trace, and by then iteratively refining that
+abstraction while the resulting model is of a practicable size.
 

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
switching is starting to mature

@@
 \bstate\ that the system is projected to be in when \action\ is executed, 
-i.e., $\exists\state\in\states
+i.e., $\exists\state\in\states\;
 \bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
@@
 
-Our approach uses sequential planning to compute a linear plan, and
-one trace of that plan's execution. The latter encapsulates
-assumptions about: (1) the true underlying state, (2) how the plan's
-execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions.
+A sequential session uses a classical planner to compute a trace. The
+latter encapsulates assumptions about: (1) the true underlying state,
+(2) how the plan's execution will progress, and (3) the possibility of
+the agent eventually holding strong beliefs about the truth values of
+specific state propositions. Here we describe the deterministic
+planning problem, derived from the DTPDDL model, that admits plans
+which correspond to traces.
 
-In a deterministic-action POMDP, all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
+In a deterministic-action POMDP all the uncertainty in state and
+action is expressed in the $(\pp{:init})$ DTPDDL declaration. Our
+approach uses the structure of that, as it occurs in the problem
+description, to define a set of state-assumptions available to
+sequential planning.
 %%
-Writing \#\ if the value of a proposition is unspecified in an
-assumption, for our $\pp{DORA}$ example we have the following
-assumptions:
+Writing \#\ if the value of a proposition is unspecified, for
+$\pp{DORA}$ we have the following assumptions:
 
@@
 
-\noindent A valid assumption in our setting corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this relaxed
-case, a conjunctive term is visited iff its atomic subterms are
-visited, and zero or more of its immediate probabilistic subterms are
-visited. The starting state, $\state_0$, for sequential planning is
-the unique abstract state with probability $1.0$ -- e.g., for
-$\pp{DORA}$:
+\noindent In detail, a valid assumption in our setting corresponds to a {\em
+relaxed} visitation of the root term of $(\pp{:init})$. In this
+relaxed case, a conjunctive term is visited iff its atomic subterms
+are visited, and zero or more of its immediate probabilistic subterms
+are visited. The starting state, $\state_0$, for sequential planning
+is an {\em abstract} state that captures the unique assumption which
+has probability $1.0$. In $\pp{DORA}$, that is:
 
@@
 $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ with atomic terms from
-$T_i$ and conflicting propositions from $s$ removed. For example, if
+$T_i$ and conflicting propositions from $s$ deleted. For example, if
 the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is a sink, where
-$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
+delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is an added sink,
+where $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$
+and $\transProb(\state^\bot, \action, \state^\bot) = 1$.
 
-Because $\state_0$ is abstract, and because actions can have
-preconditions that are not uniformly true in the underlying
-(resp. abstract) belief-state, we must make two further adjustments to
-the problem posed to the sequential planner. First, if
+For sequential planning the operators from the POMDP model are
+available with semantics that accommodate the above abstraction as
+follows. In the deterministic model a proposition $\prop$ can be
+thought to have a ternary interpretation at a state \state, as either
+{\em true}, written $\prop\in\state$, {\em false}, written
+$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
+notation written $\prop=\#\in\state$. For example, in $\pp{DORA}$
+$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
+$\action\in\actions$ in the deterministic model, if
+$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
+or negative effect of $\action$, then $\action$ is not applicable
+in \state\ if $\prop=\#\in\state$. Indeed, first an assumptive action
+must be applied to make a truth assignment to \prop, and then that
+truth assignment could be modified by POMDP operators. Lastly, we
+address the switching semantics of action execution. If
 $\action\in\actions$ has a precondition that is not true with
-probability $1$ in $\bstate_0$, then before it is executed blindly,
-the sequential continual planning switches to a decision-theoretic
-session, discussed in detail below. Overall, by scheduling $\action$
-the sequential planner makes a naive assumption about the possibility
-of the agent eventually ``committing'' to the truth values of specific
-state propositions in $\poss(\action)$. The role of decision-theoretic
-plans is to examine the validity of assumptions about the runtime
-state of sequential planning. Second, if in the current state \state\
-the value of $\prop$ is $\#$ ---because a truth value is not specified
-in $\state_0$ and no assumption is made about it in the plan
-to \state--- then if $\prop \in \poss(\action)$, or if $\prop$ is the
-subject of a positive or negative effect of $\action$, we have that
-action $\action$ is not applicable in \state.
+probability $1$ in the underlying belief $\bstate$, then before it is
+executed blindly, the system switches to a contingent session,
+discussed in detail in the following section.
 
-Finally, we come to valuing sequential plans, and therefore giving a
-criteria for the sequential planner to optimise. Where $p_i$ is the
-probability that the $i^{th}$ sequenced action does not transition to
-$s^\bot$, we value a trace as follows:
+We now describe the optimisation criteria given to the classical
+planner. Where $\prob_i$ is the probability that the $i^{th}$
+sequenced action $\action_i$ does not transition to $\state^\bot$, we
+define the value of a trace
+$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
+deterministic model to equal:
 
 \begin{equation}\label{eq:tracevalue}
-V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N-1} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
+V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
 \end{equation}
 
-\noindent The optimal trace given a plan, i.e., the non-assumptive
-actions in sequence, therefore has value:
+\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
+actions from the trace, therefore has value:
 
 \[
-V^* = \max_{p_1, .., p_n} \prod_{i=1..N-1} p_i \sum_{i=1..N-1}
+V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
 \reward(\state_i, \action_i),
@@
 
-\noindent which is equal to the maximal contribution a trace of the plan can
-make to Eq~\ref{eq:expectedvalue}. %% In other words, writing $V_i$ for
-%% the value of the $i^{th}$ trace of the plan expressed in $s_0, a_0,
-%% s_1, a_1,.., s_N$, we have
-%%
-%% \[
-%% V^* = \arg\min_{V_i} \Expect \bigg{[} 
-%% \sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0
-%% \bigg{]} - V_i.
-%% \]
-%%
-In goal directed problems, where the only non-zero reward is received
-at the first transition to a goal state, the Eq~\ref{eq:tracevalue}
-criteria corresponds exactly to that of
-\system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
+\noindent which is equal to the maximal additive contribution a trace of the plan can
+make to Eq~\ref{eq:expectedvalue}, taking an FSC that corresponds to
+the sequential plan.  In goal directed problems, where the only
+non-zero reward is received at the first transition to a goal state,
+the Eq~\ref{eq:tracevalue} criteria gives us the behaviour
+of \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
 gradient-based reinforcement learning systems, such as William's

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
the POP stuff seems now like it is a detail. I don't see what he compilation is, there is an optimisation problem there. Is there something we can cite for that?

--- 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
@@
-We now describe our continual planning approach that switches between
-sequential and decision-theoretic planning sessions.
+We describe our continual planning approach that switches between
+sequential and contingent sessions. Continual planning and execution
+proceeds in a deterministic-action POMDP described in DTPDDL. During a
+sequential session, planning is performed by a ``classical''
+system,\footnote{That is, a system designed to solve fully observable
+deterministic tasks.}  and execution proceeds according to the {\em
+trace} computed by that system. Taking the form of a classical plan,
+the {\em trace} specifies a sequence of POMDP actions that achieve the
+agent's objectives in a deterministic approximation, i.e., {\em
+determinisation}, of the problem at hand. More precisely, the trace is
+an interleaved sequence of POMDP actions and {\em assumptive}
+actions. The latter correspond to assumptions the planner makes about
+the truth value of {\em runtime state variables} -- e.g. that a box of
+cornflakes is located in the kitchen at the third plan step. Another
+kind of assumption, called an {\em applicability} assumption, is made
+when the trace includes an action $\action$ that is not applicable
+with probability $1$ at the belief-state
+\bstate\ that the system is projected to be in when \action\ is executed, 
+i.e., $\exists\state\in\states\;
+\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
+scheduling
+\action,  the serial planner makes an assumption about the
+observability of the precondition $\poss(\action)$.
 
-Sequential planning proceeds in a continual manner until the
-verification of an assumption about runtime state is scheduled. At
-that point the system switches to a decision-theoretic planner that is
-used to tailor sensory processing by the robot to test the validity of
-the assumptions made so far.
-
-In this work, the problem at hand is a POMDP described in DTPDDL. From
-hereon, we restrict our attention to POMDPs with deterministic
-actions. Therefore it is worthwhile noting that any POMDP with
-stochastic actions can be compiled into an equivalent
-deterministic-action POMDP, where all the original action uncertainty
-is expressed in the starting-state
-distribution~\cite{ng:Jordan:2000}. In our setting, that of
-finite-horizon planning, such a compilation yields a finite flat
-representation of the original POMDP, only with deterministic
-actions.
+Our approach always begins with a sequential session. Non-assumptive
+actions from the trace are executed until the applicability of the
+next scheduled action, here \action, is questionable. A contingent
+session then begins, which tailors sensory processing by the agent to
+perceive whether the assumptions made in the trace hold. For each
+active assumption in the trace, we add a {\em disconfirm} action to
+the POMDP whose execution is rewarding (resp. costly) if the
+assumption is false (resp. true). We also add one {\em confirm} action
+that is rewarding (resp. costly) to execute if $\poss(\action)$ is
+true (resp. false). If execution of the contingent plan applies a
+disconfirm action, then a new sequential session begins at the current
+underlying belief-state. If the confirm action is executed, the
+sequential session is resumed, and \action\ is executed. Otherwise,
+control rests in the continual session.
+%%
+Finally, because contingent planning is only practical in relatively
+small POMDPs, contingent sessions proceed in an abstract decision
+process determined by the current trace. That abstract process is
+constructed by first omitting all variables from the model that are
+not featured in the trace, and by then iteratively refining that
+abstraction while the resulting model is of a practicable size.
 
@@
 
-In a decision-theoretic session, the switching planner solves a sensing
-problem in an abstract processes defined in terms of the assumptive
-actions in the current sequential plan, and the $(\pp{:init})$ declaration
-from the original problem description.
+In a contingent session our switching planner solves a sensing problem
+in an abstract processes defined in terms of the assumptive actions in
+the current sequential plan, and the $(\pp{:init})$ declaration from
+the original problem description.
 
-The focus for decision-theoretic planning is on sensing, rather than
-the achievement of utility according to the given objectives. We
-therefor give the process in this session a different reward model, which
-reflects the value of performing sensing actions in the context of the
-trace proposed for the preceding sequential deliberation. In detail,
-all positive rewards from the initial POMDP are removed. Then, for
-each $\assumptiveS{i}$ action scheduled in the current sequential
-plan, we have a {\em dual} $\assumptiveDT{i}$ so that
-$\forall\state\in\states$:
+The focus for contingent planning is on sensing, we therefore give the
+process in this session an augmented reward model.  This reflects the
+value of performing sensing actions in the context of the trace
+proposed by the preceding sequential deliberation. First, all rewards
+from the original POMDP are retained. Then, for each $\assumptiveS{i}$
+action scheduled by the current trace, we have a {\em dual}
+$\assumptiveDT{i}$ so that $\forall\state\in\states$:
 
@@
 \noindent Where $\$(T_i)$ is a positive numeric quantity which
-captures the reward the agent receives for correctly jettisoning a
+captures the reward the agent receives for correctly rejecting a
 false assumption. The added dual actions are propositionally trivial
 -- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
-$\action_i$, whose scheduled execution triggered the switch to the
-decision-theoretic session, we include another propositionally trivial
-action $\actions.\poss(\action_i)$ with the reward property:
+\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
+%%
+In order to reduce the scope of possible rejections, we compile the
+serial trace into a partially ordered plan (POP). If $\action_i$ is
+the sequentially scheduled execution that triggered the switch to the
+contingent session, then we discard assumptive duals that are not
+ordered before $\action_i$ in that POP.
+%%
+Also for the action $\action_i$, we include another propositionally
+trivial action $\actions.\poss(\action_i)$ with the reward property:
 
@@
 abstraction of the environment that does not contain a $\pp{cup}$,
-where the inclusion (exclusion) of POMDP actions is the same as we
+where the applicability of POMDP actions is the same as we
 defined it for the sequential session.
@@
 practicable size. Of course, it is expensive to measure size exactly,
-and in practice we estimate this to be $2^{|\propositions^\#|}$, where
+and in practise we estimate this to be $2^{|\propositions^\#|}$, where
 $\propositions^\#$ is the set of propositions in the abstraction that

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
got new draft of switching, and Bayesian stuff for the evaluation section

% coming soon

Whether proceeding in a sequential or contingent session, our
continual planner must maintain a factored representation of the
current underlying belief-state, by performing belief revision
according to Eq~\ref{eq:revision}. That distribution corresponds to
the source of determinisations for sequential planning, the basis for
switching, and as the source for state abstraction in contingent
sessions. In our implementation we have
used~\citeauthor{king:2009}~(\citeyear{king:2009}) for this purpose.

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
got new draft of switching, and Bayesian stuff for the evaluation section

@@
 session then begins, which tailors sensory processing by the agent to
-perceive whether the assumptions made in the trace hold. For each
-active assumption in the trace, we add a {\em disconfirm} action to
+perceive whether the assumptions made in the trace hold. For each {\em
+active} assumption in the trace, we add a {\em disconfirm} action to
 the POMDP whose execution is rewarding (resp. costly) if the
@@
 control rests in the continual session.
-%%
-Finally, because contingent planning is only practical in relatively
-small POMDPs, contingent sessions proceed in an abstract decision
-process determined by the current trace. That abstract process is
-constructed by first omitting all variables from the model that are
-not featured in the trace, and by then iteratively refining that
-abstraction while the resulting model is of a practicable size.
 
-PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-HOW WE DO BELIEF REVISION ACROSS SWITCHING.
+Because contingent planning is only practical in relatively small
+POMDPs, contingent sessions proceed in an abstract decision process
+determined by the current trace. That abstract process is constructed
+by first omitting all variables from the model that are not featured
+in the trace, and by then iteratively refining that abstraction while
+the resulting model is of a practicable size.
+
+%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
+%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
 
@@
 $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-notation written $\prop=\#\in\state$. For example, in $\pp{DORA}$
+notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
 $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
@@
 or negative effect of $\action$, then $\action$ is not applicable
-in \state\ if $\prop=\#\in\state$. Indeed, first an assumptive action
+in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
 must be applied to make a truth assignment to \prop, and then that
@@
 
-In a decision-theoretic session, the switching planner solves a sensing
-problem in an abstract processes defined in terms of the assumptive
-actions in the current sequential plan, and the $(\pp{:init})$ declaration
-from the original problem description.
+In a contingent session our switching planner solves a sensing problem
+in an abstract processes defined in terms of the assumptive actions in
+the current trace, and the $(\pp{:init})$ declaration from the
+original problem description.
 
-The focus for decision-theoretic planning is on sensing, rather than
-the achievement of utility according to the given objectives. We
-therefor give the process in this session a different reward model, which
-reflects the value of performing sensing actions in the context of the
-trace proposed for the preceding sequential deliberation. In detail,
-all positive rewards from the initial POMDP are removed. Then, for
-each $\assumptiveS{i}$ action scheduled in the current sequential
-plan, we have a {\em dual} $\assumptiveDT{i}$ so that
-$\forall\state\in\states$:
+The focus for contingent planning is on sensing, we therefore give the
+process in this session an augmented reward model.  This reflects the
+value of performing sensing actions in the context of the trace
+proposed by the preceding sequential deliberation. First, all rewards
+from the original POMDP are retained. Then, for each $\assumptiveS{i}$
+action scheduled by the current trace, we have a {\em dual}
+$\assumptiveDT{i}$ so that $\forall\state\in\states$:
 
@@
 \noindent Where $\$(T_i)$ is a positive numeric quantity which
-captures the reward the agent receives for correctly jettisoning a
+captures the reward the agent receives for correctly rejecting a
 false assumption. The added dual actions are propositionally trivial
 -- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
-$\action_i$, whose scheduled execution triggered the switch to the
-decision-theoretic session, we include another propositionally trivial
-action $\actions.\poss(\action_i)$ with the reward property:
+\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
+%%
+In order to reduce the number of possible rejections we do not
+consider assumptions that are not {\em active} with respect to the
+action, $\action_i$, whose scheduled execution switch the system into
+the contingent session. A assumption $\assumptiveS{i}$ is active if
+$\action_i$ is not applicable in the sequential model unless
+$\assumptiveS{i}$ is scheduled in the trace prefix to $\action_i$. For
+example, if $\pp{DORA}$'s trace is:
+
+\[
+\begin{array}{l}
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
+\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
+(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
+(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
+(\pp{report-is-in}~\pp{box}~\pp{office}); \\
+(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
+\end{array}
+\]
+
+\noindent Taking the switching action $\action_i$ to be
+$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
+$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
+active, and therefore exclude
+$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
+POMDP posed to the contingent session.
+
+Again taking $\action_i$ to be the switching action, we include
+another propositionally trivial action $\actions.\poss(\action_i)$
+with the reward property:
 
@@
 
-For the management of switching, execution of either a dual action of
-the form $\assumptiveDT{i}$, or the $\actions.\poss(\action_i)$,
-returns control to the sequential planner. At that point, if a dual is
-executed, then the sequential planner must replan with $\bstate_0$
-equal to the current belief. Otherwise, if $\actions.\poss(\action_i)$
-is executed, then the current sequential plan is executed until
-further sensing is scheduled, or to completion. In our project the
-rewards, i.e., function $\$:2^\propositions\to\Re$, for the POMDP
-posed to decision-theoretic planning are sourced from a motivational
-component of our robotic architecture.
+For the purposes of successive switching, execution of either a dual
+action of the form $\assumptiveDT{i}$, or the
+$\actions.\poss(\action_i)$, returns control to a sequential
+session. If a dual is executed, then the sequential planner must
+replan with $\bstate_0$ equal to the underlying
+belief-state. Otherwise, if $\actions.\poss(\action_i)$ is executed,
+then the current sequential plan is executed until further sensing is
+scheduled, or to completion. In our project the rewards, i.e.,
+function $\$:2^\propositions\to\Re$, for the POMDP posed to the
+contingent session are sourced from a motivational component of our
+robotic architecture.
 
@@
 tree which only includes terms from the {\em relaxed} visitation
-characterised by the {\em assumptive} actions in the current
-sequential plan. For example, if $\pp{DORA}$'s plan is:
-
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{report-is-in}~\pp{box}~\pp{office})
-\end{array}
-\]
-
-\noindent Then the {\em relaxed} declaration will be,
+characterised by the active {\em assumptive} actions in the current
+sequential plan. Taking our example trace from earlier, we have a {\em
+relaxed} declaration:
 
@@
 
-\noindent At this point, decision-theoretic planning would proceed in an
-abstraction of the environment that does not contain a $\pp{cup}$,
-where the inclusion (exclusion) of POMDP actions is the same as we
-defined it for the sequential session.
+\noindent At this point, the session would proceed in an
+abstraction of the environment that does not contain a
+$\pp{cup}$. With regards to abstraction, the contingent and serial
+sessions apply the same applicability conditions ---i.e., that each
+effect and condition $\prop$ is specified $\prop\in\state$
+(resp. $\prop\#\in\state$)--- to actions from the DTPDDL problem
+description.
 
-In a second step, we iteratively refine the above{\em relaxed}
-declaration by adding terms from the original statement of
-$(\pp{:init})$ while the resulting state space remains of a
-practicable size. Of course, it is expensive to measure size exactly,
-and in practice we estimate this to be $2^{|\propositions^\#|}$, where
-$\propositions^\#$ is the set of propositions in the abstraction that
-were not assigned the {\em unknown} value \#. Terms are proposed for
-addition because they are probabilistic, and because they were omitted
-in the {\em relaxed} visitation characterised by {\em assumptive}
-actions in the sequential plan. In each iteration, we add one
-candidate term with the largest number of parent terms already in the
-candidate tree.
+In a second step, we iteratively refine the relaxed declaration by
+adding terms from the original statement of $(\pp{:init})$ while the
+number of abstract states that occur with non-zero probability
+according to that refined declaration remains of a practicable
+size. In detail, for each excluded atomic term, we compute the {\em
+entropy} of the corresponding proposition, {\em conditional} on the
+active assumptions. The {\em conditional entropy} measure
+is higher the more uncertain that proposition is given truth
+assignments to propositions that are the subject of active
+assumptions. Propositions that in the first place were excluded are
+iteratively added to the problem posed to the conditional session in
+increasing order according to that measure.
+
+%% In particular, we compute the {\em conditional entropy} of
+%% each excluded proposition given the active assumptions. Propositions
+%% are candidates for addition in increasing order according to that
+%% measure.
+
+%% Of course, it is expensive to measure size exactly,
+%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
+%% $\propositions^\#$ is the set of propositions in the abstraction that
+%% were not assigned the {\em unknown} value \#. Terms are proposed for
+%% addition because they are probabilistic, and because they were omitted
+%% in the {\em relaxed} visitation characterised by {\em assumptive}
+%% actions in the sequential plan. In each iteration, we add one
+%% candidate term with the largest number of parent terms already in the
+%% candidate tree.

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
New introduction added. See what you think! Lots of references added too.

@@
-We describe, \pcogx, the planning component in a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings.
+Planning for agents acting in partially observable environments with
+stochastic actions is extremely computationally
+expensive \cite{mdp-complexity}. Even relatively small partially
+observable Markov decision problems (POMDPs), represented by tens of
+propositional variables, produce state spaces far larger than existing
+optimal or close to optimal planners can handle. For the easier
+problem of planning in completely observable stochastic domains, an
+alternative approach has been to use classical sequential planners and
+replan when actions have unplanned-for
+outcomes \cite{yoon:etal:2007}. However, these approaches aren't so
+useful in partially observable domains as replanning is triggered when
+the system ends up in a state other than the desired one, and this
+state can't in general be determined.
+
+In this paper we describe, \pcogx, the planning component in a robotic
+system that continuously deliberates in a stochastic dynamic
+environment in order to achieve objectives set by the user, and
+acquire knowledge about its surroundings. This domain has the
+characteristics of a POMDP in that the state is not perfectly known,
+and state information is gained by the robot through the use of
+unreliable sensing actions. However, the state representation is
+propositional, and even for relatively simple problems is too large
+for a POMDP solver to be applied.
+
+Inspired by the replanning approaches used successfully for MDP
+planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
+that uses a replanning sequential planner to build the majority of the
+plan, but uses a decision-theoretic planner on small subproblems where
+reasoning about observations is required to find good plans. We refer
+to the complete system as a {\em switching planner}. The approach is
+not optimal, particularly as it is based on a satisficing sequential
+planner, but it performs better than the sequential planner alone and
+is fast enough to be used onboard the robot for decision-making. As an
+illustration, consider a robot searching for a box that it knows is
+often found in offices. The sequential planner might produce a plan
+that assumes that {\sc Room a} is an office, that the box is in {\sc
+Room a}, and that the box is in location 1 in {\sc Room a}, and then
+goes to {\sc Room a}, goes to location 1 and searches for the
+box. When the plan gets executed and the search begins, the
+decision-theoretic planner is called to conduct the search, and at
+this point may retract the last assumption and build a plan to search
+the whole of {\sc Room a} for the box. If this plan fails, it will do
+so by disproving one of the assumptions (for example, that {\sc Room
+a} is an office), which will then lead to replanning by the sequential
+planner based on the newly discovered knowledge.
+
+To represent planning problems of this kind, we have developed a
+first-order declarative language, called decision-theoretic planning
+domain definition language (DTPDDL), for describing domains of
+planning problems that correspond to POMDPs. We have created a system
+that automatically constructs from this a determinised planning domain
+for the sequential planner that includes actions that allow
+assumptions to be made about the values of imperfectly known state
+variables. These are then used to restrict the state space and set the
+reward function for the decision-theoretic planner, which can operate
+with the original representation of the domain.
+
+The sequential planner we use is Continual
+Planning \cite{brenner:nebel:jaamas09} (CP). CP is based
+on \fastdownward \cite{fast-downward}. CP already has the capability
+to replan, and also allows agent knowledge to be represented
+explicitly, so we can write actions that gain the agent knowledge. For
+the decision-theoretic planner we use an information-state
+\laostar\ procedure that uses the solution to the equivalent MDP as a
+heuristic. For reasons of space, we will not describe this system in
+detail here.
+
+The switching planner approach gives us some important advantages in
+our robotic domain. Firstly, the main source of uncertainty in the
+domain is unreliable observations, rather than stochasticity in the
+action effects, and this approach is particularly suited to problems
+that combine a deterministic task planning problem with observations
+as the decision-theoretic planner is only invoked to determine the
+values of variables. Secondly, the replanning makes the system very
+robust to changing goals and the discovery of new facts about the
+world, both of which are features of our domain.
+
+%To accompany
+%that language we have implemented an information-state
+%\laostar\ procedure for solving problems expressed in DTPDDL.
+
 %%
 
-Have a good understanding of the contents and function of rooms, as
-well as the linguistic referrents to rooms, widgets, and their visual
-qualities.
+%Have a good understanding of the contents and function of rooms, as
+%well as the linguistic referrents to rooms, widgets, and their visual
+%qualities.
 
-Having confidence in its beliefs about the linguistic terms for
-spaces, and the function of those spaces.
+%Having confidence in its beliefs about the linguistic terms for
+%spaces, and the function of those spaces.
 
-sometimes difficult to predict, with exogenous events, such as the
-changing of the objective,
+%sometimes difficult to predict, with exogenous events, such as the
+%changing of the objective,
 
-the speed and scalability of software for sequential planning in
-deterministic
+%the speed and scalability of software for sequential planning in
+%deterministic
 
-We developed a first-order declarative language, called DTPDDL, for
-describing domains of planning problems that correspond to POMDPs.  To
-accompany that language we have implemented an information-state
-\laostar\ procedure for solving problems expressed in DTPDDL.
+%% We developed a first-order declarative language, called DTPDDL, for
+%% describing domains of planning problems that correspond to POMDPs.  To
+%% accompany that language we have implemented an information-state
+%% \laostar\ procedure for solving problems expressed in DTPDDL.
 
-We have extended MAPSIM to parse and simulate DTPDDL problems, and
-modified \fastdownward\ so that it can find useful sequential executions
-given DTPDDL models of the problem at hand.
+%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
+%% modified \fastdownward\ so that it can find useful sequential executions
+%% given DTPDDL models of the problem at hand.
 
- called DTPDDL, along
-the lines of PDDL for the partially observable case, an \laostar\
-solution procedure, a determinisation of the DTPDDL problem in MAPL,
-and modify the \fastdownward\ system to find high quality sequential plans
-.
+%%  called DTPDDL, along
+%% the lines of PDDL for the partially observable case, an \laostar\
+%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
+%% and modify the \fastdownward\ system to find high quality sequential plans
+%% .
 
-We model the environment as a partially observable Markov decision
-process. Although planning in that model is undecidable in
-general~\cite{}, an optimal finite-horizon plan corresponds to a
-contingent plan, that is a function mapping action-observation
-histories to actions.
+%% We model the environment as a partially observable Markov decision
+%% process. Although planning in that model is undecidable in
+%% general~\cite{}, an optimal finite-horizon plan corresponds to a
+%% contingent plan, that is a function mapping action-observation
+%% histories to actions.
 
-our continual planner is reactive, replanning whenever the underlying
-domain and problem models change. For example, replanning occurs if
-the motivational component alters the objectives, and if an assumed
-outcome of a sensing action is not realised.
+%% our continual planner is reactive, replanning whenever the underlying
+%% domain and problem models change. For example, replanning occurs if
+%% the motivational component alters the objectives, and if an assumed
+%% outcome of a sensing action is not realised.
 
-brittle sensing model,
+%% brittle sensing model,
 
-the evaluation of a fluent at a state is either known. Moreover, there
-is a sensing process that run-time variables  after-which
+%% the evaluation of a fluent at a state is either known. Moreover, there
+%% is a sensing process that run-time variables  after-which
 
@@
 
-An outline of the paper is as follows. We describe POMDPs with a
+An outline of the paper is as follows. We begin by describing POMDPs with a
 propositionally factored representation of states and observations,
-and describe how to evaluate plans when they correspond to
-finite-state controllers.
+and therelationship between this and a flat POMDP representation. Next
+we describe our DTPDDL language using a running example of the robot
+looking for a box. Then we describe the switching planner, including
+how the sequential domain is produced and a plan is generated, and how
+this then provides the input for the decision-theoretic
+planner. Finally we provide an evaluation of the system on the example
+robot domain, and finish with a short survey of related work and some
+concluding remarks.

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
tiny changes.
Added to the related work section.

@@
 under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably,
+planning in deterministic problems.  Notably for example,
 \system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
@@
 robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-%%
+plan~\cite{majercik:2006}. However, as we said in the introduction,
+all these approaches struggle in partially observable domains as they
+rely on being able to determine the state at all times.
+
 Also leveraging deterministic planners in problems that feature
@@
 environments~\cite{albore:etal:2009}.
+%%
+The continual planning algorithm we use \cite{brenner:nebel:jaamas09}
+also has this characteristic, and has been applied in completely
+observable domains, particularly those featuring multple communicating
+agents. The use of knowledge operators in domains allows plans that
+act to gain knowledge, but the approach assumes that such actions are
+deterministic and reliable, an assumption that we relax in this work.
+
+The particular domain we are interested in contains a mixture of task
+planning and observation planning. There have been a number of recent
+papers representing observation planning problems as POMDPs and using
+various techniques to manage the large state
+space. In \cite{hippo-jnl} the authors take this approach in a vision
+algorithm selection problem. In their case there is a natural
+factorisation of the problem which allows them to solve large problems
+by decomposing them into a set of much smaller POMDPs. Doshi and
+Roy \cite{doshi08:pref_elic} represent a preference elicitation
+problem as a POMDP and take advantage of symmetry in the belief space
+(essentially the idea that it doesn't matter what the value of the
+variable you are trying to observe is) to exponentially shrink the
+state space. Unfortunately, although we are still exploring the Doshi
+and Roy approach, we have yet to find structure we can exploit
+similarly in our domain due to the task planning component.

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
Few updates

@@
 
-Whether proceeding in a sequential or contingent session, our
-continual planner must maintain a factored representation of the
-current underlying belief-state, by performing belief revision
-according to Eq~\ref{eq:revision}. That distribution corresponds to
-the source of determinisations for sequential planning, the basis for
-switching, and as the source for state abstraction in contingent
-sessions. In our implementation we have
-used~\citeauthor{king:2009}~(\citeyear{king:2009}) for this purpose.
+In our implementation uses the
+~\citeauthor{king:2009}~(\citeyear{king:2009}) \system{Dlib-ml} for
+successive estimation of the underlying belief-state.

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Few updates

@@
 
-\section{Switching Planner}
+\section{Switching Continual Planner}
 \input{switching.tex}

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
Few updates

@@
 Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called decision-theoretic
+we developed an extension of that language, called Decision-Theoretic
 (DT)PDDL. This can express probabilistic models of the sensing
@@
 $(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
-$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
+$(\pp{and}~\prop)$, it a conjunctive term contains a single atomic
 subterm.}

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
Few updates

--- 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
@@
 environments~\cite{albore:etal:2009}.
+%%
+The continual planning system that motivated our
+project~\cite{brenner:nebel:jaamas09} also has this characteristic,
+and has been applied in completely observable domains, particularly
+those featuring multple communicating agents. The use of knowledge
+operators in domains allows plans that act to gain knowledge, but the
+approach assumes that such actions are deterministic and reliable, an
+assumption that we relax in this work.
+
+Our work is motivated by domains that contain a mixture of task
+planning and observation planning. There have been a number of recent
+papers representing observation planning problems as POMDPs and using
+various techniques to manage the large state
+space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
+approach in a vision algorithm selection problem. In their case there
+is a natural hierarchical decomposition of the problem which allows
+them to solve large problems by breaking them into a set of small
+POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
+represent a preference elicitation problem as a POMDP and take
+advantage of symmetry in the belief space ---essentially the idea that
+it does not matter what the value of the variable you are trying to
+observe is--- to exponentially shrink the state space. Although we
+have been actively exploring the \citeauthor{doshi08:pref_elic}
+approach, we have yet to find those exploitable structures in our
+domain due to the task planning component.

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Few updates

@@
 We describe our continual planning approach that switches between
-sequential and contingent sessions. Continual planning and execution
-proceeds in a deterministic-action POMDP described in DTPDDL. During a
-sequential session, planning is performed by a ``classical''
-system,\footnote{That is, a system designed to solve fully observable
-deterministic tasks.}  and execution proceeds according to the {\em
-trace} computed by that system. Taking the form of a classical plan,
-the {\em trace} specifies a sequence of POMDP actions that achieve the
-agent's objectives in a deterministic approximation, i.e., {\em
-determinisation}, of the problem at hand. More precisely, the trace is
-an interleaved sequence of POMDP actions and {\em assumptive}
-actions. The latter correspond to assumptions the planner makes about
-the truth value of {\em runtime state variables} -- e.g. that a box of
-cornflakes is located in the kitchen at the third plan step. Another
-kind of assumption, called an {\em applicability} assumption, is made
-when the trace includes an action $\action$ that is not applicable
-with probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is executed, 
-i.e., $\exists\state\in\states\;
+sequential and contingent sessions. This proceeds by interleaved
+planning and execution in a deterministic-action POMDP described in
+DTPDDL. During a sequential session, planning is performed by a
+``classical'' system,\footnote{That is, a system designed to solve
+fully observable deterministic tasks.}  and execution proceeds
+according to the {\em trace} computed by that system. Taking the form
+of a classical plan, the trace specifies a sequence of POMDP
+actions that achieve the agent's objectives in a deterministic
+approximation, i.e., {\em determinisation}, of the problem at
+hand. More precisely, the trace is an interleaved sequence of POMDP
+actions and {\em assumptive} actions. The latter correspond to
+assumptions the planner makes about the truth value of {\em runtime
+state variables} -- e.g. that a box of cornflakes is located in the
+kitchen at the third plan step. Another kind of assumption, called an
+{\em applicability} assumption, is made when the trace includes an
+action $\action$ that is not applicable with probability $1$ at the
+belief-state
+\bstate\ that the system is projected to be in when \action\ is
+scheduled for execution, i.e., $\exists\state\in\states\;
 \bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
@@
 actions from the trace are executed until the applicability of the
-next scheduled action, here \action, is questionable. A contingent
-session then begins, which tailors sensory processing by the agent to
-perceive whether the assumptions made in the trace hold. For each
-active assumption in the trace, we add a {\em disconfirm} action to
-the POMDP whose execution is rewarding (resp. costly) if the
-assumption is false (resp. true). We also add one {\em confirm} action
-that is rewarding (resp. costly) to execute if $\poss(\action)$ is
-true (resp. false). If execution of the contingent plan applies a
+next scheduled action is uncertain.  A contingent session then begins,
+which tailors sensory processing by the agent to determine whether the
+assumptions made in the trace hold. For each {\em active} assumption
+in the trace, we add a {\em disconfirm} action to the POMDP whose
+execution is rewarding (resp. costly) if the assumption is false
+(resp. true). We also add one {\em confirm} action that is rewarding
+(resp. costly) to execute if $\poss(\action)$ is true
+(resp. false). If execution of the contingent plan applies a
 disconfirm action, then a new sequential session begins at the current
 underlying belief-state. If the confirm action is executed, the
-sequential session is resumed, and \action\ is executed. Otherwise,
-control rests in the continual session.
-%%
-Finally, because contingent planning is only practical in relatively
-small POMDPs, contingent sessions proceed in an abstract decision
-process determined by the current trace. That abstract process is
-constructed by first omitting all variables from the model that are
-not featured in the trace, and by then iteratively refining that
-abstraction while the resulting model is of a practicable size.
+sequential session is resumed, and
+\switchAction\ is executed. Otherwise, control rests in the continual
+session.
 
-PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-HOW WE DO BELIEF REVISION ACROSS SWITCHING.
+Because contingent planning is only practical in relatively small
+POMDPs, contingent sessions proceed in an abstract decision process
+determined by the current trace. That abstract process is constructed
+by first omitting all variables from the model that are not featured
+in the trace, and by then iteratively refining that abstraction while
+the resulting model is of a practicable size.
+
+Finally, whether proceeding in a sequential or contingent session, our
+continual planner maintains a factored representation of successive
+{\em belief-states}, by performing belief revision according to
+Eq~\ref{eq:revision}. Moreover, our internal representation of that
+distribution corresponds more-or-less directly to an $(\pp{:init})$
+declaration in DTPDDL. Our approach uses that {\em underlying}
+belief-state as: (1) the source of determinisations for sequential
+planning, (2) the basis for switching, and (3) the source for state
+abstraction in contingent sessions.
+
+%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
+%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
 
@@
 or negative effect of $\action$, then $\action$ is not applicable
-in \state\ if $\prop=\#\in\state$. Indeed, first an assumptive action
+in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
 must be applied to make a truth assignment to \prop, and then that
-truth assignment could be modified by POMDP operators. Lastly, we
-address the switching semantics of action execution. If
+truth assignment could be modified by POMDP operators.
+
+Addressing now the switching semantics of action execution. If
 $\action\in\actions$ has a precondition that is not true with
@@
 executed blindly, the system switches to a contingent session,
-discussed in detail in the following section.
+discussed in the following section. In detail, we halt the sequential
+session at \action\ if $\exists
+\state$ s.t. $\bstate(\state)>0$, 
+$\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
+semantics of action execution there must be a $\state'$ s.t.
+$\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
+not have been scheduled at \bstate.
+
+then $\reward(\state,\action)=10$, and
+otherwise if $\prop\not\in\state'$ then
+$\reward(\state',\action)=-10$.
 
@@
 in an abstract processes defined in terms of the assumptive actions in
-the current sequential plan, and the $(\pp{:init})$ declaration from
-the original problem description.
+the current trace, and the $(\pp{:init})$ declaration from the
+original problem description.
 
@@
 %%
-In order to reduce the scope of possible rejections, we compile the
-serial trace into a partially ordered plan (POP). If $\action_i$ is
-the sequentially scheduled execution that triggered the switch to the
-contingent session, then we discard assumptive duals that are not
-ordered before $\action_i$ in that POP.
-%%
-Also for the action $\action_i$, we include another propositionally
-trivial action $\actions.\poss(\action_i)$ with the reward property:
+In order to reduce the number of possible rejections we do not
+consider assumptions that are not {\em active} with respect to the
+action, $\action_i$, whose scheduled execution switch the system into
+the contingent session. A assumption $\assumptiveS{i}$ is active if
+$\action_i$ is not applicable in the sequential model unless
+$\assumptiveS{i}$ is scheduled in the trace prefix to $\action_i$. For
+example, if $\pp{DORA}$'s trace is:
+
+\[
+\begin{array}{l}
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
+\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
+(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
+(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
+(\pp{report-is-in}~\pp{box}~\pp{office}); \\
+(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
+\end{array}
+\]
+
+\noindent Taking the switching action $\action_i$ to be
+$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
+$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
+active, and therefore exclude
+$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
+POMDP posed to the contingent session.
+
+Again taking $\action_i$ to be the switching action, we include
+another propositionally trivial action $\actions.\poss(\action_i)$
+with the reward property:
 
@@
 
-\noindent At this point, decision-theoretic planning would proceed in an
-abstraction of the environment that does not contain a $\pp{cup}$,
-where the applicability of POMDP actions is the same as we
-defined it for the sequential session.
+\noindent At this point, the session would proceed in an
+abstraction of the environment that does not contain a
+$\pp{cup}$. With regards to abstraction, the contingent and serial
+sessions apply the same applicability conditions ---i.e., that each
+effect and condition $\prop$ is specified $\prop\in\state$
+(resp. $\prop\#\in\state$)--- to actions from the DTPDDL problem
+description.
 
-In a second step, we iteratively refine the above{\em relaxed}
-declaration by adding terms from the original statement of
-$(\pp{:init})$ while the resulting state space remains of a
-practicable size. Of course, it is expensive to measure size exactly,
-and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-$\propositions^\#$ is the set of propositions in the abstraction that
-were not assigned the {\em unknown} value \#. Terms are proposed for
-addition because they are probabilistic, and because they were omitted
-in the {\em relaxed} visitation characterised by {\em assumptive}
-actions in the sequential plan. In each iteration, we add one
-candidate term with the largest number of parent terms already in the
-candidate tree.
+In a second step, we iteratively refine the relaxed declaration by
+adding terms from the original statement of $(\pp{:init})$ while the
+number of abstract states that occur with non-zero probability
+according to that refined declaration remains of a practicable
+size. In detail, for each excluded atomic term, we compute the {\em
+entropy} of the corresponding proposition, {\em conditional} on the
+active assumptions. The {\em conditional entropy} measure
+is higher the more uncertain that proposition is given truth
+assignments to propositions that are the subject of active
+assumptions. Propositions that in the first place were excluded are
+iteratively added to the problem posed to the conditional session in
+increasing order according to that measure.
+
+%% In particular, we compute the {\em conditional entropy} of
+%% each excluded proposition given the active assumptions. Propositions
+%% are candidates for addition in increasing order according to that
+%% measure.
+
+%% Of course, it is expensive to measure size exactly,
+%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
+%% $\propositions^\#$ is the set of propositions in the abstraction that
+%% were not assigned the {\em unknown} value \#. Terms are proposed for
+%% addition because they are probabilistic, and because they were omitted
+%% in the {\em relaxed} visitation characterised by {\em assumptive}
+%% actions in the sequential plan. In each iteration, we add one
+%% candidate term with the largest number of parent terms already in the
+%% candidate tree.

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
component to requirement

@@
 approach, we have yet to find those exploitable structures in our
-domain due to the task planning component.
+domain due to the task planning requirement.

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
new conclusion

--- 2010-11-22 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
@@
 In this paper we address these challenges, developing a continual
-planner that switches between sequential and decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until a validation of an assumption about a runtime
-proposition is requested.
+planner that switches between fast sequential and expensive
+decision-theoretic planning. Given a POMDP model of the environment,
+sequential planning is used to compute an initial deterministic
+sequential plan and complementary runtime evolution of the decision
+process. That plan is executed until testing of an assumption about
+the runtime state is scheduled. At that point a contingent session
+performs some sensing which determines whether or not the sequential
+session should be allowed to continue, or otherwise acts to achieve
+the overall objectives.
 
-The sequential mode of our system always schedules decision-theoretic
-planning before executing actions that are not applicable with
-certainty. This is vastly inefficient if the utility of that plan is
-not dependent on the actions successful execution. This is the case
-where a plan is conformant. Therefore, in the future we must develop a
-cheap (heuristic) procedure to evaluate the utility of sensing at
-steps in sequential plan.
+An obvious theoretical criticism of switching continual planning
+concerns interleaved sequential and contingent sessions failing to
+make any progress towards the objectives. For example, in the worst
+case we can have each sequential session producing an identical plan,
+and each decision-theoretic session rejecting it without further
+sensing. Although this has not been an issue in our work so far, it
+must be dealt with rigorously moving forward. We suggest that a good
+way of mitigate this problem is by using a {\em motivational}
+component that maintains a {\em dynamic} reward model for sensing.
 
-The switching continual planning system we have described serves as
-the underlying planner for CogX.
+%% The sequential mode of our system always schedules decision-theoretic
+%% planning before executing actions that are not applicable with
+%% certainty. This is inefficient if the utility of the plan is not
+%% dependent on that actions successful execution. This situation always
+%% arises for conformant plans.
 
-the effects of a {\em sense} schema are perceptual, whereas the
-effects of an {\em operator} schema are over state propositions.
+%% The switching continual planning system we have described serves as
+%% the underlying planner for CogX.
 
-posted by a motivational component of the underlying robotic
-architecture.
+%% the effects of a {\em sense} schema are perceptual, whereas the
+%% effects of an {\em operator} schema are over state propositions.
 
-contingent sensory plans that are tailored to current
-objectives.
+%% posted by a motivational component of the underlying robotic
+%% architecture.
 
-In this paper we present an approach to continual planing that uses
-two planning systems. The first
+%% contingent sensory plans that are tailored to current
+%% objectives.
 
- to a distinct class of
-challenges. We suppose 
-%%
-The underlying environment is modelled as a POMDP. We use the fast
-classical satisficing system FastDownward to find a deterministic
-sequential plan and complementary runtime evolution of that
-process. This corresponds to a generalisation of replanning in
-probabilistic planning to problems with partial observability.
+%% In this paper we present an approach to continual planning that uses
+%% two planning systems. The first
 
-Interaction between the sequential planner and execution proceeds
-more-or-less analogously to popular replanning approaches
+%%  to a distinct class of
+%% challenges. We suppose 
+%% %%
+%% The underlying environment is modelled as a POMDP. We use the fast
+%% classical satisficing system FastDownward to find a deterministic
+%% sequential plan and complementary runtime evolution of that
+%% process. This corresponds to a generalisation of replanning in
+%% probabilistic planning to problems with partial observability.
 
-Addressing these challenges in a monolithic framework, we present a
-{\em switching} continual planner, that uses the fast sequential
-satisficing procedure FastDownward to perform net-benefit
-%%
-makes reasonable assumptions about the evolution of the runtime state
-given a POMDP model of the environment.
+%% Interaction between the sequential planner and execution proceeds
+%% more-or-less analogously to popular replanning approaches
 
-contingent sensory plans that are tailored to current
-objectives.
+%% Addressing these challenges in a monolithic framework, we present a
+%% {\em switching} continual planner, that uses the fast sequential
+%% satisficing procedure FastDownward to perform net-benefit
+%% %%
+%% makes reasonable assumptions about the evolution of the runtime state
+%% given a POMDP model of the environment.
 
-The decision-theoretic planner is able to tailor sensory processing on
-a robot platform to the current objective, while FastDownward  quickly
+%% contingent sensory plans that are tailored to current
+%% objectives.
 
-In this paper we develop a continual planing system that uses two
-planning systems. The first, is a state-of-the-art domain independent
-planner for deterministic problems. The second is a information-state
-contingency planning the information-state space of
+%% The decision-theoretic planner is able to tailor sensory processing on
+%% a robot platform to the current objective, while FastDownward  quickly
 
-Continual planning is a powerful technique that goes some way to
-addressing those challenges. That approach interleaves planning and
-execution, deliberately postponing planning for contingencies unless
-they eventuate during execution.
+%% In this paper we develop a continual planing system that uses two
+%% planning systems. The first, is a state-of-the-art domain independent
+%% planner for deterministic problems. The second is a information-state
+%% contingency planning the information-state space of
 
-computing a single sequential plan and
-eventuality
+%% Continual planning is a powerful technique that goes some way to
+%% addressing those challenges. That approach interleaves planning and
+%% execution, deliberately postponing planning for contingencies unless
+%% they eventuate during execution.
 
-To these challenges it seems a continual planning approach is best,
+%% computing a single sequential plan and
+%% eventuality
 
-where the runtime state evolves during plan execution
+%% To these challenges it seems a continual planning approach is best,
 
-interleaved planning and execution.
+%% where the runtime state evolves during plan execution
 
-the latter is able to tailor sensory processing on a robot platform,
-in order that it.
+%% interleaved planning and execution.
 
-{\em ad-hoc}
+%% the latter is able to tailor sensory processing on a robot platform,
+%% in order that it.
 
-, reason about degrees of belief and uncertainty about the world,
+%% {\em ad-hoc}
 
-probabilistic sequential decision making in practical sized problems
-is intractable.
+%% , reason about degrees of belief and uncertainty about the world,
 
-quantitative probabilistic models of the perception and action .
+%% probabilistic sequential decision making in practical sized problems
+%% is intractable.
+
+%% quantitative probabilistic models of the perception and action .

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
changed the related text to fix error

@@
 Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed until the applicability of the
-next scheduled action, here \action, is questionable. A contingent
-session then begins, which tailors sensory processing by the agent to
-perceive whether the assumptions made in the trace hold. For each {\em
-active} assumption in the trace, we add a {\em disconfirm} action to
-the POMDP whose execution is rewarding (resp. costly) if the
-assumption is false (resp. true). We also add one {\em confirm} action
-that is rewarding (resp. costly) to execute if $\poss(\action)$ is
-true (resp. false). If execution of the contingent plan applies a
-disconfirm action, then a new sequential session begins at the current
-underlying belief-state. If the confirm action is executed, the
-sequential session is resumed, and \action\ is executed. Otherwise,
-control rests in the continual session.
-
+actions from the trace are executed in sequence until the
+applicability of the next scheduled action is uncertain. Here, that
+action is denoted \switchAction.  A contingent session then begins,
+which tailors sensory processing by the agent to determine whether the
+assumptions made in the trace hold. For each {\em active} assumption
+in the trace, we add a {\em disconfirm} action to the POMDP whose
+execution is rewarding (resp. costly) if the assumption is false
+(resp. true). We also add one {\em confirm} action that is rewarding
+(resp. costly) to execute if $\poss(\action)$ is true
+(resp. false). If execution of the contingent plan applies a
+disconfirm action, then a new sequential session begins at the {\em
+underlying} belief-state. If the confirm action is executed, the
+sequential session is resumed, and
+\switchAction\ is executed. Otherwise, control rests in the continual
+session.
+%%
 Because contingent planning is only practical in relatively small
@@
 the resulting model is of a practicable size.
+
+Finally, whether proceeding in a sequential or contingent session, our
+continual planner maintains a factored representation of successive
+{\em belief-states}, by performing belief revision according to
+Eq~\ref{eq:revision}. Moreover, our internal representation of that
+distribution corresponds more-or-less directly to an $(\pp{:init})$
+declaration in DTPDDL. Our approach uses that underlying belief-state
+as: (1) the source of determinisations for sequential planning, (2)
+the basis for switching, and (3) the source for state abstraction in
+contingent sessions.
 
@@
 must be applied to make a truth assignment to \prop, and then that
-truth assignment could be modified by POMDP operators. Lastly, we
-address the switching semantics of action execution. If
+truth assignment could be modified by POMDP operators.
+
+Addressing now the switching semantics of action execution. If
 $\action\in\actions$ has a precondition that is not true with

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
I have to rush home now to let Katrina into the house. Will continue working on this tonight.
 - Charles

@@
 as: (1) the source of determinisations for sequential planning, (2)
-the basis for switching, and (3) the source for state abstraction in
-contingent sessions.
+the basis for switching, and (3) a mechanism to guide construction of
+an abstract process for contingent sessions.
 
@@
 
-\Omit{
-  Therefore, although our current implementation does not
-  explicitly support it, our approach generalises to POMDPs with
-  stochastic action models.
-}
+%% \Omit{
+%%   Therefore, although our current implementation does not
+%%   explicitly support it, our approach generalises to POMDPs with
+%%   stochastic action models.
+%% }
 

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
new structure, and altered for n omore FSC

@@
 
-\section{Moritz's Initial Outline}
+%% \section{Moritz's Initial Outline}
 
-\subsection{Switching Planner Outline}
+%% \subsection{Switching Planner Outline}
 
-Outline the high level approach here:
-\begin{itemize}
-\item Problem input (factored initial state, DTPDDL observations)
-\item Determinisation
-\item Problem generation for DT
-\item Belief revision
-\end{itemize}
+%% Outline the high level approach here:
+%% \begin{itemize}
+%% \item Problem input (factored initial state, DTPDDL observations)
+%% \item Determinisation
+%% \item Problem generation for DT
+%% \item Belief revision
+%% \end{itemize}
 
-\subsection{Determinisation}
+%% \subsection{Determinisation}
 
-\begin{itemize}
-\item Generation of deterministic sensing models
-\item Determinisation of the initial state
-\item Compiling reward model into (soft-)goal model
-\item Cost function in FD (maybe this could mentioned in the outline?)
-\end{itemize}
+%% \begin{itemize}
+%% \item Generation of deterministic sensing models
+%% \item Determinisation of the initial state
+%% \item Compiling reward model into (soft-)goal model
+%% \item Cost function in FD (maybe this could mentioned in the outline?)
+%% \end{itemize}
 
-\subsection{Subtask Generation}
-\begin{itemize}
-\item Goal generation from CP trace
-\item Disconfirm actions
-\item State pruning according to CP trace
-\end{itemize}
+%% \subsection{Subtask Generation}
+%% \begin{itemize}
+%% \item Goal generation from CP trace
+%% \item Disconfirm actions
+%% \item State pruning according to CP trace
+%% \end{itemize}
 
-\subsection{Belief revision}
-Not sure if this needs its own section
-\begin{itemize}
-\item Convert inital state into bayesian network
-\item Belief updates with observation nodes
-\item Create new network for next planning runs
-\end{itemize}
+%% \subsection{Belief revision}
+%% Not sure if this needs its own section
+%% \begin{itemize}
+%% \item Convert inital state into bayesian network
+%% \item Belief updates with observation nodes
+%% \item Create new network for next planning runs
+%% \end{itemize}
 

### 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
new structure, and altered for n omore FSC

@@
 planning, where the objective is to find a sequence of actions that
-achieves a goal ---I.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps, the value of a controller corresponds to
-the expected cumulative reward:
+achieves a goal ---i.e., reaches a state that satisfies a given
+Boolean condition--- with probability $1$.  Generally, whatever the
+structure of a plan, its value corresponds to the expected cumulative
+reward:
+
+%% More generally, a useful
+%% formalism for representing solutions to POMDPs is the finite-state
+%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
+%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
+%% probability that the controller prescribes \action\ when in state
+%% \node, and $\transition_\node(\action, \observation, \node') =
+%% P(\node'|\node, \action, \observation)$ gives the probability the
+%% controller transitions to state $\node'$, supposing we execute
+%% \action\ in state \node\ receiving observation \observation. When
+%% acting for finite $N$ steps from controller state $n$, the value of a
+%% controller corresponds to the expected cumulative reward:
 
 \begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, \bstate = \bstate_0 \bigg{]}
+V_{\pp{PLAN}}(\bstate, n) = \Expect \bigg{[} 
+\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
 \end{equation}
 
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{FSC}_t$ is
+\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
 the action prescribed at step $t$, and
@@
 
-\noindent Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plans.
+%% \noindent Finally, it is
+%% useful to note that there is a corresponding deterministic FSCs for
+%% both sequential (e.g., conformant) and contingent plan formats.

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
new intro to swiching section

@@
 We describe our continual planning approach that switches between
-sequential and contingent sessions. This proceeds by interleaved
+sequential and contingent sessions. It proceeds by interleaved
 planning and execution in a deterministic-action POMDP described in
 DTPDDL. During a sequential session, planning is performed by a
-``classical'' system,\footnote{That is, a system designed to solve
+``classical'' system,\footnote{That is, a planner designed to solve
 fully observable deterministic tasks.}  and execution proceeds
 according to the {\em trace} computed by that system. Taking the form
-of a classical plan, the trace specifies a sequence of POMDP
-actions that achieve the agent's objectives in a deterministic
-approximation, i.e., {\em determinisation}, of the problem at
-hand. More precisely, the trace is an interleaved sequence of POMDP
-actions and {\em assumptive} actions. The latter correspond to
-assumptions the planner makes about the truth value of {\em runtime
-state variables} -- e.g. that a box of cornflakes is located in the
-kitchen at the third plan step. Another kind of assumption, called an
-{\em applicability} assumption, is made when the trace includes an
-action $\action$ that is not applicable with probability $1$ at the
-belief-state
+of a classical plan, the trace specifies a sequence of POMDP actions
+that achieve the agent's objectives in a deterministic approximation,
+i.e., {\em determinisation}, of the problem at hand. More precisely,
+the trace is an interleaved sequence of POMDP actions and {\em
+assumptive} actions. The latter correspond to assumptions the planner
+makes about the truth value of propositions -- e.g. that a box of
+cornflakes is located in the kitchen at the third plan step. Another
+kind of assumption, called an {\em applicability} assumption, is made
+when the trace includes an action $\action$ that is not applicable
+with probability $1$ at the belief-state
 \bstate\ that the system is projected to be in when \action\ is
@@
 actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. Here, that
-action is denoted \switchAction.  A contingent session then begins,
-which tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. For each {\em active} assumption
-in the trace, we add a {\em disconfirm} action to the POMDP whose
-execution is rewarding (resp. costly) if the assumption is false
-(resp. true). We also add one {\em confirm} action that is rewarding
-(resp. costly) to execute if $\poss(\action)$ is true
-(resp. false). If execution of the contingent plan applies a
-disconfirm action, then a new sequential session begins at the {\em
-underlying} belief-state. If the confirm action is executed, the
-sequential session is resumed, and
+applicability of the next scheduled action is uncertain. We denote
+that action \switchAction.  A contingent session then begins, which
+tailors sensory processing by the agent to determine whether the
+assumptions made in the trace hold. For assumptions in the trace, we
+add a {\em disconfirm} action to the POMDP whose execution is
+rewarding (resp. costly) if the assumption is false (resp. true). We
+also add one {\em confirm} action that is rewarding (resp. costly) to
+execute if $\poss(\switchAction)$ is true (resp. false). If execution
+of the contingent plan applies a disconfirm action, then a new
+sequential session begins at the {\em underlying} belief-state. If the
+confirm action is executed, the sequential session is resumed, and
 \switchAction\ is executed. Otherwise, control rests in the continual
@@
 Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions proceed in an abstract decision process
-determined by the current trace. That abstract process is constructed
-by first omitting all variables from the model that are not featured
-in the trace, and by then iteratively refining that abstraction while
-the resulting model is of a practicable size.
+POMDPs, contingent sessions plan in an abstract decision process
+determined by the current trace. This abstraction is constructed by
+first omitting all propositions that do not featured in the trace, and
+by then iteratively refining the model while the result is of a
+practicable size.
 
@@
 continual planner maintains a factored representation of successive
-{\em belief-states}, by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of that
-distribution corresponds more-or-less directly to an $(\pp{:init})$
-declaration in DTPDDL. Our approach uses that underlying belief-state
-as: (1) the source of determinisations for sequential planning, (2)
-the basis for switching, and (3) a mechanism to guide construction of
-an abstract process for contingent sessions.
+belief-states, by performing belief revision according to
+Eq~\ref{eq:revision}. Moreover, our internal representation the
+underlying belief-state corresponds more-or-less directly to an
+$(\pp{:init})$ declaration in DTPDDL. Our approach uses that
+distribution: (1) as the source of candidate determinisations for
+sequential planning, (2) to determine when to switch, and (3) as a
+mechanism to guide construction of an abstract process for contingent
+sessions.
 

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
New notations, and corrections.

@@
 tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$,
-(2) probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{DORA}$ might
-be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
-expresses the conjunctive root of the tree -- i.e., the root node
-$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
-$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
-subterm.}
+state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
+probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
+(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
+probabilistic and atomic terms. The root term is always conjunctive,
+and the leaves are atomic. For example, the starting distribution for
+$\pp{DORA}$ might be described thusly:\footnote{In PDDL,
+$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
+i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
+$\prop$, rather than $(\pp{and}~\prop)$, it a conjunctive term
+contains a single atomic subterm.}
 
@@
 union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
+of $\prob_i$ entries on the visited subterms of probabilistic
 elements. Making these ideas concrete, our example yields the

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
New notations, and corrections.

@@
 latter encapsulates assumptions about: (1) the true underlying state,
-(2) how the plan's execution will progress, and (3) the possibility of
-the agent eventually holding strong beliefs about the truth values of
-specific state propositions. Here we describe the deterministic
-planning problem, derived from the DTPDDL model, that admits plans
-which correspond to traces.
+(2) how execution will progress, and (3) the possibility of the agent
+eventually holding strong beliefs about the truth values of specific
+state propositions. Here we describe the deterministic planning
+problem, derived from the DTPDDL model, that admits plans which
+correspond to traces.
 
@@
 
-\noindent In detail, a valid assumption in our setting corresponds to a {\em
+\noindent An assumption corresponds to a {\em
 relaxed} visitation of the root term of $(\pp{:init})$. In this
@@
 add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $p_i (T_i)$, of each probabilistic term from
+each element, $\prob_i (T_i)$, of each probabilistic term from
 $(\pp{:init})$. The physics of these actions is as
@@
 $j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
+either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct,
 or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
 %%
-Executing $\assumptiveS{i}$ in $s$ effects a transition to a successor
-state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with probability
-$1 - p_i$. Here, $s^{T_i}$ is the union of $s$ with atomic terms from
-$T_i$ and conflicting propositions from $s$ deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is an added sink,
+Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
+successor state $\state^{T_i}$ with probability $\prob_i$, and
+$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
+the union of $\state$ with atomic terms from $T_i$ and conflicting
+propositions from $s$ deleted. For example, if the action adds
+$(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must delete
+$(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added sink,
 where $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$
@@
 or negative effect of $\action$, then $\action$ is not applicable
-in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
+in \state\ if $\prop\#\in\state$. In other words, an assumptive action
 must be applied to make a truth assignment to \prop, and then that
-truth assignment could be modified by POMDP operators.
+truth assignment could be modified by the DTPDDL domain operators.
+%%
+Lastly, With regards to switching, we have that if $\action$ has a
+precondition that is not true with probability $1$ in the underlying
+belief $\bstate$, then immediately before it is executed blindly, the
+system switches to a contingent session, discussed in the following
+section.
 
-Addressing now the switching semantics of action execution. If
-$\action\in\actions$ has a precondition that is not true with
-probability $1$ in the underlying belief $\bstate$, then before it is
-executed blindly, the system switches to a contingent session,
-discussed in the following section. In detail, we halt the sequential
-session at \action\ if $\exists
-\state$ s.t. $\bstate(\state)>0$, 
-$\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-semantics of action execution there must be a $\state'$ s.t.
-$\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-not have been scheduled at \bstate.
+%% Addressing now the switching semantics of action execution. I
+%% %% In detail, we halt the sequential
+%% session at \action\ if $\exists
+%% \state$ s.t. $\bstate(\state)>0$, 
+%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
+%% semantics of action execution there must be a $\state'$ s.t.
+%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
+%% not have been scheduled at \bstate.
 
-then $\reward(\state,\action)=10$, and
-otherwise if $\prop\not\in\state'$ then
-$\reward(\state',\action)=-10$.
+%% then $\reward(\state,\action)=10$, and
+%% otherwise if $\prop\not\in\state'$ then
+%% $\reward(\state',\action)=-10$.
 
@@
 
-\noindent which is equal to the maximal additive contribution a trace of the plan can
-make to Eq~\ref{eq:expectedvalue}, taking an FSC that corresponds to
-the sequential plan.  In goal directed problems, where the only
+\noindent which is equal to the maximal 
+additive contribution a trace of the plan can make to
+Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
 non-zero reward is received at the first transition to a goal state,
@@
 in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration from the
-original problem description.
+the current trace, and the $(\pp{:init})$ declaration that models the
+underlying belief-state.
 
@@
 %%
-In order to reduce the number of possible rejections we do not
-consider assumptions that are not {\em active} with respect to the
-action, $\action_i$, whose scheduled execution switch the system into
-the contingent session. A assumption $\assumptiveS{i}$ is active if
-$\action_i$ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix to $\action_i$. For
-example, if $\pp{DORA}$'s trace is:
+In order to reduce the number of possible rejections we omit
+assumptions that are not {\em active} with respect to the action,
+\switchAction, whose scheduled execution switched the system into
+the contingent session. An assumption $\assumptiveS{i}$ is active if
+\switchAction\ is not applicable in the sequential model unless
+$\assumptiveS{i}$ is scheduled in the trace prefix
+to \switchAction. For example, if $\pp{DORA}$'s trace is:
 
@@
 
-\noindent Taking the switching action $\action_i$ to be
+\noindent Taking the switching action \switchAction\ to be
 $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
@@
 
-Again taking $\action_i$ to be the switching action, we include
-another propositionally trivial action $\actions.\poss(\action_i)$
-with the reward property:
+We also include another propositionally trivial action
+$\actions.\poss(\switchAction)$ with the reward property:
 
 \[
-\reward(\state, \actions.\poss(\action_i)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\action_i)) & \pp{if}~\;\; \poss(\action_i) \subseteq \state \\
--\$(\poss(\action_i)) & \pp{otherwise} \\
+\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
+\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
+-\$(\poss(\switchAction)) & \pp{otherwise} \\
 \end{array}
@@
 action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\action_i)$, returns control to a sequential
+$\actions.\poss(\switchAction)$, returns control to a sequential
 session. If a dual is executed, then the sequential planner must
 replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\action_i)$ is executed,
-then the current sequential plan is executed until further sensing is
-scheduled, or to completion. In our project the rewards, i.e.,
-function $\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture.
+belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
+executed, then the current sequential plan is executed until further
+sensing is scheduled, or to completion. Lastly, we note that in our
+project the rewards, i.e., function $\$:2^\propositions\to\Re$, for
+the POMDP posed to the contingent session are sourced from a
+motivational component of our robotic architecture. We make no
+assumptions about how they are derived.
 
@@
 sessions apply the same applicability conditions ---i.e., that each
-effect and condition $\prop$ is specified $\prop\in\state$
-(resp. $\prop\#\in\state$)--- to actions from the DTPDDL problem
-description.
+effect and condition $\prop$ is specified, $\prop\in\state$--- to
+actions from the DTPDDL problem description.
 

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
Dlib dlib

% coming soon

In our implementation uses the
~\citeauthor{king:2009}~(\citeyear{king:2009}) \system{dlib-ml} for
successive estimation of the underlying belief-state.

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
STRIPS section dealt with

@@
 We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. The underlying process
-dynamics are modelled in terms of a finite set of stochastic actions
-$\stochActions$, deterministic outcomes $\detActions$, and
+planning problem, with costs and rewards. We model a process
+state $\state$ as the set of propositions that are true of the
+state. Notationally, we have $\state \subseteq
+\props$. The underlying process
+dynamics are modelled in terms of a finite set of {\em probabilistic
+STRIPS operators}~\cite{boutilier:1996} $\stochActions$ over
 state-characterising propositions $\propositions$.
+%%
+Notationally, we say an action $\stochAction \in \stochActions$ is
+applicable if its precondition $\poss(\stochAction)$, a set of
+propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
+\state$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
+nature chooses a deterministic STRIPS effect $\detAct_i$, and for
+all \stochAction\ we require
+$\sum_{\detAction_i \in \detActions(\stochAction)}
+\mu_{\stochAction}(\detAction_i) = 1$.
 
-We model a process state $\state$ as the set of propositions that are
-true of the state. Notationally, we have $\state \subseteq
-\props$. State change is induced by application of actions. A
-stochastic action $\stochAction \in \stochActions$ is applicable if
-its precondition $\poss(\stochAction)$, a set of propositions, are
-satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We write $\stochActions(\state)$ for set of actions
-applicable at state $\state$.  If action $\stochAction \in
-\stochActions(\state)$ is applied at \state, nature chooses one
-element amongst a small set of deterministic outcomes
-$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-effect on the state given in terms of two lists of propositions. The
-add-list $\add(\detAction)$ and delete-list
-$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-$\detAction$, then it cannot be in the delete-list and vice versa.}
-If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-..,\prop_m]$ is chosen by nature when \stochAction\ is
-applied at state $\state$, then the resultant state is $ ( \state \cup
-\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-propositions from $\add(\detAction)$ are added to $\state$, and those
-from $\delete(\detAction)$ are removed from $\state$.
+%% We model a process state $\state$ as the set of propositions that are
+%% true of the state. Notationally, we have $\state \subseteq
+%% \props$. State change is induced by application of actions. A
+%% stochastic action $\stochAction \in \stochActions$ is applicable if
+%% its precondition $\poss(\stochAction)$, a set of propositions, are
+%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
+%% applicable at state $\state$.  If action $\stochAction \in
+%% \stochActions(\state)$ is applied at \state, nature chooses one
+%% element amongst a small set of deterministic outcomes
+%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
+%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
+%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
+%% $\sum_{\detAction_i \in \detActions(\stochAction)}
+%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
+%% effect on the state given in terms of two lists of propositions. The
+%% add-list $\add(\detAction)$ and delete-list
+%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
+%% $\detAction$, then it cannot be in the delete-list and vice versa.}
+%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
+%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
+%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
+%% applied at state $\state$, then the resultant state is $ ( \state \cup
+%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
+%% propositions from $\add(\detAction)$ are added to $\state$, and those
+%% from $\delete(\detAction)$ are removed from $\state$.
+
+%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
 
 We are concerned with problems that feature partial
-observability. These have a perceptual model given in terms of a
-finite set of stochastic {\em senses} $\stochSenses$, deterministic
-sensing outcomes $\detSenses$, and perceptual propositions
-$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
-a set of percepts $\observ \subseteq \percepts$, and we denote
-\observations\ the set of reachable observations. The underlying state
-of the process cannot be observed directly, rather, senses
-$\stochSense \in \stochSenses$ effect an observation $\observ \in
+observability. Although we could invoke {\em extended probabilistic
+STRIPS operators}~\cite{rintanen:01} to model actions and observations
+propositionally, we find it convenient for our presentation to clearly
+separate sensing and action. Therefore, we suppose a POMDP has a
+perceptual model given in terms of a finite set of stochastic {\em
+senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
+and perceptual propositions $\percepts$, called {\em percepts}. In
+detail, we take an observation $\observ$ to be a set of percepts
+$\observ \subseteq \percepts$, and denote \observations\ the set of
+reachable observations. The underlying state of the process cannot be
+observed directly, rather, senses $\stochSense \in \stochSenses$
+effect an observation $\observ \in
 \observations$ that informs what should be believed about the state a
@@
 1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active senses.
+perceptual propositions from chosen elements of active
+senses.
 
@@
 planning, where the objective is to find a sequence of actions that
-achieves a goal ---i.e., reaches a state that satisfies some given
-Boolean condition--- with probability $1$.  More generally, a useful
-formalism for representing solutions to POMDPs is the finite-state
-controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-\transition \rangle$ where: $\node \in \nodes$ is a set of controller
-states, $\selection_\node(\action) = P(\action | \node)$ gives the
-probability that the controller prescribes \action\ when in state
-\node, and $\transition_\node(\action, \observation, \node') =
-P(\node'|\node, \action, \observation)$ gives the probability the
-controller transitions to state $\node'$, supposing we execute
-\action\ in state \node\ receiving observation \observation. When
-acting for finite $N$ steps from controller state $n$, the value of a
-controller corresponds to the expected cumulative reward:
+achieves a goal ---i.e., reaches a state that satisfies a given
+Boolean condition--- with probability $1$.  Generally, whatever the
+structure of a plan, its value corresponds to the expected cumulative
+reward:
+
+%% More generally, a useful
+%% formalism for representing solutions to POMDPs is the finite-state
+%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
+%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
+%% probability that the controller prescribes \action\ when in state
+%% \node, and $\transition_\node(\action, \observation, \node') =
+%% P(\node'|\node, \action, \observation)$ gives the probability the
+%% controller transitions to state $\node'$, supposing we execute
+%% \action\ in state \node\ receiving observation \observation. When
+%% acting for finite $N$ steps from controller state $n$, the value of a
+%% controller corresponds to the expected cumulative reward:
 
 \begin{equation}\label{eq:expectedvalue}
-V_{\pp{FSC}}(\bstate, n) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{FSC}_t) \mid \pp{FSC}, n_0 = n, \bstate_0 = \bstate \bigg{]}
+V_{\pp{PLAN}}(\bstate, n) = \Expect \bigg{[} 
+\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
 \end{equation}
 
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{FSC}_t$ is
+\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
 the action prescribed at step $t$, and
@@
 
-\noindent Finally, it is
-useful to note that there is a corresponding deterministic FSCs for
-both sequential (e.g., conformant) and contingent plan formats.
+%% \noindent Finally, it is
+%% useful to note that there is a corresponding deterministic FSCs for
+%% both sequential (e.g., conformant) and contingent plan formats.

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
type

@@
 Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
+terms of propositions and percepts, in order to discuss belief
+revision and utility it is convenient to consider the underlying
+decision process in a flat format. This underlying decision process is
+given by the tuple
+$\langle \states, \bstate_0, \actions, \transProb, \reward,
 \observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
@@
 \begin{equation}\label{eq:expectedvalue}
-V_{\pp{PLAN}}(\bstate, n) = \Expect \bigg{[} 
+V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
 \sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
clean up some text in prop. section

@@
 
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss belief
-revision and utility it is convenient to consider the underlying
-decision process in a flat format. This underlying decision process is
-given by the tuple
+Until now we have discussed the POMDP in terms of propositions and
+percepts, in order to address belief revision and utility it is
+convenient to consider the underlying decision process in a flat
+format. This is given by the tuple
 $\langle \states, \bstate_0, \actions, \transProb, \reward,
@@
 
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
+Successive state estimation is available by application of Bayes'
+rule.  Taking the current belief $\bstate$ as the {\em prior}, and
+supposing action $\action$ is executed with perceptive outcome
+$\observ$, then the probability that we are in $\state$ in the
+successive belief-state $\bstate'$ is given by:
 

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
typo in prop. stuff

@@
 Until now we have discussed the POMDP in terms of propositions and
-percepts, in order to address belief revision and utility it is
+percepts. In order to address belief revision and utility it is
 convenient to consider the underlying decision process in a flat

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
new conclusion, and typos in prop. section

@@
 
-An obvious theoretical criticism of switching continual planning
-concerns interleaved sequential and contingent sessions failing to
-make any progress towards the objectives. For example, in the worst
-case we can have each sequential session producing an identical plan,
-and each decision-theoretic session rejecting it without further
-sensing. Although this has not been an issue in our work so far, it
-must be dealt with rigorously moving forward. We suggest that a good
-way of mitigate this problem is by using a {\em motivational}
-component that maintains a {\em dynamic} reward model for sensing.
+The most pressing item for future research, is to develop a scheme
+whereby the serial planner can relax {\em executability} assumptions,
+so that conformant (or semi-conformant) plans can be executed without
+interruption by a contingent session. A more general criticism of
+switching continual planning concerns interleaved sequential and
+contingent sessions failing to make any progress towards the
+objectives. For example, in the worst case we can have each sequential
+session producing an identical plan, and each decision-theoretic
+session rejecting it without further sensing. Although this has not
+been an issue in our work so far, it must be dealt with rigorously
+moving forward. We suggest that a good way of mitigate this problem is
+by developing a {\em motivational} component that maintains a {\em
+dynamic} reward model whose limiting behaviour prevents that switching
+deadlock.
 

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
new conclusion, and typos in prop. section

@@
 $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
+\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
 belief-state, \states\ is the finite set of reachable propositional

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
typo that distroyed the meanin of a sentence in the switching secion.

@@
 tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. For assumptions in the trace, we
-add a {\em disconfirm} action to the POMDP whose execution is
-rewarding (resp. costly) if the assumption is false (resp. true). We
-also add one {\em confirm} action that is rewarding (resp. costly) to
-execute if $\poss(\switchAction)$ is true (resp. false). If execution
-of the contingent plan applies a disconfirm action, then a new
-sequential session begins at the {\em underlying} belief-state. If the
-confirm action is executed, the sequential session is resumed, and
+assumptions made in the trace hold. For each assumption we add a {\em
+disconfirm} action to the POMDP whose execution is rewarding
+(resp. costly) if the assumption is false (resp. true). We also add
+one {\em confirm} action that is rewarding (resp. costly) to execute
+if $\poss(\switchAction)$ is true (resp. false). If execution of the
+contingent plan applies a disconfirm action, then a new sequential
+session begins at the {\em underlying} belief-state. If the confirm
+action is executed, the sequential session is resumed, and
 \switchAction\ is executed. Otherwise, control rests in the continual
@@
 belief-states, by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation the
+Eq~\ref{eq:revision}. Moreover, our internal representation of the
 underlying belief-state corresponds more-or-less directly to an
@@
 is an {\em abstract} state that captures the unique assumption which
-has probability $1.0$. In $\pp{DORA}$, that is:
+has probability $1$. In $\pp{DORA}$, that is:
 
@@
 $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added sink,
-where $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$
+where $\forall\action$ $\reward(\state^\bot, \action) = 0$
 and $\transProb(\state^\bot, \action, \state^\bot) = 1$.
@@
 %%
-Lastly, With regards to switching, we have that if $\action$ has a
+Lastly, with regards to switching, we have that if $\action$ has a
 precondition that is not true with probability $1$ in the underlying
@@
 non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour
-of \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
+the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
+by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
 gradient-based reinforcement learning systems, such as William's

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
spellers

@@
 
-In our implementation uses the
+Our implementation uses the
 ~\citeauthor{king:2009}~(\citeyear{king:2009}) \system{dlib-ml} for

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
changed the related text to fix error
spellers

@@
-There have been a number of developments recently towards planning
-under uncertainty using systems that were intended for sequential
-planning in deterministic problems.  Notably for example,
+There have been a number of developments recently moving towards
+planning under uncertainty using systems that were intended for
+sequential planning in deterministic problems.  Notably for example,
 \system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
@@
 robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}. However, as we said in the introduction,
-all these approaches struggle in partially observable domains as they
-rely on being able to determine the state at all times.
+plan~\cite{majercik:2006}.
+
+%%No! They simply haven't been evaluated in PO settings. They may, or
+%%may not struggle. They have sampling of traces, and that would
+%%include observations, and therefore evolutions of beliefs. SSAT was
+%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
+%%suited to POMDPs.
+%%
+%% However, as we said in the introduction,
+%% all these approaches struggle in partially observable domains as they
+%% rely on being able to determine the state at all times.
 
@@
 and has been applied in completely observable domains, particularly
-those featuring multple communicating agents. The use of knowledge
+those featuring multiple communicating agents. The use of knowledge
 operators in domains allows plans that act to gain knowledge, but the

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
spellers

@@
 Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called Decision-Theoretic
+we develop an extension of that language, called Decision-Theoretic
 (DT)PDDL. This can express probabilistic models of the sensing
@@
 and the leaves are atomic. For example, the starting distribution for
-$\pp{DORA}$ might be described thusly:\footnote{In PDDL,
+$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
 $(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
 i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
-$\prop$, rather than $(\pp{and}~\prop)$, it a conjunctive term
-contains a single atomic subterm.}
+$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
+that contain a single atomic subterm.}
 
@@
 according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
+along with the probability that it occurs. The former corresponds to the
 union of all visited atoms, and the latter corresponds to the product

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
spellers

--- 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
@@
 replan when actions have unplanned-for
-outcomes \cite{yoon:etal:2007}. However, these approaches aren't so
+outcomes \cite{yoon:etal:2007}. However, these approaches are not as
 useful in partially observable domains as replanning is triggered when
 the system ends up in a state other than the desired one, and this
-state can't in general be determined.
+state cannot in general be determined.
 
@@
 planner, but it performs better than the sequential planner alone and
-is fast enough to be used onboard the robot for decision-making. As an
+is fast enough to be used on the robot for decision-making. As an
 illustration, consider a robot searching for a box that it knows is
@@
 The switching planner approach gives us some important advantages in
-our robotic domain. Firstly, the main source of uncertainty in the
+our robotic domain. First, the main source of uncertainty in the
 domain is unreliable observations, rather than stochasticity in the
@@
 propositionally factored representation of states and observations,
-and therelationship between this and a flat POMDP representation. Next
+and the relationship between this and a flat POMDP representation. Next
 we describe our DTPDDL language using a running example of the robot

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
spellers

@@
 applicability of the next scheduled action is uncertain. We denote
-that action \switchAction.  A contingent session then begins, which
+that action \switchAction.  A contingent session then begins, that
 tailors sensory processing by the agent to determine whether the
@@
 
-\noindent Where $\$(T_i)$ is a positive numeric quantity which
+\noindent where $\$(T_i)$ is a positive numeric quantity which
 captures the reward the agent receives for correctly rejecting a
@@
 motivational component of our robotic architecture. We make no
-assumptions about how they are derived.
+assumptions here about how they are derived.
 

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
draft of the evaluation section.

@@
 % coming soon
+
+\subsection{Implementation}
+
+The implementation of the continual planning and switching parts is
+based on MAPSIM \cite{brenner:nebel:jaamas09} and is able to use several planners as
+the underlying planning system. We use a modified version of Fast
+Downward \cite{fast-downward}, which we extended with the support for
+actions with probabilities.
 
@@
 successive estimation of the underlying belief-state.
+
+The baseline approach is using the same continual planning system as
+the switching planner, but instead of creating an observation problem
+for the decision theoretic planner it will just execute one sensing
+action -- assuming that this action will confirm its assumption.
+
+\subsection{Experiments}
+For evaluation, we used a robot exploration domain similar to the one
+used on a physical robotic system. The goal of the robot is to find
+objects in a human environment and reporting back their position to a
+human.
+
+The basic building blocks of the domain consist of {\tt rooms}, {\tt
+places} and {\tt objects}. Places are grouped in rooms and objects (as
+well as the robot) are always in a place. The robot can move around
+the rooms via connections between places given by the {\tt connected}
+predicate. Each room has a (possibly unknown) {\em category/}
+(e.g. kitchen, office, living room) and depending on this category,
+objects are placed inside the rooms. An instance of this domain is
+shown in figure\ref{fig}.
+
+The robot can find out if an object is at a certain place by executing
+the {\tt look-for-object} action, which may result in a perception if
+the object is in fact there (though some objects are harder to detect
+than others -- so absence of a percept is no proof of the object's
+absence). Additionally, if the robot is in the presence of a human, it
+may simply ask what type of room they are currently in -- but
+conducting a dialogue is a bit more costly than simply running the
+vision algorithm (cost of 8 vs costs of 3).
+
+We conducted experiments on scenarios of several sizes with several
+kinds of goals (though the goal is always reporting the position of a
+certain kind of object). In order to determine the impact of sensor
+reliability on both approaches, we also ran several tests in identical
+problems with different probabilities for sensing the goal object.
+These problems are denoted by the -easy, -medium and -hard suffixes;
+in the easy case the probability of percieving the object is 0.9, in
+medium it is 0.65 and 0.4 in hard.
+
+As the initial state of the problems is stochastic, we ran 50
+simulations for each configuration. In each run the simulator
+generated the true world state by sampling from the same initial state
+distribution the planner was using. So on average the world did
+conform to the planner's expectations. Not all problems generated in
+this manner have a solution. We did not reject these problems, as we
+think that detecting the non-existance of a solution is as valuable in
+exploration problems than finding one.
+
+Fast Downward was run with the cyclic causal graph heuristic using A*
+search or weighted A* with a weight of 5, depending on the difficulty
+of the problem. We then performed multiple tests with different limits
+for the belief space size of the decision theoretic subtask. Higher
+limits should cause longer planning times but be beneficial to plan
+quality as more contingencies can be taken in to account by the POMDP
+planner.
+
+The graphs show the average costs of the agent's actions, the number
+of solved tasks as a percentage of the solvable tasks, as well as the
+planing time for the continual and decision theoretic planners.
+
+For objects that can be easily detected there is little gain in using
+a decision theoretic planner, as the greedy sensing appoach by the
+baseline continual planner is obviously sufficient here. With
+decreasing sensor reliability the more sophisticated observation
+planning pays off: while the resulting plans are still longer on
+average, the impact on the number of solved tasks was much smaller
+than for the baseline system.
+
+We think that a part of the improvement is due to the segmentation of
+the plan into several subtask, essentially performing hierarchical
+planning. Especially when the continual planner performs badly this is
+a huge gain.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "moritz_2011"
+%%% End:

### 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Incorporated some of Charles' suggestions.

@@
 replan when actions have unplanned-for
-outcomes \cite{yoon:etal:2007}. However, these approaches aren't so
-useful in partially observable domains as replanning is triggered when
-the system ends up in a state other than the desired one, and this
-state can't in general be determined.
+outcomes \cite{yoon:etal:2007}. However, these approaches rely
+crucially on being able to determine properties of the current state
+in order to decide to replan. For partially observable domains, only
+probabilistic information about these properties will in general be
+available.
 
@@
 unreliable sensing actions. However, the state representation is
-propositional, and even for relatively simple problems is too large
-for a POMDP solver to be applied.
+propositional, and even for relatively simple problems the underlying
+model is too large for a POMDP solver to be applied.
 
@@
 reasoning about observations is required to find good plans. We refer
-to the complete system as a {\em switching planner}. The approach is
-not optimal, particularly as it is based on a satisficing sequential
-planner, but it performs better than the sequential planner alone and
-is fast enough to be used onboard the robot for decision-making. As an
-illustration, consider a robot searching for a box that it knows is
-often found in offices. The sequential planner might produce a plan
+to the complete system as a {\em switching continual planner}. The
+approach is not optimal, particularly as it is based on a satisficing
+sequential planner, but it performs better than the sequential planner
+alone and is fast enough to be used in real-time on a mobile robot for
+decision-making.
+
+As an illustration, consider a robot searching for a box that it knows
+is often found in offices. The sequential planner might produce a plan
 that assumes that {\sc Room a} is an office, that the box is in {\sc
-Room a}, and that the box is in location 1 in {\sc Room a}, and then
-goes to {\sc Room a}, goes to location 1 and searches for the
-box. When the plan gets executed and the search begins, the
-decision-theoretic planner is called to conduct the search, and at
-this point may retract the last assumption and build a plan to search
-the whole of {\sc Room a} for the box. If this plan fails, it will do
-so by disproving one of the assumptions (for example, that {\sc Room
-a} is an office), which will then lead to replanning by the sequential
-planner based on the newly discovered knowledge.
+Room a}, and that the box is in location 1 in {\sc Room a}. Given
+these assumptions, the plan might be to go to {\sc Room a}, go to
+location 1 and search for the box. When the plan gets executed and the
+search begins the decision-theoretic planner is called since searching
+involves making observations. The decision-theoretic planner might
+retract the last assumption but leave the others, and build a plan to
+search the whole of {\sc Room a} for the box. If this plan fails to
+find the box, it will disprove the assumption that the box is in {\sc
+Room a}, which will then lead to replanning by the sequential planner
+based on the newly discovered knowledge.
 
@@
 planning problems that correspond to POMDPs. We have created a system
-that automatically constructs from this a determinised planning domain
+that automatically constructs from this a deterministic planning domain
 for the sequential planner that includes actions that allow
@@
 
-The sequential planner we use is Continual
-Planning \cite{brenner:nebel:jaamas09} (CP). CP is based
-on \fastdownward \cite{fast-downward}. CP already has the capability
-to replan, and also allows agent knowledge to be represented
-explicitly, so we can write actions that gain the agent knowledge. For
-the decision-theoretic planner we use an information-state
-\laostar\ procedure that uses the solution to the equivalent MDP as a
-heuristic. For reasons of space, we will not describe this system in
-detail here.
+The sequential planner we use is based
+on \fastdownward \cite{fast-downward}. We add the capability to
+replan, and also allow agent knowledge to be represented explicitly,
+so we can write actions that gain the agent knowledge. For the
+decision-theoretic planner we use forward search in the information
+state.
 
 The switching planner approach gives us some important advantages in
-our robotic domain. Firstly, the main source of uncertainty in the
+our robotic domain. First, the main source of uncertainty in the
 domain is unreliable observations, rather than stochasticity in the

### 2010-11-28 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
Added plots with results, and a bit more writing on evaluation.

@@
 
-\subsection{Implementation}
-
 The implementation of the continual planning and switching parts is
-based on MAPSIM \cite{brenner:nebel:jaamas09} and is able to use several planners as
-the underlying planning system. We use a modified version of Fast
-Downward \cite{fast-downward}, which we extended with the support for
-actions with probabilities.
-
-Our implementation uses the
-~\citeauthor{king:2009}~(\citeyear{king:2009}) \system{dlib-ml} for
-successive estimation of the underlying belief-state.
+based on MAPSIM \cite{brenner:nebel:jaamas09} and is able to use
+several planners as the underlying planning system. We use a modified
+version of Fast Downward \cite{fast-downward}, which we extended with
+the support for actions with success probabilities. Finally, we use
+\system{dlib-ml} \cite{king:2009} for successive estimation of the
+underlying belief-state.
 
@@
 
-\subsection{Experiments}
-For evaluation, we used a robot exploration domain similar to the one
-used on a physical robotic system. The goal of the robot is to find
-objects in a human environment and reporting back their position to a
-human.
+To test our approch we use a robot exploration domain similar to the
+one used on a physical robotic system. It involves a robot exploring
+an office or living environment and trying to return objects to their
+owner.
 
@@
 reliability on both approaches, we also ran several tests in identical
-problems with different probabilities for sensing the goal object.
-These problems are denoted by the -easy, -medium and -hard suffixes;
-in the easy case the probability of percieving the object is 0.9, in
-medium it is 0.65 and 0.4 in hard.
-
-As the initial state of the problems is stochastic, we ran 50
-simulations for each configuration. In each run the simulator
-generated the true world state by sampling from the same initial state
-distribution the planner was using. So on average the world did
-conform to the planner's expectations. Not all problems generated in
-this manner have a solution. We did not reject these problems, as we
-think that detecting the non-existance of a solution is as valuable in
-exploration problems than finding one.
+problems with different probabilities for sensing the goal object. The
+perception probabilities were 0.9, 9.65 and 04. As the initial state
+of the problems is stochastic, we ran 50 simulations for each
+configuration. The samples for the true initial state were drawn from
+the same distribution that was used for planning, so the planner's
+model matched its simulated world exactly. Not all problems generated
+in this manner have a solution. We did not reject these problems, as
+we believe that detecting the non-existance of a solution is as valuable
+in exploration problems than finding one.
 
@@
 
-The graphs show the average costs of the agent's actions, the number
-of solved tasks as a percentage of the solvable tasks, as well as the
-planing time for the continual and decision theoretic planners.
+\begin{figure}[h!]
+  \centering
+  \includegraphics{dora1-time}
+  \includegraphics{dora2-time}
+  \includegraphics{dora3-time}
+  \includegraphics{dora4-time}
+  \includegraphics{dora56-time}
+  \caption{Average runtime}
+  \label{fig:results-time}
+\end{figure}
 
-For objects that can be easily detected there is little gain in using
-a decision theoretic planner, as the greedy sensing appoach by the
-baseline continual planner is obviously sufficient here. With
-decreasing sensor reliability the more sophisticated observation
-planning pays off: while the resulting plans are still longer on
-average, the impact on the number of solved tasks was much smaller
-than for the baseline system.
+\begin{figure}[h!]
+  \centering
+  \includegraphics{dora1-quality}
+  \includegraphics{dora2-quality}
+  \includegraphics{dora3-quality}
+  \includegraphics{dora4-quality}
+  \includegraphics{dora56-quality}
+  \caption{Average plan costs and number of successful runs}
+  \label{fig:results-quality}
+\end{figure}
 
-We think that a part of the improvement is due to the segmentation of
+The graphs in figure \ref{fig:results-time} show the average planning
+times. For the smaller belief states, the cost for the decision
+theoretic planning is compensated by the decrease of time spent in
+Fast Downward.
+
+Figure \ref{fig:results-quality} shows the average costs of the
+executed plans as well as the percentage of solvable tasks that were
+actually solved by the planner. For objects that can be easily
+detected there is little gain in using a decision theoretic planner,
+as the greedy sensing appoach by the baseline continual planner is
+obviously sufficient here. With decreasing sensor reliability the more
+sophisticated observation planning pays off: while the resulting plans
+are still longer on average, the impact on the number of solved tasks
+was much smaller than for the baseline system.
+
+Impact of state space size: the configuration with a belief space size
+of 200 was noticably slower than the other configurations while having
+no noticable benefit.
+
+We believe that a part of the improvement is due to the segmentation of
 the plan into several subtask, essentially performing hierarchical

### 2010-11-28 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
More work on evaluation. Converted plots to pdf.

@@
 
-The basic building blocks of the domain consist of {\tt rooms}, {\tt
-places} and {\tt objects}. Places are grouped in rooms and objects (as
-well as the robot) are always in a place. The robot can move around
-the rooms via connections between places given by the {\tt connected}
-predicate. Each room has a (possibly unknown) {\em category/}
-(e.g. kitchen, office, living room) and depending on this category,
-objects are placed inside the rooms. An instance of this domain is
-shown in figure\ref{fig}.
+\begin{figure}[h]
+  \centering
+  \input{figures/dora}
+  \caption{A exploration domain with three rooms and 13 places. The
+    robot is in in the center of room A}
+\label{fig:dora2}
+\end{figure}
+An instance of this domain is shown in figure \ref{fig:dora2}. The
+basic building blocks of the domain consist of {\tt rooms}, {\tt
+  places} and {\tt objects}. Places are grouped in rooms and objects
+(as well as the robot) are always in a place. The robot can move
+around the rooms via connections between places given by the {\tt
+  connected} predicate. Each room has a (possibly unknown) {\em
+  category/} (e.g. kitchen, office, living room) and depending on this
+category, objects are placed inside the rooms.
 
@@
 
-We conducted experiments on scenarios of several sizes with several
-kinds of goals (though the goal is always reporting the position of a
-certain kind of object). In order to determine the impact of sensor
-reliability on both approaches, we also ran several tests in identical
-problems with different probabilities for sensing the goal object. The
-perception probabilities were 0.9, 9.65 and 04. As the initial state
-of the problems is stochastic, we ran 50 simulations for each
-configuration. The samples for the true initial state were drawn from
-the same distribution that was used for planning, so the planner's
-model matched its simulated world exactly. Not all problems generated
-in this manner have a solution. We did not reject these problems, as
-we believe that detecting the non-existance of a solution is as valuable
-in exploration problems than finding one.
+We conducted experiments on scenarios of several sizes. The robot's
+goal was to find one or more objects and report back their position to
+a person. In order to determine the impact of sensor reliability on
+both approaches, we also ran tests on identical problems but changed
+the sensor model for the goal objects: the probability for percieving
+the object if it was there went from 0.9 in the easiest case to 0.65
+and 0.4 in the average and hard cases. In addition, we wanted to test
+how our system performed on tasks requiring indirect sensing, so we
+gave the planner goals to visit a certain type of room (e.g. kitchen
+or office). We didn't place any humans in those scenarios, so the only
+way of determining the room category was by looking for objects
+typical for that room type. As this indirect reasoning is a tasks that
+cannot be performed by the continual planner alone, we did not run the
+baseline system on those problems.
+
+As the initial state of the problems is stochastic, we ran 50
+simulations for each configuration. The samples for the true initial
+state were drawn from the same distribution that was used for
+planning, so the planner's model matched its simulated world
+exactly. Not all problems generated in this manner have a solution. We
+did not reject these problems, as we believe that detecting the
+non-existance of a solution is important in stochastic domains.
 
@@
 
+Using a satisficing, optimistic serial planner and using continual
+planning makes optimising for the expected reward difficult. The
+continual planner will in general be overly optimistic regarding the
+remaining costs to the goal while being overly pessimistic regarding
+the probability of reaching it (as it only regards one trace). This
+means that setting the reward function to non-extremal setting had
+little effect on the resulting planner behaviour. For this reason, we
+chose not to use the expected reward as a metric for our results, as
+that information would have been of little value to the
+reader. Instead we show the average costs of the plan and the success
+rate (as a ratio between solved and solvable problems) separately.
+
 \begin{figure}[h!]
-  \centering
-  \includegraphics{dora1-time}
-  \includegraphics{dora2-time}
-  \includegraphics{dora3-time}
-  \includegraphics{dora4-time}
-  \includegraphics{dora56-time}
+  % \centering
+  % \includegraphics{dora1-time}\hfill
+  % \vspace{2mm}
+  \includegraphics{dora2-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora3-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora4-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora56-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora-cat-time}\hfill
   \caption{Average runtime}
@@
 \begin{figure}[h!]
-  \centering
-  \includegraphics{dora1-quality}
-  \includegraphics{dora2-quality}
-  \includegraphics{dora3-quality}
-  \includegraphics{dora4-quality}
-  \includegraphics{dora56-quality}
+  % \centering
+  % \includegraphics{dora1-quality}\hfill
+  % \vspace{2mm}
+  \includegraphics{dora2-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora3-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora4-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora56-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora-cat-quality}\hfill
   \caption{Average plan costs and number of successful runs}
@@
 The graphs in figure \ref{fig:results-time} show the average planning
-times. For the smaller belief states, the cost for the decision
-theoretic planning is compensated by the decrease of time spent in
-Fast Downward.
+times. Not surprisingly, the time the POMDP planner takes increases
+strongly with larger belief spaces. But for the smaller belief states,
+the cost for the decision theoretic planning is compensated by the
+decrease of time spent in Fast Downward.
 
@@
 
-Impact of state space size: the configuration with a belief space size
-of 200 was noticably slower than the other configurations while having
-no noticable benefit.
+As already mentioned, less aggressive pruning of the initial belief
+space resultes in longer runtimes. The impace of the pruning on plan
+costs and success rate were limited, though. Increasing the size limit
+beyond 50 did rarely pay off because while additional information
+might in theroy allow better plans, (add reason here). The indirect
+sensing tests show the same result, though with much higher time spent
+in the decision theoretic planner. The relatively high success rate
+even with small belief spaces might also indicate that the entropy
+heuristic for belief space pruning is effective in keeping the most
+relevant information.
 
-We believe that a part of the improvement is due to the segmentation of
-the plan into several subtask, essentially performing hierarchical
-planning. Especially when the continual planner performs badly this is
-a huge gain.
+% We believe that a part of the improvement is due to the segmentation of
+% the plan into several subtask, essentially performing hierarchical
+% planning. Especially when the continual planner performs badly this is
+% a huge gain.
 

### 2010-11-28 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Few fixes requested by Moritz

@@
 replan when actions have unplanned-for
-outcomes \cite{yoon:etal:2007}. However, these approaches are not as
-useful in partially observable domains as replanning is triggered when
-the system ends up in a state other than the desired one, and this
-state cannot in general be determined.
+outcomes \cite{yoon:etal:2007}. However, these approaches rely
+crucially on being able to determine properties of the current state
+in order to decide to replan. For partially observable domains, only
+probabilistic information about these properties will in general be
+available.
 
@@
 reasoning about observations is required to find good plans. We refer
-to the complete system as a {\em switching planner}. The approach is
-not optimal, particularly as it is based on a satisficing sequential
-planner, but it performs better than the sequential planner alone and
-is fast enough to be used on the robot for decision-making. As an
-illustration, consider a robot searching for a box that it knows is
-often found in offices. The sequential planner might produce a plan
+to the complete system as a {\em switching continual planner}. The
+approach is not optimal, particularly as it is based on a satisficing
+sequential planner, but it performs better than the sequential planner
+alone and is fast enough to be used in real-time on a mobile robot for
+decision-making.
+
+As an illustration, consider a robot searching for a box that it knows
+is often found in offices. The sequential planner might produce a plan
 that assumes that {\sc Room a} is an office, that the box is in {\sc
-Room a}, and that the box is in location 1 in {\sc Room a}, and then
-goes to {\sc Room a}, goes to location 1 and searches for the
-box. When the plan gets executed and the search begins, the
-decision-theoretic planner is called to conduct the search, and at
-this point may retract the last assumption and build a plan to search
-the whole of {\sc Room a} for the box. If this plan fails, it will do
-so by disproving one of the assumptions (for example, that {\sc Room
-a} is an office), which will then lead to replanning by the sequential
-planner based on the newly discovered knowledge.
+Room a}, and that the box is in location 1 in {\sc Room a}. Given
+these assumptions, the plan might be to go to {\sc Room a}, go to
+location 1 and search for the box. When the plan gets executed and the
+search begins the decision-theoretic planner is called since searching
+involves making observations. The decision-theoretic planner might
+retract the last assumption but leave the others, and build a plan to
+search the whole of {\sc Room a} for the box. If this plan fails to
+find the box, it will disprove the assumption that the box is in {\sc
+Room a}, which will then lead to replanning by the sequential planner
+based on the newly discovered knowledge.
 
@@
 
-The sequential planner we use is Continual
-Planning \cite{brenner:nebel:jaamas09} (CP). CP is based
-on \fastdownward \cite{fast-downward}. CP already has the capability
-to replan, and also allows agent knowledge to be represented
-explicitly, so we can write actions that gain the agent knowledge. For
-the decision-theoretic planner we use an information-state
-\laostar\ procedure that uses the solution to the equivalent MDP as a
-heuristic. For reasons of space, we will not describe this system in
-detail here.
+The sequential planner we use is based
+on \fastdownward \cite{fast-downward}. We add the capability to
+replan, and also allow agent knowledge to be represented explicitly,
+so we can write actions that gain the agent knowledge. For the
+decision-theoretic planner we use forward search in the information
+state.
 
 The switching planner approach gives us some important advantages in
-our robotic domain. First, the main source of uncertainty in the
-domain is unreliable observations, rather than stochasticity in the
-action effects, and this approach is particularly suited to problems
-that combine a deterministic task planning problem with observations
-as the decision-theoretic planner is only invoked to determine the
-values of variables. Secondly, the replanning makes the system very
-robust to changing goals and the discovery of new facts about the
-world, both of which are features of our domain.
+our robotic domain. First, the significant source of uncertainty in
+the domain is the unreliability of observations, and this approach is
+particularly suited to problems that combine a deterministic task
+planning problem with observations as the decision-theoretic planner
+is only invoked to determine a characteristic of the underlying
+state. Secondly, the replanning makes the system very robust to
+changing objectives and the discovery of new facts about the world, both of
+which are features of our domain.
+
+The remainder of the paper proceeds as follows. We begin by describing
+POMDPs with a propositionally factored representation of states and
+observations. We also outline the relationship between this and a flat
+POMDP representation. Next we describe our DTPDDL language with
+example declarations from a mobile robot exploration task. Then we
+describe the switching planner, including how the sequential domain is
+produced and a plan is generated, and how this then provides the input
+for the decision-theoretic planner. Finally we provide an evaluation
+of the system on the example robot domain, and finish with a short
+survey of related work and some concluding remarks.
 
@@
 %% quickly~\cite{littman:etal:98}
-
-An outline of the paper is as follows. We begin by describing POMDPs with a
-propositionally factored representation of states and observations,
-and the relationship between this and a flat POMDP representation. Next
-we describe our DTPDDL language using a running example of the robot
-looking for a box. Then we describe the switching planner, including
-how the sequential domain is produced and a plan is generated, and how
-this then provides the input for the decision-theoretic
-planner. Finally we provide an evaluation of the system on the example
-robot domain, and finish with a short survey of related work and some
-concluding remarks.

### 2010-11-28 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Few fixes requested by Moritz

@@
 size. In detail, for each excluded atomic term, we compute the {\em
-entropy} of the corresponding proposition, {\em conditional} on the
-active assumptions. The {\em conditional entropy} measure
-is higher the more uncertain that proposition is given truth
-assignments to propositions that are the subject of active
-assumptions. Propositions that in the first place were excluded are
-iteratively added to the problem posed to the conditional session in
-increasing order according to that measure.
+entropy} of the active assumptions, {\em conditional} on the
+corresponding proposition. Loosely speaking, the {\em conditional
+entropy} measure is higher the more uncertain those assumptions are
+given truth assignment to the proposition. Propositions that in the
+first place were excluded are iteratively added to the problem posed
+to the conditional session in increasing order according to that
+measure.
+
+%% In detail, for each excluded atomic term, we compute the {\em
+%% entropy} of the corresponding proposition, {\em conditional} on the
+%% active assumptions. The {\em conditional entropy} measure
+%% is higher the more uncertain that proposition is given truth
+%% assignments to propositions that are the subject of active
+%% assumptions. Propositions that in the first place were excluded are
+%% iteratively added to the problem posed to the conditional session in
+%% increasing order according to that measure.
 

### 2010-11-28 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
Few fixes requested by Moritz

@@
 %%
-Notationally, we say an action $\stochAction \in \stochActions$ is
+We say an action $\stochAction \in \stochActions$ is
 applicable if its precondition $\poss(\stochAction)$, a set of
@@
 \observations$ that informs what should be believed about the state a
-process is in. In detail, if $\action$ is applied effecting a
+process is in. If $\action$ is applied effecting a
 transition to a successor state $\state'$, then an observation occurs
@@
 1$. The observation received by the agent corresponds to the union of
-perceptual propositions from chosen elements of active
+perceptual propositions from the chosen elements of active
 senses.

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
MOre notes in the introduction

@@
 stochastic actions is extremely computationally
-expensive \cite{mdp-complexity}. Even relatively small partially
+expensive~\cite{mdp-complexity}. Even relatively small partially
 observable Markov decision problems (POMDPs), represented by tens of
@@
 alternative approach has been to use classical sequential planners and
-replan when actions have unplanned-for
-outcomes \cite{yoon:etal:2007}. However, these approaches rely
-crucially on being able to determine properties of the current state
-in order to decide to replan. For partially observable domains, only
-probabilistic information about these properties will in general be
-available.
+replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
+Execution monitoring tracks properties of the state that determine
+whether a new plan is necessary. For partially observable domains,
+only probabilistic information about such properties will, in general,
+be available.
 
-In this paper we describe, \pcogx, the planning component in a robotic
-system that continuously deliberates in a stochastic dynamic
-environment in order to achieve objectives set by the user, and
-acquire knowledge about its surroundings. This domain has the
-characteristics of a POMDP in that the state is not perfectly known,
-and state information is gained by the robot through the use of
-unreliable sensing actions. However, the state representation is
-propositional, and even for relatively simple problems the underlying
-model is too large for a POMDP solver to be applied.
+%% However, these approaches rely
+%% crucially on being able to determine properties of the current state
+%% in order to decide to replan. For partially observable domains, only
+%% probabilistic information about these properties will in general be
+%% available.
 
-Inspired by the replanning approaches used successfully for MDP
-planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
-that uses a replanning sequential planner to build the majority of the
-plan, but uses a decision-theoretic planner on small subproblems where
-reasoning about observations is required to find good plans. We refer
-to the complete system as a {\em switching continual planner}. The
-approach is not optimal, particularly as it is based on a satisficing
-sequential planner, but it performs better than the sequential planner
-alone and is fast enough to be used in real-time on a mobile robot for
-decision-making.
+This paper describes the planning component for a robotic system that
+continuously deliberates in a stochastic dynamic environment in order
+to achieve objectives set by the user, and acquire knowledge about its
+surroundings. This domain features partial observability, particularly
+because the state is not perfectly known, and state information is
+gained by the robot through the use of unreliable sensing
+actions. Using a propositionally factored state representation, for
+interesting tasks the corresponding POMDP model is too large for a
+POMDP solver to be applied directly. Inspired by the replanning
+approaches used successfully for MDP
+planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose an approach
+that uses a replanning system to compute a reasonably valuable trace
+in the model, and then uses a decision-theoretic planner on small
+subproblems where reasoning about observations might be useful. We
+refer to the complete system as a {\em switching continual
+planner}. The approach is not optimal, particularly as it is based on
+a satisficing sequential planner, but it performs better than the
+sequential planner alone, and is fast enough to be used in real-time
+on a mobile robot for decision-making.
 
-As an illustration, consider a robot searching for a box that it knows
-is often found in offices. The sequential planner might produce a plan
-that assumes that {\sc Room a} is an office, that the box is in {\sc
-Room a}, and that the box is in location 1 in {\sc Room a}. Given
-these assumptions, the plan might be to go to {\sc Room a}, go to
-location 1 and search for the box. When the plan gets executed and the
-search begins the decision-theoretic planner is called since searching
-involves making observations. The decision-theoretic planner might
-retract the last assumption but leave the others, and build a plan to
-search the whole of {\sc Room a} for the box. If this plan fails to
-find the box, it will disprove the assumption that the box is in {\sc
-Room a}, which will then lead to replanning by the sequential planner
-based on the newly discovered knowledge.
+%% As an illustration, consider a robot searching for a box that it knows
+%% is often found in offices. The sequential planner might produce a plan
+%% that assumes that {\sc Room a} is an office, that the box is in {\sc
+%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
+%% these assumptions, the plan might be to go to {\sc Room a}, go to
+%% location 1 and search for the box. When the plan gets executed and the
+%% search begins the decision-theoretic planner is called since searching
+%% involves making observations. The decision-theoretic planner might
+%% retract the last assumption but leave the others, and build a plan to
+%% search the whole of {\sc Room a} for the box. If this plan fails to
+%% find the box, it will disprove the assumption that the box is in {\sc
+%% Room a}, which will then lead to replanning by the sequential planner
+%% based on the newly discovered knowledge.
 
-To represent planning problems of this kind, we have developed a
-first-order declarative language, called decision-theoretic planning
-domain definition language (DTPDDL), for describing domains of
-planning problems that correspond to POMDPs. We have created a system
-that automatically constructs from this a deterministic planning domain
-for the sequential planner that includes actions that allow
-assumptions to be made about the values of imperfectly known state
-variables. These are then used to restrict the state space and set the
-reward function for the decision-theoretic planner, which can operate
-with the original representation of the domain.
+Our switching system is domain independent, taking domain and problem
+descriptions in a first-order declarative language we have developed,
+called the decision-theoretic planning domain definition language
+(DTPDDL). From those descriptions of POMDP, we automatically construct
+a deterministic model for the sequential planner, that includes
+operators which correspond to assumptions about the values of
+imperfectly known state variables. Assumptions scheduled by the
+sequential system are used to propose a pragmatic abstract
+belief-state space to the decision-theoretic system, and to modify the
+reward function for the decision-theoretic planner, so that it might
+pursue sensing related to those assumptions.
+
+%%  which can operate
+%% with the original representation of the domain.
 
@@
 replan, and also allow agent knowledge to be represented explicitly,
-so we can write actions that gain the agent knowledge. For the
-decision-theoretic planner we use forward search in the information
-state.
+so we can write actions that gain the agent knowledge. LEASE CITE
+MICHAEL's WORK AS THE BASIS FOR THIS. For the decision-theoretic
+planner we have implemented our own forward search in the belief-state
+space.
 
-The switching planner approach gives us some important advantages in
-our robotic domain. First, the significant source of uncertainty in
-the domain is the unreliability of observations, and this approach is
+Our switching planner gives us some important advantages in our
+robotic domain. First, the significant source of uncertainty in the
+domain is the unreliability of observations, and this approach is
 particularly suited to problems that combine a deterministic task
@@
 state. Secondly, the replanning makes the system very robust to
-changing objectives and the discovery of new facts about the world, both of
-which are features of our domain.
+changing objectives and the discovery of new facts about the world,
+both of which are features of our domain.
 

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
Fixed evaluation section

@@
 
-The implementation of the continual planning and switching parts is
-based on MAPSIM \cite{brenner:nebel:jaamas09} and is able to use
-several planners as the underlying planning system. We use a modified
-version of Fast Downward \cite{fast-downward}, which we extended with
-the support for actions with success probabilities. Finally, we use
-\system{dlib-ml} \cite{king:2009} for successive estimation of the
-underlying belief-state.
+We have implemented our switching continual approach in MAPSIM
+environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
+to use several underlying planning systems. We have also extended the
+MAPSIM system so that it can parse DTPDDL, and perform successive
+estimation of the underlying belief-state using the
+\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
+we have use our own version of Fast Downward~\cite{fast-downward} for
+sequential sessions. We have extended that system to support actions
+with success probabilities. In our evaluation Fast Downward is run
+with the cyclic causal graph heuristic using an A* search or weighted
+A* (with weight 5). We use A* for easy problems and WA* for more
+difficult problems where A* was ineffective. Contingent sessions use
+our own forward search procedure.
 
-The baseline approach is using the same continual planning system as
-the switching planner, but instead of creating an observation problem
-for the decision theoretic planner it will just execute one sensing
-action -- assuming that this action will confirm its assumption.
+\Omit{ We then perform multiple tests with different limits for the
+  belief-space size in contingent sessions.  Higher limits should
+  cause longer planning times but be beneficial to plan quality as
+  more contingencies can be taken in to account by the POMDP planner.
+}
 
-To test our approch we use a robot exploration domain similar to the
-one used on a physical robotic system. It involves a robot exploring
-an office or living environment and trying to return objects to their
-owner.
+In order to evaluate our system, we have also implemented a baseline
+approach in MAPSIM. Rather than invoking a contingent session when a
+switching action $\switchAction$ is scheduled for execution, the
+baseline simply executes one action that can trigger a sensing outcome
+determined by the precondition of $\switchAction$, and replans in the
+resulting belief-state.
+
+%% executes switching actions and replans in the
+%% resulting belief state.
+
+%% but instead of
+%% creating an observation problem for the decision theoretic planner it
+%% will just execute one sensing action -- assuming that this action will
+%% confirm its assumption.
+
+To test our approach we use a robot exploration domain based on a that
+from our physical robotic system. Here, a robot is exploring an office
+or living environment, and trying to report the locations of objects
+to their owner.
 
@@
   \caption{A exploration domain with three rooms and 13 places. The
-    robot is in in the center of room A}
+    robot is in in the centre of room A}
 \label{fig:dora2}
 \end{figure}
-An instance of this domain is shown in figure \ref{fig:dora2}. The
-basic building blocks of the domain consist of {\tt rooms}, {\tt
-  places} and {\tt objects}. Places are grouped in rooms and objects
-(as well as the robot) are always in a place. The robot can move
-around the rooms via connections between places given by the {\tt
-  connected} predicate. Each room has a (possibly unknown) {\em
+An instance of this domain is shown in figure \ref{fig:dora2}. Basic
+types of element are{\tt rooms}, {\tt places} and {\tt
+  objects}. Places are topological map nodes that occur in rooms. and
+Objects (including the robot) are always located at a place. The robot
+can move around the rooms via connections between places given by the
+{\tt connected} predicate. Each room has a (possibly unknown) {\em
   category/} (e.g. kitchen, office, living room) and depending on this
@@
 The robot can find out if an object is at a certain place by executing
-the {\tt look-for-object} action, which may result in a perception if
-the object is in fact there (though some objects are harder to detect
-than others -- so absence of a percept is no proof of the object's
-absence). Additionally, if the robot is in the presence of a human, it
-may simply ask what type of room they are currently in -- but
-conducting a dialogue is a bit more costly than simply running the
-vision algorithm (cost of 8 vs costs of 3).
+the {\tt look-for-object} action. Moreover, the contents of a room
+indicates its category, and certain objects are more likely to be
+located in rooms of certain categories. For example, a box of
+cornflakes is more likely to be located in the kitchen than the
+office. The {\tt look-for-object} might result in a positive
+perception if the object is there. Some objects are harder to detect
+than others, so that an instantaneous positive detection of the object
+is not proof positive that it is there. Additionally, if the robot is
+in the presence of a human, it can ask what type of room they are
+currently in, however conducting a dialogue is more costly than
+running a vision algorithm (cost of 8 vs costs of 3).
 
-We conducted experiments on scenarios of several sizes. The robot's
-goal was to find one or more objects and report back their position to
-a person. In order to determine the impact of sensor reliability on
-both approaches, we also ran tests on identical problems but changed
-the sensor model for the goal objects: the probability for percieving
-the object if it was there went from 0.9 in the easiest case to 0.65
-and 0.4 in the average and hard cases. In addition, we wanted to test
-how our system performed on tasks requiring indirect sensing, so we
-gave the planner goals to visit a certain type of room (e.g. kitchen
-or office). We didn't place any humans in those scenarios, so the only
+We perform experiments on tasks of several sizes. The robot's
+objective is to find one or more objects and report back their
+position to a person. In order to determine the impact of sensor
+reliability in a comparison with the baseline, we also ran tests on
+identical problems but changed the sensor model for the target
+objects: the probability for perceiving the object if it was there
+went from 0.9 in the easiest case to 0.65 and 0.4 in the average and
+hard cases respectively. In addition, we wanted to test how our system
+performed on tasks requiring indirect sensing, so we gave the planner
+rewards for visiting a certain type of room (e.g. kitchen or
+office). We did not place any humans in those scenarios, so the only
 way of determining the room category was by looking for objects
@@
 
-As the initial state of the problems is stochastic, we ran 50
-simulations for each configuration. The samples for the true initial
-state were drawn from the same distribution that was used for
-planning, so the planner's model matched its simulated world
-exactly. Not all problems generated in this manner have a solution. We
-did not reject these problems, as we believe that detecting the
-non-existance of a solution is important in stochastic domains.
-
-Fast Downward was run with the cyclic causal graph heuristic using A*
-search or weighted A* with a weight of 5, depending on the difficulty
-of the problem. We then performed multiple tests with different limits
-for the belief space size of the decision theoretic subtask. Higher
-limits should cause longer planning times but be beneficial to plan
-quality as more contingencies can be taken in to account by the POMDP
-planner.
+In our evaluation we run 50 simulations in each configuration. The
+samples for the true initial state are drawn from the same
+distribution that is used for planning, therefore the planer and
+simulation models are in exact correspondence. Not all problems have a
+solution, and we explore whether or not detection of the non-existence
+of a solution is important in stochastic domains.
+%%
+We also perform multiple tests
+with different limits for the belief-space size in contingent
+sessions.  Higher limits should cause longer planning times but be
+beneficial to plan quality as more contingencies can be taken in to
+account by the POMDP planner.
 
@@
 continual planner will in general be overly optimistic regarding the
-remaining costs to the goal while being overly pessimistic regarding
-the probability of reaching it (as it only regards one trace). This
+remaining costs to a goal while being overly pessimistic regarding the
+probability of reaching it (as it only considered one trace). This
 means that setting the reward function to non-extremal setting had
@@
 chose not to use the expected reward as a metric for our results, as
-that information would have been of little value to the
-reader. Instead we show the average costs of the plan and the success
-rate (as a ratio between solved and solvable problems) separately.
+that information would not have been informative. Instead we show the
+average costs of the plan and the success rate (as a ratio between
+solved and solvable problems) separately.
 
@@
 The graphs in figure \ref{fig:results-time} show the average planning
-times. Not surprisingly, the time the POMDP planner takes increases
-strongly with larger belief spaces. But for the smaller belief states,
-the cost for the decision theoretic planning is compensated by the
-decrease of time spent in Fast Downward.
+times. Not surprisingly, the time for contingent planning increases
+steeply as we have more states with non-zero probability (i.e., less
+abstract) in $\bstate_0$. For small initial configurations, the cost
+of contingent planning is compensated by the decrease of time spent in
+Fast Downward.
 
@@
 actually solved by the planner. For objects that can be easily
-detected there is little gain in using a decision theoretic planner,
-as the greedy sensing appoach by the baseline continual planner is
+detected there is little gain in using a decision-theoretic planner,
+as the greedy sensing approach by the baseline continual planner is
 obviously sufficient here. With decreasing sensor reliability the more
-sophisticated observation planning pays off: while the resulting plans
-are still longer on average, the impact on the number of solved tasks
-was much smaller than for the baseline system.
+sophisticated observation strategies of contingent planning pay off:
+while the resulting plans are still longer on average, the impact on
+the number of solved tasks was much smaller than for the baseline
+system.
 
-As already mentioned, less aggressive pruning of the initial belief
-space resultes in longer runtimes. The impace of the pruning on plan
-costs and success rate were limited, though. Increasing the size limit
-beyond 50 did rarely pay off because while additional information
-might in theroy allow better plans, (add reason here). The indirect
-sensing tests show the same result, though with much higher time spent
-in the decision theoretic planner. The relatively high success rate
-even with small belief spaces might also indicate that the entropy
-heuristic for belief space pruning is effective in keeping the most
-relevant information.
+Less aggressive abstraction of the initial belief space results in
+longer runtimes. We find that abstraction has little impact on plan
+costs and success rate in our scenario. Increasing the size of the
+initial abstract belief beyond 50 states rarely pays off, because
+while additional information may facilitate better plans, the indirect
+sensing tests report the same result. Here, we only really see an
+impact in terms of the runtime of contingent planning. The relatively
+high success rate even with small belief spaces indicates the
+effectiveness of using conditional entropy to guide abstraction
+refinement.
 

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex

@@
 
-We have implemented our switching continual approach in MAPSIM
+We have implemented our switching continual approach in the MAPSIM
 environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
-to use several underlying planning systems. We have also extended the
+to use several underlying planning systems. We have also extended
 MAPSIM system so that it can parse DTPDDL, and perform successive
-estimation of the underlying belief-state using the
+estimation of the underlying belief-state using
 \system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we have use our own version of Fast Downward~\cite{fast-downward} for
+we use our own version of Fast Downward~\cite{fast-downward} for
 sequential sessions. We have extended that system to support actions
@@
 A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* was ineffective. Contingent sessions use
+difficult problems where A* is ineffective. Contingent sessions use
 our own forward search procedure.
@@
 types of element are{\tt rooms}, {\tt places} and {\tt
-  objects}. Places are topological map nodes that occur in rooms. and
-Objects (including the robot) are always located at a place. The robot
+  objects}. Places are topological map nodes that occur in rooms.
+Objects, including the robot, are always located at a place. The robot
 can move around the rooms via connections between places given by the
-{\tt connected} predicate. Each room has a (possibly unknown) {\em
-  category/} (e.g. kitchen, office, living room) and depending on this
+{\tt connected} predicate. Each room has a, possibly unknown, {\em
+  category} (e.g. kitchen, office, living room) and depending on this
 category, objects are placed inside the rooms.
@@
 perception if the object is there. Some objects are harder to detect
-than others, so that an instantaneous positive detection of the object
-is not proof positive that it is there. Additionally, if the robot is
-in the presence of a human, it can ask what type of room they are
+than others, and an instantaneous positive detection of the object is
+not proof positive that it is there. Additionally, if the robot is in
+the presence of a human, it can ask what type of room they are
 currently in, however conducting a dialogue is more costly than
@@
 position to a person. In order to determine the impact of sensor
-reliability in a comparison with the baseline, we also ran tests on
+reliability in a comparison with the baseline, we also run tests on
 identical problems but changed the sensor model for the target
@@
 went from 0.9 in the easiest case to 0.65 and 0.4 in the average and
-hard cases respectively. In addition, we wanted to test how our system
-performed on tasks requiring indirect sensing, so we gave the planner
-rewards for visiting a certain type of room (e.g. kitchen or
-office). We did not place any humans in those scenarios, so the only
-way of determining the room category was by looking for objects
-typical for that room type. As this indirect reasoning is a tasks that
-cannot be performed by the continual planner alone, we did not run the
-baseline system on those problems.
+hard cases respectively. In addition, we test how our system performs
+in tasks requiring indirect sensing. Here, we gave the planner rewards
+for visiting a certain type of room (e.g. kitchen or office). We did
+not place any humans in those scenarios, so the only way of
+determining the room category is by looking for objects typical for
+that room type. As this indirect reasoning is a tasks that cannot be
+performed by the continual planner alone, we did not run the baseline
+system on those problems.
 
 In our evaluation we run 50 simulations in each configuration. The
-samples for the true initial state are drawn from the same
-distribution that is used for planning, therefore the planer and
-simulation models are in exact correspondence. Not all problems have a
-solution, and we explore whether or not detection of the non-existence
-of a solution is important in stochastic domains.
+planning and simulation models are in exact correspondence. Not all
+problems have a solution, so we also examine whether or not detection
+of the non-existence of a solution is important in stochastic domains.
 %% 
-We also perform multiple tests
-with different limits for the belief-space size in contingent
-sessions.  Higher limits should cause longer planning times but be
-beneficial to plan quality as more contingencies can be taken in to
-account by the POMDP planner.
+We perform multiple tests with different limits for the
+belief-space size in contingent sessions.  Higher limits result in
+longer contingent planning times and higher plan quality.
 
-Using a satisficing, optimistic serial planner and using continual
-planning makes optimising for the expected reward difficult. The
-continual planner will in general be overly optimistic regarding the
-remaining costs to a goal while being overly pessimistic regarding the
-probability of reaching it (as it only considered one trace). This
-means that setting the reward function to non-extremal setting had
-little effect on the resulting planner behaviour. For this reason, we
-chose not to use the expected reward as a metric for our results, as
-that information would not have been informative. Instead we show the
-average costs of the plan and the success rate (as a ratio between
-solved and solvable problems) separately.
+Using a satisficing, optimistic serial planner in sequential sessions
+makes optimising for the expected reward difficult. Only reasoning
+about single traces, it is optimistic wrt the remaining costs to a
+goal and pessimistic regarding the probability of reaching tat. By
+setting the rewards to non-extremal values has little effect on that
+system. Therefore, we do not consider expected utility in our
+experiments. Instead, we report average costs of the plans and the
+success rate (as a ratio between solved and solvable problems).
 
@@
 
-The graphs in figure \ref{fig:results-time} show the average planning
-times. Not surprisingly, the time for contingent planning increases
-steeply as we have more states with non-zero probability (i.e., less
-abstract) in $\bstate_0$. For small initial configurations, the cost
-of contingent planning is compensated by the decrease of time spent in
-Fast Downward.
+%% Figure \ref{fig:results-quality} shows the average costs of the
+%% executed plans as well as the percentage of solvable tasks that were
+%% actually solved by the planner.
 
-Figure \ref{fig:results-quality} shows the average costs of the
-executed plans as well as the percentage of solvable tasks that were
-actually solved by the planner. For objects that can be easily
-detected there is little gain in using a decision-theoretic planner,
-as the greedy sensing approach by the baseline continual planner is
-obviously sufficient here. With decreasing sensor reliability the more
-sophisticated observation strategies of contingent planning pay off:
-while the resulting plans are still longer on average, the impact on
-the number of solved tasks was much smaller than for the baseline
-system.
+Our experimental results are summarised in
+Figure~\ref{fig:results-time} and Figure~\ref{fig:results-quality}. We
+find that the time spent on contingent planning increases steeply as
+the abstraction $\bstate_0$ becomes more refined. For coarse initial
+configurations, the cost of contingent planning is compensated by the
+decrease of time spent in Fast Downward. Addressing success ratios and
+plan costs, for objects that can be easily detected there is little
+gain in using a decision-theoretic planner, as the greedy sensing
+approach by the baseline is sufficient. With decreasing sensor
+reliability contingent planning pays off.%% : while the resulting plans are still
+%% longer on average, the impact on the number of solved tasks was much
+%% smaller than for the baseline system.  Less aggressive abstraction
+of the initial configuration results in longer runtimes. In our
+scenario it is worthwhile being assumptively aggressive.  We suppose
+the relatively high success rate irrespective of the coarseness of the
+initial configuration indicates the effectiveness of using conditional
+entropy to guide abstraction refinement.
 
-Less aggressive abstraction of the initial belief space results in
-longer runtimes. We find that abstraction has little impact on plan
+We find that abstraction has little impact on plan
 costs and success rate in our scenario. Increasing the size of the
@@
 sensing tests report the same result. Here, we only really see an
-impact in terms of the runtime of contingent planning. The relatively
-high success rate even with small belief spaces indicates the
-effectiveness of using conditional entropy to guide abstraction
-refinement.
+impact in terms of the runtime of contingent planning.
 

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
I am stuck in Germany. Will be in Vienna at 9pm. Moritz, I will try to contact you then.

@@
 
-To test our approach we use a robot exploration domain based on a that
-from our physical robotic system. Here, a robot is exploring an office
-or living environment, and trying to report the locations of objects
-to their owner.
+To test our approach we use a robot exploration domain based on a
+scenario from our physical robotic system. Here, a robot is
+exploring an office or living environment, and trying to report the
+locations of objects to their owner.
 
@@
 office. The {\tt look-for-object} might result in a positive
-perception if the object is there. Some objects are harder to detect
-than others, and an instantaneous positive detection of the object is
-not proof positive that it is there. Additionally, if the robot is in
-the presence of a human, it can ask what type of room they are
-currently in, however conducting a dialogue is more costly than
+perception if the object is there. An instantaneous positive detection
+of the object is not proof positive that it is there. Also, some
+objects are harder to detect than others. Also, Additionally, if the
+robot is in the presence of a human, it can ask what type of room they
+are currently in, however conducting a dialogue is more costly than
 running a vision algorithm (cost of 8 vs costs of 3).

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Few new changes to the introduction.

@@
 alternative approach has been to use classical sequential planners and
-replan when actions have unplanned-for
-outcomes \cite{yoon:etal:2007}. However, these approaches rely
-crucially on being able to determine properties of the current state
-in order to decide to replan. For partially observable domains, only
-probabilistic information about these properties will in general be
-available.
+replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
+Execution monitoring tracks properties of the state that determine
+whether a new plan is necessary. For partially observable domains,
+only probabilistic information about such properties will, in general,
+be available.
 
-In this paper we describe, \pcogx, the planning component in a robotic
-system that continuously deliberates in a stochastic dynamic
-environment in order to achieve objectives set by the user, and
-acquire knowledge about its surroundings. This domain has the
-characteristics of a POMDP in that the state is not perfectly known,
-and state information is gained by the robot through the use of
-unreliable sensing actions. However, the state representation is
-propositional, and even for relatively simple problems the underlying
-model is too large for a POMDP solver to be applied.
+%% However, these approaches rely
+%% crucially on being able to determine properties of the current state
+%% in order to decide to replan. For partially observable domains, only
+%% probabilistic information about these properties will in general be
+%% available.
 
-Inspired by the replanning approaches used successfully for MDP
-planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
-that uses a replanning sequential planner to build the majority of the
-plan, but uses a decision-theoretic planner on small subproblems where
-reasoning about observations is required to find good plans. We refer
-to the complete system as a {\em switching continual planner}. The
-approach is not optimal, particularly as it is based on a satisficing
-sequential planner, but it performs better than the sequential planner
-alone and is fast enough to be used in real-time on a mobile robot for
-decision-making.
+This paper describes the planning component for a robotic system that
+continuously deliberates in a stochastic dynamic environment in order
+to achieve objectives set by the user, and acquire knowledge about its
+surroundings. This domain features partial observability, particularly
+because the state is not perfectly known, and state information is
+gained by the robot through the use of unreliable sensing
+actions. Using a propositionally factored state representation, for
+interesting tasks the corresponding POMDP model is too large for a
+POMDP solver to be applied directly. Inspired by the replanning
+approaches used successfully for MDP
+planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose an approach
+that uses a replanning system to compute a reasonably valuable trace
+in the model, and then uses a decision-theoretic planner on small
+subproblems where reasoning about observations might be useful. We
+refer to the complete system as a {\em switching continual
+planner}. The approach is not optimal, particularly as it is based on
+a satisficing sequential planner, but it performs better than the
+sequential planner alone, and is fast enough to be used for real-time
+decision-making on a mobile robot.
 
-As an illustration, consider a robot searching for a box that it knows
-is often found in offices. The sequential planner might produce a plan
-that assumes that {\sc Room a} is an office, that the box is in {\sc
-Room a}, and that the box is in location 1 in {\sc Room a}. Given
-these assumptions, the plan might be to go to {\sc Room a}, go to
-location 1 and search for the box. When the plan gets executed and the
-search begins the decision-theoretic planner is called since searching
-involves making observations. The decision-theoretic planner might
-retract the last assumption but leave the others, and build a plan to
-search the whole of {\sc Room a} for the box. If this plan fails to
-find the box, it will disprove the assumption that the box is in {\sc
-Room a}, which will then lead to replanning by the sequential planner
-based on the newly discovered knowledge.
+%% As an illustration, consider a robot searching for a box that it knows
+%% is often found in offices. The sequential planner might produce a plan
+%% that assumes that {\sc Room a} is an office, that the box is in {\sc
+%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
+%% these assumptions, the plan might be to go to {\sc Room a}, go to
+%% location 1 and search for the box. When the plan gets executed and the
+%% search begins the decision-theoretic planner is called since searching
+%% involves making observations. The decision-theoretic planner might
+%% retract the last assumption but leave the others, and build a plan to
+%% search the whole of {\sc Room a} for the box. If this plan fails to
+%% find the box, it will disprove the assumption that the box is in {\sc
+%% Room a}, which will then lead to replanning by the sequential planner
+%% based on the newly discovered knowledge.
 
-To represent planning problems of this kind, we have developed a
-first-order declarative language, called decision-theoretic planning
-domain definition language (DTPDDL), for describing domains of
-planning problems that correspond to POMDPs. We have created a system
-that automatically constructs from this a deterministic planning domain
-for the sequential planner that includes actions that allow
-assumptions to be made about the values of imperfectly known state
-variables. These are then used to restrict the state space and set the
-reward function for the decision-theoretic planner, which can operate
-with the original representation of the domain.
+Our switching system is domain independent, taking domain and problem
+descriptions in a first-order declarative language we have developed,
+called the decision-theoretic planning domain definition language
+(DTPDDL). From those descriptions of POMDPs, we automatically
+construct a deterministic model for the sequential planner. That model
+includes operators which correspond to assumptions about the values of
+imperfectly known state variables. Assumptions scheduled by the
+sequential system are used to propose a pragmatic abstract
+belief-state space to the decision-theoretic system, and to modify the
+reward function, so that system might pursue sensing related to those
+assumptions.
+
+%%  which can operate
+%% with the original representation of the domain.
 
 The sequential planner we use is based
-on \fastdownward \cite{fast-downward}. We add the capability to
+on \fastdownward~\cite{fast-downward}. We add the capability to
 replan, and also allow agent knowledge to be represented explicitly,
-so we can write actions that gain the agent knowledge. For the
-decision-theoretic planner we use forward search in the information
-state.
+so we can write actions that gain the agent knowledge. LEASE CITE
+MICHAEL's WORK AS THE BASIS FOR THIS. For the decision-theoretic
+planner we have implemented our own forward search in the belief-state
+space.
 
-The switching planner approach gives us some important advantages in
-our robotic domain. First, the main source of uncertainty in the
-domain is unreliable observations, rather than stochasticity in the
-action effects, and this approach is particularly suited to problems
-that combine a deterministic task planning problem with observations
-as the decision-theoretic planner is only invoked to determine the
-values of variables. Secondly, the replanning makes the system very
-robust to changing goals and the discovery of new facts about the
-world, both of which are features of our domain.
+Our switching planner gives us some important advantages in our
+robotic domain. First, the significant source of uncertainty in the
+domain is the unreliability of observations, and this approach is
+particularly suited to problems that combine a deterministic task
+planning problem with observations as the decision-theoretic planner
+is only invoked to determine a characteristic of the underlying
+state. Secondly, the replanning makes the system very robust to
+changing objectives and the discovery of new facts about the world,
+both of which are features of our domain.
+
+The remainder of the paper proceeds as follows. We begin by describing
+POMDPs with a propositionally factored representation of states and
+observations. We also outline the relationship between this and a flat
+POMDP representation. Next we describe our DTPDDL language with
+example declarations from a mobile robot exploration task. Then we
+describe the switching planner, including how the sequential domain is
+produced and a plan is generated, and how this then provides the input
+for the decision-theoretic planner. Finally we provide an evaluation
+of the system on the example robot domain, and finish with a short
+survey of related work and some concluding remarks.
 

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
More changes to the introduction

@@
 approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose an approach
-that uses a replanning system to compute a reasonably valuable trace
-in the model, and then uses a decision-theoretic planner on small
-subproblems where reasoning about observations might be useful. We
-refer to the complete system as a {\em switching continual
-planner}. The approach is not optimal, particularly as it is based on
-a satisficing sequential planner, but it performs better than the
-sequential planner alone, and is fast enough to be used in real-time
-on a mobile robot for decision-making.
+planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
+planning approach that uses a classical planning system to compute a
+reasonably valuable trace in the model, and then uses a
+decision-theoretic planner on small subproblems where reasoning about
+observations might be useful. We refer to the complete system as a
+{\em switching continual planner}. The approach is not optimal,
+particularly as it relies on the results of sequential planning
+directly. It does nevertheless perform better than a purely sequential
+replanner, and is fast enough to be used for real-time decision-making
+on a mobile robot.
 
@@
 called the decision-theoretic planning domain definition language
-(DTPDDL). From those descriptions of POMDP, we automatically construct
-a deterministic model for the sequential planner, that includes
-operators which correspond to assumptions about the values of
+(DTPDDL). From those descriptions of POMDPs, we automatically
+construct a deterministic model for the sequential planner. That model
+includes operators which correspond to assumptions about the values of
 imperfectly known state variables. Assumptions scheduled by the
@@
 belief-state space to the decision-theoretic system, and to modify the
-reward function for the decision-theoretic planner, so that it might
-pursue sensing related to those assumptions.
+reward function, so that system might pursue sensing related to those
+assumptions.
 
@@
 
-The sequential planner we use is based
-on \fastdownward \cite{fast-downward}. We add the capability to
-replan, and also allow agent knowledge to be represented explicitly,
-so we can write actions that gain the agent knowledge. LEASE CITE
-MICHAEL's WORK AS THE BASIS FOR THIS. For the decision-theoretic
-planner we have implemented our own forward search in the belief-state
-space.
+%% The sequential planner we use is based
+%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
+%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
+%% knowledge to be represented explicitly, so we can write actions that
+%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
+%% THIS. For the decision-theoretic planner we have implemented our own
+%% forward search in the belief-state space.
 

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
shortened the introduction somewhat

@@
 Execution monitoring tracks properties of the state that determine
-whether a new plan is necessary. For partially observable domains,
-only probabilistic information about such properties will, in general,
-be available.
+whether further planning is required. For partially observable
+domains, only probabilistic information about such properties will, in
+general, be available.
 
@@
 approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose an approach
-that uses a replanning system to compute a reasonably valuable trace
-in the model, and then uses a decision-theoretic planner on small
-subproblems where reasoning about observations might be useful. We
-refer to the complete system as a {\em switching continual
-planner}. The approach is not optimal, particularly as it is based on
-a satisficing sequential planner, but it performs better than the
-sequential planner alone, and is fast enough to be used for real-time
-decision-making on a mobile robot.
+planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
+planning approach that uses a classical planning system to compute a
+reasonably valuable trace in the model, and then uses a
+decision-theoretic planner on small subproblems where reasoning about
+observations might be useful. We refer to the complete system as a
+{\em switching continual planner}. The approach is not optimal,
+particularly as it relies on the results of sequential planning
+directly. It does nevertheless perform better than a purely sequential
+replanner, and is fast enough to be used for real-time decision-making
+on a mobile robot.
+
+%% Switching gives us some important advantages in our robotic
+%% domain. First, the significant source of uncertainty in the domain is
+%% the unreliability of observations, and this approach is particularly
+%% suited to problems that combine a deterministic task planning problem
+%% with observations as the decision-theoretic planner is often invoked
+%% to determine a characteristic of the underlying state. Secondly,
+%% replanning makes our system robust to changing objectives and the
+%% discovery of new facts about the world.
 
@@
 
-The sequential planner we use is based
-on \fastdownward~\cite{fast-downward}. We add the capability to
-replan, and also allow agent knowledge to be represented explicitly,
-so we can write actions that gain the agent knowledge. LEASE CITE
-MICHAEL's WORK AS THE BASIS FOR THIS. For the decision-theoretic
-planner we have implemented our own forward search in the belief-state
-space.
-
-Our switching planner gives us some important advantages in our
-robotic domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations, and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations as the decision-theoretic planner
-is only invoked to determine a characteristic of the underlying
-state. Secondly, the replanning makes the system very robust to
-changing objectives and the discovery of new facts about the world,
-both of which are features of our domain.
+%% The sequential planner we use is based
+%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
+%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
+%% knowledge to be represented explicitly, so we can write actions that
+%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
+%% THIS. For the decision-theoretic planner we have implemented our own
+%% forward search in the belief-state space.
 

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
Trimmed to eight awesome pages!

@@
 problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-language that facilitates factored descriptions of domains and problem
-instances. There are straightforward compilations from problems
-expressed compactly in PPDDL to propositional representations amenable
-to state-of-the-art planning procedures.
+Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
+used in all three of the International Planning Competitions since
+2004. A variation on PDDL for describing domains with stochastic
+actions and uncertain starting configurations, PPDDL is a declarative
+first-order language that facilitates factored descriptions of domains
+and problems. There are straightforward compilations from problems
+expressed in PPDDL to propositional representations amenable
+to state-of-the-art planners.
 
 Because PPDDL cannot model domains that feature partial observability,
-we develop an extension of that language, called Decision-Theoretic
-(DT)PDDL. This can express probabilistic models of the sensing
+we develop an extension we call {\em Decision-Theoretic
+(DT)PDDL}. This can express probabilistic models of the sensing
 consequences of acting, to quantitatively capture unreliability in
@@
 reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, {\em
-reward} is interpreted as accumulated, whereas in DTPDDL it is
-instantaneous.}
+{\em decreased} terms in operator schemata. In that setting, reward is
+accumulated, whereas in DTPDDL it is instantaneous.}
 
-Completing the example, we model the sensing outcome that results from
+Completing the example, we model the sensing that results from
 instantiating the $\pp{look-for-object}$ operator:
@@
 \noindent The interpretation of such an expression can be given
-according to a {\em visitation} of terms. An atom is {\em visited} iff
+by a {\em visitation} of terms: An atom is {\em visited} iff
 its conjunctive parent is visited, and a conjunctive term is visited
@@
 
-From hereon, in order to simplify the discussion (and implementation)
+From hereon, to simplify the discussion (and implementation)
 we shall restrict our attention to POMDPs with deterministic
-actions. It is worthwhile noting that any POMDP with stochastic
-actions can be compiled into an equivalent deterministic-action POMDP,
+actions. We note that POMDPs with stochastic
+actions can be compiled into equivalent deterministic-action POMDPs,
 where all the original action uncertainty is expressed in the

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Trimmed to eight awesome pages!

@@
 observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, produce state spaces far larger than existing
+propositional variables, have state spaces far larger than existing
 optimal or close to optimal planners can handle. For the easier
 problem of planning in completely observable stochastic domains, an
-alternative approach has been to use classical sequential planners and
+alternative has been to use classical sequential planners and
 replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
 Execution monitoring tracks properties of the state that determine
-whether a new plan is necessary. For partially observable domains,
-only probabilistic information about such properties will, in general,
-be available.
+whether further planning is required. For partially observable
+domains, only probabilistic information about such properties will, in
+general, be available.
 
@@
 
+%% Switching gives us some important advantages in our robotic
+%% domain. First, the significant source of uncertainty in the domain is
+%% the unreliability of observations, and this approach is particularly
+%% suited to problems that combine a deterministic task planning problem
+%% with observations as the decision-theoretic planner is often invoked
+%% to determine a characteristic of the underlying state. Secondly,
+%% replanning makes our system robust to changing objectives and the
+%% discovery of new facts about the world.
+
 %% As an illustration, consider a robot searching for a box that it knows
@@
 
-Our switching system is domain independent, taking domain and problem
+Our system is domain independent, taking domain and problem
 descriptions in a first-order declarative language we have developed,
 called the decision-theoretic planning domain definition language
-(DTPDDL). From those descriptions of POMDPs, we automatically
+(DTPDDL). From these descriptions of POMDPs, we automatically
 construct a deterministic model for the sequential planner. That model
-includes operators which correspond to assumptions about the values of
+includes actions which correspond to making assumptions about the values of
 imperfectly known state variables. Assumptions scheduled by the
@@
 
-Our switching planner gives us some important advantages in our
-robotic domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations, and this approach is
+The switching planner provides important advantages in our
+mobile robot domain. First, the significant source of uncertainty in the
+domain is the unreliability of observations and this approach is
 particularly suited to problems that combine a deterministic task
-planning problem with observations as the decision-theoretic planner
-is only invoked to determine a characteristic of the underlying
-state. Secondly, the replanning makes the system very robust to
-changing objectives and the discovery of new facts about the world,
-both of which are features of our domain.
+planning problem with observations because the decision-theoretic planner
+is only invoked to deal with state uncertainty.
+%determine a characteristic of the underlying state. 
+Secondly, replanning makes the system robust to
+changing objectives and discoveries about the world,
+both of which feature in our domain.
 

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Trimmed to eight awesome pages!

@@
 
+\usepackage{tikz}
+\usetikzlibrary{shapes}
+
 \usepackage{times}
@@
 \usepackage{courier}
-
+\usepackage{graphicx}
+\graphicspath{{plots/}}
 \include{macros}%
@@
 for realistic problems.
-
+%%
 We present a \emph{switching planner} that employs fast sequential

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
typo in value function
Trimmed to eight awesome pages!

@@
 dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:1996} $\stochActions$ over
+STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
 state-characterising propositions $\propositions$.
 %%
-Notationally, we say an action $\stochAction \in \stochActions$ is
+We say an action $\stochAction \in \stochActions$ is
 applicable if its precondition $\poss(\stochAction)$, a set of
@@
 
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
+Finally, we make some fairly standard assumptions. First, that
+action execution and sensing occur instantaneously, and that only one
+action can be applied per plan-step. Second, 
+that
+%it can arise that in
+%some propositional states an action is applicable, and in others it is
+%not. Moreover, a belief-state can assign non-zero probability to
+%states where that applicability holds, and states where it does
+%not. We differ
+%%in situations where an action exists that is applicable in some state 
+%% with non-zero probability, and not applicable in another, we differ
+%from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
+%because 
+%%in that
+%we forbid the execution of the action.
+%actions that are not applicable in
+%any state of the current belief.  
+%Also,
 unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, treating an action
+we incorporate a PPDDL-like default semantics, treating an action
 \stochAction\ executed at $\state_i$ as if it has no effect at
@@
 
+% Mucked around with this, then realised the two sentences were
+% contradictory. If we forbid actions applicable in one state but not
+% another in the belief state, then clearly we aren't using a
+% PPDDL-like semantics.
+%% Doubled comments are stuff I added, then removed.
+
 \subsection{Costs, Rewards, and Belief Revision}
 
-Whereas until now we have considered a POMDP model that is factored in
-terms of propositions and percepts, in order to discuss utilities and
-policies it is convenient to consider the underlying decision process
-in a flat format. This underlying decision process is given by the
-tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
+Until now we have discussed the POMDP in terms of propositions and
+percepts. In order to address belief revision and utility it is
+convenient to consider the underlying decision process in a flat
+format. This is given by the tuple
+$\langle \states, \bstate_0, \actions, \transProb, \reward,
+\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
 belief-state, \states\ is the finite set of reachable propositional
@@
 
-In the POMDP model successive state estimation is available by
-application of Bayes' rule.  Taking the current belief $\bstate$ as
-the {\em prior}, and supposing action $\action$ is executed with
-perceptive outcome $\observ$, then the probability that we are in
-$\state$ in the successive belief-state $\bstate'$ is given by:
+Successive state estimation is by application of Bayes'
+rule.  Taking the current belief $\bstate$ as the {\em prior}, and
+supposing action $\action$ is executed with perceptive outcome
+$\observ$, the probability that we are in $\state$ in the
+successive belief-state $\bstate'$ is:
 

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
Trimmed to eight awesome pages!

--- 2010-11-27 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/relatex.tex
@@
-There have been a number of developments recently moving towards
+There have been a number of recent papers on
 planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems.  Notably for example,
+sequential planning in deterministic problems,  for example,
 \system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
@@
 %%suited to POMDPs.
+
+% Normally if we are going to compare with related work, we do
+% actually *compare*. Why didn't you try those approaches? I think
+% it's safe to say that FFR will struggle. Why would it even include
+% in its plan an observational action that doesn't change the world?
+
 %%
@@
 approach assumes that such actions are deterministic and reliable, an
-assumption that we relax in this work.
+assumption that we relax.
 

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Trimmed to eight awesome pages!

@@
 available with semantics that accommodate the above abstraction as
-follows. In the deterministic model a proposition $\prop$ can be
+follows: In the deterministic model a proposition $\prop$ can be
 thought to have a ternary interpretation at a state \state, as either
@@
 %%
-Lastly, with regards to switching, we have that if $\action$ has a
+Lastly, we have that if $\action$ has a
 precondition that is not true with probability $1$ in the underlying

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
OKay

@@
-We describe our continual planning approach that switches between
-sequential and contingent sessions. It proceeds by interleaved
-planning and execution in a deterministic-action POMDP described in
-DTPDDL. During a sequential session, planning is performed by a
-``classical'' system,\footnote{That is, a planner designed to solve
-fully observable deterministic tasks.}  and execution proceeds
-according to the {\em trace} computed by that system. Taking the form
-of a classical plan, the trace specifies a sequence of POMDP actions
-that achieve the agent's objectives in a deterministic approximation,
-i.e., {\em determinisation}, of the problem at hand. More precisely,
-the trace is an interleaved sequence of POMDP actions and {\em
-assumptive} actions. The latter correspond to assumptions the planner
-makes about the truth value of propositions -- e.g. that a box of
-cornflakes is located in the kitchen at the third plan step. Another
-kind of assumption, called an {\em applicability} assumption, is made
-when the trace includes an action $\action$ that is not applicable
-with probability $1$ at the belief-state
+We describe our approach that switches between sequential and
+contingent sessions. As a continual planning approach, it proceeds by
+interleaving planning and execution in a deterministic-action POMDP
+described in DTPDDL. During a sequential session, planning is
+performed by a ``classical'' system,\footnote{That is, a planner
+designed to solve fully observable deterministic tasks.}  and
+execution proceeds according to the {\em trace} computed by that
+system. Taking the form of a classical plan, the trace specifies a
+sequence of POMDP actions that achieve the agent's objectives in a
+deterministic approximation, i.e., {\em determinisation}, of the
+problem at hand. More precisely, the trace is an interleaved sequence
+of POMDP actions and {\em assumptive} actions. The latter correspond
+to assumptions the planner makes about the truth value of propositions
+-- e.g. that a box of cornflakes is located in the kitchen at the
+third plan step. These are called {\em applicability} assumptions if
+the trace includes an action $\action$ that is not applicable with
+probability $1$ at the belief-state
 \bstate\ that the system is projected to be in when \action\ is
@@
 tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. For each assumption we add a {\em
-disconfirm} action to the POMDP whose execution is rewarding
-(resp. costly) if the assumption is false (resp. true). We also add
-one {\em confirm} action that is rewarding (resp. costly) to execute
-if $\poss(\switchAction)$ is true (resp. false). If execution of the
-contingent plan applies a disconfirm action, then a new sequential
-session begins at the {\em underlying} belief-state. If the confirm
-action is executed, the sequential session is resumed, and
-\switchAction\ is executed. Otherwise, control rests in the continual
-session.
+assumptions made in the trace hold. 
+%%
+We add {\em disconfirm} and {\em confirm} actions to the model, so
+that session is encouraged to judge the assumptions made in the
+trace. On execution of one of those actions, control is returned to
+the sequential session that continues at the {\em underlying}
+belief-state.
+%%
+%% For each assumption we add a {\em
+%% disconfirm} action to the POMDP whose execution is rewarding
+%% (resp. costly) if the assumption is false (resp. true). We also add
+%% one {\em confirm} action that is rewarding (resp. costly) to execute
+%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
+%% contingent plan applies a disconfirm action, then a new sequential
+%% session begins at the {\em underlying} belief-state. If the confirm
+%% action is executed, the sequential session is resumed, and
+%% \switchAction\ is executed. Otherwise, control rests in the continual
+%% session.
 %%
@@
 determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not featured in the trace, and
+first omitting all propositions that do not feature in the trace, and
 by then iteratively refining the model while the result is of a
@@
 underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration in DTPDDL. Our approach uses that
-distribution: (1) as the source of candidate determinisations for
-sequential planning, (2) to determine when to switch, and (3) as a
-mechanism to guide construction of an abstract process for contingent
-sessions.
+$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
+as the source of candidate determinisations for sequential planning,
+(2) to determine when to switch, and (3) as a mechanism to guide
+construction of an abstract process for contingent sessions.
 
@@
 $j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct,
-or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
+either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
+conjunct, or it occurs in $T_k$ for some executed
+$\assumptiveS{k}$. Also, a distinct assumption about a proposition
+$\prop$ can only be made once, and must be sequenced before any POMDP
+actions that have $\prop$ in their precondition or effect.
+%%
+In other words, an assumptive action
+must be applied to make a truth assignment to \prop, and then that
+truth assignment can be modified by the DTPDDL domain operators.
+%%
 %%
@@
 $\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ and conflicting
-propositions from $s$ deleted. For example, if the action adds
-$(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must delete
+the union of $\state$ with atomic terms from $T_i$ and the proposition
+denoting the undefined fact is deleted. For example, if the action
+adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must delete
 $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added sink,
-where $\forall\action$ $\reward(\state^\bot, \action) = 0$
-and $\transProb(\state^\bot, \action, \state^\bot) = 1$.
+where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
+$\transProb(\state^\bot, \action, \state^\bot) = 1$.
+%%
+Lastly, we have that if $\action$ has a precondition that is not true
+with probability $\theta$, a given threshold, in the underlying belief
+$\bstate$, then immediately before it is executed blindly, the system
+switches to a contingent session, discussed in the following section.
 
-For sequential planning the operators from the POMDP model are
-available with semantics that accommodate the above abstraction as
-follows: In the deterministic model a proposition $\prop$ can be
-thought to have a ternary interpretation at a state \state, as either
-{\em true}, written $\prop\in\state$, {\em false}, written
-$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-$\action\in\actions$ in the deterministic model, if
-$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-or negative effect of $\action$, then $\action$ is not applicable
-in \state\ if $\prop\#\in\state$. In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment could be modified by the DTPDDL domain operators.
-%%
-Lastly, we have that if $\action$ has a
-precondition that is not true with probability $1$ in the underlying
-belief $\bstate$, then immediately before it is executed blindly, the
-system switches to a contingent session, discussed in the following
-section.
+%% For sequential planning the operators from the POMDP model are
+%% available with semantics that accommodate the above abstraction as
+%% follows: In the deterministic model a proposition $\prop$ can be
+%% thought to have a ternary interpretation at a state \state, as either
+%% {\em true}, written $\prop\in\state$, {\em false}, written
+%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
+%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
+%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
+%% $\action\in\actions$ in the deterministic model, if
+%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
+%% or negative effect of $\action$, then $\action$ is not applicable
+%% in \state\ if $\prop\#\in\state$.
 

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
OKay

--- 2010-11-24 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
@@
 problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
-all 3 of the International Planning Competitions since 2004. A
-variation on PDDL for describing domains with stochastic actions and
-uncertain starting configurations, PPDDL is a declarative first-order
-language that facilitates factored description of domains and problem
-instances. There are straightforward compilations from problems
-expressed compactly in PPDDL to propositional representations amenable
-to state-of-the-art planning procedures.
+Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
+used in all three of the International Planning Competitions since
+2004. A variation on PDDL for describing domains with stochastic
+actions and uncertain starting configurations, PPDDL is a declarative
+first-order language that facilitates factored descriptions of domains
+and problems. There are straightforward compilations from problems
+expressed in PPDDL to propositional representations amenable
+to state-of-the-art planners.
 
 Because PPDDL cannot model domains that feature partial observability,
-we developed an extension of that language, called DTPDDL. This can
-express probabilistic models of the sensing consequences of acting, to
-quantitatively capture unreliability in perception. That expressive
-power is achieved by incorporating perceptual analogues of fluent,
-predicate, and action definitions. In detail, we have declarations of
-state characterising predicate and fluent symbols according to the
-PPDDL syntax. In addition, we allow two other declarations, for
-perceptual predicates and fluents respectively. For example, suppose
-our robot is tasked with exploring {\em locations} in order to
-identify the whereabouts of a {\em visual-object}. We must describe
-state and perceptual facts that model the {\em true}, resp. perceived,
-locations of objects. In DTPDDL, these declarations appear as:
+we develop an extension we call {\em Decision-Theoretic
+(DT)PDDL}. This can express probabilistic models of the sensing
+consequences of acting, to quantitatively capture unreliability in
+perception. That expressive power is achieved by incorporating
+perceptual analogues of fluent, predicate, and action definitions. In
+detail, we have declarations of state characterising predicate and
+fluent symbols according to the PPDDL syntax. In addition, we allow
+two other declarations, for perceptual predicates and fluents
+respectively. For example, suppose our robot is tasked with exploring
+{\em locations} in order to identify the whereabouts of a {\em
+visual-object}. We must describe state and perceptual facts that model
+the {\em true}, resp. perceived, locations of objects. In DTPDDL,
+these declarations appear as:
 
@@
 \begin{tabtt}
-(\=:action look-for-object \+ \\
+(\=:action object-search \+ \\
    :parameters (\=?r - robot ?v - visual-object\\
@@
 \noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
+$\pp{object-search}$ has an effect on the state. That is, its
 execution incurs an instantaneous penalty, $3$, that corresponds to
@@
 reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, {\em
-reward} is interpreted as accumulated (resp. instantaneous) rewards
-and penalties.}
+{\em decreased} terms in operator schemata. In that setting, reward is
+accumulated, whereas in DTPDDL it is instantaneous.}
 
-Completing the example, we model the sensing outcome that results from
-instantiating the $\pp{look-for-object}$ operator:
+Completing the example, we model the sensing that results from
+instantiating the $\pp{object-search}$ operator:
 
@@
  \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
+ :execution \> ( \> object-search ?r ?v ?l) \\
  :precondition (and (= (is-in ?r) ?l) ) \\
@@
 \noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- I.e,
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office}) =
-\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
+the sense action-precondition -- i.e,
+$(\pp{object-search}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
+$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
 we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
 order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
+if action $(\pp{object-search}~\pp{DORA}~\pp{box}~\pp{office})$ is
 executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ of perceiving it.
+the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
 
@@
 tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$,
-(2) probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
-where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
-atomic terms. The root term is always conjunctive, and the leaves are
-atomic. For example, the starting distribution for $\pp{DORA}$ might
-be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
-expresses the conjunctive root of the tree -- I.e., the root node
-$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
-$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
-subterm.}
+state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
+probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
+(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
+probabilistic and atomic terms. The root term is always conjunctive,
+and the leaves are atomic. For example, the starting distribution for
+$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
+$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
+i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
+$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
+that contain a single atomic subterm.}
 
@@
 according to this recursive definition encapsulates a starting state,
-along with the probability it occurs. The former corresponds to the
+along with the probability that it occurs. The former corresponds to the
 union of all visited atoms, and the latter corresponds to the product
-of $p_i$ entries on the visited subterms of probabilistic
+of $\prob_i$ entries on the visited subterms of probabilistic
 elements. Making these ideas concrete, our example yields the
@@
 \normalsize
+
+From hereon, to simplify the discussion (and implementation)
+we shall restrict our attention to POMDPs with deterministic
+actions. We note that POMDPs with stochastic
+actions can be compiled into equivalent deterministic-action POMDPs,
+where all the original action uncertainty is expressed in the
+starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
+that of finite-horizon planning, such a compilation yields a finite
+flat representation of the original POMDP, only with deterministic
+actions.

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
back to look-for-object

@@
 
-In our implementation uses the
-~\citeauthor{king:2009}~(\citeyear{king:2009}) \system{Dlib-ml} for
-successive estimation of the underlying belief-state.
+We have implemented our switching continual approach in the MAPSIM
+environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
+to use several underlying planning systems. We have also extended
+MAPSIM system so that it can parse DTPDDL, and perform successive
+estimation of the underlying belief-state using
+\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
+we use our own version of Fast Downward~\cite{fast-downward} for
+sequential sessions. We have extended that system to support actions
+with success probabilities. In our evaluation Fast Downward is run
+with the cyclic causal graph heuristic using an A* search or weighted
+A* (with weight 5). We use A* for easy problems and WA* for more
+difficult problems where A* is ineffective. Contingent sessions use
+our own forward search procedure.
+
+\Omit{ We then perform multiple tests with different limits for the
+  belief-space size in contingent sessions.  Higher limits should
+  cause longer planning times but be beneficial to plan quality as
+  more contingencies can be taken in to account by the POMDP planner.
+}
+
+In order to evaluate our system, we have also implemented a baseline
+approach in MAPSIM. Rather than invoking a contingent session when a
+switching action $\switchAction$ is scheduled for execution, the
+baseline simply executes one action that can trigger a sensing outcome
+determined by the precondition of $\switchAction$, and replans in the
+resulting belief-state.
+
+%% executes switching actions and replans in the
+%% resulting belief state.
+
+%% but instead of
+%% creating an observation problem for the decision theoretic planner it
+%% will just execute one sensing action -- assuming that this action will
+%% confirm its assumption.
+
+To test our approach we use a robot exploration domain based on a
+scenario from our physical robotic system. Here, a robot is
+exploring an office or living environment, and trying to report the
+locations of objects to their owner.
+
+\begin{figure}[h]
+  \centering
+  \input{figures/dora}
+  \caption{A exploration domain with three rooms and 13 places. The
+    robot is in in the centre of room A}
+\label{fig:dora2}
+\end{figure}
+An instance of this domain is shown in figure \ref{fig:dora2}. Basic
+types of element are {\tt rooms}, {\tt places} and {\tt
+  objects}. Places are topological map nodes that occur in rooms.
+Objects, including the robot, are always located at a place. The robot
+can move around the rooms via connections between places given by the
+{\tt connected} predicate. Each room has a, possibly unknown, {\em
+  category} (e.g. kitchen, office, living room) and depending on this
+category, objects are placed inside the rooms.
+
+The robot can find out if an object is at a certain place by executing
+the {\tt look-for-object} action. Moreover, the contents of a room
+indicates its category, and certain objects are more likely to be
+located in rooms of certain categories. For example, a box of
+cornflakes is more likely to be located in the kitchen than the
+office. The {\tt look-for-object} might result in a positive
+perception if the object is there. An instantaneous positive detection
+of the object is not proof positive that it is there. Also, some
+objects are harder to detect than others. Also, Additionally, if the
+robot is in the presence of a human, it can ask what type of room they
+are currently in, however conducting a dialogue is more costly than
+running a vision algorithm (cost of 8 vs costs of 3).
+
+We perform experiments on tasks of several sizes. The robot's
+objective is to find one or more objects and report back their
+position to a person. In order to determine the impact of sensor
+reliability in a comparison with the baseline, we also run tests on
+identical problems but changed the sensor model for the target
+objects: the probability for perceiving the object if it was there
+went from 0.9 in the easiest case to 0.65 and 0.4 in the average and
+hard cases respectively. In addition, we test how our system performs
+in tasks requiring indirect sensing. Here, we gave the planner rewards
+for visiting a certain type of room (e.g. kitchen or office). We did
+not place any humans in those scenarios, so the only way of
+determining the room category is by looking for objects typical for
+that room type. As this indirect reasoning is a tasks that cannot be
+performed by the continual planner alone, we did not run the baseline
+system on those problems.
+
+In our evaluation we run 50 simulations in each configuration. A run
+timeouts after 30 minutes (1800 seconds). The planning and simulation
+models are in exact correspondence. Not all problems have a solution,
+so we also examine whether or not detection of the non-existence of a
+solution is important in stochastic domains.
+%% 
+We perform multiple tests with different limits for the
+belief-space size in contingent sessions.  Higher limits result in
+longer contingent planning times and higher plan quality.
+
+Using a satisficing, optimistic serial planner in sequential sessions
+makes optimising for the expected reward difficult. Only reasoning
+about single traces, it is optimistic wrt the remaining costs to a
+goal and pessimistic regarding the probability of reaching tat. By
+setting the rewards to non-extremal values has little effect on that
+system. Therefore, we do not consider expected utility in our
+experiments. Instead, we report average costs of the plans and the
+success rate (as a ratio between solved and solvable problems).
+
+\begin{figure}[h!]
+  % \centering
+  % \includegraphics{dora1-time}\hfill
+  % \vspace{2mm}
+  \includegraphics{dora2-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora3-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora4-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora56-time}\hfill
+  \vspace{2mm}
+  \includegraphics{dora-cat-time}\hfill
+  \caption{Average runtime}
+  \label{fig:results-time}
+\end{figure}
+
+\begin{figure}[h!]
+  % \centering
+  % \includegraphics{dora1-quality}\hfill
+  % \vspace{2mm}
+  \includegraphics{dora2-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora3-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora4-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora56-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{dora-cat-quality}\hfill
+  \caption{Average plan costs and number of successful runs.}
+  \label{fig:results-quality}
+\end{figure}
+
+%% Figure \ref{fig:results-quality} shows the average costs of the
+%% executed plans as well as the percentage of solvable tasks that were
+%% actually solved by the planner.
+
+Our experimental results are summarised in
+Figure~\ref{fig:results-time} and Figure~\ref{fig:results-quality}. We
+find that the time spent on contingent planning increases steeply as
+the abstraction $\bstate_0$ becomes more refined. For coarse initial
+configurations, the cost of contingent planning is compensated by the
+decrease of time spent in Fast Downward. Addressing success ratios and
+plan costs, for objects that can be easily detected there is little
+gain in using a decision-theoretic planner, as the greedy sensing
+approach by the baseline is sufficient. With decreasing sensor
+reliability contingent planning pays off.%% : while the resulting plans are still
+%% longer on average, the impact on the number of solved tasks was much
+%% smaller than for the baseline system.  Less aggressive abstraction
+of the initial configuration results in longer runtimes. In our
+scenario it is worthwhile being assumptively aggressive.  We suppose
+the relatively high success rate irrespective of the coarseness of the
+initial configuration indicates the effectiveness of using conditional
+entropy to guide abstraction refinement.
+
+We find that abstraction has little impact on plan
+costs and success rate in our scenario. Increasing the size of the
+initial abstract belief beyond 50 states rarely pays off, because
+while additional information may facilitate better plans, the indirect
+sensing tests report the same result. Here, we only really see an
+impact in terms of the runtime of contingent planning.
+
+% We believe that a part of the improvement is due to the segmentation of
+% the plan into several subtask, essentially performing hierarchical
+% planning. Especially when the continual planner performs badly this is
+% a huge gain.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "moritz_2011"
+%%% End:

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
OKay
corrections

@@
 An instance of this domain is shown in figure \ref{fig:dora2}. Basic
-types of element are{\tt rooms}, {\tt places} and {\tt
+types of element are {\tt rooms}, {\tt places} and {\tt
   objects}. Places are topological map nodes that occur in rooms.
@@
 
-The robot can find out if an object is at a certain place by executing
-the {\tt look-for-object} action. Moreover, the contents of a room
-indicates its category, and certain objects are more likely to be
-located in rooms of certain categories. For example, a box of
-cornflakes is more likely to be located in the kitchen than the
-office. The {\tt look-for-object} might result in a positive
-perception if the object is there. An instantaneous positive detection
-of the object is not proof positive that it is there. Also, some
-objects are harder to detect than others. Also, Additionally, if the
-robot is in the presence of a human, it can ask what type of room they
-are currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
+The robot can look for an object at a place by executing a {\tt
+  look-for-object} action. Moreover, the contents of a room indicate
+its category, because certain objects are more likely to be located in
+rooms of particular categories. For example, a box of cornflakes is
+more likely to be located in the kitchen than the office. Executing
+{\tt look-for-object} results in a noisy sensing outcome, sometimes
+indicating a positive perception if the object is there. Here, some
+objects are harder to detect than others. Lastly, if the robot is in
+the presence of a human, it can ask what type of room it is currently
+in, however conducting a dialogue is more costly than running a vision
+algorithm (cost of 8 vs costs of 3).
 
-We perform experiments on tasks of several sizes. The robot's
-objective is to find one or more objects and report back their
-position to a person. In order to determine the impact of sensor
-reliability in a comparison with the baseline, we also run tests on
-identical problems but changed the sensor model for the target
-objects: the probability for perceiving the object if it was there
-went from 0.9 in the easiest case to 0.65 and 0.4 in the average and
-hard cases respectively. In addition, we test how our system performs
-in tasks requiring indirect sensing. Here, we gave the planner rewards
-for visiting a certain type of room (e.g. kitchen or office). We did
-not place any humans in those scenarios, so the only way of
-determining the room category is by looking for objects typical for
-that room type. As this indirect reasoning is a tasks that cannot be
-performed by the continual planner alone, we did not run the baseline
-system on those problems.
+%%  Therefore, an instantaneous positive detection of an object is
+%% not proof positive of it being there.
 
-In our evaluation we run 50 simulations in each configuration. The
-planning and simulation models are in exact correspondence. Not all
-problems have a solution, so we also examine whether or not detection
-of the non-existence of a solution is important in stochastic domains.
+We perform experiments on tasks of several sizes. The objective is to
+find one or more objects and report their position to the user. In
+order to determine the impact of the reliability of sensing in a
+comparison with the baseline, we run tests on identical problems but
+changed the sensor model for the target objects: the probability for
+perceiving the object if it was there went from 0.9 in the easiest
+case to 0.65 and 0.4 in the average and hard cases respectively. In
+addition, we test how our system performs in tasks requiring indirect
+sensing. Here, we gave the planner rewards for visiting a certain type
+of room (e.g. kitchen or office). We did not place any humans in those
+scenarios, so the only way of determining the room category is by
+looking for objects typical for that room type. As this indirect
+reasoning is a tasks that cannot be performed by the continual planner
+alone, we did not run the baseline system on those problems.
+
+In our evaluation we run 50 simulations in each configuration. A run
+timeouts after 30 minutes (1800 seconds). The planning and simulation
+models are in exact correspondence. Not all problems have a solution,
+so we also examine whether or not detection of the non-existence of a
+solution is important in stochastic domains.
 %% 
@@
   \includegraphics{dora-cat-quality}\hfill
-  \caption{Average plan costs and number of successful runs}
+  \caption{Average plan costs and number of successful runs.}
   \label{fig:results-quality}

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
corrections

@@
-We describe our continual planning approach that switches between
-sequential and contingent sessions. This proceeds by interleaved
-planning and execution in a deterministic-action POMDP described in
-DTPDDL. During a sequential session, planning is performed by a
-``classical'' system,\footnote{That is, a system designed to solve
-fully observable deterministic tasks.}  and execution proceeds
-according to the {\em trace} computed by that system. Taking the form
-of a classical plan, the trace specifies a sequence of POMDP
-actions that achieve the agent's objectives in a deterministic
-approximation, i.e., {\em determinisation}, of the problem at
-hand. More precisely, the trace is an interleaved sequence of POMDP
-actions and {\em assumptive} actions. The latter correspond to
-assumptions the planner makes about the truth value of {\em runtime
-state variables} -- e.g. that a box of cornflakes is located in the
-kitchen at the third plan step. Another kind of assumption, called an
-{\em applicability} assumption, is made when the trace includes an
-action $\action$ that is not applicable with probability $1$ at the
-belief-state
+We describe our approach that switches between sequential and
+contingent sessions. As a continual planning approach, it proceeds by
+interleaving planning and execution in a deterministic-action POMDP
+described in DTPDDL. During a sequential session, planning is
+performed by a ``classical'' system,\footnote{That is, a planner
+designed to solve fully observable deterministic tasks.}  and
+execution proceeds according to the {\em trace} computed by that
+system. Taking the form of a classical plan, the trace specifies a
+sequence of POMDP actions that achieve the agent's objectives in a
+deterministic approximation, i.e., {\em determinisation}, of the
+problem at hand. More precisely, the trace is an interleaved sequence
+of POMDP actions and {\em assumptive} actions. The latter correspond
+to assumptions the planner makes about the truth value of propositions
+-- e.g. that a box of cornflakes is located in the kitchen at the
+third plan step. They are called {\em applicability} assumptions if
+the trace includes an action $\action$ that is not applicable with
+probability $1$ at the belief-state
 \bstate\ that the system is projected to be in when \action\ is
@@
 Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed until the applicability of the
-next scheduled action is uncertain.  A contingent session then begins,
-which tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. For each {\em active} assumption
-in the trace, we add a {\em disconfirm} action to the POMDP whose
-execution is rewarding (resp. costly) if the assumption is false
-(resp. true). We also add one {\em confirm} action that is rewarding
-(resp. costly) to execute if $\poss(\action)$ is true
-(resp. false). If execution of the contingent plan applies a
-disconfirm action, then a new sequential session begins at the current
-underlying belief-state. If the confirm action is executed, the
-sequential session is resumed, and
-\switchAction\ is executed. Otherwise, control rests in the continual
-session.
-
+actions from the trace are executed in sequence until the
+applicability of the next scheduled action is uncertain. We denote
+that action \switchAction.  A contingent session then begins, that
+tailors sensory processing by the agent to determine whether the
+assumptions made in the trace hold. 
+%%
+We add {\em disconfirm} and {\em confirm} actions to the model, so
+that session is encouraged to judge the assumptions made in the
+trace. On execution of one of those actions, control is returned to
+the sequential session that continues at the {\em underlying}
+belief-state.
+%%
+%% For each assumption we add a {\em
+%% disconfirm} action to the POMDP whose execution is rewarding
+%% (resp. costly) if the assumption is false (resp. true). We also add
+%% one {\em confirm} action that is rewarding (resp. costly) to execute
+%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
+%% contingent plan applies a disconfirm action, then a new sequential
+%% session begins at the {\em underlying} belief-state. If the confirm
+%% action is executed, the sequential session is resumed, and
+%% \switchAction\ is executed. Otherwise, control rests in the continual
+%% session.
+%%
 Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions proceed in an abstract decision process
-determined by the current trace. That abstract process is constructed
-by first omitting all variables from the model that are not featured
-in the trace, and by then iteratively refining that abstraction while
-the resulting model is of a practicable size.
+POMDPs, contingent sessions plan in an abstract decision process
+determined by the current trace. This abstraction is constructed by
+first omitting all propositions that do not feature in the trace, and
+by then iteratively refining the model while the result is of a
+practicable size.
 
@@
 continual planner maintains a factored representation of successive
-{\em belief-states}, by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of that
-distribution corresponds more-or-less directly to an $(\pp{:init})$
-declaration in DTPDDL. Our approach uses that {\em underlying}
-belief-state as: (1) the source of determinisations for sequential
-planning, (2) the basis for switching, and (3) the source for state
-abstraction in contingent sessions.
+belief-states, by performing belief revision according to
+Eq~\ref{eq:revision}. Moreover, our internal representation of the
+underlying belief-state corresponds more-or-less directly to an
+$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
+as the source of candidate determinisations for sequential planning,
+(2) to determine when to switch, and (3) as a mechanism to guide
+construction of an abstract process for contingent sessions.
 
@@
 latter encapsulates assumptions about: (1) the true underlying state,
-(2) how the plan's execution will progress, and (3) the possibility of
-the agent eventually holding strong beliefs about the truth values of
-specific state propositions. Here we describe the deterministic
-planning problem, derived from the DTPDDL model, that admits plans
-which correspond to traces.
+(2) how execution will progress, and (3) the possibility of the agent
+eventually holding strong beliefs about the truth values of specific
+state propositions. Here we describe the deterministic planning
+problem, derived from the DTPDDL model, that admits plans which
+correspond to traces.
 
 In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ DTPDDL declaration. Our
-approach uses the structure of that, as it occurs in the problem
-description, to define a set of state-assumptions available to
-sequential planning.
+action is expressed in the $(\pp{:init})$ declaration. Our approach
+uses the structure of that, as it occurs in the problem description,
+to define a set of state-assumptions available to sequential planning.
 %%
@@
 $j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
-or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
+either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
+conjunct, or it occurs in $T_k$ for some executed
+$\assumptiveS{k}$. Also, an assumption about a proposition
+$\prop$ can only be made once, and must be sequenced before any POMDP
+actions that have $\prop$ in their precondition or effect.
 %%
-Executing $\assumptiveS{i}$ in $s$ effects a transition to a successor
-state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with probability
-$1 - p_i$. Here, $s^{T_i}$ is the union of $s$ with atomic terms from
-$T_i$ and conflicting propositions from $s$ deleted. For example, if
+In other words, an assumptive action
+must be applied to make a truth assignment to \prop, and then that
+truth assignment can be modified by the DTPDDL domain operators.
+%%
+%%
+Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
+successor state $\state^{T_i}$ with probability $\prob_i$, and
+$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
+the union of $\state$ with atomic terms from $T_i$ with the
+proposition denoting the undefined fact is deleted. For example, if
 the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is an added sink,
-where $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$
-and $\transProb(\state^\bot, \action, \state^\bot) = 1$.
+delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
+sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
+$\transProb(\state^\bot, \action, \state^\bot) = 1$.
+%%
+Lastly, we have that if $\action$ has a precondition that is not true
+with probability $\theta$, a given threshold, in the underlying belief
+$\bstate$, then immediately before it is executed blindly the system
+switches to a contingent session (described below).
 
-For sequential planning the operators from the POMDP model are
-available with semantics that accommodate the above abstraction as
-follows. In the deterministic model a proposition $\prop$ can be
-thought to have a ternary interpretation at a state \state, as either
-{\em true}, written $\prop\in\state$, {\em false}, written
-$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-$\action\in\actions$ in the deterministic model, if
-$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-or negative effect of $\action$, then $\action$ is not applicable
-in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment could be modified by POMDP operators.
+%% For sequential planning the operators from the POMDP model are
+%% available with semantics that accommodate the above abstraction as
+%% follows: In the deterministic model a proposition $\prop$ can be
+%% thought to have a ternary interpretation at a state \state, as either
+%% {\em true}, written $\prop\in\state$, {\em false}, written
+%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
+%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
+%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
+%% $\action\in\actions$ in the deterministic model, if
+%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
+%% or negative effect of $\action$, then $\action$ is not applicable
+%% in \state\ if $\prop\#\in\state$.
 
-Addressing now the switching semantics of action execution. If
-$\action\in\actions$ has a precondition that is not true with
-probability $1$ in the underlying belief $\bstate$, then before it is
-executed blindly, the system switches to a contingent session,
-discussed in the following section. In detail, we halt the sequential
-session at \action\ if $\exists
-\state$ s.t. $\bstate(\state)>0$, 
-$\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-semantics of action execution there must be a $\state'$ s.t.
-$\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-not have been scheduled at \bstate.
+%% Addressing now the switching semantics of action execution. I
+%% %% In detail, we halt the sequential
+%% session at \action\ if $\exists
+%% \state$ s.t. $\bstate(\state)>0$, 
+%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
+%% semantics of action execution there must be a $\state'$ s.t.
+%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
+%% not have been scheduled at \bstate.
 
-then $\reward(\state,\action)=10$, and
-otherwise if $\prop\not\in\state'$ then
-$\reward(\state',\action)=-10$.
+%% then $\reward(\state,\action)=10$, and
+%% otherwise if $\prop\not\in\state'$ then
+%% $\reward(\state',\action)=-10$.
 
@@
 
-\noindent which is equal to the maximal additive contribution a trace of the plan can
-make to Eq~\ref{eq:expectedvalue}, taking an FSC that corresponds to
-the sequential plan.  In goal directed problems, where the only
+\noindent which is equal to the maximal 
+additive contribution a trace of the plan can make to
+Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
 non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour
-of \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
+the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
+by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
 gradient-based reinforcement learning systems, such as William's
@@
 \$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
--\$(T_i) & \pp{otherwise} \\
+\hat\$(T_i) & \pp{otherwise} \\
 \end{array}
@@
 
-\noindent Where $\$(T_i)$ is a positive numeric quantity which
-captures the reward the agent receives for correctly rejecting a
-false assumption. The added dual actions are propositionally trivial
--- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
+\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
+(negative) numeric quantity which captures the utility the agent
+receives for correctly (incorrectly) rejecting a false assumption. The
+added dual actions are propositionally trivial -- i.e.,
+$\forall\state\in\states$ $\poss(\assumptiveDT{i})
 \subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
 %%
-In order to reduce the number of possible rejections we do not
-consider assumptions that are not {\em active} with respect to the
-action, $\action_i$, whose scheduled execution switch the system into
-the contingent session. A assumption $\assumptiveS{i}$ is active if
-$\action_i$ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix to $\action_i$. For
-example, if $\pp{DORA}$'s trace is:
+In order to reduce the number of possible rejections we omit
+assumptions that are not {\em active} with respect to the action,
+\switchAction, whose scheduled execution switched the system into
+the contingent session. An assumption $\assumptiveS{i}$ is active if
+\switchAction\ is not applicable in the sequential model unless
+$\assumptiveS{i}$ is scheduled in the trace prefix
+to \switchAction. For example, if $\pp{DORA}$'s trace is:
 
@@
 
-Again taking $\action_i$ to be the switching action, we include
-another propositionally trivial action $\actions.\poss(\action_i)$
-with the reward property:
+We also include another propositionally trivial action
+$\actions.\poss(\switchAction)$ with the reward property:
 
 \[
-\reward(\state, \actions.\poss(\action_i)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\action_i)) & \pp{if}~\;\; \poss(\action_i) \subseteq \state \\
--\$(\poss(\action_i)) & \pp{otherwise} \\
+\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
+\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
+\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
 \end{array}
@@
 action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\action_i)$, returns control to a sequential
+$\actions.\poss(\switchAction)$, returns control to a sequential
 session. If a dual is executed, then the sequential planner must
 replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\action_i)$ is executed,
-then the current sequential plan is executed until further sensing is
-scheduled, or to completion. In our project the rewards, i.e.,
-function $\$:2^\propositions\to\Re$, for the POMDP posed to the
+belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
+executed, then the current sequential plan is executed until further
+sensing is scheduled, or to completion. Lastly, we note that in our
+project the rewards, i.e., functions
+$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
 contingent session are sourced from a motivational component of our
-robotic architecture.
+robotic architecture. For this paper we take $\hat\$(x)= - \$(x)(1 -
+p) / p$, where $\$(x)$ is a given constant.
+
+%%  We make no assumptions here about how they are
+%% derived.
 
@@
 adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states that occur with non-zero probability
-according to that refined declaration remains of a practicable
-size. In detail, for each excluded atomic term, we compute the {\em
-entropy} of the corresponding proposition, {\em conditional} on the
-active assumptions. The {\em conditional entropy} measure
-is higher the more uncertain that proposition is given truth
-assignments to propositions that are the subject of active
-assumptions. Propositions that in the first place were excluded are
-iteratively added to the problem posed to the conditional session in
-increasing order according to that measure.
+number of abstract states in $\bstate_0$ that occur with non-zero
+probability according to that refined declaration remains of a
+practicable size. In detail, for each excluded atomic term, we compute
+the {\em entropy} of the active assumptions, {\em conditional} on the
+corresponding proposition. Loosely speaking, the {\em conditional
+entropy} measure is higher the more uncertain those assumptions are
+given truth assignment to the proposition. Propositions that in the
+first place were excluded are iteratively added to the problem posed
+to the conditional session in increasing order according to that
+measure.
+
+%% In detail, for each excluded atomic term, we compute the {\em
+%% entropy} of the corresponding proposition, {\em conditional} on the
+%% active assumptions. The {\em conditional entropy} measure
+%% is higher the more uncertain that proposition is given truth
+%% assignments to propositions that are the subject of active
+%% assumptions. Propositions that in the first place were excluded are
+%% iteratively added to the problem posed to the conditional session in
+%% increasing order according to that measure.
 

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
paper updates

@@
 
-\begin{figure}[h]
-  \centering
-  \input{figures/dora}
-  \caption{A exploration domain with three rooms and 13 places. The
-    robot is in in the centre of room A}
-\label{fig:dora2}
-\end{figure}
-An instance of this domain is shown in figure \ref{fig:dora2}. Basic
-types of element are {\tt rooms}, {\tt places} and {\tt
+% \begin{figure}[h]
+%   \centering
+%   \input{figures/dora}
+%   \caption{A exploration domain with three rooms and 13 places. The
+%     robot is in in the centre of room A}
+% \label{fig:dora2}
+% \end{figure}
+% An instance of this domain is shown in figure \ref{fig:dora2}. 
+Basic types of element are {\tt rooms}, {\tt places} and {\tt
   objects}. Places are topological map nodes that occur in rooms.
@@
 
-The robot can find out if an object is at a certain place by executing
-the {\tt look-for-object} action. Moreover, the contents of a room
-indicates its category, and certain objects are more likely to be
-located in rooms of certain categories. For example, a box of
-cornflakes is more likely to be located in the kitchen than the
-office. The {\tt look-for-object} might result in a positive
-perception if the object is there. An instantaneous positive detection
-of the object is not proof positive that it is there. Also, some
-objects are harder to detect than others. Also, Additionally, if the
-robot is in the presence of a human, it can ask what type of room they
-are currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
+The robot can look for an object at a place by executing a {\tt
+  look-for-object} action. Moreover, the contents of a room indicate
+its category, because certain objects are more likely to be located in
+rooms of particular categories. For example, a box of cornflakes is
+more likely to be located in the kitchen than the office. Executing
+{\tt look-for-object} results in a noisy sensing outcome, sometimes
+indicating a positive perception if the object is there. Here, some
+objects are harder to detect than others. Lastly, if the robot is in
+the presence of a human, it can ask what type of room it is currently
+in, however conducting a dialogue is more costly than running a vision
+algorithm (cost of 8 vs costs of 3).
 
-We perform experiments on tasks of several sizes. The robot's
-objective is to find one or more objects and report back their
-position to a person. In order to determine the impact of sensor
-reliability in a comparison with the baseline, we also run tests on
-identical problems but changed the sensor model for the target
-objects: the probability for perceiving the object if it was there
-went from 0.9 in the easiest case to 0.65 and 0.4 in the average and
-hard cases respectively. In addition, we test how our system performs
-in tasks requiring indirect sensing. Here, we gave the planner rewards
-for visiting a certain type of room (e.g. kitchen or office). We did
-not place any humans in those scenarios, so the only way of
-determining the room category is by looking for objects typical for
-that room type. As this indirect reasoning is a tasks that cannot be
-performed by the continual planner alone, we did not run the baseline
-system on those problems.
+%%  Therefore, an instantaneous positive detection of an object is
+%% not proof positive of it being there.
+
+We perform experiments on tasks of several sizes. The objective is to
+find one or more objects and report their position to the user. In
+order to determine the impact of the reliability of sensing in a
+comparison with the baseline, we run tests on identical problems but
+changed the sensor model for the target objects: the probability for
+perceiving the object if it was there went from 0.9 in the easiest
+case to 0.65 and 0.4 in the average and hard cases respectively. In
+addition, we test how our system performs in tasks requiring indirect
+sensing. Here, we gave the planner rewards for visiting a certain type
+of room (e.g. kitchen or office). We did not place any humans in those
+scenarios, so the only way of determining the room category is by
+looking for objects typical for that room type. As this indirect
+reasoning is a tasks that cannot be performed by the continual planner
+alone, we did not run the baseline system on those problems.
 
@@
 about single traces, it is optimistic wrt the remaining costs to a
-goal and pessimistic regarding the probability of reaching tat. By
+goal and pessimistic regarding the probability of reaching it. By
 setting the rewards to non-extremal values has little effect on that

### 2010-11-29 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
paper updates

@@
 -- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. These are called {\em applicability} assumptions if
+third plan step. They are called {\em applicability} assumptions if
 the trace includes an action $\action$ that is not applicable with
@@
 In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ DTPDDL declaration. Our
-approach uses the structure of that, as it occurs in the problem
-description, to define a set of state-assumptions available to
-sequential planning.
+action is expressed in the $(\pp{:init})$ declaration. Our approach
+uses the structure of that, as it occurs in the problem description,
+to define a set of state-assumptions available to sequential planning.
 %%
@@
 conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, a distinct assumption about a proposition
+$\assumptiveS{k}$. Also, an assumption about a proposition
 $\prop$ can only be made once, and must be sequenced before any POMDP
@@
 $\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ and the proposition
-denoting the undefined fact is deleted. For example, if the action
-adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must delete
-$(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added sink,
-where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
+the union of $\state$ with atomic terms from $T_i$ with the
+proposition denoting the undefined fact is deleted. For example, if
+the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
+delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
+sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
 $\transProb(\state^\bot, \action, \state^\bot) = 1$.
@@
 with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly, the system
-switches to a contingent session, discussed in the following section.
+$\bstate$, then immediately before it is executed blindly the system
+switches to a contingent session (described below).
 
@@
 
-\noindent where $\$(T_i)$ is a positive numeric quantity which
-captures the reward the agent receives for correctly rejecting a
-false assumption. The added dual actions are propositionally trivial
--- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
+\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
+(negative) numeric quantity which captures the utility the agent
+receives for correctly (incorrectly) rejecting a false assumption. The
+added dual actions are propositionally trivial -- i.e.,
+$\forall\state\in\states$ $\poss(\assumptiveDT{i})
 \subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
@@
 \$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
--\$(\poss(\switchAction)) & \pp{otherwise} \\
+\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
 \end{array}
@@
 sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., function $\$:2^\propositions\to\Re$, for
-the POMDP posed to the contingent session are sourced from a
-motivational component of our robotic architecture. We make no
-assumptions here about how they are derived.
+project the rewards, i.e., functions
+$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
+contingent session are sourced from a motivational component of our
+robotic architecture. For this paper, we set $\$(x)$ to be a positive
+constant, and have $\hat\$(x)= - \$(x)(1 - p) / p$ where $p$ is the
+probability that $x$ is true.
+
+%%  We make no assumptions here about how they are
+%% derived.
 
@@
 characterised by the active {\em assumptive} actions in the current
-sequential plan. Taking our example trace from earlier, we have a {\em
-relaxed} declaration:
+sequential plan. Figure \ref{fig:abstraction} shows the belief state
+of a more detailed version of our previous example. The diamonds
+represent probabilistic terms of the state description while circles
+represent conjunctive or atomic terms.
 
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen))) \\
-\end{tabtt}
-\normalsize
+\begin{figure}[h!]
+  \centering
+  \tikzstyle{tree} = [sibling distance=4mm]
+  \tikzstyle{toplevel} = [grow'=right, sibling distance=18mm]
+  \tikzstyle{seclevel} = [sibling distance=8mm]
+  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2mm]
+  \tikzstyle{cnode} = [circle, draw=black, minimum size=2.5mm]
+  \tikzstyle{assumption} = [solid, thick, draw=black]
+  \tikzstyle{selected} = [solid, draw=black]
+  \tikzstyle{unused} = [densely dashed, draw=black!40]
+  \subfloat[The initial belief state with the assumptions made by the continual
+    planner.]{
+      \label{fig:abstraction-a}
+      \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode] (cup) {} 
+          child {node [cnode] (cupp1) {}}
+          child {node [cnode] (cupp2) {}}
+        }
+      }
+      child {node[cnode] (kitchen) {}
+        child {node [pnode] (box2) {} 
+          child {node [cnode] (box2p1) {}}
+          child {node [cnode] (box2p2) {}}
+        }
+      };
+   \tiny
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of kitchen] {$\pp{kitchen}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[below=0 of cup] {$\pp{(is-in cup)}$};
+   \node[right=0 of boxp1] {$\pp{place1}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p1] {$\pp{place1}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+   \node[right=0 of cupp1] {$\pp{place1}$};
+   \node[right=0 of cupp2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+\qquad
+  \subfloat[Remove facts that are not part of an assumption.]{
+    \label{fig:abstraction-b}
+    \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode, unused] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode, unused] (cup) {} 
+          child {node [cnode, unused] (cupp1) {}}
+          child {node [cnode, unused] (cupp2) {}}
+        }
+      }
+      child {node[cnode, selected] (kitchen) {}
+        child {node [pnode, selected] (box2) {} 
+          child {node [cnode, unused] (box2p1) {}}
+          child {node [cnode, selected] (box2p2) {}}
+        }
+      };
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+   \tiny
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+\qquad
+  \subfloat[Re-add relevant facts until the belief space limit is reached.]{
+    \label{fig:abstraction-c}
+    \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode, selected] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode, unused] (cup) {} 
+          child {node [cnode, unused] (cupp1) {}}
+          child {node [cnode, unused] (cupp2) {}}
+        }
+      }
+      child {node[cnode, selected] (kitchen) {}
+        child {node [pnode, selected] (box2) {} 
+          child {node [cnode, selected] (box2p1) {}}
+          child {node [cnode, selected] (box2p2) {}}
+        }
+      };
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \draw[selected] (box) -- (boxp1);
+   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+   \draw[selected] (box2) -- (box2p1);
+   \tiny
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[right=0 of boxp1] {$\pp{place1}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p1] {$\pp{place1}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
 
-\noindent and therefore, the belief-state $\bstate_0$ is:
+  \caption{Generating the abstract belief space.}
+\label{fig:abstraction}
+\end{figure}
 
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-\hline
-\end{tabular}
-\normalsize
+% \small
+% \begin{tabtt}
+% (\=:init (= (is-in DORA) kitchen) \+ \\
+%        (probabilistic \=.8 (= (is-in box) office)  \\
+% 		      \>.2 (= (is-in box) kitchen))) \\
+% \end{tabtt}
+% \normalsize
 
-\noindent At this point, the session would proceed in an
-abstraction of the environment that does not contain a
-$\pp{cup}$. With regards to abstraction, the contingent and serial
-sessions apply the same applicability conditions ---i.e., that each
-effect and condition $\prop$ is specified, $\prop\in\state$--- to
-actions from the DTPDDL problem description.
+% \noindent and therefore, the belief-state $\bstate_0$ is:
+
+% \small
+% \begin{tabular}{cccc}
+% \hline
+% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
+% \hline
+% .8 & kitchen & office & \# \\
+% .2 & kitchen & kitchen & \# \\
+% \hline
+% \end{tabular}
+% \normalsize
+
+ % With
+% regards to abstraction, the contingent and serial sessions apply the
+% same applicability conditions ---i.e., that each effect and condition
+% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
+% problem description.
+
+In the first step, we remove all facts that are not part of an
+assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+would proceed in an abstraction of the environment where that does not
+contain $\pp{place2}$, the $\pp{cup}$ or a $\pp{kitchen}$.
 
@@
 adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states that occur with non-zero probability
+number of abstract states in $\bstate_0$ that occur with non-zero probability
 according to that refined declaration remains of a practicable
@@
   entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition. Loosely speaking, the {\em conditional
-entropy} measure is higher the more uncertain those assumptions are
-given truth assignment to the proposition. Propositions that in the
-first place were excluded are iteratively added to the problem posed
-to the conditional session in increasing order according to that
-measure.
+corresponding proposition.  Loosely speaking, the lower the {\em
+  conditional entropy} of an atom is, the more information about the
+assumptions can the contingent planner expect to derive from it. Thus,
+propositions that in the first place were excluded are iteratively
+added to the problem posed to the conditional session in increasing
+order according to that measure (Fig. \ref{fig:abstraction-c}).
 
@@
 %% candidate tree.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "moritz_2011"
+%%% End:

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
fixed discussions
Tiny changes to your example

@@
 contingent session are sourced from a motivational component of our
-robotic architecture. For this paper we take $\hat\$(x)= - \$(x)(1 -
-p) / p$, where $\$(x)$ is a given constant.
+robotic architecture. For this paper, we set $\$(x)$ to be a positive
+constant, and have $\hat\$(x)= - \$(x)(1 - p) / p$ where $p$ is the
+probability that $x$ is true.
 
@@
 characterised by the active {\em assumptive} actions in the current
-sequential plan. Taking our example trace from earlier, we have a {\em
-relaxed} declaration:
+sequential plan. Figure \ref{fig:abstraction} shows the belief-state
+of a more detailed version of our previous example. The diamonds
+represent probabilistic terms of the state description while circles
+represent conjunctive or atomic terms.
 
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen))) \\
-\end{tabtt}
-\normalsize
+\begin{figure}[h!]
+  \centering
+  \tikzstyle{tree} = [sibling distance=4mm]
+  \tikzstyle{toplevel} = [grow'=right, sibling distance=18mm]
+  \tikzstyle{seclevel} = [sibling distance=8mm]
+  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2mm]
+  \tikzstyle{cnode} = [circle, draw=black, minimum size=2.5mm]
+  \tikzstyle{assumption} = [solid, thick, draw=black]
+  \tikzstyle{selected} = [solid, draw=black]
+  \tikzstyle{unused} = [densely dashed, draw=black!40]
+  \subfloat[The initial belief-state with the assumptions made by the continual
+    planner.]{
+      \label{fig:abstraction-a}
+      \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode] (cup) {} 
+          child {node [cnode] (cupp1) {}}
+          child {node [cnode] (cupp2) {}}
+        }
+      }
+      child {node[cnode] (kitchen) {}
+        child {node [pnode] (box2) {} 
+          child {node [cnode] (box2p1) {}}
+          child {node [cnode] (box2p2) {}}
+        }
+      };
+   \tiny
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of kitchen] {$\pp{kitchen}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[below=0 of cup] {$\pp{(is-in cup)}$};
+   \node[right=0 of boxp1] {$\pp{place1}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p1] {$\pp{place1}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+   \node[right=0 of cupp1] {$\pp{place1}$};
+   \node[right=0 of cupp2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+\qquad
+  \subfloat[Remove facts that are not part of an assumption.]{
+    \label{fig:abstraction-b}
+    \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode, unused] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode, unused] (cup) {} 
+          child {node [cnode, unused] (cupp1) {}}
+          child {node [cnode, unused] (cupp2) {}}
+        }
+      }
+      child {node[cnode, selected] (kitchen) {}
+        child {node [pnode, selected] (box2) {} 
+          child {node [cnode, unused] (box2p1) {}}
+          child {node [cnode, selected] (box2p2) {}}
+        }
+      };
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+   \tiny
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+\qquad
+  \subfloat[Re-add relevant facts until the belief space limit is reached.]{
+    \label{fig:abstraction-c}
+    \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode, selected] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode, unused] (cup) {} 
+          child {node [cnode, unused] (cupp1) {}}
+          child {node [cnode, unused] (cupp2) {}}
+        }
+      }
+      child {node[cnode, selected] (kitchen) {}
+        child {node [pnode, selected] (box2) {} 
+          child {node [cnode, selected] (box2p1) {}}
+          child {node [cnode, selected] (box2p2) {}}
+        }
+      };
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \draw[selected] (box) -- (boxp1);
+   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+   \draw[selected] (box2) -- (box2p1);
+   \tiny
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[right=0 of boxp1] {$\pp{place1}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p1] {$\pp{place1}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
 
-\noindent and therefore, the belief-state $\bstate_0$ is:
+  \caption{Generating the abstract belief space.}
+\label{fig:abstraction}
+\end{figure}
 
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-\hline
-\end{tabular}
-\normalsize
+% \small
+% \begin{tabtt}
+% (\=:init (= (is-in DORA) kitchen) \+ \\
+%        (probabilistic \=.8 (= (is-in box) office)  \\
+% 		      \>.2 (= (is-in box) kitchen))) \\
+% \end{tabtt}
+% \normalsize
 
-\noindent At this point, the session would proceed in an
-abstraction of the environment that does not contain a
-$\pp{cup}$. With regards to abstraction, the contingent and serial
-sessions apply the same applicability conditions ---i.e., that each
-effect and condition $\prop$ is specified, $\prop\in\state$--- to
-actions from the DTPDDL problem description.
+% \noindent and therefore, the belief-state $\bstate_0$ is:
+
+% \small
+% \begin{tabular}{cccc}
+% \hline
+% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
+% \hline
+% .8 & kitchen & office & \# \\
+% .2 & kitchen & kitchen & \# \\
+% \hline
+% \end{tabular}
+% \normalsize
+
+ % With
+% regards to abstraction, the contingent and serial sessions apply the
+% same applicability conditions ---i.e., that each effect and condition
+% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
+% problem description.
+
+In the first step, we remove all facts that are not part of an
+assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+would proceed in an abstraction of the environment where that does not
+contain $\pp{place2}$, the $\pp{cup}$ or a $\pp{kitchen}$.
 
@@
 the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition. Loosely speaking, the {\em conditional
-entropy} measure is higher the more uncertain those assumptions are
-given truth assignment to the proposition. Propositions that in the
-first place were excluded are iteratively added to the problem posed
-to the conditional session in increasing order according to that
-measure.
+corresponding proposition.  Loosely speaking, the lower the {\em
+  conditional entropy} of an atom, the more information about the
+assumptions the contingent planner expect to derive from it. Thus,
+propositions that in the first place were excluded are iteratively
+added to the problem posed to the conditional session in increasing
+order according to that measure (Fig. \ref{fig:abstraction-c}).
 

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
changes to the bibliography and evaluation text

@@
 To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is
-exploring an office or living environment, and trying to report the
-locations of objects to their owner.
-
-% \begin{figure}[h]
-%   \centering
-%   \input{figures/dora}
-%   \caption{A exploration domain with three rooms and 13 places. The
-%     robot is in in the centre of room A}
-% \label{fig:dora2}
-% \end{figure}
-% An instance of this domain is shown in figure \ref{fig:dora2}. 
-Basic types of element are {\tt rooms}, {\tt places} and {\tt
-  objects}. Places are topological map nodes that occur in rooms.
-Objects, including the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by the
+scenario from our physical robotic system. Here, a robot is exploring
+an office or living environment, and trying to report the locations of
+objects to their owner. Basic types are {\tt rooms}, {\tt places} and
+{\tt objects}. Places are topological map nodes that occur in rooms.
+Objects, and also the robot, are always located at a place. The robot
+can move around the rooms via connections between places given by a
 {\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room) and depending on this
-category, objects are placed inside the rooms.
-
-The robot can look for an object at a place by executing a {\tt
-  look-for-object} action. Moreover, the contents of a room indicate
-its category, because certain objects are more likely to be located in
-rooms of particular categories. For example, a box of cornflakes is
-more likely to be located in the kitchen than the office. Executing
-{\tt look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. Here, some
-objects are harder to detect than others. Lastly, if the robot is in
-the presence of a human, it can ask what type of room it is currently
-in, however conducting a dialogue is more costly than running a vision
-algorithm (cost of 8 vs costs of 3).
+  category} (e.g. kitchen, office, living room). Certain objects are
+more likely to be located in rooms of a particular category.  For
+example, a box of cornflakes is more likely to be located in the
+kitchen than the office. The robot can look for an object at a place
+by executing a {\tt look-for-object} action. Executing {\tt
+  look-for-object} results in a noisy sensing outcome, sometimes
+indicating a positive perception if the object is there. We have that
+some objects are harder to detect than others. Finally, if the robot
+is in the presence of a human, it can ask what type of room it is
+currently in, however conducting a dialogue is more costly than
+running a vision algorithm (cost of 8 vs costs of 3).
 

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
minor corrections

@@
 \qquad
-  \subfloat[Remove facts that are not part of an assumption.]{
+  \subfloat[Removing facts that are not part of an assumption.]{
     \label{fig:abstraction-b}
@@
 \qquad
-  \subfloat[Re-add relevant facts until the belief space limit is reached.]{
+  \subfloat[Re-adding relevant facts until the belief-state size limit is reached.]{
     \label{fig:abstraction-c}
@@
 assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment where that does not
-contain $\pp{place2}$, the $\pp{cup}$ or a $\pp{kitchen}$.
+would proceed in an abstraction of the environment that does not
+contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$.
 

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
The graphs need updating!
new discussion section

@@
 
-We perform experiments on tasks of several sizes. The objective is to
-find one or more objects and report their position to the user. In
-order to determine the impact of the reliability of sensing in a
-comparison with the baseline, we run tests on identical problems but
-changed the sensor model for the target objects: the probability for
-perceiving the object if it was there went from 0.9 in the easiest
-case to 0.65 and 0.4 in the average and hard cases respectively. In
-addition, we test how our system performs in tasks requiring indirect
-sensing. Here, we gave the planner rewards for visiting a certain type
-of room (e.g. kitchen or office). We did not place any humans in those
-scenarios, so the only way of determining the room category is by
-looking for objects typical for that room type. As this indirect
-reasoning is a tasks that cannot be performed by the continual planner
-alone, we did not run the baseline system on those problems.
+We evaluated our switching planner against the baseline in several
+problems with the number of rooms ranging from 3 to 6. We examine the
+impact of sensing reliability, experimenting using sensor models that
+are: ({\em reliable}) $0.9$ chance the object is perceived if it is
+present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
+chance is 0.4. The continual planning times are reported in
+Figure~\ref{fig:results-time}, and the quality data in
+Figure~\ref{fig:results-quality}. Here, in switching runs we examine
+the performance of our system where $b_0$ allocates non-zero
+probability to between 20 and 200 abstract states during contingent
+sessions. For Items~$a$-$e$ the objective is to find one or more
+objects and report their position to the user. In some of the tasks we
+configured the environment so that no plan exists, by excluding the
+desired object from the environment. Item~$f$ reports for tasks
+requiring indirect sensing, where the robot must relocate to a
+particular room category. For Item~$f$ there is no human in the
+environment, and therefore the baseline cannot be used, because there
+are no sensing schemata that directly sense room categories. In our
+evaluation we run 50 simulations in each configuration, and have a
+timeout on each simulation 30 minutes (1800 seconds). In our domain,
+rewards encapsulate a goal condition, and therefore we find it
+intuitive to report average costs of the plans and the success rate in
+problems that admit a solution (i.e., positive reward scaled by a
+constant factor).
 
-In our evaluation we run 50 simulations in each configuration. A run
-timeouts after 30 minutes (1800 seconds). The planning and simulation
-models are in exact correspondence. Not all problems have a solution,
-so we also examine whether or not detection of the non-existence of a
-solution is important in stochastic domains.
-%% 
-We perform multiple tests with different limits for the
-belief-space size in contingent sessions.  Higher limits result in
-longer contingent planning times and higher plan quality.
+%% Because the switching utility model
 
-Using a satisficing, optimistic serial planner in sequential sessions
-makes optimising for the expected reward difficult. Only reasoning
-about single traces, it is optimistic wrt the remaining costs to a
-goal and pessimistic regarding the probability of reaching it. By
-setting the rewards to non-extremal values has little effect on that
-system. Therefore, we do not consider expected utility in our
-experiments. Instead, we report average costs of the plans and the
-success rate (as a ratio between solved and solvable problems).
+%% Using a satisficing, optimistic serial planner in sequential sessions
+%% makes optimising for the expected reward difficult. Only reasoning
+%% about single traces, it is optimistic wrt the remaining costs to a
+%% goal and pessimistic regarding the probability of reaching it. By
+%% setting the rewards to non-extremal values has little effect on that
+%% system. Therefore, we do not consider expected utility in our
+%% experiments. Instead, we report average costs of the plans and the
+%% success rate (as a ratio between solved and solvable problems).
 
@@
   \includegraphics{dora2-time}\hfill
-  \vspace{2mm}
+  % \vspace{2mm}
   \includegraphics{dora3-time}\hfill
-  \vspace{2mm}
+  % \vspace{2mm}
   \includegraphics{dora4-time}\hfill
@@
   \includegraphics{dora2-quality}\hfill
-  \vspace{2mm}
+  % \vspace{2mm}
   \includegraphics{dora3-quality}\hfill
-  \vspace{2mm}
+  % \vspace{2mm}
   \includegraphics{dora4-quality}\hfill
@@
 
-Our experimental results are summarised in
-Figure~\ref{fig:results-time} and Figure~\ref{fig:results-quality}. We
-find that the time spent on contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined. For coarse initial
-configurations, the cost of contingent planning is compensated by the
-decrease of time spent in Fast Downward. Addressing success ratios and
-plan costs, for objects that can be easily detected there is little
-gain in using a decision-theoretic planner, as the greedy sensing
-approach by the baseline is sufficient. With decreasing sensor
-reliability contingent planning pays off.%% : while the resulting plans are still
+Examining success ratios and plan costs, where sensing is reliable
+there is little to be gained by using the contingent planner, as the
+greedy sensing approach of the baseline is sufficient. Not
+surprisingly, as sensing degrades contingent planning pays off.  Also,
+we find that time spent in contingent planning increases steeply as
+the abstraction $\bstate_0$ becomes more refined.  That refinement
+seems to be paying off in terms of the success ratio, particularly for
+problems $d$ and $e$, where we had sequential sessions using weighted
+$A^*$ (rather than $A^*$). For less refined initial configurations, the
+increase cost of contingent planning is compensated by a decrease in
+time spent executing the serial system Fast Downward. The relatively
+high success rate irrespective of the level of refinement in the
+initial configuration indicates the effectiveness of using conditional
+entropy to guide abstraction refinement in our setting.
+
+%% :
+%%while the resulting plans are still
 %% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.  Less aggressive abstraction
-of the initial configuration results in longer runtimes. In our
-scenario it is worthwhile being assumptively aggressive.  We suppose
-the relatively high success rate irrespective of the coarseness of the
-initial configuration indicates the effectiveness of using conditional
-entropy to guide abstraction refinement.
+%% smaller than for the baseline system.
 
-We find that abstraction has little impact on plan
-costs and success rate in our scenario. Increasing the size of the
-initial abstract belief beyond 50 states rarely pays off, because
-while additional information may facilitate better plans, the indirect
-sensing tests report the same result. Here, we only really see an
-impact in terms of the runtime of contingent planning.
+%% Less aggressive abstraction of the initial configuration results in
+%% longer runtimes, with little to be gained in terms of plan quality. We
+%% conclude that in our scenario it is worthwhile being assumptively
+%% aggressive.
+
+%% We find that abstraction has little impact on plan
+%% costs and success rate in our scenario. Increasing the size of the
+%% initial abstract belief beyond 50 states rarely pays off, because
+%% while additional information may facilitate better plans, the indirect
+%% sensing tests report the same result. Here, we only really see an
+%% impact in terms of the runtime of contingent planning.
 

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
okay

@@
 environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
-to use several underlying planning systems. We have also extended
+to use several underlying planning systems. We have also extended the
 MAPSIM system so that it can parse DTPDDL, and perform successive
@@
 To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is
-exploring an office or living environment, and trying to report the
-locations of objects to their owner.
-
-\begin{figure}[h]
-  \centering
-  \input{figures/dora}
-  \caption{A exploration domain with three rooms and 13 places. The
-    robot is in in the centre of room A}
-\label{fig:dora2}
-\end{figure}
-An instance of this domain is shown in figure \ref{fig:dora2}. Basic
-types of element are {\tt rooms}, {\tt places} and {\tt
-  objects}. Places are topological map nodes that occur in rooms.
-Objects, including the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by the
+scenario from our physical robotic system. Here, a robot is exploring
+an office or living environment, and trying to report the locations of
+objects to their owner. Basic types are {\tt rooms}, {\tt places} and
+{\tt objects}. Places are topological map nodes that occur in rooms.
+Objects, and also the robot, are always located at a place. The robot
+can move around the rooms via connections between places given by a
 {\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room) and depending on this
-category, objects are placed inside the rooms.
-
-The robot can look for an object at a place by executing a {\tt
-  look-for-object} action. Moreover, the contents of a room indicate
-its category, because certain objects are more likely to be located in
-rooms of particular categories. For example, a box of cornflakes is
-more likely to be located in the kitchen than the office. Executing
-{\tt look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. Here, some
-objects are harder to detect than others. Lastly, if the robot is in
-the presence of a human, it can ask what type of room it is currently
-in, however conducting a dialogue is more costly than running a vision
-algorithm (cost of 8 vs costs of 3).
+  category} (e.g. kitchen, office, living room). Certain objects are
+more likely to be located in rooms of a particular category.  For
+example, a box of cornflakes is more likely to be located in the
+kitchen than the office. The robot can look for an object at a place
+by executing a {\tt look-for-object} action. Executing {\tt
+  look-for-object} results in a noisy sensing outcome, sometimes
+indicating a positive perception if the object is there. We have that
+some objects are harder to detect than others. Finally, if the robot
+is in the presence of a human, it can ask what type of room it is
+currently in, however conducting a dialogue is more costly than
+running a vision algorithm (cost of 8 vs costs of 3).
 
@@
 
-We perform experiments on tasks of several sizes. The objective is to
-find one or more objects and report their position to the user. In
-order to determine the impact of the reliability of sensing in a
-comparison with the baseline, we run tests on identical problems but
-changed the sensor model for the target objects: the probability for
-perceiving the object if it was there went from 0.9 in the easiest
-case to 0.65 and 0.4 in the average and hard cases respectively. In
-addition, we test how our system performs in tasks requiring indirect
-sensing. Here, we gave the planner rewards for visiting a certain type
-of room (e.g. kitchen or office). We did not place any humans in those
-scenarios, so the only way of determining the room category is by
-looking for objects typical for that room type. As this indirect
-reasoning is a tasks that cannot be performed by the continual planner
-alone, we did not run the baseline system on those problems.
+We evaluated our switching planner against the baseline in several
+problems with the number of rooms ranging from 3 to 6. We examine the
+impact of sensing reliability, experimenting using sensor models that
+are: ({\em reliable}) $0.9$ chance the object is perceived if it is
+present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
+chance is 0.4. The continual planning times are reported in
+Figure~\ref{fig:results-time}, and the quality data in
+Figure~\ref{fig:results-quality}. Here, in switching runs we examine
+the performance of our system where $b_0$ allocates non-zero
+probability to between 20 and 200 abstract states during contingent
+sessions. For Items~$a$-$e$ the objective is to find one or more
+objects and report their position to the user. In some of the tasks we
+configured the environment so that no plan exists, by excluding the
+desired object from the environment. Item~$f$ reports for tasks
+requiring indirect sensing, where the robot must relocate to a
+particular room category. For Item~$f$ there is no human in the
+environment, and therefore the baseline cannot be used, because there
+are no sensing schemata that directly sense room categories. In our
+evaluation we run 50 simulations in each configuration, and have a
+timeout on each simulation 30 minutes (1800 seconds). In our domain,
+rewards encapsulate a goal condition, and therefore we find it
+intuitive to report average costs of the plans and the success rate in
+problems that admit a solution (i.e., positive reward scaled by a
+constant factor).
 
-In our evaluation we run 50 simulations in each configuration. A run
-timeouts after 30 minutes (1800 seconds). The planning and simulation
-models are in exact correspondence. Not all problems have a solution,
-so we also examine whether or not detection of the non-existence of a
-solution is important in stochastic domains.
-%% 
-We perform multiple tests with different limits for the
-belief-space size in contingent sessions.  Higher limits result in
-longer contingent planning times and higher plan quality.
+%% Because the switching utility model
 
-Using a satisficing, optimistic serial planner in sequential sessions
-makes optimising for the expected reward difficult. Only reasoning
-about single traces, it is optimistic wrt the remaining costs to a
-goal and pessimistic regarding the probability of reaching tat. By
-setting the rewards to non-extremal values has little effect on that
-system. Therefore, we do not consider expected utility in our
-experiments. Instead, we report average costs of the plans and the
-success rate (as a ratio between solved and solvable problems).
+%% Using a satisficing, optimistic serial planner in sequential sessions
+%% makes optimising for the expected reward difficult. Only reasoning
+%% about single traces, it is optimistic wrt the remaining costs to a
+%% goal and pessimistic regarding the probability of reaching it. By
+%% setting the rewards to non-extremal values has little effect on that
+%% system. Therefore, we do not consider expected utility in our
+%% experiments. Instead, we report average costs of the plans and the
+%% success rate (as a ratio between solved and solvable problems).
 

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Some graphics adjustments

@@
   \centering
-  \tikzstyle{tree} = [sibling distance=4mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=18mm]
-  \tikzstyle{seclevel} = [sibling distance=8mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=2.5mm]
+  \tikzstyle{tree} = [sibling distance=4.5mm]
+  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
+  \tikzstyle{seclevel} = [sibling distance=9mm]
+  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
+  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
   \tikzstyle{assumption} = [solid, thick, draw=black]

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
last change fo the night

@@
 contingent session are sourced from a motivational component of our
-robotic architecture. For this paper, we set $\$(x)$ to be a positive
-constant, and have $\hat\$(x)= - \$(x)(1 - p) / p$ where $p$ is the
-probability that $x$ is true.
+robotic architecture. For this paper, for $\assumptiveDT{i}$ actions
+we set $\$(x)$ to be a positive constant, and have $\hat\$(x)= -
+\$(x)(1 - p) / p$ where $p$ is the probability that $x$ is true. For
+$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
+\$(x)p/(1-p)$.\footnote{These values were chosen to encourage sensing.}
 

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
reading over what we wrote for switching

@@
 contingent session are sourced from a motivational component of our
-robotic architecture. For this paper, we set $\$(x)$ to be a positive
-constant, and have $\hat\$(x)= - \$(x)(1 - p) / p$ where $p$ is the
-probability that $x$ is true.
+robotic architecture. In this paper, for $\assumptiveDT{i}$ actions
+we set $\$(x)$ to be a positive constant, and have $\hat\$(x)= -
+\$(x)(1 - p) / p$ where $p$ is the probability that $x$ is true. For
+$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
+\$(x)p/(1-p)$.\footnote{These values were chosen to encourage sensing.}
 
@@
 characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief state
+sequential plan. Figure \ref{fig:abstraction} shows the belief-state
 of a more detailed version of our previous example. The diamonds
@@
   \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief state with the assumptions made by the continual
-    planner.]{
+  \subfloat[The initial belief-state with the assumptions made by the continual
+    planner in bold.]{
       \label{fig:abstraction-a}
@@
 \qquad
-  \subfloat[Re-add relevant facts until the belief space limit is reached.]{
+  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
     \label{fig:abstraction-c}
@@
 assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment where that does not
-contain $\pp{place2}$, the $\pp{cup}$ or a $\pp{kitchen}$.
-
+would proceed in an abstraction of the environment that does not
+contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
+%%
 In a second step, we iteratively refine the relaxed declaration by
 adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero probability
-according to that refined declaration remains of a practicable
-size. In detail, for each excluded atomic term, we compute the {\em
-  entropy} of the active assumptions, {\em conditional} on the
+number of abstract states in $\bstate_0$ that occur with non-zero
+probability according to that refined declaration remains of a
+practicable size. In detail, for each excluded atomic term, we compute
+the {\em entropy} of the active assumptions, {\em conditional} on the
 corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom is, the more information about the
-assumptions can the contingent planner expect to derive from it. Thus,
+  conditional entropy} of an atom, the more information about the
+assumptions the contingent planner expect to derive from it. Thus,
 propositions that in the first place were excluded are iteratively

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
small change
few more corrections

@@
 We evaluated our switching planner against the baseline in several
-problems with the number of rooms ranging from 3 to 6. We examine the
+tasks with the number of rooms ranging from 3 to 6. We examine the
 impact of sensing reliability, experimenting using sensor models that
@@
 probability to between 20 and 200 abstract states during contingent
-sessions. For Items~$a$-$e$ the objective is to find one or more
+sessions. For Item~$a$-$e$ the objective is to find one or more
 objects and report their position to the user. In some of the tasks we
@@
 evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation 30 minutes (1800 seconds). In our domain,
+timeout on each simulation of 30 minutes (1800 seconds). In our domain,
 rewards encapsulate a goal condition, and therefore we find it
@@
 seems to be paying off in terms of the success ratio, particularly for
-problems $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations, the
-increase cost of contingent planning is compensated by a decrease in
-time spent executing the serial system Fast Downward. The relatively
-high success rate irrespective of the level of refinement in the
-initial configuration indicates the effectiveness of using conditional
-entropy to guide abstraction refinement in our setting.
+tasks $d$ and $e$, where we had sequential sessions using weighted
+$A^*$ (rather than $A^*$). For less refined initial configurations,
+the increase cost of contingent planning is compensated by a decrease
+in Fast Downward planning times. The relatively high success rate
+irrespective of the level of refinement in the initial configuration
+indicates the effectiveness of using conditional entropy to guide
+abstraction refinement in our setting.
 

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
new order for macros, and consitent prob notations

@@
 contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions
-we set $\$(x)$ to be a positive constant, and have $\hat\$(x)= -
-\$(x)(1 - p) / p$ where $p$ is the probability that $x$ is true. For
+robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
+set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
+- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
 $\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)p/(1-p)$.\footnote{These values were chosen to encourage sensing.}
+\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
+  sensing.}
 

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
OKay
Notations section had become slightly broken

@@
 We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process
-state $\state$ as the set of propositions that are true of the
-state. Notationally, we have $\state \subseteq
+planning problem, with costs and rewards. We model a process state
+$\state$ as the set of propositions that are true of the
+state. Notationally, $\props$ is the set of propositions, $\prop$ is
+an element from that set, and we have $\state \subseteq
 \props$. The underlying process
@@
 %%
-Function $\reward:\states \times \actions \to \Re$ is a bounded real
+The function $\reward:\states \times \actions \to \Re$ is a bounded real
 valued reward function. Therefore a finite positive constant $c$

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
few changes

@@
 
-Finally, we make some fairly standard assumptions. First, that
-action execution and sensing occur instantaneously, and that only one
-action can be applied per plan-step. Second, 
-that
-%it can arise that in
-%some propositional states an action is applicable, and in others it is
-%not. Moreover, a belief-state can assign non-zero probability to
-%states where that applicability holds, and states where it does
-%not. We differ
-%%in situations where an action exists that is applicable in some state 
+Finally, we make some fairly standard assumptions. First, that action
+execution and sensing occur instantaneously, and that only one action
+can be applied per plan-step. Secondly, it can arise that in some
+states an action is applicable, and in others it is not. Moreover, a
+belief-state can assign non-zero probability to states where that
+applicability holds, and states where it does not. %% We differ
+%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
+%% In beliefs where an action exists that is applicable in some state 
 %% with non-zero probability, and not applicable in another, we differ
-%from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%because 
-%%in that
-%we forbid the execution of the action.
-%actions that are not applicable in
-%any state of the current belief.  
-%Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
+%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
+%% because 
+%% in that
+%% we forbid the execution of the action.
+%% actions that are not applicable in
+%% any state of the current belief.  
+%% Also,
+Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
 we incorporate a PPDDL-like default semantics, treating an action

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
few requests from moritz

@@
 We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process
-state $\state$ as the set of propositions that are true of the
-state. Notationally, we have $\state \subseteq
+planning problem, with costs and rewards. We model a process state
+$\state$ as the set of propositions that are true of the
+state. Notationally, $\props$ is the set of propositions, $\prop$ is
+an element from that set, and we have $\state \subseteq
 \props$. The underlying process
 dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:1996} $\stochActions$ over
+STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
 state-characterising propositions $\propositions$.
@@
 $\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$; In the
-usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
+\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
+usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
 %%
@@
 
-Finally, we make a number of fairly standard assumptions. First, that
-action execution and sensing occurs instantaneously, and that only one
-action can be applied at a plan-step. Second, it can arise that in
-some propositional states an action is applicable, and in others it is
-not. Moreover, a belief-state can assign non-zero probability to
-states where that applicability holds, and states where it does
-not. We differ
-from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-because we forbid the execution of actions that are not applicable in
-any state of the current belief.  Also,
-unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we do incorporate a PPDDL-like default semantics, treating an action
+Finally, we make some fairly standard assumptions. First, that action
+execution and sensing occur instantaneously, and that only one action
+can be applied per plan-step. Secondly, it can arise that in some
+states an action is applicable, and in others it is not. Moreover, a
+belief-state can assign non-zero probability to states where that
+applicability holds, and states where it does not. %% We differ
+%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
+%% In beliefs where an action exists that is applicable in some state 
+%% with non-zero probability, and not applicable in another, we differ
+%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
+%% because 
+%% in that
+%% we forbid the execution of the action.
+%% actions that are not applicable in
+%% any state of the current belief.  
+%% Also,
+Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
+we incorporate a PPDDL-like default semantics, treating an action
 \stochAction\ executed at $\state_i$ as if it has no effect at
@@
 \state$.
+
+% Mucked around with this, then realised the two sentences were
+% contradictory. If we forbid actions applicable in one state but not
+% another in the belief state, then clearly we aren't using a
+% PPDDL-like semantics.
+%% Doubled comments are stuff I added, then removed.
 
@@
 
-Successive state estimation is available by application of Bayes'
+Successive state estimation is by application of Bayes'
 rule.  Taking the current belief $\bstate$ as the {\em prior}, and
 supposing action $\action$ is executed with perceptive outcome
-$\observ$, then the probability that we are in $\state$ in the
-successive belief-state $\bstate'$ is given by:
+$\observ$, the probability that we are in $\state$ in the
+successive belief-state $\bstate'$ is:
 

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
few requests from moritz

--- 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
@@
 trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the {\em underlying}
+the sequential session that continues at the current%%{\em underlying}
 belief-state.
@@
 (negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting a false assumption. The
+receives for correctly (incorrectly) rejecting an assumption. The
 added dual actions are propositionally trivial -- i.e.,
@@
 contingent session are sourced from a motivational component of our
-robotic architecture. For this paper, we set $\$(x)$ to be a positive
-constant, and have $\hat\$(x)= - \$(x)(1 - p) / p$ where $p$ is the
-probability that $x$ is true.
+robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
+set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
+- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
+$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
+\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
+  sensing.}
 
@@
   \subfloat[The initial belief-state with the assumptions made by the continual
-    planner.]{
+    planner in bold.]{
       \label{fig:abstraction-a}
@@
 \qquad
-  \subfloat[Re-adding relevant facts until the belief-state size limit is reached.]{
+  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
     \label{fig:abstraction-c}
@@
 contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-
+%%
 In a second step, we iteratively refine the relaxed declaration by

### 2010-11-30 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
Changes to graph

@@
   \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, thick, draw=black]
+  \tikzstyle{assumption} = [solid, very thick, draw=black]
   \tikzstyle{selected} = [solid, draw=black]
@@
   \end{tikzpicture}}
-\qquad
+\vspace{2mm}
   \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex

@@
 propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
+\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
 nature chooses a deterministic STRIPS effect $\detAct_i$, and for
@@
 STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for our presentation to clearly
+propositionally, we find it convenient for presentation and computation to
 separate sensing and action. Therefore, we suppose a POMDP has a
@@
 detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote \observations\ the set of
+$\observ \subseteq \percepts$, and denote by \observations\ the set of
 reachable observations. The underlying state of the process cannot be
@@
 effect an observation $\observ \in
-\observations$ that informs what should be believed about the state a
+\observations$ that informs what should be believed about the state the
 process is in. If $\action$ is applied effecting a
@@
 %%
-From $\psi$ we have that for each $\state \in \states$ and action
+From $\psi$ we have that for each $\state' \in \states$ and action
 $\action \in \actions$, an observation $\observ \in \observations$ is
 generated independently according to a probability distribution
-$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
+$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
 \action)$ the probability of getting observation $\observ$ in state
-$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state, \action) = 1$.
+$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
+\observations} \obsDist_\observ(\state', \action) = 1$.
 
@@
 supposing action $\action$ is executed with perceptive outcome
-$\observ$, the probability that we are in $\state$ in the
+$\observ$, the probability that we are in $\state'$ in the
 successive belief-state $\bstate'$ is:
@@
 \begin{equation}\label{eq:revision}
-\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
-  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
+\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
+  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
 \end{equation}

### 2011-01-19 subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
Oops, the source code was deleted

%%%% ijcai11.tex

\typeout{IJCAI-11 Instructions for Authors}

% These are the instructions for authors for IJCAI-11.
% They are the same as the ones for IJCAI-07 with superficical wording
%   changes only.

\documentclass{article}
% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
\usepackage{ijcai11}

% Use the postscript times font!
\usepackage{times}

% the following package is optional:
%\usepackage{latexsym}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

\usepackage{subfig}
%% \usepackage{tikz}
%% \usetikzlibrary{shapes}
%% \usetikzlibrary{positioning}

\usepackage{graphicx}
\graphicspath{{plots/}}
\include{macros}%

\title{A Switching Planner for Combined Task and Observation Planning}
\author{anonymous}

\begin{document}

\maketitle

\begin{abstract}

Realistic robot planning problems in uncertain environments often
require achieving tasks while also finding out about the
world. Because the world state cannot be observed directly, these
problems are naturally represented as partially observable Markov
decision problems (POMDPs). However, these are typically intractable
for realistic problems.
%%
We present a \emph{switching planner} that employs fast sequential
planning to decide on the overall strategy, and uses a
decision-theoretic planner to solve the subproblems where partial
observability will significantly impact the quality of the plan. We
demonstrate the approach in a realistic robot exploration domain.

\end{abstract}

\section{Introduction}

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive~\cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, have state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative has been to use classical sequential planners and
replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
Execution monitoring tracks properties of the state that determine
whether further planning is required. For partially observable
domains, only probabilistic information about such properties will, in
general, be available.

%% However, these approaches rely
%% crucially on being able to determine properties of the current state
%% in order to decide to replan. For partially observable domains, only
%% probabilistic information about these properties will in general be
%% available.

This paper describes the planning component for a robotic system that
continuously deliberates in a stochastic dynamic environment in order
to achieve objectives set by the user, and acquire knowledge about its
surroundings. This domain features partial observability, particularly
because the state is not perfectly known, and state information is
gained by the robot through the use of unreliable sensing
actions. 
%Using a propositionally factored state representation, for
For
interesting sized tasks the corresponding POMDP model is too large for a
POMDP solver to be applied directly. Inspired by the replanning
approaches used successfully for MDP
planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
planning approach that uses a classical planning system to compute a
reasonably valuable trace in the model, and then uses a
decision-theoretic planner on small subproblems where reasoning about
observations might be useful. We refer to the complete system as a
{\em switching continual planner}. The approach is not optimal,
particularly as it relies on the results of sequential planning
directly. It does nevertheless perform better than a purely sequential
replanner, and is fast enough to be used for real-time decision-making
on a mobile robot.

%% Switching gives us some important advantages in our robotic
%% domain. First, the significant source of uncertainty in the domain is
%% the unreliability of observations, and this approach is particularly
%% suited to problems that combine a deterministic task planning problem
%% with observations as the decision-theoretic planner is often invoked
%% to determine a characteristic of the underlying state. Secondly,
%% replanning makes our system robust to changing objectives and the
%% discovery of new facts about the world.

%% As an illustration, consider a robot searching for a box that it knows
%% is often found in offices. The sequential planner might produce a plan
%% that assumes that {\sc Room a} is an office, that the box is in {\sc
%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
%% these assumptions, the plan might be to go to {\sc Room a}, go to
%% location 1 and search for the box. When the plan gets executed and the
%% search begins the decision-theoretic planner is called since searching
%% involves making observations. The decision-theoretic planner might
%% retract the last assumption but leave the others, and build a plan to
%% search the whole of {\sc Room a} for the box. If this plan fails to
%% find the box, it will disprove the assumption that the box is in {\sc
%% Room a}, which will then lead to replanning by the sequential planner
%% based on the newly discovered knowledge.

Our system is domain independent, taking domain and problem
descriptions in a first-order declarative language we have developed,
called the decision-theoretic planning domain definition language
(DTPDDL). From these descriptions of POMDPs, we automatically
construct a deterministic model suitable for sequential planners. That
model includes actions which correspond to making assumptions about
the values of imperfectly known state variables. Assumptions scheduled
by the sequential system are used to propose a pragmatic abstract
belief-state space to the decision-theoretic system, and to modify the
reward function, so that system might pursue sensing related to those
assumptions.

%%  which can operate
%% with the original representation of the domain.

%% The sequential planner we use is based
%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
%% knowledge to be represented explicitly, so we can write actions that
%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
%% THIS. For the decision-theoretic planner we have implemented our own
%% forward search in the belief-state space.

The switching planner provides important advantages in our
mobile robot domain. First, the significant source of uncertainty in the
domain is the unreliability of observations and this approach is
particularly suited to problems that combine a deterministic task
planning problem with observations because the decision-theoretic planner
is only invoked to deal with state uncertainty.
%determine a characteristic of the underlying state. 
Secondly, replanning makes the system robust to
changing objectives and discoveries about the world,
both of which feature in our domain.

The remainder of the paper proceeds as follows. We begin by describing
POMDPs with a propositionally factored representation of states and
observations. We also outline the relationship between this and a flat
POMDP representation. Next we describe our DTPDDL language with
example declarations from a mobile robot exploration task. Then we
describe the switching planner, including how the sequential domain is
produced and a plan is generated, and how this then provides the input
for the decision-theoretic planner. Finally we provide an evaluation
of the system on the example robot domain, and finish with a short
survey of related work and some concluding remarks.

%To accompany
%that language we have implemented an information-state
%\laostar\ procedure for solving problems expressed in DTPDDL.

%%

%Have a good understanding of the contents and function of rooms, as
%well as the linguistic referrents to rooms, widgets, and their visual
%qualities.

%Having confidence in its beliefs about the linguistic terms for
%spaces, and the function of those spaces.

%sometimes difficult to predict, with exogenous events, such as the
%changing of the objective,

%the speed and scalability of software for sequential planning in
%deterministic

%% We developed a first-order declarative language, called DTPDDL, for
%% describing domains of planning problems that correspond to POMDPs.  To
%% accompany that language we have implemented an information-state
%% \laostar\ procedure for solving problems expressed in DTPDDL.

%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
%% modified \fastdownward\ so that it can find useful sequential executions
%% given DTPDDL models of the problem at hand.

%%  called DTPDDL, along
%% the lines of PDDL for the partially observable case, an \laostar\
%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
%% and modify the \fastdownward\ system to find high quality sequential plans
%% .

%% We model the environment as a partially observable Markov decision
%% process. Although planning in that model is undecidable in
%% general~\cite{}, an optimal finite-horizon plan corresponds to a
%% contingent plan, that is a function mapping action-observation
%% histories to actions.

%% our continual planner is reactive, replanning whenever the underlying
%% domain and problem models change. For example, replanning occurs if
%% the motivational component alters the objectives, and if an assumed
%% outcome of a sensing action is not realised.

%% brittle sensing model,

%% the evaluation of a fluent at a state is either known. Moreover, there
%% is a sensing process that run-time variables  after-which

%% For $\prop \in \state$ we say proposition $\prop$ characterises state
%% $\state$. There is always a unique starting state $\state_0$. The goal
%% $\goal$ is a set of propositions, and we say that state $\state$ is a
%% goal state iff $\goal \subseteq \state$.

%%  To keep this exposition
%% simple, for any two distinct actions $\stochAction_i \neq
%% \stochAction_j$, if outcome $\detAct$ is a possibility for
%% $\stochAction_i$ then it cannot also be a possibility for
%% $\stochAction_j$ -- i.e., if $\detAct \in
%% \detActions(\stochAction_i)$ then $\detAct \not\in
%% \detActions(\stochAction_j)$.

%% The solution to a probabilistic planning problem is a contingency
%% plan. This consists of an assignment of actions to states at each
%% discrete timestep up to the planning horizon $n$. The optimal
%% contingency plan is one which prescribes actions to states that
%% maximise the probability that the goal is achieved within $n$ steps
%% from the starting state $\state_0$. For the purposes of this paper
%% we say a plan fails, i.e. achieves the goal with probability $0$, in
%% situations where it does not prescribe an action. Computing the
%% optimal plan for a problem is computationally intractable, and an
%% important direction for research in the field is to develop
%% heuristic mechanisms for generating small linear plans
%% quickly~\cite{littman:etal:98}

\section{Propositional Decision-Theoretic Planning}

%% {\em runtime variables}
%% {\em decision variables}
%% {\em uninitialised variables}
%% {\em omitted variables}

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. We model a process state
$\state$ as the set of propositions that are true of the
state. Notationally, $\props$ is the set of propositions, $\prop$ is
an element from that set, and we have $\state \subseteq
\props$. The underlying process
dynamics are modelled in terms of a finite set of {\em probabilistic
STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
state-characterising propositions $\propositions$.
%%
We say an action $\stochAction \in \stochActions$ is
applicable if its precondition $\poss(\stochAction)$, a set of
propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
nature chooses a deterministic STRIPS effect $\detAct_i$, and for
all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$.

%% We model a process state $\state$ as the set of propositions that are
%% true of the state. Notationally, we have $\state \subseteq
%% \props$. State change is induced by application of actions. A
%% stochastic action $\stochAction \in \stochActions$ is applicable if
%% its precondition $\poss(\stochAction)$, a set of propositions, are
%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
%% applicable at state $\state$.  If action $\stochAction \in
%% \stochActions(\state)$ is applied at \state, nature chooses one
%% element amongst a small set of deterministic outcomes
%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
%% $\sum_{\detAction_i \in \detActions(\stochAction)}
%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
%% effect on the state given in terms of two lists of propositions. The
%% add-list $\add(\detAction)$ and delete-list
%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
%% $\detAction$, then it cannot be in the delete-list and vice versa.}
%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
%% applied at state $\state$, then the resultant state is $ ( \state \cup
%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
%% propositions from $\add(\detAction)$ are added to $\state$, and those
%% from $\delete(\detAction)$ are removed from $\state$.

%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})

We are concerned with problems that feature partial
observability. Although we could invoke {\em extended probabilistic
STRIPS operators}~\cite{rintanen:01} to model actions and observations
propositionally, we find it convenient for presentation and computation to
separate sensing and action. Therefore, we suppose a POMDP has a
perceptual model given in terms of a finite set of stochastic {\em
senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
and perceptual propositions $\percepts$, called {\em percepts}. In
detail, we take an observation $\observ$ to be a set of percepts
$\observ \subseteq \percepts$, and denote by \observations\ the set of
reachable observations. The underlying state of the process cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in
\observations$ that informs what should be believed about the state the
process is in. If $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from the chosen elements of active
senses.

A POMDP has a starting configuration that corresponds to a Bayesian
belief-state. Intuitively, this is the robot's subjective belief about
its environment. Formally, a belief-state $\bstate$ is a probability
distribution over process states. We write $\bstate(\state)$ to denote
the probability that the process is in $\state$ according to
$\bstate$, and $\bstate_0$ when discussing the starting
configuration.

Finally, we make some fairly standard assumptions. First, that action
execution and sensing occur instantaneously, and that only one action
can be applied per plan-step. Secondly, it can arise that in some
states an action is applicable, and in others it is not. Moreover, a
belief-state can assign non-zero probability to states where that
applicability holds, and states where it does not. %% We differ
%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
%% In beliefs where an action exists that is applicable in some state 
%% with non-zero probability, and not applicable in another, we differ
%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
%% because 
%% in that
%% we forbid the execution of the action.
%% actions that are not applicable in
%% any state of the current belief.  
%% Also,
Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
we incorporate a PPDDL-like default semantics, treating an action
\stochAction\ executed at $\state_i$ as if it has no effect at
states $\state$ if $\poss(\stochAction) \not\subseteq
\state$.

% Mucked around with this, then realised the two sentences were
% contradictory. If we forbid actions applicable in one state but not
% another in the belief state, then clearly we aren't using a
% PPDDL-like semantics.
%% Doubled comments are stuff I added, then removed.

\subsection{Costs, Rewards, and Belief Revision}

Until now we have discussed the POMDP in terms of propositions and
percepts. In order to address belief revision and utility it is
convenient to consider the underlying decision process in a flat
format. This is given by the tuple
$\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
The function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs as
negative rewards.
%%
From $\psi$ we have that for each $\state' \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
\action)$ the probability of getting observation $\observ$ in state
$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state', \action) = 1$.

Successive state estimation is by application of Bayes'
rule.  Taking the current belief $\bstate$ as the {\em prior}, and
supposing action $\action$ is executed with perceptive outcome
$\observ$, the probability that we are in $\state'$ in the
successive belief-state $\bstate'$ is:

\begin{equation}\label{eq:revision}
\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
\end{equation}

\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
normalising factor, that is, the probability of getting observation
$\observ$ given we execute $\action$ in $\bstate$.

\subsection{Plan Evaluation}

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---i.e., reaches a state that satisfies a given
Boolean condition--- with probability $1$.  Generally, whatever the
structure of a plan, its value corresponds to the expected cumulative
reward:

%% More generally, a useful
%% formalism for representing solutions to POMDPs is the finite-state
%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
%% probability that the controller prescribes \action\ when in state
%% \node, and $\transition_\node(\action, \observation, \node') =
%% P(\node'|\node, \action, \observation)$ gives the probability the
%% controller transitions to state $\node'$, supposing we execute
%% \action\ in state \node\ receiving observation \observation. When
%% acting for finite $N$ steps from controller state $n$, the value of a
%% controller corresponds to the expected cumulative reward:

\begin{equation}\label{eq:expectedvalue}
V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
\end{equation}

\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
the action prescribed at step $t$, and

\[\reward(\bstate, \action) = \sum_{\state \in \states}
\bstate(\state)\reward(\state, \action).\]

%% \noindent Finally, it is
%% useful to note that there is a corresponding deterministic FSCs for
%% both sequential (e.g., conformant) and contingent plan formats.

\section{Decision-Theoretic PDDL}

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
used in all three of the International Planning Competitions since
2004. A variation on PDDL for describing domains with stochastic
actions and uncertain starting configurations, PPDDL is a declarative
first-order language that facilitates factored descriptions of domains
and problems. There are straightforward compilations from problems
expressed in PPDDL to propositional representations amenable
to state-of-the-art planners.

Because PPDDL cannot model domains that feature partial observability,
we develop an extension we call {\em Decision-Theoretic
(DT)PDDL}. This can express probabilistic models of the sensing
consequences of acting, to quantitatively capture unreliability in
perception. That expressive power is achieved by incorporating
perceptual analogues of fluent, predicate, and action definitions. In
detail, we have declarations of state characterising predicate and
fluent symbols according to the PPDDL syntax. In addition, we allow
two other declarations, for perceptual predicates and fluents
respectively. For example, suppose our robot is tasked with exploring
{\em locations} in order to identify the whereabouts of a {\em
visual-object}. We must describe state and perceptual facts that model
the {\em true}, resp. perceived, locations of objects. In DTPDDL,
these declarations appear as:

\small
\begin{tabtt}
(\=:functions  ;; state fluents\\
  \> (is-in ?v - visual-object) - location )\\
(:observable-functions  ;; perceptual fluents\\
  \> (o-is-in ?v - visual-object) - location )
\end{tabtt}
\normalsize

\noindent To model the sensing capabilities of the agent, we have
operator-like declarations, with preconditions expressed using state
predicate and action function symbols, and uniformly positive effects
over perceptual predicates. A sense declaration corresponds to a
lifted description of proposition {\em senses}. Whereas the effects of
an {\em operator} schema describe how states change under application
of actions, the effects of a {\em sense} schema are perceptual,
specifying the composition of an observation following the execution
of an action.

%% Also, and without a loss of generality, we find it
%% convenient to separate the actional precondition from the state
%% precondition.

Making the above ideas concrete with an example, suppose a robot
called $\pp{DORA}$ is able to look for a visual-object, such as a
$\pp{box}$, at a given place. We can model that deterministic action
using the following operator schema:

\small
\begin{tabtt}
(\=:action look-for-object \+ \\
   :parameters (\=?r - robot ?v - visual-object\\
   \> ?l - location) \\
   :precondition (and (= (is-in ?r) ?l) ) \\
   :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
\normalsize

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect on the state. That is, its
execution incurs an instantaneous penalty, $3$, that corresponds to
the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
reward} is a {\em reserved word}, and occurs in {\em increased} and
{\em decreased} terms in operator schemata. In that setting, reward is
accumulated, whereas in DTPDDL it is instantaneous.}

Completing the example, we model the sensing that results from
instantiating the $\pp{look-for-object}$ operator:

\small
\begin{tabtt}
(\= :sense vision \+\\
 :parameters \= (\= ?r - robot ?v - visual-object\\
 \>\>  ?l - location) \\
 :execution \> ( \> look-for-object ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
   \> \> (probabilistic 0.8 \\
   \>  \>(= (o-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
   \>  \> (probabilistic 0.1 \\
   \>  \> (= (o-is-in ?v) ?l))))) \\
\end{tabtt}
\normalsize

\noindent Here, the {\em execution} declaration is a lifted description of
the sense action-precondition -- i.e,
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
order to fully demonstrate the syntax. Interpreting the above schema,
if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
executed, there is a $0.8$ chance of perceiving the box if it is in
the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.

The syntax and semantics for describing initial state distributions in
DTPDDL is taken verbatim from PPDDL. It is useful to present that
factored tree-like structure here, in order to aid us when discussing
our switching continual planner. The distribution is expressed in a
tree-like structure of terms. Each term is either: (1) atomic, e.g., a
state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
probabilistic and atomic terms. The root term is always conjunctive,
and the leaves are atomic. For example, the starting distribution for
$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
that contain a single atomic subterm.}

\small
\begin{tabtt}
(\=:init (= (is-in DORA) kitchen) \+ \\
       (probabilistic \=.8 (= (is-in box) office)  \\
		      \>.2 (= (is-in box) kitchen)) \\
       (probabilistic .3 (= (is-in cup) office)  \\
		      \>.7 (= (is-in cup) kitchen))) \\
\end{tabtt}
\normalsize

\noindent The interpretation of such an expression can be given
by a {\em visitation} of terms: An atom is {\em visited} iff
its conjunctive parent is visited, and a conjunctive term is visited
iff all its immediate subterms are visited. A probabilistic term is
visited iff its conjunctive parent is visited, and exactly one of its
subterms, $T_i$, is visited. Each visitation of the root term
according to this recursive definition encapsulates a starting state,
along with the probability that it occurs. The former corresponds to the
union of all visited atoms, and the latter corresponds to the product
of $\prob_i$ entries on the visited subterms of probabilistic
elements. Making these ideas concrete, our example yields the
following flat state distribution:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
\hline
\end{tabular}
\normalsize

From hereon, to simplify the discussion (and implementation)
we shall restrict our attention to POMDPs with deterministic
actions. We note that POMDPs with stochastic
actions can be compiled into equivalent deterministic-action POMDPs,
where all the original action uncertainty is expressed in the
starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
that of finite-horizon planning, such a compilation yields a finite
flat representation of the original POMDP, only with deterministic
actions.

\section{Switching Continual Planner}

We describe our approach that switches between sequential and
contingent sessions. As a continual planning approach, it proceeds by
interleaving planning and execution in a deterministic-action POMDP
described in DTPDDL. During a sequential session, planning is
performed by a ``classical'' system,\footnote{That is, a planner
designed to solve fully observable deterministic tasks.}  and
execution proceeds according to the {\em trace} computed by that
system. Taking the form of a classical plan, the trace specifies a
sequence of POMDP actions that achieve the agent's objectives in a
deterministic approximation, i.e., {\em determinisation}, of the
problem at hand. More precisely, the trace is an interleaved sequence
of POMDP actions and {\em assumptive} actions. The latter correspond
to assumptions the planner makes about the truth value of propositions
-- e.g. that a box of cornflakes is located in the kitchen at the
third plan step. They are called {\em applicability} assumptions if
the trace includes an action $\action$ that is not applicable with
probability $1$ at the belief-state
\bstate\ that the system is projected to be in when \action\ is
scheduled for execution, i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. We denote
that action by \switchAction.  A contingent session then begins that
tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. 
%%
We add {\em disconfirm} and {\em confirm} actions to the model, so
the session is encouraged to judge the assumptions made in the
trace. On execution of one of those actions, control is returned to
the sequential session that continues at the current%%{\em underlying}
belief-state.
%%
%% For each assumption we add a {\em
%% disconfirm} action to the POMDP whose execution is rewarding
%% (resp. costly) if the assumption is false (resp. true). We also add
%% one {\em confirm} action that is rewarding (resp. costly) to execute
%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
%% contingent plan applies a disconfirm action, then a new sequential
%% session begins at the {\em underlying} belief-state. If the confirm
%% action is executed, the sequential session is resumed, and
%% \switchAction\ is executed. Otherwise, control rests in the continual
%% session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions plan in an abstract decision process
determined by the current trace. This abstraction is constructed by
first omitting all propositions that do not feature in the trace, and
by then iteratively refining the model while the result is of a
practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation of the
underlying belief-state corresponds more-or-less directly to an
$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
as the source of candidate determinisations for sequential planning,
(2) to determine when to switch, and (3) as a mechanism to guide
construction of an abstract process for contingent sessions.

%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.

%% \Omit{
%%   Therefore, although our current implementation does not
%%   explicitly support it, our approach generalises to POMDPs with
%%   stochastic action models.
%% }

\subsection{Sequential Planning in POMDPs}

A sequential session uses a classical planner to compute a trace. The
latter encapsulates assumptions about: (1) the true underlying state,
(2) how execution will progress, and (3) the possibility of the agent
eventually holding strong beliefs about the truth values of specific
state propositions. Here we describe the deterministic planning
problem, derived from the DTPDDL model, that admits plans which
correspond to traces.

In a deterministic-action POMDP all the uncertainty in state and
action is expressed in the $(\pp{:init})$ declaration. Our approach
uses the structure of that, as it occurs in the problem description,
to define a set of state-assumptions available to sequential planning.
%%
Writing \#\ if the value of a proposition is unspecified, for
$\pp{DORA}$ we have the following assumptions:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent An assumption corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or more of its immediate probabilistic subterms
are visited. The starting state, $\state_0$, for sequential planning
is an {\em abstract} state that captures the unique assumption which
has probability $1$. In $\pp{DORA}$, that is:

\[
\begin{array}{l}
\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
\end{array}
\]

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
conjunct, or it occurs in $T_k$ for some executed
$\assumptiveS{k}$. Also, an assumption about a proposition
$\prop$ can only be made once, and must be sequenced before any POMDP
actions that have $\prop$ in their precondition or effect.
%%
In other words, an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment can be modified by the DTPDDL domain operators.
%%
%%
Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$ with the
proposition denoting the undefined fact is deleted. For example, if
the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.
%%
Lastly, we have that if $\action$ has a precondition that is not true
with probability $\theta$, a given threshold, in the underlying belief
$\bstate$, then immediately before it is executed blindly the system
switches to a contingent session (described below).

%% For sequential planning the operators from the POMDP model are
%% available with semantics that accommodate the above abstraction as
%% follows: In the deterministic model a proposition $\prop$ can be
%% thought to have a ternary interpretation at a state \state, as either
%% {\em true}, written $\prop\in\state$, {\em false}, written
%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
%% $\action\in\actions$ in the deterministic model, if
%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
%% or negative effect of $\action$, then $\action$ is not applicable
%% in \state\ if $\prop\#\in\state$.

%% Addressing now the switching semantics of action execution. I
%% %% In detail, we halt the sequential
%% session at \action\ if $\exists
%% \state$ s.t. $\bstate(\state)>0$, 
%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
%% semantics of action execution there must be a $\state'$ s.t.
%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
%% not have been scheduled at \bstate.

%% then $\reward(\state,\action)=10$, and
%% otherwise if $\prop\not\in\state'$ then
%% $\reward(\state',\action)=-10$.

We now describe the optimisation criteria given to the classical
planner. Where $\prob_i$ is the probability that the $i^{th}$
sequenced action $\action_i$ does not transition to $\state^\bot$, we
define the value of a trace
$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
deterministic model to equal:

\begin{equation}\label{eq:tracevalue}
V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
\end{equation}

\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
actions from the trace, therefore has value:

\[
V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]

\noindent which is equal to the maximal 
additive contribution a trace of the plan can make to
Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
non-zero reward is received at the first transition to a goal state,
the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

\subsection{Decision-Theoretic Planning in Abstractions}

In a contingent session our switching planner solves a sensing problem
in an abstract processes defined in terms of the assumptive actions in
the current trace, and the $(\pp{:init})$ declaration that models the
underlying belief-state.

The focus for contingent planning is on sensing, we therefore give the
process in this session an augmented reward model.  This reflects the
value of performing sensing actions in the context of the trace
proposed by the preceding sequential deliberation. First, all rewards
from the original POMDP are retained. Then, for each $\assumptiveS{i}$
action scheduled by the current trace, we have a {\em dual}
$\assumptiveDT{i}$ so that $\forall\state\in\states$:

\[
\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
\hat\$(T_i) & \pp{otherwise} \\
\end{array}
\]

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
(negative) numeric quantity which captures the utility the agent
receives for correctly (incorrectly) rejecting an assumption. The
added dual actions are propositionally trivial -- i.e.,
$\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
%%
In order to reduce the number of possible rejections we omit
assumptions that are not {\em active} with respect to the action,
\switchAction, whose scheduled execution switched the system into
the contingent session. An assumption $\assumptiveS{i}$ is active if
\switchAction\ is not applicable in the sequential model unless
$\assumptiveS{i}$ is scheduled in the trace prefix
to \switchAction. For example, if $\pp{DORA}$'s trace is:

\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
(\pp{report-is-in}~\pp{box}~\pp{office}); \\
(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
\end{array}
\]

\noindent Taking the switching action \switchAction\ to be
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
active, and therefore exclude
$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
POMDP posed to the contingent session.

We also include another propositionally trivial action
$\actions.\poss(\switchAction)$ with the reward property:

\[
\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
\end{array}
\]

For the purposes of successive switching, execution of either a dual
action of the form $\assumptiveDT{i}$, or the
$\actions.\poss(\switchAction)$, returns control to a sequential
session. If a dual is executed, then the sequential planner must
replan with $\bstate_0$ equal to the underlying
belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
executed, then the current sequential plan is executed until further
sensing is scheduled, or to completion. Lastly, we note that in our
project the rewards, i.e., functions
$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
contingent session are sourced from a motivational component of our
robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
  sensing.}

%%  We make no assumptions here about how they are
%% derived.

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the active {\em assumptive} actions in the current
sequential plan. Figure \ref{fig:abstraction} shows the belief-state
of a more detailed version of our previous example. The diamonds
represent probabilistic terms of the state description while circles
represent conjunctive or atomic terms.

%% \begin{figure}[h!]
%%   \centering
%%   \tikzstyle{tree} = [sibling distance=4.5mm]
%%   \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
%%   \tikzstyle{seclevel} = [sibling distance=9mm]
%%   \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
%%   \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
%%   \tikzstyle{assumption} = [solid, very thick, draw=black]
%%   \tikzstyle{selected} = [solid, draw=black]
%%   \tikzstyle{unused} = [densely dashed, draw=black!40]
%%   \subfloat[The initial belief-state with the assumptions made by the continual
%%     planner in bold.]{
%%       \label{fig:abstraction-a}
%%       \begin{tikzpicture}[
%%     level 3/.style={tree}]
%%     \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
%%       child[seclevel] {node[cnode, assumption] (office) {}
%%         child {node [pnode, assumption] (box) {}
%%           child {node [cnode] (boxp1) {}}
%%           child {node [cnode, assumption] (boxp2) {}}
%%         }
%%         child {node [pnode] (cup) {} 
%%           child {node [cnode] (cupp1) {}}
%%           child {node [cnode] (cupp2) {}}
%%         }
%%       }
%%       child {node[cnode] (kitchen) {}
%%         child {node [pnode] (box2) {} 
%%           child {node [cnode] (box2p1) {}}
%%           child {node [cnode] (box2p2) {}}
%%         }
%%       };
%%    \tiny
%%    \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
%%    \node[above=0 of cat] {$\pp{(category room1)}$};
%%    \node[below=0 of kitchen] {$\pp{kitchen}$};
%%    \node[below=0 of office] {$\pp{office}$};
%%    \node[below=0 of box] {$\pp{(is-in box)}$};
%%    \node[below=0 of box2] {$\pp{(is-in box)}$};
%%    \node[below=0 of cup] {$\pp{(is-in cup)}$};
%%    \node[right=0 of boxp1] {$\pp{place1}$};
%%    \node[right=0 of boxp2] {$\pp{place2}$};
%%    \node[right=0 of box2p1] {$\pp{place1}$};
%%    \node[right=0 of box2p2] {$\pp{place2}$};
%%    \node[right=0 of cupp1] {$\pp{place1}$};
%%    \node[right=0 of cupp2] {$\pp{place2}$};
%%     % \node[node] (kitchen) right of (cat) {};
%%   \end{tikzpicture}}
%% \qquad
%%   \subfloat[Removing facts that are not part of an assumption.]{
%%     \label{fig:abstraction-b}
%%     \begin{tikzpicture}[
%%     level 3/.style={tree}]
%%     \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
%%       child[seclevel] {node[cnode, assumption] (office) {}
%%         child {node [pnode, assumption] (box) {}
%%           child {node [cnode, unused] (boxp1) {}}
%%           child {node [cnode, assumption] (boxp2) {}}
%%         }
%%         child {node [pnode, unused] (cup) {} 
%%           child {node [cnode, unused] (cupp1) {}}
%%           child {node [cnode, unused] (cupp2) {}}
%%         }
%%       }
%%       child {node[cnode, selected] (kitchen) {}
%%         child {node [pnode, selected] (box2) {} 
%%           child {node [cnode, unused] (box2p1) {}}
%%           child {node [cnode, selected] (box2p2) {}}
%%         }
%%       };
%%    \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
%%    \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
%%    \tiny
%%    \node[above=0 of cat] {$\pp{(category room1)}$};
%%    \node[below=0 of office] {$\pp{office}$};
%%    \node[below=0 of box] {$\pp{(is-in box)}$};
%%    \node[below=0 of box2] {$\pp{(is-in box)}$};
%%    \node[right=0 of boxp2] {$\pp{place2}$};
%%    \node[right=0 of box2p2] {$\pp{place2}$};
%%     % \node[node] (kitchen) right of (cat) {};
%%   \end{tikzpicture}}
%% \vspace{2mm}
%%   \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
%%     \label{fig:abstraction-c}
%%     \begin{tikzpicture}[
%%     level 3/.style={tree}]
%%     \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
%%       child[seclevel] {node[cnode, assumption] (office) {}
%%         child {node [pnode, assumption] (box) {}
%%           child {node [cnode, selected] (boxp1) {}}
%%           child {node [cnode, assumption] (boxp2) {}}
%%         }
%%         child {node [pnode, unused] (cup) {} 
%%           child {node [cnode, unused] (cupp1) {}}
%%           child {node [cnode, unused] (cupp2) {}}
%%         }
%%       }
%%       child {node[cnode, selected] (kitchen) {}
%%         child {node [pnode, selected] (box2) {} 
%%           child {node [cnode, selected] (box2p1) {}}
%%           child {node [cnode, selected] (box2p2) {}}
%%         }
%%       };
%%    \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
%%    \draw[selected] (box) -- (boxp1);
%%    \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
%%    \draw[selected] (box2) -- (box2p1);
%%    \tiny
%%    \node[above=0 of cat] {$\pp{(category room1)}$};
%%    \node[below=0 of office] {$\pp{office}$};
%%    \node[below=0 of box] {$\pp{(is-in box)}$};
%%    \node[below=0 of box2] {$\pp{(is-in box)}$};
%%    \node[right=0 of boxp1] {$\pp{place1}$};
%%    \node[right=0 of boxp2] {$\pp{place2}$};
%%    \node[right=0 of box2p1] {$\pp{place1}$};
%%    \node[right=0 of box2p2] {$\pp{place2}$};
%%     % \node[node] (kitchen) right of (cat) {};
%%   \end{tikzpicture}}

%%   \caption{Generating the abstract belief space.}
%% \label{fig:abstraction}
%% \end{figure}

%%% OLDER STUFF

% \small
% \begin{tabtt}
% (\=:init (= (is-in DORA) kitchen) \+ \\
%        (probabilistic \=.8 (= (is-in box) office)  \\
% 		      \>.2 (= (is-in box) kitchen))) \\
% \end{tabtt}
% \normalsize

% \noindent and therefore, the belief-state $\bstate_0$ is:

% \small
% \begin{tabular}{cccc}
% \hline
% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
% \hline
% .8 & kitchen & office & \# \\
% .2 & kitchen & kitchen & \# \\
% \hline
% \end{tabular}
% \normalsize

 % With
% regards to abstraction, the contingent and serial sessions apply the
% same applicability conditions ---i.e., that each effect and condition
% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
% problem description.

In the first step, we remove all facts that are not part of an
assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
would proceed in an abstraction of the environment that does not
contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
%%
In a second step, we iteratively refine the relaxed declaration by
adding terms from the original statement of $(\pp{:init})$ while the
number of abstract states in $\bstate_0$ that occur with non-zero
probability according to that refined declaration remains of a
practicable size. In detail, for each excluded atomic term, we compute
the {\em entropy} of the active assumptions, {\em conditional} on the
corresponding proposition.  Loosely speaking, the lower the {\em
  conditional entropy} of an atom, the more information about the
assumptions the contingent planner expect to derive from it. Thus,
propositions that in the first place were excluded are iteratively
added to the problem posed to the conditional session in increasing
order according to that measure (Fig. \ref{fig:abstraction-c}).

%% In detail, for each excluded atomic term, we compute the {\em
%% entropy} of the corresponding proposition, {\em conditional} on the
%% active assumptions. The {\em conditional entropy} measure
%% is higher the more uncertain that proposition is given truth
%% assignments to propositions that are the subject of active
%% assumptions. Propositions that in the first place were excluded are
%% iteratively added to the problem posed to the conditional session in
%% increasing order according to that measure.

%% In particular, we compute the {\em conditional entropy} of
%% each excluded proposition given the active assumptions. Propositions
%% are candidates for addition in increasing order according to that
%% measure.

%% Of course, it is expensive to measure size exactly,
%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
%% $\propositions^\#$ is the set of propositions in the abstraction that
%% were not assigned the {\em unknown} value \#. Terms are proposed for
%% addition because they are probabilistic, and because they were omitted
%% in the {\em relaxed} visitation characterised by {\em assumptive}
%% actions in the sequential plan. In each iteration, we add one
%% candidate term with the largest number of parent terms already in the
%% candidate tree.

\section{Evaluation}

We have implemented our switching continual approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09}. 
% Our implementation is able
% to use several underlying planning systems.
We have extended MAPSIM so that it can parse DTPDDL, and perform
successive estimation of the underlying belief-state using
\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
we use our own version of Fast Downward~\cite{fast-downward} for
sequential sessions. We have extended that system to support actions
with success probabilities. In our evaluation Fast Downward is run
with the cyclic causal graph heuristic using an A* search or weighted
A* (with weight 5). We use A* for easy problems and WA* for more
difficult problems where A* is ineffective. Contingent sessions use
our own forward search procedure.

\Omit{ We then perform multiple tests with different limits for the
  belief-space size in contingent sessions.  Higher limits should
  cause longer planning times but be beneficial to plan quality as
  more contingencies can be taken in to account by the POMDP planner.
}

In order to evaluate our system, we have also implemented a baseline
approach in MAPSIM. Rather than invoking a contingent session when a
switching action $\switchAction$ is scheduled for execution, the
baseline simply executes one action that can trigger a sensing outcome
determined by the precondition of $\switchAction$, and replans in the
resulting belief-state.

%% executes switching actions and replans in the
%% resulting belief state.

%% but instead of
%% creating an observation problem for the decision theoretic planner it
%% will just execute one sensing action -- assuming that this action will
%% confirm its assumption.

To test our approach we use a robot exploration domain based on a
scenario from our physical robotic system. Here, a robot is exploring
an office or living environment, and trying to report the locations of
objects to their owner. Basic types are {\tt rooms}, {\tt places} and
{\tt objects}. Places are topological map nodes that occur in rooms.
Objects, and also the robot, are always located at a place. The robot
can move around the rooms via connections between places given by a
{\tt connected} predicate. Each room has a, possibly unknown, {\em
  category} (e.g. kitchen, office, living room). Certain objects are
more likely to be located in rooms of a particular category.  For
example, a box of cornflakes is more likely to be located in the
kitchen than the office. The robot can look for an object at a place
by executing a {\tt look-for-object} action. Executing {\tt
  look-for-object} results in a noisy sensing outcome, sometimes
indicating a positive perception if the object is there. We have that
some objects are harder to detect than others. Finally, if the robot
is in the presence of a human, it can ask what type of room it is
currently in, however conducting a dialogue is more costly than
running a vision algorithm (cost of 8 vs costs of 3).

%%  Therefore, an instantaneous positive detection of an object is
%% not proof positive of it being there.

We evaluated our switching planner against the baseline in several
tasks with the number of rooms ranging from 3 to 6. We examine the
impact of sensing reliability, experimenting using sensor models that
are: ({\em reliable}) $0.9$ chance the object is perceived if it is
present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
chance is 0.4. The continual planning times are reported in
Figure~\ref{fig:results-time}, and the quality data in
Figure~\ref{fig:results-quality}. Here, in switching runs we examine
the performance of our system where $b_0$ allocates non-zero
probability to between 20 and 200 abstract states during contingent
sessions. For Item~$a$-$e$ the objective is to find one or more
objects and report their position to the user. In some of the tasks we
configured the environment so that no plan exists, by excluding the
desired object from the environment. Item~$f$ reports for tasks
requiring indirect sensing, where the robot must relocate to a
particular room category. For Item~$f$ there is no human in the
environment, and therefore the baseline cannot be used, because there
are no sensing schemata that directly sense room categories. In our
evaluation we run 50 simulations in each configuration, and have a
timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
  core.}. In our domain, rewards encapsulate a goal condition, and
therefore we find it intuitive to report average costs of the plans
and the success rate in problems that admit a solution (i.e., positive
reward scaled by a constant factor).

%% Because the switching utility model

%% Using a satisficing, optimistic serial planner in sequential sessions
%% makes optimising for the expected reward difficult. Only reasoning
%% about single traces, it is optimistic wrt the remaining costs to a
%% goal and pessimistic regarding the probability of reaching it. By
%% setting the rewards to non-extremal values has little effect on that
%% system. Therefore, we do not consider expected utility in our
%% experiments. Instead, we report average costs of the plans and the
%% success rate (as a ratio between solved and solvable problems).

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-time}\hfill
  \vspace{2mm}
  \includegraphics{dora56-time}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-time}\hfill
  \caption{Average runtime}
  \label{fig:results-time}
\end{figure}

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

%% Figure \ref{fig:results-quality} shows the average costs of the
%% executed plans as well as the percentage of solvable tasks that were
%% actually solved by the planner.

Examining success ratios and plan costs, where sensing is reliable
there is little to be gained by using the contingent planner, as the
greedy sensing approach of the baseline is sufficient. Not
surprisingly, as sensing degrades contingent planning pays off.  Also,
we find that time spent in contingent planning increases steeply as
the abstraction $\bstate_0$ becomes more refined.  That refinement
seems to be paying off in terms of the success ratio, particularly for
tasks $d$ and $e$, where we had sequential sessions using weighted
$A^*$ (rather than $A^*$). For less refined initial configurations,
the increase cost of contingent planning is compensated by a decrease
in Fast Downward planning times. The relatively high success rate
irrespective of the level of refinement in the initial configuration
indicates the effectiveness of using conditional entropy to guide
abstraction refinement in our setting.

%% :
%%while the resulting plans are still
%% longer on average, the impact on the number of solved tasks was much
%% smaller than for the baseline system.

%% Less aggressive abstraction of the initial configuration results in
%% longer runtimes, with little to be gained in terms of plan quality. We
%% conclude that in our scenario it is worthwhile being assumptively
%% aggressive.

%% \section{Discussion}
%% \input{discussion.tex}

\section{Related Work}
There have been a number of recent papers on
planning under uncertainty using systems that were intended for
sequential planning in deterministic problems,  for example,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition.
In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
plans and corresponding execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust sequential plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}.

%%No! They simply haven't been evaluated in PO settings. They may, or
%%may not struggle. They have sampling of traces, and that would
%%include observations, and therefore evolutions of beliefs. SSAT was
%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
%%suited to POMDPs.

% Normally if we are going to compare with related work, we do
% actually *compare*. Why didn't you try those approaches? I think
% it's safe to say that FFR will struggle. Why would it even include
% in its plan an observational action that doesn't change the world?

%%
%% However, as we said in the introduction,
%% all these approaches struggle in partially observable domains as they
%% rely on being able to determine the state at all times.

Also leveraging deterministic planners in problems that feature
uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
---i.e., sequential planning in unobservable worlds--- can be modelled
as a deterministic problem, and therefore solved using sequential
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. Most
recently this research thread has been extended to contingent planning
in fully observable non-deterministic
environments~\cite{albore:etal:2009}.
%%
The continual planning system that motivated our
project~\cite{brenner:nebel:jaamas09} also has this characteristic,
and has been applied in completely observable domains, particularly
those featuring multiple communicating agents.

%% The use of knowledge
%% operators in domains allows plans that act to gain knowledge, but the
%% approach assumes that such actions are deterministic and reliable, an
%% assumption that we relax.

Our work is motivated by domains that contain a mixture of task
planning and observation planning. There have been a number of recent
papers representing observation planning problems as POMDPs and using
various techniques to manage the large state
space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
approach in a vision algorithm selection problem. In their case there
is a natural hierarchical decomposition of the problem which allows
them to solve large problems by breaking them into a set of small
POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
represent a preference elicitation problem as a POMDP and take
advantage of symmetry in the belief space ---essentially the idea that
it does not matter what the value of the variable you are trying to
observe is--- to exponentially shrink the state space. Although we
have been actively exploring the \citeauthor{doshi08:pref_elic}
approach, we have yet to find those exploitable structures in our
domain due to the task planning requirement.

\section{Concluding Remarks}

From an automated planning perspective, the problem of practical
mobile robot control poses important and contrary challenges.
%%
On the one hand, planning and execution monitoring must be
lightweight, robust, timely, and should span the lifetime of the
robot. Those processes must seamlessly accommodate exogenous events,
changing objectives, and the underlying unpredictability of the
environment.
%%
On the other hand, robot planning should perform computationally
expensive reasoning about contingencies, and possible revisions of
subjective belief according to quantitatively modelled uncertainty in
acting and sensing.

In this paper we address these challenges, developing a continual
planner that switches between fast sequential and expensive
decision-theoretic planning. Given a POMDP model of the environment,
sequential planning is used to compute an initial deterministic
sequential plan and complementary runtime evolution of the decision
process. That plan is executed until testing of an assumption about
the runtime state is scheduled. At that point a contingent session
performs some sensing which determines whether or not the sequential
session should be allowed to continue, or otherwise acts to achieve
the overall objectives.

The most pressing item for future research, is to develop a scheme
whereby the serial planner can relax {\em executability} assumptions,
so that conformant (or semi-conformant) plans can be executed without
interruption by a contingent session. A more general criticism of
switching continual planning concerns interleaved sequential and
contingent sessions failing to make any progress towards the
objectives. For example, in the worst case we can have each sequential
session producing an identical plan, and each decision-theoretic
session rejecting it without further sensing. Although this has not
been an issue in our work so far, it must be dealt with rigorously
in the future. We suggest that a good way to mitigate this problem is
by developing a {\em motivational} component that maintains a {\em
dynamic} reward model whose limiting behaviour prevents that switching
deadlock.

%% The sequential mode of our system always schedules decision-theoretic
%% planning before executing actions that are not applicable with
%% certainty. This is inefficient if the utility of the plan is not
%% dependent on that actions successful execution. This situation always
%% arises for conformant plans.

%% The switching continual planning system we have described serves as
%% the underlying planner for CogX.

%% the effects of a {\em sense} schema are perceptual, whereas the
%% effects of an {\em operator} schema are over state propositions.

%% posted by a motivational component of the underlying robotic
%% architecture.

%% contingent sensory plans that are tailored to current
%% objectives.

%% In this paper we present an approach to continual planning that uses
%% two planning systems. The first

%%  to a distinct class of
%% challenges. We suppose 
%% %%
%% The underlying environment is modelled as a POMDP. We use the fast
%% classical satisficing system FastDownward to find a deterministic
%% sequential plan and complementary runtime evolution of that
%% process. This corresponds to a generalisation of replanning in
%% probabilistic planning to problems with partial observability.

%% Interaction between the sequential planner and execution proceeds
%% more-or-less analogously to popular replanning approaches

%% Addressing these challenges in a monolithic framework, we present a
%% {\em switching} continual planner, that uses the fast sequential
%% satisficing procedure FastDownward to perform net-benefit
%% %%
%% makes reasonable assumptions about the evolution of the runtime state
%% given a POMDP model of the environment.

%% contingent sensory plans that are tailored to current
%% objectives.

%% The decision-theoretic planner is able to tailor sensory processing on
%% a robot platform to the current objective, while FastDownward  quickly

%% In this paper we develop a continual planing system that uses two
%% planning systems. The first, is a state-of-the-art domain independent
%% planner for deterministic problems. The second is a information-state
%% contingency planning the information-state space of

%% Continual planning is a powerful technique that goes some way to
%% addressing those challenges. That approach interleaves planning and
%% execution, deliberately postponing planning for contingencies unless
%% they eventuate during execution.

%% computing a single sequential plan and
%% eventuality

%% To these challenges it seems a continual planning approach is best,

%% where the runtime state evolves during plan execution

%% interleaved planning and execution.

%% the latter is able to tailor sensory processing on a robot platform,
%% in order that it.

%% {\em ad-hoc}

%% , reason about degrees of belief and uncertainty about the world,

%% probabilistic sequential decision making in practical sized problems
%% is intractable.

%% quantitative probabilistic models of the perception and action .

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ijcai11}

\end{document}

### 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
Moritz, I put your diagram back into the system

--- 2011-01-19 subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
 
-\typeout{IJCAI-11 Instructions for Authors}
+\typeout{IJCAI Switching Planner Submission}
 
@@
 \usepackage{subfig}
-%% \usepackage{tikz}
-%% \usetikzlibrary{shapes}
-%% \usetikzlibrary{positioning}
+\usepackage{tikz}
+\usetikzlibrary{shapes}
+\usetikzlibrary{positioning}
 
@@
 
-%% \begin{figure}[h!]
-%%   \centering
-%%   \tikzstyle{tree} = [sibling distance=4.5mm]
-%%   \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-%%   \tikzstyle{seclevel} = [sibling distance=9mm]
-%%   \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-%%   \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-%%   \tikzstyle{assumption} = [solid, very thick, draw=black]
-%%   \tikzstyle{selected} = [solid, draw=black]
-%%   \tikzstyle{unused} = [densely dashed, draw=black!40]
-%%   \subfloat[The initial belief-state with the assumptions made by the continual
-%%     planner in bold.]{
-%%       \label{fig:abstraction-a}
-%%       \begin{tikzpicture}[
-%%     level 3/.style={tree}]
-%%     \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-%%       child[seclevel] {node[cnode, assumption] (office) {}
-%%         child {node [pnode, assumption] (box) {}
-%%           child {node [cnode] (boxp1) {}}
-%%           child {node [cnode, assumption] (boxp2) {}}
-%%         }
-%%         child {node [pnode] (cup) {} 
-%%           child {node [cnode] (cupp1) {}}
-%%           child {node [cnode] (cupp2) {}}
-%%         }
-%%       }
-%%       child {node[cnode] (kitchen) {}
-%%         child {node [pnode] (box2) {} 
-%%           child {node [cnode] (box2p1) {}}
-%%           child {node [cnode] (box2p2) {}}
-%%         }
-%%       };
-%%    \tiny
-%%    \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-%%    \node[above=0 of cat] {$\pp{(category room1)}$};
-%%    \node[below=0 of kitchen] {$\pp{kitchen}$};
-%%    \node[below=0 of office] {$\pp{office}$};
-%%    \node[below=0 of box] {$\pp{(is-in box)}$};
-%%    \node[below=0 of box2] {$\pp{(is-in box)}$};
-%%    \node[below=0 of cup] {$\pp{(is-in cup)}$};
-%%    \node[right=0 of boxp1] {$\pp{place1}$};
-%%    \node[right=0 of boxp2] {$\pp{place2}$};
-%%    \node[right=0 of box2p1] {$\pp{place1}$};
-%%    \node[right=0 of box2p2] {$\pp{place2}$};
-%%    \node[right=0 of cupp1] {$\pp{place1}$};
-%%    \node[right=0 of cupp2] {$\pp{place2}$};
-%%     % \node[node] (kitchen) right of (cat) {};
-%%   \end{tikzpicture}}
-%% \qquad
-%%   \subfloat[Removing facts that are not part of an assumption.]{
-%%     \label{fig:abstraction-b}
-%%     \begin{tikzpicture}[
-%%     level 3/.style={tree}]
-%%     \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-%%       child[seclevel] {node[cnode, assumption] (office) {}
-%%         child {node [pnode, assumption] (box) {}
-%%           child {node [cnode, unused] (boxp1) {}}
-%%           child {node [cnode, assumption] (boxp2) {}}
-%%         }
-%%         child {node [pnode, unused] (cup) {} 
-%%           child {node [cnode, unused] (cupp1) {}}
-%%           child {node [cnode, unused] (cupp2) {}}
-%%         }
-%%       }
-%%       child {node[cnode, selected] (kitchen) {}
-%%         child {node [pnode, selected] (box2) {} 
-%%           child {node [cnode, unused] (box2p1) {}}
-%%           child {node [cnode, selected] (box2p2) {}}
-%%         }
-%%       };
-%%    \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-%%    \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-%%    \tiny
-%%    \node[above=0 of cat] {$\pp{(category room1)}$};
-%%    \node[below=0 of office] {$\pp{office}$};
-%%    \node[below=0 of box] {$\pp{(is-in box)}$};
-%%    \node[below=0 of box2] {$\pp{(is-in box)}$};
-%%    \node[right=0 of boxp2] {$\pp{place2}$};
-%%    \node[right=0 of box2p2] {$\pp{place2}$};
-%%     % \node[node] (kitchen) right of (cat) {};
-%%   \end{tikzpicture}}
-%% \vspace{2mm}
-%%   \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
-%%     \label{fig:abstraction-c}
-%%     \begin{tikzpicture}[
-%%     level 3/.style={tree}]
-%%     \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-%%       child[seclevel] {node[cnode, assumption] (office) {}
-%%         child {node [pnode, assumption] (box) {}
-%%           child {node [cnode, selected] (boxp1) {}}
-%%           child {node [cnode, assumption] (boxp2) {}}
-%%         }
-%%         child {node [pnode, unused] (cup) {} 
-%%           child {node [cnode, unused] (cupp1) {}}
-%%           child {node [cnode, unused] (cupp2) {}}
-%%         }
-%%       }
-%%       child {node[cnode, selected] (kitchen) {}
-%%         child {node [pnode, selected] (box2) {} 
-%%           child {node [cnode, selected] (box2p1) {}}
-%%           child {node [cnode, selected] (box2p2) {}}
-%%         }
-%%       };
-%%    \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-%%    \draw[selected] (box) -- (boxp1);
-%%    \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-%%    \draw[selected] (box2) -- (box2p1);
-%%    \tiny
-%%    \node[above=0 of cat] {$\pp{(category room1)}$};
-%%    \node[below=0 of office] {$\pp{office}$};
-%%    \node[below=0 of box] {$\pp{(is-in box)}$};
-%%    \node[below=0 of box2] {$\pp{(is-in box)}$};
-%%    \node[right=0 of boxp1] {$\pp{place1}$};
-%%    \node[right=0 of boxp2] {$\pp{place2}$};
-%%    \node[right=0 of box2p1] {$\pp{place1}$};
-%%    \node[right=0 of box2p2] {$\pp{place2}$};
-%%     % \node[node] (kitchen) right of (cat) {};
-%%   \end{tikzpicture}}
-
-%%   \caption{Generating the abstract belief space.}
-%% \label{fig:abstraction}
-%% \end{figure}
+\begin{figure}[h!]
+  \centering
+  \tikzstyle{tree} = [sibling distance=4.5mm]
+  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
+  \tikzstyle{seclevel} = [sibling distance=9mm]
+  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
+  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
+  \tikzstyle{assumption} = [solid, very thick, draw=black]
+  \tikzstyle{selected} = [solid, draw=black]
+  \tikzstyle{unused} = [densely dashed, draw=black!40]
+  \subfloat[The initial belief-state with the assumptions made by the continual
+    planner in bold.]{
+      \label{fig:abstraction-a}
+      \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode] (cup) {} 
+          child {node [cnode] (cupp1) {}}
+          child {node [cnode] (cupp2) {}}
+        }
+      }
+      child {node[cnode] (kitchen) {}
+        child {node [pnode] (box2) {} 
+          child {node [cnode] (box2p1) {}}
+          child {node [cnode] (box2p2) {}}
+        }
+      };
+   \tiny
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of kitchen] {$\pp{kitchen}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[below=0 of cup] {$\pp{(is-in cup)}$};
+   \node[right=0 of boxp1] {$\pp{place1}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p1] {$\pp{place1}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+   \node[right=0 of cupp1] {$\pp{place1}$};
+   \node[right=0 of cupp2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+\qquad
+  \subfloat[Removing facts that are not part of an assumption.]{
+    \label{fig:abstraction-b}
+    \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode, unused] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode, unused] (cup) {} 
+          child {node [cnode, unused] (cupp1) {}}
+          child {node [cnode, unused] (cupp2) {}}
+        }
+      }
+      child {node[cnode, selected] (kitchen) {}
+        child {node [pnode, selected] (box2) {} 
+          child {node [cnode, unused] (box2p1) {}}
+          child {node [cnode, selected] (box2p2) {}}
+        }
+      };
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+   \tiny
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+\vspace{2mm}
+  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
+    \label{fig:abstraction-c}
+    \begin{tikzpicture}[
+    level 3/.style={tree}]
+    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
+      child[seclevel] {node[cnode, assumption] (office) {}
+        child {node [pnode, assumption] (box) {}
+          child {node [cnode, selected] (boxp1) {}}
+          child {node [cnode, assumption] (boxp2) {}}
+        }
+        child {node [pnode, unused] (cup) {} 
+          child {node [cnode, unused] (cupp1) {}}
+          child {node [cnode, unused] (cupp2) {}}
+        }
+      }
+      child {node[cnode, selected] (kitchen) {}
+        child {node [pnode, selected] (box2) {} 
+          child {node [cnode, selected] (box2p1) {}}
+          child {node [cnode, selected] (box2p2) {}}
+        }
+      };
+   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+   \draw[selected] (box) -- (boxp1);
+   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+   \draw[selected] (box2) -- (box2p1);
+   \tiny
+   \node[above=0 of cat] {$\pp{(category room1)}$};
+   \node[below=0 of office] {$\pp{office}$};
+   \node[below=0 of box] {$\pp{(is-in box)}$};
+   \node[below=0 of box2] {$\pp{(is-in box)}$};
+   \node[right=0 of boxp1] {$\pp{place1}$};
+   \node[right=0 of boxp2] {$\pp{place2}$};
+   \node[right=0 of box2p1] {$\pp{place1}$};
+   \node[right=0 of box2p2] {$\pp{place2}$};
+    % \node[node] (kitchen) right of (cat) {};
+  \end{tikzpicture}}
+
+  \caption{Generating the abstract belief space.}
+\label{fig:abstraction}
+\end{figure}
 

### 2011-01-20 subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
New title and abstract

@@
 
-\title{A Switching Planner for Combined Task and Observation Planning}
+\title{Interleaved Serial and Decision-Theoretic Planning for a Mobile Robot}
+%%\title{Interleaved Serial and Decision-Theoretic Planning for
+%%Practical Robot Control}
+
+%% \title{A Switching Planner for Combined Task and Observation Planning}
 \author{anonymous}
@@
 
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
+%% Those processes must
+
+From an automated planning perspective the problem of practical mobile
+robot control in realistic environments poses many important and
+contrary challenges.  On the one hand, the planning process must be
+lightweight, robust, and timely. Over the lifetime of the robot it
+must always respond quickly with new plans that accommodate exogenous
+events, changing objectives, and the underlying unpredictability of
+the environment.  On the other hand, in order to exhibit useful
+behaviours the planning process must perform computationally expensive
+reasoning about contingencies and possible revisions of subjective
+beliefs according to quantitatively modelled uncertainty in acting and
+sensing.
+%%
+Addressing these challenges, we develop a \emph{continual planning}
+approach that {\em switches} between using a fast satisficing
+``classical'' planner, to decide on the overall strategy, and
+decision-theoretic planning to solve small subproblems where deeper
+consideration of the sensing model is practical, and can significantly
+impact overall performance. We demonstrate our approach in a realistic
+robot exploration domain.
+
+%% partial observability
+
+%% In cooperation with execution monitoring, it must perform repair
+%% and/or replanning in order to seamlessly accommodate exogenous events,
+%% changing objectives, and the underlying unpredictability of the
+%% environment.
+
+%% Realistic robot planning problems in uncertain environments often
+%% require achieving tasks while also finding out about the
+%% world. Because the world state cannot be observed directly, these
+%% problems are naturally represented as partially observable Markov
+%% decision problems (POMDPs). However, these are typically intractable
+%% for realistic problems.
+%%
 

### 2011-01-20 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
different title

@@
 
-\title{A Switching Planner for Combined Task and Observation Planning}
+%% \title{Interleaved Serial and Decision-Theoretic Planning for a Mobile Robot}
+
+\title{Interleaved Serial and Decision-Theoretic Planning for
+Practical Robot Control}
+
+%% \title{A Switching Planner for Combined Task and Observation Planning}
 \author{anonymous}

### 2011-01-20 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
added the word "large"

@@
 %%
-Addressing these challenges, we develop a \emph{continual planning}
-approach that {\em switches} between using a fast satisficing
-``classical'' planner, to decide on the overall strategy, and
-decision-theoretic planning to solve small subproblems where deeper
-consideration of the sensing model is practical, and can significantly
-impact overall performance. We demonstrate our approach in a realistic
-robot exploration domain.
+Towards addressing these challenges, we develop a \emph{continual
+planning} approach that {\em switches} between using a fast
+satisficing ``classical'' planner, to decide on the overall strategy,
+and decision-theoretic planning to solve small subproblems where
+deeper consideration of the sensing model is practical, and can
+significantly impact overall performance. We demonstrate our approach
+in large problems from a realistic robot exploration domain.
 

### 2011-01-20 subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
clarification to englishizisationXX

@@
 
-\title{Interleaved Serial and Decision-Theoretic Planning for a Mobile Robot}
-%%\title{Interleaved Serial and Decision-Theoretic Planning for
-%%Practical Robot Control}
+%% \title{Interleaved Serial and Decision-Theoretic Planning for a Mobile Robot}
+
+\title{Interleaved Serial and Decision-Theoretic Planning for
+Practical Robot Control}
 
@@
 %%
-Addressing these challenges, we develop a \emph{continual planning}
-approach that {\em switches} between using a fast satisficing
-``classical'' planner, to decide on the overall strategy, and
-decision-theoretic planning to solve small subproblems where deeper
-consideration of the sensing model is practical, and can significantly
-impact overall performance. We demonstrate our approach in a realistic
-robot exploration domain.
+Towards addressing these challenges, we develop a \emph{continual
+planning} approach that {\em switches} between using a fast
+satisficing ``classical'' planner, to decide on the overall strategy,
+and decision-theoretic planning to solve small subproblems where
+deeper consideration of the sensing model is both practical, and can
+significantly impact overall performance. We demonstrate our approach
+in large problems from a realistic robot exploration domain.
 

### 2011-01-28 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
AAAI version of the switching planner paper

--- 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
-%%%% ijcai11.tex
-
-\typeout{IJCAI Switching Planner Submission}
-
-% These are the instructions for authors for IJCAI-11.
-% They are the same as the ones for IJCAI-07 with superficical wording
-%   changes only.
-
-\documentclass{article}
-% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
-\usepackage{ijcai11}
-
-% Use the postscript times font!
-\usepackage{times}
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\usepackage{subfig}
-\usepackage{tikz}
-\usetikzlibrary{shapes}
-\usetikzlibrary{positioning}
-
-\usepackage{graphicx}
-\graphicspath{{plots/}}
-\include{macros}%
-
-\title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Planning for agents acting in partially observable environments with
-stochastic actions is extremely computationally
-expensive~\cite{mdp-complexity}. Even relatively small partially
-observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, have state spaces far larger than existing
-optimal or close to optimal planners can handle. For the easier
-problem of planning in completely observable stochastic domains, an
-alternative has been to use classical sequential planners and
-replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
-Execution monitoring tracks properties of the state that determine
-whether further planning is required. For partially observable
-domains, only probabilistic information about such properties will, in
-general, be available.
-
-%% However, these approaches rely
-%% crucially on being able to determine properties of the current state
-%% in order to decide to replan. For partially observable domains, only
-%% probabilistic information about these properties will in general be
-%% available.
-
-This paper describes the planning component for a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings. This domain features partial observability, particularly
-because the state is not perfectly known, and state information is
-gained by the robot through the use of unreliable sensing
-actions. 
-%Using a propositionally factored state representation, for
-For
-interesting sized tasks the corresponding POMDP model is too large for a
-POMDP solver to be applied directly. Inspired by the replanning
-approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
-planning approach that uses a classical planning system to compute a
-reasonably valuable trace in the model, and then uses a
-decision-theoretic planner on small subproblems where reasoning about
-observations might be useful. We refer to the complete system as a
-{\em switching continual planner}. The approach is not optimal,
-particularly as it relies on the results of sequential planning
-directly. It does nevertheless perform better than a purely sequential
-replanner, and is fast enough to be used for real-time decision-making
-on a mobile robot.
-
-%% Switching gives us some important advantages in our robotic
-%% domain. First, the significant source of uncertainty in the domain is
-%% the unreliability of observations, and this approach is particularly
-%% suited to problems that combine a deterministic task planning problem
-%% with observations as the decision-theoretic planner is often invoked
-%% to determine a characteristic of the underlying state. Secondly,
-%% replanning makes our system robust to changing objectives and the
-%% discovery of new facts about the world.
-
-%% As an illustration, consider a robot searching for a box that it knows
-%% is often found in offices. The sequential planner might produce a plan
-%% that assumes that {\sc Room a} is an office, that the box is in {\sc
-%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
-%% these assumptions, the plan might be to go to {\sc Room a}, go to
-%% location 1 and search for the box. When the plan gets executed and the
-%% search begins the decision-theoretic planner is called since searching
-%% involves making observations. The decision-theoretic planner might
-%% retract the last assumption but leave the others, and build a plan to
-%% search the whole of {\sc Room a} for the box. If this plan fails to
-%% find the box, it will disprove the assumption that the box is in {\sc
-%% Room a}, which will then lead to replanning by the sequential planner
-%% based on the newly discovered knowledge.
-
-Our system is domain independent, taking domain and problem
-descriptions in a first-order declarative language we have developed,
-called the decision-theoretic planning domain definition language
-(DTPDDL). From these descriptions of POMDPs, we automatically
-construct a deterministic model suitable for sequential planners. That
-model includes actions which correspond to making assumptions about
-the values of imperfectly known state variables. Assumptions scheduled
-by the sequential system are used to propose a pragmatic abstract
-belief-state space to the decision-theoretic system, and to modify the
-reward function, so that system might pursue sensing related to those
-assumptions.
-
-%%  which can operate
-%% with the original representation of the domain.
-
-%% The sequential planner we use is based
-%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
-%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
-%% knowledge to be represented explicitly, so we can write actions that
-%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
-%% THIS. For the decision-theoretic planner we have implemented our own
-%% forward search in the belief-state space.
-
-The switching planner provides important advantages in our
-mobile robot domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations because the decision-theoretic planner
-is only invoked to deal with state uncertainty.
-%determine a characteristic of the underlying state. 
-Secondly, replanning makes the system robust to
-changing objectives and discoveries about the world,
-both of which feature in our domain.
-
-The remainder of the paper proceeds as follows. We begin by describing
-POMDPs with a propositionally factored representation of states and
-observations. We also outline the relationship between this and a flat
-POMDP representation. Next we describe our DTPDDL language with
-example declarations from a mobile robot exploration task. Then we
-describe the switching planner, including how the sequential domain is
-produced and a plan is generated, and how this then provides the input
-for the decision-theoretic planner. Finally we provide an evaluation
-of the system on the example robot domain, and finish with a short
-survey of related work and some concluding remarks.
-
-%To accompany
-%that language we have implemented an information-state
-%\laostar\ procedure for solving problems expressed in DTPDDL.
-
-%%
-
-%Have a good understanding of the contents and function of rooms, as
-%well as the linguistic referrents to rooms, widgets, and their visual
-%qualities.
-
-%Having confidence in its beliefs about the linguistic terms for
-%spaces, and the function of those spaces.
-
-%sometimes difficult to predict, with exogenous events, such as the
-%changing of the objective,
-
-%the speed and scalability of software for sequential planning in
-%deterministic
-
-%% We developed a first-order declarative language, called DTPDDL, for
-%% describing domains of planning problems that correspond to POMDPs.  To
-%% accompany that language we have implemented an information-state
-%% \laostar\ procedure for solving problems expressed in DTPDDL.
-
-%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
-%% modified \fastdownward\ so that it can find useful sequential executions
-%% given DTPDDL models of the problem at hand.
-
-%%  called DTPDDL, along
-%% the lines of PDDL for the partially observable case, an \laostar\
-%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
-%% and modify the \fastdownward\ system to find high quality sequential plans
-%% .
-
-%% We model the environment as a partially observable Markov decision
-%% process. Although planning in that model is undecidable in
-%% general~\cite{}, an optimal finite-horizon plan corresponds to a
-%% contingent plan, that is a function mapping action-observation
-%% histories to actions.
-
-%% our continual planner is reactive, replanning whenever the underlying
-%% domain and problem models change. For example, replanning occurs if
-%% the motivational component alters the objectives, and if an assumed
-%% outcome of a sensing action is not realised.
-
-%% brittle sensing model,
-
-%% the evaluation of a fluent at a state is either known. Moreover, there
-%% is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process state
-$\state$ as the set of propositions that are true of the
-state. Notationally, $\props$ is the set of propositions, $\prop$ is
-an element from that set, and we have $\state \subseteq
-\props$. The underlying process
-dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
-state-characterising propositions $\propositions$.
-%%
-We say an action $\stochAction \in \stochActions$ is
-applicable if its precondition $\poss(\stochAction)$, a set of
-propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature chooses a deterministic STRIPS effect $\detAct_i$, and for
-all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$.
-
-%% We model a process state $\state$ as the set of propositions that are
-%% true of the state. Notationally, we have $\state \subseteq
-%% \props$. State change is induced by application of actions. A
-%% stochastic action $\stochAction \in \stochActions$ is applicable if
-%% its precondition $\poss(\stochAction)$, a set of propositions, are
-%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
-%% applicable at state $\state$.  If action $\stochAction \in
-%% \stochActions(\state)$ is applied at \state, nature chooses one
-%% element amongst a small set of deterministic outcomes
-%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-%% $\sum_{\detAction_i \in \detActions(\stochAction)}
-%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-%% effect on the state given in terms of two lists of propositions. The
-%% add-list $\add(\detAction)$ and delete-list
-%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-%% $\detAction$, then it cannot be in the delete-list and vice versa.}
-%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
-%% applied at state $\state$, then the resultant state is $ ( \state \cup
-%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-%% propositions from $\add(\detAction)$ are added to $\state$, and those
-%% from $\delete(\detAction)$ are removed from $\state$.
-
-%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
-
-We are concerned with problems that feature partial
-observability. Although we could invoke {\em extended probabilistic
-STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for presentation and computation to
-separate sensing and action. Therefore, we suppose a POMDP has a
-perceptual model given in terms of a finite set of stochastic {\em
-senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
-and perceptual propositions $\percepts$, called {\em percepts}. In
-detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote by \observations\ the set of
-reachable observations. The underlying state of the process cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in
-\observations$ that informs what should be believed about the state the
-process is in. If $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
-usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from the chosen elements of active
-senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make some fairly standard assumptions. First, that action
-execution and sensing occur instantaneously, and that only one action
-can be applied per plan-step. Secondly, it can arise that in some
-states an action is applicable, and in others it is not. Moreover, a
-belief-state can assign non-zero probability to states where that
-applicability holds, and states where it does not. %% We differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
-%% In beliefs where an action exists that is applicable in some state 
-%% with non-zero probability, and not applicable in another, we differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%% because 
-%% in that
-%% we forbid the execution of the action.
-%% actions that are not applicable in
-%% any state of the current belief.  
-%% Also,
-Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we incorporate a PPDDL-like default semantics, treating an action
-\stochAction\ executed at $\state_i$ as if it has no effect at
-states $\state$ if $\poss(\stochAction) \not\subseteq
-\state$.
-
-% Mucked around with this, then realised the two sentences were
-% contradictory. If we forbid actions applicable in one state but not
-% another in the belief state, then clearly we aren't using a
-% PPDDL-like semantics.
-%% Doubled comments are stuff I added, then removed.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Until now we have discussed the POMDP in terms of propositions and
-percepts. In order to address belief revision and utility it is
-convenient to consider the underlying decision process in a flat
-format. This is given by the tuple
-$\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-The function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs as
-negative rewards.
-%%
-From $\psi$ we have that for each $\state' \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
-\action)$ the probability of getting observation $\observ$ in state
-$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state', \action) = 1$.
-
-Successive state estimation is by application of Bayes'
-rule.  Taking the current belief $\bstate$ as the {\em prior}, and
-supposing action $\action$ is executed with perceptive outcome
-$\observ$, the probability that we are in $\state'$ in the
-successive belief-state $\bstate'$ is:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
-  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---i.e., reaches a state that satisfies a given
-Boolean condition--- with probability $1$.  Generally, whatever the
-structure of a plan, its value corresponds to the expected cumulative
-reward:
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
-%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
-%% probability that the controller prescribes \action\ when in state
-%% \node, and $\transition_\node(\action, \observation, \node') =
-%% P(\node'|\node, \action, \observation)$ gives the probability the
-%% controller transitions to state $\node'$, supposing we execute
-%% \action\ in state \node\ receiving observation \observation. When
-%% acting for finite $N$ steps from controller state $n$, the value of a
-%% controller corresponds to the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
-the action prescribed at step $t$, and
-
-\[\reward(\bstate, \action) = \sum_{\state \in \states}
-\bstate(\state)\reward(\state, \action).\]
-
-%% \noindent Finally, it is
-%% useful to note that there is a corresponding deterministic FSCs for
-%% both sequential (e.g., conformant) and contingent plan formats.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
-used in all three of the International Planning Competitions since
-2004. A variation on PDDL for describing domains with stochastic
-actions and uncertain starting configurations, PPDDL is a declarative
-first-order language that facilitates factored descriptions of domains
-and problems. There are straightforward compilations from problems
-expressed in PPDDL to propositional representations amenable
-to state-of-the-art planners.
-
-Because PPDDL cannot model domains that feature partial observability,
-we develop an extension we call {\em Decision-Theoretic
-(DT)PDDL}. This can express probabilistic models of the sensing
-consequences of acting, to quantitatively capture unreliability in
-perception. That expressive power is achieved by incorporating
-perceptual analogues of fluent, predicate, and action definitions. In
-detail, we have declarations of state characterising predicate and
-fluent symbols according to the PPDDL syntax. In addition, we allow
-two other declarations, for perceptual predicates and fluents
-respectively. For example, suppose our robot is tasked with exploring
-{\em locations} in order to identify the whereabouts of a {\em
-visual-object}. We must describe state and perceptual facts that model
-the {\em true}, resp. perceived, locations of objects. In DTPDDL,
-these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like declarations, with preconditions expressed using state
-predicate and action function symbols, and uniformly positive effects
-over perceptual predicates. A sense declaration corresponds to a
-lifted description of proposition {\em senses}. Whereas the effects of
-an {\em operator} schema describe how states change under application
-of actions, the effects of a {\em sense} schema are perceptual,
-specifying the composition of an observation following the execution
-of an action.
-
-%% Also, and without a loss of generality, we find it
-%% convenient to separate the actional precondition from the state
-%% precondition.
-
-Making the above ideas concrete with an example, suppose a robot
-called $\pp{DORA}$ is able to look for a visual-object, such as a
-$\pp{box}$, at a given place. We can model that deterministic action
-using the following operator schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs an instantaneous penalty, $3$, that corresponds to
-the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
-reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, reward is
-accumulated, whereas in DTPDDL it is instantaneous.}
-
-Completing the example, we model the sensing that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- i.e,
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
-$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-our switching continual planner. The distribution is expressed in a
-tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
-(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
-probabilistic and atomic terms. The root term is always conjunctive,
-and the leaves are atomic. For example, the starting distribution for
-$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
-$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
-i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
-$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
-that contain a single atomic subterm.}
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-by a {\em visitation} of terms: An atom is {\em visited} iff
-its conjunctive parent is visited, and a conjunctive term is visited
-iff all its immediate subterms are visited. A probabilistic term is
-visited iff its conjunctive parent is visited, and exactly one of its
-subterms, $T_i$, is visited. Each visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability that it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $\prob_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-From hereon, to simplify the discussion (and implementation)
-we shall restrict our attention to POMDPs with deterministic
-actions. We note that POMDPs with stochastic
-actions can be compiled into equivalent deterministic-action POMDPs,
-where all the original action uncertainty is expressed in the
-starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
-that of finite-horizon planning, such a compilation yields a finite
-flat representation of the original POMDP, only with deterministic
-actions.
-
-\section{Switching Continual Planner}
-
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
-
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
-%%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
-
-Finally, whether proceeding in a sequential or contingent session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
-
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
-
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
-
-\subsection{Sequential Planning in POMDPs}
-
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
-
-\noindent An assumption corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
-
-\[
-\begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
-\end{array}
-\]
-
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
-%%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
-%%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
-
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
-
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
-
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
-
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
-\[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
-\reward(\state_i, \action_i),
-\]
-
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
-
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
-\[
-\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
-\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\hat\$(T_i) & \pp{otherwise} \\
-\end{array}
-\]
-
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
-%%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
-
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
-\end{array}
-\]
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
-
-We also include another propositionally trivial action
-$\actions.\poss(\switchAction)$ with the reward property:
-
-\[
-\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
-\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
-\end{array}
-\]
-
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
-
-%%  We make no assumptions here about how they are
-%% derived.
-
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
-
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief-state with the assumptions made by the continual
-    planner in bold.]{
-      \label{fig:abstraction-a}
-      \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode] (cup) {} 
-          child {node [cnode] (cupp1) {}}
-          child {node [cnode] (cupp2) {}}
-        }
-      }
-      child {node[cnode] (kitchen) {}
-        child {node [pnode] (box2) {} 
-          child {node [cnode] (box2p1) {}}
-          child {node [cnode] (box2p2) {}}
-        }
-      };
-   \tiny
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of kitchen] {$\pp{kitchen}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[below=0 of cup] {$\pp{(is-in cup)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-   \node[right=0 of cupp1] {$\pp{place1}$};
-   \node[right=0 of cupp2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Generating the abstract belief space.}
-\label{fig:abstraction}
-\end{figure}
-
-%%% OLDER STUFF
-
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
-
-\section{Evaluation}
-
-We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. 
-% Our implementation is able
-% to use several underlying planning systems.
-We have extended MAPSIM so that it can parse DTPDDL, and perform
-successive estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
-
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
-}
-
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
-
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
-
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
-
-We evaluated our switching planner against the baseline in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
-  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
-  core.}. In our domain, rewards encapsulate a goal condition, and
-therefore we find it intuitive to report average costs of the plans
-and the success rate in problems that admit a solution (i.e., positive
-reward scaled by a constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
-  \caption{Average runtime}
-  \label{fig:results-time}
-\end{figure}
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
-  \caption{Average plan costs and number of successful runs.}
-  \label{fig:results-quality}
-\end{figure}
-
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% \section{Discussion}
-%% \input{discussion.tex}
-
-\section{Related Work}
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
-
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-%%
-The continual planning system that motivated our
-project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-and has been applied in completely observable domains, particularly
-those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.
-
-\section{Concluding Remarks}
-
 From an automated planning perspective, the problem of practical
@@
 %% quantitative probabilistic models of the perception and action .
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai11}
-
-\end{document}

### 2011-01-28 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
AAAI version of the switching planner paper

--- 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
-%%%% ijcai11.tex
-
-\typeout{IJCAI Switching Planner Submission}
-
-% These are the instructions for authors for IJCAI-11.
-% They are the same as the ones for IJCAI-07 with superficical wording
-%   changes only.
-
-\documentclass{article}
-% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
-\usepackage{ijcai11}
-
-% Use the postscript times font!
-\usepackage{times}
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\usepackage{subfig}
-\usepackage{tikz}
-\usetikzlibrary{shapes}
-\usetikzlibrary{positioning}
-
-\usepackage{graphicx}
-\graphicspath{{plots/}}
-\include{macros}%
-
-\title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Planning for agents acting in partially observable environments with
-stochastic actions is extremely computationally
-expensive~\cite{mdp-complexity}. Even relatively small partially
-observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, have state spaces far larger than existing
-optimal or close to optimal planners can handle. For the easier
-problem of planning in completely observable stochastic domains, an
-alternative has been to use classical sequential planners and
-replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
-Execution monitoring tracks properties of the state that determine
-whether further planning is required. For partially observable
-domains, only probabilistic information about such properties will, in
-general, be available.
-
-%% However, these approaches rely
-%% crucially on being able to determine properties of the current state
-%% in order to decide to replan. For partially observable domains, only
-%% probabilistic information about these properties will in general be
-%% available.
-
-This paper describes the planning component for a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings. This domain features partial observability, particularly
-because the state is not perfectly known, and state information is
-gained by the robot through the use of unreliable sensing
-actions. 
-%Using a propositionally factored state representation, for
-For
-interesting sized tasks the corresponding POMDP model is too large for a
-POMDP solver to be applied directly. Inspired by the replanning
-approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
-planning approach that uses a classical planning system to compute a
-reasonably valuable trace in the model, and then uses a
-decision-theoretic planner on small subproblems where reasoning about
-observations might be useful. We refer to the complete system as a
-{\em switching continual planner}. The approach is not optimal,
-particularly as it relies on the results of sequential planning
-directly. It does nevertheless perform better than a purely sequential
-replanner, and is fast enough to be used for real-time decision-making
-on a mobile robot.
-
-%% Switching gives us some important advantages in our robotic
-%% domain. First, the significant source of uncertainty in the domain is
-%% the unreliability of observations, and this approach is particularly
-%% suited to problems that combine a deterministic task planning problem
-%% with observations as the decision-theoretic planner is often invoked
-%% to determine a characteristic of the underlying state. Secondly,
-%% replanning makes our system robust to changing objectives and the
-%% discovery of new facts about the world.
-
-%% As an illustration, consider a robot searching for a box that it knows
-%% is often found in offices. The sequential planner might produce a plan
-%% that assumes that {\sc Room a} is an office, that the box is in {\sc
-%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
-%% these assumptions, the plan might be to go to {\sc Room a}, go to
-%% location 1 and search for the box. When the plan gets executed and the
-%% search begins the decision-theoretic planner is called since searching
-%% involves making observations. The decision-theoretic planner might
-%% retract the last assumption but leave the others, and build a plan to
-%% search the whole of {\sc Room a} for the box. If this plan fails to
-%% find the box, it will disprove the assumption that the box is in {\sc
-%% Room a}, which will then lead to replanning by the sequential planner
-%% based on the newly discovered knowledge.
-
-Our system is domain independent, taking domain and problem
-descriptions in a first-order declarative language we have developed,
-called the decision-theoretic planning domain definition language
-(DTPDDL). From these descriptions of POMDPs, we automatically
-construct a deterministic model suitable for sequential planners. That
-model includes actions which correspond to making assumptions about
-the values of imperfectly known state variables. Assumptions scheduled
-by the sequential system are used to propose a pragmatic abstract
-belief-state space to the decision-theoretic system, and to modify the
-reward function, so that system might pursue sensing related to those
-assumptions.
-
-%%  which can operate
-%% with the original representation of the domain.
-
-%% The sequential planner we use is based
-%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
-%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
-%% knowledge to be represented explicitly, so we can write actions that
-%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
-%% THIS. For the decision-theoretic planner we have implemented our own
-%% forward search in the belief-state space.
-
-The switching planner provides important advantages in our
-mobile robot domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations because the decision-theoretic planner
-is only invoked to deal with state uncertainty.
-%determine a characteristic of the underlying state. 
-Secondly, replanning makes the system robust to
-changing objectives and discoveries about the world,
-both of which feature in our domain.
-
-The remainder of the paper proceeds as follows. We begin by describing
-POMDPs with a propositionally factored representation of states and
-observations. We also outline the relationship between this and a flat
-POMDP representation. Next we describe our DTPDDL language with
-example declarations from a mobile robot exploration task. Then we
-describe the switching planner, including how the sequential domain is
-produced and a plan is generated, and how this then provides the input
-for the decision-theoretic planner. Finally we provide an evaluation
-of the system on the example robot domain, and finish with a short
-survey of related work and some concluding remarks.
-
-%To accompany
-%that language we have implemented an information-state
-%\laostar\ procedure for solving problems expressed in DTPDDL.
-
-%%
-
-%Have a good understanding of the contents and function of rooms, as
-%well as the linguistic referrents to rooms, widgets, and their visual
-%qualities.
-
-%Having confidence in its beliefs about the linguistic terms for
-%spaces, and the function of those spaces.
-
-%sometimes difficult to predict, with exogenous events, such as the
-%changing of the objective,
-
-%the speed and scalability of software for sequential planning in
-%deterministic
-
-%% We developed a first-order declarative language, called DTPDDL, for
-%% describing domains of planning problems that correspond to POMDPs.  To
-%% accompany that language we have implemented an information-state
-%% \laostar\ procedure for solving problems expressed in DTPDDL.
-
-%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
-%% modified \fastdownward\ so that it can find useful sequential executions
-%% given DTPDDL models of the problem at hand.
-
-%%  called DTPDDL, along
-%% the lines of PDDL for the partially observable case, an \laostar\
-%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
-%% and modify the \fastdownward\ system to find high quality sequential plans
-%% .
-
-%% We model the environment as a partially observable Markov decision
-%% process. Although planning in that model is undecidable in
-%% general~\cite{}, an optimal finite-horizon plan corresponds to a
-%% contingent plan, that is a function mapping action-observation
-%% histories to actions.
-
-%% our continual planner is reactive, replanning whenever the underlying
-%% domain and problem models change. For example, replanning occurs if
-%% the motivational component alters the objectives, and if an assumed
-%% outcome of a sensing action is not realised.
-
-%% brittle sensing model,
-
-%% the evaluation of a fluent at a state is either known. Moreover, there
-%% is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process state
-$\state$ as the set of propositions that are true of the
-state. Notationally, $\props$ is the set of propositions, $\prop$ is
-an element from that set, and we have $\state \subseteq
-\props$. The underlying process
-dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
-state-characterising propositions $\propositions$.
-%%
-We say an action $\stochAction \in \stochActions$ is
-applicable if its precondition $\poss(\stochAction)$, a set of
-propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature chooses a deterministic STRIPS effect $\detAct_i$, and for
-all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$.
-
-%% We model a process state $\state$ as the set of propositions that are
-%% true of the state. Notationally, we have $\state \subseteq
-%% \props$. State change is induced by application of actions. A
-%% stochastic action $\stochAction \in \stochActions$ is applicable if
-%% its precondition $\poss(\stochAction)$, a set of propositions, are
-%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
-%% applicable at state $\state$.  If action $\stochAction \in
-%% \stochActions(\state)$ is applied at \state, nature chooses one
-%% element amongst a small set of deterministic outcomes
-%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-%% $\sum_{\detAction_i \in \detActions(\stochAction)}
-%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-%% effect on the state given in terms of two lists of propositions. The
-%% add-list $\add(\detAction)$ and delete-list
-%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-%% $\detAction$, then it cannot be in the delete-list and vice versa.}
-%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
-%% applied at state $\state$, then the resultant state is $ ( \state \cup
-%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-%% propositions from $\add(\detAction)$ are added to $\state$, and those
-%% from $\delete(\detAction)$ are removed from $\state$.
-
-%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
-
-We are concerned with problems that feature partial
-observability. Although we could invoke {\em extended probabilistic
-STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for presentation and computation to
-separate sensing and action. Therefore, we suppose a POMDP has a
-perceptual model given in terms of a finite set of stochastic {\em
-senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
-and perceptual propositions $\percepts$, called {\em percepts}. In
-detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote by \observations\ the set of
-reachable observations. The underlying state of the process cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in
-\observations$ that informs what should be believed about the state the
-process is in. If $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
-usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from the chosen elements of active
-senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make some fairly standard assumptions. First, that action
-execution and sensing occur instantaneously, and that only one action
-can be applied per plan-step. Secondly, it can arise that in some
-states an action is applicable, and in others it is not. Moreover, a
-belief-state can assign non-zero probability to states where that
-applicability holds, and states where it does not. %% We differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
-%% In beliefs where an action exists that is applicable in some state 
-%% with non-zero probability, and not applicable in another, we differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%% because 
-%% in that
-%% we forbid the execution of the action.
-%% actions that are not applicable in
-%% any state of the current belief.  
-%% Also,
-Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we incorporate a PPDDL-like default semantics, treating an action
-\stochAction\ executed at $\state_i$ as if it has no effect at
-states $\state$ if $\poss(\stochAction) \not\subseteq
-\state$.
-
-% Mucked around with this, then realised the two sentences were
-% contradictory. If we forbid actions applicable in one state but not
-% another in the belief state, then clearly we aren't using a
-% PPDDL-like semantics.
-%% Doubled comments are stuff I added, then removed.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Until now we have discussed the POMDP in terms of propositions and
-percepts. In order to address belief revision and utility it is
-convenient to consider the underlying decision process in a flat
-format. This is given by the tuple
-$\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-The function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs as
-negative rewards.
-%%
-From $\psi$ we have that for each $\state' \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
-\action)$ the probability of getting observation $\observ$ in state
-$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state', \action) = 1$.
-
-Successive state estimation is by application of Bayes'
-rule.  Taking the current belief $\bstate$ as the {\em prior}, and
-supposing action $\action$ is executed with perceptive outcome
-$\observ$, the probability that we are in $\state'$ in the
-successive belief-state $\bstate'$ is:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
-  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---i.e., reaches a state that satisfies a given
-Boolean condition--- with probability $1$.  Generally, whatever the
-structure of a plan, its value corresponds to the expected cumulative
-reward:
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
-%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
-%% probability that the controller prescribes \action\ when in state
-%% \node, and $\transition_\node(\action, \observation, \node') =
-%% P(\node'|\node, \action, \observation)$ gives the probability the
-%% controller transitions to state $\node'$, supposing we execute
-%% \action\ in state \node\ receiving observation \observation. When
-%% acting for finite $N$ steps from controller state $n$, the value of a
-%% controller corresponds to the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
-the action prescribed at step $t$, and
-
-\[\reward(\bstate, \action) = \sum_{\state \in \states}
-\bstate(\state)\reward(\state, \action).\]
-
-%% \noindent Finally, it is
-%% useful to note that there is a corresponding deterministic FSCs for
-%% both sequential (e.g., conformant) and contingent plan formats.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
-used in all three of the International Planning Competitions since
-2004. A variation on PDDL for describing domains with stochastic
-actions and uncertain starting configurations, PPDDL is a declarative
-first-order language that facilitates factored descriptions of domains
-and problems. There are straightforward compilations from problems
-expressed in PPDDL to propositional representations amenable
-to state-of-the-art planners.
-
-Because PPDDL cannot model domains that feature partial observability,
-we develop an extension we call {\em Decision-Theoretic
-(DT)PDDL}. This can express probabilistic models of the sensing
-consequences of acting, to quantitatively capture unreliability in
-perception. That expressive power is achieved by incorporating
-perceptual analogues of fluent, predicate, and action definitions. In
-detail, we have declarations of state characterising predicate and
-fluent symbols according to the PPDDL syntax. In addition, we allow
-two other declarations, for perceptual predicates and fluents
-respectively. For example, suppose our robot is tasked with exploring
-{\em locations} in order to identify the whereabouts of a {\em
-visual-object}. We must describe state and perceptual facts that model
-the {\em true}, resp. perceived, locations of objects. In DTPDDL,
-these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like declarations, with preconditions expressed using state
-predicate and action function symbols, and uniformly positive effects
-over perceptual predicates. A sense declaration corresponds to a
-lifted description of proposition {\em senses}. Whereas the effects of
-an {\em operator} schema describe how states change under application
-of actions, the effects of a {\em sense} schema are perceptual,
-specifying the composition of an observation following the execution
-of an action.
-
-%% Also, and without a loss of generality, we find it
-%% convenient to separate the actional precondition from the state
-%% precondition.
-
-Making the above ideas concrete with an example, suppose a robot
-called $\pp{DORA}$ is able to look for a visual-object, such as a
-$\pp{box}$, at a given place. We can model that deterministic action
-using the following operator schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs an instantaneous penalty, $3$, that corresponds to
-the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
-reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, reward is
-accumulated, whereas in DTPDDL it is instantaneous.}
-
-Completing the example, we model the sensing that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- i.e,
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
-$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-our switching continual planner. The distribution is expressed in a
-tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
-(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
-probabilistic and atomic terms. The root term is always conjunctive,
-and the leaves are atomic. For example, the starting distribution for
-$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
-$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
-i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
-$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
-that contain a single atomic subterm.}
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-by a {\em visitation} of terms: An atom is {\em visited} iff
-its conjunctive parent is visited, and a conjunctive term is visited
-iff all its immediate subterms are visited. A probabilistic term is
-visited iff its conjunctive parent is visited, and exactly one of its
-subterms, $T_i$, is visited. Each visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability that it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $\prob_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-From hereon, to simplify the discussion (and implementation)
-we shall restrict our attention to POMDPs with deterministic
-actions. We note that POMDPs with stochastic
-actions can be compiled into equivalent deterministic-action POMDPs,
-where all the original action uncertainty is expressed in the
-starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
-that of finite-horizon planning, such a compilation yields a finite
-flat representation of the original POMDP, only with deterministic
-actions.
-
-\section{Switching Continual Planner}
-
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
-
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
-%%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
-
-Finally, whether proceeding in a sequential or contingent session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
-
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
-
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
-
-\subsection{Sequential Planning in POMDPs}
-
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
-
-\noindent An assumption corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
-
-\[
-\begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
-\end{array}
-\]
-
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
-%%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
-%%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
-
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
-
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
-
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
-
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
-\[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
-\reward(\state_i, \action_i),
-\]
-
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
-
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
-\[
-\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
-\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\hat\$(T_i) & \pp{otherwise} \\
-\end{array}
-\]
-
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
-%%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
-
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
-\end{array}
-\]
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
-
-We also include another propositionally trivial action
-$\actions.\poss(\switchAction)$ with the reward property:
-
-\[
-\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
-\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
-\end{array}
-\]
-
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
-
-%%  We make no assumptions here about how they are
-%% derived.
-
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
-
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief-state with the assumptions made by the continual
-    planner in bold.]{
-      \label{fig:abstraction-a}
-      \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode] (cup) {} 
-          child {node [cnode] (cupp1) {}}
-          child {node [cnode] (cupp2) {}}
-        }
-      }
-      child {node[cnode] (kitchen) {}
-        child {node [pnode] (box2) {} 
-          child {node [cnode] (box2p1) {}}
-          child {node [cnode] (box2p2) {}}
-        }
-      };
-   \tiny
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of kitchen] {$\pp{kitchen}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[below=0 of cup] {$\pp{(is-in cup)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-   \node[right=0 of cupp1] {$\pp{place1}$};
-   \node[right=0 of cupp2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Generating the abstract belief space.}
-\label{fig:abstraction}
-\end{figure}
-
-%%% OLDER STUFF
-
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
-
-\section{Evaluation}
-
-We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. 
-% Our implementation is able
-% to use several underlying planning systems.
-We have extended MAPSIM so that it can parse DTPDDL, and perform
-successive estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
-
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
-}
-
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
-
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
-
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
-
-We evaluated our switching planner against the baseline in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
-  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
-  core.}. In our domain, rewards encapsulate a goal condition, and
-therefore we find it intuitive to report average costs of the plans
-and the success rate in problems that admit a solution (i.e., positive
-reward scaled by a constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
-  \caption{Average runtime}
-  \label{fig:results-time}
-\end{figure}
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
-  \caption{Average plan costs and number of successful runs.}
-  \label{fig:results-quality}
-\end{figure}
-
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% \section{Discussion}
-%% \input{discussion.tex}
-
-\section{Related Work}
 There have been a number of recent papers on
@@
 domain due to the task planning requirement.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
-
-The most pressing item for future research, is to develop a scheme
-whereby the serial planner can relax {\em executability} assumptions,
-so that conformant (or semi-conformant) plans can be executed without
-interruption by a contingent session. A more general criticism of
-switching continual planning concerns interleaved sequential and
-contingent sessions failing to make any progress towards the
-objectives. For example, in the worst case we can have each sequential
-session producing an identical plan, and each decision-theoretic
-session rejecting it without further sensing. Although this has not
-been an issue in our work so far, it must be dealt with rigorously
-in the future. We suggest that a good way to mitigate this problem is
-by developing a {\em motivational} component that maintains a {\em
-dynamic} reward model whose limiting behaviour prevents that switching
-deadlock.
-
-%% The sequential mode of our system always schedules decision-theoretic
-%% planning before executing actions that are not applicable with
-%% certainty. This is inefficient if the utility of the plan is not
-%% dependent on that actions successful execution. This situation always
-%% arises for conformant plans.
-
-%% The switching continual planning system we have described serves as
-%% the underlying planner for CogX.
-
-%% the effects of a {\em sense} schema are perceptual, whereas the
-%% effects of an {\em operator} schema are over state propositions.
-
-%% posted by a motivational component of the underlying robotic
-%% architecture.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% In this paper we present an approach to continual planning that uses
-%% two planning systems. The first
-
-%%  to a distinct class of
-%% challenges. We suppose 
-%% %%
-%% The underlying environment is modelled as a POMDP. We use the fast
-%% classical satisficing system FastDownward to find a deterministic
-%% sequential plan and complementary runtime evolution of that
-%% process. This corresponds to a generalisation of replanning in
-%% probabilistic planning to problems with partial observability.
-
-%% Interaction between the sequential planner and execution proceeds
-%% more-or-less analogously to popular replanning approaches
-
-%% Addressing these challenges in a monolithic framework, we present a
-%% {\em switching} continual planner, that uses the fast sequential
-%% satisficing procedure FastDownward to perform net-benefit
-%% %%
-%% makes reasonable assumptions about the evolution of the runtime state
-%% given a POMDP model of the environment.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% The decision-theoretic planner is able to tailor sensory processing on
-%% a robot platform to the current objective, while FastDownward  quickly
-
-%% In this paper we develop a continual planing system that uses two
-%% planning systems. The first, is a state-of-the-art domain independent
-%% planner for deterministic problems. The second is a information-state
-%% contingency planning the information-state space of
-
-%% Continual planning is a powerful technique that goes some way to
-%% addressing those challenges. That approach interleaves planning and
-%% execution, deliberately postponing planning for contingencies unless
-%% they eventuate during execution.
-
-%% computing a single sequential plan and
-%% eventuality
-
-%% To these challenges it seems a continual planning approach is best,
-
-%% where the runtime state evolves during plan execution
-
-%% interleaved planning and execution.
-
-%% the latter is able to tailor sensory processing on a robot platform,
-%% in order that it.
-
-%% {\em ad-hoc}
-
-%% , reason about degrees of belief and uncertainty about the world,
-
-%% probabilistic sequential decision making in practical sized problems
-%% is intractable.
-
-%% quantitative probabilistic models of the perception and action .
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai11}
-
-\end{document}

### 2011-01-28 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
AAAI version of the switching planner paper

--- 2010-11-26 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
@@
 
+\usepackage{subfig}
+
+\usepackage{tikz}
+\usetikzlibrary{shapes}
+\usetikzlibrary{positioning}
+
 \usepackage{times}
@@
 %
-%\title{Switching in Continual Planning for Practical Robot Control}
-\title{A Switching Planner for Combined Task and Observation Planning}
-%\title{A Continual Planner that Switches Between Sequential and
-%  Decision-Theoretic Planning} 
-\author{NA}
+%% \title{Interleaved Serial and Decision-Theoretic Planning for a Mobile Robot}
+
+%% \title{Interleaved Serial and Decision-Theoretic Planning for
+%% Practical Robot Control}
+
+\title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
+
+%% \title{A Switching Planner for Combined Task and Observation Planning}
+\author{anonymous}
 \setcounter{secnumdepth}{0}
@@
 
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
+%% Those processes must
 
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
+From an automated planning perspective the problem of practical mobile
+robot control in realistic environments poses many important and
+contrary challenges.  On the one hand, the planning process must be
+lightweight, robust, and timely. Over the lifetime of the robot it
+must always respond quickly with new plans that accommodate exogenous
+events, changing objectives, and the underlying unpredictability of
+the environment.  On the other hand, in order to exhibit useful
+behaviours the planning process must perform computationally expensive
+reasoning about contingencies and possible revisions of subjective
+beliefs according to quantitatively modelled uncertainty in acting and
+sensing.
+%%
+Towards addressing these challenges, we develop a \emph{continual
+planning} approach that {\em switches} between using a fast
+satisficing ``classical'' planner, to decide on the overall strategy,
+and decision-theoretic planning to solve small subproblems where
+deeper consideration of the sensing model is both practical, and can
+significantly impact overall performance. We demonstrate our approach
+in large problems from a realistic robot exploration domain.
+
+%% partial observability
+
+%% In cooperation with execution monitoring, it must perform repair
+%% and/or replanning in order to seamlessly accommodate exogenous events,
+%% changing objectives, and the underlying unpredictability of the
+%% environment.
+
+%% Realistic robot planning problems in uncertain environments often
+%% require achieving tasks while also finding out about the
+%% world. Because the world state cannot be observed directly, these
+%% problems are naturally represented as partially observable Markov
+%% decision problems (POMDPs). However, these are typically intractable
+%% for realistic problems.
+%%
 

### 2011-02-02 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
Change title back and small change to abstract.

--- 2011-01-28 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
@@
 
-\title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
+% \title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
 
-%% \title{A Switching Planner for Combined Task and Observation Planning}
+\title{A Switching Planner for Combined Task and Observation Planning}
 \author{anonymous}
@@
 deeper consideration of the sensing model is both practical, and can
-significantly impact overall performance. We demonstrate our approach
+significantly impact overall performance. We evaluate our approach
 in large problems from a realistic robot exploration domain.

### 2011-02-03 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
Rearrange to likely structure

@@
 
-\section{Propositional Decision-Theoretic Planning}
-\input{prop-dtp.tex}
+%% \section{Propositional Decision-Theoretic Planning}
+%% \input{prop-dtp.tex}
+
+\section{Related Work}
+\input{relatex.tex}
 
@@
 %% \input{discussion.tex}
-
-\section{Related Work}
-\input{relatex.tex}
 

### 2011-02-03 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
New bibliography, and the related work section is begenning

@@
+\cite{wyattetal2010tamd}
+
+For the purpose of symbolic
+planning,~\cite{wyattetal2010tamd,talamadupula:2010,Kraft2008} treat
+sensing deterministically and beliefs qualitatively.
+%%
+ Addressing specifically this second
+challenge,~\citeauthor{talamadupula:2010} identify
+\emph{continual planning} in the presence of detailed probabilistic
+ models as an important direction for future research.
+
+Focusing now specifically on planning, our work is motivated by
+domains that contain a mixture of task planning and observation
+planning. There have been a number of recent papers representing
+observation planning problems as POMDPs and using various techniques
+to manage the large state
+space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
+approach in a vision algorithm selection problem. In their case there
+is a natural hierarchical decomposition of the problem which allows
+them to solve large problems by breaking them into a set of small
+POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
+represent a preference elicitation problem as a POMDP and take
+advantage of symmetry in the belief space ---essentially the idea that
+it does not matter what the value of the variable you are trying to
+observe is--- to exponentially shrink the state space. Although we
+have been actively exploring the \citeauthor{doshi08:pref_elic}
+approach, those exploitable structures are not present in our domain
+due to the task planning requirement.
+
+\cite{shani:etal:08}
+
+Generally There has been much recent work on scaling POMDP solution procedures
+to medium-sized instances. In the case of general domain-independent
+factored systems, the state-of-the-art scales to relatively small
+problems with $2^{22}$
+states~\cite{shani:etal:2008}.\footnote{Considering only room
+categories and distribution of objects, problems we consider in this
+paper have $\sim 10^{27}$ states. The details of view points, from
+local active visual search, and those of robot location, further
+increase that figure. Therefore, not only because they are offline,
+but also because they have limited scalability, these approaches are
+infeasible in our setting.} At their limit, these procedures take over an
+hour to converge.  For classes of POMDP that feature exploitable
+structures (e.g., no actions with negative effects), problems with as
+many as $10^{30}$ states can be targeted by offline
+procedures~\cite{brunskill:russell:2010}. Moving someway towards
+addressing all the challenges we have outlined, recent online POMDP
+solution procedures have been developed which can exploit highly
+approximate value functions -- typically computed using a point-based
+procedure -- and heuristic in forward
+search~\cite{ross:etal:2008}. Their applicability in our setting is
+limited due to the large amount of
+\emph{problem-specific} offline processing required to get useful
+search guidance. A {\em very} recent and promising online approach for
+large POMDPs employs Monte-Carlo sampling to break the curse of
+dimensionality in situations where goal reachability is
+easy~\cite{silver:veness:2010}. Although we suppose it an interesting
+item for future work to pursue that direction, it should be noted that
+ease of goal reachability is not guaranteed in the problems we face,
+and is certainly not a property to be assumed in domain independent
+planning.
+
 There have been a number of recent papers on
@@
 %% assumption that we relax.
-
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.

### 2011-02-03 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
Comple of new bits of text

--- 2011-02-03 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
@@
-\cite{wyattetal2010tamd}
+%%  and
+%% using various techniques to manage the large state
+%% space
 
-For the purpose of symbolic
-planning,~\cite{wyattetal2010tamd,talamadupula:2010,Kraft2008} treat
-sensing deterministically and beliefs qualitatively.
+%% \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
+%% approach in a
+
+Addressing task and observation planning specifically, there have been
+a number of recent developments where the underlying problem is
+modelled as a POMDP.
 %%
- Addressing specifically this second
-challenge,~\citeauthor{talamadupula:2010} identify
-\emph{continual planning} in the presence of detailed probabilistic
- models as an important direction for future research.
-
-Focusing now specifically on planning, our work is motivated by
-domains that contain a mixture of task planning and observation
-planning. There have been a number of recent papers representing
-observation planning problems as POMDPs and using various techniques
-to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
+For the vision algorithm
+selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit a
+natural hierarchical decomposition of the underlying POMDP into
+subproblems where decision-theoretic planning is reasonably
+fast. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
 represent a preference elicitation problem as a POMDP and take
@@
 have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, those exploitable structures are not present in our domain
-due to the task planning requirement.
+approach, those exploitable structures are not present in problems we
+have considered so far due to the task planning requirement.
 
-\cite{shani:etal:08}
-
-Generally There has been much recent work on scaling POMDP solution procedures
-to medium-sized instances. In the case of general domain-independent
-factored systems, the state-of-the-art scales to relatively small
-problems with $2^{22}$
+More generally, there has been much recent work on scaling POMDP
+solution procedures to medium-sized instances~\cite{shani:etal:08}. In
+the case of general domain-independent factored systems, the
+state-of-the-art scales to relatively small problems with $2^{22}$
 states~\cite{shani:etal:2008}.\footnote{Considering only room
@@
 but also because they have limited scalability, these approaches are
-infeasible in our setting.} At their limit, these procedures take over an
-hour to converge.  For classes of POMDP that feature exploitable
+infeasible in our setting.} At their limit, these procedures take over
+an hour to converge.  For classes of POMDP that feature exploitable
 structures (e.g., no actions with negative effects), problems with as
@@
 large POMDPs employs Monte-Carlo sampling to break the curse of
-dimensionality in situations where goal reachability is
-easy~\cite{silver:veness:2010}. Although we suppose it an interesting
-item for future work to pursue that direction, it should be noted that
-ease of goal reachability is not guaranteed in the problems we face,
-and is certainly not a property to be assumed in domain independent
-planning.
+dimensionality in situations where goal reachability is {\em easily}
+determined~\cite{silver:veness:2010}.
 
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
+%% Although we suppose it an interesting
+%% item for future work to pursue that direction, it should be noted that
+%% ease of goal reachability is not guaranteed in the problems we face,
+%% and is certainly not a property to be assumed in domain independent
+%% planning.
+
+In the direction of leveraging {\em classical} systems/approaches for
+planning under uncertainty, the most highlighted system to date has
+been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
+the probabilistic track of the 2004 International Planning
+Competition.  In the continual paradigm, \system{FFR$_a$} uses the
+fast satisficing procedure \system{FF}~\cite{hoffmann:nebel:2001} to
+compute sequential plans and corresponding execution traces.
 %%
@@
 plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
 
@@
 %% assumption that we relax.
+
+%%No! They simply haven't been evaluated in PO settings. They may, or
+%%may not struggle. They have sampling of traces, and that would
+%%include observations, and therefore evolutions of beliefs. SSAT was
+%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
+%%suited to POMDPs.
+
+% Normally if we are going to compare with related work, we do
+% actually *compare*. Why didn't you try those approaches? I think
+% it's safe to say that FFR will struggle. Why would it even include
+% in its plan an observational action that doesn't change the world?
+
+%%
+%% However, as we said in the introduction,
+%% all these approaches struggle in partially observable domains as they
+%% rely on being able to determine the state at all times.

### 2011-02-03 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
With a question for moritz

@@
 %%
-For the vision algorithm
+For vision algorithm
 selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit a
@@
 have considered so far due to the task planning requirement. 
+%%
+\citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010}) 
+%%
+Finally,
+our approach is in a similar vein to the more classical {\em
+dual-mode} control~\cite{cassandra96actingunder}, however in our case
+entropy reduction is targeted by planning in an abstract process which
+is informed by one execution trace ---computed by a {\em classical}
+planner--- and the underlying belief-state.
 
@@
 state-of-the-art scales to relatively small problems with $2^{22}$
-states~\cite{shani:etal:2008}.\footnote{Considering only room
-categories and distribution of objects, problems we consider in this
-paper have $\sim 10^{27}$ states. The details of view points, from
-local active visual search, and those of robot location, further
-increase that figure. Therefore, not only because they are offline,
-but also because they have limited scalability, these approaches are
-infeasible in our setting.} At their limit, these procedures take over
-an hour to converge.  For classes of POMDP that feature exploitable
+states~\cite{shani:etal:2008}. At their limit, these procedures take
+over an hour to converge, and $\sim10$ seconds on average to perform a
+single Bellman backup.  For classes of POMDP that feature exploitable
 structures (e.g., no actions with negative effects), problems with as
@@
 procedures~\cite{brunskill:russell:2010}. Moving someway towards
-addressing all the challenges we have outlined, recent online POMDP
-solution procedures have been developed which can exploit highly
-approximate value functions -- typically computed using a point-based
-procedure -- and heuristic in forward
-search~\cite{ross:etal:2008}. Their applicability in our setting is
-limited due to the large amount of
-\emph{problem-specific} offline processing required to get useful
+addressing real-time decision making, recent online POMDP solution
+procedures have been developed which can exploit highly approximate
+value functions -- typically computed using a point-based procedure --
+and heuristics in forward search~\cite{ross:etal:2008}. Their
+applicability in our setting is limited, firstly because they only
+scale to smaller problems, with thousands of states, and also due to
+the large amount of
+\emph{problem-specific} offline processing that might be required to get useful
 search guidance. A {\em very} recent and promising online approach for

### 2011-02-03 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
New intro text

@@
 %%
-\citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010}) 
 %%
@@
 solution procedures to medium-sized instances~\cite{shani:etal:08}. In
-the case of general domain-independent factored systems, the
-state-of-the-art scales to relatively small problems with $2^{22}$
+that
+setting \citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010})
+recently addressed an inefficiency of offline point-based techniques
+in problems with medium length planning horizons, however their
+approach take tens of minutes to start yielding good behaviour even in
+relatively small problems (tens of thousands of states). In the case
+of general domain-independent factored systems, the state-of-the-art
+scales to relatively small problems with $2^{22}$
 states~\cite{shani:etal:2008}. At their limit, these procedures take

### 2011-02-04 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
shrinking and growing, all at the same time.  Not sure how we are going to get this down to 6 pages yet.

--- 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
-%%%% ijcai11.tex
-
-\typeout{IJCAI Switching Planner Submission}
-
-% These are the instructions for authors for IJCAI-11.
-% They are the same as the ones for IJCAI-07 with superficical wording
-%   changes only.
-
-\documentclass{article}
-% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
-\usepackage{ijcai11}
-
-% Use the postscript times font!
-\usepackage{times}
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\usepackage{subfig}
-\usepackage{tikz}
-\usetikzlibrary{shapes}
-\usetikzlibrary{positioning}
-
-\usepackage{graphicx}
-\graphicspath{{plots/}}
-\include{macros}%
-
-\title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Planning for agents acting in partially observable environments with
-stochastic actions is extremely computationally
-expensive~\cite{mdp-complexity}. Even relatively small partially
-observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, have state spaces far larger than existing
-optimal or close to optimal planners can handle. For the easier
-problem of planning in completely observable stochastic domains, an
-alternative has been to use classical sequential planners and
-replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
-Execution monitoring tracks properties of the state that determine
-whether further planning is required. For partially observable
-domains, only probabilistic information about such properties will, in
-general, be available.
-
-%% However, these approaches rely
-%% crucially on being able to determine properties of the current state
-%% in order to decide to replan. For partially observable domains, only
-%% probabilistic information about these properties will in general be
-%% available.
-
-This paper describes the planning component for a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings. This domain features partial observability, particularly
-because the state is not perfectly known, and state information is
-gained by the robot through the use of unreliable sensing
-actions. 
-%Using a propositionally factored state representation, for
-For
-interesting sized tasks the corresponding POMDP model is too large for a
-POMDP solver to be applied directly. Inspired by the replanning
-approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
-planning approach that uses a classical planning system to compute a
-reasonably valuable trace in the model, and then uses a
-decision-theoretic planner on small subproblems where reasoning about
-observations might be useful. We refer to the complete system as a
-{\em switching continual planner}. The approach is not optimal,
-particularly as it relies on the results of sequential planning
-directly. It does nevertheless perform better than a purely sequential
-replanner, and is fast enough to be used for real-time decision-making
-on a mobile robot.
-
-%% Switching gives us some important advantages in our robotic
-%% domain. First, the significant source of uncertainty in the domain is
-%% the unreliability of observations, and this approach is particularly
-%% suited to problems that combine a deterministic task planning problem
-%% with observations as the decision-theoretic planner is often invoked
-%% to determine a characteristic of the underlying state. Secondly,
-%% replanning makes our system robust to changing objectives and the
-%% discovery of new facts about the world.
-
-%% As an illustration, consider a robot searching for a box that it knows
-%% is often found in offices. The sequential planner might produce a plan
-%% that assumes that {\sc Room a} is an office, that the box is in {\sc
-%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
-%% these assumptions, the plan might be to go to {\sc Room a}, go to
-%% location 1 and search for the box. When the plan gets executed and the
-%% search begins the decision-theoretic planner is called since searching
-%% involves making observations. The decision-theoretic planner might
-%% retract the last assumption but leave the others, and build a plan to
-%% search the whole of {\sc Room a} for the box. If this plan fails to
-%% find the box, it will disprove the assumption that the box is in {\sc
-%% Room a}, which will then lead to replanning by the sequential planner
-%% based on the newly discovered knowledge.
-
-Our system is domain independent, taking domain and problem
-descriptions in a first-order declarative language we have developed,
-called the decision-theoretic planning domain definition language
-(DTPDDL). From these descriptions of POMDPs, we automatically
-construct a deterministic model suitable for sequential planners. That
-model includes actions which correspond to making assumptions about
-the values of imperfectly known state variables. Assumptions scheduled
-by the sequential system are used to propose a pragmatic abstract
-belief-state space to the decision-theoretic system, and to modify the
-reward function, so that system might pursue sensing related to those
-assumptions.
-
-%%  which can operate
-%% with the original representation of the domain.
-
-%% The sequential planner we use is based
-%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
-%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
-%% knowledge to be represented explicitly, so we can write actions that
-%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
-%% THIS. For the decision-theoretic planner we have implemented our own
-%% forward search in the belief-state space.
-
-The switching planner provides important advantages in our
-mobile robot domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations because the decision-theoretic planner
-is only invoked to deal with state uncertainty.
-%determine a characteristic of the underlying state. 
-Secondly, replanning makes the system robust to
-changing objectives and discoveries about the world,
-both of which feature in our domain.
-
-The remainder of the paper proceeds as follows. We begin by describing
-POMDPs with a propositionally factored representation of states and
-observations. We also outline the relationship between this and a flat
-POMDP representation. Next we describe our DTPDDL language with
-example declarations from a mobile robot exploration task. Then we
-describe the switching planner, including how the sequential domain is
-produced and a plan is generated, and how this then provides the input
-for the decision-theoretic planner. Finally we provide an evaluation
-of the system on the example robot domain, and finish with a short
-survey of related work and some concluding remarks.
-
-%To accompany
-%that language we have implemented an information-state
-%\laostar\ procedure for solving problems expressed in DTPDDL.
-
-%%
-
-%Have a good understanding of the contents and function of rooms, as
-%well as the linguistic referrents to rooms, widgets, and their visual
-%qualities.
-
-%Having confidence in its beliefs about the linguistic terms for
-%spaces, and the function of those spaces.
-
-%sometimes difficult to predict, with exogenous events, such as the
-%changing of the objective,
-
-%the speed and scalability of software for sequential planning in
-%deterministic
-
-%% We developed a first-order declarative language, called DTPDDL, for
-%% describing domains of planning problems that correspond to POMDPs.  To
-%% accompany that language we have implemented an information-state
-%% \laostar\ procedure for solving problems expressed in DTPDDL.
-
-%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
-%% modified \fastdownward\ so that it can find useful sequential executions
-%% given DTPDDL models of the problem at hand.
-
-%%  called DTPDDL, along
-%% the lines of PDDL for the partially observable case, an \laostar\
-%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
-%% and modify the \fastdownward\ system to find high quality sequential plans
-%% .
-
-%% We model the environment as a partially observable Markov decision
-%% process. Although planning in that model is undecidable in
-%% general~\cite{}, an optimal finite-horizon plan corresponds to a
-%% contingent plan, that is a function mapping action-observation
-%% histories to actions.
-
-%% our continual planner is reactive, replanning whenever the underlying
-%% domain and problem models change. For example, replanning occurs if
-%% the motivational component alters the objectives, and if an assumed
-%% outcome of a sensing action is not realised.
-
-%% brittle sensing model,
-
-%% the evaluation of a fluent at a state is either known. Moreover, there
-%% is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process state
-$\state$ as the set of propositions that are true of the
-state. Notationally, $\props$ is the set of propositions, $\prop$ is
-an element from that set, and we have $\state \subseteq
-\props$. The underlying process
-dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
-state-characterising propositions $\propositions$.
-%%
-We say an action $\stochAction \in \stochActions$ is
-applicable if its precondition $\poss(\stochAction)$, a set of
-propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature chooses a deterministic STRIPS effect $\detAct_i$, and for
-all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$.
-
-%% We model a process state $\state$ as the set of propositions that are
-%% true of the state. Notationally, we have $\state \subseteq
-%% \props$. State change is induced by application of actions. A
-%% stochastic action $\stochAction \in \stochActions$ is applicable if
-%% its precondition $\poss(\stochAction)$, a set of propositions, are
-%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
-%% applicable at state $\state$.  If action $\stochAction \in
-%% \stochActions(\state)$ is applied at \state, nature chooses one
-%% element amongst a small set of deterministic outcomes
-%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-%% $\sum_{\detAction_i \in \detActions(\stochAction)}
-%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-%% effect on the state given in terms of two lists of propositions. The
-%% add-list $\add(\detAction)$ and delete-list
-%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-%% $\detAction$, then it cannot be in the delete-list and vice versa.}
-%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
-%% applied at state $\state$, then the resultant state is $ ( \state \cup
-%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-%% propositions from $\add(\detAction)$ are added to $\state$, and those
-%% from $\delete(\detAction)$ are removed from $\state$.
-
-%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
-
-We are concerned with problems that feature partial
-observability. Although we could invoke {\em extended probabilistic
-STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for presentation and computation to
-separate sensing and action. Therefore, we suppose a POMDP has a
-perceptual model given in terms of a finite set of stochastic {\em
-senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
-and perceptual propositions $\percepts$, called {\em percepts}. In
-detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote by \observations\ the set of
-reachable observations. The underlying state of the process cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in
-\observations$ that informs what should be believed about the state the
-process is in. If $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
-usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from the chosen elements of active
-senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make some fairly standard assumptions. First, that action
-execution and sensing occur instantaneously, and that only one action
-can be applied per plan-step. Secondly, it can arise that in some
-states an action is applicable, and in others it is not. Moreover, a
-belief-state can assign non-zero probability to states where that
-applicability holds, and states where it does not. %% We differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
-%% In beliefs where an action exists that is applicable in some state 
-%% with non-zero probability, and not applicable in another, we differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%% because 
-%% in that
-%% we forbid the execution of the action.
-%% actions that are not applicable in
-%% any state of the current belief.  
-%% Also,
-Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we incorporate a PPDDL-like default semantics, treating an action
-\stochAction\ executed at $\state_i$ as if it has no effect at
-states $\state$ if $\poss(\stochAction) \not\subseteq
-\state$.
-
-% Mucked around with this, then realised the two sentences were
-% contradictory. If we forbid actions applicable in one state but not
-% another in the belief state, then clearly we aren't using a
-% PPDDL-like semantics.
-%% Doubled comments are stuff I added, then removed.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Until now we have discussed the POMDP in terms of propositions and
-percepts. In order to address belief revision and utility it is
-convenient to consider the underlying decision process in a flat
-format. This is given by the tuple
-$\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-The function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs as
-negative rewards.
-%%
-From $\psi$ we have that for each $\state' \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
-\action)$ the probability of getting observation $\observ$ in state
-$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state', \action) = 1$.
-
-Successive state estimation is by application of Bayes'
-rule.  Taking the current belief $\bstate$ as the {\em prior}, and
-supposing action $\action$ is executed with perceptive outcome
-$\observ$, the probability that we are in $\state'$ in the
-successive belief-state $\bstate'$ is:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
-  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---i.e., reaches a state that satisfies a given
-Boolean condition--- with probability $1$.  Generally, whatever the
-structure of a plan, its value corresponds to the expected cumulative
-reward:
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
-%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
-%% probability that the controller prescribes \action\ when in state
-%% \node, and $\transition_\node(\action, \observation, \node') =
-%% P(\node'|\node, \action, \observation)$ gives the probability the
-%% controller transitions to state $\node'$, supposing we execute
-%% \action\ in state \node\ receiving observation \observation. When
-%% acting for finite $N$ steps from controller state $n$, the value of a
-%% controller corresponds to the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
-the action prescribed at step $t$, and
-
-\[\reward(\bstate, \action) = \sum_{\state \in \states}
-\bstate(\state)\reward(\state, \action).\]
-
-%% \noindent Finally, it is
-%% useful to note that there is a corresponding deterministic FSCs for
-%% both sequential (e.g., conformant) and contingent plan formats.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
-used in all three of the International Planning Competitions since
-2004. A variation on PDDL for describing domains with stochastic
-actions and uncertain starting configurations, PPDDL is a declarative
-first-order language that facilitates factored descriptions of domains
-and problems. There are straightforward compilations from problems
-expressed in PPDDL to propositional representations amenable
-to state-of-the-art planners.
-
-Because PPDDL cannot model domains that feature partial observability,
-we develop an extension we call {\em Decision-Theoretic
-(DT)PDDL}. This can express probabilistic models of the sensing
-consequences of acting, to quantitatively capture unreliability in
-perception. That expressive power is achieved by incorporating
-perceptual analogues of fluent, predicate, and action definitions. In
-detail, we have declarations of state characterising predicate and
-fluent symbols according to the PPDDL syntax. In addition, we allow
-two other declarations, for perceptual predicates and fluents
-respectively. For example, suppose our robot is tasked with exploring
-{\em locations} in order to identify the whereabouts of a {\em
-visual-object}. We must describe state and perceptual facts that model
-the {\em true}, resp. perceived, locations of objects. In DTPDDL,
-these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like declarations, with preconditions expressed using state
-predicate and action function symbols, and uniformly positive effects
-over perceptual predicates. A sense declaration corresponds to a
-lifted description of proposition {\em senses}. Whereas the effects of
-an {\em operator} schema describe how states change under application
-of actions, the effects of a {\em sense} schema are perceptual,
-specifying the composition of an observation following the execution
-of an action.
-
-%% Also, and without a loss of generality, we find it
-%% convenient to separate the actional precondition from the state
-%% precondition.
-
-Making the above ideas concrete with an example, suppose a robot
-called $\pp{DORA}$ is able to look for a visual-object, such as a
-$\pp{box}$, at a given place. We can model that deterministic action
-using the following operator schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs an instantaneous penalty, $3$, that corresponds to
-the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
-reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, reward is
-accumulated, whereas in DTPDDL it is instantaneous.}
-
-Completing the example, we model the sensing that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- i.e,
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
-$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-our switching continual planner. The distribution is expressed in a
-tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
-(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
-probabilistic and atomic terms. The root term is always conjunctive,
-and the leaves are atomic. For example, the starting distribution for
-$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
-$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
-i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
-$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
-that contain a single atomic subterm.}
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-by a {\em visitation} of terms: An atom is {\em visited} iff
-its conjunctive parent is visited, and a conjunctive term is visited
-iff all its immediate subterms are visited. A probabilistic term is
-visited iff its conjunctive parent is visited, and exactly one of its
-subterms, $T_i$, is visited. Each visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability that it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $\prob_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-From hereon, to simplify the discussion (and implementation)
-we shall restrict our attention to POMDPs with deterministic
-actions. We note that POMDPs with stochastic
-actions can be compiled into equivalent deterministic-action POMDPs,
-where all the original action uncertainty is expressed in the
-starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
-that of finite-horizon planning, such a compilation yields a finite
-flat representation of the original POMDP, only with deterministic
-actions.
-
-\section{Switching Continual Planner}
-
 We describe our approach that switches between sequential and
@@
 the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
+by \system{FFR$_a$}~\cite{yoon:etal:2007}. Finally, it is
 worth clarifying that although the above discussion has been in terms
@@
 artificially limits the length of candidate sequential plans.
+
+%% From the perspective of
+%% gradient-based reinforcement learning systems, such as William's
+%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
+%% under reasonable assumptions the optimal trace identifies a maximal
+%% gradient step from a uniformly random soft-max policy.
 
@@
 
-%%% OLDER STUFF
-
 % \small
@@
 
-\section{Evaluation}
-
-We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. 
-% Our implementation is able
-% to use several underlying planning systems.
-We have extended MAPSIM so that it can parse DTPDDL, and perform
-successive estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
-
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
-}
-
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
-
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
-
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
-
-We evaluated our switching planner against the baseline in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
-  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
-  core.}. In our domain, rewards encapsulate a goal condition, and
-therefore we find it intuitive to report average costs of the plans
-and the success rate in problems that admit a solution (i.e., positive
-reward scaled by a constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
-  \caption{Average runtime}
-  \label{fig:results-time}
-\end{figure}
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
-  \caption{Average plan costs and number of successful runs.}
-  \label{fig:results-quality}
-\end{figure}
-
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% \section{Discussion}
-%% \input{discussion.tex}
-
-\section{Related Work}
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
-
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-%%
-The continual planning system that motivated our
-project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-and has been applied in completely observable domains, particularly
-those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
-
-The most pressing item for future research, is to develop a scheme
-whereby the serial planner can relax {\em executability} assumptions,
-so that conformant (or semi-conformant) plans can be executed without
-interruption by a contingent session. A more general criticism of
-switching continual planning concerns interleaved sequential and
-contingent sessions failing to make any progress towards the
-objectives. For example, in the worst case we can have each sequential
-session producing an identical plan, and each decision-theoretic
-session rejecting it without further sensing. Although this has not
-been an issue in our work so far, it must be dealt with rigorously
-in the future. We suggest that a good way to mitigate this problem is
-by developing a {\em motivational} component that maintains a {\em
-dynamic} reward model whose limiting behaviour prevents that switching
-deadlock.
-
-%% The sequential mode of our system always schedules decision-theoretic
-%% planning before executing actions that are not applicable with
-%% certainty. This is inefficient if the utility of the plan is not
-%% dependent on that actions successful execution. This situation always
-%% arises for conformant plans.
-
-%% The switching continual planning system we have described serves as
-%% the underlying planner for CogX.
-
-%% the effects of a {\em sense} schema are perceptual, whereas the
-%% effects of an {\em operator} schema are over state propositions.
-
-%% posted by a motivational component of the underlying robotic
-%% architecture.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% In this paper we present an approach to continual planning that uses
-%% two planning systems. The first
-
-%%  to a distinct class of
-%% challenges. We suppose 
-%% %%
-%% The underlying environment is modelled as a POMDP. We use the fast
-%% classical satisficing system FastDownward to find a deterministic
-%% sequential plan and complementary runtime evolution of that
-%% process. This corresponds to a generalisation of replanning in
-%% probabilistic planning to problems with partial observability.
-
-%% Interaction between the sequential planner and execution proceeds
-%% more-or-less analogously to popular replanning approaches
-
-%% Addressing these challenges in a monolithic framework, we present a
-%% {\em switching} continual planner, that uses the fast sequential
-%% satisficing procedure FastDownward to perform net-benefit
-%% %%
-%% makes reasonable assumptions about the evolution of the runtime state
-%% given a POMDP model of the environment.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% The decision-theoretic planner is able to tailor sensory processing on
-%% a robot platform to the current objective, while FastDownward  quickly
-
-%% In this paper we develop a continual planing system that uses two
-%% planning systems. The first, is a state-of-the-art domain independent
-%% planner for deterministic problems. The second is a information-state
-%% contingency planning the information-state space of
-
-%% Continual planning is a powerful technique that goes some way to
-%% addressing those challenges. That approach interleaves planning and
-%% execution, deliberately postponing planning for contingencies unless
-%% they eventuate during execution.
-
-%% computing a single sequential plan and
-%% eventuality
-
-%% To these challenges it seems a continual planning approach is best,
-
-%% where the runtime state evolves during plan execution
-
-%% interleaved planning and execution.
-
-%% the latter is able to tailor sensory processing on a robot platform,
-%% in order that it.
-
-%% {\em ad-hoc}
-
-%% , reason about degrees of belief and uncertainty about the world,
-
-%% probabilistic sequential decision making in practical sized problems
-%% is intractable.
-
-%% quantitative probabilistic models of the perception and action .
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai11}
-
-\end{document}
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "moritz_2011"
+%%% End:

### 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
shrinking and growing, all at the same time.  Not sure how we are going to get this down to 6 pages yet.

@@
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
+%% From an automated planning perspective, the problem of practical
+%% mobile robot control poses important and contrary challenges.
+%% %%
+%% On the one hand, planning and execution monitoring must be
+%% lightweight, robust, timely, and should span the lifetime of the
+%% robot. Those processes must seamlessly accommodate exogenous events,
+%% changing objectives, and the underlying unpredictability of the
+%% environment.
+%% %%
+%% On the other hand, robot planning should perform computationally
+%% expensive reasoning about contingencies, and possible revisions of
+%% subjective belief according to quantitatively modelled uncertainty in
+%% acting and sensing.
 
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
+For use in an integrated continual planning and execution monitoring
+system on a mobile robot platform, we developed a base planner that
+switches between fast sequential and expensive decision-theoretic
+planning. Given a POMDP model of the environment, sequential planning
+is used to compute an initial deterministic sequential plan and
+complementary runtime evolution of the decision process. That plan is
+executed until testing of an assumption about the runtime state is
+scheduled. At that point a contingent session performs some sensing
+which determines whether or not the sequential session should be
+allowed to continue, or otherwise acts to achieve the overall
+objectives.
 

### 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
shrinking and growing, all at the same time.  Not sure how we are going to get this down to 6 pages yet.

--- 2011-01-28 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
@@
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
+%%  and
+%% using various techniques to manage the large state
+%% space
+
+%% \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
+%% approach in a
+
+Addressing task and observation planning specifically, there have been
+a number of recent developments where the underlying problem is
+modelled as a POMDP.
+%%
+For vision algorithm
+selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit a
+natural hierarchical decomposition of the underlying POMDP into
+subproblems where decision-theoretic planning is reasonably
+fast. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
+represent a preference elicitation problem as a POMDP and take
+advantage of symmetry in the belief space ---essentially the idea that
+it does not matter what the value of the variable you are trying to
+observe is--- to exponentially shrink the state space. Although we
+have been actively exploring the \citeauthor{doshi08:pref_elic}
+approach, those exploitable structures are not present in problems we
+have considered so far due to the task planning requirement. 
+%%
+%%
+Finally,
+our approach is in a similar vein to the more classical {\em
+dual-mode} control~\cite{cassandra96actingunder}, however in our case
+entropy reduction is targeted by planning in an abstract process which
+is informed by one execution trace ---computed by a {\em classical}
+planner--- and the underlying belief-state.
+
+More generally, there has been much recent work on scaling POMDP
+solution procedures to medium-sized
+instances~\cite{shani:etal:08,KurHsu08}. In that
+setting \citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010})
+recently addressed an inefficiency of offline point-based techniques
+in problems with medium length planning horizons, however their
+approach take tens of minutes to good plans, and only scales to tens
+of thousands of states. In the case of general domain-independent
+factored systems, the state-of-the-art scales to relatively small
+problems with $2^{22}$ states~\cite{shani:etal:2008}. At their limit,
+these procedures take over an hour to converge, and $\sim10$ seconds
+on average to perform a single Bellman backup.  For classes of POMDP
+that feature exploitable structures (e.g., no actions with negative
+effects), problems with as many as $10^{30}$ states can be targeted by
+offline procedures~\cite{brunskill:russell:2010}. Moving someway
+towards addressing real-time decision making, recent online POMDP
+solution procedures have been developed which can exploit highly
+approximate value functions -- typically computed using a point-based
+procedure -- and heuristics in forward
+search~\cite{ross:etal:2008}. Their applicability in our setting is
+limited, firstly because they only scale to smaller problems, with
+thousands of states, and also due to the large amount of
+\emph{problem-specific} offline processing that might be required to get useful
+search guidance. A {\em very} recent and promising online approach for
+large POMDPs employs Monte-Carlo sampling to break the curse of
+dimensionality in situations where goal reachability is {\em easily}
+determined~\cite{silver:veness:2010}.
+
+%% Although we suppose it an interesting
+%% item for future work to pursue that direction, it should be noted that
+%% ease of goal reachability is not guaranteed in the problems we face,
+%% and is certainly not a property to be assumed in domain independent
+%% planning.
+
+In the direction of leveraging {\em classical} systems/approaches for
+planning under uncertainty, the most highlighted system to date has
+been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
+the probabilistic track of the 2004 International Planning
+Competition.  In the continual paradigm, \system{FFR$_a$}
+uses \system{FF} to compute sequential plans and execution traces.
 %%
@@
 sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
+deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
+%% robust sequential plan, or contingent
+%% plan~\cite{majercik:2006}.
 
@@
 
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.
+%%No! They simply haven't been evaluated in PO settings. They may, or
+%%may not struggle. They have sampling of traces, and that would
+%%include observations, and therefore evolutions of beliefs. SSAT was
+%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
+%%suited to POMDPs.
+
+% Normally if we are going to compare with related work, we do
+% actually *compare*. Why didn't you try those approaches? I think
+% it's safe to say that FFR will struggle. Why would it even include
+% in its plan an observational action that doesn't change the world?
+
+%%
+%% However, as we said in the introduction,
+%% all these approaches struggle in partially observable domains as they
+%% rely on being able to determine the state at all times.

### 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
shrinking and growing, all at the same time.  Not sure how we are going to get this down to 6 pages yet.
completed introduction

@@
 events, changing objectives, and the underlying unpredictability of
-the environment.  On the other hand, in order to exhibit useful
+the environment.  On the other hand, in order to exhibit efficient
 behaviours the planning process must perform computationally expensive
@@
 
-\section{Propositional Decision-Theoretic Planning}
-\input{prop-dtp.tex}
+%% \section{Propositional Decision-Theoretic Planning}
+%% \input{prop-dtp.tex}
+
+\section{Background and Related Work}
+\input{relatex.tex}
 

### 2011-02-04 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
Changed intro, and removed the CP paper citation

--- 2011-02-03 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
@@
 More generally, there has been much recent work on scaling POMDP
-solution procedures to medium-sized instances~\cite{shani:etal:08}. In
-the case of general domain-independent factored systems, the
-state-of-the-art scales to relatively small problems with $2^{22}$
-states~\cite{shani:etal:2008}. At their limit, these procedures take
-over an hour to converge, and $\sim10$ seconds on average to perform a
-single Bellman backup.  For classes of POMDP that feature exploitable
-structures (e.g., no actions with negative effects), problems with as
-many as $10^{30}$ states can be targeted by offline
-procedures~\cite{brunskill:russell:2010}. Moving someway towards
-addressing real-time decision making, recent online POMDP solution
-procedures have been developed which can exploit highly approximate
-value functions -- typically computed using a point-based procedure --
-and heuristics in forward search~\cite{ross:etal:2008}. Their
-applicability in our setting is limited, firstly because they only
-scale to smaller problems, with thousands of states, and also due to
-the large amount of
+solution procedures to medium-sized
+instances~\cite{shani:etal:08,KurHsu08}. In that
+setting \citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010})
+recently addressed an inefficiency of offline point-based techniques
+in problems with medium length planning horizons, however their
+approach take tens of minutes to good plans, and only scales to tens
+of thousands of states. In the case of general domain-independent
+factored systems, the state-of-the-art scales to relatively small
+problems with $2^{22}$ states~\cite{shani:etal:2008}. At their limit,
+these procedures take over an hour to converge, and $\sim10$ seconds
+on average to perform a single Bellman backup.  For classes of POMDP
+that feature exploitable structures (e.g., no actions with negative
+effects), problems with as many as $10^{30}$ states can be targeted by
+offline procedures~\cite{brunskill:russell:2010}. Moving someway
+towards addressing real-time decision making, recent online POMDP
+solution procedures have been developed which can exploit highly
+approximate value functions -- typically computed using a point-based
+procedure -- and heuristics in forward
+search~\cite{ross:etal:2008}. Their applicability in our setting is
+limited, firstly because they only scale to smaller problems, with
+thousands of states, and also due to the large amount of
 \emph{problem-specific} offline processing that might be required to get useful
@@
 the probabilistic track of the 2004 International Planning
-Competition.  In the continual paradigm, \system{FFR$_a$} uses the
-fast satisficing procedure \system{FF}~\cite{hoffmann:nebel:2001} to
-compute sequential plans and corresponding execution traces.
+Competition.  In the continual paradigm, \system{FFR$_a$}
+uses \system{FF} to compute sequential plans and execution traces.
 %%
@@
 sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
+deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
+%% robust sequential plan, or contingent
+%% plan~\cite{majercik:2006}.
 
@@
 %%
-The continual planning system that motivated our
-project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-and has been applied in completely observable domains, particularly
-those featuring multiple communicating agents.
+%% The continual planning system that motivated our
+%% project~\cite{brenner:nebel:jaamas09} also has this characteristic,
+%% and has been applied in completely observable domains, particularly
+%% those featuring multiple communicating agents.
 

### 2011-02-04 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
New dtpddl section

@@
 
-\section{Related Work}
+\section{Background and Related Work}
 \input{relatex.tex}
 
-\section{Decision-Theoretic PDDL}
+\section{Planning Language and Notations}
 \input{dtpddl.tex}

### 2011-02-04 subarchitectures/planner.sa/trunk/src/latex/AAAI11/dtpddl.tex
New dtpddl section

%% BEGIN SCRAPPY

%% deal with POND and the Perseus language

%% For example, the robot usually knows exactly where it is located on a
%% topological map, and therefore there are tight constraints about what
%% regions of the map can be observed, and what regions of the map can be
%% moved to in an atomic step.

%% Moreover, we use this domain description to model the action physics
%% of abstract processes posed to decision-theoretic planning
%% sessions. Here, we forbid the planner from considering courses of
%% action that are not {\em known} to be possible to alleviate the
%% computational burden planning in the abstract process.

%% END SCRAPPY

We give an overview of the declarative first-order language DTPDDL, an
extension of PPDDL that can express probabilistic models of the
sensing consequences of acting, to quantitatively capture
unreliability in perception. There are straightforward compilations
from problems expressed in DTPDDL to flat (and factored)
representations of the underlying decision processes. Although similar
to Bryce's POND input language, DTPDDL distinguishes itself by
explicitly treating state and perceptual symbols separately, and by
providing distinct declarations for operators (i.e, state model) and
senses (i.e., observation model). In this last respect, DTPDDL admits
more compact domain descriptions where sensing effects are common
across multiple actions. In detail, DTPDDL has perceptual analogues of
fluent and predicate symbols. For example our {\em object search}
domains have:

\vspace{-1ex}
\small
\begin{tabtt}
(\=:functions  ;; state fluents\\
  \> (is-in ?v - visual-object) - location )\\
(:perceptual-functions  ;; perceptual fluents\\
  \> (o-is-in ?v - visual-object) - location )
\end{tabtt}
\normalsize
\vspace{-1ex}

\noindent Where the first symbol models the actual location of
objects, and the second symbol the instantaneous sensing of objects
following application of an action with sensing consequences.
%%
To model sensing capabilities, we have operator-like declarations,
with preconditions expressed using state and action symbols, and
uniformly positive effects over perceptual symbols. For example, where
{\em look-for-object} is the operator that applies an object detection
algorithm at a specific place, in a simplified {\em object search} task
we have:

\vspace{-1ex}
\small
\begin{tabtt}
(\= :sense vision  :parameters \+ \\
  (?r -robot ?v -visual-object ?l -location) \\
 :execution  (look-for-object ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :\=effect (and \+\\
    (\= when (= (is-in ?v) ?l)\\
    \>(probabilistic .8 (= (o-is-in ?v) ?l))) \\
   (when (not (= (is-in ?v) ?l)) \\
    \>(probabilistic .1 (= (o-is-in ?v) ?l))))) \\
\end{tabtt}
\normalsize
\vspace{-3ex}

We now review the DTPDDL syntax for describing an initial state
distribution, taken verbatim from PPDDL, in order to aid us in
discussing our planner. That distribution is expressed in a tree-like
structure of terms. Each term is either: (1) atomic, e.g., a state
proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
probabilistic and atomic terms. The root term is always conjunctive,
and the leaves are atomic. For example, a simplified object search
could have:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$ expresses the
conjunctive root of the tree -- i.e., the root node
$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
$(\pp{and}~\prop)$, for conjunctive terms that contain a single atomic
subterm.}

\vspace{-1ex}
\small
\begin{tabtt}
(\=:init (= (is-in DORA) kitchen) \+ \\
       (probabilistic \=.8 (= (is-in box) office)  \\
		      \>.2 (= (is-in box) kitchen)) \\
       (probabilistic .3 (= (is-in cup) office)  \\
		      \>.7 (= (is-in cup) kitchen))) \\
\end{tabtt}
\normalsize

\vspace{-3ex}

\noindent The interpretation is given
by a {\em visitation} of terms: An atom is {\em visited} iff
its conjunctive parent is visited, and a conjunctive term is visited
iff all its immediate subterms are visited. A probabilistic term is
visited iff its conjunctive parent is visited, and exactly one of its
subterms, $T_i$, is visited. Each visitation of the root term
according to this recursive definition encapsulates a starting state,
along with the probability that it occurs. The former corresponds to the
union of all visited atoms, and the latter corresponds to the product
of $\prob_i$ entries on the visited subterms of probabilistic
elements. Making this concrete, the above example yields the
following flat distribution:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
\hline
\end{tabular}
\normalsize

%% From hereon, to simplify the discussion (and implementation)
%% we shall restrict our attention to POMDPs with deterministic
%% actions. We note that POMDPs with stochastic
%% actions can be compiled into equivalent deterministic-action POMDPs,
%% where all the original action uncertainty is expressed in the
%% starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
%% that of finite-horizon planning, such a compilation yields a finite
%% flat representation of the original POMDP, only with deterministic
%% actions.

%% The syntax and semantics for describing initial state distributions in
%% DTPDDL is taken verbatim from PPDDL. It is useful to present that
%% factored tree-like structure here,

%% \noindent Here, the {\em execution} declaration is a lifted description of
%% the sense action-precondition -- i.e,
%% $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
%% $\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
%% we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
%% order to fully demonstrate the syntax. Interpreting the above schema,
%% if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
%% executed, there is a $0.8$ chance of perceiving the box if it is in
%% the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.

%% \small
%% \begin{tabtt}
%% (\= :sense vision \+\\
%%  :parameters \= (\= ?r - robot ?v - visual-object\\
%%  \>\>  ?l - location) \\
%%  :execution \> ( \> look-for-object ?r ?v ?l) \\
%%  :precondition (and (= (is-in ?r) ?l) ) \\
%%  :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
%%    \> \> (probabilistic 0.8 \\
%%    \>  \>(= (o-is-in ?v) ?l))) \\
%%   \> (when (not (= (is-in ?v) ?l)) \\
%%    \>  \> (probabilistic 0.1 \\
%%    \>  \> (= (o-is-in ?v) ?l))))) \\
%% \end{tabtt}
%% \normalsize

%% A
%% sense declaration corresponds to a lifted description of proposition
%% {\em senses}. Whereas the effects of an {\em operator} schema describe
%% how states change under application of actions, the effects of a {\em
%% sense} schema are perceptual, specifying the composition of an
%% observation following the execution of an action.

%% Also, and without a loss of generality, we find it
%% convenient to separate the actional precondition from the state
%% precondition.

%% Making the above ideas concrete with an example, suppose a robot
%% called $\pp{DORA}$ is able to look for a visual-object, such as a
%% $\pp{box}$, at a given place. We can model that deterministic action
%% using the following operator schema:

%% \small
%% \begin{tabtt}
%% (\=:action look-for-object \+ \\
%%    :parameters (\=?r - robot ?v - visual-object\\
%%    \> ?l - location) \\
%%    :precondition (and (= (is-in ?r) ?l) ) \\
%%    :effect (and (assign (reward) -3) ) ) \\
%% \end{tabtt}
%% \normalsize

%% \noindent In a slight departure from PPDDL, we suppose that
%% $\pp{look-for-object}$ has an effect on the state. That is, its
%% execution incurs an instantaneous penalty, $3$, that corresponds to
%% the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
%% reward} is a {\em reserved word}, and occurs in {\em increased} and
%% {\em decreased} terms in operator schemata. In that setting, reward is
%% accumulated, whereas in DTPDDL it is instantaneous.}

%% The modelling language of choice for planning in probabilistic
%% problems is the Probabilistic Planning Domain Definition
%% Language~\cite{younes:etal:2005}. PPDDL has been
%% used in all three of the International Planning Competitions since
%% 2004. A variation on PDDL for describing domains with stochastic
%% actions and uncertain starting configurations, PPDDL is a declarative
%% first-order language that facilitates factored descriptions of domains
%% and problems. There are straightforward compilations from problems
%% expressed in PPDDL to propositional representations amenable
%% to state-of-the-art planners.

%% Because PPDDL cannot model domains that feature partial observability,
%% we develop an extension we call {\em Decision-Theoretic
%% (DT)PDDL}. This can express probabilistic models of the sensing
%% consequences of acting, to quantitatively capture unreliability in
%% perception. That expressive power is achieved by incorporating
%% perceptual analogues of fluent, predicate, and action definitions. In
%% detail, we have declarations of state characterising predicate and
%% fluent symbols according to the PPDDL syntax. In addition, we allow
%% two other declarations, for perceptual predicates and fluents
%% respectively. For example, suppose our robot is tasked with exploring
%% {\em locations} in order to identify the whereabouts of a {\em
%% visual-object}. We must describe state and perceptual facts that model
%% the {\em true}, resp. perceived, locations of objects. In DTPDDL,
%% these declarations appear as:

### 2011-02-04 subarchitectures/planner.sa/trunk/src/latex/AAAI11/dtpddl.tex
small change

@@
 
-\noindent Where the first symbol models the actual location of
-objects, and the second symbol the instantaneous sensing of objects
+\noindent Where the first fluent symbol models the actual location of
+objects, and the second the instantaneous sensing of objects
 following application of an action with sensing consequences.

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Few small changes

@@
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
+We now describe our {\em switching} planning system that operates
+according to the continual planning paradigm, given descriptions of
+the current problem and domain in DTPDDL. Recapping, the system {\em
+switches}, in the sense that planning proceeds in interleaved
+sessions, in which the base planner is either {\em sequential}/linear
+or {\em decision-theoretic}.
+%%
+During a sequential session, a rewarding {\em trace} of a possible
+execution is computed using a cost-optimising satisficing planner
+which trades action costs, goal rewards, and determinacy.
+%%
+Formatted as a linear plan, the trace specifies a sequence of actions
+that achieves some reachable goals following a deterministic
+approximation of the problem at hand.
+%%
+Structurally, it is a sequence of elements that are either: (i) ground
+actions from the DTPDDL description of the world, or (ii) atomic {\em
+assumptions}, modelled as deterministic actions, made about the truth
+value of facts that can only be determined at runtime (e.g., that a
+box of cornflakes is located on the corner bench in the kitchen).
+%%
+The system always begins in a sequential session, so that plan
+execution proceeds by applying DTPDDL actions from the trace in
+sequence until the applicability of the next scheduled action is too
+uncertain according to a threshold parameter~(here, 95\%). In that
+eventuality a DT session commences, which tailors sensory processing
+to determine whether the assumptions made in the trace hold, or which
+otherwise acts to achieve the goals.
 
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
-%%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
+Because online DT planning in large problems is impractical (we seek
+response times in seconds), DT sessions plan in an abstract process
+determined by the current trace and underlying belief-state. This
+abstraction is constructed by first excluding all propositions (and
+related actions) that are not true of any state in the trace, then
+adding them back, using as a heuristic the entropy of the trace
+assumptions conditional on a candidate proposition. Propositions are
+added, one at a time, until the number of states in the initial
+belief-state reaches a given threshold.\footnote{Our experiments are
+performed using a number of different threshold parameters.}  To the
+resulting abstract model we also add {\em disconfirm} and {\em
+confirm} actions that the contingent session can schedule in order to
+judge an atomic assumption in the trace. In the abstract model these
+actions yield a relatively small reward if the corresponding judgement
+is true (or small penalty otherwise). If a judgement action is
+scheduled for execution the contingent session is terminated, and
+control is returned to a sequential session.
 
@@
 continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
-
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
-
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
+belief-states by performing belief revision. Our internal
+representation of the underlying belief-state corresponds more-or-less
+directly to an $(\pp{:init})$ declaration. That belief-state
+distribution is used: (1) as the source of candidate determinisations
+for sequential planning, (2) in determining when to switch to a DT
+session, and (3) as a mechanism to guide construction of an abstract
+process for contingent sessions.
 

### 2011-02-05 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
switching is starting to get written again

--- 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
@@
 
-\subsection{Sequential Planning in POMDPs}
+\subsection{Sequential Sessions}
 
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
+We consider only deterministic-action POMDs, where all the uncertainty
+in state and action is expressed in the $(\pp{:init})$ declaration. In
+sequential sessions, this declaration is used by our approach to
+define a starting state, and the set of state-assumptions available to
+sequential planning.  Writing \#\ if the value of a proposition is
+unspecified, continuing our example from the previous section, we have
+the following assumptions:
 
@@
 \hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
+%% .24 & kitchen & office & office \\
+%% .06 & kitchen & kitchen & office \\
+%% .56 & kitchen & office & kitchen \\
+%% .14 & kitchen & kitchen & kitchen \\
 .7 & kitchen & \# &  kitchen\\
@@
 
-\noindent An assumption corresponds to a {\em
+\noindent Each assumption corresponds to one distinct {\em
 relaxed} visitation of the root term of $(\pp{:init})$. In this
 relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
+are visited, and zero or one of its immediate probabilistic subterms
+are visited. For a sequential session, the starting state corresponds
+to the unique assumption with probability $1$ -- I.e., the union of
+facts from the root conjunctive term. In our example, the starting
+state is:
 
@@
 
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
+We also annotate DTPDDL model with an assumptive action
+$\assumptiveS{i}$ for each element $\prob_i (T_i)$, of each
+probabilistic term from $(\pp{:init})$. Here, $\assumptiveS{i}$ can be
+executed if no $\assumptiveS{j}$, $j \neq i$, has been executed from
+the same probabilistic term, and, either
+$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
+it occurs in $T_k$ for some executed $\assumptiveS{k}$. 
 %%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
+We add additional constraints that forbid the scheduling of
+assumptions about facts after actions with preconditions, or effects,
+that mention those facts. For example, the robot cannot assume it is
+plugged into a power source immediately after it unplugs itself.
 %%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
+Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
+to a successor state $\state^{T_i}$ with probability $\prob_i$, and
 $\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
+the union of $\state$ with atomic terms from $T_i$. State
+$\state^\bot$ is an added sink.
 
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
+%% %%
+%% Lastly, we have that if $\action$ has a precondition that is not true
+%% with probability $\theta$, a given threshold, in the underlying belief
+%% $\bstate$, then immediately before it is executed blindly the system
+%% switches to a contingent session (described below).
 
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
-
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
+We now describe the optimisation criteria used during sequential
+sessions. Where $\prob_i$ is the probability that the $i^{th}$
 sequenced action $\action_i$ does not transition to $\state^\bot$, we
@@
 
-\subsection{Decision-Theoretic Planning in Abstractions}
+\subsection{DT Sessions}
 

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/dtpddl.tex
Only one extra column to remove now. Any ideas people?!

@@
 \begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
+(\=:init (= (is-in R2D2) kitchen) \+ \\
        (probabilistic \=.8 (= (is-in box) office)  \\
@@
 \hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
 \hline
@@
 %% these declarations appear as:
+
+If a dual is executed, then the sequential planner must
+replan with $\bstate_0$ equal to the underlying
+belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
+executed, then the current sequential plan is executed until further
+sensing is scheduled, or to completion.
+
+% \small
+% \begin{tabtt}
+% (\=:init (= (is-in DORA) kitchen) \+ \\
+%        (probabilistic \=.8 (= (is-in box) office)  \\
+% 		      \>.2 (= (is-in box) kitchen))) \\
+% \end{tabtt}
+% \normalsize
+
+% \noindent and therefore, the belief-state $\bstate_0$ is:
+
+% \small
+% \begin{tabular}{cccc}
+% \hline
+% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
+% \hline
+% .8 & kitchen & office & \# \\
+% .2 & kitchen & kitchen & \# \\
+% \hline
+% \end{tabular}
+% \normalsize
+
+ % With
+% regards to abstraction, the contingent and serial sessions apply the
+% same applicability conditions ---i.e., that each effect and condition
+% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
+% problem description.

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Only one extra column to remove now. Any ideas people?!

@@
 \hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
 \hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
+%% .24 & kitchen & office & office \\
+%% .06 & kitchen & kitchen & office \\
+%% .56 & kitchen & office & kitchen \\
+%% .14 & kitchen & kitchen & kitchen \\
 .7 & kitchen & \# &  kitchen\\
@@
 
-\noindent An assumption corresponds to a {\em
+\noindent Each assumption corresponds to one distinct {\em
 relaxed} visitation of the root term of $(\pp{:init})$. In this
 relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
+are visited, and zero or one of its immediate probabilistic subterms
+are visited. For a sequential session, the starting state corresponds
+to the unique assumption with probability $1$ -- I.e., the union of
+facts from the root conjunctive term. In our example, the starting
+state is:
 
+\small
 \[
 \begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
+\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
 \;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
@@
 \]
+\normalsize
 
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
+We also annotate DTPDDL model with an assumptive action
+$\assumptiveS{i}$ for each element $\prob_i (T_i)$, of each
+probabilistic term from $(\pp{:init})$. Here, $\assumptiveS{i}$ can be
+executed if no $\assumptiveS{j}$, $j \neq i$, has been executed from
+the same probabilistic term, and, either
+$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
+it occurs in $T_k$ for some executed $\assumptiveS{k}$. 
 %%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
+We add additional constraints that forbid the scheduling of
+assumptions about facts after actions with preconditions, or effects,
+that mention those facts. For example, the robot cannot assume it is
+plugged into a power source immediately after it unplugs itself.
 %%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
+Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
+to a successor state $\state^{T_i}$ with probability $\prob_i$, and
 $\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
+the union of $\state$ with atomic terms from $T_i$. State
+$\state^\bot$ is an added sink.
 
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
+We now describe the optimisation criteria used during sequential
+sessions. Where $\prob_i$ is the probability that the $i^{th}$
+sequenced action $\action_i$ from trace,
+$\state_0, \action_0,\state_1, \action_1,.., \state_N$, does not
+transition to $\state^\bot$, we have that the optimal trace has value:
 
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
+%% define the value of a trace
+%% $\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
+%% deterministic model to equal:
 
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
+%% \small
+%% \begin{equation}\label{eq:tracevalue}
+%% V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
+%% \end{equation}
+%% \normalsize
 
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
+%% \noindent The
+%% optimal trace given a plan, i.e., the sequence of non-assumptive
+%% actions from the trace, therefore has value:
 
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
+\small
 \[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
+V^* = \max_N \max_{\state_0, \action_0,.., \state_N} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
 \reward(\state_i, \action_i),
 \]
+\normalsize
 
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
+\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
+reward ---i.e., can be negative for action costs, or positive for goal
+achievement--- received for executing action $\action_i$ in state
+$\state_i$. Finally, it is worth clarifying that in practise we do not
+artificially limit the length of traces that the sequential planner
+can consider. Moreover, in our setting there is always a finite
+optimal trace.
 
-%% From the perspective of
-%% gradient-based reinforcement learning systems, such as William's
-%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-%% under reasonable assumptions the optimal trace identifies a maximal
-%% gradient step from a uniformly random soft-max policy.
+\subsection{DT Sessions}
 
-\subsection{Decision-Theoretic Planning in Abstractions}
+The DT session plans in a {\em small} abstract processes defined
+according to the assumptive actions in the proceeding trace, and the
+$(\pp{:init})$ declaration that models the underlying
+belief-state. Targeted sensing is encouraged by augmenting the reward
+model to reflect the value of knowing the truth about assumptions in
+the trace proposed by the preceding sequential deliberation. In
+detail, all rewards from the original POMDP are
+retained. Additionally, for each $\assumptiveS{i}$ action scheduled by
+the current trace, we have a {\em dual} $\assumptiveDT{i}$ so that
+forall states $\state$:
 
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
-
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
+\small
 \[
@@
 \]
+\normalsize
 
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
+\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
+relatively\footnote{Compared to rewards from the original process
+model.} small positive (negative) numeric quantity which captures the
+utility the agent receives for correctly (incorrectly) rejecting an
+assumption. Dual action execution is modelled as a self-transition --
+I.e., the process remains in the state from which the dual is
+executed.
 %%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
+Furthermore, if \switchAction\ is the action that switched the system
+to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
+active} if
 \switchAction\ is not applicable in the sequential model unless
 $\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
+to \switchAction. We only include the duals of active assumptions in
+DT sessions. For example, if the trace is:
 
+\small
 \[
@@
 \actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
+(\pp{look}~\pp{box}~\pp{office});
+(\pp{look}~\pp{cup}~\pp{kitchen});\\
+(\pp{report}~\pp{box}~\pp{office}); 
+(\pp{report}~\pp{cup}~\pp{kitchen})
 \end{array}
 \]
+\normalsize
 
 \noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
+$(\pp{look}~\pp{box}~\pp{office})$, we have that
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
@@
 $\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
+POMDP posed to the DT session.
 
-We also include another propositionally trivial action
+Given \switchAction, we also include another self-transition action
 $\actions.\poss(\switchAction)$ with the reward property:
@@
 
-For the purposes of successive switching, execution of either a dual
+\noindent We note that in our project the dual rewards are sourced
+from a motivational component of our robotic architecture. In this
+paper, for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a positive
+constant, and have $\hat\$(x)= - \$(x)(1 - \prob) / \prob$ where
+$\prob$ is the probability that $x$ is true. For
+$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
+\$(x)\prob/(1-\prob)$
+%%
+Finally, for the purposes of successive switching, execution of either a dual
 action of the form $\assumptiveDT{i}$, or the
 $\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
+session, which then continues from the underlying belief-state. If an
+assumption was rejected, we prohibit that sequential session from
+making it again.
+%%
 
@@
 
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
 In the first step, we remove all facts that are not part of an

### 2011-02-05 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
6.5 pages

@@
 effects), problems with as many as $10^{30}$ states can be targeted by
-offline procedures~\cite{brunskill:russell:2010}. Moving someway
+such offline procedures~\cite{brunskill:russell:2010}. Moving someway
 towards addressing real-time decision making, recent online POMDP

### 2011-02-05 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
Oops. Didn't mean to do a commit then

--- 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
-%%%% ijcai11.tex
-
-\typeout{IJCAI Switching Planner Submission}
-
-% These are the instructions for authors for IJCAI-11.
-% They are the same as the ones for IJCAI-07 with superficical wording
-%   changes only.
-
-\documentclass{article}
-% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
-\usepackage{ijcai11}
-
-% Use the postscript times font!
-\usepackage{times}
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\usepackage{subfig}
-\usepackage{tikz}
-\usetikzlibrary{shapes}
-\usetikzlibrary{positioning}
-
-\usepackage{graphicx}
-\graphicspath{{plots/}}
-\include{macros}%
-
-\title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Planning for agents acting in partially observable environments with
-stochastic actions is extremely computationally
-expensive~\cite{mdp-complexity}. Even relatively small partially
-observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, have state spaces far larger than existing
-optimal or close to optimal planners can handle. For the easier
-problem of planning in completely observable stochastic domains, an
-alternative has been to use classical sequential planners and
-replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
-Execution monitoring tracks properties of the state that determine
-whether further planning is required. For partially observable
-domains, only probabilistic information about such properties will, in
-general, be available.
-
-%% However, these approaches rely
-%% crucially on being able to determine properties of the current state
-%% in order to decide to replan. For partially observable domains, only
-%% probabilistic information about these properties will in general be
-%% available.
-
-This paper describes the planning component for a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings. This domain features partial observability, particularly
-because the state is not perfectly known, and state information is
-gained by the robot through the use of unreliable sensing
-actions. 
-%Using a propositionally factored state representation, for
-For
-interesting sized tasks the corresponding POMDP model is too large for a
-POMDP solver to be applied directly. Inspired by the replanning
-approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
-planning approach that uses a classical planning system to compute a
-reasonably valuable trace in the model, and then uses a
-decision-theoretic planner on small subproblems where reasoning about
-observations might be useful. We refer to the complete system as a
-{\em switching continual planner}. The approach is not optimal,
-particularly as it relies on the results of sequential planning
-directly. It does nevertheless perform better than a purely sequential
-replanner, and is fast enough to be used for real-time decision-making
-on a mobile robot.
-
-%% Switching gives us some important advantages in our robotic
-%% domain. First, the significant source of uncertainty in the domain is
-%% the unreliability of observations, and this approach is particularly
-%% suited to problems that combine a deterministic task planning problem
-%% with observations as the decision-theoretic planner is often invoked
-%% to determine a characteristic of the underlying state. Secondly,
-%% replanning makes our system robust to changing objectives and the
-%% discovery of new facts about the world.
-
-%% As an illustration, consider a robot searching for a box that it knows
-%% is often found in offices. The sequential planner might produce a plan
-%% that assumes that {\sc Room a} is an office, that the box is in {\sc
-%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
-%% these assumptions, the plan might be to go to {\sc Room a}, go to
-%% location 1 and search for the box. When the plan gets executed and the
-%% search begins the decision-theoretic planner is called since searching
-%% involves making observations. The decision-theoretic planner might
-%% retract the last assumption but leave the others, and build a plan to
-%% search the whole of {\sc Room a} for the box. If this plan fails to
-%% find the box, it will disprove the assumption that the box is in {\sc
-%% Room a}, which will then lead to replanning by the sequential planner
-%% based on the newly discovered knowledge.
-
-Our system is domain independent, taking domain and problem
-descriptions in a first-order declarative language we have developed,
-called the decision-theoretic planning domain definition language
-(DTPDDL). From these descriptions of POMDPs, we automatically
-construct a deterministic model suitable for sequential planners. That
-model includes actions which correspond to making assumptions about
-the values of imperfectly known state variables. Assumptions scheduled
-by the sequential system are used to propose a pragmatic abstract
-belief-state space to the decision-theoretic system, and to modify the
-reward function, so that system might pursue sensing related to those
-assumptions.
-
-%%  which can operate
-%% with the original representation of the domain.
-
-%% The sequential planner we use is based
-%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
-%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
-%% knowledge to be represented explicitly, so we can write actions that
-%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
-%% THIS. For the decision-theoretic planner we have implemented our own
-%% forward search in the belief-state space.
-
-The switching planner provides important advantages in our
-mobile robot domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations because the decision-theoretic planner
-is only invoked to deal with state uncertainty.
-%determine a characteristic of the underlying state. 
-Secondly, replanning makes the system robust to
-changing objectives and discoveries about the world,
-both of which feature in our domain.
-
-The remainder of the paper proceeds as follows. We begin by describing
-POMDPs with a propositionally factored representation of states and
-observations. We also outline the relationship between this and a flat
-POMDP representation. Next we describe our DTPDDL language with
-example declarations from a mobile robot exploration task. Then we
-describe the switching planner, including how the sequential domain is
-produced and a plan is generated, and how this then provides the input
-for the decision-theoretic planner. Finally we provide an evaluation
-of the system on the example robot domain, and finish with a short
-survey of related work and some concluding remarks.
-
-%To accompany
-%that language we have implemented an information-state
-%\laostar\ procedure for solving problems expressed in DTPDDL.
-
-%%
-
-%Have a good understanding of the contents and function of rooms, as
-%well as the linguistic referrents to rooms, widgets, and their visual
-%qualities.
-
-%Having confidence in its beliefs about the linguistic terms for
-%spaces, and the function of those spaces.
-
-%sometimes difficult to predict, with exogenous events, such as the
-%changing of the objective,
-
-%the speed and scalability of software for sequential planning in
-%deterministic
-
-%% We developed a first-order declarative language, called DTPDDL, for
-%% describing domains of planning problems that correspond to POMDPs.  To
-%% accompany that language we have implemented an information-state
-%% \laostar\ procedure for solving problems expressed in DTPDDL.
-
-%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
-%% modified \fastdownward\ so that it can find useful sequential executions
-%% given DTPDDL models of the problem at hand.
-
-%%  called DTPDDL, along
-%% the lines of PDDL for the partially observable case, an \laostar\
-%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
-%% and modify the \fastdownward\ system to find high quality sequential plans
-%% .
-
-%% We model the environment as a partially observable Markov decision
-%% process. Although planning in that model is undecidable in
-%% general~\cite{}, an optimal finite-horizon plan corresponds to a
-%% contingent plan, that is a function mapping action-observation
-%% histories to actions.
-
-%% our continual planner is reactive, replanning whenever the underlying
-%% domain and problem models change. For example, replanning occurs if
-%% the motivational component alters the objectives, and if an assumed
-%% outcome of a sensing action is not realised.
-
-%% brittle sensing model,
-
-%% the evaluation of a fluent at a state is either known. Moreover, there
-%% is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process state
-$\state$ as the set of propositions that are true of the
-state. Notationally, $\props$ is the set of propositions, $\prop$ is
-an element from that set, and we have $\state \subseteq
-\props$. The underlying process
-dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
-state-characterising propositions $\propositions$.
-%%
-We say an action $\stochAction \in \stochActions$ is
-applicable if its precondition $\poss(\stochAction)$, a set of
-propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature chooses a deterministic STRIPS effect $\detAct_i$, and for
-all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$.
-
-%% We model a process state $\state$ as the set of propositions that are
-%% true of the state. Notationally, we have $\state \subseteq
-%% \props$. State change is induced by application of actions. A
-%% stochastic action $\stochAction \in \stochActions$ is applicable if
-%% its precondition $\poss(\stochAction)$, a set of propositions, are
-%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
-%% applicable at state $\state$.  If action $\stochAction \in
-%% \stochActions(\state)$ is applied at \state, nature chooses one
-%% element amongst a small set of deterministic outcomes
-%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-%% $\sum_{\detAction_i \in \detActions(\stochAction)}
-%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-%% effect on the state given in terms of two lists of propositions. The
-%% add-list $\add(\detAction)$ and delete-list
-%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-%% $\detAction$, then it cannot be in the delete-list and vice versa.}
-%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
-%% applied at state $\state$, then the resultant state is $ ( \state \cup
-%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-%% propositions from $\add(\detAction)$ are added to $\state$, and those
-%% from $\delete(\detAction)$ are removed from $\state$.
-
-%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
-
-We are concerned with problems that feature partial
-observability. Although we could invoke {\em extended probabilistic
-STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for presentation and computation to
-separate sensing and action. Therefore, we suppose a POMDP has a
-perceptual model given in terms of a finite set of stochastic {\em
-senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
-and perceptual propositions $\percepts$, called {\em percepts}. In
-detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote by \observations\ the set of
-reachable observations. The underlying state of the process cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in
-\observations$ that informs what should be believed about the state the
-process is in. If $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
-usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from the chosen elements of active
-senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make some fairly standard assumptions. First, that action
-execution and sensing occur instantaneously, and that only one action
-can be applied per plan-step. Secondly, it can arise that in some
-states an action is applicable, and in others it is not. Moreover, a
-belief-state can assign non-zero probability to states where that
-applicability holds, and states where it does not. %% We differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
-%% In beliefs where an action exists that is applicable in some state 
-%% with non-zero probability, and not applicable in another, we differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%% because 
-%% in that
-%% we forbid the execution of the action.
-%% actions that are not applicable in
-%% any state of the current belief.  
-%% Also,
-Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we incorporate a PPDDL-like default semantics, treating an action
-\stochAction\ executed at $\state_i$ as if it has no effect at
-states $\state$ if $\poss(\stochAction) \not\subseteq
-\state$.
-
-% Mucked around with this, then realised the two sentences were
-% contradictory. If we forbid actions applicable in one state but not
-% another in the belief state, then clearly we aren't using a
-% PPDDL-like semantics.
-%% Doubled comments are stuff I added, then removed.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Until now we have discussed the POMDP in terms of propositions and
-percepts. In order to address belief revision and utility it is
-convenient to consider the underlying decision process in a flat
-format. This is given by the tuple
-$\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-The function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs as
-negative rewards.
-%%
-From $\psi$ we have that for each $\state' \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
-\action)$ the probability of getting observation $\observ$ in state
-$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state', \action) = 1$.
-
-Successive state estimation is by application of Bayes'
-rule.  Taking the current belief $\bstate$ as the {\em prior}, and
-supposing action $\action$ is executed with perceptive outcome
-$\observ$, the probability that we are in $\state'$ in the
-successive belief-state $\bstate'$ is:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
-  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---i.e., reaches a state that satisfies a given
-Boolean condition--- with probability $1$.  Generally, whatever the
-structure of a plan, its value corresponds to the expected cumulative
-reward:
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
-%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
-%% probability that the controller prescribes \action\ when in state
-%% \node, and $\transition_\node(\action, \observation, \node') =
-%% P(\node'|\node, \action, \observation)$ gives the probability the
-%% controller transitions to state $\node'$, supposing we execute
-%% \action\ in state \node\ receiving observation \observation. When
-%% acting for finite $N$ steps from controller state $n$, the value of a
-%% controller corresponds to the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
-the action prescribed at step $t$, and
-
-\[\reward(\bstate, \action) = \sum_{\state \in \states}
-\bstate(\state)\reward(\state, \action).\]
-
-%% \noindent Finally, it is
-%% useful to note that there is a corresponding deterministic FSCs for
-%% both sequential (e.g., conformant) and contingent plan formats.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
-used in all three of the International Planning Competitions since
-2004. A variation on PDDL for describing domains with stochastic
-actions and uncertain starting configurations, PPDDL is a declarative
-first-order language that facilitates factored descriptions of domains
-and problems. There are straightforward compilations from problems
-expressed in PPDDL to propositional representations amenable
-to state-of-the-art planners.
-
-Because PPDDL cannot model domains that feature partial observability,
-we develop an extension we call {\em Decision-Theoretic
-(DT)PDDL}. This can express probabilistic models of the sensing
-consequences of acting, to quantitatively capture unreliability in
-perception. That expressive power is achieved by incorporating
-perceptual analogues of fluent, predicate, and action definitions. In
-detail, we have declarations of state characterising predicate and
-fluent symbols according to the PPDDL syntax. In addition, we allow
-two other declarations, for perceptual predicates and fluents
-respectively. For example, suppose our robot is tasked with exploring
-{\em locations} in order to identify the whereabouts of a {\em
-visual-object}. We must describe state and perceptual facts that model
-the {\em true}, resp. perceived, locations of objects. In DTPDDL,
-these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like declarations, with preconditions expressed using state
-predicate and action function symbols, and uniformly positive effects
-over perceptual predicates. A sense declaration corresponds to a
-lifted description of proposition {\em senses}. Whereas the effects of
-an {\em operator} schema describe how states change under application
-of actions, the effects of a {\em sense} schema are perceptual,
-specifying the composition of an observation following the execution
-of an action.
-
-%% Also, and without a loss of generality, we find it
-%% convenient to separate the actional precondition from the state
-%% precondition.
-
-Making the above ideas concrete with an example, suppose a robot
-called $\pp{DORA}$ is able to look for a visual-object, such as a
-$\pp{box}$, at a given place. We can model that deterministic action
-using the following operator schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs an instantaneous penalty, $3$, that corresponds to
-the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
-reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, reward is
-accumulated, whereas in DTPDDL it is instantaneous.}
-
-Completing the example, we model the sensing that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- i.e,
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
-$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-our switching continual planner. The distribution is expressed in a
-tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
-(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
-probabilistic and atomic terms. The root term is always conjunctive,
-and the leaves are atomic. For example, the starting distribution for
-$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
-$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
-i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
-$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
-that contain a single atomic subterm.}
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-by a {\em visitation} of terms: An atom is {\em visited} iff
-its conjunctive parent is visited, and a conjunctive term is visited
-iff all its immediate subterms are visited. A probabilistic term is
-visited iff its conjunctive parent is visited, and exactly one of its
-subterms, $T_i$, is visited. Each visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability that it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $\prob_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-From hereon, to simplify the discussion (and implementation)
-we shall restrict our attention to POMDPs with deterministic
-actions. We note that POMDPs with stochastic
-actions can be compiled into equivalent deterministic-action POMDPs,
-where all the original action uncertainty is expressed in the
-starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
-that of finite-horizon planning, such a compilation yields a finite
-flat representation of the original POMDP, only with deterministic
-actions.
-
-\section{Switching Continual Planner}
-
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
-
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
-%%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
-
-Finally, whether proceeding in a sequential or contingent session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
-
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
-
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
-
-\subsection{Sequential Planning in POMDPs}
-
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
-
-\noindent An assumption corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
-
-\[
-\begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
-\end{array}
-\]
-
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
-%%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
-%%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
-
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
-
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
-
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
-
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
-\[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
-\reward(\state_i, \action_i),
-\]
-
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
-
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
-\[
-\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
-\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\hat\$(T_i) & \pp{otherwise} \\
-\end{array}
-\]
-
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
-%%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
-
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
-\end{array}
-\]
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
-
-We also include another propositionally trivial action
-$\actions.\poss(\switchAction)$ with the reward property:
-
-\[
-\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
-\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
-\end{array}
-\]
-
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
-
-%%  We make no assumptions here about how they are
-%% derived.
-
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
-
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief-state with the assumptions made by the continual
-    planner in bold.]{
-      \label{fig:abstraction-a}
-      \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode] (cup) {} 
-          child {node [cnode] (cupp1) {}}
-          child {node [cnode] (cupp2) {}}
-        }
-      }
-      child {node[cnode] (kitchen) {}
-        child {node [pnode] (box2) {} 
-          child {node [cnode] (box2p1) {}}
-          child {node [cnode] (box2p2) {}}
-        }
-      };
-   \tiny
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of kitchen] {$\pp{kitchen}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[below=0 of cup] {$\pp{(is-in cup)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-   \node[right=0 of cupp1] {$\pp{place1}$};
-   \node[right=0 of cupp2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Generating the abstract belief space.}
-\label{fig:abstraction}
-\end{figure}
-
-%%% OLDER STUFF
-
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
-
-\section{Evaluation}
+% coming soon
 
 We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. 
-% Our implementation is able
-% to use several underlying planning systems.
-We have extended MAPSIM so that it can parse DTPDDL, and perform
-successive estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
-
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
-}
-
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
+environment~\cite{brenner:nebel:jaamas09},
+using \system{dlib-ml}~\cite{king:2009} for belief
+revision. Sequential sessions use a modified version of Fast
+Downward~\cite{fast-downward}, and DT sessions use our own fast
+contingent procedure. We also implemented a simple dual-mode
+replanning {\em baseline} approach in MAPSIM. Here, a DT session
+executes a single greedy entropy reduction action, wrt to active
+assumptions, when a switching action is scheduled for execution,
+returning control to a new sequential session.
 
@@
 
-We evaluated our switching planner against the baseline in several
+We evaluated our switching planner against the {\em baseline} in several
 tasks with the number of rooms ranging from 3 to 6. We examine the
@@
 particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
+environment, and therefore the {\em baseline} cannot be used, because there
 are no sensing schemata that directly sense room categories. In our
@@
 
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
 Examining success ratios and plan costs, where sensing is reliable
 there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
+greedy sensing approach of the {\em baseline} is sufficient. Not
 surprisingly, as sensing degrades contingent planning pays off.  Also,
@@
 abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% \section{Discussion}
-%% \input{discussion.tex}
-
-\section{Related Work}
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
-
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-%%
-The continual planning system that motivated our
-project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-and has been applied in completely observable domains, particularly
-those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
-
-The most pressing item for future research, is to develop a scheme
-whereby the serial planner can relax {\em executability} assumptions,
-so that conformant (or semi-conformant) plans can be executed without
-interruption by a contingent session. A more general criticism of
-switching continual planning concerns interleaved sequential and
-contingent sessions failing to make any progress towards the
-objectives. For example, in the worst case we can have each sequential
-session producing an identical plan, and each decision-theoretic
-session rejecting it without further sensing. Although this has not
-been an issue in our work so far, it must be dealt with rigorously
-in the future. We suggest that a good way to mitigate this problem is
-by developing a {\em motivational} component that maintains a {\em
-dynamic} reward model whose limiting behaviour prevents that switching
-deadlock.
-
-%% The sequential mode of our system always schedules decision-theoretic
-%% planning before executing actions that are not applicable with
-%% certainty. This is inefficient if the utility of the plan is not
-%% dependent on that actions successful execution. This situation always
-%% arises for conformant plans.
-
-%% The switching continual planning system we have described serves as
-%% the underlying planner for CogX.
-
-%% the effects of a {\em sense} schema are perceptual, whereas the
-%% effects of an {\em operator} schema are over state propositions.
-
-%% posted by a motivational component of the underlying robotic
-%% architecture.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% In this paper we present an approach to continual planning that uses
-%% two planning systems. The first
-
-%%  to a distinct class of
-%% challenges. We suppose 
-%% %%
-%% The underlying environment is modelled as a POMDP. We use the fast
-%% classical satisficing system FastDownward to find a deterministic
-%% sequential plan and complementary runtime evolution of that
-%% process. This corresponds to a generalisation of replanning in
-%% probabilistic planning to problems with partial observability.
-
-%% Interaction between the sequential planner and execution proceeds
-%% more-or-less analogously to popular replanning approaches
-
-%% Addressing these challenges in a monolithic framework, we present a
-%% {\em switching} continual planner, that uses the fast sequential
-%% satisficing procedure FastDownward to perform net-benefit
-%% %%
-%% makes reasonable assumptions about the evolution of the runtime state
-%% given a POMDP model of the environment.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% The decision-theoretic planner is able to tailor sensory processing on
-%% a robot platform to the current objective, while FastDownward  quickly
-
-%% In this paper we develop a continual planing system that uses two
-%% planning systems. The first, is a state-of-the-art domain independent
-%% planner for deterministic problems. The second is a information-state
-%% contingency planning the information-state space of
-
-%% Continual planning is a powerful technique that goes some way to
-%% addressing those challenges. That approach interleaves planning and
-%% execution, deliberately postponing planning for contingencies unless
-%% they eventuate during execution.
-
-%% computing a single sequential plan and
-%% eventuality
-
-%% To these challenges it seems a continual planning approach is best,
-
-%% where the runtime state evolves during plan execution
-
-%% interleaved planning and execution.
-
-%% the latter is able to tailor sensory processing on a robot platform,
-%% in order that it.
-
-%% {\em ad-hoc}
-
-%% , reason about degrees of belief and uncertainty about the world,
-
-%% probabilistic sequential decision making in practical sized problems
-%% is intractable.
-
-%% quantitative probabilistic models of the perception and action .
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai11}
-
-\end{document}

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
experimental results section is still shit! I don't get the task f problem yet

--- 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
@@
-%% From an automated planning perspective, the problem of practical
-%% mobile robot control poses important and contrary challenges.
-%% %%
-%% On the one hand, planning and execution monitoring must be
-%% lightweight, robust, timely, and should span the lifetime of the
-%% robot. Those processes must seamlessly accommodate exogenous events,
-%% changing objectives, and the underlying unpredictability of the
-%% environment.
-%% %%
-%% On the other hand, robot planning should perform computationally
-%% expensive reasoning about contingencies, and possible revisions of
-%% subjective belief according to quantitatively modelled uncertainty in
-%% acting and sensing.
-
 For use in an integrated continual planning and execution monitoring
@@
 switches between fast sequential and expensive decision-theoretic
-planning. Given a POMDP model of the environment, sequential planning
+planning. Given a DTPDDL model of the environment, sequential planning
 is used to compute an initial deterministic sequential plan and
@@
 allowed to continue, or otherwise acts to achieve the overall
-objectives.
+objectives. We find our approach performs quickly and relatively
+efficiently in large real-world task and observation planning
+problems.
 
+\Omit{
 The most pressing item for future research, is to develop a scheme
@@
 deadlock.
+}
 

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
experimental results section is still shit! I don't get the task f problem yet

% coming soon

We have implemented our switching continual approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09},
using \system{dlib-ml}~\cite{king:2009} for belief
revision. Sequential sessions use a modified version of Fast
Downward~\cite{fast-downward}, and DT sessions use our own contingent
procedure. We also implemented a simple dual-mode replanning {\em
baseline} approach. Here, when a switching action is scheduled for
execution the DT session executes a single entropy reduction action,
that directly senses an active assumption -- I.e,. We say a fact can
be directly sensed, if there is a DTPDDL {\em sense} declaration that
mentions the fact in its precondition.  Control is then immediately
returned to a new sequential session.

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consists of {\em rooms}
(office/kitchen/etc), and an underlying topological map that models
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological space, the latter indicate the category of space they
inhabit -- E.g., Spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask them about the location of
objects.

We compare our switching planner against the {\em baseline} in several
realistic tasks, with the number of rooms ranging from 3 (12-places,
13-objects, $|$states$|$$>\sim10^{48}$) to 6 (26-places, 21-objects,
$|$states$|$ $>10^{48}$). We experiment with a sensing model that is:
({\em reliable}) $0.1$ probability of false negative, ({\em
semi-reliable}) chance is $0.35$, and ({\em noisy}) chance is $0.4$.
We evaluate with DT sessions taking $b_0$ admitting between 20 and 200
abstract states with non-zero probability. We run 50 simulations in
each configuration, and have a timeout on each simulation of 30
minutes (1800 seconds)\footnote{All experiments were conducted on a
2.3GHz AMD Opteron using one CPU core.}. The continual planning times
are reported in Figure~\ref{fig:results-time}, and the quality data in
Figure~\ref{fig:results-quality}.
%%
For Items~$a$-$e$ the goal is to find objects and report their
position to a user. In some cases there is a non-zero probability that
no plan exists, as the desired object might not be present in the
environment. Item~$f$ reports for tasks requiring indirect sensing,
where the robot must relocate to a room with a particular category
(e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
so is omitted from that reporting.  Because reward is only allocated
once, on the achievement of the goal,  we find it
intuitive to report average plan costs and the success rates in
problems that admit a solution (i.e., positive reward scaled by a
constant factor).

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-time}\hfill
  \vspace{2mm}
  \includegraphics{dora56-time}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-time}\hfill
  \caption{Average runtime}
  \label{fig:results-time}
\end{figure}

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

We find that if sensing is reliable there is little to be gained
through DT sessions, as the greedy approach of the {\em baseline} is
sufficient. As sensing degrades contingent planning pays off.  Time
spent in DT planning increases steeply as the abstraction $\bstate_0$
becomes more refined.  That refinement seems to be paying off in terms
of the success rate, particularly for tasks $d$ and $e$. For less
refined initial configurations, the increase cost of DT
planning is compensated by a decrease in Fast Downward planning
times. The relatively high success rate irrespective of the level of
refinement in the initial configuration indicates the effectiveness of
using conditional entropy to guide abstraction refinement in our
setting.

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
6-pages!
Moritz, can you add the new experimetns, and perhaps throw away some of the older ones.

@@
 
-\noindent We note that in our project the dual rewards are sourced
+\noindent Usually dual rewards are sourced
 from a motivational component of our robotic architecture. In this
@@
 %%
-Finally, for the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
+Finally, execution of either a dual action $\assumptiveDT{i}$, or the
 $\actions.\poss(\switchAction)$, returns control to a sequential
@@
 making it again.
+
+The process posed to a DT session is abstract. The abstraction is
+defined by first constructing an $(\pp{:init})$ declaration without
+any-active assumptions from the trace -- E.g.,
+(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
+probabilistic terms, and circles are atomic and/or conjunctive.
 %%
+For each excluded atom, we compute the {\em entropy} of the active
+assumptions, {\em conditional} on that atom. Intuitively, lower
+entropy indicates an atom gives better information about
+assumptions. Facts are iteratively added to the belief-state in
+increasing order according to that measure
+(Fig.\ref{fig:abstraction-c}).
 
-%%  We make no assumptions here about how they are
-%% derived.
-
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
+%% In the first step, we remove all facts that are not part of an
+%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+%% would proceed in an abstraction of the environment that does not
+%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
+%% %%
+%% In a second step, we iteratively refine the relaxed declaration by
+%% adding terms from the original statement of $(\pp{:init})$ while the
+%% number of abstract states in $\bstate_0$ that occur with non-zero
+%% probability according to that refined declaration remains of a
+%% practicable size. In detail,
 
@@
   \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief-state with the assumptions made by the continual
-    planner in bold.]{
-      \label{fig:abstraction-a}
-      \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode] (cup) {} 
-          child {node [cnode] (cupp1) {}}
-          child {node [cnode] (cupp2) {}}
-        }
-      }
-      child {node[cnode] (kitchen) {}
-        child {node [pnode] (box2) {} 
-          child {node [cnode] (box2p1) {}}
-          child {node [cnode] (box2p2) {}}
-        }
-      };
-   \tiny
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of kitchen] {$\pp{kitchen}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[below=0 of cup] {$\pp{(is-in cup)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-   \node[right=0 of cupp1] {$\pp{place1}$};
-   \node[right=0 of cupp2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
+  %% \subfloat[The initial belief-state with the assumptions made by the continual
+  %%   planner in bold.]{
+  %%     \label{fig:abstraction-a}
+  %%     \begin{tikzpicture}[
+  %%   level 3/.style={tree}]
+  %%   \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
+  %%     child[seclevel] {node[cnode, assumption] (office) {}
+  %%       child {node [pnode, assumption] (box) {}
+  %%         child {node [cnode] (boxp1) {}}
+  %%         child {node [cnode, assumption] (boxp2) {}}
+  %%       }
+  %%       child {node [pnode] (cup) {} 
+  %%         child {node [cnode] (cupp1) {}}
+  %%         child {node [cnode] (cupp2) {}}
+  %%       }
+  %%     }
+  %%     child {node[cnode] (kitchen) {}
+  %%       child {node [pnode] (box2) {} 
+  %%         child {node [cnode] (box2p1) {}}
+  %%         child {node [cnode] (box2p2) {}}
+  %%       }
+  %%     };
+  %%  \tiny
+  %%  \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
+  %%  \node[above=0 of cat] {$\pp{(category room1)}$};
+  %%  \node[below=0 of kitchen] {$\pp{kitchen}$};
+  %%  \node[below=0 of office] {$\pp{office}$};
+  %%  \node[below=0 of box] {$\pp{(is-in box)}$};
+  %%  \node[below=0 of box2] {$\pp{(is-in box)}$};
+  %%  \node[below=0 of cup] {$\pp{(is-in cup)}$};
+  %%  \node[right=0 of boxp1] {$\pp{place1}$};
+  %%  \node[right=0 of boxp2] {$\pp{place2}$};
+  %%  \node[right=0 of box2p1] {$\pp{place1}$};
+  %%  \node[right=0 of box2p2] {$\pp{place2}$};
+  %%  \node[right=0 of cupp1] {$\pp{place1}$};
+  %%  \node[right=0 of cupp2] {$\pp{place2}$};
+  %%   % \node[node] (kitchen) right of (cat) {};
+  %% \end{tikzpicture}}
 \qquad
@@
 \vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
+  \subfloat[Refinement, by adding relevant facts.]{
     \label{fig:abstraction-c}
@@
 
-  \caption{Generating the abstract belief space.}
+  \caption{Abstract belief-state, and refinement.}
 \label{fig:abstraction}
 \end{figure}
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
 

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
shrinking related work section

@@
 
-More generally, there has been much recent work on scaling POMDP
-solution procedures to medium-sized
-instances~\cite{shani:etal:08,KurHsu08}. In that
-setting \citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010})
-recently addressed an inefficiency of offline point-based techniques
-in problems with medium length planning horizons, however their
-approach take tens of minutes to good plans, and only scales to tens
-of thousands of states. In the case of general domain-independent
-factored systems, the state-of-the-art scales to relatively small
-problems with $2^{22}$ states~\cite{shani:etal:2008}. At their limit,
-these procedures take over an hour to converge, and $\sim10$ seconds
-on average to perform a single Bellman backup.  For classes of POMDP
-that feature exploitable structures (e.g., no actions with negative
-effects), problems with as many as $10^{30}$ states can be targeted by
-such offline procedures~\cite{brunskill:russell:2010}. Moving someway
-towards addressing real-time decision making, recent online POMDP
-solution procedures have been developed which can exploit highly
-approximate value functions -- typically computed using a point-based
+There has also been much recent work on scaling offline approximate
+POMDP solution procedures to medium-sized instances. Recent
+contributions propose more efficient belief-point sampling
+schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and structured
+representations with procedures that can efficiently exploit
+them~\cite{brunskill:russell:2010,shani:etal:2008}. Offline domain
+independent systems scale to {\em logistics} problems with $2^{22}$
+states~\cite{shani:etal:2008}, taking over an hour to converge, and
+around 10 seconds on average to perform each Bellman
+backup. \citeauthor{brunskill:russell:2010} are able to solve problems
+with as many as $10^{30}$ states, by further exploiting certain
+problem features -- E.g., no actions with negative effects.
+%%
+Moving someway towards supporting real-time decision making, recent
+online POMDP solution procedures have been developed which leverage
+highly approximate value functions -- computed using an offline
 procedure -- and heuristics in forward
-search~\cite{ross:etal:2008}. Their applicability in our setting is
-limited, firstly because they only scale to smaller problems, with
-thousands of states, and also due to the large amount of
-\emph{problem-specific} offline processing that might be required to get useful
-search guidance. A {\em very} recent and promising online approach for
+search~\cite{ross:etal:2008}. These approaches are applicable in
+relatively small problems, and can require
+expensive \emph{problem-specific} offline processing in order to yield
+good behaviours.
+%%
+A {\em very} recent and promising online approach for
 large POMDPs employs Monte-Carlo sampling to break the curse of
@@
 determined~\cite{silver:veness:2010}.
+
+%% Their applicability in our setting is
+%% limited, firstly because they only scale to smaller problems, with
+%% thousands of states, and also due to the large amount of
+%% \emph{problem-specific} offline processing that might be required to get useful
+%% search guidance.
+
+%%  In that
+%% setting \citeauthor{kurniawati:etal:2010}~(\citeyear{})
+%% recently addressed an inefficiency of offline point-based techniques
+%% in problems with medium length planning horizons, however their
+%% approach take tens of minutes to good plans, and only scales to tens
+%% of thousands of states. In the case of general domain-independent
+%% factored systems, the state-of-the-art scales to relatively small
+%% problems with $2^{22}$ states~\cite{shani:etal:2008}.
+
+%% At their limit,
+%% these procedures take over an hour to converge, and $\sim10$ seconds
+%% on average to perform a single Bellman backup.  For classes of POMDP
+%% that feature exploitable structures (e.g., no actions with negative
+%% effects), problems with as many as $10^{30}$ states can be targeted by
+%% such offline procedures~\cite{brunskill:russell:2010}.
 

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
new related text

@@
 
+%%  ---essentially the idea that
+%% it does not matter what the value of the variable you are trying to
+%% observe is---
+
 Addressing task and observation planning specifically, there have been
@@
 For vision algorithm
-selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit a
-natural hierarchical decomposition of the underlying POMDP into
-subproblems where decision-theoretic planning is reasonably
-fast. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
+selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit an
+explicitly modelled hierarchical decomposition of the underlying
+POMDP. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
 represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, those exploitable structures are not present in problems we
-have considered so far due to the task planning requirement. 
+advantage of symmetry in the belief space to exponentially shrink the
+state space. Although we have been actively exploring
+the \citeauthor{doshi08:pref_elic} approach, those exploitable
+symmetries are not present in problems we have considered so far due
+to the task planning requirement.
 %%
-%%
-Finally,
-our approach is in a similar vein to the more classical {\em
-dual-mode} control~\cite{cassandra96actingunder}, however in our case
-entropy reduction is targeted by planning in an abstract process which
-is informed by one execution trace ---computed by a {\em classical}
-planner--- and the underlying belief-state.
+Finally, our approach is in a similar vein to {\em dual-mode}
+control~\cite{cassandra96actingunder}, in our case we use a {\em
+classical} planner to provide guidance for reducing entropy.
+
+%% entropy reduction is targeted by planning in an abstract process which
+%% is informed by one execution trace ---computed by a {\em classical}
+%% planner--- and the underlying belief-state.
 

### 2011-02-07 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
shrinking

@@
 For vision algorithm
-selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit a
-natural hierarchical decomposition of the underlying POMDP into
-subproblems where decision-theoretic planning is reasonably
-fast. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
+selection,~\citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) exploit an
+explicitly modelled hierarchical decomposition of the underlying
+POMDP. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
 represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, those exploitable structures are not present in problems we
-have considered so far due to the task planning requirement. 
+advantage of symmetry in the belief-space to exponentially shrink the
+state-space. Although we have been actively exploring
+the \citeauthor{doshi08:pref_elic} approach, those exploitable
+symmetries are not present in problems we consider due to the task
+planning requirement.
 %%
+Finally, our approach is in a similar vein to {\em dual-mode}
+control~\cite{cassandra96actingunder}, in our case we use a {\em
+classical} planner to provide guidance for reducing entropy.
+
+%% entropy reduction is targeted by planning in an abstract process which
+%% is informed by one execution trace ---computed by a {\em classical}
+%% planner--- and the underlying belief-state.
+
+There has also been much recent work on scaling offline approximate
+POMDP solution procedures to medium-sized instances. Recent
+contributions propose more efficient belief-point sampling
+schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and structured
+representations with procedures that can efficiently exploit
+them~\cite{brunskill:russell:2010,shani:etal:2008}. Offline domain
+independent systems scale to {\em logistics} problems with $2^{22}$
+states~\cite{shani:etal:2008}, taking over an hour to converge, and
+around 10 seconds on average to perform each Bellman
+backup. \citeauthor{brunskill:russell:2010} are able to solve problems
+with as many as $10^{30}$ states, by further exploiting certain
+problem features -- E.g., no actions with negative effects.
 %%
-Finally,
-our approach is in a similar vein to the more classical {\em
-dual-mode} control~\cite{cassandra96actingunder}, however in our case
-entropy reduction is targeted by planning in an abstract process which
-is informed by one execution trace ---computed by a {\em classical}
-planner--- and the underlying belief-state.
-
-More generally, there has been much recent work on scaling POMDP
-solution procedures to medium-sized
-instances~\cite{shani:etal:08,KurHsu08}. In that
-setting \citeauthor{kurniawati:etal:2010}~(\citeyear{kurniawati:etal:2010})
-recently addressed an inefficiency of offline point-based techniques
-in problems with medium length planning horizons, however their
-approach take tens of minutes to good plans, and only scales to tens
-of thousands of states. In the case of general domain-independent
-factored systems, the state-of-the-art scales to relatively small
-problems with $2^{22}$ states~\cite{shani:etal:2008}. At their limit,
-these procedures take over an hour to converge, and $\sim10$ seconds
-on average to perform a single Bellman backup.  For classes of POMDP
-that feature exploitable structures (e.g., no actions with negative
-effects), problems with as many as $10^{30}$ states can be targeted by
-offline procedures~\cite{brunskill:russell:2010}. Moving someway
-towards addressing real-time decision making, recent online POMDP
-solution procedures have been developed which can exploit highly
-approximate value functions -- typically computed using a point-based
+Moving someway towards supporting real-time decision making, recent
+online POMDP solution procedures have been developed which leverage
+highly approximate value functions -- computed using an offline
 procedure -- and heuristics in forward
-search~\cite{ross:etal:2008}. Their applicability in our setting is
-limited, firstly because they only scale to smaller problems, with
-thousands of states, and also due to the large amount of
-\emph{problem-specific} offline processing that might be required to get useful
-search guidance. A {\em very} recent and promising online approach for
+search~\cite{ross:etal:2008}. These approaches are applicable in
+relatively small problems, and can require
+expensive \emph{problem-specific} offline processing in order to yield
+good behaviours.
+%%
+A {\em very} recent and promising online approach for
 large POMDPs employs Monte-Carlo sampling to break the curse of

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
more edits to the frontmatter

@@
 represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space to exponentially shrink the
-state space. Although we have been actively exploring
+advantage of symmetry in the belief-space to exponentially shrink the
+state-space. Although we have been actively exploring
 the \citeauthor{doshi08:pref_elic} approach, those exploitable
-symmetries are not present in problems we have considered so far due
-to the task planning requirement.
+symmetries are not present in problems we consider due to the task
+planning requirement.
 %%
 Finally, our approach is in a similar vein to {\em dual-mode}
-control~\cite{cassandra96actingunder}, in our case we use a {\em
-classical} planner to provide guidance for reducing entropy.
+control~\cite{cassandra96actingunder}, where planning switches between
+entropy and utility focuses.
+
+%% serial planning in
+%% problem determinisations underlines strategies to goal achievement,
+%% and
+
+%% in our case we use serial
+%% planning in a determinisation of the underlying process to guide DT
+%% sessions, which typically act to reducing entropy.
 
@@
 with as many as $10^{30}$ states, by further exploiting certain
-problem features -- E.g., no actions with negative effects.
+problem features -- E.g., problems where no actions with negative
+effects.
 %%
@@
 determined~\cite{silver:veness:2010}.
+
+In the direction of leveraging {\em classical} systems/approaches for
+planning under uncertainty, the most highlighted system to date has
+been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
+the probabilistic track of the 2004 International Planning
+Competition.  In the continual paradigm, \system{FFR$_a$}
+uses \system{FF} to compute sequential plans and execution traces.
+%%
+More computationally expensive approaches in this vein combine
+sampling strategies on valuations over {\em runtime variables} with
+deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
+%% robust sequential plan, or contingent
+%% plan~\cite{majercik:2006}.
+
+Also leveraging deterministic planners in problems that feature
+uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
+---i.e., sequential planning in unobservable worlds--- can be modelled
+as a deterministic problem, and therefore solved using sequential
+systems. In this conformant setting, advances have been towards
+compact representations of beliefs amenable to existing best-first
+search planning procedures, and lazy evaluations of beliefs. Most
+recently this research thread has been extended to contingent planning
+in fully observable non-deterministic
+environments~\cite{albore:etal:2009}.
 
@@
 
-In the direction of leveraging {\em classical} systems/approaches for
-planning under uncertainty, the most highlighted system to date has
-been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
-the probabilistic track of the 2004 International Planning
-Competition.  In the continual paradigm, \system{FFR$_a$}
-uses \system{FF} to compute sequential plans and execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
-%% robust sequential plan, or contingent
-%% plan~\cite{majercik:2006}.
-
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
 %%

### 2011-02-07 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
attached

@@
 Finally, our approach is in a similar vein to {\em dual-mode}
-control~\cite{cassandra96actingunder}, in our case we use a {\em
-classical} planner to provide guidance for reducing entropy.
+control~\cite{cassandra96actingunder}, where planning switches between
+entropy and utility focuses.
+
+%% serial planning in
+%% problem determinisations underlines strategies to goal achievement,
+%% and
+
+%% in our case we use serial
+%% planning in a determinisation of the underlying process to guide DT
+%% sessions, which typically act to reducing entropy.
 
@@
 contributions propose more efficient belief-point sampling
-schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and structured
+schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and factored
 representations with procedures that can efficiently exploit
-them~\cite{brunskill:russell:2010,shani:etal:2008}. Offline domain
-independent systems scale to {\em logistics} problems with $2^{22}$
-states~\cite{shani:etal:2008}, taking over an hour to converge, and
-around 10 seconds on average to perform each Bellman
+structures in those
+representations~\cite{brunskill:russell:2010,shani:etal:2008}. Offline
+domain independent systems scale to {\em logistics} problems with
+$2^{22}$ states~\cite{shani:etal:2008}, taking over an hour to
+converge, and around 10 seconds on average to perform each Bellman
 backup. \citeauthor{brunskill:russell:2010} are able to solve problems
-with as many as $10^{30}$ states, by further exploiting certain
-problem features -- E.g., no actions with negative effects.
+with approximately $10^{30}$ states, by further exploiting certain
+problem features -- E.g., problems where no actions with negative
+effects.
 %%
@@
 A {\em very} recent and promising online approach for
-large POMDPs employs Monte-Carlo sampling to break the curse of
+larger POMDPs employs Monte-Carlo sampling to break the curse of
 dimensionality in situations where goal reachability is {\em easily}
 determined~\cite{silver:veness:2010}.
+
+In the direction of leveraging {\em classical} systems/approaches for
+planning under uncertainty, the most highlighted system to date has
+been \system{FFR$_a$}~\cite{yoon:etal:2007}; The winning entry from
+the probabilistic track of the 2004 International Planning
+Competition.  In the continual paradigm, \system{FFR$_a$}
+uses \system{FF} to compute sequential plans and execution traces.
+%%
+More computationally expensive approaches in this vein combine
+sampling strategies on valuations over {\em runtime variables} with
+deterministic planning procedures~\cite{yoon:etal:2008}. %% The outcome is typically a more
+%% robust sequential plan, or contingent
+%% plan~\cite{majercik:2006}.
+
+Also leveraging deterministic planners in problems that feature
+uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
+$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
+---i.e., sequential planning in unobservable worlds--- can be modelled
+as a deterministic problem, and therefore solved using sequential
+systems. In this conformant setting, advances have been towards
+compact representations of beliefs amenable to existing best-first
+search planning procedures, and lazy evaluations of beliefs. Most
+recently this research thread has been extended to contingent planning
+in fully observable non-deterministic
+environments~\cite{albore:etal:2009}.
 

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
corrections

@@
 %%
-Structurally, it is a sequence of elements that are either: (i) ground
+Structurally, a trace is a sequence of elements that are either: (i) ground
 actions from the DTPDDL description of the world, or (ii) atomic {\em
@@
 abstraction is constructed by first excluding all propositions (and
-related actions) that are not true of any state in the trace, then
-adding them back, using as a heuristic the entropy of the trace
-assumptions conditional on a candidate proposition. Propositions are
-added, one at a time, until the number of states in the initial
-belief-state reaches a given threshold.\footnote{Our experiments are
-performed using a number of different threshold parameters.}  To the
-resulting abstract model we also add {\em disconfirm} and {\em
-confirm} actions that the contingent session can schedule in order to
-judge an atomic assumption in the trace. In the abstract model these
-actions yield a relatively small reward if the corresponding judgement
-is true (or small penalty otherwise). If a judgement action is
-scheduled for execution the contingent session is terminated, and
-control is returned to a sequential session.
+related actions) that are not true, or {\em assumed } true, of states
+in the trace, then adding them back, using as a heuristic the entropy
+of the trace assumptions conditional on a candidate
+proposition. Propositions are added, one at a time, until the number
+of states in the initial belief-state reaches a given
+threshold.\footnote{Our experiments are performed using a number of
+different threshold parameters.}  To the resulting abstract model we
+also add {\em disconfirm} and {\em confirm} actions that the
+DT session can schedule in order to judge an atomic assumption
+in the trace. In the abstract model these actions yield a relatively
+small reward if the corresponding judgement is true (or small penalty
+otherwise). If a judgement action is scheduled for execution the
+DT session is terminated, and control is returned to a
+sequential session.
 
-Finally, whether proceeding in a sequential or contingent session, our
+Finally, whether proceeding in a sequential or DT session, our
 continual planner maintains a factored representation of successive
@@
 session, and (3) as a mechanism to guide construction of an abstract
-process for contingent sessions.
+process for DT sessions.
 
@@
 
+\small
 \[
 \begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
+\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
 \;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
@@
 \]
+\normalsize
 
@@
 
-%% %%
-%% Lastly, we have that if $\action$ has a precondition that is not true
-%% with probability $\theta$, a given threshold, in the underlying belief
-%% $\bstate$, then immediately before it is executed blindly the system
-%% switches to a contingent session (described below).
-
 We now describe the optimisation criteria used during sequential
 sessions. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
+sequenced action, $\action_i$, from a trace
+$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
+does not transition to $\state^\bot$, we have that the optimal trace
+has value:
 
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
+%% define the value of a trace
+%% $\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
+%% deterministic model to equal:
 
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
+%% \small
+%% \begin{equation}\label{eq:tracevalue}
+%% V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
+%% \end{equation}
+%% \normalsize
 
+%% \noindent The
+%% optimal trace given a plan, i.e., the sequence of non-assumptive
+%% actions from the trace, therefore has value:
+
+\small
 \[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
+V^* = \max_N \max_{\state_0, \action_0,.., \state_N} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
 \reward(\state_i, \action_i),
 \]
+\normalsize
 
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
-
-%% From the perspective of
-%% gradient-based reinforcement learning systems, such as William's
-%% REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-%% under reasonable assumptions the optimal trace identifies a maximal
-%% gradient step from a uniformly random soft-max policy.
+\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
+reward ---i.e., can be negative for action costs, or positive for goal
+achievement--- received for executing action $\action_i$ in state
+$\state_i$. Finally, it is worth clarifying that in practise we do not
+artificially limit the length of traces that the sequential planner
+can consider. Moreover, in our setting there is always a finite
+optimal trace.
 
@@
 
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
+The DT session plans in a {\em small} abstract processes defined
+according to the assumptive actions in the proceeding trace, and the
+$(\pp{:init})$ declaration that models the underlying
+belief-state. Targeted sensing is encouraged by augmenting the reward
+model to reflect the value of knowing the truth about assumptions in
+the trace proposed by the preceding sequential deliberation. In
+detail, all rewards from the original POMDP are
+retained. Additionally, for each $\assumptiveS{i}$ action scheduled by
+the current trace, we have a {\em dual} $\assumptiveDT{i}$ so that
+forall states $\state$:
 
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
+\small
 \[
@@
 
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
+\noindent Usually dual rewards are sourced
+from a motivational component of our robotic architecture. In this
+paper, for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a positive
+constant, and have $\hat\$(x)= - \$(x)(1 - \prob) / \prob$ where
+$\prob$ is the probability that $x$ is true. For
+$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
+\$(x)\prob/(1-\prob)$
+%%
+Finally, execution of either a dual action $\assumptiveDT{i}$, or the
 $\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
+session, which then continues from the underlying belief-state. If an
+assumption was rejected, we prohibit that sequential session from
+making it again.
 
-%%  We make no assumptions here about how they are
-%% derived.
+The process posed to a DT session is abstract. The abstraction is
+defined by first constructing an $(\pp{:init})$ declaration without
+any-active assumptions from the trace -- E.g.,
+(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
+probabilistic terms, and circles are atomic and/or conjunctive.
+%%
+For each excluded atom, we compute the {\em entropy} of the active
+assumptions, {\em conditional} on that atom. Intuitively, lower
+entropy indicates an atom gives better information about
+assumptions. Facts are iteratively added to the belief-state in
+increasing order according to that measure
+(Fig.\ref{fig:abstraction-c}).
 
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
+%% In the first step, we remove all facts that are not part of an
+%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+%% would proceed in an abstraction of the environment that does not
+%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
+%% %%
+%% In a second step, we iteratively refine the relaxed declaration by
+%% adding terms from the original statement of $(\pp{:init})$ while the
+%% number of abstract states in $\bstate_0$ that occur with non-zero
+%% probability according to that refined declaration remains of a
+%% practicable size. In detail,
 
@@
 
-  \caption{Generating the abstract belief space.}
+  \caption{Abstract belief-state, and refinement.}
 \label{fig:abstraction}
 \end{figure}
-
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
 

### 2011-02-07 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
6-pages!
Moritz, can you add the new experimetns, and perhaps throw away some of the older ones.
Also made the correction in the concluding remarks

@@
+\vspace{-1ex}
+
 For use in an integrated continual planning and execution monitoring
@@
 executed until testing of an assumption about the runtime state is
-scheduled. At that point a contingent session performs some sensing
+scheduled. At that point a DT session performs some sensing
 which determines whether or not the sequential session should be
@@
 objectives. We find our approach performs quickly and relatively
-efficiently in large real-world task and observation planning
-problems.
+efficiently in a number of large real-world task and observation
+planning problems.
 
@@
 so that conformant (or semi-conformant) plans can be executed without
-interruption by a contingent session. A more general criticism of
+interruption by a DT session. A more general criticism of
 switching continual planning concerns interleaved sequential and
-contingent sessions failing to make any progress towards the
+DT sessions failing to make any progress towards the
 objectives. For example, in the worst case we can have each sequential

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
corrections

--- 2011-01-28 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
@@
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
+\vspace{-1ex}
 
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
+For use in an integrated continual planning and execution monitoring
+system on a mobile robot platform, we developed a base planner that
+switches between fast sequential and expensive DT planning. Given a
+DTPDDL model of the environment, sequential planning is used to
+compute an initial deterministic sequential plan and complementary
+runtime evolution of the decision process. That plan is executed until
+testing of an assumption about the runtime state is scheduled. At that
+point a DT session performs some sensing which determines whether or
+not the sequential session should be allowed to continue, or otherwise
+acts to achieve the overall objectives. We find our approach performs
+quickly and relatively efficiently in a number of large real-world
+task and observation planning problems.
 
+\Omit{
 The most pressing item for future research, is to develop a scheme

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
spellers

@@
 
-We consider only deterministic-action POMDs, where all the uncertainty
-in state and action is expressed in the $(\pp{:init})$ declaration. In
-sequential sessions, this declaration is used by our approach to
-define a starting state, and the set of state-assumptions available to
-sequential planning.  Writing \#\ if the value of a proposition is
-unspecified, continuing our example from the previous section, we have
-the following assumptions:
+We consider only deterministic-action POMDPs, where all state
+uncertainty is expressed in the $(\pp{:init})$ declaration. This
+declaration is used by our approach to define a starting state, and
+the set of state-assumptions available to sequential planning.
+Writing \#\ if the value of a proposition is unspecified, continuing
+our example from the previous section, we have the following
+assumptions:
 
@@
 are visited. For a sequential session, the starting state corresponds
-to the unique assumption with probability $1$ -- I.e., the union of
-facts from the root conjunctive term. In our example, the starting
-state is:
+to an assumption (usually unique) with probability $1$. In our
+example, the starting state is:
 

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
not sure if its getting shorter, but it's not getting longer

@@
 
-We consider only deterministic-action POMDPs, where all state
+We only consider deterministic-action POMDPs, where all state
 uncertainty is expressed in the $(\pp{:init})$ declaration. This
-declaration is used by our approach to define a starting state, and
-the set of state-assumptions available to sequential planning.
-Writing \#\ if the value of a proposition is unspecified, continuing
-our example from the previous section, we have the following
-assumptions:
+declaration is used by our approach to define the starting state for
+sequential sessions, and the set of state-assumptions available to
+sequential planning.  Writing \#\ if the value of a proposition is
+unspecified, taking the $(\pp{:init})$ example from the previous
+section, we have the following assumptions:
 
@@
 \noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or one of its immediate probabilistic subterms
-are visited. For a sequential session, the starting state corresponds
-to an assumption (usually unique) with probability $1$. In our
-example, the starting state is:
+relaxed} visitation of the root. Here, a conjunctive term is visited
+iff its atomic subterms are visited, and zero or one of its immediate
+probabilistic subterms are visited. For a sequential session, the
+starting state corresponds to an assumption (usually unique) with
+probability $1$. In our example, that starting state is:
 
@@
 
-We also annotate DTPDDL model with an assumptive action
-$\assumptiveS{i}$ for each element $\prob_i (T_i)$, of each
-probabilistic term from $(\pp{:init})$. Here, $\assumptiveS{i}$ can be
-executed if no $\assumptiveS{j}$, $j \neq i$, has been executed from
-the same probabilistic term, and, either
+Encapsulating state-assumptions, we annotate the problem posed during
+a sequential session with an assumptive action $\assumptiveS{i}$ for
+each element $\prob_i (T_i)$, of each probabilistic term from
+$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
+$\assumptiveS{j}$, $j \neq i$, has been executed from the same
+probabilistic term, and, either
 $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
@@
 artificially limit the length of traces that the sequential planner
-can consider. Moreover, in our setting there is always a finite
-optimal trace.
+can consider. Moreover, for problems we consider there is always a
+finite optimal trace.
 
@@
 belief-state. Targeted sensing is encouraged by augmenting the reward
-model to reflect the value of knowing the truth about assumptions in
-the trace proposed by the preceding sequential deliberation. In
-detail, all rewards from the original POMDP are
-retained. Additionally, for each $\assumptiveS{i}$ action scheduled by
-the current trace, we have a {\em dual} $\assumptiveDT{i}$ so that
-forall states $\state$:
+model to reflect a heuristic value of knowing the truth about
+assumptions in the trace. In detail, all rewards from the original
+POMDP are retained. Additionally, for each $\assumptiveS{i}$ action
+scheduled by the current trace, we have a {\em dual}
+$\assumptiveDT{i}$ so that forall states $\state$:
 
@@
 utility the agent receives for correctly (incorrectly) rejecting an
-assumption. Dual action execution is modelled as a self-transition --
-I.e., the process remains in the state from which the dual is
+assumption. A dual action can only be executed once, and corresponds
+otherwise to a self-transition -- I.e., except in the applicability of
+the dual, the process remains in the state from which a dual is
 executed.
@@
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
+active, and therefore exclude it from the
 POMDP posed to the DT session.

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
shorter I think

@@
 to \switchAction. We only include the duals of active assumptions in
-DT sessions. For example, if the trace is:
+DT sessions. Continuing our simplified example, if the trace is:
 
@@
 
-\noindent Usually dual rewards are sourced
-from a motivational component of our robotic architecture. In this
-paper, for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a positive
+Execution of either a dual action $\assumptiveDT{i}$, or the
+$\actions.\poss(\switchAction)$, returns control to a sequential
+session, which then continues from the underlying
+belief-state. Turning to the detail of dual rewards, in our integrated
+system these are sourced from a motivational subsystem. In this paper,
+for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a small positive
 constant, and have $\hat\$(x)= - \$(x)(1 - \prob) / \prob$ where
@@
 \$(x)\prob/(1-\prob)$
+
+\Omit{
 %%
-Finally, execution of either a dual action $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session, which then continues from the underlying belief-state. If an
+Finally, 
+If an
 assumption was rejected, we prohibit that sequential session from
 making it again.
+}
 

### 2011-02-07 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
more change

@@
 For use in an integrated continual planning and execution monitoring
-system on a mobile robot platform, we developed a base planner that
-switches between fast sequential and expensive decision-theoretic
-planning. Given a DTPDDL model of the environment, sequential planning
-is used to compute an initial deterministic sequential plan and
-complementary runtime evolution of the decision process. That plan is
-executed until testing of an assumption about the runtime state is
-scheduled. At that point a DT session performs some sensing
-which determines whether or not the sequential session should be
-allowed to continue, or otherwise acts to achieve the overall
-objectives. We find our approach performs quickly and relatively
-efficiently in a number of large real-world task and observation
-planning problems.
+system on a mobile robot platform, we developed a planner that
+switches between fast sequential and expensive DT planning. Given a
+DTPDDL model of the environment, sequential planning is used to
+compute an initial deterministic sequential plan and complementary
+runtime evolution of the decision process. That plan is executed until
+testing of an assumption about the runtime state is scheduled. At that
+point a DT session performs some sensing which determines whether or
+not the sequential session should be allowed to continue, or otherwise
+acts to achieve the overall objectives. We find our approach performs
+quickly and relatively efficiently in a number of large real-world
+task and observation planning problems.
 

### 2011-02-07 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
exhibit was the wrong word

@@
 events, changing objectives, and the underlying unpredictability of
-the environment.  On the other hand, in order to exhibit efficient
+the environment.  On the other hand, in order to promote efficient
 behaviours the planning process must perform computationally expensive

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
New experimental data (not yet complete)

@@
 revision. Sequential sessions use a modified version of Fast
-Downward~\cite{fast-downward}, and DT sessions use our own fast
-contingent procedure. We also implemented a simple dual-mode
-replanning {\em baseline} approach in MAPSIM. Here, a DT session
-executes a single greedy entropy reduction action, wrt to active
-assumptions, when a switching action is scheduled for execution,
-returning control to a new sequential session.
+Downward~\cite{fast-downward}, and DT sessions use our own contingent
+procedure. We also implemented a simple dual-mode replanning {\em
+baseline} approach. Here, when a switching action is scheduled for
+execution the DT session executes a single entropy reduction action,
+that directly senses an active assumption -- I.e,. We say a fact can
+be directly sensed, if there is a DTPDDL {\em sense} declaration that
+mentions the fact in its precondition.  Control is then immediately
+returned to a new sequential session.
 
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
+We evaluate our approaches in robot exploration tasks from home and
+office environments. Spatially, these consists of {\em rooms}
+(office/kitchen/etc), and an underlying topological map that models
+smaller areas of space, called {\em places}, and connectivity between
+those. The mobile {\em robot} and {\em visual objects} inhabit the
+topological space, the latter indicate the category of space they
+inhabit -- E.g., Spoons are likely to be in kitchens. By examining
+view cones at places for particular objects, the robot is able to: (1)
+categorise space at high (room) and low (place) levels, and (2) find
+objects for the user, exploiting information about object
+co-occurrence and room categories for efficiency. Also, in the
+presence of a person, the robot can ask them about the location of
+objects.
 
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
+We compare our switching planner against the {\em baseline} in several
+realistic tasks, with the number of rooms ranging from 3 (12-places,
+16-objects, $|$states$|$$>\sim10^{21}$) to 6 (26-places, 21-objects,
+$|$states$|$ $>10^{36}$). For or experiments we model three sensors:
+{\em Reliable} sensors have a $0.1$ probability of a false negative,
+{\em semi-reliable} have a chance of $0.3$ of false negative and $0.1$
+of false positive , and finally {\em noisy} sensors with probabilities
+of $0.5$ and $0.2$ respectively. Each object class is assigned one
+sensor model, so e.g. cornflakes may be harder to detect than
+refrigeratos. To determine the influence of the sensor model, we
+performed several experiments where we only changed the difficulty of
+the target object(s). All non-target objects remain unchanged so the
+planner may exploit object co-occurances with reliable objects to
+compensate for the increased sensor noise.
 
-We evaluated our switching planner against the {\em baseline} in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the {\em baseline} cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
-  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
-  core.}. In our domain, rewards encapsulate a goal condition, and
-therefore we find it intuitive to report average costs of the plans
-and the success rate in problems that admit a solution (i.e., positive
-reward scaled by a constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
+We evaluate with DT sessions taking $b_0$ admitting between 20 and 200
+abstract states with non-zero probability. We run 50 simulations in
+each configuration, and have a timeout on each simulation of 30
+minutes (1800 seconds)\footnote{All experiments were conducted on a
+  2.3GHz AMD Opteron using one CPU core.}. The continual planning
+times are reported in Figure~\ref{fig:results-time}, and the quality
+data in Figure~\ref{fig:results-quality}.
+%%
+For Items~$a$-$e$ the goal is to find objects and report their
+position to a user. In some cases there is a non-zero probability that
+no plan exists, as the desired object might not be present in the
+environment. Item~$f$ reports for tasks requiring indirect sensing,
+where the robot must relocate to a room with a particular category
+(e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
+so is omitted from that reporting.  Because reward is only allocated
+once, on the achievement of the goal,  we find it
+intuitive to report average plan costs and the success rates in
+problems that admit a solution (i.e., positive reward scaled by a
+constant factor).
 
@@
   \vspace{2mm}
-  \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
+  % \includegraphics{dora56-time}\hfill
+  % \vspace{2mm}
+  % \includegraphics{dora-cat-time}\hfill
   \caption{Average runtime}
@@
   % \vspace{2mm}
+  \includegraphics{pomdp-quality}\hfill
   \includegraphics{dora2-quality}\hfill
@@
   \vspace{2mm}
-  \includegraphics{dora56-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
+  % \includegraphics{dora56-quality}\hfill
+  % \vspace{2mm}
+  % \includegraphics{dora-cat-quality}\hfill
   \caption{Average plan costs and number of successful runs.}
@@
 
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the {\em baseline} is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
+We find that if sensing is reliable there is little to be gained
+through DT sessions, as the greedy approach of the {\em baseline} is
+sufficient. As sensing degrades contingent planning pays off.  Time
+spent in DT planning increases steeply as the abstraction $\bstate_0$
+becomes more refined.  That refinement seems to be paying off in terms
+of the success rate, particularly for tasks $d$ and $e$. For less
+refined initial configurations, the increase cost of DT
+planning is compensated by a decrease in Fast Downward planning
+times. The relatively high success rate irrespective of the level of
+refinement in the initial configuration indicates the effectiveness of
+using conditional entropy to guide abstraction refinement in our
+setting.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "aaai11"
+%%% End:

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
spellers

@@
 execution the DT session executes a single entropy reduction action,
-that directly senses an active assumption -- I.e,. We say a fact can
+that directly senses an active assumption -- i.e,. We say a fact can
 be directly sensed, if there is a DTPDDL {\em sense} declaration that
@@
 We evaluate our approaches in robot exploration tasks from home and
-office environments. Spatially, these consists of {\em rooms}
+office environments. Spatially, these consist of {\em rooms}
 (office/kitchen/etc), and an underlying topological map that models
@@
 topological space, the latter indicate the category of space they
-inhabit -- E.g., Spoons are likely to be in kitchens. By examining
+inhabit -- e.g., Spoons are likely to be in kitchens. By examining
 view cones at places for particular objects, the robot is able to: (1)
@@
 
-We compare our switching planner against the {\em baseline} in several
+We compare our switching planner with the {\em baseline} in several
 realistic tasks, with the number of rooms ranging from 3 (12-places,
-13-objects, $|$states$|$$>\sim10^{48}$) to 6 (26-places, 21-objects,
-$|$states$|$ $>10^{48}$). We experiment with a sensing model that is:
-({\em reliable}) $0.1$ probability of false negative, ({\em
-semi-reliable}) chance is $0.35$, and ({\em noisy}) chance is $0.4$.
+16-objects, $|$states$|$$>\sim10^{21}$) to 6 (26-places, 21-objects,
+$|$states$|$ $>10^{36}$). For or experiments we model three sensors:
+{\em Reliable} sensors have a $0.1$ probability of a false negative,
+{\em semi-reliable} have a chance of $0.3$ of false negative and $0.1$
+of false positive , and finally {\em noisy} sensors with probabilities
+of $0.5$ and $0.2$ respectively. Each object class is assigned one
+sensor model, so e.g. cornflakes may be harder to detect than
+refrigeratos. To determine the influence of the sensor model, we
+performed several experiments where we only changed the difficulty of
+the target object(s). All non-target objects remain unchanged so the
+planner may exploit object co-occurances with reliable objects to
+compensate for the increased sensor noise.
+
 We evaluate with DT sessions taking $b_0$ admitting between 20 and 200
@@
 minutes (1800 seconds)\footnote{All experiments were conducted on a
-2.3GHz AMD Opteron using one CPU core.}. The continual planning times
-are reported in Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}.
+  2.3GHz AMD Opteron using one CPU core.}. The continual planning
+times are reported in Figure~\ref{fig:results-time}, and the quality
+data in Figure~\ref{fig:results-quality}.
 %%
@@
 of the success rate, particularly for tasks $d$ and $e$. For less
-refined initial configurations, the increase cost of DT
-planning is compensated by a decrease in Fast Downward planning
+refined initial configurations, the increased cost of DT
+planning is compensated for by a decrease in Fast Downward planning
 times. The relatively high success rate irrespective of the level of
@@
 setting.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "aaai11"
+%%% End:

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
spellers

@@
 related actions) that are not true, or {\em assumed } true, of states
-in the trace, then adding them back, using as a heuristic the entropy
+in the trace, then adding them back, using as an heuristic the entropy
 of the trace assumptions conditional on a candidate
@@
 scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that forall states $\state$:
+$\assumptiveDT{i}$ so that for all states $\state$:
 
@@
 assumption. A dual action can only be executed once, and corresponds
-otherwise to a self-transition -- I.e., except in the applicability of
+otherwise to a self-transition -- i.e., except in the applicability of
 the dual, the process remains in the state from which a dual is
@@
 defined by first constructing an $(\pp{:init})$ declaration without
-any-active assumptions from the trace -- E.g.,
+any-active assumptions from the trace -- e.g.,
 (Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
New experimental data

@@
 
-We compare our switching planner against the {\em baseline} in several
+We compare our switching planner with the {\em baseline} in several
 realistic tasks, with the number of rooms ranging from 3 (12-places,
@@
   \vspace{2mm}
-  % \includegraphics{dora56-time}\hfill
+  \includegraphics{dora56-time}\hfill
   % \vspace{2mm}
@@
   \includegraphics{pomdp-quality}\hfill
+  \includegraphics{pomdp-solvable-quality}\hfill
   \includegraphics{dora2-quality}\hfill
@@
   \vspace{2mm}
-  % \includegraphics{dora56-quality}\hfill
+  \includegraphics{dora56-quality}\hfill
   % \vspace{2mm}

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
changed order, etc

@@
 
-\section{Background and Related Work}
-\input{relatex.tex}
-
-\section{Decision-Theoretic PDDL}
+\section{Planning Language and Notations}
 \input{dtpddl.tex}
@@
 \input{evaluation.tex}
+
+\section{Related Work}
+\input{relatex.tex}
 

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
changed order, etc

@@
 contributions propose more efficient belief-point sampling
-schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and structured
+schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and factored
 representations with procedures that can efficiently exploit
-them~\cite{brunskill:russell:2010,shani:etal:2008}. Offline domain
-independent systems scale to {\em logistics} problems with $2^{22}$
-states~\cite{shani:etal:2008}, taking over an hour to converge, and
-around 10 seconds on average to perform each Bellman
+structures in those
+representations~\cite{brunskill:russell:2010,shani:etal:2008}. Offline
+domain independent systems scale to {\em logistics} problems with
+$2^{22}$ states~\cite{shani:etal:2008}, taking over an hour to
+converge, and around 10 seconds on average to perform each Bellman
 backup. \citeauthor{brunskill:russell:2010} are able to solve problems
-with as many as $10^{30}$ states, by further exploiting certain
-problem features -- E.g., problems where no actions with negative
+with approximately $10^{30}$ states, by further exploiting certain
+problem features -- E.g., problems where no actions have negative
 effects.
@@
 A {\em very} recent and promising online approach for
-large POMDPs employs Monte-Carlo sampling to break the curse of
+larger POMDPs employs Monte-Carlo sampling to break the curse of
 dimensionality in situations where goal reachability is {\em easily}

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
small changes

@@
 satisficing ``classical'' planner, to decide on the overall strategy,
-and decision-theoretic planning to solve small subproblems where
+and decision-theoretic planning to solve small abstract subproblems where
 deeper consideration of the sensing model is both practical, and can

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
small changes

@@
 For use in an integrated continual planning and execution monitoring
-system on a mobile robot platform, we developed a base planner that
-switches between fast sequential and expensive DT planning. Given a
-DTPDDL model of the environment, sequential planning is used to
-compute an initial deterministic sequential plan and complementary
-runtime evolution of the decision process. That plan is executed until
-testing of an assumption about the runtime state is scheduled. At that
-point a DT session performs some sensing which determines whether or
-not the sequential session should be allowed to continue, or otherwise
-acts to achieve the overall objectives. We find our approach performs
-quickly and relatively efficiently in a number of large real-world
-task and observation planning problems.
+system on a mobile robot platform, we developed a planner that
+switches between fast sequential and expensive DT planning in small
+abstractions of the problem at hand. Given a DTPDDL model of the
+environment, sequential planning is used to compute an initial
+deterministic sequential plan and complementary runtime evolution of
+the decision process. That plan is executed until testing of an
+assumption about the runtime state is scheduled. At that point a DT
+session performs some sensing which determines whether or not the
+sequential session should be allowed to continue, or otherwise acts to
+achieve the overall objectives. We find our approach performs quickly
+and relatively efficiently in a number of large real-world task and
+observation planning problems.
 

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
small changes

@@
 with approximately $10^{30}$ states, by further exploiting certain
-problem features -- E.g., problems where no actions with negative
+problem features -- E.g., problems where no actions have negative
 effects.
@@
 determined~\cite{silver:veness:2010}. 
+%%
+Our approach can also be thought of as an online POMDP solver that
+uses a sequential plan to guide the search, rather than (e.g.,
+Monte-Carlo) sampling. Also, compared to most online POMDP procedures, which
+replan at each step, our approach involves relatively little replanning.
+
+%% Our approach requires minimal DT replanning,
+%% preferring
+
+%% Also, our approach will execute a few
+%% actions (as much as it can before an assumption is found to be true or
+%% false) before needing to plan again, rather than planning after each
+%% step.
+
+%% Moving someway towards supporting real-time decision making, recent
+%% online POMDP solution procedures have been developed which leverage
+%% highly approximate value functions -- computed using an offline
+%% procedure -- and heuristics in forward
+%% search~\cite{ross:etal:2008}. These approaches are applicable in
+%% larger problems (\cite{ross:etal:2008} includes experiments in a domain
+%% with around 12 thousand states), but the fact they often require
+%% expensive \emph{problem-specific} offline processing in order to yield
+%% good behaviours means they are impractical in problems as large as ours.
+%% %%
+%% A recent and promising online approach for
+%% larger POMDPs employs Monte-Carlo sampling to break the curse of
+%% dimensionality in situations where goal reachability is {\em easily}
+%% determined~\cite{silver:veness:2010}. Our approach can also be thought of
+%% as an online POMDP solver, but uses the sequential plan to guide the
+%% search
+%% for a decision-theoretic plan, rather than Monte-Carlo sampling. In
+%% addition, our approach will execute a few actions (as much as it can
+%% before
+%% an assumption is found to be true or false) before needing to plan again,
+%% rather than planning after each step.
+
+%% [have a look at the bit in parentheses above, I'm not sure I quite
+%% characterised how much plan it executes before replanning satisfactorally]
 

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
small changes

@@
 according to the continual planning paradigm, given descriptions of
-the current problem and domain in DTPDDL. Recapping, the system {\em
+the current problem and domain in DTPDDL. The system {\em
 switches}, in the sense that planning proceeds in interleaved
@@
 sessions. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action, $\action_i$, from a trace
+sequenced action, $\action_i$, from a trace of state-action pairs
 $\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
english

@@
 We now describe our {\em switching} planning system that operates
-according to the continual planning paradigm, given descriptions of
-the current problem and domain in DTPDDL. Recapping, the system {\em
+according to the continual planning paradigm, given a description of
+the current problem and domain in DTPDDL. The system {\em
 switches}, in the sense that planning proceeds in interleaved
@@
 abstraction is constructed by first excluding all propositions (and
-related actions) that are not true of any state in the trace, then
-adding them back, using as a heuristic the entropy of the trace
-assumptions conditional on a candidate proposition. Propositions are
-added, one at a time, until the number of states in the initial
-belief-state reaches a given threshold.\footnote{Our experiments are
-performed using a number of different threshold parameters.}  To the
-resulting abstract model we also add {\em disconfirm} and {\em
-confirm} actions that the contingent session can schedule in order to
-judge an atomic assumption in the trace. In the abstract model these
-actions yield a relatively small reward if the corresponding judgement
-is true (or small penalty otherwise). If a judgement action is
-scheduled for execution the contingent session is terminated, and
-control is returned to a sequential session.
+related actions) that are not true, or {\em assumed } true, of states
+in the trace, then adding them back, using as an heuristic the entropy
+of the trace assumptions conditional on a candidate
+proposition. Propositions are added, one at a time, until the number
+of states in the initial belief-state reaches a given
+threshold.\footnote{Our experiments are performed using a number of
+different threshold parameters.}  To the resulting abstract model we
+also add {\em disconfirm} and {\em confirm} actions that the
+DT session can schedule in order to judge an atomic assumption
+in the trace. In the abstract model these actions yield a relatively
+small reward if the corresponding judgement is true (or small penalty
+otherwise). If a judgement action is scheduled for execution the
+DT session is terminated, and control is returned to a
+sequential session.
 
-Finally, whether proceeding in a sequential or contingent session, our
+Finally, whether proceeding in a sequential or DT session, our
 continual planner maintains a factored representation of successive
@@
 
-We consider only deterministic-action POMDs, where all the uncertainty
-in state and action is expressed in the $(\pp{:init})$ declaration. In
-sequential sessions, this declaration is used by our approach to
-define a starting state, and the set of state-assumptions available to
+We only consider deterministic-action POMDPs, where all state
+uncertainty is expressed in the $(\pp{:init})$ declaration. This
+declaration is used by our approach to define the starting state for
+sequential sessions, and the set of state-assumptions available to
 sequential planning.  Writing \#\ if the value of a proposition is
-unspecified, continuing our example from the previous section, we have
-the following assumptions:
+unspecified, taking the $(\pp{:init})$ example from the previous
+section, we have the following assumptions:
 
@@
 \noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or one of its immediate probabilistic subterms
-are visited. For a sequential session, the starting state corresponds
-to the unique assumption with probability $1$ -- I.e., the union of
-facts from the root conjunctive term. In our example, the starting
-state is:
+relaxed} visitation of the root. Here, a conjunctive term is visited
+iff its atomic subterms are visited, and zero or one of its immediate
+probabilistic subterms are visited. For a sequential session, the
+starting state corresponds to an assumption (usually unique) with
+probability $1$. In our example, that starting state is:
 
@@
 sessions. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ from trace,
-$\state_0, \action_0,\state_1, \action_1,.., \state_N$, does not
-transition to $\state^\bot$, we have that the optimal trace has value:
+sequenced action, $\action_i$, from a trace of state-action pairs
+$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
+does not transition to $\state^\bot$, we have that the optimal trace
+has value:
 
@@
 belief-state. Targeted sensing is encouraged by augmenting the reward
-model to reflect the value of knowing the truth about assumptions in
-the trace proposed by the preceding sequential deliberation. In
-detail, all rewards from the original POMDP are
-retained. Additionally, for each $\assumptiveS{i}$ action scheduled by
-the current trace, we have a {\em dual} $\assumptiveDT{i}$ so that
-forall states $\state$:
+model to reflect a heuristic value of knowing the truth about
+assumptions in the trace. In detail, all rewards from the original
+POMDP are retained. Additionally, for each $\assumptiveS{i}$ action
+scheduled by the current trace, we have a {\em dual}
+$\assumptiveDT{i}$ so that for all states $\state$:
 
@@
 utility the agent receives for correctly (incorrectly) rejecting an
-assumption. Dual action execution is modelled as a self-transition --
-I.e., the process remains in the state from which the dual is
+assumption. A dual action can only be executed once, and corresponds
+otherwise to a self-transition -- i.e., except in the applicability of
+the dual, the process remains in the state from which a dual is
 executed.

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
grammar

@@
 We now describe our {\em switching} planning system that operates
-according to the continual planning paradigm, given descriptions of
+according to the continual planning paradigm, given a description of
 the current problem and domain in DTPDDL. The system {\em
@@
 Because online DT planning in large problems is impractical (we seek
-response times in seconds), DT sessions plan in an abstract process
+response times in seconds), a DT session plans in an abstract process
 determined by the current trace and underlying belief-state. This

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
spellers and typos

@@
 topological space, the latter indicate the category of space they
-inhabit -- e.g., Spoons are likely to be in kitchens. By examining
+inhabit -- e.g., spoons are likely to be in kitchens. By examining
 view cones at places for particular objects, the robot is able to: (1)

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
few extra corrections, and a reformatting of hte introduction

@@
 
-The DT session plans in a {\em small} abstract processes defined
+A DT session plans in a {\em small} abstract processes defined
 according to the assumptive actions in the proceeding trace, and the
@@
 utility the agent receives for correctly (incorrectly) rejecting an
-assumption. A dual action can only be executed once, and corresponds
-otherwise to a self-transition -- i.e., except in the applicability of
-the dual, the process remains in the state from which a dual is
-executed.
+assumption. A dual action can only be executed once, and except in the
+applicability of that dual corresponds to a self-transformation
 %%

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
new conclusion

@@
 system on a mobile robot platform, we developed a planner that
-switches between fast sequential and expensive DT planning. Given a
-DTPDDL model of the environment, sequential planning is used to
-compute an initial deterministic sequential plan and complementary
-runtime evolution of the decision process. That plan is executed until
-testing of an assumption about the runtime state is scheduled. At that
-point a DT session performs some sensing which determines whether or
-not the sequential session should be allowed to continue, or otherwise
-acts to achieve the overall objectives. We find our approach performs
-quickly and relatively efficiently in a number of large real-world
-task and observation planning problems.
+switches between: (a) fast sequential planning, and (b) expensive DT
+planning in small abstractions of the problem at hand. Given a DTPDDL
+model of the environment, sequential planning is used to compute an
+initial sequential plan and complementary runtime evolution of the
+decision process. That plan is executed until the outcome of the next
+action scheduled for execution is too uncertain. At that point a DT
+session performs sensing that can help determine how the sequential
+plan might evolve, or otherwise acts to achieve the goals. We find our
+approach performs quickly and relatively efficiently in a number of
+large real-world task and observation planning problems.
 

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
new conclusion

@@
 %%
-Finally, our approach is in a similar vein to {\em dual-mode}
+Also, our approach is in a similar vein to {\em dual-mode}
 control~\cite{cassandra96actingunder}, where planning switches between

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
applicability, is now outcome determinicity

@@
 execution proceeds by applying DTPDDL actions from the trace in
-sequence until the applicability of the next scheduled action is too
-uncertain according to a threshold parameter~(here, 95\%). In that
+sequence until the outcome of executing the next scheduled action is
+too uncertain according to a threshold parameter~(here, 95\%). In that
 eventuality a DT session commences, which tailors sensory processing

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
changes to switching section

@@
 determined by the current trace and underlying belief-state. This
-abstraction is constructed by first excluding all propositions (and
-related actions) that are not true, or {\em assumed } true, of states
-in the trace, then adding them back, using as an heuristic the entropy
-of the trace assumptions conditional on a candidate
-proposition. Propositions are added, one at a time, until the number
-of states in the initial belief-state reaches a given
-threshold.\footnote{Our experiments are performed using a number of
-different threshold parameters.}  To the resulting abstract model we
+abstract problem is created by selecting only those propositions that
+are \emph{relevant} to the subproblem the DT session is asked to
+solve.
+% MOG: Describe this in detail in the appropriate subsection
+%  This
+% abstraction is constructed by first excluding all propositions (and
+% related actions) that are not true, or {\em assumed } true, of states
+% in the trace, then adding them back, using as an heuristic the entropy
+% of the trace assumptions conditional on a candidate
+% proposition. Propositions are added, one at a time, until the number
+% of states in the initial belief-state reaches a given
+% threshold.\footnote{Our experiments are performed using a number of
+% different threshold parameters.} 
+To the resulting abstract model we
 also add {\em disconfirm} and {\em confirm} actions that the
@@
 otherwise). If a judgement action is scheduled for execution the
-DT session is terminated, and control is returned to a
-sequential session.
+DT session is terminated, and control is returned to the
+sequential planner.
 
@@
 continual planner maintains a factored representation of successive
-belief-states by performing belief revision. Our internal
-representation of the underlying belief-state corresponds more-or-less
-directly to an $(\pp{:init})$ declaration. That belief-state
-distribution is used: (1) as the source of candidate determinisations
-for sequential planning, (2) in determining when to switch to a DT
-session, and (3) as a mechanism to guide construction of an abstract
-process for DT sessions.
+belief-states by performing belief revision. For that, we transform
+the $(\pp{:init})$ declaration into a tree-shaped bayesian network
+which gets updated whenever an actions is performed or an observation
+is received. That belief-state representation is used:
+(1) as the source of candidate determinisations for sequential
+planning, (2) in determining when to switch to a DT session, and (3)
+as a mechanism to guide construction of an abstract process for DT
+sessions.
 
@@
 
-We only consider deterministic-action POMDPs, where all state
+As we only consider deterministic-action POMDPs, all state
 uncertainty is expressed in the $(\pp{:init})$ declaration. This
@@
 probabilistic subterms are visited. For a sequential session, the
-starting state corresponds to an assumption (usually unique) with
+starting state corresponds to an assumption (usually unique) which
 probability $1$. In our example, that starting state is:
@@
 Encapsulating state-assumptions, we annotate the problem posed during
-a sequential session with an assumptive action $\assumptiveS{i}$ for
+a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
 each element $\prob_i (T_i)$, of each probabilistic term from
@@
 
-A DT session plans in a {\em small} abstract processes defined
-according to the assumptive actions in the proceeding trace, and the
-$(\pp{:init})$ declaration that models the underlying
-belief-state. Targeted sensing is encouraged by augmenting the reward
-model to reflect a heuristic value of knowing the truth about
+When an action that requires sensing is encountered, the planner
+switches to a decision theoretic session. That session plans in a {\em
+  small} abstract processes defined by the action that triggered the
+DT session, the assumptive actions in the proceeding trace and the
+current belief state. Targeted sensing is encouraged by augmenting the
+reward model to reflect a heuristic value of knowing the truth about
 assumptions in the trace. In detail, all rewards from the original
-POMDP are retained. Additionally, for each $\assumptiveS{i}$ action
-scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that for all states $\state$:
+POMDP are retained. Additionally, for each relevant assumptive action
+$\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
+  action} $\assumptiveDT{i}$ so that for all states $\state$:
 
@@
 utility the agent receives for correctly (incorrectly) rejecting an
-assumption. A dual action can only be executed once, and except in the
+assumption. A disconfirm action can only be executed once, and except in the
 applicability of that dual corresponds to a self-transformation
 %%
-Furthermore, if \switchAction\ is the action that switched the system
-to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
-active} if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. We only include the duals of active assumptions in
-DT sessions. Continuing our simplified example, if the trace is:
+If \switchAction\ is the action that switched the system to a DT
+session, then an assumption $\assumptiveS{i}$ is {\em relevant} if it
+is directly or indirectly required for \switchAction\ to be
+applicable. Continuing our simplified example, if the trace is:
+
+% Furthermore, if \switchAction\ is the action that switched the system
+% to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
+% active} if
+% \switchAction\ is not applicable in the sequential model unless
+% $\assumptiveS{i}$ is scheduled in the trace prefix
+% to \switchAction. We only include the duals of active assumptions in
+% DT sessions. Continuing our simplified example, if the trace is:
 
@@
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude it from the
+relevant for this session, and therefore exclude it from the
 POMDP posed to the DT session.
@@
 Given \switchAction, we also include another self-transition action
-$\actions.\poss(\switchAction)$ with the reward property:
+$\actions.\poss(\switchAction)$, the \emph{goal action} with the
+reward property:
 
@@
 
-Execution of either a dual action $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session, which then continues from the underlying
-belief-state. Turning to the detail of dual rewards, in our integrated
-system these are sourced from a motivational subsystem. In this paper,
-for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a small positive
-constant, and have $\hat\$(x)= - \$(x)(1 - \prob) / \prob$ where
-$\prob$ is the probability that $x$ is true. For
+Execution of either a disconfirmation or the goal action returns
+control to a sequential session, which then continues from the
+underlying belief-state. Turning to the detail of rewards, in our
+integrated system these are sourced from a motivational subsystem. In
+this paper, for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a
+small positive constant, and have $\hat\$(x)= - \$(x)(1 - \prob) /
+\prob$ where $\prob$ is the probability that $x$ is true. For
 $\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$
+\$(x)\prob/(1-\prob)$.
 
@@
 
-The process posed to a DT session is abstract. The abstraction is
-defined by first constructing an $(\pp{:init})$ declaration without
-any-active assumptions from the trace -- e.g.,
-(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
-probabilistic terms, and circles are atomic and/or conjunctive.
+As are aiming to keep the decision theoretic session fast, we try to
+keep the state space of the abstract problem as small as possible. In
+a deterministic-actio POMDP, the state space size is determined by two
+factors: the number of non-zero states in the initial belief-state and
+the number of executable actions. We thus want to keep the number of
+uncertain facts as low as possible without losing important
+information about the sensing goal and keep the number of actions down
+without crippling the ability of the DT planner to gather the required
+information.
+
+We perform the abstraction of the belief space by first constructing
+an $(\pp{:init})$ declaration which only contains relevant assumptions
+from the trace -- e.g., (Fig.\ref{fig:abstraction-b}) gives an
+example, where diamonds are probabilistic terms, and circles are
+atomic and/or conjunctive.
 %%
@@
 assumptions. Facts are iteratively added to the belief-state in
-increasing order according to that measure
-(Fig.\ref{fig:abstraction-c}).
+increasing order according to that measure until the belief space size
+reaches a predefined limit (Fig.\ref{fig:abstraction-c}).
+
+We then proceed to limit the actions by only including those that are
+possibly required to a) reach the goal or b) to directly observe an
+atom that is part of the abstract belief state.
+
+TODO: relaxed planning graph
 
@@
 %%% mode: latex
-%%% TeX-master: "moritz_2011"
+%%% TeX-master: "aaai11"
 %%% End:

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
changes to switching section

@@
 minutes (1800 seconds)\footnote{All experiments were conducted on a
-  2.3GHz AMD Opteron using one CPU core.}. The continual planning
+ 2.66GHz Intel Xeon X5355 CPU using one CPU core.}. The continual planning
 times are reported in Figure~\ref{fig:results-time}, and the quality

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
remove paragraph from evaluation

@@
 no plan exists, as the desired object might not be present in the
-environment. Item~$f$ reports for tasks requiring indirect sensing,
-where the robot must relocate to a room with a particular category
-(e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
-so is omitted from that reporting.  Because reward is only allocated
+environment. 
+% MOG: I didn't do  evaluations for those, so let's remove them
+% Item~$f$ reports for tasks requiring indirect sensing,
+% where the robot must relocate to a room with a particular category
+% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
+% so is omitted from that reporting.
+
+Because reward is only allocated
 once, on the achievement of the goal,  we find it
@@
   \includegraphics{pomdp-quality}\hfill
-  \includegraphics{pomdp-solvable-quality}\hfill
+  % \includegraphics{pomdp-solvable-quality}\hfill
   \includegraphics{dora2-quality}\hfill

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Revised text about state space abstraction

@@
 
-The process posed to a DT session is abstract. The abstraction is
-defined by first constructing an $(\pp{:init})$ declaration without
-any-active assumptions from the trace -- e.g.,
-(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
-probabilistic terms, and circles are atomic and/or conjunctive.
+As we are aiming to keep the decision theoretic session fast, we try
+to keep the state space of the abstract problem as small as
+possible. In a deterministic-action POMDP, the state space size is
+determined by two factors: the number of non-zero states in the
+initial belief-state and the number of executable actions. As
+determining the relevance of an action for a given goal in in general
+a hard problem, we concentrate on keeping the number of uncertain
+facts as low as possible without losing important information about
+the sensing goal.
+
+We perform the abstraction of the belief space by first constructing
+an $(\pp{:init})$ declaration which only contains relevant assumptions
+from the trace -- e.g., (Fig.\ref{fig:abstraction-b}) gives an
+example, where diamonds are probabilistic terms, and circles are
+atomic and/or conjunctive.
 %%
@@
 assumptions. Facts are iteratively added to the belief-state in
-increasing order according to that measure
-(Fig.\ref{fig:abstraction-c}).
-
-%% In the first step, we remove all facts that are not part of an
-%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-%% would proceed in an abstraction of the environment that does not
-%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%% %%
-%% In a second step, we iteratively refine the relaxed declaration by
-%% adding terms from the original statement of $(\pp{:init})$ while the
-%% number of abstract states in $\bstate_0$ that occur with non-zero
-%% probability according to that refined declaration remains of a
-%% practicable size. In detail,
+increasing order according to that measure until the belief space size
+reaches a predefined limit or no more atoms that reduce the entropy
+can be added (Fig.\ref{fig:abstraction-c}).
 
@@
 
+% When limiting the number of actions, we aim to include those that are
+% possibly required to a) reach the goal or b) to directly observe an
+% atom that is part of the abstract belief state. In general, checking
+% this for an action is \textsc{PSPACE} complete for \textsc{STRIPS}
+% planning.
+
+% We approximate the set of relevant actions as follows: Let
+% \relevantAtoms\ be the set of atoms that either occur in the goal
+% condition \poss(\switchAction) or in a precondition \poss($a$) of an
+% action $a$ that can directly sense any atom in the abstract belief
+% state. We construct a \emph{relaxed planning graph} with
+% \relevantAtoms\ as goal atoms from which we extract a relaxed
+% plan.
+
+%% In the first step, we remove all facts that are not part of an
+%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+%% would proceed in an abstraction of the environment that does not
+%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
+%% %%
+%% In a second step, we iteratively refine the relaxed declaration by
+%% adding terms from the original statement of $(\pp{:init})$ while the
+%% number of abstract states in $\bstate_0$ that occur with non-zero
+%% probability according to that refined declaration remains of a
+%% practicable size. In detail,
+
 %%% Local Variables: 
 %%% mode: latex
-%%% TeX-master: "moritz_2011"
+%%% TeX-master: "aaai11"
 %%% End:

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
Fixed graphs + small changes to evaluation section

@@
 co-occurrence and room categories for efficiency. Also, in the
-presence of a person, the robot can ask them about the location of
-objects.
+presence of a person, the robot can ask them about the category of the
+current room.
 
@@
 realistic tasks, with the number of rooms ranging from 3 (12-places,
-16-objects, $|$states$|$$>\sim10^{21}$) to 6 (26-places, 21-objects,
+16-objects, $|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects,
 $|$states$|$ $>10^{36}$). For or experiments we model three sensors:
@@
 
-We evaluate with DT sessions taking $b_0$ admitting between 20 and 200
+We evaluate with DT sessions taking $b_0$ admitting between 20 and 100
 abstract states with non-zero probability. We run 50 simulations in
@@
 minutes (1800 seconds)\footnote{All experiments were conducted on a
- 2.66GHz Intel Xeon X5355 CPU using one CPU core.}. The continual planning
+ 2.66GHz Intel Xeon X5355 using one CPU core.}. The continual planning
 times are reported in Figure~\ref{fig:results-time}, and the quality
@@
 %%
-For Items~$a$-$e$ the goal is to find objects and report their
-position to a user. In some cases there is a non-zero probability that
+For each task, the goal is to find one or several objects and report
+their position to a user. Usually there is a non-zero probability that
 no plan exists, as the desired object might not be present in the
@@
 % so is omitted from that reporting.  
-
 Because reward is only allocated
@@
 refined initial configurations, the increased cost of DT
-planning is compensated for by a decrease in Fast Downward planning
+planning is partially compensated for by a decrease in Fast Downward planning
 times. The relatively high success rate irrespective of the level of

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
changes requested by moritz

@@
 co-occurrence and room categories for efficiency. Also, in the
-presence of a person, the robot can ask them about the location of
-objects.
+presence of a person, the robot can ask them about the category of the
+current room.
 
-We compare our switching planner with the {\em baseline} in several
-realistic tasks, with the number of rooms ranging from 3 (12-places,
-16-objects, $|$states$|$$>\sim10^{21}$) to 6 (26-places, 21-objects,
-$|$states$|$ $>10^{36}$). For or experiments we model three sensors:
-{\em Reliable} sensors have a $0.1$ probability of a false negative,
-{\em semi-reliable} have a chance of $0.3$ of false negative and $0.1$
-of false positive , and finally {\em noisy} sensors with probabilities
-of $0.5$ and $0.2$ respectively. Each object class is assigned one
-sensor model, so e.g. cornflakes may be harder to detect than
-refrigeratos. To determine the influence of the sensor model, we
-performed several experiments where we only changed the difficulty of
-the target object(s). All non-target objects remain unchanged so the
-planner may exploit object co-occurances with reliable objects to
-compensate for the increased sensor noise.
+We compare switching with the {\em baseline} in several realistic
+tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
+$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
+$>10^{36}$). We also compare those systems with an approximately
+optimal policy computed using Smith's {\sc zmdp} for small 2 room
+problems (4-places, 3-objects, $|$states$|$ $\simeq 5000$). For or
+experiments we model three sensors: {\em Reliable} sensors have a
+$0.1$ probability of a false negative, {\em semi-reliable} have a
+chance of $0.3$ of false negative and $0.1$ of false positive , and
+finally {\em noisy} sensors with probabilities of $0.5$ and $0.2$
+respectively. Each object class is assigned one sensor model, so
+e.g. cornflakes may be harder to detect than refrigeratos. To
+determine the influence of the sensor model, we performed several
+experiments where we only changed the difficulty of the target
+object(s). All non-target objects remain unchanged so the planner may
+exploit object co-occurances with reliable objects to compensate for
+the increased sensor noise.
 
-We evaluate with DT sessions taking $b_0$ admitting between 20 and 200
+We evaluate with DT sessions taking $b_0$ admitting between 20 and 100
 abstract states with non-zero probability. We run 50 simulations in
@@
 %%
-For Items~$a$-$e$ the goal is to find objects and report their
-position to a user. In some cases there is a non-zero probability that
+For each task, the goal is to find one or several objects and report
+their position to a user. Usually there is a non-zero probability that
 no plan exists, as the desired object might not be present in the
-environment. Item~$f$ reports for tasks requiring indirect sensing,
-where the robot must relocate to a room with a particular category
-(e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
-so is omitted from that reporting.  Because reward is only allocated
+environment.
+% MOG: I didn't do  evaluations for those, so let's remove them
+% Item~$f$ reports for tasks requiring indirect sensing,
+% where the robot must relocate to a room with a particular category
+% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
+% so is omitted from that reporting.  
+Because reward is only allocated
 once, on the achievement of the goal,  we find it

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
changes requested by moritz

@@
 The system always begins in a sequential session, so that plan
-execution proceeds by applying DTPDDL actions from the trace in
-sequence until the outcome of executing the next scheduled action is
+execution proceeds by applying DTPDDL actions from the trace, in
+sequence, until the outcome of executing the next scheduled action is
 too uncertain according to a threshold parameter~(here, 95\%). In that
@@
 Because online DT planning in large problems is impractical (we seek
-response times in seconds), a DT session plans in an abstract process
-determined by the current trace and underlying belief-state. This
-abstract problem is created by selecting only those propositions that
-are \emph{relevant} to the subproblem the DT session is asked to
-solve.
+response times in seconds), a DT session plans for an abstract process
+determined by the current trace and underlying belief-state. That
+abstract problem is characterised by a limited number of propositions,
+chosen due to their {\em relevance} to the action whose scheduled
+execution triggered the switch to the DT session.
+%%the subproblem the DT session is asked tosolve.
 % MOG: Describe this in detail in the appropriate subsection
@@
 % different threshold parameters.} 
-To the resulting abstract model we
-also add {\em disconfirm} and {\em confirm} actions that the
-DT session can schedule in order to judge an atomic assumption
-in the trace. In the abstract model these actions yield a relatively
-small reward if the corresponding judgement is true (or small penalty
-otherwise). If a judgement action is scheduled for execution the
-DT session is terminated, and control is returned to the
-sequential planner.
+To that abstract model we also add {\em disconfirm} and {\em confirm}
+actions that the DT session can schedule in order to judge an atomic
+assumption in the trace. Those actions yield a relatively small reward
+if the corresponding judgement is true (or small penalty
+otherwise). If a judgement action is scheduled for execution the DT
+session is terminated, and control is returned to a new sequential
+session.
 
-Finally, whether proceeding in a sequential or DT session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision. For that, we transform
-the $(\pp{:init})$ declaration into a tree-shaped bayesian network
-which gets updated whenever an actions is performed or an observation
-is received. That belief-state representation is used:
-(1) as the source of candidate determinisations for sequential
-planning, (2) in determining when to switch to a DT session, and (3)
-as a mechanism to guide construction of an abstract process for DT
-sessions.
+Whatever the session type, our continual planner maintains a factored
+representation of successive belief-states by performing belief
+revision. As an internal representation of the $(\pp{:init})$
+declaration, we keep a tree-shaped Bayesian network which gets updated
+whenever an action is performed, or an observation received. That
+belief-state representation is used: (1) as the source of candidate
+determinisations for sequential planning, (2) in determining when to
+switch to a DT session, and (3) as a mechanism to guide construction
+of an abstract process for DT sessions.
 
@@
 \noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root. Here, a conjunctive term is visited
+relaxed} visitation of the root term. Here, a conjunctive term is visited
 iff its atomic subterms are visited, and zero or one of its immediate
 probabilistic subterms are visited. For a sequential session, the
-starting state corresponds to an assumption (usually unique) which
+starting state corresponds to an assumption (usually unique) with
 probability $1$. In our example, that starting state is:
@@
 %%
-We add additional constraints that forbid the scheduling of
-assumptions about facts after actions with preconditions, or effects,
+We also add constraints that forbid scheduling of
+assumptions about facts after actions with preconditions or effects
 that mention those facts. For example, the robot cannot assume it is
@@
 has value:
-
-%% define the value of a trace
-%% $\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-%% deterministic model to equal:
-
-%% \small
-%% \begin{equation}\label{eq:tracevalue}
-%% V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-%% \end{equation}
-%% \normalsize
-
-%% \noindent The
-%% optimal trace given a plan, i.e., the sequence of non-assumptive
-%% actions from the trace, therefore has value:
 
@@
 
-When an action that requires sensing is encountered, the planner
-switches to a decision theoretic session. That session plans in a {\em
-  small} abstract processes defined by the action that triggered the
-DT session, the assumptive actions in the proceeding trace and the
-current belief state. Targeted sensing is encouraged by augmenting the
-reward model to reflect a heuristic value of knowing the truth about
-assumptions in the trace. In detail, all rewards from the original
-POMDP are retained. Additionally, for each relevant assumptive action
+When an action is scheduled whose outcome is uncertain according to
+the underlying belief-state, the planner switches to a DT
+session. That plans for {\em small} abstract processes defined
+according to the action that triggered the DT session, the assumptive
+actions in the proceeding trace, and the current
+belief-state. Targeted sensing is encouraged by augmenting the reward
+model to reflect a heuristic value of knowing the truth about
+assumptions. In detail, all rewards from the underlying problem are
+retained. Additionally, for each relevant assumptive action
 $\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
@@
 \noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
-relatively\footnote{Compared to rewards from the original process
-model.} small positive (negative) numeric quantity which captures the
-utility the agent receives for correctly (incorrectly) rejecting an
-assumption. A disconfirm action can only be executed once, and except in the
-applicability of that dual corresponds to a self-transformation
+relatively small positive (negative) numeric quantity\footnote{Small
+relative to rewards from the original process model.} which captures
+the utility the agent receives for correctly (incorrectly) rejecting
+an assumption. A disconfirm action can only be executed once, and
+except in the applicability of that dual corresponds to a
+self-transformation
 %%
-If \switchAction\ is the action that switched the system to a DT
-session, then an assumption $\assumptiveS{i}$ is {\em relevant} if it
-is directly or indirectly required for \switchAction\ to be
-applicable. Continuing our simplified example, if the trace is:
+We only consider {\em relevant} assumptions and their duals when
+constructing the abstract model.  If \switchAction\ is the action that
+switched the system to a DT session, then an assumption
+$\assumptiveS{i}$ is {\em relevant} if it is necessary for the outcome
+of \switchAction\ to be determined. Continuing our simplified example,
+if the trace is:
 
@@
 relevant for this session, and therefore exclude it from the
-POMDP posed to the DT session.
+abstract decision process.
 

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
more small graph changes
few suggestions from moritz

@@
 
-We have implemented our switching continual approach in the MAPSIM
+We have implemented our switching approach in the MAPSIM
 environment~\cite{brenner:nebel:jaamas09},
@@
 baseline} approach. Here, when a switching action is scheduled for
-execution the DT session executes a single entropy reduction action,
-that directly senses an active assumption -- i.e,. We say a fact can
-be directly sensed, if there is a DTPDDL {\em sense} declaration that
-mentions the fact in its precondition.  Control is then immediately
+execution the DT session applies a single entropy reduction action,
+whose execution can provide evidence regarding the truth value of an
+active assumption.
+%%
+%%  -- i.e,. we say a fact can
+%% be directly sensed, if there is a DTPDDL {\em sense} declaration that
+%% mentions the fact in its precondition. 
+%%
+Control is then immediately
 returned to a new sequential session.
@@
 office environments. Spatially, these consist of {\em rooms}
-(office/kitchen/etc), and an underlying topological map that models
+(office/kitchen/etc), and an underlying topological map over
 smaller areas of space, called {\em places}, and connectivity between
 those. The mobile {\em robot} and {\em visual objects} inhabit the
-topological space, the latter indicate the category of space they
+topological places. Objects indicate the category of space they
 inhabit -- e.g., spoons are likely to be in kitchens. By examining
@@
 co-occurrence and room categories for efficiency. Also, in the
-presence of a person, the robot can ask them about the category of the
+presence of a person, the robot can ask about the category of the
 current room.
 
-We compare switching with the {\em baseline} in several realistic
+We compare switching to the {\em baseline} in several realistic
 tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
 $|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
-$>10^{36}$). We also compare those systems with an approximately
-optimal policy computed using Smith's {\sc zmdp} for small 2 room
-problems (4-places, 3-objects, $|$states$|$ $\simeq 5000$). For or
-experiments we model three sensors: {\em Reliable} sensors have a
-$0.1$ probability of a false negative, {\em semi-reliable} have a
-chance of $0.3$ of false negative and $0.1$ of false positive , and
-finally {\em noisy} sensors with probabilities of $0.5$ and $0.2$
-respectively. Each object class is assigned one sensor model, so
-e.g. cornflakes may be harder to detect than refrigeratos. To
-determine the influence of the sensor model, we performed several
-experiments where we only changed the difficulty of the target
-object(s). All non-target objects remain unchanged so the planner may
-exploit object co-occurances with reliable objects to compensate for
-the increased sensor noise.
+$>10^{36}$). We also compare those systems with near optimal policies
+computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
+3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
+levels of reliability in sensing: {\em reliable} sensors have a $.1$
+probability of a false negative, {\em semi-reliable} have a chance of
+$0.3$ of false negative and $0.1$ of false positive , and {\em noisy}
+sensors with probabilities of $0.5$ and $0.2$ respectively. Each
+object class is assigned one sensor model -- e.g. cornflakes may be
+harder to detect than refrigerators. We performed several experiments
+with different levels of reliability for sensing the target object(s),
+while keeping sensing models for non-target objects constant.
 
-We evaluate with DT sessions taking $b_0$ admitting between 20 and 100
-abstract states with non-zero probability. We run 50 simulations in
-each configuration, and have a timeout on each simulation of 30
-minutes (1800 seconds)\footnote{All experiments were conducted on a
- 2.66GHz Intel Xeon X5355 using one CPU core.}. The continual planning
-times are reported in Figure~\ref{fig:results-time}, and the quality
-data in Figure~\ref{fig:results-quality}.
+\Omit{
+All non-target objects remain unchanged so the planner can
+exploit object co-occurances using reliably sensed objects to
+compensate for the increased sensor noise.
+}
+
+We experimented with DT sessions with initial belief-states admitting
+between 20 and 100 abstract states with non-zero probability. We run
+50 simulations in each configuration, and have a timeout on each
+simulation of 30 minutes (1800 seconds)\footnote{All experiments were
+conducted on a 2.66GHz Intel Xeon X5355 using one CPU core.}. The
+continual planning times are reported in
+Figure~\ref{fig:results-time}, and the quality data in
+Figure~\ref{fig:results-quality}.
 %%
-For each task, the goal is to find one or several objects and report
+For each task, the goal is to find one or more objects and report
 their position to a user. Usually there is a non-zero probability that
@@
 % so is omitted from that reporting.  
-Because reward is only allocated
-once, on the achievement of the goal,  we find it
-intuitive to report average plan costs and the success rates in
-problems that admit a solution (i.e., positive reward scaled by a
+Because reward is only allocated on the achievement of a goal, we
+find it intuitive to report average plan costs and the success rates
+in problems that admit a solution (i.e., positive reward scaled by a
 constant factor).
@@
   % \vspace{2mm}
-  \includegraphics{pomdp-quality}\hfill
   % \includegraphics{pomdp-solvable-quality}\hfill
@@
   \includegraphics{dora56-quality}\hfill
+  \vspace{2mm}
+  \includegraphics{pomdp-quality}\hfill
   % \vspace{2mm}
@@
 
-We find that if sensing is reliable there is little to be gained
-through DT sessions, as the greedy approach of the {\em baseline} is
-sufficient. As sensing degrades contingent planning pays off.  Time
-spent in DT planning increases steeply as the abstraction $\bstate_0$
-becomes more refined.  That refinement seems to be paying off in terms
-of the success rate, particularly for tasks $d$ and $e$. For less
-refined initial configurations, the increased cost of DT
-planning is partially compensated for by a decrease in Fast Downward planning
-times. The relatively high success rate irrespective of the level of
-refinement in the initial configuration indicates the effectiveness of
-using conditional entropy to guide abstraction refinement in our
-setting.
+We find that if sensing is reliable, then little is gained using DT
+sessions, as the greedy approach of the {\em baseline} is
+sufficient. As sensing degrades DT sessions prove more useful. Here,
+time spent on DT planning increases steeply as the abstraction becomes
+more refined, which is compensated for by fewer planning sessions
+overall. More detailed abstractions lead to a better overall
+success rate, particularly for tasks $d$ and $e$.
+%% Moreover,
+%% for less refined abstractions, the increased cost of DT planning is
+%% partially compensated for by a decrease in Fast Downward planning
+%% times. 
+Speaking to the effectiveness of our entropy heuristic for abstraction
+refinement, we witness relatively high success rates irrespective of
+the exact level of refinement. Comparing finally to the best {\sc
+zmdp} policy, although producing relatively costly plans, the
+continual planners performed quite well, especially in terms success
+rate. A key source of inefficiency here, is due to sequential sessions
+always being optimistic, and refusing to abandon the search.
 

### 2011-02-08 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
not sure what I changed in related

@@
 co-occurrence and room categories for efficiency. Also, in the
-presence of a person, the robot can ask them about the category of the
+presence of a person, the robot can ask about the category of the
 current room.
 
-We compare our switching planner with the {\em baseline} in several
-realistic tasks, with the number of rooms ranging from 3 (12-places,
-16-objects, $|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects,
-$|$states$|$ $>10^{36}$). For or experiments we model three sensors:
-{\em Reliable} sensors have a $0.1$ probability of a false negative,
-{\em semi-reliable} have a chance of $0.3$ of false negative and $0.1$
-of false positive , and finally {\em noisy} sensors with probabilities
-of $0.5$ and $0.2$ respectively. Each object class is assigned one
-sensor model, so e.g. cornflakes may be harder to detect than
-refrigeratos. To determine the influence of the sensor model, we
-performed several experiments where we only changed the difficulty of
-the target object(s). All non-target objects remain unchanged so the
-planner may exploit object co-occurances with reliable objects to
+We compare switching to the {\em baseline} in several realistic
+tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
+$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
+$>10^{36}$). We also compare those systems with near optimal policies
+computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
+3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
+levels of reliability in sensing: {\em reliable} sensors have a $.1$
+probability of a false negative, {\em semi-reliable} have a chance of
+$0.3$ of false negative and $0.1$ of false positive, and {\em noisy}
+sensors with probabilities of $0.5$ and $0.2$ respectively. Each
+object class is assigned one sensor model -- e.g. cornflakes may be
+harder to detect than refrigerators. We performed several experiments
+with different levels of reliability for sensing the target object(s),
+while keeping sensing models for non-target objects constant.
+
+\Omit{
+All non-target objects remain unchanged so the planner can
+exploit object co-occurances using reliably sensed objects to
 compensate for the increased sensor noise.
+}
 
-We evaluate with DT sessions taking $b_0$ admitting between 20 and 100
-abstract states with non-zero probability. We run 50 simulations in
-each configuration, and have a timeout on each simulation of 30
-minutes (1800 seconds)\footnote{All experiments were conducted on a
- 2.66GHz Intel Xeon X5355 using one CPU core.}. The continual planning
-times are reported in Figure~\ref{fig:results-time}, and the quality
-data in Figure~\ref{fig:results-quality}.
+We experimented with DT sessions with initial belief-states admitting
+between 20 and 100 abstract states with non-zero probability. We run
+50 simulations in each configuration, and have a timeout on each
+simulation of 30 minutes (1800 seconds)\footnote{All experiments were
+conducted on a 2.66GHz Intel Xeon X5355 using one CPU core.}. The
+continual planning times are reported in
+Figure~\ref{fig:results-time}, and the quality data in
+Figure~\ref{fig:results-quality}.
 %%
-For each task, the goal is to find one or several objects and report
+For each task, the goal is to find one or more objects and report
 their position to a user. Usually there is a non-zero probability that
@@
 % so is omitted from that reporting.  
-Because reward is only allocated
-once, on the achievement of the goal,  we find it
-intuitive to report average plan costs and the success rates in
-problems that admit a solution (i.e., positive reward scaled by a
-constant factor).
+In these experiments we only allocate reward on the achievement of all
+goals, therefore we find it intuitive to report average plan costs and
+the success rates in problems that admit a complete solution (i.e.,
+positive reward scaled by a constant factor).
 
@@
 
-We find that if sensing is reliable there is little to be gained
-through DT sessions, as the greedy approach of the {\em baseline} is
-sufficient. As sensing degrades contingent planning pays off.  Time
-spent in DT planning increases steeply as the abstraction $\bstate_0$
-becomes more refined.  That refinement seems to be paying off in terms
-of the success rate, particularly for tasks $d$ and $e$. For less
-refined initial configurations, the increased cost of DT
-planning is partially compensated for by a decrease in Fast Downward planning
-times. The relatively high success rate irrespective of the level of
-refinement in the initial configuration indicates the effectiveness of
-using conditional entropy to guide abstraction refinement in our
-setting.
+We find that if sensing is reliable, then little is gained using DT
+sessions, as the greedy approach of the {\em baseline} is
+sufficient. As sensing degrades DT sessions prove more useful. Here,
+time spent on DT planning increases steeply as the abstraction becomes
+more refined, which is compensated for by fewer planning sessions
+overall. More detailed abstractions lead to a better overall
+success rate, particularly for tasks $d$ and $e$.
+%% Moreover,
+%% for less refined abstractions, the increased cost of DT planning is
+%% partially compensated for by a decrease in Fast Downward planning
+%% times. 
+Speaking to the effectiveness of our entropy heuristic for abstraction
+refinement, we see relatively high success rates irrespective of
+the level of refinement. Comparing finally to the best {\sc
+zmdp} policy, although producing relatively costly plans, the
+continual planners performed quite well, especially in terms success
+rate. A key source of inefficiency here, is due to sequential sessions
+always being optimistic, and refusing to abandon the search.
 

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/dtpddl.tex
removed too large footnote removed comments from pddl examples

@@
-%% BEGIN SCRAPPY
-
-%% deal with POND and the Perseus language
-
-%% For example, the robot usually knows exactly where it is located on a
-%% topological map, and therefore there are tight constraints about what
-%% regions of the map can be observed, and what regions of the map can be
-%% moved to in an atomic step.
-
-%% Moreover, we use this domain description to model the action physics
-%% of abstract processes posed to decision-theoretic planning
-%% sessions. Here, we forbid the planner from considering courses of
-%% action that are not {\em known} to be possible to alleviate the
-%% computational burden planning in the abstract process.
-
-%% END SCRAPPY
-
 We give an overview of the declarative first-order language DTPDDL, an
@@
 from problems expressed in DTPDDL to flat (and factored)
-representations of the underlying decision processes. Although similar
+representations of the underlying decision process. Although similar
 to Bryce's POND input language, DTPDDL distinguishes itself by
@@
 more compact domain descriptions where sensing effects are common
-across multiple actions. In detail, DTPDDL has perceptual analogues of
-fluent and predicate symbols. For example our {\em object search}
-domains have:
+across multiple operators. In detail, DTPDDL has perceptual analogues of
+fluent and predicate symbols. For example, a simple {\em object search}
+domain would have:
 
@@
 \begin{tabtt}
-(\=:functions  ;; state fluents\\
+(\=:functions\\
   \> (is-in ?v - visual-object) - location )\\
-(:perceptual-functions  ;; perceptual fluents\\
+(:perceptual-functions\\
   \> (o-is-in ?v - visual-object) - location )
@@
 %%
-To model sensing capabilities, we have operator-like declarations,
-with preconditions expressed using state and action symbols, and
-uniformly positive effects over perceptual symbols. For example, where
-{\em look-for-object} is the operator that applies an object detection
-algorithm at a specific place, in a simplified {\em object search} task
-we have:
+To model sensing capabilities, we have operator-like ``sense''
+declarations, with preconditions expressed using state and action
+symbols, and uniformly positive effects over perceptual symbols. For
+example, where {\em look-for-object} is the operator that applies an
+object detection algorithm at a specific place, an {\em
+object search} task will have:
 
@@
 \vspace{-3ex}
+
+\noindent I.e., there is a 10\% {\em false positive} rate, and 20\%
+  probability of a {\em false negative}.
 
@@
 
-If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion.
+%% If a dual is executed, then the sequential planner must
+%% replan with $\bstate_0$ equal to the underlying
+%% belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
+%% executed, then the current sequential plan is executed until further
+%% sensing is scheduled, or to completion.
 
@@
 % problem description.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "aaai11"
+%%% End:

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
more switching text

@@
 the utility the agent receives for correctly (incorrectly) rejecting
-an assumption. A disconfirm action can only be executed once, and
-except in the applicability of that dual corresponds to a
-self-transformation
+an assumption.
 %%
-We only consider {\em relevant} assumptions and their duals when
-constructing the abstract model.  If \switchAction\ is the action that
-switched the system to a DT session, then an assumption
-$\assumptiveS{i}$ is {\em relevant} if it is necessary for the outcome
-of \switchAction\ to be determined. Continuing our simplified example,
-if the trace is:
-
-% Furthermore, if \switchAction\ is the action that switched the system
-% to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
-% active} if
-% \switchAction\ is not applicable in the sequential model unless
-% $\assumptiveS{i}$ is scheduled in the trace prefix
-% to \switchAction. We only include the duals of active assumptions in
-% DT sessions. Continuing our simplified example, if the trace is:
+In terms of action physics, a disconfirm action can only be executed
+once, and otherwise is modelled as self-transformation.
+%%
+We only consider {\em relevant} assumptions when constructing the
+abstract model.  If \switchAction\ is the action that switched the
+system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
+relevant} if it is necessary for the outcome of \switchAction\ to be
+determined.  Continuing our simplified example, if the trace is:
 
@@
 
-Given \switchAction, we also include another self-transition action
-$\actions.\poss(\switchAction)$, the \emph{goal action} with the
-reward property:
+Given \switchAction, we also include another once-only self-transition
+action $\actions.\poss(\switchAction)$, the \emph{confirmation action}
+with the reward property:
 
@@
 
-Execution of either a disconfirmation or the goal action returns
-control to a sequential session, which then continues from the
-underlying belief-state. Turning to the detail of rewards, in our
-integrated system these are sourced from a motivational subsystem. In
-this paper, for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a
-small positive constant, and have $\hat\$(x)= - \$(x)(1 - \prob) /
+Execution of either a disconfirmation or the confirmation action
+returns control to a sequential session, which then continues from the
+underlying belief-state.
+
+Turning to the detail of (dis-)confirmation rewards, in our integrated
+system these are sourced from a motivational subsystem. In this paper,
+for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a small positive
+constant, and have $\hat\$(x)= - \$(x)(1 - \prob) /
 \prob$ where $\prob$ is the probability that $x$ is true. For
@@
 
-As we are aiming to keep the decision theoretic session fast, we try
-to keep the state space of the abstract problem as small as
-possible. In a deterministic-action POMDP, the state space size is
-determined by two factors: the number of non-zero states in the
-initial belief-state and the number of executable actions. As
-determining the relevance of an action for a given goal in in general
-a hard problem, we concentrate on keeping the number of uncertain
-facts as low as possible without losing important information about
-the sensing goal.
+%%  the size of the state space
+%% is determined by two factors: the number of non-zero states in the
+%% initial belief-state, and the number of actions. Determining
+%% action reachability is intractable, as is determining their relevance
+%% to goal achievement.
 
-We perform the abstraction of the belief space by first constructing
-an $(\pp{:init})$ declaration which only contains relevant assumptions
-from the trace -- e.g., (Fig.\ref{fig:abstraction-b}) gives an
-example, where diamonds are probabilistic terms, and circles are
-atomic and/or conjunctive.
+%% As we are aiming to keep the decision theoretic session fast, we try
+%% to keep the state space of the abstract problem as small as
+%% possible. In a deterministic-action POMDP, the state space size is
+%% determined by two factors: the number of non-zero states in the
+%% initial belief-state and the number of executable actions. As
+%% determining the relevance of an action for a given goal in in general
+%% a hard problem, we concentrate on keeping the number of uncertain
+%% facts as low as possible without losing important information about
+%% the sensing goal.
+
+In order to guarantee fast DT sessions, planning occurs for a small
+abstract problem. In our setting, where the underlying decision
+process has a deterministic action model, a significant cause of
+problem largeness is due to the number of states that occur with
+non-zero probability in the underlying belief-state. In posing a small
+abstract process for a DT session, we construct an abstract
+belief-state that is chiefly characterised by facts related to the
+current trace. In that abstraction, instances of operator or sense
+declarations whose specification mentions an excluded fact are not
+available.
+%%
+In detail, we first construct an $(\pp{:init})$ declaration which only
+contains relevant assumptions from the trace -- e.g.,
+(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
+probabilistic terms, and circles are atomic and/or conjunctive.
 %%
@@
 increasing order according to that measure until the belief space size
-reaches a predefined limit or no more atoms that reduce the entropy
-can be added (Fig.\ref{fig:abstraction-c}).
+reaches a predefined limit (Fig.\ref{fig:abstraction-c}).
+
+%%  or no more atoms that reduce the entropy
+%% can be added
 

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
small footnote

@@
 \noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root term. Here, a conjunctive term is visited
-iff its atomic subterms are visited, and zero or one of its immediate
-probabilistic subterms are visited. For a sequential session, the
-starting state corresponds to an assumption (usually unique) with
-probability $1$. In our example, that starting state is:
+relaxed} visitation of the root term. Here, a conjunctive term is
+visited iff its atomic subterms are visited, and zero or one of its
+immediate probabilistic subterms are visited. For a sequential
+session, the starting state corresponds to an assumption (usually
+unique) with probability $1$.\footnote{To simplify the discussion, we
+suppose that the root term of $(\pp{:init})$ lists all atoms that are
+necessarily true.} In our example, that starting state is:
 

### 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
typo

@@
 \noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root. Here, a conjunctive term is visited
-iff its atomic subterms are visited, and zero or one of its immediate
-probabilistic subterms are visited. For a sequential session, the
-starting state corresponds to an assumption (usually unique) which
-probability $1$. In our example, that starting state is:
+relaxed} visitation of the root term. Here, a conjunctive term is
+visited iff its atomic subterms are visited, and zero or one of its
+immediate probabilistic subterms are visited. For a sequential
+session, the starting state corresponds to an assumption (usually
+unique) with probability $1$.\footnote{To simplify the discussion, we
+suppose that the root term of $(\pp{:init})$ lists all atoms that are
+necessarily true.} In our example, that starting state is:
 
@@
 \noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
-relatively\footnote{Compared to rewards from the original process
-model.} small positive (negative) numeric quantity which captures the
-utility the agent receives for correctly (incorrectly) rejecting an
-assumption. A disconfirm action can only be executed once, and except in the
-applicability of that dual corresponds to a self-transformation
+relatively small positive (negative) numeric quantity\footnote{Small
+relative to rewards from the original process model.} which captures
+the utility the agent receives for correctly (incorrectly) rejecting
+an assumption.
 %%
-If \switchAction\ is the action that switched the system to a DT
-session, then an assumption $\assumptiveS{i}$ is {\em relevant} if it
-is directly or indirectly required for \switchAction\ to be
-applicable. Continuing our simplified example, if the trace is:
-
-% Furthermore, if \switchAction\ is the action that switched the system
-% to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
-% active} if
-% \switchAction\ is not applicable in the sequential model unless
-% $\assumptiveS{i}$ is scheduled in the trace prefix
-% to \switchAction. We only include the duals of active assumptions in
-% DT sessions. Continuing our simplified example, if the trace is:
+In terms of action physics, a disconfirm action can only be executed
+once, and otherwise is modelled as self-transformation.
+%%
+We only consider {\em relevant} assumptions when constructing the
+abstract model.  If \switchAction\ is the action that switched the
+system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
+relevant} if it is necessary for the outcome of \switchAction\ to be
+determined.  Continuing our simplified example, if the trace is:
 
@@
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-relevant for this session, and therefore exclude it from the
-POMDP posed to the DT session.
+relevant for this session, and therefore we exclude the corresponding
+disconfirm action from the abstract decision process.
 
-Given \switchAction, we also include another self-transition action
-$\actions.\poss(\switchAction)$, the \emph{goal action} with the
-reward property:
+Given \switchAction, we also include another once-only self-transition
+action $\actions.\poss(\switchAction)$, the \emph{confirmation action}
+with the reward property:
 
@@
 
-As are aiming to keep the decision theoretic session fast, we try to
-keep the state space of the abstract problem as small as possible. In
-a deterministic-actio POMDP, the state space size is determined by two
-factors: the number of non-zero states in the initial belief-state and
-the number of executable actions. We thus want to keep the number of
-uncertain facts as low as possible without losing important
-information about the sensing goal and keep the number of actions down
-without crippling the ability of the DT planner to gather the required
-information.
+%%  the size of the state space
+%% is determined by two factors: the number of non-zero states in the
+%% initial belief-state, and the number of actions. Determining
+%% action reachability is intractable, as is determining their relevance
+%% to goal achievement.
 
-We perform the abstraction of the belief space by first constructing
-an $(\pp{:init})$ declaration which only contains relevant assumptions
-from the trace -- e.g., (Fig.\ref{fig:abstraction-b}) gives an
-example, where diamonds are probabilistic terms, and circles are
-atomic and/or conjunctive.
+%% As we are aiming to keep the decision theoretic session fast, we try
+%% to keep the state space of the abstract problem as small as
+%% possible. In a deterministic-action POMDP, the state space size is
+%% determined by two factors: the number of non-zero states in the
+%% initial belief-state and the number of executable actions. As
+%% determining the relevance of an action for a given goal in in general
+%% a hard problem, we concentrate on keeping the number of uncertain
+%% facts as low as possible without losing important information about
+%% the sensing goal.
+
+In order to guarantee fast DT sessions, planning occurs for a small
+abstract problem. In our setting, where the underlying decision
+process has a deterministic action model, a significant cause of
+problem largeness is due to the number of states that occur with
+non-zero probability in the underlying belief-state. In posing a small
+abstract process for a DT session, we construct an abstract
+belief-state that is chiefly characterised by facts related to the
+current trace. In that abstraction, instances of operator or sense
+declarations whose specification mentions an excluded fact are not
+available.
+%%
+In detail, we first construct an $(\pp{:init})$ declaration which only
+contains relevant assumptions from the trace -- e.g.,
+(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
+probabilistic terms, and circles are atomic and/or conjunctive.
 %%
@@
 
-We then proceed to limit the actions by only including those that are
-possibly required to a) reach the goal or b) to directly observe an
-atom that is part of the abstract belief state.
-
-TODO: relaxed planning graph
-
-%% In the first step, we remove all facts that are not part of an
-%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-%% would proceed in an abstraction of the environment that does not
-%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%% %%
-%% In a second step, we iteratively refine the relaxed declaration by
-%% adding terms from the original statement of $(\pp{:init})$ while the
-%% number of abstract states in $\bstate_0$ that occur with non-zero
-%% probability according to that refined declaration remains of a
-%% practicable size. In detail,
+%%  or no more atoms that reduce the entropy
+%% can be added
 
@@
 
+% When limiting the number of actions, we aim to include those that are
+% possibly required to a) reach the goal or b) to directly observe an
+% atom that is part of the abstract belief state. In general, checking
+% this for an action is \textsc{PSPACE} complete for \textsc{STRIPS}
+% planning.
+
+% We approximate the set of relevant actions as follows: Let
+% \relevantAtoms\ be the set of atoms that either occur in the goal
+% condition \poss(\switchAction) or in a precondition \poss($a$) of an
+% action $a$ that can directly sense any atom in the abstract belief
+% state. We construct a \emph{relaxed planning graph} with
+% \relevantAtoms\ as goal atoms from which we extract a relaxed
+% plan.
+
+%% In the first step, we remove all facts that are not part of an
+%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+%% would proceed in an abstraction of the environment that does not
+%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
+%% %%
+%% In a second step, we iteratively refine the relaxed declaration by
+%% adding terms from the original statement of $(\pp{:init})$ while the
+%% number of abstract states in $\bstate_0$ that occur with non-zero
+%% probability according to that refined declaration remains of a
+%% practicable size. In detail,
+
 %%% Local Variables: 

### 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
typo

@@
 Given \switchAction, we also include another once-only self-transition
-action $\actions.\poss(\switchAction)$, the \emph{confirmation action}
+action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
 with the reward property:

### 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
few small changes

@@
 
+\vspace{-1ex}
+
 \small
@@
 \normalsize
+
+\vspace{-1ex}
 
@@
 
+\vspace{-1ex}
 \small
@@
 
+\vspace{-1ex}
+
 \noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a
-relatively small positive (negative) numeric quantity\footnote{Small
-relative to rewards from the original process model.} which captures
+small positive (negative) numeric quantity which captures
 the utility the agent receives for correctly (incorrectly) rejecting
@@
 %%
-In detail, we first construct an $(\pp{:init})$ declaration which only
-contains relevant assumptions from the trace -- e.g.,
-(Fig.\ref{fig:abstraction-b}) gives an example, where diamonds are
-probabilistic terms, and circles are atomic and/or conjunctive.
+In detail, we first construct an $(\pp{:init})$ declaration which
+contains the root term, and relevant assumptions from the trace --
+e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
+diamonds are probabilistic terms, and circles are atomic and/or
+conjunctive.
 %%

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
few small changes

@@
 probability of a false negative, {\em semi-reliable} have a chance of
-$0.3$ of false negative and $0.1$ of false positive , and {\em noisy}
+$0.3$ of false negative and $0.1$ of false positive, and {\em noisy}
 sensors with probabilities of $0.5$ and $0.2$ respectively. Each
@@
 All non-target objects remain unchanged so the planner can
-exploit object co-occurances using reliably sensed objects to
+exploit object co-occurrences using reliably sensed objects to
 compensate for the increased sensor noise.
@@
 
-We experimented with DT sessions with initial belief-states admitting
-between 20 and 100 abstract states with non-zero probability. We run
-50 simulations in each configuration, and have a timeout on each
-simulation of 30 minutes (1800 seconds)\footnote{All experiments were
-conducted on a 2.66GHz Intel Xeon X5355 using one CPU core.}. The
-continual planning times are reported in
+Our evaluation examines DT sessions with initial belief-states
+admitting between 20 and 100 abstract states with non-zero
+probability. We run 50 simulations in each configuration, and have a
+timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
+experiments were conducted on a 2.66GHz Intel Xeon X5355 using one CPU
+core.}. The continual planning times are reported in
 Figure~\ref{fig:results-time}, and the quality data in
@@
 % so is omitted from that reporting.  
-Because reward is only allocated on the achievement of a goal, we
-find it intuitive to report average plan costs and the success rates
-in problems that admit a solution (i.e., positive reward scaled by a
-constant factor).
+In these experiments we only allocate reward on the achievement of all
+goals, therefore we find it intuitive to report average plan costs and
+the success rates in problems that admit a complete solution (i.e.,
+positive reward scaled by a constant factor). The exception occurs for
+items f and g of Figure~\ref{fig:results-quality}, where we report the
+overall success rate of the respective approaches.
 
@@
 Speaking to the effectiveness of our entropy heuristic for abstraction
-refinement, we witness relatively high success rates irrespective of
-the exact level of refinement. Comparing finally to the best {\sc
+refinement, we see relatively high success rates irrespective of
+the level of refinement. Comparing finally to the best {\sc
 zmdp} policy, although producing relatively costly plans, the

### 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
was a spacing problem

--- 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
@@
-\vspace{-1ex}
+%\vspace{-1ex}
 
 For use in an integrated continual planning and execution monitoring
-system on a mobile robot platform, we developed a base planner that
-switches between fast sequential and expensive DT planning. Given a
-DTPDDL model of the environment, sequential planning is used to
-compute an initial deterministic sequential plan and complementary
-runtime evolution of the decision process. That plan is executed until
-testing of an assumption about the runtime state is scheduled. At that
-point a DT session performs some sensing which determines whether or
-not the sequential session should be allowed to continue, or otherwise
-acts to achieve the overall objectives. We find our approach performs
-quickly and relatively efficiently in a number of large real-world
-task and observation planning problems.
+system on a mobile robot platform, we developed a planner that
+switches between: (a) fast sequential planning, and (b) expensive DT
+planning in small abstractions of the problem at hand. Given a DTPDDL
+model of the environment, sequential planning is used to compute an
+initial sequential plan and complementary runtime evolution of the
+decision process. That plan is executed until the outcome of the next
+action scheduled for execution is too uncertain. At that point a DT
+session performs sensing that can help determine how the sequential
+plan might evolve, or otherwise acts to achieve the goals. We find our
+approach performs quickly and relatively efficiently in a number of
+large real-world task and observation planning problems.
 

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
added footnote

--- 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
@@
 \noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root term. Here, a conjunctive term is visited
-iff its atomic subterms are visited, and zero or one of its immediate
-probabilistic subterms are visited. For a sequential session, the
-starting state corresponds to an assumption (usually unique) with
-probability $1$. In our example, that starting state is:
+relaxed} visitation of the root term. Here, a conjunctive term is
+visited iff its atomic subterms are visited, and zero or one of its
+immediate probabilistic subterms are visited. For a sequential
+session, the starting state corresponds to an assumption (usually
+unique) with probability $1$.\footnote{To simplify the discussion, we
+suppose that conjunctive terms include all atoms that are true with
+probability 1 according to their visitation.} In our example, that starting state is:
+
+\vspace{-1ex}
 
@@
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-relevant for this session, and therefore exclude it from the
-abstract decision process.
+relevant for this session, and therefore we exclude the corresponding
+disconfirm action from the abstract decision process.
 
 Given \switchAction, we also include another once-only self-transition
-action $\actions.\poss(\switchAction)$, the \emph{confirmation action}
+action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
 with the reward property:

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
worry gone

@@
 session, the starting state corresponds to an assumption (usually
-unique) with probability $1$.\footnote{To simplify the discussion, we
-suppose that conjunctive terms include all atoms that are true with
-probability 1 according to their visitation.} In our example, that starting state is:
+unique) with probability $1$.\footnote{Without a loss of generality,
+we assume conjunctive terms list atoms that are true with
+probability 1 according to a visitation, and that actions do not
+have negative preconditions.}  In our example, that starting state is:
 

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
active is no longer a word

@@
 whose execution can provide evidence regarding the truth value of an
-active assumption.
+{\em relevant} assumption.
 %%

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
active is no longer a word

@@
 In terms of action physics, a disconfirm action can only be executed
-once, and otherwise is modelled as self-transformation.
+once, and otherwise is modelled as a self-transformation.
 %%
@@
 %%
-For each excluded atom, we compute the {\em entropy} of the active
+For each excluded atom, we compute the {\em entropy} of the {\em
+relevant}
 assumptions, {\em conditional} on that atom. Intuitively, lower

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
grammar

--- 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
@@
 execution the DT session applies a single entropy reduction action,
-whose execution can provide evidence regarding the truth value of an
-active assumption.
+whose execution can provide evidence regarding the truth value of a
+{\em relevant} assumption.
 %%

### 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
oooops

@@
 the success rates in problems that admit a complete solution (i.e.,
-positive reward scaled by a constant factor).
+positive reward scaled by a constant factor). The exception occurs for
+items f and g of Figure~\ref{fig:results-quality}, where we report
+expected discounted rewards (not plan costs).
 

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
new conclusion

--- 2011-02-07 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
@@
-\vspace{-1ex}
+%\vspace{-1ex}
 
-For use in an integrated continual planning and execution monitoring
-system on a mobile robot platform, we developed a planner that
-switches between fast sequential and expensive DT planning. Given a
-DTPDDL model of the environment, sequential planning is used to
-compute an initial deterministic sequential plan and complementary
-runtime evolution of the decision process. That plan is executed until
-testing of an assumption about the runtime state is scheduled. At that
-point a DT session performs some sensing which determines whether or
-not the sequential session should be allowed to continue, or otherwise
-acts to achieve the overall objectives. We find our approach performs
-quickly and relatively efficiently in a number of large real-world
-task and observation planning problems.
+We have addressed a key challenge, specifically that of high-level
+continual planning and execution monitoring for efficient
+deliberations according to rich probabilistic models afforded by
+recent integrated robotic systems. We developed a system that can plan
+quickly given large realistic probabilistic models, by switching
+between: (a) fast sequential planning, and (b) expensive DT planning
+in small abstractions of the problem at hand. Sequential and DT
+planning is interleaved, the former identifying a rewarding sequential
+plan in the underlying process, and the latter solving small sensing
+problems posed during sequential plan execution.
+%%
+We have evaluated our system in large real-world task and
+observation planning problems, finding that it performs quickly
+and relatively efficiently.
+
+%% We developed a continual planning and execution monitoring
+%% system for use on a mobile robot platform. .
+%% %%
+%% Our planner switches between: (a) fast sequential planning, and (b)
+%% expensive DT planning in small abstractions of the problem at
+%% hand. Our approach addresses a key challenge put to the community, by
+%% Planning for a DTPDDL model of the environment, sequential and DT
+%% sessions are interleaved, the former identifying a rewarding plan, and
+%% the latter solving practically sized sensing problems posed during
+%% serial plan executions execution.
+%% %%
+%% Our system answers a key challenge, with {\em probabilistic} --i.e., {\em quantitative}--
+%% models of noisy sensing and state.
+
+%% is used to compute an
+%% sequential plans and complementary runtime evolutions of the
+%% decision process. That plans are executed until the outcome of the next
+%% action scheduled for execution is too uncertain. At that point a DT
+%% session performs sensing that can help determine how the sequential
+%% plan might evolve, or otherwise acts to achieve the goals. We find our
+%% approach performs quickly and relatively efficiently in a number of
+%% large real-world task and observation planning problems.
 

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
Fix the title in the pdfinfo

@@
 \pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
+/Title (A Switching Planner for Combined Task and Observation Planning)
+% /Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
 /Author (NA)}
@@
 
-\section{Background and Related Work}
-\input{relatex.tex}
-
 \section{Planning Language and Notations}

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
new evaluation text from RIchard

@@
 Downward~\cite{fast-downward}, and DT sessions use our own contingent
-procedure. We also implemented a simple dual-mode replanning {\em
-baseline} approach. Here, when a switching action is scheduled for
-execution the DT session applies a single entropy reduction action,
-whose execution can provide evidence regarding the truth value of a
-{\em relevant} assumption.
+procedure. Since most of the problems we consider are much larger than
+any available DT planner can solve, for comparison purposes we also
+implemented a simple dual-mode replanning {\em baseline}
+approach. Here, when a switching action is scheduled for execution the
+DT session applies a single entropy reduction action, whose execution
+can provide evidence regarding the truth value of a {\em relevant}
+assumption.
 %%
@@
 positive reward scaled by a constant factor). The exception occurs for
-items f and g of Figure~\ref{fig:results-quality}, where we report the
-overall success rate of the respective approaches.
+items f and g of Figure~\ref{fig:results-quality}, where we report
+expected discounted rewards (not plan costs).
 

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/dtpddl.tex
Typos and minor edits in a number of places. Switching continual planner section clarified and shortened.

@@
 \noindent I.e., there is a 10\% {\em false positive} rate, and 20\%
-  probability of a {\em false negative}.
+  probability of a {\em false negative}. This representation allows us to represent actions that have multiple independent observational effects.
 
@@
 subterms, $T_i$, is visited. Each visitation of the root term
-according to this recursive definition encapsulates a starting state,
+according to this recursive definition defines a starting state,
 along with the probability that it occurs. The former corresponds to the

### 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
Typos and minor edits in a number of places. Switching continual planner section clarified and shortened.

@@
 procedure. Since most of the problems we consider are much larger than
-any available DT planner can solve, for comparison purposes we also
-implemented a simple dual-mode replanning {\em baseline}
+any available DT planner can solve directly, for comparison purposes
+we also implemented a simple dual-mode replanning baseline
 approach. Here, when a switching action is scheduled for execution the
 DT session applies a single entropy reduction action, whose execution
-can provide evidence regarding the truth value of a {\em relevant}
+can provide evidence regarding the truth value of a relevant
 assumption.
@@
 topological places. Objects indicate the category of space they
-inhabit -- e.g., spoons are likely to be in kitchens. By examining
+inhabit---e.g., spoons are likely to be in kitchens. By examining
 view cones at places for particular objects, the robot is able to: (1)
@@
 
-We compare switching to the {\em baseline} in several realistic
+We compare switching to the baseline in several realistic
 tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
@@
 sensors with probabilities of $0.5$ and $0.2$ respectively. Each
-object class is assigned one sensor model -- e.g. cornflakes may be
+object class is assigned one sensor model---e.g. cornflakes may be
 harder to detect than refrigerators. We performed several experiments
@@
 We find that if sensing is reliable, then little is gained using DT
-sessions, as the greedy approach of the {\em baseline} is
+sessions, as the greedy approach of the baseline is
 sufficient. As sensing degrades DT sessions prove more useful. Here,

### 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Typos and minor edits in a number of places. Switching continual planner section clarified and shortened.

@@
 We now describe our {\em switching} planning system that operates
-according to the continual planning paradigm, given a description of
+according to the continual planning paradigm given a description of
 the current problem and domain in DTPDDL. The system {\em
-switches}, in the sense that planning proceeds in interleaved
-sessions, in which the base planner is either {\em sequential}/linear
+switches} in the sense that planning proceeds in interleaved
+sessions in which the base planner is either {\em sequential}/linear
 or {\em decision-theoretic}.
 %%
-During a sequential session, a rewarding {\em trace} of a possible
-execution is computed using a cost-optimising satisficing planner
-which trades action costs, goal rewards, and determinacy.
-%%
-Formatted as a linear plan, the trace specifies a sequence of actions
-that achieves some reachable goals following a deterministic
-approximation of the problem at hand.
-%%
-Structurally, a trace is a sequence of elements that are either: (i) ground
-actions from the DTPDDL description of the world, or (ii) atomic {\em
-assumptions}, modelled as deterministic actions, made about the truth
-value of facts that can only be determined at runtime (e.g., that a
-box of cornflakes is located on the corner bench in the kitchen).
-%%
-The system always begins in a sequential session, so that plan
-execution proceeds by applying DTPDDL actions from the trace, in
-sequence, until the outcome of executing the next scheduled action is
-too uncertain according to a threshold parameter~(here, 95\%). In that
-eventuality a DT session commences, which tailors sensory processing
-to determine whether the assumptions made in the trace hold, or which
-otherwise acts to achieve the goals.
+The system begins in a sequential session during which it runs a
+cost-based satisficing planner which trades off action costs, goal
+rewards and determinacy. This planner generates a reward-giving
+sequence of ground actions from the DTPDDL domain which may also
+include {\em assumptions}: actions that assume the truth of some fact
+that is unknown at plan time (e.g. that a box of cornflakes is located
+on the corner bench in the kitchen). We call this sequence a {\em
+trace}. Execution of the trace proceeds by applying the actions until
+the probability of the next action being applicable falls below some
+threshold (here 95\%), at which point the planner switches to a DT session.
 
-Because online DT planning in large problems is impractical (we seek
-response times in seconds), a DT session plans for an abstract process
-determined by the current trace and underlying belief-state. That
-abstract problem is characterised by a limited number of propositions,
+Because online DT planning is impractical for the size of problem we
+are interested in, when a DT session begins we use the trace from the
+preceeding sequential session to restrict the state space. The state space is characterised by a limited number of propositions,
 chosen due to their {\em relevance} to the action whose scheduled
 execution triggered the switch to the DT session.
-%%the subproblem the DT session is asked tosolve.
-% MOG: Describe this in detail in the appropriate subsection
-%  This
-% abstraction is constructed by first excluding all propositions (and
-% related actions) that are not true, or {\em assumed } true, of states
-% in the trace, then adding them back, using as an heuristic the entropy
-% of the trace assumptions conditional on a candidate
-% proposition. Propositions are added, one at a time, until the number
-% of states in the initial belief-state reaches a given
-% threshold.\footnote{Our experiments are performed using a number of
-% different threshold parameters.} 
-To that abstract model we also add {\em disconfirm} and {\em confirm}
-actions that the DT session can schedule in order to judge an atomic
-assumption in the trace. Those actions yield a relatively small reward
-if the corresponding judgement is true (or small penalty
-otherwise). If a judgement action is scheduled for execution the DT
-session is terminated, and control is returned to a new sequential
-session.
+%The state
+%space is constructed by first removing all propositions (and related
+%actions) that are not true or assumed true of states in the trace,
+%then adding them back as long as the number of states does not exceed
+%a size threshold\footnote{Our experiments are performed using a number
+%of different threshold parameters.}  Propositions to add back are
+%chosen using as a heuristic the entropy of the trace assumptions
+%conditional on the candidate proposition. 
+To allow the DT planner to
+determine the truth of the assumptions in the trace, we add {\em
+disconfirm} and {\em confirm} actions to the problem for each of
+them. These actions yield a relatively small reward/penalty if the
+corresponding judgement is true/false. If a judgement action is
+scheduled for execution the DT session is terminated, and control is
+returned to a sequential session.
+
+%% We now describe our {\em switching} planning system that operates
+%% according to the continual planning paradigm, given a description of
+%% the current problem and domain in DTPDDL. The system {\em
+%% switches}, in the sense that planning proceeds in interleaved
+%% sessions, in which the base planner is either {\em sequential}/linear
+%% or {\em decision-theoretic}.
+%% %%
+%% During a sequential session, a rewarding {\em trace} of a possible
+%% execution is computed using a cost-optimising satisficing planner
+%% which trades action costs, goal rewards, and determinacy.
+%% %%
+%% Formatted as a linear plan, the trace specifies a sequence of actions
+%% that achieves some reachable goals following a deterministic
+%% approximation of the problem at hand.
+%% %%
+%% Structurally, a trace is a sequence of elements that are either: (i) ground
+%% actions from the DTPDDL description of the world, or (ii) atomic {\em
+%% assumptions}, modelled as deterministic actions, made about the truth
+%% value of facts that can only be determined at runtime (e.g., that a
+%% box of cornflakes is located on the corner bench in the kitchen).
+%% %%
+%% The system always begins in a sequential session, so that plan
+%% execution proceeds by applying DTPDDL actions from the trace, in
+%% sequence, until the outcome of executing the next scheduled action is
+%% too uncertain according to a threshold parameter~(here, 95\%). In that
+%% eventuality a DT session commences, which tailors sensory processing
+%% to determine whether the assumptions made in the trace hold, or which
+%% otherwise acts to achieve the goals.
+
+%% Because online DT planning in large problems is impractical (we seek
+%% response times in seconds), a DT session plans for an abstract process
+%% determined by the current trace and underlying belief-state. That
+%% abstract problem is characterised by a limited number of propositions,
+%% chosen due to their {\em relevance} to the action whose scheduled
+%% execution triggered the switch to the DT session.
+%% %%the subproblem the DT session is asked tosolve.
+%% % MOG: Describe this in detail in the appropriate subsection
+%% %  This
+%% % abstraction is constructed by first excluding all propositions (and
+%% % related actions) that are not true, or {\em assumed } true, of states
+%% % in the trace, then adding them back, using as an heuristic the entropy
+%% % of the trace assumptions conditional on a candidate
+%% % proposition. Propositions are added, one at a time, until the number
+%% % of states in the initial belief-state reaches a given
+%% % threshold.\footnote{Our experiments are performed using a number of
+%% % different threshold parameters.} 
+%% To that abstract model we also add {\em disconfirm} and {\em confirm}
+%% actions that the DT session can schedule in order to judge an atomic
+%% assumption in the trace. Those actions yield a relatively small reward
+%% if the corresponding judgement is true (or small penalty
+%% otherwise). If a judgement action is scheduled for execution the DT
+%% session is terminated, and control is returned to a new sequential
+%% session.
 
 Whatever the session type, our continual planner maintains a factored
-representation of successive belief-states by performing belief
-revision. As an internal representation of the $(\pp{:init})$
+representation of successive belief-states.
+% by performing belief revision. 
+As an internal representation of the $(\pp{:init})$
 declaration, we keep a tree-shaped Bayesian network which gets updated
@@
 session, the starting state corresponds to an assumption (usually
-unique) with probability $1$.\footnote{To simplify the discussion, we
-suppose that the root term of $(\pp{:init})$ lists all atoms that are
-necessarily true.} In our example, that starting state is:
+unique) with probability $1$.\footnote{Without a loss of generality,
+we assume conjunctive terms list atoms that are true with
+probability 1 according to a visitation, and that actions do not
+have negative preconditions.}  In our example, that starting state is:
 
@@
 
-Encapsulating state-assumptions, we annotate the problem posed during
+To represent state assumptions we augment the problem posed during
 a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
@@
 We now describe the optimisation criteria used during sequential
-sessions. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action, $\action_i$, from a trace of state-action pairs
+sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
+action, $\action_i$, from a trace of state-action pairs
 $\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
-does not transition to $\state^\bot$, we have that the optimal trace
-has value:
+does not transition to $\state^\bot$, and
+$\reward(\state_i, \action_i)$ be the instantaneous reward received
+for executing action $\action_i$ in state $\state_i$. The optimal
+trace has value:
+
+%We now describe the optimisation criteria used during sequential
+%sessions. Where $\prob_i$ is the probability that the $i^{th}$
+%sequenced action, $\action_i$, from a trace of state-action pairs
+%$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
+%does not transition to $\state^\bot$, we have that the optimal trace
+%has value:
 
@@
 
-\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
-reward ---i.e., can be negative for action costs, or positive for goal
-achievement--- received for executing action $\action_i$ in state
-$\state_i$. Finally, it is worth clarifying that in practise we do not
-artificially limit the length of traces that the sequential planner
-can consider. Moreover, for problems we consider there is always a
-finite optimal trace.
+%\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
+%reward ---i.e., can be negative for action costs, or positive for goal
+%achievement--- received for executing action $\action_i$ in state
+%$\state_i$. Finally, it is worth clarifying that in practise we do not
+%artificially limit the length of traces that the sequential planner
+%can consider. Moreover, for problems we consider there is always a
+%finite optimal trace.
 
@@
 In detail, we first construct an $(\pp{:init})$ declaration which
-contains the root term, and relevant assumptions from the trace --
-e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
+contains the root term, and relevant assumptions from the trace---e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
 diamonds are probabilistic terms, and circles are atomic and/or

### 2011-02-16 code/subarchitectures/dialogue.sa/branches/spring-school-2011/resources/grammars/openccg/prosodic_moloko.v6.0.1/docs/comsys.moloko/event-modifier-restriction.moloko.tex
branch for 2011 spring school

--- 2010-07-20 code/subarchitectures/dialogue.sa/trunk/resources/grammars/openccg/moloko.v6/docs/comsys.moloko/event-modifier.tex
@@
- \section{Event Modifier Restriction} \label{sec-Event-Modifier-Restriction}
-We mentioned earlier that the Moloko is quite promiscous when it comes to  semantic or pragmatic 'un-acceptability', notions which must be handled grammar externally.
+ \section{Event Modifier Restriction}
+As discussed in () above, the Moloko is quite promiscous when it comes to  semantic or pragmatic 'un-acceptability', notions which must be handled grammar externally.
 There is, however, one major exception to this design principle: verbs are able to lexically specify what kinds of event modifiers they can recieve.
@@
 
-The restriction of event modifiers is handled almost identically to the way the combinatorial behaviour of modifiers was handled (see \ref{sec-Dual-Relation-Words} above).
-It is done through the use of
+In fact, the restriction of event modifiers is handled almost identically to the way the combinatorial behaviour of modifiers was handled (see .. above). It is done through the use of 
 \begin{enumerate}
@@
 \item  the inclusion of these features within the syntactic category responsible for 'attaching' the  modifier to its modifee.
-This includes the type-changing rules used to handle  the modifier function of adverbs and prepositions (XXXXX), and also subordinate clause modifiers.\footnote{although event modifier restriction is surely a matter of semantics and not syntax, there are two reasons why it is better handled via feature.
+This includes the type-changing rules used to handle  the modifier function of adverbs and prepositions (), and also subordinate clause modifiers.\footnote{although event modifier restriction is surely a matter of semantics and not syntax, there are two reasons why it is better handled via feature.
 First, OpenCCG  much better at 'accessing' syntactic features than semantic ones.
 In fact, once semantic content has been added the semantic representation of a parse, it is no longer 'visible' to the parsing at all.
-Second, as discussed above (XXXXX), one of the design features for Moloko was to keep the semantic sorts as grammar external as possible.
+Second, as discussed above (), one of the design features for Moloko was to keep the semantic sorts as grammar external as possible.
 Using the event-type to control modification would violate this.}

### 2011-02-16 code/subarchitectures/dialogue.sa/branches/spring-school-2011/apis/dialogue/src/java/de/dfki/lt/tr/dialogue/cplan/RuleTracer.java

@@ -1,0 +1,16 @@
+package de.dfki.lt.tr.dialogue.cplan;
+
+public interface RuleTracer {
+  public static final int DISPLAY_MATCHING = 1;
+  public static final int DISPLAY_MODIFICATION = DISPLAY_MATCHING << 1;
+
+  public abstract void setTracing(int bitmask);
+
+  public abstract void traceMatch(DagEdge current, Rule r,
+      Bindings bindings);
+
+  public abstract void traceAfterApplication(DagEdge current, Rule r);
+
+  public abstract void traceBeforeApplication(DagEdge current, Rule r);
+
+}

### 2011-03-09 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/dtpddl.tex
Our workshop attempt.

--- 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
-%%%% ijcai11.tex
-
-\typeout{IJCAI Switching Planner Submission}
-
-% These are the instructions for authors for IJCAI-11.
-% They are the same as the ones for IJCAI-07 with superficical wording
-%   changes only.
-
-\documentclass{article}
-% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
-\usepackage{ijcai11}
-
-% Use the postscript times font!
-\usepackage{times}
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\usepackage{subfig}
-\usepackage{tikz}
-\usetikzlibrary{shapes}
-\usetikzlibrary{positioning}
-
-\usepackage{graphicx}
-\graphicspath{{plots/}}
-\include{macros}%
-
-\title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Planning for agents acting in partially observable environments with
-stochastic actions is extremely computationally
-expensive~\cite{mdp-complexity}. Even relatively small partially
-observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, have state spaces far larger than existing
-optimal or close to optimal planners can handle. For the easier
-problem of planning in completely observable stochastic domains, an
-alternative has been to use classical sequential planners and
-replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
-Execution monitoring tracks properties of the state that determine
-whether further planning is required. For partially observable
-domains, only probabilistic information about such properties will, in
-general, be available.
-
-%% However, these approaches rely
-%% crucially on being able to determine properties of the current state
-%% in order to decide to replan. For partially observable domains, only
-%% probabilistic information about these properties will in general be
-%% available.
-
-This paper describes the planning component for a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings. This domain features partial observability, particularly
-because the state is not perfectly known, and state information is
-gained by the robot through the use of unreliable sensing
-actions. 
-%Using a propositionally factored state representation, for
-For
-interesting sized tasks the corresponding POMDP model is too large for a
-POMDP solver to be applied directly. Inspired by the replanning
-approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
-planning approach that uses a classical planning system to compute a
-reasonably valuable trace in the model, and then uses a
-decision-theoretic planner on small subproblems where reasoning about
-observations might be useful. We refer to the complete system as a
-{\em switching continual planner}. The approach is not optimal,
-particularly as it relies on the results of sequential planning
-directly. It does nevertheless perform better than a purely sequential
-replanner, and is fast enough to be used for real-time decision-making
-on a mobile robot.
-
-%% Switching gives us some important advantages in our robotic
-%% domain. First, the significant source of uncertainty in the domain is
-%% the unreliability of observations, and this approach is particularly
-%% suited to problems that combine a deterministic task planning problem
-%% with observations as the decision-theoretic planner is often invoked
-%% to determine a characteristic of the underlying state. Secondly,
-%% replanning makes our system robust to changing objectives and the
-%% discovery of new facts about the world.
-
-%% As an illustration, consider a robot searching for a box that it knows
-%% is often found in offices. The sequential planner might produce a plan
-%% that assumes that {\sc Room a} is an office, that the box is in {\sc
-%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
-%% these assumptions, the plan might be to go to {\sc Room a}, go to
-%% location 1 and search for the box. When the plan gets executed and the
-%% search begins the decision-theoretic planner is called since searching
-%% involves making observations. The decision-theoretic planner might
-%% retract the last assumption but leave the others, and build a plan to
-%% search the whole of {\sc Room a} for the box. If this plan fails to
-%% find the box, it will disprove the assumption that the box is in {\sc
-%% Room a}, which will then lead to replanning by the sequential planner
-%% based on the newly discovered knowledge.
-
-Our system is domain independent, taking domain and problem
-descriptions in a first-order declarative language we have developed,
-called the decision-theoretic planning domain definition language
-(DTPDDL). From these descriptions of POMDPs, we automatically
-construct a deterministic model suitable for sequential planners. That
-model includes actions which correspond to making assumptions about
-the values of imperfectly known state variables. Assumptions scheduled
-by the sequential system are used to propose a pragmatic abstract
-belief-state space to the decision-theoretic system, and to modify the
-reward function, so that system might pursue sensing related to those
-assumptions.
-
-%%  which can operate
-%% with the original representation of the domain.
-
-%% The sequential planner we use is based
-%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
-%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
-%% knowledge to be represented explicitly, so we can write actions that
-%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
-%% THIS. For the decision-theoretic planner we have implemented our own
-%% forward search in the belief-state space.
-
-The switching planner provides important advantages in our
-mobile robot domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations because the decision-theoretic planner
-is only invoked to deal with state uncertainty.
-%determine a characteristic of the underlying state. 
-Secondly, replanning makes the system robust to
-changing objectives and discoveries about the world,
-both of which feature in our domain.
-
-The remainder of the paper proceeds as follows. We begin by describing
-POMDPs with a propositionally factored representation of states and
-observations. We also outline the relationship between this and a flat
-POMDP representation. Next we describe our DTPDDL language with
-example declarations from a mobile robot exploration task. Then we
-describe the switching planner, including how the sequential domain is
-produced and a plan is generated, and how this then provides the input
-for the decision-theoretic planner. Finally we provide an evaluation
-of the system on the example robot domain, and finish with a short
-survey of related work and some concluding remarks.
-
-%To accompany
-%that language we have implemented an information-state
-%\laostar\ procedure for solving problems expressed in DTPDDL.
-
-%%
-
-%Have a good understanding of the contents and function of rooms, as
-%well as the linguistic referrents to rooms, widgets, and their visual
-%qualities.
-
-%Having confidence in its beliefs about the linguistic terms for
-%spaces, and the function of those spaces.
-
-%sometimes difficult to predict, with exogenous events, such as the
-%changing of the objective,
-
-%the speed and scalability of software for sequential planning in
-%deterministic
-
-%% We developed a first-order declarative language, called DTPDDL, for
-%% describing domains of planning problems that correspond to POMDPs.  To
-%% accompany that language we have implemented an information-state
-%% \laostar\ procedure for solving problems expressed in DTPDDL.
-
-%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
-%% modified \fastdownward\ so that it can find useful sequential executions
-%% given DTPDDL models of the problem at hand.
-
-%%  called DTPDDL, along
-%% the lines of PDDL for the partially observable case, an \laostar\
-%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
-%% and modify the \fastdownward\ system to find high quality sequential plans
-%% .
-
-%% We model the environment as a partially observable Markov decision
-%% process. Although planning in that model is undecidable in
-%% general~\cite{}, an optimal finite-horizon plan corresponds to a
-%% contingent plan, that is a function mapping action-observation
-%% histories to actions.
-
-%% our continual planner is reactive, replanning whenever the underlying
-%% domain and problem models change. For example, replanning occurs if
-%% the motivational component alters the objectives, and if an assumed
-%% outcome of a sensing action is not realised.
-
-%% brittle sensing model,
-
-%% the evaluation of a fluent at a state is either known. Moreover, there
-%% is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-\section{Propositional Decision-Theoretic Planning}
-
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
-We describe the partially observable propositional probabilistic
-planning problem, with costs and rewards. We model a process state
-$\state$ as the set of propositions that are true of the
-state. Notationally, $\props$ is the set of propositions, $\prop$ is
-an element from that set, and we have $\state \subseteq
-\props$. The underlying process
-dynamics are modelled in terms of a finite set of {\em probabilistic
-STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
-state-characterising propositions $\propositions$.
-%%
-We say an action $\stochAction \in \stochActions$ is
-applicable if its precondition $\poss(\stochAction)$, a set of
-propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
-\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
-nature chooses a deterministic STRIPS effect $\detAct_i$, and for
-all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
-\mu_{\stochAction}(\detAction_i) = 1$.
-
-%% We model a process state $\state$ as the set of propositions that are
-%% true of the state. Notationally, we have $\state \subseteq
-%% \props$. State change is induced by application of actions. A
-%% stochastic action $\stochAction \in \stochActions$ is applicable if
-%% its precondition $\poss(\stochAction)$, a set of propositions, are
-%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
-%% applicable at state $\state$.  If action $\stochAction \in
-%% \stochActions(\state)$ is applied at \state, nature chooses one
-%% element amongst a small set of deterministic outcomes
-%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-%% $\sum_{\detAction_i \in \detActions(\stochAction)}
-%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-%% effect on the state given in terms of two lists of propositions. The
-%% add-list $\add(\detAction)$ and delete-list
-%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-%% $\detAction$, then it cannot be in the delete-list and vice versa.}
-%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
-%% applied at state $\state$, then the resultant state is $ ( \state \cup
-%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-%% propositions from $\add(\detAction)$ are added to $\state$, and those
-%% from $\delete(\detAction)$ are removed from $\state$.
-
-%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
-
-We are concerned with problems that feature partial
-observability. Although we could invoke {\em extended probabilistic
-STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for presentation and computation to
-separate sensing and action. Therefore, we suppose a POMDP has a
-perceptual model given in terms of a finite set of stochastic {\em
-senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
-and perceptual propositions $\percepts$, called {\em percepts}. In
-detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote by \observations\ the set of
-reachable observations. The underlying state of the process cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in
-\observations$ that informs what should be believed about the state the
-process is in. If $\action$ is applied effecting a
-transition to a successor state $\state'$, then an observation occurs
-according to the active senses $\stochSenses(\stochAction, \state')
-\subseteq \stochSenses$. A sense $\stochSense$ is active, written
-$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
-action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
-$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
-usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
-%%
-When a sense is active, nature must choose exactly one outcome amongst
-a small set of deterministic choices $\detSenses(\stochSense)
-\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
-$i$ we have $\detSense_i \subseteq \percepts$. The probability of
-the $i^{th}$ element being chosen is given by
-$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
-\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
-1$. The observation received by the agent corresponds to the union of
-perceptual propositions from the chosen elements of active
-senses.
-
-A POMDP has a starting configuration that corresponds to a Bayesian
-belief-state. Intuitively, this is the robot's subjective belief about
-its environment. Formally, a belief-state $\bstate$ is a probability
-distribution over process states. We write $\bstate(\state)$ to denote
-the probability that the process is in $\state$ according to
-$\bstate$, and $\bstate_0$ when discussing the starting
-configuration.
-
-Finally, we make some fairly standard assumptions. First, that action
-execution and sensing occur instantaneously, and that only one action
-can be applied per plan-step. Secondly, it can arise that in some
-states an action is applicable, and in others it is not. Moreover, a
-belief-state can assign non-zero probability to states where that
-applicability holds, and states where it does not. %% We differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
-%% In beliefs where an action exists that is applicable in some state 
-%% with non-zero probability, and not applicable in another, we differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%% because 
-%% in that
-%% we forbid the execution of the action.
-%% actions that are not applicable in
-%% any state of the current belief.  
-%% Also,
-Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we incorporate a PPDDL-like default semantics, treating an action
-\stochAction\ executed at $\state_i$ as if it has no effect at
-states $\state$ if $\poss(\stochAction) \not\subseteq
-\state$.
-
-% Mucked around with this, then realised the two sentences were
-% contradictory. If we forbid actions applicable in one state but not
-% another in the belief state, then clearly we aren't using a
-% PPDDL-like semantics.
-%% Doubled comments are stuff I added, then removed.
-
-\subsection{Costs, Rewards, and Belief Revision}
-
-Until now we have discussed the POMDP in terms of propositions and
-percepts. In order to address belief revision and utility it is
-convenient to consider the underlying decision process in a flat
-format. This is given by the tuple
-$\langle \states, \bstate_0, \actions, \transProb, \reward,
-\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
-belief-state, \states\ is the finite set of reachable propositional
-states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
-\state')$ giving the probability of a transition from state $s$ to
-$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
-$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
-%%
-The function $\reward:\states \times \actions \to \Re$ is a bounded real
-valued reward function. Therefore a finite positive constant $c$
-exists so that for all $\state \in \states$ and $\action \in
-\actions$, $|\reward(\state, \action)| < c$. We model costs as
-negative rewards.
-%%
-From $\psi$ we have that for each $\state' \in \states$ and action
-$\action \in \actions$, an observation $\observ \in \observations$ is
-generated independently according to a probability distribution
-$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
-\action)$ the probability of getting observation $\observ$ in state
-$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
-\observations} \obsDist_\observ(\state', \action) = 1$.
-
-Successive state estimation is by application of Bayes'
-rule.  Taking the current belief $\bstate$ as the {\em prior}, and
-supposing action $\action$ is executed with perceptive outcome
-$\observ$, the probability that we are in $\state'$ in the
-successive belief-state $\bstate'$ is:
-
-\begin{equation}\label{eq:revision}
-\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
-  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
-\end{equation}
-
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
-
-\subsection{Plan Evaluation}
-
-An optimal solution to a finite-horizon POMDP is a contingent plan,
-and can be expressed as a mapping from observation histories to
-actions. Although suboptimal in general, useful plans can also take a
-classical sequential format. This is the case in {\em conformant}
-planning, where the objective is to find a sequence of actions that
-achieves a goal ---i.e., reaches a state that satisfies a given
-Boolean condition--- with probability $1$.  Generally, whatever the
-structure of a plan, its value corresponds to the expected cumulative
-reward:
-
-%% More generally, a useful
-%% formalism for representing solutions to POMDPs is the finite-state
-%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
-%% \transition \rangle$ where: $\node \in \nodes$ is a set of controller
-%% states, $\selection_\node(\action) = P(\action | \node)$ gives the
-%% probability that the controller prescribes \action\ when in state
-%% \node, and $\transition_\node(\action, \observation, \node') =
-%% P(\node'|\node, \action, \observation)$ gives the probability the
-%% controller transitions to state $\node'$, supposing we execute
-%% \action\ in state \node\ receiving observation \observation. When
-%% acting for finite $N$ steps from controller state $n$, the value of a
-%% controller corresponds to the expected cumulative reward:
-
-\begin{equation}\label{eq:expectedvalue}
-V_{\pp{PLAN}}(\bstate) = \Expect \bigg{[} 
-\sum_{t=0}^{N-1}  \reward(\bstate_t, \pp{PLAN}_t) \mid \pp{PLAN}, \bstate_0 = \bstate \bigg{]}
-\end{equation}
-
-\noindent Where $b_t$ is the belief-state at step $t$, $\pp{PLAN}_t$ is
-the action prescribed at step $t$, and
-
-\[\reward(\bstate, \action) = \sum_{\state \in \states}
-\bstate(\state)\reward(\state, \action).\]
-
-%% \noindent Finally, it is
-%% useful to note that there is a corresponding deterministic FSCs for
-%% both sequential (e.g., conformant) and contingent plan formats.
-
-\section{Decision-Theoretic PDDL}
-
 The modelling language of choice for planning in probabilistic
@@
 actions.
-
-\section{Switching Continual Planner}
-
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
-
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
-%%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
-
-Finally, whether proceeding in a sequential or contingent session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
-
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
-
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
-
-\subsection{Sequential Planning in POMDPs}
-
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
-
-\noindent An assumption corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
-
-\[
-\begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
-\end{array}
-\]
-
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
-%%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
-%%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
-
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
-
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
-
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
-
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
-\[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
-\reward(\state_i, \action_i),
-\]
-
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
-
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
-\[
-\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
-\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\hat\$(T_i) & \pp{otherwise} \\
-\end{array}
-\]
-
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
-%%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
-
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
-\end{array}
-\]
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
-
-We also include another propositionally trivial action
-$\actions.\poss(\switchAction)$ with the reward property:
-
-\[
-\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
-\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
-\end{array}
-\]
-
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
-
-%%  We make no assumptions here about how they are
-%% derived.
-
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
-
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief-state with the assumptions made by the continual
-    planner in bold.]{
-      \label{fig:abstraction-a}
-      \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode] (cup) {} 
-          child {node [cnode] (cupp1) {}}
-          child {node [cnode] (cupp2) {}}
-        }
-      }
-      child {node[cnode] (kitchen) {}
-        child {node [pnode] (box2) {} 
-          child {node [cnode] (box2p1) {}}
-          child {node [cnode] (box2p2) {}}
-        }
-      };
-   \tiny
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of kitchen] {$\pp{kitchen}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[below=0 of cup] {$\pp{(is-in cup)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-   \node[right=0 of cupp1] {$\pp{place1}$};
-   \node[right=0 of cupp2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Generating the abstract belief space.}
-\label{fig:abstraction}
-\end{figure}
-
-%%% OLDER STUFF
-
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
-
-\section{Evaluation}
-
-We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. 
-% Our implementation is able
-% to use several underlying planning systems.
-We have extended MAPSIM so that it can parse DTPDDL, and perform
-successive estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
-
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
-}
-
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
-
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
-
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
-
-We evaluated our switching planner against the baseline in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
-  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
-  core.}. In our domain, rewards encapsulate a goal condition, and
-therefore we find it intuitive to report average costs of the plans
-and the success rate in problems that admit a solution (i.e., positive
-reward scaled by a constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
-  \caption{Average runtime}
-  \label{fig:results-time}
-\end{figure}
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
-  \caption{Average plan costs and number of successful runs.}
-  \label{fig:results-quality}
-\end{figure}
-
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% \section{Discussion}
-%% \input{discussion.tex}
-
-\section{Related Work}
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
-
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-%%
-The continual planning system that motivated our
-project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-and has been applied in completely observable domains, particularly
-those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
-
-The most pressing item for future research, is to develop a scheme
-whereby the serial planner can relax {\em executability} assumptions,
-so that conformant (or semi-conformant) plans can be executed without
-interruption by a contingent session. A more general criticism of
-switching continual planning concerns interleaved sequential and
-contingent sessions failing to make any progress towards the
-objectives. For example, in the worst case we can have each sequential
-session producing an identical plan, and each decision-theoretic
-session rejecting it without further sensing. Although this has not
-been an issue in our work so far, it must be dealt with rigorously
-in the future. We suggest that a good way to mitigate this problem is
-by developing a {\em motivational} component that maintains a {\em
-dynamic} reward model whose limiting behaviour prevents that switching
-deadlock.
-
-%% The sequential mode of our system always schedules decision-theoretic
-%% planning before executing actions that are not applicable with
-%% certainty. This is inefficient if the utility of the plan is not
-%% dependent on that actions successful execution. This situation always
-%% arises for conformant plans.
-
-%% The switching continual planning system we have described serves as
-%% the underlying planner for CogX.
-
-%% the effects of a {\em sense} schema are perceptual, whereas the
-%% effects of an {\em operator} schema are over state propositions.
-
-%% posted by a motivational component of the underlying robotic
-%% architecture.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% In this paper we present an approach to continual planning that uses
-%% two planning systems. The first
-
-%%  to a distinct class of
-%% challenges. We suppose 
-%% %%
-%% The underlying environment is modelled as a POMDP. We use the fast
-%% classical satisficing system FastDownward to find a deterministic
-%% sequential plan and complementary runtime evolution of that
-%% process. This corresponds to a generalisation of replanning in
-%% probabilistic planning to problems with partial observability.
-
-%% Interaction between the sequential planner and execution proceeds
-%% more-or-less analogously to popular replanning approaches
-
-%% Addressing these challenges in a monolithic framework, we present a
-%% {\em switching} continual planner, that uses the fast sequential
-%% satisficing procedure FastDownward to perform net-benefit
-%% %%
-%% makes reasonable assumptions about the evolution of the runtime state
-%% given a POMDP model of the environment.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% The decision-theoretic planner is able to tailor sensory processing on
-%% a robot platform to the current objective, while FastDownward  quickly
-
-%% In this paper we develop a continual planing system that uses two
-%% planning systems. The first, is a state-of-the-art domain independent
-%% planner for deterministic problems. The second is a information-state
-%% contingency planning the information-state space of
-
-%% Continual planning is a powerful technique that goes some way to
-%% addressing those challenges. That approach interleaves planning and
-%% execution, deliberately postponing planning for contingencies unless
-%% they eventuate during execution.
-
-%% computing a single sequential plan and
-%% eventuality
-
-%% To these challenges it seems a continual planning approach is best,
-
-%% where the runtime state evolves during plan execution
-
-%% interleaved planning and execution.
-
-%% the latter is able to tailor sensory processing on a robot platform,
-%% in order that it.
-
-%% {\em ad-hoc}
-
-%% , reason about degrees of belief and uncertainty about the world,
-
-%% probabilistic sequential decision making in practical sized problems
-%% is intractable.
-
-%% quantitative probabilistic models of the perception and action .
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai11}
-
-\end{document}

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/ijcai-workshop.tex
Our workshop attempt.

--- 2011-02-04 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
@@
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
+\typeout{IJCAI Switching Planner Submission}
+
+\documentclass{article}
+\usepackage{ijcai11}
+\usepackage{times}
 
 \usepackage{subfig}
-
 \usepackage{tikz}
@@
 
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
 \usepackage{graphicx}
@@
 
-\nocopyright
-
 \pdfinfo{
 /Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
+/Subject (IJCAI Workshop on Decision Making in Partially Observable, Uncertain Worlds: Exploring Insights from Multiple Communities)
 /Author (NA)}
@@
 
-% \title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
+\title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
 
-\title{A Switching Planner for Combined Task and Observation Planning}
+%% \title{A Switching Planner for Combined Task and Observation Planning}
 \author{anonymous}
@@
 events, changing objectives, and the underlying unpredictability of
-the environment.  On the other hand, in order to exhibit efficient
+the environment.  On the other hand, in order to exhibit useful
 behaviours the planning process must perform computationally expensive
@@
 deeper consideration of the sensing model is both practical, and can
-significantly impact overall performance. We evaluate our approach
+significantly impact overall performance. We demonstrate our approach
 in large problems from a realistic robot exploration domain.
@@
 %% partial observability
+
+For example, the robot usually knows exactly where it is located on a
+topological map, and therefore there are tight constraints about what
+regions of the map can be observed, and what regions of the map can be
+moved to in an atomic step.
+
+Moreover, we use this domain description to model the action physics
+of abstract processes posed to decision-theoretic planning
+sessions. Here, we forbid the planner from considering courses of
+action that are not {\em known} to be possible to alleviate the
+computational burden planning in the abstract process.
+
+The initial abstract process, let's suppose, has two states. The state
+where everything assumed is true, and the null state. This has a KL
+divergence from the :init term, because we can suppose it uniformly
+chooses one concrete state, given the abstract :init states.
+
+As you add propositions to the abstract :init term, then the KL
+divergence decreases.
 
@@
 
-%% \section{Propositional Decision-Theoretic Planning}
-%% \input{prop-dtp.tex}
-
-\section{Background and Related Work}
-\input{relatex.tex}
+\section{Propositional Decision-Theoretic Planning}
+\input{prop-dtp.tex}
 
@@
 %% \input{discussion.tex}
+
+\section{Related Work}
+\input{relatex.tex}
 

### 2011-03-09 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/evaluation.tex
forgot one of the files.

--- 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
@@
 We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
-to use several underlying planning systems. We have also extended the
-MAPSIM system so that it can parse DTPDDL, and perform successive
-estimation of the underlying belief-state using
+environment~\cite{brenner:nebel:jaamas09}. 
+% Our implementation is able
+% to use several underlying planning systems.
+We have extended MAPSIM so that it can parse DTPDDL, and perform
+successive estimation of the underlying belief-state using
 \system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
@@
 evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds). In our domain,
-rewards encapsulate a goal condition, and therefore we find it
-intuitive to report average costs of the plans and the success rate in
-problems that admit a solution (i.e., positive reward scaled by a
-constant factor).
+timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
+  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
+  core.}. In our domain, rewards encapsulate a goal condition, and
+therefore we find it intuitive to report average costs of the plans
+and the success rate in problems that admit a solution (i.e., positive
+reward scaled by a constant factor).
 

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/conc.tex
New versions

--- 2011-02-09 subarchitectures/planner.sa/trunk/src/latex/AAAI11/conc.tex
@@
 We have addressed a key challenge, specifically that of high-level
-continual planning and execution monitoring for efficient
-deliberations according to rich probabilistic models afforded by
-recent integrated robotic systems. We developed a system that can plan
-quickly given large realistic probabilistic models, by switching
-between: (a) fast sequential planning, and (b) expensive DT planning
-in small abstractions of the problem at hand. Sequential and DT
-planning is interleaved, the former identifying a rewarding sequential
-plan in the underlying process, and the latter solving small sensing
-problems posed during sequential plan execution.
+continual planning for efficient deliberations according to rich
+probabilistic models afforded by recent integrated robotic systems. We
+developed a system that can plan quickly given large realistic
+probabilistic models, by switching between: (a) fast sequential
+planning, and (b) expensive DT planning in small abstractions of the
+problem at hand. Sequential and DT planning is interleaved, the former
+identifying a rewarding sequential plan for the underlying process, and
+the latter solving small sensing problems posed during sequential plan
+execution.
 %%

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/ijcai-workshop.tex
New versions

@@
 
-\title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
-
-%% \title{A Switching Planner for Combined Task and Observation Planning}
+\title{A Switching Planner for Combined Task and Observation Planning}
 \author{anonymous}
@@
 events, changing objectives, and the underlying unpredictability of
-the environment.  On the other hand, in order to exhibit useful
+the environment.  On the other hand, in order to promote efficient
 behaviours the planning process must perform computationally expensive
@@
 satisficing ``classical'' planner, to decide on the overall strategy,
-and decision-theoretic planning to solve small subproblems where
+and decision-theoretic planning to solve small abstract subproblems where
 deeper consideration of the sensing model is both practical, and can
-significantly impact overall performance. We demonstrate our approach
+significantly impact overall performance. We evaluate our approach
 in large problems from a realistic robot exploration domain.
@@
 
-\section{Propositional Decision-Theoretic Planning}
-\input{prop-dtp.tex}
+%% \section{Propositional Decision-Theoretic Planning}
+%% \input{prop-dtp.tex}
 
-\section{Decision-Theoretic PDDL}
+\section{Planning Language and Notations}
 \input{dtpddl.tex}
@@
 
+\section{Related Work}
+\input{relatex.tex}
+
 %% \section{Discussion}
 %% \input{discussion.tex}
-
-\section{Related Work}
-\input{relatex.tex}
 
@@
 
-\bibliographystyle{aaai}
+\bibliographystyle{named}
 \bibliography{papers}
@@
 stet the goal condition is satisfied.
+
+\title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}
+
+%% \title{A Switching Planner for Combined Task and Observation Planning}
+\author{anonymous}
+\setcounter{secnumdepth}{0}
+
+\begin{document} 
+\maketitle
+
+\begin{abstract}
+
+%% Those processes must
+
+From an automated planning perspective the problem of practical mobile
+robot control in realistic environments poses many important and
+contrary challenges.  On the one hand, the planning process must be
+lightweight, robust, and timely. Over the lifetime of the robot it
+must always respond quickly with new plans that accommodate exogenous
+events, changing objectives, and the underlying unpredictability of
+the environment.  On the other hand, in order to exhibit useful
+behaviours the planning process must perform computationally expensive
+reasoning about contingencies and possible revisions of subjective
+beliefs according to quantitatively modelled uncertainty in acting and
+sensing.
+%%
+Towards addressing these challenges, we develop a \emph{continual
+planning} approach that {\em switches} between using a fast
+satisficing ``classical'' planner, to decide on the overall strategy,
+and decision-theoretic planning to solve small subproblems where
+deeper consideration of the sensing model is both practical, and can
+significantly impact overall performance. We demonstrate our approach
+in large problems from a realistic robot exploration domain.
+
+%% partial observability
+
+For example, the robot usually knows exactly where it is located on a
+topological map, and therefore there are tight constraints about what
+regions of the map can be observed, and what regions of the map can be
+moved to in an atomic step.
+
+Moreover, we use this domain description to model the action physics
+of abstract processes posed to decision-theoretic planning
+sessions. Here, we forbid the planner from considering courses of
+action that are not {\em known} to be possible to alleviate the
+computational burden planning in the abstract process.
+
+The initial abstract process, let's suppose, has two states. The state
+where everything assumed is true, and the null state. This has a KL
+divergence from the :init term, because we can suppose it uniformly
+chooses one concrete state, given the abstract :init states.
+
+As you add propositions to the abstract :init term, then the KL
+divergence decreases.
+
+%% In cooperation with execution monitoring, it must perform repair
+%% and/or replanning in order to seamlessly accommodate exogenous events,
+%% changing objectives, and the underlying unpredictability of the
+%% environment.
+
+%% Realistic robot planning problems in uncertain environments often
+%% require achieving tasks while also finding out about the
+%% world. Because the world state cannot be observed directly, these
+%% problems are naturally represented as partially observable Markov
+%% decision problems (POMDPs). However, these are typically intractable
+%% for realistic problems.
+%%
+
+\end{abstract}
+
+%% \section{Moritz's Initial Outline}
+
+%% \subsection{Switching Planner Outline}
+
+%% Outline the high level approach here:
+%% \begin{itemize}
+%% \item Problem input (factored initial state, DTPDDL observations)
+%% \item Determinisation
+%% \item Problem generation for DT
+%% \item Belief revision
+%% \end{itemize}
+
+%% \subsection{Determinisation}
+
+%% \begin{itemize}
+%% \item Generation of deterministic sensing models
+%% \item Determinisation of the initial state
+%% \item Compiling reward model into (soft-)goal model
+%% \item Cost function in FD (maybe this could mentioned in the outline?)
+%% \end{itemize}
+
+%% \subsection{Subtask Generation}
+%% \begin{itemize}
+%% \item Goal generation from CP trace
+%% \item Disconfirm actions
+%% \item State pruning according to CP trace
+%% \end{itemize}
+
+%% \subsection{Belief revision}
+%% Not sure if this needs its own section
+%% \begin{itemize}
+%% \item Convert inital state into bayesian network
+%% \item Belief updates with observation nodes
+%% \item Create new network for next planning runs
+%% \end{itemize}
+
+\section{Introduction}
+\input{intro.tex}
+
+\section{Propositional Decision-Theoretic Planning}
+\input{prop-dtp.tex}
+
+\section{Decision-Theoretic PDDL}
+\input{dtpddl.tex}
+
+\section{Switching Continual Planner}
+\input{switching.tex}
+
+\section{Evaluation}
+\input{evaluation.tex}
+
+%% \section{Discussion}
+%% \input{discussion.tex}
+
+\section{Related Work}
+\input{relatex.tex}
+
+\section{Concluding Remarks}
+\input{conc.tex}
+
+\bibliographystyle{named}
+\bibliography{papers}
+
+\end{document}
+
+\subsection{PPDDL}
+
+We briefly discuss the description language PPDDL, a fairly
+straight-forward extension of the Planning Domain Definition Language
+for describing fully observable Markov decision processes.
+
+\small{
+\begin{tabtt}
+(\= :action look-at-object \\
+  \> :parameters \=(?r - robot ?v - visual-object\\
+  \> \> ?l - label ?l - location) \\
+  \> :precondition (and (= (is-in ?r) ?l) \\
+  \> \> (= (label ?o) ?l) ) \\
+  \> :effect (and (assign (reward) -3) ) ) \\
+\end{tabtt}
+}
+
+\small{
+\begin{tabtt}
+(\= :observe visual-object \\
+  \> :parameters \=(?r - robot ?v - visual-object\\
+  \> \> ?l - label ?l - location) \\
+  \> :execution (look-at-object ?r ?v ?l ?l) \\
+  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
+\end{tabtt}
+}
+
+\small{
+\begin{tabtt}
+(\= :sensor look-for-object \\
+  \> :agent (?a - robot) \\
+  \> :parameters (?l - label ?l - location) \\
+  \> :variables (?o - visual-object) \\
+  \> :precondition (\=and (= (is-in ?r) ?l) \\
+  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
+  \> :effect () \\
+  \> :sense (= (is-in ?o) ?l) \\
+\end{tabtt}
+}
+
+\subsection{MAPL}
+
+%% \scriptsize
+%% \begin{nopagebreak}\begin{tabtt}
+%% <observation-def> \=::= \=(:sense <observation symbol> \\
+%%                         \> \> :parameters (<typed list (variable)>)  \\
+%%                         \> \> <o-def body>) \\
+%%   <o-symbol> \> ::= <name> \\
+%%   <o-def body> \> ::= [:precondition <GD>] \\
+%%   \> \> [:execution <atomic action(term)> ] \\
+%%     \> \> [:effect <o-effect>] \\
+%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
+%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
+%%   <o-effect> \> ::= <c-o-effect> \\
+%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
+%%   <c-o-effect> \> ::= <p-o-effect> \\
+%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
+%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
+%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
+%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
+%%   <o-f-exp> \> ::= <number>\\
+%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
+%%   <o-f-exp> \> ::= <o-f-head>\\
+%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
+%%   <o-f-head> \> ::= <o-function-symbol>\\
+%% \end{tabtt}\end{nopagebreak}
+%% \noteme{<structure-def> ::= <attach-def>}
+%% \normalsize
+
+%% \footnote{We have simplified the schemata for illustrative
+%% purposes. In practice, the probabilities of a particular observation
+%% here should be parametrised by the type of visual object the robot is
+%% searching for, and category of the location.}
+
+We suppose $\bstate_0$ is given in a factored tree format of the
+form:
+
+%% \small{
+%% \begin{tabtt}
+%% (:init  (and \=f_{11} f_{12} ..
+%%  \> (probabilistic p_{}) 
+%%  \> (probabilistic ..) ..) \\
+%% \end{tabtt}
+%% }
+
+%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]
+
+planning in practical sized POMDPs is intractable. One approach is to
+described the task hierarchically, thereby constrain and simplify the
+policy search task. The approach ~\cite{}.
+
+%% More generally, a useful
+%% formalism for representing solutions to POMDPs is the finite-state
+%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
+%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
+%% $\selection_\node(\action) = P(\action | \node)$, and
+%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
+%% \action, \observation)$. Where the value of acting is the discounted
+%% accumulated reward over an infinite horizon, then each controller can
+%% be assigned a value, written $V_\node(\state)$, expressed in the usual
+%% way according to the Bellman equation:
+
+%% \begin{equation}\label{eq:evaluation}
+%% \begin{array}{lcl}
+%% V_\node(\state) & = & \sum_{\action \in \actions}
+%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\
+
+%% && \hspace{-10ex} \beta \sum_{\action, \observation,
+%% \state', \node'} \transition_\node(\action, \observation, \node')
+%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
+%% \action) V_{\node'}(\state')
+%% \end{array}
+%% \end{equation}
+
+%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
+%% then:
+
+%% \begin{equation} \label{eq:valueBelief}
+%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
+%% \end{equation}
+
+%% \noindent There is a corresponding deterministic FSCs for both
+%% sequential (i.e., conformant) and contingent
+%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
+%% case, if we take $\beta = 1$, and alter the process dynamics so there
+%% is a compulsory transition to a zero utility sink state when that
+%% horizon is reached.
+
+ ---i.e., the values of {\em
+runtime variables} in the continual paradigm--- are treated , and
+therefore plan for a specific eventuality, replanning from scratch if
+something
+
+Sequential planning in our approach proceeds in two phases. In the
+first phase, the goal is to identify a plan prefix
+
+In sequential planning, the plan prefix corresponds to choosing a
+valid abstract starting state.
+
+moreover, for the purpose of decision-theoretic abstractions,
+
+A relaxed form of {\em visitation} we call {\em abstract
+  visitation}. In the abstract case a conjunctive term is visited
+  provided all its atomic subterms are visited, and zero or more of
+  its probabilistic terms are visited.
+
+The planner has a compulsory starting action, that visits the root
+term, and then has to execute visitation actions until that plan
+prefix corresponds to a valid {\em abstract visitation} of the {\em
+:init} decleration. Beyond such a prefix, the process actions and senses are
+available to the planner
+
+Goal expressions in the problem description ---for example
+$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
+location of the {\em box}--- are treated as control-knowledge by the
+planner. Here, a plan is not considered valid unless in the finial
+stet the goal condition is satisfied.

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/intro.tex
New versions

%% LONG PLANNING HORIZON. MORITZ, CAN YOU PLEASE GIVE ME THE NUMBER OF
%% STEPS IN THE BIGGER PLANS.

A number of recent integrated robotic systems incorporate a
high-level {\em continual planning} and execution monitoring
subsystem~\cite{wyattetal2010tamd,talamadupula:2010,Kraft2008}.
%%
For the purpose of planning, sensing is modelled
deterministically, and beliefs about the underlying state are
modelled qualitatively.
%%
Both~\citeauthor{talamadupula:2010} and~\citeauthor{wyattetal2010tamd}
identify
continual planning with probabilistic
models of noisy sensing and state as an important challenge for future
research.
%%
Motivating that sentiment, planning according to accurate
stochastic models should yield more efficient and robust
deliberations.
%%
%% In the first place, that means moving from deterministic to
%% probabilistic models of noisy sensing, and from qualitative to
%% quantitative models of state uncertainty. 
%%
In essence, the challenge is to develop a planner that
exhibits speed and scalability similar to planners employed in
existing robotic systems---e.g.,~\citeauthor{wyattetal2010tamd} use a
satisficing {\em classical} procedure---and which is also able to
synthesise relatively efficient deliberations according to detailed
probabilistic models of the environment.

This paper describes a {\em switching} domain independent planning
approach we have developed to address this challenge. 
%%
Our Planner is {\em continual} in the usual sense that plans are
adapted and rebuilt online in reaction to changes to the model of the
underlying problem and/or domain---e.g., when goals are
modified, or when the topological map is altered by a door being
closed.
%%
It is integrated on a mobile robot platform that continuously
deliberates in a stochastic dynamic environment in order to achieve
goals set by the user, and acquire knowledge about its surroundings.
%%
Our planner takes problem and domain descriptions expressed in a novel
extension of PPDDL~\cite{younes:etal:2005}, called {\em
Decision-Theoretic DTPDDL}, for modelling stochastic decision
problems that feature partial observability.  In this paper we
restrict our attention to problem models that correspond to
deterministic-action multi-goal-oriented POMDPs 
% \footnote{We note that
% POMDPs with stochastic actions can be compiled into equivalent
% deterministic-action POMDPs, where all the original action uncertainty
% is expressed in the starting-state
% distribution~\cite{ng:Jordan:2000}.}
in which all actions have
non-zero cost---i.e., an optimal policy can be formatted as a finite
horizon contingent plan. Moreover, we target problems of a size and
complexity that is challenging to state-of-the-art sequential
satisficing planners, and which are too large to be solved directly by
decision-theoretic (DT) systems.

Our planner {\em switches}, in the sense that the base planning
procedure changes depending on our robot's subjective degrees of
belief, and on progress in plan execution. When the underlying planner is
a fast (satisficing) {\em classical} planner, we say planning is in a
{\em sequential} session, and otherwise it is in a DT session.
%%
A {\em sequential} session plans, and then pursues a high-level
strategy---e.g., go to the kitchen bench, and then observe the
cornflakes on it.
%%
A DT session proceeds in a practically sized {\em abstract} process,
determined according to the current sequential strategy and underlying
belief-state.
%%

We evaluate our approach in simulation on problems posed by object search and room categorisation tasks that our indoor
robot undertakes. Those feature a deterministic task planning aspect
with an active sensing problem. The larger of these problems features
$6$ rooms, $25$ topological places, and $21$ active sensing
actions. The corresponding decision process has a number of states
exceeding $10^{36}$, and high-quality plans require very long planning
horizons.
%%
Although our approach is not optimal, particularly as it relies on the
results of satisficing sequential planning directly, we find that it
does nevertheless perform better than a purely sequential replanning
baseline. Moreover, it is fast enough to be used for real-time
decision making on a mobile robot.

### 2011-03-09 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/evaluation.tex
New version for IJCAI workshop

--- 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
@@
 
-We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
-to use several underlying planning systems. We have also extended the
-MAPSIM system so that it can parse DTPDDL, and perform successive
-estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
+We have implemented our switching approach in the MAPSIM
+environment~\cite{brenner:nebel:jaamas09},
+using \system{dlib-ml}~\cite{king:2009} for belief
+revision. Sequential sessions use a modified version of Fast
+Downward~\cite{fast-downward}, and DT sessions use our own contingent
+procedure. Since most of the problems we consider are much larger than
+any available DT planner can solve directly, for comparison purposes
+we also implemented a simple dual-mode replanning baseline
+approach. Here, when a switching action is scheduled for execution the
+DT session plans to a single entropy reduction action, whose execution
+can provide evidence regarding the truth value of a relevant
+assumption.
+%%
+%%  -- i.e,. we say a fact can
+%% be directly sensed, if there is a DTPDDL {\em sense} declaration that
+%% mentions the fact in its precondition. 
+%%
+Control is then immediately
+returned to a new sequential session.
 
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
+We evaluate our approaches in robot exploration tasks from home and
+office environments. Spatially, these consist of {\em rooms}
+(office/kitchen/etc), and an underlying topological map over
+smaller areas of space, called {\em places}, and connectivity between
+those. The mobile {\em robot} and {\em visual objects} inhabit the
+topological places. Objects indicate the category of space they
+inhabit---e.g., spoons are likely to be in kitchens. By examining
+view cones at places for particular objects, the robot is able to: (1)
+categorise space at high (room) and low (place) levels, and (2) find
+objects for the user, exploiting information about object
+co-occurrence and room categories for efficiency. Also, in the
+presence of a person, the robot can ask about the category of the
+current room.
+
+We compare switching to the baseline in several realistic
+tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
+$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
+$>10^{36}$). We also compare those systems with near optimal policies
+computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
+3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
+levels of reliability in sensing: {\em reliable} sensors have a $.1$
+probability of a false negative, {\em semi-reliable} have a chance of
+$0.3$ of false negative and $0.1$ of false positive, and {\em noisy}
+sensors with probabilities of $0.5$ and $0.2$ respectively. Each
+object class is assigned one sensor model---e.g. cornflakes may be
+harder to detect than refrigerators. We performed several experiments
+with different levels of reliability for sensing the target object(s),
+while keeping sensing models for non-target objects constant.
+
+\Omit{
+All non-target objects remain unchanged so the planner can
+exploit object co-occurrences using reliably sensed objects to
+compensate for the increased sensor noise.
 }
 
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
-
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
-
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
-
-We evaluated our switching planner against the baseline in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
+Our evaluation examines DT sessions with initial belief-states
+admitting between 20 and 100 abstract states with non-zero
+probability. We run 50 simulations in each configuration, and have a
+timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
+experiments were conducted on a 2.66GHz Intel Xeon X5355 using one CPU
+core.}. The continual planning times are reported in
 Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds). In our domain,
-rewards encapsulate a goal condition, and therefore we find it
-intuitive to report average costs of the plans and the success rate in
-problems that admit a solution (i.e., positive reward scaled by a
-constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
+Figure~\ref{fig:results-quality}.
+%%
+For each task, the goal is to find one or more objects and report
+their position to a user. Usually there is a non-zero probability that
+no plan exists, as the desired object might not be present in the
+environment.
+% MOG: I didn't do  evaluations for those, so let's remove them
+% Item~$f$ reports for tasks requiring indirect sensing,
+% where the robot must relocate to a room with a particular category
+% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
+% so is omitted from that reporting.  
+In these experiments we only allocate reward on the achievement of all
+goals, therefore we find it intuitive to report average plan costs and
+the success rates in problems that admit a complete solution (i.e.,
+positive reward scaled by a constant factor). The exception occurs for
+items f and g of Figure~\ref{fig:results-quality}, where we report
+expected discounted rewards (not plan costs).
 
@@
   \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
+  % \vspace{2mm}
+  % \includegraphics{dora-cat-time}\hfill
   \caption{Average runtime}
@@
   % \vspace{2mm}
+  % \includegraphics{pomdp-solvable-quality}\hfill
   \includegraphics{dora2-quality}\hfill
@@
   \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
+  \includegraphics{pomdp-quality}\hfill
+  % \vspace{2mm}
+  % \includegraphics{dora-cat-quality}\hfill
   \caption{Average plan costs and number of successful runs.}
@@
 
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% We find that abstraction has little impact on plan
-%% costs and success rate in our scenario. Increasing the size of the
-%% initial abstract belief beyond 50 states rarely pays off, because
-%% while additional information may facilitate better plans, the indirect
-%% sensing tests report the same result. Here, we only really see an
-%% impact in terms of the runtime of contingent planning.
-
-% We believe that a part of the improvement is due to the segmentation of
-% the plan into several subtask, essentially performing hierarchical
-% planning. Especially when the continual planner performs badly this is
-% a huge gain.
+We find that if sensing is reliable, then little is gained using DT
+sessions, as the greedy approach of the baseline is
+sufficient. As sensing degrades DT sessions prove more useful. Here,
+time spent on DT planning increases steeply as the abstraction becomes
+more refined, which is compensated for by fewer planning sessions
+overall. More detailed abstractions lead to a better overall
+success rate, particularly for tasks $d$ and $e$.
+%% Moreover,
+%% for less refined abstractions, the increased cost of DT planning is
+%% partially compensated for by a decrease in Fast Downward planning
+%% times. 
+Speaking to the effectiveness of our entropy heuristic for abstraction
+refinement, we see relatively high success rates irrespective of
+the level of refinement. Comparing finally to the best {\sc
+zmdp} policy, although producing relatively costly plans, the
+continual planners performed quite well, especially in terms success
+rate. A key source of inefficiency here, is due to sequential sessions
+always being optimistic, and refusing to abandon the search.
 

### 2011-03-09 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/relatex.tex
New version for IJCAI workshop

--- 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
@@
 online POMDP solution procedures have been developed which leverage
-highly approximate value functions -- computed using an offline
-procedure -- and heuristics in forward
+highly approximate value functions----computed using an offline
+procedure----and heuristics in forward
 search~\cite{ross:etal:2008}. These approaches are applicable in
@@
 compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
+search planning procedures, and lazy evaluations of beliefs. We
+consider it an appealing future direction to pursue conformant
+reasoning during the sequential sessions we proposed. Most
 recently this research thread has been extended to contingent planning
@@
 environments~\cite{albore:etal:2009}.
-
-%% Their applicability in our setting is
-%% limited, firstly because they only scale to smaller problems, with
-%% thousands of states, and also due to the large amount of
-%% \emph{problem-specific} offline processing that might be required to get useful
-%% search guidance.
-
-%%  In that
-%% setting \citeauthor{kurniawati:etal:2010}~(\citeyear{})
-%% recently addressed an inefficiency of offline point-based techniques
-%% in problems with medium length planning horizons, however their
-%% approach take tens of minutes to good plans, and only scales to tens
-%% of thousands of states. In the case of general domain-independent
-%% factored systems, the state-of-the-art scales to relatively small
-%% problems with $2^{22}$ states~\cite{shani:etal:2008}.
-
-%% At their limit,
-%% these procedures take over an hour to converge, and $\sim10$ seconds
-%% on average to perform a single Bellman backup.  For classes of POMDP
-%% that feature exploitable structures (e.g., no actions with negative
-%% effects), problems with as many as $10^{30}$ states can be targeted by
-%% such offline procedures~\cite{brunskill:russell:2010}.
-
-%% Although we suppose it an interesting
-%% item for future work to pursue that direction, it should be noted that
-%% ease of goal reachability is not guaranteed in the problems we face,
-%% and is certainly not a property to be assumed in domain independent
-%% planning.
-
-%%
-%% The continual planning system that motivated our
-%% project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-%% and has been applied in completely observable domains, particularly
-%% those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/intro.tex
New version for IJCAI workshop

@@
 extension of PPDDL~\cite{younes:etal:2005}, called {\em
-Decision-Theoretic DTPDDL}, for modelling stochastic decision
-problems that feature partial observability.  In this paper we
-restrict our attention to problem models that correspond to
-deterministic-action multi-goal-oriented POMDPs 
-% \footnote{We note that
-% POMDPs with stochastic actions can be compiled into equivalent
-% deterministic-action POMDPs, where all the original action uncertainty
-% is expressed in the starting-state
-% distribution~\cite{ng:Jordan:2000}.}
-in which all actions have
-non-zero cost---i.e., an optimal policy can be formatted as a finite
-horizon contingent plan. Moreover, we target problems of a size and
-complexity that is challenging to state-of-the-art sequential
-satisficing planners, and which are too large to be solved directly by
-decision-theoretic (DT) systems.
+Decision-Theoretic DTPDDL}, for modelling stochastic decision problems
+that feature partial observability.  In this paper we restrict our
+attention to problem models that correspond to deterministic-action
+multi-goal-oriented POMDPs in which all actions have non-zero
+cost\footnote{We note that POMDPs with stochastic actions can be
+compiled into equivalent deterministic-action POMDPs, where all the
+original action uncertainty is expressed in the starting-state
+distribution~\cite{ng:Jordan:2000}.} -- i.e., an optimal policy can be
+formatted as a finite horizon contingent plan. Moreover, we target
+problems of a size and complexity that is challenging to
+state-of-the-art sequential satisficing planners, and which are too
+large to be solved directly by decision-theoretic (DT) systems.
 

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/switching.tex
New version for IJCAI workshop

@@
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
+We now describe our {\em switching} planning system that operates
+according to the continual planning paradigm. The system {\em
+switches} in the sense that planning and plan execution proceed in
+interleaved sessions in which the {\em base planner} is either {\em
+sequential} or {\em decision-theoretic}.
+%%
+The first session is sequential, and begins when a DTPDDL description
+of the current problem and domain are posted to the system.  During a
+sequential session a serial plan is computed that corresponds to one
+execution-{\em trace} in the underlying decision-process. That trace
+is a reward-giving sequence of process actions and {\em assumptive}
+actions. Each {\em assumptive} action corresponds to an assertion
+about the truth of a fact that is unknown at plan time -- e.g. that a
+box of cornflakes is located on the corner bench in the kitchen. The
+trace specifies a plan and characterises a {\em deterministic
+approximation} (see~\cite{yoon:etal:2008}) of the underlying process
+in which that plan is valuable. Traces are computed by a
+cost-optimising classical planner which trades off action costs, goal
+rewards, and determinacy. Execution of a trace proceeds according to
+the process actions in the order that they appear in the trace. If,
+according to the underlying belief-state, the outcome of the next
+action scheduled for execution is not predetermined above a threshold
+(here 95\%), then the system switches to a DT session.
 
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
+Because online DT planning is impractical for the size of problem we
+are interested in, DT sessions plan in a small abstract problem
+defined in terms of the trace from the proceeding sequential
+session. This abstract state-space is characterised by a limited
+number of propositions, chosen due to their {\em relevance} to the
+action whose scheduled execution triggered the switch to the DT
+session.
 %%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
+To allow the DT planner to judge assumptions from the trace, we add
+{\em disconfirm} and {\em confirm} actions to the problem for each of
+them. Those yield a relatively small reward/penalty if the
+corresponding judgement is true/false. If a judgement action is
+scheduled for execution, then the DT session is terminated, and
+a new sequential session begins.
 
-Finally, whether proceeding in a sequential or contingent session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
+%% Because online DT planning is impractical for the size of problem we
+%% are interested in, when a DT session begins we use the trace from the
+%% preceeding sequential session to restrict the state space. The state space is characterised by a limited number of propositions,
+%% chosen due to their {\em relevance} to the action whose scheduled
+%% execution triggered the switch to the DT session.
+%% %The state
+%% %space is constructed by first removing all propositions (and related
+%% %actions) that are not true or assumed true of states in the trace,
+%% %then adding them back as long as the number of states does not exceed
+%% %a size threshold\footnote{Our experiments are performed using a number
+%% %of different threshold parameters.}  Propositions to add back are
+%% %chosen using as a heuristic the entropy of the trace assumptions
+%% %conditional on the candidate proposition. 
+%% To allow the DT planner to
+%% determine the truth of the assumptions in the trace, we add {\em
+%% disconfirm} and {\em confirm} actions to the problem for each of
+%% them. These actions yield a relatively small reward/penalty if the
+%% corresponding judgement is true/false. If a judgement action is
+%% scheduled for execution the DT session is terminated, and control is
+%% returned to a sequential session.
 
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
+Whatever the session type, our continual planner maintains a factored
+representation of successive belief-states.
+% by performing belief revision. 
+As an internal representation of the $(\pp{:init})$
+declaration, we keep a tree-shaped Bayesian network which gets updated
+whenever an action is performed, or an observation received. That
+belief-state representation is used: (1) as the source of candidate
+determinisations for sequential planning, (2) in determining when to
+switch to a DT session, and (3) as a mechanism to guide construction
+of an abstract process for DT sessions.
 
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
+\subsection{Sequential Sessions}
 
-\subsection{Sequential Planning in POMDPs}
+As we only consider deterministic-action POMDPs, all state
+uncertainty is expressed in the $(\pp{:init})$ declaration. This
+declaration is used by our approach to define the starting state for
+sequential sessions, and the set of state-assumptions available to
+sequential planning.  Writing \#\ if the value of a proposition is
+unspecified, taking the $(\pp{:init})$ example from the previous
+section, we have the following assumptions:
 
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
-
-\small
+\tiny
 \begin{tabular}{cccc}
 \hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
+Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
 \hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
+%% .24 & kitchen & office & office \\
+%% .06 & kitchen & kitchen & office \\
+%% .56 & kitchen & office & kitchen \\
+%% .14 & kitchen & kitchen & kitchen \\
 .7 & kitchen & \# &  kitchen\\
@@
 
-\noindent An assumption corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
+\noindent Each assumption corresponds to one distinct {\em
+relaxed} visitation of the root term. Here, a conjunctive term is
+visited iff its atomic subterms are visited, and zero or one of its
+immediate probabilistic subterms are visited. For a sequential
+session, the starting state corresponds to an assumption (usually
+unique) with probability $1$.\footnote{Without a loss of generality,
+we assume conjunctive terms list atoms that are true with
+probability 1 according to a visitation, and that actions do not
+have negative preconditions.}  In our example, that starting state is:
 
+\vspace{-1ex}
+
+\tiny
 \[
 \begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
+\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
 \;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
@@
 \]
+\normalsize
 
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
+\vspace{-1ex}
+
+To represent state assumptions we augment the problem posed during
+a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
+each element $\prob_i (T_i)$, of each probabilistic term from
+$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
+$\assumptiveS{j}$, $j \neq i$, has been executed from the same
+probabilistic term, and, either
+$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
+it occurs in $T_k$ for some executed $\assumptiveS{k}$.
 %%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
+We also add constraints that forbid scheduling of
+assumptions about facts after actions with preconditions or effects
+that mention those facts. For example, the robot cannot assume it is
+plugged into a power source immediately after it unplugs itself.
 %%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
+Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
+to a successor state $\state^{T_i}$ with probability $\prob_i$, and
+$\state^\bot$ with probability $1 - \prob_i$. Here in addition to
+respecting the above constraints, $\state^{T_i}$ is the union of
+$\state$ with atomic terms from $T_i$. State $\state^\bot$ is an added
+sink.
 
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
+We now describe the optimisation criteria used during sequential
+sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
+action, $\action_i$, from a trace of state-action pairs
+$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
+does not transition to $\state^\bot$, and
+$\reward(\state_i, \action_i)$ be the instantaneous reward received
+for executing action $\action_i$ in state $\state_i$. The optimal
+trace has value:
 
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
+%We now describe the optimisation criteria used during sequential
+%sessions. Where $\prob_i$ is the probability that the $i^{th}$
+%sequenced action, $\action_i$, from a trace of state-action pairs
+%$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
+%does not transition to $\state^\bot$, we have that the optimal trace
+%has value:
 
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
-
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
+\tiny
 \[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
+V^* = \max_N \max_{\state_0, \action_0,.., \state_N} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
 \reward(\state_i, \action_i),
 \]
+\normalsize
 
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
+%\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
+%reward ---i.e., can be negative for action costs, or positive for goal
+%achievement--- received for executing action $\action_i$ in state
+%$\state_i$. Finally, it is worth clarifying that in practise we do not
+%artificially limit the length of traces that the sequential planner
+%can consider. Moreover, for problems we consider there is always a
+%finite optimal trace.
 
-\subsection{Decision-Theoretic Planning in Abstractions}
+\subsection{DT Sessions}
 
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
+During a sequential session, when an action is scheduled whose outcome
+is uncertain according to the underlying belief-state, the planner
+switches to a DT session. That plans for {\em small} abstract
+processes defined according to the action that triggered the DT
+session, the assumptive actions in the proceeding trace, and the
+current belief-state. Targeted sensing is encouraged by augmenting the
+reward model to reflect a heuristic value of knowing the truth about
+assumptions. In detail, all rewards from the underlying problem are
+retained. Additionally, for each relevant assumptive action
+$\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
+action} $\assumptiveDT{i}$ so that for all states $\state$:
 
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
+\vspace{-1ex}
+\tiny
 \[
@@
 \]
+\normalsize
 
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
+\vspace{-1ex}
+
+\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a
+small positive (negative) numeric quantity which captures
+the utility the agent receives for correctly (incorrectly) rejecting
+an assumption.
 %%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
+In terms of action physics, a disconfirm action can only be executed
+once, and otherwise is modelled as a self-transformation.
+%%
+We only consider {\em relevant} assumptions when constructing the
+abstract model.  If \switchAction\ is the action that switched the
+system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
+relevant} if it is necessary for the outcome of \switchAction\ to be
+determined.  Continuing our simplified example, if the trace is:
 
+\tiny
 \[
@@
 \actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
+(\pp{look}~\pp{box}~\pp{office});
+(\pp{look}~\pp{cup}~\pp{kitchen});\\
+(\pp{report}~\pp{box}~\pp{office}); 
+(\pp{report}~\pp{cup}~\pp{kitchen})
 \end{array}
 \]
+\normalsize
 
 \noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
+$(\pp{look}~\pp{box}~\pp{office})$, we have that
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
+relevant for this session, and therefore we exclude the corresponding
+disconfirm action from the abstract decision process.
 
-We also include another propositionally trivial action
-$\actions.\poss(\switchAction)$ with the reward property:
+Given \switchAction, we also include another once-only self-transition
+action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
+with the reward property:
 
@@
 
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
+\noindent Here, \poss(\switchAction) is the set of positive literals
+that comprise the precondition of \switchAction. Execution of
+either a disconfirmation or the confirmation action returns control to
+a sequential session, which then continues from the underlying
+belief-state.
+
+Turning to the detail of (dis-)confirmation rewards, in our integrated
+system these are sourced from a motivational subsystem. In this paper,
+for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a small positive
+constant, and have $\hat\$(x)= - \$(x)(1 - \prob) /
+\prob$ where $\prob$ is the probability that $x$ is true. For
 $\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
+\$(x)\prob/(1-\prob)$.
 
-%%  We make no assumptions here about how they are
-%% derived.
+\Omit{
+%%
+Finally, 
+If an
+assumption was rejected, we prohibit that sequential session from
+making it again.
+}
 
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
+%%  the size of the state space
+%% is determined by two factors: the number of non-zero states in the
+%% initial belief-state, and the number of actions. Determining
+%% action reachability is intractable, as is determining their relevance
+%% to goal achievement.
 
+%% As we are aiming to keep the decision theoretic session fast, we try
+%% to keep the state space of the abstract problem as small as
+%% possible. In a deterministic-action POMDP, the state space size is
+%% determined by two factors: the number of non-zero states in the
+%% initial belief-state and the number of executable actions. As
+%% determining the relevance of an action for a given goal in in general
+%% a hard problem, we concentrate on keeping the number of uncertain
+%% facts as low as possible without losing important information about
+%% the sensing goal.
+
+In order to guarantee fast DT sessions, planning occurs for a small
+abstract problem. In our setting, where the underlying decision
+process has a deterministic action model, a significant cause of
+problem largeness is due to the number of states that occur with
+non-zero probability in the underlying belief-state. We construct an
+abstract belief-state that is chiefly characterised by facts related
+to the current trace. In that abstraction, instances of operator or
+sense declarations whose specification mentions an excluded fact are
+not available.
+%%
+In detail, we first construct an $(\pp{:init})$ declaration which
+contains the root term, and relevant assumptions from the trace---e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
+diamonds are probabilistic terms, and circles are atomic and/or
+conjunctive.
+%%
+For each excluded atom, we compute the {\em entropy} of the {\em
+relevant}
+assumptions, {\em conditional} on that atom. Intuitively, lower
+entropy indicates an atom gives better information about
+assumptions. Facts are iteratively added to the belief-state in
+increasing order according to that measure until the number of
+abstract states in the initial belief reach predefined limit (Fig.\ref{fig:abstraction-c}).
+
+%%  or no more atoms that reduce the entropy
+%% can be added
+
+\vspace{-.5cm}
 \begin{figure}[h!]
@@
 
-  \caption{Generating the abstract belief space.}
+  \caption{Abstract belief-state, and refinement.}
 \label{fig:abstraction}
@@
 
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
+% When limiting the number of actions, we aim to include those that are
+% possibly required to a) reach the goal or b) to directly observe an
+% atom that is part of the abstract belief state. In general, checking
+% this for an action is \textsc{PSPACE} complete for \textsc{STRIPS}
+% planning.
 
-% \noindent and therefore, the belief-state $\bstate_0$ is:
+% We approximate the set of relevant actions as follows: Let
+% \relevantAtoms\ be the set of atoms that either occur in the goal
+% condition \poss(\switchAction) or in a precondition \poss($a$) of an
+% action $a$ that can directly sense any atom in the abstract belief
+% state. We construct a \emph{relaxed planning graph} with
+% \relevantAtoms\ as goal atoms from which we extract a relaxed
+% plan.
 
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
+%% In the first step, we remove all facts that are not part of an
+%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
+%% would proceed in an abstraction of the environment that does not
+%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
+%% %%
+%% In a second step, we iteratively refine the relaxed declaration by
+%% adding terms from the original statement of $(\pp{:init})$ while the
+%% number of abstract states in $\bstate_0$ that occur with non-zero
+%% probability according to that refined declaration remains of a
+%% practicable size. In detail,
 

### 2011-03-16 subarchitectures/vision.sa/branches/post-review-yr2-11533/src/c++/vision/components/VisualLearner/test/gy2report/gy2article.tex
gy2article: report about testing

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
% \usepackage{amsmath}

\begin{document}

\section{Automated testing}

In this test we are teaching the robot about shapes and colours. The automated
tutor places objects on the table one by one. When an object is on the table
the tutor first describes the shape and then the color of the object. The tutor
waits until the learning task for each description is completed. If the tutor
doesn't know an attribute, he says "hi" and waits until the robot responds.
The automated tutor is implemented with a state machine.

\subsection{The first automated tutor}

The basic state machine is shown in figure \ref{fig:simpleStates}. There are
several problems related to this state machine. They are mostly related to the
presence of entries in various working memories.

\begin{figure}
\includegraphics[width=0.8\textwidth]{xdata/simple_states.pdf}
\label{fig:simpleStates}
\caption{Basic state machine for the automated tutor}
\end{figure}

\subsubsection*{\em opDescribe, stWaitForResponse }

The tutor describes the object that is currently in the scene and waits for the
robot to respond. Because of a too short timeout in the {\em Abducer} component
the response may never come. When the {\em Abducer} fails this way, it never
works again and the system has to be restarted. We solved this by increasing
the timeouts from $250ms$ to $5s$ in the following dialogue components:
refresolve, cint-recognizer, cint-realizer.

\subsubsection*{\em stTeaching, stWaitForResponse, stWaitForLearningTask}

VisualObjects appear and disappear while the robot is processing tutor's
instructions. When a VisualObjects disappears from the WM, the planner usually
creates a plan that references an invalid belief. Such a plan remains in a {\tt
PENDING} state. The valid new plans that are created afterwards are in {\tt
SURFACED} state but are never {\tt ACTIVATED}.

The VisualObjects appear and disappear even though the visual subsystem is
looking at a constant image pair. The effect is most likely related to the
RANSAC algorithm in the {\em PlanePopout} component.

\subsubsection*{\em opUnloadScene, opLoadScene}

After an object has been learned, the VisualLearningTask is overwritten. This
is a signal to the tutor that he can continue with the next object. The system
presents an empty scene, waits that all VisualObjects disappear. Then it
presents a scene with the next object.

Although the previous learning task has been completed, the change in the
contents of the set of VisualObjects in the WM locks the planner as in the
previous case.

\subsection{The improved tutor}

To improve the automated tutor we added some more states to the state machine
and most importantly we added some timeouts and pauses. The complete state
machine is shown in figure \ref{fig:allStates}. We also extended the system so
that it doesn't change the set of VisualObjects in the WM while the object
descriptions are being processed. In the state machine this is the part between
the states {\em stObjectOn} and {\em stUnloadScene}.

\begin{figure}
\includegraphics[width=1.0\textwidth]{xdata/states.pdf}
\label{fig:allStates}
\caption{Complete state machine for the automated tutor}
\end{figure}

With this improved state machine we can run the system for about 2 hours while
we present approximately 200 objects to the robot. After two hours the robot
just stops responding although all the components are still running. Most often
the system gets stuck in the {\em stWaitForLearningTask} state or presents new
objects without ever receiving a LearningTask.  We suspect that this might be
linked to the increasing memory requirements of the system which eventually
runs out of it.

We have 1200 image pairs in the database. Some (less than 10) of the image
pairs cause the components {\em PlanePopout} or {\em SOIFilter} to crash. In
average we can run the system for about 100 objects before we have to restart
it.

We also made an interesting observation in the {\em stWaitForLearningTask}
state. This state is reached $2$ times while the object is on the scene. After
the tutor describes the shape, the first learning task is created after $3-5$
seconds. When the learning is complete the tutor then waits for $2s$ and
describes the color of the object.  The interesting thing about the second
LearningTask task is that it's time to appear increases with the system's
running time. Initially it is the same as for the first LearningTask. After
$4000s$ it increases to $7s$ and after $7000s$ it takes $12s$ for it to appear.
Finally the system stops resonding as described above.

\end{document}

### 2011-03-16 code/subarchitectures/vision.sa/branches/post-review-yr2-11533/src/c++/vision/components/VisualLearner/test/gy2report/gy2article.tex
gy2 test report update

--- 2011-03-16 subarchitectures/vision.sa/branches/post-review-yr2-11533/src/c++/vision/components/VisualLearner/test/gy2report/gy2article.tex
@@
 
+When the state machine comes into the state {\em stObjectOn} we wait for $5s$
+so that the scene can become stable. Sometimes more than one VisualObject is
+created for a presented object. In some other cases the object is not well
+defined and can disappear. We want to make sure the scene is stable before we
+prevent the changes of the set of VisualObjects in the WM.
+
+In the state {\em stUnloadScene} the system waits for another $5s$ before it
+replaces the scene with an empty one. Without this timeout, the planner would
+lock again.
+
 \begin{figure}
@@
 
-We have 1200 image pairs in the database. Some (less than 10) of the image
-pairs cause the components {\em PlanePopout} or {\em SOIFilter} to crash. In
-average we can run the system for about 100 objects before we have to restart
-it.
-
 We also made an interesting observation in the {\em stWaitForLearningTask}
@@
 $4000s$ it increases to $7s$ and after $7000s$ it takes $12s$ for it to appear.
-Finally the system stops resonding as described above.
+Finally the system stops resonding as described above.  The increasing time
+could be related to the increasing number of messages from the planner:
+``Pointer to belief xxx cannot be resolved''.
+
+\subsection{Other crashes}
+
+The c++ server sometimes crashes right after the message ``Planner has
+attempted to overwrite an outdated WM entry. WMA:
+XX:goal.scheduler:data:planner.sa.  Local version: 5.  WM version: 6''.
+
+Some of the image pairs (less than 10 out of $1200$ in the database) cause the
+components {\em PlanePopout} or {\em SOIFilter} to crash.
 

### 2011-03-16 code/subarchitectures/vision.sa/branches/post-review-yr2-11533/src/c++/vision/components/VisualLearner/test/gy2report/gy2article.tex
gy2 report, output log added

@@
 
+\subsection{Example log}
+
+This is a part of the log file that shows the state transitions while the robot
+is learning one object (approximately $50s$).
+
+\begin{verbatim}
+[INFO gy2article: 4218.43 NEW STATE stTableEmpty]
+[INFO gy2article: 4220.43 SCENE 357: 1081vl]
+[INFO gy2article: 4220.44 **********]
+[INFO gy2article: 4220.44 NEW STATE stWaitToAppear (to=20s)]
+[INFO gy2article: 4222.45 STATE stWaitToAppear]
+[INFO gy2article: 4226.47 **********]
+[INFO gy2article: 4226.47 NEW STATE stObjectOn]
+[INFO analyzer: blockVisualObjectUpdates enable=true]
+[INFO gy2article: 4231.47 **********]
+[INFO gy2article: 4231.47 NEW STATE stTeaching (to=10s)]
+[INFO SOI.filter: blockProtoObjectUpdates enable=true]
+[INFO gy2article: 4232.47 Tutor: this is elongated]
+[INFO gy2article: 4232.47 **********]
+[INFO gy2article: 4232.47 NEW STATE stWaitForResponse (to=15s)]
+[INFO gy2article: 4234.48 STATE stWaitForResponse]
+[INFO gy2article: 4234.48 Robot says: thank you]
+[INFO gy2article: 4234.48 **********]
+[INFO gy2article: 4234.48 NEW STATE stWaitForLearningTask (to=120s)]
+[INFO gy2article: 4237.06 Learnig task added: [WMA: vision.sa:o2:J]]
+[INFO gy2article: 4240.21 Learnig task changed: [WMA: vision.sa:o2:J]]
+[INFO gy2article: 4240.21 A Learning Task was detected.]
+[INFO gy2article: 4240.21 **********]
+[INFO gy2article: 4240.21 NEW STATE stTeaching (to=10s)]
+[INFO gy2article: 4241.22 Tutor: this is black]
+[INFO gy2article: 4241.22 **********]
+[INFO gy2article: 4241.22 NEW STATE stWaitForResponse (to=15s)]
+[INFO gy2article: 4243.23 Robot says: thank you]
+[INFO gy2article: 4243.23 **********]
+[INFO gy2article: 4243.23 NEW STATE stWaitForLearningTask (to=120s)]
+[INFO gy2article: 4249.55 Learnig task added: [WMA: vision.sa:p2:J]]
+[INFO gy2article: 4256.04 Learnig task changed: [WMA: vision.sa:p2:J]]
+[INFO gy2article: 4256.05 A Learning Task was detected.]
+[INFO gy2article: 4256.05 **********]
+[INFO gy2article: 4256.05 NEW STATE stTeaching (to=10s)]
+[INFO gy2article: 4257.06 **********]
+[INFO gy2article: 4257.06 NEW STATE stEndOfTeaching (to=15s)]
+[INFO gy2article: 4257.08 **********]
+[INFO gy2article: 4257.08 NEW STATE stUnloadScene]
+[INFO analyzer: blockVisualObjectUpdates enable=false]
+[INFO gy2article: 4262.10 SCENE: empty]
+[INFO SOI.filter: blockProtoObjectUpdates enable=false]
+[INFO gy2article: 4262.10 **********]
+[INFO gy2article: 4262.10 NEW STATE stWaitToDisappear (to=20s)]
+[INFO gy2article: 4264.11 STATE stWaitToDisappear]
+[INFO gy2article: 4268.13 **********]
+[INFO gy2article: 4268.13 NEW STATE stTableEmpty]
+\end{verbatim}
+
 \end{document}

### 2011-03-16 subarchitectures/vision.sa/branches/post-review-yr2-11533/src/c++/vision/components/VisualLearner/test/gy2report/gy2article.tex
report corrected

@@
 creates a plan that references an invalid belief. Such a plan remains in a {\tt
-PENDING} state. The valid new plans that are created afterwards are in {\tt
-SURFACED} state but are never {\tt ACTIVATED}.
+PENDING} state. The new goals that are created afterwards are in {\tt SURFACED}
+state but are never {\tt ACTIVATED}. No new plans are created.
 
@@
 $4000s$ it increases to $7s$ and after $7000s$ it takes $12s$ for it to appear.
-Finally the system stops resonding as described above.
+Finally the system stops resonding as described above.  The increasing time
+could be related to the increasing number of messages from the planner:
+``Pointer to belief xxx cannot be resolved''.
+
+\subsection{Other crashes}
+
+The c++ server sometimes crashes right after the message ``Planner has
+attempted to overwrite an outdated WM entry. WMA:
+XX:goal.scheduler:data:planner.sa.  Local version: 5.  WM version: 6''.
+
+Some of the image pairs (less than 10 out of $1200$ in the database) cause the
+components {\em PlanePopout} or {\em SOIFilter} to crash.
+
+\subsection{Example log}
+
+This is a part of the log file that shows the state transitions while the robot
+is learning one object (approximately $50s$).
+
+\begin{verbatim}
+[INFO gy2article: 4218.43 NEW STATE stTableEmpty]
+[INFO gy2article: 4220.43 SCENE 357: 1081vl]
+[INFO gy2article: 4220.44 **********]
+[INFO gy2article: 4220.44 NEW STATE stWaitToAppear (to=20s)]
+[INFO gy2article: 4222.45 STATE stWaitToAppear]
+[INFO gy2article: 4226.47 **********]
+[INFO gy2article: 4226.47 NEW STATE stObjectOn]
+[INFO analyzer: blockVisualObjectUpdates enable=true]
+[INFO gy2article: 4231.47 **********]
+[INFO gy2article: 4231.47 NEW STATE stTeaching (to=10s)]
+[INFO SOI.filter: blockProtoObjectUpdates enable=true]
+[INFO gy2article: 4232.47 Tutor: this is elongated]
+[INFO gy2article: 4232.47 **********]
+[INFO gy2article: 4232.47 NEW STATE stWaitForResponse (to=15s)]
+[INFO gy2article: 4234.48 STATE stWaitForResponse]
+[INFO gy2article: 4234.48 Robot says: thank you]
+[INFO gy2article: 4234.48 **********]
+[INFO gy2article: 4234.48 NEW STATE stWaitForLearningTask (to=120s)]
+[INFO gy2article: 4237.06 Learnig task added: [WMA: vision.sa:o2:J]]
+[INFO gy2article: 4240.21 Learnig task changed: [WMA: vision.sa:o2:J]]
+[INFO gy2article: 4240.21 A Learning Task was detected.]
+[INFO gy2article: 4240.21 **********]
+[INFO gy2article: 4240.21 NEW STATE stTeaching (to=10s)]
+[INFO gy2article: 4241.22 Tutor: this is black]
+[INFO gy2article: 4241.22 **********]
+[INFO gy2article: 4241.22 NEW STATE stWaitForResponse (to=15s)]
+[INFO gy2article: 4243.23 Robot says: thank you]
+[INFO gy2article: 4243.23 **********]
+[INFO gy2article: 4243.23 NEW STATE stWaitForLearningTask (to=120s)]
+[INFO gy2article: 4249.55 Learnig task added: [WMA: vision.sa:p2:J]]
+[INFO gy2article: 4256.04 Learnig task changed: [WMA: vision.sa:p2:J]]
+[INFO gy2article: 4256.05 A Learning Task was detected.]
+[INFO gy2article: 4256.05 **********]
+[INFO gy2article: 4256.05 NEW STATE stTeaching (to=10s)]
+[INFO gy2article: 4257.06 **********]
+[INFO gy2article: 4257.06 NEW STATE stEndOfTeaching (to=15s)]
+[INFO gy2article: 4257.08 **********]
+[INFO gy2article: 4257.08 NEW STATE stUnloadScene]
+[INFO analyzer: blockVisualObjectUpdates enable=false]
+[INFO gy2article: 4262.10 SCENE: empty]
+[INFO SOI.filter: blockProtoObjectUpdates enable=false]
+[INFO gy2article: 4262.10 **********]
+[INFO gy2article: 4262.10 NEW STATE stWaitToDisappear (to=20s)]
+[INFO gy2article: 4264.11 STATE stWaitToDisappear]
+[INFO gy2article: 4268.13 **********]
+[INFO gy2article: 4268.13 NEW STATE stTableEmpty]
+\end{verbatim}
 

### 2011-03-31 code/subarchitectures/vision.sa/branches/post-review-yr2-11533/src/c++/vision/components/VisualLearner/test/gy2report/gy2article.tex
title, figure size
george tester parameters

@@
 \begin{document}
+
+\title{Evaluation of the George Year 2 System}
+\author{Marko Mahnič\\ University of Ljubljana}
+\date{2011-03-16}
+
+\maketitle
 
@@
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{xdata/simple_states.pdf}
+\includegraphics[width=0.6\textwidth]{xdata/simple_states.pdf}
 \label{fig:simpleStates}
@@
 
-Although the previous learning task has been completed, the change in the
-contents of the set of VisualObjects in the WM locks the planner as in the
-previous case.
+The change in the contents of the set of VisualObjects in the WM locks the
+planner again: when a teaching goal for the new object appears, the planner
+creates a plan that includes a belief about an object that has just
+disappeared. This results in a plan in a {\tt PENDING} state.
 
@@
 components {\em PlanePopout} or {\em SOIFilter} to crash.
+
+On rare occasions Matlab VisualRecognizer crashes when rendering the
+recognition results.
 

### 2011-04-21 subarchitectures/planner.sa/trunk/src/latex/AAAI11/prop-dtp.tex
Hi Moritz, I have merged the material I think we should have added to the paper...

--- 2011-01-19 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI/ijcai11.tex
@@
-%%%% ijcai11.tex
-
-\typeout{IJCAI Switching Planner Submission}
-
-% These are the instructions for authors for IJCAI-11.
-% They are the same as the ones for IJCAI-07 with superficical wording
-%   changes only.
-
-\documentclass{article}
-% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
-\usepackage{ijcai11}
-
-% Use the postscript times font!
-\usepackage{times}
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\usepackage{subfig}
-\usepackage{tikz}
-\usetikzlibrary{shapes}
-\usetikzlibrary{positioning}
-
-\usepackage{graphicx}
-\graphicspath{{plots/}}
-\include{macros}%
-
-\title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-Realistic robot planning problems in uncertain environments often
-require achieving tasks while also finding out about the
-world. Because the world state cannot be observed directly, these
-problems are naturally represented as partially observable Markov
-decision problems (POMDPs). However, these are typically intractable
-for realistic problems.
-%%
-We present a \emph{switching planner} that employs fast sequential
-planning to decide on the overall strategy, and uses a
-decision-theoretic planner to solve the subproblems where partial
-observability will significantly impact the quality of the plan. We
-demonstrate the approach in a realistic robot exploration domain.
-
-\end{abstract}
-
-\section{Introduction}
-
-Planning for agents acting in partially observable environments with
-stochastic actions is extremely computationally
-expensive~\cite{mdp-complexity}. Even relatively small partially
-observable Markov decision problems (POMDPs), represented by tens of
-propositional variables, have state spaces far larger than existing
-optimal or close to optimal planners can handle. For the easier
-problem of planning in completely observable stochastic domains, an
-alternative has been to use classical sequential planners and
-replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
-Execution monitoring tracks properties of the state that determine
-whether further planning is required. For partially observable
-domains, only probabilistic information about such properties will, in
-general, be available.
-
-%% However, these approaches rely
-%% crucially on being able to determine properties of the current state
-%% in order to decide to replan. For partially observable domains, only
-%% probabilistic information about these properties will in general be
-%% available.
-
-This paper describes the planning component for a robotic system that
-continuously deliberates in a stochastic dynamic environment in order
-to achieve objectives set by the user, and acquire knowledge about its
-surroundings. This domain features partial observability, particularly
-because the state is not perfectly known, and state information is
-gained by the robot through the use of unreliable sensing
-actions. 
-%Using a propositionally factored state representation, for
-For
-interesting sized tasks the corresponding POMDP model is too large for a
-POMDP solver to be applied directly. Inspired by the replanning
-approaches used successfully for MDP
-planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose a continual
-planning approach that uses a classical planning system to compute a
-reasonably valuable trace in the model, and then uses a
-decision-theoretic planner on small subproblems where reasoning about
-observations might be useful. We refer to the complete system as a
-{\em switching continual planner}. The approach is not optimal,
-particularly as it relies on the results of sequential planning
-directly. It does nevertheless perform better than a purely sequential
-replanner, and is fast enough to be used for real-time decision-making
-on a mobile robot.
-
-%% Switching gives us some important advantages in our robotic
-%% domain. First, the significant source of uncertainty in the domain is
-%% the unreliability of observations, and this approach is particularly
-%% suited to problems that combine a deterministic task planning problem
-%% with observations as the decision-theoretic planner is often invoked
-%% to determine a characteristic of the underlying state. Secondly,
-%% replanning makes our system robust to changing objectives and the
-%% discovery of new facts about the world.
-
-%% As an illustration, consider a robot searching for a box that it knows
-%% is often found in offices. The sequential planner might produce a plan
-%% that assumes that {\sc Room a} is an office, that the box is in {\sc
-%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
-%% these assumptions, the plan might be to go to {\sc Room a}, go to
-%% location 1 and search for the box. When the plan gets executed and the
-%% search begins the decision-theoretic planner is called since searching
-%% involves making observations. The decision-theoretic planner might
-%% retract the last assumption but leave the others, and build a plan to
-%% search the whole of {\sc Room a} for the box. If this plan fails to
-%% find the box, it will disprove the assumption that the box is in {\sc
-%% Room a}, which will then lead to replanning by the sequential planner
-%% based on the newly discovered knowledge.
-
-Our system is domain independent, taking domain and problem
-descriptions in a first-order declarative language we have developed,
-called the decision-theoretic planning domain definition language
-(DTPDDL). From these descriptions of POMDPs, we automatically
-construct a deterministic model suitable for sequential planners. That
-model includes actions which correspond to making assumptions about
-the values of imperfectly known state variables. Assumptions scheduled
-by the sequential system are used to propose a pragmatic abstract
-belief-state space to the decision-theoretic system, and to modify the
-reward function, so that system might pursue sensing related to those
-assumptions.
-
-%%  which can operate
-%% with the original representation of the domain.
-
-%% The sequential planner we use is based
-%% on \fastdownward~\cite{fast-downward}. We add the capability to replan
-%% ~\cite{brenner:nebel:jaamas09}, and also allow agent
-%% knowledge to be represented explicitly, so we can write actions that
-%% gain the agent knowledge. LEASE CITE MICHAEL's WORK AS THE BASIS FOR
-%% THIS. For the decision-theoretic planner we have implemented our own
-%% forward search in the belief-state space.
-
-The switching planner provides important advantages in our
-mobile robot domain. First, the significant source of uncertainty in the
-domain is the unreliability of observations and this approach is
-particularly suited to problems that combine a deterministic task
-planning problem with observations because the decision-theoretic planner
-is only invoked to deal with state uncertainty.
-%determine a characteristic of the underlying state. 
-Secondly, replanning makes the system robust to
-changing objectives and discoveries about the world,
-both of which feature in our domain.
-
-The remainder of the paper proceeds as follows. We begin by describing
-POMDPs with a propositionally factored representation of states and
-observations. We also outline the relationship between this and a flat
-POMDP representation. Next we describe our DTPDDL language with
-example declarations from a mobile robot exploration task. Then we
-describe the switching planner, including how the sequential domain is
-produced and a plan is generated, and how this then provides the input
-for the decision-theoretic planner. Finally we provide an evaluation
-of the system on the example robot domain, and finish with a short
-survey of related work and some concluding remarks.
-
-%To accompany
-%that language we have implemented an information-state
-%\laostar\ procedure for solving problems expressed in DTPDDL.
-
-%%
-
-%Have a good understanding of the contents and function of rooms, as
-%well as the linguistic referrents to rooms, widgets, and their visual
-%qualities.
-
-%Having confidence in its beliefs about the linguistic terms for
-%spaces, and the function of those spaces.
-
-%sometimes difficult to predict, with exogenous events, such as the
-%changing of the objective,
-
-%the speed and scalability of software for sequential planning in
-%deterministic
-
-%% We developed a first-order declarative language, called DTPDDL, for
-%% describing domains of planning problems that correspond to POMDPs.  To
-%% accompany that language we have implemented an information-state
-%% \laostar\ procedure for solving problems expressed in DTPDDL.
-
-%% We have extended MAPSIM to parse and simulate DTPDDL problems, and
-%% modified \fastdownward\ so that it can find useful sequential executions
-%% given DTPDDL models of the problem at hand.
-
-%%  called DTPDDL, along
-%% the lines of PDDL for the partially observable case, an \laostar\
-%% solution procedure, a determinisation of the DTPDDL problem in MAPL,
-%% and modify the \fastdownward\ system to find high quality sequential plans
-%% .
-
-%% We model the environment as a partially observable Markov decision
-%% process. Although planning in that model is undecidable in
-%% general~\cite{}, an optimal finite-horizon plan corresponds to a
-%% contingent plan, that is a function mapping action-observation
-%% histories to actions.
-
-%% our continual planner is reactive, replanning whenever the underlying
-%% domain and problem models change. For example, replanning occurs if
-%% the motivational component alters the objectives, and if an assumed
-%% outcome of a sensing action is not realised.
-
-%% brittle sensing model,
-
-%% the evaluation of a fluent at a state is either known. Moreover, there
-%% is a sensing process that run-time variables  after-which
-
-%% For $\prop \in \state$ we say proposition $\prop$ characterises state
-%% $\state$. There is always a unique starting state $\state_0$. The goal
-%% $\goal$ is a set of propositions, and we say that state $\state$ is a
-%% goal state iff $\goal \subseteq \state$.
-
-%%  To keep this exposition
-%% simple, for any two distinct actions $\stochAction_i \neq
-%% \stochAction_j$, if outcome $\detAct$ is a possibility for
-%% $\stochAction_i$ then it cannot also be a possibility for
-%% $\stochAction_j$ -- i.e., if $\detAct \in
-%% \detActions(\stochAction_i)$ then $\detAct \not\in
-%% \detActions(\stochAction_j)$.
-
-%% The solution to a probabilistic planning problem is a contingency
-%% plan. This consists of an assignment of actions to states at each
-%% discrete timestep up to the planning horizon $n$. The optimal
-%% contingency plan is one which prescribes actions to states that
-%% maximise the probability that the goal is achieved within $n$ steps
-%% from the starting state $\state_0$. For the purposes of this paper
-%% we say a plan fails, i.e. achieves the goal with probability $0$, in
-%% situations where it does not prescribe an action. Computing the
-%% optimal plan for a problem is computationally intractable, and an
-%% important direction for research in the field is to develop
-%% heuristic mechanisms for generating small linear plans
-%% quickly~\cite{littman:etal:98}
-
-\section{Propositional Decision-Theoretic Planning}
-
 %% {\em runtime variables}
@@
 Boolean condition--- with probability $1$.  Generally, whatever the
-structure of a plan, its value corresponds to the expected cumulative
-reward:
+plan format, its value corresponds to the expected reward:
 
@@
 %% both sequential (e.g., conformant) and contingent plan formats.
-
-\section{Decision-Theoretic PDDL}
-
-The modelling language of choice for planning in probabilistic
-problems is the Probabilistic Planning Domain Definition
-Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
-used in all three of the International Planning Competitions since
-2004. A variation on PDDL for describing domains with stochastic
-actions and uncertain starting configurations, PPDDL is a declarative
-first-order language that facilitates factored descriptions of domains
-and problems. There are straightforward compilations from problems
-expressed in PPDDL to propositional representations amenable
-to state-of-the-art planners.
-
-Because PPDDL cannot model domains that feature partial observability,
-we develop an extension we call {\em Decision-Theoretic
-(DT)PDDL}. This can express probabilistic models of the sensing
-consequences of acting, to quantitatively capture unreliability in
-perception. That expressive power is achieved by incorporating
-perceptual analogues of fluent, predicate, and action definitions. In
-detail, we have declarations of state characterising predicate and
-fluent symbols according to the PPDDL syntax. In addition, we allow
-two other declarations, for perceptual predicates and fluents
-respectively. For example, suppose our robot is tasked with exploring
-{\em locations} in order to identify the whereabouts of a {\em
-visual-object}. We must describe state and perceptual facts that model
-the {\em true}, resp. perceived, locations of objects. In DTPDDL,
-these declarations appear as:
-
-\small
-\begin{tabtt}
-(\=:functions  ;; state fluents\\
-  \> (is-in ?v - visual-object) - location )\\
-(:observable-functions  ;; perceptual fluents\\
-  \> (o-is-in ?v - visual-object) - location )
-\end{tabtt}
-\normalsize
-
-\noindent To model the sensing capabilities of the agent, we have
-operator-like declarations, with preconditions expressed using state
-predicate and action function symbols, and uniformly positive effects
-over perceptual predicates. A sense declaration corresponds to a
-lifted description of proposition {\em senses}. Whereas the effects of
-an {\em operator} schema describe how states change under application
-of actions, the effects of a {\em sense} schema are perceptual,
-specifying the composition of an observation following the execution
-of an action.
-
-%% Also, and without a loss of generality, we find it
-%% convenient to separate the actional precondition from the state
-%% precondition.
-
-Making the above ideas concrete with an example, suppose a robot
-called $\pp{DORA}$ is able to look for a visual-object, such as a
-$\pp{box}$, at a given place. We can model that deterministic action
-using the following operator schema:
-
-\small
-\begin{tabtt}
-(\=:action look-for-object \+ \\
-   :parameters (\=?r - robot ?v - visual-object\\
-   \> ?l - location) \\
-   :precondition (and (= (is-in ?r) ?l) ) \\
-   :effect (and (assign (reward) -3) ) ) \\
-\end{tabtt}
-\normalsize
-
-\noindent In a slight departure from PPDDL, we suppose that
-$\pp{look-for-object}$ has an effect on the state. That is, its
-execution incurs an instantaneous penalty, $3$, that corresponds to
-the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
-reward} is a {\em reserved word}, and occurs in {\em increased} and
-{\em decreased} terms in operator schemata. In that setting, reward is
-accumulated, whereas in DTPDDL it is instantaneous.}
-
-Completing the example, we model the sensing that results from
-instantiating the $\pp{look-for-object}$ operator:
-
-\small
-\begin{tabtt}
-(\= :sense vision \+\\
- :parameters \= (\= ?r - robot ?v - visual-object\\
- \>\>  ?l - location) \\
- :execution \> ( \> look-for-object ?r ?v ?l) \\
- :precondition (and (= (is-in ?r) ?l) ) \\
- :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
-   \> \> (probabilistic 0.8 \\
-   \>  \>(= (o-is-in ?v) ?l))) \\
-  \> (when (not (= (is-in ?v) ?l)) \\
-   \>  \> (probabilistic 0.1 \\
-   \>  \> (= (o-is-in ?v) ?l))))) \\
-\end{tabtt}
-\normalsize
-
-\noindent Here, the {\em execution} declaration is a lifted description of
-the sense action-precondition -- i.e,
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
-$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
-we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
-order to fully demonstrate the syntax. Interpreting the above schema,
-if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
-executed, there is a $0.8$ chance of perceiving the box if it is in
-the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.
-
-The syntax and semantics for describing initial state distributions in
-DTPDDL is taken verbatim from PPDDL. It is useful to present that
-factored tree-like structure here, in order to aid us when discussing
-our switching continual planner. The distribution is expressed in a
-tree-like structure of terms. Each term is either: (1) atomic, e.g., a
-state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
-probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
-(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
-probabilistic and atomic terms. The root term is always conjunctive,
-and the leaves are atomic. For example, the starting distribution for
-$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
-$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
-i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
-$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
-that contain a single atomic subterm.}
-
-\small
-\begin{tabtt}
-(\=:init (= (is-in DORA) kitchen) \+ \\
-       (probabilistic \=.8 (= (is-in box) office)  \\
-		      \>.2 (= (is-in box) kitchen)) \\
-       (probabilistic .3 (= (is-in cup) office)  \\
-		      \>.7 (= (is-in cup) kitchen))) \\
-\end{tabtt}
-\normalsize
-
-\noindent The interpretation of such an expression can be given
-by a {\em visitation} of terms: An atom is {\em visited} iff
-its conjunctive parent is visited, and a conjunctive term is visited
-iff all its immediate subterms are visited. A probabilistic term is
-visited iff its conjunctive parent is visited, and exactly one of its
-subterms, $T_i$, is visited. Each visitation of the root term
-according to this recursive definition encapsulates a starting state,
-along with the probability that it occurs. The former corresponds to the
-union of all visited atoms, and the latter corresponds to the product
-of $\prob_i$ entries on the visited subterms of probabilistic
-elements. Making these ideas concrete, our example yields the
-following flat state distribution:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-\hline
-\end{tabular}
-\normalsize
-
-From hereon, to simplify the discussion (and implementation)
-we shall restrict our attention to POMDPs with deterministic
-actions. We note that POMDPs with stochastic
-actions can be compiled into equivalent deterministic-action POMDPs,
-where all the original action uncertainty is expressed in the
-starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
-that of finite-horizon planning, such a compilation yields a finite
-flat representation of the original POMDP, only with deterministic
-actions.
-
-\section{Switching Continual Planner}
-
-We describe our approach that switches between sequential and
-contingent sessions. As a continual planning approach, it proceeds by
-interleaving planning and execution in a deterministic-action POMDP
-described in DTPDDL. During a sequential session, planning is
-performed by a ``classical'' system,\footnote{That is, a planner
-designed to solve fully observable deterministic tasks.}  and
-execution proceeds according to the {\em trace} computed by that
-system. Taking the form of a classical plan, the trace specifies a
-sequence of POMDP actions that achieve the agent's objectives in a
-deterministic approximation, i.e., {\em determinisation}, of the
-problem at hand. More precisely, the trace is an interleaved sequence
-of POMDP actions and {\em assumptive} actions. The latter correspond
-to assumptions the planner makes about the truth value of propositions
--- e.g. that a box of cornflakes is located in the kitchen at the
-third plan step. They are called {\em applicability} assumptions if
-the trace includes an action $\action$ that is not applicable with
-probability $1$ at the belief-state
-\bstate\ that the system is projected to be in when \action\ is
-scheduled for execution, i.e., $\exists\state\in\states\;
-\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
-scheduling
-\action,  the serial planner makes an assumption about the
-observability of the precondition $\poss(\action)$.
-
-Our approach always begins with a sequential session. Non-assumptive
-actions from the trace are executed in sequence until the
-applicability of the next scheduled action is uncertain. We denote
-that action by \switchAction.  A contingent session then begins that
-tailors sensory processing by the agent to determine whether the
-assumptions made in the trace hold. 
-%%
-We add {\em disconfirm} and {\em confirm} actions to the model, so
-the session is encouraged to judge the assumptions made in the
-trace. On execution of one of those actions, control is returned to
-the sequential session that continues at the current%%{\em underlying}
-belief-state.
-%%
-%% For each assumption we add a {\em
-%% disconfirm} action to the POMDP whose execution is rewarding
-%% (resp. costly) if the assumption is false (resp. true). We also add
-%% one {\em confirm} action that is rewarding (resp. costly) to execute
-%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
-%% contingent plan applies a disconfirm action, then a new sequential
-%% session begins at the {\em underlying} belief-state. If the confirm
-%% action is executed, the sequential session is resumed, and
-%% \switchAction\ is executed. Otherwise, control rests in the continual
-%% session.
-%%
-Because contingent planning is only practical in relatively small
-POMDPs, contingent sessions plan in an abstract decision process
-determined by the current trace. This abstraction is constructed by
-first omitting all propositions that do not feature in the trace, and
-by then iteratively refining the model while the result is of a
-practicable size.
-
-Finally, whether proceeding in a sequential or contingent session, our
-continual planner maintains a factored representation of successive
-belief-states by performing belief revision according to
-Eq~\ref{eq:revision}. Moreover, our internal representation of the
-underlying belief-state corresponds more-or-less directly to an
-$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
-as the source of candidate determinisations for sequential planning,
-(2) to determine when to switch, and (3) as a mechanism to guide
-construction of an abstract process for contingent sessions.
-
-%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
-%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.
-
-%% \Omit{
-%%   Therefore, although our current implementation does not
-%%   explicitly support it, our approach generalises to POMDPs with
-%%   stochastic action models.
-%% }
-
-\subsection{Sequential Planning in POMDPs}
-
-A sequential session uses a classical planner to compute a trace. The
-latter encapsulates assumptions about: (1) the true underlying state,
-(2) how execution will progress, and (3) the possibility of the agent
-eventually holding strong beliefs about the truth values of specific
-state propositions. Here we describe the deterministic planning
-problem, derived from the DTPDDL model, that admits plans which
-correspond to traces.
-
-In a deterministic-action POMDP all the uncertainty in state and
-action is expressed in the $(\pp{:init})$ declaration. Our approach
-uses the structure of that, as it occurs in the problem description,
-to define a set of state-assumptions available to sequential planning.
-%%
-Writing \#\ if the value of a proposition is unspecified, for
-$\pp{DORA}$ we have the following assumptions:
-
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-\hline
-.24 & kitchen & office & office \\
-.06 & kitchen & kitchen & office \\
-.56 & kitchen & office & kitchen \\
-.14 & kitchen & kitchen & kitchen \\
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
-
-\noindent An assumption corresponds to a {\em
-relaxed} visitation of the root term of $(\pp{:init})$. In this
-relaxed case, a conjunctive term is visited iff its atomic subterms
-are visited, and zero or more of its immediate probabilistic subterms
-are visited. The starting state, $\state_0$, for sequential planning
-is an {\em abstract} state that captures the unique assumption which
-has probability $1$. In $\pp{DORA}$, that is:
-
-\[
-\begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
-\end{array}
-\]
-
-In order to make assumptions available to the sequential planner, we
-add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
-each element, $\prob_i (T_i)$, of each probabilistic term from
-$(\pp{:init})$. The physics of these actions is as
-follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
-$j \neq i$, has been executed from the same probabilistic term, and,
-either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
-conjunct, or it occurs in $T_k$ for some executed
-$\assumptiveS{k}$. Also, an assumption about a proposition
-$\prop$ can only be made once, and must be sequenced before any POMDP
-actions that have $\prop$ in their precondition or effect.
-%%
-In other words, an assumptive action
-must be applied to make a truth assignment to \prop, and then that
-truth assignment can be modified by the DTPDDL domain operators.
-%%
-%%
-Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
-successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$ with the
-proposition denoting the undefined fact is deleted. For example, if
-the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
-delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
-sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
-$\transProb(\state^\bot, \action, \state^\bot) = 1$.
-%%
-Lastly, we have that if $\action$ has a precondition that is not true
-with probability $\theta$, a given threshold, in the underlying belief
-$\bstate$, then immediately before it is executed blindly the system
-switches to a contingent session (described below).
-
-%% For sequential planning the operators from the POMDP model are
-%% available with semantics that accommodate the above abstraction as
-%% follows: In the deterministic model a proposition $\prop$ can be
-%% thought to have a ternary interpretation at a state \state, as either
-%% {\em true}, written $\prop\in\state$, {\em false}, written
-%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
-%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
-%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
-%% $\action\in\actions$ in the deterministic model, if
-%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
-%% or negative effect of $\action$, then $\action$ is not applicable
-%% in \state\ if $\prop\#\in\state$.
-
-%% Addressing now the switching semantics of action execution. I
-%% %% In detail, we halt the sequential
-%% session at \action\ if $\exists
-%% \state$ s.t. $\bstate(\state)>0$, 
-%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
-%% semantics of action execution there must be a $\state'$ s.t.
-%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
-%% not have been scheduled at \bstate.
-
-%% then $\reward(\state,\action)=10$, and
-%% otherwise if $\prop\not\in\state'$ then
-%% $\reward(\state',\action)=-10$.
-
-We now describe the optimisation criteria given to the classical
-planner. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action $\action_i$ does not transition to $\state^\bot$, we
-define the value of a trace
-$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
-deterministic model to equal:
-
-\begin{equation}\label{eq:tracevalue}
-V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
-\end{equation}
-
-\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
-actions from the trace, therefore has value:
-
-\[
-V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
-\reward(\state_i, \action_i),
-\]
-
-\noindent which is equal to the maximal 
-additive contribution a trace of the plan can make to
-Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
-non-zero reward is received at the first transition to a goal state,
-the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
-by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
-gradient-based reinforcement learning systems, such as William's
-REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
-under reasonable assumptions the optimal trace identifies a maximal
-gradient step from a uniformly random soft-max policy. Finally, it is
-worth clarifying that although the above discussion has been in terms
-of a trace of length $N$, there is nothing in our approach that
-artificially limits the length of candidate sequential plans.
-
-\subsection{Decision-Theoretic Planning in Abstractions}
-
-In a contingent session our switching planner solves a sensing problem
-in an abstract processes defined in terms of the assumptive actions in
-the current trace, and the $(\pp{:init})$ declaration that models the
-underlying belief-state.
-
-The focus for contingent planning is on sensing, we therefore give the
-process in this session an augmented reward model.  This reflects the
-value of performing sensing actions in the context of the trace
-proposed by the preceding sequential deliberation. First, all rewards
-from the original POMDP are retained. Then, for each $\assumptiveS{i}$
-action scheduled by the current trace, we have a {\em dual}
-$\assumptiveDT{i}$ so that $\forall\state\in\states$:
-
-\[
-\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
-\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
-\hat\$(T_i) & \pp{otherwise} \\
-\end{array}
-\]
-
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
-(negative) numeric quantity which captures the utility the agent
-receives for correctly (incorrectly) rejecting an assumption. The
-added dual actions are propositionally trivial -- i.e.,
-$\forall\state\in\states$ $\poss(\assumptiveDT{i})
-\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
-%%
-In order to reduce the number of possible rejections we omit
-assumptions that are not {\em active} with respect to the action,
-\switchAction, whose scheduled execution switched the system into
-the contingent session. An assumption $\assumptiveS{i}$ is active if
-\switchAction\ is not applicable in the sequential model unless
-$\assumptiveS{i}$ is scheduled in the trace prefix
-to \switchAction. For example, if $\pp{DORA}$'s trace is:
-
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
-(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
-(\pp{report-is-in}~\pp{box}~\pp{office}); \\
-(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
-\end{array}
-\]
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-active, and therefore exclude
-$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
-POMDP posed to the contingent session.
-
-We also include another propositionally trivial action
-$\actions.\poss(\switchAction)$ with the reward property:
-
-\[
-\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
-\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
-\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
-\end{array}
-\]
-
-For the purposes of successive switching, execution of either a dual
-action of the form $\assumptiveDT{i}$, or the
-$\actions.\poss(\switchAction)$, returns control to a sequential
-session. If a dual is executed, then the sequential planner must
-replan with $\bstate_0$ equal to the underlying
-belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
-executed, then the current sequential plan is executed until further
-sensing is scheduled, or to completion. Lastly, we note that in our
-project the rewards, i.e., functions
-$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
-contingent session are sourced from a motivational component of our
-robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
-set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
-- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
-$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
-\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
-  sensing.}
-
-%%  We make no assumptions here about how they are
-%% derived.
-
-The starting belief-state for the abstract process is given in terms
-of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
-tree which only includes terms from the {\em relaxed} visitation
-characterised by the active {\em assumptive} actions in the current
-sequential plan. Figure \ref{fig:abstraction} shows the belief-state
-of a more detailed version of our previous example. The diamonds
-represent probabilistic terms of the state description while circles
-represent conjunctive or atomic terms.
-
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  \subfloat[The initial belief-state with the assumptions made by the continual
-    planner in bold.]{
-      \label{fig:abstraction-a}
-      \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode] (cup) {} 
-          child {node [cnode] (cupp1) {}}
-          child {node [cnode] (cupp2) {}}
-        }
-      }
-      child {node[cnode] (kitchen) {}
-        child {node [pnode] (box2) {} 
-          child {node [cnode] (box2p1) {}}
-          child {node [cnode] (box2p2) {}}
-        }
-      };
-   \tiny
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of kitchen] {$\pp{kitchen}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[below=0 of cup] {$\pp{(is-in cup)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-   \node[right=0 of cupp1] {$\pp{place1}$};
-   \node[right=0 of cupp2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts until the $\bstate_0$ size limit is reached.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Generating the abstract belief space.}
-\label{fig:abstraction}
-\end{figure}
-
-%%% OLDER STUFF
-
-% \small
-% \begin{tabtt}
-% (\=:init (= (is-in DORA) kitchen) \+ \\
-%        (probabilistic \=.8 (= (is-in box) office)  \\
-% 		      \>.2 (= (is-in box) kitchen))) \\
-% \end{tabtt}
-% \normalsize
-
-% \noindent and therefore, the belief-state $\bstate_0$ is:
-
-% \small
-% \begin{tabular}{cccc}
-% \hline
-% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
-% \hline
-% .8 & kitchen & office & \# \\
-% .2 & kitchen & kitchen & \# \\
-% \hline
-% \end{tabular}
-% \normalsize
-
- % With
-% regards to abstraction, the contingent and serial sessions apply the
-% same applicability conditions ---i.e., that each effect and condition
-% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
-% problem description.
-
-In the first step, we remove all facts that are not part of an
-assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-would proceed in an abstraction of the environment that does not
-contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%%
-In a second step, we iteratively refine the relaxed declaration by
-adding terms from the original statement of $(\pp{:init})$ while the
-number of abstract states in $\bstate_0$ that occur with non-zero
-probability according to that refined declaration remains of a
-practicable size. In detail, for each excluded atomic term, we compute
-the {\em entropy} of the active assumptions, {\em conditional} on the
-corresponding proposition.  Loosely speaking, the lower the {\em
-  conditional entropy} of an atom, the more information about the
-assumptions the contingent planner expect to derive from it. Thus,
-propositions that in the first place were excluded are iteratively
-added to the problem posed to the conditional session in increasing
-order according to that measure (Fig. \ref{fig:abstraction-c}).
-
-%% In detail, for each excluded atomic term, we compute the {\em
-%% entropy} of the corresponding proposition, {\em conditional} on the
-%% active assumptions. The {\em conditional entropy} measure
-%% is higher the more uncertain that proposition is given truth
-%% assignments to propositions that are the subject of active
-%% assumptions. Propositions that in the first place were excluded are
-%% iteratively added to the problem posed to the conditional session in
-%% increasing order according to that measure.
-
-%% In particular, we compute the {\em conditional entropy} of
-%% each excluded proposition given the active assumptions. Propositions
-%% are candidates for addition in increasing order according to that
-%% measure.
-
-%% Of course, it is expensive to measure size exactly,
-%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
-%% $\propositions^\#$ is the set of propositions in the abstraction that
-%% were not assigned the {\em unknown} value \#. Terms are proposed for
-%% addition because they are probabilistic, and because they were omitted
-%% in the {\em relaxed} visitation characterised by {\em assumptive}
-%% actions in the sequential plan. In each iteration, we add one
-%% candidate term with the largest number of parent terms already in the
-%% candidate tree.
-
-\section{Evaluation}
-
-We have implemented our switching continual approach in the MAPSIM
-environment~\cite{brenner:nebel:jaamas09}. 
-% Our implementation is able
-% to use several underlying planning systems.
-We have extended MAPSIM so that it can parse DTPDDL, and perform
-successive estimation of the underlying belief-state using
-\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
-we use our own version of Fast Downward~\cite{fast-downward} for
-sequential sessions. We have extended that system to support actions
-with success probabilities. In our evaluation Fast Downward is run
-with the cyclic causal graph heuristic using an A* search or weighted
-A* (with weight 5). We use A* for easy problems and WA* for more
-difficult problems where A* is ineffective. Contingent sessions use
-our own forward search procedure.
-
-\Omit{ We then perform multiple tests with different limits for the
-  belief-space size in contingent sessions.  Higher limits should
-  cause longer planning times but be beneficial to plan quality as
-  more contingencies can be taken in to account by the POMDP planner.
-}
-
-In order to evaluate our system, we have also implemented a baseline
-approach in MAPSIM. Rather than invoking a contingent session when a
-switching action $\switchAction$ is scheduled for execution, the
-baseline simply executes one action that can trigger a sensing outcome
-determined by the precondition of $\switchAction$, and replans in the
-resulting belief-state.
-
-%% executes switching actions and replans in the
-%% resulting belief state.
-
-%% but instead of
-%% creating an observation problem for the decision theoretic planner it
-%% will just execute one sensing action -- assuming that this action will
-%% confirm its assumption.
-
-To test our approach we use a robot exploration domain based on a
-scenario from our physical robotic system. Here, a robot is exploring
-an office or living environment, and trying to report the locations of
-objects to their owner. Basic types are {\tt rooms}, {\tt places} and
-{\tt objects}. Places are topological map nodes that occur in rooms.
-Objects, and also the robot, are always located at a place. The robot
-can move around the rooms via connections between places given by a
-{\tt connected} predicate. Each room has a, possibly unknown, {\em
-  category} (e.g. kitchen, office, living room). Certain objects are
-more likely to be located in rooms of a particular category.  For
-example, a box of cornflakes is more likely to be located in the
-kitchen than the office. The robot can look for an object at a place
-by executing a {\tt look-for-object} action. Executing {\tt
-  look-for-object} results in a noisy sensing outcome, sometimes
-indicating a positive perception if the object is there. We have that
-some objects are harder to detect than others. Finally, if the robot
-is in the presence of a human, it can ask what type of room it is
-currently in, however conducting a dialogue is more costly than
-running a vision algorithm (cost of 8 vs costs of 3).
-
-%%  Therefore, an instantaneous positive detection of an object is
-%% not proof positive of it being there.
-
-We evaluated our switching planner against the baseline in several
-tasks with the number of rooms ranging from 3 to 6. We examine the
-impact of sensing reliability, experimenting using sensor models that
-are: ({\em reliable}) $0.9$ chance the object is perceived if it is
-present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
-chance is 0.4. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}. Here, in switching runs we examine
-the performance of our system where $b_0$ allocates non-zero
-probability to between 20 and 200 abstract states during contingent
-sessions. For Item~$a$-$e$ the objective is to find one or more
-objects and report their position to the user. In some of the tasks we
-configured the environment so that no plan exists, by excluding the
-desired object from the environment. Item~$f$ reports for tasks
-requiring indirect sensing, where the robot must relocate to a
-particular room category. For Item~$f$ there is no human in the
-environment, and therefore the baseline cannot be used, because there
-are no sensing schemata that directly sense room categories. In our
-evaluation we run 50 simulations in each configuration, and have a
-timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
-  experiments were conducted on a 2.3GHz AMD Opteron using one CPU
-  core.}. In our domain, rewards encapsulate a goal condition, and
-therefore we find it intuitive to report average costs of the plans
-and the success rate in problems that admit a solution (i.e., positive
-reward scaled by a constant factor).
-
-%% Because the switching utility model
-
-%% Using a satisficing, optimistic serial planner in sequential sessions
-%% makes optimising for the expected reward difficult. Only reasoning
-%% about single traces, it is optimistic wrt the remaining costs to a
-%% goal and pessimistic regarding the probability of reaching it. By
-%% setting the rewards to non-extremal values has little effect on that
-%% system. Therefore, we do not consider expected utility in our
-%% experiments. Instead, we report average costs of the plans and the
-%% success rate (as a ratio between solved and solvable problems).
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-time}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-time}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-time}\hfill
-  \caption{Average runtime}
-  \label{fig:results-time}
-\end{figure}
-
-\begin{figure}[h!]
-  % \centering
-  % \includegraphics{dora1-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora2-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora3-quality}\hfill
-  % \vspace{2mm}
-  \includegraphics{dora4-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora56-quality}\hfill
-  \vspace{2mm}
-  \includegraphics{dora-cat-quality}\hfill
-  \caption{Average plan costs and number of successful runs.}
-  \label{fig:results-quality}
-\end{figure}
-
-%% Figure \ref{fig:results-quality} shows the average costs of the
-%% executed plans as well as the percentage of solvable tasks that were
-%% actually solved by the planner.
-
-Examining success ratios and plan costs, where sensing is reliable
-there is little to be gained by using the contingent planner, as the
-greedy sensing approach of the baseline is sufficient. Not
-surprisingly, as sensing degrades contingent planning pays off.  Also,
-we find that time spent in contingent planning increases steeply as
-the abstraction $\bstate_0$ becomes more refined.  That refinement
-seems to be paying off in terms of the success ratio, particularly for
-tasks $d$ and $e$, where we had sequential sessions using weighted
-$A^*$ (rather than $A^*$). For less refined initial configurations,
-the increase cost of contingent planning is compensated by a decrease
-in Fast Downward planning times. The relatively high success rate
-irrespective of the level of refinement in the initial configuration
-indicates the effectiveness of using conditional entropy to guide
-abstraction refinement in our setting.
-
-%% :
-%%while the resulting plans are still
-%% longer on average, the impact on the number of solved tasks was much
-%% smaller than for the baseline system.
-
-%% Less aggressive abstraction of the initial configuration results in
-%% longer runtimes, with little to be gained in terms of plan quality. We
-%% conclude that in our scenario it is worthwhile being assumptively
-%% aggressive.
-
-%% \section{Discussion}
-%% \input{discussion.tex}
-
-\section{Related Work}
-There have been a number of recent papers on
-planning under uncertainty using systems that were intended for
-sequential planning in deterministic problems,  for example,
-\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
-probabilistic track of the 2004 International Planning Competition.
-In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
-procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
-plans and corresponding execution traces.
-%%
-More computationally expensive approaches in this vein combine
-sampling strategies on valuations over {\em runtime variables} with
-deterministic planning procedures. The outcome is typically a more
-robust sequential plan~\cite{yoon:etal:2008}, or contingent
-plan~\cite{majercik:2006}.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
-
-Also leveraging deterministic planners in problems that feature
-uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
-$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
-as a deterministic problem, and therefore solved using sequential
-systems. In this conformant setting, advances have been towards
-compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-%%
-The continual planning system that motivated our
-project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-and has been applied in completely observable domains, particularly
-those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-Our work is motivated by domains that contain a mixture of task
-planning and observation planning. There have been a number of recent
-papers representing observation planning problems as POMDPs and using
-various techniques to manage the large state
-space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
-approach in a vision algorithm selection problem. In their case there
-is a natural hierarchical decomposition of the problem which allows
-them to solve large problems by breaking them into a set of small
-POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
-represent a preference elicitation problem as a POMDP and take
-advantage of symmetry in the belief space ---essentially the idea that
-it does not matter what the value of the variable you are trying to
-observe is--- to exponentially shrink the state space. Although we
-have been actively exploring the \citeauthor{doshi08:pref_elic}
-approach, we have yet to find those exploitable structures in our
-domain due to the task planning requirement.
-
-\section{Concluding Remarks}
-
-From an automated planning perspective, the problem of practical
-mobile robot control poses important and contrary challenges.
-%%
-On the one hand, planning and execution monitoring must be
-lightweight, robust, timely, and should span the lifetime of the
-robot. Those processes must seamlessly accommodate exogenous events,
-changing objectives, and the underlying unpredictability of the
-environment.
-%%
-On the other hand, robot planning should perform computationally
-expensive reasoning about contingencies, and possible revisions of
-subjective belief according to quantitatively modelled uncertainty in
-acting and sensing.
-
-In this paper we address these challenges, developing a continual
-planner that switches between fast sequential and expensive
-decision-theoretic planning. Given a POMDP model of the environment,
-sequential planning is used to compute an initial deterministic
-sequential plan and complementary runtime evolution of the decision
-process. That plan is executed until testing of an assumption about
-the runtime state is scheduled. At that point a contingent session
-performs some sensing which determines whether or not the sequential
-session should be allowed to continue, or otherwise acts to achieve
-the overall objectives.
-
-The most pressing item for future research, is to develop a scheme
-whereby the serial planner can relax {\em executability} assumptions,
-so that conformant (or semi-conformant) plans can be executed without
-interruption by a contingent session. A more general criticism of
-switching continual planning concerns interleaved sequential and
-contingent sessions failing to make any progress towards the
-objectives. For example, in the worst case we can have each sequential
-session producing an identical plan, and each decision-theoretic
-session rejecting it without further sensing. Although this has not
-been an issue in our work so far, it must be dealt with rigorously
-in the future. We suggest that a good way to mitigate this problem is
-by developing a {\em motivational} component that maintains a {\em
-dynamic} reward model whose limiting behaviour prevents that switching
-deadlock.
-
-%% The sequential mode of our system always schedules decision-theoretic
-%% planning before executing actions that are not applicable with
-%% certainty. This is inefficient if the utility of the plan is not
-%% dependent on that actions successful execution. This situation always
-%% arises for conformant plans.
-
-%% The switching continual planning system we have described serves as
-%% the underlying planner for CogX.
-
-%% the effects of a {\em sense} schema are perceptual, whereas the
-%% effects of an {\em operator} schema are over state propositions.
-
-%% posted by a motivational component of the underlying robotic
-%% architecture.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% In this paper we present an approach to continual planning that uses
-%% two planning systems. The first
-
-%%  to a distinct class of
-%% challenges. We suppose 
-%% %%
-%% The underlying environment is modelled as a POMDP. We use the fast
-%% classical satisficing system FastDownward to find a deterministic
-%% sequential plan and complementary runtime evolution of that
-%% process. This corresponds to a generalisation of replanning in
-%% probabilistic planning to problems with partial observability.
-
-%% Interaction between the sequential planner and execution proceeds
-%% more-or-less analogously to popular replanning approaches
-
-%% Addressing these challenges in a monolithic framework, we present a
-%% {\em switching} continual planner, that uses the fast sequential
-%% satisficing procedure FastDownward to perform net-benefit
-%% %%
-%% makes reasonable assumptions about the evolution of the runtime state
-%% given a POMDP model of the environment.
-
-%% contingent sensory plans that are tailored to current
-%% objectives.
-
-%% The decision-theoretic planner is able to tailor sensory processing on
-%% a robot platform to the current objective, while FastDownward  quickly
-
-%% In this paper we develop a continual planing system that uses two
-%% planning systems. The first, is a state-of-the-art domain independent
-%% planner for deterministic problems. The second is a information-state
-%% contingency planning the information-state space of
-
-%% Continual planning is a powerful technique that goes some way to
-%% addressing those challenges. That approach interleaves planning and
-%% execution, deliberately postponing planning for contingencies unless
-%% they eventuate during execution.
-
-%% computing a single sequential plan and
-%% eventuality
-
-%% To these challenges it seems a continual planning approach is best,
-
-%% where the runtime state evolves during plan execution
-
-%% interleaved planning and execution.
-
-%% the latter is able to tailor sensory processing on a robot platform,
-%% in order that it.
-
-%% {\em ad-hoc}
-
-%% , reason about degrees of belief and uncertainty about the world,
-
-%% probabilistic sequential decision making in practical sized problems
-%% is intractable.
-
-%% quantitative probabilistic models of the perception and action .
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
-\bibliography{ijcai11}
-
-\end{document}

### 2011-04-21 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
few suggestions from moritz
not sure what I changed in related
Hi Moritz, I have merged the material I think we should have added to the paper...

@@
 $T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
----i.e., sequential planning in unobservable worlds--- can be modelled
+-- i.e., sequential planning in unobservable worlds -- can be modelled
 as a deterministic problem, and therefore solved using sequential
@@
 compact representations of beliefs amenable to existing best-first
-search planning procedures, and lazy evaluations of beliefs. Most
-recently this research thread has been extended to contingent planning
-in fully observable non-deterministic
-environments~\cite{albore:etal:2009}.
-
-%% Their applicability in our setting is
-%% limited, firstly because they only scale to smaller problems, with
-%% thousands of states, and also due to the large amount of
-%% \emph{problem-specific} offline processing that might be required to get useful
-%% search guidance.
-
-%%  In that
-%% setting \citeauthor{kurniawati:etal:2010}~(\citeyear{})
-%% recently addressed an inefficiency of offline point-based techniques
-%% in problems with medium length planning horizons, however their
-%% approach take tens of minutes to good plans, and only scales to tens
-%% of thousands of states. In the case of general domain-independent
-%% factored systems, the state-of-the-art scales to relatively small
-%% problems with $2^{22}$ states~\cite{shani:etal:2008}.
-
-%% At their limit,
-%% these procedures take over an hour to converge, and $\sim10$ seconds
-%% on average to perform a single Bellman backup.  For classes of POMDP
-%% that feature exploitable structures (e.g., no actions with negative
-%% effects), problems with as many as $10^{30}$ states can be targeted by
-%% such offline procedures~\cite{brunskill:russell:2010}.
-
-%% Although we suppose it an interesting
-%% item for future work to pursue that direction, it should be noted that
-%% ease of goal reachability is not guaranteed in the problems we face,
-%% and is certainly not a property to be assumed in domain independent
-%% planning.
-
-%%
-%% The continual planning system that motivated our
-%% project~\cite{brenner:nebel:jaamas09} also has this characteristic,
-%% and has been applied in completely observable domains, particularly
-%% those featuring multiple communicating agents.
-
-%% The use of knowledge
-%% operators in domains allows plans that act to gain knowledge, but the
-%% approach assumes that such actions are deterministic and reliable, an
-%% assumption that we relax.
-
-%%No! They simply haven't been evaluated in PO settings. They may, or
-%%may not struggle. They have sampling of traces, and that would
-%%include observations, and therefore evolutions of beliefs. SSAT was
-%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
-%%suited to POMDPs.
-
-% Normally if we are going to compare with related work, we do
-% actually *compare*. Why didn't you try those approaches? I think
-% it's safe to say that FFR will struggle. Why would it even include
-% in its plan an observational action that doesn't change the world?
-
-%%
-%% However, as we said in the introduction,
-%% all these approaches struggle in partially observable domains as they
-%% rely on being able to determine the state at all times.
+search planning procedures, and lazy evaluations of beliefs. We
+consider it an appealing future direction to pursue conformant
+reasoning during the sequential sessions we proposed. Most recently
+this research thread has been extended to contingent planning in fully
+observable non-deterministic environments~\cite{albore:etal:2009}.

### 2011-04-21 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
Hi Moritz, I have merged the material I think we should have added to the paper...

@@
 \usepackage{graphicx}
+\usepackage{tabularx}
 \graphicspath{{plots/}}
@@
 
-\nocopyright
+%%\nocopyright
 
 \pdfinfo{
-/Title (Switching in Continual Planning for Practical Robot Control)
-/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
+/Title (A Switching Planner for Combined Task and Observation Planning)
+% /Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
+/Author ()}
 % The file aaai.sty is the style file for AAAI Press 
@@
 \title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
+\author{
+Moritz G{\"o}belbecker \\ University of Freiburg, Germany\\
+\{goebelbe\}@informatik.uni-freiburg.de
+\And
+Charles O. Gretton, Richard Dearden,\\
+University of Birmingham, United Kingdom\\
+\{c.gretton,R.W.Dearden\}@cs.bham.ac.uk} 
 \setcounter{secnumdepth}{0}
@@
 in large problems from a realistic robot exploration domain.
-
-%% partial observability
-
-%% In cooperation with execution monitoring, it must perform repair
-%% and/or replanning in order to seamlessly accommodate exogenous events,
-%% changing objectives, and the underlying unpredictability of the
-%% environment.
-
-%% Realistic robot planning problems in uncertain environments often
-%% require achieving tasks while also finding out about the
-%% world. Because the world state cannot be observed directly, these
-%% problems are naturally represented as partially observable Markov
-%% decision problems (POMDPs). However, these are typically intractable
-%% for realistic problems.
-%%
 
@@
 
-%% \section{Propositional Decision-Theoretic Planning}
-%% \input{prop-dtp.tex}
+\section{Propositionally Factored Decision-Theoretic Planning}
+\input{prop-dtp.tex}
 
@@
 \input{conc.tex}
+
+\textbf{Acknowledgements}: The research leading to these results has
+received funding from the European Community's Seventh Framework
+Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
 

### 2011-04-21 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
typo
Hi Moritz, I have merged the material I think we should have added to the paper...

@@
 Downward~\cite{fast-downward}, and DT sessions use our own contingent
-procedure. We also implemented a simple dual-mode replanning {\em
-baseline} approach. Here, when a switching action is scheduled for
-execution the DT session applies a single entropy reduction action,
-whose execution can provide evidence regarding the truth value of a
-{\em relevant} assumption.
+procedure. Since most of the problems we consider are much larger than
+any available DT planner can solve directly, for comparison purposes
+we also implemented a simple dual-mode replanning baseline
+approach. Here, when a switching action is scheduled for execution the
+DT session applies a single entropy reduction action, whose execution
+can provide evidence regarding the truth value of a relevant
+assumption.
 %%
@@
 core.}. The continual planning times are reported in
-Figure~\ref{fig:results-time}, and the quality data in
-Figure~\ref{fig:results-quality}.
+Fig.~\ref{fig:results-time}, and the quality data in
+Fig.~\ref{fig:results-quality}.
 %%
@@
 positive reward scaled by a constant factor). The exception occurs for
-items f and g of Figure~\ref{fig:results-quality}, where we report
+items f and g of Fig.~\ref{fig:results-quality}, where we report
 expected discounted rewards (not plan costs).

### 2011-04-21 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Hi Moritz, I have merged the material I think we should have added to the paper...

--- 2011-02-09 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
@@
+\newcommand{\entropy}{\ensuremath{\mathrm{H}}}
+
 We now describe our {\em switching} planning system that operates
-according to the continual planning paradigm, given a description of
-the current problem and domain in DTPDDL. The system {\em
-switches}, in the sense that planning proceeds in interleaved
-sessions, in which the base planner is either {\em sequential}/linear
-or {\em decision-theoretic}.
-%%
-During a sequential session, a rewarding {\em trace} of a possible
-execution is computed using a cost-optimising satisficing planner
-which trades action costs, goal rewards, and determinacy.
-%%
-Formatted as a linear plan, the trace specifies a sequence of actions
-that achieves some reachable goals following a deterministic
-approximation of the problem at hand.
-%%
-Structurally, a trace is a sequence of elements that are either: (i) ground
-actions from the DTPDDL description of the world, or (ii) atomic {\em
-assumptions}, modelled as deterministic actions, made about the truth
-value of facts that can only be determined at runtime (e.g., that a
-box of cornflakes is located on the corner bench in the kitchen).
-%%
-The system always begins in a sequential session, so that plan
-execution proceeds by applying DTPDDL actions from the trace, in
-sequence, until the outcome of executing the next scheduled action is
-too uncertain according to a threshold parameter~(here, 95\%). In that
-eventuality a DT session commences, which tailors sensory processing
-to determine whether the assumptions made in the trace hold, or which
-otherwise acts to achieve the goals.
+according to the continual planning paradigm. The system {\em
+switches} in the sense that planning and plan execution proceed in
+interleaved sessions in which the {\em base planner} is either {\em
+sequential} or {\em decision-theoretic}.  The first session is
+sequential, and begins when a DTPDDL description of the current
+problem and domain are posted to the system.  During a sequential
+session a serial plan is computed that corresponds to one
+execution-{\em trace} in the underlying decision-process. That trace
+is a reward-giving sequence of process actions and {\em assumptive}
+actions. Each {\em assumptive} action corresponds to an assertion
+about the truth of a fact that is unknown at plan time -- e.g. that a
+box of cornflakes is located on the corner bench in the kitchen. The
+trace specifies a plan and characterises a {\em deterministic
+approximation} (see~\cite{yoon:etal:2008}) of the underlying process
+in which that plan is valuable. Traces are computed by a
+cost-optimising classical planner which trades off action costs, goal
+rewards, and determinacy. Execution of a trace proceeds according to
+the process actions in the order that they appear in the trace. If,
+according to the underlying belief-state, the outcome of the next
+action scheduled for execution is not predetermined above a threshold
+(here 95\%), then the system switches to a DT session.
 
-Because online DT planning in large problems is impractical (we seek
-response times in seconds), a DT session plans for an abstract process
-determined by the current trace and underlying belief-state. That
-abstract problem is characterised by a limited number of propositions,
-chosen due to their {\em relevance} to the action whose scheduled
-execution triggered the switch to the DT session.
-%%the subproblem the DT session is asked tosolve.
-% MOG: Describe this in detail in the appropriate subsection
-%  This
-% abstraction is constructed by first excluding all propositions (and
-% related actions) that are not true, or {\em assumed } true, of states
-% in the trace, then adding them back, using as an heuristic the entropy
-% of the trace assumptions conditional on a candidate
-% proposition. Propositions are added, one at a time, until the number
-% of states in the initial belief-state reaches a given
-% threshold.\footnote{Our experiments are performed using a number of
-% different threshold parameters.} 
-To that abstract model we also add {\em disconfirm} and {\em confirm}
-actions that the DT session can schedule in order to judge an atomic
-assumption in the trace. Those actions yield a relatively small reward
-if the corresponding judgement is true (or small penalty
-otherwise). If a judgement action is scheduled for execution the DT
-session is terminated, and control is returned to a new sequential
-session.
+Because online DT planning is impractical for the size of problem we
+are interested in, DT sessions plan in a small abstract problem
+defined in terms of the trace from the proceeding sequential
+session. This abstract state-space is characterised by a limited
+number of propositions, chosen due to their {\em relevance} to the
+action whose scheduled execution triggered the switch to the DT
+session.  To allow the DT planner to judge assumptions from the trace,
+we add {\em disconfirm} and {\em confirm} actions to the problem for
+each of them. Those yield a relatively small reward/penalty if the
+corresponding judgement is true/false. If a judgement action is
+scheduled for execution, then the DT session is terminated, and a new
+sequential session begins.
 
 Whatever the session type, our continual planner maintains a factored
-representation of successive belief-states by performing belief
-revision. As an internal representation of the $(\pp{:init})$
-declaration, we keep a tree-shaped Bayesian network which gets updated
-whenever an action is performed, or an observation received. That
-belief-state representation is used: (1) as the source of candidate
+representation of successive belief-states.  As an internal
+representation of the $(\pp{:init})$ declaration, we keep a
+tree-shaped Bayesian network which gets updated whenever an action is
+performed, or an observation received. That belief-state
+representation is used: (1) as the source of candidate
 determinisations for sequential planning, (2) in determining when to
@@
 \hline
-%% .24 & kitchen & office & office \\
-%% .06 & kitchen & kitchen & office \\
-%% .56 & kitchen & office & kitchen \\
-%% .14 & kitchen & kitchen & kitchen \\
 .7 & kitchen & \# &  kitchen\\
@@
 We now describe the optimisation criteria used during sequential
-sessions. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action, $\action_i$, from a trace of state-action pairs
+sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
+action, $\action_i$, from a trace of state-action pairs
 $\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
-does not transition to $\state^\bot$, we have that the optimal trace
-has value:
+does not transition to $\state^\bot$, and
+$\reward(\state_i, \action_i)$ be the instantaneous reward received
+for executing action $\action_i$ in state $\state_i$. The optimal
+trace has value:
 
@@
 \normalsize
-
-\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
-reward ---i.e., can be negative for action costs, or positive for goal
-achievement--- received for executing action $\action_i$ in state
-$\state_i$. Finally, it is worth clarifying that in practise we do not
-artificially limit the length of traces that the sequential planner
-can consider. Moreover, for problems we consider there is always a
-finite optimal trace.
 
@@
 \begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
-(\pp{look}~\pp{box}~\pp{office});
-(\pp{look}~\pp{cup}~\pp{kitchen});\\
-(\pp{report}~\pp{box}~\pp{office}); 
-(\pp{report}~\pp{cup}~\pp{kitchen})
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{kitchen}));\\
+\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}));\\
+(\pp{look}~\pp{box}~\pp{kitchen});
+(\pp{look}~\pp{cup}~\pp{office});\\
+(\pp{report}~\pp{box}~\pp{kitchen}); 
+(\pp{report}~\pp{cup}~\pp{office})
 \end{array}
@@
 
-\Omit{
+%%%%
+
+In order to guarantee fast DT sessions, those plan in an abstract
+process determined by the current trace and underlying belief-state.
 %%
-Finally, 
-If an
-assumption was rejected, we prohibit that sequential session from
-making it again.
-}
+The abstract process posed to the DT planner is constructed by first
+constraining as statically false all propositions except those which
+are true with probability 1, or which are the subject of {\em
+relevant} assumptions. For example, taking the above trace and
+switching action {\texttt (look~box~office)}, the underlying
+belief in Fig.~\ref{fig:beliefs}B would determine a fully
+constrained belief given by Fig.~\ref{fig:beliefs}C.
+%%
+Next, static constraints are removed, one proposition at a time, until
+the number of states that can be true with non-zero probability in the
+initial belief of the abstract process reaches a given threshold.
+%%
+In detail, for each statically-false proposition we compute the {\em
+entropy} of the state assumptions of the current trace {\em
+conditional} on that proposition. 
+%%
+Let $X$ be a set of propositions and $2^X$ the powerset of $X$, then taking
 
-%%  the size of the state space
-%% is determined by two factors: the number of non-zero states in the
-%% initial belief-state, and the number of actions. Determining
-%% action reachability is intractable, as is determining their relevance
-%% to goal achievement.
+\small
+\[
+\chi = \{\bigwedge_{x \in X'
+  \cap X}x \; \land \bigwedge_{x \in X \setminus X'}\neg x \;|\; X' \in 2^X\},
+\]
+\normalsize
 
-%% As we are aiming to keep the decision theoretic session fast, we try
-%% to keep the state space of the abstract problem as small as
-%% possible. In a deterministic-action POMDP, the state space size is
-%% determined by two factors: the number of non-zero states in the
-%% initial belief-state and the number of executable actions. As
-%% determining the relevance of an action for a given goal in in general
-%% a hard problem, we concentrate on keeping the number of uncertain
-%% facts as low as possible without losing important information about
-%% the sensing goal.
+\noindent we have that $\chi$ is a set of conjunctions each of which
+corresponds to one truth assignment to elements in $X$. Where
+$p(\phi)$ gives the probability that a conjunction $\phi$ holds in the
+belief-state of the DTPDDL process, the entropy of $X$
+\emph{conditional} on a proposition $y$, written $\entropy(X|y)$, is
+given by Eq.~\ref{eq:condent}.
 
-In order to guarantee fast DT sessions, planning occurs for a small
-abstract problem. In our setting, where the underlying decision
-process has a deterministic action model, a significant cause of
-problem largeness is due to the number of states that occur with
-non-zero probability in the underlying belief-state. In posing a small
-abstract process for a DT session, we construct an abstract
-belief-state that is chiefly characterised by facts related to the
-current trace. In that abstraction, instances of operator or sense
-declarations whose specification mentions an excluded fact are not
-available.
-%%
-In detail, we first construct an $(\pp{:init})$ declaration which
-contains the root term, and relevant assumptions from the trace --
-e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
-diamonds are probabilistic terms, and circles are atomic and/or
-conjunctive.
-%%
-For each excluded atom, we compute the {\em entropy} of the active
-assumptions, {\em conditional} on that atom. Intuitively, lower
-entropy indicates an atom gives better information about
-assumptions. Facts are iteratively added to the belief-state in
-increasing order according to that measure until the belief space size
-reaches a predefined limit (Fig.\ref{fig:abstraction-c}).
+\vspace{-1ex}
+\small
+\begin{equation}\label{eq:condent}
+  \entropy(X|y) = \sum_{x \in \chi, y' \in \{y, \neg y\}} p(x \land y') \log_2
+  \frac{p(y')}{p(x \land y')}
+\end{equation}
+\normalsize
 
-%%  or no more atoms that reduce the entropy
-%% can be added
+A low $\entropy(X|y)$ value suggests that knowing the truth value of
+$y$ is useful for determining whether or not some assumptions $X$
+hold. When removing a static constraint on propositions during the
+abstract process construction, $y_i$ is considered before $y_j$ if
+$\entropy(X|y_i)~<~\entropy(X|y_j)$. For example, if the serial plan
+assumes the robot is in a kitchen, then propositions about the
+contents of kitchens, e.g. {\texttt (=(is-in~cup)kitchen)}, are
+added to characterise the abstract process' states. Taking
+assumption $X$ to be {\texttt (=(is-in~box)kitchen)}, in relaxing
+static constraints the following entropies are calculated:
 
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  %% \subfloat[The initial belief-state with the assumptions made by the continual
-  %%   planner in bold.]{
-  %%     \label{fig:abstraction-a}
-  %%     \begin{tikzpicture}[
-  %%   level 3/.style={tree}]
-  %%   \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-  %%     child[seclevel] {node[cnode, assumption] (office) {}
-  %%       child {node [pnode, assumption] (box) {}
-  %%         child {node [cnode] (boxp1) {}}
-  %%         child {node [cnode, assumption] (boxp2) {}}
-  %%       }
-  %%       child {node [pnode] (cup) {} 
-  %%         child {node [cnode] (cupp1) {}}
-  %%         child {node [cnode] (cupp2) {}}
-  %%       }
-  %%     }
-  %%     child {node[cnode] (kitchen) {}
-  %%       child {node [pnode] (box2) {} 
-  %%         child {node [cnode] (box2p1) {}}
-  %%         child {node [cnode] (box2p2) {}}
-  %%       }
-  %%     };
-  %%  \tiny
-  %%  \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-  %%  \node[above=0 of cat] {$\pp{(category room1)}$};
-  %%  \node[below=0 of kitchen] {$\pp{kitchen}$};
-  %%  \node[below=0 of office] {$\pp{office}$};
-  %%  \node[below=0 of box] {$\pp{(is-in box)}$};
-  %%  \node[below=0 of box2] {$\pp{(is-in box)}$};
-  %%  \node[below=0 of cup] {$\pp{(is-in cup)}$};
-  %%  \node[right=0 of boxp1] {$\pp{place1}$};
-  %%  \node[right=0 of boxp2] {$\pp{place2}$};
-  %%  \node[right=0 of box2p1] {$\pp{place1}$};
-  %%  \node[right=0 of box2p2] {$\pp{place2}$};
-  %%  \node[right=0 of cupp1] {$\pp{place1}$};
-  %%  \node[right=0 of cupp2] {$\pp{place2}$};
-  %%   % \node[node] (kitchen) right of (cat) {};
-  %% \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+\small
+\begin{tabtt}
+.47 \= = \entropy(\=X|(=(is-in~milk)office))\+ \\
+ = \entropy(X|(=(is-in~milk)kitchen)) \-\\
+.97 = \entropy(X|(=(is-in~cup)office))\+\\
+ = \entropy(X|(=(is-in~cup)kitchen))
+\end{tabtt}
+\normalsize
+
+\noindent Therefore, the first static constraint to be relaxed is for
+\texttt{(=(is-in~milk)office)}, or
+equivalently \texttt{(=(is-in~milk)kitchen)}, giving a refined
+abstract belief state depicted in
+Fig.~\ref{fig:beliefs}A. Summarising, if for Fig.\ref{fig:beliefs}B
+the DT session is restricted to belief-states with fewer than $8$
+elements, then the starting belief-state of the DT session does not
+mention a ``cup''.
+
+\begin{figure}
 \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Abstract belief-state, and refinement.}
-\label{fig:abstraction}
+\begin{tabularx}{\columnwidth}{XX}
+\hline
+\vspace{1ex}
+\scriptsize (A) Partially constrained abstract \mbox{\hspace{4ex}belief-state} & 
+\vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
+\vspace{-2ex} 
+\begin{tabtt}
+(\=:init (=(is-in~R2D2)kitchen) \+ \\
+       (\=.6(and\=(=(is-in~box)kitchen) \+\+ \\
+        (\=.9(=(is-in~milk)~kitchen))\+\\
+         .1(=(is-in~milk)office))\-\-  \\
+       .4(and(=(is-in~box)office) \+\\
+        (.1(=(is-in~milk)kitchen))\+\\
+        .9(=(is-in~milk)office)))
+\end{tabtt} &
+\vspace{-2ex} 
+\begin{tabtt}
+\hspace{-6ex}(\=:init (=(is-in~R2D2)office) \+ \\
+       (\=.6(and\=(=(is-in~box)kitchen) \+\+ \\
+        (\=.9(=(is-in~milk)kitchen))\+\\
+         .1(=(is-in~milk)office))\-\-  \\
+       .4(and(=(is-in~box)office) \+\\
+        (.1(=(is-in~milk)kitchen))\+\\
+        .9(=(is-in~milk)office))) \-\-\-\\
+       (\=.6(=(is-in~cup)office) \+ \\
+      .4(=(is-in~cup)kitchen))) \\
+\end{tabtt} \\
+\vspace{-6ex}\scriptsize(C) Fully constrained abstract \mbox{\hspace{5ex}belief-state} & \\
+\vspace{-5ex} \begin{tabtt}
+(\=:init (=(is-in~R2D2)kitchen) \+ \\
+       (\=.6(=(is-in~box)kitchen))) \\
+\end{tabtt} & \\
+\hline
+\end{tabularx}
+\caption{Simplified examples of abstract belief-states from
+       DT sessions. \label{fig:beliefs}}
+\normalsize
 \end{figure}
-
-% When limiting the number of actions, we aim to include those that are
-% possibly required to a) reach the goal or b) to directly observe an
-% atom that is part of the abstract belief state. In general, checking
-% this for an action is \textsc{PSPACE} complete for \textsc{STRIPS}
-% planning.
-
-% We approximate the set of relevant actions as follows: Let
-% \relevantAtoms\ be the set of atoms that either occur in the goal
-% condition \poss(\switchAction) or in a precondition \poss($a$) of an
-% action $a$ that can directly sense any atom in the abstract belief
-% state. We construct a \emph{relaxed planning graph} with
-% \relevantAtoms\ as goal atoms from which we extract a relaxed
-% plan.
-
-%% In the first step, we remove all facts that are not part of an
-%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-%% would proceed in an abstraction of the environment that does not
-%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%% %%
-%% In a second step, we iteratively refine the relaxed declaration by
-%% adding terms from the original statement of $(\pp{:init})$ while the
-%% number of abstract states in $\bstate_0$ that occur with non-zero
-%% probability according to that refined declaration remains of a
-%% practicable size. In detail,
 

### 2011-04-21 subarchitectures/planner.sa/trunk/src/latex/AAAI11/prop-dtp.tex
Got rid of one of the paragraph

@@
-%% {\em runtime variables}
-%% {\em decision variables}
-%% {\em uninitialised variables}
-%% {\em omitted variables}
-
 We describe the partially observable propositional probabilistic
@@
 \mu_{\stochAction}(\detAction_i) = 1$.
-
-%% We model a process state $\state$ as the set of propositions that are
-%% true of the state. Notationally, we have $\state \subseteq
-%% \props$. State change is induced by application of actions. A
-%% stochastic action $\stochAction \in \stochActions$ is applicable if
-%% its precondition $\poss(\stochAction)$, a set of propositions, are
-%% satisfied in the current state. We write $\stochActions(\state)$ for set of actions
-%% applicable at state $\state$.  If action $\stochAction \in
-%% \stochActions(\state)$ is applied at \state, nature chooses one
-%% element amongst a small set of deterministic outcomes
-%% $\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
-%% \}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
-%% nature takes outcome $\detAct_i$, and for all \stochAction\ we require
-%% $\sum_{\detAction_i \in \detActions(\stochAction)}
-%% \mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
-%% effect on the state given in terms of two lists of propositions. The
-%% add-list $\add(\detAction)$ and delete-list
-%% $\delete(\detAction)$.\footnote{If a proposition is in the add-list of
-%% $\detAction$, then it cannot be in the delete-list and vice versa.}
-%% If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
-%% ..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
-%% ..,\prop_m]$ is chosen by nature when \stochAction\ is
-%% applied at state $\state$, then the resultant state is $ ( \state \cup
-%% \add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
-%% propositions from $\add(\detAction)$ are added to $\state$, and those
-%% from $\delete(\detAction)$ are removed from $\state$.
-
-%% \citeauthor{rintanen:01}~(\citeyear{rintanen:01})
 
@@
 
-Finally, we make some fairly standard assumptions. First, that action
-execution and sensing occur instantaneously, and that only one action
-can be applied per plan-step. Secondly, it can arise that in some
-states an action is applicable, and in others it is not. Moreover, a
-belief-state can assign non-zero probability to states where that
-applicability holds, and states where it does not. %% We differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
-%% In beliefs where an action exists that is applicable in some state 
-%% with non-zero probability, and not applicable in another, we differ
-%% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
-%% because 
-%% in that
-%% we forbid the execution of the action.
-%% actions that are not applicable in
-%% any state of the current belief.  
-%% Also,
-Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
-we incorporate a PPDDL-like default semantics, treating an action
-\stochAction\ executed at $\state_i$ as if it has no effect at
-states $\state$ if $\poss(\stochAction) \not\subseteq
-\state$.
+%% Finally, we make some fairly standard assumptions. First, that action
+%% execution and sensing occur instantaneously, and that only one action
+%% can be applied per plan-step. Secondly, it can arise that in some
+%% states an action is applicable, and in others it is not. Moreover, a
+%% belief-state can assign non-zero probability to states where that
+%% applicability holds, and states where it does not. %% We differ
+%% %% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04})
+%% %% In beliefs where an action exists that is applicable in some state 
+%% %% with non-zero probability, and not applicable in another, we differ
+%% %% from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
+%% %% because 
+%% %% in that
+%% %% we forbid the execution of the action.
+%% %% actions that are not applicable in
+%% %% any state of the current belief.  
+%% %% Also,
+%% Unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
+%% we incorporate a PPDDL-like default semantics, treating an action
+%% \stochAction\ executed at $\state_i$ as if it has no effect at
+%% states $\state$ if $\poss(\stochAction) \not\subseteq
+%% \state$.
 

### 2011-04-21 subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
Bryce paper is now cited, and we are over the 7-page limit. Any help saving space would be much appreciated...

@@
 
-In the direction of leveraging {\em classical} systems/approaches for
+In the direction of leveraging {\em classical} approaches for
 planning under uncertainty, the most highlighted system to date has

### 2011-04-26 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
Some latex-fixes Reduce fontsize of bibliography

@@
 
-\nocopyright
+%%\nocopyright
+
+\special{! /pdfmark where
+{pop} {userdict /pdfmark /cleartomark load put} ifelse
+[ /Author (Moritz Goebelbecker, Charles O. Gretton, Richard Dearden)
+/Title (A Switching Planner for Combined Task and Observation Planning)
+/Subject (Proceedings of the Twenty-Fifth Conference on Artificial Intelligence)
+/Keywords (planning under uncertainty)
+/DOCINFO pdfmark
+userdict /pdfmark /cleartomark load put}
 
@@
 /Title (A Switching Planner for Combined Task and Observation Planning)
-% /Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author (NA)}
+/Subject (Proceedings of the Twenty-Fifth Conference on Artificial Intelligence)
+/Author (Moritz Goebelbecker, Charles O. Gretton, Richard Dearden)
+}
+
 % The file aaai.sty is the style file for AAAI Press 
@@
 \title{A Switching Planner for Combined Task and Observation Planning}
-\author{anonymous}
+\author{
+Moritz G{\"o}belbecker \\ Albert-Ludwigs-Universit{\"a}t Freiburg, Germany\\
+goebelbe@informatik.uni-freiburg.de
+\And
+Charles O. Gretton, Richard Dearden,\\
+University of Birmingham, United Kingdom\\
+\{c.gretton,R.W.Dearden\}@cs.bham.ac.uk} 
 \setcounter{secnumdepth}{0}
@@
 %% \section{Discussion}
-%% \input{discussion.tex}
+%% \noinput{discussion.tex}
 
@@
 \input{conc.tex}
+
+\small
+\paragraph{Acknowledgements:} The research leading to these results has
+received funding from the European Community's Seventh Framework
+Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
 

### 2011-04-27 code/subarchitectures/dialogue.sa/branches/spring-school-2011-DEPRECATED/resources/grammars/openccg/moloko.v6/docs/comsys.moloko/intro.moloko.tex
deprecated

\section{Introduction}

This is where a nice sexy introduction will soon appear. Until then, my hopes is that these few words plus the table of contents will provide enough of a guide.

Basically, \textbf{chapter 1} discusses the overall 'themes' of the semantics, though I am still VERY unhappy with the intro and sections 2.1 and 2.2. So don't put too much stock in em.

Chapters 3 to 5 discuss major and novel design decisions. \textbf{Chapter 3} describes why adjectives lexically receive ONLY \cxx{adj}{M} form (i.e why \cx{n}{T} \fwdsl{} \cx{n}{T}  is generated via rule). \textbf{Chapter 4} discusses how we have managed to constrain event modifiers lexically. This is the counterpart to not having every imaginable adjunct attached to every verbal argument structure (something which Chapter 1 currently does a crappy job of explaining). \textbf{Chapter 5} discusses how we have essentially added clause-level mood handling constructions to the grammar, and why this is a good thing. It also overviews question semantics.

\textbf{Chapter 6} discusses the fricking copular verb in detail. Including lots of examples of questions.

\textbf{Chapter 7} discusses how the grammar was designed to be incremental, including a few nice step-by-step examples.

\textbf{Chapter 8} goes in detail through all of the families, syntactic features, rules, dictionary macros, etc.

Finally \textbf{Chapter 9} gives a brief tour of the grammar files (this will be useful for the next Moloko master,  whoever the poor schmuck may be).

### 2011-04-27 code/subarchitectures/dialogue.sa/branches/spring-school-2011-DEPRECATED/resources/grammars/openccg/prosodic_moloko.v6.0.1/docs/comsys.moloko/incrementality.moloko.tex
deprecated

\sect{Incrementality}

A driving factor motivating the development of Moloko was the need for incrementality in utterance parsing. This is essential for its use as a component in the broader context of incremental situated dialogue processing. At each 'step' in the flow of interaction, we want to be able to get as much meaning out of our linguistic representation as possible. In our context, these 'steps' boil down to words and thus what we want are word-by-word linguistic representations which are both integrated with what came before, and projective (or predicative) of what could come next.\footnote{as the moloko grammar is a 'pure' generative grammar, it cannot make any predictions beyond what could possibly come. To get at what is more (or most) likely to come, the possible readings outputted by Moloko must be augmented with other sources of knowledge (statistical expectations, predictions/preferences based on some element of physical or  discourse/interactional context  etc)}.These two dimensions of incrementality, 'back-integration' and 'forward-projection' , will be discussed in turn. Following this, a number of illustrative parses will be given, discussing step-by-step how incrementality is achieved. This will hopefully give a taste of how these principles have impacted the design of the grammar.

\subsect{Back-Integration}

The integration of the 'now' with the 'before' can be separated into issues of syntactic integration and issues of semantic integration. The first are concerned with parsability--we never want the parser to complain because it 'doesn't know yet' if a particular reading could be possible. Instead, the grammar should be designed so that at each point, all the possible readings can parse. The second is concerned with the structure of the current semantic representation. We do not want multiple semantic chunks which must 'wait' to be combined. Instead, whenever possible, we want semantic readings which consist of only a single semantic structure, with each sub-structure connected in a way licensed by the grammar.

\subsect{Forward-Projection}

Moloko was designed to optimize the following maxim: 'words should tell you what they expect as soon as they can';  in other words 'project early'. Translated into a grammar design principle, whenever possible, the first word in a construction projects its dependencies. This provides the maximal amount of information in terms of what could come next both syntactically (in terms of 'rightward' dependents) and semantically (in terms of 'unfilled roles'). We have already seen how this impacted the treatment of mood. Note, that again we see a crucial difference between dependents and modifiers. Modifiers are not projected, and when encountered may lead to an increase in the number of parses instead of the typical decrease in parses as more 'information' becomes incrementally available (see e.g. step 6 below).

\subsection{Some Illustrations}

\subsubsection{Example 1: \emph{the man put a ball on the table}}
First, consider the incremental parsing for this straight forward indicative sentence: \\

\begin{verbatim}
Step 1:  the

Parse 1: np/^n : 
  @x1(
      <Delimination>unique ^ 
      <Num>sg ^ 
      <Quantification>specific)

Parse 2: s/(s\!np/np)/^np/^n : 
  @x1:event(
            <Mood>ind ^ 
            <Fronted>(x2 ^ 
                      <Delimination>unique ^ 
                      <Num>sg ^ 
                      <Quantification>specific) ^ 
            <Subject>x3)

Parse 3: s/(s\!np)/^n : 
  @x1:event(
            <Mood>ind ^ 
            <Subject>(x2 ^ 
                      <Delimination>unique ^ 
                      <Num>sg ^ 
                      <Quantification>specific))
\end{verbatim}
In addition to these 3 parses, there are those corresponding to the plural readings of \emph{the}.\footnote{ although we can implement the 'filling-in' of underspecifed values at the level of syntactic features and semantic sorts, this cannot be done properly at the level of semantic features.}Clearly, as this is the first word in the utterance there is nothing to back-integrate into.\footnote{this of course says nothing about incremental discourse parsing, which is handled in a separate module outside of Moloko.}However, from even this first word there is a massive amount of projection. First, at the immediate level, \emph{the}  begins building a \catg{np}  encoding an entity which we know is unique and specific. It projects an \catg{n\id{T}} , with T expected to add the proposition(the 'semantic head' of this modification relation).  Moving upwards, the \catg{np} combines with the two indicative clause rules to begin building a clause/event with T filling the role of \diam{Subject} or \diam{Fronted}. The appropriate clausal elements (verb-phrase and subject then verb-phrase-'missing'-a-rightward-np-dependent) are projected as well, though critically, after this nominal.
\begin{verbatim}

Step 2: the man

Parse 1: np: 
  @m1:person(man ^
      <Delimination>unique ^ 
      <Num>sg ^ 
      <Quantification>specific)

Parse 2: s/(s\!np/np)/^np: 
  @x1:event(
            <Mood>ind ^ 
            <Fronted>(m1:person ^ man ^ 
                      <Delimination>unique ^ 
                      <Num>sg ^ 
                      <Quantification>specific) ^ 
            <Subject>x3)

Parse 3: s/(s\!np) : 
  @x1:event(
            <Mood>ind ^ 
            <Subject>(m1:person ^ man ^ 
                      <Delimination>unique ^ 
                      <Num>sg ^ 
                      <Quantification>specific))
\end{verbatim}
First, because \emph{man} is singular, all three of the plural readings are lost. In the remaining three singular readings, we see that \emph{man} has satisfied the projected noun and added its prop to the entity, back-integrating perfectly both syntactically and semantically. Note that at this point because \emph{man}  has no dependents it make no projections and we have a potentially complete np and its corresponding entity. This, of course, does not prohibit further rightward modification (the man on the table, the man that I told you about, etc) but as this is optional, it does not figure into the parse of the utterance.\footnote{in this way, CCG is quite different from PS based formalisms, where we would in fact get a number of 'np-bar' readings at this point, projecting these possible modifications. If we want somehow to model this 'linguistic knowledge' about the possibility of the rightward modification of nps, it would have to be handled somehow}
\begin{verbatim}
Step 3: the man put

Parse: s/pp/np : 
  @p1:action(put ^ 
             <Mood>ind ^ 
             <Tense>past ^ 
             <Actor>(m1:person ^ man ^ 
                     <Delimination>unique ^ 
                     <Num>sg ^ 
                     <Quantification>specific) ^ 
             <Patient>x1:entity ^ 
             <Result>x2:m-whereto ^ 
             <Subject>m1:person)
\end{verbatim}
Note that all but the standard indicative parses have been lost, in both cases because \emph{put} did not fit with what was expected: nothing for the \catg{np}  parse and a second \catg{np} (i.e. the \diam{Subject} ) for the fronted parse. The verb \emph{put} has exerted its control over the entire clause, massively increasing the syntactic and semantic complexity of the parse.\footnote{these properties are no doubt what have lead to the traditional treatment of the main verb as \emph{the} clausal head, both syntactically and semantically (including mood). In section \ref{sec-Mood} we motivate our decision to consider clausal constructions as an additional 'major contributor' to the clause.} It has filled in the proposition of the event and added the semantic \diam{Tense} feature. It has also added its participant roles. The \diam{Actor} role has been filled by the man (m1) via co-indexing, and the \diam{Patient} and \diam{Result} slots will be filled by the semantic objects corresponding to the projected syntactic complements \catg{\fwdsl{}np} and {\fwdsl{}pp} correspondingly. Note that the semantic sorts of these semantic objects are available (entity and whereto). At this point then, we have some very useful information about what we expect to come in the remainder of the utterance. This is, of course, still provisional. We may end up receiving less information (a fragment) or more likely, more information (additional modifications \emph{the man put it on the table \underline{yesterday} \underline{when you weren't here}}, etc)
\begin{verbatim}
Step 4: the man put a

Parse: s/pp/^n : 
  @p1:action(put ^ 
             <Mood>ind ^ 
             <Tense>past ^ 
             <Actor>(m1:person ^ man ^ 
                     <Delimination>unique ^ 
                     <Num>sg ^ 
                     <Quantification>specific) ^ 
             <Patient>(x1:entity ^ 
                       <Delimination>existential ^ 
                       <Num>sg ^ 
                       <Quantification>specific) ^ 
             <Result>x2:m-whereto ^ 
             <Subject>m1:person)
\end{verbatim}
The determiner \emph{a} begins building the \catg{np} which was projected in step 3. It is integrated syntactically ( {\fwdsl{}np} is 'replaced' with  {\fwdsl{\wedge}n}) and semantically (the entity filling the \diam{Patient} role is now marked to be existential, specific and singular). The only change in projection is that we are now expect a \cxx{n}{T}{sg} instead of a \cxx{np}{T}{sg-or-pl}.
 \begin{verbatim}
Step 5: the man put a ball

s/pp	.....
 \end{verbatim}
Nothing much interesting happens here. The patient entity is (provisionally) full, and the amount of projected material decreases, with now the pp expected 'next'.
\begin{verbatim}
Step 6: the man put a ball on

Parse 1: s/pp/^np : 
  @p1:action(put ^ 
             <Mood>ind ^ 
             <Tense>past ^ 
             <Actor>(m1:person ^ man ^ ...)
             <Patient>(b1:thing ^ ball ^ 
                       <Delimination>existential ^ 
                       <Num>sg ^ 
                       <Quantification>specific ^ 
                       <Modifier>(o1:m-location ^ on ^ 
                                  <Anchor>x1)) ^ 
             <Result>x2:m-whereto ^ 
             <Subject>m1:person)
Parse 2: s/^np : 
  @p1:action(put ^ 
             <Mood>ind ^ \id{T}
             <Tense>past ^ 
             <Actor>(m1:person ^ man ^ ...)
             <Patient>(b1:thing ^ ball ^ 
                       <Delimination>existential ^ 
                       <Num>sg ^ 
                       <Quantification>specific) ^ 
             <Result>(o1:m-whereto ^ on ^ 
                      <Anchor>x1) ^ 
             <Subject>m1:person)
 \end{verbatim}
Encountering the word \emph{on} leads to two parses. This is due to a combination of factors. The preposition \emph{on} is lexically ambiguous between a dynamic whereto and a static locational reading and also between its function as dependent and modifier. The \diam{Result} role requires a whereto dependent (\catg{pp\id{M:m-whereto} \fwdsl{\wedge} np\id{A}} ) and the immediately preceding nominal (i.e. ball) is allowed to be locationally post-modified(\catg{pp \id{M:m-location} \fwdsl{\wedge} np\id{A} } which is transformed via rule into \catg{ n\id{T} \fwdsl{\wedge} np\id{A} \backsl{\star}n\id{T} }, see next illustration for more details) . These combinatorial possibilities result in parse 1 and 2 respectively. In the first case, the projected \catg{pp\id{R}} comp is saturated with R filling \emph{put}'s \diam{Result} role. In the latter, this expectation has been is temporally 'pushed back' in priority by the presence of this modifier. This reading is available for sentences like \emph{the man put \underline{the ball on the table} into the box}. In both cases, what is expected next is the \catg{np\id{A}} corresponding to \emph{on}'s \diam{Anchor} role
\begin{verbatim}
Step 7 and 8: the man put a ball on the table

Parse 1: s/pp : 
  @p1:action(put ^ 
             <Mood>ind ^ 
             <Tense>past ^ 
             <Actor>(m1:person ^ man ^ ...)
             <Patient>(b1:thing ^ ball ^ 
                       <Delimination>existential ^ 
                       <Num>sg ^ 
                       <Quantification>specific ^ 
                       <Modifier>(o1:m-location ^ on ^ 
                                  <Anchor>(t1:thing ^ table ^ 
                                           <Delimination>unique ^ 
                                           <Num>sg ^ 
                                           <Quantification>specific))) ^ 
             <Result>x1:m-whereto ^ 
             <Subject>m1:person)
Parse 2: s : 
  @p1:action(put ^ 
             <Mood>ind ^ 
             <Tense>past ^ 
             <Actor>(m1:person ^ man ^ ...)
             <Patient>(b1:thing ^ ball ^ 
                       <Delimination>existential ^ 
                       <Num>sg ^ 
                       <Quantification>specific) ^ 
             <Result>(o1:m-whereto ^ on ^ 
                      <Anchor>(t1:thing ^ table ^ 
                               <Delimination>unique ^ 
                               <Num>sg ^ 
                               <Quantification>specific)) ^ 
             <Subject>m1:person)
\end{verbatim}
The words \emph{the} and \emph{table} combine to create an \catg{np\id{A}} with A satisfying this  \diam{Anchor} role. In the first parse, this finishes the projected dependencies resulting in a (potentially) complete \catg{s} reading. In the latter case, we essentially return to the parse in step 5, though this time with a modified patient.

\subsubsection{Example 2: \emph{ this big ball on the table}}
As a second example, we will consider a parse involving the pre and post modification of a noun. As clause level projection has already been discussed, I will focus only on the parses which relate to the local building of this \catg{np}

\begin{verbatim}
step 1: this

Parse 1: np/^n : 
  @x1(
      <Delimination>unique ^ 
      <Num>sg ^ 
      <Proximity>proximal ^ 
      <Quantification>specific)

Parse 2: np : 
  @c1:entity(context ^ 
             <Delimination>unique ^ 
             <Num>sg ^ 
             <Proximity>proximal ^ 
             <Quantification>specific)

Parse 3: s : 
  @c1:event(context ^ 
            <Proximity>proximal)
\end{verbatim}      
The word \emph{this} has a number of readings. The first is its determiner reading (c.f. step 1 parse 1 above). The second and third are its context entity and event readings (see section \ref{sec-Contextualized}). 
\begin{verbatim}
step 2: the big

Parse 1: np/^n : 
  @x1:entity(
             <Delimination>unique ^ 
             <Num>sg ^ 
             <Proximity>proximal ^ 
             <Quantification>specific ^ 
             <Modifier>(b1:q-size ^ big))
\end{verbatim} 
We can see that both in terms of back-integration and forward-projection this reading is behaving as we want it. The projected entity is further specified (modified) and we 'still' project a noun along with its associated nominal variable T and proposition.

In order to understand exactly how this reading is built, we must look a big more carefully at how adjectives like \emph{big} work. As discussed in () all adjectives begin as atomic categories. So, we have:\\

\catg{adj[M]}   :  \atsign{b1}{size} (\prop{big} ) \\ \\
This is then 'transformed' into a complex category, capable of modification, 
via the following type-changing rule:\\

\catg{adj[M]}   \arrow \catg{n\id{T} \fwdsl{} n\id{T} }  : \atsign{t1}{entity}(  \diam{Modifier}(m1) ) ) \\ \\
This result is thus \\

\catg{n\id{T} \fwdsl{} n\id{T} }  : \atsign{t1}{entity}(  \diam{Modifier}(b1:size \wand\ \prop{big}) )\\\\
which via simple forward composition combines with step 1's \catg{np \fwdsl{hat} n} resulting in the step 2 above.
\begin{verbatim}
Step 3: the big ball

Parse 1: np : 
  @b1:thing(ball ^ 
            <Delimination>unique ^ 
            <Num>sg ^ 
            <Proximity>proximal ^ 
            <Quantification>specific ^ 
            <Modifier>(b2:q-size ^ big))
\end{verbatim}
The readings here are parallel to those in step 2 of example 1. Again, the important thing is that we have a (potentially) complete np and entity.
\begin{verbatim}
Step 4: the big ball on

Parse 1: np/^np : 
  @b1:thing(ball ^ 
            <Delimination>unique ^ 
            <Num>sg ^ 
            <Proximity>proximal ^ 
            <Quantification>specific ^ 
            <Modifier>(b2:q-size ^ big) ^ 
            <Modifier>(o1:m-location ^ on ^ 
                       <Anchor>x1))

\end{verbatim}
Before investigating exactly how the complete reading from 3 was 'opened up' to allow further modification, we should notice that this reading is perfectly incremental. The preposition \emph{on} has attached itself to the entity as a locational modifier, and has appended its own syntactic projection creating the complex cat \catg{np\fwdsl{hat}np}. We have thus gone from a 'complete' reading back to an 
syntactically and semantically 'unfinished' reading.

To explain what has happened here, it is best to look at the derivational history of this reading. The word in parantheses specify where the derivational step comes from. This can be a lexical entry (lex), a type changing grammatical rule (gram) or a composition rule, e.g. (\ensuremath{>}) or (\ensuremath{>}B). 
\begin{verbatim}
   (lex)             this :- np/^n
   (lex)             big :- adj
   (gram)            tcr1: adj$1 => n/^n$1
\end{verbatim}
The \ensuremath{\$_{1}} here represents any potential dependencies that the \catg{adj} may bring along with it. We will see this in action a few steps down with \emph{on}
\begin{verbatim}
   (tcr1)           big :- n/^n
   (lex)            ball :- n
   (>)              big ball :- n
\end{verbatim}
Note that at this point, we have not actually integrated \emph{the} into the derivation yet. This does not violate incrementality, however, because we are actually 'in the middle' of a parse step. Incrementality is imposed on the output of a parse step, not on the process itself 
\begin{verbatim}
   (lex)          on :- pp/^np
   (gram)         tcr2: pp$1 => n$1\*n
   (tcr2)         on :- n/np\*n
\end{verbatim}
First, we have the lexical entry for \emph{on}, which as discussed above is \catg{pp\fwdsl{\wedge}np}, i.e. the atomic category plus its nominal dependency. Then, the rule named tcr2 comes into play, transforming the basic category into the complex category \catg{n\fwdsl{}np\backsl{\star}n} . The \ensuremath{\$_{1}}works here to take all of preposition's syntactic dependencies and place them in the correct location in the new complex category. In this case, as this is a post modifier, this is directly after the modifee (hence the ordering \catg{\fwdsl{}np\backsl{\star}n}).\footnote{remember due to the currying of CCG categories, the order in syntactic categories is the reverse of their linear realization.}This category should understood as: 'I want an \catg{n} directly (*) before me (this is the modifee), followed by a \catg{np} ( \emph{on}'s dependent), and then I'll give you back another \catg{n} (corresponding to the resulting now modified entity)
\begin{verbatim}
   (<)               big ball on :- n/np
\end{verbatim}
This first  requirement is met by \emph{big ball}, leaving the np.
\begin{verbatim}
   (>B)              the big ball on :- np/np
\end{verbatim}
Finally, \emph{the} comes back into play:  \catg{np\fwdsl{hat}n} +  \catg{n\fwdsl{}np} = \catg{np\fwdsl{}np} via \ensuremath{>}B
\begin{verbatim}
Step 5 and 6 : this big ball on the table

Parse 3: np : 
  @b1:thing(ball ^ 
            <Delimination>unique ^ 
            <Num>sg ^ 
            <Proximity>proximal ^ 
            <Quantification>specific ^ 
            <Modifier>(b2:q-size ^ big) ^ 
            <Modifier>(o1:m-location ^ on ^ 
                       <Anchor>(t1:thing ^ table ^ 
                                <Delimination>unique ^ 
                                <Num>sg ^ 
                                <Quantification>specific)))
\end{verbatim}
The \diam{Anchor} role and \catg{\fwdsl{} np} dependent are filled by \emph{the table} resulting in yet another (potentially) complete \catg{np} / entity.

### 2011-04-27 subarchitectures/dialogue.sa/branches/spring-school-2011-DEPRECATED/resources/grammars/openccg/moloko.v6/docs/comsys.moloko/semantics.moloko.tex
deprecated

\section{Semantics}

Moloko's semantics can divided into three broad categories: entities, events and modifiers. These correspond to the three propositional acts of reference, predication and modification (Croft). Of course, these categories must be treated in the broadest, most general of terms. Entities (T) correspond to not only to concrete objects, but any facet of experience which can be construed, or reified as a thing (Talmy, Langacker). Events (E) cover all types of dynamic processes, as well as ascriptions and states which endure over time. Finally, modifiers can be applied to any of the three basic categories (T, E, M) and they need not be inherit properties like size and color for T, or location, time and polarity for E--they also include evaluations, judgements and many other types of textual and interpersonal relationships. 
\subsection{The Semantic/Ontological Hierarchy} \label{sec-Ontology}
In order for the grammar to be of any use, the three broad semantic categories, T, E and M, require much finer levels of granularity. The Moloko grammar includes a richly sorted ontological type heirarchy for each. Consider the following semantic representation for the utterance \emph{quickly give the red ball to GJ}: 
\vspace{-1.5em} %% some negative vertical spacing

\begin{align*}
\atsign{g1}{action-non-motion}(& \prop{give}  \\ 
      & \wand\ \diam{Mood}(imp) \\
      & \wand\ \diam{Actor}(a1:animate \wand\ \prop{addressee})  \\ 
      & \wand\ \diam{Patient}(b1:thing  \wand\ \prop{ball})\\
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Num}(sg)\\
      & \wand\ \diam{Modifier}(r1:color  \wand\ \prop{red}) )\\
      & \wand\ \diam{Recipient}(g2:person  \wand\ \prop{GJ}) )\\
      & \wand\ \diam{Modifier}(r1:manner \wand\ \prop{quickly}) ))
\end{align*}
At the top level, we see that the event \emph{give} (g1) is of type  'action-non-motion'. At the next level, the three entities, i.e. the event participants (see below), \emph{addressee, ball, GJ} are of type 'animate', 'thing' and 'person' respectively. Finally, the event modifier \emph{quickly} is of type 'manner' and the entity modifier \emph{red} is of type 'color'. An important note concerning this semantic type heirarchy is that, for the most part, it is 'external' to the grammar itself. 
What that means is the semantic sort given to a word does not dictate its treatment within the grammar proper. For example, the fact that kick is specified to represent an 'action-motion' event does not dictate its combinatorial possibilities. This is done, instead, by specifiying the appropriate lexical family (here a 'dative' ditranstive), and by specifying its modifier group to allow for the appropriate modifications (manner, place, time, etc) and block others (e.g. dynamic locationals like whereto, wherefrom) (see... below). This is done primarily to allow easy, modular extension to the semantic heirarchy. So for example, if finer grain distinctions are needed amongst verb sorts, the grammar need not be internally modified. There are, however, some exceptions  to this principle, and will be covered in ()

\subsection{Semantic Relations: Modifiers vs Dependents } \label{sec-Semantic-Relations}

I NEED TO FIX THIS. IT WAS WRITTEN QUICKLY, A LONG TIME AGO, AND IT STINKS.
CAN CONNECT IT WITH VERBAL ROLE/MODIFIER

Given the quite broad definition of modifiers given in section X, one could view nearly any relationship as a modification. For example, a determiner like \emph{the} modifies the entity of its head noun by delimiting and quantifying it. Likewise, a negation marker, as in \emph{I did not put it there}, can be seen as an event modifier (negative propositional semantics) and a 'intention' modifier (specifying that, counter some expectation, the speaker claims they DIDN'T do it).\footnote{currently, these two levels of semantics (ideational and interpersonal) are not separated within the semantic representation.}Essentially then, whenever a word or construction 'adds' an 'extra' semantic dependency relation or feature to another 'independent' semantic 'head', it can be viewed as a modifier. This stands in contrast to head-dependent relations, where a word 'fills in' a 'missing'  role within the semantic head. This is the case, for example, with verbs and their complements, prepositions and their 'anchors', etc.

Although not much rides on this distinction between 'modifier' and 'dependent', there is one case where this has greatly impacted the structure of the moloko grammar--this is in the cases of the major 'open class' modifiers: adjectives,  prepositions, and to a lesser degree adverbs. Words of these type need to be able to play either role . Consider the following sentence pairs:

\begin{itemize}
\item I wrote the letter \underline{on the table}
\item I put the picture on the table        (Caused Motion)
\item I wanted the \underline{bigger} picture
\item I made the picture bigger             (Resultant Verb)
\end{itemize}
In the first sentence of each pair, the underlined constituents are acting as a modifier. The sentences would be semantically and syntactically 'complete' without them:
\begin{itemize}
\item	I wrote the letter
\item	I wanted the picture
\end{itemize}
Compare this second set of sentences. Here, the same words are somehow 'essential' to the sentences. Removing them leads to syntactic (and semantic) incompleteness.
\begin{itemize}
\item	I put the letter xxx
\item	I wanted the picture xxx
\end{itemize}
To handle the dual functionality of such words they must be able to behave in two grammatically distinct ways. This will be discussed in section \ref{sec-Dual-Relation-Words}.

\subsection{Entity Semantics}

In addition to their propositional head and its corresponding ontological sort, all entities are given the following levels of semantic structure: 
\begin{enumerate}
\item Number
\item Specification
\item Delimination
\end{enumerate}
Where appropriate, entities also receive this additional semantic structure
\begin{enumerate}
\item Ownership and  Compounding
\item Groups
\end{enumerate}
We will also discuss pronouns in this section.

\subsubsection{Number/Specification and Delimination}
See Moloko document for WP9. Only change is removal of the singular and non-singular distinction from \diam{Quantification}. As this information is specified by the feature \diam{Num}, this was redundant.

For the sake of incrementality (see section \ref{sec-Incrementality}), all of this information is encoded in determiners and not in the nouns themselves. Thus, the full \catg{np} \emph{a cat} is formed by combining \atsign{c1}{animate}( \prop{cat} \ )  with : 
\begin{align*}
\atsign{x1}{entity}(&  \\ 
      & \wand\ \diam{Delimination}(existential) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Number}(sg) \ ) \\ \\
\end{align*}

\subsubsection{Ownership and Compounding}
The semantic relationships which exist between the two objects underlying compounding and possessive constructions are rich and varied, and must be handled grammar externally. In each case, we have included only a single dependency relation to mark this connection: \diam{Owner} and \diam{Compound} respectively. We see both of these relations contributing to  the semantic structure of \emph{the edge of GJ's coffee mug}: 
\vspace{-1.5em}

\begin{align*}
\atsign{e1}{location}(& \prop{edge}  \\ 
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Num}(sg)\\
      & \wand\ \diam{Owner}(t2:thing \wand\ \prop{mug}\\
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Num}(sg)\\
      & \wand\ \diam{Compound}(c2:thing \wand\ \prop{coffee})\\
      & \wand\ \diam{Owner}(t2:person \wand\ \prop{GJ})))
\end{align*}

\subsubsect{Groups}
We handle the semantics of expressions like \emph{some of GJ's money} and  \emph{the first of the three balls}, i.e. subsets or groups, using the following structure:
\vspace{-1.5em}

\begin{align*}
\atsign{g1}{entity}(& \prop{group}  \\ 
      & subset-description \\
      & \diam{Set}( set-description ))
\end{align*}
So, for example \emph{some of the balls} , we have:
\begin{align*}
\atsign{g1}{entity}(& \prop{group}  \\ 
      & \wand\ \diam{Delimination}(existential) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Set}(b2:thing \wand\ \prop{ball}\\
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Num}(pl)
\end{align*}
Note the difference in semantics between \emph{I want \underline{three of the books}} and {I want \underline{three books}}:
\vspace{-1.5em}

\begin{align*}
\atsign{g1}{entity}(& \prop{group}  \\ 
      & \wand\ \diam{Delimination}(variable) \\
      & \wand\ \diam{Quantification}(unspecific) \\
      & \wand\ \diam{Number}(pl) \\
      & \wand\ \diam{Modifier}(t1:number-cardinal \wand\ \prop{three}) \\
      & \wand\ \diam{Set}(b2:thing \wand\ \prop{book}\\
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(unspecific) \\
      & \wand\ \diam{Num}(pl)\\
\end{align*}
\vspace{-2.5em}
\begin{align*}
\atsign{b1}{thing}(& \prop{book}  \\ 
      & \wand\ \diam{Delimination}(existential) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Number}(pl) \\
      & \wand\ \diam{Modifier}(t1:number-cardinal \wand\ \prop{three}) 
\end{align*}

\subsubsect{Pronouns}

Here are some illustrations of the various forms and functions of pronouns:
\begin{enumerate}
\item \emph{\ul{I} am happy, \ul{they} brought GJ the ball }
\item \emph{ give \ul{it} to \ul{me}, the robot picked \ul{it} up and put \ul{it} with \ul{them} }
\item \emph{\ul{my} ball, what is \ul{your} name}
\item \emph{\ul{mine} are over there, the girl already ate \ul{hers}}
\end{enumerate}
1 and 2 are full nominal uses, nominative and non-nominative respectively. All of these uses receive the same semantics: the nominative, singular form as \prop{semantic head} with the number specified by the \diam{Num} feature. These entities do not receive \diam{Delimination} and \diam{Quantification} \footnote{though they should.}. Thus, the 1st person singular pronoun in any of these uses receives: \emph{ \atsign{i1}{person}( \prop{I } \wand\ \diam{Num}(sg) ) }. Note that \emph{you} is ambiguous between singular and plural and thus receives two readings.

3 are ownership marking determiners (possessive pronouns) and 4 are owned entities. Both make use of the \diam{Owner} role discussed above. Here are \emph{their ball} and \emph{theirs}:
\vspace{-1.5em}

\begin{align*}
\atsign{b1}{thing}(& \prop{ball}  \\ 
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Number}(sg) \\
      & \wand\ \diam{Owner}(t1:person \wand\ \prop{they} \wand\ \diam{Number}(pl) ) 
\end{align*}
\begin{align*}
\atsign{c1}{entity}(& \prop{context}  \\ 
      & \wand\ \diam{Delimination}(unique) \\
      & \wand\ \diam{Quantification}(specific) \\
      & \wand\ \diam{Number}(sg) \\
      & \wand\ \diam{Owner}(t1:person \wand\ \prop{they} \wand\ \diam{Number}(pl) ) 
\end{align*}
Two comments about \emph{theirs}. First, this form is ambiguous between the singular reading given and a variable, unspecific plural reading. Second, it makes use of a contextualized semantic head (see section \ref{sec-Contextualized}).

\subsection{Event Semantics}

In addition their propositional head and its corresponding ontological sort, events are given the following levels of semantic structure: 
\begin{enumerate}
\item Participant Roles 
\item TAM (tense, aspect, modality) and Voice
\item Mood
\item Modality
\end{enumerate}
\subsubsection{Participant Roles}

First, roles vs modifiers.

Second, the number of event role labels used within the grammar are quite small (Actor, Patient, Recipient, Result, Event). Again, the idea is that the specifics of participant role interpretation, i.e. what it 'means' to be kicked, or to run, or to be \emph{put on the table}, are handled outside the grammar. As the grammar was designed for use in embodied agents, this is of course fully intentional. Note, however, that because of this decision, we do not get any meaningful level of semantics internal inferencing, like one does within a richer frame-based role structure. In our case, these  types of inferences must be handled outside the grammar.

Third, events specify only the number and (top-level) semantic type (E, T or M) of their participants. Only in a few cases do they subcategorize based on finer grained semantic levels. For example, the verb \emph{put} specifies that it's result is a dynamic whereto location but ditranstive recipients are not constrained to animate referents. The  grammar was built with the expectation that these 'appropriateness' restrictions are handled outside the grammar. The main, and notable, exception to this design principle is the grammar internal restriction of event modifiers (see...)

\subsubsection{Tense/Aspect/Polarity} \label{sec-TAM}

Events are currently categorized along the traditional lines of tense (past, present, future), grammatical (as compared to lexical) aspect (the continuous and perfect dimensions), polarity (positive, negative) and voice (active, passive) using semantic features. Tense is always marked, whereas aspect, polarity and voice rely on the idea of unspecified defaults: only the 'marked' forms are marked. For polarity, if no negative particle/auxillary occurs the event is 'implicitly' postive but does not receive a feature marking this. The same is true for voice - if a passive construction is not used, the event is 'implicitly' active- and for aspect-if the event is not marked for progressive or for perfective, it simply receives no aspectual feature, instead receiving the default (imperfective, non-progressive) interpretation. So, for example, here are \emph{it got taken}\footnote{ Both \emph{get} and \emph{be} passives receive the same semantic treatment. The \emph{Actor} has been filled using a contextualized semantic object (see section \ref{sec-Contextualized} below).} (passive, non-continuous, imperfective, past, positive) and \emph{she isn't coming} (active, continuous, imperfective, present, negative):
\begin{verbatim}
  @t1:action-non-motion(take ^ 
                        <Mood>ind ^ 
                        <Tense>past ^ 
                        <Voice>passive ^ 
                        <Actor>(c1:entity ^ context) ^ 
                        <Patient>(i1:thing ^ it ^ <Num>sg) )

  @c1:action-motion(come ^ 
                    <Aspect>continuous ^ 
                    <Mood>ind ^ 
                    <Polarity>neg ^ 
                    <Tense>pres ^ 
                    <Actor>(s1:person ^ she ^ <Num>sg) ) 
\end{verbatim}

\subsubsection{Mood}

Mood--the classification of utterances into imperatives, interrogatives and indicatives--is not a property of events, and hence not a component of what is typically referred to as 'propositional semantics' (or ideational meaning). Nevertheless, as the various layers of meaning, e.g. ideational, interpersonal, textual etc. have not been separated out in the Moloko grammar, this interpersonal feature is attached as a feature to the event. For open interrogatives, a dependency relation \diam{Wh-Rest} is added at this level. It specifies the nature and scope of the question.  This will be discussed in detail in section \ref{sec-Mood}.

\subsubsect{Modality}

Depending on what kind of word is contributing this meaning, the encoding of modality is handled in one of two ways.  For pure modal-auxilliary verbs (\emph{can, should, must} etc.) this is marked by adding a \diam{Modifier} dependency relation with appropriate contents to the main event. Here is semantic representation for \emph{can you walk}.

\vspace{-1.5em} 
\begin{align*}
\atsign{w1}{action-motion}(& \prop{walk}  \\ 
      & \wand\ \diam{Mood}(int) \\
      & \wand\ \diam{Actor}(y1:animate \wand\ \prop{you}\wand \diam{Num}(sg))  \\ 
      & \wand\ \diam{Modifier}(c1:ability \wand\ \prop{can}) ))
\end{align*} 
Most modal or modal-like words (\emph{want, need, try, have(to), continue, keep, try, stop, be able, be willing, etc.}), however, have been handled as main events with their scoped over event occupying the role \diam{Event} (note the co-indexing of I).

\vspace{-1.5em} 
\begin{align*}
\atsign{w1}{????}(& \prop{want}  \\ 
      & \wand\ \diam{Mood}(ind) \\
      & \wand\ \diam{Actor}(i1:animate \wand\ \prop{I}\wand \diam{Num}(sg))  \\ 
      & \wand\ \diam{Event}(w1:action-motion \wand\ \prop{walk} \\
      & \wand\ \diam{Actor}(i1:animate)) 
\end{align*} 
This was done for two reasons. First, whereas no other modal can scope over a pure modal, they can scope over semi-modals(e.g. \emph{* I want to can , * I tried to should } but \emph{I can keep running}. This can consequently lead to major differences in propositional meaning based on changes in word order, i.e. \emph{I wanted to keep running} vs. \emph{I kept wanting to run}). Second, many of these verbs also have 'object controlled' readings which require an \diam{Actor}, thus further motivating this treatment. e.g. \emph{I want him to walk}.
\vspace{-1.5em}

\begin{align*}
\atsign{w1}{????}(& \prop{want}  \\ 
      & \wand\ \diam{Mood}(ind) \\
      & \wand\ \diam{Actor}(i1:animate \wand\ \prop{I}\wand \diam{Num}(sg))  \\ 
      & \wand\ \diam{Patient}(h1:animate \wand\ \prop{him}\wand \diam{Num}(sg))  \\ 
      & \wand\ \diam{Event}(w1:action-motion \wand\ \prop{walk} \\
      & \wand\ \diam{Actor}(h1:animate))
\end{align*} 
Note that in each of these examples the modality has, like all semantic objects, been ontologically subcategorized.

\subsection{Modifier Semantics}

Consider some examples of modifiers.

\begin{enumerate}

\item \diam{Modifier}(i1:whereto \wand\  \prop{into}  \wand\ \diam{Anchor}(m1:entity \wand\ \prop{mug} ) ) 
\item \diam{Modifier}(y1:time-point \wand\ \prop{now}) ))
\item \diam{Modifier}(o1:time-point \wand\  \prop{on} \diam{Anchor}(t1:day \wand\ \prop{Tuesday}) ))
\item \diam{Modifier}(w1:instrumental \wand\  \prop{with} \diam{Anchor}(t1:thing \wand\ \prop{ball}) ))
\item \diam{Modifier}(q1:manner \wand\ \prop{quickly}) ))
\item \diam{Modifier}(a1:frequency \wand\ \prop{always}) ))\\

\item \diam{Modifier}(r1:color \wand\ \prop{red}) ))
\item \diam{Modifier}(b1:size \wand\ \prop{big}) ))
\item \diam{Modifier}(f1:number-cardinal \wand\ \prop{five}) ))\\

\item \diam{Modifier}(i1:location \wand\  \prop{on}  \wand\ \diam{Anchor}(t1:entity \wand\ \prop{table} ) ) 
\item \diam{Modifier}(w1:accompaniment \wand\  \prop{with} \diam{Anchor}(g1:person \wand\ \prop{GJ}) ))\\

\item \diam{Modifier}(r1:degree \wand\ \prop{really}) ))
\item \diam{Modifier}(m1:degree \wand\ \prop{much}) ))

\end{enumerate}
Most modifiers in Moloko are subsumed under one dependency relation named \diam{Modifier}. Functional subcategorization has been moved to the semantic/ontological sort of the propositional head of the modifier. This is true for event modifiers (1-6), entity modifiers (7-9), event/entity modifiers (10-11) and modifier modifiers (12-13). This is preferable to specifically labeled relations ( \diam{Location} , \diam{Time} , \diam{Property} , \diam{ Instrument} etc. ) for a number of reasons. First, it provides a uniform  treatment of a wide variety of phenomena. Second, it is easily extendable: all that is required is the modification of the semantic/ontological hierarchy. Third, it allows for the hierarchic grouping of relations.

In addition to their propositional head and sort, when appropriate modifiers receive a \diam{Degree} value of comparative or superlative. The 'default' value is the base value.

\subsection{Miscellaneous Semantic Issues}
\subsubsection{Quantifier and Modifier Scope}

The scope of quantifiers and modifiers is given a quite naive treatment in the Moloko grammar. In general, these operators scope directly over the appropriate semantic object, i.e they attach to it. Thus, clause level negation attaches directly to the main event, nominal quantifiers attach to entities, etc. Moreover, the differences in scope and/or information structure associated with the ordering of such operators is not treated either, e.g. \emph{I am normally happy} and \emph{normally I am happy} receive the same semantic representation.

\subsubsection{Proximity}

Both the determiner and deictic uses of \emph{this} and \emph{that}, as well as \emph{here} and \emph{there} are handled by contributing the semantic feature \diam{Proximity} with values \fv{proximal} and \fv{distal} respectively. Due to issues with \textsc{DotCCG}, we have three separate sets of values corresponding to entities, modifiers and events.

\subsubsection{Contextualized Semantic Objects} \label{sec-Contextualized}

The Moloko grammar has been designed for use in situated interaction. If it is to handle language in such environments, clearly it must be sensitive to the relevant phenomena. In this section, we will mention a few related cases, all of which have been handled by 'filling in' a semantic role with the head \prop{context}.

To begin, people often 'leave out' linguistic material which traditional grammars mark as necessary. This is particularly true in both situated and highly interactional settings. To handle such grammatically fragmentary but interactionally complete utterances, we have added a number readings which 'contextualize' the relevant compliment slots. \\ \\
Consider this simple hypothetical exchange. 
\begin{itemize}
\item 1 \ul{User}:  \ \ \emph{go get me a coffee}
\item 2 \ul{Robot}: \emph{Sure} \\  \\ Robot goes to kitchen and returns 
\item 3  \ul{User}: \ \ \emph{Hi Robot did you get that coffee for me} 
\item 4  \ul{Robot}: \emph{um I tried to but I couldn't} 
\item 5  \ul{Robot}: \emph{I didn't see any mugs}
\item 6  \ul{User}:  \ \ \emph{didn't you}  
\item 7  \ul{Robot}: \emph{No} 
\item 8  \ul{User}: \ \ \emph{well try to do it again please} 
\item 9  \ul{Robot}: \emph{ok I will} 
\end{itemize}
The event \prop{Robot \ get \  coffee \ for \ User} is first introduced as an imperative request by the user at line 1. Beginning at line 3, it is re-envoked and then tossed back in forth between the two participants in lines 4, 8 and 9, i.e. it is 'interpersonally argued' with only the mood, tense, aspect, polarity, and other modality like features changing from line to line. Similarly with \prop{Robot \ didn't \ see \ mugs } in 5 and 6 (and 7??). What is crucially important here is that many of the resulting utterances do not contain 'full clauses', i.e. the  verbal compliments corresponding to this event are not expressed. As an illustration of how we have handled this, here are the semantic structures corresponding to the various 'minor clauses' used in the dialogue above: \footnote{One important exception to this pattern is in the case of modal-like verbs whose event comp  either cannot or has not been marked by an infinite \emph{to}. Compare e.g., \emph{I started to} and \emph{I started}. In the latter case, the semantic structure does not receive an \diam{Event} compliment and hence does not receive a \prop{context} head.  }  \\ \\
\emph{I tried to but I couldn't}
\begin{verbatim}
   @b1:event(but ^ 
            <First>(t1:modal ^ try ^ 
                    <Mood>ind ^ 
                    <Tense>past ^ 
                    <Actor>(i1:person ^ I ^ <Num>sg) ^ 
                    <Event>(c1:event ^ context) ^ 
                    <Subject>i1:person) ^ 
            <Next>(c2:event ^ context ^ 
                   <Mood>ind ^ 
                   <Polarity>neg ^ 
                   <Modifier>(c3:modal ^ could) ^ 
                   <Subject>(i2:person ^ I ^<Num>sg)))
\end{verbatim}
\emph{I will}
\begin{verbatim}
  @c1:event(context ^ 
            <Mood>ind ^ 
            <Tense>fut ^ 
            <Modifier>(w1:modal ^ will) ^ 
            <Subject>(i1:person ^ I ^ <Num>sg))
\end{verbatim}
\emph{didn't you}
\begin{verbatim}
  @c1:event(context ^ 
            <Mood>int ^ 
            <Tense>past ^ 
            <Subject>(y1:person ^ you ^ <Num>sg))
\end{verbatim}
In addition to those cases where an event compliment is fully unexpressed, it can also be referred to using a pronoun, as in \emph{try to do it again please}. This has also been handled using a \prop{context} head: 
\begin{verbatim}
    @t1:modal(try ^ 
            <Mood>imp ^ 
            <Actor>(a1:entity ^ addressee) ^ 
            <Event>(d1:event ^ do ^ 
                    <Actor>a1:entity ^ 
                    <Event>(c1:event ^ context)) ^ 
            <Modifier>(a2:m-frequency ^ again) ^ 
            <Modifier>(p1:m-comment ^ please) ^ 
            <Subject>a1:entity)
\end{verbatim}
Similarly, there are number of constructions in which \prop{entities} also have either 'unexpressed' or pronoun/deictic contexualized semantic heads. We have already seen such usage for ownership prnouns (section \ref{sec-Pronouns}) and the  \diam{Actor} in passive constructions (section \ref{sec-TAM}). Here are a few additional examples: \emph{this}, \emph{these two} and \emph{the green one}.
\begin{verbatim}
  @c1:entity(context ^   
             <Delimitation>unique ^ 
             <Num>sg ^ 
             <Proximity>proximal ^ 
             <Quantification>specific)

   @c1:entity(context ^ 
             <Delimitation>unique ^ 
             <Num>pl ^ 
             <Proximity>proximal ^ 
             <Quantification>specific ^ 
             <Modifier>(t1:number-cardinal ^ two))

    @c1:entity(context ^ 
             <Delimitation>unique ^ 
             <Num>sg ^ 
             <Quantification>specific ^ 
             <Modifier>(g1:q-color ^ green))  
 \end{verbatim}       
The same holds for \prop{locational} and \prop{temporal} deictics. Consider the readings for \emph{put it there}, \emph{the table in here} and \emph{I haven't seen it since then}. \footnote{ \emph{then} and \emph{now} could be marked for and hence distinguished by \diam{Proximity} }
\begin{verbatim}
 @p1:action-non-motion(put ^ 
                        <Mood>imp ^ 
                        <Actor>(a1:entity ^ addressee) ^ 
                        <Patient>(i1:thing ^ it ^ <Num>sg) ^ 
                        <Result>(c1:m-whereto ^ context ^ 
                                 <Proximity>m-distal) ^ 
                        <Subject>a1:entity)

  @t1:thing(table ^ 
            <Delimitation>unique ^ 
            <Num>sg ^ 
            <Quantification>specific ^ 
            <Modifier>(i1:m-location ^ in ^ 
                       <Anchor>(c1:e-location ^ context ^ 
                                <Delimitation>unique ^ 
                                <Num>sg ^ 
                                <Proximity>proximal ^ 
                                <Quantification>specific )))

  @s1:perception(see ^ 
                 <Polarity>neg ^  <Mood>ind ^ 
                 <Aspect>perfect ^  <Tense>pres ^ 
                 <Actor>(i1:person ^ I ^  <Num>sg) ^ 
                 <Modifier>(s2:m-time-point ^ since ^ 
                            <Anchor>(c1:e-time-unit ^ context ^ 
                                     <Delimitation>unique ^ 
                                     <Num>sg ^ 
                                     <Quantification>specific) ^ 
                 <Patient>(i2:thing ^ it ^ <Num>sg) ^ 
                 <Subject>i1:person)
\end{verbatim}

\subsubsection{Role Defined entities and modifiers}
We have treated relative clauses by adding a specific modifier relation \diam{Role-In} to the entity which contains the semantics of the clause, i.e. the event that this entity plays a role in. The restricted entity is co-indexed with the appropriate event role. Here is \emph{ball that I took} and \emph{ball that I wanted to take}.
\begin{verbatim}
  @b1:thing(ball ^ 
            <Role-in>(t1:action-non-motion ^ take ^ 
                      <Tense>past ^ 
                      <Actor>(i1:person ^ I ^ 
                              <Num>sg) ^ 
                      <Patient>b1:thing ^ 
                      <Subject>i1:person)

  @b1:thing(ball ^ 
            <Role-in>(w1:cognition ^ want ^ 
                      <Tense>past ^ 
                      <Actor>(i1:person ^ I ^ 
                              <Num>sg) ^ 
                      <Event>(t1:action-non-motion ^ take ^ 
                              <Actor>i1:person ^ 
                              <Patient>b1:thing) ^ 
                      <Subject>i1:person)
\end{verbatim}                                            
Note that these 'minor clauses' have no mood marking. They do however have a \diam{Subject}. This is to insure full semantic integration during their incremental parsing.

They are not fully functioning (and hence 'turned off' in the grammar) but we are near to having similar readings for modifiers like \emph{where I wanted to walk}:
\begin{verbatim}
  @w1:m-location(where ^ 
                 <Scope>(w2:action-motion ^ walk ^ 
                         <Actor>(i1:person ^ I ^ 
                                 <Num>sg) ^ 
                         <Role-in>(w3:cognition ^ want ^ 
                                   <Tense>past ^ 
                                   <Actor>i1:person ^ 
                                   <Event>w2:action-motion ^ 
                                   <Subject>i1:person)))
\end{verbatim}
This could then allow sentences like \emph{put it \ul{where I told you to put it}}, \emph{I went \ul{where you wanted me to go}, I don't know \ul{where it is}, I saw \ul{what you picked up}}, etc. 
\subsection{Location}

e, q, m location: q and m should be collapsed, in fact all should \\
static vs. dynamic \\
chains of locations: the cup is in the box in the room vs. the cup is in the room, in the box. Also chains of dynamics.  go out the door around the corner up the stairs ....  All handled by forward projecting rules \ref{sec-Prepositions}. The correct ordering must be sorted out grammar externally \\
specifying location (prepositions on events, entities) \\
questioning location: where, which place, which room \\

\subsection{Time}

e, q, m time \\
sequence, point, interval \\
specifying location (after you came, after that day, after then, afterwards) \\
questioning time: when, how long \\

### 2011-04-29 code/systems/css-2011-teams/red/subarchitectures/vision.sa/src/c++/vision/components/StereoDetector/vs3/VisionCore.cc
created team

@@ -1,0 +1,566 @@
+/**
+ * @file VisionCore.cc
+ * @author Andreas Richtsfeld, Michael Zillich
+ * @date 2006, 2010
+ * @version 0.1
+ * @brief Vision Core for perceptual grouping.
+ */
+
+#include <time.h>
+#include <signal.h>
+#include <assert.h>
+#include <cstdio>
+#include <stdexcept>
+
+#include "Math.hh"
+#include "Draw.hh"
+
+#include "FormSegments.hh"
+#include "FormLines.hh"
+#include "FormArcs.hh"
+//#include "FormParallelLineGroups.hh"
+#include "FormConvexArcGroups.hh"
+#include "FormEllipses.hh"
+#include "FormCircles.hh"
+#include "FormJunctions.hh"
+#include "FormCorners.hh"
+#include "FormArcJunctions.hh"
+#include "FormEJunctions.hh"
+#include "FormCylinders.hh"
+#include "FormClosures.hh"
+#include "FormRectangles.hh"
+#include "FormFlaps.hh"
+#include "FormFlapsAri.hh"
+#include "FormCubes.hh"
+#include "FormCones.hh"
+
+#include "Line.hh"
+#include "Collinearity.hh"
+#include "LJunction.hh"
+#include "TJunction.hh"
+#include "Closure.hh"
+#include "Rectangle.hh"
+#include "Ellipse.hh"
+#include "VisionCore.hh"
+
+namespace Z
+{
+
+#ifdef DEBUG_CHECKINDEX
+/**
+ * HACK: this must be a non-inline, non-template function, otherwise the
+ * debugger cannot put a breakpoint at throw.
+ */
+void CheckIndex(unsigned i, unsigned size)
+{
+  if(i >= size)
+    throw Except(__HERE__, "invalid index %u, valid range [0,%u)", i, size);
+}
+#endif
+
+
+/**
+ * @brief Constructor of class VisionCore.
+ * @param config_name Name of config file.
+ */
+VisionCore::VisionCore(const string &config_name)
+{
+  InitMath();
+  img = 0;
+  vote_img = 0;
+  p_e = 0.;
+  p_ee = 0.;
+  roi_center = Vector2(0., 0.);
+  roi_sigma = 0.;
+  use_masking = false;
+  InitGestaltPrinciples();
+  Configure(config_name);
+}
+
+/**
+ * @brief Destructor of class VisionCore.
+ */
[…]

### 2011-04-29 systems/css-2011-teams/red/subarchitectures/planner.sa/src/c++/dtplanner/dtpddl/turnstyle.cc
created team

@@ -1,0 +1,385 @@
+/* Copyright (C) 2010 Charles Gretton (charles.gretton@gmail.com)
+ *
+ *    This program is free software: you can redistribute it and/or modify
+ *    it under the terms of the GNU General Public License as published by
+ *    the Free Software Foundation, either version 3 of the License, or
+ *    (at your option) any later version.
+ *
+ *    This program is distributed in the hope that it will be useful,
+ *    but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *    GNU General Public License for more details.
+ *
+ *    You should have received a copy of the GNU General Public License
+ *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ *
+ * Dear CogX team member :: Please email (charles.gretton@gmail.com)
+ * if you make a change to this and commit that change to SVN. In that
+ * email, can you please attach the source files you changed as they
+ * appeared before you changed them (i.e., fresh out of SVN), and the
+ * diff files (*). Alternatively, you could not commit your changes,
+ * but rather post me a patch (**) which I will test and then commit
+ * if it does not cause any problems under testing.
+ *
+ * (*) see http://www.gnu.org/software/diffutils/diffutils.html --
+ * GNU-09/2009
+ *
+ * (**) see http://savannah.gnu.org/projects/patch -- GNU-09/2009
+ * 
+ */
+
+#include "turnstyle.hh"
+
+using namespace Turnstyle;
+
+CNF::CNF(Problem_Data& _problem_Data)
+{
+    satisfiability = unknown;
+    
+    problem_Data = _problem_Data;
+
+    if(!check__no_empty_clause())satisfiability = no;
+    
+    bool different = false;
+    do{
+        auto  problem_Data__tmp = simplify__unit_prop(problem_Data);
+        different = (problem_Data__tmp != problem_Data);
+        if(different)problem_Data = problem_Data__tmp;
+        
+        if(!check__no_empty_clause()){satisfiability = no;return;};
+    }while(different);
+
+    problem_Data = simplify__subsumption(problem_Data);
+    problem_Data = simplify__remove_A_and_NOT_A(problem_Data);
+}
+
+bool CNF::clause_trivial(const Clause& A) const
+{
+    for(auto p = A.begin(); p != A.end(); p++){
[…]

### 2011-04-29 systems/css-2011-teams/red/subarchitectures/vision.sa/src/matlab/basicL/Maggot3/auxtools/fitSingleGaussL2Approx.m

@@ -1,0 +1,173 @@
+%%
+% Originally a part of: Maggot (developed within EU project CogX)
+% Author: Matej Kristan, 2009 (matej.kristan@fri.uni-lj.si; http://vicos.fri.uni-lj.si/matejk/)
+% Last revised: 2009
+%%
+function f0 = fitSingleGaussL2Approx( pdf, varargin ) 
+% For theory, see: Zhang and Kwok, "Simplifying mixture models through
+% function approximation".
+%
+ 
+
+minValueAtZero = 1e-5 ; 
+C_atError = 1 ; 
+maxIterations = 10 ;
+percentIt = 1e-3 ;
+idx_selection = [] ;
+initialPdf = [] ;
+allowWeightOptimization = 1 ;
+
+% process arguments
+args = varargin;
+nargs = length(args);
+for i = 1:2:nargs
+    switch args{i}
+        case 'idx_selection', idx_selection = args{i+1} ; 
+        case 'initialPdf', initialPdf = args{i+1} ; 
+        case 'allowWeightOptimization', allowWeightOptimization = args{i+1} ; 
+        otherwise 
+            msg = sprintf('Unknown parameter: %s !', args{i}) ;
+            error(msg) ;
+    end
+end
+
+if ~isempty(idx_selection)
+   pdf.w = pdf.w(idx_selection) ;
+   pdf.Mu = pdf.Mu(:,idx_selection) ;
+   pdf.Cov = pdf.Cov(idx_selection) ;
+end
+
+% modify allocation of covariance matrices
+d = rows(pdf.Mu) ;
+l_refs = cols(pdf.w) ;
+% target pdf
+len_target = l_refs ;
+
+% allocate the covariance matrix structure
+C_inv = {} ; 
+inv_sqrt_C_det = zeros(1,l_refs) ;
+ 
+% initialize mean, covariance and weight of the new component
+if isempty(initialPdf)
+    [t1, C1, w1] = momentMatchPdf( pdf.Mu, pdf.Cov, pdf.w) ;
+else
+    t1 = initialPdf.Mu ;
+    C1 = initialPdf.Cov{1} ; 
+    w1 = initialPdf.w ;
+end
+ 
+% initialize precalculated arrays
+for i = 1 : l_refs    
+    C_inv = horzcat(C_inv, inv( C1 + pdf.Cov{i} ) ) ;
[…]

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
bunch of changes to smaller details.

@@
 topological places. Objects indicate the category of space they
-inhabit---e.g., spoons are likely to be in kitchens. By examining
+inhabit -- e.g., spoons are likely to be in kitchens. By examining
 view cones at places for particular objects, the robot is able to: (1)
@@
 sensors with probabilities of $0.5$ and $0.2$ respectively. Each
-object class is assigned one sensor model---e.g. cornflakes may be
+object class is assigned one sensor model -- e.g. cornflakes may be
 harder to detect than refrigerators. We performed several experiments

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/prop-dtp.tex
bunch of changes to smaller details.

--- 2011-04-21 subarchitectures/planner.sa/trunk/src/latex/AAAI11/prop-dtp.tex
@@
 all \stochAction\ we require
-$\sum_{\detAction_i \in \detActions(\stochAction)}
+$\sum_{\detAction_i}
 \mu_{\stochAction}(\detAction_i) = 1$.
@@
 STRIPS operators}~\cite{rintanen:01} to model actions and observations
-propositionally, we find it convenient for presentation and computation to
-separate sensing and action. Therefore, we suppose a POMDP has a
-perceptual model given in terms of a finite set of stochastic {\em
-senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
-and perceptual propositions $\percepts$, called {\em percepts}. In
-detail, we take an observation $\observ$ to be a set of percepts
-$\observ \subseteq \percepts$, and denote by \observations\ the set of
-reachable observations. The underlying state of the process cannot be
-observed directly, rather, senses $\stochSense \in \stochSenses$
-effect an observation $\observ \in
+propositionally, we find it convenient for presentation and
+computation to separate sensing and action. Therefore, we suppose a
+POMDP has a perceptual model given in terms of a finite set of
+stochastic {\em senses} $\stochSenses$, deterministic sensing outcomes
+$\detSenses$, and perceptual propositions $\percepts$, called {\em
+percepts}. In detail, we take an observation $\observ$ to be a set of
+percepts , i.e., $\observ \subseteq \percepts$, and denote
+by \observations\ the set of \Omit{reachable} observations. The underlying
+state of the process cannot be observed directly, rather, senses
+$\stochSense \in \stochSenses$ effect an observation $\observ \in
 \observations$ that informs what should be believed about the state the
-process is in. If $\action$ is applied effecting a
+process is in. If action $\action$ is applied effecting a
 transition to a successor state $\state'$, then an observation occurs
@@
 $\stochAction$, and the state-precondition $\poss_\states(\stochSense)
-\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
-usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
+\subseteq \propositions$ is satisfied by the state $\state'$, i.e., \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
 %%
@@
 states, \actions\ is the finite set of actions, and \observations\ is
-the finite set of reachable observations (i.e., perceptual states).
-Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
-have a state transition function $\transProb(\state, \action,
+the finite set of reachable observations.  Where
+$\state, \state' \in \states$, $a \in \actions$, from $\mu$ we have a
+state transition function $\transProb(\state, \action,
 \state')$ giving the probability of a transition from state $s$ to
@@
 
-\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
-normalising factor, that is, the probability of getting observation
-$\observ$ given we execute $\action$ in $\bstate$.
+\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is a
+normalising factor, giving the probability of getting observation
+$\observ$ if $\action$ is applied to $\bstate$.
 

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
bunch of changes to smaller details.

@@
 with approximately $10^{30}$ states, by further exploiting certain
-problem features -- E.g., problems where no actions have negative
+problem features -- e.g., problems where no actions have negative
 effects.
@@
 determined~\cite{silver:veness:2010}. 
+%%
+Our approach can also be thought of as an online POMDP solver that
+uses a sequential plan to guide the search, rather than (e.g.,
+Monte-Carlo) sampling. Also, compared to most online POMDP procedures, which
+replan at each step, our approach involves relatively little replanning.
 
-In the direction of leveraging {\em classical} systems/approaches for
+%% Our approach requires minimal DT replanning,
+%% preferring
+
+%% Also, our approach will execute a few
+%% actions (as much as it can before an assumption is found to be true or
+%% false) before needing to plan again, rather than planning after each
+%% step.
+
+%% Moving someway towards supporting real-time decision making, recent
+%% online POMDP solution procedures have been developed which leverage
+%% highly approximate value functions -- computed using an offline
+%% procedure -- and heuristics in forward
+%% search~\cite{ross:etal:2008}. These approaches are applicable in
+%% larger problems (\cite{ross:etal:2008} includes experiments in a domain
+%% with around 12 thousand states), but the fact they often require
+%% expensive \emph{problem-specific} offline processing in order to yield
+%% good behaviours means they are impractical in problems as large as ours.
+%% %%
+%% A recent and promising online approach for
+%% larger POMDPs employs Monte-Carlo sampling to break the curse of
+%% dimensionality in situations where goal reachability is {\em easily}
+%% determined~\cite{silver:veness:2010}. Our approach can also be thought of
+%% as an online POMDP solver, but uses the sequential plan to guide the
+%% search
+%% for a decision-theoretic plan, rather than Monte-Carlo sampling. In
+%% addition, our approach will execute a few actions (as much as it can
+%% before
+%% an assumption is found to be true or false) before needing to plan again,
+%% rather than planning after each step.
+
+%% [have a look at the bit in parentheses above, I'm not sure I quite
+%% characterised how much plan it executes before replanning satisfactorally]
+
+In the direction of leveraging {\em classical} approaches for
 planning under uncertainty, the most highlighted system to date has

### 2011-05-01 subarchitectures/planner.sa/trunk/src/latex/AAAI11/dtpddl.tex
bunch of changes to smaller details.

@@
 unreliability in perception. There are straightforward compilations
-from problems expressed in DTPDDL to flat (and factored)
-representations of the underlying decision process. Although similar
-to Bryce's POND input language, DTPDDL distinguishes itself by
-explicitly treating state and perceptual symbols separately, and by
-providing distinct declarations for operators (i.e, state model) and
-senses (i.e., observation model). In this last respect, DTPDDL admits
-more compact domain descriptions where sensing effects are common
-across multiple operators. In detail, DTPDDL has perceptual analogues of
-fluent and predicate symbols. For example, a simple {\em object search}
-domain would have:
+from problems expressed in DTPDDL to flat state-based (and
+propositionally factored) representations of the underlying decision
+process. Although similar to the POND input
+language~\cite{Bryce:2008}, DTPDDL distinguishes itself by explicitly
+treating state and perceptual symbols separately, and by providing
+distinct declarations for operators (i.e, state model) and senses
+(i.e., observation model). In this last respect, DTPDDL admits more
+compact domain descriptions where sensing effects are common across
+multiple operators. In detail, DTPDDL has perceptual analogues of
+fluent and predicate symbols. For example, a simple {\em object
+search} domain would have:
 
@@
 structure of terms. Each term is either: (1) atomic, e.g., a state
-proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
+proposition such as $(=(\pp{is-in}~\pp{box})\pp{office})$, (2)
 probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
@@
 \begin{tabtt}
-(\=:init (= (is-in R2D2) kitchen) \+ \\
+(\=:init (= (is-in Robot) kitchen) \+ \\
        (probabilistic \=.8 (= (is-in box) office)  \\
@@
 \hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+Probability & (is-in Robot)  & (is-in box)  & (is-in cup) \\
 \hline

### 2011-05-01 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
bunch of changes to smaller details.

@@
 \hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
+Probability & (is-in Robot)  & (is-in box)  & (is-in cup) \\
 \hline
-%% .24 & kitchen & office & office \\
-%% .06 & kitchen & kitchen & office \\
-%% .56 & kitchen & office & kitchen \\
-%% .14 & kitchen & kitchen & kitchen \\
 .7 & kitchen & \# &  kitchen\\
@@
 \begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
+\state_0 \equiv \{(=(\pp{is-in}~\pp{Robot})\pp{kitchen}),\\
+\;\;(=(\pp{is-in}~\pp{box})\#), (=(\pp{is-in}~\pp{cup})\#)\}.
 \end{array}
@@
 We now describe the optimisation criteria used during sequential
-sessions. Where $\prob_i$ is the probability that the $i^{th}$
-sequenced action, $\action_i$, from a trace of state-action pairs
+sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
+action, $\action_i$, from a trace of state-action pairs
 $\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
-does not transition to $\state^\bot$, we have that the optimal trace
-has value:
+does not transition to $\state^\bot$, then the optimal
+trace has value:
 
@@
 
-\Omit{
+%%%%
+
+In order to guarantee fast DT sessions, those plan in an abstract
+process determined by the current trace and underlying belief-state.
 %%
-Finally, 
-If an
-assumption was rejected, we prohibit that sequential session from
-making it again.
-}
+The abstract process posed to the DT planner is constructed by first
+constraining as statically false all propositions except those which
+are true with probability 1, or which are the subject of {\em
+relevant} assumptions. For example, taking the above trace and
+switching action {\texttt (look~box~office)}, the underlying
+belief in Fig.~\ref{fig:beliefs}B would determine a fully
+constrained belief given by Fig.~\ref{fig:beliefs}C.
+%%
+Next, static constraints are removed, one proposition at a time, until
+the number of states that can be true with non-zero probability in the
+initial belief of the abstract process reaches a given threshold.
+%%
+In detail, for each statically-false proposition we compute the {\em
+entropy} of the state assumptions of the current trace {\em
+conditional} on that proposition. 
+%%
+Let $X$ be a set of propositions and $2^X$ the powerset of $X$, then taking
 
-%%  the size of the state space
-%% is determined by two factors: the number of non-zero states in the
-%% initial belief-state, and the number of actions. Determining
-%% action reachability is intractable, as is determining their relevance
-%% to goal achievement.
+\small
+\[
+\chi = \{\bigwedge_{x \in X'
+  \cap X}x \; \land \bigwedge_{x \in X \setminus X'}\neg x \;|\; X' \in 2^X\},
+\]
+\normalsize
 
-%% As we are aiming to keep the decision theoretic session fast, we try
-%% to keep the state space of the abstract problem as small as
-%% possible. In a deterministic-action POMDP, the state space size is
-%% determined by two factors: the number of non-zero states in the
-%% initial belief-state and the number of executable actions. As
-%% determining the relevance of an action for a given goal in in general
-%% a hard problem, we concentrate on keeping the number of uncertain
-%% facts as low as possible without losing important information about
-%% the sensing goal.
+\noindent we have that $\chi$ is a set of conjunctions each of which
+corresponds to one truth assignment to elements in $X$. Where
+$p(\phi)$ gives the probability that a conjunction $\phi$ holds in the
+belief-state of the DTPDDL process, the entropy of $X$
+\emph{conditional} on a proposition $y$, written $\entropy(X|y)$, is
+given by Eq.~\ref{eq:condent}.
 
-In order to guarantee fast DT sessions, planning occurs for a small
-abstract problem. In our setting, where the underlying decision
-process has a deterministic action model, a significant cause of
-problem largeness is due to the number of states that occur with
-non-zero probability in the underlying belief-state. In posing a small
-abstract process for a DT session, we construct an abstract
-belief-state that is chiefly characterised by facts related to the
-current trace. In that abstraction, instances of operator or sense
-declarations whose specification mentions an excluded fact are not
-available.
-%%
-In detail, we first construct an $(\pp{:init})$ declaration which
-contains the root term, and relevant assumptions from the trace --
-e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
-diamonds are probabilistic terms, and circles are atomic and/or
-conjunctive.
-%%
-For each excluded atom, we compute the {\em entropy} of the {\em
-relevant}
-assumptions, {\em conditional} on that atom. Intuitively, lower
-entropy indicates an atom gives better information about
-assumptions. Facts are iteratively added to the belief-state in
-increasing order according to that measure until the belief space size
-reaches a predefined limit (Fig.\ref{fig:abstraction-c}).
+\vspace{-1ex}
+\small
+\begin{equation}\label{eq:condent}
+  \entropy(X|y) = \sum_{x \in \chi, y' \in \{y, \neg y\}} p(x \land y') \log_2
+  \frac{p(y')}{p(x \land y')}
+\end{equation}
+\normalsize
 
-%%  or no more atoms that reduce the entropy
-%% can be added
+A low $\entropy(X|y)$ value suggests that knowing the truth value of
+$y$ is useful for determining whether or not some assumptions $X$
+hold. When removing a static constraint on propositions during the
+abstract process construction, $y_i$ is considered before $y_j$ if
+$\entropy(X|y_i)~<~\entropy(X|y_j)$. For example, if the serial plan
+assumes the box is in the kitchen, then propositions about the
+contents of kitchens containing a box, e.g. {\texttt
+  (=(is-in~milk)kitchen)}, are added to characterise the abstract
+process' states. Taking a relevant assumption $X$ to be {\texttt
+  (=(is-in~box)kitchen)}, in relaxing static constraints the following
+entropies are calculated:
 
-\begin{figure}[h!]
-  \centering
-  \tikzstyle{tree} = [sibling distance=4.5mm]
-  \tikzstyle{toplevel} = [grow'=right, sibling distance=22mm]
-  \tikzstyle{seclevel} = [sibling distance=9mm]
-  \tikzstyle{pnode} = [diamond, draw=black, minimum size=2.5mm]
-  \tikzstyle{cnode} = [circle, draw=black, minimum size=3mm]
-  \tikzstyle{assumption} = [solid, very thick, draw=black]
-  \tikzstyle{selected} = [solid, draw=black]
-  \tikzstyle{unused} = [densely dashed, draw=black!40]
-  %% \subfloat[The initial belief-state with the assumptions made by the continual
-  %%   planner in bold.]{
-  %%     \label{fig:abstraction-a}
-  %%     \begin{tikzpicture}[
-  %%   level 3/.style={tree}]
-  %%   \node[pnode, assumption] (cat) at (1,1) {} [toplevel]
-  %%     child[seclevel] {node[cnode, assumption] (office) {}
-  %%       child {node [pnode, assumption] (box) {}
-  %%         child {node [cnode] (boxp1) {}}
-  %%         child {node [cnode, assumption] (boxp2) {}}
-  %%       }
-  %%       child {node [pnode] (cup) {} 
-  %%         child {node [cnode] (cupp1) {}}
-  %%         child {node [cnode] (cupp2) {}}
-  %%       }
-  %%     }
-  %%     child {node[cnode] (kitchen) {}
-  %%       child {node [pnode] (box2) {} 
-  %%         child {node [cnode] (box2p1) {}}
-  %%         child {node [cnode] (box2p2) {}}
-  %%       }
-  %%     };
-  %%  \tiny
-  %%  \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-  %%  \node[above=0 of cat] {$\pp{(category room1)}$};
-  %%  \node[below=0 of kitchen] {$\pp{kitchen}$};
-  %%  \node[below=0 of office] {$\pp{office}$};
-  %%  \node[below=0 of box] {$\pp{(is-in box)}$};
-  %%  \node[below=0 of box2] {$\pp{(is-in box)}$};
-  %%  \node[below=0 of cup] {$\pp{(is-in cup)}$};
-  %%  \node[right=0 of boxp1] {$\pp{place1}$};
-  %%  \node[right=0 of boxp2] {$\pp{place2}$};
-  %%  \node[right=0 of box2p1] {$\pp{place1}$};
-  %%  \node[right=0 of box2p2] {$\pp{place2}$};
-  %%  \node[right=0 of cupp1] {$\pp{place1}$};
-  %%  \node[right=0 of cupp2] {$\pp{place2}$};
-  %%   % \node[node] (kitchen) right of (cat) {};
-  %% \end{tikzpicture}}
-\qquad
-  \subfloat[Removing facts that are not part of an assumption.]{
-    \label{fig:abstraction-b}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, unused] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, unused] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
+\small
+\begin{tabtt}
+.47 \= = \entropy(\=X|(=(is-in~milk)office))\+ \\
+ = \entropy(X|(=(is-in~milk)kitchen)) \-\\
+.97 = \entropy(X|(=(is-in~cup)office))\+\\
+ = \entropy(X|(=(is-in~cup)kitchen))
+\end{tabtt}
+\normalsize
+
+\noindent Therefore, the first static constraint to be relaxed is for
+\texttt{(=(is-in~milk)office)}, or
+equivalently \texttt{(=(is-in~milk)kitchen)}, giving a refined
+abstract belief state depicted in
+Fig.~\ref{fig:beliefs}A. Summarising, if for Fig.\ref{fig:beliefs}B
+the DT session is restricted to belief-states with fewer than $8$
+elements, then the starting belief-state of the DT session does not
+mention a ``cup''.
+
+\begin{figure}
 \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-\vspace{2mm}
-  \subfloat[Refinement, by adding relevant facts.]{
-    \label{fig:abstraction-c}
-    \begin{tikzpicture}[
-    level 3/.style={tree}]
-    \node[pnode, assumption] (cat) at (1,1) {} [toplevel, unused]
-      child[seclevel] {node[cnode, assumption] (office) {}
-        child {node [pnode, assumption] (box) {}
-          child {node [cnode, selected] (boxp1) {}}
-          child {node [cnode, assumption] (boxp2) {}}
-        }
-        child {node [pnode, unused] (cup) {} 
-          child {node [cnode, unused] (cupp1) {}}
-          child {node [cnode, unused] (cupp2) {}}
-        }
-      }
-      child {node[cnode, selected] (kitchen) {}
-        child {node [pnode, selected] (box2) {} 
-          child {node [cnode, selected] (box2p1) {}}
-          child {node [cnode, selected] (box2p2) {}}
-        }
-      };
-   \draw[assumption] (cat) -- (office) -- (box) -- (boxp2);
-   \draw[selected] (box) -- (boxp1);
-   \draw[selected] (cat) -- (kitchen) -- (box2) -- (box2p2);
-   \draw[selected] (box2) -- (box2p1);
-   \tiny
-   \node[above=0 of cat] {$\pp{(category room1)}$};
-   \node[below=0 of office] {$\pp{office}$};
-   \node[below=0 of box] {$\pp{(is-in box)}$};
-   \node[below=0 of box2] {$\pp{(is-in box)}$};
-   \node[right=0 of boxp1] {$\pp{place1}$};
-   \node[right=0 of boxp2] {$\pp{place2}$};
-   \node[right=0 of box2p1] {$\pp{place1}$};
-   \node[right=0 of box2p2] {$\pp{place2}$};
-    % \node[node] (kitchen) right of (cat) {};
-  \end{tikzpicture}}
-
-  \caption{Abstract belief-state, and refinement.}
-\label{fig:abstraction}
+\begin{tabularx}{\columnwidth}{XX}
+\hline
+\vspace{1ex}
+\scriptsize (A) Partially constrained belief & %abstract belief-state & 
+\vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
+\vspace{-2ex} 
+\begin{tabtt}\hspace{-4ex}
+(\=:init (=(is-in~Robot)kitchen) \+ \\
+       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
+        (\=.9(=(is-in~milk)~kitchen))\+\\
+         .1(=(is-in~milk)office))\-\-  \\
+       .4(and(=(is-in~box)office) \+\\
+        (.1(=(is-in~milk)kitchen))\+\\
+        .9(=(is-in~milk)office)))
+\end{tabtt} &
+\vspace{-2ex} 
+\begin{tabtt}
+\hspace{-4ex}(\=:init (=(is-in~Robot)office) \+ \\
+       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
+        (\=.9(=(is-in~milk)kitchen))\+\\
+         .1(=(is-in~milk)office))\-\-  \\
+       .4(and(=(is-in~box)office) \+\\
+        (.1(=(is-in~milk)kitchen))\+\\
+        .9(=(is-in~milk)office))) \-\-\-\\
+       (\=.6(=(is-in~cup)office) \+ \\
+      .4(=(is-in~cup)kitchen))) \\
+\end{tabtt} \\
+\vspace{-8ex}\scriptsize(C) Fully constrained belief & \\
+\vspace{-7.5ex} \begin{tabtt}
+(\=:init (=(is-in~Robot)kitchen) \+ \\
+       (\=.6(=(is-in~box)kitchen))) \\
+\end{tabtt} & \\
+\hline
+\end{tabularx}
+\caption{Simplified examples of abstract belief-states from
+       DT sessions. \label{fig:beliefs}}
+\normalsize
 \end{figure}
-
-% When limiting the number of actions, we aim to include those that are
-% possibly required to a) reach the goal or b) to directly observe an
-% atom that is part of the abstract belief state. In general, checking
-% this for an action is \textsc{PSPACE} complete for \textsc{STRIPS}
-% planning.
-
-% We approximate the set of relevant actions as follows: Let
-% \relevantAtoms\ be the set of atoms that either occur in the goal
-% condition \poss(\switchAction) or in a precondition \poss($a$) of an
-% action $a$ that can directly sense any atom in the abstract belief
-% state. We construct a \emph{relaxed planning graph} with
-% \relevantAtoms\ as goal atoms from which we extract a relaxed
-% plan.
-
-%% In the first step, we remove all facts that are not part of an
-%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
-%% would proceed in an abstraction of the environment that does not
-%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
-%% %%
-%% In a second step, we iteratively refine the relaxed declaration by
-%% adding terms from the original statement of $(\pp{:init})$ while the
-%% number of abstract states in $\bstate_0$ that occur with non-zero
-%% probability according to that refined declaration remains of a
-%% practicable size. In detail,
 

### 2011-05-01 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
Middle initial

@@
 {pop} {userdict /pdfmark /cleartomark load put} ifelse
-[ /Author (Moritz Goebelbecker, Charles O. Gretton, Richard Dearden)
+[ /Author (Moritz Goebelbecker, Charles Gretton, Richard Dearden)
 /Title (A Switching Planner for Combined Task and Observation Planning)

### 2011-05-01 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
missed a couple of them

@@
 
+\special{! /pdfmark where
+{pop} {userdict /pdfmark /cleartomark load put} ifelse
+[ /Author (Moritz Goebelbecker, Charles Gretton, Richard Dearden)
+/Title (A Switching Planner for Combined Task and Observation Planning)
+/Subject (Proceedings of the Twenty-Fifth Conference on Artificial Intelligence)
+/Keywords (planning under uncertainty)
+/DOCINFO pdfmark
+userdict /pdfmark /cleartomark load put}
+
 \pdfinfo{
 /Title (A Switching Planner for Combined Task and Observation Planning)
-% /Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
-/Author ()}
+/Subject (Proceedings of the Twenty-Fifth Conference on Artificial Intelligence)
+/Author (Moritz Goebelbecker, Charles Gretton, Richard Dearden)
+}
+
 % The file aaai.sty is the style file for AAAI Press 
@@
 \author{
-Moritz G{\"o}belbecker \\ University of Freiburg, Germany\\
-\{goebelbe\}@informatik.uni-freiburg.de
+Moritz G{\"o}belbecker \\ Albert-Ludwigs-Universit{\"a}t Freiburg, Germany\\
+goebelbe@informatik.uni-freiburg.de
 \And
-Charles O. Gretton, Richard Dearden,\\
+Charles Gretton, Richard Dearden,\\
 University of Birmingham, United Kingdom\\
@@
 
-\textbf{Acknowledgements}: The research leading to these results has
+\small
+\paragraph{Acknowledgements:} The research leading to these results has
 received funding from the European Community's Seventh Framework

### 2011-05-01 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
changes requested by Moritz

@@
 actions. Each {\em assumptive} action corresponds to an assertion
-about the truth of a fact that is unknown at plan time -- e.g. that a
-box of cornflakes is located on the corner bench in the kitchen. The
-trace specifies a plan and characterises a {\em deterministic
-approximation} (see~\cite{yoon:etal:2008}) of the underlying process
-in which that plan is valuable. Traces are computed by a
-cost-optimising classical planner which trades off action costs, goal
-rewards, and determinacy. Execution of a trace proceeds according to
-the process actions in the order that they appear in the trace. If,
-according to the underlying belief-state, the outcome of the next
-action scheduled for execution is not predetermined above a threshold
-(here 95\%), then the system switches to a DT session.
+about some facts that are unknown at plan time -- e.g. that a box of
+cornflakes is located on the corner bench in the kitchen. The trace
+specifies a plan and characterises a {\em deterministic approximation}
+(see~\cite{yoon:etal:2008}) of the underlying process in which that
+plan is valuable. Traces are computed by a cost-optimising classical
+planner which trades off action costs, goal rewards, and
+determinacy. Execution of a trace proceeds according to the process
+actions in the order that they appear in the trace. If, according to
+the underlying belief-state, the outcome of the next action scheduled
+for execution is not predetermined above a threshold (here 95\%), then
+the system switches to a DT session.
 
@@
 
-As we only consider deterministic-action POMDPs, all state
-uncertainty is expressed in the $(\pp{:init})$ declaration. This
-declaration is used by our approach to define the starting state for
-sequential sessions, and the set of state-assumptions available to
-sequential planning.  Writing \#\ if the value of a proposition is
-unspecified, taking the $(\pp{:init})$ example from the previous
-section, we have the following assumptions:
+%% Writing \#\ if the value of a proposition is unspecified,
+%% taking the $(\pp{:init})$ example from the previous section, we have
+%% the following assumptions:
 
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in Robot)  & (is-in box)  & (is-in cup) \\
-\hline
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
+%% \small
+%% \begin{tabular}{cccc}
+%% \hline
+%% Probability & (is-in Robot)  & (is-in box)  & (is-in cup) \\
+%% \hline
+%% .7 & kitchen & \# &  kitchen\\
+%% .3 & kitchen & \# & office \\
+%% .8 & kitchen & office & \# \\
+%% .2 & kitchen & kitchen & \# \\
+%% 1.0 & kitchen & \# & \# \\
+%% \hline
+%% \end{tabular}
+%% \normalsize
 
-\noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root term. Here, a conjunctive term is
-visited iff its atomic subterms are visited, and zero or one of its
-immediate probabilistic subterms are visited. For a sequential
-session, the starting state corresponds to an assumption (usually
-unique) with probability $1$.\footnote{Without a loss of generality,
-we assume conjunctive terms list atoms that are true with
-probability 1 according to a visitation, and that actions do not
-have negative preconditions.}  In our example, that starting state is:
+%% \noindent Each assumption corresponds to one distinct {\em
+%% relaxed} visitation of the root term. Here, a conjunctive term is
+%% visited iff its atomic subterms are visited, and zero or one of its
+%% immediate probabilistic subterms are visited.
 
-\vspace{-1ex}
+As we only consider deterministic-action POMDPs, all state uncertainty
+is expressed in the $(\pp{:init})$ declaration. This declaration is
+used by our approach to define the starting state for sequential
+sessions, and the set of assumptive actions available to sequential
+planning. Without a loss of generality we suppose conjunctive terms
+list atoms that are true with probability 1 according to a visitation,
+and that actions do not have negative preconditions. For a sequential
+session the starting state corresponds to the set of facts that are
+true with probability $1$. Continuing our example, that starting state
+is the singleton:
 
@@
 \begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{Robot})\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})\#), (=(\pp{is-in}~\pp{cup})\#)\}.
+\state_0 \equiv \{(=(\pp{is-in}~\pp{Robot})\pp{kitchen})\}.
 \end{array}
@@
 
-\vspace{-1ex}
-
-To represent state assumptions we augment the problem posed during
-a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
-each element $\prob_i (T_i)$, of each probabilistic term from
+To represent state assumptions we augment the problem posed during a
+sequential session with an \emph{assumptive action} $\assumptiveS{i}$
+for each element, $\prob_i (T_i)$, of each probabilistic term from
 $(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
@@
 $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
-it occurs in $T_k$ for some executed $\assumptiveS{k}$.
-%%
-We also add constraints that forbid scheduling of
-assumptions about facts after actions with preconditions or effects
-that mention those facts. For example, the robot cannot assume it is
-plugged into a power source immediately after it unplugs itself.
-%%
-Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
-to a successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$. State
-$\state^\bot$ is an added sink.
+it occurs in $T_k$ for some executed $\assumptiveS{k}$.  We also add
+constraints that forbid scheduling of assumptions about facts after
+actions with preconditions or effects that mention those facts. For
+example, the robot cannot assume it is plugged into a power source
+immediately after it unplugs itself.  Executing $\assumptiveS{i}$ in a
+state $\state$ effects a transition to a successor state
+$\state^{T_i}$, the union of $\state$ with atomic terms from $T_i$.
+For example, consider the following sequential plan:
 
-We now describe the optimisation criteria used during sequential
-sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
-action, $\action_i$, from a trace of state-action pairs
-$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
-does not transition to $\state^\bot$, then the optimal
-trace has value:
+\small
+\[
+\begin{array}{l}
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{kitchen}));\\
+\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}));\\
+(\pp{look}~\pp{box}~\pp{kitchen});
+(\pp{look}~\pp{cup}~\pp{office});\\
+(\pp{report}~\pp{box}~\pp{kitchen}); 
+(\pp{report}~\pp{cup}~\pp{office})
+\end{array}
+\]
+\normalsize
+
+\noindent Applying the first action in $\state_0$ yields:
+
+\small
+\[
+\hspace{-1ex}\begin{array}{l}
+\{(=(\pp{is-in}~\pp{Robot})\pp{kitchen}),(=(\pp{is-in}~\pp{box})\pp{kitchen})\}
+\end{array}
+\]
+\normalsize
+
+\noindent The assumed state before the scheduled execution of
+action $(\pp{look}~\pp{box}~\pp{kitchen})$ is:
+
+\small
+\[
+\hspace{-1ex}\begin{array}{l}
+\{(=(\pp{is-in}~\pp{Robot})\pp{kitchen}),
+(=(\pp{is-in}~\pp{box})\pp{kitchen}), \\\;\;(=(\pp{is-in}~\pp{cup})\pp{office})\}
+\end{array}
+ \]
+\normalsize
+
+To describe the optimisation criteria used during sequential sessions
+we model $\assumptiveS{i}$ probabilistically, supposing that its
+application in state $\state$ effects a transition to $\state^{T_i}$
+with probability $\prob_i$, and to $\state^\bot$ with probability $1 -
+\prob_i$. State $\state^\bot$ is an added sink. Now taking $\prob_i$
+be the probability that the $i^{th}$ sequenced action, $\action_i$,
+from a trace of state-action pairs $\langle \state_0,
+\action_0,\state_1, \action_1,.., \state_N \rangle$ does not
+transition to $\state^\bot$, then the optimal sequential plan has
+value:
 
@@
 
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a
-small positive (negative) numeric quantity which captures
-the utility the agent receives for correctly (incorrectly) rejecting
-an assumption.
-%%
-In terms of action physics, a disconfirm action can only be executed
-once, and otherwise is modelled as a self-transformation.
-%%
-We only consider {\em relevant} assumptions when constructing the
-abstract model.  If \switchAction\ is the action that switched the
-system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
-relevant} if it is necessary for the outcome of \switchAction\ to be
-determined.  Continuing our simplified example, if the trace is:
-
-\small
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{kitchen}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}));\\
-(\pp{look}~\pp{box}~\pp{kitchen});
-(\pp{look}~\pp{cup}~\pp{office});\\
-(\pp{report}~\pp{box}~\pp{kitchen}); 
-(\pp{report}~\pp{cup}~\pp{office})
-\end{array}
-\]
-\normalsize
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-relevant for this session, and therefore we exclude the corresponding
-disconfirm action from the abstract decision process.
-
-Given \switchAction, we also include another once-only self-transition
-action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
-with the reward property:
+\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a small positive
+(negative) numeric quantity which captures the utility the agent
+receives for correctly (incorrectly) rejecting an assumption.  In
+terms of action physics, a disconfirm action can only be executed
+once, and otherwise is modelled as a self-transformation.  We only
+consider {\em relevant} assumptions when constructing the abstract
+model.  If \switchAction\ is the action that switched the system to a
+DT session, then an assumption $\assumptiveS{i}$ is {\em relevant} if
+it is necessary for the outcome of \switchAction\ to be
+determined. For example, taking the switching action \switchAction\ to
+be $(\pp{look}~\pp{box}~\pp{kitchen})$ from the example sequential
+plan, we have that
+$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}))$ is not
+relevant, and therefore we exclude the corresponding disconfirm action
+from the abstract decision process. Given \switchAction, we also
+include another once-only self-transition action
+$\actions.\poss(\switchAction)$, a \emph{confirmation action} with the
+reward property:
 
@@
 
-Execution of either a disconfirmation or the confirmation action
+\noindent Execution of either a disconfirmation or the confirmation action
 returns control to a sequential session, which then continues from the
@@
 
-%%%%
-
 In order to guarantee fast DT sessions, those plan in an abstract
 process determined by the current trace and underlying belief-state.
-%%
 The abstract process posed to the DT planner is constructed by first
@@
   relevant} assumptions. For example, taking the above trace and
-switching action {\texttt (look~box~office)}, the underlying
-belief in Fig.~\ref{fig:beliefs}B would determine a fully
-constrained belief given by Fig.~\ref{fig:beliefs}C.
-%%
-Next, static constraints are removed, one proposition at a time, until
-the number of states that can be true with non-zero probability in the
-initial belief of the abstract process reaches a given threshold.
-%%
-In detail, for each statically-false proposition we compute the {\em
-entropy} of the state assumptions of the current trace {\em
-conditional} on that proposition. 
-%%
-Let $X$ be a set of propositions and $2^X$ the powerset of $X$, then taking
+switching action {\texttt \mbox{(look~box~office)}}, the underlying
+belief in Fig.~\ref{fig:beliefs}B would determine a fully constrained
+belief given by Fig.~\ref{fig:beliefs}A.  Next, static constraints are
+removed, one proposition at a time, until the number of states that
+can be true with non-zero probability in the initial belief of the
+abstract process reaches a given threshold.  In detail, for each
+statically-false proposition we compute the {\em entropy} of the state
+assumptions of the current trace {\em conditional} on that
+proposition.  Let $X$ be a set of propositions and $2^X$ the powerset
+of $X$, then taking
 
@@
 assumes the box is in the kitchen, then propositions about the
-contents of kitchens containing a box, e.g. {\texttt
-  (=(is-in~milk)kitchen)}, are added to characterise the abstract
-process' states. Taking a relevant assumption $X$ to be {\texttt
-  (=(is-in~box)kitchen)}, in relaxing static constraints the following
-entropies are calculated:
+contents of kitchens containing a box,
+e.g. $(=(\pp{is-in}~\pp{milk})\pp{kitchen})$, are added to
+characterise the abstract process' states. Taking a relevant
+assumption $X$ to be $(=(\pp{is-in}~\pp{box})\pp{kitchen})$, in
+relaxing static constraints the following entropies are calculated:
 
@@
 abstract belief state depicted in
-Fig.~\ref{fig:beliefs}A. Summarising, if for Fig.\ref{fig:beliefs}B
+Fig.~\ref{fig:beliefs}C. Summarising, if for Fig.\ref{fig:beliefs}B
 the DT session is restricted to belief-states with fewer than $8$
@@
 \hline
-\vspace{1ex}
-\scriptsize (A) Partially constrained belief & %abstract belief-state & 
-\vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
-\vspace{-2ex} 
-\begin{tabtt}\hspace{-4ex}
+\scriptsize\vspace{0.5ex}(A) Fully constrained belief & 
+\vspace{0.5ex}\hspace{-4.0ex}(C) Partially constrained belief\\
+\vspace{-3.0ex}\begin{tabtt}
 (\=:init (=(is-in~Robot)kitchen) \+ \\
+       (\=.6(=(is-in~box)kitchen))) \\
+\end{tabtt} & 
+\vspace{-3.0ex}\begin{tabtt}
+\hspace{-5.0ex}(\=:init (=(is-in~Robot)kitchen) \+ \\
        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
@@
         .9(=(is-in~milk)office)))
-\end{tabtt} &
-\vspace{-2ex} 
-\begin{tabtt}
-\hspace{-4ex}(\=:init (=(is-in~Robot)office) \+ \\
+\end{tabtt}\\
+\vspace{-16.5ex}(B) Underlying DTPDDL belief & \\ 
+\vspace{-17.0ex}\begin{tabtt}
+(\=:init (=(is-in~Robot)office) \+ \\
        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
@@
       .4(=(is-in~cup)kitchen))) \\
-\end{tabtt} \\
-\vspace{-8ex}\scriptsize(C) Fully constrained belief & \\
-\vspace{-7.5ex} \begin{tabtt}
-(\=:init (=(is-in~Robot)kitchen) \+ \\
-       (\=.6(=(is-in~box)kitchen))) \\
 \end{tabtt} & \\
@@
 
+%% \begin{figure}
+%% \tiny
+%% \begin{tabularx}{\columnwidth}{XX}
+%% \hline
+%% \vspace{1ex}
+%% \scriptsize (A) Partially constrained belief & %abstract belief-state & 
+%% \vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
+%% \vspace{-2ex} 
+%% \begin{tabtt}\hspace{-4ex}
+%% (\=:init (=(is-in~Robot)kitchen) \+ \\
+%%        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
+%%         (\=.9(=(is-in~milk)~kitchen))\+\\
+%%          .1(=(is-in~milk)office))\-\-  \\
+%%        .4(and(=(is-in~box)office) \+\\
+%%         (.1(=(is-in~milk)kitchen))\+\\
+%%         .9(=(is-in~milk)office)))
+%% \end{tabtt} &
+%% \vspace{-2ex} 
+%% \begin{tabtt}
+%% \hspace{-4ex}(\=:init (=(is-in~Robot)office) \+ \\
+%%        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
+%%         (\=.9(=(is-in~milk)kitchen))\+\\
+%%          .1(=(is-in~milk)office))\-\-  \\
+%%        .4(and(=(is-in~box)office) \+\\
+%%         (.1(=(is-in~milk)kitchen))\+\\
+%%         .9(=(is-in~milk)office))) \-\-\-\\
+%%        (\=.6(=(is-in~cup)office) \+ \\
+%%       .4(=(is-in~cup)kitchen))) \\
+%% \end{tabtt} \\
+%% \vspace{-8ex}\scriptsize(C) Fully constrained belief & \\
+%% \vspace{-7.5ex} \begin{tabtt}
+%% (\=:init (=(is-in~Robot)kitchen) \+ \\
+%%        (\=.6(=(is-in~box)kitchen))) \\
+%% \end{tabtt} & \\
+%% \hline
+%% \end{tabularx}
+%% \caption{Simplified examples of abstract belief-states from
+%%        DT sessions. \label{fig:beliefs}}
+%% \normalsize
+%% \end{figure}
+
 %%% Local Variables: 

### 2011-05-01 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Bunch of errors in some of the sentences in the switching section. Now seem to be fixed.

@@
 
+%% due to their {\em relevance} to the
+%% action whose scheduled execution triggered the switch to the DT
+%% session.
+
 Because online DT planning is impractical for the size of problem we
 are interested in, DT sessions plan in a small abstract problem
-defined in terms of the trace from the proceeding sequential
-session. This abstract state-space is characterised by a limited
-number of propositions, chosen due to their {\em relevance} to the
-action whose scheduled execution triggered the switch to the DT
-session.  To allow the DT planner to judge assumptions from the trace,
-we add {\em disconfirm} and {\em confirm} actions to the problem for
-each of them. Those yield a relatively small reward/penalty if the
-corresponding judgement is true/false. If a judgement action is
-scheduled for execution, then the DT session is terminated, and a new
-sequential session begins.
+defined in terms of the trace from the proceeding sequential session.
+This abstract state-space is characterised by a limited number of
+propositions, chosen because they relate evidence about assumptions
+in the trace.  To allow the DT planner to judge {\em relevant}
+assumptions from the trace, we add {\em disconfirm} and {\em confirm}
+actions to the problem for each of them. Those yield a relatively
+small reward/penalty if the corresponding judgement is true/false. If
+a judgement action is scheduled for execution, then the DT session is
+terminated, and a new sequential session begins.
 
@@
 with probability $\prob_i$, and to $\state^\bot$ with probability $1 -
-\prob_i$. State $\state^\bot$ is an added sink. Now taking $\prob_i$
-be the probability that the $i^{th}$ sequenced action, $\action_i$,
-from a trace of state-action pairs $\langle \state_0,
-\action_0,\state_1, \action_1,.., \state_N \rangle$ does not
-transition to $\state^\bot$, then the optimal sequential plan has
-value:
+\prob_i$. State $\state^\bot$ is an added sink. Taking $\prob_i$ to be
+the probability that the $i^{th}$ sequenced action, $\action_i$, from
+a trace of state-action pairs $\langle \state_0, \action_0,\state_1,
+\action_1,.., \state_N \rangle$ does not transition to $\state^\bot$,
+then the optimal sequential plan has value:
 
@@
 determined. For example, taking the switching action \switchAction\ to
-be $(\pp{look}~\pp{box}~\pp{kitchen})$ from the example sequential
-plan, we have that
+be $(\pp{look}~\pp{box}~\pp{kitchen})$ from our earlier sequential
+plan example, we have that
 $\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}))$ is not
@@
 abstract process reaches a given threshold.  In detail, for each
-statically-false proposition we compute the {\em entropy} of the state
-assumptions of the current trace {\em conditional} on that
+statically-false proposition we compute the {\em entropy} of the
+relevant assumptions of the current trace {\em conditional} on that
 proposition.  Let $X$ be a set of propositions and $2^X$ the powerset
@@
 \vspace{-17.0ex}\begin{tabtt}
-(\=:init (=(is-in~Robot)office) \+ \\
+(\=:init (=(is-in~Robot)kitchen) \+ \\
        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
@@
 \end{tabularx}
-\caption{Simplified examples of abstract belief-states from
+\caption{Simplified examples of belief-states from
        DT sessions. \label{fig:beliefs}}

### 2011-05-03 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
small changes

@@
 actions. Each {\em assumptive} action corresponds to an assertion
-about the truth of a fact that is unknown at plan time -- e.g. that a
-box of cornflakes is located on the corner bench in the kitchen. The
-trace specifies a plan and characterises a {\em deterministic
-approximation} (see~\cite{yoon:etal:2008}) of the underlying process
-in which that plan is valuable. Traces are computed by a
-cost-optimising classical planner which trades off action costs, goal
-rewards, and determinacy. Execution of a trace proceeds according to
-the process actions in the order that they appear in the trace. If,
-according to the underlying belief-state, the outcome of the next
-action scheduled for execution is not predetermined above a threshold
-(here 95\%), then the system switches to a DT session.
+about some facts that are unknown at plan time -- e.g. that a box of
+cornflakes is located on the corner bench in the kitchen. The trace
+specifies a plan and characterises a {\em deterministic approximation}
+(see~\cite{yoon:etal:2008}) of the underlying process in which that
+plan is valuable. Traces are computed by a cost-optimising classical
+planner which trades off action costs, goal rewards, and
+determinacy. Execution of a trace proceeds according to the process
+actions in the order that they appear in the trace. If, according to
+the underlying belief-state, the outcome of the next action scheduled
+for execution is not predetermined above a threshold (here 95\%), then
+the system switches to a DT session.
+
+%% due to their {\em relevance} to the
+%% action whose scheduled execution triggered the switch to the DT
+%% session.
 
@@
 are interested in, DT sessions plan in a small abstract problem
-defined in terms of the trace from the proceeding sequential
-session. This abstract state-space is characterised by a limited
-number of propositions, chosen due to their {\em relevance} to the
-action whose scheduled execution triggered the switch to the DT
-session.  To allow the DT planner to judge assumptions from the trace,
-we add {\em disconfirm} and {\em confirm} actions to the problem for
-each of them. Those yield a relatively small reward/penalty if the
-corresponding judgement is true/false. If a judgement action is
-scheduled for execution, then the DT session is terminated, and a new
-sequential session begins.
+defined in terms of the trace from the proceeding sequential session.
+This abstract state-space is characterised by a limited number of
+propositions, chosen because they relate evidence about assumptions
+in the trace.  To allow the DT planner to judge
+assumptions from the trace, we add {\em disconfirm} and {\em confirm}
+actions to the problem for each of them. Those yield a relatively
+small reward/penalty if the corresponding judgement is true/false. If
+a judgement action is scheduled for execution, then the DT session is
+terminated, and a new sequential session begins.
 
@@
 
-As we only consider deterministic-action POMDPs, all state
-uncertainty is expressed in the $(\pp{:init})$ declaration. This
-declaration is used by our approach to define the starting state for
-sequential sessions, and the set of state-assumptions available to
-sequential planning.  Writing \#\ if the value of a proposition is
-unspecified, taking the $(\pp{:init})$ example from the previous
-section, we have the following assumptions:
+%% Writing \#\ if the value of a proposition is unspecified,
+%% taking the $(\pp{:init})$ example from the previous section, we have
+%% the following assumptions:
 
-\small
-\begin{tabular}{cccc}
-\hline
-Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
-\hline
-.7 & kitchen & \# &  kitchen\\
-.3 & kitchen & \# & office \\
-.8 & kitchen & office & \# \\
-.2 & kitchen & kitchen & \# \\
-1.0 & kitchen & \# & \# \\
-\hline
-\end{tabular}
-\normalsize
+%% \small
+%% \begin{tabular}{cccc}
+%% \hline
+%% Probability & (is-in Robot)  & (is-in box)  & (is-in cup) \\
+%% \hline
+%% .7 & kitchen & \# &  kitchen\\
+%% .3 & kitchen & \# & office \\
+%% .8 & kitchen & office & \# \\
+%% .2 & kitchen & kitchen & \# \\
+%% 1.0 & kitchen & \# & \# \\
+%% \hline
+%% \end{tabular}
+%% \normalsize
 
-\noindent Each assumption corresponds to one distinct {\em
-relaxed} visitation of the root term. Here, a conjunctive term is
-visited iff its atomic subterms are visited, and zero or one of its
-immediate probabilistic subterms are visited. For a sequential
-session, the starting state corresponds to an assumption (usually
-unique) with probability $1$.\footnote{Without a loss of generality,
-we assume conjunctive terms list atoms that are true with
-probability 1 according to a visitation, and that actions do not
-have negative preconditions.}  In our example, that starting state is:
+%% \noindent Each assumption corresponds to one distinct {\em
+%% relaxed} visitation of the root term. Here, a conjunctive term is
+%% visited iff its atomic subterms are visited, and zero or one of its
+%% immediate probabilistic subterms are visited.
 
-\vspace{-1ex}
+As we only consider deterministic-action POMDPs, all state uncertainty
+is expressed in the $(\pp{:init})$ declaration. This declaration is
+used by our approach to define the starting state for sequential
+sessions, and the set of assumptive actions available to sequential
+planning. Without a loss of generality we also suppose that actions do not have negative preconditions. For a sequential
+session the starting state corresponds to the set of facts that are
+true with probability $1$. Continuing our example, that starting state
+is the singleton:
 
@@
 \begin{array}{l}
-\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
-\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
+\state_0 \equiv \{(=(\pp{is-in}~\pp{Robot})\pp{kitchen})\}.
 \end{array}
@@
 $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
-it occurs in $T_k$ for some executed $\assumptiveS{k}$.
-%%
-We also add constraints that forbid scheduling of
-assumptions about facts after actions with preconditions or effects
-that mention those facts. For example, the robot cannot assume it is
-plugged into a power source immediately after it unplugs itself.
-%%
-Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
-to a successor state $\state^{T_i}$ with probability $\prob_i$, and
-$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
-the union of $\state$ with atomic terms from $T_i$. State
-$\state^\bot$ is an added sink.
+it occurs in $T_k$ for some executed $\assumptiveS{k}$.  We also add
+constraints that forbid scheduling of assumptions about facts after
+actions with preconditions or effects that mention those facts. For
+example, the robot cannot assume it is plugged into a power source
+immediately after it unplugs itself.  Executing $\assumptiveS{i}$ in a
+state $\state$ effects a transition to a successor state
+$\state^{T_i}$, the union of $\state$ with atomic terms from $T_i$.
+For example, consider the following sequential plan:
 
-We now describe the optimisation criteria used during sequential
-sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
-action, $\action_i$, from a trace of state-action pairs
-$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
-does not transition to $\state^\bot$, and
-$\reward(\state_i, \action_i)$ be the instantaneous reward received
-for executing action $\action_i$ in state $\state_i$. The optimal
-trace has value:
+\small
+\[
+\begin{array}{l}
+\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{kitchen}));\\
+\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}));\\
+(\pp{look}~\pp{box}~\pp{kitchen});
+(\pp{look}~\pp{cup}~\pp{office});\\
+(\pp{report}~\pp{box}~\pp{kitchen}); 
+(\pp{report}~\pp{cup}~\pp{office})
+\end{array}
+\]
+\normalsize
+
+\noindent Applying the first action in $\state_0$ yields:
+
+\small
+\[
+\hspace{-1ex}\begin{array}{l}
+\{(=(\pp{is-in}~\pp{Robot})\pp{kitchen}),(=(\pp{is-in}~\pp{box})\pp{kitchen})\}
+\end{array}
+\]
+\normalsize
+
+\noindent with a probability of $0.8$. The assumed state before the
+scheduled execution of action $(\pp{look}~\pp{box}~\pp{kitchen})$ is:
+
+\small
+\[
+\hspace{-1ex}\begin{array}{l}
+\{(=(\pp{is-in}~\pp{Robot})\pp{kitchen}),
+(=(\pp{is-in}~\pp{box})\pp{kitchen}), \\\;\;(=(\pp{is-in}~\pp{cup})\pp{office})\}
+\end{array}
+ \]
+\normalsize
+
+and has a probability of $0.24$.
+
+To describe the optimisation criteria used during sequential sessions
+we model $\assumptiveS{i}$ probabilistically, supposing that its
+application in state $\state$ effects a transition to $\state^{T_i}$
+with probability $\prob_i$, and to $\state^\bot$ with probability $1 -
+\prob_i$. State $\state^\bot$ is an added sink. Taking $\prob_i$ to be
+the probability that the $i^{th}$ sequenced action, $\action_i$, from
+a trace of state-action pairs $\langle \state_0, \action_0,\state_1,
+\action_1,.., \state_N \rangle$ does not transition to $\state^\bot$,
+then the optimal sequential plan has value:
 
@@
 
-\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a
-small positive (negative) numeric quantity which captures
-the utility the agent receives for correctly (incorrectly) rejecting
-an assumption.
-%%
-In terms of action physics, a disconfirm action can only be executed
-once, and otherwise is modelled as a self-transformation.
-%%
-We only consider {\em relevant} assumptions when constructing the
-abstract model.  If \switchAction\ is the action that switched the
-system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
-relevant} if it is necessary for the outcome of \switchAction\ to be
-determined.  Continuing our simplified example, if the trace is:
-
-\small
-\[
-\begin{array}{l}
-\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{kitchen}));\\
-\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}));\\
-(\pp{look}~\pp{box}~\pp{kitchen});
-(\pp{look}~\pp{cup}~\pp{office});\\
-(\pp{report}~\pp{box}~\pp{kitchen}); 
-(\pp{report}~\pp{cup}~\pp{office})
-\end{array}
-\]
-\normalsize
-
-\noindent Taking the switching action \switchAction\ to be
-$(\pp{look}~\pp{box}~\pp{office})$, we have that
-$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
-relevant for this session, and therefore we exclude the corresponding
-disconfirm action from the abstract decision process.
-
-Given \switchAction, we also include another once-only self-transition
-action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
-with the reward property:
+\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a small positive
+(negative) numeric quantity which captures the utility the agent
+receives for correctly (incorrectly) rejecting an assumption.  In
+terms of action physics, a disconfirm action can only be executed
+once, and otherwise is modelled as a self-transformation.  We only
+consider {\em relevant} assumptions when constructing the abstract
+model.  If \switchAction\ is the action that switched the system to a
+DT session, then an assumption $\assumptiveS{i}$ is {\em relevant} if
+it is necessary for the outcome of \switchAction\ to be
+determined. For example, taking the switching action \switchAction\ to
+be $(\pp{look}~\pp{box}~\pp{kitchen})$ from our earlier sequential
+plan example, we have that
+$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}))$ is not
+relevant, and therefore we exclude the corresponding disconfirm action
+from the abstract decision process. Given \switchAction, we also
+include another once-only self-transition action
+$\actions.\poss(\switchAction)$, a \emph{confirmation action} with the
+reward property:
 
@@
   relevant} assumptions. For example, taking the above trace and
-switching action {\texttt (look~box~office)}, the underlying
-belief in Fig.~\ref{fig:beliefs}B would determine a fully
-constrained belief given by Fig.~\ref{fig:beliefs}C.
-%%
-Next, static constraints are removed, one proposition at a time, until
-the number of states that can be true with non-zero probability in the
-initial belief of the abstract process reaches a given threshold.
-%%
-In detail, for each statically-false proposition we compute the {\em
-entropy} of the state assumptions of the current trace {\em
-conditional} on that proposition. 
-%%
-Let $X$ be a set of propositions and $2^X$ the powerset of $X$, then taking
+switching action {\texttt \mbox{(look~box~office)}}, the underlying
+belief in Fig.~\ref{fig:beliefs}B would determine a fully constrained
+belief given by Fig.~\ref{fig:beliefs}A.  Next, static constraints are
+removed, one proposition at a time, until the number of states that
+can be true with non-zero probability in the initial belief of the
+abstract process reaches a given threshold.  In detail, for each
+statically-false proposition we compute the {\em entropy} of the
+relevant assumptions of the current trace {\em conditional} on that
+proposition.  Let $X$ be a set of propositions and $2^X$ the powerset
+of $X$, then taking
 
@@
 $\entropy(X|y_i)~<~\entropy(X|y_j)$. For example, if the serial plan
-assumes the robot is in a kitchen, then propositions about the
-contents of kitchens, e.g. {\texttt (=(is-in~cup)kitchen)}, are
-added to characterise the abstract process' states. Taking
-assumption $X$ to be {\texttt (=(is-in~box)kitchen)}, in relaxing
-static constraints the following entropies are calculated:
+assumes the box is in the kitchen, then propositions about the
+contents of kitchens containing a box,
+e.g. $(=(\pp{is-in}~\pp{milk})\pp{kitchen})$, are added to
+characterise the abstract process' states. Taking a relevant
+assumption $X$ to be $(=(\pp{is-in}~\pp{box})\pp{kitchen})$, in
+relaxing static constraints the following entropies are calculated:
 
@@
 \hline
-\vspace{1ex}
-\scriptsize (A) Partially constrained abstract \mbox{\hspace{4ex}belief-state} & 
-\vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
-\vspace{-2ex} 
-\begin{tabtt}
-(\=:init (=(is-in~R2D2)kitchen) \+ \\
-       (\=.6(and\=(=(is-in~box)kitchen) \+\+ \\
+\scriptsize\vspace{0.5ex}(A) Fully constrained belief & 
+\vspace{0.5ex}\hspace{-4.0ex}(C) Partially constrained belief\\
+\vspace{-3.0ex}\begin{tabtt}
+(\=:init (=(is-in~Robot)kitchen) \+ \\
+       (\=.6(=(is-in~box)kitchen))) \\
+\end{tabtt} & 
+\vspace{-3.0ex}\begin{tabtt}
+\hspace{-5.0ex}(\=:init (=(is-in~Robot)kitchen) \+ \\
+       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
         (\=.9(=(is-in~milk)~kitchen))\+\\
@@
         .9(=(is-in~milk)office)))
-\end{tabtt} &
-\vspace{-2ex} 
-\begin{tabtt}
-\hspace{-6ex}(\=:init (=(is-in~R2D2)office) \+ \\
-       (\=.6(and\=(=(is-in~box)kitchen) \+\+ \\
+\end{tabtt}\\
+\vspace{-16.5ex}(B) Underlying DTPDDL belief & \\ 
+\vspace{-17.0ex}\begin{tabtt}
+(\=:init (=(is-in~Robot)kitchen) \+ \\
+       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
         (\=.9(=(is-in~milk)kitchen))\+\\
@@
       .4(=(is-in~cup)kitchen))) \\
-\end{tabtt} \\
-\vspace{-6ex}\scriptsize(C) Fully constrained abstract \mbox{\hspace{5ex}belief-state} & \\
-\vspace{-5ex} \begin{tabtt}
-(\=:init (=(is-in~R2D2)kitchen) \+ \\
-       (\=.6(=(is-in~box)kitchen))) \\
 \end{tabtt} & \\
@@
 \end{figure}
+
+%% \begin{figure}
+%% \tiny
+%% \begin{tabularx}{\columnwidth}{XX}
+%% \hline
+%% \vspace{1ex}
+%% \scriptsize (A) Partially constrained belief & %abstract belief-state & 
+%% \vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
+%% \vspace{-2ex} 
+%% \begin{tabtt}\hspace{-4ex}
+%% (\=:init (=(is-in~Robot)kitchen) \+ \\
+%%        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
+%%         (\=.9(=(is-in~milk)~kitchen))\+\\
+%%          .1(=(is-in~milk)office))\-\-  \\
+%%        .4(and(=(is-in~box)office) \+\\
+%%         (.1(=(is-in~milk)kitchen))\+\\
+%%         .9(=(is-in~milk)office)))
+%% \end{tabtt} &
+%% \vspace{-2ex} 
+%% \begin{tabtt}
+%% \hspace{-4ex}(\=:init (=(is-in~Robot)office) \+ \\
+%%        (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
+%%         (\=.9(=(is-in~milk)kitchen))\+\\
+%%          .1(=(is-in~milk)office))\-\-  \\
+%%        .4(and(=(is-in~box)office) \+\\
+%%         (.1(=(is-in~milk)kitchen))\+\\
+%%         .9(=(is-in~milk)office))) \-\-\-\\
+%%        (\=.6(=(is-in~cup)office) \+ \\
+%%       .4(=(is-in~cup)kitchen))) \\
+%% \end{tabtt} \\
+%% \vspace{-8ex}\scriptsize(C) Fully constrained belief & \\
+%% \vspace{-7.5ex} \begin{tabtt}
+%% (\=:init (=(is-in~Robot)kitchen) \+ \\
+%%        (\=.6(=(is-in~box)kitchen))) \\
+%% \end{tabtt} & \\
+%% \hline
+%% \end{tabularx}
+%% \caption{Simplified examples of abstract belief-states from
+%%        DT sessions. \label{fig:beliefs}}
+%% \normalsize
+%% \end{figure}
 

### 2011-05-03 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
remove comma

@@
 /Subject (Proceedings of the Twenty-Fifth Conference on Artificial Intelligence)
-/Keywords (planning under uncertainty)
+/Keywords (planning under uncertainty, continual planning, robot planning)
 /DOCINFO pdfmark
@@
 \And
-Charles Gretton, Richard Dearden,\\
+Charles Gretton, Richard Dearden\\
 University of Birmingham, United Kingdom\\

### 2011-05-19 code/systems/dora-yr3-interactive/subarchitectures/dialogue.sa/apis/dialogue/test/java/de/dfki/lt/tr/dialmanagement/components/DialogueManagerYr2Test.java
branched for new interactive dora

@@ -1,0 +1,232 @@
+package de.dfki.lt.tr.dialmanagement.components;
+
+import static org.junit.Assert.*;
+
+import java.util.Arrays;
+
+import org.junit.Before;
+import org.junit.Test;
+
+import de.dfki.lt.tr.beliefs.slice.intentions.CommunicativeIntention;
+import de.dfki.lt.tr.beliefs.slice.intentions.Intention;
+import de.dfki.lt.tr.beliefs.slice.intentions.IntentionalContent;
+import de.dfki.lt.tr.beliefs.slice.logicalcontent.ModalFormula;
+import de.dfki.lt.tr.beliefs.slice.logicalcontent.dFormula;
+import de.dfki.lt.tr.dialmanagement.arch.DialogueException;
+import de.dfki.lt.tr.dialmanagement.data.policies.DialoguePolicy;
+import de.dfki.lt.tr.dialmanagement.data.policies.PolicyAction;
+import de.dfki.lt.tr.dialmanagement.utils.EpistemicObjectUtils;
+import de.dfki.lt.tr.dialmanagement.utils.FormulaUtils;
+import de.dfki.lt.tr.dialmanagement.utils.XMLPolicyReader;
+
+/**
+ * Tests for the yr2 CogX review meeting
+ * 
+ * @author Pierre Lison (plison@dfki.de)
+ * @version 11/10/2010
+ *
+ */
+public class DialogueManagerYr2Test {
+
+
+	// logging and debugging
+	public static boolean LOGGING = true;
+	public static boolean DEBUG = true;
+	
+	// the configuration files
+	public static String POLICYFILE = "subarchitectures/dialogue.sa/config/policies/yr2/fullpolicy.xml";
+
+	// the dialogue manager
+	public DialogueManager manager;
+	
+	
+	
+	/**
+	 * Construct the policy based on the configuration files
+	 * 
+	 * @throws DialogueException if the files are ill-formatted
+	 */
+	@Before
+	public void constructPolicy() throws DialogueException {
+		
+		DialoguePolicy policy = XMLPolicyReader.constructPolicy(POLICYFILE);	
+		policy.ensureWellFormedPolicy();	
+		manager = new DialogueManager(policy);
+	}
+	
+	/**
+	 * where forwarding shouldn't fire
+	 * 
+	 * @throws DialogueException
+	 */
+	@Test
+	public void blockforwarding() throws DialogueException {
+		
+		dFormula formula = FormulaUtils.constructFormula("<belief>(anything ^ here)");
+		
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/aldoma/papers/learning_scale/learning_scale.tex
TUW V4R (vision for robotics) software suite, used by various TUW components

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage[pdftex]{graphicx}
\usepackage[usenames,dvips,pdftex]{color}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{amsmath, amsfonts}
\usepackage[nice]{nicefrac}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{cite}
\usepackage{units}
\usepackage{url}
%\usepackage[noend]{sty/algorithmic}

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Learning Real Scale of Object Instances And Categories Scale Distributions from CAD Models
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Aitor Aldoma, Walter Wohlkinger and Markus Vincze% <-this % stops a space
\thanks{This work was not supported by any organization}% <-this % stops a space
\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
        University of Twente, 7500 AE Enschede, The Netherlands
        {\tt\small h.kwakernaak@autsubmit.com}}%
\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
        Dayton, OH 45435, USA
        {\tt\small pmisra@cs.wright.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

This paper presents a method to automatically learn a scaled CAD model representation
of object instances seen by a depth sensor. The CAD models representing objects belonging to different categories
are downloaded automatically from the Internet. When an object is placed in front of the camera, the system tries
to recognize it and estimate their pose against the already learned CAD models using a fast scale-dependant 3D feature. Upon failure, a learning stage is triggered that
categorizes the object using a scale invariant view based descriptor to retrieve CAD models from the categories database with
similar geometries as seen from the sensor viewpoint. The retrieved CAD models are fitted against the partial view of the object, 
the model that best fits the partial view is selected and a scaled version of model that fits the object instance is stored for future recognition. Moreover, after
a certain amount of object instances has been learned, the system learns a scale distribution for each category that can be used to filter false
positives based on the already-learned distribution to improve the efficiency of the learning stage. This approach has several advantages over: (i) high-precission scanners as they
have a high cost and the scanning process is cumbersome and (ii) mesh reconstruction methods based on images or fusion of partial views regarding accuracy and completeness of the
CAD model. We demonstrate the use of this approach to learn a set of object instances and present recognition and pose estimations results using the scale invariant semi-global
3D feature CVFH.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{FRAMEWORK}

\onecolumn
\begin{figure}[htb]
  \begin{center}
    \includegraphics[angle=270,width=\columnwidth]{images/framework.pdf}
    \caption{Framework}
    \label{fig:framework}
  \end{center}
\end{figure}

\twocolumn

\section{LEARNING STAGE}

\subsection{Categorization}

\subsection{CAD model fitting}

\section{CATEGORIES SCALE DISTRIBUTION}

\section{SCALE DEPENDANT RECOGNITION AND POSE ESTIMATION}

\section{EVALUATION}

The CAD model database from the Internet contains approx. $8000$ objects belonging
to $200$ categories published online ...

We evaluate the system in two ways: (i) we select $20$ real objects. We manually search for each real object  and at the same time learn them using categorization and stable pose
alignment obtaining . We can then evaluate if

(ii) we try to learn $50$ objects
belonging to $5$ different categories using the scale invariant categorization
method and stable pose alignment presented befofe and we use the learned models
to perform recognition and pose estimation on several scenes containing this objects using CVFH.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ACKNOWLEDGMENTS}

The authors gratefully acknowledge the contribution of National Research Organization and reviewers' comments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1}
J.G.F. Francis, The QR Transformation I, {\it Comput. J.}, vol. 4, 1961, pp 265-271.

\bibitem{c2}
H. Kwakernaak and R. Sivan, {\it Modern Signals and Systems}, Prentice Hall, Englewood Cliffs, NJ; 1991.

\bibitem{c3}
D. Boley and R. Maier, "A Parallel QR Algorithm for the Non-Symmetric Eigenvalue Algorithm", {\it in Third SIAM Conference on Applied Linear Algebra}, Madison, WI, 1988, pp. A20.

\end{thebibliography}

\end{document}

### 2011-09-01 code/tools/v4r/trunk/v4r/Tracker/Filter.h

@@ -1,0 +1,74 @@
+
+
+#ifndef _TRACKING_SMOOTH_FILTER_
+#define _TRACKING_SMOOTH_FILTER_
+
+ 
+namespace Tracking{
+
+class FloatFilter
+{
+public:
+    FloatFilter(float a=0.0f){ this->a = a; b = 1.0f - a; z = 0.0f; }
+
+	inline void Set(const float &z) { this->z = z; }
+
+	operator float() const { return z; }
+	FloatFilter& operator=(const float &in){
+		if(!isnan(in))
+			z = (in * b) + (z * a);
+		return (*this);
+	}
+
+private:
+    float a, b, z;
+};
+
+class FloatFilterRise
+{
+public:
+    FloatFilterRise(float a=0.0f){ this->a = a; b = 1.0f - a; z = 0.0f; }
+
+	inline void Set(const float &z) { this->z = z; }
+
+	operator float() const { return z; }
+	FloatFilterRise& operator=(const float &in){
+		if(!isnan(in)){
+			if(in>=z)
+				z = (in * b) + (z * a);
+			else
+				z = in;
+		}
+		return (*this);
+	}
+
+private:
+    float a, b, z;
+};
+
+class FloatFilterFall
+{
+public:
+    FloatFilterFall(float a=0.0f){ this->a = a; b = 1.0f - a; z = 0.0f; }
+
+	inline void Set(const float &z) { this->z = z; }
+
+	operator float() const { return z; }
+	FloatFilterFall& operator=(const float &in){
+		if(!isnan(in)){
+			if(in<=z)
+				z = (in * b) + (z * a);
+			else
+				z = in;
+		}
+		return (*this);
+	}
+
+private:
+    float a, b, z;
+};
+
+
+} 
+ 
+#endif

### 2011-09-01 code/tools/v4r/trunk/v4r/aldoma/generic_trainer_recognizer/src/estimators/cvfh_estimator.cpp

@@ -1,0 +1,63 @@
+#include "estimators/cvfh_estimator.h"
+#include "pcl/visualization/pcl_visualizer.h"
+
+template<typename PointInT, typename PointOutT, typename FeatureT>
+void
+CVFHEstimator<PointInT, PointOutT, FeatureT>::computeCVFH (
+    PointOutTPtr cloud_normals,
+                                                           std::vector<pcl::PointCloud<FeatureT>,
+                                                               Eigen::aligned_allocator<pcl::PointCloud<
+                                                               FeatureT> > > & vfh_signatures,
+                                                           std::vector<Eigen::Vector3f> & centroids_dominant_orientations)
+{
+  //compute CVFH
+  typedef typename pcl::CVFHEstimation<PointOutT, PointOutT, pcl::VFHSignature308> CVFHEstimation;
+  pcl::PointCloud < FeatureT > cvfh_signatures;
+
+  CVFHEstimation cvfh;
+  cvfh.setClusterTolerance(leaf_size_ * leaf_size_factor_);
+  cvfh.setLeafSize(leaf_size_);
+  cvfh.setEPSAngleThreshold(eps_angle_threshold_);
+  cvfh.setCurvatureThreshold(curvature_threshold_);
+  cvfh.setNormalizeBins(normalize_bins_);
+
+  typename pcl::KdTree<PointOutT>::Ptr cvfh_tree = boost::make_shared<pcl::KdTreeFLANN<PointOutT> > ();
+  cvfh.setSearchMethod (cvfh_tree);
+  cvfh.setInputCloud (cloud_normals);
+  cvfh.setInputNormals (cloud_normals);
+  cvfh.compute (cvfh_signatures);
+
+
+  //pcl::io::savePCDFile("clusters_colored.pcd", *cvfh.clusters_colored_);
+  std::cout << "thresholds:" << eps_angle_threshold_ << " " << curvature_threshold_ << "normalize_bins:" << normalize_bins_ << std::endl;
+  //std::cout << "compute CVFH:" << cvfh_signatures.size() << std::endl;
+
+  for (size_t i = 0; i < cvfh_signatures.points.size (); i++)
+  {
+    pcl::PointCloud < FeatureT > vfh_signature;
+    vfh_signature.points.resize (1);
+    vfh_signature.width = vfh_signature.height = 1;
+    for (int d = 0; d < 308; ++d)
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/aldoma/overlap_likelihood/include/overlap.h

@@ -1,0 +1,42 @@
+/*
+ * overlap.h
+ *
+ *  Created on: Jun 6, 2011
+ *      Author: aitor
+ */
+
+#ifndef OVERLAP_H_
+#define OVERLAP_H_
+
+#include "pcl/common/common.h"
+#include "vtkSmartPointer.h"
+#include "vtkPolyData.h"
+//#include "vtkSmartPointer.h"
+
+class OverlapLikelihood {
+  private:
+    vtkSmartPointer<vtkPolyData> poly_input_;
+    pcl::PointCloud<pcl::PointXYZ>::Ptr input_;
+    vtkSmartPointer<vtkPolyData> poly_output_;
+    float MAX_DIST_;
+    std::string path_for_vtk_files_;
+    void getVerticesAsPointCloud(vtkSmartPointer<vtkPolyData> & poly, pcl::PointCloud<pcl::PointXYZ>::Ptr & vertices);
+
+  public:
+    OverlapLikelihood(std::string path_for_vtk_files, float MAX_DIST=0.01);
+    ~OverlapLikelihood();
+
+    //Input
+    void setModel(vtkSmartPointer<vtkPolyData> & poly);
+    void setModel(std::string path_to_ply, Eigen::Matrix4f & transform, float scale = 1.0);
+    void setPartialPointCloud(pcl::PointCloud<pcl::PointXYZ>::Ptr & input);
+
+    //Computation
+    void compute();
+
+    //Output
+    void getOverlapLikelihoodPolydata(vtkSmartPointer<vtkPolyData> & poly_likelihood);
+    void generateIVModel(std::string & path_to_iv);
+};
+
+#endif /* OVERLAP_H_ */

### 2011-09-01 code/tools/v4r/trunk/v4r/daisy/matrix_manipulation.h

@@ -1,0 +1,142 @@
+////////////////////////////////////////////////////////////////////////////////
+// This program is free software; you can redistribute it                     //
+// and/or modify it under the terms of the GNU General Public License         //
+// version 2 (or higher) as published by the Free Software Foundation.        //
+//                                                                            //
+// This program is distributed in the hope that it will be useful, but        //
+// WITHOUT ANY WARRANTY; without even the implied warranty of                 //
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU          //
+// General Public License for more details.                                   //
+//                                                                            //
+// Written and (C) by                                                         //
+// Engin Tola                                                                 //
+//                                                                            //
+// web   : http://cvlab.epfl.ch/~tola/                                        //
+// email : engin.tola@epfl.ch                                                 //
+//                                                                            //
+// If you use this code for research purposes, please refer to the            //
+// webpage above                                                              //
+////////////////////////////////////////////////////////////////////////////////
+
+#ifndef KUTILITY_MATRIX_MANIPULATION_H
+#define KUTILITY_MATRIX_MANIPULATION_H
+
+#include <opencv/cv.h>
+
+namespace kutility
+{
+   template<class T>
+   inline void transpose(T* m, int rn, int cn, T* mt)
+   {
+      for( int r=0; r<rn; r++ )
+         for( int c=0; c<cn; c++ )
+            mt[ c*rn+r ] = m[ r*cn+c ];
+   }
+
+   template<typename T1, typename T2, typename T3> inline
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/edgelinker/contour.cpp

@@ -1,0 +1,737 @@
+/**
+ * @file contour
+ * @author Markus Bader
+ * @date Mo Jul 9 2007
+ * @version 0.1
+ * @brief
+ *
+ * @see
+ **/
+#include "contour.h"
+#include <iostream>
+#include <stack>
+#include <cstdio>
+#include "opencv/highgui.h"
+
+namespace V4R {
+
+const int Contour::ANGLE_8U = 8;
+const int Contour::ANGLE_32F = 32;
+const int Contour::ANGLE_64F = 64;
+
+
+const int Contour::MODE_SIMPLE = 1;
+const int Contour::MODE_CONTOUR = 2;
+const int Contour::MODE_GRAIDENT = 3;
+const int Contour::MODE_COMPLEX = 4;
+
+const float Contour::fPi  = 3.1415926535897932384626433832795;
+const double Contour::dPi = 3.1415926535897932384626433832795;
+
+const int Contour::pppDirectionWeightsField[4][3][3] = {
+    { {2,1,2}, // 0 0 0
+        {4,0,4}, // 1 1 1
+        {2,1,2}  // 0 0 0
+    },
+    { {1,2,4}, // 0 0 1
+        {2,0,2}, // 0 1 0
+        {4,2,1}  // 1 0 0
+    },
+    { {2,4,2}, // 0 1 0
+        {1,0,1}, // 0 1 0
+        {2,4,4}  // 0 1 0
+    },
+    { {4,2,1}, // 1 0 0
+        {2,0,2}, // 0 1 0
+        {1,2,4}  // 0 0 1
+    }
+};
+
+const int Contour::ppGradientWeightsField[4][9] = {
+    {3, 5, 6, 8, 2, 0, 7, 1, 4},
+    {2, 6, 1, 3, 7, 5, 0, 8, 4},
+    {1, 7, 0, 6, 8, 2, 3, 5, 4},
+    {0, 8, 1, 7, 5, 3, 2, 6, 4}
+};
+
+const int Contour::pppCommingFromEdgeWeightsField[9][3][3] = {
+    { {0,1,2}, // 1 0 0
+        {1,0,3}, // 0 0 0
+        {2,3,4}  // 0 0 0
+    },
+    { {1,0,1}, // 0 1 0
+        {2,0,2}, // 0 0 0
+        {3,4,3}  // 0 0 0
+    },
+    { {2,1,0}, // 0 0 1
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/elldetect/syFormClosedSegmentGroups.hpp

@@ -1,0 +1,52 @@
+//
+// (C) 2010, Aitor Aldoma Buchaca,
+//           Johann Prankl,
+//           Gerhard Obernberger <gerhard@obernberger.at>
+//
+//
+
+#ifndef SYFORMSEGMENTGROUPS_HPP
+#define SYFORMSEGMENTGROUPS_HPP
+
+#include "syArray.hpp"
+#include "syEdgel.hpp"
+#include "sySegment.hpp"
+#include "syVoteImage.hpp"
+
+NAMESPACE_CLASS_BEGIN( RTE )
+
+//begin class///////////////////////////////////////////////////////////////////
+//
+//    CLASS DESCRIPTION:
+//       RTE form closed segment groups
+//
+//    FUNCTION DESCRIPTION:
+//
+////////////////////////////////////////////////////////////////////////////////
+class CzFormClosedSegmentGroups
+{
+private:
+
+  CzVoteImage *voteImg;
+
+public:
+   CzFormClosedSegmentGroups();
+   ~CzFormClosedSegmentGroups(); 
+   
+   // Release vote image
+   void Reset(IplImage *img);
+   
+   // Find lines
+   void Operate(CzArray<CzSegment*> &segments, int iLinkingGrowCount);
+   // Draw the vote image onto the image img (for debugging purposes)
+	void DrawVoteImg(IplImage *img, bool use_colour);
+	// Close a contour
+	void CloseContour(CzSegment * seg, CzArray<int> &idsUsed, CzSegment * joined);
+};
+//end class/////////////////////////////////////////////////////////////////////
+
+ 
+NAMESPACE_CLASS_END()
+
+#endif
+

### 2011-09-01 code/tools/v4r/trunk/v4r/elldetect/sySegment.cpp

@@ -1,0 +1,85 @@
+//
+// (C) 2010, Aitor Aldoma Buchaca,
+//           Johann Prankl,
+//           Michael Zillich,
+//           Gerhard Obernberger <gerhard@obernberger.at>
+//
+//
+
+#include "syDebug.hpp"
+
+#include <algorithm>
+#include "sySegment.hpp"
+
+NAMESPACE_CLASS_BEGIN( RTE )
+
+
+////////////////////////////////////////////////////////////////////////////////
+CzSegment::CzSegment()
+{
+	end = NULL;
+	start = NULL;
+	s = e = false;
+	closable = false;
+	closed = false;
+}
+
+////////////////////////////////////////////////////////////////////////////////
+// Construct a segment with a given number of edgels.
+CzSegment::CzSegment(const CzArray<CzEdgel> &arr) 
+   : edgels(arr.Size())
+{
+   edgels = arr;
+   end = NULL;
+   start = NULL;
+   s = e = false;
+   closable = false;
+   closed = false;
+}
+
+////////////////////////////////////////////////////////////////////////////////
+// Returns true if any part of the Segment is at pixel position (x,y).
+bool CzSegment::IsAtPosition(int x, int y)
+{
+   double xd = (double)x, 
+          yd = (double)y;
+   for (unsigned i = 0; i < edgels.Size(); i++) {
+      if (IsEqual(xd, edgels[i].p.x) && IsEqual(yd, edgels[i].p.y)) {
+         return true;
+      }
+   }
+   return false;
+}
+
+////////////////////////////////////////////////////////////////////////////////
+// Returns the normalised tangent direction at the specified pixel.
+// Range of points over which tangent is estimated is adapted automatically.
+// Starting with points i-1 to i+1, extends range until line through first and
+// last point no longer fits the points between.
+// Note: direction points "forward" in terms of counting pixels on the edge.
+CzVector2 CzSegment::Tangent(int i, int l, int u)
+{
+   const double MAX_DIST = 1.; // nasty threshold
+   if (u == (int)UNDEF_ID)
+      u = edgels.Size() - 1;
+   int j, start = max(l, i - 1);
+   int end = min(i + 1, u);
+   bool on_line = true;
+   CzVector2 d, d_out;
+   d_out = Normalise(edgels[end].p - edgels[start].p);
+   while (on_line && (start >= l && end <= u)) {
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/figure/figure.h

@@ -1,0 +1,85 @@
+/*
+    <one line to give the program's name and a brief idea of what it does.>
+    Copyright (C) 2011  Markus Bader <email>
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+
+#ifndef V4R_FIGURE_H
+#define V4R_FIGURE_H
+#include <opencv/cv.h>
+#include <opencv/cv.hpp>
+
+namespace V4R {
+
+class Figure
+{
+public:
+    class Parameters {
+    public:
+        enum CoordinateSystem {
+            left_handed = 0,
+            right_handed = 1 // Austrian-Kangaroos System
+        };
+        Parameters();
+        Parameters(const Parameters& param);
+        std::string name;
+        cv::Size size;
+        cv::Rect_<double> roi;
+        double alpha;
+        CoordinateSystem system;
+        double grid;
+    };
+    Figure();
+    void init();
+    void init(const Parameters& param);
+    void clear();
+    void show(int ms = 10);
+    void plotOrigin(bool write = true);
+    void setGrid(double boxSize, uchar r = 200, uchar g = 200, uchar b = 200);
+    void setColor(uchar r, uchar g, uchar b, double opacity = 1);
+    void setBackground(uchar r, uchar g, uchar b);
+    void setFont(int fontFace);
+    void plotLine(double x1, double y1, double x2, double y2, int thickness=1, int lineType=8);
+    void plotArrow(double x1, double y1, double x2, double y2, int thickness=1, int lineType=8);
+    void plotDot(double x, double y);
+    void plotGrid();
+    void plotText(double x, double y, const std::string& text, int ox = 0, int oy = 0);
+    void plotText(double x, double y, double nr, int ox = 0, int oy = 0);
+    void plotText(double x, double y, int nr, int ox = 0, int oy = 0);
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/geometry/quaterion.h

@@ -1,0 +1,101 @@
+/*
+    Copyright (c) <year>, <copyright holder>
+    All rights reserved.
+
+    Redistribution and use in source and binary forms, with or without
+    modification, are permitted provided that the following conditions are met:
+        * Redistributions of source code must retain the above copyright
+        notice, this list of conditions and the following disclaimer.
+        * Redistributions in binary form must reproduce the above copyright
+        notice, this list of conditions and the following disclaimer in the
+        documentation and/or other materials provided with the distribution.
+        * Neither the name of the <organization> nor the
+        names of its contributors may be used to endorse or promote products
+        derived from this software without specific prior written permission.
+
+    THIS SOFTWARE IS PROVIDED BY <copyright holder> ''AS IS'' AND ANY
+    EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
+    WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
+    DISCLAIMED. IN NO EVENT SHALL <copyright holder> BE LIABLE FOR ANY
+    DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES
+    (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES;
+    LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND
+    ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
+    (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS
+    SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+
+*/
+
+#ifndef V4R_QUATERION_H
+#define V4R_QUATERION_H
+
+#include <opencv/cv.h>
+
+namespace V4R {
+
+#ifndef V4R_QUATERION
+#define V4R_QUATERION
+template <typename T>
+class Quaterion  : public cv::Vec<T,4> {
+public:
+    Quaterion() {};
[…]

### 2011-09-01 code/tools/v4r/trunk/v4r/rosextensions/param.h

@@ -1,0 +1,39 @@
+/*
+ * param.h
+ *
+ *  Created on: Feb 13, 2011
+ *  Author: Markus Bader
+ */
+
+#ifndef PARAM_H_
+#define PARAM_H_
+
+namespace ROS{
+
+void updateParameter(ros::NodeHandle n, const std::string &name, std::vector<std::string> &value, std::string defaults) {
+  std::string tmp = defaults;
+  n.getParam(name.c_str(), tmp);
+  ROS_INFO("\t%s: ->  %s", name.c_str(), tmp.c_str());
+  boost::erase_all(tmp, " ");
+  boost::split(value, tmp, boost::is_any_of(";,"));
+}
+
+void updateParameter(ros::NodeHandle n, const std::string &name, int &value) {
+  int tmp = value;
+  n.getParam(name.c_str(), tmp);
+  if (tmp != value) {
+    ROS_INFO("\t%s: %i ->  %i", name.c_str(), value, tmp);
+    value = tmp;
+  }
+}
+
+void updateParameter(ros::NodeHandle n, const std::string &name, double &value) {
+  double tmp = value;
+  n.getParam(name.c_str(), tmp);
+  if (tmp != value) {
+    ROS_INFO("\t%s: %5.2f ->  %5.2f", name.c_str(), value, tmp);
+    value = tmp;
+  }
+}
+}
+#endif /* PARAM_H_ */

### 2011-09-01 tools/v4r/trunk/v4r/aldoma/papers/affordances/affordances.tex
TUW V4R (vision for robotics) software suite, used by various TUW components

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Supervised Learning Of Hidden and Non-Hidden $0$-order Affordances
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

%\author{Huibert Kwakernaak and Pradeep Misra% <-this % stops a space
%\thanks{This work was not supported by any organization}% <-this % stops a space
%\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
%        University of Twente, 7500 AE Enschede, The Netherlands
%        {\tt\small h.kwakernaak@autsubmit.com}}%
%\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
%        Dayton, OH 45435, USA
%        {\tt\small pmisra@cs.wright.edu}}%
%}

\author{Aitor Aldoma, Markus Vincze\\
ACIN - Vienna University of Technology\\
{\tt\small {aldoma,vincze}@tuwien.ac.at}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Federico Tombari\\
Computer Vision Lab\\
DEIS - ARCES, University of Bologna\\
{\tt\small {federico.tombari}@unibo.it}
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

%This paper focuses on the learning and detection of $0$-order affordances.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

From a robotic perspective being able to understand a
scene and moreover, understanding which are the interaction
possibilities that are provided in a specific environment, are
a key capability for a task-guided robotic agent. What an
environment affords depends strongly on two factors: (i) the
objects and their configuration in the world and second,
the interaction capabilities embodied on a specific agent.
The combination of both factors is coined under the term
affordance in the literature:

\begin{quote} "Affordances relate the utility of things, events, and
places to the needs of animals and their actions in fulfilling
them [...]. Affordances themselves are perceived and, in
fact, are the essence of what we perceive." \end{quote}

This quote provides an intuitive definition of the term affordance
and moreover, poses an opportunity to motivate our analysis on
affordances from a robotic perspective.

The affordances of an object - what we term under $0$-order affordances -
are supported by geometrical -local or global- properties of the object. For instance, objects like chairs or sofas
can be used for sitting as they provide a parallel surface to the ground and a perpendicular one used to lean back and mugs, bowls
and in general containers, are used for liquid-containment because they provide a closed concavity. $0$-order affordances
do not depend solely on the geometry of the objects but also on their configuration in the world. Liquid containers can only
fulfill their function if they are in an upright pose or objects like sofas and chair can be used for sitting when found
in a specific pose.

The following NEEDS TO BE STRUCTURED AND PHRASED PROPERLY!!

\hspace{5 mm}

Moreover, if we think about perceiving affordances directly from a certain viewpoint, 
things get even more interesting as the ... TODO: for example a chair seen from behind does not afford sittable ...

\hspace{2 mm}

From a robotics perspective, do we really want to be able to ONLY perceive affordances when directly perceivable.... We can manipulate
the environment ... Contradict Gibson!!

\hspace{2 mm}

Object recognition based on CAD models...

\hspace{2 mm}

Supervised-Learning of affordances based on geometry for CAD models...

\hspace{2 mm}

Talk about why stable poses can used for this things because we live in an structured environment and man-made objects have been
designed to fulfill their functions usually when found in a certain stable pose... Objects that fulfill the same function share
at least one stable plane!!

\hspace{2 mm}

If we solve object recognition => we get a direct mapping for affordances

\hspace{2 mm}

%Gibson says "[...] affordances themselves are perceived [...] "
%\hspace{5 mm}
%\begin{quote} "There’s little we can find in common to all chairs – except
%for their intended use." \end{quote}

%This quotes extracted from ancient literature about affordances emphasize tha

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RELATED WORK}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{LEARNING $0$-ORDER AFFORDANCES}

Which affordances

Why use stable planes?

\subsection{Affordance labelling on CAD models}

Stable planes

Picture of the tool

\subsection{Classifiers}

Evaluation of different classifiers for one or two affordances.

\section{OBJECT RECOGNITION}

Evaluation of different descriptors

\section{DETECTING HIDDEN AND NON-HIDDEN AFFORDANCES}

Explain how once the pose is retrieved from the object recognition module,
how the stable pose is found if any.

\section{EVALUATION}

Evaluation of the whole process.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS AND FUTURE WORKS}

\subsection{Conclusions}

\subsection{Future Works}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ACKNOWLEDGMENTS}

The authors gratefully acknowledge the contribution of National Research Organization and reviewers' comments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1}
J.G.F. Francis, The QR Transformation I, {\it Comput. J.}, vol. 4, 1961, pp 265-271.

\bibitem{c2}
H. Kwakernaak and R. Sivan, {\it Modern Signals and Systems}, Prentice Hall, Englewood Cliffs, NJ; 1991.

\bibitem{c3}
D. Boley and R. Maier, "A Parallel QR Algorithm for the Non-Symmetric Eigenvalue Algorithm", {\it in Third SIAM Conference on Applied Linear Algebra}, Madison, WI, 1988, pp. A20.

\end{thebibliography}

\end{document}

### 2011-09-01 tools/v4r/trunk/ak/shm/shmimage.hpp

@@ -1,0 +1,163 @@
+/***************************************************************************
+ *   Copyright (C) 2010 by Markus Bader and Bernhard Miller                *
+ *   markus.bader@tuwien.ac.at                                             *
+ *                                                                         *
+ *   This program is free software; you can redistribute it and/or modify  *
+ *   it under the terms of the GNU General Public License as published by  *
+ *   the Free Software Foundation; either version 2 of the License, or     *
+ *   (at your option) any later version.                                   *
+ *                                                                         *
+ *   This program is distributed in the hope that it will be useful,       *
+ *   but WITHOUT ANY WARRANTY; without even the implied warranty of        *
+ *   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the         *
+ *   GNU General Public License for more details.                          *
+ *                                                                         *
+ *   You should have received a copy of the GNU General Public License     *
+ *   along with this program; if not, write to the                         *
+ *   Free Software Foundation, Inc.,                                       *
+ *   59 Temple Place - Suite 330, Boston, MA  02111-1307, USA.             *
+ ***************************************************************************/
+
+
+/**
+ * @file shmimage
+ * @author Markus Bader
+ * @date Aug 4 2011
+ * @version 0.1
+ * @brief
+ *
+ * @see
+ **/
+
+
+
+#ifndef AKSHAREDAKIMAGE_HPP
+#define AKSHAREDAKIMAGE_HPP
+
+#include <ak/cv/akimage.h>
+#include <ak/shm/shmvar.hpp>
+#include <stdint.h>
+#include <string.h>
+
+namespace ak {
+
+
+/**
+* @brief Image class base on the opencv IplImage
+* @author Markus Bader <markus.bader@austrian-kangaroos.com>
+*/
+class ShmImage : public ShmArray<char, Shm::VAR_IMAGE> {
+    static const int HEADER_SIZE = 0x400;
+public:
+    /**
+    * @brief constructor<br>
+    * all data pointer will be NULL
+    **/
+    ShmImage() {
+    }
+
+    /**
+      * @brief creates a the shared memory if it does not yet exist<br>
[…]

### 2011-09-01 tools/v4r/trunk/v4r/CModelRecogniser/CModel.cc

@@ -1,0 +1,154 @@
+/**
+ * $Id$
+ */
+
+
+#include "CModel.hh"
+
+
+
+
+namespace P 
+{
+
+
+
+/************************************************************************************
+ * Constructor/Destructor
+ */
+
+CModel::CModel(int subdivHist)
+{
+  if (subdivHist>0)
+    viewHist = new SphereHistogram(subdivHist);
+}
+
+CModel::~CModel()
+{
+}
+
+
+/***************************************************************************************/
+/**
+ * Deep copy a model
+ */
+void CModel::copyTo(cv::Ptr<CModel> &dst)
+{
+  if (dst.empty()) dst = new CModel();
+
+  dst->id = id;
+
+  dst->views.resize(views.size());
+
+  for (unsigned i=0; i<views.size(); i++)
+    views[i]->copyTo(dst->views[i]);
+
+  if (!viewHist.empty())
+  {
+    dst->viewHist = new SphereHistogram();
+    *dst->viewHist = *viewHist;
+  }
+
+  dst->center = center;
+  dst->pose = pose;
+}
+
+/**
+ * clear()
+ */
+void CModel::clear()
+{
+  views.clear();
+  if (!viewHist.empty()) viewHist->Clear();  
+}
+
+
+/**
+ * stream to file
+ */
+void CModel::save(ofstream &os)
+{
+  os<<id<<'\n';
+  os<<center.x<<' '<<center.y<<' '<<center.z<<'\n';
+  os<<views.size()<<'\n';
+  for (unsigned i=0; i<views.size(); i++)
+  {
+    views[i]->id = i;
+    views[i]->save(os);
+  }
+
+    
+  //save links
+  for (unsigned j=0; j<views.size(); j++)
+  {
+    View &view = *views[j];
+    set<PKeypoint*>::iterator it;
+    for (unsigned i=0; i<view.keys.size(); i++)
+    {
+      os<<view.keys[i]->links.size()<<' ';
+      for (it=view.keys[i]->links.begin(); it!=view.keys[i]->links.end(); it++)
+      {
+        if ((*it)->HaveView())
+          os<<(*it)->view->id<<' '<<(*it)->id<<' ';
+        else
+          os<<UINT_MAX<<' ';
+      }
+      os<<'\n';
+    }
+  }
+  
+}
+
+/**
+ * stream from file
+ */
+void CModel::load(ifstream &is)
+{
[…]

### 2011-09-01 tools/v4r/trunk/v4r/CModelRecogniser/IMCluster.cc

@@ -1,0 +1,21 @@
+/**
+ * $Id$
+ */
+
+
+#include "IMCluster.hh"
+
+namespace P 
+{
+
+void DeleteIMClusters(std::vector<IMCluster *> &clusters)
+{
+  for (unsigned i=0; i<clusters.size(); i++)
+    delete clusters[i];
+  clusters.clear();
+}
+
+
+
+}
+

### 2011-09-01 tools/v4r/trunk/v4r/CModelRecogniser/SphereHistogram.cc

@@ -1,0 +1,538 @@
+/**
+ * $Id$
+ */
+
+
+#include "SphereHistogram.hh"
+
+
+namespace P 
+{
+
+
+
+
+
+
+/********************** SphereHistogram ************************
+ * Constructor/Destructor
+ */
+SphereHistogram::SphereHistogram(unsigned subdevisions)
+ : sumWeight(0.)
+{
+  Init(subdevisions);
+}
+
+SphereHistogram::~SphereHistogram()
+{
+  Release();
+}
+
+/**
+ * Init icosahedron
+ */
+void SphereHistogram::InitIcosahedron()
+{
+  double t = (1.+sqrt(5.))/2.;
+  double tau = t/sqrt(1.+t*t);
+  double one = 1./sqrt(1.+t*t);
+
+  vertices.resize(12);
+
+  vertices[0] = cv::Point3d(tau, one, 0.0);
+  vertices[1] = cv::Point3d(-tau, one, 0.0);
+  vertices[2] = cv::Point3d(-tau, -one, 0.0);
+  vertices[3] = cv::Point3d(tau, -one, 0.0);
+  vertices[4] = cv::Point3d(one, 0.0 ,  tau);
+  vertices[5] = cv::Point3d(one, 0.0 , -tau);
+  vertices[6] = cv::Point3d(-one, 0.0 , -tau);
+  vertices[7] = cv::Point3d(-one, 0.0 , tau);
+  vertices[8] = cv::Point3d(0.0 , tau, one);
+  vertices[9] = cv::Point3d(0.0 , -tau, one);
+  vertices[10] = cv::Point3d(0.0 , -tau, -one);
+  vertices[11] = cv::Point3d(0.0 , tau, -one);
+
+  unsigned icosahedron_faces[] = {
+              4, 8, 7,
+              4, 7, 9,
+              5, 6, 11,
+              5, 10, 6,
+              0, 4, 3,
+              0, 3, 5,
+              2, 7, 1,
+              2, 1, 6,
+              8, 0, 11,
+              8, 11, 1,
+              9, 10, 3,
+              9, 2, 10,
+              8, 4, 0,
+              11, 0, 5,
+              4, 9, 3,
+              5, 3, 10,
+              7, 8, 1,
+              6, 1, 11,
+              7, 2, 9,
+              6, 10, 2};
+
+  for (unsigned i=0; i<20; i++)
+  {
+    startFaces.push_back(new SFace());
[…]

### 2011-09-01 tools/v4r/trunk/v4r/PGeometry/Pose.cc

@@ -1,0 +1,156 @@
+/**
+ * $Id$
+ * Johann Prankl, 2010-01-27 
+ * prankl@acin.tuwien.ac.at
+ */
+
+
+#include "Pose.hh"
+
+
+namespace P 
+{
+
+
+
+/***************************** Pose ******************************
+ * Constructor/Destructor
+ */
+Pose::Pose()
+{
+}
+
+Pose::Pose(bool init)
+{
+  if (init)
+  {
+    InitializePose(*this);
+  }
+}
+
+Pose::Pose(const cv::Mat& matR, const cv::Mat& matt)
+{
+  if (matR.type() != CV_64F) matR.convertTo(R, CV_64F);
+  else matR.copyTo(R);
+  if (matt.type() != CV_64F) matt.convertTo(t, CV_64F);
+  else matt.copyTo(t);
+}
+
+Pose::Pose(const Pose &pose)
+{
+  CopyPose((Pose&)pose, *this);
+}
+
+
+Pose::~Pose()
+{
+}
+
+/**************************** PRIVATE *************************************/
+
+
+
+/**************************** PUBLIC *************************************/
+/**
+ * copy operator
+ */
+Pose& Pose::operator=(const Pose &pose)
+{
+  CopyPose((Pose&)pose, *this); 
+  return *this;
+}
+
+
+/**
+ * Print a pose to a stream
+ */
+ostream& operator<<(ostream &os, const Pose &pose)
+{
+  //if (pose.empty()) 
+  //  return os<<'['<<']'<<'\n'<<'['<<']'<<'\n';
+
+  const double *R = pose.R.ptr<double>(0);
+  const double *t = pose.t.ptr<double>(0);
+
+  return os<<'['<< R[0]<<' '<<R[1]<<' '<<R[2]<<';'
+                << R[3]<<' '<<R[4]<<' '<<R[5]<<';'
+                << R[6]<<' '<<R[7]<<' '<<R[8]<< ']'
+           <<'\n'
+           <<'['<< t[0]<<' '<<t[1]<<' '<<t[2]<<']'
+           <<'\n';
+}
+
+/**
+ * Read a vector from a stream.
+ */
+istream& operator>>(istream &is, Pose &pose)
+{
+  if (pose.empty()) pose.create();
+
[…]

### 2011-09-01 tools/v4r/trunk/v4r/TomGine/tgMaterial.h

@@ -1,0 +1,41 @@
+ /**
+ * @file tgMaterial.h
+ * @author Thomas Mörwald
+ * @date January 2010
+ * @version 0.1
+ * @brief Specify material parameters for the OpenGL lighting model.
+ */
+
+#ifndef TG_MATERIAL_MODEL
+#define TG_MATERIAL_MODEL
+
+#include "tgMathlib.h"
+
+namespace TomGine{
+
+/** @brief Specify material parameters for the OpenGL lighting model. */
+class tgMaterial
+{
+public:
+	vec4 ambient;		///< ambient RGBA reflectance of the material
+	vec4 diffuse;		///< diffuse RGBA reflectance of the material
+	vec4 specular;		///< specular RGBA reflectance of the material
+	vec4 color;			///< RGBA color (used directly if lighting disabled)
+	float shininess;	///< RGBA specular exponent of the material
+
+	/** @brief Create grey material. */
+	tgMaterial();
+
+	/** @brief Enables lighting and applies material to OpenGL. */
+	void Activate() const;
+
+	/** @brief Create material with a specific color reflectance RGBA */
+	void Color(float r, float g, float b, float a=1.0f);
+
+	/** @brief Create material with random color reflectance RGBA. */
+	void Random();
+};
+
+} // namespace TomGine
+
+#endif

### 2011-09-01 tools/v4r/trunk/v4r/TomGine/tgNurbsVolume.cpp

@@ -1,0 +1,216 @@
+
+#include "tgNurbsVolume.h"
+#include <string>
+
+using namespace TomGine;
+
+tgNurbsVolume::tgNurbsVolume(){
+	std::string nurbs_vert = std::string(V4R_DIR) + "/v4r/TomGine/shader/nurbsvolume.vert";
+	std::string cox_head = std::string(V4R_DIR) + "/v4r/TomGine/shader/coxdeboor.c";
+	std::string color_frag = std::string(V4R_DIR) + "/v4r/TomGine/shader/color.frag";
+	shNurbsSurface = new tgShader(nurbs_vert.c_str(), color_frag.c_str(), cox_head.c_str());
+}
+
+tgNurbsVolume::tgNurbsVolume( const tgNurbsVolumePatch &data )
+{
+	// setup NURBS shader
+	std::string nurbs_vert = std::string(V4R_DIR) + "/v4r/TomGine/shader/nurbsvolume.vert";
+	std::string cox_head = std::string(V4R_DIR) + "/v4r/TomGine/shader/coxdeboor.c";
+	std::string color_frag = std::string(V4R_DIR) + "/v4r/TomGine/shader/color.frag";
+	shNurbsSurface = new tgShader(nurbs_vert.c_str(), color_frag.c_str(), cox_head.c_str());
+
+    Set(data);
+
+}
+
+tgNurbsVolume::~tgNurbsVolume(){
+	delete shNurbsSurface;
+}
+
+void tgNurbsVolume::Set( const tgNurbsVolumePatch &data )
+{
+	nurbsData = data;
+
+	texKnotsU.Load(&nurbsData.knotsU[0], nurbsData.knotsU.size(), GL_R32F, GL_RED, GL_FLOAT);
+    texKnotsV.Load(&nurbsData.knotsV[0], nurbsData.knotsV.size(), GL_R32F, GL_RED, GL_FLOAT);
+    texKnotsW.Load(&nurbsData.knotsW[0], nurbsData.knotsW.size(), GL_R32F, GL_RED, GL_FLOAT);
+    texCP.Load(&nurbsData.cps[0].x, nurbsData.ncpsU, nurbsData.ncpsV, nurbsData.ncpsW, GL_RGBA32F, GL_RGBA, GL_FLOAT);
+
+    // setup NURBS shader
[…]

### 2011-09-01 tools/v4r/trunk/v4r/TomGine/tgNurbsVolume.h

@@ -1,0 +1,125 @@
+/**
+* @file tgNurbsVolume
+* @author Thomas Moerwald
+* @date May 2011
+* @version 0.1
+* @brief NURBS volume representation and its GPU (GLSL) implementation "nurbsvolume.vert", "coxdeboor.c"
+*/
+
+#ifndef _TG_NURBS_VOLUME_H_
+#define _TG_NURBS_VOLUME_H_
+
+#include <vector>
+
+#include <v4r/TomGine/tgMathlib.h>
+#include <v4r/TomGine/tgTexture.h>
+#include <v4r/TomGine/tgShader.h>
+#include <v4r/TomGine/tgShapeCreator.h>
+#include <v4r/TomGine/tgRenderModel.h>
+
+namespace TomGine{
+
+/** @brief NURBS volume representation for one continuous patch. */
+struct tgNurbsVolumePatch{
+	std::vector<float> knotsU, knotsV, knotsW;	///< Knot vectors.
+    std::vector<vec4> cps;						///< Control point vector. A control point exists of a 3D vector and the weight as 4th entry.
+    unsigned ncpsU, ncpsV, ncpsW;				///< Number of control points
+    unsigned orderU, orderV, orderW;			///< Order of NURBS
+    unsigned resU, resV, resW;					///< Resolution of mesh grid.
+    bool sync;									///< Indicating memory synchronization with GPU.
+    /** @brief Create empty NURBS volume. */
+    tgNurbsVolumePatch(){
+    	ncpsU = 0; ncpsV = 0; ncpsW = 0;
+    	orderU = 2; orderV = 2; orderW = 0;
+    	resU = 16; resV = 16; resW = 16;
+    	sync = false;
+    }
+    /** @brief Print NURBS data to console. */
+    void Print(){
+    	printf("tgNurbsVolumePatch: \n");
+    	printf("orderU: %d, orderV: %d, orderW: %d, resU: %d, resV: %d, resW: %d\n",
+    			orderU, orderV, orderW, resU, resV, resW);
+    	printf("Knot vector U: ");
+    	for(unsigned i=0; i<knotsU.size(); i++)
+    		printf("%f ", knotsU[i]);
+    	printf("\n");
+    	printf("Knot vector V: ");
+    	for(unsigned i=0; i<knotsV.size(); i++)
+    		printf("%f ", knotsV[i]);
+    	printf("\n");
+    	printf("Knot vector W: ");
[…]

### 2011-09-01 tools/v4r/trunk/v4r/aldoma/alignment/src/align.cpp

@@ -1,0 +1,45 @@
+//#include "stable_planes.hpp"
+//#include <sampling.hpp>
+//#include <Eigen/StdVector>
+#include <alignment/stable_planes_alignment.h>
+#include <omp.h>
+
+int
+main (int argc, char ** argv)
+{
+  //ALIGN TWO 3D MESHES
+  vtkSmartPointer < vtkPLYReader > readerQuery = vtkSmartPointer<vtkPLYReader>::New ();
+  readerQuery->SetFileName (argv[1]);
+  vtkSmartPointer < vtkPolyData > polydata1 = readerQuery->GetOutput ();
+  polydata1->Update ();
+
+  vtkSmartPointer < vtkPLYReader > readerTarget = vtkSmartPointer<vtkPLYReader>::New ();
+  readerTarget->SetFileName (argv[2]);
+  vtkSmartPointer < vtkPolyData > polydata2 = readerTarget->GetOutput ();
+  polydata2->Update ();
+
+  bool parallel=false;
+  if(argc > 3) {
+    if (atoi(argv[3]) == 1) {
+      parallel = true;
+    }
+  }
+
+  bool gpu=false;
+    if(argc > 5) {
+      if (atoi(argv[5]) == 1) {
+        gpu = true;
+      }
+    }
+
+  std::string name_1 = std::string(argv[1]);
+  std::string name_2 = std::string(argv[2]);
+
+  StablePlanesAlignment spa(15000,atoi(argv[4]), parallel, gpu);
+  Eigen::Matrix4f transform;
+  spa.setName1(name_1);
+  spa.setName2(name_2);
+  spa.alignMeshes(polydata1, polydata2,transform);
+
+  return 0;
+}

### 2011-09-01 tools/v4r/trunk/v4r/aldoma/learning_object_physical_attributes/include/scale_distribution.h

@@ -1,0 +1,267 @@
+/*
+ * scale_distribution.h
+ *
+ *  Created on: Aug 9, 2011
+ *      Author: aitor
+ */
+
+#ifndef SCALE_DISTRIBUTION_H_
+#define SCALE_DISTRIBUTION_H_
+
+float read_scale_from_file(std::string file) {
+
+  std::cout << "read_scale_from_file:" << file << std::endl;
+
+  //read scale and push it into the vector
+  std::ifstream in;
+  in.open (file.c_str (), std::ifstream::in);
+
+  char linebuf[256];
+  in.getline (linebuf, 256);
+  std::string line (linebuf);
+  std::vector<std::string> strs;
+  boost::split (strs, line, boost::is_any_of (" "));
+
+  return atof (strs[0].c_str ());
+}
+
+void
+load_scale_samples (std::string directory, std::map<std::string, std::vector<float> > & category_scale_samples)
+{
+  bf::path dir = directory;
+  bf::directory_iterator end_itr;
+  for (bf::directory_iterator itr (dir); itr != end_itr; ++itr)
+  {
+    //check if its a directory, then get models in it
+    if (bf::is_directory (*itr))
+    {
+      //this is a category...
+      bf::path curr_path = itr->path ();
+      std::vector<float> scales_for_category;
+
+      bf::directory_iterator end_itr2;
+      for (bf::directory_iterator itr (curr_path); itr != end_itr2; ++itr)
+      {
+        bf::path curr_path2 = itr->path ();
+        std::string file = curr_path2.filename ();
+        std::vector<std::string> strs;
+        boost::split (strs, file, boost::is_any_of ("."));
+        if (strs[strs.size () - 1] == "txt" && strs[strs.size () - 2] == "scale")
+        {
+          //std::cout << file << std::endl;
+          std::stringstream path_to_scale_file;
+          path_to_scale_file << directory << "/" << curr_path.filename () << "/" << file;
+          float s = read_scale_from_file(path_to_scale_file.str());
+          scales_for_category.push_back (s);
+        }
+      }
+
+      category_scale_samples[curr_path.filename ()] = scales_for_category;
+      //std::cout << "Number of scales read:" << scales_for_category.size() << std::endl;
+    }
+    else
+    {
+      //ignore...
+    }
+  }
+
+}
+
+void
+meanShift (std::vector<float> & scale_samples, float bandwidth, float threshold,
+           std::vector<std::vector<int> > & cluster_indices)
+{
+
[…]

### 2011-09-01 tools/v4r/trunk/v4r/daisy/daisy.h

@@ -1,0 +1,416 @@
+////////////////////////////////////////////////////////////////////////////////
+// This program is free software; you can redistribute it                     //
+// and/or modify it under the terms of the GNU General Public License         //
+// version 2 (or higher) as published by the Free Software Foundation.        //
+//                                                                            //
+// This program is distributed in the hope that it will be useful, but        //
+// WITHOUT ANY WARRANTY; without even the implied warranty of                 //
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU          //
+// General Public License for more details.                                   //
+//                                                                            //
+// Written and (C) by                                                         //
+// Engin Tola                                                                 //
+//                                                                            //
+// web   : http://cvlab.epfl.ch/~tola/                                        //
+// email : engin.tola@epfl.ch                                                 //
+//                                                                            //
+// If you use this code for research purposes, please refer to the            //
+// webpage above                                                              //
+////////////////////////////////////////////////////////////////////////////////
+
+#ifndef DAISY_H
+#define DAISY_H
+
+#include <cmath>
+#include "assert.h"
+
+#include <v4r/daisy/general.h>
+#include <v4r/daisy/math.h>
+
+const float g_sigma_0 = 1;
+const float g_sigma_1 = sqrt(2);
+const float g_sigma_2 = 8;
+const float g_sigma_step = std::pow(2,1.0/2);
+const int   g_scale_st = int( (log(g_sigma_1/g_sigma_0)) / log(g_sigma_step) );
+static int  g_scale_en = 1;
+
+const float g_sigma_init = 1.6;
+static int g_grid_orientation_resolution = 360;
+
+static int g_cube_number = 3;
+static int g_grid_radius[10]; // m_rad_q_no < 10 --> must make this a member variable
+
+class daisy
+{
+public:
+   daisy();
+   ~daisy();
+
+   /// maximum radius of the descriptor region.
+   float m_rad;
+
+   /// the number of quantizations of the radius.
+   int m_rad_q_no;
+
[…]

### 2011-09-01 tools/v4r/trunk/v4r/geometry/stereovision.h

@@ -1,0 +1,61 @@
+/*
+    <one line to give the library's name and an idea of what it does.>
+    Copyright (C) 2010  Markus Bader
+
+    This library is free software; you can redistribute it and/or
+    modify it under the terms of the GNU Lesser General Public
+    License as published by the Free Software Foundation; either
+    version 2.1 of the License, or (at your option) any later version.
+
+    This library is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+    Lesser General Public License for more details.
+
+    You should have received a copy of the GNU Lesser General Public
+    License along with this library; if not, write to the Free Software
+    Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA  02110-1301  USA
+
+*/
+
+#ifndef STEREOVISION_H
+#define STEREOVISION_H
+
+
+#include <v4r/geometry/structs.h>
+
+namespace V4R {
+class CameraGeometry;
+
+class StereoVision
+{
+public:
+	
+    enum Type {
+	IMG_A = 0,
+	IMG_B = 1
+    };
+    StereoVision();
+    StereoVision(CameraGeometry &rGeoA, CameraGeometry &rGeoB);
+    ~StereoVision();
+    void set(CameraGeometry &rGeoA, CameraGeometry &rGeoB);
+    void updateEssential();
+		cv::Point3_<double> triangulatePoint(const cv::Point_<double> &rPointA, const cv::Point_<double> &rPointB);
+    cv::Mat triangulatePoints(const cv::Mat &pPointsA, const cv::Mat &pPointsB);
+
+    const cv::Mat_<double> &MEssA() const {
+        return mEssA;
+    };
+    const cv::Mat_<double> &MEssB() const {
+        return mEssB;
+    };
+private:
+    CameraGeometry *mpGeoA;
+    CameraGeometry *mpGeoB;
+    cv::Mat_<double> mEssA;
+    cv::Mat_<double> mEssB;
+
+};
+};
+
+#endif // STEREO_H

### 2012-01-17 scenarios/cdsr/trunk/papers/ACS2011/understanding-human-spatial-regions.tex

\documentclass[letterpaper]{article}

\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{graphicx}
\usepackage{stfloats}
\usepackage{color}

\usepackage{float}
\floatstyle{boxed} 
\restylefloat{figure}

\usepackage[noend,linesnumbered,algoruled]{algorithm2e}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
% \pdfinfo{
% /Title (Input Your Title Here)
% /Subject (Input The Proceedings Title Here)
% /Author (First Name, Last Name;
% First Name, Last Name;
% First Name, Last Name;)
% }
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\from}[2]{\textcolor{red}{\noindent\textbf{//}\textbf{Note
      from #1:}\textsc{ #2}\textbf{//}}}

      \newcommand{\fw}[1]{\texttt{#1}}

\begin{document}
% \title{Transferring Human Spatial Regions from Sensor Data via Analogy}
\title{Representing and Reasoning About Spatial Regions Defined by Context}
% \title{Towards a Cognitive System Than Can Represent and Reason about Spatial Regions Defined by Context}
% \title{Context-Dependent Spatial Regions}
% \title{Qualitative and Analogical Cognition for Spatial Regions Defined by Context}

\author{Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com \And Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk \And Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  }

 \maketitle
 \begin{abstract}
In order to collaborate with people in the real world, cognitive systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and propose a method for a cognitive system to learn them incrementally by combining qualitative spatial representations, semantic labels, and analogy. Using data from a mobile robot, we generate a relational representation of semantically labeled objects and their configuration. Next, we show how the boundary of a context-dependent spatial region can be defined using \textit{anchor points}. Finally, we demonstrate how an existing computational model of analogy can be used to transfer this region to a new situation. 
 \end{abstract}

 \section{Introduction}

Consider a janitorial robot cleaning a classroom. While performing this task, it encounters a teacher working with a student. The teacher tells the robot to ``start at the front of the classroom'', expecting it to go to the front of the classroom and begin cleaning that area. This response requires that the robot is able to \emph{determine the spatial region in the environment that satisfies this concept}.

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not easily perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a chalkboard), their role in this context (seats for students to watch a teacher who writes on the chalkboard) and their configuration in space (the seats point toward the chalkboard). In this paper, we will refer to such regions as \textit{context-dependent spatial regions}.

% Pat:
% Finally, let me urge you to include specific claims in your paper and
% to provide evidence that supports them. Such evidence can take many
% forms, from experimental studies to proofs to compelling arguments,
% but making this explicit will improve your chances.

Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:
\begin{itemize}

\item{If cognitive systems are to collaborate with humans in everyday environments, then they must be able to represent and reason about the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.}

\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. The identification of context-dependent spatial regions provides a suitable task for studying this integration. Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.}

\item{Cognitive systems must adapt their knowledge and experience to new situations. While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf. one-shot learning~\cite{Fei-Fei/etal:2006}). This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system. For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g. theaters) without additional tutoring.}
\end{itemize}

\begin{figure*}[t!]
  \includegraphics[width=\textwidth]{example1.jpg}
  \caption{
This image demonstrates the behavior we desire of our cognitive system. The robot is able to identify the front of the classroom by analogy to a previously labelled example. To do this it must identify the correspondences between the example and its current environment (the desks in the front row and chalkboards) and then transfer the spatial region (the front of the room is between these entities).}
  \label{fig:example}
\end{figure*}

Figure \ref{fig:example} illustrates our approach for understanding context-dependent spatial regions via analogy. Our first assumption is that these regions can be defined using qualitative spatial representations~\cite{Cohn:2001}. Our second assumption is that geometrically and semantically similar areas will feature similar context-dependent spatial regions, and that these can be transferred through analogy.

In the context of the symposium on ``Advances in Cognitive Systems'' we are making two specific claims:

\begin{enumerate}
\item Understanding context-dependent spatial regions is an essential ability for a cognitive system that must interact with humans.
\item This ability can be realized in an artificial cognitive system through the combination of a sensor-based spatial model, qualitative spatial representations, and analogy.  
\end{enumerate}

\noindent We appeal to the reader's experience and intuition to evidence the first claim. To support this, we have already provided a number of examples of situations in which humans rely on context-dependent spatial regions, but we have not done any systematic collection or analysis of real-world situations in which they occur. To evidence the second claim, we present the results of developing an extension to an existing cognitive system that enables it to learn context-dependent spatial regions by example.

The rest of the paper is structured as follows. First, we describe how we generate qualitative spatial representations from the spatial model of an existing system. Second, we describe how we define context-dependent spatial regions using \textit{anchor points} \cite{Klenk/etal2005}, i.e., symbolic descriptions that tie conceptual entities to geometric points. Next, we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer knowledge from an existing context-dependent spatial region to a new situation. Finally, we discuss some limitations of our current implementation, and position it with respect to previous work and cognitive systems in general.

\section{From Sensors to Semantics}

Our approach to representing and reasoning about context-dependent spatial regions requires the existence of qualitative spatial representations~\cite{Cohn:2001} describing the relationships between the entities a cognitive system perceives (including objects, groups, and regions). By adding semantic labels, or types, to each entity, we are able to reason about the geometric \emph{and} conceptual characteristics of the system's environment. The following sections describe how we generate appropriate qualitative spatial representations using the spatial model of an existing, state-of-the-art, cognitive system.

\subsection{The Dora System}

\begin{figure}[t!]
  \includegraphics[width=\columnwidth]{images/classroom-peekabot-annotated.pdf}
  \caption{A visualization of Dora's existing spatial model taken from the running system. Walls are visualized where flat surfaces (walls, desks and other objects) are detected during SLAM.}
  \label{fig:dora-spatial}
\end{figure}

We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial representation~\cite{Hawes/etal:2011}. In this paper, we focus on the basic elements of this spatial model. For more information on Dora's other competences, see recent papers, e.g.~\cite{Hawes/etal:2011,Hanheide/etal:2011}.

In the simplest terms, Dora's spatial model can be treated as two layers: a \emph{metric layer} and a \emph{place layer}. A visualization of these from the running system is shown in Figure~\ref{fig:dora-spatial}.
% Metric layer
The metric layer is a collection of lines in a 2D global coordinate frame. These lines are generated by a  process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM). Lines in this SLAM map represent features extracted from laser scans wherever a straight line is present for long enough to be considered permanent. In practice, lines are generated at the positions of walls and any other objects that are flat at the height of the laser (e.g. foot stools, bins, closed doors etc.). The robot's location in the metric layer is represented as a 2D position plus an orientation.
% Place layer
As the robot moves around in the world, the free space it passes through is discretized into a graph of connected nodes $1m$ apart. We refer to these nodes as \emph{places} and the graph of connected places forms the spatial model's place layer. The robot's location in the place layer is represented by a pointer to a place. Places allow Dora to reason about space in basic qualitative ways, e.g. by clustering places into rooms based on the presence of doorways, and provide a representation of space capable of facilitating behavior planning.
% Objects
Dora is capable of using vision to recognize predefined objects. Recognition can either be triggered through autonomous visual search or a a user's command. When an object is detected it is represented in the place layer relative to Dora's current position. The recognizer associates each object with a label that was provided during a training phase. In our current work, these labels are used directly in analogical processing. We use a convex hull around the places in a room to represent its area, and the 2D locations of perceived objects to create the qualitative spatial representations in the next section.

% simulation
As the purpose of this paper is to explore the possibility of using qualitative spatial reasoning and analogy on top of Dora's spatial model, we decided to perform our research using the Player/Stage hardware abstraction layer and simulation tools~\cite{GerkeyVaughanHoward03} rather than on a real robot. This allows us to run the Dora system \emph{mostly unchanged} in a controlled simulation environment, rather than deal with some of the unrelated challenges of running the robot in the real world. The key difference is that the sensor readings are free of noise in the simulated world. This is particularly significant where vision is concerned, as Dora's object recognizer is replaced entirely by a simulated blob detector which is always correct. Beyond this, the functionality of the system remains unchanged, and, crucially, the representations our work builds on will be the same when the system is run in the real world.

\subsection{Qualitative Spatial Representation Extraction}

Given data from Dora's spatial model, we compute the following qualitative spatial representations. First, we extract the object and room entities from the sensor data. Next, we determine the qualitative spatial relationships between these entities. For the room, we compute its region as a convex hull around the set of points. For the objects, we consider only their centroids. For each pair of adjacent objects, we compute topological and positional relationships. Adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}. For the topological relationships, we use the region connection calculus, RCC8, \cite{Cohn:2001}. For positional relations, we use two predicates \fw{leftOf} and \fw{below}. An entity is \fw{leftOf} another if the x coordinate of its centroid is less than that of the other entity's. An entity is \fw{below} another if the y coordinate of its centroid is less than that of the other entity's. Objects whose x or y coordinates are within a threshold of one another will not be related by \fw{leftOf} or \fw{below} positional relations.

%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
%Desk 1 and Desk 2 are not labeled in the Figure.

Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}. Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint. To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}. The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate. Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}. These qualitative spatial relationships provide the structure necessary for analogical processing.

\subsection{Representing Context-Dependent Spatial Regions}

To enable analogical transfer, it is necessary to represent context-dependent spatial regions symbolically. We do this using \textit{anchor points} \cite{Klenk/etal2005} which are symbolic descriptions linking a conceptual entity to a particular point on a perceived entity.  A context-dependent spatial region is a conceptual entity, and the objects in the room and the room itself are perceived entities. As stated previously, the room is represented by a set of points. Anchor points are created using unary functions over a percieved entity. For example, \fw{(CentroidFn Area1)} represents the centroid of geometric extent of \fw{Area1}. We use anchor points to specify the boundary of the context-dependent spatial region using the \fw{regionBoundary} relation. For a particular classroom, we are able to define the front of the room as the region bounded by the leftmost top point of the room, the whiteboard, the leftmost bottom point of the room, and each of the desks in the front row. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each context-dependent spatial region has one type and a variable number of boundary points.
%\from{Nick}{This is crying out for a diagram, perhaps on the stage map of classroom 1.}
%\from{Nick}{You mention chalkboard above, but it is not anywhere else in the below example.}
%\from{Klenk}{We will do for Figure 5.}

\begin{figure}[h]
	{\fontsize{8}{8} %%Because aaai.sty overrides usual font size commands

\fw{(regionType CDSR1 FrontRegion)\\
(regionBoundary CDSR1 (LeftmostTopFn Room1))\\
(regionBoundary CDSR1 (LeftmostBottomFn Room1))\\
(regionBoundary CDSR1 Desk1)\\
(regionBoundary CDSR1 Desk2)\\
(regionBoundary CDSR1 Desk3)\\
(regionBoundary CDSR1 Desk4)\\
(regionBoundary CDSR1 Whiteboard0)\\
}}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR1}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top left of Figure~\ref{fig:dora-maps}. The anchor points (connected in green in Figure~\ref{fig:dora-maps}) define the region. \fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively. The sets of points representing the top and bottom of the classroom are found by selecting the points defining the area within a threshold of the area's maximum or minimum y coordinate. The other corners of the context-dependent spatial region are defined by the four desks and the whiteboard. Because perceived objects (e.g. the desks) are defined as single points, they can serve as anchor points directly. The semantic label \fw{FrontRegion} ties the geometric extent of the boundary points to a conceptual region. This definition for the front of the classroom is specific to \fw{Room1} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were closer to the whiteboard then the region would be smaller).

% In the next section, we describe how to transfer these environment-specific representations to new situations via analogy.

\section{Analogical Transfer of Spatial Regions}

% While this work focuses on how analogy enables inferences from specific examples to new situations, but its role in the larger cognitive system this work envisions cannot be ignored.

We assume that a cognitive system will have a way of initially acquiring examples of context-dependent spatial regions, e.g., by being taught by a human through dialogue, sketching, or hand-coding. As stated previously, to avoid burdening potential tutors with the task of teaching the system every context-dependent spatial region individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after its initial training. For example, after training a janitorial robot in just one classroom, it should be able to identify similar regions in the other classrooms in the building. Motivated by the learning and generalization abilities of humans, we have chosen analogy to tackle this problem.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.

The Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} performs analogical matching. Given two structured representations as input (the base and target), SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target.  Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{analogy2.png}
  \caption{Analogical mapping between four base expressions and two target expressions. Ovals and squares represent relations and entities respectively. Purple dashed bi-directional arrows are the correspondences between the base and target, and the green dashed arrow, oval, and square represent a candidate inference projected into the target.}
  \label{fig:analogy}
\end{figure}

Figure \ref{fig:analogy} illustrates a sample mapping between four base expressions and two target ones. Each oval represents a binary predicate, and its two entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(rcc8-DC Desk1 Desk2)} and \fw{(below Desk1 Desk2)}, and the  target expressions \fw{(rcc8-DC Desk20 Desk22)} and \fw{(below Desk20 Desk22)}. This is done in the following manner. First, the predicates of these expressions are placed in correspondence by the identicality constraint. Next, parallel connectivity aligns their arguments, \fw{Desk1} with \fw{Desk20} and \fw{Desk2} with \fw{Desk22}. While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint. In Figure \ref{fig:analogy}, these correspondences are highlighted by the purple hashed bi-directional arrows. Next, SME creates a candidate inference for the expression \fw{(regionBoundary CSDR1 Desk1)}, because \fw{Desk1} participates in the mapping. The candidate inference is stated as \fw{(regionBoundary (:skolem CSDR1) Desk20)}. \fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences. In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes \fw{Desk20}. Note that inference is selective, with no candidate inferences generated for the entirely unmapped \fw{below} expression. Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

% It feels odd having this as a separate subsection
%\subsection{Initial Approach}\label{sec:approach}

In our system, the base and target representations consist of the identified objects from Dora's spatial model drawn from two different runs in two different environment, plus the qualitative spatial relationships computed on top of these. The base also contains a labeled context-dependent spatial region that is sought in target environment. We use SME to create an analogical mapping between the base and target. The result of this mapping is a set of correspondences and a set of candidate inferences. We transfer context-dependent spatial regions as follows. First, we identify the context-dependent spatial region entity of the sought semantic type in the base. This entity should not participate in the mapping, but should be included as \fw{:skolem} expressions in the candidate inferences. To transfer this entity to the target, we collect the candidate inferences resulting from the \fw{regionBoundary} statements in the base that mention the base entity. The second arguments of these candidate inferences are anchor points in the target environment. Therefore, to identify the context-dependent spatial region in the target, we create a new entity, assign it the semantic type of the sought region, and define its boundary using the anchor points from the candidate inferences.

\section{Example}

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{images/classroom-maps-axes-annotation.png}
  \caption{The top two images show the Player/Stage classroom simulations used in the example, with the base on the left and the target on the right. The desks are simulated as red boxes, and the chalkboard and lectern (where present) as blue and green boxes respectively. Furthermore, the solid green line in the base connects the boundary points of the front of the room. The lower two images show the metric and place maps generated by Dora during the example runs.}
  \label{fig:dora-maps}
\end{figure}

Using the previously described techniques, we have implemented a system capable of performing an inference similar to the one presented in Figure \ref{fig:example}. The base and target cases are created by running the Dora system on the two different simulated classroom worlds shown in Figure~\ref{fig:dora-maps}. In each case, Dora is manually driven around the room to allow it to create metric and place maps. Once the map is created, Dora is then driven in front of each object and the visual recognition system is run, resulting in the objects being added to the place layer.

The data from Dora's spatial model is then passed on to the qualitative spatial representation generator. In the base case, there are 16 desks and a chalkboard. There are two other entities in the case: the room area and the context-specific spatial region representing the front of the room. The case includes a total of 96 expressions relating the 19 entities. Nine of these expressions are used to define the boundary of a context-specific spatial region representing the front of the room: the four desks in the front row, the chalkboard, and the leftmost bottom and leftmost top of the room. The target case includes only 12 desks arranged in three columns of four, a chalkboard and a lectern. The target case includes 73 statements and 14 entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation. The resulting analogy includes 62 correspondences between the entities and expressions in the base and target and 53 candidate inferences.  All seven \fw{regionBoundary} statements in the base appear as candidate inferences. Five of the anchor points are exactly correct, the leftmost corners of the room, the chalkboard and two of the desks in the front row. The other two analogically transferred anchor points are to the lectern and a desk in the third row. Figure \ref{fig:success} illustrates the region created by the seven boundary points in the target environment. While this region is not identical to how a human would specify it, our system is able to produce a fairly accurate context-dependent spatial region in a new environment from a single example.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/target-success.png}
  \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region. The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
  \label{fig:success}
\end{figure}

\section{Limitations of Initial Approach}

The spatial transfer algorithm described in here has a number of limitations to be addressed in future work.  First, we have not incorporated a memory retrieval mechanism, such as MAC/FAC \cite{forbus/etal1995}.  As a cognitive system gains experience in different environments, it will have access to a range of context-dependent spatial regions defined in a variety of task situations. To retrieve an analogous example, the agent must select the most similar environment that contains a labeled example of the sought after region type. This extension is fairly straightforward, but beyond the scope of this paper. Second, our initial approach ignores a number of potentially relevant pieces of information. After the context-dependent spatial region has been identified in the target environment, its qualitative spatial relationships could be compared with those from the base to either refine the region or provide a confidence measure for the inference. In our example, the transferred region contains the bottom desk in the second row, but the corresponding region does not contain any desks topologically. Third, our algorithm uses an absolute reference frame. Therefore, the environments must be oriented in a similar fashion to successfully transfer a context-dependent spatial region. We can overcome this limitation by performing a series of analogies between the base and the target in which we rotate the base and select the analogy with the best mapping. Finally, more advanced qualitative spatial relationships should improve performance. For example, recognizing all the desks as one region, or grouping them into rows or columns, would provide a better level of abstraction for this task.

\section{Related Work}

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g. door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}. While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions. While steps have been taken toward incorporating this knowledge  (i.e. the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work. For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g. the areas near doors \cite{Zender2008a}).

% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
 % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.

There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.  The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics. 
% Therefore, we hope and expect to see more research combining these three fields.

\section{Discussion}

% Nick: All this reptetition and probably unnecessary
% Klenk: I like the repetition. I think it is really important to drive home the cognitive systems story here.

Our first claim for this work is that in order for artificial cognitive systems to collaborate with human users, they must be able to understand and reason about the types of spatial regions that humans use. In particular we have identified context-dependent spatial regions, i.e., regions defined by their geometric, functional, and semantic relations to surrounding entities, as an important type of region that cannot be represented by current systems. Addressing this problem is important as such regions are ubiquitous in human environments and communication, ranging from regions in everyday experience (e.g. aisles, marketplaces) to domain-specific regions (e.g. the ``pocket'' in American Football).

In addition to addressing a specific need in human-robot interaction, developing a cognitive system able to represent and reason about context-dependent spatial regions is important for two further reasons. First, as these regions are not defined solely by their geometry, we must investigate approaches that integrate different types of knowledge. This is a key problem in cognitive systems work in general. Second, in order to reduce the burden on developers and users, a cognitive system that understands one type of context-dependent spatial region should be able to use this understanding to recognize similar regions in other areas. Developing such a generalization ability (particularly one that can work incrementally from a small number of examples) is another general challenge for the cognitive systems community.

Our second claim in this work is that the combination of a sensor-based spatial model, qualitative spatial representations, and analogy provides a potential solution to the aforementioned problems. We presented results from an initial proof-of-concept implementation that supports this claim. This system is able to transfer, via analogy, context-dependent spatial regions defined using anchor points over logical representations consisting of  object types and qualitative spatial relationships inferred from sensor data. Whilst the system has a number of limitations, and the transferred context-dependent spatial region presented does not perfectly match human expectations, it validates further development of our approach.

To support the definition of new spatial regions which the robot should be aware of, it is desirable to integrate a sketching interface on top of a visualization of the spatial knowledge the robot already has. Sketching \cite{Forbus/etal:2009a} is often used by humans to convey new ideas and concepts, making this a natural interface for human-robot interaction. When faced with an unknown context-dependent spatial region, the mobile robot could then ask a human to define the region using \textit{sketch annotations} over its representation \cite{Klenk/etal2005}.

For mobile robots, spatial understanding is required to enable the pursuit of spatially-defined goals (e.g the goal that the front of the classroom is clean). Consequently, spatial understanding should be evaluated with respect to a cognitive system's actions in its environment. The minimal performance for spatial understanding is that the cognitive system is able to navigate to context-dependent spatial regions. A longer term research project involves reasoning about potentially conflicting goals with overlapping spatial extents. A cognitive system should be able to combine them to generate new goals to pursue. The scenario in Figure \ref{fig:example} exemplifies this process of combining goals through spatial extent, when the robot generates the goal of cleaning the front of the classroom. With the proliferation of mobile robots, sensors and computation, new ideas, such as, context-dependent spatial regions, have the potential to reshape how users interact with cognitive systems.

\section{Acknowledgments}

The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams. The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.

\bibliography{understanding-human-spatial-regions} \bibliographystyle{aaai} \end{document}

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2011/evaluation.tex

\documentclass[letterpaper]{article}

\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{graphicx}
\usepackage{stfloats}
\usepackage{color}

\begin{document}

 \section{Evaluation}

To evaluate our progress toward building a cognitive system capabile of reasoning about these regions, we conducted the an experiment focusing on the following questions:
\begin{itemize}
\item{How well do anchor points capture context-dependent spatial regions?}
\item{When provided with an example definition of a CDSR, how well does our appraoch identify the CDSR in the new situation?}
\end{itemize}
 Our materials are six classrooms (two simulated and four real) and two simulated studio apartments. To determine how people consider CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons representing a list of region types for seach room, users were also shown a polygon drawn by another and asked to determine if it was an acceptable depiction of the region. All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.

We consider a \textit{problem instance} to be a room and a sought CDSR type (e.g., "front"). For each room containing a CDSR of the sought type, we generate an \textit{inferred region} our analogical approach. To assess the quality of the transfer, we use the following measures:

\begin{equation}
	precision=\frac{area(infered region \cap target region)}{area(infered region)}
\end{equation}
\begin{equation}
	recall=\frac{area(infered region \cap target region)}{area(target region)}
\end{equation}

As a baseline, we will compare our results against (1) the manually encoded anchors for the problem instance and (2) the entire room.

\end{document}

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2011/evaluation.tex

@@
 \item{How well do anchor points capture context-dependent spatial regions?}
-\item{When provided with an example definition of a CDSR, how well does our appraoch identify the CDSR in the new situation?}
+\item{When provided with an example definition of a CDSR, how well does our approach identify the CDSR in the new situation?}
+\item{Is the integration of geometric and semantic knowledge necessary?}
+\item{Does similarity correlate with performance?}
 \end{itemize}
- Our materials are six classrooms (two simulated and four real) and two simulated studio apartments.
+
+I'm not sure how many of the above questions we want to talk about.
+The last one is fairly easy and won't take up much space in the paper.
+The integration of knowledge question is could take some more work as the grouping is determined by semantic type, so in testing just positional knowledge,  I would have to create an alternate set of anchor points.
+
+\subsection{Materials}
+We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments.
 To determine how people consider CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons representing a list of region types for seach room, users were also shown a polygon drawn by another and asked to determine if it was an acceptable depiction of the region.
@@
 \begin{equation}
-	precision=\frac{area(infered region \cap target region)}{area(infered region)}
+	p=\frac{area(inferred region \cap target region)}{area(inferred region)}
 \end{equation}
 \begin{equation}
-	recall=\frac{area(infered region \cap target region)}{area(target region)}
+	r=\frac{area(inferred region \cap target region)}{area(target region)}
 \end{equation}
 
-As a baseline, we will compare our results against (1) the manually encoded anchors for the problem instance and (2) the entire room.
+As a baseline, we will compare our results against the manually encoded CDSRS for the problem instance and the entire room.
+The precision and recall for considering the entire room provides a baseline from which to assess each of the questions.
+The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
+
+\subsection{Results}
+\begin{table}
+\caption{Performance by Room}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Room & Transfer & Manually Encoded & Entire Room \\
+\hline
+Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Studio-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+\hline
+\end{tabular}
+\end{table}
+
+If we are just going to report means, we should standard deviations as well.
+
+\begin{table}
+\caption{Performance by Region Type}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Region & Transfer & Manually Encoded & Entire Room \\
+\hline
+Front & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Back & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Front Rows & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Back Rows & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Kitchen & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Office & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+\hline
+\end{tabular}
+\end{table}
+
+Or we could report all of the results in a single entry.
+
+\begin{table}
+\caption{Performance against Target Region}
+\begin{tabular}{|c|c|c|}
+\hline
+Transfer & Manually Encoded & Entire Room \\
+\hline
+p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+\hline
+\end{tabular}
+\end{table}
+
+Overall the transferred region was significantly better than the entire room (p<Blah), and not statistically different than the manually encoded region (p=Blah).
+There was a positive correlation (r=Blah) between the structural evaluation score of the analogy and f-measure of the inference.
+
+\subsection{Discussion}
+Our results
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
JK: added qsr-extraction.tex and modified main-*.tex to include it

--- 2012-01-17 scenarios/cdsr/trunk/papers/ACS2011/understanding-human-spatial-regions.tex
@@
 
-\subsection{Qualitative Spatial Representation Extraction}
+\input{qsr-extraction}
+%\subsection{Qualitative Spatial Representation Extraction}
 
-Given data from Dora's spatial model, we compute the following qualitative spatial representations.
+%Given data from Dora's spatial model, we compute the following qualitative spatial representations.
 First, we extract the object and room entities from the sensor data.
@@
 
-%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
-%Desk 1 and Desk 2 are not labeled in the Figure.
+%%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
+%%Desk 1 and Desk 2 are not labeled in the Figure.
 
-Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}.
+%Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}.
 Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint.
@@
 
-\bibliography{understanding-human-spatial-regions} \bibliographystyle{aaai} \end{document}
+\bibliography{aaai12-cdsr}
+\bibliographystyle{aaai}
+\end{document}

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
JK: added qsr-extraction.tex and modified main-*.tex to include it

--- 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
@@
-\documentclass[letterpaper]{article}
+\subsection{Qualitative Spatial Representation Extraction}
 
-\usepackage{aaai}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{graphicx}
-\usepackage{stfloats}
-\usepackage{color}
-
-\usepackage{float}
-\floatstyle{boxed} 
-\restylefloat{figure}
-
-\usepackage[noend,linesnumbered,algoruled]{algorithm2e}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PDFMARK for TeX and GhostScript
-% Uncomment and complete the following for metadata if
-% your paper is typeset using TeX and GhostScript (e.g
-% if you use .ps or .eps files in your paper):
-% \special{! /pdfmark where
-% {pop} {userdict /pdfmark /cleartomark load put} ifelse
-% [ /Author (John Doe, Jane Doe)
-% /Title (Paper Title)
-% /Keywords (AAAI, artificial intelligence)
-% /DOCINFO pdfmark}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PDFINFO for PDFTeX
-% Uncomment and complete the following for metadata if
-% your paper is typeset using PDFTeX
-% \pdfinfo{
-% /Title (Input Your Title Here)
-% /Subject (Input The Proceedings Title Here)
-% /Author (First Name, Last Name;
-% First Name, Last Name;
-% First Name, Last Name;)
-% }
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Uncomment only if you need to use section numbers
-% and change the 0 to a 1 or 2
-% \setcounter{secnumdepth}{0}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\from}[2]{\textcolor{red}{\noindent\textbf{//}\textbf{Note
-      from #1:}\textsc{ #2}\textbf{//}}}
-
-      \newcommand{\fw}[1]{\texttt{#1}}
-
-\begin{document}
-% \title{Transferring Human Spatial Regions from Sensor Data via Analogy}
-\title{Representing and Reasoning About Spatial Regions Defined by Context}
-% \title{Towards a Cognitive System Than Can Represent and Reason about Spatial Regions Defined by Context}
-% \title{Context-Dependent Spatial Regions}
-% \title{Qualitative and Analogical Cognition for Spatial Regions Defined by Context}
-
-\author{Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com \And Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk \And Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  }
-
- \maketitle
- \begin{abstract}
-In order to collaborate with people in the real world, cognitive systems must be able to represent and reason about spatial regions in human environments.
-Consider the command \emph{``go to the front of the classroom''}.
-The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone.
-Instead it is defined by its functional use, implied by nearby objects and their configuration.
-In this paper, we define such areas as \textit{context-dependent spatial regions} and propose a method for a cognitive system to learn them incrementally by combining qualitative spatial representations, semantic labels, and analogy.
-Using data from a mobile robot, we generate a relational representation of semantically labeled objects and their configuration.
-Next, we show how the boundary of a context-dependent spatial region can be defined using \textit{anchor points}.
-Finally, we demonstrate how an existing computational model of analogy can be used to transfer this region to a new situation.
- \end{abstract}
-
- \section{Introduction}
-
-Consider a janitorial robot cleaning a classroom.
-While performing this task, it encounters a teacher working with a student.
-The teacher tells the robot to ``start at the front of the classroom'', expecting it to go to the front of the classroom and begin cleaning that area.
-This response requires that the robot is able to \emph{determine the spatial region in the environment that satisfies this concept}.
-
-The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments.
-Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g.
-walls and doors).
-However, many regions to which humans routinely refer are not so easily defined.
-Consider, for example, the aforementioned region \textit{the front of the classroom}.
-This region is not easily perceivable using just the geometry of the environment.
-Instead, it is defined by the objects present in the room (chairs, a desk, a chalkboard), their role in this context (seats for students to watch a teacher who writes on the chalkboard) and their configuration in space (the seats point toward the chalkboard).
-In this paper, we will refer to such regions as \textit{context-dependent spatial regions}.
-
-% Pat:
-% Finally, let me urge you to include specific claims in your paper and
-% to provide evidence that supports them. Such evidence can take many
-% forms, from experimental studies to proofs to compelling arguments,
-% but making this explicit will improve your chances.
-
-Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:
-\begin{itemize}
-
-\item{If cognitive systems are to collaborate with humans in everyday environments, then they must be able to represent and reason about the spatial regions humans refer to.
-Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.}
-
-\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-The identification of context-dependent spatial regions provides a suitable task for studying this integration.
-Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.}
-
-\item{Cognitive systems must adapt their knowledge and experience to new situations.
-While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf.
-one-shot learning~\cite{Fei-Fei/etal:2006}).
-This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system.
-For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g.
-theaters) without additional tutoring.}
-\end{itemize}
-
-\begin{figure*}[t!]
-  \includegraphics[width=\textwidth]{example1.jpg}
-  \caption{
-This image demonstrates the behavior we desire of our cognitive system.
-The robot is able to identify the front of the classroom by analogy to a previously labelled example.
-To do this it must identify the correspondences between the example and its current environment (the desks in the front row and chalkboards) and then transfer the spatial region (the front of the room is between these entities).}
-  \label{fig:example}
-\end{figure*}
-
-Figure \ref{fig:example} illustrates our approach for understanding context-dependent spatial regions via analogy.
-Our first assumption is that these regions can be defined using qualitative spatial representations~\cite{Cohn:2001}.
-Our second assumption is that geometrically and semantically similar areas will feature similar context-dependent spatial regions, and that these can be transferred through analogy.
-
-In the context of the symposium on ``Advances in Cognitive Systems'' we are making two specific claims:
-
-\begin{enumerate}
-\item Understanding context-dependent spatial regions is an essential ability for a cognitive system that must interact with humans.
-\item This ability can be realized in an artificial cognitive system through the combination of a sensor-based spatial model, qualitative spatial representations, and analogy.  
-\end{enumerate}
-
-\noindent We appeal to the reader's experience and intuition to evidence the first claim.
-To support this, we have already provided a number of examples of situations in which humans rely on context-dependent spatial regions, but we have not done any systematic collection or analysis of real-world situations in which they occur.
-To evidence the second claim, we present the results of developing an extension to an existing cognitive system that enables it to learn context-dependent spatial regions by example.
-
-The rest of the paper is structured as follows.
-First, we describe how we generate qualitative spatial representations from the spatial model of an existing system.
-Second, we describe how we define context-dependent spatial regions using \textit{anchor points} \cite{Klenk/etal2005}, i.e., symbolic descriptions that tie conceptual entities to geometric points.
-Next, we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer knowledge from an existing context-dependent spatial region to a new situation.
-Finally, we discuss some limitations of our current implementation, and position it with respect to previous work and cognitive systems in general.
-
-\section{From Sensors to Semantics}
-
-Our approach to representing and reasoning about context-dependent spatial regions requires the existence of qualitative spatial representations~\cite{Cohn:2001} describing the relationships between the entities a cognitive system perceives (including objects, groups, and regions).
-By adding semantic labels, or types, to each entity, we are able to reason about the geometric \emph{and} conceptual characteristics of the system's environment.
-The following sections describe how we generate appropriate qualitative spatial representations using the spatial model of an existing, state-of-the-art, cognitive system.
-
-\subsection{The Dora System}
-
-\begin{figure}[t!]
-  \includegraphics[width=\columnwidth]{images/classroom-peekabot-annotated.pdf}
-  \caption{A visualization of Dora's existing spatial model taken from the running system. Walls are visualized where flat surfaces (walls, desks and other objects) are detected during SLAM.}
-  \label{fig:dora-spatial}
-\end{figure}
-
-We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial representation~\cite{Hawes/etal:2011}.
-In this paper, we focus on the basic elements of this spatial model.
-For more information on Dora's other competences, see recent papers, e.g.~\cite{Hawes/etal:2011,Hanheide/etal:2011}.
-
-In the simplest terms, Dora's spatial model can be treated as two layers: a \emph{metric layer} and a \emph{place layer}.
-A visualization of these from the running system is shown in Figure~\ref{fig:dora-spatial}.
-% Metric layer
-The metric layer is a collection of lines in a 2D global coordinate frame.
-These lines are generated by a  process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
-Lines in this SLAM map represent features extracted from laser scans wherever a straight line is present for long enough to be considered permanent.
-In practice, lines are generated at the positions of walls and any other objects that are flat at the height of the laser (e.g.
-foot stools, bins, closed doors etc.).
-The robot's location in the metric layer is represented as a 2D position plus an orientation.
-% Place layer
-As the robot moves around in the world, the free space it passes through is discretized into a graph of connected nodes $1m$ apart.
-We refer to these nodes as \emph{places} and the graph of connected places forms the spatial model's place layer.
-The robot's location in the place layer is represented by a pointer to a place.
-Places allow Dora to reason about space in basic qualitative ways, e.g.
-by clustering places into rooms based on the presence of doorways, and provide a representation of space capable of facilitating behavior planning.
-% Objects
-Dora is capable of using vision to recognize predefined objects.
-Recognition can either be triggered through autonomous visual search or a a user's command.
-When an object is detected it is represented in the place layer relative to Dora's current position.
-The recognizer associates each object with a label that was provided during a training phase.
-In our current work, these labels are used directly in analogical processing.
-We use a convex hull around the places in a room to represent its area, and the 2D locations of perceived objects to create the qualitative spatial representations in the next section.
-
-% simulation
-As the purpose of this paper is to explore the possibility of using qualitative spatial reasoning and analogy on top of Dora's spatial model, we decided to perform our research using the Player/Stage hardware abstraction layer and simulation tools~\cite{GerkeyVaughanHoward03} rather than on a real robot.
-This allows us to run the Dora system \emph{mostly unchanged} in a controlled simulation environment, rather than deal with some of the unrelated challenges of running the robot in the real world.
-The key difference is that the sensor readings are free of noise in the simulated world.
-This is particularly significant where vision is concerned, as Dora's object recognizer is replaced entirely by a simulated blob detector which is always correct.
-Beyond this, the functionality of the system remains unchanged, and, crucially, the representations our work builds on will be the same when the system is run in the real world.
-
-\input{qsr-extraction}
-%\subsection{Qualitative Spatial Representation Extraction}
+JK working on this!
 
@@
 These qualitative spatial relationships provide the structure necessary for analogical processing.
-
-\subsection{Representing Context-Dependent Spatial Regions}
-
-To enable analogical transfer, it is necessary to represent context-dependent spatial regions symbolically.
-We do this using \textit{anchor points} \cite{Klenk/etal2005} which are symbolic descriptions linking a conceptual entity to a particular point on a perceived entity.
-A context-dependent spatial region is a conceptual entity, and the objects in the room and the room itself are perceived entities.
-As stated previously, the room is represented by a set of points.
-Anchor points are created using unary functions over a percieved entity.
-For example, \fw{(CentroidFn Area1)} represents the centroid of geometric extent of \fw{Area1}.
-We use anchor points to specify the boundary of the context-dependent spatial region using the \fw{regionBoundary} relation.
-For a particular classroom, we are able to define the front of the room as the region bounded by the leftmost top point of the room, the whiteboard, the leftmost bottom point of the room, and each of the desks in the front row.
-After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation.
-Therefore, each context-dependent spatial region has one type and a variable number of boundary points.
-%\from{Nick}{This is crying out for a diagram, perhaps on the stage map of classroom 1.}
-%\from{Nick}{You mention chalkboard above, but it is not anywhere else in the below example.}
-%\from{Klenk}{We will do for Figure 5.}
-
-\begin{figure}[h]
-	{\fontsize{8}{8} %%Because aaai.sty overrides usual font size commands
-
-\fw{(regionType CDSR1 FrontRegion)\\
-(regionBoundary CDSR1 (LeftmostTopFn Room1))\\
-(regionBoundary CDSR1 (LeftmostBottomFn Room1))\\
-(regionBoundary CDSR1 Desk1)\\
-(regionBoundary CDSR1 Desk2)\\
-(regionBoundary CDSR1 Desk3)\\
-(regionBoundary CDSR1 Desk4)\\
-(regionBoundary CDSR1 Whiteboard0)\\
-}}
-  \caption{Representing the front of the classroom context-dependent region \fw{CDSR1}}
-  \label{fig:cdsr-reps}
-\end{figure}
-
-Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top left of Figure~\ref{fig:dora-maps}.
-The anchor points (connected in green in Figure~\ref{fig:dora-maps}) define the region.
-\fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively.
-The sets of points representing the top and bottom of the classroom are found by selecting the points defining the area within a threshold of the area's maximum or minimum y coordinate.
-The other corners of the context-dependent spatial region are defined by the four desks and the whiteboard.
-Because perceived objects (e.g.
-the desks) are defined as single points, they can serve as anchor points directly.
-The semantic label \fw{FrontRegion} ties the geometric extent of the boundary points to a conceptual region.
-This definition for the front of the classroom is specific to \fw{Room1} and its entities.
-It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary.
-If the desks were in a different position then the region would cover a different extent (e.g.
-if they were closer to the whiteboard then the region would be smaller).
-
-% In the next section, we describe how to transfer these environment-specific representations to new situations via analogy.
-
-\section{Analogical Transfer of Spatial Regions}
-
-% While this work focuses on how analogy enables inferences from specific examples to new situations, but its role in the larger cognitive system this work envisions cannot be ignored.
-
-We assume that a cognitive system will have a way of initially acquiring examples of context-dependent spatial regions, e.g., by being taught by a human through dialogue, sketching, or hand-coding.
-As stated previously, to avoid burdening potential tutors with the task of teaching the system every context-dependent spatial region individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after its initial training.
-For example, after training a janitorial robot in just one classroom, it should be able to identify similar regions in the other classrooms in the building.
-Motivated by the learning and generalization abilities of humans, we have chosen analogy to tackle this problem.
-
-Analogy is an essential cognitive process.
-In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}.
-The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}.
-This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}.
-The identicality constraint provides a strong preference for only allowing identical predicates to match.
-Parallel connectivity states that if two predicates are matched then their arguments must also match.
-The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa.
-To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
-
-The Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} performs analogical matching.
-Given two structured representations as input (the base and target), SME produces one or more mappings.
-Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target.
-Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences.
-SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{analogy2.png}
-  \caption{Analogical mapping between four base expressions and two target expressions.
-Ovals and squares represent relations and entities respectively.
-Purple dashed bi-directional arrows are the correspondences between the base and target, and the green dashed arrow, oval, and square represent a candidate inference projected into the target.}
-  \label{fig:analogy}
-\end{figure}
-
-Figure \ref{fig:analogy} illustrates a sample mapping between four base expressions and two target ones.
-Each oval represents a binary predicate, and its two entity arguments are represented by squares.
-SME generates a mapping between the base expressions \fw{(rcc8-DC Desk1 Desk2)} and \fw{(below Desk1 Desk2)}, and the  target expressions \fw{(rcc8-DC Desk20 Desk22)} and \fw{(below Desk20 Desk22)}.
-This is done in the following manner.
-First, the predicates of these expressions are placed in correspondence by the identicality constraint.
-Next, parallel connectivity aligns their arguments, \fw{Desk1} with \fw{Desk20} and \fw{Desk2} with \fw{Desk22}.
-While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint.
-In Figure \ref{fig:analogy}, these correspondences are highlighted by the purple hashed bi-directional arrows.
-Next, SME creates a candidate inference for the expression \fw{(regionBoundary CSDR1 Desk1)}, because \fw{Desk1} participates in the mapping.
-The candidate inference is stated as \fw{(regionBoundary (:skolem CSDR1) Desk20)}.
-\fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences.
-In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes \fw{Desk20}.
-Note that inference is selective, with no candidate inferences generated for the entirely unmapped \fw{below} expression.
-Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.
-
-% It feels odd having this as a separate subsection
-%\subsection{Initial Approach}\label{sec:approach}
-
-In our system, the base and target representations consist of the identified objects from Dora's spatial model drawn from two different runs in two different environment, plus the qualitative spatial relationships computed on top of these.
-The base also contains a labeled context-dependent spatial region that is sought in target environment.
-We use SME to create an analogical mapping between the base and target.
-The result of this mapping is a set of correspondences and a set of candidate inferences.
-We transfer context-dependent spatial regions as follows.
-First, we identify the context-dependent spatial region entity of the sought semantic type in the base.
-This entity should not participate in the mapping, but should be included as \fw{:skolem} expressions in the candidate inferences.
-To transfer this entity to the target, we collect the candidate inferences resulting from the \fw{regionBoundary} statements in the base that mention the base entity.
-The second arguments of these candidate inferences are anchor points in the target environment.
-Therefore, to identify the context-dependent spatial region in the target, we create a new entity, assign it the semantic type of the sought region, and define its boundary using the anchor points from the candidate inferences.
-
-\section{Example}
-
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{images/classroom-maps-axes-annotation.png}
-  \caption{The top two images show the Player/Stage classroom simulations used in the example, with the base on the left and the target on the right.
-The desks are simulated as red boxes, and the chalkboard and lectern (where present) as blue and green boxes respectively.
-Furthermore, the solid green line in the base connects the boundary points of the front of the room.
-The lower two images show the metric and place maps generated by Dora during the example runs.}
-  \label{fig:dora-maps}
-\end{figure}
-
-Using the previously described techniques, we have implemented a system capable of performing an inference similar to the one presented in Figure \ref{fig:example}.
-The base and target cases are created by running the Dora system on the two different simulated classroom worlds shown in Figure~\ref{fig:dora-maps}.
-In each case, Dora is manually driven around the room to allow it to create metric and place maps.
-Once the map is created, Dora is then driven in front of each object and the visual recognition system is run, resulting in the objects being added to the place layer.
-
-The data from Dora's spatial model is then passed on to the qualitative spatial representation generator.
-In the base case, there are 16 desks and a chalkboard.
-There are two other entities in the case: the room area and the context-specific spatial region representing the front of the room.
-The case includes a total of 96 expressions relating the 19 entities.
-Nine of these expressions are used to define the boundary of a context-specific spatial region representing the front of the room: the four desks in the front row, the chalkboard, and the leftmost bottom and leftmost top of the room.
-The target case includes only 12 desks arranged in three columns of four, a chalkboard and a lectern.
-The target case includes 73 statements and 14 entities.
-
-SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation.
-The resulting analogy includes 62 correspondences between the entities and expressions in the base and target and 53 candidate inferences.
-All seven \fw{regionBoundary} statements in the base appear as candidate inferences.
-Five of the anchor points are exactly correct, the leftmost corners of the room, the chalkboard and two of the desks in the front row.
-The other two analogically transferred anchor points are to the lectern and a desk in the third row.
-Figure \ref{fig:success} illustrates the region created by the seven boundary points in the target environment.
-While this region is not identical to how a human would specify it, our system is able to produce a fairly accurate context-dependent spatial region in a new environment from a single example.
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{images/target-success.png}
-  \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region.
-The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
-  \label{fig:success}
-\end{figure}
-
-\section{Limitations of Initial Approach}
-
-The spatial transfer algorithm described in here has a number of limitations to be addressed in future work.
-First, we have not incorporated a memory retrieval mechanism, such as MAC/FAC \cite{forbus/etal1995}.
-As a cognitive system gains experience in different environments, it will have access to a range of context-dependent spatial regions defined in a variety of task situations.
-To retrieve an analogous example, the agent must select the most similar environment that contains a labeled example of the sought after region type.
-This extension is fairly straightforward, but beyond the scope of this paper.
-Second, our initial approach ignores a number of potentially relevant pieces of information.
-After the context-dependent spatial region has been identified in the target environment, its qualitative spatial relationships could be compared with those from the base to either refine the region or provide a confidence measure for the inference.
-In our example, the transferred region contains the bottom desk in the second row, but the corresponding region does not contain any desks topologically.
-Third, our algorithm uses an absolute reference frame.
-Therefore, the environments must be oriented in a similar fashion to successfully transfer a context-dependent spatial region.
-We can overcome this limitation by performing a series of analogies between the base and the target in which we rotate the base and select the analogy with the best mapping.
-Finally, more advanced qualitative spatial relationships should improve performance.
-For example, recognizing all the desks as one region, or grouping them into rows or columns, would provide a better level of abstraction for this task.
-
-\section{Related Work}
-
-Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}.
-When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g.
-door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}.
-While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions.
-While steps have been taken toward incorporating this knowledge  (i.e.
-the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
-For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g.
-the areas near doors \cite{Zender2008a}).
-
-% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
- % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.
-
-There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks.
-Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}.
-A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations.
-This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user.
-The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning.
-For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}.
-In that work, the anchor points defined the endpoints of lines.
-We go beyond that result to use anchor points to specify 2D regions.
-The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g.
-edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g.
-\cite{Lockwood/etal2008}).
-Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics.
-% Therefore, we hope and expect to see more research combining these three fields.
-
-\section{Discussion}
-
-% Nick: All this reptetition and probably unnecessary
-% Klenk: I like the repetition. I think it is really important to drive home the cognitive systems story here.
-
-Our first claim for this work is that in order for artificial cognitive systems to collaborate with human users, they must be able to understand and reason about the types of spatial regions that humans use.
-In particular we have identified context-dependent spatial regions, i.e., regions defined by their geometric, functional, and semantic relations to surrounding entities, as an important type of region that cannot be represented by current systems.
-Addressing this problem is important as such regions are ubiquitous in human environments and communication, ranging from regions in everyday experience (e.g.
-aisles, marketplaces) to domain-specific regions (e.g.
-the ``pocket'' in American Football).
-
-In addition to addressing a specific need in human-robot interaction, developing a cognitive system able to represent and reason about context-dependent spatial regions is important for two further reasons.
-First, as these regions are not defined solely by their geometry, we must investigate approaches that integrate different types of knowledge.
-This is a key problem in cognitive systems work in general.
-Second, in order to reduce the burden on developers and users, a cognitive system that understands one type of context-dependent spatial region should be able to use this understanding to recognize similar regions in other areas.
-Developing such a generalization ability (particularly one that can work incrementally from a small number of examples) is another general challenge for the cognitive systems community.
-
-Our second claim in this work is that the combination of a sensor-based spatial model, qualitative spatial representations, and analogy provides a potential solution to the aforementioned problems.
-We presented results from an initial proof-of-concept implementation that supports this claim.
-This system is able to transfer, via analogy, context-dependent spatial regions defined using anchor points over logical representations consisting of  object types and qualitative spatial relationships inferred from sensor data.
-Whilst the system has a number of limitations, and the transferred context-dependent spatial region presented does not perfectly match human expectations, it validates further development of our approach.
-
-To support the definition of new spatial regions which the robot should be aware of, it is desirable to integrate a sketching interface on top of a visualization of the spatial knowledge the robot already has.
-Sketching \cite{Forbus/etal:2009a} is often used by humans to convey new ideas and concepts, making this a natural interface for human-robot interaction.
-When faced with an unknown context-dependent spatial region, the mobile robot could then ask a human to define the region using \textit{sketch annotations} over its representation \cite{Klenk/etal2005}.
-
-For mobile robots, spatial understanding is required to enable the pursuit of spatially-defined goals (e.g the goal that the front of the classroom is clean).
-Consequently, spatial understanding should be evaluated with respect to a cognitive system's actions in its environment.
-The minimal performance for spatial understanding is that the cognitive system is able to navigate to context-dependent spatial regions.
-A longer term research project involves reasoning about potentially conflicting goals with overlapping spatial extents.
-A cognitive system should be able to combine them to generate new goals to pursue.
-The scenario in Figure \ref{fig:example} exemplifies this process of combining goals through spatial extent, when the robot generates the goal of cleaning the front of the classroom.
-With the proliferation of mobile robots, sensors and computation, new ideas, such as, context-dependent spatial regions, have the potential to reshape how users interact with cognitive systems.
-
-\section{Acknowledgments}
-
-The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams.
-The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.
-215181, CogX.
-
-\bibliography{aaai12-cdsr}
-\bibliographystyle{aaai}
-\end{document}

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
JK: qsr-extraction.tex updated

--- 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
@@
 
-JK working on this!
+Given data from Dora's spatial model we compute a set of spatial relations between each pair of adjacent objects.
+For each object in the room we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}.
+The relations we compute between a given object and one of the objects adjacent to it are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
+The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.
 
-%Given data from Dora's spatial model, we compute the following qualitative spatial representations.
-First, we extract the object and room entities from the sensor data.
-Next, we determine the qualitative spatial relationships between these entities.
-For the room, we compute its region as a convex hull around the set of points.
-For the objects, we consider only their centroids.
-For each pair of adjacent objects, we compute topological and positional relationships.
-Adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}.
-For the topological relationships, we use the region connection calculus, RCC8, \cite{Cohn:2001}.
-For positional relations, we use two predicates \fw{leftOf} and \fw{below}.
-An entity is \fw{leftOf} another if the x coordinate of its centroid is less than that of the other entity's.
-An entity is \fw{below} another if the y coordinate of its centroid is less than that of the other entity's.
-Objects whose x or y coordinates are within a threshold of one another will not be related by \fw{leftOf} or \fw{below} positional relations.
+To compute these relations we first, extract the object and room entities from the sensor data.
+Next, for the room, we compute its region as a convex hull.
+Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid.
+This translation results in the coordinates of the all the other objects in the room being translated into a frame of references whose origin is the centroid of the landmark.
+For each object adjacent to the landmark we compute the 8 relations by computing the inner angle between the vector from the origin (the landmark's centroid) and the object's location and the direction vector of the relation, as defined above.
+This angle is computed by taking the dot product of these two vectors.
+$\dots$
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
working through

@@
 
-\author{Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com \And Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk \And Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  }
+\author{}
 
@@
 Next, we show how the boundary of a context-dependent spatial region can be defined using \textit{anchor points}.
-Finally, we demonstrate how an existing computational model of analogy can be used to transfer this region to a new situation.
+We then demonstrate how an existing computational model of analogy can be used to transfer this region to a new situation.
+To evaluate this process we compare transferred regions to annotations made on maps of real rooms by human volunteers.
 \end{abstract}
@@
 Instead, it is defined by the objects present in the room (chairs, a desk, a chalkboard), their role in this context (seats for students to watch a teacher who writes on the chalkboard) and their configuration in space (the seats point toward the chalkboard).
-In this paper, we will refer to such regions as \textit{context-dependent spatial regions}.
+In this paper, we will refer to such regions as \textit{context-dependent spatial regions} (CDSRs).
 
-% Pat:
-% Finally, let me urge you to include specific claims in your paper and
-% to provide evidence that supports them. Such evidence can take many
-% forms, from experimental studies to proofs to compelling arguments,
-% but making this explicit will improve your chances.
+% nah: These are the main claims from the ACS paper, but they can be reduced/dropped for aaai
 
 Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:
+
 \begin{itemize}
@@
 
-\begin{figure*}[t!]
-  \includegraphics[width=\textwidth]{example1.jpg}
-  \caption{
-This image demonstrates the behavior we desire of our cognitive system.
-The robot is able to identify the front of the classroom by analogy to a previously labelled example.
-To do this it must identify the correspondences between the example and its current environment (the desks in the front row and chalkboards) and then transfer the spatial region (the front of the room is between these entities).}
-  \label{fig:example}
-\end{figure*}
+% nah: removed large motivating picture, hoping we can make do with just our result pictures, or replace it with images used for ground truth
 
-Figure \ref{fig:example} illustrates our approach for understanding context-dependent spatial regions via analogy.
-Our first assumption is that these regions can be defined using qualitative spatial representations~\cite{Cohn:2001}.
-Our second assumption is that geometrically and semantically similar areas will feature similar context-dependent spatial regions, and that these can be transferred through analogy.
+This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs.
+Our approach is founded on two assumptions.
+The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
 
-In the context of the symposium on ``Advances in Cognitive Systems'' we are making two specific claims:
+The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.
 
-\begin{enumerate}
-\item Understanding context-dependent spatial regions is an essential ability for a cognitive system that must interact with humans.
-\item This ability can be realized in an artificial cognitive system through the combination of a sensor-based spatial model, qualitative spatial representations, and analogy.  
-\end{enumerate}
+The rest of the paper is structured following these assumptions.
+Section~\ref{sec:qsr-gen} describes how we generate QSRs from the spatial model of an existing, state-of-the-art, cognitive system and use these to define CDSRs.
+Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
+Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.
 
-\noindent We appeal to the reader's experience and intuition to evidence the first claim.
-To support this, we have already provided a number of examples of situations in which humans rely on context-dependent spatial regions, but we have not done any systematic collection or analysis of real-world situations in which they occur.
-To evidence the second claim, we present the results of developing an extension to an existing cognitive system that enables it to learn context-dependent spatial regions by example.
+\section{From Sensors to Semantics}\label{sec:qsr-gen}
 
-The rest of the paper is structured as follows.
-First, we describe how we generate qualitative spatial representations from the spatial model of an existing system.
-Second, we describe how we define context-dependent spatial regions using \textit{anchor points} \cite{Klenk/etal2005}, i.e., symbolic descriptions that tie conceptual entities to geometric points.
-Next, we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer knowledge from an existing context-dependent spatial region to a new situation.
-Finally, we discuss some limitations of our current implementation, and position it with respect to previous work and cognitive systems in general.
-
-\section{From Sensors to Semantics}
-
-Our approach to representing and reasoning about context-dependent spatial regions requires the existence of qualitative spatial representations~\cite{Cohn:2001} describing the relationships between the entities a cognitive system perceives (including objects, groups, and regions).
+% Our approach to representing and reasoning about context-dependent spatial regions requires the existence of qualitative spatial representations~\cite{Cohn:2001} describing the relationships between the entities a cognitive system perceives (including objects, groups, and regions).
 By adding semantic labels, or types, to each entity, we are able to reason about the geometric \emph{and} conceptual characteristics of the system's environment.
@@
 
-We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial representation~\cite{Hawes/etal:2011}.
-In this paper, we focus on the basic elements of this spatial model.
+We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial model~\cite{Hawes/etal:2011}.
+In this paper, we draw on the metric layer of this model.
 For more information on Dora's other competences, see recent papers, e.g.~\cite{Hawes/etal:2011,Hanheide/etal:2011}.
 
-In the simplest terms, Dora's spatial model can be treated as two layers: a \emph{metric layer} and a \emph{place layer}.
-A visualization of these from the running system is shown in Figure~\ref{fig:dora-spatial}.
-% Metric layer
-The metric layer is a collection of lines in a 2D global coordinate frame.
+Dora's metric layer is a collection of lines in a 2D global coordinate frame.
 These lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
@@
 In practice, lines are generated at the positions of walls and any other objects that are flat at the height of the laser (e.g.
-foot stools, bins, closed doors etc.).
+bins, closed doors etc.).
 The robot's location in the metric layer is represented as a 2D position plus an orientation.
-% Place layer
-As the robot moves around in the world, the free space it passes through is discretized into a graph of connected nodes $1m$ apart.
-We refer to these nodes as \emph{places} and the graph of connected places forms the spatial model's place layer.
-The robot's location in the place layer is represented by a pointer to a place.
-Places allow Dora to reason about space in basic qualitative ways, e.g.
-by clustering places into rooms based on the presence of doorways, and provide a representation of space capable of facilitating behavior planning.
+
 % Objects
@@
 We use a convex hull around the places in a room to represent its area, and the 2D locations of perceived objects to create the qualitative spatial representations in the next section.
+
+lines = walls
 
@@
 \input{qsr-extraction}
-%\subsection{Qualitative Spatial Representation Extraction}
-
-%Given data from Dora's spatial model, we compute the following qualitative spatial representations.
-First, we extract the object and room entities from the sensor data.
-Next, we determine the qualitative spatial relationships between these entities.
-For the room, we compute its region as a convex hull around the set of points.
-For the objects, we consider only their centroids.
-For each pair of adjacent objects, we compute topological and positional relationships.
-Adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}.
-For the topological relationships, we use the region connection calculus, RCC8, \cite{Cohn:2001}.
-For positional relations, we use two predicates \fw{leftOf} and \fw{below}.
-An entity is \fw{leftOf} another if the x coordinate of its centroid is less than that of the other entity's.
-An entity is \fw{below} another if the y coordinate of its centroid is less than that of the other entity's.
-Objects whose x or y coordinates are within a threshold of one another will not be related by \fw{leftOf} or \fw{below} positional relations.
-
-%%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
-%%Desk 1 and Desk 2 are not labeled in the Figure.
-
-%Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}.
-Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint.
-To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}.
-The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate.
-Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}.
-These qualitative spatial relationships provide the structure necessary for analogical processing.
 
@@
 
-\section{Analogical Transfer of Spatial Regions}
+\section{Analogical Transfer of Spatial Regions}\label{sec:analogy}
 
@@
 
-\section{Acknowledgments}
-
-The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams.
+% \section{Acknowledgments}
+% 
+% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams.
 The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.
 215181, CogX.
+
+% 7th page may only contain references
+\newpage
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
dora section updated

@@
 We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial model~\cite{Hawes/etal:2011}.
-In this paper, we draw on the metric layer of this model.
+In this paper, we draw on the metric map from this model.
 For more information on Dora's other competences, see recent papers, e.g.~\cite{Hawes/etal:2011,Hanheide/etal:2011}.
 
-Dora's metric layer is a collection of lines in a 2D global coordinate frame.
+Dora's metric map is a collection of lines in a 2D global coordinate frame.
 These lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
@@
 
-% Objects
-Dora is capable of using vision to recognize predefined objects.
-Recognition can either be triggered through autonomous visual search or a a user's command.
-When an object is detected it is represented in the place layer relative to Dora's current position.
-The recognizer associates each object with a label that was provided during a training phase.
-In our current work, these labels are used directly in analogical processing.
-We use a convex hull around the places in a room to represent its area, and the 2D locations of perceived objects to create the qualitative spatial representations in the next section.
+Dora is capable of using vision to recognize pre-trained 3D object models.
+Recognition can either be triggered through autonomous visual search or at a user's command.
+When an object is detected it is represented in the metric map
+by placing a copy of the model at the detected pose. The recognizer associates each object with a type label that was provided during a training phase.
 
-lines = walls
-
-% simulation
-As the purpose of this paper is to explore the possibility of using qualitative spatial reasoning and analogy on top of Dora's spatial model, we decided to perform our research using the Player/Stage hardware abstraction layer and simulation tools~\cite{GerkeyVaughanHoward03} rather than on a real robot.
-This allows us to run the Dora system \emph{mostly unchanged} in a controlled simulation environment, rather than deal with some of the unrelated challenges of running the robot in the real world.
-The key difference is that the sensor readings are free of noise in the simulated world.
-This is particularly significant where vision is concerned, as Dora's object recognizer is replaced entirely by a simulated blob detector which is always correct.
-Beyond this, the functionality of the system remains unchanged, and, crucially, the representations our work builds on will be the same when the system is run in the real world.
+To enable us to generate a range of different evaluation situations in a reasonable length of time, we have generated data from Dora in both real rooms and in simulation.
+Simulation is performed using the Player/Stage hardware abstraction layer~\cite{GerkeyVaughanHoward03} allowing us to run the system \emph{mostly unchanged} in a pre-defined environment.
+Also, to enable us to detect a wider range of objects than is usually possible (from armchairs to whiteboards), we used a simulated object recogniser in all runs.
+The recogniser was configured with types and positions of objects in the environment and was guaranteed to detected them when the robot was orientated towards them.
+This eliminated any errors from the recognition process, but was still influenced by errors in robot localisation.
 

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
minor edits

@@
 
-JK working on this!
+For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+The relations we compute between each given \emph{landmark} object and its adjacent neighbour are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
+The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.
 
-%Given data from Dora's spatial model, we compute the following qualitative spatial representations.
-First, we extract the object and room entities from the sensor data.
-Next, we determine the qualitative spatial relationships between these entities.
-For the room, we compute its region as a convex hull around the set of points.
-For the objects, we consider only their centroids.
-For each pair of adjacent objects, we compute topological and positional relationships.
-Adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}.
-For the topological relationships, we use the region connection calculus, RCC8, \cite{Cohn:2001}.
-For positional relations, we use two predicates \fw{leftOf} and \fw{below}.
-An entity is \fw{leftOf} another if the x coordinate of its centroid is less than that of the other entity's.
-An entity is \fw{below} another if the y coordinate of its centroid is less than that of the other entity's.
-Objects whose x or y coordinates are within a threshold of one another will not be related by \fw{leftOf} or \fw{below} positional relations.
+We generate these relations as follows.
+Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid.
+This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
+For each neighbour object adjacent to the landmark we compute the 8 relations by calculating the inner angle between the vector from the origin (the landmark's centroid) and the neighbour's location and the direction vector of the relation, as defined above.
+This angle is computed by taking the dot product of these two vectors.
+$\dots$
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
done sect 2

@@
 
+The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e.
+what can be done there and where.
+In this work we determine this using the type of objects present in a room and their location, both relative to each other and relative to the room itself.
+The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.
+
 \subsection{The Dora System}
@@
 
-To enable analogical transfer, it is necessary to represent context-dependent spatial regions symbolically.
-We do this using \textit{anchor points} \cite{Klenk/etal2005} which are symbolic descriptions linking a conceptual entity to a particular point on a perceived entity.
-A context-dependent spatial region is a conceptual entity, and the objects in the room and the room itself are perceived entities.
-As stated previously, the room is represented by a set of points.
-Anchor points are created using unary functions over a percieved entity.
-For example, \fw{(CentroidFn Area1)} represents the centroid of geometric extent of \fw{Area1}.
-We use anchor points to specify the boundary of the context-dependent spatial region using the \fw{regionBoundary} relation.
-For a particular classroom, we are able to define the front of the room as the region bounded by the leftmost top point of the room, the whiteboard, the leftmost bottom point of the room, and each of the desks in the front row.
+We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs.
+Anchor points are symbolic descriptions which link a conceptual entity (i.e.
+something that can be reasoned about but not directly perceived, such as a CDSR) to a particular point on a perceived entity.
+The perceived entities we use are the objects recognised by Dora, and the room itself.
+The room representation is created by putting a convex hull around the lines in Dora's SLAM map.
+Anchor points are created from perceived entities using unary functions, e.g.
+\fw{(CentroidFn Desk1)} represents the centroid of the geometric extent of \fw{Desk1}, and are linked to a particular CDSR using a \fw{regionBoundary} relation.
 After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation.
-Therefore, each context-dependent spatial region has one type and a variable number of boundary points.
-%\from{Nick}{This is crying out for a diagram, perhaps on the stage map of classroom 1.}
-%\from{Nick}{You mention chalkboard above, but it is not anywhere else in the below example.}
-%\from{Klenk}{We will do for Figure 5.}
+Therefore, each CDSR has one type and a variable number of boundary points.
+
+\textbf{TODO: Need to update example, including new fns, when we have updated pictures in.}
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
analogy section edited

@@
 
-\begin{figure}[t!]
-  \includegraphics[width=\columnwidth]{images/classroom-peekabot-annotated.pdf}
-  \caption{A visualization of Dora's existing spatial model taken from the running system. Walls are visualized where flat surfaces (walls, desks and other objects) are detected during SLAM.}
-  \label{fig:dora-spatial}
-\end{figure}
-
 We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial model~\cite{Hawes/etal:2011}.
@@
 
-\textbf{TODO: Need to update example, including new fns, when we have updated pictures in.}
+\from{Nick}{TODO: Need to update example, including new fns, when we have updated pictures in.}
 
@@
 
-% In the next section, we describe how to transfer these environment-specific representations to new situations via analogy.
-
 \section{Analogical Transfer of Spatial Regions}\label{sec:analogy}
 
-% While this work focuses on how analogy enables inferences from specific examples to new situations, but its role in the larger cognitive system this work envisions cannot be ignored.
-
-We assume that a cognitive system will have a way of initially acquiring examples of context-dependent spatial regions, e.g., by being taught by a human through dialogue, sketching, or hand-coding.
-As stated previously, to avoid burdening potential tutors with the task of teaching the system every context-dependent spatial region individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after its initial training.
-For example, after training a janitorial robot in just one classroom, it should be able to identify similar regions in the other classrooms in the building.
-Motivated by the learning and generalization abilities of humans, we have chosen analogy to tackle this problem.
+We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding.
+To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training.
+For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
+Motivated by the learning and generalization abilities of humans, we have chosen \emph{analogy} as the approach our system will use to solve this problem.
 
@@
 
-The Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} performs analogical matching.
+We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system.
 Given two structured representations as input (the base and target), SME produces one or more mappings.
@@
 \end{figure}
+
+\from{Nick}{TODO: Update Figure \ref{fig:analogy} with correct representation for current version.}
 
@@
 
-In our system, the base and target representations consist of the identified objects from Dora's spatial model drawn from two different runs in two different environment, plus the qualitative spatial relationships computed on top of these.
-The base also contains a labeled context-dependent spatial region that is sought in target environment.
-We use SME to create an analogical mapping between the base and target.
-The result of this mapping is a set of correspondences and a set of candidate inferences.
-We transfer context-dependent spatial regions as follows.
-First, we identify the context-dependent spatial region entity of the sought semantic type in the base.
-This entity should not participate in the mapping, but should be included as \fw{:skolem} expressions in the candidate inferences.
-To transfer this entity to the target, we collect the candidate inferences resulting from the \fw{regionBoundary} statements in the base that mention the base entity.
+In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, plus the QSRs computed on top of these.
+The base also contains a labeled CDSR which we assume is sought in target.
+The result of result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target.
+We use these to transfer the CDSR from base to target (i.e.
+recognizing the CDSR in the target) as follows.
+First, we identify the CDSR of the sought type in the base.
+This entity should not participate in the mapping, but should be included in \fw{:skolem} expressions in the candidate inferences.
+To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{regionBoundary} statements mentioning the base CDSR.
 The second arguments of these candidate inferences are anchor points in the target environment.
-Therefore, to identify the context-dependent spatial region in the target, we create a new entity, assign it the semantic type of the sought region, and define its boundary using the anchor points from the candidate inferences.
+We use these to define the boundary of the CDSR in the target, linking them to a new CDSR entity using \fw{regionBoundary} statements.
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
analogy section edited

@@
 
-Given data from Dora's spatial model we compute a set of spatial relations between each pair of adjacent objects.
-For each object in the room we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}.
-The relations we compute between a given object and one of the objects adjacent to it are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
+\from{Nick}{Do these 8 relations have names? We need them here so we can refer to the predicates later.}
+
+For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+The relations we compute between each given \emph{landmark} object and its adjacent neighbour are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
 The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.
 
-To compute these relations we first, extract the object and room entities from the sensor data.
-Next, for the room, we compute its region as a convex hull.
+We generate these relations as follows.
 Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid.
-This translation results in the coordinates of the all the other objects in the room being translated into a frame of references whose origin is the centroid of the landmark.
-For each object adjacent to the landmark we compute the 8 relations by computing the inner angle between the vector from the origin (the landmark's centroid) and the object's location and the direction vector of the relation, as defined above.
+This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
+For each neighbour object adjacent to the landmark we compute the 8 relations by calculating the inner angle between the vector from the origin (the landmark's centroid) and the neighbour's location and the direction vector of the relation, as defined above.
 This angle is computed by taking the dot product of these two vectors.

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
example looked at. but needs actual data

@@
 
+\from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
+
 We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding.
@@
 
-\section{Example}
+\section{Example System Run}\label{sec:example}
 
@@
 
-Using the previously described techniques, we have implemented a system capable of performing an inference similar to the one presented in Figure \ref{fig:example}.
-The base and target cases are created by running the Dora system on the two different simulated classroom worlds shown in Figure~\ref{fig:dora-maps}.
-In each case, Dora is manually driven around the room to allow it to create metric and place maps.
-Once the map is created, Dora is then driven in front of each object and the visual recognition system is run, resulting in the objects being added to the place layer.
+\from{Nick}{I'd like to keep an example like this in the paper, but we need to choose carefully so we can minimise the number of space-consuming images we need}
 
-The data from Dora's spatial model is then passed on to the qualitative spatial representation generator.
-In the base case, there are 16 desks and a chalkboard.
-There are two other entities in the case: the room area and the context-specific spatial region representing the front of the room.
+To elucidate the workings of our system, we now present an example of how it can transfer a CDSR describing the front of a known classroom (the base) to a new classroom (the target).
+
+We first create the base and target representations by running Dora in the two different classrooms \textbf{TODO:image+ref here}.
+In each case, Dora is manually driven around the room to allow it to create a metric map.
+Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run.
+The map and object data that results from this is then passed on to the QSR generator.
+\from{Nick}{TODO: update for actual example} In the base case, there are 16 desks and a chalkboard.
+There are two other entities in the case: the room area and the CDSR representing the front of the room.
 The case includes a total of 96 expressions relating the 19 entities.
-Nine of these expressions are used to define the boundary of a context-specific spatial region representing the front of the room: the four desks in the front row, the chalkboard, and the leftmost bottom and leftmost top of the room.
+Nine of these expressions are used to define the boundary of a CDSR representing the front of the room: the four desks in the front row, the chalkboard, and the leftmost bottom and leftmost top of the room.
 The target case includes only 12 desks arranged in three columns of four, a chalkboard and a lectern.
@@
 Figure \ref{fig:success} illustrates the region created by the seven boundary points in the target environment.
-While this region is not identical to how a human would specify it, our system is able to produce a fairly accurate context-dependent spatial region in a new environment from a single example.
+
+% While this region is not identical to how a human would specify it, our system is able to produce a fairly accurate context-dependent spatial region in a new environment from a single example.
 
@@
 
-\section{Limitations of Initial Approach}
-
-The spatial transfer algorithm described in here has a number of limitations to be addressed in future work.
+% We can put back in if desired
+%
+% \section{Limitations of Initial Approach}
+% 
+% The spatial transfer algorithm described in here has a number of limitations to be addressed in future work.
 First, we have not incorporated a memory retrieval mechanism, such as MAC/FAC \cite{forbus/etal1995}.

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
included eval in main
added comments

--- 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2011/evaluation.tex
@@
-\documentclass[letterpaper]{article}
+\section{Evaluation}\label{sec:evaluation}
 
-\usepackage{aaai}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{graphicx}
-\usepackage{stfloats}
-\usepackage{color}
-
-\begin{document}
-
- \section{Evaluation}
-
-To evaluate our progress toward building a cognitive system capabile of reasoning about these regions, we conducted the an experiment focusing on the following questions:
+To evaluate our progress toward building a cognitive system capable of reasoning about these CDSRs, we conducted the an experiment focusing on the following questions:
 \begin{itemize}
 \item{How well do anchor points capture context-dependent spatial regions?}
-\item{When provided with an example definition of a CDSR, how well does our approach identify the CDSR in the new situation?}
+\item{When provided with a base representation containing a labelled CDSR, how well does our approach identify the CDSR in a given target?}
 \item{Is the integration of geometric and semantic knowledge necessary?}
@@
 We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments.
-To determine how people consider CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons representing a list of region types for seach room, users were also shown a polygon drawn by another and asked to determine if it was an acceptable depiction of the region.
+To determine how people define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
+Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
 All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.
 
-We consider a \textit{problem instance} to be a room and a sought CDSR type (e.g., "front").
-For each room containing a CDSR of the sought type, we generate an \textit{inferred region} our analogical approach.
+\from{Nick}{What is the base CDSR in a problem instance?}
+
+We consider a \textit{problem instance} to be a room and a sought CDSR type (e.g., ``front'').
+For each room for which we have a human-labelled target CDSR of the sought type, we generate an \textit{inferred region} using our analogical approach.
 To assess the quality of the transfer, we use the following measures:
@@
 
-As a baseline, we will compare our results against the manually encoded CDSRS for the problem instance and the entire room.
+As a baseline, we will compare our results against the manually encoded CDSRs for the problem instance and the entire room.
 The precision and recall for considering the entire room provides a baseline from which to assess each of the questions.
 The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
+
+\from{Nick}{What are manually encoded CDSRs? Are these the ones you do and use for the base case?}
 
@@
 
-If we are just going to report means, we should standard deviations as well.
+\from{Nick}{Would this be just a single region type or averaging across all applicable ones?}
+
+If we are just going to report means, we should standard deviations as well. %yup!
 
@@
 
+% this would be lovely!
+
 Overall the transferred region was significantly better than the entire room (p<Blah), and not statistically different than the manually encoded region (p=Blah).
@@
 \subsection{Discussion}
+
 Our results
-
-\end{document}

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
renaming
related work edited a little

--- 2012-01-17 scenarios/cdsr/trunk/papers/ACS2011/understanding-human-spatial-regions.tex
@@
 
+The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e.
+what can be done there and where.
+In this work we determine this using the type of objects present in a room and their location, both relative to each other and relative to the room itself.
+The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.
+
 \subsection{The Dora System}
 
-\begin{figure}[t!]
-  \includegraphics[width=\columnwidth]{images/classroom-peekabot-annotated.pdf}
-  \caption{A visualization of Dora's existing spatial model taken from the running system. Walls are visualized where flat surfaces (walls, desks and other objects) are detected during SLAM.}
-  \label{fig:dora-spatial}
-\end{figure}
-
-We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial representation~\cite{Hawes/etal:2011}.
-In this paper, we focus on the basic elements of this spatial model.
+We base our work on Dora, a mobile cognitive robot with a pre-existing multi-layered spatial model~\cite{Hawes/etal:2011}.
+In this paper, we draw on the metric map from this model.
 For more information on Dora's other competences, see recent papers, e.g.~\cite{Hawes/etal:2011,Hanheide/etal:2011}.
 
-In the simplest terms, Dora's spatial model can be treated as two layers: a \emph{metric layer} and a \emph{place layer}.
-A visualization of these from the running system is shown in Figure~\ref{fig:dora-spatial}.
-% Metric layer
-The metric layer is a collection of lines in a 2D global coordinate frame.
+Dora's metric map is a collection of lines in a 2D global coordinate frame.
 These lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
@@
 In practice, lines are generated at the positions of walls and any other objects that are flat at the height of the laser (e.g.
-foot stools, bins, closed doors etc.).
+bins, closed doors etc.).
 The robot's location in the metric layer is represented as a 2D position plus an orientation.
-% Place layer
-As the robot moves around in the world, the free space it passes through is discretized into a graph of connected nodes $1m$ apart.
-We refer to these nodes as \emph{places} and the graph of connected places forms the spatial model's place layer.
-The robot's location in the place layer is represented by a pointer to a place.
-Places allow Dora to reason about space in basic qualitative ways, e.g.
-by clustering places into rooms based on the presence of doorways, and provide a representation of space capable of facilitating behavior planning.
-% Objects
-Dora is capable of using vision to recognize predefined objects.
-Recognition can either be triggered through autonomous visual search or a a user's command.
-When an object is detected it is represented in the place layer relative to Dora's current position.
-The recognizer associates each object with a label that was provided during a training phase.
-In our current work, these labels are used directly in analogical processing.
-We use a convex hull around the places in a room to represent its area, and the 2D locations of perceived objects to create the qualitative spatial representations in the next section.
 
-% simulation
-As the purpose of this paper is to explore the possibility of using qualitative spatial reasoning and analogy on top of Dora's spatial model, we decided to perform our research using the Player/Stage hardware abstraction layer and simulation tools~\cite{GerkeyVaughanHoward03} rather than on a real robot.
-This allows us to run the Dora system \emph{mostly unchanged} in a controlled simulation environment, rather than deal with some of the unrelated challenges of running the robot in the real world.
-The key difference is that the sensor readings are free of noise in the simulated world.
-This is particularly significant where vision is concerned, as Dora's object recognizer is replaced entirely by a simulated blob detector which is always correct.
-Beyond this, the functionality of the system remains unchanged, and, crucially, the representations our work builds on will be the same when the system is run in the real world.
+Dora is capable of using vision to recognize pre-trained 3D object models.
+Recognition can either be triggered through autonomous visual search or at a user's command.
+When an object is detected it is represented in the metric map
+by placing a copy of the model at the detected pose. The recognizer associates each object with a type label that was provided during a training phase.
 
-\subsection{Qualitative Spatial Representation Extraction}
+To enable us to generate a range of different evaluation situations in a reasonable length of time, we have generated data from Dora in both real rooms and in simulation.
+Simulation is performed using the Player/Stage hardware abstraction layer~\cite{GerkeyVaughanHoward03} allowing us to run the system \emph{mostly unchanged} in a pre-defined environment.
+Also, to enable us to detect a wider range of objects than is usually possible (from armchairs to whiteboards), we used a simulated object recogniser in all runs.
+The recogniser was configured with types and positions of objects in the environment and was guaranteed to detected them when the robot was orientated towards them.
+This eliminated any errors from the recognition process, but was still influenced by errors in robot localisation.
 
-Given data from Dora's spatial model, we compute the following qualitative spatial representations.
-First, we extract the object and room entities from the sensor data.
-Next, we determine the qualitative spatial relationships between these entities.
-For the room, we compute its region as a convex hull around the set of points.
-For the objects, we consider only their centroids.
-For each pair of adjacent objects, we compute topological and positional relationships.
-Adjacency is determined by creating a voronoi diagram as is standard geometric reasoning \cite{Forbus/etal2003}.
-For the topological relationships, we use the region connection calculus, RCC8, \cite{Cohn:2001}.
-For positional relations, we use two predicates \fw{leftOf} and \fw{below}.
-An entity is \fw{leftOf} another if the x coordinate of its centroid is less than that of the other entity's.
-An entity is \fw{below} another if the y coordinate of its centroid is less than that of the other entity's.
-Objects whose x or y coordinates are within a threshold of one another will not be related by \fw{leftOf} or \fw{below} positional relations.
-
-%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
-%Desk 1 and Desk 2 are not labeled in the Figure.
-
-Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}.
-Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint.
-To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}.
-The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate.
-Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}.
-These qualitative spatial relationships provide the structure necessary for analogical processing.
+\input{qsr-extraction}
 
@@
 
-To enable analogical transfer, it is necessary to represent context-dependent spatial regions symbolically.
-We do this using \textit{anchor points} \cite{Klenk/etal2005} which are symbolic descriptions linking a conceptual entity to a particular point on a perceived entity.
-A context-dependent spatial region is a conceptual entity, and the objects in the room and the room itself are perceived entities.
-As stated previously, the room is represented by a set of points.
-Anchor points are created using unary functions over a percieved entity.
-For example, \fw{(CentroidFn Area1)} represents the centroid of geometric extent of \fw{Area1}.
-We use anchor points to specify the boundary of the context-dependent spatial region using the \fw{regionBoundary} relation.
-For a particular classroom, we are able to define the front of the room as the region bounded by the leftmost top point of the room, the whiteboard, the leftmost bottom point of the room, and each of the desks in the front row.
+We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs.
+Anchor points are symbolic descriptions which link a conceptual entity (i.e.
+something that can be reasoned about but not directly perceived, such as a CDSR) to a particular point on a perceived entity.
+The perceived entities we use are the objects recognised by Dora, and the room itself.
+The room representation is created by putting a convex hull around the lines in Dora's SLAM map.
+Anchor points are created from perceived entities using unary functions, e.g.
+\fw{(CentroidFn Desk1)} represents the centroid of the geometric extent of \fw{Desk1}, and are linked to a particular CDSR using a \fw{regionBoundary} relation.
 After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation.
-Therefore, each context-dependent spatial region has one type and a variable number of boundary points.
-%\from{Nick}{This is crying out for a diagram, perhaps on the stage map of classroom 1.}
-%\from{Nick}{You mention chalkboard above, but it is not anywhere else in the below example.}
-%\from{Klenk}{We will do for Figure 5.}
+Therefore, each CDSR has one type and a variable number of boundary points.
+
+\from{Nick}{TODO: Need to update example, including new fns, when we have updated pictures in.}
 
@@
 
-% In the next section, we describe how to transfer these environment-specific representations to new situations via analogy.
+\section{Analogical Transfer of Spatial Regions}\label{sec:analogy}
 
-\section{Analogical Transfer of Spatial Regions}
+\from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
 
-% While this work focuses on how analogy enables inferences from specific examples to new situations, but its role in the larger cognitive system this work envisions cannot be ignored.
-
-We assume that a cognitive system will have a way of initially acquiring examples of context-dependent spatial regions, e.g., by being taught by a human through dialogue, sketching, or hand-coding.
-As stated previously, to avoid burdening potential tutors with the task of teaching the system every context-dependent spatial region individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after its initial training.
-For example, after training a janitorial robot in just one classroom, it should be able to identify similar regions in the other classrooms in the building.
-Motivated by the learning and generalization abilities of humans, we have chosen analogy to tackle this problem.
+We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding.
+To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training.
+For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
+Motivated by the learning and generalization abilities of humans, we have chosen \emph{analogy} as the approach our system will use to solve this problem.
 
@@
 
-In our system, the base and target representations consist of the identified objects from Dora's spatial model drawn from two different runs in two different environment, plus the qualitative spatial relationships computed on top of these.
-The base also contains a labeled context-dependent spatial region that is sought in target environment.
-We use SME to create an analogical mapping between the base and target.
-The result of this mapping is a set of correspondences and a set of candidate inferences.
-We transfer context-dependent spatial regions as follows.
-First, we identify the context-dependent spatial region entity of the sought semantic type in the base.
-This entity should not participate in the mapping, but should be included as \fw{:skolem} expressions in the candidate inferences.
-To transfer this entity to the target, we collect the candidate inferences resulting from the \fw{regionBoundary} statements in the base that mention the base entity.
+In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, plus the QSRs computed on top of these.
+The base also contains a labeled CDSR which we assume is sought in target.
+The result of result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target.
+We use these to transfer the CDSR from base to target (i.e.
+recognizing the CDSR in the target) as follows.
+First, we identify the CDSR of the sought type in the base.
+This entity should not participate in the mapping, but should be included in \fw{:skolem} expressions in the candidate inferences.
+To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{regionBoundary} statements mentioning the base CDSR.
 The second arguments of these candidate inferences are anchor points in the target environment.
-Therefore, to identify the context-dependent spatial region in the target, we create a new entity, assign it the semantic type of the sought region, and define its boundary using the anchor points from the candidate inferences.
+We use these to define the boundary of the CDSR in the target, linking them to a new CDSR entity using \fw{regionBoundary} statements.
 
-\section{Example}
+\section{Example System Run}\label{sec:example}
 
@@
 
-\section{Limitations of Initial Approach}
+\input{evaluation}
 
-The spatial transfer algorithm described in here has a number of limitations to be addressed in future work.
+% We can put back in if desired
+%
+% \section{Limitations of Initial Approach}
+% 
+% The spatial transfer algorithm described in here has a number of limitations to be addressed in future work.
 First, we have not incorporated a memory retrieval mechanism, such as MAC/FAC \cite{forbus/etal1995}.
@@
 door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}.
-While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions.
-While steps have been taken toward incorporating this knowledge  (i.e.
-the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
-For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g.
+While capable of recognizing spatial regions, these approaches do not consider the semantic information necessary for identifying \emph{context-dependent} regions.
+While steps have been taken towards incorporating this knowledge  into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
+For example, systems usually only refer to the categories of whole rooms, \cite{Hawes/etal:2010b,Vasudevan/etal:2008,Galindo/etal:2005}, whilst others are able to temporarily mediate their behavior when certain functional contexts are detected (e.g.
 the areas near doors \cite{Zender2008a}).
-
-% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
- % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.
 
@@
 A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations.
-This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user.
+This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about CDSRs by identifying the same salient features as its human user.
 The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning.
@@
 Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics.
-% Therefore, we hope and expect to see more research combining these three fields.
 
-\section{Discussion}
+\section{Conclusion}
 
-% Nick: All this reptetition and probably unnecessary
-% Klenk: I like the repetition. I think it is really important to drive home the cognitive systems story here.
+\from{Nick}{TODO!}
 
-Our first claim for this work is that in order for artificial cognitive systems to collaborate with human users, they must be able to understand and reason about the types of spatial regions that humans use.
-In particular we have identified context-dependent spatial regions, i.e., regions defined by their geometric, functional, and semantic relations to surrounding entities, as an important type of region that cannot be represented by current systems.
-Addressing this problem is important as such regions are ubiquitous in human environments and communication, ranging from regions in everyday experience (e.g.
-aisles, marketplaces) to domain-specific regions (e.g.
-the ``pocket'' in American Football).
-
-In addition to addressing a specific need in human-robot interaction, developing a cognitive system able to represent and reason about context-dependent spatial regions is important for two further reasons.
-First, as these regions are not defined solely by their geometry, we must investigate approaches that integrate different types of knowledge.
-This is a key problem in cognitive systems work in general.
-Second, in order to reduce the burden on developers and users, a cognitive system that understands one type of context-dependent spatial region should be able to use this understanding to recognize similar regions in other areas.
-Developing such a generalization ability (particularly one that can work incrementally from a small number of examples) is another general challenge for the cognitive systems community.
-
-Our second claim in this work is that the combination of a sensor-based spatial model, qualitative spatial representations, and analogy provides a potential solution to the aforementioned problems.
-We presented results from an initial proof-of-concept implementation that supports this claim.
-This system is able to transfer, via analogy, context-dependent spatial regions defined using anchor points over logical representations consisting of  object types and qualitative spatial relationships inferred from sensor data.
-Whilst the system has a number of limitations, and the transferred context-dependent spatial region presented does not perfectly match human expectations, it validates further development of our approach.
-
-To support the definition of new spatial regions which the robot should be aware of, it is desirable to integrate a sketching interface on top of a visualization of the spatial knowledge the robot already has.
-Sketching \cite{Forbus/etal:2009a} is often used by humans to convey new ideas and concepts, making this a natural interface for human-robot interaction.
-When faced with an unknown context-dependent spatial region, the mobile robot could then ask a human to define the region using \textit{sketch annotations} over its representation \cite{Klenk/etal2005}.
-
-For mobile robots, spatial understanding is required to enable the pursuit of spatially-defined goals (e.g the goal that the front of the classroom is clean).
-Consequently, spatial understanding should be evaluated with respect to a cognitive system's actions in its environment.
-The minimal performance for spatial understanding is that the cognitive system is able to navigate to context-dependent spatial regions.
-A longer term research project involves reasoning about potentially conflicting goals with overlapping spatial extents.
-A cognitive system should be able to combine them to generate new goals to pursue.
-The scenario in Figure \ref{fig:example} exemplifies this process of combining goals through spatial extent, when the robot generates the goal of cleaning the front of the classroom.
-With the proliferation of mobile robots, sensors and computation, new ideas, such as, context-dependent spatial regions, have the potential to reshape how users interact with cognitive systems.
-
-\section{Acknowledgments}
-
-The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams.
+% \section{Acknowledgments}
+% 
+% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams.
 The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.
@@
 
-\bibliography{understanding-human-spatial-regions} \bibliographystyle{aaai} \end{document}
+% 7th page may only contain references
+\clearpage
+
+\bibliography{aaai12-cdsr}
+\bibliographystyle{aaai}
+\end{document}

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
minor

@@
 
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{images/target-success.png}
-  \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region.
-The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
-  \label{fig:success}
-\end{figure}
+\input{evaluation}
 
@@
 Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics.
-% Therefore, we hope and expect to see more research combining these three fields.
 
-\section{Discussion}
+\section{Conclusion}
 
-% Nick: All this reptetition and probably unnecessary
-% Klenk: I like the repetition. I think it is really important to drive home the cognitive systems story here.
-
-Our first claim for this work is that in order for artificial cognitive systems to collaborate with human users, they must be able to understand and reason about the types of spatial regions that humans use.
-In particular we have identified context-dependent spatial regions, i.e., regions defined by their geometric, functional, and semantic relations to surrounding entities, as an important type of region that cannot be represented by current systems.
-Addressing this problem is important as such regions are ubiquitous in human environments and communication, ranging from regions in everyday experience (e.g.
-aisles, marketplaces) to domain-specific regions (e.g.
-the ``pocket'' in American Football).
-
-In addition to addressing a specific need in human-robot interaction, developing a cognitive system able to represent and reason about context-dependent spatial regions is important for two further reasons.
-First, as these regions are not defined solely by their geometry, we must investigate approaches that integrate different types of knowledge.
-This is a key problem in cognitive systems work in general.
-Second, in order to reduce the burden on developers and users, a cognitive system that understands one type of context-dependent spatial region should be able to use this understanding to recognize similar regions in other areas.
-Developing such a generalization ability (particularly one that can work incrementally from a small number of examples) is another general challenge for the cognitive systems community.
-
-Our second claim in this work is that the combination of a sensor-based spatial model, qualitative spatial representations, and analogy provides a potential solution to the aforementioned problems.
-We presented results from an initial proof-of-concept implementation that supports this claim.
-This system is able to transfer, via analogy, context-dependent spatial regions defined using anchor points over logical representations consisting of  object types and qualitative spatial relationships inferred from sensor data.
-Whilst the system has a number of limitations, and the transferred context-dependent spatial region presented does not perfectly match human expectations, it validates further development of our approach.
-
-To support the definition of new spatial regions which the robot should be aware of, it is desirable to integrate a sketching interface on top of a visualization of the spatial knowledge the robot already has.
-Sketching \cite{Forbus/etal:2009a} is often used by humans to convey new ideas and concepts, making this a natural interface for human-robot interaction.
-When faced with an unknown context-dependent spatial region, the mobile robot could then ask a human to define the region using \textit{sketch annotations} over its representation \cite{Klenk/etal2005}.
-
-For mobile robots, spatial understanding is required to enable the pursuit of spatially-defined goals (e.g the goal that the front of the classroom is clean).
-Consequently, spatial understanding should be evaluated with respect to a cognitive system's actions in its environment.
-The minimal performance for spatial understanding is that the cognitive system is able to navigate to context-dependent spatial regions.
-A longer term research project involves reasoning about potentially conflicting goals with overlapping spatial extents.
-A cognitive system should be able to combine them to generate new goals to pursue.
-The scenario in Figure \ref{fig:example} exemplifies this process of combining goals through spatial extent, when the robot generates the goal of cleaning the front of the classroom.
-With the proliferation of mobile robots, sensors and computation, new ideas, such as, context-dependent spatial regions, have the potential to reshape how users interact with cognitive systems.
+\from{Nick}{TODO!}
 
@@
 % 7th page may only contain references
-\newpage
+\clearpage
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
JK: qsr-extraction.tex updated, spatial relation images added, bib file updated

@@
 
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{images/target-success.png}
+  \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region.
+The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
+  \label{fig:success}
+\end{figure}
+
 \input{evaluation}
@@
 door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}.
-While capable of recognizing spatial regions, these approaches do not consider the semantic information necessary for identifying \emph{context-dependent} regions.
-While steps have been taken towards incorporating this knowledge  into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
-For example, systems usually only refer to the categories of whole rooms, \cite{Hawes/etal:2010b,Vasudevan/etal:2008,Galindo/etal:2005}, whilst others are able to temporarily mediate their behavior when certain functional contexts are detected (e.g.
+While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions.
+While steps have been taken toward incorporating this knowledge  (i.e.
+the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
+For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g.
 the areas near doors \cite{Zender2008a}).
+
+% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
+ % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.
 
@@
 A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations.
-This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about CDSRs by identifying the same salient features as its human user.
+This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user.
 The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning.
@@
 Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics.
+% Therefore, we hope and expect to see more research combining these three fields.
 

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
JK: qsr-extraction.tex updated, spatial relation images added, bib file updated

@@
 
-\from{Nick}{Do these 8 relations have names? We need them here so we can refer to the predicates later.}
+For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms, \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}.
+The model accommodates both direction and distance as factors in the relative position of objects.
 
-For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
-The relations we compute between each given \emph{landmark} object and its adjacent neighbour are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
-The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.
+The relations we compute between each given \emph{landmark} object and its adjacent neighbours are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
+The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$; and, the predicates used to denote these relations are named accordingly, e.g.
+\textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYMinOne}, etc.
 
-We generate these relations as follows.
-Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid.
+We generate the strengths of these spatial relations as follows.
+First we compute the maximum distance between any two points in the room, this value is used to normalize the distances between objects in the room.
+Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid.
 This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
-For each neighbour object adjacent to the landmark we compute the 8 relations by calculating the inner angle between the vector from the origin (the landmark's centroid) and the neighbour's location and the direction vector of the relation, as defined above.
+We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the the distance between the landmark's centroid and the adjacent object's location, and (b) the inner angle between the vector from the origin (the landmark's centroid) to the neighbour's location and the direction vector of the relation.
 This angle is computed by taking the dot product of these two vectors.
-$\dots$
+These two spatial components, distance and angular deviation, are then integrated to compute the strength of a given relationships as follows:
 
-%%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
-%%Desk 1 and Desk 2 are not labeled in the Figure.
+\begin{equation}
+strength = (innerAngle > 90^{\circ}? 0 : (1- \frac{innerAngle}{90^{\circ}}) * (1- \frac{distance}{maxDistance}))
+\end{equation}
 
-%Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}.
-Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint.
-To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}.
-The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate.
-Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}.
-These qualitative spatial relationships provide the structure necessary for analogical processing.
+Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region. These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+
+\begin{figure}
+\centerline{
+\includegraphics[width=2in]{./images/spatialRelations/xOneYZeroWhite.png}
+}
+\caption{A visualisation of a the strength of a spatial relation across a region.
+In this figure the landmark is located at the center of the image and the direction vector used was $<1,0>$ (i.e.
+right of the landmark).
+The darker the pixel the stronger the spatial relation is deemed to be at that point.}
+\label{fig:spr}
+\end{figure}

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
edited qsr text a bit

@@
 For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
-The relations we compute between each given \emph{landmark} object and its adjacent neighbour are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
-The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.
+The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}.
+The model accommodates both direction and distance as factors in the relative position of objects.
 
-We generate these relations as follows.
-Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid.
+The relations we compute between each given \emph{landmark} object and its adjacent neighbours are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
+The canonical directions of these relations are defined using the following vectors: $\langle0,1\rangle$, $\langle1,1\rangle$, $\langle1,0\rangle$, $\langle1,-1\rangle$, $\langle0,-1\rangle$, $\langle-1,-1\rangle$, $\langle-1,0\rangle$, $\langle-1,1\rangle$.
+The predicates used to denote these relations are named accordingly, e.g.
+\textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYMinOne}, etc.
+
+We generate the strengths of these spatial relations as follows.
+First we compute the maximum distance between any two points in the room, this value is used to normalize the distances between objects.
+Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid.
 This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
-For each neighbour object adjacent to the landmark we compute the 8 relations by calculating the inner angle between the vector from the origin (the landmark's centroid) and the neighbour's location and the direction vector of the relation, as defined above.
-This angle is computed by taking the dot product of these two vectors.
-$\dots$
+We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance between the landmark's centroid and the adjacent object's location, and (b) the inner angle between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location.
+These two spatial components are integrated to compute the strength of a given relationships using Equation~\ref{eq:spatial}.
+Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.
 
-%%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
-%%Desk 1 and Desk 2 are not labeled in the Figure.
+\from{Nick}{TODO: The following equation breaks the Matrix}
 
-%Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}.
-Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint.
-To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}.
-The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate.
-Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}.
-These qualitative spatial relationships provide the structure necessary for analogical processing.
+\begin{equation}
+strength = (innerAngle > 90^{\circ}? 0 : (1- \frac{innerAngle}{90^{\circ}}) * (1- \frac{distance}{maxDistance}))
+\label{eq:spatial}
+\end{equation}
+
+% These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+
+\begin{figure}
+\centerline{
+\includegraphics[width=2in]{./images/spatialRelations/xZeroYOne.png}
+}
+\caption{A visualisation of a the strength of a spatial relation across a region.
+In this figure the landmark is located at the center of the image and the direction vector used was $\langle0,1\rangle$ (i.e.
+above of the landmark).
+The darker the pixel the stronger the spatial relation is deemed to be at that point.}
+\label{fig:spr}
+\end{figure}

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
fixed equation and replaced MinOne with NegOne

@@
 The predicates used to denote these relations are named accordingly, e.g.
-\textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYMinOne}, etc.
+\textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYNegOne}, etc.
 
 We generate the strengths of these spatial relations as follows.
-First we compute the maximum distance between any two points in the room, this value is used to normalize the distances between objects.
+First we compute the maximum distance $d_{max}$ between any two points in the room, this value is used to normalize the distances between objects.
 Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid.
 This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
-We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance between the landmark's centroid and the adjacent object's location, and (b) the inner angle between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location.
-These two spatial components are integrated to compute the strength of a given relationships using Equation~\ref{eq:spatial}.
+We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance $d$ between the landmark's centroid and the adjacent object's location, and (b) the inner angle $\theta$ between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location.
+These two spatial components are integrated to compute the strength $s$ of a given relationships using Equation~\ref{eq:spatial}.
 Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.
 
-\from{Nick}{TODO: The following equation breaks the Matrix}
-
 \begin{equation}
-strength = (innerAngle > 90^{\circ}? 0 : (1- \frac{innerAngle}{90^{\circ}}) * (1- \frac{distance}{maxDistance}))
+s = \left\{
+\begin{array}{l l}
+\left(1- \frac{\theta}{90}\right) * \left(1- \frac{d}{d_{max}}\right) & \mathrm{if} \quad \theta \leq 90^{\circ}\\
+0 & \mathrm{otherwise}
+\end{array} \right.
+%strength = (innerAngle > 90^{\circ}? 0 : \left(1- \frac{innerAngle}{90^{\circ}}\right) * \left(1- \frac{distance}{maxDistance}\right))
 \label{eq:spatial}

### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
fixed one typo and made example predicate names correspond to the given order of relations

@@
 For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
-The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms, \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}.
+The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}.
 The model accommodates both direction and distance as factors in the relative position of objects.
@@
 The relations we compute between each given \emph{landmark} object and its adjacent neighbours are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object.
-The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$; and, the predicates used to denote these relations are named accordingly, e.g.
-\textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYMinOne}, etc.
+The canonical directions of these relations are defined using the following vectors: $\langle0,1\rangle$, $\langle1,1\rangle$, $\langle1,0\rangle$, $\langle1,-1\rangle$, $\langle0,-1\rangle$, $\langle-1,-1\rangle$, $\langle-1,0\rangle$, $\langle-1,1\rangle$.
+The predicates used to denote these relations are named accordingly, e.g.
+\textit{xZeroYOne}, \textit{xOneYOne}, \textit{xOneYZero}, \textit{xOneYNegOne}, etc.
 
 We generate the strengths of these spatial relations as follows.
-First we compute the maximum distance between any two points in the room, this value is used to normalize the distances between objects in the room.
+First we compute the maximum distance $d_{max}$ between any two points in the room, this value is used to normalize the distances between objects.
 Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid.
 This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
-We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the the distance between the landmark's centroid and the adjacent object's location, and (b) the inner angle between the vector from the origin (the landmark's centroid) to the neighbour's location and the direction vector of the relation.
-This angle is computed by taking the dot product of these two vectors.
-These two spatial components, distance and angular deviation, are then integrated to compute the strength of a given relationships as follows:
+We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance $d$ between the landmark's centroid and the adjacent object's location, and (b) the inner angle $\theta$ between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location.
+These two spatial components are integrated to compute the strength $s$ of a given relationship using Equation~\ref{eq:spatial}.
+Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.
 
 \begin{equation}
-strength = (innerAngle > 90^{\circ}? 0 : (1- \frac{innerAngle}{90^{\circ}}) * (1- \frac{distance}{maxDistance}))
+s = \left\{
+\begin{array}{l l}
+\left(1- \frac{\theta}{90}\right) * \left(1- \frac{d}{d_{max}}\right) & \mathrm{if} \quad \theta \leq 90^{\circ}\\
+0 & \mathrm{otherwise}
+\end{array} \right.
+%strength = (innerAngle > 90^{\circ}? 0 : \left(1- \frac{innerAngle}{90^{\circ}}\right) * \left(1- \frac{distance}{maxDistance}\right))
+\label{eq:spatial}
 \end{equation}
 
-Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region. These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+% These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
 
@@
 \centerline{
-\includegraphics[width=2in]{./images/spatialRelations/xOneYZeroWhite.png}
+\includegraphics[width=2in]{./images/spatialRelations/xZeroYOne.png}
 }
 \caption{A visualisation of a the strength of a spatial relation across a region.
-In this figure the landmark is located at the center of the image and the direction vector used was $<1,0>$ (i.e.
-right of the landmark).
+In this figure the landmark is located at the center of the image and the direction vector used was $\langle0,1\rangle$ (i.e.
+above of the landmark).
 The darker the pixel the stronger the spatial relation is deemed to be at that point.}

### 2012-01-21 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
baseline data

@@
 The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
+\from{Klenk}{I think that I can do away with all of the stuff about the union from each of the subjects.
+Especially since we are averaging them all together anyways.
+We can have a variation in which we control for everything else and vary the user.
+Perhaps even another table...}
 
 \from{Nick}{What are manually encoded CDSRs? Are these the ones you do and use for the base case?}
+\from{Klenk}{Yes.}
 
@@
 \hline
-Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Studio-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Classroom-1 & p=nil,r=nil & p=.93,r=.60 & p=.23,r=1 \\
+Classroom-2 & p=nil,r=nil & p=.90, r=.69 & p=.19,r=1 \\
+Classroom-3 & p=nil,r=nil & p=.94,r=.64 & p=.17,r=1 \\
+Classroom-4 & p=nil,r=nil & p=85,r=.89 & p=.19,r=1 \\  %This is 222-Classroom
+Classroom-5 & p=nil,r=nil & p=.87,r=.71 & p=.24,r=1 \\ % This is 222-Seminar
+Classroom-6 & p=nil,r=nil & p=.76,r=.66 & p=.20,r=.99 \\ % This is UG40
 \hline
@@
 \from{Nick}{Would this be just a single region type or averaging across all applicable ones?}
+\from{Klenk}{This is averaging across all applicable ones.}
 
@@
 \hline
-Front & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Back & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Front Rows & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Back Rows & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Kitchen & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
-Office & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+Front & p=nil,r=nil & p=.89,r=.79 & p=.23,r=1 \\
+Back & p=nil,r=nil & p=.85,r=.82 & p=.13,r=1 \\
+Front Rows & p=nil,r=nil & p=.85,r=.48 & p=.22,r=1 \\
+Back Rows & p=nil,r=nil & p=.94,r=.40 & p=.7,r=.9 \\
+Kitchen & p=nil,r=nil & p=nil,r=nil & p=nil,r=nil \\
+Office & p=nil,r=nil & p=nil,r=nil & p=nil,r=nil \\
 \hline
@@
 \hline
-p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
+p=nil,r=nil & p=.88,r=.66 & p=.20,r=.99 \\
 \hline

### 2012-01-21 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
added some notes and text

--- 2012-01-21 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
@@
 \subsection{Materials}
+
+\from{Nick}{TODO: Perhaps list manually encoded CDSRs}
+
 We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments.
-To determine how people define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
+The simulated rooms were based on real-life counterparts.
+For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
+These were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.
+
+To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
 Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
 All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.
+The user-provided data provides the ground truth against which we evaluate our transfers.
 
-\from{Nick}{What is the base CDSR in a problem instance?}
-
-We consider a \textit{problem instance} to be a room and a sought CDSR type (e.g., ``front'').
-For each room for which we have a human-labelled target CDSR of the sought type, we generate an \textit{inferred region} using our analogical approach.
-To assess the quality of the transfer, we use the following measures:
+We consider a \textit{problem instance} to be a room and a sought CDSR type.
+For each room for which we have a human-labelled target region of the sought type, we generate an \textit{inferred region} using analogical transfer.
+To assess the quality of the transfer, we calculate precision ($p$) and recall ($r$) as follows:
 
@@
 
-As a baseline, we will compare our results against the manually encoded CDSRs for the problem instance and the entire room.
-The precision and recall for considering the entire room provides a baseline from which to assess each of the questions.
+As a baseline, we compare the transfer results against our manually encoded CDSRs for the target and a region that represents the entire room.
+Results on the entire room provide a baseline performance measure.
 The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
+
+\from{Nick}{I'm confused here.
+Are we comparing the inferred region against our manual region for the target room, or are we comparing our manual region for the target room against the user-provided region for the target room?}
+
 \from{Klenk}{I think that I can do away with all of the stuff about the union from each of the subjects.
@@
 
-\from{Nick}{What are manually encoded CDSRs? Are these the ones you do and use for the base case?}
-\from{Klenk}{Yes.}
+\from{Nick}{Which approach makes the explanation easier, union or averaging?
+The union makes our target the broadest acceptable region (not the average region though).
+Averaging shows how well we match an individual's conception of the region on average.
+Is one of these more relevant to our story than the other?}
 
@@
 \begin{table}
+\small
 \caption{Performance by Room}
@@
 \begin{table}
+\small
 \caption{Performance by Region Type}
@@
 \begin{table}
+\small
 \caption{Performance against Target Region}

### 2012-01-23 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
adding images

@@
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{images/classroom-maps-axes-annotation.png}
-  \caption{The top two images show the Player/Stage classroom simulations used in the example, with the base on the left and the target on the right.
-The desks are simulated as red boxes, and the chalkboard and lectern (where present) as blue and green boxes respectively.
-Furthermore, the solid green line in the base connects the boundary points of the front of the room.
-The lower two images show the metric and place maps generated by Dora during the example runs.}
-  \label{fig:dora-maps}
+  \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
+  \caption{\textbf{TODO}}
+  \label{fig:rooms}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{images/ug40.png}
+  \caption{\textbf{TODO}}
+  \label{fig:ug40}
 \end{figure}
@@
 
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{images/target-success.png}
-  \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region.
+% \begin{figure}[h]
+%   \includegraphics[width=\columnwidth]{images/target-success.png}
+%   \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region.
 The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
-  \label{fig:success}
-\end{figure}
+%   \label{fig:success}
+% \end{figure}
 

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
adding images

@@
 \subsection{Materials}
+
+\from{Nick}{TODO: Perhaps list manually encoded CDSRs}
+
 We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments.
-To determine how people define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
+The simulated rooms were based on real-life counterparts.
+For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
+These were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.
+
+To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
 Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
 All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.
+The user-provided data provides the ground truth against which we evaluate our transfers.
 
-\from{Nick}{What is the base CDSR in a problem instance?}
-
-We consider a \textit{problem instance} to be a room and a sought CDSR type (e.g., ``front'').
-For each room for which we have a human-labelled target CDSR of the sought type, we generate an \textit{inferred region} using our analogical approach.
-To assess the quality of the transfer, we use the following measures:
+We consider a \textit{problem instance} to be a room and a sought CDSR type.
+For each room for which we have a human-labelled target region of the sought type, we generate an \textit{inferred region} using analogical transfer.
+To assess the quality of the transfer, we calculate precision ($p$, the proportion of the inferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the inferred region) as follows:
 

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
adding images

@@
 
+\begin{figure*}[t]
+  \includegraphics[width=\textwidth]{images/worked-example.png}
+  \caption{\textbf{TODO}}
+  \label{fig:rooms-and-pic}
+\end{figure*}
+
 \begin{figure}[t]
-  \includegraphics[width=\columnwidth]{images/classroom-maps-axes-annotation.png}
-  \caption{The top two images show the Player/Stage classroom simulations used in the example, with the base on the left and the target on the right.
-The desks are simulated as red boxes, and the chalkboard and lectern (where present) as blue and green boxes respectively.
-Furthermore, the solid green line in the base connects the boundary points of the front of the room.
-The lower two images show the metric and place maps generated by Dora during the example runs.}
-  \label{fig:dora-maps}
+  \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
+  \caption{\textbf{TODO}}
+  \label{fig:rooms}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{images/ug40.png}
+  \caption{\textbf{TODO}}
+  \label{fig:ug40}
 \end{figure}

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
trying to sanitise images and missing bits

@@
 
-\from{Nick}{TODO: Perhaps list manually encoded CDSRs}
-
 We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments.
@@
 For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
-These were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.
+For the classrooms these were the front and back, and the front and back rows of desk.
+For the studios these were the kitchen, office and living areas.
+These manually encoded regions were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.
 

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
trying to sanitise images and missing bits

@@
 
-\usepackage{float}
-\floatstyle{boxed} 
-\restylefloat{figure}
+% \usepackage{float}
+% \floatstyle{boxed} 
+% \restylefloat{figure}
 
@@
 
-Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top left of Figure~\ref{fig:dora-maps}.
-The anchor points (connected in green in Figure~\ref{fig:dora-maps}) define the region.
+Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top of Figure~\ref{fig:rooms}.
+The anchor points (connected in orange in Figure~\ref{fig:rooms}) define the region.
 \fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively.
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=\textwidth]{images/worked-example.png}
-  \caption{\textbf{TODO}}
-  \label{fig:rooms-and-pic}
-\end{figure*}
+% all images in a single figure -- pretty ugly and a waste of space
+% \begin{figure*}[t]
+%   \includegraphics[width=\textwidth]{images/worked-example.png}
+%   \caption{\textbf{TODO}}
+%   \label{fig:rooms-and-pic}
+% \end{figure*}
 
-\begin{figure}[t]
+\begin{figure}
   \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
-  \caption{\textbf{TODO}}
+  \caption{Maps of 2 real classrooms generated by our system, plus an expert-annotated CDSR (orange, top image), subject-annotated CDSR (red, bottom image) and a CDSR transferred by analogy (green bottom image).
+The classroom used to generate the top classroom is pictured in Figure~\ref{fig:ug40}}
   \label{fig:rooms}
@@
 
-\begin{figure}[t]
+\begin{figure}
   \includegraphics[width=\columnwidth]{images/ug40.png}
-  \caption{\textbf{TODO}}
+  \caption{One of the classrooms used in our evaluation. This image was presented to subjects who were  asked to annotate a copy of the image in the bottom half of Figure~\ref{fig:rooms}.}
   \label{fig:ug40}
@@
 
-\from{Nick}{I'd like to keep an example like this in the paper, but we need to choose carefully so we can minimise the number of space-consuming images we need}
-
 To elucidate the workings of our system, we now present an example of how it can transfer a CDSR describing the front of a known classroom (the base) to a new classroom (the target).
 
-We first create the base and target representations by running Dora in the two different classrooms \textbf{TODO:image+ref here}.
+We first create the base and target representations by running Dora in the two different classrooms.
+These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively.
 In each case, Dora is manually driven around the room to allow it to create a metric map.
@@
 The map and object data that results from this is then passed on to the QSR generator.
-\from{Nick}{TODO: update for actual example} In the base case, there are 16 desks and a chalkboard.
-There are two other entities in the case: the room area and the CDSR representing the front of the room.
-The case includes a total of 96 expressions relating the 19 entities.
-Nine of these expressions are used to define the boundary of a CDSR representing the front of the room: the four desks in the front row, the chalkboard, and the leftmost bottom and leftmost top of the room.
-The target case includes only 12 desks arranged in three columns of four, a chalkboard and a lectern.
-The target case includes 73 statements and 14 entities.
+In the base case Dora perceives 2 desks, a whiteboard and the room area.
+To this we add the CDSR representing the front of the room.
+The case includes a total of \textbf{SOMENUMBER} expressions relating the \textbf{SOMENUMBER} entities.
+\textbf{SOMENUMBER} of these expressions are used to define the boundary of a CDSR representing the front of the room: the \textbf{SOMENUMBER} desks in the front row, the whiteboard, and the leftmost bottom and leftmost top of the room.
+The target case includes \textbf{SOMENUMBER}.
+The target case includes \textbf{SOMENUMBER} statements and \textbf{SOMENUMBER} entities.
 
 SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation.
-The resulting analogy includes 62 correspondences between the entities and expressions in the base and target and 53 candidate inferences.
-All seven \fw{regionBoundary} statements in the base appear as candidate inferences.
-Five of the anchor points are exactly correct, the leftmost corners of the room, the chalkboard and two of the desks in the front row.
-The other two analogically transferred anchor points are to the lectern and a desk in the third row.
-Figure \ref{fig:success} illustrates the region created by the seven boundary points in the target environment.
-
-% While this region is not identical to how a human would specify it, our system is able to produce a fairly accurate context-dependent spatial region in a new environment from a single example.
-
-% \begin{figure}[h]
-%   \includegraphics[width=\columnwidth]{images/target-success.png}
-%   \caption{Successful analogical transfer results in the identification of the front of the room as the green dashed trapezoidal region.
-The red boxes in rows are the identified desks, the blue box along the wall is the chalkboard, the green box is the lectern, and the green and red object is Dora.}
-%   \label{fig:success}
-% \end{figure}
+The resulting analogy includes \textbf{SOMENUMBER} correspondences between the entities and expressions in the base and target and \textbf{SOMENUMBER} candidate inferences.
+All \textbf{SOMENUMBER} \fw{regionBoundary} statements in the base appear as candidate inferences.
+\textbf{SOMENUMBER} of the anchor points are exactly correct, the \textbf{\dots}.
+The other \textbf{SOMENUMBER} analogically transferred anchor points are to the \textbf{\dots}.
+The green region in the lower image of Figure \ref{fig:rooms} illustrates the CDSR created by the \textbf{SOMENUMBER} boundary points in the target environment.
 

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
JK: qsr-extraction.tex updated (typo correction)
trying to sanitise images and missing bits

@@
 \centerline{
-\includegraphics[width=2in]{./images/spatialRelations/xZeroYOne.png}
+\includegraphics[width=\columnwidth]{./images/spatialRelations/xZeroYOne.png}
 }
@@
 above of the landmark).
-The darker the pixel the stronger the spatial relation is deemed to be at that point.}
+The lighter the pixel the stronger the spatial relation is deemed to be at that point.}
 \label{fig:spr}
 \end{figure}
+
+\from{Nick}{TODO: Generating QSRs is this way allows us to\dots}

### 2012-01-23 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
applied john's comments and more edits

@@
 The predicates used to denote these relations are named accordingly, e.g.
-\textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYNegOne}, etc.
+\textit{xZeroYOne}, \textit{xOneYOne}, \textit{xOneYZero}, \textit{xOneYNegOne}, etc.
 
@@
 We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance $d$ between the landmark's centroid and the adjacent object's location, and (b) the inner angle $\theta$ between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location.
-These two spatial components are integrated to compute the strength $s$ of a given relationships using Equation~\ref{eq:spatial}.
+These two spatial components are integrated to compute the strength $s$ of a given relationship using Equation~\ref{eq:spatial}.
 Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.
@@
 
-% These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+Generating the relationships in this way (as opposed to, for example, simple coordinate-based thresholding) has the advantage that the presence and absence of relationships is represented on a continuous scale.
+This provides our representations with the flexibility necessary to manage the variation in perceptual information (i.e.
+the position of walls and objects) inevitable on a robot.
 
@@
 \centerline{
-\includegraphics[width=2in]{./images/spatialRelations/xZeroYOne.png}
+\includegraphics[width=\columnwidth]{./images/spatialRelations/xZeroYOne.png}
 }
 \caption{A visualisation of a the strength of a spatial relation across a region.
-In this figure the landmark is located at the center of the image and the direction vector used was $\langle0,1\rangle$ (i.e.
+The landmark is the red square and the direction vector used was $\langle0,1\rangle$ (i.e.
 above of the landmark).
-The darker the pixel the stronger the spatial relation is deemed to be at that point.}
+The lighter the pixel the stronger the spatial relation is deemed to be at that point.}
 \label{fig:spr}

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
applied john's comments and more edits

@@
 \begin{document}
-% \title{Transferring Human Spatial Regions from Sensor Data via Analogy}
-\title{Representing and Reasoning About Spatial Regions Defined by Context}
-% \title{Towards a Cognitive System Than Can Represent and Reason about Spatial Regions Defined by Context}
+
+\title{Towards a Cognitive System That Can Recognize Spatial Regions Based on Context}
+% \title{Representing and Reasoning About Spatial Regions Defined by Context}
 % \title{Context-Dependent Spatial Regions}
-% \title{Qualitative and Analogical Cognition for Spatial Regions Defined by Context}
 
@@
 The identification of context-dependent spatial regions provides a suitable task for studying this integration.
-Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.}
+% Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.
+}
 
@@
 The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
-
-The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.
-
+The second assumption is that semantically similar areas (e.g.
+two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.
 The rest of the paper is structured following these assumptions.
-Section~\ref{sec:qsr-gen} describes how we generate QSRs from the spatial model of an existing, state-of-the-art, cognitive system and use these to define CDSRs.
+Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs.
 Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
@@
 
-\section{From Sensors to Semantics}\label{sec:qsr-gen}
+\section{From Sensors to Qualitative Representations}\label{sec:qsr-gen}
 
@@
 what can be done there and where.
-In this work we determine this using the type of objects present in a room and their location, both relative to each other and relative to the room itself.
+In this work we implicitly represent context using the types of objects present in a room and their location, both relative to each other and relative to the room itself.
 The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.
@@
 
-\from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
+% \from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
+% done
 
@@
 For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
-Motivated by the learning and generalization abilities of humans, we have chosen \emph{analogy} as the approach our system will use to solve this problem.
+We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
 
@@
 The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}.
-This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}.
+% This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}.
 The identicality constraint provides a strong preference for only allowing identical predicates to match.
@@
 To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
-
 We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system.
-Given two structured representations as input (the base and target), SME produces one or more mappings.
-Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target.
+Given base and target representations as input, SME produces one or more mappings.
+Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures.
 Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences.
@@
 the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
-For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g.
+For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b} (largely ignore location), and others are able to mediate their behavior when certain functional contexts are detected (e.g.
 the areas near doors \cite{Zender2008a}).
 
-% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
- % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.
-
-There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks.
+There is mounting evidence that analogy, operating over structured qualitative representations, can be used to simulate a number of spatial reasoning tasks.
 Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}.
@@
 We go beyond that result to use anchor points to specify 2D regions.
-The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g.
+
+% The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g.
 edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g.
@@
 
-\from{Nick}{TODO!}
+In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions.
+The system is able to use analogy to recognise CDSRs in previously unseen environments through a process of analogy with a manually annotated known room.
+The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy.
+In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with in future work.
+These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g.
+MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g.
+comparing the QSRs present in both base and transferred regions) to improve results.
+Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.
 

### 2012-01-23 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
changed predicate names

@@
 For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
-These were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.
+For the classrooms these were the front and back, and the front and back rows of desk.
+For the studios these were the kitchen, office and living areas.
+These manually encoded regions were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.
 
-To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
+To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for each region type for each room.
+This task was performed using a webpage (link removed for blind review) on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
 Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
@@
 For each room for which we have a human-labelled target region of the sought type, we generate an \textit{inferred region} using analogical transfer.
-To assess the quality of the transfer, we calculate precision ($p$) and recall ($r$) as follows:
+To assess the quality of the transfer, we calculate precision ($p$, the proportion of the inferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the inferred region) as follows:
 
@@
 The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
-
-\from{Nick}{I'm confused here.
-Are we comparing the inferred region against our manual region for the target room, or are we comparing our manual region for the target room against the user-provided region for the target room?}
-
-\from{Klenk}{I think that I can do away with all of the stuff about the union from each of the subjects.
-Especially since we are averaging them all together anyways.
-We can have a variation in which we control for everything else and vary the user.
-Perhaps even another table...}
-
-\from{Nick}{Which approach makes the explanation easier, union or averaging?
-The union makes our target the broadest acceptable region (not the average region though).
-Averaging shows how well we match an individual's conception of the region on average.
-Is one of these more relevant to our story than the other?}
 

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
updated robot related work

@@
 
+% robots and space
+
 Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}.
-When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g.
-door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}.
-While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions.
-While steps have been taken toward incorporating this knowledge  (i.e.
-the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
-For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b} (largely ignore location), and others are able to mediate their behavior when certain functional contexts are detected (e.g.
-the areas near doors \cite{Zender2008a}).
+When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}.
+This paper follows in this tradition, adding CDSRs to his qualitative topological representations.
+Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms.
+This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}.
+The need for an autonomous system to ground references to human-generated descriptions of space has been recognised in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.
 

### 2012-01-23 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
hacked intro

@@
 
-% nah: These are the main claims from the ACS paper, but they can be reduced/dropped for aaai
+Representing and reasoning about CDSRs is beyond the capabilities of current cognitive systems, yet it is an important ability.
+If cognitive systems are to collaborate with humans in everyday environments then they must be able to represent and reason about the spatial regions humans refer to.
+Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.
+In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
+Thus, context-dependent spatial regions provide a suitable context for studying this integration.
 
-Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:
-
-\begin{itemize}
-
-\item{If cognitive systems are to collaborate with humans in everyday environments, then they must be able to represent and reason about the spatial regions humans refer to.
-Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.}
-
-\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-The identification of context-dependent spatial regions provides a suitable task for studying this integration.
-Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.}
-
-\item{Cognitive systems must adapt their knowledge and experience to new situations.
+% Cognitive systems must adapt their knowledge and experience to new situations.
 While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf.
@@
 For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g.
-theaters) without additional tutoring.}
-\end{itemize}
+theaters) without additional tutoring.
 
@@
 
-\section{From Sensors to Semantics}\label{sec:qsr-gen}
+\section{Metric to Qualitative Representations}\label{sec:qsr-gen}
 
@@
 
-\subsection{Representing Context-Dependent Spatial Regions}
+\subsection{Representing CDSRs}
 
@@
 
-\begin{figure}[t]
+% all images in a single figure -- pretty ugly and a waste of space
+% \begin{figure*}[t]
+%   \includegraphics[width=\textwidth]{images/worked-example.png}
+%   \caption{\textbf{TODO}}
+%   \label{fig:rooms-and-pic}
+% \end{figure*}
+
+\begin{figure}
   \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
-  \caption{\textbf{TODO}}
+  \caption{Maps of 2 real classrooms generated by our system, plus an expert-annotated CDSR (orange, top image), subject-annotated CDSR (red, bottom image) and a CDSR transferred by analogy (green bottom image).
+The classroom used to generate the top classroom is pictured in Figure~\ref{fig:ug40}}
   \label{fig:rooms}
@@
 
+% robots and space
+
 Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}.
-When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g.
-door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}.
-While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions.
-While steps have been taken toward incorporating this knowledge  (i.e.
-the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work.
-For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g.
-the areas near doors \cite{Zender2008a}).
+When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}.
+This paper follows in this tradition, adding CDSRs to his qualitative topological representations.
+Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms.
+This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}.
+The need for an autonomous system to ground references to human-generated descriptions of space has been recognised in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.
 
-% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
- % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.
-
-There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks.
+There is mounting evidence that analogy, operating over structured qualitative representations, can be used to simulate a number of spatial reasoning tasks.
 Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}.

### 2012-01-23 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
hacked intro

@@
 This region is not easily perceivable using just the geometry of the environment.
-Instead, it is defined by the objects present in the room (chairs, a desk, a chalkboard), their role in this context (seats for students to watch a teacher who writes on the chalkboard) and their configuration in space (the seats point toward the chalkboard).
-In this paper, we will refer to such regions as \textit{context-dependent spatial regions} (CDSRs).
+Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard).
+We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).
 
-Representing and reasoning about CDSRs is beyond the capabilities of current cognitive systems, yet it is an important ability.
-If cognitive systems are to collaborate with humans in everyday environments then they must be able to represent and reason about the spatial regions humans refer to.
+Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability.
+If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to.
 Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.
 In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-Thus, context-dependent spatial regions provide a suitable context for studying this integration.
+Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.
 
-% Cognitive systems must adapt their knowledge and experience to new situations.
-While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf.
-one-shot learning~\cite{Fei-Fei/etal:2006}).
-This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system.
-For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g.
-theaters) without additional tutoring.
-
-% nah: removed large motivating picture, hoping we can make do with just our result pictures, or replace it with images used for ground truth
-
-This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs.
+This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs.
 Our approach is founded on two assumptions.
-The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
+The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
 The second assumption is that semantically similar areas (e.g.
-two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.
+two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}.
 The rest of the paper is structured following these assumptions.
 Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs.
-Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
+Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
 Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

### 2012-01-24 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
abstract tweaked, inset added to image

@@
 Instead it is defined by its functional use, implied by nearby objects and their configuration.
-In this paper, we define such areas as \textit{context-dependent spatial regions} and propose a method for a cognitive system to learn them incrementally by combining qualitative spatial representations, semantic labels, and analogy.
-Using data from a mobile robot, we generate a relational representation of semantically labeled objects and their configuration.
-Next, we show how the boundary of a context-dependent spatial region can be defined using \textit{anchor points}.
-We then demonstrate how an existing computational model of analogy can be used to transfer this region to a new situation.
-To evaluate this process we compare transferred regions to annotations made on maps of real rooms by human volunteers.
+In this paper, we define such areas as \textit{context-dependent spatial regions} and present a cognitive system able to learn them by combining qualitative spatial representations, semantic labels, and analogy.
+The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world.
+It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.
+We then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms.
+To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
 \end{abstract}
@@
   \includegraphics[width=\columnwidth]{images/ug40.png}
-  \caption{One of the classrooms used in our evaluation. This image was presented to subjects who were  asked to annotate a copy of the image in the bottom half of Figure~\ref{fig:rooms}.}
+  \caption{One of the classrooms used in our evaluation.
+This image was presented to subjects who were asked to annotate a copy of the image in the bottom half of Figure~\ref{fig:rooms}.
+The inset shows a screenshot from the data collection webpage.}
   \label{fig:ug40}

### 2012-01-24 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
updated robot related work
abstract tweaked, inset added to image

@@
 To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for each region type for each room.
-This task was performed using a webpage (link removed for blind review) on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
+This task was performed using a webpage (link removed for blind review) shown in the inset in Figure~\ref{fig:ug40} on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
 Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
@@
 \end{table}
-
-\from{Nick}{Would this be just a single region type or averaging across all applicable ones?}
-\from{Klenk}{This is averaging across all applicable ones.}
 

### 2012-01-24 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
added results...

@@
 \begin{itemize}
-\item{How well do anchor points capture context-dependent spatial regions?}
+\item{Can anchor points capture context-dependent spatial regions?}
 \item{When provided with a base representation containing a labelled CDSR, how well does our approach identify the CDSR in a given target?}
-\item{Is the integration of geometric and semantic knowledge necessary?}
-\item{Does similarity correlate with performance?}
 \end{itemize}
-
-I'm not sure how many of the above questions we want to talk about.
-The last one is fairly easy and won't take up much space in the paper.
-The integration of knowledge question is could take some more work as the grouping is determined by semantic type, so in testing just positional knowledge,  I would have to create an alternate set of anchor points.
 
@@
 
-To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for each region type for each room.
+To determine how people define CDRSs, we asked three users to draw polygons for each region type for each room.
 This task was performed using a webpage (link removed for blind review) shown in the inset in Figure~\ref{fig:ug40} on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
-Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
-All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.
-The user-provided data provides the ground truth against which we evaluate our transfers.
+We call the each user-defined polygon a \textit{target region} against which we evaluate our transfers.
 
@@
 
-As a baseline, we compare the transfer results against our manually encoded CDSRs for the target and a region that represents the entire room.
+We compare the transfer results against our manually encoded CDSRs for the target region and against a region that represents the entire room.
 Results on the entire room provide a baseline performance measure.
-The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
+The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and provides a ceiling for the inferred region's performance.
 
@@
 \small
-\caption{Performance by Room}
-\begin{tabular}{|c|c|c|c|}
+\caption{Overall Performance}
+  \label{tab:overall}	
+\begin{tabular}{|c|c|c|}
 \hline
-Room & Transfer & Manually Encoded & Entire Room \\
+Inferred & Manually Encoded & Entire Room \\
 \hline
-Classroom-1 & p=nil,r=nil & p=.93,r=.60 & p=.23,r=1 \\
-Classroom-2 & p=nil,r=nil & p=.90, r=.69 & p=.19,r=1 \\
-Classroom-3 & p=nil,r=nil & p=.94,r=.64 & p=.17,r=1 \\
-Classroom-4 & p=nil,r=nil & p=85,r=.89 & p=.19,r=1 \\  %This is 222-Classroom
-Classroom-5 & p=nil,r=nil & p=.87,r=.71 & p=.24,r=1 \\ % This is 222-Seminar
-Classroom-6 & p=nil,r=nil & p=.76,r=.66 & p=.20,r=.99 \\ % This is UG40
+$\bar{p}$=.45 $\sigma$=.37, $\bar{r}$=.45 $\sigma$=.38 & $\bar{p}$=.71 $\sigma$=.29, $\bar{r}$=.66 $\sigma$=.24  & $\bar{p}$=.17 $\sigma$=.11, $\bar{r}$=.98 $\sigma$=.05  \\
 \hline
@@
 
-If we are just going to report means, we should standard deviations as well. %yup!
+\from{klenk}{I guess we should rotate the above table. And it has been awhile since I wrote academic, so feel free to cut the boring text below}
+
+Table~\ref{tab:overall} summarizes the results by combining each problem instance and user-defined target region.
+The mean precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
+When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.29) and recall of .66 ($\sigma$=.24).
+The region defined by the room results in a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
 
@@
 \caption{Performance by Region Type}
+  \label{tab:region}	
 \begin{tabular}{|c|c|c|c|}
 \hline
-Region & Transfer & Manually Encoded & Entire Room \\
+Region & Inferred & Manually Encoded & Entire Room \\
 \hline
-Front & p=nil,r=nil & p=.89,r=.79 & p=.23,r=1 \\
-Back & p=nil,r=nil & p=.85,r=.82 & p=.13,r=1 \\
-Front Rows & p=nil,r=nil & p=.85,r=.48 & p=.22,r=1 \\
-Back Rows & p=nil,r=nil & p=.94,r=.40 & p=.7,r=.9 \\
-Kitchen & p=nil,r=nil & p=nil,r=nil & p=nil,r=nil \\
-Office & p=nil,r=nil & p=nil,r=nil & p=nil,r=nil \\
+Front & $\bar{p}$=.33 $\sigma$=.34, $\bar{r}$=.52 $\sigma$=.40 & $\bar{p}$=.60 $\sigma$=.29, $\bar{r}$=.82 $\sigma$=.19  & $\bar{p}$=.15 $\sigma$=.09, $\bar{r}$=1 $\sigma$=0  \\
+\hline
+Back & $\bar{p}$=.45 $\sigma$=.35, $\bar{r}$=.56 $\sigma$=.40 & $\bar{p}$=.66 $\sigma$=.25, $\bar{r}$=.84 $\sigma$=.17  & $\bar{p}$=.11 $\sigma$=.06, $\bar{r}$=.99 $\sigma$=.03  \\
+\hline
+Front Rows & $\bar{p}$=.62 $\sigma$=.36, $\bar{r}$=.22 $\sigma$=.21 & $\bar{p}$=.82 $\sigma$=.30, $\bar{r}$=.50 $\sigma$=.11  & $\bar{p}$=.22 $\sigma$=.08, $\bar{r}$=1 $\sigma$=0  \\
+\hline
+Back Rows & $\bar{p}$=.74 $\sigma$=.28, $\bar{r}$=.43 $\sigma$=.23 & $\bar{p}$=.80 $\sigma$=.29, $\bar{r}$=.43 $\sigma$=.26  & $\bar{p}$=.19 $\sigma$=.06, $\bar{r}$=1 $\sigma$=0  \\
+\hline
+Kitchen & $\bar{p}$=.23 $\sigma$=.03, $\bar{r}$=.09 $\sigma$=.02 & $\bar{p}$=.78 $\sigma$=.20, $\bar{r}$=.71 $\sigma$=.13  & $\bar{p}$=.16 $\sigma$=.02, $\bar{r}$=.92 $\sigma$=.13  \\
+\hline
+Office & $\bar{p}$=.05 $\sigma$=.01, $\bar{r}$=.06 $\sigma$=.06 & $\bar{p}$=.78 $\sigma$=.29, $\bar{r}$=.55 $\sigma$=.20  & $\bar{p}$=.08 $\sigma$=.03, $\bar{r}$=.93 $\sigma$=.06  \\
+Living Room & $\bar{p}$=0 $\sigma$=0, $\bar{r}$=0 $\sigma$=0 & $\bar{p}$=.63 $\sigma$=.34, $\bar{r}$=.54 $\sigma$=.13  & $\bar{p}$=.35 $\sigma$=.22, $\bar{r}$=.96 $\sigma$=.06  \\
 \hline
@@
 
-Or we could report all of the results in a single entry.
+Table~\ref{tab:region} separates the results by the region type of the context-dependent spatial region.
+The mean precision for the transferred regions ranged from .74 for the back rows to 0 for the living room of the studio apartment.
+Comparing the manually encoded against the target regions resulted in a minimum mean precision of .60, which occurred for the front of the classroom.
+The room precision, which is directly proportionally to the size of the target region, varied from .08 for the office to .35 for the living area.
 
-\begin{table}
-\small
-\caption{Performance against Target Region}
-\begin{tabular}{|c|c|c|}
-\hline
-Transfer & Manually Encoded & Entire Room \\
-\hline
-p=nil,r=nil & p=.88,r=.66 & p=.20,r=.99 \\
-\hline
-\end{tabular}
-\end{table}
-
-% this would be lovely!
-
-Overall the transferred region was significantly better than the entire room (p<Blah), and not statistically different than the manually encoded region (p=Blah).
-There was a positive correlation (r=Blah) between the structural evaluation score of the analogy and f-measure of the inference.
+%(3 :transfer-p (0.5649091160864536d0 . 0.3824746399624735d0)  :transfer-r (0.4264200799557943d0 . 0.34846457388485863d0) :self-p   (0.8290901572528112d0 . 0.19630847005535487d0) :self-r
+%  (0.6113880445364026d0 . 0.22432810023132582d0) :room-p
+%  (0.20903207305733315d0 . 0.10291204447091876d0) :room-r
+%  (0.9960679285191474d0 . 0.011022662977174953d0))
+% (2 :transfer-p (0.4496053675639215d0 . 0.36845133415546233d0)
+%  :transfer-r (0.4174204919721308d0 . 0.3697742376406773d0) :self-p
+%  (0.7556540963495391d0 . 0.2672139547143468d0) :self-r
+%  (0.6252647119318083d0 . 0.23778540411202373d0) :room-p
+%  (0.17891943994602752d0 . 0.11273973943366206d0) :room-r
+%  (0.95742701954372d0 . 0.08432884603590181d0))
+% (1 :transfer-p (0.3360804531954755d0 . 0.3364697224078454d0)
+%  :transfer-r (0.5074731366878895d0 . 0.4231938085019586d0) :self-p
+%  (0.5493333659045643d0 . 0.33190756014975015d0) :self-r
+%  (0.7596761147032691d0 . 0.25387231566349d0) :room-p
+%  (0.12277614149421091d0 . 0.10953121808467986d0) :room-r
+%  (0.9946881982552835d0 . 0.011992196508589958d0))
 
@@
 
-Our results
+These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitated learning from a single example through analogical transfer.
+The relatively high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
+The lower scores typically resulted from target regions which were farther away from objects.
+While the inferred regions did not score as well as the manually encoded regions, they were considerably more precise than the room regions.
+The only case were transfer was not possible was between the two studio's living room cases.
+The objects which define the living area in one studio (sofa, TV, and bookshelves) are significantly different than the objects in the other (desk, bed, and sofa), as are the studios in general.

### 2012-01-24 code/subarchitectures/dialogue.sa/branches/george-y3-postreview/resources/grammars/openccg/moloko.v6/docs/comsys.moloko/mood.moloko.tex
branches/george-y3-postreview: code/subarchitectures/dialogue.sa/trunk

 
\sect{Mood}

The 'interpersonal' feature of \diam{Mood} and its syntactic encoding have been given a quite detailed and slightly unorthodox treatment in the Moloko grammar: it is not uniformly specified by the first/main verb, but  has instead been distributed throughout the grammar, with each mood value receiving a different treatment. Before detailing how this has been done, I will briefly describe the motivation behind this choice.

\subsection{Motivation: Mood in English}

At least in English, the syntactic marking of mood is best thought of as clausal-level feature specified by the presence of and/or sequential ordering of the following core clausal elements: 
\begin{enumerate}
\item the grammatical subject
\item the main verb
\item the auxilliary verb
\item a question (wh) word
\end{enumerate}
In Construction Grammar terms, the mood of a clause is specified by a number of Clause-Level Word-Order Constructions. Consider these simple clauses:
\begin{itemize}
\item I picked up the ball.
\item pick up the ball.
\item did you pick up the ball
\item what did you pick up OR who picked up the ball
\end{itemize}

\begin{table}[htdp]
\caption{Identifying Mood}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\textbf{Mood}&\textbf{Initial Element}&\textbf{Example}\\ \hline
Indicative&Noun-Phrase (in nominal case)&I\\ \hline
Imperative&Verb (in base form)&pick\\ \hline
Y/N-Interrogative&Auxiliary Verb(in finite form)&did\\ \hline
Wh-Interrogative&Wh-Word&what \\ 
\hline
\end{tabular}
\end{center}
\label{tab-Identifying-Mood }
\end{table}%
Now consider the mood and initial core element of each clause (Table \ref{tab-Identifying-Mood}). Amazingly, in all 4 cases, we can identify the mood of the clause by considering only its first core element. In other words, the first core element of a clause serves as a cue, projecting the mood of the clause to come. This is important for increasing the effectiveness of incremental processing in two ways.

First, when the mood of a clause is known, a lot is known about its core behavior both syntactically and functionally. This is much like the case of verbs and their arguments: when we encounter the verb \emph{put}, we know that a \catg{np} (the patient) and a prepositional phrase (the result) are to follow. In the case of mood, an indicative clause, for example,  consists of the subject-np followed by a fine main verb (either an auxilliary or a lexical verb and its arguments). Thus, when we encounter an initial (nominative) \catg{np} , if it is the subject-np of an indicative clause, then necessarily the next expected (core) element is the 'verb phrase'. This fits with the general 'early projection' principle (see section \ref{sec-Forward-Projection}).

Secondly, the mood of a clause is highly connected to the discourse function it may be playing (assertion, command, question, etc). Thus, this information should be integrated into the semantic representation as early as possible, allowing the possibility 1) greater expectation/predication at the discourse level and 2) for discourse expectation based pruning.

\subsection{Implementing Mood}

A number of organizational choices were required to allow for the early mood projection described above.  First, as the 'cue elements' described above are non-homogenous, the 'control' of mood had to be distributed throughout the grammar instead of residing solely with the main verb or auxiliary. This also lead to some rather unorthodox treatments. The specifics will be described below.

Second, because mood control has been moved away from the verbs themselves, they had to be blocked from creating their own moodless clauses. Thus, the subject-\catg{np} of verbs has been 'shut-off' or made inert. As an example of this, consider the verb \emph{put}. Here is what we could consider the 'indicative past-tense' entry in standard CCG:
\begin{verbatim}
   s \np /pp /np : 
  @p1:action(put ^ 
             <Mood>ind ^ 
             <Tense>past ^ 
             <Actor>x1:entity ^ 
             <Patient>x2:entity ^ 
             <Result>x3:m-whereto)
\end{verbatim}
Compare this to the corresponding Moloko entry.
\begin{verbatim}
   s \!np /pp /np : 
  @p1:action(put ^ 
             <Tense>past ^ 
             <Actor>x1:entity ^ 
             <Patient>x2:entity ^ 
             <Result>x3:m-whereto)
\end{verbatim}
Instead of \catg{\backslash{} np} we have \catg{\backslash{!} np}. Whereas in the first case this lexical item could combine with a preceding \catg{np} to form an indicative clause, this is not possible in Moloko. Instead, this verb is selected for syntactically and semantically as a compliment by one of the mood-controlling constructions (which will also add the appropriate \diam{Mood} value). In essence, the Moloko grammar does not follow the standard treatment of verbs as the (only) syntactic and semantic head of the clause.

Third, a 'semantic slot' was added to house the grammatical subject. This slot was attached to the top-level (i.e. clausal) event via a dependency relation named \diam{Subject}.\footnote{this has intentionally been removed from all of the semantic representations presented so far.}Despite its seemingly ad-hoc nature, this relation actually serves a number of important functions, though not all at  the ideational(propositional) level. First, in terms of Information Structure, the subject is nearly always the Primary Topic, both in active and passive clauses.  Second, at the interpersonal level, the subject along with the finite (see below) plays an important role in arguing the current proposition 'under negotiation' (see Mood analysis in Systemic Functional Grammar). e.g. \emph{\underline{I} am hungry, No \underline{you}  are not, Yes \underline{I} am, Well \underline{you} shouldn't be, Oh \underline{I} shouldn't, should \underline{I} ? How could \underline{you} possibly be?} Third, the subject serves as the semantic and syntactic controller of chained clauses etc.  \emph{\underline{he} walk\underline{s} into the room, pick\underline{s} up a ball and give\underline{s} it to GJ} (see section \ref{sec-Coordination} and also the the example of imperatives below). In all three cases, explicitly marking that this entity is playing all of these roles could very useful. Finally, and arguably most importantly this slot allows for the semantic integration of what would otherwise be two chunks (the projected clause/event and the np/entity), satisfying the 'one semantic chunk' principle (see section \ref{sec-Back-Integration}).

\subsubsection{Indicatives}
Currently, three types of indicative clauses are handled by the grammar:
\begin{enumerate}
\item Standard: \emph{he put the ball on the table}  
\item Fronted/Topicalized np: \emph{the ball he put on the table} 
\item Dropped/contextualized subject: \emph{put it on the table}
\end{enumerate}
Standard indicative clauses (1) have the following structure: subject-np + finite-verb. In essence, this has been handled by extending the standard syntactic type raising rule used for subject-verb integration, i.e.  \\

\catg{np\id{T} } \arrow \catg{s\id{E} \enspace  \fwdsl{} (s\id{E} \backsl{} np\id{T} ) } \\\\
has become \\
\vspace{-1.5em} %% some negative vertical spacing
\begin{align*}
\catg{np\id{T} }  \arrow  \catg{s\id{E} \fwdsl{} (s\id{E} \backsl{!} np\id{T} )}  :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(ind) \\
      & \wand\ \diam{Subject}(t1:entity ) )
\end{align*}
Syntactically, this rule performs a number of functions. First, it incrementally creates the proper clausal expectations: the clause initial \catg{np} becomes a \catg{s} missing its verb (i.e. \catg{ s \backsl{!} np }). Second, it properly distributes syntactic features among the various elements (e.g. the number and person of the \catg{np} are inherited by the vp, forcing agreement. see \ref{sec-Mood-Rules} for details). Semantically, we see first that the expected event  E has been built into the semantic structure explicitly, although of course in a fully general way (i.e. there are no expectations concerning its sub-sort, proposition, tense, thematic-roles, etc.). Second, this event is marked as indicative. Third, T, the index of the initial \catg{np} fills the \diam{Subject} role of this event. It is also co-indexed with the verb's own subject-\catg{np} (i.e. its  \backsl{!} np ) meaning that it will also play whatever thematic role the verb creates for it (e.g. \diam{Actor} or \diam{Cop-Restr}).
Thus, for example, this rule will combine with the pronoun \emph{I} to create the following parse: \\

parse: \catg{s \fwdsl{}  (s \backsl{!} np)  } 
\vspace{-1.5em} %% some negative vertical spacing

\begin{align*}
\atsign{x1}{event}(&   \\ 
      & \wand\ \diam{Mood}(ind) \\
      & \wand\ \diam{Subject}(i1:person \wand\ \prop{I} \wand\ \diam{Num}(sg)) )
\end{align*}
Again, although syntactic features are not shown the verb compliment is restricted to be in 1st person singular form.

Fronted NP clauses (2) have the following structure: fronted-np + subject-np + finite-verb-with-'missing'-np. This has also been handled by the following type-changing rule:
\vspace{-1.5em} %% some negative vertical spacing

\begin{align*}
\catg{np\id{T} }  \arrow  \catg{s\id{E} \enspace \fwdsl{} (s\id{E} \backsl{!} np\id{S} \fwdsl{} np\id{T}) \enspace \fwdsl{}np\id{S}}  :  
\atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(ind) \\
      & \wand\ \diam{Fronted}(t1:entity ) )\\
      & \wand\ \diam{Subject}(s1:entity ) )
\end{align*}
Note the ordering and co-indexing of the \catg{np}s. The initial \catg{np}, the one that undergoes this type-change, fills a construction specific role  \diam{Fronted} and co-indexes with the 'missing' \catg{np}. The second \catg{np}, the first complement of this construction, is the clausal subject and behaves identically to the case above. Thus, the clause \emph{the ball I haven't seen} (as in \emph{the mug is over on the table, but the ball I haven't seen} receives this parse:
\begin{verbatim}
  @s1:perception(see ^ 
                 <Mood>ind ^ 
                 <Polarity>neg ^ 
                 <Tense>past ^ 
                 <Aspect>perfect ^ 
                 <Actor>(i1:person ^ I ^ 
                         <Num>sg) ^ 
                 <Patient>(b1:thing ^ ball ^ 
                           <Delimination>unique ^ 
                           <Num>sg ^ 
                           <Quantification>specific) ^ 
                 <Fronted>b1:thing ^ 
                 <Subject>i1:person)
\end{verbatim}
Finally, dropped-subject indicatives have this structure:  finite-verb. As this structure is nearly identical to standard imperative clauses, I will delay discussing how it was syntactically handled until then. Semantically, this has been treated as a case of a contextualized semantic object (see section \ref{sec-contextualized}, in this case, the entity filling the \diam{Subject} role. Thus, the clause \emph {walks into the office} receives this parse:
\begin{verbatim}
 @w1:action-motion(walk ^ 
                    <Mood>ind ^ 
                    <Tense>pres ^ 
                    <Actor>(c1:entity ^ context) ^ 
                    <Modifier>(i1:m-whereto ^ into ^ 
                               <Anchor>(o1:e-location ^ office ^ 
                                        <Delimination>unique ^ 
                                        <Num>sg ^ 
                                        <Quantification>specific)) ^ 
                    <Subject>c1:entity)
\end{verbatim}

\subsubsection{Imperatives}
Currently, three types of imperative clauses are handled by the grammar:
\begin{enumerate}
\item Standard Addressee positive: \emph{put the ball on the table, be quiet}  
\item Addressee negative: \emph{don't put the ball on the table, don't be so loud} 
\item Speaker + Addressee: \emph{let's put it on the table, let's be quieter}
\end{enumerate}
The first standard imperative has this structure: base-form verb.\footnote{In fact, they must be of the verbal form 'vf-to-imp'. This allows verbs to be specified lexically whether or not they allow imperative readings. This is discussed in section \ref{sec-Mood-Rules}.} We have handled this via a series subject-removing rules.\footnote{The need for having multiple rules will be discussed in \ref{sec-Mood-Rules}.}As an example of such a rule, here is transitive clause version:
\vspace{-1.5em} %% some negative vertical spacing

\begin{align*}
\catg{s\id{E} \enspace \fwdsl{}np\id{X} \enspace \backsl{!}np\id{S}}  \arrow \catg{s\id{E} \enspace \fwdsl{}np\id{X} \enspace}   :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(imp) \\
      & \wand\ \diam{Subject}(a1:entity \wand\ \prop{addressee}  ) )
\end{align*}
It is crucial that the verb be in base form: only \emph{walk}, but not \emph{walks} or \emph{walked} can be used in imperative clauses. Dropped-subject indicatives (see above) are handled in the same manner, except in those cases the verbal form is constrained to being finite.

An alternate way of handling these two moods would be to do so lexically: each verb would have a subject-less entry which would receive, in the case of imperatives, the feature \diam{Mood}(imp) and vform value \fv{vf-base}. Although this seems a priori more attractive, there are (at least) two reasons for preferring the rule approach instead. Both of these can be illustrated by considering Moloko's semantic representation for \emph{go to the office and wait there}: 
\begin{verbatim}
  @a1:event(and ^ 
            <Mood>imp ^ 
            <Subject>(a2:entity ^ addressee) ^
            <First>(g1:action-motion ^ go ^ 
                    <Actor>a2:entity ^ 
                    <Modifier>(t1:m-whereto ^ to ^ 
                               <Anchor>(o1:e-place ^ office ^ 
                                        <Delimitation>unique ^ 
                                        <Num>sg ^ 
                                        <Quantification>specific))) ^ 
            <Next>(w1:action-non-motion ^ wait ^ 
                   <Actor>a2:entity ^ 
                   <Modifier>(c1:m-location ^ context ^ 
                              <Proximity>m-distal)))
\end{verbatim}
First, although imperative clauses do not have syntactic (i.e. 'surface realized') subjects, they still have an underlying semantic subject. In CCG, this can only be accessed (and hence controlled) by reference to the encoding \catg{np}. Consequently, the proper assignment of the referent \prop{a2} (the 'addressee') to both \diam{Actor} roles is possible only by first coordinating these two clauses (co-indexing their individual subjects) and only then applying the imperative rules to this resulting clause. Second, a lexical treatment would require treating this type of utterance as the coordination of two 'full' sentences, resulting in two separate \diam{Subject} roles and mood values. Instead, by allowing a single rule application, we receive this much more elegant representation, i.e. a single \diam{Subject} and a single mood value 'scoping over' the entire clausal chain (see section \ref{sec-Coordination} for more details).

Negative imperatives have the structure: \emph{don't} + base-form-vp. Because of the lexical item \emph{don't} this was able to be handled lexically. Thus, \emph{don't} receives a lexical entry of this form:
\vspace{-1.5em} %% some negative vertical spacing

\begin{align*}
\catg{s\id{E} \enspace \fwdsl{} (s\id{E} \backsl{!}np\id{S} )} :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(imp) \\
      & \wand\ \diam{Polarity}(neg) \\
      & \wand\ \diam{Subject}(a1:entity \wand\ \prop{addressee}  ) )
\end{align*}
Similarly Speaker+Addressee imperatives have the structure: \emph{let's} + base-form-vp, allowing a simple lexical entry for \emph{let's} of the form:
\vspace{-1.5em}

\begin{align*}
\catg{s\id{E} \enspace \fwdsl{} (s\id{E} \backsl{!}np\id{S} )} :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(imp) \\
      & \wand\ \diam{Subject}(a1:entity \wand\ \prop{speaker+addressee}  ) )
\end{align*}

\subsubsection{Closed Interrogatives}

Closed interrogative clauses are of the form:  aux-verb + subject-np + main-verb. These have been handled by giving all relevant verbs (copula,  auxiliary and modal) a lexical entry corresponding to this clause type. Here is the closed-int entry for the progressive auxiliary \emph{be} (as in \emph{I am sleeping} ) and the negative modal auxiliary \emph{can't} (see section \ref{sec-Modality} for the semantic treatment of modality):
\vspace{-1.5em}

\begin{align*}
\catg{s\id{E} \enspace \fwdsl{} (s\id{E} \backsl{!}np\id{S} ) \enspace \fwdsl{} np\id{S} } :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(int) \\
      & \wand\ \diam{Aspect}(continuous) \\
      & \wand\ \diam{Tense}(pres) \\
      & \wand\ \diam{Subject}(s1:entity) )
\end{align*}

\begin{align*}
\catg{s\id{E} \enspace \fwdsl{} (s\id{E} \backsl{!}np\id{S} ) \enspace \fwdsl{} np\id{S} } :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(int) \\
      & \wand\ \diam{Polarity}(neg) \\
      & \wand\ \diam{Tense}(pres) \\
      & \wand\ \diam{Modifier}(c1:ability \wand\ \prop{can}) ) \\
      & \wand\ \diam{Subject}(s1:entity) )
\end{align*}
In addition to these semantic differences, the vp compliment is restricted according to the lexical requirements of the particular auxiliary verb ( \emph{be} selects for the ing form and \emph{can't} selects for the base form)

\subsubsection{Open Interrogatives} \label{sec-Open-Interrogatives}

Open interrogative clauses are the most varied semantically and syntactically among the mood types. Consequently, the particulars of each type of question will not be discussed until section \ref{sec-Wh-Words}. In this section, we will reserve ourselves to briefly discussing some of the general principles. To begin, table \ref{tab-Open-Int} offers a few illustrating examples. Where relevant, long distance dependencies/extracted constituents are marked with an x.

\begin{table}[htdp]
\caption{Open Interrogative Examples}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Question-Item}&\textbf{Auxilliary}&\textbf{Subject} &\textbf{Verbal Group} \\ \hline
who & & " & picked up the ball \\ \hline
who & & " & wants some coffee \\ \hline
what & did & the robot & see x\\ \hline
where & can & I & put it x\\ \hline
what & are & you & doing x over there \\ \hline
which ball & does & he & want x on the table\\ \hline
what color ball & does & he & want me to have x\\ \hline
how many balls & should & I & pick up x for you \\ \hline
how big & should & I & make it x \\ \hline
where & are & you & going \\ \hline
where & did & he & want me to sit \\ \hline
how quickly & did & he & walk \\ \hline
\hline
\end{tabular}
\end{center}
\label{tab-Open-Int}
\end{table}%
The initial core element of all of these clauses is a question item. This can be as simple as a single word (e.g. \emph{where} or \emph{who} ) or involve a head word plus restricting compliments  (e.g. \emph{\underline{what} color ball} or \emph{\underline{how} big}). Next, in all but the cases where the subject is being questioned, this is followed first by an auxiliary and then by the subject-\catg{np}. Finally, in all cases the verbal group is the final element. Typically, it is 'missing' the compliment  corresponding to the question item, except in those cases where it is a modifier which is being questioned.
All of these examples have been handled by creating an appropriate lexical entry for the corresponding question item head. Thus, it is the 'wh-word' which controls the building of these clauses. Semantically, all open interrogative clauses have been given a dependency relation \diam{Wh-Restr} (abbr. of restrictor) which specifies the nature and the scope of the questioned item. As a first example, consider the entry for \emph{what} used in the question \emph{what did the robot see}.

\vspace{-1.5em} 
\begin{align*}
\catg{s\id{E} \enspace \fwdsl{} (s\id{E} \backsl{!}np\id{S} \fwdsl{} np\id{W}) \enspace \fwdsl{} np\id{S} \enspace \fwdsl{} aux\id{A} } :  \atsign{e1}{event}(&   \\ 
      & \wand\ \diam{Mood}(int) \\
      & \wand\ \diam{Subject}(s1:entity)\\
      & \wand\ \diam{Wh-Restr}(w1:entity \wand\ \prop{what}) )
\end{align*}
This construction has three complements, expected in the following order: an auxiliary, the subject-\catg{np} and finally a verb 'missing' a \catg{np}. The various syntactic and semantic features of the clause (agreement, tense, etc) are properly selected from and imposed by the subject and the auxiliary elements. In this case, the the \diam{Wh-Restr}  is quite simple:  it is an entity co-indexed with the missing \catg{np}, signifying that what is being questioned is an entity which is playing some thematic role in the event being built. As a slightly more complex example, consider the parse for \emph{which ball does he want on the table}.\footnote{\emph{what ball does he want on the table} receives an identical representation.} 
\begin{verbatim}
  @w1:cognition(want ^ 
                <Mood>int ^ 
                <Tense>pres ^ 
                <Actor>(h1:person ^ he ^ 
                        <Num>sg) ^ 
                <Patient>(b1:thing ^ ball) ^ 
                <Result>(o1:m-location ^ on ^ 
                         <Anchor>(t1:thing ^ table ^ 
                                  <Delimination>unique ^ 
                                  <Num>sg ^ 
                                  <Quantification>specific)) ^ 
                <Subject>h1:person ^ 
                <Wh-Restr>(w2:specifier ^ which ^ 
                           <Scope>b1:thing))
\end{verbatim}
The entry for the question word \emph{which} is identical to the one above, except that it has a fourth complement: a noun specifying the class of the entity being questioned. The index corresponding to this noun fills a dependency relation which gives the scope of the \diam{Wh-Restr}. Finally, it is this index, not the index of the \diam{Wh-Restr} as a whole, which is co-indexed with a thematic role in the event. \emph{which} also allows a contextualized semantic head reading (see \ref{sec-Contextualized}), e.g.  \emph{which does he want on the table} would be identical except \diam{Scope}(c1:entity \wand\ \prop{context}). Note that this contrasts with \emph{what does he want on the table}.

Some \diam{Wh-Restr} receive two levels of \diam{Scope}. For example, here is one of the semantic representations for \emph{what color ball do you want}:
\begin{verbatim}
  @w1:cognition(want ^ 
                <Mood>int ^ 
                <Tense>pres ^ 
                <Actor>(y1:person ^ you ^ 
                        <Num>sg) ^ 
                <Patient>(b1:thing ^ ball ^ 
                          <Delimitation>unique ^ 
                          <Num>sg ^ 
                          <Quantification>specific) ^ 
                <Subject>y1:person ^ 
                <Wh-Restr>(w2:specifier ^ what ^ 
                           <Scope>(c1:quality ^ color ^ 
                                   <Scope>b1:thing)))
\end{verbatim}
As one final example, consider the parse for \emph{where did he want me to sit}:
\begin{verbatim}
 @w1:cognition(want ^ 
                <Mood>int ^ 
                <Tense>past ^ 
                <Actor>(h1:person ^ he ^ 
                        <Num>sg) ^ 
                <Event>(s1:action-non-motion ^ sleep ^ 
                        <Actor>i1:person) ^ 
                <Patient>(i1:person ^ I ^ 
                          <Num>sg) ^ 
                <Subject>h1:person ^ 
                <Wh-Restr>(w2:m-location ^ where ^ 
                           <Scope>s1:action-non-motion))
\end{verbatim} 
The syntax required to handle this is complex and will be discussed in \ref{sec-Questioning-an-Event-Modifier}. Semantically, notice that there is no 'missing' element: the \diam{Wh-Restr} is not 'taking the place' of anything. Instead, it is a functional operator which scopes over one of the thematic roles of the event, in this case the \diam{Event} role of the verb \emph{want}. What is crucially important in examples like this is that the wh-word itself is able to semantically restrict what kinds of events(verbs) it is allowed to scope over, i.e. to question. This is done using the modifier classes described in section \ref{sec-Event-Modifier-Restriction}.  In this case, the static locational entry for \emph{where} specifies that it can only question events which allow this type of modifier. Consequently, \emph{* where did he want me to believe} would not parse.

### 2012-01-24 code/subarchitectures/dialogue.sa/branches/george-y3-postreview/apis/dialogue/src/java/de/dfki/lt/tr/dialogue/parse/GrammarDataFcn.java

@@ -1,0 +1,93 @@
+// =================================================================
+// Copyright (C) 2007 Geert-Jan M. Kruijff (gj@dfki.de)
+//
+// This library is free software; you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License
+// as published by the Free Software Foundation; either version 2.1 of
+// the License, or (at your option) any later version.
+//
+// This library is distributed in the hope that it will be useful, but
+// WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU
+// Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU Lesser General Public
+// License along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA
+// 02111-1307, USA.
+// =================================================================
+
+// =================================================================
+// PACKAGE DEFINITION
+package de.dfki.lt.tr.dialogue.parse;
+
+//=================================================================
+// IMPORTS
+
+import de.dfki.lt.tr.dialogue.util.DialogueException;
+
+//-----------------------------------------------------------------
+//JAVA IMPORTS
+//-----------------------------------------------------------------
+import java.util.Iterator;
+import java.util.Vector;
+
+/**
+The abstract class <b>GrammarDataFcn</b> implements the basic 
+functionality for the GrammarData interface. The data is internally 
+stored (as <b>Object</b>s) on a Vector <tt>a_data</tt>, which 
+is accessed through an iterator. 
+
+@version 070813
+@since   070813
+@author	 Geert-Jan M. Kruijff (gj@dfki.de)
+*/
+
+public class GrammarDataFcn 
+implements GrammarData
+{
+
+	// =================================================================
+    // CONSTRUCTOR METHODS
+    // =================================================================	
+	
+	// A list of Object data items
+	private Vector<Object> a_data; 
+
+	// =================================================================
+    // CONSTRUCTOR METHODS
+    // =================================================================	
+
+	public GrammarDataFcn () { 
+		init (); 
+	} // end GrammarDataFcn
+
+	private void init () { 
+		a_data = new Vector<Object>();
+	} // end init
+
+	// =================================================================
+    // ACCESS METHODS
+    // =================================================================	
+
+	/** Adds the provided data item as object to the list of grammar data 
+		
[…]

### 2012-01-24 code/subarchitectures/dialogue.sa/branches/george-y3-postreview/apis/dialogue/src/java/opennlp/ccg/grammar/ForwardTypeRaising.java

@@ -1,0 +1,42 @@
+///////////////////////////////////////////////////////////////////////////////
+// Copyright (C) 2003 Jason Baldridge and University of Edinburgh (Michael White)
+// 
+// This library is free software; you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public
+// License as published by the Free Software Foundation; either
+// version 2.1 of the License, or (at your option) any later version.
+// 
+// This library is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU Lesser General Public License for more details.
+// 
+// You should have received a copy of the GNU Lesser General Public
+// License along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA  02111-1307, USA.
+//////////////////////////////////////////////////////////////////////////////
+
+package opennlp.ccg.grammar;
+
+import opennlp.ccg.synsem.*;
+
+/**
+ * Forward type-raising: X => Y/(Y\X).
+ *
+ * @author  Jason Baldridge
+ * @author  Michael White
+ * @version $Revision: 1.4 $, $Date: 2005/10/18 22:20:15 $
+ */
+public class ForwardTypeRaising extends AbstractTypeRaisingRule {
+    
+    /** Creates a forward type raising rule with the given parameters. */
+    public ForwardTypeRaising (boolean useDollar, Category arg, Category result) {
+        super(">T", new Slash('/', new VarModality("i")), new Slash('\\', new VarModality("i")),
+              useDollar, arg, result);
+    }
+    
+    public String toString() {
+        return "X => Y/(Y\\X)";
+    }
+}
+

### 2012-01-24 subarchitectures/dialogue.sa/branches/george-y3-postreview/apis/cast/src/java/de/dfki/lt/tr/cast/dialogue/IntentionUnpacker.java
branches/george-y3-postreview: code/subarchitectures/dialogue.sa/trunk

@@ -1,0 +1,158 @@
+package de.dfki.lt.tr.cast.dialogue;
+
+import cast.AlreadyExistsOnWMException;
+import cast.DoesNotExistOnWMException;
+import cast.SubarchitectureComponentException;
+import cast.UnknownSubarchitectureException;
+import cast.architecture.ChangeFilterFactory;
+import cast.architecture.ManagedComponent;
+import cast.architecture.WorkingMemoryChangeReceiver;
+import cast.cdl.WorkingMemoryAddress;
+import cast.cdl.WorkingMemoryChange;
+import cast.cdl.WorkingMemoryOperation;
+import de.dfki.lt.tr.beliefs.slice.intentions.InterpretedIntention;
+import de.dfki.lt.tr.beliefs.slice.intentions.PossibleInterpretedIntentions;
+import de.dfki.lt.tr.beliefs.slice.sitbeliefs.dBelief;
+import de.dfki.lt.tr.dialogue.interpret.InterpretedUserIntention;
+import de.dfki.lt.tr.dialogue.util.BeliefIntentionUtils;
+
+public class IntentionUnpacker extends AbstractDialogueComponent {
+
+	@Override
+	public void onStart() {
+		super.onStart();
+
+		addChangeFilter(
+				ChangeFilterFactory.createLocalTypeFilter(
+						PossibleInterpretedIntentions.class,
+						WorkingMemoryOperation.ADD),
+				new WorkingMemoryChangeReceiver() {
+					@Override
+					public void workingMemoryChanged(WorkingMemoryChange _wmc) {
+						handlePossibleInterpretedIntentions(_wmc);
+					}
+				});
+	}
+
+	private void handlePossibleInterpretedIntentions(WorkingMemoryChange _wmc) {
+		addTask(new ProcessingTaskWithDataAndComponent<WorkingMemoryAddress, IntentionUnpacker>(
+				_wmc.address, this) {
+
+			@Override
+			public void execute(WorkingMemoryAddress addr) {
+				try {
+					unpackMostConfidentIntention(getComponent(), addr);
+				} catch (SubarchitectureComponentException ex) {
+					getLogger().error("component exception", ex);
+				}
+			}
+
+		});
+	}
+
+	/**
+	 * Unpack the most confident intention.
+	 * 
+	 * @param _component
+	 * @param addr
+	 * @throws DoesNotExistOnWMException
+	 * @throws UnknownSubarchitectureException
+	 * @throws AlreadyExistsOnWMException
+	 */
[…]

### 2012-01-24 code/subarchitectures/visualization.sa/branches/george-y3-postreview/src/c++/core/object/CRasterImage.hpp
branches/george-y3-postreview: code/subarchitectures/visualization.sa/trunk

@@ -1,0 +1,62 @@
+/*
+ * Author: Marko Mahnič
+ * Created: 2010-03-31
+ *
+ * © Copyright 2010 Marko Mahnič. 
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 2 of the License, or
+ * (at your option) any later version.
+ * 
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ */
+#ifndef CRASTERIMAGE_4RYC90L
+#define CRASTERIMAGE_4RYC90L
+
+#include "../Model.hpp"
+#include <QImage>
+
+namespace cogx { namespace display {
+
+class CRasterImage: public CDisplayObject
+{
+   friend class CRasterImage_Render2D;
+   static std::auto_ptr<CRenderer> render2D;
+   friend class CRasterImage_RenderScene;
+   static std::auto_ptr<CRenderer> renderScene;
+
+public:
+   QImage* m_pImage;
+
+public:
+   CRasterImage();
+   ~CRasterImage();
+   bool isBitmap(); /*override*/
+   virtual CRenderer* getRenderer(ERenderContext context); /*override*/
+   virtual ERenderContext getPreferredContext()
+   {
+      return ContextGraphics;
+   }
+   virtual bool removePart(const std::string& partId, CPtrVector<CDisplayObjectPart>& parts) /*override*/
+   {
+      return false;
+   }
+};
+
+class CRasterImage_Render2D: public CRenderer
+{
+public:
+   virtual void draw(CDisplayView *pView, CDisplayObject *pObject, void *pContext); /*override*/
+};
+
+class CRasterImage_RenderScene: public CRenderer
+{
+public:
+   virtual void draw(CDisplayView *pView, CDisplayObject *pObject, void *pContext); /*override*/
+};
+
+}} // namespace
+#endif /* end of include guard: CRASTERIMAGE_4RYC90L */

### 2012-01-24 code/subarchitectures/visualization.sa/branches/george-y3-postreview/src/c++/core/object/gllua/gluaux.h

@@ -1,0 +1,20 @@
+/* gluaux.c
+** Auxiliary function to bind GLU to Lua.
+** Waldemar Celes
+** Sep 1998
+*/
+
+#ifndef gluaux_h
+#define gluaux_h
+
+#include "tolua++.h"
+
+GLUtriangulatorObj* gluauxNewTess (void);
+void gluauxDeleteTess (GLUtriangulatorObj* tessobj);
+void gluauxTessCallback(GLUtriangulatorObj* tessobj, GLenum type, lua_Number fn);
+void gluauxBeginPolygon (GLUtriangulatorObj* tessobj);
+// void gluauxTessVertex (GLUtriangulatorObj* tessobj, GLdouble v[3], lua_Number data);
+void gluauxNextContour (GLUtriangulatorObj* tessobj, GLenum type);
+void gluauxEndPolygon (GLUtriangulatorObj* tessobj);
+
+#endif

### 2012-01-24 subarchitectures/visualization.sa/branches/george-y3-postreview/doc/v11n.tex
branches/george-y3-postreview: code/subarchitectures/visualization.sa/trunk

--- 2010-07-14 subarchitectures/visualization.sa/trunk/doc/v11n.tex
@@
 as the object.
-\todo{(TODO: It will be posible to define additional views that display multiple objects)}
+\todo{(TODO: It will be possible to define additional views that display multiple objects)}
 The display server creates rendering contexts in which the views are displayed.
@@
 and a checkbox).
+
+\subsection{CAST Configuration}
+
+The display server can run either as a component inside a CAST c++ process or
+as a standalone server. If it is used as a component it can be configured to
+redirect the clients to a standalone display server. In this case the display
+server will not create any GUI. If multiple display server components are
+created inside a c++ process, only one can be configured to create a GUI,
+otherwise the system will most likely crash during an attempt to create a
+second Qt application object. On some systems it is not possible to run the
+display server as a CAST component because the Qt implementation may require
+that the application object is created in the main thread. In this case a
+standalone display server should be used and possibly a display server
+component that will redirect the clients to the standalone server.
+
+When a display server runs as a CAST component it accepts the parameter
+\code{--redirect\--to\--host "hostname"}. When this parameter is not empty the
+server will redirect the clients that connect to it to a standalone display
+server running on the computer named \code{hostname}.
+
+Two parameters are used to configure the clients. The parameter
+\code{--displayserver "id"} takes an id of the display server component to
+connect to. The parameter \code{--standalone\--display\--host "hostname"} can
+be used to connect to a standalone display server.
+
+Wherever \code{"hostname"} is required an empty value can be used or the value
+\code{"/no"} which is translated to an empty string internally. Using an empty
+value or the equivalent \code{"/no"} has the same effect as not specifying the
+parameter at all.
+
+The preferred way to configure the use of standalone display servers is by
+using the parameters \code{--displayserver} and \code{--redirect\--to\--host
+"hostname"}. This way only one change has to be made to \code{.cast} filed when
+the location of the display server has to be changed.
+
+It is possible to configure the clients so that each client connects to a
+different display server with the parameter \code{--standalone\--display\--host
+"hostname"}. A better alternative is to define multiple display server
+components (each with a different id and possibly a different host in
+\code{--redirect\--to\--host}) and use only \code{--displayserver} for clients.
+
+\subsubsection*{Case 1: Use a standalone display server}
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   CPP MG display.srv DisplayServer --redirect-to-host localhost
+   CPP MG visualization.test VisualizationTest --displayserver "display.srv" ...
+\end{Verbatim}
+
+The redirection is configured in the DisplayServer component. The client
+connects to the server "display.srv", obtains remote server parameters (ie
+localhost), drops the connection to "display.srv" and creates a new connection
+to the standalone display server running on "localhost".
+
+\subsubsection*{Case 2: Use a standalone display server for a component}
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   CPP MG visualization.test VisualizationTest --displayserver "display.srv"
+         --standalone-display-host localhost
+\end{Verbatim}
+
+The redirection is configured in the client component. Although the parameter
+\code{--displayserver} is used, the client doesn't try to connect to a
+component named "display.srv". Instead it uses the parameters for the
+standalone display server and tries to connect directly. In a sense
+\code{--standalone\--display\--host} takes precedence over
+\code{--redirect\--to\--host}.
+
+\subsubsection*{Case 3: use the display server as a component}
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   CPP MG display.srv DisplayServer
+   CPP MG visualization.test VisualizationTest --displayserver "display.srv" ...
+\end{Verbatim}
+
+Equivalent alternatives using the parameter --redirect-to-host:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   CPP MG display.srv DisplayServer --redirect-to-host ""
+   CPP MG display.srv DisplayServer --redirect-to-host "/no"
+\end{Verbatim}
+
+With these two forms it is easy to configure the display with an environment
+variable:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   CPP MG display.srv DisplayServer --redirect-to-host $V11N_STANDALONE_HOST
+\end{Verbatim}
+
+Set the environment variable \code{V11N\_STANDALONE\_HOST} as desired:
+
+\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
+   export V11N_STANDALONE_HOST=localhost
+   export V11N_STANDALONE_HOST=/no
+   ...
+\end{Verbatim}
 

### 2012-01-24 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
editing eval section

@@
 This region is not easily perceivable using just the geometry of the environment.
-Instead, it is defined by the objects present in the room (chairs, a desk, a chalkboard), their role in this context (seats for students to watch a teacher who writes on the chalkboard) and their configuration in space (the seats point toward the chalkboard).
-In this paper, we will refer to such regions as \textit{context-dependent spatial regions} (CDSRs).
+Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard).
+We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).
 
-% nah: These are the main claims from the ACS paper, but they can be reduced/dropped for aaai
+Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability.
+If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to.
+Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.
+In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
+Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.
 
-Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:
-
-\begin{itemize}
-
-\item{If cognitive systems are to collaborate with humans in everyday environments, then they must be able to represent and reason about the spatial regions humans refer to.
-Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.}
-
-\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-The identification of context-dependent spatial regions provides a suitable task for studying this integration.
-% Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.
-}
-
-\item{Cognitive systems must adapt their knowledge and experience to new situations.
-While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf.
-one-shot learning~\cite{Fei-Fei/etal:2006}).
-This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system.
-For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g.
-theaters) without additional tutoring.}
-\end{itemize}
-
-% nah: removed large motivating picture, hoping we can make do with just our result pictures, or replace it with images used for ground truth
-
-This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs.
+This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs.
 Our approach is founded on two assumptions.
-The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
+The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
 The second assumption is that semantically similar areas (e.g.
-two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.
+two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}.
 The rest of the paper is structured following these assumptions.
 Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs.
-Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
+Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
 Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.
 
-\section{From Sensors to Qualitative Representations}\label{sec:qsr-gen}
+\section{Metric to Qualitative Representations}\label{sec:qsr-gen}
 
@@
 \begin{figure}[h]
-  \includegraphics[width=\columnwidth]{analogy2.png}
+  \includegraphics[width=\columnwidth]{images/AnalogyDiagramv4.png}
   \caption{Analogical mapping between four base expressions and two target expressions.

### 2012-01-24 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
editing eval section

@@
 
-To evaluate our progress toward building a cognitive system capable of reasoning about these CDSRs, we conducted the an experiment focusing on the following questions:
+To evaluate our progress toward building a cognitive system capable of reasoning about CDSRs, we conducted the an experiment focusing on the following questions:
 \begin{itemize}
-\item{Can anchor points capture context-dependent spatial regions?}
+\item{Are anchor points able to encode context-dependent spatial regions?}
 \item{When provided with a base representation containing a labelled CDSR, how well does our approach identify the CDSR in a given target?}
@@
 
-To determine how people define CDRSs, we asked three users to draw polygons for each region type for each room.
+To determine how people define CDRSs, we asked three na\"ive users to draw polygons for each region type for each room.
 This task was performed using a webpage (link removed for blind review) shown in the inset in Figure~\ref{fig:ug40} on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
-We call the each user-defined polygon a \textit{target region} against which we evaluate our transfers.
+The user-defined polygons define the \textit{target regions} against which we evaluate our transfers.
 
@@
 
-We compare the transfer results against our manually encoded CDSRs for the target region and against a region that represents the entire room.
-Results on the entire room provide a baseline performance measure.
-The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and provides a ceiling for the inferred region's performance.
+Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually annotated CDSR in the target and appropriate target region; and the region for the whole room and the target region.
+ Results comparing transferred and target regions measure how well the output of our system matches user expectations.
+Results comparing the manual annotations to the target regions measure how well the anchor point-based regions we chose match the users' regions (who were not constrained to anchor points).
+Results comparing whole rooms to target regions provide a baseline performance for comparison.
 
@@
 \small
-\caption{Overall Performance}
-  \label{tab:overall}	
 \begin{tabular}{|c|c|c|}
@@
 \end{tabular}
+\caption{Overall Performance}
+  \label{tab:overall}
+
 \end{table}
@@
 \small
-\caption{Performance by Region Type}
-  \label{tab:region}	
 \begin{tabular}{|c|c|c|c|}
@@
 Office & $\bar{p}$=.05 $\sigma$=.01, $\bar{r}$=.06 $\sigma$=.06 & $\bar{p}$=.78 $\sigma$=.29, $\bar{r}$=.55 $\sigma$=.20  & $\bar{p}$=.08 $\sigma$=.03, $\bar{r}$=.93 $\sigma$=.06  \\
+\hline
 Living Room & $\bar{p}$=0 $\sigma$=0, $\bar{r}$=0 $\sigma$=0 & $\bar{p}$=.63 $\sigma$=.34, $\bar{r}$=.54 $\sigma$=.13  & $\bar{p}$=.35 $\sigma$=.22, $\bar{r}$=.96 $\sigma$=.06  \\
@@
 \end{tabular}
+\caption{Performance by Region Type}
+  \label{tab:region}	
 \end{table}

### 2012-01-24 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Kate: Updated the analogy diagram and text

@@
 (regionBoundary CDSR1 Desk4)\\
-(regionBoundary CDSR1 Whiteboard0)\\
+(regionBoundary CDSR1 Whiteboard0)
 }}
@@
 \begin{figure}[h]
-  \includegraphics[width=\columnwidth]{images/AnalogyDiagramv4.png}
-  \caption{Analogical mapping between four base expressions and two target expressions.
-Ovals and squares represent relations and entities respectively.
-Purple dashed bi-directional arrows are the correspondences between the base and target, and the green dashed arrow, oval, and square represent a candidate inference projected into the target.}
+  \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
+  \caption{Analogical mapping between five base expressions and two target expressions. }
+  %Ovals and squares represent relations and entities respectively.
+Black dashed bi-directional arrows are the correspondences between the base and target, and the red dashed objects represent a candidate inference projected into the target.
+
   \label{fig:analogy}
@@
 
-Figure \ref{fig:analogy} illustrates a sample mapping between four base expressions and two target ones.
+Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and two target ones.
 Each oval represents a binary predicate, and its two entity arguments are represented by squares.
-SME generates a mapping between the base expressions \fw{(rcc8-DC Desk1 Desk2)} and \fw{(below Desk1 Desk2)}, and the  target expressions \fw{(rcc8-DC Desk20 Desk22)} and \fw{(below Desk20 Desk22)}.
+SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)}.
 This is done in the following manner.
-First, the predicates of these expressions are placed in correspondence by the identicality constraint.
-Next, parallel connectivity aligns their arguments, \fw{Desk1} with \fw{Desk20} and \fw{Desk2} with \fw{Desk22}.
-While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint.
-In Figure \ref{fig:analogy}, these correspondences are highlighted by the purple hashed bi-directional arrows.
-Next, SME creates a candidate inference for the expression \fw{(regionBoundary CSDR1 Desk1)}, because \fw{Desk1} participates in the mapping.
-The candidate inference is stated as \fw{(regionBoundary (:skolem CSDR1) Desk20)}.
+First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME.
+Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11} and \fw{Desk2} with \fw{Desk12}.
+While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
+In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows.
+Next, SME creates a candidate inference for the region boundary expression, because \fw{Desk1} participates in the mapping.
+The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}.
 \fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences.
-In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes \fw{Desk20}.
-Note that inference is selective, with no candidate inferences generated for the entirely unmapped \fw{below} expression.
-Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.
+In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group.
+Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions.
+%Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.
 

### 2012-01-24 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
typos and tables

@@
 \begin{equation}
-	p=\frac{area(inferred region \cap target region)}{area(inferred region)}
+	p=\frac{area(inferred\ region \cap target\ region)}{area(inferred\ region)}
 \end{equation}
 \begin{equation}
-	r=\frac{area(inferred region \cap target region)}{area(target region)}
+	r=\frac{area(inferred\ region \cap target\ region)}{area(target\ region)}
 \end{equation}
@@
 \subsection{Results}
-\begin{table}
+\begin{table*}
 \small
@@
 
-\end{table}
+\end{table*}
 
@@
 Table~\ref{tab:overall} summarizes the results by combining each problem instance and user-defined target region.
-The mean precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
+Both precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
 When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.29) and recall of .66 ($\sigma$=.24).
@@
 
-\begin{table}
-\small
+% these are sensible results, nice
+
+\begin{table*}
 \begin{tabular}{|c|c|c|c|}
@@
   \label{tab:region}	
-\end{table}
+\end{table*}
 
-Table~\ref{tab:region} separates the results by the region type of the context-dependent spatial region.
+Table~\ref{tab:region} separates the results by CDSR type.
 The mean precision for the transferred regions ranged from .74 for the back rows to 0 for the living room of the studio apartment.
-Comparing the manually encoded against the target regions resulted in a minimum mean precision of .60, which occurred for the front of the classroom.
-The room precision, which is directly proportionally to the size of the target region, varied from .08 for the office to .35 for the living area.
+Comparing manually encoded against  target regions resulted in a minimum mean precision of .60.
+This  occurred for the front of the classroom.
+The whole room precision, which is directly proportionally to the size of the target region, varied from .08 for the office to .35 for the living area.
 
@@
 
-These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitated learning from a single example through analogical transfer.
+These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitate learning from a single example through analogical transfer.
 The relatively high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.

### 2012-01-24 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
typos and tables

@@
 
-\from{Nick}{TODO: Update Figure \ref{fig:analogy} with correct representation for current version.}
+\from{Nick}{Kate, please update figure caption, plus descriptive paragraph below}
 
@@
 
-% all images in a single figure -- pretty ugly and a waste of space
-% \begin{figure*}[t]
-%   \includegraphics[width=\textwidth]{images/worked-example.png}
-%   \caption{\textbf{TODO}}
-%   \label{fig:rooms-and-pic}
-% \end{figure*}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
-  \caption{Maps of 2 real classrooms generated by our system, plus an expert-annotated CDSR (orange, top image), subject-annotated CDSR (red, bottom image) and a CDSR transferred by analogy (green bottom image).
-The classroom used to generate the top classroom is pictured in Figure~\ref{fig:ug40}}
+	% can swap these 2 lines for different presentations - and colours!
+ \includegraphics[width=\columnwidth]{images/worked-example.png}
+  % \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
+  \caption{Maps of 2 real classrooms generated by our system, plus an expert-annotated CDSR (red, top image), subject-annotated CDSR (blue, bottom image) and a (smaller) CDSR transferred by analogy (green bottom image).
+The classroom used to generate the bottom classroom is pictured in Figure~\ref{fig:ug40}.}
   \label{fig:rooms}

### 2012-01-24 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
centering tables

@@
 
-To evaluate our progress toward building a cognitive system capable of reasoning about these CDSRs, we conducted the an experiment focusing on the following questions:
+To evaluate our progress toward building a cognitive system capable of reasoning about CDSRs, we conducted the an experiment focusing on the following questions:
 \begin{itemize}
-\item{How well do anchor points capture context-dependent spatial regions?}
+\item{Are anchor points able to encode context-dependent spatial regions?}
 \item{When provided with a base representation containing a labelled CDSR, how well does our approach identify the CDSR in a given target?}
-\item{Is the integration of geometric and semantic knowledge necessary?}
-\item{Does similarity correlate with performance?}
 \end{itemize}
-
-I'm not sure how many of the above questions we want to talk about.
-The last one is fairly easy and won't take up much space in the paper.
-The integration of knowledge question is could take some more work as the grouping is determined by semantic type, so in testing just positional knowledge,  I would have to create an alternate set of anchor points.
 
@@
 
-To determine how non-experts define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room.
-Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region.
-All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.
-The user-provided data provides the ground truth against which we evaluate our transfers.
+To determine how people define CDRSs, we asked three na\"ive users to draw polygons for each region type for each room.
+This task was performed using a webpage (link removed for blind review) shown in the inset in Figure~\ref{fig:ug40} on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
+The user-defined polygons define the \textit{target regions} against which we evaluate our transfers.
 
@@
 \begin{equation}
-	p=\frac{area(inferred region \cap target region)}{area(inferred region)}
+	p=\frac{area(inferred\ region \cap target\ region)}{area(inferred\ region)}
 \end{equation}
 \begin{equation}
-	r=\frac{area(inferred region \cap target region)}{area(target region)}
+	r=\frac{area(inferred\ region \cap target\ region)}{area(target\ region)}
 \end{equation}
 
-As a baseline, we compare the transfer results against our manually encoded CDSRs for the target and a region that represents the entire room.
-Results on the entire room provide a baseline performance measure.
-The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.
-
-\from{Nick}{I'm confused here.
-Are we comparing the inferred region against our manual region for the target room, or are we comparing our manual region for the target room against the user-provided region for the target room?}
-
-\from{Klenk}{I think that I can do away with all of the stuff about the union from each of the subjects.
-Especially since we are averaging them all together anyways.
-We can have a variation in which we control for everything else and vary the user.
-Perhaps even another table...}
-
-\from{Nick}{Which approach makes the explanation easier, union or averaging?
-The union makes our target the broadest acceptable region (not the average region though).
-Averaging shows how well we match an individual's conception of the region on average.
-Is one of these more relevant to our story than the other?}
+Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually annotated CDSR in the target and appropriate target region; and the region for the whole room and the target region.
+ Results comparing transferred and target regions measure how well the output of our system matches user expectations.
+Results comparing the manual annotations to the target regions measure how well the anchor point-based regions we chose match the users' regions (who were not constrained to anchor points).
+Results comparing whole rooms to target regions provide a baseline performance for comparison.
 
 \subsection{Results}
-\begin{table}
-\small
-\caption{Performance by Room}
+\begin{table*}
+	\center
+\begin{tabular}{|c|c|c|}
+\hline
+Inferred & Manually Encoded & Entire Room \\
+\hline
+$\bar{p}$=.45 $\sigma$=.37, $\bar{r}$=.45 $\sigma$=.38 & $\bar{p}$=.71 $\sigma$=.29, $\bar{r}$=.66 $\sigma$=.24  & $\bar{p}$=.17 $\sigma$=.11, $\bar{r}$=.98 $\sigma$=.05  \\
+\hline
+\end{tabular}
+\caption{Overall Performance}
+  \label{tab:overall}
+
+\end{table*}
+
+\from{klenk}{I guess we should rotate the above table. And it has been awhile since I wrote academic, so feel free to cut the boring text below}
+
+Table~\ref{tab:overall} summarizes the results by combining each problem instance and user-defined target region.
+Both precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
+When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.29) and recall of .66 ($\sigma$=.24).
+The region defined by the room results in a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
+
+% these are sensible results, nice
+
+\begin{table*}
+	\center
 \begin{tabular}{|c|c|c|c|}
 \hline
-Room & Transfer & Manually Encoded & Entire Room \\
+Region & Inferred & Manually Encoded & Entire Room \\
 \hline
-Classroom-1 & p=nil,r=nil & p=.93,r=.60 & p=.23,r=1 \\
-Classroom-2 & p=nil,r=nil & p=.90, r=.69 & p=.19,r=1 \\
-Classroom-3 & p=nil,r=nil & p=.94,r=.64 & p=.17,r=1 \\
-Classroom-4 & p=nil,r=nil & p=85,r=.89 & p=.19,r=1 \\  %This is 222-Classroom
-Classroom-5 & p=nil,r=nil & p=.87,r=.71 & p=.24,r=1 \\ % This is 222-Seminar
-Classroom-6 & p=nil,r=nil & p=.76,r=.66 & p=.20,r=.99 \\ % This is UG40
+Front & $\bar{p}$=.33 $\sigma$=.34, $\bar{r}$=.52 $\sigma$=.40 & $\bar{p}$=.60 $\sigma$=.29, $\bar{r}$=.82 $\sigma$=.19  & $\bar{p}$=.15 $\sigma$=.09, $\bar{r}$=1 $\sigma$=0  \\
+\hline
+Back & $\bar{p}$=.45 $\sigma$=.35, $\bar{r}$=.56 $\sigma$=.40 & $\bar{p}$=.66 $\sigma$=.25, $\bar{r}$=.84 $\sigma$=.17  & $\bar{p}$=.11 $\sigma$=.06, $\bar{r}$=.99 $\sigma$=.03  \\
+\hline
+Front Rows & $\bar{p}$=.62 $\sigma$=.36, $\bar{r}$=.22 $\sigma$=.21 & $\bar{p}$=.82 $\sigma$=.30, $\bar{r}$=.50 $\sigma$=.11  & $\bar{p}$=.22 $\sigma$=.08, $\bar{r}$=1 $\sigma$=0  \\
+\hline
+Back Rows & $\bar{p}$=.74 $\sigma$=.28, $\bar{r}$=.43 $\sigma$=.23 & $\bar{p}$=.80 $\sigma$=.29, $\bar{r}$=.43 $\sigma$=.26  & $\bar{p}$=.19 $\sigma$=.06, $\bar{r}$=1 $\sigma$=0  \\
+\hline
+Kitchen & $\bar{p}$=.23 $\sigma$=.03, $\bar{r}$=.09 $\sigma$=.02 & $\bar{p}$=.78 $\sigma$=.20, $\bar{r}$=.71 $\sigma$=.13  & $\bar{p}$=.16 $\sigma$=.02, $\bar{r}$=.92 $\sigma$=.13  \\
+\hline
+Office & $\bar{p}$=.05 $\sigma$=.01, $\bar{r}$=.06 $\sigma$=.06 & $\bar{p}$=.78 $\sigma$=.29, $\bar{r}$=.55 $\sigma$=.20  & $\bar{p}$=.08 $\sigma$=.03, $\bar{r}$=.93 $\sigma$=.06  \\
+\hline
+Living Room & $\bar{p}$=0 $\sigma$=0, $\bar{r}$=0 $\sigma$=0 & $\bar{p}$=.63 $\sigma$=.34, $\bar{r}$=.54 $\sigma$=.13  & $\bar{p}$=.35 $\sigma$=.22, $\bar{r}$=.96 $\sigma$=.06  \\
 \hline
 \end{tabular}
-\end{table}
+\caption{Performance by Region Type}
+  \label{tab:region}	
+\end{table*}
 
-\from{Nick}{Would this be just a single region type or averaging across all applicable ones?}
-\from{Klenk}{This is averaging across all applicable ones.}
+Table~\ref{tab:region} separates the results by CDSR type.
+The mean precision for the transferred regions ranged from .74 for the back rows to 0 for the living room of the studio apartment.
+Comparing manually encoded against  target regions resulted in a minimum mean precision of .60.
+This  occurred for the front of the classroom.
+The whole room precision, which is directly proportionally to the size of the target region, varied from .08 for the office to .35 for the living area.
 
-If we are just going to report means, we should standard deviations as well. %yup!
-
-\begin{table}
-\small
-\caption{Performance by Region Type}
-\begin{tabular}{|c|c|c|c|}
-\hline
-Region & Transfer & Manually Encoded & Entire Room \\
-\hline
-Front & p=nil,r=nil & p=.89,r=.79 & p=.23,r=1 \\
-Back & p=nil,r=nil & p=.85,r=.82 & p=.13,r=1 \\
-Front Rows & p=nil,r=nil & p=.85,r=.48 & p=.22,r=1 \\
-Back Rows & p=nil,r=nil & p=.94,r=.40 & p=.7,r=.9 \\
-Kitchen & p=nil,r=nil & p=nil,r=nil & p=nil,r=nil \\
-Office & p=nil,r=nil & p=nil,r=nil & p=nil,r=nil \\
-\hline
-\end{tabular}
-\end{table}
-
-Or we could report all of the results in a single entry.
-
-\begin{table}
-\small
-\caption{Performance against Target Region}
-\begin{tabular}{|c|c|c|}
-\hline
-Transfer & Manually Encoded & Entire Room \\
-\hline
-p=nil,r=nil & p=.88,r=.66 & p=.20,r=.99 \\
-\hline
-\end{tabular}
-\end{table}
-
-% this would be lovely!
-
-Overall the transferred region was significantly better than the entire room (p<Blah), and not statistically different than the manually encoded region (p=Blah).
-There was a positive correlation (r=Blah) between the structural evaluation score of the analogy and f-measure of the inference.
+%(3 :transfer-p (0.5649091160864536d0 . 0.3824746399624735d0)  :transfer-r (0.4264200799557943d0 . 0.34846457388485863d0) :self-p   (0.8290901572528112d0 . 0.19630847005535487d0) :self-r
+%  (0.6113880445364026d0 . 0.22432810023132582d0) :room-p
+%  (0.20903207305733315d0 . 0.10291204447091876d0) :room-r
+%  (0.9960679285191474d0 . 0.011022662977174953d0))
+% (2 :transfer-p (0.4496053675639215d0 . 0.36845133415546233d0)
+%  :transfer-r (0.4174204919721308d0 . 0.3697742376406773d0) :self-p
+%  (0.7556540963495391d0 . 0.2672139547143468d0) :self-r
+%  (0.6252647119318083d0 . 0.23778540411202373d0) :room-p
+%  (0.17891943994602752d0 . 0.11273973943366206d0) :room-r
+%  (0.95742701954372d0 . 0.08432884603590181d0))
+% (1 :transfer-p (0.3360804531954755d0 . 0.3364697224078454d0)
+%  :transfer-r (0.5074731366878895d0 . 0.4231938085019586d0) :self-p
+%  (0.5493333659045643d0 . 0.33190756014975015d0) :self-r
+%  (0.7596761147032691d0 . 0.25387231566349d0) :room-p
+%  (0.12277614149421091d0 . 0.10953121808467986d0) :room-r
+%  (0.9946881982552835d0 . 0.011992196508589958d0))
 
@@
 
-Our results
+These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitate learning from a single example through analogical transfer.
+The relatively high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
+The lower scores typically resulted from target regions which were farther away from objects.
+While the inferred regions did not score as well as the manually encoded regions, they were considerably more precise than the room regions.
+The only case were transfer was not possible was between the two studio's living room cases.
+The objects which define the living area in one studio (sofa, TV, and bookshelves) are significantly different than the objects in the other (desk, bed, and sofa), as are the studios in general.

### 2012-01-24 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Kate: updating the image/text for the analogy again

@@
 \begin{figure}[h]
-  \includegraphics[width=\columnwidth]{analogy2.png}
-  \caption{Analogical mapping between four base expressions and two target expressions.
-Ovals and squares represent relations and entities respectively.
-Purple dashed bi-directional arrows are the correspondences between the base and target, and the green dashed arrow, oval, and square represent a candidate inference projected into the target.}
+  \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
+  \caption{Analogical mapping between five base expressions and three target expressions. }
+  %Ovals and squares represent relations and entities respectively.
+Black dashed bi-directional arrows are the correspondences between the base and target, and the red dashed objects represent a candidate inference projected into the target.
+
   \label{fig:analogy}
@@
 
-\from{Nick}{TODO: Update Figure \ref{fig:analogy} with correct representation for current version.}
-
-Figure \ref{fig:analogy} illustrates a sample mapping between four base expressions and two target ones.
+Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and three target ones.
 Each oval represents a binary predicate, and its two entity arguments are represented by squares.
-SME generates a mapping between the base expressions \fw{(rcc8-DC Desk1 Desk2)} and \fw{(below Desk1 Desk2)}, and the  target expressions \fw{(rcc8-DC Desk20 Desk22)} and \fw{(below Desk20 Desk22)}.
+SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case .
 This is done in the following manner.
-First, the predicates of these expressions are placed in correspondence by the identicality constraint.
-Next, parallel connectivity aligns their arguments, \fw{Desk1} with \fw{Desk20} and \fw{Desk2} with \fw{Desk22}.
-While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint.
-In Figure \ref{fig:analogy}, these correspondences are highlighted by the purple hashed bi-directional arrows.
-Next, SME creates a candidate inference for the expression \fw{(regionBoundary CSDR1 Desk1)}, because \fw{Desk1} participates in the mapping.
-The candidate inference is stated as \fw{(regionBoundary (:skolem CSDR1) Desk20)}.
+First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME.
+Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
+While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
+In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows.
+Next, SME creates a candidate inference for the region boundary expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping.
+The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}.
 \fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences.
-In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes \fw{Desk20}.
-Note that inference is selective, with no candidate inferences generated for the entirely unmapped \fw{below} expression.
-Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.
+In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group.
+Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions.
+%Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.
 

### 2012-01-24 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

@@
 
-\from{klenk}{I guess we should rotate the above table. And it has been awhile since I wrote academic, so feel free to cut the boring text below}
-
-Table~\ref{tab:overall} summarizes the results by combining each problem instance and user-defined target region.
+To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users.
 Both precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
@@
 The region defined by the room results in a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
-
-% these are sensible results, nice
 
@@
 
-Table~\ref{tab:region} separates the results by CDSR type.
-The mean precision for the transferred regions ranged from .74 for the back rows to 0 for the living room of the studio apartment.
+\note{Klenk}{UPDATE TABLE and specific numbers in these sections.}
+
+To identify how our approach fairs under different conditions, Table~\ref{tab:region} separates the results by CDSR type.
+The mean precision for the transferred regions ranged from .74 for the back rows of classrooms to 0 for the office in studio apartments.
 Comparing manually encoded against  target regions resulted in a minimum mean precision of .60.
@@
 
-%(3 :transfer-p (0.5649091160864536d0 . 0.3824746399624735d0)  :transfer-r (0.4264200799557943d0 . 0.34846457388485863d0) :self-p   (0.8290901572528112d0 . 0.19630847005535487d0) :self-r
-%  (0.6113880445364026d0 . 0.22432810023132582d0) :room-p
-%  (0.20903207305733315d0 . 0.10291204447091876d0) :room-r
-%  (0.9960679285191474d0 . 0.011022662977174953d0))
-% (2 :transfer-p (0.4496053675639215d0 . 0.36845133415546233d0)
-%  :transfer-r (0.4174204919721308d0 . 0.3697742376406773d0) :self-p
-%  (0.7556540963495391d0 . 0.2672139547143468d0) :self-r
-%  (0.6252647119318083d0 . 0.23778540411202373d0) :room-p
-%  (0.17891943994602752d0 . 0.11273973943366206d0) :room-r
-%  (0.95742701954372d0 . 0.08432884603590181d0))
-% (1 :transfer-p (0.3360804531954755d0 . 0.3364697224078454d0)
-%  :transfer-r (0.5074731366878895d0 . 0.4231938085019586d0) :self-p
-%  (0.5493333659045643d0 . 0.33190756014975015d0) :self-r
-%  (0.7596761147032691d0 . 0.25387231566349d0) :room-p
-%  (0.12277614149421091d0 . 0.10953121808467986d0) :room-r
-%  (0.9946881982552835d0 . 0.011992196508589958d0))
-
 \subsection{Discussion}
@@
 These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitate learning from a single example through analogical transfer.
-The relatively high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
-The lower scores typically resulted from target regions which were farther away from objects.
-While the inferred regions did not score as well as the manually encoded regions, they were considerably more precise than the room regions.
-The only case were transfer was not possible was between the two studio's living room cases.
-The objects which define the living area in one studio (sofa, TV, and bookshelves) are significantly different than the objects in the other (desk, bed, and sofa), as are the studios in general.
+Collaboration with human users requires a high precision and recall, because cognitive systems must be able to understand as well as refer to these regions in human environments.
+Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
+Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within a tasks with human users.
+
+The transferred regions were considerably more precise (.47) when compared to the room as whole (.17), and their recalls (.45) indicate that they captured almost half of the area indicated by the human user.
+Our approach performed the best when the extent of the context dependent spatial region is closely tied to objects as is the case in the front rows of the classroom with p's of .76 and .82 for the inferred and the manually encoded regions respectively.
+The system performs worst when the extent of the CDSR is defined as an unbounded area near or around particular objects.
+The office of a studio apartment is loosely defined as the region around the desk.
+This motivates one direction of future work to expand the vocabulary of anchor points to better capture these notions of space.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

@@
 \begin{table*}
-\small
+	\center
 \begin{tabular}{|c|c|c|}
@@
 \hline
-$\bar{p}$=.45 $\sigma$=.37, $\bar{r}$=.45 $\sigma$=.38 & $\bar{p}$=.71 $\sigma$=.29, $\bar{r}$=.66 $\sigma$=.24  & $\bar{p}$=.17 $\sigma$=.11, $\bar{r}$=.98 $\sigma$=.05  \\
+$\bar{p}$=.47 $\sigma$=.37, $\bar{r}$=.46 $\sigma$=.38 & $\bar{p}$=.71 $\sigma$=.30, $\bar{r}$=.67 $\sigma$=.25 & $\bar{p}$=.17 $\sigma$=.11, $\bar{r}$=.98 $\sigma$=.05  \\
 \hline
@@
 
-\from{klenk}{I guess we should rotate the above table. And it has been awhile since I wrote academic, so feel free to cut the boring text below}
-
-Table~\ref{tab:overall} summarizes the results by combining each problem instance and user-defined target region.
-Both precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
-When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.29) and recall of .66 ($\sigma$=.24).
+To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users.
+Both precision and recall for the inferred regions is .47 and .37 with a standard deviation of .37 and .38 respectively.
+When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25).
 The region defined by the room results in a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
 
-% these are sensible results, nice
-
 \begin{table*}
+	\center
 \begin{tabular}{|c|c|c|c|}
@@
 \hline
-Front & $\bar{p}$=.33 $\sigma$=.34, $\bar{r}$=.52 $\sigma$=.40 & $\bar{p}$=.60 $\sigma$=.29, $\bar{r}$=.82 $\sigma$=.19  & $\bar{p}$=.15 $\sigma$=.09, $\bar{r}$=1 $\sigma$=0  \\
+Front & $\bar{p}$=.32 $\sigma$=.33, $\bar{r}$=.49 $\sigma$=.41 & $\bar{p}$=.60 $\sigma$=.29, $\bar{r}$=.83 $\sigma$=.19  & $\bar{p}$=.16 $\sigma$=.10, $\bar{r}$=1 $\sigma$=0  \\
 \hline
-Back & $\bar{p}$=.45 $\sigma$=.35, $\bar{r}$=.56 $\sigma$=.40 & $\bar{p}$=.66 $\sigma$=.25, $\bar{r}$=.84 $\sigma$=.17  & $\bar{p}$=.11 $\sigma$=.06, $\bar{r}$=.99 $\sigma$=.03  \\
+Back & $\bar{p}$=.44 $\sigma$=.37, $\bar{r}$=.56 $\sigma$=.41 & $\bar{p}$=.66 $\sigma$=.25, $\bar{r}$=.84 $\sigma$=.17  & $\bar{p}$=.11 $\sigma$=.06, $\bar{r}$=.99 $\sigma$=.03  \\
 \hline
-Front Rows & $\bar{p}$=.62 $\sigma$=.36, $\bar{r}$=.22 $\sigma$=.21 & $\bar{p}$=.82 $\sigma$=.30, $\bar{r}$=.50 $\sigma$=.11  & $\bar{p}$=.22 $\sigma$=.08, $\bar{r}$=1 $\sigma$=0  \\
+Front Rows & $\bar{p}$=.76 $\sigma$=.27, $\bar{r}$=.28 $\sigma$=.21 & $\bar{p}$=.83 $\sigma$=.31, $\bar{r}$=.50 $\sigma$=.11  & $\bar{p}$=.22 $\sigma$=.08, $\bar{r}$=1 $\sigma$=0  \\
 \hline
-Back Rows & $\bar{p}$=.74 $\sigma$=.28, $\bar{r}$=.43 $\sigma$=.23 & $\bar{p}$=.80 $\sigma$=.29, $\bar{r}$=.43 $\sigma$=.26  & $\bar{p}$=.19 $\sigma$=.06, $\bar{r}$=1 $\sigma$=0  \\
+Back Rows & $\bar{p}$=.72 $\sigma$=.30, $\bar{r}$=.42 $\sigma$=.26 & $\bar{p}$=.80 $\sigma$=.29, $\bar{r}$=.43 $\sigma$=.26  & $\bar{p}$=.19 $\sigma$=.06, $\bar{r}$=1 $\sigma$=0  \\
 \hline
-Kitchen & $\bar{p}$=.23 $\sigma$=.03, $\bar{r}$=.09 $\sigma$=.02 & $\bar{p}$=.78 $\sigma$=.20, $\bar{r}$=.71 $\sigma$=.13  & $\bar{p}$=.16 $\sigma$=.02, $\bar{r}$=.92 $\sigma$=.13  \\
+Kitchen & $\bar{p}$=.60 $\sigma$=.05, $\bar{r}$=.59 $\sigma$=.34 & $\bar{p}$=.78 $\sigma$=.20, $\bar{r}$=.71 $\sigma$=.13  & $\bar{p}$=.16 $\sigma$=.02, $\bar{r}$=.92 $\sigma$=.13  \\
 \hline
-Office & $\bar{p}$=.05 $\sigma$=.01, $\bar{r}$=.06 $\sigma$=.06 & $\bar{p}$=.78 $\sigma$=.29, $\bar{r}$=.55 $\sigma$=.20  & $\bar{p}$=.08 $\sigma$=.03, $\bar{r}$=.93 $\sigma$=.06  \\
+Office & $\bar{p}$=.00 $\sigma$=.00, $\bar{r}$=.00 $\sigma$=.00 & $\bar{p}$=.78 $\sigma$=.29, $\bar{r}$=.55 $\sigma$=.20  & $\bar{p}$=.08 $\sigma$=.03, $\bar{r}$=.94 $\sigma$=.06  \\
 \hline
-Living Room & $\bar{p}$=0 $\sigma$=0, $\bar{r}$=0 $\sigma$=0 & $\bar{p}$=.63 $\sigma$=.34, $\bar{r}$=.54 $\sigma$=.13  & $\bar{p}$=.35 $\sigma$=.22, $\bar{r}$=.96 $\sigma$=.06  \\
+Living Room & $\bar{p}$=.40 $\sigma$=.39, $\bar{r}$=.01 $\sigma$=.01 & $\bar{p}$=.63 $\sigma$=.34, $\bar{r}$=.54 $\sigma$=.13  & $\bar{p}$=.35 $\sigma$=.22, $\bar{r}$=.96 $\sigma$=.06  \\
 \hline

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

@@
 To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users.
-Both precision and recall for the inferred regions is .47 and .37 with a standard deviation of .37 and .38 respectively.
-When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25).
-The region defined by the room results in a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
+The inferred regions achieved a precision of .47 ($\sigma$=.37) and a recall of .46 ($\sigma$=.38).
+Comparing the manually encoded regions against each target region results in a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25).
+The region defined by the room corresponds to the target region with a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
 
@@
 
-\note{Klenk}{UPDATE TABLE and specific numbers in these sections.}
-
 To identify how our approach fairs under different conditions, Table~\ref{tab:region} separates the results by CDSR type.
-The mean precision for the transferred regions ranged from .74 for the back rows of classrooms to 0 for the office in studio apartments.
+The mean precision for the transferred regions ranged from .76 for the front rows of classrooms to 0 for the office in studio apartments.
 Comparing manually encoded against target regions resulted in a minimum mean precision of .60.
@@
 
-The transferred regions were considerably more precise (.47) when compared to the room as whole (.17), and their recalls (.45) indicate that they captured almost half of the area indicated by the human user.
+The transferred regions were considerably more precise (.47) when compared to the room as whole (.17), and their recalls (.46) indicate that they captured almost half of the area indicated by the human user.
 Our approach performed the best when the extent of the context dependent spatial region is closely tied to objects as is the case in the front rows of the classroom with p's of .76 and .82 for the inferred and the manually encoded regions respectively.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
changed predicate names

@@
 
-For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+For each object that Dora detects we compute the strengths of 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+The relation's strength for a given pair of objects represents the applicability of the relation to the pair of objects ranging from 0 to 1 with 0 being unsuitable.
 The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}.
@@
 The predicates used to denote these relations are named accordingly, e.g.
-\textit{xZeroYOne}, \textit{xOneYOne}, \textit{xOneYZero}, \textit{xOneYNegOne}, etc.
+\textit{xZeroYPlus}, \textit{xPlusYPlus}, \textit{xPlusYZero}, \textit{xPlusYMinus}, etc.
 
@@
 This provides our representations with the flexibility necessary to manage the variation in perceptual information (i.e.
-the position of walls and objects) inevitable on a robot.
+the position of walls and objects) inevitable in human environments.
+
+The last piece of symbolic representation extracted from the sensor data is grouping entities.
+We create a group entity anytime there is a set of adjacent objects of the same type.
+For example, a classroom would likely have a group entity created in which all of the students' desks were members.
 

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

@@
 Consider, for example, the aforementioned region \textit{the front of the classroom}.
-This region is not easily perceivable using just the geometry of the environment.
+This region is not perceivable using just the geometry of the environment.
 Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard).
@@
 Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability.
-If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to.
+If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do.
 Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.
@@
 
+\from{klenk}{Cite recent Langley work}
+
 This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs.
 Our approach is founded on two assumptions.
-The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
-The second assumption is that semantically similar areas (e.g.
+The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) corresponding to sensor data of the system~\cite{Cohn:2001}.
+The second assumption is that semantically and geometrically similar areas (e.g.
 two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}.
@@
 what can be done there and where.
-In this work we implicitly represent context using the types of objects present in a room and their location, both relative to each other and relative to the room itself.
+In this work we implicitly represent context using the types of objects present in a room and their location relative to each other.
 The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.
@@
 When an object is detected it is represented in the metric map
-by placing a copy of the model at the detected pose. The recognizer associates each object with a type label that was provided during a training phase.
+by placing a copy of the model at the detected pose. The recognizer associates each object with a semantic type that was provided during a training phase.
 
@@
 We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs.
-Anchor points are symbolic descriptions which link a conceptual entity (i.e.
-something that can be reasoned about but not directly perceived, such as a CDSR) to a particular point on a perceived entity.
+Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity.
 The perceived entities we use are the objects recognised by Dora, and the room itself.
@@
 Anchor points are created from perceived entities using unary functions, e.g.
-\fw{(CentroidFn Desk1)} represents the centroid of the geometric extent of \fw{Desk1}, and are linked to a particular CDSR using a \fw{regionBoundary} relation.
+\fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate.
+Anchor points are linked to particular CDSRs using a \fw{regionBoundary} relation.
 After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation.
-Therefore, each CDSR has one type and a variable number of boundary points.
-
-\from{Nick}{TODO: Need to update example, including new fns, when we have updated pictures in.}
+Therefore, each CDSR has one type and a variable number of boundary segments.
 
@@
 
-\fw{(regionType CDSR1 FrontRegion)\\
-(regionBoundary CDSR1 (LeftmostTopFn Room1))\\
-(regionBoundary CDSR1 (LeftmostBottomFn Room1))\\
-(regionBoundary CDSR1 Desk1)\\
-(regionBoundary CDSR1 Desk2)\\
-(regionBoundary CDSR1 Desk3)\\
-(regionBoundary CDSR1 Desk4)\\
-(regionBoundary CDSR1 Whiteboard0)
-}}
-  \caption{Representing the front of the classroom context-dependent region \fw{CDSR1}}
+\fw{
+(regionType CDSR9 FrontRegion) \\
+(regionBoundarySegment CDSR9 \\
+\hspace*{2em}(YMaxXFewestFn Room-Base) \\
+\hspace*{2em}(YMinXFewestFn Room-Base)) \\
+(regionBoundarySegment CDSR9 \\
+\hspace*{2em}(YMinXFewestFn Room-Base) \\
+\hspace*{2em}(YMinXFewestFn Group1))\\
+...
+}
+}
+  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
   \label{fig:cdsr-reps}
@@
   \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
-  \caption{Analogical mapping between five base expressions and two target expressions. }
+  \caption{Analogical mapping between five base expressions and three target expressions. }
   %Ovals and squares represent relations and entities respectively.
@@
 
-\from{Nick}{Kate, please update figure caption, plus descriptive paragraph below}
-
-Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and two target ones.
+Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and three target ones.
 Each oval represents a binary predicate, and its two entity arguments are represented by squares.
-SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)}.
+SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case .
 This is done in the following manner.
 First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME.
-Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11} and \fw{Desk2} with \fw{Desk12}.
-While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
+Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
+While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
 In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows.
-Next, SME creates a candidate inference for the region boundary expression, because \fw{Desk1} participates in the mapping.
+Next, SME creates a candidate inference for the region boundary expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping.
 The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}.
@@
 % 
-% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams.
+% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams, Scott Friedman for assistance on probablistic SME, and Danny Bobrow comments on an earlier draft of this paper.
 The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Kate: Added Langly ref, Redid Analogy Diagram.  Minor tweaking of conclusion

@@
 Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability.
-If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to.
+If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do.
 Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.
 In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.
+Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm \cite{langley:inpress}.
 
@@
 Our approach is founded on two assumptions.
-The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.
-The second assumption is that semantically similar areas (e.g.
+The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) corresponding to sensor data of the system~\cite{Cohn:2001}.
+The second assumption is that semantically and geometrically similar areas (e.g.
 two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}.
@@
 This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}.
-The need for an autonomous system to ground references to human-generated descriptions of space has been recognised in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.
+The need for an autonomous system to ground references to human-generated descriptions of space has been recognized in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.
 
@@
 In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions.
-The system is able to use analogy to recognise CDSRs in previously unseen environments through a process of analogy with a manually annotated known room.
+The system identifies CDSRs in previously unseen environments through analogy with a manually annotated known room.
 The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy.
-In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with in future work.
+In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work.
 These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

@@
   \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
-  \caption{Analogical mapping between five base expressions and three target expressions. }
+  \caption{Analogical mapping between six base expressions and three target expressions. }
   %Ovals and squares represent relations and entities respectively.
@@
 
-Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and three target ones.
+Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones.
 Each oval represents a binary predicate, and its two entity arguments are represented by squares.

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

@@
 In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.
-
-\from{klenk}{Cite recent Langley work}
+Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm \cite{langley:inpress}.
 
@@
 \fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate.
-Anchor points are linked to particular CDSRs using a \fw{regionBoundary} relation.
+Anchor points are linked to particular CDSRs using a \fw{boundarySegment} ternary relation.
 After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation.
@@
 (regionType CDSR9 FrontRegion) \\
-(regionBoundarySegment CDSR9 \\
-\hspace*{2em}(YMaxXFewestFn Room-Base) \\
-\hspace*{2em}(YMinXFewestFn Room-Base)) \\
-(regionBoundarySegment CDSR9 \\
-\hspace*{2em}(YMinXFewestFn Room-Base) \\
+(boundarySegment CDSR9 \\
+\hspace*{2em}(YMaxXFewestFn Room3) \\
+\hspace*{2em}(YMinXFewestFn Room3)) \\
+(boundarySegment CDSR9 \\
+\hspace*{2em}(YMinXFewestFn Room3) \\
 \hspace*{2em}(YMinXFewestFn Group1))\\
@@
 
-Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top of Figure~\ref{fig:rooms}.
-The anchor points (connected in orange in Figure~\ref{fig:rooms}) define the region.
-\fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively.
-The sets of points representing the top and bottom of the classroom are found by selecting the points defining the area within a threshold of the area's maximum or minimum y coordinate.
-The other corners of the context-dependent spatial region are defined by the four desks and the whiteboard.
-Because perceived objects (e.g.
-the desks) are defined as single points, they can serve as anchor points directly.
-The semantic label \fw{FrontRegion} ties the geometric extent of the boundary points to a conceptual region.
-This definition for the front of the classroom is specific to \fw{Room1} and its entities.
+Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}.
+The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region.
+\fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}.
+The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks.
+There are two more boundary segments completing a polygon for this region.
+The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room".
+This definition for the front of the room is specific to \fw{Room3} and its entities.
 It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary.
 If the desks were in a different position then the region would cover a different extent (e.g.
-if they were closer to the whiteboard then the region would be smaller).
+if they were further to the left then the region would be smaller).
 
@@
 
-% \from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
-% done
-
 We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding.
-To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training.
+To avoid burdening potential users with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training.
 For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
@@
 Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures.
+Mappings are defined by expressions with an identical relation and corresponding arguments.
+When provided with expression strengths, such as, our spatial relationships, SME prefers mappings with closely aligned fact strengths.
+SME can be given \textit{pragmatic constraints} that require certain entities in the base to be included in the mapping.
 Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences.
@@
 
-In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, plus the QSRs computed on top of these.
-The base also contains a labeled CDSR which we assume is sought in target.
-The result of result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target.
+In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, the QSRs between them and any groups that have been identified.
+The base also contains a labeled CDSR of the type sought in target.
+The result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target.
 We use these to transfer the CDSR from base to target (i.e.
 recognizing the CDSR in the target) as follows.
-First, we identify the CDSR of the sought type in the base.
-This entity should not participate in the mapping, but should be included in \fw{:skolem} expressions in the candidate inferences.
-To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{regionBoundary} statements mentioning the base CDSR.
-The second arguments of these candidate inferences are anchor points in the target environment.
-We use these to define the boundary of the CDSR in the target, linking them to a new CDSR entity using \fw{regionBoundary} statements.
+First, we identify the CDSR of the sought type in the base and use SME's pragmatic constraints to ensure that the entities referred to its anchor points participate in the mapping.
+To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{boundarySegment} statements mentioning the base CDSR.
+The second and third arguments of these candidate inferences are anchor points in the target environment.
+We use these to define the boundary of the CDSR in the target.
 
@@
 The map and object data that results from this is then passed on to the QSR generator.
-In the base case Dora perceives 2 desks, a whiteboard and the room area.
+In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area.
 To this we add the CDSR representing the front of the room.
-The case includes a total of \textbf{SOMENUMBER} expressions relating the \textbf{SOMENUMBER} entities.
-\textbf{SOMENUMBER} of these expressions are used to define the boundary of a CDSR representing the front of the room: the \textbf{SOMENUMBER} desks in the front row, the whiteboard, and the leftmost bottom and leftmost top of the room.
-The target case includes \textbf{SOMENUMBER}.
-The target case includes \textbf{SOMENUMBER} statements and \textbf{SOMENUMBER} entities.
+The case includes a total of 50 expression relating the 20 entities.
+Six of these expressions are used to define the boundary segments and CDSR representing the front of the room.
+The target case includes 26 expressions and 11 entities.
 
-SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation.
-The resulting analogy includes \textbf{SOMENUMBER} correspondences between the entities and expressions in the base and target and \textbf{SOMENUMBER} candidate inferences.
-All \textbf{SOMENUMBER} \fw{regionBoundary} statements in the base appear as candidate inferences.
-\textbf{SOMENUMBER} of the anchor points are exactly correct, the \textbf{\dots}.
-The other \textbf{SOMENUMBER} analogically transferred anchor points are to the \textbf{\dots}.
-The green region in the lower image of Figure \ref{fig:rooms} illustrates the CDSR created by the \textbf{SOMENUMBER} boundary points in the target environment.
+SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation requiring \fw{Room3} and \fw{Group1} participate in the mapping as they are referenced by the anchor points in the base.
+The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences.
+Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target.
+The green region in the lower image of Figure \ref{fig:rooms} illustrates the inferred CDSR.
 

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Updated analogy section

@@
 For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
-We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
+We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic and geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
 
@@
 
-Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and three target ones.
-Each oval represents a binary predicate, and its two entity arguments are represented by squares.
+Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones.
+Each oval represents a predicate, and the entity arguments are represented by squares.
 SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case .
@@
 First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME.
-Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
+Then SME will align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
 While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
 In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows.
-Next, SME creates a candidate inference for the region boundary expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping.
-The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}.
-\fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences.
-In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group.
+Next, SME creates a candidate inference for the boundary segment expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping.
+The candidate inference is shown in red in the figure.
 Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions.

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
Kate: fixed typos

@@
 To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users.
-Both precision and recall for the inferred regions is .45 with a standard deviation of .37 and .38 respectively.
-When comparing the manually encoded regions against each target instance, we calculate a mean precision of .71 ($\sigma$=.29) and recall of .66 ($\sigma$=.24).
-The region defined by the room results in a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
+The inferred regions achieved a precision of .47 ($\sigma$=.37) and a recall of .46 ($\sigma$=.38).
+Comparing the manually encoded regions against each target region results in a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25).
+The region defined by the room corresponds to the target region with a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).
 
@@
 Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
-Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within a tasks with human users.
+Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within tasks with human users.
 
-The transferred regions were considerably more precise (.47) when compared to the room as whole (.17), and their recalls (.45) indicate that they captured almost half of the area indicated by the human user.
+The transferred regions were considerably more precise (.47) when compared to the room as whole (.17), and their recalls (.46) indicate that they captured almost half of the area indicated by the human user.
 Our approach performed the best when the extent of the context dependent spatial region is closely tied to objects as is the case in the front rows of the classroom with p's of .76 and .82 for the inferred and the manually encoded regions respectively.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

@@
 For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
-For the classrooms these were the front and back, and the front and back rows of desk.
+For the classrooms these were the front and back, and the front and back rows of desks.
 For the studios these were the kitchen, office and living areas.
@@
 We consider a \textit{problem instance} to be a room and a sought CDSR type.
-For each room for which we have a human-labelled target region of the sought type, we generate an \textit{inferred region} using analogical transfer.
-To assess the quality of the transfer, we calculate precision ($p$, the proportion of the inferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the inferred region) as follows:
+For each room containing a manually encoded CDSR of the sought type, we generate an \textit{transferred region} using analogical transfer.
+To assess the quality of the transfer, we calculate precision ($p$, the proportion of the transferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the transferred region) as follows:
 
 \begin{equation}
-	p=\frac{area(inferred\ region \cap target\ region)}{area(inferred\ region)}
+	p=\frac{area(transferred\ region \cap target\ region)}{area(transferred\ region)}
 \end{equation}
 \begin{equation}
-	r=\frac{area(inferred\ region \cap target\ region)}{area(target\ region)}
+	r=\frac{area(transferred\ region \cap target\ region)}{area(transferred\ region)}
 \end{equation}
 
-Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually annotated CDSR in the target and appropriate target region; and the region for the whole room and the target region.
- Results comparing transferred and target regions measure how well the output of our system matches user expectations.
-Results comparing the manual annotations to the target regions measure how well the anchor point-based regions we chose match the users' regions (who were not constrained to anchor points).
-Results comparing whole rooms to target regions provide a baseline performance for comparison.
+Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually encoded CDSR and target region; and the region for the whole room and the target region.
+Results comparing transferred and target regions measure how well our system is able apply a single example to new situations.
+The comparisons between the manual annotations to the target regions measure how well the anchor points we chose capture the users' regions (who were not constrained to anchor points).
+Results from the whole room regions provide a baseline performance for comparison.
 
@@
 \hline
-Inferred & Manually Encoded & Entire Room \\
+Transferred & Manually Encoded & Entire Room \\
 \hline
@@
 To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users.
-The inferred regions achieved a precision of .47 ($\sigma$=.37) and a recall of .46 ($\sigma$=.38).
+The transferred regions achieved a precision of .47 ($\sigma$=.37) and a recall of .46 ($\sigma$=.38).
 Comparing the manually encoded regions against each target region results in a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25).
@@
 Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
-Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within a tasks with human users.
+Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within tasks with human users.
 

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

@@
 
-\fw{
-(regionType CDSR9 FrontRegion) \\
-(regionBoundarySegment CDSR9 \\
-\hspace*{2em}(YMaxXFewestFn Room-Base) \\
-\hspace*{2em}(YMinXFewestFn Room-Base)) \\
-(regionBoundarySegment CDSR9 \\
-\hspace*{2em}(YMinXFewestFn Room-Base) \\
-\hspace*{2em}(YMinXFewestFn Group1))\\
-...
+\fw{(regionType CDSR9 FrontRegion) \\
+(boundarySegment CDSR9 \\
+\hspace*{2em}(YMaxXFewestFn Room3) \\
+\hspace*{2em}(YMinXFewestFn Room3)) \\
+(boundarySegment CDSR9 \\
+\hspace*{2em}(YMinXFewestFn Room3) \\
+\hspace*{2em}(YMinXFewestFn Group1))
 }
 }
-  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
+  \caption{Three of the five expressions representing the front of the classroom context-dependent region \fw{CDSR9}}
   \label{fig:cdsr-reps}
@@
 
-Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top of Figure~\ref{fig:rooms}.
-The anchor points (connected in orange in Figure~\ref{fig:rooms}) define the region.
-\fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively.
-The sets of points representing the top and bottom of the classroom are found by selecting the points defining the area within a threshold of the area's maximum or minimum y coordinate.
-The other corners of the context-dependent spatial region are defined by the four desks and the whiteboard.
-Because perceived objects (e.g.
-the desks) are defined as single points, they can serve as anchor points directly.
-The semantic label \fw{FrontRegion} ties the geometric extent of the boundary points to a conceptual region.
-This definition for the front of the classroom is specific to \fw{Room1} and its entities.
+Figure \ref{fig:cdsr-reps} contains three of the five expressions defining the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}.
+The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region.
+\fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}.
+The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks.
+There are two more boundary segments completing a polygon for this region.
+The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room".
+This definition for the front of the room is specific to \fw{Room3} and its entities.
 It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary.
 If the desks were in a different position then the region would cover a different extent (e.g.
-if they were closer to the whiteboard then the region would be smaller).
+if they were further to the left then the region would be smaller).
 
@@
 
-% \from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
-% done
-
 We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding.
-To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training.
+To avoid burdening potential users with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training.
 For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
-We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
+We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic and geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
 
@@
 Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones.
-Each oval represents a binary predicate, and its two entity arguments are represented by squares.
-SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case .
-This is done in the following manner.
+Each oval represents a predicate, and the entity arguments are represented by squares.
+SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case in the following manner.
 First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME.
-Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
+Then SME aligns the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
 While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
 In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows.
-Next, SME creates a candidate inference for the region boundary expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping.
-The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}.
-\fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences.
-In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group.
+Next, SME creates a candidate inference for the boundary segment expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping.
+The candidate inference is shown in red in the figure.
 Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions.
@@
 Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run.
-The map and object data that results from this is then passed on to the QSR generator.
-In the base case Dora perceives 2 desks, a whiteboard and the room area.
+The map and object data that result from this are then passed on to the QSR generator.
+In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area.
 To this we add the CDSR representing the front of the room.
-The case includes a total of \textbf{SOMENUMBER} expressions relating the \textbf{SOMENUMBER} entities.
-\textbf{SOMENUMBER} of these expressions are used to define the boundary of a CDSR representing the front of the room: the \textbf{SOMENUMBER} desks in the front row, the whiteboard, and the leftmost bottom and leftmost top of the room.
-The target case includes \textbf{SOMENUMBER}.
-The target case includes \textbf{SOMENUMBER} statements and \textbf{SOMENUMBER} entities.
+The case includes a total of 50 expression relating the 20 entities.
+Six of these expressions are used to define the boundary segments and CDSR representing the front of the room.
+The target case includes 26 expressions and 11 entities.
 
-SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation.
-The resulting analogy includes \textbf{SOMENUMBER} correspondences between the entities and expressions in the base and target and \textbf{SOMENUMBER} candidate inferences.
-All \textbf{SOMENUMBER} \fw{regionBoundary} statements in the base appear as candidate inferences.
-\textbf{SOMENUMBER} of the anchor points are exactly correct, the \textbf{\dots}.
-The other \textbf{SOMENUMBER} analogically transferred anchor points are to the \textbf{\dots}.
-The green region in the lower image of Figure \ref{fig:rooms} illustrates the CDSR created by the \textbf{SOMENUMBER} boundary points in the target environment.
+SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation requiring \fw{Room3} and \fw{Group1} participate in the mapping as they are referenced by the anchor points in the base.
+The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences.
+Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target.
+The green region in the lower image of Figure \ref{fig:rooms} illustrates the transferred CDSR.
 
@@
 In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions.
-The system identifies CDSRs in previously unseen environments through analogy with a manually annotated known room.
-The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy.
+The system identifies CDSRs in previously unseen environments through analogy with a single example.
+This is a difficult cognitive systems task requiring integration of semantic and geometric knowledge to identify regions which can be as small as 8\% of the room.
+Our system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy to achieve this task.
 In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work.
@@
 comparing the QSRs present in both base and transferred regions) to improve results.
-Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.
+In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them.
+Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to produce CDSRs, which overlap with user-defined region, from a single example for 6 out of 7 region types.
 

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Kate: fixed typos
minor edits page 1

@@
 It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.
-We then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms.
+From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms.
 To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
@@
 In this work we implicitly represent context using the types of objects present in a room and their location relative to each other.
-The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.
+The following sections describe how we construct symbolic representations of these ingredients of context from robot sensor data.
 
@@
 
-\fw{
-(regionType CDSR9 FrontRegion) \\
+\fw{(regionType CDSR9 FrontRegion) \\
 (boundarySegment CDSR9 \\
@@
 \hspace*{2em}(YMinXFewestFn Room3) \\
-\hspace*{2em}(YMinXFewestFn Group1))\\
-...
+\hspace*{2em}(YMinXFewestFn Group1))
 }
 }
-  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
+  \caption{Three of the five expressions representing the front of the classroom context-dependent region \fw{CDSR9}}
   \label{fig:cdsr-reps}
@@
 
-Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}.
+Figure \ref{fig:cdsr-reps} contains three of the five expressions defining the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}.
 The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region.
@@
 Each oval represents a predicate, and the entity arguments are represented by squares.
-SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case .
-This is done in the following manner.
+SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case in the following manner.
 First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME.
-Then SME will align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
+Then SME aligns the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}.
 While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa.
@@
 Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run.
-The map and object data that results from this is then passed on to the QSR generator.
+The map and object data that result from this are then passed on to the QSR generator.
 In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area.
@@
 Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target.
-The green region in the lower image of Figure \ref{fig:rooms} illustrates the inferred CDSR.
+The green region in the lower image of Figure \ref{fig:rooms} illustrates the transferred CDSR.
 

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
minor edits page 2 and fig 4

@@
 Dora's metric map is a collection of lines in a 2D global coordinate frame.
+Two example maps are pictured in Figure~\ref{fig:rooms}.
 These lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
@@
   % \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
-  \caption{Maps of 2 real classrooms generated by our system, plus an expert-annotated CDSR (red, top image), subject-annotated CDSR (blue, bottom image) and a (smaller) CDSR transferred by analogy (green bottom image).
+  \caption{Maps of 2 real classrooms generated by our system.
+The lines around the perimeter are walls, the unfilled polygons are the outlines of objects and the filled polygons are CDSRs.
+The maps show an expert-annotated CDSR (red, top image), a subject-annotated CDSR (blue, bottom image) and a CDSR transferred by analogy (green bottom image).
 The classroom used to generate the bottom classroom is pictured in Figure~\ref{fig:ug40}.}

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
minor edits page 2 and fig 4

@@
 Dora's metric map is a collection of lines in a 2D global coordinate frame.
-These lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
+Two example maps are pictured in Figure~\ref{fig:rooms}.
+Map lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
 Lines in this SLAM map represent features extracted from laser scans wherever a straight line is present for long enough to be considered permanent.

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
minor edits page 2

@@
 Two example maps are pictured in Figure~\ref{fig:rooms}.
-These lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
+Map lines are generated by a process which uses input from the robot's odometry and laser scanner to perform simultaneous localization and mapping (SLAM).
 Lines in this SLAM map represent features extracted from laser scans wherever a straight line is present for long enough to be considered permanent.
@@
 To enable us to generate a range of different evaluation situations in a reasonable length of time, we have generated data from Dora in both real rooms and in simulation.
-Simulation is performed using the Player/Stage hardware abstraction layer~\cite{GerkeyVaughanHoward03} allowing us to run the system \emph{mostly unchanged} in a pre-defined environment.
+Simulation is performed using the Player/Stage hardware abstraction layer~\cite{GerkeyVaughanHoward03} allowing us to run the system mostly unchanged in a pre-defined environment.
 Also, to enable us to detect a wider range of objects than is usually possible (from armchairs to whiteboards), we used a simulated object recogniser in all runs.
-The recogniser was configured with types and positions of objects in the environment and was guaranteed to detected them when the robot was orientated towards them.
+The recogniser was configured with types and positions of objects in the environment and was guaranteed to detect them when the robot was orientated towards them.
 This eliminated any errors from the recognition process, but was still influenced by errors in robot localisation.

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
qsr

@@
 
-For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+For each object that Dora detects we compute the strengths of 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}.
+The strength of a computed relation for a given pair of objects represents the applicability of that relation to the pair.
+Strength ranges from 0 to 1, with 0 being unsuitable.
 The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}.
@@
 
-% These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+These spatial relationships between adjacent objects provide the structure necessary for analogical processing.
+Generating the relationships in this way (as opposed to, for example, simple coordinate-based thresholding) has the advantage that the presence and absence of relationships is represented on a continuous scale.
+This provides our representations with the flexibility necessary to manage the variation in perceptual information (i.e.
+the position of walls and objects) inevitable in human environments and robot perception.
+
+In addition to spatial relations, we also create \emph{grouping entities}
+from the robot sensor data.
+Grouping entities collect together sets of adjacent objects of the same type.
+For example, a classroom would likely have a group entity created in which all of the students' desks were members.
 
@@
 \caption{A visualisation of a the strength of a spatial relation across a region.
-In this figure the landmark is located at the center of the image and the direction vector used was $\langle0,1\rangle$ (i.e.
+The landmark is the red square and the direction vector used was $\langle0,1\rangle$ (i.e.
 above of the landmark).
@@
 \end{figure}
-
-\from{Nick}{TODO: Generating QSRs is this way allows us to\dots}

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
sect 4

@@
 \fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}.
-The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks.
+The next segment connects the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks.
 There are two more boundary segments completing a polygon for this region.
-The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room".
+The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, ``the front of the room''.
 This definition for the front of the room is specific to \fw{Room3} and its entities.
@@
 For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building.
-We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic and geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
+Our system uses \emph{analogy} to solve this problem.
+We chose this approach because analogy has been previously used to successfully combine semantic and geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.
 
@@
 We first create the base and target representations by running Dora in the two different classrooms.
-These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively.
 In each case, Dora is manually driven around the room to allow it to create a metric map.
@@
 The map and object data that result from this are then passed on to the QSR generator.
+The base and target maps are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively.
 In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area.
@@
 The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences.
-Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target.
+Four of these candidate inferences define the CDSR in the target with anchor points defined on the room and the group of desks in the target.
 The green region in the lower image of Figure \ref{fig:rooms} illustrates the transferred CDSR.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
eval

@@
 To determine how people define CDRSs, we asked three na\"ive users to draw polygons for each region type for each room.
-This task was performed using a webpage (link removed for blind review) shown in the inset in Figure~\ref{fig:ug40} on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
+This task was performed using a webpage on which each user was presented with an image of the real room plus an image of the map data generated by the robot onto which the drawing could be done.
+The webpage (link removed for blind review) is shown in the inset in Figure~\ref{fig:ug40}.
 The user-defined polygons define the \textit{target regions} against which we evaluate our transfers.
@@
 We consider a \textit{problem instance} to be a room and a sought CDSR type.
-For each room containing a manually encoded CDSR of the sought type, we generate an \textit{transferred region} using analogical transfer.
+For each room containing a manually encoded CDSR of the sought type, we generate a \textit{transferred region} using analogical transfer.
 To assess the quality of the transfer, we calculate precision ($p$, the proportion of the transferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the transferred region) as follows:
@@
 
-Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually encoded CDSR and target region; and the region for the whole room and the target region.
+Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the CDSR we manually encoded for the target room and target region; and the region for the whole room and the target region.
 Results comparing transferred and target regions measure how well our system is able apply a single example to new situations.
@@
 Collaboration with human users requires a high precision and recall, because cognitive systems must be able to understand as well as refer to these regions in human environments.
-Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation.
+Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting point for a symbolic representation.
 Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within tasks with human users.
@@
 The transferred regions were considerably more precise (.47) when compared to the room as whole (.17), and their recalls (.46) indicate that they captured almost half of the area indicated by the human user.
-Our approach performed the best when the extent of the context dependent spatial region is closely tied to objects as is the case in the front rows of the classroom with p's of .76 and .82 for the inferred and the manually encoded regions respectively.
+As we create CDSRs using anchor points defined on perceived entities, our approach performs best when the boundary of the target CDSR is closely tied to such entities.
+This is the case in the front rows of the classroom, with $p$ of .76 and .82 for the inferred and the manually encoded regions respectively.
 The system performs worst when the extent of the CDSR is defined as an unbounded area near or around particular objects.
 The office of a studio apartment is loosely defined as the region around the desk.
-This motivates one direction of future work to expand the vocabulary of anchor points to better capture these notions of space.
+This motivates one direction of future work:  expanding the vocabulary of anchor points to better capture these notions of space.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
klenk's edits done

@@
 The simulated rooms were based on real-life counterparts.
-For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
+For each room we manually encoded appropriate CDSRs that could be represented by our approach.
 For the classrooms these were the front and back, and the front and back rows of desks.
@@
 \end{tabular}
-\caption{Overall Performance}
+\caption{Overall Performance Compared Against Target Regions Defined by Three Users}
   \label{tab:overall}
@@
 \hline
-Region & Inferred & Manually Encoded & Entire Room \\
+Region & Transferred & Manually Encoded & Entire Room \\
 \hline

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
klenk's edits done

@@
 The system identifies CDSRs in previously unseen environments through analogy with a single example.
-This is a difficult cognitive systems task requiring integration of semantic and geometric knowledge to identify regions which can be as small as 8\% of the room.
+This is a difficult cognitive systems task requiring integration of semantic and geometric knowledge to identify regions as small as 8\% of the room.
 Our system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy to achieve this task.
@@
 In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them.
-Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to produce CDSRs, which overlap with user-defined region, from a single example for 6 out of 7 region types.
+Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to transfer CDSRs that overlap with user-defined regions for 6 out of 7 region types.
 

### 2012-04-03 code/scenarios/cdsr/trunk/papers/AAAI2012/aaai12-cdsr-rebuttal.txt
added rebuttal text

We thank the reviewers for their great feedback. We would like to
use this space to clarify some aspects of our work. The primary criticism
from Reviewer 2 seems to be that we are not explicitly representing
function/orientation information and just using "spatial templates".

First, to explicitly represent function and recognise it from a mobile robot,
would require a number of advances in computer vision that are outside of our
expertise. It is easier to include orientation information of recognised
objects, and we plan to do so in the future. However object orientation, like object category, is just
a proxy for function (i.e. it represents function implicitly) so it would not
significantly alter our approach, just add more discriminatory power. Rather
than use an explicit representation of function, our approach assumes that
"semantically and geometrically similar areas (e.g. two different classrooms)
will feature similar CDSRs, and that these similarities can be recognised
through analogy." Therefore, we avoid explicitly representing function, but
assume that the functional properties of a region can be extracted through
analogy to a previously labelled (functional) region. If, at some point,
function representations become easily obtainable on a robot, we would be
happy to include them in our system as it should provide a better similarity
measure and assist with analogical transfer.

Second, our use of spatial templates extends only to determining the weights
on different positional relationships between two objects. CDSRs are
represented using anchor points, which are considerably different from
templates. Because anchor points are symbolic, this enables their analogical
transfer which is one of the main contributions of this work. This is a
cognitive systems approach because it allows for new CDSRs to be learned from
a single example. It is not clear how one could use spatial templates to
perform the same task, as the variation across rooms and regions would defeat a template matching approach.

The reviewer raises an important point that there may be gray area in how
these regions are used by people. The results presented in paper are averaged
across each subject. The results by subject were rather similar, although the
regions constructed were a bit different. To address this, we intend to
perform a thorough human study and also embed our system in a navigation
communication task. But the work in the paper shows that our approach can
represent CDSRs and transfer them analogically to semantically and
geometrically similar situations as determined by sensor data.

### 2012-04-03 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
added authors

@@
 \documentclass[letterpaper]{article}
+\usepackage{aaai}
 
-\usepackage{aaai}
+\setlength\titlebox{2.5in}
+
 \usepackage{times}
@@
 \title{Towards a Cognitive System That Can Recognize Spatial Regions Based on Context}
-% \title{Representing and Reasoning About Spatial Regions Defined by Context}
-% \title{Context-Dependent Spatial Regions}
 
-\author{}
+\author{%
+Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk 
+\And % 
+Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com 
+\And % 
+Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  
+\AND % 
+Graham S. Horn \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ gsh148@cs.bham.ac.uk 
+\And % 
+John Kelleher \\ Applied Intelligence Research Centre \\ Dublin Institute of Technology \\ john.d.kelleher@dit.ie}
 
@@
 
-% \section{Acknowledgments}
-% 
+\section{Acknowledgments}
+
 % The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams, Scott Friedman for assistance on probablistic SME, and Danny Bobrow comments on an earlier draft of this paper.
-The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.
-215181, CogX.
 
-% 7th page may only contain references
-\clearpage
+The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
 

### 2012-04-04 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
added john's initial

@@
 \title{Towards a Cognitive System That Can Recognize Spatial Regions Based on Context}
-% \title{Representing and Reasoning About Spatial Regions Defined by Context}
-% \title{Context-Dependent Spatial Regions}
 
-\author{}
+\author{%
+Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk 
+\And % 
+Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com 
+\And % 
+Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  
+\AND % 
+Graham S. Horn \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ gsh148@cs.bham.ac.uk 
+\And % 
+John D. Kelleher \\ Applied Intelligence Research Centre \\ Dublin Institute of Technology \\ john.d.kelleher@dit.ie}
 
@@
 In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them.
-Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to produce CDSRs, which overlap with user-defined region, from a single example for 6 out of 7 region types.
+Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to transfer CDSRs that overlap with user-defined regions for 6 out of 7 region types.
 
-% \section{Acknowledgments}
-% 
+\section{Acknowledgments}
+
 % The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams, Scott Friedman for assistance on probablistic SME, and Danny Bobrow comments on an earlier draft of this paper.
-The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.
-215181, CogX.
 
-% 7th page may only contain references
-\clearpage
+The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
 

### 2012-04-13 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

@@
 
-To determine how people define CDRSs, we asked three na\"ive users to draw polygons for each region type for each room.
+To determine how people define CDSRs, we asked three na\"ive users to draw polygons for each region type for each room.
 This task was performed using a webpage on which each user was presented with an image of the real room plus an image of the map data generated by the robot onto which the drawing could be done.
-The webpage (link removed for blind review) is shown in the inset in Figure~\ref{fig:ug40}.
+The webpage\footnote{http://home.csumb.edu/k/katherinelockwood/world/} is shown in the inset in Figure~\ref{fig:ug40}.
 The user-defined polygons define the \textit{target regions} against which we evaluate our transfers.

### 2012-04-13 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

@@
 \And % 
-John Kelleher \\ Applied Intelligence Research Centre \\ Dublin Institute of Technology \\ john.d.kelleher@dit.ie}
+John D. Kelleher \\ Applied Intelligence Research Centre \\ Dublin Institute of Technology \\ john.d.kelleher@dit.ie}
 
@@
 
-% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams, Scott Friedman for assistance on probablistic SME, and Danny Bobrow comments on an earlier draft of this paper.
-
-The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
+The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No.
+215181, CogX.
+The authors would like to thank Jeffery Usher for providing the code to compute voronoi diagrams, Scott Friedman for assistance on probabilistic SME, and Danny Bobrow comments on an earlier draft of this paper.
 

### 2012-04-13 code/scenarios/cdsr/trunk/papers/QR2012/main-QR12-cdsr.tex

--- 2012-04-13 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
@@
 
-\title{Towards a Cognitive System That Can Recognize Spatial Regions Based on Context}
+\title{Using Anchor Points to Define and Transfer Spatial Regions Based on Context}
 
 \author{%
+Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com 
+\And % 
 Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk 
-\And % 
-Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com 
 \And % 
@@
 \begin{abstract}
-In order to collaborate with people in the real world, cognitive systems must be able to represent and reason about spatial regions in human environments.
+In order to collaborate with people in the real world, AI systems must be able to represent and reason about spatial regions in human environments.
 Consider the command \emph{``go to the front of the classroom''}.
@@
 Instead it is defined by its functional use, implied by nearby objects and their configuration.
-In this paper, we define such areas as \textit{context-dependent spatial regions} and present a cognitive system able to learn them by combining qualitative spatial representations, semantic labels, and analogy.
+In this paper, we define such areas as \textit{context-dependent spatial regions} and present a system able to learn them by combining qualitative spatial representations, semantic labels, and analogy.
 The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world.
@@
 This response requires that the robot is able to \emph{determine the spatial region in the environment that satisfies this concept}.
+Moving from the metric space of the robot's sensors to the symbolic space required for reasoning and language is an important problem for qualitative reasoning.
 
-The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments.
+The ability to understand and reason about \textit{spatial regions} is essential for AI systems performing tasks for humans in everyday environments.
 Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g.
@@
 
-Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability.
-If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do.
+Current AI systems are not capable of representing and reasoning about CDSRs, yet it is an important ability.
+If AI systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do.
 Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.
 In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge.
-Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm \cite{langley:inpress}.
+Qualitative representations provide a symbolic abstraction that is a natural method for integrating these different types of knowledge in reasoning tasks.
 
@@
 
-We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs.
+We use \textit{anchor points}~\cite{Klenk/etal2011} to define the boundaries of CDSRs.
 Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity.
@@
 Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}.
-When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}.
-This paper follows in this tradition, adding CDSRs to his qualitative topological representations.
+When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy (SSH) \cite{Kuipers:2000}.
+In that work, Kuipers distinguishes between large-scale and small scale space based on the robots sensor horizon.
+This distinction has proven useful for performing localization and map building \cite{Beeson01042010}.
+CDSRs differ from this approach by incorporating non-geometric knowledge into the definition of regions.
 Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms.
@@
 The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning.
-For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}.
+For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2011}.
 In that work, the anchor points defined the endpoints of lines.
@@
 
-In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions.
+In this paper, we illustrate how qualitative representations enable the integration of geometric and semantic knowledge required to reason about CDSRs.
 The system identifies CDSRs in previously unseen environments through analogy with a single example.
-This is a difficult cognitive systems task requiring integration of semantic and geometric knowledge to identify regions as small as 8\% of the room.
 Our system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy to achieve this task.
@@
 These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g.
-MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g.
-comparing the QSRs present in both base and transferred regions) to improve results.
+MAC/FAC \cite{forbus/etal1995}), and manually-encoding the CDSRs instead of extracting them from sketched input.
 In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them.
@@
 
-\bibliography{aaai12-cdsr}
+\bibliography{QR12-cdsr}
 \bibliographystyle{aaai}

### 2012-04-13 code/scenarios/cdsr/trunk/papers/QR2012/qsr-extraction.tex

--- 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
@@
 Generating the relationships in this way (as opposed to, for example, simple coordinate-based thresholding) has the advantage that the presence and absence of relationships is represented on a continuous scale.
-This provides our representations with the flexibility necessary to manage the variation in perceptual information (i.e.
+While similar to the relationships used in fuzzy qualitative trigonometry \cite{Liu200971}, our approach includes distance as well as angle into the weight computation.
+Weighted relationships provide our representations with the flexibility necessary to manage the variation in perceptual information (i.e.
 the position of walls and objects) inevitable in human environments and robot perception.

### 2012-05-15 code/scenarios/cdsr/trunk/papers/QR2012/main-QR12-cdsr.tex

@@
 \usepackage{aaai}
+\nocopyright
 
@@
 It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.
-From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms.
+From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms by transferring the necessary anchor points.
 To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
@@
 This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs.
+Central to our approach is the use of \emph{anchor points}, symbolic expressions which link conceptual entities (e.g., CDSRs) to perceived entities (e.g., objects in the environment).
 Our approach is founded on two assumptions.
@@
 The rest of the paper is structured following these assumptions.
-Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs.
+Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs with anchor points.
 Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation.
@@
 recognizing the CDSR in the target) as follows.
-First, we identify the CDSR of the sought type in the base and use SME's pragmatic constraints to ensure that the entities referred to its anchor points participate in the mapping.
+First, we identify the CDSR of the sought type in the base and use SME's pragmatic constraints to ensure that the entities referred to by its anchor points participate in the mapping.
+For example, if the CDSR in the base included the anchor point \fw{(YMaxXMostFn Desk1)}, then we would include a constraint requiring \fw{Desk1} participate in the mapping.
 To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{boundarySegment} statements mentioning the base CDSR.
 The second and third arguments of these candidate inferences are anchor points in the target environment.
-We use these to define the boundary of the CDSR in the target.
+We use these anchor points to define the boundary of the CDSR in the target.
 

### 2012-05-15 code/scenarios/cdsr/trunk/papers/QR2012/qsr-extraction.tex

@@
 Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid.
-This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
+This results in the coordinates of all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.
 We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance $d$ between the landmark's centroid and the adjacent object's location, and (b) the inner angle $\theta$ between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location.

### 2012-06-05 scenarios/cdsr/trunk/papers/QR2012/evaluation.tex

--- 2012-04-13 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
@@
 \begin{equation}
-	r=\frac{area(transferred\ region \cap target\ region)}{area(transferred\ region)}
+	r=\frac{area(transferred\ region \cap target\ region)}{area(target\ region)}
 \end{equation}
@@
 Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting point for a symbolic representation.
-Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within tasks with human users.
+Our future work seeks to further evaluate the utility of this representation by employing it in collaborative tasks with human users.
 

### 2012-06-05 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

@@
 The simulated rooms were based on real-life counterparts.
-For each room we manually encoded the appropriate CDSRs that could be represented by our approach.
-For the classrooms these were the front and back, and the front and back rows of desk.
+For each room we manually encoded appropriate CDSRs that could be represented by our approach.
+For the classrooms these were the front and back, and the front and back rows of desks.
 For the studios these were the kitchen, office and living areas.
@@
 
-To determine how people define CDRSs, we asked three na\"ive users to draw polygons for each region type for each room.
-This task was performed using a webpage (link removed for blind review) shown in the inset in Figure~\ref{fig:ug40} on which each user was presented with an image of the real room plus the map data generated by the robot (as in Figure~\ref{fig:rooms}) onto which the drawing could be done.
+To determine how people define CDSRs, we asked three na\"ive users to draw polygons for each region type for each room.
+This task was performed using a webpage on which each user was presented with an image of the real room plus an image of the map data generated by the robot onto which the drawing could be done.
+The webpage\footnote{http://home.csumb.edu/k/katherinelockwood/world/} is shown in the inset in Figure~\ref{fig:ug40}.
 The user-defined polygons define the \textit{target regions} against which we evaluate our transfers.
@@
 We consider a \textit{problem instance} to be a room and a sought CDSR type.
-For each room for which we have a human-labelled target region of the sought type, we generate an \textit{inferred region} using analogical transfer.
-To assess the quality of the transfer, we calculate precision ($p$, the proportion of the inferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the inferred region) as follows:
+For each room containing a manually encoded CDSR of the sought type, we generate a \textit{transferred region} using analogical transfer.
+To assess the quality of the transfer, we calculate precision ($p$, the proportion of the transferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the transferred region) as follows:
 
 \begin{equation}
-	p=\frac{area(inferred\ region \cap target\ region)}{area(inferred\ region)}
+	p=\frac{area(transferred\ region \cap target\ region)}{area(transferred\ region)}
 \end{equation}
 \begin{equation}
-	r=\frac{area(inferred\ region \cap target\ region)}{area(target\ region)}
+	r=\frac{area(transferred\ region \cap target\ region)}{area(target\ region)}
 \end{equation}
 
-Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually annotated CDSR in the target and appropriate target region; and the region for the whole room and the target region.
- Results comparing transferred and target regions measure how well the output of our system matches user expectations.
-Results comparing the manual annotations to the target regions measure how well the anchor point-based regions we chose match the users' regions (who were not constrained to anchor points).
-Results comparing whole rooms to target regions provide a baseline performance for comparison.
+Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the CDSR we manually encoded for the target room and target region; and the region for the whole room and the target region.
+Results comparing transferred and target regions measure how well our system is able apply a single example to new situations.
+The comparisons between the manual annotations to the target regions measure how well the anchor points we chose capture the users' regions (who were not constrained to anchor points).
+Results from the whole room regions provide a baseline performance for comparison.
 

### 2013-08-07 scenarios/cdsr/trunk/papers/ACS2013/main.tex
empty paper

\documentclass[11pt,letterpaper]{article}
\usepackage{cogsys}
\usepackage{cogsysapa}
% \usepackage{apacite}
% \usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage[pdftex]{graphicx} % use this when importing PDF files

% Put back in

 % First page headings for accepted submissions.
% \cogsysheading{1}{2012}{1-18}{9/2012}{12/2012}
 % First page headings for poster submissions.
%\cogsysposterheading{First}{2012}{1-18}

% \ShortHeadings{Formatting Instructions}
              % {P.\ Langley, G.\ Hunt, and D.\ G.\ Shapiro}

\begin{document}

\title{CDSR Human Data/Modelling Paper}

% \author{Pat Langley}{langley@asu.edu}
% \author{Glen Hunt}{glen.hunt@asu.edu}
% \address{Computing Science and Engineering, Arizona State University, 
%          Tempe, AZ 85287 USA}
% \author{Daniel G.\ Shapiro}{dgs@isle.org}
% \address{Institute for the Study of Learning and Expertise, 
%          2164 Staunton Court, Palo Alto, CA 94306 USA}
\vskip 0.2in

\begin{abstract}
Abstract goes here
\end{abstract}

\section{Introductions}

All submissions should follow the same format to ensure that the 
printers can reproduce them without problems and to let readers 
more easily find the information that they desire.

\subsection{Paper Dimensions}

The text of the paper should be formatted in one column, with an
overall width of 6.0~inches (15.24~cm) and length of 8.0~inches 
(20.32~cm). The left margin should be 1.25~inches (3.175~cm) and the 
top margin 1.5~inches (3.81~cm). The right and bottom margins will
depend on whether you print on US letter or A4 paper.

The paper body should be set in 11~point type with a vertical spacing 
of 12~points. Please use Times Roman typeface throughout the text.

\subsection{Title and Author Information}

The paper title should be set in 14~point bold type and centered
between two horizontal rules that are 1~point thick, with 1.5~inches
between the top rule and the top edge of the page. Capitalize the
first letter of each content word and put the rest of the title in 
lower case.

Author information should start 0.3~inches (0.762~cm) below the bottom
rule surrounding the title. The authors' names should appear in
11~point bold type, electronic mail addresses in 11~point small
capitals, and physical addresses in ordinary 11~point type.

Each author's name should be flush left, whereas the email address
should be flush right on the same line. The author's physical address
should appear flush left on the ensuing line, on a single line if
possible. If successive authors have the same affiliation and address, 
then give this information only once for them.

\subsection{Partitioning the Text}

You should organize your paper into sections and paragraphs to help 
readers place a structure on the material and understand its 
contributions.

% Use \vspace for fine control of spacing above and below headings. 
\vspace{-0.018in}
\subsubsection{Sections and Subsections}
\vspace{-0.015in}

Section headings should be numbered, flush left, and set in 12~point 
bold type with the content words capitalized. Leave about 0.25~inches
(0.635~cm) of space before the heading and approximately 0.15~inches 
(0.33~cm) after the heading.

Similarly, subsection headings should be numbered, flush left, and 
set in 11~point bold type with the content words capitalized. Leave
approximately 0.2~inches (0.51~cm) of space before the heading and 
0.13~inches (0.33~cm) afterward.

Finally, subsubsection headings should be numbered, flush left, and
set in 11~point italics with the content words beginning with capitals. 
Leave about 0.18~inches (0.457~cm) of space before the heading and
about 0.1~inches (0.254~cm) after the heading. Please use no more 
than three levels of headings in a paper.

\subsubsection{Paragraphs, Quotations, and Footnotes}

Within each section or subsection, you should further partition 
the paper into paragraphs. You should not indent the first line 
of a section or subsection's initial paragraph, but you should
indent successive paragraphs by 0.25 inches (0.635~cm). There 
should be no extra spacing between paragraphs within a section.

You may want to quote an excerpt of relevant text that appears in 
the published work of some other authors:

\begin{quote}
Such a quotation should begin 0.25 inches (0.635 cm) below the preceding 
line of regular text and end 0.25 inches (0.635 cm) before it begins
again. The left and right margins should be 0.35 inches (0.889 cm) larger 
than for the regular text, but the quoted text should have the same
font as regular text, with a vertical spacing of 12~points.
\end{quote}

\noindent
You should remember to cite the source of the quotation, and you should 
close off each quotation with some regular text, which should not be
indented.

You can use footnotes\footnote{For the sake of readability, footnotes
should be complete sentences.} to provide readers with additional
information about a topic without interrupting the flow of the paper.
Indicate footnotes with a number in the text where the point is most
relevant. Place the footnote in 10~point type at the bottom of the
page on which it appears. Precede the first footnote on a page 
with a horizontal rule of 2.0~inches (5.08~cm) to separate it from 
the main text.\footnote{Multiple footnotes can appear on each page, 
in the same order as they appear in the text, but spread them across
pages if possible.}

\subsubsection{Itemized and Enumerated Lists}

When you have a number of related but distinct points, you may want
to organize them in an itemized list that:

\vskip 0.05in

\cbullet 
begins 0.15 inches (0.381 cm) below the preceding line of regular text 
and ends 0.15 inches (0.381 cm) before the regular text begins;

\cbullet 
uses a left margin for each item that is 0.25 inches (0.635 cm) to the 
right of the left margin for regular text;

\cbullet 
includes a bullet at the beginning of the first line of each item,
separated from the item's text by 0.075 inches (0.19 cm); and

\cbullet 
uses a vertical spacing of 12~points for lines within each item 
and separates successive items by the same amount.

\vskip 0.05in
\noindent
In general, you should close off each itemized list with some sentences 
of regular text, which should not be indented. In some cases, you may
instead want to use an enumerated list that replaces bullets with numbers.

\begin{figure}[t]
\vskip 0.05in
\begin{center}
% \setlength{\epsfxsize}{3.5in}
% \centerline{\epsfbox{cascade.eps}}
\includegraphics[width=3.5in]{cascade.png}
% \vskip 0.1in
\caption{Modules in the {\sc Icarus} cognitive architecture and their
         cascaded organization.} 
\label{sample-figure}
\end{center}
\vskip -0.2in
\end{figure}

\subsection{Figures}

You may want to include figures in the paper to help readers visualize
your approach and your results. Such artwork should be centered,
legible, and separated from the text. Lines should be dark and at
least 0.5~points thick for clear display and printing, and the text 
should not appear on a dark background.

% Use \newpage to insert a page break between paragraphs.

% \newpage

Label all distinct components of each figure. If the figure takes
the form of a graph, then give a name for each axis and include a 
legend that briefly describes each curve. However, do {\it not\/} 
include a title above the figure, as the caption already serves
this function.

Number figures sequentially, placing the figure number and caption
{\it after\/} the graphics, with at least 0.2~inches (0.508~cm) of 
space before the caption and 0.3~inches (0.762~cm) after it, as in
Figure~\ref{sample-figure}. The figure caption should be set in
10~point type and centered unless it runs two or more lines, in which
case it should be flush left. You may float a figure to the top or
bottom of a page, but ideally it should appear on the page that
first mentions it or on the page immediately afterward.

\subsection{Tables}

You may also want to include tables that summarize material. Like 
figures, these should be centered, legible, and numbered consecutively. 
However, place the title {\it above\/} the table with at least 
0.1~inches (0.254~cm) of space before the title and the same after 
it, as in Table~\ref{sample-table}. The table title should be set in
11~point type and centered unless it runs two or more lines, in which
case it should be flush left.

Tables contain textual material that can be typeset, as contrasted 
with figures, which contain graphical material that must be drawn. 
Specify the contents of each column in the table's topmost row. Again,
you may float a table to the top or bottom of a page, but attempt to 
place each table on the page that first mentions it or on the one
immediately afterward.

% Note use of \abovespace and \belowspace to get reasonable spacing 
% above and below tabular lines.

\def\w{$\oplus$}
\def\b{$\ominus$}
\def\h{$\odot$}

\begin{table}[t]
\vskip -0.15in
\caption{Adequacy of four models of learning in problem solving in 
         terms of whether they account (\w), fail to account (\b), 
         or partially account (\h) for phenomena from \cite{in-book-vanlehn}
         [{\tt *}] and Jones (1989) [$\diamond$].}
\label{sample-table}
\begin{small}
\begin{center}
% \vskip -0.10in
\begin{tabular}{lcccc}
\hline
\abovespace\belowspace
                       & ~~~ACT-R~~~ & ~~~Soar~~~ & {\sc ~~Eureka~~} & {\sc Daedalus} \\
\hline\abovespace
Means-ends analysis$^{\diamond}$  &   \h  &     \w    &      \w   &      \w \\
\hbox{Nonsystematicity$^{\diamond}$\hskip -0.2in} &   \b  &     \b    &      \w   &      \b \\
Problem isomorphs$^*$   &   \h  &     \h    &      \h   &      \h \\
Reduced search$^*$ &   \w  &     \w    &      \w   &      \w \\
Asymmetrical transfer$^*$    &   \w  &     \w    &      \w   &      \w \\
Einstellung effects$^*$    &   \w  &     \w    &      \w   &      \w \\
% Strategy differences   &   \h  &     \h    &      \h   &      \h \\
Reduced verbalization$^*$  &   \w  &     \w    &      \b   &      \b \\
Reduced solution time$^*$   &   \w  &     \w    &      \b   &      \b \\
% Self-monitoring        &   \h  &     \h    &      \b   &      \b \\
Rare analogies$^*$   &   \b  &     \b    &      \h   &      \h \\
\belowspace
% \belowstrut{0.10in}
Superficial analogies$^*$ &   \b  &     \b    &      \w   &      \w \\
\hline 
\end{tabular}
\end{center}
\vskip -0.10in
\end{small}
\end{table}

\subsection{Citations and References}

Please use APA reference format regardless of your formatter or word
processor. If you rely on the \LaTeX\/ bibliographic facility, use the 
files {\tt\small cogsysapa.sty} and {\tt\small cogsysapa.bst}, which
are available on the Web site, to obtain this format. These require
you to place your references in a separate file with a {\tt\small bib}
extension, as explained in the \LaTeX\/ manual.

Citations within the text should include the authors' last names and
year. If the authors' names are included as part of the sentence, place 
only the year in parentheses, as in Jones and VanLehn (1994), but 
otherwise place the entire reference in parentheses with the authors
and year separated by a comma (Jones \& VanLehn, 1984). 
List multiple references alphabetically and separate them by semicolons 
\cite{paper-laird-etal,book-newell-simon}. Use the 
`et~al.'  construction only for citations with four or more authors or
after listing all authors to a publication in an earlier reference.

Use an unnumbered first-level section heading for the references, and 
use a hanging indent style, with the first line of the reference flush
against the left margin and subsequent lines indented by 10 points. 
The references at the end of this document include examples for journal
articles \singlecite{article-forbus}, conference publications \singlecite{paper-laird-etal},
book chapters \singlecite{in-book-vanlehn}, books \singlecite{book-newell-simon}, 
edited volumes \singlecite{book-shrager-langley}, technical reports 
\singlecite{techreport-shapiro-etal}, and dissertations \singlecite{phd-choi}.

Alphabetize references in both the text and at its end by the surnames
of the first authors, with single author entries preceding multiple
author entries. Order references for the same authors by year of
publication, with the earliest first.

\section{Electronic Formatting and Submission}

{\it Advances in Cognitive Systems\/} will rely on electronic submission 
of papers for review and publication. We assume that nearly all authors 
will have access to \LaTeX\/ or Word to format their documents and can
use a Web browser to download style files and upload their papers. 
Authors who do not have such access should send email with their
concerns to {\sl acs@cogsys.org}.

\subsection{Paper Length and Format}

The title {\it Advances in Cognitive Systems\/} is associated with
both an electronic journal and an annual conference, with many of the
articles appearing in the former having been submitted and reviewed
under the auspices of the latter.

Papers submitted to the annual conference must not exceed sixteen (16)
pages, including all figures, tables, and references. We will return
to the authors any submissions that exceed this page limit or that
diverge significantly from the format specified herein.
Papers submitted for publication in the journal but not for presentation
at the meeting are not subject to this constraint. However, we encourage 
authors to keep their papers the same length as those for the conference.

Submissions may be accompanied by online appendices that contain data,
demonstrations, instructions for obtaining source code, or the source
code itself. We encourage authors to include such appendices when they
submit papers. This material will not count in a submission's page
length.

Electronic templates for producing the camera-ready copy are available
for \LaTeX\/ and Microsoft Word. Style files and sample papers are 
available on the Web at: 
\vskip 0.1in
\begin{small}
\centerline{{\tt http://www.cogsys.org/formats/} .}
\end{small}
\vskip 0.1in
\noindent
Authors who have questions about these electronic formats should direct
them to {\sl acs@cogsys.org}.

To ensure the ability to preview and print submissions, authors must
provide their manuscripts in pdf format. Papers prepared in Word
should be saved as pdf files and submitted in this format. To support
the review process, each submission must be accompanied by information
about the paper's title and abstract, as well as the authors' names
and physical addresses.

\subsection{Submitting Initial and Final Papers}

Submission of initial and final papers will take place in a purely
electronic manner through the World Wide Web. Authors who intend to
submit a paper for presentation at the annual conference should upload
their file to the submission repository at
\vskip 0.1in
\begin{small}
\centerline{{\tt http://www.cogsys.org/submit/} .}
\end{small}
\vskip 0.12in
\noindent
Submission should be completed no later than 11:59 PM Pacific time 
on the conference due date, which will be stated clearly on the 
meeting Web site. If a submission is late, then it will not be 
considered for inclusion at the meeting.

{\it Advances in Cognitive Systems\/} will also consider submissions
that are not intended for presentation at the annual conference. 
Authors who wish to submit papers of this sort should upload their
file to the submission repository at
\vskip 0.1in
\begin{small}
\centerline{{\tt http://www.cogsys.org/journal/submit/} .}
\end{small}
\vskip 0.08in
\noindent
Such submissions may occur at any time, unless they are related to 
special issues of the journal, which will have their own deadlines.

% You are welcome to submit a paper simultaneously to another meeting 
% or publication, but only if you indicate this fact clearly on the
% submission form. Simultaneous submissions that are not clearly
% specified as such will be rejected on discovery.

Final versions of papers accepted for publication, and resubmissions 
of conditionally accepted papers, should follow the same format 
as that for initial submissions. Authors should upload their revised 
papers to the same Web repository as used for the originals. 
For the annual conference, final versions should be deposited no 
later than 11:59 PM Pacific time on the due date. Authors of late 
submissions will not be provided with a slot to present their work 
at the meeting and their papers will not appear in the journal.
Papers not related to the conference may be uploaded at any time.

In both cases, papers that have been conditionally accepted will be
examined for compliance with conditions placed on their acceptance. 
Final papers must follow the content, format and length restrictions 
that are specified in these instructions.

% \newpage

\begin{acknowledgements} 
\noindent
STRANDS etc.
\end{acknowledgements}

\vspace{-0.25in}

{\parindent -10pt\leftskip 10pt\noindent
\bibliographystyle{cogsysapa}
\bibliography{cdsr}

}

% Leave a blank line before the closing brace to ensure the final 
% reference has the proper indentation.

\end{document}

### 2013-08-07 scenarios/cdsr/trunk/papers/ACS2013/main.tex
Dump of skeleton from the Google doc.

@@
 
-\section{Introductions}
+\section{Introduction - KLENK}
 
-All submissions should follow the same format to ensure that the 
-printers can reproduce them without problems and to let readers 
-more easily find the information that they desire.
+\section{Context-Dependent Spatial Regions - KLENK}
 
-\subsection{Paper Dimensions}
+\begin{itemize}
+	\item Explanation of the basic idea.
+	\item The desiderata of the model - context, extent, generalises to unseen examples
 
-The text of the paper should be formatted in one column, with an
-overall width of 6.0~inches (15.24~cm) and length of 8.0~inches 
-(20.32~cm). The left margin should be 1.25~inches (3.175~cm) and the 
-top margin 1.5~inches (3.81~cm). The right and bottom margins will
-depend on whether you print on US letter or A4 paper.
+\end{itemize}
 
-The paper body should be set in 11~point type with a vertical spacing 
-of 12~points. Please use Times Roman typeface throughout the text.
+\section{Human Studies - John}
+Research Questions: 
+\begin{enumerate}
+	\item Does context matter?
+	\item Do people agree? 
+\end{enumerate}
 
-\subsection{Title and Author Information}
+\section{Modeling - Nick}
+If we want to capture this data in a generative model, what is our hypothesis space. 
+Related work?
 
-The paper title should be set in 14~point bold type and centered
-between two horizontal rules that are 1~point thick, with 1.5~inches
-between the top rule and the top edge of the page. Capitalize the
-first letter of each content word and put the rest of the title in 
-lower case.
+\subsection{Symbolic Boundary Method - KLENK}
 
-Author information should start 0.3~inches (0.762~cm) below the bottom
-rule surrounding the title. The authors' names should appear in
-11~point bold type, electronic mail addresses in 11~point small
-capitals, and physical addresses in ordinary 11~point type.
+Overview
+Example
+Results?
+Discussion
 
-Each author's name should be flush left, whereas the email address
-should be flush right on the same line. The author's physical address
-should appear flush left on the ensuing line, on a single line if
-possible. If successive authors have the same affiliation and address, 
-then give this information only once for them.
+\subsection{Kernel Methods - NICK}
 
-\subsection{Partitioning the Text}
+Overview - either Chris’s approach or something similar
+Example
+Results?
+Discussion 
+Where do the points come from
 
-You should organize your paper into sections and paragraphs to help 
-readers place a structure on the material and understand its 
-contributions.
-
-% Use \vspace for fine control of spacing above and below headings. 
-\vspace{-0.018in}
-\subsubsection{Sections and Subsections}
-\vspace{-0.015in}
-
-Section headings should be numbered, flush left, and set in 12~point 
-bold type with the content words capitalized. Leave about 0.25~inches
-(0.635~cm) of space before the heading and approximately 0.15~inches 
-(0.33~cm) after the heading.
-
-Similarly, subsection headings should be numbered, flush left, and 
-set in 11~point bold type with the content words capitalized. Leave
-approximately 0.2~inches (0.51~cm) of space before the heading and 
-0.13~inches (0.33~cm) afterward.
-
-Finally, subsubsection headings should be numbered, flush left, and
-set in 11~point italics with the content words beginning with capitals. 
-Leave about 0.18~inches (0.457~cm) of space before the heading and
-about 0.1~inches (0.254~cm) after the heading. Please use no more 
-than three levels of headings in a paper.
-
-\subsubsection{Paragraphs, Quotations, and Footnotes}
-
-Within each section or subsection, you should further partition 
-the paper into paragraphs. You should not indent the first line 
-of a section or subsection's initial paragraph, but you should
-indent successive paragraphs by 0.25 inches (0.635~cm). There 
-should be no extra spacing between paragraphs within a section.
-
-You may want to quote an excerpt of relevant text that appears in 
-the published work of some other authors:
-
-\begin{quote}
-Such a quotation should begin 0.25 inches (0.635 cm) below the preceding 
-line of regular text and end 0.25 inches (0.635 cm) before it begins
-again. The left and right margins should be 0.35 inches (0.889 cm) larger 
-than for the regular text, but the quoted text should have the same
-font as regular text, with a vertical spacing of 12~points.
-\end{quote}
-
-\noindent
-You should remember to cite the source of the quotation, and you should 
-close off each quotation with some regular text, which should not be
-indented.
-
-You can use footnotes\footnote{For the sake of readability, footnotes
-should be complete sentences.} to provide readers with additional
-information about a topic without interrupting the flow of the paper.
-Indicate footnotes with a number in the text where the point is most
-relevant. Place the footnote in 10~point type at the bottom of the
-page on which it appears. Precede the first footnote on a page 
-with a horizontal rule of 2.0~inches (5.08~cm) to separate it from 
-the main text.\footnote{Multiple footnotes can appear on each page, 
-in the same order as they appear in the text, but spread them across
-pages if possible.}
-
-\subsubsection{Itemized and Enumerated Lists}
-
-When you have a number of related but distinct points, you may want
-to organize them in an itemized list that:
-
-\vskip 0.05in
-
-\cbullet 
-begins 0.15 inches (0.381 cm) below the preceding line of regular text 
-and ends 0.15 inches (0.381 cm) before the regular text begins;
-
-\cbullet 
-uses a left margin for each item that is 0.25 inches (0.635 cm) to the 
-right of the left margin for regular text;
-
-\cbullet 
-includes a bullet at the beginning of the first line of each item,
-separated from the item's text by 0.075 inches (0.19 cm); and
-
-\cbullet 
-uses a vertical spacing of 12~points for lines within each item 
-and separates successive items by the same amount.
-
-\vskip 0.05in
-\noindent
-In general, you should close off each itemized list with some sentences 
-of regular text, which should not be indented. In some cases, you may
-instead want to use an enumerated list that replaces bullets with numbers.
-
-\begin{figure}[t]
-\vskip 0.05in
-\begin{center}
-% \setlength{\epsfxsize}{3.5in}
-% \centerline{\epsfbox{cascade.eps}}
-\includegraphics[width=3.5in]{cascade.png}
-% \vskip 0.1in
-\caption{Modules in the {\sc Icarus} cognitive architecture and their
-         cascaded organization.} 
-\label{sample-figure}
-\end{center}
-\vskip -0.2in
-\end{figure}
-
-\subsection{Figures}
-
-You may want to include figures in the paper to help readers visualize
-your approach and your results. Such artwork should be centered,
-legible, and separated from the text. Lines should be dark and at
-least 0.5~points thick for clear display and printing, and the text 
-should not appear on a dark background.
-
-% Use \newpage to insert a page break between paragraphs.
-
-% \newpage
-
-Label all distinct components of each figure. If the figure takes
-the form of a graph, then give a name for each axis and include a 
-legend that briefly describes each curve. However, do {\it not\/} 
-include a title above the figure, as the caption already serves
-this function.
-
-Number figures sequentially, placing the figure number and caption
-{\it after\/} the graphics, with at least 0.2~inches (0.508~cm) of 
-space before the caption and 0.3~inches (0.762~cm) after it, as in
-Figure~\ref{sample-figure}. The figure caption should be set in
-10~point type and centered unless it runs two or more lines, in which
-case it should be flush left. You may float a figure to the top or
-bottom of a page, but ideally it should appear on the page that
-first mentions it or on the page immediately afterward.
-
-\subsection{Tables}
-
-You may also want to include tables that summarize material. Like 
-figures, these should be centered, legible, and numbered consecutively. 
-However, place the title {\it above\/} the table with at least 
-0.1~inches (0.254~cm) of space before the title and the same after 
-it, as in Table~\ref{sample-table}. The table title should be set in
-11~point type and centered unless it runs two or more lines, in which
-case it should be flush left.
-
-Tables contain textual material that can be typeset, as contrasted 
-with figures, which contain graphical material that must be drawn. 
-Specify the contents of each column in the table's topmost row. Again,
-you may float a table to the top or bottom of a page, but attempt to 
-place each table on the page that first mentions it or on the one
-immediately afterward.
-
-% Note use of \abovespace and \belowspace to get reasonable spacing 
-% above and below tabular lines.
-
-\def\w{$\oplus$}
-\def\b{$\ominus$}
-\def\h{$\odot$}
-
-\begin{table}[t]
-\vskip -0.15in
-\caption{Adequacy of four models of learning in problem solving in 
-         terms of whether they account (\w), fail to account (\b), 
-         or partially account (\h) for phenomena from \cite{in-book-vanlehn}
-         [{\tt *}] and Jones (1989) [$\diamond$].}
-\label{sample-table}
-\begin{small}
-\begin{center}
-% \vskip -0.10in
-\begin{tabular}{lcccc}
-\hline
-\abovespace\belowspace
-                       & ~~~ACT-R~~~ & ~~~Soar~~~ & {\sc ~~Eureka~~} & {\sc Daedalus} \\
-\hline\abovespace
-Means-ends analysis$^{\diamond}$  &   \h  &     \w    &      \w   &      \w \\
-\hbox{Nonsystematicity$^{\diamond}$\hskip -0.2in} &   \b  &     \b    &      \w   &      \b \\
-Problem isomorphs$^*$   &   \h  &     \h    &      \h   &      \h \\
-Reduced search$^*$ &   \w  &     \w    &      \w   &      \w \\
-Asymmetrical transfer$^*$    &   \w  &     \w    &      \w   &      \w \\
-Einstellung effects$^*$    &   \w  &     \w    &      \w   &      \w \\
-% Strategy differences   &   \h  &     \h    &      \h   &      \h \\
-Reduced verbalization$^*$  &   \w  &     \w    &      \b   &      \b \\
-Reduced solution time$^*$   &   \w  &     \w    &      \b   &      \b \\
-% Self-monitoring        &   \h  &     \h    &      \b   &      \b \\
-Rare analogies$^*$   &   \b  &     \b    &      \h   &      \h \\
-\belowspace
-% \belowstrut{0.10in}
-Superficial analogies$^*$ &   \b  &     \b    &      \w   &      \w \\
-\hline 
-\end{tabular}
-\end{center}
-\vskip -0.10in
-\end{small}
-\end{table}
-
-\subsection{Citations and References}
-
-Please use APA reference format regardless of your formatter or word
-processor. If you rely on the \LaTeX\/ bibliographic facility, use the 
-files {\tt\small cogsysapa.sty} and {\tt\small cogsysapa.bst}, which
-are available on the Web site, to obtain this format. These require
-you to place your references in a separate file with a {\tt\small bib}
-extension, as explained in the \LaTeX\/ manual.
-
-Citations within the text should include the authors' last names and
-year. If the authors' names are included as part of the sentence, place 
-only the year in parentheses, as in Jones and VanLehn (1994), but 
-otherwise place the entire reference in parentheses with the authors
-and year separated by a comma (Jones \& VanLehn, 1984). 
-List multiple references alphabetically and separate them by semicolons 
-\cite{paper-laird-etal,book-newell-simon}. Use the 
-`et~al.'  construction only for citations with four or more authors or
-after listing all authors to a publication in an earlier reference.
-
-Use an unnumbered first-level section heading for the references, and 
-use a hanging indent style, with the first line of the reference flush
-against the left margin and subsequent lines indented by 10 points. 
-The references at the end of this document include examples for journal
-articles \singlecite{article-forbus}, conference publications \singlecite{paper-laird-etal},
-book chapters \singlecite{in-book-vanlehn}, books \singlecite{book-newell-simon}, 
-edited volumes \singlecite{book-shrager-langley}, technical reports 
-\singlecite{techreport-shapiro-etal}, and dissertations \singlecite{phd-choi}.
-
-Alphabetize references in both the text and at its end by the surnames
-of the first authors, with single author entries preceding multiple
-author entries. Order references for the same authors by year of
-publication, with the earliest first.
-
-\section{Electronic Formatting and Submission}
-
-{\it Advances in Cognitive Systems\/} will rely on electronic submission 
-of papers for review and publication. We assume that nearly all authors 
-will have access to \LaTeX\/ or Word to format their documents and can
-use a Web browser to download style files and upload their papers. 
-Authors who do not have such access should send email with their
-concerns to {\sl acs@cogsys.org}.
-
-\subsection{Paper Length and Format}
-
-The title {\it Advances in Cognitive Systems\/} is associated with
-both an electronic journal and an annual conference, with many of the
-articles appearing in the former having been submitted and reviewed
-under the auspices of the latter.
-
-Papers submitted to the annual conference must not exceed sixteen (16)
-pages, including all figures, tables, and references. We will return
-to the authors any submissions that exceed this page limit or that
-diverge significantly from the format specified herein.
-Papers submitted for publication in the journal but not for presentation
-at the meeting are not subject to this constraint. However, we encourage 
-authors to keep their papers the same length as those for the conference.
-
-Submissions may be accompanied by online appendices that contain data,
-demonstrations, instructions for obtaining source code, or the source
-code itself. We encourage authors to include such appendices when they
-submit papers. This material will not count in a submission's page
-length.
-
-Electronic templates for producing the camera-ready copy are available
-for \LaTeX\/ and Microsoft Word. Style files and sample papers are 
-available on the Web at: 
-\vskip 0.1in
-\begin{small}
-\centerline{{\tt http://www.cogsys.org/formats/} .}
-\end{small}
-\vskip 0.1in
-\noindent
-Authors who have questions about these electronic formats should direct
-them to {\sl acs@cogsys.org}.
-
-To ensure the ability to preview and print submissions, authors must
-provide their manuscripts in pdf format. Papers prepared in Word
-should be saved as pdf files and submitted in this format. To support
-the review process, each submission must be accompanied by information
-about the paper's title and abstract, as well as the authors' names
-and physical addresses.
-
-\subsection{Submitting Initial and Final Papers}
-
-Submission of initial and final papers will take place in a purely
-electronic manner through the World Wide Web. Authors who intend to
-submit a paper for presentation at the annual conference should upload
-their file to the submission repository at
-\vskip 0.1in
-\begin{small}
-\centerline{{\tt http://www.cogsys.org/submit/} .}
-\end{small}
-\vskip 0.12in
-\noindent
-Submission should be completed no later than 11:59 PM Pacific time 
-on the conference due date, which will be stated clearly on the 
-meeting Web site. If a submission is late, then it will not be 
-considered for inclusion at the meeting.
-
-{\it Advances in Cognitive Systems\/} will also consider submissions
-that are not intended for presentation at the annual conference. 
-Authors who wish to submit papers of this sort should upload their
-file to the submission repository at
-\vskip 0.1in
-\begin{small}
-\centerline{{\tt http://www.cogsys.org/journal/submit/} .}
-\end{small}
-\vskip 0.08in
-\noindent
-Such submissions may occur at any time, unless they are related to 
-special issues of the journal, which will have their own deadlines.
-
-% You are welcome to submit a paper simultaneously to another meeting 
-% or publication, but only if you indicate this fact clearly on the
-% submission form. Simultaneous submissions that are not clearly
-% specified as such will be rejected on discovery.
-
-Final versions of papers accepted for publication, and resubmissions 
-of conditionally accepted papers, should follow the same format 
-as that for initial submissions. Authors should upload their revised 
-papers to the same Web repository as used for the originals. 
-For the annual conference, final versions should be deposited no 
-later than 11:59 PM Pacific time on the due date. Authors of late 
-submissions will not be provided with a slot to present their work 
-at the meeting and their papers will not appear in the journal.
-Papers not related to the conference may be uploaded at any time.
-
-In both cases, papers that have been conditionally accepted will be
-examined for compliance with conditions placed on their acceptance. 
-Final papers must follow the content, format and length restrictions 
-that are specified in these instructions.
-
-% \newpage
+\section{Discussion}
 

### 2013-08-07 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
fleshing out modelling section

\documentclass[11pt,letterpaper]{article}
\usepackage{cogsys}
\usepackage{cogsysapa}
% \usepackage{apacite}
% \usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage[pdftex]{graphicx} % use this when importing PDF files

% Put back in

 % First page headings for accepted submissions.
% \cogsysheading{1}{2012}{1-18}{9/2012}{12/2012}
 % First page headings for poster submissions.
%\cogsysposterheading{First}{2012}{1-18}

% \ShortHeadings{Formatting Instructions}
              % {P.\ Langley, G.\ Hunt, and D.\ G.\ Shapiro}

\begin{document}

\title{CDSR Human Data/Modelling Paper}

% \author{Pat Langley}{langley@asu.edu}
% \author{Glen Hunt}{glen.hunt@asu.edu}
% \address{Computing Science and Engineering, Arizona State University, 
%          Tempe, AZ 85287 USA}
% \author{Daniel G.\ Shapiro}{dgs@isle.org}
% \address{Institute for the Study of Learning and Expertise, 
%          2164 Staunton Court, Palo Alto, CA 94306 USA}
\vskip 0.2in

\begin{abstract}
Abstract goes here
\end{abstract}

\section{Introduction - KLENK}

\section{Context-Dependent Spatial Regions - KLENK}

\subsection{Explanation of the basic idea}

\subsubsection{Model Desiderata} % (fold)
\label{sec:desiderata}

The desiderata of the model - context, extent, generalises to unseen examples

% subsubsection subsubsection_name (end)

\section{Human Studies - John}
Research Questions: 
\begin{enumerate}
	\item Does context matter?
	\item Do people agree? 
\end{enumerate}

\section{Modeling - Nick}

As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g. understanding spatially-situated task utterances), they must have access to internal models which can represent CDSRs and recognise them in new environments. Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to work. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain code for image region representations~\cite{Freeman:1961}. This is an approach we followed in previous work, where boundary segments connected vertices defined by contextually-appropriate anchor points~\cite{Hawes:2012}.  Whilst our previous approach provided basic performance, the crisp boundaries it generated caused mismatches with human data over the extent of regions. Methods based on underlying spatial functions may be more flexible in this regard, as they can be selectively thresholded depending on context.

% Basic approaches

\textbf{TODO:} Perhaps add related work on robot spatial representations, e.g. maps, rooms.

In the following sections we explore whether two possible CDSR models (one based on boundary segments, one based on a spatial function) are able to satisfy the previously described desiderata for a cognitive systems approach to representing and recognising CDSRs. The surrounding discussion is intended to explore the strengths and weaknesses of these models, providing directions for future research.

\subsection{Symbolic Boundary Method - KLENK}

Overview
Example
Results?
Discussion

\subsection{Kernel Methods - NICK}

Overview - either Chris's approach or something similar
Example
Results?
Discussion 
Where do the points come from

\section{Discussion}

\begin{acknowledgements} 
\noindent
STRANDS etc.
\end{acknowledgements}

\vspace{-0.25in}

{\parindent -10pt\leftskip 10pt\noindent
\bibliographystyle{cogsysapa}
\bibliography{cdsr}

}

% Leave a blank line before the closing brace to ensure the final 
% reference has the proper indentation.

\end{document}

### 2013-08-07 scenarios/cdsr/trunk/papers/ACS2013/main.tex
fleshing out modelling section

@@
 
-\begin{itemize}
-	\item Explanation of the basic idea.
-	\item The desiderata of the model - context, extent, generalises to unseen examples
+\subsection{Explanation of the basic idea}
 
-\end{itemize}
+\subsubsection{Model Desiderata} % (fold)
+\label{sec:desiderata}
+
+The desiderata of the model - context, extent, generalises to unseen examples
+
+% subsubsection subsubsection_name (end)
 
@@
 \section{Modeling - Nick}
-If we want to capture this data in a generative model, what is our hypothesis space. 
-Related work?
+
+As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g.
+understanding spatially-situated task utterances), they must have access to internal models which can represent CDSRs and recognise them in new environments.
+Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
+Regions are central to QSR work~\cite{Cohn:2001}.
+Regions can either have crisp boundaries, usually defined by geometric shapes (e.g.
+the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g.
+the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g.
+those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}).
+The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to work.
+However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain code for image region representations~\cite{Freeman:1961}.
+This is an approach we followed in previous work, where boundary segments connected vertices defined by contextually-appropriate anchor points~\cite{Hawes:2012}.
+Whilst our previous approach provided basic performance, the crisp boundaries it generated caused mismatches with human data over the extent of regions.
+Methods based on underlying spatial functions may be more flexible in this regard, as they can be selectively thresholded depending on context.
+
+% Basic approaches
+
+\textbf{TODO:} Perhaps add related work on robot spatial representations, e.g. maps, rooms.
+
+In the following sections we explore whether two possible CDSR models (one based on boundary segments, one based on a spatial function) are able to satisfy the previously described desiderata for a cognitive systems approach to representing and recognising CDSRs.
+The surrounding discussion is intended to explore the strengths and weaknesses of these models, providing directions for future research.
 
@@
 
-\subsection{Kernel Methods - NICK}
+\subsection{Spatial Function Method - NICK}
 
-Overview - either Chris’s approach or something similar
+Overview - either Chris's approach or something similar
 Example

### 2013-08-08 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
Abstract Draft

@@
 \begin{abstract}
-Abstract goes here
+Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.
+Consider a safe area in a military engagement.
+The size, shape, and location of this area depend the the capabilities of the other agents and their capabilities in the environment.
+We call these regions \textit{context-dependent spatial regions}(CDSRs).
+We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment.
+We conducted a human study to explore how people interpret three types of CDSRs that supports the knowledge integration claim.
+We evaluate two representation techniques against the human data and discuss their limitations as a cognitive systems solution to this problem.
 \end{abstract}
@@
 
-\subsection{Kernel Methods - NICK}
+\subsection{Spatial Function Method - NICK}
 

### 2013-08-10 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 \section{Introduction - KLENK} 
+Natural communication between humans and cognitive systems requires both participants to communicate with spatial language.
+Many spatial terms are defined not only be their geometry, but also their context.
+That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration and functional roles.
+We call such regions \textit{context-dependent spatial regions}(CDSRs).
+
+We begin by discussing CDSRs with examples and two claims about these regions.
+Next, we discuss desired properties of a cognitive system solution.
+The contribution of this paper is a human study exploring our claims about CDSRs and an empirical exploration of current models and their shortcomings.
+We close with a discussion of related and future work.
 
 \section{Context-Dependent Spatial Regions - KLENK}
-
-\subsection{Explanation of the basic idea}
+Consider the following regions: the front of a classroom (\ref{fig:classroom}), safety in military engagement (\ref{fig:safety}), or the geographical feature of a bottleneck (\ref{fig:bottleneck}).
+To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agent's intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
+These regions exists at different levels of spatial resolution.
+Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses
 

### 2013-08-20 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK: added overview description of trial analysis

@@
 \end{enumerate}
+
+Method and Materials 
+Trial 1
+Trial 2
+Trial 3
+
+-----------------
+
+Task 1: The participant draws a polygon denoting the region
+Analysis of results:
+(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+Within the data for each trial,
+-- For each participant, generate the intersection and union polygon of all the other participants
+---- Generate N random points in the area of the full stimulus
+---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
+---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
+Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
+(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
+
+This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
+
+(2) Does context matter?
+I understand that all the rooms stimuli have the same extent.
+So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
+We are hoping for a low kappa score
+
+-----------------
+
+Task 2: The participant selected a point denoting the sweetspot of the region
+Analysis of results:
+(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
+We would hope that the variance of the subjects responses is lower than that of randomly generated points.
+After the initial comparison with the random point sets we may exclude participants based on extreme variance.
+
+(2) Does context matter?
+Do the clusters of participants responses move?
+- for each experimental stimulus compute the mean x and y location for the responses for that trial
+- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.
+
+-----------------
+
+Task 3: The participant grades a point with respect to membership of the region on a likert scale
+I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
+I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
+Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
+(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
+(b) distance from the mean sweet spot for that stimulus from task 2
+
+-----------------
 

### 2013-08-26 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-We begin by discussing CDSRs with examples and two claims about these regions.
+We begin by discussing CDSRs with examples.
+We claim that (1) people can effectively communicate with these regions, and (2) the size and shape of these regions change with the context.
 Next, we discuss desired properties of a cognitive system solution.
@@
 \section{Context-Dependent Spatial Regions - KLENK}
-Consider the following regions: the front of a classroom (\ref{fig:classroom}), safety in military engagement (\ref{fig:safety}), or the geographical feature of a bottleneck (\ref{fig:bottleneck}).
+Consider the following regions (shown in Figure \ref{examples}): the front of a classroom (\ref{fig:classroom}), safety in military engagement (\ref{fig:safety}), or the geographical feature of a bottleneck (\ref{fig:bottleneck}).
 To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agent's intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
-These regions exists at different levels of spatial resolution.
-Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses
+Furthermore, these regions exists at different levels of spatial resolution.
+Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.
 
-\subsubsection{Model Desiderata} % (fold)
-\label{sec:desiderata}
+%Combine these figures into a single figure across
+\begin{figure}
+  \includegraphics[width=\columnwidth*.3]{figures/classroom.png}
+  \caption{The front of the classroom is the area between the desks and the whiteboard where people give presentations.}
+  \label{fig:classroom}
+\end{figure}
 
-The desiderata of the model - context, extent, generalises to unseen examples
+\begin{figure}
+  \includegraphics[width=\columnwidth*.3]{figures/safety-3.jpg}
+  \caption{Safety for the scout is defined by the location of the friendly turrets and the capabilities of the opposing force.}
+  \label{fig:safety}
+\end{figure}
 
-% subsubsection subsubsection_name (end)
+\begin{figure}
+  \includegraphics[width=\columnwidth*.3]{figures/bottleneck.JPG}
+  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
+  \label{fig:safety}
+\end{figure}
+
+Understanding these regions is crucial to joint intelligent action with humans.
+Not only are these regions frequently tied to goals and commands: "go to the front of the classroom", "defend the bottleneck", and "get to safety", but they are also central to understanding human activities: e.g., "Is there a class in session?".
+
+Representing and reasoning about these regions is a Cognitive Systems problem. for the following reasons:
+\begin{itemize}
+\item Understanding context requires integration of different types of knowledge including geometric, semantic, and functional knowledge about the environment.
+\item The representation should be transferable to new situations.
+While humans do not demonstrate mastery after a single example of concept, they are able to begin using it immediately refining it over time.
+\item The representation should be applicable in different reasoning tasks.  The functional nature of these regions makes them important for many tasks from action planning to activity recognition.
+\end{itemize}
 
@@
 \subsection{Symbolic Boundary Method - KLENK}
+In our previous work \cite{Hawes:2012}, we presented an approach on a mobile robot that combined qualitative spatial relations, object recognition, and analogy to transfer CDSRs represented as polygons defined by \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11}.
+Anchor points are symbolic descriptions that link a conceptual entity (e.g., the front of a classroom) to perceived entities (e.g., the lower left extent of a desk).
+An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.
+In the know example, a human defines the front of the classroom as a polygon consisting of four anchor points.
+Analogical transfer works by aligning structured representations and transferring partially aligned statements from the known example to the new environment.
+In this case, the anchor points are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.
 
-Overview
-Example
-Results?
-Discussion
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
+  \caption{The known example defines the front of the room with four anchor points.
+By transferring the anchor points to the new environment, the system is able to ground the room's front in its perception.}
+  \label{fig:anchor-point-transfer}
+\end{figure}
+
+This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task.
+
+Applying this model to the points and regions indicated in our experiment illustrates two limitations.
+Regions do not always physically bounded by objects.
+In the classroom examples, people frequently did not extend the front of the room all way into the corners.
+In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
+This indicates that people consider the dynamics of the situation when evaluating these terms, which is not accounted for in this model.
 

### 2013-08-29 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 \usepackage[pdftex]{graphicx} % use this when importing PDF files
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
-We begin by discussing CDSRs with examples and two claims about these regions.
-Next, we discuss desired properties of a cognitive system solution.
-The contribution of this paper is a human study exploring our claims about CDSRs and an empirical exploration of current models and their shortcomings.
-We close with a discussion of related and future work.
+We begin by discussing CDSRs with examples.
+We claim that (1) people can effectively communicate with these regions, and (2) the size and shape of these regions change with the context.
+We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.
+To explore these claims, we present the results of a human study, in which, people marked and rated points and regions for three different regions in 13 different contexts.
+In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
+Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also, to enable robust intelligent assistants in human environments.
 
-\section{Context-Dependent Spatial Regions - KLENK}
-Consider the following regions: the front of a classroom (\ref{fig:classroom}), safety in military engagement (\ref{fig:safety}), or the geographical feature of a bottleneck (\ref{fig:bottleneck}).
+\section{Context-Dependent Spatial Regions}
+Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.
 To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agent's intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
-These regions exists at different levels of spatial resolution.
-Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses
+Furthermore, these regions exists at different levels of spatial resolution.
+Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.
 
-\subsubsection{Model Desiderata} % (fold)
-\label{sec:desiderata}
+%Combine these figures into a single figure across
+\begin{figure}
+\centering
+%% Change the classroom to be a picture stimuli
+  \begin{subfigure}[b]{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/classroom.png}
+  \caption{The front of the classroom is the area between the desks and the whiteboard where people give presentations.}
+  \end{subfigure}
+\begin{subfigure}[b]{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/safety-3.jpg}
+  \caption{Safety for the scout is defined by the location of the friendly turrets and the capabilities of the opposing force.}
+\end{subfigure}
+\begin{subfigure}[b]{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
+  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
+  \label{fig:safety}
+  \end{subfigure}
 
-The desiderata of the model - context, extent, generalises to unseen examples
+  \caption{Examples of spatial regions used as stimuli in our experiments.}
+  \label{fig:examples}
+\end{figure}
 
-% subsubsection subsubsection_name (end)
+Understanding these regions is crucial to joint intelligent action with humans.
+Not only are these regions frequently tied to goals and commands: "go to the front of the classroom", "defend the bottleneck", and "get to safety", but they are also central to understanding human activities: e.g., "Is there a class in session?".
+
+The long-term goal of this work is to create of agents that can naturally interact with people in human environments.
+While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.
+First, the representational approach should be transferable to new situations.
+While humans do not demonstrate mastery after a single example of concept, they are able to begin using it immediately refining it over time.
+Second, the representation should be applicable in different reasoning tasks.
+The functional nature of these regions makes them important for many tasks from action planning to activity recognition.
 

### 2013-09-03 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.
-Consider a safe area in a military engagement.
-The size, shape, and location of this area depend the the capabilities of the other agents and their capabilities in the environment.
-We call these regions \textit{context-dependent spatial regions}(CDSRs).
-We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment.
-We conducted a human study to explore how people interpret three types of CDSRs that supports the knowledge integration claim.
+Consider an agent asked to remain "in the safe zone" during a military engagement.
+The size, shape, and location of this area depend the the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
+We call these types of regions \textit{context-dependent spatial regions}(CDSRs) due to the context-dependent nature of their extent.
+We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.
+We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.
 We evaluate two representation techniques against the human data and discuss their limitations as a cognitive systems solution to this problem.
@@
 \section{Introduction - KLENK} 
-Natural communication between humans and cognitive systems requires both participants to communicate with spatial language.
-Many spatial terms are defined not only be their geometry, but also their context.
+Natural collaboration between humans and cognitive systems requires both participants to communicate using a shared understanding of spatial language.
+Cognitive systems can do quite well with spatial terms defined in terms of coordinates on a map, or that can be specified using known landmarks.
+However, many spatial terms are defined not only be their geometry, but also their context and their relevance to current goals.
 That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration and functional roles.
@@
 
-We begin by discussing CDSRs with examples.
-We claim that (1) people can effectively communicate with these regions, and (2) the size and shape of these regions change with the context.
-Next, we discuss desired properties of a cognitive system solution.
-The contribution of this paper is a human study exploring our claims about CDSRs and an empirical exploration of current models and their shortcomings.
-We close with a discussion of related and future work.
+We begin by discussing the features of CDSRs using illustrative examples.
+We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.
+We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.
+To explore these claims, we present the results of a human study, in which, people marked and rated points and regions for three different environments in 13 different contexts.
+In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
+Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.
 
-\section{Context-Dependent Spatial Regions - KLENK}
-Consider the following regions (shown in Figure \ref{examples}): the front of a classroom (\ref{fig:classroom}), safety in military engagement (\ref{fig:safety}), or the geographical feature of a bottleneck (\ref{fig:bottleneck}).
-To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agent's intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
-Furthermore, these regions exists at different levels of spatial resolution.
+\section{Context-Dependent Spatial Regions}
+Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.
+To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
+Furthermore, these regions exist at different levels of spatial resolution.
 Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth*.3]{figures/classroom.png}
+\centering
+%% Change the classroom to be a picture stimuli
+  \begin{subfigure}[b]{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/classroom.png}
   \caption{The front of the classroom is the area between the desks and the whiteboard where people give presentations.}
-  \label{fig:classroom}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth*.3]{figures/safety-3.jpg}
+  \end{subfigure}
+\begin{subfigure}[b]{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/safety-3.jpg}
   \caption{Safety for the scout is defined by the location of the friendly turrets and the capabilities of the opposing force.}
-  \label{fig:safety}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth*.3]{figures/bottleneck.JPG}
+\end{subfigure}
+\begin{subfigure}[b]{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
   \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
   \label{fig:safety}
+  \end{subfigure}
+
+  \caption{Examples of spatial regions used as stimuli in our experiments.}
+  \label{fig:examples}
 \end{figure}
@@
 
-Representing and reasoning about these regions is a Cognitive Systems problem. for the following reasons:
-\begin{itemize}
-\item Understanding context requires integration of different types of knowledge including geometric, semantic, and functional knowledge about the environment.
-\item The representation should be transferable to new situations.
+The long-term goal of this work is to create of agents that can naturally interact with people in human environments.
+While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.
+First, the representational approach should be transferable to new situations.
 While humans do not demonstrate mastery after a single example of concept, they are able to begin using it immediately refining it over time.
-\item The representation should be applicable in different reasoning tasks.  The functional nature of these regions makes them important for many tasks from action planning to activity recognition.
-\end{itemize}
+Second, the representation should be applicable in different reasoning tasks.
+The functional nature of these regions makes them important for many tasks from action planning to activity recognition.
 

### 2013-09-06 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK - added design, stimuli and trial 1 analsysis sections

@@
 \section{Human Studies - John}
-Research Questions: 
+
+\subsection{Research Question, Design and Stimuli}
+
+The study was designed to examine the following questions about spatial regions:
 \begin{enumerate}
-	\item Does context matter?
-	\item Do people agree? 
+	\item Given a particular situation how well do people agree on the location and extent of a named spatial region?
+	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named spatial region?
 \end{enumerate}
 
-Method and Materials 
-Trial 1
+To examine these questions we developed three different trials.
+In all three trial types the subject was presented with a visual stimuli (such as a room floor plan or a map) and an accompanying linguistic description (such as \emph{go to the front of the room}, however the way the task the subject was set varied across the trials:
+\begin{enumerate}
+	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the spatial region described in the description or to select a no region option.
+	\item Trial type 2: a location was marked on the visual stimuli (using a red x) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
+Three points were pre-selected for each stimuli such that one point was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location.
+The selection of these points was informed by the results of a pilot study.
+	\item Trial type 3: subjects were asked to mark on the visual stimuli, by clicking with the mouse, the location that they felt was most prototypical of the spatial description. 
+\end{enumerate}
+
+All three trial types use the same visual stimuli and accompanying spatial descriptions.
+Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment.
+The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.
+
+The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations.
+For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom.
+We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a prototypical layout with the chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and a prototypical layout with the chairs far away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
+To control for left-right bias effects two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts.
+Conseqeuntly, there were 7 classroom stimuli (circle, U $\times 2$, close $\times 2$, far $\times 2$).
+In designing these stimuli we deliberately chose some layouts that were very prototypical or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}.
+By varying the prototypicality of the layouts we hoped to get some variation across the between subject agreement on a per stimuli basis, for example we expected to get a higher between subject agreement for the stimuli that were more prototypical.
+This would provide insight into research question 1, in particular it would point to the effect of prototypicality on the semantics of contextually defined spatial regions.
+Also, because the prototypicality of the scenes was varied by adjusting the location and orientation of objects in the scenes, a variation in agreement across the stimuli would provide evidence that the orientation and location of objects do effect the location and extent of spatial regions.
+
+The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
+Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
+Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.
+
+\begin{figure}
+%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+\centerline{
+\begin{tabular}{cccc}
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle.pdf}\caption{}\label{fig:classroom-a}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-photo.jpg}\caption{}\label{fig:classroom-b}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-shaped.pdf}\caption{}\label{fig:classroom-c}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-photo.jpg}\caption{}\label{fig:classroom-d}\end{subfigure}\\
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-close.pdf}\caption{}\label{fig:classroom-f}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-far-away.pdf}\caption{}\label{fig:classroom-e}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-1.jpg}\caption{}\label{fig:bottleneck-a}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2.jpg}\caption{}\label{fig:bottleneck-b}\end{subfigure}\\
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close.jpg}\caption{}\label{fig:tanks-a}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-far.jpg}\caption{}\label{fig:tanks-b}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets.jpeg}\caption{}\label{fig:turrets-a}\end{subfigure}&
+\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets-plus-one.jpeg}\caption{}\label{fig:turrets-b}\end{subfigure}
+\end{tabular}
+}
+\caption{Experiment Stimuli}
+\label{fig:exp_stimuli}
+\end{figure}
+
+\subsection{Participants}
+
+43 undergraduate students from two different third level institutes took part in the study.
+
+\subsection{Analysis}
+
+%Task 1: The participant draws a polygon denoting the region
+%Analysis of results:
+%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+%Within the data for each trial,
+%-- For each participant, generate the intersection and union polygon of all the other participants
+%---- Generate N random points in the area of the full stimulus
+%---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
+%---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
+Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
+%(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
+%This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
+%(2) Does context matter?
+%I understand that all the rooms stimuli have the same extent.
+%So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
+%We are hoping for a low kappa score
+
+%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
+%\begin{equation*}
+%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
+%\end{equation*}
+%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators. 
+In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
+In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
+
+One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$.
+Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes.
+$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
+In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distribute across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
+After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
+In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
+Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.
+It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$.
+A $\kappa$ score between $0.3-0.4$ would be considered as being poor to fair.
+This would indicate that there was some agreement between subjects but less that would be expected.
+However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
+Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
+In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no regrion responses for each stimuli.
+It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also relationship between typicality and the number of no region responses.
+Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
+Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
+Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
+$\kappa$ scores in this range can to interpreted as fair to good.
+
+\begin{figure}
+%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+\centerline{
+\begin{tabular}{ccc}
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
+\end{tabular}
+}
+\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
+\label{fig:type-1-hists}
+\end{figure}
+
+\begin{table}
+\centerline{
+\begin{tabular}{l c c}
+Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+\hline
+\hline
+(h) & 0.5034 & 0\\ % 1_2
+(g) & 0.4585 & 1\\ % 1_1
+(k) & 0.4548 & 0\\ % 3_3
+(f) & 0.3930 & 0\\ % 2_4
+(i) & 0.3906 & 0\\ % 3_1
+(e-flipped) & 0.3813 & 2\\ % 2_3
+(j) & 0.3787 & 1\\ % 3_2
+(f-flipped) & 0.3720 & 2\\ % 2_5
+(e) & 0.3629 & 1\\ %2_2
+(l) & 0.3164& 0\\ % 3_4
+(c) & 0.2438 & 2\\ % 2_6
+(c-flipped) & 0.2414 & 4\\ %2_7
+(a) & 0.0996 & 10 \\ % 2_1
+\hline
+\end{tabular}
+}
+\label{tab:stimuli-kappa-scores}
+\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.
+The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
+\end{table}
+
+The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
+Furthermore, there results support the hypothesis that the defintion of spatial regions in a given context is effected by the location and orientation of objects in the context.
+For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}.
+Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.
+
 Trial 2
@@
 -----------------
-
-Task 1: The participant draws a polygon denoting the region
-Analysis of results:
-(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
-Within the data for each trial,
--- For each participant, generate the intersection and union polygon of all the other participants
----- Generate N random points in the area of the full stimulus
----- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
----- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
-Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
-(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
-
-This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
-
-(2) Does context matter?
-I understand that all the rooms stimuli have the same extent.
-So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
-We are hoping for a low kappa score
 

### 2013-09-06 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk fixed reference

@@
 
-We begin by discussing CDSRs with examples.
-We claim that (1) people can effectively communicate with these regions, and (2) the size and shape of these regions change with the context.
+We begin by discussing the features of CDSRs using illustrative examples.
+We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.
 We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.
-To explore these claims, we present the results of a human study, in which, people marked and rated points and regions for three different regions in 13 different contexts.
+To explore these claims, we present the results of a human study, in which, people marked and rated points and regions for three different environments in 13 different contexts.
 In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
-Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also, to enable robust intelligent assistants in human environments.
+Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.
 
@@
 Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.
-To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agent's intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
-Furthermore, these regions exists at different levels of spatial resolution.
+To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
+Furthermore, these regions exist at different levels of spatial resolution.
 Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.

### 2013-09-06 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk fixed some typos

@@
 This would indicate that there was some agreement between subjects but less that would be expected.
-However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
+However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
 In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no regrion responses for each stimuli.
-It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also relationship between typicality and the number of no region responses.
+It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
 Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.

### 2013-09-07 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK extended discussion of likert scale experiment

@@
 \usepackage{subcaption}
+\usepackage{multirow}
 
@@
 	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the spatial region described in the description or to select a no region option.
-	\item Trial type 2: a location was marked on the visual stimuli (using a red x) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-Three points were pre-selected for each stimuli such that one point was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location.
-The selection of these points was informed by the results of a pilot study.
+	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
 	\item Trial type 3: subjects were asked to mark on the visual stimuli, by clicking with the mouse, the location that they felt was most prototypical of the spatial description. 
@@
 
-Trial 2
-Trial 3
+The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
+For each stimuli domain (classroom, map, military engagement) three points were pre-selected such that for each particular stimuli one of the points was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location.
+For clarity it is worth emphaising that, although the same three points were used for all the stimuli from a particular domain, the particular point that acted as the sweetspot for a particular stimuli within a domain set would vary from stimuli to stimuli.
+The selection of these points was informed by the results of a pilot study.
+Figure \ref{fig:locations} illustrates the points selected for each domain and overlays them on one of the visual stimuli from each domain.
+Table \ref{tab:point-likert-scores} list the average likert scores for the different points in each domain.
+In the map domain the points are named from top-left to bottom-right (e.g., $Pt1$ is the point in the top left of Figure \ref{fig:bottleneck-points}); in the classroom domain the points are named in order from left to right (e.g.
+$Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}).
+
+\begin{figure}
+%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+\centerline{
+\begin{tabular}{ccc}
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
+\end{tabular}
+}
+\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
+\label{fig:locations}
+\end{figure}
+
+\begin{table}
+\centerline{
+\begin{tabular}{l c c c l}
+\hline
+Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
+Stim ID & P1 & P2 & P3 & Significance\\
+(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & ToDo\\ % 1_1
+(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & ToDo\\ % 1_2
+\hline
+Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
+Stim ID & P1 & P2 & P3 & Significance\\
+(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & ToDo\\ % 2_1
+(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & ToDo\\ % 2_6
+(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & ToDo\\ %2_7
+(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & ToDo\\ %2_2
+(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & ToDo\\ % 2_3
+(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & ToDo\\ % 2_4
+(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & ToDo\\\\ % 2_5
+\hline
+Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
+Stim ID & P1 & P2 & P3 & Significance\\
+(i) & $1.1951$ & $4.3902$ & $\mathbf{4.6428}$ & ToDo\\ % 3_1
+(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & ToDo\\ % 3_2
+(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & ToDo\\ % 3_3
+(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & ToDo\\ % 3_4
+\hline
+\hline
+\end{tabular}
+}
+\label{tab:point-likert-scores}
+\caption{Need to put in caption!!!}
+\end{table}
 

### 2013-09-07 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK added to likert scale section

@@
 $Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}).
+For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.
 
@@
 Stim ID & P1 & P2 & P3 & Significance\\
-(i) & $1.1951$ & $4.3902$ & $\mathbf{4.6428}$ & ToDo\\ % 3_1
+(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & ToDo\\ % 3_1
 (j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & ToDo\\ % 3_2

### 2013-09-08 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK added results of t-tests for likert scale results

@@
 
-Trial 2
-Trial 3
+The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
+For each stimuli domain (classroom, map, military engagement) three points were pre-selected such that for each particular stimuli one of the points was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location.
+For clarity it is worth emphaising that, although the same three points were used for all the stimuli from a particular domain, the particular point that acted as the sweetspot for a particular stimuli within a domain set would vary from stimuli to stimuli.
+The selection of these points was informed by the results of a pilot study.
+Figure \ref{fig:locations} illustrates the points selected for each domain and overlays them on one of the visual stimuli from each domain.
+Table \ref{tab:point-likert-scores} list the average likert scores for the different points in each domain.
+In the map domain the points are named from top-left to bottom-right (e.g., $Pt1$ is the point in the top left of Figure \ref{fig:bottleneck-points}); in the classroom domain the points are named in order from left to right (e.g.
+$Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}).
+For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.
+
+\begin{figure}
+%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+\centerline{
+\begin{tabular}{ccc}
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
+\end{tabular}
+}
+\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
+\label{fig:locations}
+\end{figure}
+
+\begin{table}
+\centerline{
+\begin{tabular}{l c c c l}
+\hline
+Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
+Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
+(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
+\hline
+Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
+Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
+(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
+(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
+(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
+(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
+(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
+(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\\\ % 2_5
+\hline
+Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
+Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
+(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
+(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
+(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
+\hline
+\hline
+\end{tabular}
+}
+\label{tab:point-likert-scores}
+\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
+\end{table}
+
+\begin{table}
+\centerline{
+\begin{tabular}{c c c c }
+\hline
+Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
+$H_0$ & $H_A$ & $\alpha$ & p-value\\
+$\mu(c.P1)=\mu(c-flipped.P3)$ & $\mu(c.P1)!=\mu(c-flipped.P3)$ & $0.05$ & $0.6315$\\
+$\mu(e.P1)=\mu(e-flipped.P3)$ & $\mu(e.P1)!=\mu(e-flipped.P3)$ & $0.05$ & $0.7358$\\
+$\mu(f.P1)=\mu(f-flipped.P3)$ & $\mu(f.P1)!=\mu(f-flipped.P3)$ & $0.05$ & $0.2592$\\
+\hline
+\hline
+\end{tabular}
+}
+\label{tab:flipped-point-likert-scores}
+\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
+\end{table}
 

### 2013-09-09 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK updated the text introducing the likert scale results and the comparision of the sweetspot across the flipped trials

@@
 Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment.
+Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
 The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.
@@
 
-\subsection{Participants}
+\subsection{Participants and Procedure}
 
 43 undergraduate students from two different third level institutes took part in the study.
+To control for sequence effects the presentation order of the trials were randomised each participant.
+Also in order to minimise fatigue of the subjects for the Type 2 Likert trials each subject only ranked a randomly selected subset of the trials.
 
@@
 The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-For each stimuli domain (classroom, map, military engagement) three points were pre-selected such that for each particular stimuli one of the points was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location.
-For clarity it is worth emphaising that, although the same three points were used for all the stimuli from a particular domain, the particular point that acted as the sweetspot for a particular stimuli within a domain set would vary from stimuli to stimuli.
-The selection of these points was informed by the results of a pilot study.
-Figure \ref{fig:locations} illustrates the points selected for each domain and overlays them on one of the visual stimuli from each domain.
-Table \ref{tab:point-likert-scores} list the average likert scores for the different points in each domain.
-In the map domain the points are named from top-left to bottom-right (e.g., $Pt1$ is the point in the top left of Figure \ref{fig:bottleneck-points}); in the classroom domain the points are named in order from left to right (e.g.
-$Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}).
+For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
+
+The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
+One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped.
+In the classroom domain the expected sweetspot varied between the points depending on the stimuli.
+In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets.
+P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+
+Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain.
 For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.
+In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
+For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests.
+There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
+Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli.
+In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
+
+As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis.
+These flipped stimuli included the classroom stimuli (c), (e) and (f).
+This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials.
+Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3.
+To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
+Table \ref{tab:flipped-point-likert-scores} lists the results of this trial.
+In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
 
@@
 Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
-Stim ID & P1 & P2 & P3 & Significance\\
-(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & ToDo\\ % 1_1
-(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & ToDo\\ % 1_2
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
+Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
+(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
+\\
 \hline
 Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
-Stim ID & P1 & P2 & P3 & Significance\\
-(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & ToDo\\ % 2_1
-(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & ToDo\\ % 2_6
-(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & ToDo\\ %2_7
-(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & ToDo\\ %2_2
-(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & ToDo\\ % 2_3
-(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & ToDo\\ % 2_4
-(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & ToDo\\\\ % 2_5
+ ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
+Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
+(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
+(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
+(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
+(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
+(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
+(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
+\\
 \hline
 Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
-Stim ID & P1 & P2 & P3 & Significance\\
-(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & ToDo\\ % 3_1
-(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & ToDo\\ % 3_2
-(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & ToDo\\ % 3_3
-(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & ToDo\\ % 3_4
+ ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
+Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
+(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
+(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
+(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
+\\
 \hline
@@
 \label{tab:point-likert-scores}
-\caption{Need to put in caption!!!}
+\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
 \end{table}
 
------------------
+\begin{table}
+\centerline{
+\begin{tabular}{c c c c }
+\hline
+Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
+$H_0$ & $H_A$ & $\alpha$ & p-value\\
+$\mu(c.P1)=\mu(c-flipped.P3)$ & $\mu(c.P1)!=\mu(c-flipped.P3)$ & $0.05$ & $0.6315$\\
+$\mu(e.P1)=\mu(e-flipped.P3)$ & $\mu(e.P1)!=\mu(e-flipped.P3)$ & $0.05$ & $0.7358$\\
+$\mu(f.P1)=\mu(f-flipped.P3)$ & $\mu(f.P1)!=\mu(f-flipped.P3)$ & $0.05$ & $0.2592$\\
+\hline
+\hline
+\end{tabular}
+}
+\label{tab:flipped-point-likert-scores}
+\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
+\end{table}
 
------------------
+Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular situation depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted situation was consistent across similar situations, as shown by the fact that the sweetspot flipped from P1 to P3 when the situation was flipped.
 
-Task 2: The participant selected a point denoting the sweetspot of the region
-Analysis of results:
-(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
-Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
-We would hope that the variance of the subjects responses is lower than that of randomly generated points.
-After the initial comparison with the random point sets we may exclude participants based on extreme variance.
+%-----------------
+%
+%Task 2: The participant selected a point denoting the sweetspot of the region
+%Analysis of results:
+%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+%Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
+%We would hope that the variance of the subjects responses is lower than that of randomly generated points.
+%After the initial comparison with the random point sets we may exclude participants based on extreme variance.
+% 
+%(2) Does context matter?
+%Do the clusters of participants responses move?
+%- for each experimental stimulus compute the mean x and y location for the responses for that trial
+%- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.
 
-(2) Does context matter?
-Do the clusters of participants responses move?
-- for each experimental stimulus compute the mean x and y location for the responses for that trial
-- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.
-
------------------
-
-Task 3: The participant grades a point with respect to membership of the region on a likert scale
-I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
-I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
-Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
-(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
-(b) distance from the mean sweet spot for that stimulus from task 2
-
------------------
+%-----------------
+%
+%
+%Task 3: The participant grades a point with respect to membership of the region on a likert scale
+%I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
+%I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
+%Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
+%(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
+%(b) distance from the mean sweet spot for that stimulus from task 2
+%
+%-----------------
 

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 Consider an agent asked to remain "in the safe zone" during a military engagement.
-The size, shape, and location of this area depend the the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
+The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
 We call these types of regions \textit{context-dependent spatial regions}(CDSRs) due to the context-dependent nature of their extent.
@@
 We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.
-We evaluate two representation techniques against the human data and discuss their limitations as a cognitive systems solution to this problem.
+We discss the implications for these results on cognitive systems development.
 \end{abstract}
@@
 We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.
-To explore these claims, we present the results of a human study, in which, people marked and rated points and regions for three different environments in 13 different contexts.
+To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
 In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
@@
 Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.
-To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
+To identify these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
 Furthermore, these regions exist at different levels of spatial resolution.
@@
   \includegraphics[width=\textwidth]{figures/classroom.png}
-  \caption{The front of the classroom is the area between the desks and the whiteboard where people give presentations.}
+  \caption{\textit{The front of the classroom} could be defined as the area between the desks and the whiteboard where people give presentations.}
   \end{subfigure}
@@
   \includegraphics[width=\textwidth]{figures/safety-3.jpg}
-  \caption{Safety for the scout is defined by the location of the friendly turrets and the capabilities of the opposing force.}
+  \caption{\textit{Safety for the scout} is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
 \end{subfigure}
@@
   \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
-  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
+  \caption{A \textit{bottleneck} occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
   \label{fig:safety}
@@
 
-The long-term goal of this work is to create of agents that can naturally interact with people in human environments.
+The long-term goal of this work is to create intelligent agents that can naturally interact with people in human environments.
 While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.
 First, the representational approach should be transferable to new situations.
-While humans do not demonstrate mastery after a single example of concept, they are able to begin using it immediately refining it over time.
+While humans do not demonstrate mastery after a single example of a concept, they are able to begin using it immediately refining it over time.
 Second, the representation should be applicable in different reasoning tasks.
@@
 \begin{enumerate}
-	\item Given a particular situation how well do people agree on the location and extent of a named spatial region?
-	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named spatial region?
+	\item Given a particular situation how well do people agree on the location and extent of a named context dependent spatial region?
+	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named context dependent spatial region?
 \end{enumerate}
 
-To examine these questions we developed three different trials.
-In all three trial types the subject was presented with a visual stimuli (such as a room floor plan or a map) and an accompanying linguistic description (such as \emph{go to the front of the room}, however the way the task the subject was set varied across the trials:
-\begin{enumerate}
-	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the spatial region described in the description or to select a no region option.
-	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-	\item Trial type 3: subjects were asked to mark on the visual stimuli, by clicking with the mouse, the location that they felt was most prototypical of the spatial description. 
-\end{enumerate}
-
-All three trial types use the same visual stimuli and accompanying spatial descriptions.
+To examine these questions, we developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).
 Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment.
-The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.
+Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
+The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and (3) the military engagement stimuli were always accompanied by the description \emph{safety}.
 
@@
 For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom.
-We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a prototypical layout with the chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and a prototypical layout with the chairs far away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
+We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
 To control for left-right bias effects two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts.
 Conseqeuntly, there were 7 classroom stimuli (circle, U $\times 2$, close $\times 2$, far $\times 2$).
-In designing these stimuli we deliberately chose some layouts that were very prototypical or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}.
-By varying the prototypicality of the layouts we hoped to get some variation across the between subject agreement on a per stimuli basis, for example we expected to get a higher between subject agreement for the stimuli that were more prototypical.
-This would provide insight into research question 1, in particular it would point to the effect of prototypicality on the semantics of contextually defined spatial regions.
-Also, because the prototypicality of the scenes was varied by adjusting the location and orientation of objects in the scenes, a variation in agreement across the stimuli would provide evidence that the orientation and location of objects do effect the location and extent of spatial regions.
-
-The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
-Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
-Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.
+In designing these stimuli we deliberately chose some layouts that were very common or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}.
+By varying the typicality of the layouts we hoped to get some variation across the between subject agreement on a per stimuli basis, for example we expected to get a higher between subject agreement for the stimuli that were more common.
+This would provide insight into research question 1, in particular it would point to the effect of typicality on the semantics of contextually defined spatial regions.
+Also, because the typicality of the scenes was varied by adjusting the location and orientation of objects in the scenes, a variation in agreement across the stimuli would provide evidence that the orientation and location of objects do effect the location and extent of spatial regions.
 
@@
 
-\subsection{Participants}
+The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
+Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
+Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.
+
+Using these stimuli pairs, we developed an online experiment with two different types of trials.
+\begin{enumerate}
+	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
+The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
+Subjects drew polygons by clicking to position the verticies with lines being automatically generated between each point indicated as the subject drew.
+There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
+	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
+The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
+One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped.
+In the classroom domain the expected sweetspot varied between the points depending on the stimuli.
+In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets.
+P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+\end{enumerate}
+
+\subsection{Participants and Procedure}
 
 43 undergraduate students from two different third level institutes took part in the study.
+To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.
 
@@
 $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
-In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distribute across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
+In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
 After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
@@
 The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
-Furthermore, there results support the hypothesis that the defintion of spatial regions in a given context is effected by the location and orientation of objects in the context.
+Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
 For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}.
@@
 The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-For each stimuli domain (classroom, map, military engagement) three points were pre-selected such that for each particular stimuli one of the points was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location.
-For clarity it is worth emphaising that, although the same three points were used for all the stimuli from a particular domain, the particular point that acted as the sweetspot for a particular stimuli within a domain set would vary from stimuli to stimuli.
-The selection of these points was informed by the results of a pilot study.
-Figure \ref{fig:locations} illustrates the points selected for each domain and overlays them on one of the visual stimuli from each domain.
-Table \ref{tab:point-likert-scores} list the average likert scores for the different points in each domain.
-In the map domain the points are named from top-left to bottom-right (e.g., $Pt1$ is the point in the top left of Figure \ref{fig:bottleneck-points}); in the classroom domain the points are named in order from left to right (e.g.
-$Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}).
+For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
+
+Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain.
 For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.
+In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
+For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests.
+There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
+Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli.
+In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
+
+As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis.
+These flipped stimuli included the classroom stimuli (c), (e) and (f).
+This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials.
+Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3.
+To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
+Table \ref{tab:flipped-point-likert-scores} lists the results of this trial.
+In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
 
@@
 (h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
+\\
 \hline
@@
 (f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
-(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\\\ % 2_5
+(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
+\\
 \hline
@@
 (l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
+\\
 \hline
@@
 
------------------
+Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular situation depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted situation was consistent across similar situations, as shown by the fact that the sweetspot flipped from P1 to P3 when the situation was flipped.
 
------------------
+%-----------------
+%
+%Task 2: The participant selected a point denoting the sweetspot of the region
+%Analysis of results:
+%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+%Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
+%We would hope that the variance of the subjects responses is lower than that of randomly generated points.
+%After the initial comparison with the random point sets we may exclude participants based on extreme variance.
+% 
+%(2) Does context matter?
+%Do the clusters of participants responses move?
+%- for each experimental stimulus compute the mean x and y location for the responses for that trial
+%- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.
 
-Task 2: The participant selected a point denoting the sweetspot of the region
-Analysis of results:
-(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
-Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
-We would hope that the variance of the subjects responses is lower than that of randomly generated points.
-After the initial comparison with the random point sets we may exclude participants based on extreme variance.
-
-(2) Does context matter?
-Do the clusters of participants responses move?
-- for each experimental stimulus compute the mean x and y location for the responses for that trial
-- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.
-
------------------
-
-Task 3: The participant grades a point with respect to membership of the region on a likert scale
-I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
-I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
-Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
-(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
-(b) distance from the mean sweet spot for that stimulus from task 2
-
------------------
+%-----------------
+%
+%
+%Task 3: The participant grades a point with respect to membership of the region on a likert scale
+%I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
+%I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
+%Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
+%(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
+%(b) distance from the mean sweet spot for that stimulus from task 2
+%
+%-----------------
 

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-Using these stimuli pairs, we developed an online experiment with two different types of trials.
-\begin{enumerate}
-	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
+\subsection{Participants and Procedure}
+Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
+43 undergraduate students from two different third level institutes took part in the study.
+To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.
+
+\subsection{Trial Type 1 - Design}
+In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
 The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
@@
 There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
-	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
-In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
-One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the classroom domain the expected sweetspot varied between the points depending on the stimuli.
-In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
-In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets.
-P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
-\end{enumerate}
 
-\subsection{Participants and Procedure}
-
-43 undergraduate students from two different third level institutes took part in the study.
-To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.
-
-\subsection{Analysis}
-
-%Task 1: The participant draws a polygon denoting the region
-%Analysis of results:
-%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
-%Within the data for each trial,
-%-- For each participant, generate the intersection and union polygon of all the other participants
-%---- Generate N random points in the area of the full stimulus
-%---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
-%---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
-Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
-%(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
-%This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
-%(2) Does context matter?
-%I understand that all the rooms stimuli have the same extent.
-%So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
-%We are hoping for a low kappa score
-
-%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
-%\begin{equation*}
-%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
-%\end{equation*}
-%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators. 
-In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
-In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
-
+\subsection{Trial Type 1 - Results}
 One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$.
@@
 
-The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
+\subsection{Trial Type 2 - Design} 
+In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
+The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
+One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped.
+In the classroom domain the expected sweetspot varied between the points depending on the stimuli.
+In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets.
+P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+
+\subsection{Trial Type 2 - Results}
+
+%Task 1: The participant draws a polygon denoting the region
+%Analysis of results:
+%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+%Within the data for each trial,
+%-- For each participant, generate the intersection and union polygon of all the other participants
+%---- Generate N random points in the area of the full stimulus
+%---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
+%---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
+Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
+%(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
+%This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
+%(2) Does context matter?
+%I understand that all the rooms stimuli have the same extent.
+%So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
+%We are hoping for a low kappa score
+
+%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
+%\begin{equation*}
+%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
+%\end{equation*}
+%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators. 
+In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
+In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
+
+%The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched %the described spatial region.
 For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.

### 2013-09-11 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK updated table 2 by removing t-test results

@@
 
-\section{Human Studies - John}
+\section{Human Studies}
 
@@
 \begin{enumerate}
-	\item Given a particular situation how well do people agree on the location and extent of a named spatial region?
-	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named spatial region?
+	\item Given a particular situation how well do people agree on the location and extent of a named context dependent spatial region?
+	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named context dependent spatial region?
 \end{enumerate}
 
-To examine these questions we developed three different trials.
-In all three trial types the subject was presented with a visual stimuli (such as a room floor plan or a map) and an accompanying linguistic description (such as \emph{go to the front of the room}, however the way the task the subject was set varied across the trials:
-\begin{enumerate}
-	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the spatial region described in the description or to select a no region option.
-	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-	\item Trial type 3: subjects were asked to mark on the visual stimuli, by clicking with the mouse, the location that they felt was most prototypical of the spatial description. 
-\end{enumerate}
-
-All three trial types use the same visual stimuli and accompanying spatial descriptions.
+To examine these questions, we developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).
 Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment.
 Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
-The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.
+The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and (3) the military engagement stimuli were always accompanied by the description \emph{safety}.
 
@@
 
+The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
+Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
+Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.
+
 \subsection{Participants and Procedure}
+Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
+43 undergraduate students from two different third level institutes took part in the study.
+To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.
 
-43 undergraduate students from two different third level institutes took part in the study.
-To control for sequence effects the presentation order of the trials were randomised each participant.
-Also in order to minimise fatigue of the subjects for the Type 2 Likert trials each subject only ranked a randomly selected subset of the trials.
+\subsection{Trial Type 1 - Design}
+In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
+The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
+Subjects drew polygons by clicking to position the verticies with lines being automatically generated between each point indicated as the subject drew.
+There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
 
-\subsection{Analysis}
-
-%Task 1: The participant draws a polygon denoting the region
-%Analysis of results:
-%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
-%Within the data for each trial,
-%-- For each participant, generate the intersection and union polygon of all the other participants
-%---- Generate N random points in the area of the full stimulus
-%---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
-%---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
-Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
-%(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
-%This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
-%(2) Does context matter?
-%I understand that all the rooms stimuli have the same extent.
-%So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
-%We are hoping for a low kappa score
-
-%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
-%\begin{equation*}
-%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
-%\end{equation*}
-%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators. 
-In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
-In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
-
+\subsection{Trial Type 1 - Results}
 One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$.
@@
 
-The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
-
+\subsection{Trial Type 2 - Design} 
+In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
 The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
@@
 P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+
+\subsection{Trial Type 2 - Results}
+
+%Task 1: The participant draws a polygon denoting the region
+%Analysis of results:
+%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
+%Within the data for each trial,
+%-- For each participant, generate the intersection and union polygon of all the other participants
+%---- Generate N random points in the area of the full stimulus
+%---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
+%---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
+Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
+%(see http://en.wikipedia.org/wiki/Cohen%27s_kappa )
+%This process will result in for each trial a list of kappa scores for each participant indicating the extent to which they agree with everyone else. We are hoping for a high kappa score
+%(2) Does context matter?
+%I understand that all the rooms stimuli have the same extent.
+%So one way of evaluating whether context matters is to use the kapp score between the intersection/union polygon generated using all the response (less the excluded partitipants) for one trial and the equivalent polygon for another trial.
+%We are hoping for a low kappa score
+
+%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
+%\begin{equation*}
+%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
+%\end{equation*}
+%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators. 
+In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
+In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
+
+%The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched %the described spatial region.
+For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
 
@@
 \centerline{
-\begin{tabular}{l c c c l}
+\begin{tabular}{l c c c }
 \hline
-Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
-Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
-(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
-(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
+Domain & \multicolumn{3}{c}{Map/Bottleneck} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} \\
+Stim ID & P1 & P2 & P3 \\
+(g) & $1.1707$ & $\mathbf{4.6667}$ \\ % 1_1
+(h) & $1.1707$ & $\mathbf{4.7561}$ \\ % 1_2
 \\
 \hline
-Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
-Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
-(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
-(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
-(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
-(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
-(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
-(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
-(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
+Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} \\
+Stim ID & P1 & P2 & P3 \\
+(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ \\ % 2_1
+(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ \\ % 2_6
+(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ \\ %2_7
+(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ \\ %2_2
+(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ \\ % 2_3
+(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ \\ % 2_4
+(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ \\ % 2_5
 \\
 \hline
-Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
-Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
-(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
-(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
-(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
-(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
+Domain & \multicolumn{3}{c}{Military Engagement/Safe Region} \\
+ ~ & \multicolumn{3}{c}{Average Likert Rating} \\
+Stim ID & P1 & P2 & P3\\
+(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ \\ % 3_1
+(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ \\ % 3_2
+(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ \\ % 3_3
+(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$\\ % 3_4
 \\
@@
 \label{tab:point-likert-scores}
-\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
+\caption{Average Likert Ratings per Point by Domain.}
 \end{table}
+
+%\begin{table}
+%\centerline{
+%\begin{tabular}{l c c c l}
+%\hline
+%Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
+% ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
+%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+%(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
+%(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
+%\\
+%\hline
+%Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
+% ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
+%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+%(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
+%(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
+%(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
+%(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
+%(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
+%(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
+%(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
+%\\
+%\hline
+%Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
+% ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
+%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+%(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
+%(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
+%(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
+%(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
+%\\
+%\hline
+%\hline
+%\end{tabular}
+%}
+%\label{tab:point-likert-scores}
+%\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
+%\end{table}
 

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK updated the text in the analysis section and resized the bottleneck1.jpg

@@
 
-The study was designed to examine the following questions about spatial regions:
+The study was designed to examine the following questions about CDSRs:
 \begin{enumerate}
@@
 
-To examine these questions, we developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).
+Both of these research questions address the fundamental claim that CDSRs are contextually defines.
+Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects.
+Research question 2 address the contextual nature of CDSRs.
+Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+
+To examine these questions, we developed two experimental tasks:
+\begin{enumerate}
+	\item Task 1: subjects were presented with a visual representation of a spatial scene and an linguistic description containing a CDSR and were tasked with drawing a polygon a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene or to select a no region opt
+	\item Task 2: subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were taked with rating on a five point likert scale how well the marked location fitted with their understanding of the CDSR.
+\end{enumerate}
+
+Both Task 1 and 2 addressed both research questions.
+Between subject agreement metrics can be computed for both Task 1 and Task 2 and, consequently, both tasks address research question 1.
+In particular, a variation in between subject agreement across the different contexts for a particular CDSR could be taken as evidence that some contexts are more amenable to a particular CDSR than others and hence CDSRs are contextually dependent.
+By shifting the focus of the analysis to examining the variation in responses for a given CDSR in context that differed only with respect to the location of objects both the results from both tasks can address research question 2.
+In the case of Task 1 the variation we will examine is the area of the polygons, which in Task 2 the variation we will examine is the scoring of the marked locations.
+
+We developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).
 Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment.
+These stimuli were used in both Task 1 and 2.
 Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
@@
 The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
+We hoped this variation woudd result in a variation in the size of the polygons for these different visual contexts.
 Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
@@
 %where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators.
-In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
+
+%In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
 In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
@@
 In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
-For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests.
-There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
+For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points.
+%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
 Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli.
@@
 To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
-Table \ref{tab:flipped-point-likert-scores} lists the results of this trial.
+%Table \ref{tab:flipped-point-likert-scores} lists the results of this trial.
 In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
+In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
 
-\begin{figure}
-%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
-\centerline{
-\begin{tabular}{ccc}
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
-\end{tabular}
-}
-\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
-\label{fig:locations}
-\end{figure}
+%\begin{figure}
+%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+%\centerline{
+%\begin{tabular}{ccc}
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
+%\end{tabular}
+%}
+%\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
+%\label{fig:locations}
+%\end{figure}
 

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK updated the text and removed table 3

@@
 To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
-%Table \ref{tab:flipped-point-likert-scores} lists the results of this trial.
-In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
+In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
 In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
@@
 
-\begin{table}
-\centerline{
-\begin{tabular}{c c c c }
-\hline
-Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
-$H_0$ & $H_A$ & $\alpha$ & p-value\\
-$\mu(c.P1)=\mu(c-flipped.P3)$ & $\mu(c.P1)!=\mu(c-flipped.P3)$ & $0.05$ & $0.6315$\\
-$\mu(e.P1)=\mu(e-flipped.P3)$ & $\mu(e.P1)!=\mu(e-flipped.P3)$ & $0.05$ & $0.7358$\\
-$\mu(f.P1)=\mu(f-flipped.P3)$ & $\mu(f.P1)!=\mu(f-flipped.P3)$ & $0.05$ & $0.2592$\\
-\hline
-\hline
-\end{tabular}
-}
-\label{tab:flipped-point-likert-scores}
-\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
-\end{table}
+%\begin{table}
+%\centerline{
+%\begin{tabular}{c c c c }
+%\hline
+%Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
+%$H_0$ & $H_A$ & $\alpha$ & p-value\\
+%$\mu(c.P1)=\mu(c-flipped.P3)$ & $\mu(c.P1)!=\mu(c-flipped.P3)$ & $0.05$ & $0.6315$\\
+%$\mu(e.P1)=\mu(e-flipped.P3)$ & $\mu(e.P1)!=\mu(e-flipped.P3)$ & $0.05$ & $0.7358$\\
+%$\mu(f.P1)=\mu(f-flipped.P3)$ & $\mu(f.P1)!=\mu(f-flipped.P3)$ & $0.05$ & $0.2592$\\
+%\hline
+%\hline
+%\end{tabular}
+%}
+%\label{tab:flipped-point-likert-scores}
+%\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
+%\end{table}
 

### 2013-09-11 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK added a discussion of the P2 ratings into the paper

@@
 In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
-For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests.
-There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
+For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points using a confidence level of $95\%$.
+%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
 Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli.
@@
 To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
-Table \ref{tab:flipped-point-likert-scores} lists the results of this trial.
-In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
+In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$.
+In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
+Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l).
+In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land.
+Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j).
+As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j).
+This higher rating is statiscially significant at a $95\%$ confidence level.
+Stimuli (k) and (l) present and analagous situation with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
+Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
+Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high.
+This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR.
+Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
 
-\begin{figure}
-%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
-\centerline{
-\begin{tabular}{ccc}
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
-\end{tabular}
-}
-\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
-\label{fig:locations}
-\end{figure}
+%\begin{figure}
+%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+%\centerline{
+%\begin{tabular}{ccc}
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
+%\end{tabular}
+%}
+%\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
+%\label{fig:locations}
+%\end{figure}
 

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex
changed bits to related work

@@
 In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
-For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points.
+For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points using a confidence level of $95\%$.
 %There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
@@
 To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
-In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.
+In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$.
 In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
+Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l).
+In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land.
+Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j).
+As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j).
+This higher rating is statiscially significant at a $95\%$ confidence level.
+Stimuli (k) and (l) present and analagous situation with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
+Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
+Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high.
+This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR.
+Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
 
@@
 
-\section{Modeling - Nick}
+\section{Related Work}
 
 As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g.
-understanding spatially-situated task utterances), they must have access to internal models which can represent CDSRs and recognise them in new environments.
+understanding spatially-situated task utterances), they must have access to internal models which can both represent CDSRs and recognise them in new environments.
 Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
@@
 The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
-Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to work.
-However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain code for image region representations~\cite{Freeman:1961}.
-This is an approach we followed in previous work, where boundary segments connected vertices defined by contextually-appropriate anchor points~\cite{Hawes:2012}.
-Whilst our previous approach provided basic performance, the crisp boundaries it generated caused mismatches with human data over the extent of regions.
-Methods based on underlying spatial functions may be more flexible in this regard, as they can be selectively thresholded depending on context.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to model the regions with sufficient accuracy.
+However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
 
-% Basic approaches
-
-\textbf{TODO:} Perhaps add related work on robot spatial representations, e.g. maps, rooms.
-
-In the following sections we explore whether two possible CDSR models (one based on boundary segments, one based on a spatial function) are able to satisfy the previously described desiderata for a cognitive systems approach to representing and recognising CDSRs.
-The surrounding discussion is intended to explore the strengths and weaknesses of these models, providing directions for future research.
-
-\subsection{Symbolic Boundary Method - KLENK}
-In our previous work \cite{Hawes:2012}, we presented an approach on a mobile robot that combined qualitative spatial relations, object recognition, and analogy to transfer CDSRs represented as polygons defined by \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11}.
+This is an approach we followed in previous work \cite{Hawes:2012}, where we developed an approach that defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon.
 Anchor points are symbolic descriptions that link a conceptual entity (e.g., the front of a classroom) to perceived entities (e.g., the lower left extent of a desk).
+Once a particular representation has been modelled in this manner, we were able to find the same type of region in previously unseen rooms using analogical transfer.
 An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.
@@
 This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task.
-
 Applying this model to the points and regions indicated in our experiment illustrates two limitations.
@@
 
-\subsection{Spatial Function Method - NICK}
-
-Overview - either Chris's approach or something similar
-Example
-Results?
-Discussion 
-Where do the points come from
+The use of spatial regions connected to function has been explored in other robotics work, e.g.
+~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
+Such fields can only vary with respect to context in terms of their size, which does not make them particularly well suited to modelling CDSRs.
 
@@
 \noindent
-STRANDS etc.
+The research leading to these results has received funding from the European Union Seventh Framework Program (FP7/2007-2013) under the project STRANDS (grant agreement No 600623), and from the EPSRC under grant EP/K014293/1.
 \end{acknowledgements}

### 2013-09-11 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
minor edits

@@
 Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.
-Consider an agent asked to remain "in the safe zone" during a military engagement.
+Consider an agent asked to remain ``in the safe zone'' during a military engagement.
 The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
@@
 
-\section{Introduction - KLENK} 
+\section{Introduction} 
 Natural collaboration between humans and cognitive systems requires both participants to communicate using a shared understanding of spatial language.
-Cognitive systems can do quite well with spatial terms defined in terms of coordinates on a map, or that can be specified using known landmarks.
-However, many spatial terms are defined not only be their geometry, but also their context and their relevance to current goals.
+Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.
+However, many spatial terms are defined not only by their geometry, but also their context and their relevance to current goals.
 That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration and functional roles.
-We call such regions \textit{context-dependent spatial regions}(CDSRs).
+We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).
 
-We begin by discussing the features of CDSRs using illustrative examples.
+Section~\ref{sec:cdsrs} discusses the features of CDSRs using illustrative examples.
 We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.
@@
 
-\section{Context-Dependent Spatial Regions}
-Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.
+\section{Context-Dependent Spatial Regions}\label{sec:cdsrs}
+
+Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, or the geographical feature of a bottleneck.
 To identify these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
 Furthermore, these regions exist at different levels of spatial resolution.
-Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.
+Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.
 
@@
 Understanding these regions is crucial to joint intelligent action with humans.
-Not only are these regions frequently tied to goals and commands: "go to the front of the classroom", "defend the bottleneck", and "get to safety", but they are also central to understanding human activities: e.g., "Is there a class in session?".
+Not only are these regions frequently tied to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety'', but they are also central to understanding human activities: e.g., ``Is there a class in session?''.
 
-The long-term goal of this work is to create intelligent agents that can naturally interact with people in human environments.
+The long-term goal of our work is to create cognitive systems that can naturally interact with people in human environments.
 While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.
@@
 
-\subsection{Research Question, Design and Stimuli}
+\subsection{Research Questions, Design and Stimuli}
 
-The study was designed to examine the following questions about CDSRs:
+We developed the study described below in order to examine the following questions about CDSRs:
 \begin{enumerate}
-	\item Given a particular situation how well do people agree on the location and extent of a named context dependent spatial region?
-	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named context dependent spatial region?
+	\item Given a particular situation how well do people agree on the location and extent of a named CDSR?
+	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named CDSR?
 \end{enumerate}
 
-Both of these research questions address the fundamental claim that CDSRs are contextually defines.
-Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects.
+Both of these research questions address the fundamental claim that CDSRs are contextually defined.
+Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects.
 Research question 2 address the contextual nature of CDSRs.
@@
 To examine these questions, we developed two experimental tasks:
-\begin{enumerate}
-	\item Task 1: subjects were presented with a visual representation of a spatial scene and an linguistic description containing a CDSR and were tasked with drawing a polygon a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene or to select a no region opt
-	\item Task 2: subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were taked with rating on a five point likert scale how well the marked location fitted with their understanding of the CDSR.
-\end{enumerate}
+\begin{itemize}
+	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were asked to rate how well the marked location fitted with their understanding of the CDSR on a five point likert scale.
+\end{itemize}
 
-Both Task 1 and 2 addressed both research questions.
-Between subject agreement metrics can be computed for both Task 1 and Task 2 and, consequently, both tasks address research question 1.
+Both Task 1 and 2 address both research questions.
+Between subject agreement metrics can be computed for both Task 1 and Task 2.
 In particular, a variation in between subject agreement across the different contexts for a particular CDSR could be taken as evidence that some contexts are more amenable to a particular CDSR than others and hence CDSRs are contextually dependent.
 By shifting the focus of the analysis to examining the variation in responses for a given CDSR in context that differed only with respect to the location of objects both the results from both tasks can address research question 2.
-In the case of Task 1 the variation we will examine is the area of the polygons, which in Task 2 the variation we will examine is the scoring of the marked locations.
+In the case of Task 1 the variation we will examine is the area of the polygons, while in Task 2 the variation we will examine is the scoring of the marked locations.
 
-We developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).
-Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment.
+We developed a set of visual stimuli depicting different environments, and paired each with an accompanying linguistic description of a CDSR easily understood within the context of that environment (see Figure \ref{fig:exp_stimuli}).
 These stimuli were used in both Task 1 and 2.
 Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
-The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and (3) the military engagement stimuli were always accompanied by the description \emph{safety}.
+The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the military engagement stimuli were always accompanied by the description \emph{safety}; and (3) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}.
 
@@
 We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
-To control for left-right bias effects two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts.
-Conseqeuntly, there were 7 classroom stimuli (circle, U $\times 2$, close $\times 2$, far $\times 2$).
+To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts.
+Consequently, there were 7 classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).
 In designing these stimuli we deliberately chose some layouts that were very common or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}.
@@
 The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
-We hoped this variation woudd result in a variation in the size of the polygons for these different visual contexts.
+We hoped this variation would result in a variation in the size of the polygons for these different visual contexts.
 Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
@@
 The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
-Subjects drew polygons by clicking to position the verticies with lines being automatically generated between each point indicated as the subject drew.
+Subjects drew polygons by clicking to position the vertices with lines being automatically generated between each point indicated as the subject drew.
 There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
@@
 \subsection{Trial Type 1 - Results}
+
 One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$.
@@
 It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$.
-A $\kappa$ score between $0.3-0.4$ would be considered as being poor to fair.
-This would indicate that there was some agreement between subjects but less that would be expected.
+A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e.
+there was some agreement between subjects but less that would be expected.
 However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
-In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no regrion responses for each stimuli.
+
+In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
 It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
@@
 Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
-$\kappa$ scores in this range can to interpreted as fair to good.
+$\kappa$ scores in this range can be interpreted as fair to good.
 
@@
 In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+The location of these points was informed by the results of a pilot study.
+In the final study they were positioned such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
 In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
@@
 In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
+
 Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l).

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-\subsection{Trial Type 1 - Design}
+\subsection{Task 1 - Design}
 In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
 The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
-Subjects drew polygons by clicking to position the verticies with lines being automatically generated between each point indicated as the subject drew.
+Subjects drew polygons by clicking to position the vertices with lines being automatically generated between each point indicated as the subject drew.
 There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
 
-\subsection{Trial Type 1 - Results}
-One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$.
-Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes.
-$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
-In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
+\subsection{Task 1 - Results}
+We use two methods to analyze the results of region drawing task.
+First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous situations (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).
+Second, we use the extent points of the regions to assess changes between trials.
+That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.
+
+We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.
+We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
 After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
 In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
+$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
+
+% Klenk: This is where I left off.  I think we can drop the histogram and talk.
+
 Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.
 It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$.
-A $\kappa$ score between $0.3-0.4$ would be considered as being poor to fair.
-This would indicate that there was some agreement between subjects but less that would be expected.
+A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e.
+there was some agreement between subjects but less that would be expected.
 However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
-In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no regrion responses for each stimuli.
-It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
-Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
-Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
-Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
-$\kappa$ scores in this range can to interpreted as fair to good.
-
-\begin{figure}
-%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
-\centerline{
-\begin{tabular}{ccc}
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
-\end{tabular}
-}
-\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
-\label{fig:type-1-hists}
-\end{figure}
 
@@
 
-The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
+In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
+It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
+Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
+Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
+Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
+$\kappa$ scores in this range can be interpreted as fair to good.
+
+The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
 Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
@@
 
-\subsection{Trial Type 2 - Design} 
+%\begin{figure}
+%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+%\centerline{
+%\begin{tabular}{ccc}
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
+%\end{tabular}
+%}
+%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
+%\label{fig:type-1-hists}
+%\end{figure}
+
+\begin{table}
+\centerline{
+\begin{tabular}{c c}
+Expected Relationship & Human Results  \\
+\hline
+\hline
+area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
+maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
+minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
+maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
+maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
+maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
+\hline
+\end{tabular}
+}
+\label{tab:region-location-tests}
+\caption{Comparison of the expected relationships to the mean human responses.}
+\end{table}
+
+\subsection{Task 2 - Design} 
 In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+The location of these points was informed by the results of a pilot study.
+In the final study they were positioned such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
 In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
@@
 
-\subsection{Trial Type 2 - Results}
+\subsection{Task 2 - Results}
 

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, or the geographical feature of a bottleneck.
-To identify these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
+To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
 Furthermore, these regions exist at different levels of spatial resolution.
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects.
+Question 1 addresses this claim by examining variation in CDSR placement within different situtations.
+In a situation exhibiting clear contextual clues (e.g.
+determining a safe zone in a battle situation where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR.
+In a less well-defined context, one might expect little to no agreement between human communicators.
+Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+
+%Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextually dependent then one would expect a variation %across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a %high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement %between subjects.
 Research question 2 address the contextual nature of CDSRs.
-Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+Research question 2 addresses the contextual nature of CDSRs %in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
 
@@
 	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were asked to rate how well the marked location fitted with their understanding of the CDSR on a five point likert scale.
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the marked location fit with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
@@
 In the case of Task 1 the variation we will examine is the area of the polygons, while in Task 2 the variation we will examine is the scoring of the marked locations.
+
+% kate: vocabulary clarification: situation vs. context??
 
@@
 
-\subsection{Trial Type 1 - Design}
+\subsection{Task 1 - Design}
 In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
@@
 
-\subsection{Trial Type 1 - Results}
+\subsection{Task 1 - Results}
+We use two methods to analyze the results of region drawing task.
+First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous situations (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).
+Second, we use the extent points of the regions to assess changes between trials.
+That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.
 
-One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$.
-Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes.
-$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
-In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
+%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
+% but want to talk to John about it during the telecon tomorrow first
+
+We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.
+We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
 After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
 In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
+$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
+
+% Klenk: This is where I left off.  I think we can drop the histogram and talk.
+
 Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.
@@
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
-
-In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
-It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
-Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
-Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
-Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
-$\kappa$ scores in this range can be interpreted as fair to good.
-
-\begin{figure}
-%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
-\centerline{
-\begin{tabular}{ccc}
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
-\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
-\end{tabular}
-}
-\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
-\label{fig:type-1-hists}
-\end{figure}
 
@@
 
-The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
+In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
+It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
+Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
+Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
+Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
+$\kappa$ scores in this range can be interpreted as fair to good.
+
+%Kate: this is a place (above) that I think we could combine/cut.  I think we can talk about the non-typical classrooms more concisely.  Let's
+% say something like "some of the layouts were non-typical -> less experience with the context -> we expect lower agreement -> this is what 
+% we see" then we can say that if we remove these examples, agreement on the more typical scence is good.  So suggestions are
+% (1) condense discussion (2) make it clear exactly what we expect for the non typical scenes and WHY this is tied back to our claim
+% about the role of context in CDSRs.  I am happy to make these changes if John agrees that I am not horribly misinterpreting his fine results
+
+The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
 Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
@@
 
-\subsection{Trial Type 2 - Design} 
+%\begin{figure}
+%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
+%\centerline{
+%\begin{tabular}{ccc}
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
+%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
+%\end{tabular}
+%}
+%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
+%\label{fig:type-1-hists}
+%\end{figure}
+
+\begin{table}
+\centerline{
+\begin{tabular}{c c}
+Expected Relationship & Human Results  \\
+\hline
+\hline
+area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
+maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
+minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
+maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
+maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
+maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
+\hline
+\end{tabular}
+}
+\label{tab:region-location-tests}
+\caption{Comparison of the expected relationships to the mean human responses.}
+\end{table}
+
+\subsection{Task 2 - Design} 
 In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
@@
 
-\subsection{Trial Type 2 - Results}
+%Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though
+
+\subsection{Task 2 - Results}
 
@@
 In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
+%lending support for question 2 ...
 

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK references to situation changed to context

@@
 We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.
-We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.
+We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
 To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
@@
 While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.
-First, the representational approach should be transferable to new situations.
+First, the representational approach should be transferable to new contexts.
 While humans do not demonstrate mastery after a single example of a concept, they are able to begin using it immediately refining it over time.
@@
 \begin{enumerate}
-	\item Given a particular situation how well do people agree on the location and extent of a named CDSR?
-	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named CDSR?
+	\item Given a particular context how well do people agree on the location and extent of a named CDSR?
+	\item Does the orientation and location of objects (with relevant functional semantics) in a given context effect the location and extent of a named CDSR?
 \end{enumerate}
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects.
+Question 1 addresses this claim by examining variation in CDSR placement within different situtations.
+In a context exhibiting clear contextual clues (e.g.
+determining a safe zone in a battle context where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR.
+In a less well-defined context, one might expect little to no agreement between human communicators.
+Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+
+%Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextually dependent then one would expect a variation %across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a %high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement %between subjects.
 Research question 2 address the contextual nature of CDSRs.
-Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+Research question 2 addresses the contextual nature of CDSRs %in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
 
@@
 We use two methods to analyze the results of region drawing task.
-First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous situations (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).
+First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).
 Second, we use the extent points of the regions to assess changes between trials.
 That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.
+
+%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
+% but want to talk to John about it during the telecon tomorrow first
 
@@
 there was some agreement between subjects but less that would be expected.
-However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
+However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects.
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
@@
 
+%\begin{table}
+%\centerline{
+%\begin{scriptsize}
+%\begin{tabular}{l c c c c c c c c c c c c c}
+%% (h) - 1_2; (g) - 1_1; (k) - 3_3; (f) - 2_4; (i) - 3_1; (e-flipped) - 2_3; (j) - 3_2; (f-flipped) - 2_5; (e) - 2_2;  (l) - 3_4;  (c) -  2_6;  (c-flipped) - 2_7;  (a) - 2_1
+%Stim ID & (h) & (g) & (k) & (f) & (i) & (e-flipped) & (j) & (f-flipped) & (e) & (l) & (c) & (c-flipped) & (a)\\ 
+%Average Pairwise $\kappa$ & 0.5034 & 0.4585 & 0.4548 & 0.3930 & 0.3906 & 0.3813 & 0.3787 & 0.3720 & 0.3629 & 0.3164 & 0.2438 & 0.2414 & 0.0996 \\ 
+%No Region Responses & 0 & 1 & 0& 0& 0& 2&1&2& 1& 0&2&4& 10\\
+%\hline
+%\end{tabular}
+%\end{scriptsize}
+%}
+%\label{tab:stimuli-kappa-scores}
+%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimuli.
+The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
+%\end{table}
+
 In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
 It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
-Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
+Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
 Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
@@
 
+%Kate: this is a place (above) that I think we could combine/cut.  I think we can talk about the non-typical classrooms more concisely.  Let's
+% say something like "some of the layouts were non-typical -> less experience with the context -> we expect lower agreement -> this is what 
+% we see" then we can say that if we remove these examples, agreement on the more typical scence is good.  So suggestions are
+% (1) condense discussion (2) make it clear exactly what we expect for the non typical scenes and WHY this is tied back to our claim
+% about the role of context in CDSRs.  I am happy to make these changes if John agrees that I am not horribly misinterpreting his fine results
+
 The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
@@
 For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}.
-Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.
+Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision context $a$ had an average pairwise agreement of $0.0996$.
 
@@
 P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+
+%Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though
 
@@
 This higher rating is statiscially significant at a $95\%$ confidence level.
-Stimuli (k) and (l) present and analagous situation with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
+Stimuli (k) and (l) present and analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
 Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
@@
 
-Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular situation depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted situation was consistent across similar situations, as shown by the fact that the sweetspot flipped from P1 to P3 when the situation was flipped.
+Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
 
@@
 In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-This indicates that people consider the dynamics of the situation when evaluating these terms, which is not accounted for in this model.
+This indicates that people consider the dynamics of the context when evaluating these terms, which is not accounted for in this model.
 

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 \section{Discussion}
+Cognitive systems do not face problems in isolation.
+Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
+In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.
+We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment resulted predicted changes in the location, size, and shape of the spatial region.
+We outline requirements for a cognitive systems solution and discuss limitations in current approaches from cognitive systems and robotics.
+Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.
 

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk fixed error in table 3

@@
 Stim ID & P1 & P2 & P3 \\
-(g) & $1.1707$ & $\mathbf{4.6667}$ \\ % 1_1
-(h) & $1.1707$ & $\mathbf{4.7561}$ \\ % 1_2
+(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$\\ % 1_1
+(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ \\ % 1_2
 \\

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk flipped table 3

--- 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
@@
 \centerline{
-\begin{tabular}{l c c c }
+\begin{tabular}{l c c c c c c c }
 \hline
-Domain & \multicolumn{3}{c}{Map/Bottleneck} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} \\
-Stim ID & P1 & P2 & P3 \\
-(g) & $1.1707$ & $\mathbf{4.6667}$ \\ % 1_1
-(h) & $1.1707$ & $\mathbf{4.7561}$ \\ % 1_2
-\\
 \hline
-Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} \\
-Stim ID & P1 & P2 & P3 \\
-(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ \\ % 2_1
-(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ \\ % 2_6
-(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ \\ %2_7
-(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ \\ %2_2
-(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ \\ % 2_3
-(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ \\ % 2_4
-(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ \\ % 2_5
-\\
+%Domain & \multicolumn{7}{l}{Map/Bottleneck} \\
+\textbf{Map} & g & h & ~ & ~ & ~ & ~ & ~ \\ 
+P1 & $1.1707$ & $1.1707$ & ~ & ~ & ~ & ~ & ~ \\
+P2 & $\mathbf{4.6667}$ & $\mathbf{4.7561}$ & ~ & ~ & ~ & ~ & ~ \\
+P3 & $1.85$ & $1.585$ & ~ & ~ & ~ & ~ & ~ \\
 \hline
-Domain & \multicolumn{3}{c}{Military Engagement/Safe Region} \\
- ~ & \multicolumn{3}{c}{Average Likert Rating} \\
-Stim ID & P1 & P2 & P3\\
-(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ \\ % 3_1
-(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ \\ % 3_2
-(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ \\ % 3_3
-(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$\\ % 3_4
-\\
+%Domain & \multicolumn{7}{l}{Classrom/Front of Room} \\
+\textbf{Classroom}& a & c & c-flipped & e & e-flipped & f & f-flipped \\
+P1 & $\mathbf{3.5122}$ & $\mathbf{4.7073}$ & $1.244$  & $\mathbf{4.7561}$ & $1.2439$ & $\mathbf{4.8292}$ & $1.1463$\\
+P2 & $2.6341$ & $2.9024$ & $2.9268$ & $2.0000$ & $1.8500$ & $2.0976$ & $2.3171$\\
+P3 & $1.9024$ & $1.3571$ & $\mathbf{4.6429}$ & $1.2439$ & $\mathbf{4.7073}$ & $1.2750$ & $\mathbf{4.6585}$\\
 \hline
+%Domain & \multicolumn{7}{l}{Military Engagement/Safe Region} \\
+\textbf{Military} & i & j & k & l & ~ & ~ & ~ \\
+P1 & $1.1951$ & $1.0952$ & $1.0714$ & $1.2327$ & ~ & ~ & ~ \\
+P2 & $4.3902$ & $2.2927$ & $2.7804$ & $3.1667$& ~ & ~ & ~ \\
+P3 & $\mathbf{4.9024}$ & $\mathbf{4.5122}$ & $\mathbf{4.6585}$ & $\mathbf{4.7073}$ & ~ & ~ & ~ \\
 \hline
@@
 \label{tab:point-likert-scores}
-\caption{Average Likert Ratings per Point by Domain.}
+\caption{Average Likert Ratings per Point by Domain and stimulus.}
 \end{table}
+
+%\begin{table}
+%\centerline{
+%\begin{tabular}{l c c c }
+%\hline
+%\hline
+%Domain & \multicolumn{3}{c}{Map/Bottleneck} \\
+%% ~ & \multicolumn{3}{c}{Average Likert Rating} \\
+%Stim ID & P1 & P2 & P3 \\
+%(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$\\ % 1_1
+%(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ \\ % 1_2
+%\\
+%\hline
+%Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
+%% ~ & \multicolumn{3}{c}{Average Likert Rating} \\
+%Stim ID & P1 & P2 & P3 \\
+%(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ \\ % 2_1
+%(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ \\ % 2_6
+%(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ \\ %2_7
+%(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ \\ %2_2
+%(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ \\ % 2_3
+%(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ \\ % 2_4
+%(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ \\ % 2_5
+%\\
+%\hline
+%Domain & \multicolumn{3}{c}{Military Engagement/Safe Region} \\
+%% ~ & \multicolumn{3}{c}{Average Likert Rating} \\
+%Stim ID & P1 & P2 & P3\\
+%(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ \\ % 3_1
+%(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ \\ % 3_2
+%(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ \\ % 3_3
+%(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$\\ % 3_4
+%\\
+%\hline
+%\end{tabular}
+%}
+%\label{tab:point-likert-scores}
+%\caption{Average Likert Ratings per Point by Domain and stimulus.}
+%\end{table}
 

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, or the geographical feature of a bottleneck.
-To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).
-Furthermore, these regions exist at different levels of spatial resolution.
+To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
 Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.
@@
 
-\subsection{Research Questions, Design and Stimuli}
-
 We developed the study described below in order to examine the following questions about CDSRs:
 \begin{enumerate}
-	\item Given a particular situation how well do people agree on the location and extent of a named CDSR?
-	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named CDSR?
+	\item Given a particular context how well do people agree on the location and extent of a named CDSR?
+	\item Does the orientation and location of objects (with relevant functional semantics) in a given context effect the location and extent of a named CDSR?
 \end{enumerate}
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Question 1 addresses this claim by examining variation in CDSR placement within different situtations.
-In a situation exhibiting clear contextual clues (e.g.
-determining a safe zone in a battle situation where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR.
+Question 1 addresses this claim by examining variation in CDSR placement within different situations.
+In a context exhibiting clear contextual clues (e.g.
+determining a safe zone in a battle context where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR.
 In a less well-defined context, one might expect little to no agreement between human communicators.
@@
 \begin{itemize}
-	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the marked location fit with their understanding of the CDSR on a five point likert scale.
+	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the extent of the CDSR in the scene (or stated there was no such region).
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well it fit with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
@@
 
-% kate: vocabulary clarification: situation vs. context??
-
 We developed a set of visual stimuli depicting different environments, and paired each with an accompanying linguistic description of a CDSR easily understood within the context of that environment (see Figure \ref{fig:exp_stimuli}).
@@
 For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom.
-We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
+We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
 To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts.
-Consequently, there were 7 classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).
-In designing these stimuli we deliberately chose some layouts that were very common or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}.
-By varying the typicality of the layouts we hoped to get some variation across the between subject agreement on a per stimuli basis, for example we expected to get a higher between subject agreement for the stimuli that were more common.
-This would provide insight into research question 1, in particular it would point to the effect of typicality on the semantics of contextually defined spatial regions.
-Also, because the typicality of the scenes was varied by adjusting the location and orientation of objects in the scenes, a variation in agreement across the stimuli would provide evidence that the orientation and location of objects do effect the location and extent of spatial regions.
+Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).
+The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.
+Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
 
@@
 
-The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region.
-We hoped this variation would result in a variation in the size of the polygons for these different visual contexts.
-Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
-Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.
-
 \subsection{Participants and Procedure}
@@
 We use two methods to analyze the results of region drawing task.
-First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous situations (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).
+First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 Second, we use the extent points of the regions to assess changes between trials.
@@
 $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
+Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.
 
@@
 
-Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.
-It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$.
-A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e.
-there was some agreement between subjects but less that would be expected.
-However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects.
+%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects.
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
@@
 
-In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
-It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
-Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
-Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
+%\begin{table}
+%\centerline{
+%\begin{scriptsize}
+%\begin{tabular}{l c c c c c c c c c c c c c}
+%% (h) - 1_2; (g) - 1_1; (k) - 3_3; (f) - 2_4; (i) - 3_1; (e-flipped) - 2_3; (j) - 3_2; (f-flipped) - 2_5; (e) - 2_2;  (l) - 3_4;  (c) -  2_6;  (c-flipped) - 2_7;  (a) - 2_1
+%Stim ID & (h) & (g) & (k) & (f) & (i) & (e-flipped) & (j) & (f-flipped) & (e) & (l) & (c) & (c-flipped) & (a)\\ 
+%Average Pairwise $\kappa$ & 0.5034 & 0.4585 & 0.4548 & 0.3930 & 0.3906 & 0.3813 & 0.3787 & 0.3720 & 0.3629 & 0.3164 & 0.2438 & 0.2414 & 0.0996 \\ 
+%No Region Responses & 0 & 1 & 0& 0& 0& 2&1&2& 1& 0&2&4& 10\\
+%\hline
+%\end{tabular}
+%\end{scriptsize}
+%}
+%\label{tab:stimuli-kappa-scores}
+%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimuli.
+The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
+%\end{table}
+
+Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
+It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.
+Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35.
+Looking at the classroom results as a whole, these results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
+This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.
+
+%Given this relatively large variation across stimuli, which we attribute to the typicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
+%Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
 Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
@@
 
-The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
+%The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
 Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
 For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}.
-Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.
+Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision context $a$ had an average pairwise agreement of $0.0996$.
 
@@
 
-\section{Related Work}
+\section{Applicability of Previous Approaches to Cognitive Systems }
 
-As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g.
-understanding spatially-situated task utterances), they must have access to internal models which can both represent CDSRs and recognise them in new environments.
-Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
+Recall from section \ref{} that in addition to being able to represent these regions for a single environment, the cognitive system (1) must recognize regions of the same type in new environments with new contexts, and (2) use the region for multiple reasoning tasks.
+While spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
 Regions are central to QSR work~\cite{Cohn:2001}.
@@
 The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
-Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to model the regions with sufficient accuracy.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
+
 However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
+This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
+While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiemnt illustrates two limitations.
+First, Regions are not always physically bounded by objects.
+In the classroom examples, people frequently did not extend the front of the room all way to the desks.
+Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
 
-This is an approach we followed in previous work \cite{Hawes:2012}, where we developed an approach that defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon.
-Anchor points are symbolic descriptions that link a conceptual entity (e.g., the front of a classroom) to perceived entities (e.g., the lower left extent of a desk).
-Once a particular representation has been modelled in this manner, we were able to find the same type of region in previously unseen rooms using analogical transfer.
+The connection between spatial regions and function has been explored in other robotics work, e.g.
+~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
+Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+
+%Once a particular representation has been modelled in this manner, we were able to find the same type of region in previously unseen rooms using analogical transfer.
 An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
-  \caption{The known example defines the front of the room with four anchor points.
+%\begin{figure}
+%  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
+%  \caption{The known example defines the front of the room with four anchor points.
 By transferring the anchor points to the new environment, the system is able to ground the room's front in its perception.}
-  \label{fig:anchor-point-transfer}
-\end{figure}
-
-This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task.
-Applying this model to the points and regions indicated in our experiment illustrates two limitations.
-Regions do not always physically bounded by objects.
-In the classroom examples, people frequently did not extend the front of the room all way into the corners.
-In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-This indicates that people consider the dynamics of the situation when evaluating these terms, which is not accounted for in this model.
-
-The use of spatial regions connected to function has been explored in other robotics work, e.g.
-~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
-Such fields can only vary with respect to context in terms of their size, which does not make them particularly well suited to modelling CDSRs.
+%  \label{fig:anchor-point-transfer}
+%\end{figure}
 
 \section{Discussion}
+Cognitive systems do not face problems in isolation.
+Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
+In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.
+We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment resulted predicted changes in the location, size, and shape of the spatial region.
+We outline requirements for a cognitive systems solution and discuss limitations in current approaches from cognitive systems and robotics.
+Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.
 

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
tables 1 and 2 together

@@
 \usepackage{multirow}
+% \usepackage{subfig}
 
@@
 
-\subsection{Research Questions, Design and Stimuli}
-
 We developed the study described below in order to examine the following questions about CDSRs:
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Question 1 addresses this claim by examining variation in CDSR placement within different situtations.
+Question 1 addresses this claim by examining variation in CDSR placement within different situations.
 In a context exhibiting clear contextual clues (e.g.
@@
 We use two methods to analyze the results of region drawing task.
-First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).
+First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 Second, we use the extent points of the regions to assess changes between trials.
@@
 
-Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.
-It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$.
-A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e.
-there was some agreement between subjects but less that would be expected.
-However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects.
+%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects.
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
 
-\begin{table}
-\centerline{
+\begin{table}[t]
+\parbox{.35\linewidth}{
 \begin{tabular}{l c c}
-Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+ & Average & No \\
+Stim ID & Pairwise $\kappa$ & Reg.\\
 \hline
@@
 \end{tabular}
+\label{tab:stimuli-kappa-scores}
+\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
 }
-\label{tab:stimuli-kappa-scores}
-\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.
-The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
+\hfill
+\parbox{.65\linewidth}{
+\begin{tabular}{c c}
+Expected Relationship & Human Results  \\
+\hline
+\hline
+area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
+maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
+% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
+minimumX(e-flipped)  &  \\
+ > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
+maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
+maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
+maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
+\hline
+\end{tabular}
+\label{tab:region-location-tests}
+\caption{Comparison of the expected relationships to the mean human responses.}
+}
 \end{table}
@@
 
-In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
-It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses.
-Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
-Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
+Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
+It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.
+Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35.
+Looking at the classroom results as a whole, these results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
+This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.
+
+%Given this relatively large variation across stimuli, which we attribute to the typicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded.
+%Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject.
 Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$.
@@
 
-The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
+%The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree.
 Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
@@
 %\end{figure}
-
-\begin{table}
-\centerline{
-\begin{tabular}{c c}
-Expected Relationship & Human Results  \\
-\hline
-\hline
-area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
-maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
-minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
-maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
-maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
-maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
-\hline
-\end{tabular}
-}
-\label{tab:region-location-tests}
-\caption{Comparison of the expected relationships to the mean human responses.}
-\end{table}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
-  \caption{The known example defines the front of the room with four anchor points.
+%\begin{figure}
+%  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
+%  \caption{The known example defines the front of the room with four anchor points.
 By transferring the anchor points to the new environment, the system is able to ground the room's front in its perception.}
-  \label{fig:anchor-point-transfer}
-\end{figure}
-
-This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task.
-Applying this model to the points and regions indicated in our experiment illustrates two limitations.
-Regions do not always physically bounded by objects.
-In the classroom examples, people frequently did not extend the front of the room all way into the corners.
-In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-This indicates that people consider the dynamics of the context when evaluating these terms, which is not accounted for in this model.
-
-The use of spatial regions connected to function has been explored in other robotics work, e.g.
-~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
-Such fields can only vary with respect to context in terms of their size, which does not make them particularly well suited to modelling CDSRs.
+%  \label{fig:anchor-point-transfer}
+%\end{figure}
 

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
footnote size refs

@@
 
-\begin{table}
-\centerline{
+\begin{table}[t]
+\parbox{.35\linewidth}{
 \begin{tabular}{l c c}
-Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+ & Average & No \\
+Stim ID & Pairwise $\kappa$ & Reg.\\
 \hline
@@
 {\parindent -10pt\leftskip 10pt\noindent
+\footnotesize
 \bibliographystyle{cogsysapa}

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK fixed up figure one and the resized the images files

@@
 Section~\ref{sec:cdsrs} discusses the features of CDSRs using illustrative examples.
-We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.
+We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.
 We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
@@
 %% Change the classroom to be a picture stimuli
-  \begin{subfigure}[b]{0.3\textwidth}
+\begin{tabular}{c c c} 
+  \begin{subfigure}[b]{0.32\textwidth}
   \includegraphics[width=\textwidth]{figures/classroom.png}
-  \caption{\textit{The front of the classroom} could be defined as the area between the desks and the whiteboard where people give presentations.}
-  \end{subfigure}
-\begin{subfigure}[b]{0.3\textwidth}
+%  \caption{\textit{The front of the classroom} could be defined as the area between the desks and the whiteboard where people give presentations.}
+  \caption{\textit{The front of the classroom}.}
+  \end{subfigure} &
+\begin{subfigure}[b]{0.32\textwidth}
   \includegraphics[width=\textwidth]{figures/safety-3.jpg}
-  \caption{\textit{Safety for the scout} is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
-\end{subfigure}
-\begin{subfigure}[b]{0.3\textwidth}
+%  \caption{\textit{Safety for the scout} is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
+  \caption{\textit{The safe region for the scout}.}
+\end{subfigure} & 
+\begin{subfigure}[b]{0.32\textwidth}
   \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
-  \caption{A \textit{bottleneck} occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
+%  \caption{A \textit{bottleneck} occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
+ \caption{\textit{The bottleneck}.}
   \label{fig:safety}
-  \end{subfigure}
-
+  \end{subfigure}\\
+\end{tabular}
   \caption{Examples of spatial regions used as stimuli in our experiments.}

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk revised abstract, intro and section 2

@@
 
-Understanding these regions is crucial to joint intelligent action with humans.
-Not only are these regions frequently tied to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety'', but they are also central to understanding human activities: e.g., ``Is there a class in session?''.
+Understanding these regions is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.
+In the rest of the paper we present a study that examined the contextual dependence of these regions and dicuss the implications that these t.
+Following this we discuss the implications of these regions for the development of cognitive systems.
 
-The long-term goal of our work is to create cognitive systems that can naturally interact with people in human environments.
+%JK Removed the following text and replaced with text above primarily to save space, feel free to switch back.
+%  Not only are these regions frequently tied to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety'', but they are also central to understanding human activities: e.g., ``Is there a class in session?''.
+%The long-term goal of our work is to create cognitive systems that can naturally interact with people in human environments.
 While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk revised section 3

@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Question 1 addresses this claim by examining variation in CDSR placement within different situations.
-In a context exhibiting clear contextual clues (e.g.
+Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (more bluntly, does a classroom always have a front or is this dependent on the configuration of objects in the room), and (b) the variation in CDSR placement and size within different context.
+Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+
+%In a context exhibiting clear contextual clues (e.g.
 determining a safe zone in a battle context where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR.
-In a less well-defined context, one might expect little to no agreement between human communicators.
-Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+However, in the same environment (classroom, geographic location) but with less well-defined context, one might expect little to no agreement between human communicators.
 
@@
 \begin{itemize}
-	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the extent of the CDSR in the scene (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well it fit with their understanding of the CDSR on a five point likert scale.
+	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fitted with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
@@
 
-\subsection{Participants and Procedure}
-Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
+%\subsection{Participants and Procedure}
+%Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
 43 undergraduate students from two different third level institutes took part in the study.
 To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.
+
+\textbf{Participants and Procedure:} The experiment was delivered online.
+43 undergraduate students from two third level institutes took part in the study.
+To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.
 

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk new section 4

@@
 
-Recall from section \ref{} that in addition to being able to represent these regions for a single environment, the cognitive system (1) must recognize regions of the same type in new environments with new contexts, and (2) use the region for multiple reasoning tasks.
+%Recall from section \ref{} that in addition to being able to represent these regions for a single environment, the cognitive system (1) must recognize regions of the same type in new environments with new contexts, and (2) use the region for multiple reasoning tasks.
 While spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
@@
 Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
-
-However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
+%
+%However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
 This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
@@
 Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-
-The connection between spatial regions and function has been explored in other robotics work, e.g.
+%
+%The connection between spatial regions and function has been explored in other robotics work, e.g.
 ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
 Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+
+%JK Previous version above I replaced it with this shorter version below
+
+Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
+Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
+A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
+However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines).
+This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
+While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiemnt illustrates two limitations.
+First, Regions are not always physically bounded by objects.
+In the classroom examples, people frequently did not extend the front of the room all way to the desks.
+Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
+Finally, the connection between spatial regions and function has been explored in other robotics work, e.g.
+~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
+Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+
+%--------------------------
 

### 2013-09-13 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-% \author{Pat Langley}{langley@asu.edu}
-% \author{Glen Hunt}{glen.hunt@asu.edu}
-% \address{Computing Science and Engineering, Arizona State University, 
-%          Tempe, AZ 85287 USA}
-% \author{Daniel G.\ Shapiro}{dgs@isle.org}
-% \address{Institute for the Study of Learning and Expertise, 
-%          2164 Staunton Court, Palo Alto, CA 94306 USA}
+\author{Kate Lockwood}{klockwood@csumb.edu}
+\address{ITCD Department, California State University - Monterey Bay}
+\author{John D. Kelleher}{john.d.kelleher@dit.ie}
+\address{Applied Intelligence Research Centre, Dublin Institute of Technology }
+\author{Matthew Klenk}{klenk@parc.com}
+\address{Palo Alto Research Center, Palo Alto CA }
+\author{Nick Hawes}{n.a.hawes@cs.bham.ac.uk}
+\address{Intelligent Robotics Lab, University of Birmingham, UK }
 \vskip 0.2in
@@
 The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
-We call these types of regions \textit{context-dependent spatial regions}(CDSRs) due to the context-dependent nature of their extent.
+We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.
 We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.
@@
 
-Section~\ref{sec:cdsrs} discusses the features of CDSRs using illustrative examples.
-We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.
+We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.
 We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
@@
 \begin{enumerate}
-	\item Given a particular context how well do people agree on the location and extent of a named CDSR?
-	\item Does the orientation and location of objects (with relevant functional semantics) in a given context effect the location and extent of a named CDSR?
+	\item Given a particular context, do people agree on the location and extent of a named CDSR?
+	\item Does changing the context in a given scenario predictably effect the CDSR?
 \end{enumerate}
@@
 
-To examine these questions, we developed two experimental tasks:
+We developed two experimental tasks that each address both research questions:
 \begin{itemize}
-	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the extent of the CDSR in the scene (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well it fit with their understanding of the CDSR on a five point likert scale.
+	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fitted with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
 
-Both Task 1 and 2 address both research questions.
-Between subject agreement metrics can be computed for both Task 1 and Task 2.
-In particular, a variation in between subject agreement across the different contexts for a particular CDSR could be taken as evidence that some contexts are more amenable to a particular CDSR than others and hence CDSRs are contextually dependent.
-By shifting the focus of the analysis to examining the variation in responses for a given CDSR in context that differed only with respect to the location of objects both the results from both tasks can address research question 2.
-In the case of Task 1 the variation we will examine is the area of the polygons, while in Task 2 the variation we will examine is the scoring of the marked locations.
-
-We developed a set of visual stimuli depicting different environments, and paired each with an accompanying linguistic description of a CDSR easily understood within the context of that environment (see Figure \ref{fig:exp_stimuli}).
+We developed a set of visual stimuli depicting different contexts, and paired each with an accompanying linguistic description of a CDSR (see Figure \ref{fig:exp_stimuli}).
 These stimuli were used in both Task 1 and 2.
 Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
-The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the military engagement stimuli were always accompanied by the description \emph{safety}; and (3) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}.
+The stimuli can be broken into three groups based on the situation: (1) the classroom stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the military engagement stimuli were always accompanied by the description \emph{safety}; and (3) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}.
 
@@
 
-\subsection{Participants and Procedure}
-Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
+%\subsection{Participants and Procedure}
+%Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
 43 undergraduate students from two different third level institutes took part in the study.
@@
 
+\textbf{Participants and Procedure:} The experiment was delivered online.
+43 undergraduate students from two third level institutes took part in the study.
+To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.
+
 \subsection{Task 1 - Design}
-In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
-The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
-Subjects drew polygons by clicking to position the vertices with lines being automatically generated between each point indicated as the subject drew.
-There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
+In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
+The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.
+Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.
+There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.
 
@@
 \begin{table}[t]
-\parbox{.35\linewidth}{
+\parbox{.50\linewidth}{
+\centering 
 \begin{tabular}{l c c}
@@
 \end{tabular}
+\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
 \label{tab:stimuli-kappa-scores}
-\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
 }
 \hfill
-\parbox{.65\linewidth}{
+\parbox{.50\linewidth}{
+\centering
 \begin{tabular}{c c}
@@
 \hline
-area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
-maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
+area(g) > area(h) & 291.2 > 276.2* \\
+maxX(f) > maxX(e) & 176.1 > 167.1 \\
 % minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
-minimumX(e-flipped)  &  \\
- > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
-maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
-maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
-maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
+minX(e-flipped)  &  \\
+ > minX(f-flipped) & 336.1 > 321.3* \\
+maxY(i) > maxY(j) & 304.64 > 265.04* \\
+maxY(j) > maxY(k) & 265.04 > 251.71 \\
+maxY(k) = maxY(l) & 265.04 != 290.03 \\
 \hline
 \end{tabular}
+\caption{Comparison of the expected relationships to the mean human responses.}
 \label{tab:region-location-tests}
-\caption{Comparison of the expected relationships to the mean human responses.}
 }
@@
 Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35.
-Looking at the classroom results as a whole, these results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
+Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
 This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.
@@
 
+Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of the same region type between different contexts.
+In five of the six comparisons, the expected relationships were found in human data supporting our claim that changes of context systematically change the extents of the regions.
+The comparison of the safety area with tanks (j) versus turrets (k) is interesting because it provides some evidence that people are simulating actions of other agents in the scene.
+That is, because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
+The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).
+Our expectation was that people would reason that a lone turret would be no match for three tanks and indicate the safe region.
+A possible explanation for this is that there is a domain expertise component to understanding contexts.
+
 \subsection{Task 2 - Design} 
-In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-The location of these points was informed by the results of a pilot study.
-In the final study they were positioned such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
-In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
-One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the classroom domain the expected sweetspot varied between the points depending on the stimuli.
-In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
-In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets.
-P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+In task two, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region, or select not applicable.
+The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli.
+This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
+In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows.
+In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.
+P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).
 
@@
 
-Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain.
-For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.
-In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
-For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points using a confidence level of $95\%$.
+Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context.
+In all context, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.
+To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.
+There was no statistical difference between the stimuli and their flipped variants.
+These likert ratings on individual stimuli supports the claim that people are able to agree on the locations of CDSRs with changing context.
+
 %There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
@@
 
-As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis.
+%As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis.
 These flipped stimuli included the classroom stimuli (c), (e) and (f).
@@
 %lending support for question 2 ...
+
+% The following paragraph should end by saying this about question 1.
 
@@
 \label{tab:point-likert-scores}
-\caption{Average Likert Ratings per Point by Domain and stimulus.}
+\caption{Average Likert Ratings per Point by Sitaution and Stimulus.}
 \end{table}
@@
 Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-
-The connection between spatial regions and function has been explored in other robotics work, e.g.
+%
+%The connection between spatial regions and function has been explored in other robotics work, e.g.
 ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
 Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+
+%JK Previous version above I replaced it with this shorter version below
+
+Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
+Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
+A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
+However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines).
+This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
+While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiment illustrates two limitations.
+First, Regions are not always physically bounded by objects.
+In the classroom examples, people frequently did not extend the front of the room all way to the desks.
+Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
+Finally, the connection between spatial regions and function has been explored in other robotics work, e.g.
+~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
+Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).
+While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.
+
+%--------------------------
 

### 2013-09-13 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-\title{CDSR Human Data/Modelling Paper}
+\title{The Role of Context in Spatial Region Identification}
 
-% \author{Pat Langley}{langley@asu.edu}
-% \author{Glen Hunt}{glen.hunt@asu.edu}
-% \address{Computing Science and Engineering, Arizona State University, 
-%          Tempe, AZ 85287 USA}
-% \author{Daniel G.\ Shapiro}{dgs@isle.org}
-% \address{Institute for the Study of Learning and Expertise, 
-%          2164 Staunton Court, Palo Alto, CA 94306 USA}
+\author{Kate Lockwood}{klockwood@csumb.edu}
+\address{ITCD Department, California State University - Monterey Bay}
+\author{John D. Kelleher}{john.d.kelleher@dit.ie}
+\address{Applied Intelligence Research Centre, Dublin Institute of Technology }
+\author{Matthew Klenk}{klenk@parc.com}
+\address{Palo Alto Research Center, Palo Alto CA }
+\author{Nick Hawes}{n.a.hawes@cs.bham.ac.uk}
+\address{Intelligent Robotics Lab, University of Birmingham, UK }
 \vskip 0.2in
@@
 
-Section~\ref{sec:cdsrs} discusses the features of CDSRs using illustrative examples.
 We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.
@@
 
-Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, or the geographical feature of a bottleneck.
+Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.
 To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
@@
 
-Understanding these regions is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.
-In the rest of the paper we present a study that examined the contextual dependence of these regions and dicuss the implications that these t.
+Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.
+In the rest of the paper we present a study that examines the contextual dependence of these regions.
 Following this we discuss the implications of these regions for the development of cognitive systems.
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (more bluntly, does a classroom always have a front or is this dependent on the configuration of objects in the room), and (b) the variation in CDSR placement and size within different context.
+Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g.
+does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts.
 Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
@@
 
-To examine these questions, we developed two experimental tasks:
+We developed two experimental tasks that each address both research questions:
 \begin{itemize}
 	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fitted with their understanding of the CDSR on a five point likert scale.
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
 
-Both Task 1 and 2 address both research questions.
-Between subject agreement metrics can be computed for both Task 1 and Task 2.
-In particular, a variation in between subject agreement across the different contexts for a particular CDSR could be taken as evidence that some contexts are more amenable to a particular CDSR than others and hence CDSRs are contextually dependent.
-By shifting the focus of the analysis to examining the variation in responses for a given CDSR in context that differed only with respect to the location of objects both the results from both tasks can address research question 2.
-In the case of Task 1 the variation we will examine is the area of the polygons, while in Task 2 the variation we will examine is the scoring of the marked locations.
-
-We developed a set of visual stimuli depicting different environments, and paired each with an accompanying linguistic description of a CDSR easily understood within the context of that environment (see Figure \ref{fig:exp_stimuli}).
+We developed a set of visual stimuli depicting different contexts, and paired each with an accompanying linguistic description of a CDSR (see Figure \ref{fig:exp_stimuli}).
 These stimuli were used in both Task 1 and 2.
 Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
-The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the military engagement stimuli were always accompanied by the description \emph{safety}; and (3) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}.
+The stimuli can be broken into three groups based on the situation: (1) the classroom stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the military engagement stimuli were always accompanied by the description \emph{safety}; and (3) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}.
 
@@
 We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
-To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts.
+To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images.
 Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).
@@
 \subsection{Task 1 - Design}
-In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
-The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.
-Subjects drew polygons by clicking to position the vertices with lines being automatically generated between each point indicated as the subject drew.
-There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.
+In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
+The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.
+Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.
+There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.
 
 \subsection{Task 1 - Results}
-We use two methods to analyze the results of region drawing task.
-First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
+We use two methods to analyze the results of the region drawing task.
+First, we use interannotator agreement for two reasons: (1) fair to good overall agreement, would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 Second, we use the extent points of the regions to assess changes between trials.
-That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.
+That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.
 
-%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
-% but want to talk to John about it during the telecon tomorrow first
+%Kate:  (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with %context,  I changed this SUBSTANTIALLY to something that I think sounds better, but makes a different argument.
+Let me know if it is ok.
+Really %agreement would indicated shared understanding while variation between contexts would lend support to the importance of context.
+Not sure I %captured this
 
@@
 It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.
-Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35.
-Looking at the classroom results as a whole, these results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
+Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35.
+Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
 This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.
@@
 
+Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.
+In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.
+The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.
+Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
+The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).
+A possible explanation for this is that there is a domain expertise component to understanding contexts.
+
+%Kate: I took out "Our expectation was that people would reason that a lone turret would be no match for three tanks and indicate the safe region. " because it didn't make sense to me.
+
 \subsection{Task 2 - Design} 
-In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.
-The location of these points was informed by the results of a pilot study.
-In the final study they were positioned such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
-In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}).
-One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the classroom domain the expected sweetspot varied between the points depending on the stimuli.
-In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
-In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets.
-P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.
+In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.
+The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR.
+In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli.
+This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
+In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows.
+In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.
+P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).
 
@@
 
-Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain.
-For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.
-In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli.
-For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points using a confidence level of $95\%$.
+Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context.
+In all contexts, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.
+To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.
+There was no statistical difference between the stimuli and their flipped variants.
+These likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.
+
 %There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
@@
 
-Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l).
+% The following paragraph should end by saying this about question 1.
+
+Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l).
 In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land.
-Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j).
-As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j).
-This higher rating is statiscially significant at a $95\%$ confidence level.
-Stimuli (k) and (l) present and analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
+Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
+As the results in Table \ref{tab:point-likert-scores} show this is the case with this higher rating statiscially significant at a $95\%$ confidence level.
+Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
 Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
@@
 However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines).
-This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiemnt illustrates two limitations.
+This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
+While satisfying the requirements of a cognitive systems solution, applying this model to the results of our experiment illustrates two limitations.
 First, Regions are not always physically bounded by objects.
 In the classroom examples, people frequently did not extend the front of the room all way to the desks.
-Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
+Second, people consider dynamics of a context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
 Finally, the connection between spatial regions and function has been explored in other robotics work, e.g.
@@
 Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).
+While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.
 
@@
 In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.
-We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment resulted predicted changes in the location, size, and shape of the spatial region.
+We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.
 We outline requirements for a cognitive systems solution and discuss limitations in current approaches from cognitive systems and robotics.

### 2013-09-13 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk fixed some typos

@@
 
-\title{CDSR Human Data/Modelling Paper}
+\title{The Role of Context in Spatial Region Identification}
 
@@
 	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fitted with their understanding of the CDSR on a five point likert scale.
+	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
@@
 \subsection{Task 1 - Results}
-We use two methods to analyze the results of region drawing task.
-First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
+We use two methods to analyze the results of the region drawing task.
+First, we use interannotator agreement for two reasons: (1) fair to good overall agreement, would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 Second, we use the extent points of the regions to assess changes between trials.
-That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.
+That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.
 
-%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
-% but want to talk to John about it during the telecon tomorrow first
+%Kate:  (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with %context,  I changed this SUBSTANTIALLY to something that I think sounds better, but makes a different argument.
+Let me know if it is ok.
+Really %agreement would indicated shared understanding while variation between contexts would lend support to the importance of context.
+Not sure I %captured this
 
@@
 It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.
-Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout score have kappa scores of less than 0.35.
+Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35.
 Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
@@
 
-Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of the same region type between different contexts.
-In five of the six comparisons, the expected relationships were found in human data supporting our claim that changes of context systematically change the extents of the regions.
-The comparison of the safety area with tanks (j) versus turrets (k) is interesting because it provides some evidence that people are simulating actions of other agents in the scene.
-That is, because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
+Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.
+In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.
+The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.
+Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
 The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).
-Our expectation was that people would reason that a lone turret would be no match for three tanks and indicate the safe region.
 A possible explanation for this is that there is a domain expertise component to understanding contexts.
 
+%Kate: I took out "Our expectation was that people would reason that a lone turret would be no match for three tanks and indicate the safe region. " because it didn't make sense to me.
+
 \subsection{Task 2 - Design} 
-In task two, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region, or select not applicable.
-The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region.
+In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.
+The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR.
 In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli.
@@
 Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context.
-In all context, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.
+In all contexts, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.
 To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.
 There was no statistical difference between the stimuli and their flipped variants.
-These likert ratings on individual stimuli supports the claim that people are able to agree on the locations of CDSRs with changing context.
+These likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.
 
@@
 
-Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l).
+Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l).
 In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land.
-Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j).
-As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j).
-This higher rating is statiscially significant at a $95\%$ confidence level.
-Stimuli (k) and (l) present and analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
+Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
+As the results in Table \ref{tab:point-likert-scores} show this is the case with this higher rating statistically significant $(p<.05)$.
+Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
 Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
@@
 This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiment illustrates two limitations.
+While satisfying the requirements of a cognitive systems solution, applying this model to the results of our experiment illustrates two limitations.
 First, Regions are not always physically bounded by objects.
 In the classroom examples, people frequently did not extend the front of the room all way to the desks.
-Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
+Second, people consider dynamics of a context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
 Finally, the connection between spatial regions and function has been explored in other robotics work, e.g.

### 2013-09-13 scenarios/cdsr/trunk/papers/ACS2013/main.tex
minor fixes

@@
 \author{Nick Hawes}{n.a.hawes@cs.bham.ac.uk}
-\address{Intelligent Robotics Lab, University of Birmingham, UK }
+\address{School of Computer Science, University of Birmingham, UK }
 \vskip 0.2in
@@
 This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli.
 In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows.
@@
 }
+\caption{Average Likert Ratings per Point by Sitaution and Stimulus.}
 \label{tab:point-likert-scores}
-\caption{Average Likert Ratings per Point by Sitaution and Stimulus.}
 \end{table}

### 2013-09-13 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.
-We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.
-We discss the implications for these results on cognitive systems development.
+We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing a cognitive systems with this ability.
 \end{abstract}
@@
 Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.
-However, many spatial terms are defined not only by their geometry, but also their context and their relevance to current goals.
-That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration and functional roles.
+However, many spatial terms are defined not only by their geometry, but also their context.
+That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration, functional roles, and goals.
 We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).
@@
 We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.
-We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
 To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
 In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
+We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
 Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.
@@
 Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.
-To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
+To determine the extent of these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
 Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.
@@
 In the rest of the paper we present a study that examines the contextual dependence of these regions.
-Following this we discuss the implications of these regions for the development of cognitive systems.
+Following this we discuss how the results align with previous research and requirements for a cognitive systems model of this ability.
 
@@
 does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts.
-Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
 
@@
 \begin{itemize}
-	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
+	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stating there was no such region).
+	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
 \end{itemize}
@@
 The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.
-Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
+Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).
 
@@
 \subsection{Task 1 - Design}
-In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
+In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a "no region" option.
 The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.
@@
 We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
-After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
-In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
-$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
+After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
+In this calculation, if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
+$\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.
 Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.
@@
 Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+%\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
 
@@
 \hline
-area(g) > area(h) & 291.2 > 276.2* \\
+area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
 maxX(f) > maxX(e) & 176.1 > 167.1 \\
@@
 minX(e-flipped)  &  \\
- > minX(f-flipped) & 336.1 > 321.3* \\
-maxY(i) > maxY(j) & 304.64 > 265.04* \\
+ > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
+maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
 maxY(j) > maxY(k) & 265.04 > 251.71 \\
@@
 \end{table}
+
+\footnotetext[2]{Significance value from T-Test (p<0.1)}
 
@@
 Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35.
-Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
-This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.
+Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
+This dependence is observed by the fact that just moving the desks into a circle configuration the shared understanding of the front of the room no longer exists.
 
@@
 The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).
+That is, a single turret does not provide safety from three tanks.
 A possible explanation for this is that there is a domain expertise component to understanding contexts.
@@
 This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).
+In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli.
 In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows.
-In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.
-P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).
+In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it lay behind the lone turret but in front of the row turrets.
+P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.
 
@@
 Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
-As the results in Table \ref{tab:point-likert-scores} show this is the case with this higher rating statiscially significant at a $95\%$ confidence level.
+This is confirmed by the results in Table \ref{tab:point-likert-scores} and the higher rating is statistically significant $(p<.05)$.
 Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
 Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
+Thus, it is unclear the extent to which particiapnts were simulating the outcome of the battle.
 Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high.
@@
 
-Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
+Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
 
@@
 
-Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
+Cognitive systems seek to reproduce the full range of human behavior.
+In addition to reproducing these results, a cognitive systems solution must learn about new regions and apply this knowledge in multiple tasks.
+While spatial reasoning is an established area in AI and robotics, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
 Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
-A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
+A fundamental decision when creating a computational model of a CDSR is the representation of its boundary.
 Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
@@
 This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-While satisfying the requirements of a cognitive systems solution, applying this model to the results of our experiment illustrates two limitations.
-First, Regions are not always physically bounded by objects.
+One limitation of the anchor point approach is that regions are not always physically bounded by objects.
 In the classroom examples, people frequently did not extend the front of the room all way to the desks.
-Second, people consider dynamics of a context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-Finally, the connection between spatial regions and function has been explored in other robotics work, e.g.
+The connection between spatial regions and function has been explored in other robotics work, e.g.
 ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
@@
 McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).
-While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.
+While this approach incorporates functional aspects of the geography, it is unclear how it would apply to the other situations in our experiment.
 
@@
 We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.
-We outline requirements for a cognitive systems solution and discuss limitations in current approaches from cognitive systems and robotics.
+We posit that a cognitive system should be able to learn new regions, account for changing contexts, and apply this knoweldge in multiple reasoning tasks.
 Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.

### 2013-11-09 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk: correct small typos

@@
 We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.
-We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing a cognitive systems with this ability.
+We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing cognitive systems with this ability.
 \end{abstract}
@@
 Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.
-However, many spatial terms are defined not only by their geometry, but also their context.
-That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration, functional roles, and goals.
+However, many spatial terms are defined not only by their geometry, but also by their context.
+That is, to understand this language requires considering the other objects and agents in the environment as well as their configuration, functional roles, and goals.
 We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).
 
-We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.
-To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
+We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.
+To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
 In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
@@
 
-Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.
+Consider the following regions shown in Figure \ref{fig:examples}: the front of a classroom, safety in a military engagement, and a geographic bottleneck.
 To determine the extent of these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
@@
 	\item Given a particular context, do people agree on the location and extent of a named CDSR?
-	\item Does changing the context in a given scenario predictably effect the CDSR?
+	\item Does changing the context in a given scenario predictably affect the CDSR?
 \end{enumerate}
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g.
-does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts.
-Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given environment (e.g., does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts.
+Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follows the variations  in the given context.
 
@@
 \begin{itemize}
-	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stating there was no such region).
-	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
+	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimulus that defined the CDSR (or stating there was no such region).
+	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point Likert scale.
 \end{itemize}
@@
 For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom.
-We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}.
+We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli.}.
 To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images.
@@
 The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.
-Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).
+Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:turrets-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).
 
@@
 }
-\caption{Experiment Stimuli}
+\caption{The stimuli used in the experiment.}
 \label{fig:exp_stimuli}
@@
 %Using these stimuli pairs, we developed an online experiment with two different types of trials to explore our initial questions.
-43 undergraduate students from two different third level institutes took part in the study.
-To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.
+43 undergraduate students from two different third-level institutes took part in the study.
+To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimulus presented within each trial.
 
 \textbf{Participants and Procedure:} The experiment was delivered online.
-43 undergraduate students from two third level institutes took part in the study.
+43 undergraduate students from two third-level institutes took part in the study.
 To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.
@@
 \subsection{Task 1 - Design}
-In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a "no region" option.
+In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
 The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.
 Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.
-There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.
+There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.
 
@@
 We use two methods to analyze the results of the region drawing task.
-First, we use interannotator agreement for two reasons: (1) fair to good overall agreement, would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
+First, we use interannotator agreement for two reasons: (1) fair to good overall agreement would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 Second, we use the extent points of the regions to assess changes between trials.
@@
 We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.
-We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
-After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
-In this calculation, if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
+We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon.
+After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimulus basis.
+In this calculation, if either of the subjects in a pair entered a no region response for a particular stimulus a $\kappa$ value of 0 was used.
 $\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.
@@
 %However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects.
-Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
+Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimulus, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimulus.
 \renewcommand{\thefootnote}{\fnsymbol{footnote}}
@@
 \end{tabular}
-\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
+\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimulus.}
 \label{tab:stimuli-kappa-scores}
@@
 
-\footnotetext[2]{Significance value from T-Test (p<0.1)}
+\footnotetext[2]{Significance value from t-test (p<0.1)}
 
@@
 %\label{tab:stimuli-kappa-scores}
-%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimuli.
+%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimulus.
 The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
@@
 
-Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
+Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimulus used in the experiment and the number of no region responses for each stimulus.
 It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.
-Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35.
-Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
+Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flipped), and the circle layout have kappa scores of less than 0.35.
+Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is affected by the location and orientation of objects in the environment.
 This dependence is observed by the fact that just moving the desks into a circle configuration the shared understanding of the front of the room no longer exists.
@@
 %}
-%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
+%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimulus, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
 %\label{fig:type-1-hists}
@@
 Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
-The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).
+The case that did not demonstrate the expected relationship was the safety region between the row of turrets (k) and turrets +1(l).
 That is, a single turret does not provide safety from three tanks.
@@
 \subsection{Task 2 - Design} 
-In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.
+In task two, a location was marked on the stimulus (using a red dot) and the subject was asked to rank on a five point Likert scale how well the location matched the given CDSR, or select not applicable.
 The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR.
-In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli.
+In the classroom domain, the three points were located along the horizontal line that bisected the stimulus with P1 positioned on the left of the stimulus, P2 in the center and P3 on the right of the stimulus.
 This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli.
-In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows.
-In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it lay behind the lone turret but in front of the row turrets.
+In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.
+In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image as follows.
+In stimuli (j) and (k) the point lay between all forces, in stimulus (i), it lay behind the friendly tanks, and in stimulus (l) , it lay behind the lone turret but in front of the row turrets.
 P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.
@@
 
-%In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
+%In the type 1 trials subjects drew a polygon on the visual stimulus that defined the location and extent of the relevant spatial region.
 In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
 
-%The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched %the described spatial region.
-For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
+%The type 2 trials the same visual stimulus and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimulus (using a red dot) and the subject was asked to rank on a five point Likert scale their judgement of how well the location matched %the described spatial region.
+For each stimulus domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
 
-Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context.
-In all contexts, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.
-To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.
+Table \ref{tab:point-likert-scores} lists the average Likert scores for the different points in each context.
+In all contexts, the expected sweetspot received the maximum average Likert rating that was statistically significant using one-tailed two-sample t-tests $(p<.05)$.
+To control for left--right bias, we compared the ratings of sweetspots in the classroom stimuli (c), (e), and (f) with their flipped variants.
 There was no statistical difference between the stimuli and their flipped variants.
-These likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.
+These Likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.
 
-%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
-Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli.
+%There were two of these t-tests carried out for each stimulus, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
+Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimulus.
 In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
@@
 Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3.
-To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
+To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two-sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
 In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$.
@@
 Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l).
-In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land.
+In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in no-man's land.
 Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
 This is confirmed by the results in Table \ref{tab:point-likert-scores} and the higher rating is statistically significant $(p<.05)$.
-Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
-Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
-Thus, it is unclear the extent to which particiapnts were simulating the outcome of the battle.
+Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimulus (l) but in no-man's land in stimulus (k).
+Again, the average rating of P2 for stimulus (l) is higher than that of (k), although this difference is not statistically significant.
+Thus, it is unclear the extent to which participants were simulating the outcome of the battle.
 Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high.
@@
 }
-\caption{Average Likert Ratings per Point by Sitaution and Stimulus.}
+\caption{Average Likert ratings per point by situation and stimulus.}
 \label{tab:point-likert-scores}
@@
 % ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
-%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample t-test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
 %(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
@@
 % ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
-%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample t-test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
 %(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
@@
 % ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
-%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
+%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample t-test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
 %(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
@@
 %\label{tab:point-likert-scores}
-%\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
+%\caption{Average Likert Ratings per Point by Domain with t-test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimulus}
 %\end{table}
@@
 %\label{tab:flipped-point-likert-scores}
-%\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
+%\caption{Two Tailed Two Sample t-tests for flipped stimuli pair trials.}
 %\end{table}
 
-Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
+Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimulus (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
 
@@
 %
-%Task 3: The participant grades a point with respect to membership of the region on a likert scale
+%Task 3: The participant grades a point with respect to membership of the region on a Likert scale
 %I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
 %I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
-%Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
+%Assuming that there is a variation across the Likert scores for a particular stimulus then check if their is a correlation between the scores and:
 %(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
@@
 First, Regions are not always physically bounded by objects.
-In the classroom examples, people frequently did not extend the front of the room all way to the desks.
+In the classroom examples, people frequently did not extend the front of the room all the way to the desks.
 Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
@@
 Cognitive systems seek to reproduce the full range of human behavior.
-In addition to reproducing these results, a cognitive systems solution must learn about new regions and apply this knowledge in multiple tasks.
+In addition to reproducing these results, a cognitive systems solution must learn about new regions and apply this knowledge to multiple tasks.
 While spatial reasoning is an established area in AI and robotics, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
-Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
+Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robotics (see, \emph{inter alia,} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
 A fundamental decision when creating a computational model of a CDSR is the representation of its boundary.
@@
 One limitation of the anchor point approach is that regions are not always physically bounded by objects.
-In the classroom examples, people frequently did not extend the front of the room all way to the desks.
+In the classroom examples, people frequently did not extend the front of the room all the way to the desks.
 The connection between spatial regions and function has been explored in other robotics work, e.g.
@@
 McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).
-While this approach incorporates functional aspects of the geography, it is unclear how it would apply to the other situations in our experiment.
+Although this approach incorporates functional aspects of the geography, it is unclear how it would apply to the other situations in our experiment.
 
@@
 Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
-In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.
-We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.
-We posit that a cognitive system should be able to learn new regions, account for changing contexts, and apply this knoweldge in multiple reasoning tasks.
+In this paper, we presented a human study to develop requirements for reasoning about context-dependent spatial regions for cognitive systems.
+We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.
+We posit that a cognitive system should be able to learn new regions, account for changing contexts, and apply this knowledge in multiple reasoning tasks.
 Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.

### 2013-11-10 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk fixed more typos

@@
   \includegraphics[width=\textwidth]{figures/classroom.png}
-%  \caption{\textit{The front of the classroom} could be defined as the area between the desks and the whiteboard where people give presentations.}
-  \caption{\textit{The front of the classroom}.}
+%  \caption{``The front of the classroom'' could be defined as the area between the desks and the whiteboard where people give presentations.}
+  \caption{``The front of the classroom''.}
   \end{subfigure} &
@@
   \includegraphics[width=\textwidth]{figures/safety-3.jpg}
-%  \caption{\textit{Safety for the scout} is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
-  \caption{\textit{The safe region for the scout}.}
+%  \caption{Safety for the scout is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
+  \caption{``The safe region for the scout''.}
 \end{subfigure} & 
@@
   \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
-%  \caption{A \textit{bottleneck} occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
- \caption{\textit{The bottleneck}.}
+%  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
+ \caption{``The bottleneck''.}
   \label{fig:safety}
@@
 	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimulus that defined the CDSR (or stating there was no such region).
-	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point Likert scale.
+	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point \textit{Likert scale}.
 \end{itemize}
@@
 Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli.
-The stimuli can be broken into three groups based on the situation: (1) the classroom stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the military engagement stimuli were always accompanied by the description \emph{safety}; and (3) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}.
+The stimuli can be broken into three groups based on the situation: (1) the classroom stimuli were always accompanied by the spatial description ``the front of the classroom''; (2) the military engagement stimuli were always accompanied by the description ``safety''; and (3) the geographical map stimuli were always accompanied by the description ``the bottleneck''.
 
@@
 
-We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.
+We use \textit{Cohen's kappa} coefficient, $\kappa$, to measure the agreement between two regions as follows.
 We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon.
@@
 
-\begin{table}[t]
-\parbox{.50\linewidth}{
-\centering 
+%\begin{table}[t]
+%\parbox{.50\linewidth}{
+%\centering 
+%\begin{tabular}{l c c}
+%% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+% & Average & No \\
+%Stim ID & Pairwise $\kappa$ & Reg.\\
+%\hline
+%\hline
+%(h) & 0.5034 & 0\\ % 1_2
+%(g) & 0.4585 & 1\\ % 1_1
+%(k) & 0.4548 & 0\\ % 3_3
+%(f) & 0.3930 & 0\\ % 2_4
+%(i) & 0.3906 & 0\\ % 3_1
+%(e-flipped) & 0.3813 & 2\\ % 2_3
+%(j) & 0.3787 & 1\\ % 3_2
+%(f-flipped) & 0.3720 & 2\\ % 2_5
+%(e) & 0.3629 & 1\\ %2_2
+%(l) & 0.3164& 0\\ % 3_4
+%(c) & 0.2438 & 2\\ % 2_6
+%(c-flipped) & 0.2414 & 4\\ %2_7
+%(a) & 0.0996 & 10 \\ % 2_1
+%\hline
+%\end{tabular}
+%\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimulus.}
+%\label{tab:stimuli-kappa-scores}
+%}
+%\hfill
+%\parbox{.50\linewidth}{
+%\centering
+%\begin{tabular}{c c}
+%Expected Relationship & Human Results  \\
+%\hline
+%\hline
+%area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
+%maxX(f) > maxX(e) & 176.1 > 167.1 \\
+%% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
+%minX(e-flipped)  &  \\
+% > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
+%maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
+%maxY(j) > maxY(k) & 265.04 > 251.71 \\
+%maxY(k) = maxY(l) & 265.04 != 290.03 \\
+%\hline
+%\end{tabular}
+%\caption{Comparison of the expected relationships to the mean human responses.}
+%\label{tab:region-location-tests}
+%}
+%\end{table}
+%\footnotetext[2]{Significance value from t-test (p<0.1)}
+
+\begin{table}[htb]
+\begin{center}
 \begin{tabular}{l c c}
 % Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
- & Average & No \\
-Stim ID & Pairwise $\kappa$ & Reg.\\
+Stim ID & Average Pairwise $\kappa$ & No Reg.\\
 \hline
@@
 \end{tabular}
+\end{center}
 \caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimulus.}
 \label{tab:stimuli-kappa-scores}
-}
-\hfill
-\parbox{.50\linewidth}{
-\centering
-\begin{tabular}{c c}
-Expected Relationship & Human Results  \\
+\end{table}
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{r c l r c l}
+\multicolumn{3}{c}{Expected Relationship} & \multicolumn{3}{c}{Human Results}  \\
 \hline
 \hline
-area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
-maxX(f) > maxX(e) & 176.1 > 167.1 \\
-% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
-minX(e-flipped)  &  \\
- > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
-maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
-maxY(j) > maxY(k) & 265.04 > 251.71 \\
-maxY(k) = maxY(l) & 265.04 != 290.03 \\
+area(g) &>& area(h) & 291.2 & >& 276.2* \\
+maxX(f) &>& maxX(e) & 176.1& >& 167.1 \\
+% minimumX(e-flipped) &>& minimumX(f-flipped) & 336.1& >& 321.3 (p<.12) \\
+minX(e-flipped) &>& minX(f-flipped) & 336.1& >& 321.3* \\
+maxY(i) &>& maxY(j) & 304.64 &>& 265.04* \\
+maxY(j) &>& maxY(k) & 265.04 &>& 251.71 \\
+maxY(k) &=& maxY(l) & 265.04 &!=& 290.03 \\
 \hline
 \end{tabular}
-\caption{Comparison of the expected relationships to the mean human responses.}
+\end{center}
+\caption{Comparison of the expected relationships to the mean human responses. *Significance value from t-test (p<0.1).}
 \label{tab:region-location-tests}
-}
 \end{table}
-
-\footnotetext[2]{Significance value from t-test (p<0.1)}
 
@@
 Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is affected by the location and orientation of objects in the environment.
-This dependence is observed by the fact that just moving the desks into a circle configuration the shared understanding of the front of the room no longer exists.
+An extreme example of this relationship between the configuration of objects in a context and the meaning of a CDSR is that our results indicate that the shared understanding of ``the front of the room'' breaks down in the classroom setting when the chairs in the room are moved into a circle configuration as indicated by the fact that context (a) (Figure \ref{fig:classroom-a}) had an extremly low average pairwise agreement of  $0.0996$).
 
@@
 Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment.
-For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}.
+For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely ``the front of the room''.
 Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision context $a$ had an average pairwise agreement of $0.0996$.
@@
 \subsection{Task 2 - Design} 
-In task two, a location was marked on the stimulus (using a red dot) and the subject was asked to rank on a five point Likert scale how well the location matched the given CDSR, or select not applicable.
-The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR.
+In task two, a location was marked on the stimulus (using a red dot) and the subject was asked to rank on a five point \textit{Likert scale} how well the location matched the given CDSR, or select not applicable.
+The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textit{sweetspot}, was located near the expected prototypical center of the CDSR.
 In the classroom domain, the three points were located along the horizontal line that bisected the stimulus with P1 positioned on the left of the stimulus, P2 in the center and P3 on the right of the stimulus.
 This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.
+In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected \textit{sweetspot} and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.
 In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image as follows.
 In stimuli (j) and (k) the point lay between all forces, in stimulus (i), it lay behind the friendly tanks, and in stimulus (l) , it lay behind the lone turret but in front of the row turrets.
-P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.
+P3 was the expected \textit{sweetspot} in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.
 
@@
 %---- For each of these N points check whether the participant's polygon and the [intersection/union] polygon both include it (agree++) or if one includes it and the other doesn't (disagree++)
-%---- Then we calculate a Cohen Kappa score as a measure of agreement between the participant and eveyone else.
+%---- Then we calculate a \textit{Cohen kappa} score as a measure of agreement between the participant and eveyone else.
 Note we need to calculate the chance of random agreement but we can do this as a function over the relative size of the participant's polygon in the total area of the stimulus and the relative size of the intersection/union polygon
@@
 
-%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
+%\textit{Cohen's kappa} measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
 %\begin{equation*}
@@
 
-%The type 2 trials the same visual stimulus and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimulus (using a red dot) and the subject was asked to rank on a five point Likert scale their judgement of how well the location matched %the described spatial region.
+%The type 2 trials the same visual stimulus and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimulus (using a red dot) and the subject was asked to rank on a five point \textit{Likert scale} their judgement of how well the location matched %the described spatial region.
 For each stimulus domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
@@
 Table \ref{tab:point-likert-scores} lists the average Likert scores for the different points in each context.
-In all contexts, the expected sweetspot received the maximum average Likert rating that was statistically significant using one-tailed two-sample t-tests $(p<.05)$.
-To control for left--right bias, we compared the ratings of sweetspots in the classroom stimuli (c), (e), and (f) with their flipped variants.
+In all contexts, the expected \textit{sweetspot} received the maximum average Likert rating that was statistically significant using one-tailed two-sample t-tests $(p<.05)$.
+To control for left--right bias, we compared the ratings of \textit{sweetspots} in the classroom stimuli (c), (e), and (f) with their flipped variants.
 There was no statistical difference between the stimuli and their flipped variants.
@@
 
-%There were two of these t-tests carried out for each stimulus, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
+%There were two of these t-tests carried out for each stimulus, each one tested whether the \textit{sweetspot} ranking was statistically higher than one of the other points.
 Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimulus.
-In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
+In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the \textit{sweetspot} was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
 
@@
 These flipped stimuli included the classroom stimuli (c), (e) and (f).
-This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials.
-Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3.
-To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two-sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
-In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$.
-In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
+This flipping of stimuli gave us the opportunity to compare the results of the \textit{sweetspots} rankings between this pairs of stimuli trials.
+Given the horizontal flipping of the stimuli it was expected that the \textit{sweetspot} for the stimuli should also be flipped between P1 and P3.
+To test whether or not the expected shift in the \textit{sweetspots} occurred we ran two-tailed two-sample t-tests between the \textit{sweetspots} of the flipped and non-flipped stimuli pairs.
+In all cases the results indicated that there was no statistical difference between the rating of the \textit{sweetspot} for the non-flipped stimuli and the rating of the \textit{sweetspot} for the flipped stimuli using a confidence level of $95\%$.
+In other words, the \textit{sweetspot} flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
 %lending support for question 2 ...
@@
 Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l).
-In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in no-man's land.
+In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.
 Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
@@
 Thus, it is unclear the extent to which participants were simulating the outcome of the battle.
-Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high.
+Notice, that the rating of the \textit{sweetspot} for the stimuli (i) and (l) remain quite high.
 This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR.
-Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
+Overall, the analysis of the ratings of the different \textit{sweetspots} across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
 
@@
 
-Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimulus (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
+Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the \textit{sweetspot} was determined by the particular context depicted in the visual stimulus (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the \textit{sweetspot} flipped from P1 to P3 when the context was flipped.
 
@@
 %
-%Task 2: The participant selected a point denoting the sweetspot of the region
+%Task 2: The participant selected a point denoting the \textit{sweetspot} of the region
 %Analysis of results:
@@
 %
-%Task 3: The participant grades a point with respect to membership of the region on a Likert scale
+%Task 3: The participant grades a point with respect to membership of the region on a \textit{Likert scale}
 %I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
@@
 While spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
-Regions are central to QSR work~\cite{Cohn:2001}.
+Regions are central to qualiative spatial reasoning (QSR) work~\cite{Cohn:2001}.
 Regions can either have crisp boundaries, usually defined by geometric shapes (e.g.
@@
 Cognitive systems seek to reproduce the full range of human behavior.
-In addition to reproducing these results, a cognitive systems solution must learn about new regions and apply this knowledge to multiple tasks.
+Spatial reasoning is an important element of human behavior.
 While spatial reasoning is an established area in AI and robotics, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
-Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robotics (see, \emph{inter alia,} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
+Regions are central to qualitative spatial reasoning (QSR) work~\cite{Cohn:2001} and also in the area of cognitive robotics (see, \emph{inter alia,} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
 A fundamental decision when creating a computational model of a CDSR is the representation of its boundary.
@@
 This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-One limitation of the anchor point approach is that regions are not always physically bounded by objects.
+One limitation of the \textit{anchor point} approach is that regions are not always physically bounded by objects.
 In the classroom examples, people frequently did not extend the front of the room all the way to the desks.
@@
 An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.
-In the know example, a human defines the front of the classroom as a polygon consisting of four anchor points.
+In the know example, a human defines the front of the classroom as a polygon consisting of four \textit{anchor points}.
 Analogical transfer works by aligning structured representations and transferring partially aligned statements from the known example to the new environment.
-In this case, the anchor points are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.
+In this case, the \textit{anchor points} are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.
 
@@
 %  \includegraphics[width=\columnwidth]{figures/classroom-transfer.png}
-%  \caption{The known example defines the front of the room with four anchor points.
-By transferring the anchor points to the new environment, the system is able to ground the room's front in its perception.}
+%  \caption{The known example defines the front of the room with four \textit{anchor points}.
+By transferring the \textit{anchor points} to the new environment, the system is able to ground the room's front in its perception.}
 %  \label{fig:anchor-point-transfer}

### 2013-11-13 scenarios/cdsr/trunk/papers/ACS2013/main.tex
edited 4 and 5

@@
 We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.
-We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.
-We discss the implications for these results on cognitive systems development.
+We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing cognitive systems with this ability.
 \end{abstract}
@@
 Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.
-However, many spatial terms are defined not only by their geometry, but also their context and their relevance to current goals.
-That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration and functional roles.
+However, many spatial terms are defined not only by their geometry, but also by their context.
+That is, to understand this language requires considering the other objects and agents in the environment as well as their configuration, functional roles, and goals.
 We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).
 
-We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.
-We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
-To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
+We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.
+To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
 In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
+We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
 Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.
@@
 
-Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.
-To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
+Consider the following regions shown in Figure \ref{fig:examples}: the front of a classroom, safety in a military engagement, and a geographic bottleneck.
+To determine the extent of these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.
 Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.
@@
 Both of these research questions address the fundamental claim that CDSRs are contextually defined.
-Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g.
-does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts.
-Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.
+Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given environment (e.g., does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts.
+Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follows the variations  in the given context.
 
@@
 \begin{itemize}
-	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
-	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
+	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimulus that defined the CDSR (or stating there was no such region).
+	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point \textit{Likert scale}.
 \end{itemize}
@@
 The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.
-Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).
+Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:turrets-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).
 
@@
 Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.
-There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.
+There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.
 
@@
 
-We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.
-We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon.
-After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis.
-In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.
-$\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.
+We use \textit{Cohen's kappa} coefficient, $\kappa$, to measure the agreement between two regions as follows.
+We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon.
+After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimulus basis.
+In this calculation, if either of the subjects in a pair entered a no region response for a particular stimulus a $\kappa$ value of 0 was used.
+$\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.
 Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.
@@
 %However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects.
-Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.
+Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimulus, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimulus.
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+%\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
 
-\begin{table}[t]
-\parbox{.50\linewidth}{
-\centering 
+%\begin{table}[t]
+%\parbox{.50\linewidth}{
+%\centering 
+%\begin{tabular}{l c c}
+%% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
+% & Average & No \\
+%Stim ID & Pairwise $\kappa$ & Reg.\\
+%\hline
+%\hline
+%(h) & 0.5034 & 0\\ % 1_2
+%(g) & 0.4585 & 1\\ % 1_1
+%(k) & 0.4548 & 0\\ % 3_3
+%(f) & 0.3930 & 0\\ % 2_4
+%(i) & 0.3906 & 0\\ % 3_1
+%(e-flipped) & 0.3813 & 2\\ % 2_3
+%(j) & 0.3787 & 1\\ % 3_2
+%(f-flipped) & 0.3720 & 2\\ % 2_5
+%(e) & 0.3629 & 1\\ %2_2
+%(l) & 0.3164& 0\\ % 3_4
+%(c) & 0.2438 & 2\\ % 2_6
+%(c-flipped) & 0.2414 & 4\\ %2_7
+%(a) & 0.0996 & 10 \\ % 2_1
+%\hline
+%\end{tabular}
+%\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimulus.}
+%\label{tab:stimuli-kappa-scores}
+%}
+%\hfill
+%\parbox{.50\linewidth}{
+%\centering
+%\begin{tabular}{c c}
+%Expected Relationship & Human Results  \\
+%\hline
+%\hline
+%area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
+%maxX(f) > maxX(e) & 176.1 > 167.1 \\
+%% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
+%minX(e-flipped)  &  \\
+% > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
+%maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
+%maxY(j) > maxY(k) & 265.04 > 251.71 \\
+%maxY(k) = maxY(l) & 265.04 != 290.03 \\
+%\hline
+%\end{tabular}
+%\caption{Comparison of the expected relationships to the mean human responses.}
+%\label{tab:region-location-tests}
+%}
+%\end{table}
+%\footnotetext[2]{Significance value from t-test (p<0.1)}
+
+\begin{table}[htb]
+\begin{center}
 \begin{tabular}{l c c}
 % Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
- & Average & No \\
-Stim ID & Pairwise $\kappa$ & Reg.\\
+Stim ID & Average Pairwise $\kappa$ & No Reg.\\
 \hline
@@
 \end{tabular}
-\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
+\end{center}
+\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimulus.}
 \label{tab:stimuli-kappa-scores}
-}
-\hfill
-\parbox{.50\linewidth}{
-\centering
-\begin{tabular}{c c}
-Expected Relationship & Human Results  \\
+\end{table}
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{r c l r c l}
+\multicolumn{3}{c}{Expected Relationship} & \multicolumn{3}{c}{Human Results}  \\
 \hline
 \hline
-area(g) > area(h) & 291.2 > 276.2* \\
-maxX(f) > maxX(e) & 176.1 > 167.1 \\
-% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
-minX(e-flipped)  &  \\
- > minX(f-flipped) & 336.1 > 321.3* \\
-maxY(i) > maxY(j) & 304.64 > 265.04* \\
-maxY(j) > maxY(k) & 265.04 > 251.71 \\
-maxY(k) = maxY(l) & 265.04 != 290.03 \\
+area(g) &>& area(h) & 291.2 & >& 276.2* \\
+maxX(f) &>& maxX(e) & 176.1& >& 167.1 \\
+% minimumX(e-flipped) &>& minimumX(f-flipped) & 336.1& >& 321.3 (p<.12) \\
+minX(e-flipped) &>& minX(f-flipped) & 336.1& >& 321.3* \\
+maxY(i) &>& maxY(j) & 304.64 &>& 265.04* \\
+maxY(j) &>& maxY(k) & 265.04 &>& 251.71 \\
+maxY(k) &=& maxY(l) & 265.04 &!=& 290.03 \\
 \hline
 \end{tabular}
-\caption{Comparison of the expected relationships to the mean human responses.}
+\end{center}
+\caption{Comparison of the expected relationships to the mean human responses. *Significance value from t-test (p<0.1).}
 \label{tab:region-location-tests}
-}
 \end{table}
@@
 
-Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.
+Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimulus used in the experiment and the number of no region responses for each stimulus.
 It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.
-Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35.
-Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.
-This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.
+Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flipped), and the circle layout have kappa scores of less than 0.35.
+Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is affected by the location and orientation of objects in the environment.
+An extreme example of this relationship between the configuration of objects in a context and the meaning of a CDSR is that our results indicate that the shared understanding of ``the front of the room'' breaks down in the classroom setting when the chairs in the room are moved into a circle configuration as indicated by the fact that context (a) (Figure \ref{fig:classroom-a}) had an extremly low average pairwise agreement of  $0.0996$).
 
@@
 Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
-The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).
+The case that did not demonstrate the expected relationship was the safety region between the row of turrets (k) and turrets +1(l).
+That is, a single turret does not provide safety from three tanks.
 A possible explanation for this is that there is a domain expertise component to understanding contexts.
@@
 \subsection{Task 2 - Design} 
-In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.
-The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR.
-In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli.
+In task two, a location was marked on the stimulus (using a red dot) and the subject was asked to rank on a five point \textit{Likert scale} how well the location matched the given CDSR, or select not applicable.
+The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textit{sweetspot}, was located near the expected prototypical center of the CDSR.
+In the classroom domain, the three points were located along the horizontal line that bisected the stimulus with P1 positioned on the left of the stimulus, P2 in the center and P3 on the right of the stimulus.
 This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.
-In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli.
-In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows.
-In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.
-P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).
+In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected \textit{sweetspot} and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.
+In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image as follows.
+In stimuli (j) and (k) the point lay between all forces, in stimulus (i), it lay behind the friendly tanks, and in stimulus (l) , it lay behind the lone turret but in front of the row turrets.
+P3 was the expected \textit{sweetspot} in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.
 
@@
 
-%In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region.
+%In the type 1 trials subjects drew a polygon on the visual stimulus that defined the location and extent of the relevant spatial region.
 In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.
 
-%The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched %the described spatial region.
-For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
+%The type 2 trials the same visual stimulus and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on %the visual stimulus (using a red dot) and the subject was asked to rank on a five point \textit{Likert scale} their judgement of how well the location matched %the described spatial region.
+For each stimulus domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.
 
-Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each context.
-In all contexts, the expected sweetspot received the maximum average likert rating that was statistically significant using one-tailed two sample t-tests $(p<.05)$.
-To control for left-right bias, we compared the ratings of sweetspots in the classroom stimuli (c,e and f) with their flipped variants.
+Table \ref{tab:point-likert-scores} lists the average Likert scores for the different points in each context.
+In all contexts, the expected \textit{sweetspot} received the maximum average Likert rating that was statistically significant using one-tailed two-sample t-tests $(p<.05)$.
+To control for left--right bias, we compared the ratings of \textit{sweetspots} in the classroom stimuli (c), (e), and (f) with their flipped variants.
 There was no statistical difference between the stimuli and their flipped variants.
-These likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.
+These Likert ratings on individual stimuli support the claim that people are able to agree on the locations of CDSRs with changing context.
 
-%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points.
-Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli.
-In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
+%There were two of these t-tests carried out for each stimulus, each one tested whether the \textit{sweetspot} ranking was statistically higher than one of the other points.
+Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimulus.
+In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the \textit{sweetspot} was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.
 
@@
 These flipped stimuli included the classroom stimuli (c), (e) and (f).
-This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials.
-Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3.
-To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs.
-In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$.
-In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
+This flipping of stimuli gave us the opportunity to compare the results of the \textit{sweetspots} rankings between this pairs of stimuli trials.
+Given the horizontal flipping of the stimuli it was expected that the \textit{sweetspot} for the stimuli should also be flipped between P1 and P3.
+To test whether or not the expected shift in the \textit{sweetspots} occurred we ran two-tailed two-sample t-tests between the \textit{sweetspots} of the flipped and non-flipped stimuli pairs.
+In all cases the results indicated that there was no statistical difference between the rating of the \textit{sweetspot} for the non-flipped stimuli and the rating of the \textit{sweetspot} for the flipped stimuli using a confidence level of $95\%$.
+In other words, the \textit{sweetspot} flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.
 %lending support for question 2 ...
@@
 Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l).
-In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land.
+In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.
 Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
-As the results in Table \ref{tab:point-likert-scores} show this is the case with this higher rating statistically significant $(p<.05)$.
-Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k).
-Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant.
-Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high.
+This is confirmed by the results in Table \ref{tab:point-likert-scores} and the higher rating is statistically significant $(p<.05)$.
+Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimulus (l) but in no-man's land in stimulus (k).
+Again, the average rating of P2 for stimulus (l) is higher than that of (k), although this difference is not statistically significant.
+Thus, it is unclear the extent to which participants were simulating the outcome of the battle.
+Notice, that the rating of the \textit{sweetspot} for the stimuli (i) and (l) remain quite high.
 This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR.
-Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
+Overall, the analysis of the ratings of the different \textit{sweetspots} across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
 
@@
 %\label{tab:flipped-point-likert-scores}
-%\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
+%\caption{Two Tailed Two Sample t-tests for flipped stimuli pair trials.}
 %\end{table}
 
-Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.
+Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the \textit{sweetspot} was determined by the particular context depicted in the visual stimulus (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the \textit{sweetspot} flipped from P1 to P3 when the context was flipped.
 
@@
 %
-%Task 3: The participant grades a point with respect to membership of the region on a likert scale
+%Task 3: The participant grades a point with respect to membership of the region on a \textit{Likert scale}
 %I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
 %I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
-%Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
+%Assuming that there is a variation across the Likert scores for a particular stimulus then check if their is a correlation between the scores and:
 %(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
@@
 
-\section{Applicability of Previous Approaches to Cognitive Systems }
+\section{Applicability of Previous Work to CDSRs}
 
-%Recall from section \ref{} that in addition to being able to represent these regions for a single environment, the cognitive system (1) must recognize regions of the same type in new environments with new contexts, and (2) use the region for multiple reasoning tasks.
-While spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
-Regions are central to QSR work~\cite{Cohn:2001}.
+Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e.
+that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work, and thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs.
+The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.
+
+Regions are central to qualiative spatial reasoning (QSR)~\cite{Cohn:2001}.
 Regions can either have crisp boundaries, usually defined by geometric shapes (e.g.
 the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g.
-the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g.
+the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g.
 those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}).
 The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
-Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
-%
-%However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
-This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiemnt illustrates two limitations.
-First, Regions are not always physically bounded by objects.
-In the classroom examples, people frequently did not extend the front of the room all way to the desks.
-Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-%
-%The connection between spatial regions and function has been explored in other robotics work, e.g.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points seem unlikely to model the regions with sufficient flexibility.
+
+However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
+This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of region polygons, thereby linking the regions to perceived objects.
+While satisfying the requirements of a cognitive systems solution, considering the application of this model to the  results of our experiment illustrates two limitations.
+First, regions are not always physically bounded by objects.
+In the classroom examples, people frequently did not extend the front of the room all the way to the desks.
+Second, people consider \emph{dynamics} part of the context, e.g.
+in stimuli (l), people did not indicate that it was safe behind the lone turret in front of the others.
+
+The connection between spatial regions and function has been explored in other robotics work, e.g.
 ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
-Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+Such a representation cannot account for the changes in context shown in this experiment, changes which depend on multiple objects, agents, and function.
 
 %JK Previous version above I replaced it with this shorter version below
+% Nick: now reinstating for longer version and editing
 
-Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
-Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
-A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
+% Cognitive systems seek to reproduce the full range of human behavior.
+Spatial reasoning is an important element of human behavior.
+While spatial reasoning is an established area in AI and robotics, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
+Regions are central to qualitative spatial reasoning (QSR) work~\cite{Cohn:2001} and also in the area of cognitive robotics (see, \emph{inter alia,} ~\cite{brenneretal07ijcai,kelleher/costello:09}).
+A fundamental decision when creating a computational model of a CDSR is the representation of its boundary.
 Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
@@
 This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-While satisfying the requirements of a cognitive systems solution, applying this model to the results of our experiment illustrates two limitations.
-First, Regions are not always physically bounded by objects.
-In the classroom examples, people frequently did not extend the front of the room all way to the desks.
-Second, people consider dynamics of a context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-Finally, the connection between spatial regions and function has been explored in other robotics work, e.g.
+One limitation of the \textit{anchor point} approach is that regions are not always physically bounded by objects.
+In the classroom examples, people frequently did not extend the front of the room all the way to the desks.
+The connection between spatial regions and function has been explored in other robotics work, e.g.
 ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
@@
 McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).
-While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.
+Although this approach incorporates functional aspects of the geography, it is unclear how it would apply to the other situations in our experiment.
 
@@
 
-\section{Discussion}
-Cognitive systems do not face problems in isolation.
-Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
-In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.
-We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.
-We outline requirements for a cognitive systems solution and discuss limitations in current approaches from cognitive systems and robotics.
+\section{Conclusion}
+
+% Cognitive systems do not face problems in isolation.  Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
+
+In this paper, we presented a human study which explored the requirements for reasoning about context-dependent spatial regions in a cognitive system.
+We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.
+We posit that a cognitive system should be able to learn new regions, account for changing contexts, and apply this knowledge in multiple reasoning tasks.
 Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.

### 2013-11-13 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.
-To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
-In discussing the results, we consider a number of approaches from current cognitive systems and robotics.
-We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
-Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.
-
-\section{Context-Dependent Spatial Regions}\label{sec:cdsrs}
-
 Consider the following regions shown in Figure \ref{fig:examples}: the front of a classroom, safety in a military engagement, and a geographic bottleneck.
@@
 Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.
-In the rest of the paper we present a study that examines the contextual dependence of these regions.
-Following this we discuss how the results align with previous research and requirements for a cognitive systems model of this ability.
+We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.
+To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.
+In discussing the results, we consider the applicability of approaches from cognitive systems and robotics.
+We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.
+Current approaches have a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.
 
@@
 \centerline{
-\begin{tabular}{cccc}
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle.pdf}\caption{}\label{fig:classroom-a}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-photo.jpg}\caption{}\label{fig:classroom-b}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-shaped.pdf}\caption{}\label{fig:classroom-c}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-photo.jpg}\caption{}\label{fig:classroom-d}\end{subfigure}\\
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-close.pdf}\caption{}\label{fig:classroom-f}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/classroom-far-away.pdf}\caption{}\label{fig:classroom-e}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-1.jpg}\caption{}\label{fig:bottleneck-a}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2.jpg}\caption{}\label{fig:bottleneck-b}\end{subfigure}\\
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close.jpg}\caption{}\label{fig:tanks-a}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/tanks-far.jpg}\caption{}\label{fig:tanks-b}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets.jpeg}\caption{}\label{fig:turrets-a}\end{subfigure}&
-\begin{subfigure}[b]{0.18\textwidth}\includegraphics[width=\textwidth]{figures/turrets-plus-one.jpeg}\caption{}\label{fig:turrets-b}\end{subfigure}
+\begin{tabular}{ccc}
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle.pdf}\caption{}\label{fig:classroom-a}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-photo.jpg}\caption{}\label{fig:classroom-b}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-shaped.pdf}\caption{}\label{fig:classroom-c}\end{subfigure}\\
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/classroom-u-photo.jpg}\caption{}\label{fig:classroom-d}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/classroom-close.pdf}\caption{}\label{fig:classroom-f}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/classroom-far-away.pdf}\caption{}\label{fig:classroom-e}\end{subfigure}\\
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-1.jpg}\caption{}\label{fig:bottleneck-a}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2.jpg}\caption{}\label{fig:bottleneck-b}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close.jpg}\caption{}\label{fig:tanks-a}\end{subfigure}\\
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/tanks-far.jpg}\caption{}\label{fig:tanks-b}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/turrets.jpeg}\caption{}\label{fig:turrets-a}\end{subfigure}&
+\begin{subfigure}[b]{0.33\textwidth}\includegraphics[width=\textwidth]{figures/turrets-plus-one.jpeg}\caption{}\label{fig:turrets-b}\end{subfigure}
 \end{tabular}
@@
 
-\subsection{Task 1 - Design}
+\subsection{Task 1 - Marking Region}
 In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.
@@
 There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.
+We used two methods to analyze the results inter-annotator agreement and comparing features of CDSRs between stimuli.
 
-\subsection{Task 1 - Results}
-We use two methods to analyze the results of the region drawing task.
-First, we use interannotator agreement for two reasons: (1) fair to good overall agreement would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
-Second, we use the extent points of the regions to assess changes between trials.
+Second, we use features of the regions (e.g., area, maximum x-coordinate) to assess changes between trials.
 That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.
+
+\subsubsection{Inter-Annotator Agreement}
 
@@
 Not sure I %captured this
+We use two methods to analyze the results of the region drawing task.
+First, we use inter-annotator agreement for two reasons: (1) fair to good overall agreement would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 
-We use \textit{Cohen's kappa} coefficient, $\kappa$, to measure the agreement between two regions as follows.
+To measure the agreement between two regions, we use \textit{Cohen's kappa} coefficient, $\kappa$, as follows.
 We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon.
@@
 \begin{center}
-\begin{tabular}{l c c}
+\begin{tabular}{l c c c}
 % Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
-Stim ID & Average Pairwise $\kappa$ & No Reg.\\
+Stim ID & Ambiguity Expected? & Average Pairwise $\kappa$ & No Reg.\\
 \hline
 \hline
-(h) & 0.5034 & 0\\ % 1_2
-(g) & 0.4585 & 1\\ % 1_1
-(k) & 0.4548 & 0\\ % 3_3
-(f) & 0.3930 & 0\\ % 2_4
-(i) & 0.3906 & 0\\ % 3_1
-(e-flipped) & 0.3813 & 2\\ % 2_3
-(j) & 0.3787 & 1\\ % 3_2
-(f-flipped) & 0.3720 & 2\\ % 2_5
-(e) & 0.3629 & 1\\ %2_2
-(l) & 0.3164& 0\\ % 3_4
-(c) & 0.2438 & 2\\ % 2_6
-(c-flipped) & 0.2414 & 4\\ %2_7
-(a) & 0.0996 & 10 \\ % 2_1
+(h) & N & 0.5034 & 0\\ % 1_2
+(g) & N & 0.4585 & 1\\ % 1_1
+(k) & N & 0.4548 & 0\\ % 3_3
+(f) & N & 0.3930 & 0\\ % 2_4
+(i) & N & 0.3906 & 0\\ % 3_1
+(e-flipped) & N & 0.3813 & 2\\ % 2_3
+(j) & N & 0.3787 & 1\\ % 3_2
+(f-flipped) & N & 0.3720 & 2\\ % 2_5
+(e) & N & 0.3629 & 1\\ %2_2
+(l) & Y & 0.3164& 0\\ % 3_4
+(c) & Y & 0.2438 & 2\\ % 2_6
+(c-flipped) & Y & 0.2414 & 4\\ %2_7
+(a) & Y & 0.0996 & 10 \\ % 2_1
 \hline
@@
 \label{tab:stimuli-kappa-scores}
-\end{table}
-
-\begin{table}[htb]
-\begin{center}
-\begin{tabular}{r c l r c l}
-\multicolumn{3}{c}{Expected Relationship} & \multicolumn{3}{c}{Human Results}  \\
-\hline
-\hline
-area(g) &>& area(h) & 291.2 & >& 276.2* \\
-maxX(f) &>& maxX(e) & 176.1& >& 167.1 \\
-% minimumX(e-flipped) &>& minimumX(f-flipped) & 336.1& >& 321.3 (p<.12) \\
-minX(e-flipped) &>& minX(f-flipped) & 336.1& >& 321.3* \\
-maxY(i) &>& maxY(j) & 304.64 &>& 265.04* \\
-maxY(j) &>& maxY(k) & 265.04 &>& 251.71 \\
-maxY(k) &=& maxY(l) & 265.04 &!=& 290.03 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Comparison of the expected relationships to the mean human responses. *Significance value from t-test (p<0.1).}
-\label{tab:region-location-tests}
 \end{table}
@@
 
-Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.
+\subsubsection{Predicted Region Changes and Simulation}
+\label{sec:simulation}
+To assess the claim that changing contexts predictably affects the CDSRs, we identified six expected relationships of region features between stimuli.
+For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-x coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).
+CDSRs are dynamic, in the move to safety scenario, we expect that people will simulate the actions of the agents.
+Therefore, we expect the maximum-Y to be larger in the tanks scenario (j) than the turrets (k) because the friendly tanks will pursue the enemy tanks.
+Similarly, we expect maximum-Y coordinate to the be same in the two different turrets scenarios (k) and (l), because people would simulate the results of the battle and realize that a single turret will not provide safety against three tanks.
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{r c l r c l}
+\multicolumn{3}{c}{Expected Relationship} & \multicolumn{3}{c}{Human Results}  \\
+\hline
+\hline
+area(g) &>& area(h) & 291.2 & >& 276.2* \\
+maxX(f) &>& maxX(e) & 176.1& >& 167.1 \\
+% minimumX(e-flipped) &>& minimumX(f-flipped) & 336.1& >& 321.3 (p<.12) \\
+minX(e-flipped) &>& minX(f-flipped) & 336.1& >& 321.3* \\
+maxY(i) &>& maxY(j) & 304.64 &>& 265.04* \\
+maxY(j) &>& maxY(k) & 265.04 &>& 251.71 \\
+maxY(k) &=& maxY(l) & 265.04 &!=& 290.03 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Comparison of the expected relationships to the mean human responses. *Significance value from t-test (p<0.1).}
+\label{tab:region-location-tests}
+\end{table}
+
+Table \ref{tab:region-location-tests} compares the expected relationships between the CDSR features between stimuli.
 In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.
-The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.
-Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.
-The case that did not demonstrate the expected relationship was the safety region between the row of turrets (k) and turrets +1(l).
-That is, a single turret does not provide safety from three tanks.
-A possible explanation for this is that there is a domain expertise component to understanding contexts.
+The last two rows test our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.
+The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that people are simulating actions of other agents in the scene.
+In the other simulation case, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).
+The average maximum-Y was greater for the turret +1 stimuli indicating that people felt that the single turret provided protection from the enemy tanks.
+As discussed in the previous section, we expected some ambiguity in this trial as people likely have different models for the capabilities of the different units in this scenario.
 
@@
 
-\subsection{Task 2 - Design} 
-In task two, a location was marked on the stimulus (using a red dot) and the subject was asked to rank on a five point \textit{Likert scale} how well the location matched the given CDSR, or select not applicable.
+\subsection{Task 2 - Rating Points} 
+In task two, the subject was asked to rank three points on a five point \textit{Likert scale} how well the location matched the given CDSR, or select not applicable.
 The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textit{sweetspot}, was located near the expected prototypical center of the CDSR.
@@
 In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected \textit{sweetspot} and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.
-In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image as follows.
-In stimuli (j) and (k) the point lay between all forces, in stimulus (i), it lay behind the friendly tanks, and in stimulus (l) , it lay behind the lone turret but in front of the row turrets.
-P3 was the expected \textit{sweetspot} in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.
-
+In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image.
+The locations are shown in Figure \ref{fig:locations}.
 %Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though
 
-\subsection{Task 2 - Results}
+\begin{figure}
+\centerline{
+\begin{tabular}{ccc}
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
+\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
+\end{tabular}
+}
+\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
+\label{fig:locations}
+\end{figure}
+
+\subsubsection{Sweetspots}
 
@@
 
+Our hypothesis was that the expected \textit{sweetspots} would score highest among the three points in all situations, and this would demonstrate broad agreement between participants concerning the locations of the CDSRs.
 Table \ref{tab:point-likert-scores} lists the average Likert scores for the different points in each context.
@@
 % The following paragraph should end by saying this about question 1.
-
-Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l).
-In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.
-Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
-This is confirmed by the results in Table \ref{tab:point-likert-scores} and the higher rating is statistically significant $(p<.05)$.
-Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimulus (l) but in no-man's land in stimulus (k).
-Again, the average rating of P2 for stimulus (l) is higher than that of (k), although this difference is not statistically significant.
-Thus, it is unclear the extent to which participants were simulating the outcome of the battle.
-Notice, that the rating of the \textit{sweetspot} for the stimuli (i) and (l) remain quite high.
-This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR.
-Overall, the analysis of the ratings of the different \textit{sweetspots} across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.
-
-%\begin{figure}
-%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
-%\centerline{
-%\begin{tabular}{ccc}
-%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
-%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
-%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
-%\end{tabular}
-%}
-%\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
-%\label{fig:locations}
-%\end{figure}
 
@@
 \end{table}
+
+\subsubsection{Between Stimuli Comparisons}
+
+To assess the claim that changing contexts predictably affects the CDSRs, we can compare the ratings of points between stimuli.
+In the classroom domain, we assume that part of the definition of the front of the room is that where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the u-shaped scenario (c) than the rows scenarios (e) and (f).
+This is confirmed in the results from Table \ref{tab:point-likert-scores}.
+In the military domain,  In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.
+Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
+This is confirmed in Table \ref{tab:point-likert-scores} and is statistically significant $(p<.05)$.
+To assess the extent of participants use of simulation in determining the context, we compare the scores the \textit{no-man's land} point between the tanks (j) to the turrets (k) stimuli and the turrent (k) and turrets +1 (l).
+The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimuli (j) than (k).
+The single turret should offer only temporary protection against the energy tank force, therefore it should not increase the safety rating.
+Neither of these results were found in Table \ref{tab:point-likert-scores}.
+As discussed in Section \ref{sec:simulation}, we expected ambiguity with the turrets +1 stimuli.
+The difference in the rating of P2 between the tanks and turrets scenario provides two alternative explanations regarding how participants interpreted the context:  (1) the turrets were viewed as superior defensive units and, thus, better able to provide safety, or (2) the turrets being fixed could not move away, and thus, establish a more permanent CDSR for safety.
+Further experiments are needed to explore these hypotheses.
 
@@
 
-Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the \textit{sweetspot} was determined by the particular context depicted in the visual stimulus (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the \textit{sweetspot} flipped from P1 to P3 when the context was flipped.
-
 %-----------------
@@
 
-\section{Applicability of Previous Approaches to Cognitive Systems }
+\subsection{discussion}
+Taken together, these results support the following conclusions.
+First, people are able to agree on location and area of CDSRs.
+This agreement is reduced in situations where the domain knowledge of the context is ambiguous, such as, the circular desk arrangement or the defensive capabilities of a single turret.
+Second, changing the context predictably changes the subjects assessments of location, shape and size of the CDSR.
+These results were supported by tasks involving people labeling regions as well as rating individual points.
+On the other hand, our experiments produced mixed results concerning our hypothesis that people use simulation to evaluate dynamic changes to CDSRs.
+Further studies are required to control for the domain knowledge of the participants exploring this question.
 
-%Recall from section \ref{} that in addition to being able to represent these regions for a single environment, the cognitive system (1) must recognize regions of the same type in new environments with new contexts, and (2) use the region for multiple reasoning tasks.
-While spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art.
-Regions are central to qualiative spatial reasoning (QSR) work~\cite{Cohn:2001}.
+\section{Applicability of Previous Work to CDSRs}
+
+Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e.
+that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work, and thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs.
+The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.
+
+Regions are central to qualiative spatial reasoning (QSR)~\cite{Cohn:2001}.
 Regions can either have crisp boundaries, usually defined by geometric shapes (e.g.
 the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g.
-the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g.
+the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g.
 those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}).
 The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.
-Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy.
-%
-%However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
-This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.
-While satisfying the requirements of a cognitive systems solution, applying this model to our results of our experiemnt illustrates two limitations.
-First, Regions are not always physically bounded by objects.
+Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points seem unlikely to model the regions with sufficient flexibility.
+
+However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.
+This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of region polygons, thereby linking the regions to perceived objects.
+While satisfying the requirements of a cognitive systems solution, considering the application of this model to the  results of our experiment illustrates two limitations.
+First, regions are not always physically bounded by objects.
 In the classroom examples, people frequently did not extend the front of the room all the way to the desks.
-Second, people consider dynamics part of the context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others.
-%
-%The connection between spatial regions and function has been explored in other robotics work, e.g.
+Second, people consider \emph{dynamics} part of the context, e.g.
+in stimuli (l), people did not indicate that it was safe behind the lone turret in front of the others.
+
+The connection between spatial regions and function has been explored in other robotics work, e.g.
 ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object.
-Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.
+Such a representation cannot account for the changes in context shown in this experiment, changes which depend on multiple objects, agents, and function.
 
 %JK Previous version above I replaced it with this shorter version below
+% Nick: now reinstating for longer version and editing
 
-Cognitive systems seek to reproduce the full range of human behavior.
+% Cognitive systems seek to reproduce the full range of human behavior.
 Spatial reasoning is an important element of human behavior.
@@
 
-\section{Discussion}
-Cognitive systems do not face problems in isolation.
-Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
-In this paper, we presented a human study to develop requirements for reasoning about context-dependent spatial regions for cognitive systems.
+\section{Conclusion}
+
+% Cognitive systems do not face problems in isolation.  Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.
+
+In this paper, we presented a human study which explored the requirements for reasoning about context-dependent spatial regions in a cognitive system.
 We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.

### 2013-11-13 scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
   \begin{subfigure}[b]{0.32\textwidth}
-  \includegraphics[width=\textwidth]{figures/classroom.png}
+  \includegraphics[width=\textwidth]{figures/classroom-label.png}
 %  \caption{``The front of the classroom'' could be defined as the area between the desks and the whiteboard where people give presentations.}
@@
 \begin{subfigure}[b]{0.32\textwidth}
-  \includegraphics[width=\textwidth]{figures/safety-3.jpg}
+  \includegraphics[width=\textwidth]{figures/tanks-close-label.jpg}
 %  \caption{Safety for the scout is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
@@
 \begin{subfigure}[b]{0.32\textwidth}
-  \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
+  \includegraphics[width=\textwidth]{figures/bottleneck-label.JPG}
 %  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
@@
 \end{tabular}
-  \caption{Examples of spatial regions used as stimuli in our experiments.}
+  \caption{Examples of the spatial regions used as stimuli in our experiments.}
   \label{fig:examples}
@@
 %-----------------
+
+\subsection{discussion}
+Taken together, these results support the following conclusions.
+First, people are able to agree on location and area of CDSRs.
+This agreement is reduced in situations where the domain knowledge of the context is ambiguous, such as, the circular desk arrangement or the defensive capabilities of a single turret.
+Second, changing the context predictably changes the subjects assessments of location, shape and size of the CDSR.
+These results were supported by tasks involving people labeling regions as well as rating individual points.
+On the other hand, our experiments produced mixed results concerning our hypothesis that people use simulation to evaluate dynamic changes to CDSRs.
+Further studies are required to control for the domain knowledge of the participants exploring this question.
 

### 2013-11-15 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 
-Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.
+Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tied to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.
 We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.
@@
 For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom.
-We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli.}.
-To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images.
+We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U-shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli.}.
+To control for left-right bias effects, two versions of the U-shaped layout (c), close rows layout (e) and far rows layout (f) were created by horizontally flipping the images.
 Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).
@@
 \textbf{Participants and Procedure:} The experiment was delivered online.
-43 undergraduate students from two third-level institutes took part in the study.
+Forty-three undergraduate students from two third-level institutes took part in the study.
 To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.
@@
 For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-x coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).
-CDSRs are dynamic, in the move to safety scenario, we expect that people will simulate the actions of the agents.
+CDSRs can be dynamic.
+Thus, in the move to safety scenario, we expect that people will simulate the actions of the agents.
 Therefore, we expect the maximum-Y to be larger in the tanks scenario (j) than the turrets (k) because the friendly tanks will pursue the enemy tanks.
@@
 Table \ref{tab:region-location-tests} compares the expected relationships between the CDSR features between stimuli.
-In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.
-The last two rows test our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.
-The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that people are simulating actions of other agents in the scene.
-In the other simulation case, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).
+In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically affect the perceived extent of a CDSR.
+The last two rows of the table evaluate our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.
+The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that subjects were simulating actions of other agents in the scene.
+In the last row, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).
 The average maximum-Y was greater for the turret +1 stimuli indicating that people felt that the single turret provided protection from the enemy tanks.
@@
 To assess the claim that changing contexts predictably affects the CDSRs, we can compare the ratings of points between stimuli.
-In the classroom domain, we assume that part of the definition of the front of the room is that where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the u-shaped scenario (c) than the rows scenarios (e) and (f).
+In the classroom domain, we assume that part of the definition of the front of the room is where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the U-shaped scenario (c) than the rows scenarios (e) and (f).
 This is confirmed in the results from Table \ref{tab:point-likert-scores}.
-In the military domain,  In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.
+In the military domain, in stimulus (i), the friendly tanks are positioned between the P2 point and the enemy tanks, whereas, in stimulus (j), the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.
 Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j).
 This is confirmed in Table \ref{tab:point-likert-scores} and is statistically significant $(p<.05)$.
-To assess the extent of participants use of simulation in determining the context, we compare the scores the \textit{no-man's land} point between the tanks (j) to the turrets (k) stimuli and the turrent (k) and turrets +1 (l).
+To assess the extent of participants use of simulation in determining the context, we compare the subject's ratings of the \textit{no-man's land} point between the tanks (j) and the turrets (k) stimuli as well as between the turrent (k) and turrets +1 (l).
 The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimuli (j) than (k).
-The single turret should offer only temporary protection against the energy tank force, therefore it should not increase the safety rating.
+The single turret should offer only temporary protection against the enemy tank force, therefore it should not increase the safety rating.
 Neither of these results were found in Table \ref{tab:point-likert-scores}.
@@
 
-\subsection{discussion}
+\subsection{Discussion}
 Taken together, these results support the following conclusions.
@@
 This agreement is reduced in situations where the domain knowledge of the context is ambiguous, such as, the circular desk arrangement or the defensive capabilities of a single turret.
-Second, changing the context predictably changes the subjects assessments of location, shape and size of the CDSR.
+Second, changing the context predictably changes the subjects assessments of location, shape, and size of the CDSR.
 These results were supported by tasks involving people labeling regions as well as rating individual points.
@@
 Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e.
-that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work, and thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs.
+that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work.
+Thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs.
 The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.

### 2013-11-15 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk: fixed some typos

@@
 Consider an agent asked to remain ``in the safe zone'' during a military engagement.
-The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
+The size, shape, and location of this area depends on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.
 We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.
@@
 \subsection{Task 2 - Rating Points} 
-In task two, the subject was asked to rank three points on a five point \textit{Likert scale} how well the location matched the given CDSR, or select not applicable.
+In task two, the subject was asked to rank three points on a five point \textit{Likert scale} with respect to how well the location of the point matched the given CDSR, or select not applicable.
 The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textit{sweetspot}, was located near the expected prototypical center of the CDSR.
@@
 the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g.
-those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}).
+those created by potential fields~\cite{brenneretal07ijcai,kelleher/costello:09} or by Gaussian kernels~\cite{burbridge-dearden12}).
 The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary.

### 2013-11-15 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

@@
 There was no upper bound placed on the number of line segments drawn, but subjects were required to provide at least three line segments or to choose no region before continuing.
-We used two methods to analyze the results inter-annotator agreement and comparing features of CDSRs between stimuli.
-
-Second, we use features of the regions (e.g., area, maximum x-coordinate) to assess changes between trials.
-That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase showing that region extent varies systematically with context changes.
+We analyzed the results in two ways: measuring inter-annotator agreement  and comparing features of CDSRs between stimuli.
 
@@
 Not sure I %captured this
-We use two methods to analyze the results of the region drawing task.
-First, we use inter-annotator agreement for two reasons: (1) fair to good overall agreement would provide evidence that people have a shared interpretation of CDSRs in the given contexts, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
+We tested two hypotheses concerning inter-annotator agreement.
+First, we expect fair to good overall agreement indicating that subjects share the interpretation of the CDSRs in the given contexts.
+Second, we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) desk arrangements as well as the turret + 1 (l) military engagement configuration).
 
-To measure the agreement between two regions, we use \textit{Cohen's kappa} coefficient, $\kappa$, as follows.
+To measure the inter-annotator agreement between two subjects, we use \textit{Cohen's kappa} coefficient, $\kappa$, as follows.
 We generated a set of 47,000 test points that were uniformly distributed across each stimulus and used each subject's polygon to classify the test points as being inside or outside the polygon.
@@
 
-\subsubsection{Predicted Region Changes and Simulation}
+\subsubsection{Predicted Region Changes Across Stimuli}
 \label{sec:simulation}
 To assess the claim that changing contexts predictably affects the CDSRs, we identified six expected relationships of region features between stimuli.
-For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-x coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).
+For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-X coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).
 CDSRs can be dynamic.
@@
 The last two rows of the table evaluate our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.
-The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that subjects were simulating actions of other agents in the scene.
+The comparison between the maximum-Y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that subjects were simulating actions of other agents in the scene.
 In the last row, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).
@@
 To assess the extent of participants use of simulation in determining the context, we compare the subject's ratings of the \textit{no-man's land} point between the tanks (j) and the turrets (k) stimuli as well as between the turrent (k) and turrets +1 (l).
-The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimuli (j) than (k).
+The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimulus (j) than stimulus (k).
 The single turret should offer only temporary protection against the enemy tank force, therefore it should not increase the safety rating.
